### 2023-09-29 main.tex
Initial Overleaf Import

\documentclass
  [ acmsmall
  , pdftex
  , dvipsnames
  , review
  , nonacm
  , screen
  , authordraft
  ]{acmart}

\usepackage[utf8]{inputenc}

% quiver is for quiver figures
\usepackage{quiver}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{xargs}
\usepackage[
  colorinlistoftodos,
  prependcaption,
  % disable,
]{todonotes}
\usepackage{fontenc}

\usepackage{minted}

\usepackage{subcaption}

\usepackage{hyperref}

\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}

\usepackage{listings}
\lstset{
  columns=flexible,
  keepspaces=true,
  showstringspaces=false,
  stringstyle=\slshape\color{green!40!black},
  basicstyle=\ttfamily\small,
  language=Python,
  morekeywords={*, self},
  commentstyle=\slshape\color{black!60},
  tabsize=2,
}

\usepackage{adjustbox}

\input{preamble}
\input{macros}
\input{sample.tikzstyles}
\input{hypergraph.tikzstyles}
\input{hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\usepackage{mathtools,mathrsfs}
\usepackage{tikz-cd}
\usepackage{tikzit}

\usepackage{import}

% category names
\newcommand{\Ecospans}{{\catname{ECsp_{D}(EHyp_{\Sigma})}}}
\newcommand{\MdaEcospans}{{\catname{MAECsp_{D}(EHyp_{\Sigma})}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{\overline{MAECsp_{D}(EHyp_{\Sigma})}}}}

% There is already a remark in the preamble, but it does
% not work for some reason
\newtheorem{remark}{Remark}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

\title{Equivalence Hypergraphs: Semantics of E-Graphs and their Generalization to Monoidal Theories}
\author{Chris Barrett, Dan Ghica, Aleksei Tiurin}

\begin{document}

\begin{abstract}
Abstract.
\end{abstract}

\maketitle

% some sections might be commented out as overleaf compilation times out.

% \input{sections/introduction}
\input{sections/categorical_semantics}
\input{sections/combinatorial_semantics}

\input{sections/soundness_and_completeness}

\input{sections/e-graph-translation}
\input{sections/conclusion}

\subsection*{Acknowledgements}
Sadrzadeh Mehrnoosh

\clearpage
\bibliographystyle{acm}
\bibliography{bibliography}

\end{document}

### 2023-09-29 sections/categorical_semantics.tex
Initial Overleaf Import

\newtheorem*{notation*}{Notation}

\section{Categorical Semantics}

\subsection{Symmetric Monoidal Categories and String Diagrams}

We will assume familiarity with the definition of category,  functor and natural transformation.
\begin{notation*}
We use the following notation: given a catgeory $\mathbb{C}$, we write $ob(\mathbb{C})$ for the set of objects $\mathbb{C}(A,B)$ for hom-sets and $f: A \to B$ for $f \in \mathbb{C}(A,B)$. We write $\textbf{id}_A$ for the identity on $A$ and commonly write $f;g$ for composition in diagram order.  Composition in the usual order is written $g \circ f$. 
We will often omit subscripts on identities and natural transformations where it can be inferred. 
\end{notation*}
\begin{definition}
A \textit{strict symmetric monoidal category (SMC)} $(\mathbb{C}, \otimes, 1)$ is given by a category $\mathbb{C}$ with the following equipment: 
\begin{itemize}
\item a \textit{tensor product} functor $\otimes: \mathbb{C} \times \mathbb{C} \to \mathbb{C}$,
\item a \textit{tensor unit} object $1 \in ob(\mathbb{C})$, 
\item the \textit{symmetry} natural isomorphism: $\sigma: A \otimes B \to B \otimes A$
\item and such that: $A \otimes (B \otimes C) = (A \otimes B) \otimes C$ ,  $1 \otimes A = A$ ,  $A \otimes 1 = A$
\end{itemize}
satisfying certain coherence conditions [MacLane].  A \textit{permutations and products category (PROP)} is a symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$. In this case, we will enote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$.

A \textit{strict symmetric monoidal functor} $F: (\mathbb{C}, \otimes_\mathbb{C}, 1_\mathbb{C}) \to (\mathbb{D}, \otimes_\mathbb{D}, 1_\mathbb{D})$ is given by an underlying functor $F: \mathbb{C} \to \mathbb{D}$ such that 
\[
 1_{\mathbb{D}} = F(1_{\mathbb{C}})
 \qquad F(A \otimes_{\mathbb{C}} B) = F(A) \otimes_{\mathbb{D}} F(B)\ ,
\]
satisfying certain coherence conditions. 
We adopt the convention that $\otimes$ binds more tightly than $(;\!)$. 
A \textit{PROP functor} is additionally identity-on-objects. 
\end{definition}

String diagrams for {SMCs} are constructed as in Figure 1. Note, $\textsf{id}_0$ is denoted by the empty diagram.

\subsection{Symmetric Monoidal Theories and PROPs. }

A presentation of an algebraic theory is traditionally given by a set together with a \textit{signature} of $n$-ary operations and a set of \textit{equations} -- formally,  pairs of terms freely generated over the signature.

For example, the \textit{semilattice signature} consists of a single binary operation $+$.  Given a set $A$, this can be used to freely generate terms according to the following grammar.
\begin{align*}
	\infer{x_1, \ldots, x_n \vdash x_j}{x_i \in A} \qquad \infer{\Gamma \vdash t_1 + t_2}{\Gamma \vdash t_1 && \Gamma \vdash t_2} 
\end{align*}
Such terms may be organized into a Cartesian category, with morphisms given by tuples of terms of the following form,  and composition is given by substitution of terms for free variables.  
\[
	x_1, \ldots, x_n \vdash (t_1, \ldots, t_m) : n \to m\ .
\]
A set of equations can then be given as a set of parallel pairs of morphisms.  
For example, the equations for the theory of semilattices are the following.
\[
	x,y,z \vdash x+(y+z) = (x+y)+z : 3 \to 1\ \qquad  
    x,y \vdash x+y = y+x \qquad 
    x \vdash x+x = x
\]
Now, quotienting the category by the least congruence identifying every such parallel pair, we recover a category whose morphisms are \textit{equivalence classes} of terms modulo the expected equations. 
This gives a categorical presentation of an algebraic theory, known as its \textit{Lawvere category}. 
Importantly,  its category of models (\textit{i.e.}, category of functors to \textbf{Set}) is equivalent to the expected category of models of the algebraic theory in question.

Presenting algebraic theories as Lawvere categories gives a framework in which the concept is amenable to generalization.  We are interested here in symmetric monoidal theories, which are generalizations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.  Lawvere categories will be generalized to particular symmetric monoidal categories called PROPs, but the ideas are otherwise as described above. 
\begin{definition}
A \textit{monoidal signature} $\Sigma$ is given by a set of \textit{generators} $f: n \to m$,  with \textit{arity} $n$ and \textit{coarity} $m$.  A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements. 
\end{definition}
Analogous to freely generating terms over an algebraic signature,  we now consider terms freely generated over a monoidal signature.  These are naturally arranged into a particular form of symmetric monoidal category: a PROP.

We can form a category $\textbf{Sig}$ of monoidal signatures and signature homomorphisms.  There is an evident forgetful functor $U: \textsf{PROP} \to \textsf{Sig}$ from the category of PROPs, which considers every morphism of a PROP as an operation of a signature, forgetting the remaining structure of the category.   Its left adjoint $F$, pictured below, left,  generates a free PROP over a monoidal signature.
% https://q.uiver.app/#q=WzAsNixbMCwxLCJcXHRleHRiZntTaWd9Il0sWzIsMSwiXFx0ZXh0YmZ7UFJPUH0iXSxbMSwxLCJcXHJvdGF0ZWJveHs5MH17XFx2ZGFzaH0iXSxbMywwLCJcXFNpZ21hIl0sWzUsMCwiVUYoXFxTaWdtYSkiXSxbNSwyLCJVKFxcbWF0aGJie0N9KSJdLFswLDEsIkYiLDAseyJjdXJ2ZSI6LTJ9XSxbMSwwLCJVIiwwLHsiY3VydmUiOi0yfV0sWzMsNCwiaSIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzMsNSwiXFxmb3JhbGwgaiIsMix7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzQsNSwiXFxleGlzdHMgVShcXGxsYnJhY2tldCAtIFxccnJicmFja2V0KSJdXQ==
\[\begin{tikzcd}
	&&& \Sigma && {UF(\Sigma)} \\
	{\textsf{Sig}} & {\ \  \perp} & {\textsf{PROP}} \\
	&&&&& {U(\mathbb{C})}
	\arrow["F", bend left, from=2-1, to=2-3]
	\arrow["U", bend left, from=2-3, to=2-1]
	\arrow["i", hook, from=1-4, to=1-6]
	\arrow["{\forall j}", from=1-4, to=3-6]
	\arrow["{\exists U(\llbracket - \rrbracket)}", from=1-6, to=3-6]
\end{tikzcd}\]
The corresponding universal property is given above, right.  The signature morphism $i:\Sigma \to UF(\Sigma)$ is given by the obvious inclusion.  Roughly,  for every PROP $\mathbb{C}$, a signature homomorphism $f: \Sigma \to U(\mathbb{C})$ , which chooses a term of $\mathbb{C}$ for each generator $\Sigma$,  uniquely determines a functor $\llbracket - \rrbracket: F(\Sigma) \to \mathbb{C}$ which preserves this choice.  In the sequel, we will name the free functor $F$ as $\textsf{PROP}(-): \textsf{Sig} \to \textsf{PROP}$.

\begin{definition}
The free \textit{products and permutations category} over a monoidal signature $\Sigma$ is $\textsf{PROP}(\Sigma)$.  
\end{definition}
The free category $\textsf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.

We can now specify an equation associated with a monoidal signature as a pair of parallel morphisms of the form above, leading to the following definition.
\begin{definition}
A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$. 
\end{definition}
Finally, analogous to a Lawvere theory, we define a \textit{symmetric monoidal theory}. 
\begin{definition}
A \textit{symmetric monoidal theory} $\textsf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textsf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
\end{definition}

It will often be convenient to present SMTs in terms of string diagrams [which are technically equiv classes of arrows?!] 
\begin{example}
The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
\[
    \tikzfig{categorical-semantics/Cartesian-equipment}
\]
together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
\[
    \tikzfig{categorical-semantics/Cartesian-theory}
\]
\end{example}

\begin{example}
The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
\[
    \tikzfig{categorical-semantics/Cartesian-naturality}
\]
By Fox's Theoem [cite],  $\textsf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
\end{example}

\subsection{Semilattice Enriched Monoidal Categories}

We will use enrichment to axiomatize the e-box structure as the ``join" or``sum" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.  More precisely our hom-sets will be given the structure of a semilattice.

Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{Semilattice}).  In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
\[
	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
\]
is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.  Technically, this amounts to asking that the tensor product is an enriched functor. 
We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{Semilattice}$.

%\begin{definition}
%A \textit{semigroup} is a pair $(A, +)$ of a set $A$ together with an associative binary operation $+: A \times A \to A$.  We say a semigroup is \textit{commutative} when $+$ is {commutative} and \textit{idempotent} when $+$ is idempotent.  
%Because we will only be concerned with commutative, idempotent semigroups in this paper, we will now often abbreviate the former to simply ``semigroup". 
%\end{definition}

\begin{definition}
A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms. 
\begin{align*}
f ; (g+h) &= f;g + f;h &
(f+g) ; h &= f;h + g;h \\
f \otimes (g+h) &= f \otimes g + f \otimes h & 
(f+g) \otimes h &= f \otimes h + g \otimes h,
%\item A set $ob(\mathbb{C})$ of objects;
%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
\end{align*}
Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.

A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$. 
\end{definition}
Note, in particular,  that semigroups \update{semilattice?} do not require a unit for $+$. {This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct -- a degeneracy we wish to avoid.}

As before, there is an evident forgetful functor $U: \textsf{PROP}^+ \to \textsf{PROP}$ from the category of semilattice enriched PROPs, which forgets the semilattice structure on hom-sets. Its left adjoint $F$, pictured below, generates the free enrichmenet of a PROP. 
% https://q.uiver.app/#q=WzAsNixbMCwxLCJcXHRleHRzZntQUk9QfSJdLFsyLDEsIlxcdGV4dHNme1BST1B9XisiXSxbMSwxLCJcXHBlcnAiXSxbMywwLCJcXFNpZ21hIl0sWzUsMCwiVUYoXFxTaWdtYSkiXSxbNSwyLCJVKFxcbWF0aGJie0N9KSJdLFswLDEsIkYiLDAseyJjdXJ2ZSI6LTJ9XSxbMSwwLCJVIiwwLHsiY3VydmUiOi0yfV0sWzMsNCwiaSIsMCx7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzMsNSwiXFxmb3JhbGwgaiIsMix7InN0eWxlIjp7InRhaWwiOnsibmFtZSI6Imhvb2siLCJzaWRlIjoidG9wIn19fV0sWzQsNSwiXFxleGlzdHMgVShcXGxsYnJhY2tldCAtIFxccnJicmFja2V0KSJdXQ==
\[\begin{tikzcd}
	&&& \mathbb{C} && {UF(\mathbb{C})} \\
	{\textsf{PROP}} & \perp & {\textsf{PROP}^+} \\
	&&&&& {U(\mathbb{D})}
	\arrow["F", bend left, from=2-1, to=2-3]
	\arrow["U", bend left, from=2-3, to=2-1]
	\arrow["i", hook, from=1-4, to=1-6]
	\arrow["{\forall j}"', hook, from=1-4, to=3-6]
	\arrow["{\exists U(\llbracket - \rrbracket)}", from=1-6, to=3-6]
\end{tikzcd}
\]
The corresponding universal property is given above, right. The symmetric monoidal functor $i: \mathbb{C} \to UF(\mathbb{C})$ is given by the obvious inclusion of symmetric monoidal categories.  In the sequel, we will name the free functor $F$ as $(-)^+: \textsf{PROP} \to \textsf{PROP}^+$.  We further name the composition of the functors $\textsf{PROP}(-)$ and $\textsf{SMT}(-)$ with $(-)^+$ as $\textsf{PROP}^+(-), \textsf{SMT}^+(-): \textsf{Sig} \to \textsf{PROP}^+$: these generate the free enriched PROP and freely enriched SMT over a monoidal signature and presentation of a symmetric monidal theory, respectively.

\begin{definition}
The free enriched PROP over a monoidal signature $\Sigma$ is $\textsf{PROP}^+(\Sigma)$. 
The freely enriched SMT over a monoidal signature $\Sigma$ is $\textsf{SMT}^+(\Sigma, \mathcal{E})$. 
\end{definition}
The free category $\textsf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.

To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the binary operation on morphisms.  This will later be used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.

Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.

We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. Indeed, the diagrammatic language itself is not really novel and is similar to \textit{e.g.}, \cite{duncan_generalised_2009}.

\begin{figure}
\[  
    \scalebox{0.75}{
	\tikzfig{categorical-semantics/egraph-strings-equations}
    }
\]
\caption{Equations for a  semilattice enriched symmetric monoidal category}
\label{fig:string-equations}
\end{figure}

### 2023-09-29 sections/combinatorial_semantics.tex
Initial Overleaf Import

\section{Combinatorial semantics}

% TODO: fix union to Kleene star, introduce notation, put explanation out of def environmentnt, fix notation for in(e) and out(e), conflict relation is not transitive

\update{Below I will assume that we work with a simple PROP (not coloured)}

Before we begin we first fix some notation. When $V$ is a set, we denote with $V^{*}$ a set of all finite sequences of the elements of $V$, i.e. $^*$ is a Kleene star. When $\phi : V \to V'$ is a function, we denote with $\phi^{*}$ an extension of $\phi$ to sequences, i.e. $\phi^{*} : V^* \to V'^{*}$ and it applies $\phi$ element-wise. Similarly, in the case when $\psi \subset V \times V$ is a relation: we denote with $\psi^{*}: \subset V^* \times V^*$ the element-wise relation on ordered sequences.

There was a correspondence between hypergraphs and terms in $\textsf{PROP}(\Sigma)$ developed in~\cite{Frobenius, Frobenius2} which we will refer to throughout this section. We first recall the notion of a hypergraph from there.

\Aleksei{Actually, in~\cite{Frobenius} a hypergraph is defined as an object of the corresponding category and labelled hypergraphs are defined as a slice category. This is needed to show adhesiveness. As our category is not adhesive, I think we can follow a more straightforward definition of a tuple.}

\begin{definition}[Hypergraph]
\label{def:hypergraph}

A hypergraph $\mathcal{G}_{\Sigma}$ over a monoidal signature $\Sigma$ is a tuple 
\[
(V_{\mathcal{G}},E_{\mathcal{G}},s,t,l)
\]

where $V_{\mathcal{G}}$ is the set of nodes, $E_{\mathcal{G}}$ is a set of edges, $s : E_{\mathcal{G}} \to V_{\mathcal{G}}^{*}$ is a source function that for each edge returns the ordered list of the edge's source nodes, $t : E_{\mathcal{G}} \to V_{\mathcal{G}}^{*}$ is a target function that for each edge returns the ordered list of the edge's target nodes, $l : E_{\mathcal{G}} \to \Sigma$ is a label function that labels each edge with a generator from monoidal signature $\Sigma$. The label function should satisfy the following conditions: $\forall e,\; \text{s.t.}\; l(e) : m \to n\;. \;|s(e)| = m$ and $\forall e,\; \text{s.t.}\; l(e) : m \to n \;. \;|t(e)| = n$

We will omit the subscripts when they are clear from the context.
\end{definition}

\begin{definition}[Hypergraph homomorphism]

A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that

\begin{enumerate}
        \item $(s_{\mathcal{F}};\phi_V^*)(e) = (\phi_E;s_{\mathcal{G}})(e)$
    \item $(t_{\mathcal{F}};\phi_V^*)(e) = (\phi_E;t_{\mathcal{G}})(e)$
    \item $l_{\mathcal{F}}(e) = \phi_E;l_{\mathcal{G}}(e)$
\end{enumerate}

where $\phi^{*}$ is an element-wise application of $\phi$.
\end{definition}

\begin{definition}[Category $\catname{Hyp_{\Sigma}}$]

Hypergraphs over a monoidal signature $\Sigma$ together with hypergraph homomorphisms form a category $\catname{Hyp_{\Sigma}}$. 
\end{definition}

\begin{definition}[Definition 2.10~\cite{Frobenius}] Let $\mathbb{C}$ be a category with all finite colimits. A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.

A morphism $\alpha : (X \xrightarrow{} A \xleftarrow{} Y) \Rightarrow (X \xrightarrow{} B \xleftarrow{} Y)$ between cospans is an arrow $\alpha : A \to B$ in $\mathbb{C}$ such that the following diagram commutes.

\[
\begin{tikzcd}
                                               & A \arrow[dd, "\alpha"] &                                                \\
X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
                                               & B                      &                                               
\end{tikzcd}
\]

Cospans are isomorphic if $\alpha$ is an isomorphism. For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$. The composition of cospans $X \xrightarrow{} A \xleftarrow{f} Y$ and $Y \xrightarrow{g} B \xleftarrow{} Z$ is $X \xrightarrow{} A +_{f,g} B \xleftarrow{} Z$, constructed by taking a pushout of $f,g$. This data is the category $\catname{Csp(\mathbb{C})}$: the objects are those of $\mathbb{C}$ and the arrows are isomorphism classes of cospans. Finally, $\catname{Csp(\mathbb{C})}$ is a monoidal category: monoidal product is given by a coproduct (denoted with $+$) in $\mathbb{C}$ and the unit is given by the initial object of $\mathcal{C}$ as can be seen below.

\[
\adjustbox{scale=0.75,center}{
% https://q.uiver.app/#q=WzAsMTQsWzAsMCwibiJdLFsxLDAsIlxcbWF0aGNhbHtHfSJdLFsyLDAsIm0iXSxbMSwxLCJcXG90aW1lcyJdLFswLDIsIm4nIl0sWzIsMiwibSciXSxbMSwyLCJcXG1hdGhjYWx7Ryd9Il0sWzMsMSwiOj0iXSxbNCwxLCJuK24nIl0sWzYsMSwiXFxtYXRoY2Fse0d9K1xcbWF0aGNhbHtHJ30iXSxbOCwxLCJtK20nIl0sWzEwLDEsIjAiXSxbMTEsMSwiMCJdLFsxMiwxLCIwIl0sWzAsMSwiZiJdLFsyLDEsImciLDJdLFs0LDYsImYnIl0sWzUsNiwiZyciLDJdLFs4LDksImYrZiciXSxbMTAsOSwiZytnJyIsMl0sWzExLDEyLCJpZF8wIl0sWzEzLDEyLCJpZF8wIiwyXV0=
\begin{tikzcd}
	n & {\mathcal{G}} & m \\
	& \otimes && {:=} & {n+n'} && {\mathcal{G}+\mathcal{G'}} && {m+m'} && 0 & 0 & 0 \\
	{n'} & {\mathcal{G'}} & {m'}
	\arrow["f", from=1-1, to=1-2]
	\arrow["g"', from=1-3, to=1-2]
	\arrow["{f'}", from=3-1, to=3-2]
	\arrow["{g'}"', from=3-3, to=3-2]
	\arrow["{f+f'}", from=2-5, to=2-7]
	\arrow["{g+g'}"', from=2-9, to=2-7]
	\arrow["{id_0}", from=2-11, to=2-12]
	\arrow["{id_0}"', from=2-13, to=2-12]
\end{tikzcd}
}
\]

Symmetry is inherited from $\mathcal{C}$ as well as the coproduct is commutative and is given below

\[
n + m \xrightarrow{\sigma_{n,m}} m + n \xleftarrow{id + id} m + n
\]

Naturality is implied by $+$ being a coproduct~\cite{Frobenius}.

\end{definition}

\begin{remark}

Taking isomorphism classes of cospans as morphisms turns $\catname{Csp(\mathbb{C})}$ into a proper category since pushouts are defined up to isomorphism and associative.
\end{remark}

When $\mathbb{C} = Hyp_{\Sigma}$ we get a category of cospans of hypergraphs. An empty hypergraph is the initial object in $Hyp_{\Sigma}$ and coproduct and pushout are defined on the underlying sets of nodes and edges.
Following~\cite{Frobenius2} morphisms in an SMC with a signature $\Sigma$ can be interpreted as particular cospans of hypergraphs.

\begin{definition}[Category $\;\catname{Csp_{D}(Hyp_{\Sigma})}$]
\label{def:cspd}
Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
\end{definition}

In particular, this means that any morphism in $\catname{Csp_{D}(Hyp_{\Sigma})}$ is of the form

\[
n \xrightarrow{} \mathcal{F} \xleftarrow{} m
\]

where $n,\;m$ are discrete hypergraphs and $\mathcal{F}$ is an arbitrary hypergraph. When string diagrams are interpreted as cospans of hypergraphs $n,m$ correspond to input and output interfaces of a string diagram respectively.

A pictorial representation of such a cospan is depicted below in Figure~\ref{fig:hypergraph_and_string_diagram}. We use node labels to define mappings $n \to \mathcal{F}$ and $m \to \mathcal{F}$.

\begin{figure}
    \centering
    \begin{subfigure}{0.45\linewidth}
    \[
    \scalebox{0.75}{
	\tikzfig{combinatorial_semantics/mda-hypergraph-example}
   }
   \]
    \end{subfigure}
    \begin{subfigure}{0.45\linewidth}
        \[
        \scalebox{0.75}{
        \tikzfig{combinatorial_semantics/mda-hypergraph-example-string}
        }
        \]
    \end{subfigure}
    \caption{Cospan of hypergraphs and corresponding string diagram}
    \label{fig:hypergraph_and_string_diagram}
\end{figure}

\begin{definition}[Degree of a node] 
The in-degree of a node $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in s(e)$  Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in t(e)$.

\end{definition}

\begin{definition}[Path]
    A path from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{F}$ is an ordered sequence of edges $e_i \in E_{F}$ : $[e_0, e_1, \ldots, e_{n-1}]$ such that for all $i < n - 1$ some node $v \in t(e_i)$ is also in $s(e_{i+1})$.
\end{definition}

\begin{definition}[Acyclicity]
    A hypergraph is acyclic if it contains no paths of length $n > 0$ that start and end in the same edge $e$.
\end{definition}

\begin{definition}[Monogamy]
\label{def:monogamy_hyp}

We call a cospan $n \xrightarrow{f} \mathcal{F} \xleftarrow{g} m$ monogamous directed acyclic if

\begin{enumerate}
    \item $\mathcal{F}$ is a directed acyclic hypergraph
    \item $f$ and $g$ are monos
    \item the in-degree and out-degree of every vertex of $\mathcal{F}$ is at most 1
    \item vertices of $\mathcal{F}$ with in-degree 0 are precisely the image of $f$
    \item vertices of $\mathcal{F}$ with out-degree 0 are precisely the image of $g$
\end{enumerate}

\end{definition}

\begin{definition}[Category $\catname{MACsp_{D}(Hyp_{\Sigma})}$]
    When we narrow our attention to cospans (morphisms) of $\catname{Csp_{D}(Hyp_{\Sigma})}$ which are monogamous we get the category of monogamous directed acyclic (mda-) cospans of hypegraphs $\catname{MACsp_{D}(Hyp_{\Sigma})}$. Vertical and horizontal compositions of mda-cospans are mda-cospans, identities and symmetries are mda-cospans, and hence this is a proper category~\cite{Frobenius2}.
\end{definition}

\begin{theorem}[Corollary 26~\cite{Frobenius2}]
    $\textsf{PROP}(\Sigma) \cong \catname{MACsp_{D}(Hyp_{\Sigma})}$
\end{theorem}
This theorem essentially states a correspondence between morphisms in SMC (string diagrams) and mda-cospans of hypergraphs.
% This section presents a combinatorial structure that defines our e-hypergraphs.

% We closely follow a set of definitions designed in~\cite{fscd}. We extend this set to support an additional e-structure.

Now we are ready to define e-hypergraphs which are a combinatorial representation of enriched PROPs.

\begin{definition}[E-hypergraph]

    An e-hypergraph $\mathcal{G}$ over a monoidal signature $\Sigma$ is a tuple \[(V_{\mathcal{G}},E_{\mathcal{G}},s,t,l,<_c,\#)\] where $(V_{\mathcal{G}},E_{\mathcal{G}},s,t)$ are the same as in~\ref{def:hypergraph}.
    The labelling $l$ is modified to include an extra value $\bot \;$ $l :E_{\mathcal{G}} \to \Sigma + 1$. With an abuse of notation, we will confuse $\mathcal{G}$ with $V_{\mathcal{G}} \cup E_{\mathcal{G}}$, when treating it as a set.

    $<_c \subset \mathcal{G} \times \mathcal{G}$ is a child relation imposing a partial order on $\mathcal{G}$.
    We require that for a given $x \in \mathcal{G}$ a set of all predecessors of $x\;$ $[x) = \{x' \in \mathcal{G} \;||\; x' <_c x\; \}$ is finite and consists of edges $e \in E_{\mathcal{G}}$ such that $l(e) = \bot$ only. Next, each $x$ has at most one immediate predecessor. Such order was called a \textbf{well-founded arborescent} strict partial order in~\cite{orders_and_events}.
    Similarly, a set of all successors is $(x] = \{x' \in \mathcal{G} \; ||\; x < x'\}$ where $x'$ can be a node or an edge. 
    % An edge $e \in E_{\mathcal{G}}$ is unlabelled, i.e. $l(e) = \bot$ iff $(e] \not = \varnothing$.
    Finally, we require that $<_c$ respects connectivity, i.e. if $v \in s(e)$ then $e' < e$ iff $e' <_c v$.

    $\langle \mathcal{G}, <_c, \# \rangle$ is an event-structure where $\# \subset \mathcal{G} \times \mathcal{G}$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
    \begin{enumerate}
        \item 
        \[
    \forall x_1, x_2, x_3 \in \mathcal{G} \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
    \] 
    if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
    \item $<_c$ and $\#$ are mutually exclusive
    \item $\#$ is defined \textit{only} for $x \in \mathcal{G}$ such that $[x) \not \eq \varnothing$
    \item $\#$ should respect connectivity
    \item \update{Complement of $\#$ should be transitive}
    \item for all distinct $x_1, x_2, x_3 \in V \cup E$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
    \item for all $x, x' \in \mathcal{G}$ $x \#_{\mu} x'$ implies $[x) = [x')$
        \end{enumerate}

The last two requirements are known as \textit{confusion freeness}~\cite{ConfusionFreeEvents}.
\end{definition}

Intuitively, the child and conflict relation together make it possible to model e-classes in our e-hypergraphs: an unlabelled edge $e$ represents an e-class where the children of $e$ that are in immediate conflict correspond to nodes and edges in different e-class instances.
Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$. Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.

\begin{lemma}[Complement of $\#$ is equivalence relation]
\begin{proof}
As $\#$ is irreflexive then its complement $\bar{\#}$ is reflexive because $\forall x \; . \; (x,x) \not \in \#$. Because $\#$ is symmetric so is $\bar{\#}$, otherwise if $(x_1, x_2) \in \bar{\#}$ and $(x_2,x_1) \not \in \bar{\#}$ then $(x_2,x_1) \in {\#}$ which is symmetric and hence $(x_1,x_2)$ should be in ${\#}$. Finally, the complement is transitive because of (5).

\end{proof}
\end{lemma}

\begin{remark}
    We will further call edges $e$ such that $l(e) = \bot$ \textit{hierarchical}. We will also refer to edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ as top-level.
\end{remark}

% We associate an e-class with a hierarchical edge, i.e. an edge with children. 
% The parent functions satisfy some conditions. First, an edge and any of its source and target vertices must have the same parent: $p_V (v) = p_E(e) = p_V (v')$ for all $v \in s(e)$ and $v' \in t(e)$, respectively.
% Second, the parent relation must be acyclic.
% More precisely, we assume for all $e \in E$ there is some $k \geq 1$ such that $(p_E,\bot)k(e) = \bot$ where $\bot$ is the element of $1$ and $p_{E,\bot} :E+1 \to E+1$ is the extension of $p_E$ adding $p_{E,\bot}(\bot)=\bot$.

% $\#_{V}$ and $\#_{E}$ are conflict relations that tell whether nodes and edges belong to different instances within an e-class: if the elements are in conflict, they belong to different instances. \update{E-classes are represented by edges $e : l_E(e) = \bot$. Hence, the conflict relation $\#_{V}$ between $v_1$ and $v_2$ is defined only when $p_V(e_1) \not \eq \bot$ and $p_V(e_2) \not \eq \bot$. Likewise the conflict relation $\#_{E}$ between $e_1$ and $e_2$ is defined only when $p_{E}(e_1) \not \eq \bot$ and $p_{E}(e_2) \not \eq \bot$}

% The relation should respect connectivity, that is, if $e_1 \#_E e_2$ then $s(e_1) \#_V^* s(e_2)$ and $t(e_1) \#_V^* t(e_2)$, where $\#_V^*$ is element-wise relation. 
% % This guarantees that related nodes and edges are contained within the same e-class.
% \update{
% We also require that if $v_1 \#_{V} v_2$ then $\forall e_1,e_2 : v_1 \in s(e_1)$ and $v_2 \in s(e_2)\; e_1 \#_{E} e_2$  and the same for the case when $v_i \in t(e_1),\;v_j \in t(e_2)$. Finally, the relation should be \textit{hereditary}: if $e_1 \#_E e_2 \text{ and } \forall n,m,k,q > 0, \; v_1,v_2 \in V_{\mathcal{G}} \text{ and } e_1', e_2' \in E_{\mathcal{G}} \text{ s.t. } p^n(v_1) = e_1, \; p^m(v_2) = e_2, \; p^k(e_1') = e_1, \; p^q(e_2') = e_2\; . \;
%  v_1 \#_V v_2 \text{ and } e_1' \#_E e_2'$. Where $p^n(v),\; n > 0$ is $n^{th}$ parent of $v$} The relation is also irreflexive and symmetric.

\begin{example}
An example of an e-hypergraph can be found in Figure~\ref{fig:e-cospan}. The e-hypergraph is depicted in the area shaded grey. Edges $g$ and $h$ are in conflict (as well as their incoming and outcoming nodes) and they share the same predecessor --- the edge shaded with lighter colour. Vertical line delimits the edges and nodes which are in conflict. Intuitively, such lines delimit e-class instances.
\end{example}

\begin{definition}[Convex subgraph]
    We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ convex if for all $v_i, v_j \in V_{\mathcal{H}}$ all $v_k \in path(v_i, v_j)$ are also in $\mathcal{H}$.
\end{definition}

\begin{definition}[Down-closed subgraph]
    We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ down-closed if for all $e \in E_{\mathcal{H}}$ all elements of $(e]$ are also in $\mathcal{H}$.
\end{definition}

 % \begin{remark}
 %    We will use $p_{E}^n(e)$ as a syntactic sugar for $p_{E,\bot} ; p_{E,\bot} \ldots p_{E,\bot}(e)$ to denote the n$^{th}$ parent of an edge $e$. Similarly
 %     we will use $p_{V}^n(v)$ as a syntactic sugar for $p_{E}^{n-1};p_{V}(v)$ to denote the n$^{th}$ parent of a node $v$.
 % \end{remark}

\begin{definition}{E-hypergraphs homomorphism}

A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that

\begin{enumerate}
    \item $\phi$ is hypergraph homomorphism

    \item if $x_1$ is the immediate predecessor of $x_2$ in $\mathcal{F}$ then $\phi(x_1)$ is the immediate predecessor of $\phi(x_2)$ in $\mathcal{G}$

    % \item
    % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
    \question{
        \item
    if $x_1 \#_{\mathcal{F}} x_2$ then $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
    }
\end{enumerate}

\update{Let's denote the fact of $x_1$ being the immediate predecessor of $x_2$ by $x_1 <_{\mu} x_2$. Also, note that preserving the immediate predecessor implies preserving order. Homomorphism definition seems to be fine: by not requiring the \textit{only if} direction in bullet 2 we basically allow either of two things

\begin{itemize}
    \item $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 <_{\mu} x_2$
    \item $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1$ is unrelated to $x_2$
\end{itemize}

We can not have $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 < x_2$, i.e. the case where $x_1$ is not the immediate predecessor of $x_2$ because then it would mean that there exists at least $x_3$ such that $x_1 <_{\mu} x_2 <_{\mu} x_3$ and because $\phi$ should preserve them it would mean that $\phi(x_1) <_{\mu} \phi(x_3) <_{\mu} \phi(x_2)$ implying that $\phi(x_2)$ has two immediate predecessors: $\phi(x_1)$ and $\phi(x_3)$. Of course, if $\phi$ does not identify $x_1$ and $x_3$.

The first bullet is fine because it means that $\phi$ reflects the immediate predecessor. The second bullet leaves us with two options depicted below.

}

\begin{figure}
    \centering
\begin{subfigure}{0.4\linewidth}

\scalebox{0.5}{
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node, label={above:$x_1$}] (19) at (-3.5, -4.5) {};
		\node [style=node, label={above:$x_2$}] (20) at (-4.5, -6.5) {};
		\node [style=node, label={above:$x_3$}] (21) at (-2.5, -6.5) {};
		\node [style=node, label={above:$x_4$}] (22) at (-3.5, -8.5) {};
		\node [style=node, label={above:$x_5$}] (23) at (-1.5, -8.5) {};
		\node [style=node, label={right:$x_2$}] (24) at (4.5, -5.5) {};
		\node [style=node, label={above:$x_1$}] (25) at (4.5, -2.5) {};
		\node [style=node, label={below:$x_5$}] (26) at (4.5, -11.5) {};
		\node [style=none] (27) at (4.5, -7.5) {};
		\node [style=none] (28) at (4.5, -9.5) {};
		\node [style=none] (29) at (4.5, -8.5) {$\vdots$};
		\node [style=none] (30) at (0.5, -6.5) {};
		\node [style=none] (31) at (2.5, -6.5) {};
		\node [style=none] (32) at (1.5, -5.5) {$\phi$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (19) to (20);
		\draw (19) to (21);
		\draw (21) to (22);
		\draw (21) to (23);
		\draw (24) to (27.center);
		\draw (26) to (28.center);
		\draw (25) to (24);
		\draw [style=diredge] (30.center) to (31.center);
	\end{pgfonlayer}
\end{tikzpicture}

}

\caption{Case 1}
\end{subfigure}
\hfill
\begin{subfigure}{0.4\linewidth}

\scalebox{0.5}{
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node, label={above:$x_1$}] (19) at (-6.5, -4.5) {};
		\node [style=node, label={above:$x_2$}] (20) at (-7.5, -6.5) {};
		\node [style=node, label={above:$x_3$}] (21) at (-5.5, -6.5) {};
		\node [style=node, label={above:$x_4$}] (30) at (-3, -4.5) {};
		\node [style=node, label={above:$x_1$}] (31) at (2.5, -4.5) {};
		\node [style=node, label={above:$x_2$}] (32) at (1.5, -6.5) {};
		\node [style=node, label={above:$x_3$}] (33) at (2.75, -6.5) {};
		\node [style=node, label={above:$x_4$}] (34) at (4, -6.5) {};
		\node [style=none] (35) at (-1.5, -5.5) {};
		\node [style=none] (36) at (-0.5, -5.5) {};
		\node [style=none] (37) at (-1, -5) {$\phi$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (19) to (20);
		\draw (19) to (21);
		\draw (31) to (32);
		\draw (31) to (33);
		\draw (31) to (34);
		\draw [style=diredge] (35.center) to (36.center);
	\end{pgfonlayer}
\end{tikzpicture}

}

\caption{Case 2}

\end{subfigure}
\end{figure}

\update{Case (a) above is impossible on two occasions: for the image of $x_5$ to be a successor of the image of $x_2$, $x_5$ itself should be hierarchical. Then, because $\phi$ should preserve the immediate predecessor, the image of $x_3$ should be the immediate predecessor of the image of $x_5$ and the image of $x_1$ should be the immediate predecessor of the image of $x_3$ and of the image of $x_2$. Since the image of $x_2$ is a successor of the image of $x_5$ on the right there should be two paths from the image of $x_5$ to the image of $x_1$ one through the image of $x_3$ and another through the image of $x_2$, this effectively means that some node on this path has two immediate predecessors.

The case (b) above typechecks (for typed e-hypergraphs) only if $x_4$ is of type $0 \to 0$. This case is fine because we always can append, e.g. an edge, to a hierarchical edge using the semilattice equations (the one for tensoring the box). This also assumes that the image is down-closed.
}

\question{There is an issue with requiring that $\phi$ should reflect the conflict though. When we consider e-cospans and the maps from the interfaces to the carrier we map some nodes from the interface to conflicted nodes in the carrier. However, these nodes in the interface can not be in conflict as they have no predecessors. We can ask that this is a property of a matching morphism $m$ rather than arbitrary homomorphism. The same was done for down-closed-ness, for example.
}

If we further require that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we will call such homomorphism \textit{strict}.
\end{definition}

% The last 2 conditions say that if two elements are from different instances of an e-class, then, under a homomorphism they shall remain in different instances. that is, homomorphism should preserve and reflect conflicts.

% \question{If we do not introduce $\bot$ as a parent, we do not need a notion of a weaker homomorphism. That is point 2 above suffices. Is is enough or shall we require the preservation of the immediate predecessor?}

In other words, homomorphism should preserve immediate predecessors and preserve and reflect the conflict relation.

\begin{remark}
    When we consider an image of $\mathcal{F}$ under $\phi$ in $\mathcal{G}$ as an e-hypergraph in its own rights we assign $[x) = \varnothing$ for all $x$ whose immediate predecessor $x'$ not in the image of $\phi$.
\end{remark}

\begin{definition}[Category $\catname{EHyp_{\Sigma}}$]
Such e-hypergraphs form a category $\catname{EHyp_{\Sigma}}$ where morphisms are e-hypergraph homomorphisms defined above.
\end{definition}

\begin{remark}
    Similarly to $\catname{Hyp_{\Sigma}}$ pushouts in $\catname{EHyp_{\Sigma}}$ are computed on the underlying sets of nodes and edges. However, unlike $\catname{Hyp_{\Sigma}}$ $\catname{EHyp_{\Sigma}}$ is not \textit{adhesive} meaning that not all pushouts along monos exist. Consider an example span below in Figure~\ref{fig:pushout_fail}.

\begin{figure}[h!]
    \centering
    \[    
    \scalebox{0.5}{\tikzfig{combinatorial_semantics/pushout_fail_example}
    }
    \]
    \caption{Pushout failure}
    \label{fig:pushout_fail}
\end{figure}

    In order to build a commuting square which is a pushout the immediate predecessors of $f$ should be identified in the pushout because any element of an e-hypergraph has only one immediate predecessor. However, such identification is not possible as these predecessors have different types. Luckily, we do not need arbitrary pushouts and pushouts that we actually need do exist.
\end{remark}

To build a proper correspondence between terms in an enriched category and e-hypergraphs we introduce e-hypergraphs with interfaces.

\update{I will build below $\catname{ECsp_{D}(EHyp_{\Sigma})}$ (E for extended) instead of $ECsp(\mathcal{C})$ for arbitrary $\mathcal{C}$ with colimits to avoid talking about pushouts in general below, because we use set difference.}
\begin{definition}[Category $\catname{ECsp_{D}(EHyp_{\Sigma})}$]
    We will build the category of \textit{extended} cospans of e-hypergraphs as follows.
    Let the objects of $\catname{ECsp_{D}(EHyp_{\Sigma})}$ be only discrete e-hypergraphs of $\catname{EHyp_{\Sigma}}$ and morphisms from $ext(X) \to ext(Y)$ be the cospans (ext stands for external)

    \[
    ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
    \] where $A \in obj(EHyp_{\Sigma})$.

    \update{Can I say that a morphism between two such cospans is 3 arrows?}

    Two cospans

% https://q.uiver.app/#q=WzAsOCxbMCwxLCJleHQoWCkiXSxbMSwwLCJYIl0sWzIsMCwiQSJdLFszLDAsIn5ZIl0sWzQsMSwiZXh0KFkpIl0sWzEsMiwiWCciXSxbMiwyLCJBJyJdLFszLDIsIlknIl0sWzAsMSwiIiwyLHsiY3VydmUiOi0yfV0sWzEsMl0sWzMsMl0sWzQsMywiIiwwLHsiY3VydmUiOjJ9XSxbNCw3LCIiLDIseyJjdXJ2ZSI6LTJ9XSxbNyw2XSxbNSw2XSxbMCw1LCIiLDAseyJjdXJ2ZSI6Mn1dLFsxLDUsIlxcYWxwaGEiLDFdLFsyLDYsIlxcYmV0YSIsMV0sWzMsNywiXFxnYW1tYSIsMV1d

\[
\begin{tikzcd}
	& X & A & {~Y} \\
	{ext(X)} &&&& {ext(Y)} \\
	& {X'} & {A'} & {Y'}
	\arrow[curve={height=-12pt}, from=2-1, to=1-2]
	\arrow[from=1-2, to=1-3]
	\arrow[from=1-4, to=1-3]
	\arrow[curve={height=12pt}, from=2-5, to=1-4]
	\arrow[curve={height=-12pt}, from=2-5, to=3-4]
	\arrow[from=3-4, to=3-3]
	\arrow[from=3-2, to=3-3]
	\arrow[curve={height=12pt}, from=2-1, to=3-2]
	\arrow["\alpha"{description}, from=1-2, to=3-2]
	\arrow["\beta"{description}, from=1-3, to=3-3]
	\arrow["\gamma"{description}, from=1-4, to=3-4]
\end{tikzcd}
\]

are isomorphic if $\alpha$, $\beta$ and $\gamma$ are isomorphisms and the diagram above commutes.

An identity cospan is

\[
X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
\]

Composition of two morphisms $f : ext(X) \to ext(Y)$ and $g : ext(Y) \to ext(Z)$ is computed in two stages. First, $A +_{f_{ext};f, g_{ext};g} B$ is computed which is a pushout below

% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
\[\begin{tikzcd}
	{ext(X)} &&& Y & {ext(Y)} & {Y'} &&& {ext(Z)} \\
	& X &&&&&& Z \\
	&& A &&&& B \\
	&&&& {^{\ulcorner}A+_{f_{ext};f,g_{ext};g}B}
	\arrow[from=1-1, to=2-2]
	\arrow[from=2-2, to=3-3]
	\arrow["f", from=1-4, to=3-3]
	\arrow["{f_{ext}}", from=1-5, to=1-4]
	\arrow["{g_{ext}}"', from=1-5, to=1-6]
	\arrow["g"', from=1-6, to=3-7]
	\arrow[from=2-8, to=3-7]
	\arrow[from=1-9, to=2-8]
	\arrow[from=3-3, to=4-5]
	\arrow[from=3-7, to=4-5]
\end{tikzcd}
\]

Then, new legs for $A +_{f_{ext};f,g_{ext};g} B$ are computed as below

\[
ext(X) \xrightarrow{f_{ext}} X \uplus (Y' \setminus ext(Y)) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus ext(Y)) \xleftarrow{g_{ext}} ext(Z)
\]

since $X,Y,Z,Y', ext(Y)$ are all discrete, $\uplus$ and $\setminus$ are standard set operations. $\setminus$ could be also formalised further as the following pushout complement

\[
\begin{tikzcd}
ext(Y) \arrow[d] & 0 \arrow[l] \arrow[d]    \\
Y'          & Y' \setminus ext(Y) \arrow[l]
\end{tikzcd}
\]

and then the disjoint union is just a coproduct. Defining cospans up to isomorphism, associativity of coproducts and pushouts, and the presence of an id-cospan turns this construction into a valid category.

Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into a SMC. More precisely, below are cospans for tensor product, its unit and symmetry.

% https://q.uiver.app/#q=WzAsMTcsWzAsMCwiZXh0KFgpIl0sWzEsMCwiWCJdLFsyLDAsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzAsMiwiZXh0KFgnKSJdLFsxLDIsIlgnIl0sWzIsMiwiQiJdLFszLDIsIlknIl0sWzQsMiwiZXh0KFknKSJdLFsyLDEsIlxcb3RpbWVzIl0sWzUsMSwiOj0iXSxbNiwxLCJleHQoWCkrZXh0KFgnKSJdLFs4LDEsIlgrWCciXSxbMTAsMSwiQStCIl0sWzEyLDEsIlkrWSciXSxbMTQsMSwiZXh0KFkpK2V4dChZJykiXSxbMCwxLCJmX3tleHR9Il0sWzQsMywiZ197ZXh0fSIsMl0sWzMsMiwiZyIsMl0sWzEsMiwiZiJdLFs1LDYsImYnX3tleHR9Il0sWzYsNywiZiciXSxbOCw3LCJnJyIsMl0sWzksOCwiZydfe2V4dH0iLDJdLFsxMiwxMywiZl97ZXh0fStmJ197ZXh0fSJdLFsxMywxNCwiZitmJyJdLFsxNSwxNCwiZytnJyIsMl0sWzE2LDE1LCJnX3tleHR9ICsgZydfe2V4dH0iLDJdXQ==
\[
\adjustbox{scale=0.6, center}{
\begin{tikzcd}
	{ext(X)} & X & A & Y & {ext(Y)} \\
	&& \otimes &&& {:=} & {ext(X)+ext(X')} && {X+X'} && {A+B} && {Y+Y'} && {ext(Y)+ext(Y')} \\
	{ext(X')} & {X'} & B & {Y'} & {ext(Y')}
	\arrow["{f_{ext}}", from=1-1, to=1-2]
	\arrow["{g_{ext}}"', from=1-5, to=1-4]
	\arrow["g"', from=1-4, to=1-3]
	\arrow["f", from=1-2, to=1-3]
	\arrow["{f'_{ext}}", from=3-1, to=3-2]
	\arrow["{f'}", from=3-2, to=3-3]
	\arrow["{g'}"', from=3-4, to=3-3]
	\arrow["{g'_{ext}}"', from=3-5, to=3-4]
	\arrow["{f_{ext}+f'_{ext}}", from=2-7, to=2-9]
	\arrow["{f+f'}", from=2-9, to=2-11]
	\arrow["{g+g'}"', from=2-13, to=2-11]
	\arrow["{g_{ext} + g'_{ext}}"', from=2-15, to=2-13]
\end{tikzcd}
}
\]

% https://q.uiver.app/#q=WzAsNSxbMCwwLCIwIl0sWzEsMCwiMCJdLFsyLDAsIjAiXSxbMywwLCIwIl0sWzQsMCwiMCJdLFswLDEsImlkXzAiXSxbMSwyLCJpZF8wIl0sWzMsMiwiaWRfMCIsMl0sWzQsMywiaWRfMCIsMl1d
\[
\adjustbox{scale=0.75, center}{
\begin{tikzcd}
	0 & 0 & 0 & 0 & 0
	\arrow["{id_0}", from=1-1, to=1-2]
	\arrow["{id_0}", from=1-2, to=1-3]
	\arrow["{id_0}"', from=1-4, to=1-3]
	\arrow["{id_0}"', from=1-5, to=1-4]
\end{tikzcd}
}
\]

% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
\[
\adjustbox{scale=0.75, center}{
\begin{tikzcd}
	{X + Y} && {X + Y} && {Y + X} && {Y + X} && {Y + X}
	\arrow["{id_X + id_Y}", from=1-1, to=1-3]
	\arrow["{\sigma_{X,Y}}", from=1-3, to=1-5]
	\arrow["{id_Y + id_X}"', from=1-9, to=1-7]
	\arrow["{id_Y + id_X}"', from=1-7, to=1-5]
\end{tikzcd}
}
\]

\end{definition}

\begin{remark}
    When writing things like $Y' \setminus ext(Y)$ we mean $Y' \setminus g(Y')$ and so on.
\end{remark}

\begin{remark}
An example of such a morphism in $\catname{ECsp_{D}(EHyp_{\Sigma})}$ can be found in Figure~\ref{fig:e-cospan}. The legs of a cospan are depicted on top and at the bottom from the grey-shaded area. These rectangles correspond to $ext(X)$, $X$, $Y$, and $ext(Y)$ respectively. In order to not clutter the figures we will collapse the legs, e.g. we will 
 collapse $ext(X)$ and $X$ into one rectangle and delimit the image of the former within the latter with a vertical dashed line.
\end{remark}

\begin{figure}[h!]
    \centering
    \[
    \scalebox{0.5}{
    \tikzfig{combinatorial_semantics/e-hypergraph-example}
    }
    \]
    \caption{Morphism in $\catname{ECsp_{D}(EHyp_{\Sigma})}$}
    \label{fig:e-cospan}
\end{figure}

Consider a cospan

\[
    ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
\]
\label{mda-cospan}

\begin{definition}[Interface]

When talking about cospans we will call $X$ and $Y$ input and output interfaces respectively. We will also call $ext(X)$ and $ext(Y)$ the outermost or external input and output interfaces and $X \setminus ext(X)$ and $Y \setminus ext(Z)$ the internal input and output interfaces. By abusing the notation we will also refer to the images of $ext(X)$, $X$, $Y$, $ext(Y)$, $X \setminus ext(X)$, $Y \setminus ext(Y)$ as to interfaces of the corresponding type. Graphically we distinguish the outermost part of the interface with a vertical line.
\end{definition}

\begin{remark}
    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y$. Consider an example in Figure~\ref{fig:e-hyp_composition_example} below. The pushout is depicted in Figure~\ref{fig:e-hyp_composition_example_2}. New interfaces are shown in Figure~\ref{fig:e-hyp_composition_example_3}. They are computed as shown above: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken. Similarly for the output interface. In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$. Note that pushout for such composition always exists as the nodes that are being identified are unrelated: they are not in conflict with anything and the sets of their predecessors are empty.
\end{remark}

\begin{figure}
    \centering

\begin{subfigure}{0.55\linewidth}

\centering
\[
    \scalebox{0.35}{

    \tikzfig{combinatorial_semantics/e-hypergraph-composition_1}
    }
    \]

\caption{Composition (nodes in the image of $Y$ are highlighted)}
\label{fig:e-hyp_composition_example_1}
\end{subfigure}
\hfill
\begin{subfigure}{0.35\linewidth}
\centering
\[
    \scalebox{0.35}{

    \tikzfig{combinatorial_semantics/e-hypergraph-composition_2}
    }
    \]

\caption{Coproduct of the carriers with the nodes in the image of $Y$ identified}
\label{fig:e-hyp_composition_example_2}
\end{subfigure}
\hfill
\begin{subfigure}{\linewidth}

\centering
\[
    \scalebox{0.35}{

    \tikzfig{combinatorial_semantics/e-hypergraph-composition_3}
    }
\]

\caption{Composition result}
\label{fig:e-hyp_composition_example_3}

\end{subfigure}

\caption{$\catname{ECsp(EHyp_{\Sigma})}$ composition example}
\label{fig:e-hyp_composition_example}
\end{figure}

Next, we will define $+$ operation on morphisms of $\Ecospans$. This will help us make this category into a semilattice-enriched SMC but for now, we will not require that $+$ should satisfy any equations.

\update{below definition depends on whether we require the homomorphism to reflect the conflicts}

\begin{definition}
    $+$ of two morphisms 
    \[
    ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
    \]
    \[
    ext(X) \xrightarrow{f'_{ext}} X' \xrightarrow{f'} B \xleftarrow{g'} Y' \xleftarrow{g'_{ext}} ext(Y)
    \]

    is defined in the Figure~\ref{fig:A+B}. And is a cospan

    \[
    ext(X) \xrightarrow{f_{ext}} ext(X) + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} ext(Y) + Y + Y' \xleftarrow{g_{ext}} ext(Y)
    \]

    $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|ext(X)|$ sources and $|ext(Y)|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v)$. And such that its sources and targets are top-level nodes. $h$ and $k$ are morphisms that map $ext(X)$ and $ext(Y)$ to these sources and targets respectively. Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
\end{definition}

\begin{figure}
    \[
    \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/f_plus_g_new}
    }
    \]
    \caption{$+$ of two morphisms in $\Ecospans$}
    \label{fig:A+B}
\end{figure}

\begin{definition}[Mda-cospan in $\catname{ECsp_{D}(EHyp_{\Sigma})}$]

Consider a cospan

\[
ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
\]

We will call such cospan \textit{monogamous directed acyclic} (mda-) if

\begin{enumerate}
        \item $A$ is a directed acyclic e-hypergraph.
        \item $f_{ext}$, $f$, $g$, $g_{ext}$ are monomorphic.
        \item The in-degree and out-degree of every node in $A$ are at most 1.
        \item Nodes with the in(out)-degree of 0 are exactly the nodes in the image of $f$ ($g$).
        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ ($g$) which pre-image is not in $ext(X)$ ($ext(Y)$). 
        \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
\end{enumerate}

\end{definition}

The cospan in Figure~\ref{fig:e-cospan} is also an mda-cospan.

\update{We can probably omit the proofs below because they are obvious. But keeping them here serves as a sanity check.}

\begin{proposition}
    Id morphism

    \[
    X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
    \]

    is an mda-cospan.
    \begin{proof}
        Proof by checking all the constraints above. Discrete e-hypergraphs is obviously directed and acyclic. Id-morphisms are monos. The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$. Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
    \end{proof}
\end{proposition}

\update{Anything about symmetry?}

\begin{proposition}
    Tensor unit

    \[
    0 \xrightarrow{id_{0}} 0 \xrightarrow{id_{0}} 0 \xleftarrow{id_{0}} 0 \xleftarrow{id_0} 0
    \]

    is an mda-cospan.
    \begin{proof}
        The empty e-hypergraph is obviously directed and acyclic. Id-morphisms are monos. The bullets (3) --- (6) are satisfied by the absence of nodes.
    \end{proof}
\end{proposition}

\begin{proposition}
    The tensor product of two mda-cospans is again an mda-cospan.

    \begin{proof}
        Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$. Hence, everything is preserved component-wise by copairing and coproduct.
    \end{proof}
\end{proposition}

\begin{proposition}
    Composition of two mda-cospans 
    \[
ext(X) \xrightarrow{f_{ext}} X \uplus (Y' \setminus ext(Y)) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus ext(Y)) \xleftarrow{g_{ext}} ext(Z)
\]

is again an mda-cospan.

    \begin{proof}
        Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles. This also does not increase their degree to more than 1. $f_{ext}$ is mono because its image $X$ is unchanged. The image of $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono. Pushout is then obtained by identifying the nodes with the pre-image in $ext(Y)$ which are removed in $Y' \setminus ext(Y)$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus ext(Y)$. $g';i_2$ yields $h$ and therefore $h$ is mono. Likewise, $k$ is mono.
    \end{proof}
    \update{Is this proof ok?}
\end{proposition}

\begin{proposition}
    $+$ of two mda-cospans

        \[
    ext(X) \xrightarrow{} X \xrightarrow{} A \xleftarrow{} Y \xleftarrow{} ext(Y)
    \]
    \[
    ext(X) \xrightarrow{} X' \xrightarrow{} B \xleftarrow{} Y' \xleftarrow{} ext(Y)
    \]

    is an mda-cospan.

    \begin{proof}
        By construction in Figure~\ref{fig:A+B}.
    \end{proof}
\end{proposition}

\begin{definition}[Category $\catname{MAECsp_{D}(EHyp_{\Sigma})}$]

    Taking into account all the propositions above gives us a subcategory of $\catname{ECsp_{D}(EHyp_{\Sigma})}$ where each morphism is an mda-cospan. This category is also symmetric monoidal.
\end{definition}

% show that EHyp is symmetric monoidal

% To show the equivalence between terms in an enriched category and e-hypergraphs we introduce the notion of e-hypergraphs with interfaces. An e-hypergraph $\mathcal{G}$ with input and output interfaces is a cospan in the category $\catname{EHyp_\Sigma}$

% where $n,m$ are ordered discrete e-hypergraphs. The objects of $\catname{EHyp_{\Sigma}}$ and such cospans constitute a monoidal category $Csp(\catname{EHyp_{\Sigma}})$. Objects of $Csp(\catname{EHyp_{\Sigma}})$ are the same as of $\catname{EHyp_{\Sigma}}$ and morphisms are isomorphism classes of cospans. Cospans are isomorphic if their legs are the same and the carriers are isomorphic.

% Consider a cospan 
% \[
% n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m
% \]

% \begin{definition}[Interfaces]
% \label{interfaces}
% The input interface $in(e)$ of an edge $e \in E_{\mathcal{G}}$ is the subgraph of vertices $v$ of $n$ such that $(f;p_{V,\mathcal{G}})(v) = e$. Similarly, the output interface $out(e)$ of an edge $e$ is the subgraph of $m$ satisfying $(g; p_{V,\mathcal{G}})(v) = e$. The outermost (external) input $in_{ext}(e)$ and output $out_{ext}(e)$ interfaces are the subgraphs of the interfaces whose image has parent $\bot$. 
% \end{definition}

% By slightly overloading the notation we denote with $in(\mathcal{G})$ ($out(\mathcal{G})$) the image of $f$ ($g$) and with $in_{ext}(\mathcal{G})$ ($out_{ext}(\mathcal{G})$) the set of nodes $v_i$ in the image with $p_V(v_i) = \bot$.

% An example could be seen below in Figure~\ref{fig:e-cospan}. We delimit external and internal interfaces with a vertical line. In this particular example $in(\mathcal{G}) = \{0, 1, 2, 3\}$, $in_{ext}(\mathcal{G}) = \{0,1\}$, and $out(\mathcal{G}) = \{6,4,5\}$, $out_{ext}(\mathcal{G}) = \{6\}$

% Say that D just picks discrete ehypergraphs, no arrow from PROP+

% \begin{definition}[Category of e-hypergraphs with interfaces]
%     We call a full subcategory $\catname{Csp_{D}(EHyp_{\Sigma})}$ of $\catname{Csp(EHyp_{\Sigma})}$ as a category of e-hypergraphs with interfaces. Functor $D : \text{PROP}^{+} \to EHyp_{\Sigma}$ chooses only discrete e-hypergraphs from $\catname{EHyp_{\Sigma}}$.
% \end{definition}

% say that H_e inherits ehypergraphneess

\begin{definition}[Input and output nodes of a hierarchical edge]

Consider an mda-cospan~\ref{mda-cospan} and an edge $e$ in $E_{A}$ such that $l(e) = \bot$. Then we take all nodes that are immediate successors of $e$. We call the nodes that have a pre-image in $X$ input nodes of $e$ and those that have a pre-image in $Y$ output nodes.
\end{definition}

% \begin{definition}[Set of inner subgraphs of an edge $e \in E_\mathcal{G}$]

% Consider an edge $e \in \mathcal{G}$ such that $l(e) = \bot$. We denote with $\mathcal{H}_{e}$ a set of inner subgraphs of the edge $e$, that is $\mathcal{H}_e \triangleq \{e' \in E_\mathcal{G} \text{ and } v' \in V_{\mathcal{G}} \; | \; \exists n,m > 0 \; . \; p^n(e') = e \text{ and } p^m(v') = e\}$. \update{And such that $\mathcal{H}_{e}$ inherits all the relations from $\mathcal{H}$ for the corresponding nodes and edges. What happens to the parent of e.g. $v$ for which $p_{V}(v) = e$?}
% \end{definition}

% add sketch of a proof for saying that a complement is an equivalence, define partition
% \begin{definition}[Instance subgraph of an edge $e \in E_\mathcal{G}$]

% Consider a partition of $\mathcal{H}_e$ according to a complement of the relation $[\#_V,\#_E]$ denoted as $[\bar{\#}_{V},\bar{\#}_{E}]$.
% Note that these relations are equivalence relations as the original relation is irreflexive. We call each $i^{th}$ subset of this partition as $i^{th}$ instance subgraph of $e$, that is, every $\mathcal{H}^i_e \subseteq \mathcal{H}_{e,\bar{\#}}$ is an instance subgraph of $e$.

% \end{definition}

\begin{definition}[Well-typed edge]

    Consider a hierarchical edge $e$ in $E_{A}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively. Then consider partitions $V_{e, in}^{\bar{\#}}$ and $V_{e,out}^{\bar{\#}}$  of the nodes according to $\bar{\#}$. We call $e$ well-typed if for all $v \in V_{e,in}^{\bar{\#}}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\bar{\#}}$ $|v| = |t(e)|$. We will call an mda-cospan well-typed if all hierarchical edges in its carrier are well-typed.
\end{definition}

\update{I have run out of possible names for a category}

\begin{definition}[Category $\WellTypedMdaEcospans$]

The objects of $\MdaEcospans$ and well-typed morphisms (cospans) form the category $\WellTypedMdaEcospans$
\end{definition}

\begin{proposition}
    Id-cospan

    \[
    X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
    \]

    is well-typed.
    \begin{proof}
        By definition, we only include well-typed morphisms.
    \end{proof}
\end{proposition}

\begin{proposition}
    Tensor unit

    \[
    0 \xrightarrow{id_0} 0 \xrightarrow{id_0} 0 \xleftarrow{id_0} 0 \xleftarrow{id_0} 0
    \]

    is well-typed.
    \begin{proof}
        The empty hypergraph does not contain any hierarchical edges and hence the cospan above is well-typed.
    \end{proof}
\end{proposition}

\begin{proposition}[The tensor product of two mda-cospans is again an mda-cospan]
    The tensor product does not introduce any new hierarchical edges and images of the interfaces are
preserved by copairing.
\end{proposition}

\begin{proposition}[Composition of two well-typed mda-cospans is a well-typed mda-cospan]
     Composition does not introduce any new hierarchical edges.
Furthermore, as composition identifies only nodes in the image of the outermost interfaces, input and
output nodes of hierarchical edges are not affected.
\end{proposition}

\begin{proposition}
    $+$ of two \textit{well-typed} mda-cospans

        \[
    ext(X) \xrightarrow{} X \xrightarrow{} A \xleftarrow{} Y \xleftarrow{} ext(Y)
    \]
    \[
    ext(X) \xrightarrow{} X' \xrightarrow{} B \xleftarrow{} Y' \xleftarrow{} ext(Y)
    \]

    is a \textit{well-typed} mda-cospan.

    \begin{proof}
        By construction in Figure~\ref{fig:A+B}.
    \end{proof}
\end{proposition}

The propositions above make $\WellTypedMdaEcospans$ into a proper symmetric monoidal category. There is an obvious inclusion

\[
\mathcal{I} : \catname{MACsp_{D}(Hyp_\Sigma)}
\xhookrightarrow{} \WellTypedMdaEcospans
\]\label{functor:I}

For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\catname{MACsp_{D}(Hyp_{\Sigma})}$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\# = \varnothing$, and $<_{c} = \varnothing$. It is well-typed on the nose as it has no hierarchical edges.

% \update{Two cospans $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$, $p \xrightarrow{} \mathcal{G} \xleftarrow{} q$ in $\catname{MACsp_{D}(Hyp_{\Sigma})}$ are composable when $out^{-1}_{out}(\mathcal{G})$ and $in^{-1}_{out}(\mathcal{H})$ are isomorphic.
% The result of such composition is depicted below
% }
% \[
% \adjustbox{scale=0.75,center}{
% \begin{tikzcd}
%             &                         &                                   & k \arrow[ld] \arrow[rd] \arrow[lld] \arrow[rrd] &                                  &                         &             \\
% n \arrow[r] & \mathcal{G} \arrow[rrd] & m \arrow[l]                       &                                                 & p \arrow[r]                      & \mathcal{H} \arrow[lld] & q \arrow[l] \\
%             &                         & n \cup (p \setminus k)  \arrow[r] & \mathcal{G};\mathcal{H}                         & q \cup (m \setminus k) \arrow[l] &                         &            
% \end{tikzcd}
% }
% \]

% where $k = out^{-1}_{out}(\mathcal{G})$ and $\mathcal{G;H}$ is a pushout of $\mathcal{G} \xleftarrow{} k \xrightarrow{} \mathcal{H}$. New input interface for the composition is obtained by adding the input interface of $\mathcal{H}$ to $n$ and removing $p$ (and likewise for the output interface). $p \setminus k$ can be defined as a following pushout complement (and likewise for $m \setminus k$):

% \[
% \adjustbox{scale=0.75,center}{
% \begin{tikzcd}
% k \arrow[d] & 0 \arrow[l] \arrow[d]   \\
% p           & p \setminus k \arrow[l]
% \end{tikzcd}
% }
% \]

% then $n \cup (p \setminus k)$ is just a coproduct, since $n$ and $(p \setminus k)$ are disjoint.

% We will also use the notion of a boundary complement to preserve monogomousness when rewriting.

% \begin{definition}[Boundary complement]
%     For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
% and mono $f : L \to G$, a pushout complement as depicted below

% \begin{figure}[h!]
%     \centering
%     \includegraphics[width=0.3\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
%     \label{fig:boundary_complement}
% \end{figure}

% is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to L^{\bot}$ and $d_2 : m \to L^{\bot}$ making the above triangle commute and such that
% \[
%     n + j \xrightarrow{[d_1,c_2]} L^{\bot} \xleftarrow{[d_2,c_1]} m + i
% \]
% is a monogamous cospan
% \end{definition}

\subsection{Syntactic rewriting of PROPs}

\Aleksei{Should this be a separate section rather than a subsection?}

\begin{definition}[\cite{Frobenius}]
A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e. pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities. Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as follows, for some rule $\langle l,r \rangle \in \mathcal{R}$    
\end{definition}

\begin{figure}[h!]

    \[
    \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
    \]
\end{figure}

\begin{definition}
Given a rewriting system $\mathcal{R}$ for a $\textsf{PROP}(\Sigma)$ we define 
 a rewriting system $\mathcal{R}^{+}$ in a $\textsf{PROP}^{+}(\Sigma)$  as a set of pairs $\langle l , r \rangle$ for each pair $\langle l, r \rangle$ in $\mathcal{R}$. However, the rewriting happens modulo semilattice enriched laws rather than SMC laws. We define what $a \Rightarrow_{\mathcal{R}^+} b$ means for some rule $\langle l , r\rangle $ based on the structure of $a$ and $b$.

If they are decomposable as above then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$. Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as below.

\begin{figure}

    \[
    \scalebox{0.6}{
    \tikzfig{combinatorial_semantics/prop_plus_rewriting}
    }
    \]
\end{figure}

\end{definition}

% \update{

% \begin{definition}
%     Functor $\llbracket \cdot \rrbracket : \text{PROP}^{+} \to \catname{MACsp_{D}(EHyp_{\Sigma})}$.

%     We will build this functor inductively. As noted above, $\catname{MACsp_{D}(Hyp_{\Sigma})}$ is a full subcategory of $\catname{MACsp_{D}(EHyp_{\Sigma})}$ and the acting of this functor on generators can be found in~\cite{Frobenius}. We need to define what $+$ from $\text{PROP}^{+}$ in $\catname{MACsp_{D}(EHyp_{\Sigma})}$ is. The definition could be found in Figure~\ref{fig:f+g} where well-typed mda-cospans for images of arbitrary morphisms $f, g : m \to n$ are depicted on the left and the image of $f + g$ is depicted on the right. A new edge is introduced which is the immediate predecessor of arbitrary e-hypergraphs $f$ and all edges and nodes of $f$ and $g$ are in conflict. The nodes are added to the interfaces and to the new edge to preserve well-typedness.

% \end{definition}

% }

We will later mimic rewrites in $\textsf{PROP}^{+}(\Sigma)$ at the level of $\WellTypedMdaEcospans$ by utilising a generalisation of double-pushout with interfaces (DPOI) approach defined in~\cite{Frobenius}. First, we will recall standard double-pushout (DPO) rewriting.

DPO is a technique for formalising rewriting in graph-like categories. For a given rewrite rule $\langle l, r \rangle$ it corresponds to an intuitive notion of removing an occurrence of $l$ within a graph and then plugging $r$ in there. More specifically, a DPO rule is a span $L \xleftarrow{} K \xrightarrow{} R$ in some category $\mathcal{C}$ with pushouts, where $L$ stands for the $l$ part of a rewrite rule, $R$ stands for the $r$, and $K$ is the invariant that should be maintained during rewriting.

A typical DPO square looks like this

\[
\begin{tikzcd}
L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H
\end{tikzcd}
\]

where the two marked squares are pushouts. Morphisms $K \to C \to G$ are called a \textit{pushout complement} and morphism $m$ is called a \textit{match} and is defined when the corresponding pushout complement exists. When $f,g$ are injective, constructing $C$ can be intuitively understood as finding an occurrence of $L$ in $G$ and removing from this occurrence everything that has no pre-image in $K$ and then inserting in there everything from $R$ that has no pre-image in $K$ (hence the two pushouts: one for deletion and one for insertion). Then we say that $G$ is rewritten into $H$ under a rewrite rule $L \xleftarrow{} K \xrightarrow{} R$.

However, this notion of DPO does not account for hypergraphs with interfaces, i.e. those from $\catname{MACsp_{D}(Hyp_{\Sigma})}$ where $l,r$ from $\langle l, r \rangle$ are cospans. To support this we first assume that we only need to maintain the interfaces of $i \xrightarrow{} L \xleftarrow{} j$ and $i \xrightarrow{} R \xleftarrow{} j$ during rewriting, then we can take $K \cong i + j$, i.e. take $K$ to be the coproduct of $i,j$ (remark 3.14~\cite{Frobenius}). What remains is to account for the interfaces in $G$, $C$, and $H$, which brings us to the following modified DPOI square. $J \cong n + m$ where $n$ and $m$ are the input and output interfaces of $G$, $C$, and $H$ respectively. The only difference from DPO square above is that we now explicitly require that the initial object $G$, intermediate object $C$ and final object $H$ all share the same interfaces.

\[
\begin{tikzcd}
L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H \\
                  & J \arrow[lu] \arrow[u] \arrow[ru]          &              
\end{tikzcd}
\]

To mimic rewrites in $\textsf{PROP}(\Sigma)$ we need to put some restrictions on pushout complements as not all of them lead to mda-cospans.

\begin{definition}[Boundary complement for $\catname{MACsp_{D}(Hyp_{\Sigma}))}$]
\label{def:boundary_original}
For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
and mono $f : L \to G$, a pushout complement as depicted below

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.3\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
\end{figure}

is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to L^{\bot}$ and $d_2 : m \to L^{\bot}$ making the above triangle commute and such that
\[
    n + j \xrightarrow{[d_1,c_2]} L^{\bot} \xleftarrow{[d_2,c_1]} m + i
\]
is a monogamous cospan. 
\end{definition}

Intuitively, such a pushout complement requires that inputs get only glued to outputs and vice-verse in the two pushout squares.

\begin{definition}{Convex match}

We call a match $m : L \to G$ convex if it is mono and its image is a convex subgraph of $G$.

\end{definition}

\begin{definition}{Rewriting in $\catname{MACsp_{D}({Hyp_{\Sigma}})}$}

Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ in $\catname{Hyp_{\Sigma}}$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold

\begin{itemize}
    \item $m : L \to G$ is a convex match;
    \item $i + j \to C \to G$ is a boundary complement in the leftmost pushout.
\end{itemize}

\end{definition}

\begin{theorem}[Theorem 35~\cite{Frobenius2}]
    Let $R$ be any rewriting system on $\textsf{PROP}(\Sigma)$. Then,
    \[
    g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llbracket ^{\ulcorner}g^{\urcorner} \rrbracket \Rrightarrow_{\llbracket \mathcal{R} \rrbracket} \llbracket ^{\ulcorner}h^{\urcorner} \rrbracket 
    \]

    where $\llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \catname{MACsp_{D}(Hyp_{\Sigma})}$ is a faithful monoidal functor.
\end{theorem}

\begin{remark}
    Note that even though $n \xrightarrow{} G \xleftarrow{} m$ and $0 \xrightarrow{} G \xleftarrow{} n + m$ are essentially isomorphic \update{(they are not technically isomorphic because they have different legs)}, the functor $\llbracket \cdot \rrbracket$ deals with arbitrary string diagrams (that have multiple inputs and multiple outputs) in $\textsf{PROP}(\Sigma)$ and interprets them as cospans $n \xrightarrow{} G \xleftarrow{} m$. However, there is a syntactic operation $\ulcorner \cdot \urcorner$ on $\textsf{PROP}(\Sigma)$ to translate between string diagrams with multiple inputs and outputs and string diagrams with only outputs.

    \[
\tikzfig{combinatorial_semantics/bending_wires}
    \]

    Now the left-hand and right-hand sides are equivalent in the sense that $f \Rightarrow_{\langle l, r \rangle} g$ if and only if $\ulcorner f \urcorner \Rrightarrow_{\langle \ulcorner l \urcorner, \ulcorner r \urcorner \rangle} 
 \ulcorner g \urcorner$ and the latter string diagram can be interpreted as $0 \xrightarrow{} G \xleftarrow{} n + m$ for which we defined rewriting above.
\end{remark}

\Aleksei{TODO: discuss this}

\Aleksei{This holds for boxes with bent wires, which require a compact closed structure. The proof for the "only if" direction refers to the result for rewriting in PROP + Frob and for the converse direction it uses decomposition lemma and isomorphism-ness of a functor $\llbracket \cdot \rrbracket$ (Theorem 25~\cite{Frobenius2})}

\Aleksei{At some point I forgot to make a grey shaded area for cospans. Need to fix this.}

We have a more general notion of interface and hence we need to adjust DPOI rewriting a little bit. The first issue is that interfaces are not preserved during rewriting, consider an example above which corresponds to a rewrite rule of $\langle f, f+g \rangle$.

\begin{figure}[t]
    \centering
    \[
    \scalebox{0.75}{\tikzfig{combinatorial_semantics/f_to_f_plus_g_example}}
    \]
\end{figure}

The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$. Note that the outermost interfaces are preserved.

Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces. Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.

\begin{figure}
    \centering
    \[
    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
    \label{fig:interface_change_example}
\end{figure}

The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule. Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule. The entry in the middle of the span depicts the outermost interfaces which are preserved. The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right. The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it. Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.

\Aleksei{TODO: show that there exists a pushout for each of the axioms and lifted rewrites}

\question{Shall I replace $i,j$ etc. with $ext(X)$?}

To account for such changes in the interfaces we put some additional constraints on DPOI rewriting. Consider a DPOI diagram with extra constraints below.

% \[\begin{tikzcd}
% 	&&&& {i + j} \\
% 	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
% 	\\
% 	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
% 	\\
% 	&&&& {n+m}
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
% 	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
% 	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
% 	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
% 	\arrow[from=2-1, to=2-3]
% 	\arrow["{[f_i,f_o]}", from=2-1, to=4-1]
% 	\arrow[from=4-1, to=4-3]
% 	\arrow[from=2-9, to=2-7]
% 	\arrow[from=4-9, to=4-7]
% 	\arrow["{[h_i,h_o]}"{description}, from=2-9, to=4-9]
% 	\arrow[from=6-5, to=4-9]
% 	\arrow[from=6-5, to=4-1]
% 	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
% 	\arrow["{[i_r,j_r]}"{description}, from=1-5, to=2-9]
% \end{tikzcd}\]

% https://q.uiver.app/#q=WzAsMTEsWzIsMSwiTCJdLFs0LDAsImkgKyBqIl0sWzIsMywiR157XFx1cmNvcm5lcn0iXSxbNCwzLCJDIl0sWzQsNSwibittIl0sWzAsMSwibF9pK2xfbyJdLFswLDMsImdfaStnX28iXSxbNiwxLCJSIl0sWzYsMywiXntcXHVsY29ybmVyfUgiXSxbOCwzLCJoX2kraF9vIl0sWzgsMSwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw2XSxbMSw1LCJbaV9sLGpfbF0iLDFdLFs0LDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzMsOCwiIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNyw4LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxLDcsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzEwLDddLFs5LDhdLFsxLDEwXSxbNCw5XV0=
\[\begin{tikzcd}
	&&&& {i + j} \\
	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
	\\
	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
	\\
	&&&& {n+m}
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
	\arrow[from=2-1, to=2-3]
	\arrow[from=4-1, to=4-3]
	\arrow[from=6-5, to=4-1]
	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
	\arrow[from=2-9, to=2-7]
	\arrow[from=4-9, to=4-7]
	\arrow[from=1-5, to=2-9]
	\arrow[from=6-5, to=4-9]
\end{tikzcd}\]

The original DPOI square is shown in red. The extra bits simply account for more complicated interfaces.
For example, the left-hand side $L$ of a rewrite rule $\langle L, R \rangle$ which originally was some cospan $i \xrightarrow{} L \xleftarrow{} j$ in $\catname{MACsp_{D}(Hyp_{\Sigma})}$ now is a cospan $i \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} j$ from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
The top half of this diagram says that the outermost interfaces should coincide for $L$ and $R$. The bottom half puts the same restrictions on $G$ and $H$. 
% The extra squares on both sides say that there should exist a matching of the interfaces together with a matching of e-hypergraphs. This requirement was implicit in the case of $\catname{MACsp_{D}(Hyp_{\Sigma}}$ because the whole interfaces and the outermost interfaces coincided. Because now the interfaces can change, we need to keep track of which parts of the interfaces are subjected to changes.

We also adjust the definition of a boundary complement a bit.

\begin{definition}[Boundary complement in $\WellTypedMdaEcospans$]
\label{def:boundary_new}

% https://q.uiver.app/#q=WzAsNyxbMiwxLCJMIl0sWzQsMCwiaSArIGoiXSxbMiwzLCJHXntcXHVyY29ybmVyfSJdLFs0LDMsIkMiXSxbNCw1LCJuK20iXSxbMCwxLCJsX2krbF9vIl0sWzAsMywiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw2XSxbMSw1LCJbaV9sLGpfbF0iLDFdXQ==
\[\begin{tikzcd}
	&&&& {i + j} \\
	{l_i+l_o} && L \\
	\\
	{g_i+g_o} && {G^{\urcorner}} && C \\
	\\
	&&&& {n+m}
	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
	\arrow[from=2-1, to=2-3]
	\arrow[from=4-1, to=4-3]
	\arrow[from=6-5, to=4-1]
	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
\end{tikzcd}\]
Consider the diagram above where the marked square is a pushout and $i + j$ and $n + m$ are the \textit{outermost} interfaces for cospans $i \xrightarrow{} l_i \xrightarrow{} L \xleftarrow{} l_o \xleftarrow{} j$ and $n \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m$. We say that $i+j \to C \to G^{\urcorner}$ is a \textit{boundary} complement if $[i_c, i_j]$ is mono and there exist $n_c : n \to C$ and $m_c : m \to C$ making the above triangle commute and such that there exists a \textit{not necessarily well-typed} mda-cospan

\[
n \xrightarrow{} c_i + j \xrightarrow{} C \xleftarrow{} c_o + i \xleftarrow{} m
\]

where $c_i$ and $c_o$ are computed as $c_i = g_i \setminus (l_i \setminus i)$ and $c_o = g_o \setminus (l_o \setminus j)$.

% \Aleksei{The pushout interpretation is wrong because there might not be a morphism between $i$ and $c_i$ and similarly for $j$ and $c_o$. Any other categorical interpretation? The outermost interface of the rewrite span may not be a part of the interface of $G$}

% where $l_i + l_o$ is the whole interface for $L$, $g_i + g_o$ is the whole interface for $G$.
\end{definition}

% Intuitively, these pushout squares mean that when we delete the occurrence of $L$ from $G$ we also remove the occurrence of $L$'s interface within $G$'s interface preserving the outermost interfaces $i + j$. The highlighted arrows form the pushout square for the original boundary complement.

% We call morphisms $i + j \to C \to G$ a boundary complement if

% \begin{enumerate}
%     \item The square above commutes and $L \xrightarrow{} G \xleftarrow{} C$ is a pushout
%     \item There exists an mda cospan $g_i \setminus (l_i \setminus i) + j \xrightarrow{} C \xleftarrow{} g_o \setminus (l_o \setminus j) + i$. \question{I am unsure about categorical notation for this set difference. Maybe it can be paraphrased in the following way?}
%     \item There exists an mda cospan $c_i + j \xrightarrow{} C \xleftarrow{} c_o + i$ where $c_i$ and $c_o$ are the following pushout complements in the category of discrete hypergraphs:
% \end{enumerate}

% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$. \question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}

\begin{remark}
Note that cospan $n \xrightarrow{} c_i + j \xrightarrow{} C \xleftarrow{} c_o + i \xleftarrow{} m$ is not well-typed. The cospan's input interface contains nodes that previously were a part of the output interface.
\end{remark}

First, let's check that this notion generalises the ordinary boundary complement from~\ref{def:boundary_original}. For this, we assume that child and conflict relations are empty in which case $i + j$ and $l_i + l_o$ coincide as well as $g_i + g_o$ and $n + m$. Hence, $g_i \setminus (l_i \setminus i) + j = g_i + j = n + j$ and the same for the output interfaces. Thus we have $n + j \xrightarrow{} C \xleftarrow{} m + i$, which corresponds to the definition from~\ref{def:boundary_original}.

% Next, consider an example from figure~\ref{fig:dpoir}. The outermost interfaces of $L$ coincide with the whole interfaces, i.e. $l_i + l_o = i + j$. However, $n + m = \{0'', 1''\}$ and $g_i + g_o = \{0'',0,2,1,3,1''\}$. According to the definition of a boundary complement the following mda cospan should exist: $g_i \setminus (l_i \setminus i) + j = \{0'', 0, 2\} \setminus \varnothing \cup \{1\} = \{0'',0,2,1\} \xrightarrow{} C \xleftarrow{} \{0,1,3,1''\} = \text{same procedure for output interfaces}$.

% \question{We need to show that this boundary complement is unique and that after performing a pushout on the right the result is also an mda hypergraph}

% We can now present a complete DPOI square

% \[
% \begin{tikzcd}
%                    &                    & i+j \arrow[ld] \arrow[lld] \arrow[ddd, "f"] \arrow[rrd] \arrow[rd] &                    &              &  &                       &               &                       &               \\
% L \arrow[dd, "m"'] & l_i+l_o \arrow[l]  &                                                                    & r_i+r_j \arrow[r]  & R \arrow[dd] &  & i \arrow[r] \arrow[d] & r_i \arrow[d] & j \arrow[d] \arrow[r] & r_o \arrow[d] \\
%                    &                    &                                                                    &                    &              &  & c_i \arrow[r]         & h_i           & c_o \arrow[r]         & h_o           \\
% G                  &                    & C \arrow[ll] \arrow[rr]                                            &                    & H            &  &                       &               &                       &               \\
%                    & g_i+g_o \arrow[lu] & n+m \arrow[u] \arrow[l] \arrow[llu] \arrow[r] \arrow[rru]          & h_i+h_o \arrow[ru] &              &  &                       &               &                       &              
% \end{tikzcd}
% \]

% \begin{remark}
%     Note that in all pushout squares for the interfaces all the morphisms are monos. \question{Does it follow from anything above? Because the corresponding triangles commute?}
% \end{remark}

\begin{proposition}
    The boundary complement in~\ref{def:boundary_new} when exists is unique
    \begin{proof}
        \update{TODO}
    \end{proof}
\end{proposition}

\question{
\begin{definition}[Convex down-closed conflict reflecting match] 
    We call $m : L \to G$ in $\catname{EHyp_{\Sigma}}$ a convex down-closed conflict reflecting match if

    \begin{enumerate}
        \item $m$ is monomorphism
        \item $m(L)$ is a convex down-closed subgraph of $G$
        \item if $m(x_1) \# m(x_2)$ then $x_1 \# x_2$
    \end{enumerate}
\end{definition}
}

\begin{definition}[Rewriting in $\WellTypedMdaEcospans$]
    Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n + m$ and $H \xleftarrow{} h_i + h_o \xleftarrow{} n+m$ in $\WellTypedMdaEcospans$ with the outermost interfaces $n + m$, $G$ rewrites convexly into $H$ with the outermost interface $n + m$ --- notation $( G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}} (E \xleftarrow{} n + m )$ --- if there exists a rule $l_i + l_o \xrightarrow{} L \xleftarrow{} i + j \xrightarrow{} R \xleftarrow{} r_i + r_o$ with the outermost interfaces $i+j$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts

% https://q.uiver.app/#q=WzAsMTEsWzIsMSwiTCJdLFs0LDAsImkgKyBqIl0sWzIsMywiR157XFx1cmNvcm5lcn0iXSxbNCwzLCJDIl0sWzQsNSwibittIl0sWzAsMSwibF9pK2xfbyJdLFswLDMsImdfaStnX28iXSxbNiwxLCJSIl0sWzYsMywiXntcXHVsY29ybmVyfUgiXSxbOCwzLCJoX2kraF9vIl0sWzgsMSwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw2XSxbMSw1LCJbaV9sLGpfbF0iLDFdLFs0LDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzMsOCwiIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNyw4LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxLDcsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzEwLDddLFs5LDhdLFsxLDEwXSxbNCw5XV0=
\[\begin{tikzcd}
	&&&& {i + j} \\
	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
	\\
	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
	\\
	&&&& {n+m}
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
	\arrow[from=2-1, to=2-3]
	\arrow[from=4-1, to=4-3]
	\arrow[from=6-5, to=4-1]
	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
	\arrow[from=2-9, to=2-7]
	\arrow[from=4-9, to=4-7]
	\arrow[from=1-5, to=2-9]
	\arrow[from=6-5, to=4-9]
\end{tikzcd}\]
    and the following conditions hold

    \begin{enumerate}
        \item $m : L \to G$ is a convex down-closed conflict reflecting match;
        \item $i + j \to C \to G$ is a boundary complement in the leftmost pushout.
    \end{enumerate}
\end{definition}

% \Aleksei{We may need the following proofs:}
% \begin{itemize}
%     \item The existence of pushout complements (for rewriting)
%     \item Soundness and completeness, that is the correspondence between enriched terms and e-hypergraphs.
%     \item Rewriting
%     \item Anything else?
% \end{itemize}

\Aleksei{Do we need to introduce substitution? This is needed if our rewrite rules are going to have variables as labels.}

The whole interfaces of the resulting e-hypergraph $H$ from the cospan $n \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m$ are constructed as $h_i = (r_i \setminus i) + c_i$ and $h_o = (r_o \setminus j) + c_o$. $h_i \to H$ is defined as a copairing of $f_1 : r_i \setminus i \to R \to H$ and $f_2 : c_i \to C \to H$. Since both $f_1$ and $f_2$ are mono, their copairing is also a mono. Similarly for $h_o$.

% using the two pushout squares below.
% Intuitively, these squares combine the interfaces of $R$ and $C$ and identify those bits that share the same pre-image in the outermost interface $i + j$.

% \[\begin{tikzcd}
% i \arrow[r] \arrow[d] & r_i \arrow[d] & j \arrow[r] \arrow[d] & r_o \arrow[d] \\
% c_i \arrow[r]         & h_i           & c_o \arrow[r]         & h_o          
% \end{tikzcd}\]

% Going back to the example from the figure~\ref{fig:dpoir}, $h_i = \{0'',0,0',2',2\}$, $r_i = \{0,0',2'\}$, $r_o = \{1',3',1\}$, $h_o = \{1',3',3,1,1''\}$, $c_i = \{0'',2,0\}$, $c_o = \{1'',1,3\}$, $i = \{0\}$, $j = \{1\}$, where $h_i$ is exactly $c_i \coprod_{\equiv_{i}} r_i = \{0'',2,0\} \coprod_{\equiv_{i}} \{0,0',2'\} = \{0'',0,0',2',2\}$ and the same for $h_o$.

\begin{proposition}

$n \xrightarrow{n_h} h_i \xrightarrow{f} H \xleftarrow{g} h_o \xleftarrow{m_h} m$ is a well-typed mda-cospan.

\begin{proof}
$f$ and $g$ are monos by construction. $n + m \to h_i + h_o$ is a mono by commutativity because $n + m \to H$ is a mono (also by commutativity). 
The only in-degree 0 nodes have a pre-image in $c_i$ and $r_i \setminus i$  and end up being in the image of $h_i$ (the nodes with the pre-image in $j$ get are glued together). Well-typedness is restored once the right-hand side of the rewrite rule is glued into $C$.  \question{Uniqueness of this pushout complement? Is it a presheaf? As a presheaf category is adhesive, $\catname{Csp_{D}(EHyp)}$ can not be a presheaf as it is not adhesive.}
\end{proof}
\end{proposition}

\Aleksei{Check the proof
}

\begin{lemma}(Decomposition). Let $m \xrightarrow{} G \xleftarrow{} n$ be a monogamous acyclic cospan and $L$ a convex and down-closed sub-hypergraph of $G$. Then there exists $k \in \mathcal{N}$ and a unique cospan $i \xrightarrow{} L \xleftarrow{} j$ such that G factors as $\ldots$

\update{TODO: this lemma is needed to prove soundness of rewriting}
\end{lemma}

\Aleksei{Can we reuse anything from~\cite{Frobenius}?}

Consider an example below that illustrates the introduced constraints.

\begin{remark}
    For now, we will informally denote a combinatorial interpretation of a morphism $f$ in $\textsf{PROP}^{+}(\Sigma)$ as $\llbracket f \rrbracket$.
\end{remark}

\begin{example}

A DPOI square for rewriting in $\WellTypedMdaEcospans$ is depicted in Figure~\ref{fig:refined_dpoi}.
A projection rule of $\llbracket f + g \rrbracket \to \llbracket f \rrbracket$ is shown in the top half of the diagram.
The image of the matching $m$ is shown in red and its image is a convex down-closed e-hypergraph.
Pushout complement is then computed by removing the image of $m$ from the target e-hypergraph by keeping intact the nodes that have a pre-image in the carrier of the rewriting span. The corresponding mda-cospan is depicted at the bottom: $c_i = g_i \setminus (l_i \setminus i) = \{18, 19, 1, 2, 13, 14, 5, 6, 7, 8\} \setminus (\{1,2,5,6,7,8\} \setminus \{1,2\}) = \{18, 19, 1, 2, 13, 14, 5, 6, 7, 8\} \setminus \{5,6,7,8\} = \{18,19,1,2,13,14\}$ and $c_i + j = \{18,19, 1, 2, 13, 14, 3, 4\}$ which is the input interface for $C$. Similar computations are performed for $c_o$. Then pushout glues $R$ and $C$ together along $i + j$ and the resulting interface $h_i = c_i + (r_i \setminus i) = \{18,19, 1, 2, 13, 14\} + \varnothing = \{18,19, 1, 2, 13, 14\}$ and likewise for $h_o$.

    \begin{figure}
        \centering
\[
  \scalebox{0.3}{      \tikzfig{combinatorial_semantics/refined_dpoi_example}
}
        \]
    \caption{Refined DPOI rewriting}
    \label{fig:refined_dpoi}
    \end{figure}

\end{example}

Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we add them as structural rewrite rules into $\llbracket \mathcal{R}^+ \rrbracket$. Recall that the category $\catname{EHyp_{\Sigma}}$ does not have all pushouts and hence we need to explicitly show that the pushouts that we need exist.

\begin{proposition}
Pushouts for all semilattice enrichment equations exist.

\begin{proof}
    Let us give a combinatorial interpretation for an instance of the first equation (distributivity) in~\ref{fig:string-equations}. It can be found in Figure~\ref{fig:distributivity_interpretation}. After we identify a convex down-closed image of the left-hand side of this rule within a graph $G$ we remove the image by keeping only the outermost nodes. Then we simply glue the right-hand side of the rule into the hole. Since we glue along the outermost nodes this does not introduce any conflict with regard to child relation. Such equations are not absorbed by e-hypergraph representation unlike SMC equations because the right-hand side contains more edges.

\end{proof}
\end{proposition}

\update{I think I haven't paid any attention to this fact previously, but it seems that in~\cite{Frobenius} the functor $\llbracket - \rrbracket$ is between $PROP_{\Sigma_{Frob}}$ and $Csp_{D}(Hyp_{\Sigma})$, i.e. PROP is quotiented by Frobenius equations, but they are absorbed in hypergraphs domain. We will have a similar thing: $\llbracket - \rrbracket: PROP^{+}_{\Sigma^{+}} \to \WellTypedMdaEcospans_{R^+}$, i.e. we will quotient PROP$^{+}$ by semilattice equations and quotient e-hypergraphs by the interpretation of these equations as they are not absorbed. The quotienting on the left is always implicit?}

\begin{figure}
    \centering
    \[
    \scalebox{0.45}{
    \tikzfig{combinatorial_semantics/structural_rule_pushout}}\]  
    \caption{Distributivity rewrite rule interpretation}
    \label{fig:distributivity_interpretation}
\end{figure}

\begin{definition}
    We will denote $\llbracket \mathcal{E} \rrbracket$ the interpretation of the equations in~\ref{fig:string-equations} as rewrite rules in $\WellTypedMdaEcospans$. In particular, for all $l = r$ in $\mathcal{E}$ $\llbracket \mathcal{E} \rrbracket$ will contain $\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle$ and $\langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle$. \update{Technically we add those for all input and output interfaces.}
\end{definition}

\begin{remark}
    There is a minor limitation for DPOI rewriting above. It is unable to express rewriting for rules that contain an edge with no inputs and outputs on their right-hand side. This is because there is no way to impose both conflict and child relation of such an edge as it has no incident nodes in its context. An example of this could be seen in Figure~\ref{fig:non_expressible}. After an image of a match is removed there is no way of plugging $k$ into the hierarchical edge as it has no connection with this edge via the shared interface. However, we think that this is only a minor limitation because in practice such edges with no incident nodes correspond to `scalars' if we consider for example a category of vector spaces where a rewrite rule with a scalar could be

\begin{alignat*}{3}
\begin{pmatrix}
 1 &  2 \\
-1 &  1 \\
\end{pmatrix}
&\!\begin{aligned}
\hspace{1em} & \otimes &
\end{aligned}
&\!\begin{aligned}
2 &
\end{aligned}
&\!\begin{aligned}
 & \xrightarrow{} & 
\end{aligned}
&\!\begin{aligned}
\begin{pmatrix}
 2 &  4 \\
-2 &  2 \\
\end{pmatrix}    
\end{aligned}
\end{alignat*}
\end{remark}

That is, typically such scalars are absorbed and not present on the right-hand side of a rewrite rule.

\begin{figure}
    \centering
\[
\scalebox{0.5}{
\tikzfig{combinatorial_semantics/non_expressible_rewrite}
}
\]
    \caption{Non expressible rewrite}
    \label{fig:non_expressible}
\end{figure}

\newpage

### 2023-09-29 sections/conclusion.tex
Initial Overleaf Import

\section{Conclusion and Further Work}

\begin{itemize}
\item implementation for ZX
\item computational linguistics
\item traced categories, hypergraphs for traced comonoids and full e-graphs
\item closed categories, hierarchical hypergraphs and lambda-calculus (example)
\end{itemize}

### 2023-09-29 sections/e-graph-translation.tex
Initial Overleaf Import

\section{E-Rewriting for E-Hypergraphs}

The (convex) \textit{matching} of one e-hypergraph $L$ (\textit{e.g.}, the left-hand side of a rewrite rule) into another e-hypergraph $G$ was defined in the previous section as a monic morphism $L \to G$ (whose image is additionally convex, \update{consistent}). In other words, it corresponds to finding a subgraph isomorphic to $L$ within $G$ (subject to conditions respecing the hierarchical and conflict structures).

However, in general we will have to explode the e-hypergraph, by applying the structural rules for (horizontal, vertical) congruence first, in order to find such a match. In this section, we define the  e-matching (and subsequent e-rewriting) as the problem of finding matches (and applying rewrites) within an e-hypergraph \textit{without} first inducing duplication of subgraphs. We begin first with definitions, before proceeding to prove them “correct” with respect to the plain notion of matching (rewriting). \update{Example}

\begin{definition}[Projection]
The set of projections of an e-hypergraph $G$ are \update{project each e-box}.    
\end{definition}

\begin{definition}[E-matching]
We define the \textit{e-matching problem} as follows: given two e-hypergraphs \update{(with interfaces??)} $L$ and $G$, we say that a convex subgraph $G' \subseteq G$ is a \textit{e-match} for $L$ if there exists a projection of $G'$ which isomorphic to $L$.
\end{definition}

\update{What about refolding graph?}
\begin{lemma}[Correctness of e-matching]
$G' \subseteq G$ is an e-match for $L$ if and only if there is a sequence of structural rewrites $G \to^* H$ such that $L$ is a match in $H$.
\end{lemma}
\begin{proof}
The implication

The converse direction is trivial. 
\end{proof}

\begin{definition}[E-matching]
Given an e-match $G' \subset G$ of $L$, 
define its e-rewriting as the application of the DPO rewriting rule $G' \to G' + L$  to the match $L \to G$ induced by the ematching. \update{diagram? interfaces?}
\end{definition}

\begin{definition}[E-rewriting]

Given a rewrite rule $L \to R$ and an e-match $G' \subset G$ of $L$ in $G$, define the \textit{e-rewriting} of $L \to R$ in $G$ to $H$, written $G \Rightarrow_{L \to R}^e H$, where $H$ is defined as the result of applying DPO-rewriting of the rewrite $G \to G+R$ to the match $L$ in $G$. \update{interfaces.}
\end{definition}

In order to state the correctness of e-rewriting, we first require the notion of \textit{maximal unfolding} of an e-hypergraph to a normal form. 
\begin{definition}
Given an e-hypergraph $G$, there exists a sequence of structural rewrites transforming it to the form $G_1 + \ldots + G_n$, where each $G_i$ is a plain hypergraph. This is called its \textit{sharing normal form}.
\end{definition}
\update{Above should be lemma then defn.}

\begin{lemma}[Correctness of e-rewriting]
Statement?
\end{lemma}

In the case of plain string diagrams represented as (mda-)hypergraphs, the problem of finding a matching of the left-hand side of a rewrite rule % is essentially the (induced??) sub-hypergraph isomorphism problem [up to interfaces, and convexity?]

Informally, for every e-box $e = \phi_1 + \ldots + \phi_n$ appearing in a string diagram $\psi$ as $\psi\{\phi_1 + \ldots + \phi_n\}$ we can define the $i$th projection of $e$ in $\psi$ as $\pi^e_i = \phi\{\psi_i\}$.

### 2023-09-29 sections/introduction.tex
Initial Overleaf Import

\newcommand\mylet[2]{\textsf{let } #1 = #2 \textsf{ in }}

\cite{heijltjes_functional_2023}
\cite{barrett_functional_2023}
\cite{griggio_proceedings_2022}
\cite{flatt_small_2022}
\cite{ghica_operational_2021}
 % digital circuits
\cite{ghica_hierarchical_2023}
\cite{ghica_rewriting_2023} % traced comonoid
\cite{wilson_string_2023} % non-strict categories

\cite{zhang_relational_2022} % relational ematching 
\cite{alvarez-picallo_rewriting_2022} % hierarchical
\cite{alvarez-picallo_functorial_2021} % hierarchical

\cite{bonchi_tape_nodate} % tape diagrams
\cite{baldan_categorical_2014} % signal flow
\cite{dpo}
\cite{maclane}
\cite{singher2023colored}
\cite{koehler2022sketchguided} % latest egraph

\section{Introduction}

Rewrite-driven program optimization involves applying a sequence of semantics-preserving rewrites which may improve speed of execution, resource usage, \textit{etc}. Because early application of some rewrites can enable or block the subsequent application of others, the choice of application order hugely impacts the quality of the resulting optimization.  This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites, for example,  those implementing constant propogation, loop unfolding, or unreachable code elimination. Typical approaches to this problem use heuristics to make an ad-hoc choice of ordering.

Another recently proposed solution to this problem is \textit{equality saturation} 
\cite{10.1145/1594834.1480915,EggPaper}: instead of maintaining a \textit{single},  putatively optimized program which is rewritten \textit{destructively} at each step, we can maintain a \textit{set} of equivalent programs, where each rewrite step \textit{non-destructively} grows the set.  Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can be extracted% from the set,  which now represents all equivalent ways to express the program with respect to the given rewrites.

The use of \textit{equality graphs (e-graphs)} to compactly represent the set of equivalent programs is what makes the technique tractable in practise, where the naive approach is clearly unfeasible.

While equality saturation is a state-of-the-art optimization technique, the formal mathematical theory of e-graphs is perhaps underdeveloped. In this paper, we contribute a categorical understanding of e-graphs. Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{enriched} Cartesian categories, where the enriched structure allows us to consider formal sets of terms (morphisms).

However, we do not develop the categorical perspective purely for its own sake: rephrasing the formalism of e-graphs in terms of algebraic theories and Cartesian categories suggests a natural generalization to monoidal theories. Therefore, our theoretical development leads to new potential domains of application, including optimziation of string diagrams representing digital \cite{ghica_compositional_2023} and quantum circuits
\cite{coecke_interacting_2011,ZX} and potentially to natural language \cite{wazni_quantum_2022,coecke_lambek_2013}.

Towards implementation, we 
additionally provide a combinatorial representation of our enriched string diagrams in terms of hypergraphs. We give a specification of hypergraph rewriting in this novel setting via a suitable extension of the double pushout (DPO) rewriting framework 
\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.

\subsection*{E-graphs}

\begin{figure}\label{fig:egraph-strings}
\[
    \tikzfig{categorical-semantics/egraph-strings}
\]
\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
\end{figure}

\begin{figure}\label{fig:e-graph-example}
\[
    \scalebox{0.5}{
    \tikzfig{categorical-semantics/egraph-translation}
    }
\]
\caption{Example translation of e-graphs into string diagrams for semilattice enriched (traced) Cartesian categories.}
\end{figure}

\begin{figure}\label{fig:let-calculus}
\begin{align*}
    (a*2)/2 \\
    \mylet{y}{a*2} y/2 \\ 
    \mylet{y}{\{a*2,a<\!\!<1\}} y/2 \\
    \mylet{y}{\{a*2,a<\!\!<1\}} \{y/2, a*(2/2)\} \\
    \mylet{x}{\{2/2, 1\}} \mylet{y}{\{a*2, a <\!\!< 1\}} \{y/2, a*x\} \\
    \textsf{letrec } z = (\mylet{x}{\{2/2, 1\}} \mylet{y}{\{z*2, z <\!\!< 1\}} \{a, y/2, z*x\}) \textsf{ in } z
\end{align*}
\caption{Example term calculus corresponding to the string diagrams of Figure \ref{fig:e-graph-example}}
\end{figure}

E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory. They generalize the typical directed acylcic graph (DAG) representation of terms to include equivalence classes. 
A sequence of examples, taken from \cite{EggPaper}, are given in the first row of Figure \ref{fig:e-graph-example}. Here, nodes (solid boxes) have children given by equivalence classes (dashed boxes) of nodes instead of just a single node. Note, the equivalence classes relate nodes whose corresponding \textit{sub-trees} denote equivalent terms.

Beginning with the e-graph (a), corresponding to the directed acyclic graph (DAG) representation of the term $(a*2)/2$, each subsequent e-graph (b)-(e) uses the given rewrite rule to add information: creating new nodes and edges, or merging equivalence classes. The final e-graph, (e), contains a cycle, and represents infinitely many terms: $a, a*1, (a*1)*1, $...

\subsection*{Semantics via Semilattice Enriched Categories}

One can easily see the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category by considering string diagrams 
\cite{noauthor_09083347_nodate,joyal_geometry_1991, mellies_functorial_2006}. In Figure \ref{fig:egraph-strings} we have extended the usual string diagrammatic notation for symmetric monoidal categories with an additional generator which has the  typing rule below, left.
The ability to take formal sums of morphisms is used to model the equivalence class structure of e-graphs.  
\[
\infer{\phi_1 + \ldots + \phi_n: A \to B}{\{\phi_i: A \to B\}_{i \in \{1,\ldots,n\}}}
\qquad
\infer{\Gamma \vdash \{t_1, \ldots, t_n\}: A}{\{\Gamma \vdash t_i: A\}_{i \in \{1, \ldots, n\}}}
\]
To aid understanding, in Figure \ref{fig:let-calculus} we give an example of a representation of the string diagrams in Figure \ref{fig:e-graph-example} using an informal let-calculus, which is standard except for its augmentation with the typing rule above, right, corresponding to semilattice enrichment. The new construct forms a set of terms and satisfies the following additional congruence: $K\{\{t_1, \ldots, t_n\}\} = \{K\{t_1\}, \ldots, K\{t_n\}\}$.

String diagrams are read from bottom-to-top and we consider additional generators for the duplication and deletion transformations of a Cartesian category.

The translation from e-graphs to our enriched string diagrams is illustrated informally below. 
Note how the typing constraints of the $+$ constructor are satisfied.
\[
    \tikzfig{categorical-semantics/egraph-translation-1}
\]

Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}. In particular, note that the translation of the \textit{cyclic} e-graph (e)
requires the use of a categorical \textit{trace}, generating a cycle. In this paper, we focus on \textit{acyclic} e-graphs and their corresponding catgeories; but the above example illustrates that the extension to the cyclic case appears routine.

Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal sums. Thus we generalize e-graphs from algebraic to monoidal theories.

\subsection*{Combinatorial Representation of Enriched String Diagrams}

In order to implement our generalized version of e-graphs, we require to perform rewriting not on algebraic terms, but on string diagrams. String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo "connectivity") or as a 2-dimensional syntax quotiented by the equations of an SMC. For example, we have the following equivalences between diagrams. 
\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
\[
	\tikzfig{categorical-semantics/interchange}
\]

Unfortunately, this makes string diagrams relatively unamenable to efficient implementation as a data structure, due to the difficulty of calculating the quotient.

Nevertheless, there is a body of work dealing with this issue by representing string diagrams as combinatorial objects -- namely, (appropriately generalized) hypergraphs [cite]. Here, the generators $c_i$ become vertices which are connected via hyper-edges. Thus the expected quotient becomes simply hypergraph isomorphism. With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterize} the free symmetric monoidal category generated by some $c_i$ [Kissinger].

We are interested in not just the free SMC over a set of generators, but rather in \textit{symmetric monoidal theories} which involve considering extra equations between morphisms: for instance, the rewrite rules $(a)-(e)$ of Figure \ref{fig:e-graph-example}. These can be seen as equations between string diagrams, and thus between their hypergraph representations. Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identity $l$ in any hypergraph $G$ and replace it with $r$ (and vice-versa). 
The standard approach to this is known as\textit{double pushout (DPO) rewriting}. [Example].

Our application requires a generalization of these concepts from hypergraphs to \textit{e-hypergraphs}, which we define as a hypergraph with an additional relation denoting the hierarchical structure introduced by "e-boxes", \textit{i.e.,} the generator for semilattice enrichment.

[Example of combinatorial representation]

\subsection*{Soundness and Completeness}
Given the setup above, we prove soundness and completeness of the combinatorial representation for semilattice enriched categories, extending the results of [Kissinger] for symmetric monoidal categories. While the structural equalities of SMCs are factored out in the representation, the structural equalities arising from the enrichment are not, and should not be: they represent the un/sharing of subdiagrams via the enriched structure, which is precisely what allows for the compact representation of equivalence classes [see the equations of Figure 4]. Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating symmetric monoidal theory (which do not). Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched categories. Indeed, we have $f = g$ in the free category if and only if there exists a sequence of DPO-rewrites between their translations: $\llbracket f \rrbracket \rightsquigarrow \llbracket g \rrbracket$.

[Example equations]
a
\subsection*{E-matching and E-rewriting}

Having defined e-hypergraphs and shown them to correctly model rewriting string diagrams for SMCs, equipped with a notion of "equivalence class", we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph. In general, this involves unsharing the e-box structure before a redex becomes available. Thus, we define a notion of e-matching for e-hypergraphs: that is, to find redexes working modulo the e-structure. In particular, we wish to locate the smallest subgraph $G' \subset G$ "containing" the redex $L$. Given this, the analogue of e-rewriting $L \to R$ is simple to define, due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.

[Example diagrams]

\subsection*{Future Applications}

Example diagrams for circuits, trace (loop unrolling, PEGs), LC, ZX here.

\subsection{Related Work}
Q: Related work - at end? \\

### 2023-09-29 sections/soundness_and_completeness.tex
Initial Overleaf Import

\section{Soundness and Completeness}

\update{Below I kind of use $\llbracket - \rrbracket$ to define itself when using $\llbracket \mathcal{E} \rrbracket$.}
In this section, we will build a correspondence between terms in a semilattice enriched $\textsf{PROP}^{+}(\Sigma)$ and morphisms in $\WellTypedMdaEcospans$. Similarly to $\textsf{PROP}^{+}(\Sigma)$ where terms are defined up to semilattice enrichment and SMC equations we will say that two terms $G$ and $F$ in $\WellTypedMdaEcospans$ are equal when $G \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} F$. As SMC equations are absorbed by e-hypergraph representation we only quotient by $\llbracket \mathcal{E} \rrbracket$. We are now ready to build a semilattice-enriched SMC functor

\[
\llbracket - \rrbracket : \textsf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans
\]

by freeness, it suffices to define how the functor acts on generators. As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \catname{MACsp_{D}(Hyp_{\Sigma})}$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}. Everything else follows by defining

\begin{enumerate}
    \item $\llbracket f;g \rrbracket = \llbracket f \rrbracket ; \llbracket g \rrbracket$  for $f,g \in \Sigma$;
    \item $\llbracket f \otimes g \rrbracket = \llbracket f \rrbracket \otimes \llbracket g \rrbracket$  for $f,g \in \Sigma$;
    \item $\llbracket f + g \rrbracket = \llbracket f \rrbracket + \llbracket g \rrbracket$  for $f,g \in \Sigma$;
\end{enumerate}

this also makes the functor into a semilattice-enriched SMC one on the nose.

\update{Anything about soundness? I.e. if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$? Because $f = g$ is modulo semilattice-enriched SMC equations we have to show that $\llbracket f \rrbracket \Rrightarrow \llbracket g \rrbracket$. Doesn't it require the soundness of rewriting?
}

\begin{theorem}[$\llbracket - \rrbracket$ is faithful]

Essentially we need to show that if $\llbracket f \rrbracket = \llbracket g \rrbracket$ then $f = g$ which is equivalent to if $\llbracket f \rrbracket \not = \llbracket g \rrbracket$ then $f \not = g$. First, we note that we have a \question{weak (modulo commutativity)} normal form for any cospan in $\WellTypedMdaEcospans$ which is achieved by merging all hierarchical edges into a single hierarchical edge making it top-level, that is, every cospan is representable as below

\[
\scalebox{0.75}{
\tikzfig{soundness_and_completeness/normal_form}
}
\]

where all $A_1\; \ldots \;A_n$ do not contain any hierarchical edges. We will write such form as $A_1 + \ldots A_n$ Then, $\llbracket f \rrbracket \not = \llbracket g \rrbracket$ leads to two options

\begin{itemize}
    \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_m$ where $n \not = m$. By functoriality and weak normal form in $\textsf{PROP}^{+}(\Sigma)$, $f = a_1 + \ldots + a_n$, $g = b_1 + \ldots b_m$ and $\llbracket f \rrbracket = \llbracket a_1 \rrbracket + \ldots + \llbracket a_n \rrbracket$, $\llbracket g \rrbracket = \llbracket b_1 \rrbracket + \ldots \llbracket b_m \rrbracket$. This essentially means that $f \not = g$ as it has different number of summands. 
    \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_n$. Then there exists a permutation $\sigma$ and an index $i$ such that $ A_i  \not =  B_{\sigma(i)}$. By functoriality there exists $a_i$ such that $\llbracket a_i \rrbracket = A_i$ and $b_j$ such that $\llbracket b_j \rrbracket = B_{\sigma(i)}$. As $A_i$ does not have any hierarchical edges it effectively lives in $\mathcal{I}(\catname{MACsp_{D}(Hyp_{\Sigma})})$. By faithfulness of $\llbracket \cdot \rrbracket;\mathcal{I}$ it follows that $a_i \not = b_j$ and $f \not = g$. \update{Prove that $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful?}
\end{itemize}

\end{theorem}

### 2023-10-01 main.tex
test
some clean-up

@@
   , screen
-  , authordraft
+  % , authordraft
   ]{acmart}
@@
 % quiver is for quiver figures
-\usepackage{quiver}
+% \usepackage{quiver}
 \usepackage{tikz}
 \usepackage{tikz-network}
-\usepackage{amsthm}
 \usepackage{xspace}
@@
 
-\usepackage{mathtools,mathrsfs}
+% \usepackage{mathtools,mathrsfs}
 \usepackage{tikz-cd}
@@
 
-\title{Equivalence Hypergraphs: Semantics of E-Graphs and their Generalization to Monoidal Theories}
+\title{Equivalence Hypergraphs:\\ Semantics of E-Graphs and their Generalization to 
+Monoidal Theories}
 \author{Chris Barrett, Dan Ghica, Aleksei Tiurin}
@@
 
-% some sections might be commented out as overleaf compilation times out.
-
-% \input{sections/introduction}
+\input{sections/introduction}
 \input{sections/categorical_semantics}
 \input{sections/combinatorial_semantics}
-
 \input{sections/soundness_and_completeness}
-
 \input{sections/e-graph-translation}

### 2023-10-09 sections/introduction.tex
Fixesing inaccuracies for combinatorial semantics
Update introduction.tex

@@
 \cite{singher2023colored}
-\cite{koehler2022sketchguided} % latest egraph
 
@@
 
-Rewrite-driven program optimization involves applying a sequence of semantics-preserving rewrites which may improve speed of execution, resource usage, \textit{etc}.
+Rewrite-driven program optimization consists of applying a sequence of semantics-preserving rewrites which may improve speed of execution, resource usage, \textit{etc}.
 Because early application of some rewrites can enable or block the subsequent application of others, the choice of application order hugely impacts the quality of the resulting optimization.
@@
 
-Another recently proposed solution to this problem is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915,EggPaper}: instead of maintaining a \textit{single},  putatively optimized program which is rewritten \textit{destructively} at each step, we can maintain a \textit{set} of equivalent programs, where each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can be extracted% from the set,  which now represents all equivalent ways to express the program with respect to the given rewrites.
+A recently proposed,  alternative solution to this problem is \textit{equality saturation} 
+\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimized program which is rewritten \textit{destructively} at each step, we can maintain a \textit{set} of equivalent programs, where each rewrite step \textit{non-destructively} grows the set.
+Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted.% from the set,  which now represents all equivalent ways to express the program with respect to the given rewrites.
 
-The use of \textit{equality graphs (e-graphs)} to compactly represent the set of equivalent programs is what makes the technique tractable in practise, where the naive approach is clearly unfeasible.
+The use of \textit{equality graphs (e-graphs)} \cite{EggPaper} -- a generalization of the typical directed acylcic graph (DAG) representation of terms to include equivalence classes -- to compactly represent the set of equivalent programs is what makes the technique tractable in practise, where the naive approach is clearly unfeasible.
 
 While equality saturation is a state-of-the-art optimization technique, the formal mathematical theory of e-graphs is perhaps underdeveloped.
-In this paper, we contribute a categorical understanding of e-graphs.
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{enriched} Cartesian categories, where the enriched structure allows us to consider formal sets of terms (morphisms).
+In this paper, we contribute a categorical understanding of e-graphs and their rewriting.
+Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories, where the semilattice \textit{join} allows us to consider formal sets of terms (string diagrams, morphisms).
 
@@
 \cite{coecke_interacting_2011,ZX} and potentially to natural language \cite{wazni_quantum_2022,coecke_lambek_2013}.
+We also suggest how our approach will allow future  generalization to include rewriting for the $\lambda$-calculus
+\cite{koehler2022sketchguided} % latest egraph.
 
 Towards implementation, we 
-additionally provide a combinatorial representation of our enriched string diagrams in terms of hypergraphs.
+additionally provide a combinatorial representation of our enriched string diagrams in terms of (hierarchical)  hypergraphs.
 We give a specification of hypergraph rewriting in this novel setting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
+\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  thus also formalizing the rewriting theory of standard e-graphs.
 
@@
 
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.
-They generalize the typical directed acylcic graph (DAG) representation of terms to include equivalence classes.
+E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory. They generalize the typical DAG representation of terms to include equivalence classes. 
 A sequence of examples, taken from \cite{EggPaper}, are given in the first row of Figure \ref{fig:e-graph-example}.
@@
 
-Beginning with the e-graph (a), corresponding to the directed acyclic graph (DAG) representation of the term $(a*2)/2$, each subsequent e-graph (b)-(e) uses the given rewrite rule to add information: creating new nodes and edges, or merging equivalence classes.
+Beginning with the e-graph (a), corresponding to the DAG representation of the term $(a*2)/2$, each subsequent e-graph (b)-(e) uses the given rewrite rule to add information: creating new nodes and edges, or merging equivalence classes.
 The final e-graph, (e), contains a cycle, and represents infinitely many terms: $a, a*1, (a*1)*1, $...
@@
 String diagrams are read from bottom-to-top and we consider additional generators for the duplication and deletion transformations of a Cartesian category.
+Note, further, that in the Cartesian case we can restrict to generators $c$ with a single output.
 
 The translation from e-graphs to our enriched string diagrams is illustrated informally below. 
-Note how the typing constraints of the $+$ constructor are satisfied.
+Note how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component unnecessary inputs.
 \[
@@
 
-In order to implement our generalized version of e-graphs, we require to perform rewriting not on algebraic terms, but on string diagrams.
+In order to \textit{implement} our generalized version of e-graphs, we will  perform rewriting not on algebraic terms, but on combinatorial representations of the corresponding string diagrams.
 String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo "connectivity") or as a 2-dimensional syntax quotiented by the equations of an SMC.
 For example, we have the following equivalences between diagrams.
-\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
+%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
 \[
@@
 
-Nevertheless, there is a body of work dealing with this issue by representing string diagrams as combinatorial objects -- namely, (appropriately generalized) hypergraphs [cite].
+Nevertheless, there is a body of work dealing with this issue by representing string diagrams as combinatorial objects -- namely, (appropriately generalized) hypergraphs 
+\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
 Here, the generators $c_i$ become vertices which are connected via hyper-edges.
 Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterize} the free symmetric monoidal category generated by some $c_i$ [Kissinger].
+With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterize} the free symmetric monoidal category generated by some $c_i$.
 
@@
 These can be seen as equations between string diagrams, and thus between their hypergraph representations.
-Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identity $l$ in any hypergraph $G$ and replace it with $r$ (and vice-versa).
-The standard approach to this is known as\textit{double pushout (DPO) rewriting}. [Example].
+Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ in any hypergraph $G$ and replace it with (a sub-hypergraph corresponding to) $r$.
+The standard approach to this is known as\textit{double pushout (DPO) rewriting}.
 
-Our application requires a generalization of these concepts from hypergraphs to \textit{e-hypergraphs}, which we define as a hypergraph with an additional relation denoting the hierarchical structure introduced by "e-boxes", \textit{i.e.,} the generator for semilattice enrichment.
-
-[Example of combinatorial representation]
+Our application requires a generalization of these concepts from hypergraphs to \textit{e-hypergraphs}, which we define as a hypergraph with two additional relations denoting the hierarchical structure introduced by "e-boxes" ( \textit{i.e.,} the generator for semilattice enrichment) and the separation of the components of each e-box.
+Giving the appropriate definitions is slightly delicate matter, and constitutes the main body of the paper.
 
 \subsection*{Soundness and Completeness}
-Given the setup above, we prove soundness and completeness of the combinatorial representation for semilattice enriched categories, extending the results of [Kissinger] for symmetric monoidal categories.
-While the structural equalities of SMCs are factored out in the representation, the structural equalities arising from the enrichment are not, and should not be: they represent the un/sharing of subdiagrams via the enriched structure, which is precisely what allows for the compact representation of equivalence classes [see the equations of Figure 4].
+The main technical result of the paper is a proof soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs.
+While the structural equalities of SMCs are factored out in the representation,  important structural equalities arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be: they represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
 Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating symmetric monoidal theory (which do not).
+
 Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched categories.
-Indeed, we have $f = g$ in the free category if and only if there exists a sequence of DPO-rewrites between their translations: $\llbracket f \rrbracket \rightsquigarrow \llbracket g \rrbracket$.
+That is, we exhibit a translation functor $\llbracket-\rrbracket$ from the free semilattice enriched SMC over a symmetric monoidal theory to our (quotiented) category of e-hypergraphs such that
+morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow \llbracket_{DPO} g \rrbracket$.
 
-[Example equations]
-a
 \subsection*{E-matching and E-rewriting}

### 2023-10-09 sections/combinatorial_semantics.tex
some clean-up

@@
+% !TEX root = ../main.tex
+
 \section{Combinatorial semantics}
 
-% TODO: fix union to Kleene star, introduce notation, put explanation out of def environmentnt, fix notation for in(e) and out(e), conflict relation is not transitive
-
 \update{Below I will assume that we work with a simple PROP (not coloured)}
@@
 Before we begin we first fix some notation.
-When $V$ is a set, we denote with $V^{*}$ a set of all finite sequences of the elements of $V$, i.e.
-$^*$ is a Kleene star.
-When $\phi : V \to V'$ is a function, we denote with $\phi^{*}$ an extension of $\phi$ to sequences, i.e.
-$\phi^{*} : V^* \to V'^{*}$ and it applies $\phi$ element-wise.
+When $V$ is a set, we denote with $V^{*}$ a set of all finite sequences of the elements of $V$, i.e. $^*$ is a free monoid.
+When $\phi : V \to V'$ is a function, we denote with $\phi^{*}$ an extension of $\phi$ to sequences, i.e. $\phi^{*} : V^* \to V'^{*}$ and it applies $\phi$ element-wise.
 Similarly, in the case when $\psi \subset V \times V$ is a relation: we denote with $\psi^{*}: \subset V^* \times V^*$ the element-wise relation on ordered sequences.
+We will denote with $A +_{f,g} B$ the pushout of $A \xleftarrow{f} C \xrightarrow{g} B$.
 
@@
 
-\begin{definition}[Hypergraph]
+\begin{definition}[Category $\catname{Hyp_{\Sigma}}$]
+
+A hypergraph $\mathcal{G}_{\Sigma}$ over a monoidal signature $\Sigma$ is a tuple 
+\[
+(V_{\mathcal{G}},E_{\mathcal{G}},s,t,l)
+\]
 \label{def:hypergraph}
-
-A hypergraph $\mathcal{G}_{\Sigma}$ over a monoidal signature $\Sigma$ is a tuple 
-\[
-(V_{\mathcal{G}},E_{\mathcal{G}},s,t,l)
-\]
 
@@
 We will omit the subscripts when they are clear from the context.
-\end{definition}
-
-\begin{definition}[Hypergraph homomorphism]
 
@@
 
-where $\phi^{*}$ is an element-wise application of $\phi$.
-\end{definition}
-
-\begin{definition}[Category $\catname{Hyp_{\Sigma}}$]
-
 Hypergraphs over a monoidal signature $\Sigma$ together with hypergraph homomorphisms form a category $\catname{Hyp_{\Sigma}}$. 
@@
 
-\begin{definition}[Definition 2.10~\cite{Frobenius}] Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-
-A morphism $\alpha : (X \xrightarrow{} A \xleftarrow{} Y) \Rightarrow (X \xrightarrow{} B \xleftarrow{} Y)$ between cospans is an arrow $\alpha : A \to B$ in $\mathbb{C}$ such that the following diagram commutes.
+\begin{definition}[Definition 2.10~\cite{Frobenius}]
+Let $\mathbb{C}$ be a category with all finite colimits. A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
+
+Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
 
@@
 
-Cospans are isomorphic if $\alpha$ is an isomorphism.
 For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
@@
 
-\[
-\adjustbox{scale=0.75,center}{
-% https://q.uiver.app/#q=WzAsMTQsWzAsMCwibiJdLFsxLDAsIlxcbWF0aGNhbHtHfSJdLFsyLDAsIm0iXSxbMSwxLCJcXG90aW1lcyJdLFswLDIsIm4nIl0sWzIsMiwibSciXSxbMSwyLCJcXG1hdGhjYWx7Ryd9Il0sWzMsMSwiOj0iXSxbNCwxLCJuK24nIl0sWzYsMSwiXFxtYXRoY2Fse0d9K1xcbWF0aGNhbHtHJ30iXSxbOCwxLCJtK20nIl0sWzEwLDEsIjAiXSxbMTEsMSwiMCJdLFsxMiwxLCIwIl0sWzAsMSwiZiJdLFsyLDEsImciLDJdLFs0LDYsImYnIl0sWzUsNiwiZyciLDJdLFs4LDksImYrZiciXSxbMTAsOSwiZytnJyIsMl0sWzExLDEyLCJpZF8wIl0sWzEzLDEyLCJpZF8wIiwyXV0=
-\begin{tikzcd}
-	n & {\mathcal{G}} & m \\
-	& \otimes && {:=} & {n+n'} && {\mathcal{G}+\mathcal{G'}} && {m+m'} && 0 & 0 & 0 \\
-	{n'} & {\mathcal{G'}} & {m'}
-	\arrow["f", from=1-1, to=1-2]
-	\arrow["g"', from=1-3, to=1-2]
-	\arrow["{f'}", from=3-1, to=3-2]
-	\arrow["{g'}"', from=3-3, to=3-2]
-	\arrow["{f+f'}", from=2-5, to=2-7]
-	\arrow["{g+g'}"', from=2-9, to=2-7]
-	\arrow["{id_0}", from=2-11, to=2-12]
-	\arrow["{id_0}"', from=2-13, to=2-12]
-\end{tikzcd}
-}
-\]
-
-Symmetry is inherited from $\mathcal{C}$ as well as the coproduct is commutative and is given below
+% \[
+% \adjustbox{scale=0.75,center}{
+% https://q.uiver.app/#q=WzAsMTEsWzAsMCwiQSJdLFsxLDAsIlxcbWF0aGNhbHtHfSJdLFsyLDAsIkIiXSxbMywwLCJcXG90aW1lcyJdLFs0LDAsIkEnIl0sWzYsMCwiQiciXSxbNSwwLCJcXG1hdGhjYWx7Ryd9Il0sWzcsMCwiOj0iXSxbOCwwLCJBK0EnIl0sWzEwLDAsIlxcbWF0aGNhbHtHfStcXG1hdGhjYWx7Ryd9Il0sWzEyLDAsIkIrQiciXSxbMCwxLCJmIl0sWzIsMSwiZyIsMl0sWzQsNiwiZiciXSxbNSw2LCJnJyIsMl0sWzgsOSwiZitmJyJdLFsxMCw5LCJnK2cnIiwyXV0=
+% \begin{tikzcd}
+% 	A & {\mathcal{G}} & B & \otimes & {A'} & {\mathcal{G'}} & {B'} & {:=} & {A+A'} && {\mathcal{G}+\mathcal{G'}} && {B+B'}
+% 	\arrow["f", from=1-1, to=1-2]
+% 	\arrow["g"', from=1-3, to=1-2]
+% 	\arrow["{f'}", from=1-5, to=1-6]
+% 	\arrow["{g'}"', from=1-7, to=1-6]
+% 	\arrow["{f+f'}", from=1-9, to=1-11]
+% 	\arrow["{g+g'}"', from=1-13, to=1-11]
+% \end{tikzcd}
+\[
+    A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B \otimes A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B' := A + A' \xrightarrow{f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g'} B + B'
+\]
+
+Symmetry is inherited from $\mathcal{C}$ as the coproduct is commutative and is given below.
 
@@
 
-When $\mathbb{C} = Hyp_{\Sigma}$ we get a category of cospans of hypergraphs.
+When $\mathbb{C} = \catname{Hyp_{\Sigma}}$ we get a category of cospans of hypergraphs.
+$\catname{Hyp_{\Sigma}}$ has all finite colimits and, in particular, coproduct is a disjoint union of two hypergraphs.
 An empty hypergraph is the initial object in $Hyp_{\Sigma}$ and coproduct and pushout are defined on the underlying sets of nodes and edges.
@@
 \label{def:cspd}
-Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{Csp_{D}(Hyp_{\Sigma})}$ is of the form
-
-\[
-n \xrightarrow{} \mathcal{F} \xleftarrow{} m
-\]
-
-where $n,\;m$ are discrete hypergraphs and $\mathcal{F}$ is an arbitrary hypergraph.
+Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a \textit{full} sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
+\end{definition}
+
+In particular, this means that any morphism in $\catname{Csp_{D}(Hyp_{\Sigma})}$ is of the form $n \xrightarrow{} \mathcal{G} \xleftarrow{} m$,
+where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
 When string diagrams are interpreted as cospans of hypergraphs $n,m$ correspond to input and output interfaces of a string diagram respectively.
-
-A pictorial representation of such a cospan is depicted below in Figure~\ref{fig:hypergraph_and_string_diagram}. We use node labels to define mappings $n \to \mathcal{F}$ and $m \to \mathcal{F}$.
+A pictorial representation of such a cospan is depicted below in Figure~\ref{fig:hypergraph_and_string_diagram}.
+We use node labels to define mappings $n \to \mathcal{F}$ and $m \to \mathcal{F}$.
 
@@
     \[
-    \scalebox{0.75}{
+    \scalebox{0.6}{
 	\tikzfig{combinatorial_semantics/mda-hypergraph-example}
@@
     \end{subfigure}
+    \hfill
     \begin{subfigure}{0.45\linewidth}
         \[
+        \raisebox{-0.6cm}{
         \scalebox{0.75}{
         \tikzfig{combinatorial_semantics/mda-hypergraph-example-string}
-        }
+        }}
         \]
@@
 \begin{definition}[Path]
-    A path from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{F}$ is an ordered sequence of edges $e_i \in E_{F}$ : $[e_0, e_1, \ldots, e_{n-1}]$ such that for all $i < n - 1$ some node $v \in t(e_i)$ is also in $s(e_{i+1})$.
+    A path from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $e_i \in E_{G}$ : $[e_0, e_1, \ldots, e_{n-1}]$ such that for all $i < n - 1$ some node $v \in t(e_i)$ is also in $s(e_{i+1})$.
 \end{definition}
@@
 
-We call a cospan $n \xrightarrow{f} \mathcal{F} \xleftarrow{g} m$ monogamous directed acyclic if
+We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\catname{Csp_{D}(Hyp_{\Sigma})}$ monogamous directed acyclic if
 
 \begin{enumerate}
-    \item $\mathcal{F}$ is a directed acyclic hypergraph
+    \item $\mathcal{G}$ is a directed acyclic hypergraph
     \item $f$ and $g$ are monos
-    \item the in-degree and out-degree of every vertex of $\mathcal{F}$ is at most 1
-    \item vertices of $\mathcal{F}$ with in-degree 0 are precisely the image of $f$
-    \item vertices of $\mathcal{F}$ with out-degree 0 are precisely the image of $g$
+    \item the in-degree and out-degree of every vertex of $\mathcal{G}$ is at most 1
+    \item vertices of $\mathcal{G}$ with in-degree 0 are precisely the image of $f$
+    \item vertices of $\mathcal{G}$ with out-degree 0 are precisely the image of $g$
 \end{enumerate}
@@
 
-\begin{definition}[Category $\catname{MACsp_{D}(Hyp_{\Sigma})}$]
-    When we narrow our attention to cospans (morphisms) of $\catname{Csp_{D}(Hyp_{\Sigma})}$ which are monogamous we get the category of monogamous directed acyclic (mda-) cospans of hypegraphs $\catname{MACsp_{D}(Hyp_{\Sigma})}$.
-Vertical and horizontal compositions of mda-cospans are mda-cospans, identities and symmetries are mda-cospans, and hence this is a proper category~\cite{Frobenius2}.
-\end{definition}
+When we narrow our attention to cospans (morphisms) of $\catname{Csp_{D}(Hyp_{\Sigma})}$ which are monogamous we get the category of monogamous directed acyclic (mda-) cospans of hypegraphs $\MdaCospans$.
+
+\begin{definition}[Category $\MdaCospans$]
+    We define $\MdaCospans$ as a subcategory of $\catname{Csp_{D}(Hyp_{\Sigma})}$.
+\end{definition}
+
+Vertical and horizontal compositions of mda-cospans are mda-cospans, identities and symmetries are mda-cospans, tensor unit is discrete, and hence this is a proper symmetric monoidal category~\cite{Frobenius2}.
 
 \begin{theorem}[Corollary 26~\cite{Frobenius2}]
-    $\textsf{PROP}(\Sigma) \cong \catname{MACsp_{D}(Hyp_{\Sigma})}$
+    $\textsf{PROP}(\Sigma) \cong \MdaCospans$
 \end{theorem}
 This theorem essentially states a correspondence between morphisms in SMC (string diagrams) and mda-cospans of hypergraphs.
-% This section presents a combinatorial structure that defines our e-hypergraphs.
-
-% We closely follow a set of definitions designed in~\cite{fscd}. We extend this set to support an additional e-structure.
 
@@
 \end{definition}
+
+\Aleksei{Switch to the complent in the definition of the relation and call it a \textit{consistency} relation}
 
@@
 
-In other words, homomorphism should preserve immediate predecessors and preserve and reflect the conflict relation.
+In other words, homomorphism should preserve immediate predecessors and the conflict relation.
 
@@
 Such e-hypergraphs form a category $\catname{EHyp_{\Sigma}}$ where morphisms are e-hypergraph homomorphisms defined above.
+In particular, it has a coproduct which is a disjoint union of e-hypergraphs and a tensor unit which is an empty e-hypergraphs with empty relations which makes it into a monoidal category.
 \end{definition}
@@
     Consider an example span below in Figure~\ref{fig:pushout_fail}.
-
 \begin{figure}[h!]
@@
     Luckily, we do not need arbitrary pushouts and pushouts that we actually need do exist.
+    \update{In general, in the pushout $A \xrightarrow{f} D \xleftarrow{g} B$, $D$ would inherit the relations from $A$ and $B$ when the relations are not mutually exclusive.}
 \end{remark}
@@
 
-\update{I will build below $\catname{ECsp_{D}(EHyp_{\Sigma})}$ (E for extended) instead of $ECsp(\mathcal{C})$ for arbitrary $\mathcal{C}$ with colimits to avoid talking about pushouts in general below, because we use set difference.}
-\begin{definition}[Category $\catname{ECsp_{D}(EHyp_{\Sigma})}$]
+\update{I will build below $\Ecospans$ (E for extended) instead of $ECsp(\mathcal{C})$ for arbitrary $\mathcal{C}$ with colimits to avoid talking about pushouts in general below, because we use set difference.}
+\begin{definition}[Category $\Ecospans$]
     We will build the category of \textit{extended} cospans of e-hypergraphs as follows.
-    Let the objects of $\catname{ECsp_{D}(EHyp_{\Sigma})}$ be only discrete e-hypergraphs of $\catname{EHyp_{\Sigma}}$ and morphisms from $ext(X) \to ext(Y)$ be the cospans (ext stands for external)
-
-    \[
-    ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
-    \] where $A \in obj(EHyp_{\Sigma})$.
-
-    \update{Can I say that a morphism between two such cospans is 3 arrows?}
+    Let the objects of $\Ecospans$ be only discrete e-hypergraphs (sets of edges and nodes are empty as well as both child and conflict relations) of $\catname{EHyp_{\Sigma}}$ and morphisms from $W \to Z$ be the cospans
+
+    \[
+    W \xrightarrow{f} X \xrightarrow{g} A \xleftarrow{h} Y \xleftarrow{k} Z
+    \] 
+    where $A \in obj(\catname{EHyp_{\Sigma}})$, $W, X, Y, Z$ are discrete and $f,g,h,k$ are monos.
+    Since $f$ and $k$ are monos, $W$ and $Z$ are essentially subgraphs of $X$ and $Y$ and hence we will introduce the following notation which better depicts that such cospans are effectively e-hypergraphs with \textit{external} interfaces
+
+    \[
+    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
+    \]
 
@@
 
+    \[
+    \scalebox{0.75}{
+        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
+    }
+    \]
+
 % https://q.uiver.app/#q=WzAsOCxbMCwxLCJleHQoWCkiXSxbMSwwLCJYIl0sWzIsMCwiQSJdLFszLDAsIn5ZIl0sWzQsMSwiZXh0KFkpIl0sWzEsMiwiWCciXSxbMiwyLCJBJyJdLFszLDIsIlknIl0sWzAsMSwiIiwyLHsiY3VydmUiOi0yfV0sWzEsMl0sWzMsMl0sWzQsMywiIiwwLHsiY3VydmUiOjJ9XSxbNCw3LCIiLDIseyJjdXJ2ZSI6LTJ9XSxbNyw2XSxbNSw2XSxbMCw1LCIiLDAseyJjdXJ2ZSI6Mn1dLFsxLDUsIlxcYWxwaGEiLDFdLFsyLDYsIlxcYmV0YSIsMV0sWzMsNywiXFxnYW1tYSIsMV1d
 
-\[
-\begin{tikzcd}
-	& X & A & {~Y} \\
-	{ext(X)} &&&& {ext(Y)} \\
-	& {X'} & {A'} & {Y'}
-	\arrow[curve={height=-12pt}, from=2-1, to=1-2]
-	\arrow[from=1-2, to=1-3]
-	\arrow[from=1-4, to=1-3]
-	\arrow[curve={height=12pt}, from=2-5, to=1-4]
-	\arrow[curve={height=-12pt}, from=2-5, to=3-4]
-	\arrow[from=3-4, to=3-3]
-	\arrow[from=3-2, to=3-3]
-	\arrow[curve={height=12pt}, from=2-1, to=3-2]
-	\arrow["\alpha"{description}, from=1-2, to=3-2]
-	\arrow["\beta"{description}, from=1-3, to=3-3]
-	\arrow["\gamma"{description}, from=1-4, to=3-4]
-\end{tikzcd}
-\]
+\Aleksei{Redraw this diagram in tikzit because importing 'quiver' introduces error}
+
+% \[
+% \begin{tikzcd}
+% 	& X & A & {~Y} \\
+% 	{X_{ext}} &&&& {Y_{ext}} \\
+% 	& {X'} & {A'} & {Y'}
+% 	\arrow[curve={height=-12pt}, from=2-1, to=1-2]
+% 	\arrow[from=1-2, to=1-3]
+% 	\arrow[from=1-4, to=1-3]
+% 	\arrow[curve={height=12pt}, from=2-5, to=1-4]
+% 	\arrow[curve={height=-12pt}, from=2-5, to=3-4]
+% 	\arrow[from=3-4, to=3-3]
+% 	\arrow[from=3-2, to=3-3]
+% 	\arrow[curve={height=12pt}, from=2-1, to=3-2]
+% 	\arrow["\alpha"{description}, from=1-2, to=3-2]
+% 	\arrow["\beta"{description}, from=1-3, to=3-3]
+% 	\arrow["\gamma"{description}, from=1-4, to=3-4]
+% \end{tikzcd}
+% \]
 
@@
 
-Composition of two morphisms $f : ext(X) \to ext(Y)$ and $g : ext(Y) \to ext(Z)$ is computed in two stages. First, $A +_{f_{ext};f, g_{ext};g} B$ is computed which is a pushout below
+Composition of two morphisms $f : X_{ext} \to Y_{ext}$ and $g : Y_{ext} \to Z_{ext}$ is computed in two stages
+First, $A +_{f_{ext};f, g_{ext};g} B$ is computed which is a pushout below
 
 % https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[\begin{tikzcd}
-	{ext(X)} &&& Y & {ext(Y)} & {Y'} &&& {ext(Z)} \\
-	& X &&&&&& Z \\
-	&& A &&&& B \\
-	&&&& {^{\ulcorner}A+_{f_{ext};f,g_{ext};g}B}
-	\arrow[from=1-1, to=2-2]
-	\arrow[from=2-2, to=3-3]
-	\arrow["f", from=1-4, to=3-3]
-	\arrow["{f_{ext}}", from=1-5, to=1-4]
-	\arrow["{g_{ext}}"', from=1-5, to=1-6]
-	\arrow["g"', from=1-6, to=3-7]
-	\arrow[from=2-8, to=3-7]
-	\arrow[from=1-9, to=2-8]
-	\arrow[from=3-3, to=4-5]
-	\arrow[from=3-7, to=4-5]
-\end{tikzcd}
-\]
-
-Then, new legs for $A +_{f_{ext};f,g_{ext};g} B$ are computed as below
-
-\[
-ext(X) \xrightarrow{f_{ext}} X \uplus (Y' \setminus ext(Y)) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus ext(Y)) \xleftarrow{g_{ext}} ext(Z)
-\]
-
-since $X,Y,Z,Y', ext(Y)$ are all discrete, $\uplus$ and $\setminus$ are standard set operations. $\setminus$ could be also formalised further as the following pushout complement
-
-\[
-\begin{tikzcd}
-ext(Y) \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-Y'          & Y' \setminus ext(Y) \arrow[l]
-\end{tikzcd}
-\]
-
-and then the disjoint union is just a coproduct.
+
+\[
+\scalebox{0.9}
+{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
+\]
+
+Then, new feet for $A +_{f_{ext};f,g_{ext};g} B$ are computed as below
+
+\[
+X_{ext} \xrightarrow{f_{ext}} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext}} Z_{ext}
+\]
+
+since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
+
+% \[
+% \begin{tikzcd}
+% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
+% Y'          & Y' \setminus Y_{ext} \arrow[l]
+% \end{tikzcd}
+% \]
+
+% and then the disjoint union is just a coproduct. 
 Defining cospans up to isomorphism, associativity of coproducts and pushouts, and the presence of an id-cospan turns this construction into a valid category.
@@
 Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into a SMC.
-More precisely, below are cospans for tensor product, its unit and symmetry.
+More precisely, below are cospans for tensor product and symmetry. Tensor unit is given by the unit in $\catname{EHyp_{\Sigma}}$.
 
@@
 \begin{tikzcd}
-	{ext(X)} & X & A & Y & {ext(Y)} \\
-	&& \otimes &&& {:=} & {ext(X)+ext(X')} && {X+X'} && {A+B} && {Y+Y'} && {ext(Y)+ext(Y')} \\
-	{ext(X')} & {X'} & B & {Y'} & {ext(Y')}
+	{X_{ext}} & X & A & Y & {Y_{ext}} \\
+	&& \otimes &&& {:=} & {X_{ext}+ext(X')} && {X+X'} && {A+B} && {Y+Y'} && {Y_{ext}+Y'_{ext}} \\
+	{ext(X')} & {X'} & B & {Y'} & {Y'_{ext}}
 	\arrow["{f_{ext}}", from=1-1, to=1-2]
@@
 
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCIwIl0sWzEsMCwiMCJdLFsyLDAsIjAiXSxbMywwLCIwIl0sWzQsMCwiMCJdLFswLDEsImlkXzAiXSxbMSwyLCJpZF8wIl0sWzMsMiwiaWRfMCIsMl0sWzQsMywiaWRfMCIsMl1d
-\[
-\adjustbox{scale=0.75, center}{
-\begin{tikzcd}
-	0 & 0 & 0 & 0 & 0
-	\arrow["{id_0}", from=1-1, to=1-2]
-	\arrow["{id_0}", from=1-2, to=1-3]
-	\arrow["{id_0}"', from=1-4, to=1-3]
-	\arrow["{id_0}"', from=1-5, to=1-4]
-\end{tikzcd}
-}
-\]
-
 % https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
@@
 \begin{remark}
-    When writing things like $Y' \setminus ext(Y)$ we mean $Y' \setminus g(Y')$ and so on.
+    When writing things like $Y' \setminus Y_{ext}$ we mean $Y' \setminus g(Y')$ and so on.
 \end{remark}
@@
 \begin{remark}
-An example of such a morphism in $\catname{ECsp_{D}(EHyp_{\Sigma})}$ can be found in Figure~\ref{fig:e-cospan}.
-The legs of a cospan are depicted on top and at the bottom from the grey-shaded area.
-These rectangles correspond to $ext(X)$, $X$, $Y$, and $ext(Y)$ respectively.
-In order to not clutter the figures we will collapse the legs, e.g.
-we will
- collapse $ext(X)$ and $X$ into one rectangle and delimit the image of the former within the latter with a vertical dashed line.
+An example of such a morphism in $\Ecospans$ can be found in Figure~\ref{fig:e-cospan}.
+The feet of a cospan are depicted on top and at the bottom from the grey-shaded area.
+These rectangles correspond to $X_{ext}$, $X$, $Y$, and $Y_{ext}$ respectively.
 \end{remark}
@@
     \]
-    \caption{Morphism in $\catname{ECsp_{D}(EHyp_{\Sigma})}$}
+    \caption{Morphism in $\Ecospans$}
     \label{fig:e-cospan}
@@
 \[
-    ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
+    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
 \]
@@
 When talking about cospans we will call $X$ and $Y$ input and output interfaces respectively.
-We will also call $ext(X)$ and $ext(Y)$ the outermost or external input and output interfaces and $X \setminus ext(X)$ and $Y \setminus ext(Z)$ the internal input and output interfaces.
-By abusing the notation we will also refer to the images of $ext(X)$, $X$, $Y$, $ext(Y)$, $X \setminus ext(X)$, $Y \setminus ext(Y)$ as to interfaces of the corresponding type.
-Graphically we distinguish the outermost part of the interface with a vertical line.
+We will also call $X_{ext}$ and $Y_{ext}$ the outermost or external input and output interfaces and $X \setminus X_{ext}$ and $Y \setminus Z_{ext}$ the internal input and output interfaces.
+By abusing the notation we will also refer to the images of $f_{ext}$, $g_{ext}$ and $f$, $g$ (as well as their versions narrowed to $X \setminus X_{ext}$ and $Y \setminus Y_{ext}$), as to interfaces of the corresponding type.
 \end{definition}
@@
     In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$.
+    The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
+    \update{We should probably require that the image of $f_{ext};f{}$ has no predecessors and has no incoming edges}
     Note that pushout for such composition always exists as the nodes that are being identified are unrelated: they are not in conflict with anything and the sets of their predecessors are empty.
@@
 
-\caption{$\catname{ECsp(EHyp_{\Sigma})}$ composition example}
+\caption{$\Ecospans$ composition example}
 \label{fig:e-hyp_composition_example}
@@
 
+\Aleksei{what is $h$? $k$? $f_ext: ext(X) -> X$, so how is $f_ext$ also of type $ext(X) -> ext(X) + X + X'$ ? }
 \begin{definition}
@@
     \[
-    ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
-    \]
-    \[
-    ext(X) \xrightarrow{f'_{ext}} X' \xrightarrow{f'} B \xleftarrow{g'} Y' \xleftarrow{g'_{ext}} ext(Y)
+    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
+    \]
+    \[
+    X_{ext} \xrightarrow{f'_{ext}} X' \xrightarrow{f'} B \xleftarrow{g'} Y' \xleftarrow{g'_{ext}} Y_{ext}
     \]
@@
     \[
-    ext(X) \xrightarrow{f_{ext}} ext(X) + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} ext(Y) + Y + Y' \xleftarrow{g_{ext}} ext(Y)
-    \]
+    X_{ext} \xrightarrow{f_{ext}} X_{ext} + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} Y_{ext} + Y + Y' \xleftarrow{g_{ext}} Y_{ext}
+    \]
+    \label{morphism:composition_ecospans}
 
     $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
-    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|ext(X)|$ sources and $|ext(Y)|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v)$.
+    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v)$.
 And such that its sources and targets are top-level nodes.
-$h$ and $k$ are morphisms that map $ext(X)$ and $ext(Y)$ to these sources and targets respectively.
+$h$ and $k$ are morphisms that map $X_{ext}$ and $Y_{ext}$ to these sources and targets respectively.
 Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
@@
 
-\begin{definition}[Mda-cospan in $\catname{ECsp_{D}(EHyp_{\Sigma})}$]
+\begin{definition}[Mda-cospan in $\Ecospans$]
 
@@
 \[
-ext(X) \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} ext(Y)
+X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
 \]
@@
         \item The in-degree and out-degree of every node in $A$ are at most 1.
-        \item Nodes with the in(out)-degree of 0 are exactly the nodes in the image of $f$ ($g$).
-        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ ($g$) which pre-image is not in $ext(X)$ ($ext(Y)$). 
+        \item Nodes with the in(out)-degree of 0 are exactly the nodes in the image of $f$ (respectively, $g$).
+        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ ($g$) which pre-image is not in $X_{ext}$ ($Y_{ext}$). 
         \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
 \end{enumerate}
+\label{def:monogamy_ehyp}
 
@@
 
-\begin{proposition}
-    Id morphism
-
-    \[
-    X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
-    \]
-
-    is an mda-cospan.
+\begin{lemma}
+    \begin{enumerate}
+        \item The tensor product of two mda-cospans is an mda-cospan.
+        \item $+$ of two mda-cospans is an mda-cospan.
+    \end{enumerate}
     \begin{proof}
-        Proof by checking all the constraints above.
+        Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
+        Hence, everything is preserved component-wise by copairing and coproduct.
+    \end{proof}
+    \begin{proof}
+        By construction in Figure~\ref{fig:A+B}.
+    \end{proof}
+\end{lemma}
+
+% \begin{proposition}
+%     Id morphism
+
+%     \[
+%     X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
+%     \]
+
+%     is an mda-cospan.
+%     \begin{proof}
+%         Proof by checking all the constraints above.
 Discrete e-hypergraphs is obviously directed and acyclic.
@@
 Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
-    \end{proof}
-\end{proposition}
+%     \end{proof}
+% \end{proposition}
 
@@
 
-\begin{proposition}
-    Tensor unit
-
-    \[
-    0 \xrightarrow{id_{0}} 0 \xrightarrow{id_{0}} 0 \xleftarrow{id_{0}} 0 \xleftarrow{id_0} 0
-    \]
-
-    is an mda-cospan.
-    \begin{proof}
-        The empty e-hypergraph is obviously directed and acyclic. Id-morphisms are monos. The bullets (3) --- (6) are satisfied by the absence of nodes.
-    \end{proof}
-\end{proposition}
-
-\begin{proposition}
-    The tensor product of two mda-cospans is again an mda-cospan.
-
-    \begin{proof}
-        Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$. Hence, everything is preserved component-wise by copairing and coproduct.
-    \end{proof}
-\end{proposition}
-
-\begin{proposition}
-    Composition of two mda-cospans 
-    \[
-ext(X) \xrightarrow{f_{ext}} X \uplus (Y' \setminus ext(Y)) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus ext(Y)) \xleftarrow{g_{ext}} ext(Z)
-\]
-
-is again an mda-cospan.
-
-    \begin{proof}
+% \begin{proposition}
+%     Composition of two mda-cospans 
+%     \[
+% X_{ext} \xrightarrow{f_{ext}} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext}} Z_{ext}
+% \]
+
+% is again an mda-cospan.
+
+%     \begin{proof}
+%         Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
+This also does not increase their degree to more than 1.
+$f_{ext}$ is mono because its image $X$ is unchanged.
+The image of $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
+Pushout is then obtained by identifying the nodes with the pre-image in $Y_{ext}$ which are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
+$g';i_2$ yields $h$ and therefore $h$ is mono.
+Likewise, $k$ is mono.
+%     \end{proof}
+%     \update{Is this proof ok?}
+% \end{proposition}
+
+% \begin{proposition}
+%     $+$ of two mda-cospans
+
+%         \[
+%     X_{ext} \xrightarrow{} X \xrightarrow{} A \xleftarrow{} Y \xleftarrow{} Y_{ext}
+%     \]
+%     \[
+%     X_{ext} \xrightarrow{} X' \xrightarrow{} B \xleftarrow{} Y' \xleftarrow{} Y_{ext}
+%     \]
+
+%     is an mda-cospan.
+
+%     \begin{proof}
+%         By construction in Figure~\ref{fig:A+B}.
+%     \end{proof}
+% \end{proposition}
+
+\begin{definition}[Category $\MdaEcospans$]
+
+    Narrowing the morphisms of $\Ecospans$ to mda-cospans gives us a subcategory $\MdaEcospans$.
+    Id morphims is given by the following cospan
+
+    \[
+    X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
+    \]
+
+    which is an mda-cospan by checking all the necessary constraints in~\ref{def:monogamy_ehyp}.
+    Discrete e-hypergraph $X$ is obviously directed and acyclic.
+    Id-morphisms are monos.
+    The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$.
+    Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
+
+    Composition is given in~\ref{morphism:composition_ecospans} and composing two mda-cospans yields an mda-cospan.
     Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
@@
     The image of $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
-Pushout is then obtained by identifying the nodes with the pre-image in $ext(Y)$ which are removed in $Y' \setminus ext(Y)$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus ext(Y)$.
+    Pushout is then obtained by identifying the nodes with the pre-image in $Y_{ext}$ which are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
 $g';i_2$ yields $h$ and therefore $h$ is mono.
     Likewise, $k$ is mono.
-    \end{proof}
-    \update{Is this proof ok?}
-\end{proposition}
-
-\begin{proposition}
-    $+$ of two mda-cospans
-
-        \[
-    ext(X) \xrightarrow{} X \xrightarrow{} A \xleftarrow{} Y \xleftarrow{} ext(Y)
-    \]
-    \[
-    ext(X) \xrightarrow{} X' \xrightarrow{} B \xleftarrow{} Y' \xleftarrow{} ext(Y)
-    \]
-
-    is an mda-cospan.
-
-    \begin{proof}
-        By construction in Figure~\ref{fig:A+B}.
-    \end{proof}
-\end{proposition}
-
-\begin{definition}[Category $\catname{MAECsp_{D}(EHyp_{\Sigma})}$]
-
-    Taking into account all the propositions above gives us a subcategory of $\catname{ECsp_{D}(EHyp_{\Sigma})}$ where each morphism is an mda-cospan. This category is also symmetric monoidal.
+
+    This category is also symmetric monoidal and has $+$ by the lemma above.
 \end{definition}
@@
 
-\begin{proposition}
-    Tensor unit
-
-    \[
-    0 \xrightarrow{id_0} 0 \xrightarrow{id_0} 0 \xleftarrow{id_0} 0 \xleftarrow{id_0} 0
-    \]
-
-    is well-typed.
-    \begin{proof}
-        The empty hypergraph does not contain any hierarchical edges and hence the cospan above is well-typed.
-    \end{proof}
-\end{proposition}
-
 \begin{proposition}[The tensor product of two mda-cospans is again an mda-cospan]
@@
         \[
-    ext(X) \xrightarrow{} X \xrightarrow{} A \xleftarrow{} Y \xleftarrow{} ext(Y)
-    \]
-    \[
-    ext(X) \xrightarrow{} X' \xrightarrow{} B \xleftarrow{} Y' \xleftarrow{} ext(Y)
+    X_{ext} \xrightarrow{} X \xrightarrow{} A \xleftarrow{} Y \xleftarrow{} Y_{ext}
+    \]
+    \[
+    X_{ext} \xrightarrow{} X' \xrightarrow{} B \xleftarrow{} Y' \xleftarrow{} Y_{ext}
     \]
@@
 \mathcal{I} : \catname{MACsp_{D}(Hyp_\Sigma)}
-\xhookrightarrow{} \WellTypedMdaEcospans
+\xrightarrow{} \WellTypedMdaEcospans
 \]\label{functor:I}
 
-For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\catname{MACsp_{D}(Hyp_{\Sigma})}$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\# = \varnothing$, and $<_{c} = \varnothing$.
+For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\MdaCospans$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\# = \varnothing$, and $<_{c} = \varnothing$.
 It is well-typed on the nose as it has no hierarchical edges.
 
-% \update{Two cospans $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$, $p \xrightarrow{} \mathcal{G} \xleftarrow{} q$ in $\catname{MACsp_{D}(Hyp_{\Sigma})}$ are composable when $out^{-1}_{out}(\mathcal{G})$ and $in^{-1}_{out}(\mathcal{H})$ are isomorphic.
+% \update{Two cospans $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$, $p \xrightarrow{} \mathcal{G} \xleftarrow{} q$ in $\MdaCospans$ are composable when $out^{-1}_{out}(\mathcal{G})$ and $in^{-1}_{out}(\mathcal{H})$ are isomorphic.
 % The result of such composition is depicted below
@@
 %     We will build this functor inductively.
-As noted above, $\catname{MACsp_{D}(Hyp_{\Sigma})}$ is a full subcategory of $\catname{MACsp_{D}(EHyp_{\Sigma})}$ and the acting of this functor on generators can be found in~\cite{Frobenius}.
+As noted above, $\MdaCospans$ is a full subcategory of $\catname{MACsp_{D}(EHyp_{\Sigma})}$ and the acting of this functor on generators can be found in~\cite{Frobenius}.
 We need to define what $+$ from $\text{PROP}^{+}$ in $\catname{MACsp_{D}(EHyp_{\Sigma})}$ is.
@@
 However, this notion of DPO does not account for hypergraphs with interfaces, i.e.
-those from $\catname{MACsp_{D}(Hyp_{\Sigma})}$ where $l,r$ from $\langle l, r \rangle$ are cospans.
+those from $\MdaCospans$ where $l,r$ from $\langle l, r \rangle$ are cospans.
 To support this we first assume that we only need to maintain the interfaces of $i \xrightarrow{} L \xleftarrow{} j$ and $i \xrightarrow{} R \xleftarrow{} j$ during rewriting, then we can take $K \cong i + j$, i.e.
@@
 
-    where $\llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \catname{MACsp_{D}(Hyp_{\Sigma})}$ is a faithful monoidal functor.
+    where $\llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \MdaCospans$ is a faithful monoidal functor.
 \end{theorem}
@@
 
-\question{Shall I replace $i,j$ etc. with $ext(X)$?}
+\question{Shall I replace $i,j$ etc. with $X_{ext}$?}
 
@@
 The original DPOI square is shown in red. The extra bits simply account for more complicated interfaces.
-For example, the left-hand side $L$ of a rewrite rule $\langle L, R \rangle$ which originally was some cospan $i \xrightarrow{} L \xleftarrow{} j$ in $\catname{MACsp_{D}(Hyp_{\Sigma})}$ now is a cospan $i \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} j$ from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
+For example, the left-hand side $L$ of a rewrite rule $\langle L, R \rangle$ which originally was some cospan $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ now is a cospan $i \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} j$ from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
 The top half of this diagram says that the outermost interfaces should coincide for $L$ and $R$. The bottom half puts the same restrictions on $G$ and $H$. 
@@
 
-    \begin{figure}
-        \centering
-\[
-  \scalebox{0.3}{      \tikzfig{combinatorial_semantics/refined_dpoi_example}
-}
-        \]
-    \caption{Refined DPOI rewriting}
-    \label{fig:refined_dpoi}
-    \end{figure}
+%     \begin{figure}
+%         \centering
+% \[
+%   \scalebox{0.3}{      \tikzfig{combinatorial_semantics/refined_dpoi_example}
+% }
+%         \]
+%     \caption{Refined DPOI rewriting}
+%     \label{fig:refined_dpoi}
+%     \end{figure}
 
 \end{example}
+
+\Aleksei{This figure yeilds "Dimension too large". Need to fix it.}
 

### 2023-10-17 sections/introduction.tex
Fixed figures, brushed up proofs, bit restructured

--- 2023-09-29 sections/introduction.tex
@@
+% !TEX root = ../main.tex
+
 \newcommand\mylet[2]{\textsf{let } #1 = #2 \textsf{ in }}
@@
 
+\update{Move focus from semantics of e-graphs to their extension to monoidal theories?}
 While equality saturation is a state-of-the-art optimization technique, the formal mathematical theory of e-graphs is perhaps underdeveloped.

### 2023-10-17 sections/soundness_and_completeness.tex
Fixesing inaccuracies for combinatorial semantics
Fixed figures, brushed up proofs, bit restructured

@@
+% !TEX root = ../main.tex
+
 \section{Soundness and Completeness}
@@
 In this section, we will build a correspondence between terms in a semilattice enriched $\textsf{PROP}^{+}(\Sigma)$ and morphisms in $\WellTypedMdaEcospans$.
-Similarly to $\textsf{PROP}^{+}(\Sigma)$ where terms are defined up to semilattice enrichment and SMC equations we will say that two terms $G$ and $F$ in $\WellTypedMdaEcospans$ are equal when $G \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} F$.
-As SMC equations are absorbed by e-hypergraph representation we only quotient by $\llbracket \mathcal{E} \rrbracket$.
+Similarly to $\textsf{PROP}^{+}(\Sigma)$ where terms are defined up to semilattice enrichment and SMC equations we will say that two morphisms $G$ and $F$ in $\WellTypedMdaEcospans$ are equal when $G \Rrightarrow{}_{\mathcal{S}} F$.
+As SMC equations are absorbed by e-hypergraph representation we only quotient by $\mathcal{S}$. 
 We are now ready to build a semilattice-enriched SMC functor
@@
 \[
-\llbracket - \rrbracket : \textsf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans
+[\cdot] : \textsf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans
 \]
@@
 by freeness, it suffices to define how the functor acts on generators.
-As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \catname{MACsp_{D}(Hyp_{\Sigma})}$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}.
+As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \MdaCospans$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}.
 Everything else follows by defining
@@
 \begin{enumerate}
-    \item $\llbracket f;g \rrbracket = \llbracket f \rrbracket ; \llbracket g \rrbracket$  for $f,g \in \Sigma$;
-    \item $\llbracket f \otimes g \rrbracket = \llbracket f \rrbracket \otimes \llbracket g \rrbracket$  for $f,g \in \Sigma$;
-    \item $\llbracket f + g \rrbracket = \llbracket f \rrbracket + \llbracket g \rrbracket$  for $f,g \in \Sigma$;
+    \item $[ f;g ] = [ f ] ; [ g ]$  for $f,g \in \Sigma$;
+    \item $[ f \otimes g ] = [ f ] \otimes [ g ]$  for $f,g \in \Sigma$;
+    \item $[ f + g ] = [ f ] + [ g ]$  for $f,g \in \Sigma$;
 \end{enumerate}
@@
 this also makes the functor into a semilattice-enriched SMC one on the nose.
+% \update{This essentially means that $[\cdot] = \llbracket \cdot \rrbracket ; \mathcal{I}$ plus three equations above.}
 
@@
 
+\begin{theorem}
+    If $f = g$ in $PROP^{+}(\Sigma)$ then $[f] \Rrightarrow_{\mathcal{S}} [g]$ in $\WellTypedMdaEcospans$.
+    \begin{proof}
+        \update{
+            By induction on the number of applied equations to turn $f$ into $g$?
+            E.g. $f = g$ in one step essentially means that $f = l$ for some $l$ in~\ref{fig:string-equations} and $g = r$ (or the other way around).
+            Then we need to show that $[f] \Rrightarrow_{\langle l, r \rangle \in \mathcal{S}} [g']$ and $[g'] = [g]$.
+            }
+    \end{proof}
+\end{theorem}
+
+\begin{corollary}
+    $[\cdot] : PROP^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is sound in a sense that
+    if $f = g$ in $PROP^{+}(\Sigma)$ then $[f] = [g]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$
+\end{corollary}
+
+\begin{proposition}
+    Functor $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful.
+    \begin{proof}
+        Consider the commutative diagram below.
+
+        % https://q.uiver.app/#q=WzAsNCxbMCwwLCJcXHRleHR7UFJPUH0oXFxTaWdtYSkiXSxbMiwwLCJcXHRleHR7UFJPUH1eeyt9KFxcU2lnbWEpIl0sWzAsMiwiXFxNZGFDb3NwYW5zIl0sWzIsMiwiXFxzZnJhY3tcXFdlbGxUeXBlZE1kYUVjb3NwYW5zfXtcXG1hdGhjYWx7U319Il0sWzAsMSwiXFxtYXRoY2Fse0p9Il0sWzAsMiwiXFxsbGJyYWNrZXQgXFxjZG90IFxccnJicmFja2V0IiwyXSxbMSwzLCJbXFxjZG90XSJdLFsyLDMsIlxcbWF0aGNhbHtJfSIsMl1d
+        \[
+        \begin{tikzcd}
+            {\text{PROP}(\Sigma)} && {\text{PROP}^{+}(\Sigma)} \\
+            \\
+            \MdaCospans && {\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}}
+            \arrow["{\mathcal{J}}", from=1-1, to=1-3]
+            \arrow["{\llbracket \cdot \rrbracket}"', from=1-1, to=3-1]
+            \arrow["{[\cdot]}", from=1-3, to=3-3]
+            \arrow["{\mathcal{I}}"', from=3-1, to=3-3]
+        \end{tikzcd}
+        \]
+    \end{proof}
+
+    $\llbracket \cdot \rrbracket$ is faithfull by Corollary 4.3 in~\cite{Frobenius}.
+    $\mathcal{J}$ is an inclusion that maps terms in $\textsf{PROP}(\Sigma)$ to SMC fragment of $\textsf{PROP}^{+}(\Sigma)$ and is automatically faithful. 
+    Similarly, $\mathcal{I}$ is faithful because it is inclusion.
+    Then, by composition, $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful. 
+\end{proposition}
+
 \begin{theorem}[$\llbracket - \rrbracket$ is faithful]
+\begin{proof}
+Essentially we need to show that if $\llbracket f \rrbracket = \llbracket g \rrbracket$ then $f = g$ which is equivalent to if $f \not = g$ then $\llbracket f \rrbracket \not = \llbracket g \rrbracket$.
+First, we note that we have a \question{weak (modulo commutativity)} normal form for any term in $PROP^{+}(\Sigma)$ which is achieved by applying distributivity for tensor and composition as well as unpacking equations, that is, every term is representable as in Figure~\ref{fig:normal_form_string} where $\phi_1,\;\ldots\;,\phi_n$ are pure SMC terms.
+\update{It is not actaully a `normal' form because we can still apply SMC equations, so it should be normal modulo SMC and commutativity}.
 
-Essentially we need to show that if $\llbracket f \rrbracket = \llbracket g \rrbracket$ then $f = g$ which is equivalent to if $\llbracket f \rrbracket \not = \llbracket g \rrbracket$ then $f \not = g$.
-First, we note that we have a \question{weak (modulo commutativity)} normal form for any cospan in $\WellTypedMdaEcospans$ which is achieved by merging all hierarchical edges into a single hierarchical edge making it top-level, that is, every cospan is representable as below
+\begin{figure}
+    \[
+    \scalebox{0.75}{
+        \tikzfig{soundness_and_completeness/normal_form_string}
+    }    
+    \]
+    \caption{Weak normal form of a term in $PROP^{+}(\Sigma)$}
+    \label{fig:normal_form_string}
+\end{figure}
+
+Then we get a corresponding cospan for such weak normal form in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ as in Figure~\ref{fig:normal_form_ehyp}
+
+\begin{figure}
 
@@
 \]
+\caption{Interpretation of the weak normal form in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$}
+\label{fig:normal_form_ehyp}
+\end{figure}
 
-where all $A_1\; \ldots \;A_n$ do not contain any hierarchical edges. We will write such form as $A_1 + \ldots A_n$ Then, $\llbracket f \rrbracket \not = \llbracket g \rrbracket$ leads to two options
+where all $A_1\; \ldots \;A_n$ do not contain any hierarchical edges.
+We will write such forms as $\phi_1 + \ldots \phi_n$ and $A_1 + \ldots A_n$ respectively.
+Then, $\llbracket f \rrbracket \not = \llbracket g \rrbracket$ leads to two options
 
-\begin{itemize}
-    \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_m$ where $n \not = m$.
+% \begin{itemize}
+%     \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_m$ where $n \not = m$.
 By functoriality and weak normal form in $\textsf{PROP}^{+}(\Sigma)$, $f = a_1 + \ldots + a_n$, $g = b_1 + \ldots b_m$ and $\llbracket f \rrbracket = \llbracket a_1 \rrbracket + \ldots + \llbracket a_n \rrbracket$, $\llbracket g \rrbracket = \llbracket b_1 \rrbracket + \ldots \llbracket b_m \rrbracket$.
 This essentially means that $f \not = g$ as it has different number of summands.
-    \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_n$.
+%     \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_n$.
 Then there exists a permutation $\sigma$ and an index $i$ such that $ A_i  \not =  B_{\sigma(i)}$.
@@
 \update{Prove that $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful?}
+% \end{itemize}
+
+Now, assume that $f = \phi_1 + \ldots + \phi_n$ and $g = \psi_1 + \ldots + \psi_m$.
+Then, by functoriality, $[f] = [A_1] + \ldots + [A_n]$ and $g = [B_1] + \ldots + [B_m]$. 
+$f \not g$ gives us two options.
+
+\begin{itemize}
+    \item $n \not = m$. Then $[f] \not = [g]$ as they are simply not isomorphic cospans of e-hypegraphs as they have different number of summands.
+    \item $n = m$.
+Then there exists a permutation $\sigma$ and index $i$ such that $\phi_i \not = \psi_{\sigma(i)}$.
+Remember that these are just SMC terms and by faithfulness of $\llbracket \cdot \rrbracket;\mathcal{I}$ we get $A_i \not = B_{\sigma(i)}$ and $[f] \not = [g]$.
 \end{itemize}
 
+\end{proof}
 \end{theorem}
+
+\question{Is proof ok?}
+
+\begin{theorem}
+
+    Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$.
+    $f \Rightarrow{}_{\mathcal{R}^{+}}$ iff $[f] \Rrightarrow{}_{[\mathcal{R}^{+}]} [g]$.
+    \begin{proof}
+        \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.
+            But we only have this result for $\llangle \cdot \rrangle$ for terms with bent wires which is a functor to $PROP + Frob$.
+        }
+    \end{proof}
+\end{theorem}

### 2023-10-18 sections/introduction.tex
update intro

--- 2023-10-09 sections/introduction.tex
@@
 Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal sums.
-Thus we generalize e-graphs from algebraic to monoidal theories.
+Thus we generalize e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications.
 
@@
 That is, we exhibit a translation functor $\llbracket-\rrbracket$ from the free semilattice enriched SMC over a symmetric monoidal theory to our (quotiented) category of e-hypergraphs such that
-morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow \llbracket_{DPO} g \rrbracket$.
+morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow_{DPO} \llbracket g \rrbracket$.
 
@@
 Thus, we define a notion of e-matching for e-hypergraphs: that is, to find redexes working modulo the e-structure.
-In particular, we wish to locate the smallest subgraph $G' \subset G$ "containing" the redex $L$.
+In particular, we wish to locate the smallest subgraph $G' \subset G$ "containing" (in an appropriate sense) the redex $L$.
 Given this, the analogue of e-rewriting $L \to R$ is simple to define, due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
 
-[Example diagrams]
+\begin{figure}
+\[
+	\tikzfig{combinatorial_semantics/example-rewriting}
+\]
+\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
+REDO THIS}
+\end{figure}
 
-\subsection*{Future Applications}
+\subsection*{Future Work and Applications}
 
-Example diagrams for circuits, trace (loop unrolling, PEGs), LC, ZX here.
+We have already alluded to the fact we deal with only \textit{acyclic} e-graphs in this paper.
+Nevertheless,  within this setting we present a generalization from Cartesian to (symmetric) monoidal theories,  giving rise to a host of potential new domains of application.
+Further,  the string diagrammatic and graph rewriting framework we set up is ameenable to extension, including lifting the acyclicity restriction.
+
+Symmetric monoidal theories have received significant attention over recent years: the ZX-calculus can be used to characterize quantum processes, with applications to optimization of quantum circuits [cite]; similar theories also characterize signal flow diagrams [cite], digital circuits [cite], as well as having application to computational linguistics [cite].
+In future work, we hope to apply our generalization of e-graphs to optimization problems in these domains (e.g.
+Figure \ref{})
+
+The theories described above often have more structure than simply being symmetric monoidal: for example, they may contain comonoids, Frobenius algebras or traced structure (as in the case of e-graphs with cycles),  and this structure can be exploited to give more natural (in the sense of factoring out more equations) combinatorial representations.
+In the above cases, such representations and their DPO-rewriting theory is well-studied [cite, cite, cite].
+We consider that our development of hierarchical "e-box" structure within string diagrams and their associated (e-)hypergraphs ( categorically,  semilattice enrichment) is essentially orthogonal to these developments, and so ought to be easily combined to give notions of e-hypergraphs specifically suited to the optimization of these particular monoidal theories.
+
+A particularly interesting extension that we intend to consider is the combination of our "e-structure" with (Cartesian) \textit{closed} categories, allowing the natural combination of e-graph optimization techniques with the $\lambda$-calculus.
+The corresponding combinatorial representation would build on the recent work [reverse diff,FSCD].
+
+Example diagrams for circuits (loop unrolling, PEGs)
+
+\begin{figure}
+\[
+	\scalebox{0.75}{
+	\tikzfig{combinatorial_semantics/example-e-graphs}
+	}
+\]
+\caption{Example applications of e-hypergraphs.  Left, the ZX-calculus.  Right, the $\lambda$-calculus as represented by string diagrams for Cartesian closed categories.  }
+\end{figure}
 

### 2023-10-18 sections/combinatorial_semantics.tex

@@
 Similarly, in the case when $\psi \subset V \times V$ is a relation: we denote with $\psi^{*}: \subset V^* \times V^*$ the element-wise relation on ordered sequences.
-We will denote with $A +_{f,g} B$ the pushout of $A \xleftarrow{f} C \xrightarrow{g} B$.
+We will denote with $A +_{f,g} B$ the pushout of $A \xleftarrow{f} C \xrightarrow{g} B$ and will use $i_j$ to denote the $j^{\text{th}}$ injection into a coproduct of the form $X_{j} \xrightarrow{i_{j}} X_{1} + \ldots X_{j} + \ldots X_{n}$ where $+$ denotes the coproduct.
 
@@
 We can not have $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 < x_2$, i.e.
-the case where $x_1$ is not the immediate predecessor of $x_2$ because then it would mean that there exists at least $x_3$ such that $x_1 <_{\mu} x_2 <_{\mu} x_3$ and because $\phi$ should preserve them it would mean that $\phi(x_1) <_{\mu} \phi(x_3) <_{\mu} \phi(x_2)$ implying that $\phi(x_2)$ has two immediate predecessors: $\phi(x_1)$ and $\phi(x_3)$.
+the case where $x_1$ is not the immediate predecessor of $x_2$ because then it would mean that there exists at least $x_3$ such that $x_1 <_{\mu} x_3 <_{\mu} x_2$ and because $\phi$ should preserve them it would mean that $\phi(x_1) <_{\mu} \phi(x_3) <_{\mu} \phi(x_2)$ implying that $\phi(x_2)$ has two immediate predecessors: $\phi(x_1)$ and $\phi(x_3)$.
 Of course, if $\phi$ does not identify $x_1$ and $x_3$.
-
-The first bullet is fine because it means that $\phi$ reflects the immediate predecessor. The second bullet leaves us with two options depicted below.
+The first bullet is fine because it means that $\phi$ reflects the immediate predecessor.
+The second bullet leaves us with two options depicted above.
 
@@
     \]
-    where $A \in obj(\catname{EHyp_{\Sigma}})$, $W, X, Y, Z$ are discrete and $f,g,h,k$ are monos.
+
+    where
+    \begin{enumerate}
+        \item $A \in \text{obj}(\catname{EHyp_{\Sigma}})$
+        \item $W, X, Y, Z$ are discrete and $f,g,h,k$ are monos
+        \item \update{The image of $f;g$ (respectively, $k;h$) constitutes of nodes $v$ such that $[v) = \varnothing$}
+    \end{enumerate}
     Since $f$ and $k$ are monos, $W$ and $Z$ are essentially subgraphs of $X$ and $Y$ and hence we will introduce the following notation which better depicts that such cospans are effectively e-hypergraphs with \textit{external} interfaces
@@
 % https://q.uiver.app/#q=WzAsOCxbMCwxLCJleHQoWCkiXSxbMSwwLCJYIl0sWzIsMCwiQSJdLFszLDAsIn5ZIl0sWzQsMSwiZXh0KFkpIl0sWzEsMiwiWCciXSxbMiwyLCJBJyJdLFszLDIsIlknIl0sWzAsMSwiIiwyLHsiY3VydmUiOi0yfV0sWzEsMl0sWzMsMl0sWzQsMywiIiwwLHsiY3VydmUiOjJ9XSxbNCw3LCIiLDIseyJjdXJ2ZSI6LTJ9XSxbNyw2XSxbNSw2XSxbMCw1LCIiLDAseyJjdXJ2ZSI6Mn1dLFsxLDUsIlxcYWxwaGEiLDFdLFsyLDYsIlxcYmV0YSIsMV0sWzMsNywiXFxnYW1tYSIsMV1d
-
-\Aleksei{Redraw this diagram in tikzit because importing 'quiver' introduces error}
 
@@
 X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
-\]
+\]\label{morphism:identity}
 
@@
 since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
+$h = a;p_1 + g|_{Y' \setminus Y_{ext}};p_2$, $k = b;p_2 + f|_{Y \setminus Y_{ext}};p_1$.
+\question{Is this a standard syntax for a narrowed function?}
 
@@
 
-Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into a SMC.
+Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC.
 More precisely, below are cospans for tensor product and symmetry. Tensor unit is given by the unit in $\catname{EHyp_{\Sigma}}$.
@@
 }
-\]
+\]\label{morphism:tensoring}
 
@@
 }
-\]
+\]\label{morphism:symmetry}
 
@@
 \begin{remark}
-    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y$.
+    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y_{ext}$.
     Consider an example in Figure~\ref{fig:e-hyp_composition_example} below.
@@
     The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
-    \update{We should probably require that the image of $f_{ext};f{}$ has no predecessors and has no incoming edges}
     Note that pushout for such composition always exists as the nodes that are being identified are unrelated: they are not in conflict with anything and the sets of their predecessors are empty.
@@
 
-\Aleksei{what is $h$? $k$? $f_ext: ext(X) -> X$, so how is $f_ext$ also of type $ext(X) -> ext(X) + X + X'$ ? }
 \begin{definition}
@@
     \[
-    X_{ext} \xrightarrow{f_{ext}} X_{ext} + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} Y_{ext} + Y + Y' \xleftarrow{g_{ext}} Y_{ext}
+    X_{ext} \xrightarrow{i_{1}} X_{ext} + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} Y_{ext} + Y + Y' \xleftarrow{i_1} Y_{ext}
     \]
@@
     $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
-    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v)$.
+    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v) = \varnothing$.
     And such that its sources and targets are top-level nodes.
     $h$ and $k$ are morphisms that map $X_{ext}$ and $Y_{ext}$ to these sources and targets respectively.
+In particular, in Figure~\ref{fig:A+B}, $h = \{c_1 \mapsto c_1, \ldots, c_n \mapsto c_n\}$ and $k = \{d_1 \mapsto d_1, \ldots, d_l \mapsto d_l\}$.
     Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
 \end{definition}
+
+\begin{remark}
+    We can generalise the above definition to a plus of $n$ cospans in an obvious way.
+\end{remark}
 
@@
         \item Nodes with the in(out)-degree of 0 are exactly the nodes in the image of $f$ (respectively, $g$).
-        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ ($g$) which pre-image is not in $X_{ext}$ ($Y_{ext}$). 
-        \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
+        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ (respectively, $g$) which pre-image is not in $X_{ext}$ (respectively, $Y_{ext}$). 
+        % the item below is not assumed by the definition of a cospan of e-hypergraphs
+        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
 \end{enumerate}
@@
 
-\begin{lemma}
-    \begin{enumerate}
-        \item The tensor product of two mda-cospans is an mda-cospan.
-        \item $+$ of two mda-cospans is an mda-cospan.
-    \end{enumerate}
-    \begin{proof}
-        Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
-        Hence, everything is preserved component-wise by copairing and coproduct.
-    \end{proof}
-    \begin{proof}
-        By construction in Figure~\ref{fig:A+B}.
-    \end{proof}
-\end{lemma}
-
-% \begin{proposition}
-%     Id morphism
-
-%     \[
-%     X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
-%     \]
-
-%     is an mda-cospan.
+% \begin{lemma}
+%     \begin{enumerate}
+%         \item The tensor product of two mda-cospans is an mda-cospan.
+%         \item $+$ of two mda-cospans is an mda-cospan.
+%     \end{enumerate}
 %     \begin{proof}
-%         Proof by checking all the constraints above.
-Discrete e-hypergraphs is obviously directed and acyclic.
-Id-morphisms are monos.
-The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$.
-Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
+%         Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
+%         Hence, everything is preserved component-wise by copairing and coproduct.
 %     \end{proof}
-% \end{proposition}
-
-\update{Anything about symmetry?}
+%     \begin{proof}
+%         By construction in Figure~\ref{fig:A+B}.
+%     \end{proof}
+% \end{lemma}
 
@@
     Narrowing the morphisms of $\Ecospans$ to mda-cospans gives us a subcategory $\MdaEcospans$.
-    Id morphims is given by the following cospan
-
-    \[
-    X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
-    \]
-
-    which is an mda-cospan by checking all the necessary constraints in~\ref{def:monogamy_ehyp}.
+
+    This category is also symmetric monoidal and has $+$ by the lemma above.
+\end{definition}
+
+\begin{proposition}
+    $\MdaEcospans$ is an SMC.
+\begin{proof}
+To show the statement above we need to first show that $\MdaEcospans$ is a proper category and then show that it has a symmetry, tensor product and the tensor unit.
+In particular we need to show that the cospans that correspond to particular morphisms are mda-ones according to~\ref{def:monogamy_ehyp}.
+\begin{enumerate}
+    \item Id cospan is an mda-cospan. Recall the id cospan in~\ref{morphism:identity}.
     Discrete e-hypergraph $X$ is obviously directed and acyclic.
@@
     Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
-
-    Composition is given in~\ref{morphism:composition_ecospans} and composing two mda-cospans yields an mda-cospan.
+    \item Composition~\ref{morphism:composition_ecospans} of two mda-cospans yields an mda-cospan.
     Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
-    This also does not increase their degree to more than 1.
-    $f_{ext}$ is mono because its image $X$ is unchanged.
-    The image of $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
-    Pushout is then obtained by identifying the nodes with the pre-image in $Y_{ext}$ which are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
-$g';i_2$ yields $h$ and therefore $h$ is mono.
+    This also does not increase their degrees to more than 1.
+    $f_{ext}$ (respectively, $g_{ext}$) is mono because its image $X$ (respectively, $Z$) is unchanged.
+    The morphism $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
+    Also note that $p_2$ narrowed to $B \setminus Y_{ext}$ is exactly $i_2$ and similarly for $p_1$ and $i_1$.
+    Pushout is then obtained by identifying the nodes in $A + B$ that share their pre-image in $Y_{ext}$.
+    These nodes are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
+    $g';i_2$ is equal to $g';p_2$ because the image of $g'$ does not contain any nodes with pre-image in $Y_{ext}$.
+    Next, $a$ is mono and $a;p_1$ is mono, because $p_1$ only identifies nodes with pre-image in $Y_{ext}$.
+    $h = a;p_1 + g';p_2$ and hence $h$ is mono.
     Likewise, $k$ is mono.
-
-    This category is also symmetric monoidal and has $+$ by the lemma above.
-\end{definition}
-
-% show that EHyp is symmetric monoidal
+    \item Tensor product~\ref{morphism:tensoring} of two mda-cospans is an mda-cospan.
+          Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
+          Hence, everything is preserved component-wise by copairing and coproduct.
+    \item Symmetry~\ref{morphism:symmetry} is an mda-cospan.
+          $X$ and $Y$ are both discrete as well as their coproduct and $\sigma_{X,Y}$ is obviously mono.
+          Furthermore, all nodes of $X + Y$ are in the image of $\sigma_{X,Y}$ and $id_{X} + id_{Y}$.
+    \item The tensor unit is discrete because it is an empty e-hypergraph and hence does not break monogamous property.
+\end{enumerate}
+
+The bullets above makes $\MdaEcospans$ into a proper SMC.
+\end{proof}    
+\end{proposition}
+
+\question{Proofs ok?}
+
+\begin{proposition}
+    $+$ of two mda-cospans is an mda-cospan.
+    \begin{proof}
+        By construction in Figure~\ref{fig:A+B}.
+    \end{proof}
+\end{proposition}
 
@@
 
-\begin{definition}[Well-typed edge]
+\begin{definition}[Well-typed cospan]
 
@@
     We call $e$ well-typed if for all $v \in V_{e,in}^{\bar{\#}}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\bar{\#}}$ $|v| = |t(e)|$.
-We will call an mda-cospan well-typed if all hierarchical edges in its carrier are well-typed.
+    We will call an mda-cospan in $\MdaEcospans$ well-typed if all hierarchical edges in its carrier are well-typed.
 \end{definition}
@@
 \begin{proposition}
-    Id-cospan
-
-    \[
-    X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
-    \]
-
-    is well-typed.
+    $\WellTypedMdaEcospans$ is a proper SMC.
 \begin{proof}
-        By definition, we only include well-typed morphisms.
-    \end{proof}
-\end{proposition}
-
-\begin{proposition}[The tensor product of two mda-cospans is again an mda-cospan]
-    The tensor product does not introduce any new hierarchical edges and images of the interfaces are
-preserved by copairing.
-\end{proposition}
-
-\begin{proposition}[Composition of two well-typed mda-cospans is a well-typed mda-cospan]
+    As before, we need to show the well-typedness of the key ingredients of an SMC.
+    \begin{enumerate}
+        \item Id-cospan~\ref{morphism:identity} is well-typed.
+              Discrete e-hypergraphs do not have any hierarchical edges and hence are well-typed.
+        \item Composition~\ref{morphism:composition_ecospans} of two well-typed mda-cospans is a well-typed mda-cospan.
               Composition does not introduce any new hierarchical edges.
               Furthermore, as composition identifies only nodes in the image of the outermost interfaces, input and
-output nodes of hierarchical edges are not affected.
+              output nodes of hierarchical edges are not affected and the number of sources and targets of an edge remains unchanged as well.
+        \item The tensor product of two mda-cospans is again an mda-cospan.
+              The tensor product does not introduce any new hierarchical edges and does not affect the existing hierarchical edges.
+        \item The tensor unit is descrete and hence tensoring with it does not break well-typedness.
+    \end{enumerate}
+    All the items above show that $\WellTypedMdaEcospans$ is indeed an SMC.
+\end{proof}
 \end{proposition}
@@
 
-The propositions above make $\WellTypedMdaEcospans$ into a proper symmetric monoidal category. There is an obvious inclusion
+There is an obvious inclusion
 
@@
 pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities.
-Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as follows, for some rule $\langle l,r \rangle \in \mathcal{R}$
+Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as in Figure~\ref{fig:prop_rewriting}, for some rule $\langle l,r \rangle \in \mathcal{R}$
 \end{definition}
@@
     \]
+    \caption{Rewriting in $\textsf{PROP}(\Sigma)$}
+    \label{fig:prop_rewriting}
 \end{figure}
@@
 
-If they are decomposable as above then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
-Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as below.
+If they are decomposable as in Figure~\ref{fig:prop_rewriting} then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
+Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above ($\Rightarrow_{\mathcal{R}})$)) and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as in Figure~\ref{fig:enriched_prop_rewriting}.
 
@@
     \]
+    \caption{Rewriting in $\textsf{PROP}^{+}(\Sigma)$}
+    \label{fig:enriched_prop_rewriting}
 \end{figure}
@@
 
+\textbf{Update: 4 Oct}
+
 \begin{theorem}[Theorem 35~\cite{Frobenius2}]
+\label{theorem:rewriting_no_frob}
     Let $R$ be any rewriting system on $\textsf{PROP}(\Sigma)$. Then,
     \[
-    g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llbracket ^{\ulcorner}g^{\urcorner} \rrbracket \Rrightarrow_{\llbracket \mathcal{R} \rrbracket} \llbracket ^{\ulcorner}h^{\urcorner} \rrbracket 
-    \]
-
-    where $\llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \MdaCospans$ is a faithful monoidal functor.
+    g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llangle ^{\ulcorner}g^{\urcorner} \rrangle \Rrightarrow_{\llangle \mathcal{R} \rrangle} \llangle ^{\ulcorner}h^{\urcorner} \rrangle 
+    \]
+
+    where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + Frob \to \catname{Csp_{D}(Hyp_{\Sigma})}$ is an isomorphism of props.
+    \update{
+        Last time I confused functors $\llangle \cdot \rrangle$ and $\llbracket \cdot \rrbracket$.
+        Note that the result above is a weaker version of the same result for rewriting in $\textsf{PROP}(\Sigma) + Frob$.
+        $Frob$ is an SMT with the signature $\Sigma_{Frob}$ which constitutes of \textit{only} Frobenius generators and terms are quotiented by both SMC laws and Frobenius equations.
+        So, essentially, $Frob = \textsf{PROP}(\Sigma_{Frob}, \mathcal{E}_{Frob})$. 
+        \question{Is it true that by quotienting $\textsf{PROP}(\Sigma)$ by SMC laws and quotienting $Frob$ by SMC and Frob laws we automatically quotient $\textsf{PROP}(\Sigma) + Frob$ by SMC and Frob laws?}
+    }
 \end{theorem}
 
+\begin{definition}[$\ulcorner \cdot \urcorner$]
+
+    This is a syntactic operation which is defined for terms in $\textsf{PROP}(\Sigma) + Frob$ and is given below
+    \[
+        \tikzfig{combinatorial_semantics/bending_wires}
+    \]
+
+\end{definition}
+
 \begin{remark}
-    Note that even though $n \xrightarrow{} G \xleftarrow{} m$ and $0 \xrightarrow{} G \xleftarrow{} n + m$ are essentially isomorphic \update{(they are not technically isomorphic because they have different legs)}, the functor $\llbracket \cdot \rrbracket$ deals with arbitrary string diagrams (that have multiple inputs and multiple outputs) in $\textsf{PROP}(\Sigma)$ and interprets them as cospans $n \xrightarrow{} G \xleftarrow{} m$.
-However, there is a syntactic operation $\ulcorner \cdot \urcorner$ on $\textsf{PROP}(\Sigma)$ to translate between string diagrams with multiple inputs and outputs and string diagrams with only outputs.
-
-    \[
-\tikzfig{combinatorial_semantics/bending_wires}
-    \]
-
+    We need the operation above because we defined the rewriting for cospans of hypergraphs only for cospans of the form $0 \to G \xleftarrow{} n + m$ (for $\catname{Csp_{D}(Hyp_{\Sigma})},\; \MdaCospans,\; \MdaEcospans$), but the functors take arbitrary terms (string diagrams) as input.
+    This operation turns an arbitrary term into a term with no inputs.
     Now the left-hand and right-hand sides are equivalent in the sense that $f \Rightarrow_{\langle l, r \rangle} g$ if and only if $\ulcorner f \urcorner \Rrightarrow_{\langle \ulcorner l \urcorner, \ulcorner r \urcorner \rangle}
  \ulcorner g \urcorner$ and the latter string diagram can be interpreted as $0 \xrightarrow{} G \xleftarrow{} n + m$ for which we defined rewriting above.
+
 \end{remark}
 
+\begin{proposition}
+
+    $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$ is mda-cospan in $\MdaCospans$ iff $0 \xrightarrow{} \mathcal{G} \xleftarrow{} n + m$ is an mda-cospan.
+\end{proposition}
+
+In particular, the above proposition means that by fullness of $\llbracket \cdot \rrbracket$ there exists a corresponding SMC morphism for some $0 \xrightarrow{} \mathcal{G}\xleftarrow{} n + m$.
+Theorem above essentially makes two routes below. In forward direction:
+
+\[
+\textsf{PROP}(\Sigma) \xrightarrow{i_1} \textsf{PROP}(\Sigma) + Frob \xrightarrow{\ulcorner \cdot \urcorner} \textsf{PROP}(\Sigma) + Frob \xrightarrow{\llangle \cdot \rrangle} \catname{Csp_{D}(Hyp_{\Sigma})}
+\]
+
+The proof of the theorem above in the forward direction then relies on the proof for rewriting in $\textsf{PROP}(\Sigma) + Frob$ and functoriality of $\llangle \cdot \rrangle$ and checking that a match is convex.
+
+In backwards direction
+
+\[
+  \catname{Csp_{D}(Hyp_{\Sigma})} \xrightarrow{\llbracket \cdot \rrbracket}^{-1} \textsf{PROP}(\Sigma)  
+\]
+
+relies on decomposition of a hypergraph given a convex match and on fullness of
+
+\[
+    \llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \MdaCospans
+\]
+
+\question{What options we have?
+\begin{enumerate}
+    \item Define $PROP(\Sigma)^{+}$ and $PROP(\Sigma)^{+} + Frob$
+    \item Say that $n \xrightarrow{} \mathcal{G} \xleftarrow{} m$ rewrites into $n \xrightarrow{} \mathcal{F} \xleftarrow{} m$ iff $0 \xrightarrow{} \mathcal{G} \xrightarrow{} n + m$ rewrites into $0 \xrightarrow{} \mathcal{F} \xrightarrow{} n + m$.
+          This gives the following route
+          \[
+            \textsf{PROP}(\Sigma)^{+} \xrightarrow{\llbracket \cdot \rrbracket} \MdaEcospans \xrightarrow{\ulcorner \cdot \urcorner} \MdaEcospans
+          \]
+
+          and the same in the other direction with flipped arrows.
+\end{enumerate}
+}
 \Aleksei{TODO: discuss this}
 
-\Aleksei{This holds for boxes with bent wires, which require a compact closed structure.
-The proof for the "only if" direction refers to the result for rewriting in PROP + Frob and for the converse direction it uses decomposition lemma and isomorphism-ness of a functor $\llbracket \cdot \rrbracket$ (Theorem 25~\cite{Frobenius2})}
-
-\Aleksei{At some point I forgot to make a grey shaded area for cospans. Need to fix this.}
+If we quotient $\MdaCospans$ by $\llangle \mathcal{R} \rrangle$ by the theorem~\ref{theorem:rewriting_no_frob} and faithfulness and fullness of $\llbracket \cdot \rrbracket$ we get the following result.
+$f = g$ in $\textsf{PROP}(\Sigma,\mathcal{E})$ iff $\llangle ^{\ulcorner} f ^{\urcorner} \rrangle = \llangle ^{\ulcorner} f ^{\urcorner} \rrangle$ in $\MdaCospans$.
+
+\question{Do we want the same result for $\textsf{PROP}(\Sigma, \mathcal{E})^{+}$ and $\WellTypedMdaEcospans$?}
+
+\textbf{End of update}
 
 We have a more general notion of interface and hence we need to adjust DPOI rewriting a little bit.
-The first issue is that interfaces are not preserved during rewriting, consider an example above which corresponds to a rewrite rule of $\langle f, f+g \rangle$.
+The first issue is that whole interfaces are not preserved during rewriting, consider an example in Figure~\ref{fig:example_rewrite_ecospans} which corresponds to a rewrite rule of $\langle f, f+g \rangle$.
 
@@
     \[
-    \scalebox{0.75}{\tikzfig{combinatorial_semantics/f_to_f_plus_g_example}}
-    \]
+    \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_to_f_plus_g_example}}
+    \]
+    \caption{Example rewrite rule in $\WellTypedMdaEcospans$}
+    \label{fig:example_rewrite_ecospans}
 \end{figure}
@@
 
-\Aleksei{TODO: show that there exists a pushout for each of the axioms and lifted rewrites}
-
-\question{Shall I replace $i,j$ etc. with $X_{ext}$?}
-
-To account for such changes in the interfaces we put some additional constraints on DPOI rewriting. Consider a DPOI diagram with extra constraints below.
+\update{$i,j$ are replaced}
+
+To account for such changes in the interfaces we put some additional constraints on DPOI rewriting.
+Consider a DPOI diagram with extra constraints below which encodes
+ the notion that cospan 
+ $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$
+ rewrites into a cospan 
+ $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$ via a rewrite rule $\langle l, r \rangle$,
+ where
+ $l := x_{ext} \xrightarrow{} l_i \xrightarrow{} L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
+ and
+ $r := x_{ext} \xrightarrow{} r_i \xrightarrow{} L \xleftarrow{} r_o \xleftarrow{} y_{ext}$.
+ Note there are arrows from $x_{ext} + y_{ext} \to l_i + l_o$ (respectively, $r_i + r_o$) from the corresponding cospans which are not shown in the diagram.
+ Similarly for the arrows from $n_{ext} + m_{ext}$.
 
@@
 
-% https://q.uiver.app/#q=WzAsMTEsWzIsMSwiTCJdLFs0LDAsImkgKyBqIl0sWzIsMywiR157XFx1cmNvcm5lcn0iXSxbNCwzLCJDIl0sWzQsNSwibittIl0sWzAsMSwibF9pK2xfbyJdLFswLDMsImdfaStnX28iXSxbNiwxLCJSIl0sWzYsMywiXntcXHVsY29ybmVyfUgiXSxbOCwzLCJoX2kraF9vIl0sWzgsMSwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw2XSxbMSw1LCJbaV9sLGpfbF0iLDFdLFs0LDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzMsOCwiIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNyw4LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxLDcsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzEwLDddLFs5LDhdLFsxLDEwXSxbNCw5XV0=
+% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
 \[\begin{tikzcd}
-	&&&& {i + j} \\
-	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
+	{l_i+l_o} && L && {x_{ext} + y_{ext}} && R && {r_i+r_o} \\
 	\\
@@
 	\\
-	&&&& {n+m}
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
-	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
-	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
-	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
-	\arrow[from=2-1, to=2-3]
-	\arrow[from=4-1, to=4-3]
-	\arrow[from=6-5, to=4-1]
-	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
-	\arrow[from=2-9, to=2-7]
-	\arrow[from=4-9, to=4-7]
-	\arrow[from=1-5, to=2-9]
-	\arrow[from=6-5, to=4-9]
+	&&&& {n_{ext}+m_{ext}}
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
+	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
+	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
+	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
+	\arrow[from=1-1, to=1-3]
+	\arrow[from=3-1, to=3-3]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-7]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-7]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-7, to=3-7]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-7]
+	\arrow[from=1-9, to=1-7]
+	\arrow[from=3-9, to=3-7]
 \end{tikzcd}\]
 
-The original DPOI square is shown in red. The extra bits simply account for more complicated interfaces.
-For example, the left-hand side $L$ of a rewrite rule $\langle L, R \rangle$ which originally was some cospan $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ now is a cospan $i \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} j$ from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
-The top half of this diagram says that the outermost interfaces should coincide for $L$ and $R$. The bottom half puts the same restrictions on $G$ and $H$. 
+The original DPOI square is shown in red.
+The extra bits simply account for more complicated interfaces.
+For example, the left-hand side $L$ of a rewrite rule
+ $\langle L, R \rangle$ which originally was some cospan
+ $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
+ now is a cospan
+ $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
+from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
+The top half of this diagram says that the outermost interfaces
+ should coincide for $L$ and $R$.
+The bottom half puts the same restrictions on $G$ and $H$. 
 % The extra squares on both sides say that there should exist a matching of the interfaces together with a matching of e-hypergraphs.
@@
 
-% https://q.uiver.app/#q=WzAsNyxbMiwxLCJMIl0sWzQsMCwiaSArIGoiXSxbMiwzLCJHXntcXHVyY29ybmVyfSJdLFs0LDMsIkMiXSxbNCw1LCJuK20iXSxbMCwxLCJsX2krbF9vIl0sWzAsMywiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw2XSxbMSw1LCJbaV9sLGpfbF0iLDFdXQ==
+% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
 \[\begin{tikzcd}
-	&&&& {i + j} \\
-	{l_i+l_o} && L \\
+	{l_i+l_o} && L && {x_{ext} + y_{ext}} \\
 	\\
@@
 	\\
-	&&&& {n+m}
-	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
-	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
-	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
-	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
-	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
-	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
-	\arrow[from=2-1, to=2-3]
-	\arrow[from=4-1, to=4-3]
-	\arrow[from=6-5, to=4-1]
-	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
+	&&&& {n_{ext}+m_{ext}}
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
+	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
+	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
+	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
+	\arrow[from=1-1, to=1-3]
+	\arrow[from=3-1, to=3-3]
 \end{tikzcd}\]
-Consider the diagram above where the marked square is a pushout and $i + j$ and $n + m$ are the \textit{outermost} interfaces for cospans $i \xrightarrow{} l_i \xrightarrow{} L \xleftarrow{} l_o \xleftarrow{} j$ and $n \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m$.
-We say that $i+j \to C \to G^{\urcorner}$ is a \textit{boundary} complement if $[i_c, i_j]$ is mono and there exist $n_c : n \to C$ and $m_c : m \to C$ making the above triangle commute and such that there exists a \textit{not necessarily well-typed} mda-cospan
-
-\[
-n \xrightarrow{} c_i + j \xrightarrow{} C \xleftarrow{} c_o + i \xleftarrow{} m
-\]
-
-where $c_i$ and $c_o$ are computed as $c_i = g_i \setminus (l_i \setminus i)$ and $c_o = g_o \setminus (l_o \setminus j)$.
-
-% \Aleksei{The pushout interpretation is wrong because there might not be a morphism between $i$ and $c_i$ and similarly for $j$ and $c_o$.
-Any other categorical interpretation?
-The outermost interface of the rewrite span may not be a part of the interface of $G$}
-
-% where $l_i + l_o$ is the whole interface for $L$, $g_i + g_o$ is the whole interface for $G$.
+
+Consider the diagram above where the marked square is a pushout
+ and $x_{ext} + y_{ext}$ and $n_{ext} + m_{ext}$ are 
+ the \textit{outermost} interfaces for the corresponding 
+ cospans.
+We say that $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ is 
+ a \textit{boundary} complement if $[i_c, i_j]$ is mono 
+ and there exist $n_c : n_{ext} \to C$ and $m_c : m_{ext} \to C$ 
+ making the above triangle commute and such that
+there exists a \textit{not necessarily well-typed} mda-cospan
+
+\[
+n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}
+\]
+
+where $c_i$ and $c_o$ are computed as
+ $c_i = g_i \setminus (l_i \setminus x_{ext})$ and
+ $c_o = g_o \setminus (l_o \setminus y_{ext})$.
 \end{definition}
@@
 \begin{remark}
-Note that cospan $n \xrightarrow{} c_i + j \xrightarrow{} C \xleftarrow{} c_o + i \xleftarrow{} m$ is not well-typed.
+Note that cospan $n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}$ is not well-typed.
 The cospan's input interface contains nodes that previously were a part of the output interface.
@@
 First, let's check that this notion generalises the ordinary boundary complement from~\ref{def:boundary_original}.
-For this, we assume that child and conflict relations are empty in which case $i + j$ and $l_i + l_o$ coincide as well as $g_i + g_o$ and $n + m$.
-Hence, $g_i \setminus (l_i \setminus i) + j = g_i + j = n + j$ and the same for the output interfaces.
-Thus we have $n + j \xrightarrow{} C \xleftarrow{} m + i$, which corresponds to the definition from~\ref{def:boundary_original}.
+For this, we assume that child and conflict relations are empty in which case $x_{ext} + y_{ext}$ and $l_i + l_o$ coincide as well as $g_i + g_o$ and $n_{ext} + m_{ext}$.
+Hence, $g_i \setminus (l_i \setminus x_{ext}) + y_{ext} = g_i + y_{ext} = n_{ext} + y_{ext}$ and the same for the output interfaces.
+Thus we have $n_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} m_{ext} + x_{ext}$, which corresponds to the definition from~\ref{def:boundary_original}.
 
@@
 \begin{definition}[Rewriting in $\WellTypedMdaEcospans$]
-    Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n + m$ and $H \xleftarrow{} h_i + h_o \xleftarrow{} n+m$ in $\WellTypedMdaEcospans$ with the outermost interfaces $n + m$, $G$ rewrites convexly into $H$ with the outermost interface $n + m$ --- notation $( G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}} (E \xleftarrow{} n + m )$ --- if there exists a rule $l_i + l_o \xrightarrow{} L \xleftarrow{} i + j \xrightarrow{} R \xleftarrow{} r_i + r_o$ with the outermost interfaces $i+j$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
-
-% https://q.uiver.app/#q=WzAsMTEsWzIsMSwiTCJdLFs0LDAsImkgKyBqIl0sWzIsMywiR157XFx1cmNvcm5lcn0iXSxbNCwzLCJDIl0sWzQsNSwibittIl0sWzAsMSwibF9pK2xfbyJdLFswLDMsImdfaStnX28iXSxbNiwxLCJSIl0sWzYsMywiXntcXHVsY29ybmVyfUgiXSxbOCwzLCJoX2kraF9vIl0sWzgsMSwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw2XSxbMSw1LCJbaV9sLGpfbF0iLDFdLFs0LDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzMsOCwiIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNyw4LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxLDcsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzEwLDddLFs5LDhdLFsxLDEwXSxbNCw5XV0=
+    Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n_{ext} + m_{ext}$ 
+    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\WellTypedMdaEcospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $l_i + l_o \xrightarrow{} L \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} R \xleftarrow{} r_i + r_o$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
+
+% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
 \[\begin{tikzcd}
-	&&&& {i + j} \\
-	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
+	{l_i+l_o} && L && {x_{ext} + y_{ext}} && R && {r_i+r_o} \\
 	\\
@@
 	\\
-	&&&& {n+m}
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
-	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
-	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
-	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
-	\arrow[from=2-1, to=2-3]
-	\arrow[from=4-1, to=4-3]
-	\arrow[from=6-5, to=4-1]
-	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
-	\arrow[from=2-9, to=2-7]
-	\arrow[from=4-9, to=4-7]
-	\arrow[from=1-5, to=2-9]
-	\arrow[from=6-5, to=4-9]
+	&&&& {n_{ext}+m_{ext}}
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
+	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
+	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
+	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
+	\arrow[from=1-1, to=1-3]
+	\arrow[from=3-1, to=3-3]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-7]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-7]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-7, to=3-7]
+	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-7]
+	\arrow[from=1-9, to=1-7]
+	\arrow[from=3-9, to=3-7]
 \end{tikzcd}\]
@@
         \item $m : L \to G$ is a convex down-closed conflict reflecting match;
-        \item $i + j \to C \to G$ is a boundary complement in the leftmost pushout.
+        \item $x_{ext} + y_{ext} \to C \to G$ is a boundary complement in the leftmost pushout.
     \end{enumerate}
@@
 
-% \Aleksei{We may need the following proofs:}
-% \begin{itemize}
-%     \item The existence of pushout complements (for rewriting)
-%     \item Soundness and completeness, that is the correspondence between enriched terms and e-hypergraphs.
-%     \item Rewriting
-%     \item Anything else?
-% \end{itemize}
-
 \Aleksei{Do we need to introduce substitution? This is needed if our rewrite rules are going to have variables as labels.}
 
-The whole interfaces of the resulting e-hypergraph $H$ from the cospan $n \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m$ are constructed as $h_i = (r_i \setminus i) + c_i$ and $h_o = (r_o \setminus j) + c_o$.
-$h_i \to H$ is defined as a copairing of $f_1 : r_i \setminus i \to R \to H$ and $f_2 : c_i \to C \to H$.
+The whole interfaces of the resulting e-hypergraph $H$ from the cospan
+ $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$
+ are constructed as $h_i = (r_i \setminus x_{ext}) + c_i$ 
+ and
+ $h_o = (r_o \setminus y_{ext}) + c_o$.
+ $h_i \to H$ is defined as a copairing of 
+ $f_1 : r_i \setminus x_{ext} \to R \to H$ 
+ and 
+ $f_2 : c_i \to C \to H$. 
 Since both $f_1$ and $f_2$ are mono, their copairing is also a mono. 
@@
 
-$n \xrightarrow{n_h} h_i \xrightarrow{f} H \xleftarrow{g} h_o \xleftarrow{m_h} m$ is a well-typed mda-cospan.
+$n_{ext} \xrightarrow{n_h} h_i \xrightarrow{f} H \xleftarrow{g} h_o \xleftarrow{m_h} m_{ext}$ is a well-typed mda-cospan.
 
 \begin{proof}
-$f$ and $g$ are monos by construction. $n + m \to h_i + h_o$ is a mono by commutativity because $n + m \to H$ is a mono (also by commutativity). 
-The only in-degree 0 nodes have a pre-image in $c_i$ and $r_i \setminus i$  and end up being in the image of $h_i$ (the nodes with the pre-image in $j$ get are glued together).
+$f$ and $g$ are monos by construction.
+$n_{ext} + m_{ext} \to h_i + h_o$ is a mono by commutativity because $n_{ext} + m_{ext} \to H$ is a mono (also by commutativity). 
+The only in-degree 0 nodes have a pre-image in $c_i$ and $r_i \setminus x_{ext}$  and end up being in the image of $h_i$ (the nodes with the pre-image in $y_{ext}$ get glued together).
 Well-typedness is restored once the right-hand side of the rewrite rule is glued into $C$.
-\question{Uniqueness of this pushout complement?
-Is it a presheaf?
-As a presheaf category is adhesive, $\catname{Csp_{D}(EHyp)}$ can not be a presheaf as it is not adhesive.}
+\question{Uniqueness of this pushout complement? Is it a presheaf? As a presheaf category is adhesive, $\catname{Csp_{D}(EHyp)}$ can not be a presheaf as it is not adhesive.}
 \end{proof}
@@
 
-\Aleksei{Check the proof
-}
+\Aleksei{Check the proof}
 
@@
 
+\update{I've switched to $[ \cdot ]$ for our functor to be able to differenciate between $\llbracket \cdot \rrbracket$, $\llangle \cdot \rrangle$}
 \begin{remark}
-    For now, we will informally denote a combinatorial interpretation of a morphism $f$ in $\textsf{PROP}^{+}(\Sigma)$ as $\llbracket f \rrbracket$.
+    For now, we will informally denote a combinatorial interpretation of a morphism $f$ in $\textsf{PROP}^{+}(\Sigma)$ as $[ f ]$.
 \end{remark}
@@
 
-Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we add them as structural rewrite rules into $\llbracket \mathcal{R}^+ \rrbracket$.
+Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we add them as structural rewrite rules.
 Recall that the category $\catname{EHyp_{\Sigma}}$ does not have all pushouts and hence we need to explicitly show that the pushouts that we need exist.
@@
     Since we glue along the outermost nodes this does not introduce any conflict with regard to child relation.
-Such equations are not absorbed by e-hypergraph representation unlike SMC equations because the right-hand side contains more edges.
+    Such equations are not absorbed by e-hypergraph representation unlike SMC equations because the right-hand side contains more edges, i.e. graphs $G$ and $H$ are not isomorphic.
+    Similar argument applies to other rules as well.
 
@@
 \begin{definition}
-    We will denote $\llbracket \mathcal{E} \rrbracket$ the interpretation of the equations in~\ref{fig:string-equations} as rewrite rules in $\WellTypedMdaEcospans$.
-In particular, for all $l = r$ in $\mathcal{E}$ $\llbracket \mathcal{E} \rrbracket$ will contain $\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle$ and $\langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle$.
-\update{Technically we add those for all input and output interfaces.}
+    We will denote with $\mathcal{S}$ the interpretation of the equations in~\ref{fig:string-equations} as rewrite rules in $\WellTypedMdaEcospans$.
+    In particular, for all $l = r$ in~\ref{fig:string-equations}, $\mathcal{S}$ will contain $\langle [l], [r]\rangle$ and $\langle [r], [l] \rangle$.
+    \update{Technically we add those for all possible instances of $l$ and $r$ in~\ref{fig:string-equations}.}
 \end{definition}

### 2023-10-18 sections/categorical_semantics.tex
Fixesing inaccuracies for combinatorial semantics
intro/conclusion

@@
+% !TEX root = ../main.tex
+
 \newtheorem*{notation*}{Notation}
@@
 \end{itemize}
-satisfying certain coherence conditions [MacLane].
+satisfying certain coherence conditions 
+\cite{maclane}.
 A \textit{permutations and products category (PROP)} is a symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$.
@@
 \begin{align*}
-	\infer{x_1, \ldots, x_n \vdash x_j}{x_i \in A} \qquad \infer{\Gamma \vdash t_1 + t_2}{\Gamma \vdash t_1 && \Gamma \vdash t_2} 
+	\infer{\Gamma \vdash x}{x \in \Gamma} \qquad \infer{\Gamma \vdash a}{a \in A} \qquad \infer{\Gamma \vdash t_1 + t_2}{\Gamma \vdash t_1 && \Gamma \vdash t_2} 
 \end{align*}
@@
 
-It will often be convenient to present SMTs in terms of string diagrams [which are technically equiv classes of arrows?!] 
+It will often be convenient to give presentations of SMTs in terms of string diagrams.
 \begin{example}
@@
 \end{definition}
-Note, in particular,  that semigroups \update{semilattice?} do not require a unit for $+$.
+Note, in particular,  that semilattices do not require a unit for $+$.
 {This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct -- a degeneracy we wish to avoid.}
@@
 To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the binary operation on morphisms.
-This will later be used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
+It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
 
@@
 We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics.
-Indeed, the diagrammatic language itself is not really novel and is similar to \textit{e.g.}, \cite{duncan_generalised_2009}.
+Indeed, the diagrammatic language itself is not truly novel and is similar to \textit{e.g.}, \cite{duncan_generalised_2009}.
 

### 2023-10-18 sections/conclusion.tex
Fixesing inaccuracies for combinatorial semantics
update intro
intro/conclusion

@@
-\section{Conclusion and Further Work}
+% !TEX root = ../main.tex
 
-\begin{itemize}
-\item implementation for ZX
-\item computational linguistics
-\item traced categories, hypergraphs for traced comonoids and full e-graphs
-\item closed categories, hierarchical hypergraphs and lambda-calculus (example)
-\end{itemize}
+\section{Conclusion}
+
+We have described a categorical semantics for e-graphs over algebraic theories in terms of Cartesian categories enriched over a semilattice, and defined a natural generalization to account for monoidal theories in terms of SMCs,  similarly enriched.
+Having described a translation of e-graphs into our formalism, we developed a combinatorial representation of morphisms in our free category which factor out the structural SMC equations, but remain sensitive to the complexity-relevant equations induced by the enrichment.
+The design of such a representation and its corresponding  rewrite theory proved delicate, and -- together with its proof of completeness -- constitute the main technical contributions of the paper.
+Finally, we considered an analogue of e-rewriting in our setting, pushing towards the practical considerations necessary for an implementation of our ideas.
+Although we believe our work forms a solid basis for a formal mathematical theory of (acyclic) e-graphs,  there are a number of extensions which we intend to investigate in the future.
+
+\subsection{Future Work and Applications}
+
+We have already alluded to the fact we deal with only \textit{acyclic} e-graphs in this paper.
+Nevertheless,  within this setting we present a generalization from Cartesian to (symmetric) monoidal theories,  giving rise to a host of potential new domains of application.
+Further,  the string diagrammatic and graph rewriting framework we set up is ameenable to extension, including lifting the acyclicity restriction.
+
+\begin{figure}\label{fig:applications}
+\[
+	\scalebox{0.75}{
+	\tikzfig{combinatorial_semantics/example-e-graphs}
+	}
+\]
+\caption{Example applications of e-hypergraphs.  Left, the ZX-calculus.  Right, the $\lambda$-calculus as represented by string diagrams for Cartesian closed categories.  }
+\end{figure}
+
+Symmetric monoidal theories have received significant attention over recent years: the ZX-calculus can be used to characterize quantum processes, with applications to optimization of quantum circuits [cite] ; similar theories also characterize signal flow diagrams
+\cite{baldan_categorical_2014}, digital circuits 
+\cite{ghica_operational_2021}, as well as having application to computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.
+In future work, we hope to apply our generalization of e-graphs to optimization problems in these domains (e.g.
+Figure \ref{fig:applications})
+
+The theories described above often have more structure than simply being symmetric monoidal: for example, they may contain comonoids, Frobenius algebras or traced structure (as in the case of e-graphs with cycles),  and this structure can be exploited to give more natural (in the sense of factoring out more equations) combinatorial representations.
+In the above cases, such representations and their DPO-rewriting theory is well-studied
+\cite{ghica_rewriting_2023} % traced comonoid
+[cite, cite, cite].
+We consider that our development of hierarchical "e-box" structure within string diagrams and their associated (e-)hypergraphs ( categorically,  semilattice enrichment) is essentially orthogonal to these developments, and so ought to be easily combined to give notions of e-hypergraphs specifically suited to the optimization of these particular monoidal theories.
+
+A particularly interesting extension that we intend to consider is the combination of our "e-structure" with (Cartesian) \textit{closed} categories, allowing the natural combination of e-graph optimization techniques with the $\lambda$-calculus.
+The corresponding combinatorial representation would build on the recent work \cite{ghica_hierarchical_2023, alvarez-picallo_rewriting_2022, alvarez-picallo_functorial_2021}.
+% hierarchical
+
+In Figure \ref{fig:let-calculus} we give an example term calculus corresponding to our categorical semantics.
+Although undeveloped here, the first author intends to develop this idea, extending the calculus of \cite{heijltjes_functional_2023,barrett_functional_2023} into a natural term language for such string diagrams.
+We also mention that our developments would also naturally combine with recent work on string diagrams for \textit{non-strict} symmetric monoidal categories \cite{wilson_string_2023}.
+% non-strict categories
+Finally, we intend to develop implementations of the theory developed here,  with an ultimate aim of providing a tool for optimization of monoidal theories and assessing its practicality on real-world problems, such as those considered in the various domains mentioned above.

### 2023-10-18 sections/introduction.tex
intro/conclusion

--- 2023-10-17 sections/introduction.tex
@@
 \newcommand\mylet[2]{\textsf{let } #1 = #2 \textsf{ in }}
-
-\cite{heijltjes_functional_2023}
-\cite{barrett_functional_2023}
-\cite{griggio_proceedings_2022}
-\cite{flatt_small_2022}
-\cite{ghica_operational_2021}
- % digital circuits
-\cite{ghica_hierarchical_2023}
-\cite{ghica_rewriting_2023} % traced comonoid
-\cite{wilson_string_2023} % non-strict categories
-
-\cite{zhang_relational_2022} % relational ematching 
-\cite{alvarez-picallo_rewriting_2022} % hierarchical
-\cite{alvarez-picallo_functorial_2021} % hierarchical
-
-\cite{bonchi_tape_nodate} % tape diagrams
-\cite{baldan_categorical_2014} % signal flow
-\cite{dpo}
-\cite{maclane}
-\cite{singher2023colored}
 
@@
 \cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimized program which is rewritten \textit{destructively} at each step, we can maintain a \textit{set} of equivalent programs, where each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted.% from the set,  which now represents all equivalent ways to express the program with respect to the given rewrites.
-
+Upon reaching a fixed point ({saturation}),  a \textit{globally} optimal program can then be extracted.% from the set,  which now represents all equivalent ways to express the program with respect to the given rewrites.
 The use of \textit{equality graphs (e-graphs)} \cite{EggPaper} -- a generalization of the typical directed acylcic graph (DAG) representation of terms to include equivalence classes -- to compactly represent the set of equivalent programs is what makes the technique tractable in practise, where the naive approach is clearly unfeasible.
 
-\update{Move focus from semantics of e-graphs to their extension to monoidal theories?}
 While equality saturation is a state-of-the-art optimization technique, the formal mathematical theory of e-graphs is perhaps underdeveloped.
@@
 Therefore, our theoretical development leads to new potential domains of application, including optimziation of string diagrams representing digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX} and potentially to natural language \cite{wazni_quantum_2022,coecke_lambek_2013}.
-We also suggest how our approach will allow future  generalization to include rewriting for the $\lambda$-calculus
-\cite{koehler2022sketchguided} % latest egraph.
+\cite{coecke_interacting_2011,ZX} and potentially to computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.
+We also suggest how our approach will allow future  generalization to include rewriting for the $\lambda$-calculus,  in a novel approach \textit{cf.} \textit{e.g.,}
+\cite{koehler2022sketchguided}. % latest egraph.
 
-Towards implementation, we 
-additionally provide a combinatorial representation of our enriched string diagrams in terms of (hierarchical)  hypergraphs.
-We give a specification of hypergraph rewriting in this novel setting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  thus also formalizing the rewriting theory of standard e-graphs.
+With consideration of future implementations, we 
+additionally provide a combinatorial representation of our enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
+We give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
+\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}, and prove the representation sound and complete with respect to our categorical semantics.
+As a corollary, we also formalize the rewriting theory of standard e-graphs in this way.
+Finally, we touch on the question of efficient implementation of rewriting for e-hypergraphs.
 
-\subsection*{E-graphs}
+\subsection{E-graphs}
 
@@
 
-\subsection*{Semantics via Semilattice Enriched Categories}
+\subsection{Semantics via Semilattice Enriched Categories}
 
-One can easily see the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category by considering string diagrams 
+Recalling the correspondence between DAG representations of algebraic terms and the corresponding string diagrammatic representation of morphisms of a suitable Cartesian category, 
+one can easily see the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category by considering string diagrams 
 \cite{noauthor_09083347_nodate,joyal_geometry_1991, mellies_functorial_2006}.
-In Figure \ref{fig:egraph-strings} we have extended the usual string diagrammatic notation for symmetric monoidal categories with an additional generator which has the  typing rule below, left.
-The ability to take formal sums of morphisms is used to model the equivalence class structure of e-graphs.  
+In Figure \ref{fig:egraph-strings} we have extended the usual string diagrammatic notation for symmetric monoidal categories with an additional generator which has the  typing rule below, left, corresponding to taking formal sums of morphisms.
+The ability to take formal sums is used to model the equivalence class structure of e-graphs.  
 \[
@@
 Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal sums.
-Thus we generalize e-graphs from algebraic to monoidal theories.
+Thus we generalize e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we outline in the conclusion of this paper.
 
-\subsection*{Combinatorial Representation of Enriched String Diagrams}
+\subsection{Combinatorial Representation of Enriched String Diagrams}
 
-In order to \textit{implement} our generalized version of e-graphs, we will  perform rewriting not on algebraic terms, but on combinatorial representations of the corresponding string diagrams.
+In order to \textit{implement} our generalized version of e-graphs, we will  perform rewriting not on terms representing morphisms of a category, but on combinatorial representations of the corresponding string diagrams.
 String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo "connectivity") or as a 2-dimensional syntax quotiented by the equations of an SMC.
@@
 Unfortunately, this makes string diagrams relatively unamenable to efficient implementation as a data structure, due to the difficulty of calculating the quotient. 
-
 Nevertheless, there is a body of work dealing with this issue by representing string diagrams as combinatorial objects -- namely, (appropriately generalized) hypergraphs 
@@
 
-We are interested in not just the free SMC over a set of generators, but rather in \textit{symmetric monoidal theories} which involve considering extra equations between morphisms: for instance, the rewrite rules $(a)-(e)$ of Figure \ref{fig:e-graph-example}.
+We are interested in not just the free SMC over a set of generators, but rather in \textit{(symmetric) monoidal theories} which involve considering extra equations between morphisms: for instance, the rewrite rules $(a)-(e)$ of Figure \ref{fig:e-graph-example}.
 These can be seen as equations between string diagrams, and thus between their hypergraph representations.
 Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ in any hypergraph $G$ and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard approach to this is known as\textit{double pushout (DPO) rewriting}.
+We make use of the standard approach to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting}.
 
-Our application requires a generalization of these concepts from hypergraphs to \textit{e-hypergraphs}, which we define as a hypergraph with two additional relations denoting the hierarchical structure introduced by "e-boxes" ( \textit{i.e.,} the generator for semilattice enrichment) and the separation of the components of each e-box.
+Our application requires a generalization of these concepts from hypergraphs to \textit{e-hypergraphs}, which we define as a hypergraph with two additional relations denoting the hierarchical structure introduced by \textit{e-boxes} (\textit{i.e.,} the generator for semilattice enrichment) and the separation of the components of each e-box.
 Giving the appropriate definitions is slightly delicate matter, and constitutes the main body of the paper.
 
-\subsection*{Soundness and Completeness}
+\subsection{Soundness and Completeness}
 The main technical result of the paper is a proof soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs.
 While the structural equalities of SMCs are factored out in the representation,  important structural equalities arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be: they represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating symmetric monoidal theory (which do not).
+Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
 
-Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched categories.
+Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched SMCs.
 That is, we exhibit a translation functor $\llbracket-\rrbracket$ from the free semilattice enriched SMC over a symmetric monoidal theory to our (quotiented) category of e-hypergraphs such that
-morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow \llbracket_{DPO} g \rrbracket$.
+morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the monoidal theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow_{DPO} \llbracket g \rrbracket$.
 
-\subsection*{E-matching and E-rewriting}
+\subsection{E-matching and E-rewriting}
 
-Having defined e-hypergraphs and shown them to correctly model rewriting string diagrams for SMCs, equipped with a notion of "equivalence class", we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
+Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
 In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of e-matching for e-hypergraphs: that is, to find redexes working modulo the e-structure.
-In particular, we wish to locate the smallest subgraph $G' \subset G$ "containing" the redex $L$.
-Given this, the analogue of e-rewriting $L \to R$ is simple to define, due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-
-[Example diagrams]
-
-\subsection*{Future Applications}
-
-Example diagrams for circuits, trace (loop unrolling, PEGs), LC, ZX here.
+Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
+In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
+Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
+\begin{figure}
+\[
+	\tikzfig{combinatorial_semantics/example-rewriting}
+\]
+\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
+REDO THIS}
+\end{figure}
 
 \subsection{Related Work}
-Q: Related work - at end? \\
+
+Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimizers and synthesizers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
+This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries ("relational e-matching") \cite{zhang_relational_2022}.
+We are hopeful that our novel perspective on e-graphs can be extended to give foundations to these investigations.
+Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
+
+Our use of string diagrams is also central to the intuition behind our work.
+The string diagrams we work with are left essentially as informal reasoning devices, however there exists a breadth of work on similar diagrammatic formalisms for similar classes of categories.
+We mention here work on hierarchical string diagrams \cite{ghica_hierarchical_2023},  functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.

### 2023-10-19 sections/soundness_and_completeness.tex
Theorem about rewriting

@@
             Then we need to show that $[f] \Rrightarrow_{\langle l, r \rangle \in \mathcal{S}} [g']$ and $[g'] = [g]$.
+            }
+        \question{Can we use normal form here as well? $f = g$ means there exists $h$ such that $f =_{SMC^{+}} h$ and $h =_{SMC^{+}}$ and $h$ is of the form $\phi_1 + \ldots + \phi_n$?
+                  Then we need to show that $[f] \Rrightarrow_{\mathcal{S}} [h]$ and $[h] \Rrightarrow_{\mathcal{S}} [g]$.
         }
@@
 
-    Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$.
-    $f \Rightarrow{}_{\mathcal{R}^{+}}$ iff $[f] \Rrightarrow{}_{[\mathcal{R}^{+}]} [g]$.
+    Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$ and $\langle l, r \rangle \in \mathcal{R}$.
+    $a \Rightarrow{}^{+}_{\langle l, r \rangle} b$ iff $[a] \Rrightarrow{}^{+}_{\langle [l], [r] \rangle} [b]$.
     \begin{proof}
+        $a \Rrightarrow^{+}_{\langle l, r \rangle} b$ means that either
+        \begin{itemize}
+            \item $a$ and $b$ are decomposable as
+                  \[
+                    \scalebox{0.75}{
+                        \tikzfig{combinatorial_semantics/prop_rewriting}
+                    }
+                  \]
+                  \question{Here we can probably just reuse the proof going to $\ulcorner a \urcorner$ and then the result follows by functoriality of $[\cdot]$ (or $\llangle \cdot \rrangle$)}
+            \item or
+                 \[
+                    \scalebox{0.75}{
+                        \tikzfig{combinatorial_semantics/prop_plus_rewriting}
+                    }   
+                 \]
+                 and $a' \Rightarrow_{\langle l,r \rangle} b'$. This gives us the following decomposition in cospans
+                 % https://q.uiver.app/#q=WzAsMjcsWzAsNCwibSJdLFsxLDQsIkFfMSJdLFsyLDQsImkgKyBrIl0sWzMsNCwiOyJdLFs0LDAsImkiXSxbNSwwLCJYXzEiXSxbNiwwLCJqIl0sWzUsMiwiXFx2ZG90cyJdLFs1LDQsIkEnIl0sWzQsNCwiaSJdLFs2LDQsImoiXSxbNSwxLCIrIl0sWzUsMywiKyJdLFs1LDUsIisiXSxbNSw2LCJcXHZkb3RzIl0sWzUsNywiKyJdLFs0LDgsImkiXSxbNSw4LCJYX3ciXSxbNiw4LCJqIl0sWzcsNCwiXFxvdGltZXMiXSxbOCw0LCJrIl0sWzksNCwiayJdLFsxMCw0LCJrIl0sWzExLDQsIjsiXSxbMTIsNCwiaiArIGsiXSxbMTMsNCwiQV8yIl0sWzE0LDQsIm4iXSxbMCwxXSxbMiwxXSxbNCw1XSxbNiw1XSxbOSw4XSxbMTAsOF0sWzE4LDE3XSxbMTYsMTddLFsyMCwyMV0sWzIyLDIxXSxbMjYsMjVdLFsyNCwyNV1d
+                \[
+                    \adjustbox{scale=0.5, center}{
+                % \scalebox{0.75}{
+                    \begin{tikzcd}
+                    &&&& i & {X_1} & j \\
+                    &&&&& {+} \\
+                    &&&&& \vdots \\
+                    &&&&& {+} \\
+                    m & {A_1} & {i + k} & {;} & i & {A'} & j & \otimes & k & k & k & {;} & {j + k} & {A_2} & n \\
+                    &&&&& {+} \\
+                    &&&&& \vdots \\
+                    &&&&& {+} \\
+                    &&&& i & {X_w} & j
+                    \arrow[from=5-1, to=5-2]
+                    \arrow[from=5-3, to=5-2]
+                    \arrow[from=1-5, to=1-6]
+                    \arrow[from=1-7, to=1-6]
+                    \arrow[from=5-5, to=5-6]
+                    \arrow[from=5-7, to=5-6]
+                    \arrow[from=9-7, to=9-6]
+                    \arrow[from=9-5, to=9-6]
+                    \arrow[from=5-9, to=5-10]
+                    \arrow[from=5-11, to=5-10]
+                    \arrow[from=5-15, to=5-14]
+                    \arrow[from=5-13, to=5-14]
+                \end{tikzcd}}
+                \]
+                and there is a convex match of $L$ in $A'$.
+                It is then also down-closed and conflict reflecting as $A'$ is not hierarchical.
+                Need to show that arising pushout complement is a boundary complement.
+                \question{A problem is that the above terms does not decompose into a form of $A';C$ where $C$ is a pushout complement.
+                This happens because we take composition along the external interfaces and this $A'$ is within a box, so it is not a part of an external interface.
+                }
+                \Aleksei{discuss this. We can simplify PROP+ rewriting to just a box, but still have a problem with context C}
+        \end{itemize}
         \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.

### 2023-10-19 main.tex
Fixesing inaccuracies for combinatorial semantics

@@
 
+% for quotienting
+% \usepackage{faktor}
+\usepackage{xfrac}
+
 \usepackage{import}
@@
 % category names
+\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
+
 \newcommand{\Ecospans}{{\catname{ECsp_{D}(EHyp_{\Sigma})}}}
@@
 
-\title{Equivalence Hypergraphs:\\ Semantics of E-Graphs and their Generalization to 
-Monoidal Theories}
-\author{Chris Barrett, Dan Ghica, Aleksei Tiurin}
+\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories }
+\author{Chris Barrett}
+\author{Dan Ghica}
+\author{Aleksei Tiurin}
 
@@
 \begin{abstract}
-Abstract.
+Equivalence graphs (e-graphs), together with a technique called equality saturation, have recently proved a powerful and practical tool for optimization of programs.
+We describe a categorical semantics for e-graphs over algebraic theories in terms of Cartesian categories enriched over a semilattice,  and define a natural generalization in terms of symmetric monoidal categories to account for monoidal theories,  with the aim of allowing optimization of a larger class of programs based on string diagrams.
+We further develop a sound and complete combinatorial representation of morphisms in our category based on a generalization of hypergraphs,  which we call e-hypergraphs,  which factor out many structural equations.
+Finally, we consider the question of efficient rewriting in our setting.
 \end{abstract}
@@
 \subsection*{Acknowledgements}
-Sadrzadeh Mehrnoosh
+We are grateful to Sadrzadeh Mehrnoosh for enlightening discussions on potential applications of e-hypergraphs to computational linguistics.
 
-\clearpage
 \bibliographystyle{acm}

### 2023-10-25 sections/conclusion.tex
fixed clashed figure
conclusion

@@
 
-Symmetric monoidal theories have received significant attention over recent years: the ZX-calculus can be used to characterize quantum processes, with applications to optimization of quantum circuits [cite] ; similar theories also characterize signal flow diagrams
+Symmetric monoidal theories have received significant attention over recent years: the ZX-calculus can be used to characterize quantum processes, with applications to optimization of quantum circuits \cite{coecke_interacting_2011,ZX} ; similar theories also characterize signal flow diagrams
 \cite{baldan_categorical_2014}, digital circuits 

### 2023-10-25 sections/e-graph-translation.tex
Fixesing inaccuracies for combinatorial semantics
conclusion

@@
+% !TEX root = ../main.tex
+
 \section{E-Rewriting for E-Hypergraphs}
@@
 The (convex) \textit{matching} of one e-hypergraph $L$ (\textit{e.g.}, the left-hand side of a rewrite rule) into another e-hypergraph $G$ was defined in the previous section as a monic morphism $L \to G$ (whose image is additionally convex, \update{consistent}).
-In other words, it corresponds to finding a subgraph isomorphic to $L$ within $G$ (subject to conditions respecing the hierarchical and conflict structures).
+In other words, it corresponds to finding a subgraph isomorphic to $L$ within $G$ (subject to conditions respecting the hierarchical and conflict structures).
 
@@
 In this section, we define the  e-matching (and subsequent e-rewriting) as the problem of finding matches (and applying rewrites) within an e-hypergraph \textit{without} first inducing duplication of subgraphs.
-We begin first with definitions, before proceeding to prove them “correct” with respect to the plain notion of matching (rewriting).
-\update{Example}
+We begin first with definitions, before proceeding to prove them correct with respect to the plain notion of matching (rewriting).
 

### 2023-10-25 sections/combinatorial_semantics.tex

@@
 
+\begin{remark}[Pushout in $\catname{Hyp_{\Sigma}}$]
+    Consider a diagram below
+    % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAC0QBfU9TXfIRQBGclVqMWbABrdeIDNjwEiZYePrNWiEAE05fJYKKj11TVJ0AdK3ABmAJzoBjYNIDUursBvYAtgAEAEpcBgr8ykLIAEyk0RqS2iAAitziMFAA5vBEoI4QfkhkIDgQSKISWmx2YfmFiMWlSLGVliCZINQMdABGMAwAChHGOg5YmQAWOLUOBeXUTYgAzOaJbL6OLsBYAPrC3r5YgSEzc4gtiyutSRtOrrvRB7ZHwaE8ebP1F2WIACzUfTAUCQAFolsULEkAFY7aKdEDdPqDYYqUbjKaneoVRb-ECA4HLCFrHQw4SYpBXHGrKo6JhpLhAA
+
+\[
+    \adjustbox{scale=1}{
+        \begin{tikzcd}
+        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{i_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
+        Y \arrow[r, "\sfrac{i_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+                                                                        &                                                                  & Q
+        \end{tikzcd}
+    }
+\]
+
+Pushout of two hypergraphs $X = \{V_{X}, E_{X}, s_{X}, t_{X}\}$ and $Y = \{V_{Y}, E_{Y}, s_{Y}, t_{Y}\}$ (we will omit labelling functions here since they are irrelevant) along $Z$ is computed in two steps.
+First, a coproduct of $X$ and $Y$ is taken which is $X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X} + s_{Y}, t_{X} + t_{Y} \}$.
+Then, consider a relation 
+\[
+    S = \{(x_i,y_j), (y_j, x_i) \in (X+Y) \times (X+Y)\; | \; \exists z \in Z \; . \; x_i = f(z) \text{ and } y_j = g(z) \text{where $x_i \in X$ and $y_j \in Y$ }\}
+% \cup \{(x,x) \in (X+Y) \times (X+Y) \text{ where $x \in X$}\} \cup \{(y,y) \in (X+Y) \times (X+Y) \text{ where $y \in Y$}\}$ 
+\]
+and let relation $R$ be its reflexive and transitive closure. $R$ is then an equivalence relation since $S$ is symmetric.
+
+A quotient of $X+Y$ by this relation denoted by 
+\[
+    \sfrac{X+Y}{\sim R} = \{\sfrac{V_{X} + V_{Y}}{\sim R}, \sfrac{E_{X} + E_{Y}}{\sim R}, \sfrac{s_{X} + s_{Y}}{\sim R}, \sfrac{t_{X} + t_{Y}}{\sim R}\}
+\]
+where $\sfrac{s_{X} + s_{Y}}{\sim R} : \sfrac{E_{X} + E_{Y}}{\sim R} \to \sfrac{V_{X} + V_{Y}}{\sim R}$ now maps equivalence classes to equivalence classes.
+Note that if $s : e_1 \mapsto \{v_1,v_2\}$ and $s : e_2 \mapsto \{v_3,v_4\}$  and $e_1 \sim e_2$ then $\sfrac{s}{\sim R} : [e_1,e_2] \to \{[v_1,v_3],[v_2,v_4]\}$ because $f$ and $g$ are homomorphisms.
+Then, $\sfrac{i_1}{\sim R}$ maps $y \in Y$ to its equivalence class and likewise for $\sfrac{i_2}{\sim R}$.
+we will denote the equivalence classes in $\sfrac{X + Y}{\sim R}$ as $[x]$ and $[y]$ depending on their pre-image being either in $X$ or $Y$.
+Before we procced, we need to show that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are homomorphisms.
+Let $e_{x}$ be an edge in $X$ and $v_x$ is its source. We need to check that $\sfrac{i_1}{\sim R}(v_x) = (\sfrac{i_1}{\sim R};\sfrac{s_{X}}{\sim R})(e_x)$.
+$\sfrac{i_1}{\sim R}(v_x) = [v_x]$ and $(\sfrac{i_1}{\sim R};\sfrac{s_{X}}{\sim R})(e_x) = \sfrac{s_{X}}{\sim R}([e_x]) = [v_x]$ and hence $\sfrac{i_1}{\sim R}$ is a homomorphism, simialrly $\sfrac{i_2}{\sim R}$ is.
+Now, the square obviously commutes.
+To show that $\sfrac{X + Y}{\sim R}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1$ and $j_2$ such that the above square commutes there is a unique $u$ such that $j_1 = \sfrac{i_1}{\sim R} ; u$ and $j_2 = \sfrac{i_2}{\sim R} ; u$.
+We will define such $u$ as $u([x]) = j_1(x)$ and $u([y]) = j_2(y)$ and claim that is it a well-defined function by showing that if $t_1 = t_2$ then $u(t_1) = u(t_2)$.
+There are three options
+\begin{itemize}
+    \item $t_1 = [x]$ and $t_2 = [x]$, where $[x]$ is a singleton equivalence class.
+          The property then holds on the nose.
+    \item $t_1 = [y]$ and $t_2 = [y]$, where $[y]$ is a singleton equivalence class.
+    The property then holds on the nose.
+    \item $t_1 = [x]$ and $t_2 = [y]$ and there exists at least one $z$ in $Z$ such that $x = f(z)$ and $y = g(z)$.
+          By applying $j_1$ to the first we get $j_1(f(z)) = j_1(x)$ and similarly for $j_2$: $j_2(g(z)) = j_2(y)$ and $j_1(x) = j_2(y)$ by commutativity.
+          Hence, $u([x]) = u([y])$. 
+\end{itemize}
+
+Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.
+Finally, $u$ is unique.
+Assume, that there exists $w : \sfrac{X+Y}{\sim R} \to Q$ for the same $j_1$ and $j_2$.
+In order for them to commute $w([x]) = j_1(x)$ and $w([y]) = j_2(y)$ which is the same as $u$.
+That is, $j_1$ and $j_2$ determine $u$ uniquely.
+
+\end{remark}
+
 Now we are ready to define e-hypergraphs which are a combinatorial representation of enriched PROPs.
@@
     Next, each $x$ has at most one immediate predecessor.
+    \update{We will denote the fact that $e$ is an immediate predecessor of $v$ as $e <_c^{\mu} v$.}
+    % , which essentially makes $<_{c}$ into a function that for any given $x$ returns its immediate predecessor.
     Such order was called a \textbf{well-founded arborescent} strict partial order in~\cite{orders_and_events}.
@@
 \begin{definition}{E-hypergraphs homomorphism}
-
+\label{def:e-homo}    
 A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
@@
 \begin{remark}
-    Similarly to $\catname{Hyp_{\Sigma}}$ pushouts in $\catname{EHyp_{\Sigma}}$ are computed on the underlying sets of nodes and edges.
-    However, unlike $\catname{Hyp_{\Sigma}}$ $\catname{EHyp_{\Sigma}}$ is not \textit{adhesive} meaning that not all pushouts along monos exist.
+
+    We will compute pushouts in $\catname{EHyp_{\Sigma}}$ similarly to $\catname{EHyp_{\Sigma}}$.
+    Recall the diagram below.
+    \[
+    \adjustbox{scale=1}{
+        \begin{tikzcd}
+        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{i_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
+        Y \arrow[r, "\sfrac{i_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+                                                                        &                                                                  & Q
+        \end{tikzcd}
+    }
+    \]
+
+    The pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
+    First, a coproduct of $X+Y$ is computed which is 
+    \[
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X} + s_{Y}, t_{X} + t_{Y}, \#_{X} + \#_{Y}, <_{c,X} + <_{c,Y} \}
+    \] 
+    once again omitting the labelling function.
+    And consider a relation $S = \{(x,y), (y,x) \in (X+Y) \times (X+Y) \; | \;\exists z \in Z \; . \; f(z) = x \text{ and } g(z) = y \}$ and let $R$ be its reflexive and transitive closure.
+    We then quotient $X + Y$ by $R$ and claim that it is a pushout.
+    \[
+        \sfrac{X + Y}{\sim R} = \{\sfrac{V_{X} + V_{Y}}{\sim R}, \sfrac{E_{X} + E_{Y}}{\sim R}, \sfrac{s_{X} + s_{Y}}{\sim R}, \sfrac{t_{X} + t_{Y}}{\sim R}, \sfrac{\#_{X} + \#_{Y}}{\sim R}, \sfrac{<_{c,X} + <_{c,Y}}{\sim R} \}
+    \]
+    where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$.
+    \[
+    \sfrac{(<_{c,X} + <_{c,Y})}{\sim R} \subseteq \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(E_{X} + E_{Y})}{\sim R} \to \sfrac{(E_{X} + E_{Y})}{\sim R}
+    \]
+    \question{Below $e$ and $v$ do not necessarily mean edges and nodes, $v$ can be an edge as well.}
+    \question{I also kind of intuitively inventing the quotient of a relation, is there any accepted notion for this?}
+    and, whenever $e_{X} <_{c,X}^{\mu} v_{X}$ and $e_{Y} <_{c,Y}^{\mu} v_{Y}$ and $v_{x} R v_{y}$ then $e_{X} R e_{Y}$.
+    The latter condition means two things
+    \begin{itemize}
+        \label{cond:essential_for_pushout}
+        \item if $[z) = \varnothing$ then either $[f(z)) = \varnothing$ or $[g(z)) = \varnothing$.
+              \question{This point formally defines what it means to 'inherit' a relation.
+                        Let's say $v_1 = f(z)$ and $v_2 = g(z)$ and $e_1 <_{c,X}^{\mu} v_1$ and $[v_2) = \varnothing$,
+                        then, having $e_1 <_{c} [v_1,v_2]$ is fine as it does not break the condition of a predecessor being unique.}
+              \update{This point also says when we can compute a pushout: either non of X and Y should have a parent for $f(z)$ and $g(z)$ or only one of them should.}
+        \item if $[z) \not = \varnothing$ then there exists $z'$ such that $z' <_{c}^{\mu} z$ and $[g(z'), f(z')] <_{c}^{\mu} [g(z), f(z)]$
+    \end{itemize}
+    Then,
+    \[
+        \sfrac{\conflict_{X} + \conflict_{Y}}{\sim R} \subseteq \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R}
+    \]
+    meaning that whenever $e_1 \conflict e_2$ and $e_1' \conflict e_2'$ and $e_1 R e_1'$ then $[e_1, e_1'] \conflict e_2$ and $[e_1, e_1'] \conflict e_2'$.
+    $\sfrac{i_1}{\sim R}$ then maps $x \to [x]$ and $\sfrac{i_2}{\sim R}$ maps $y \to [y]$.
+    Before we proceed we need to check that $\sfrac{(X + Y)}{\sim R}$ is indeed an e-hypergraph.
+    Since $X + Y$ is clearly an e-hypergraph, we need to check whether identification breaks any of the requirements (1-7) above.
+    \begin{enumerate}
+        \item Assume $[x_1,y_1] <_{c} [x_2,y_2]$ and $[x_1, y_1] \conflict [x_3]$ which means that $x_1 \conflict x_3$.
+              Hence $x_2 \conflict x_3$ (because $X$ is an e-hypergraph) and $[x_2,y_2] \conflict [x_3]$. Similarly, if $[x_1, y_1] \conflict [y_3]$ and $y_1 \conflict y_3$.
+        \item $\sfrac{\conflict}{\sim R}$ and $\sfrac{<_{c}}{\sim R}$ are disjoint because we quotient only inside the relations.
+        \item Again, we only quotient $\conflict$ on $x$ and $y$ for which original relation is defined.
+        \item This follows from $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ being homomorphisms.
+        \item We just perform transitive closure of the complement.
+\update{Maybe we can get rid of this point and instead later just take transitive closure of the complement when talking about interfaces}
+        \item Assume $[x_1, y_1] \conflict [x_2,y_2]$ and $[x_2,y_2] \conflict [x_3,y_3]$ and all of them are distinct.
+              Then $[x_1,y_1] \conflict [x_3,y_3]$ because either $x_1 \conflict x_3$ or $y_1 \conflict y_3$ (otherwise they would not have been merged into the same equivalence class).
+        \item We never merge $x_1$ and $x_2$ such that $x_1 <_{c} x_2$ into the same equivalence class because morphisms should preserve the order.
+    \end{enumerate}
+
+    Then we need to check that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are indeed homomorphisms.
+    For these we need to check that conditions (2-3) in~\ref{def:e-homo} are satisfied.
+    Let's check for $\sfrac{i_1}{\sim R}$.
+    We identify elements from $X$ with elements from $Y$ which does not affect (2) and the same reasoning applies to (3).
+    (1) was proved earlier for $\catname{Hyp_{\Sigma}}$.
+
+    Finally, by following essentially the same argument as for pushout in $\catname{Hyp_{\Sigma}}$ we can conclude that the structure defined above is indeed a pushout in $\catname{EHyp_{\Sigma}}$ and we know precise conditions under which this concrete pushout exists.
+\end{remark}
+
+\begin{example}
     Consider an example span below in Figure~\ref{fig:pushout_fail}.
@@
 
-    In order to build a commuting square which is a pushout the immediate predecessors of $f$ should be identified in the pushout because any element of an e-hypergraph has only one immediate predecessor.
-    However, such identification is not possible as these predecessors have different types.
+    The pushout defined above does not exist for this span because neither of the set of predecessors of the images of $f$ are $\varnothing$ which violates~\ref{cond:essential_for_pushout}.
     Luckily, we do not need arbitrary pushouts and pushouts that we actually need do exist.
-    \update{In general, in the pushout $A \xrightarrow{f} D \xleftarrow{g} B$, $D$ would inherit the relations from $A$ and $B$ when the relations are not mutually exclusive.}
-\end{remark}
+\end{example}
 
@@
     In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$.
-    The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
-    Note that pushout for such composition always exists as the nodes that are being identified are unrelated: they are not in conflict with anything and the sets of their predecessors are empty.
+    % The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
+    \update{Note that pushout for such composition always exists as the interface in the middle is descrete (meaning nodes in it have no predecessors) and nodes in the images of the middle interface are at the outermost level as well.
+    Thus the span satisfies~\ref{cond:essential_for_pushout}.
+    }
 \end{remark}

### 2023-10-26 sections/combinatorial_semantics.tex
Pushout for Hyp rewritten

@@
 Pushout of two hypergraphs $X = \{V_{X}, E_{X}, s_{X}, t_{X}\}$ and $Y = \{V_{Y}, E_{Y}, s_{Y}, t_{Y}\}$ (we will omit labelling functions here since they are irrelevant) along $Z$ is computed in two steps.
-First, a coproduct of $X$ and $Y$ is taken which is $X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X} + s_{Y}, t_{X} + t_{Y} \}$.
-Then, consider a relation 
-\[
-    S = \{(x_i,y_j), (y_j, x_i) \in (X+Y) \times (X+Y)\; | \; \exists z \in Z \; . \; x_i = f(z) \text{ and } y_j = g(z) \text{where $x_i \in X$ and $y_j \in Y$ }\}
-% \cup \{(x,x) \in (X+Y) \times (X+Y) \text{ where $x \in X$}\} \cup \{(y,y) \in (X+Y) \times (X+Y) \text{ where $y \in Y$}\}$ 
-\]
-and let relation $R$ be its reflexive and transitive closure. $R$ is then an equivalence relation since $S$ is symmetric.
-
-A quotient of $X+Y$ by this relation denoted by 
-\[
-    \sfrac{X+Y}{\sim R} = \{\sfrac{V_{X} + V_{Y}}{\sim R}, \sfrac{E_{X} + E_{Y}}{\sim R}, \sfrac{s_{X} + s_{Y}}{\sim R}, \sfrac{t_{X} + t_{Y}}{\sim R}\}
-\]
-where $\sfrac{s_{X} + s_{Y}}{\sim R} : \sfrac{E_{X} + E_{Y}}{\sim R} \to \sfrac{V_{X} + V_{Y}}{\sim R}$ now maps equivalence classes to equivalence classes.
-Note that if $s : e_1 \mapsto \{v_1,v_2\}$ and $s : e_2 \mapsto \{v_3,v_4\}$  and $e_1 \sim e_2$ then $\sfrac{s}{\sim R} : [e_1,e_2] \to \{[v_1,v_3],[v_2,v_4]\}$ because $f$ and $g$ are homomorphisms.
-Then, $\sfrac{i_1}{\sim R}$ maps $y \in Y$ to its equivalence class and likewise for $\sfrac{i_2}{\sim R}$.
-we will denote the equivalence classes in $\sfrac{X + Y}{\sim R}$ as $[x]$ and $[y]$ depending on their pre-image being either in $X$ or $Y$.
-Before we procced, we need to show that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are homomorphisms.
-Let $e_{x}$ be an edge in $X$ and $v_x$ is its source. We need to check that $\sfrac{i_1}{\sim R}(v_x) = (\sfrac{i_1}{\sim R};\sfrac{s_{X}}{\sim R})(e_x)$.
-$\sfrac{i_1}{\sim R}(v_x) = [v_x]$ and $(\sfrac{i_1}{\sim R};\sfrac{s_{X}}{\sim R})(e_x) = \sfrac{s_{X}}{\sim R}([e_x]) = [v_x]$ and hence $\sfrac{i_1}{\sim R}$ is a homomorphism, simialrly $\sfrac{i_2}{\sim R}$ is.
+First, a coproduct of $X$ and $Y$ is taken which is 
+\[
+    X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X + Y}, t_{X+Y}\}
+\]
+where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $\langle s'_{X}, s'_{Y} \rangle : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};i_{1,V}^{*}$ and $s'_{Y} = s_{Y};i_{2,V}^{*}$ where $i_1,i_2$ are corresponding coproduct injections, similarly for $t_{X+Y}$.
+Recall that each arrow in $\catname{Hyp_{\Sigma}}$ is a pair of arrows for edges and nodes and we will further use subscripts $V$ and $E$ when referring to these functions.
+Then, consider relations 
+\[
+    S_{V} = \{(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \; \exists z \in V_{Z} \; .
+\; x_i = f_{V};i_{1,V}(z) \text{ and } y_j = g_{V};i_{2,V}(z) \text{where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\}
+\]
+\[
+    S_{E} = \{(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; | \; \exists z \in E_{Z} \; .
+\; x_i = f_{E};i_{1,E}(z) \text{ and } y_j = g_{E};i_{2,E}(z) \text{where $x_i \in E_{X}$ and $y_j \in E_{Y}$ }\}
+\]
+and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.
+
+\begin{lemma}
+    $x R y$ implies that either $x = y$ or there exists a sequence $(a_1,\ldots,a_{n-1},{a_n})$ such that $x = a_1$ and $y = a_{n}$ such that $a_k S a_{k+1}$ or $a_{k + 1} S a_{k}$ for $k < n$.
+\end{lemma}
+
+A quotient of $X+Y$ by these relation is 
+\[
+    \sfrac{X+Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X+Y}}{\sim (R_{V},R_{E})}\}
+\]
+where $\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})} : \sfrac{E_{X} + E_{Y}}{\sim R_{E}} \to (\sfrac{V_{X} + V_{Y}}{\sim R_{V}})^{*}$.
+In particular, there is an obvious surjective function $s_1 : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim R_{V}})$ and $s_1^{*}$ is its extension to sequences. 
+Let $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim R_{V}})^{*} = s_{X+Y};s_1^{*}$ and then $\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}([e]) = s_{X+Y};s_1^{*}(e)$ where $[e]$ denotes the equivalence class of $e$ in $E_{X} + E_{Y}$.
+We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has in image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
+We need to check that if $t_1 = t_2$ then $\sfrac{s_{X+Y}}{\sim (R_{V,E})}(t_1) = \sfrac{s_{X+Y}}{\sim (R_{V,E})}(t_2)$. 
+There are three options
+\begin{enumerate}
+    \item $t_1 = [e_{y}^{1}]$ and $t_2 = [e_{y}^{2}]$
+    \item $t_1 = [e_{x}^{1}]$ and $t_2 = [e_{x}^{2}]$
+    \item $t_1 = [e_{x}]$ and $t_2 = [e_{y}]$ 
+\end{enumerate}
+We will prove the last point (all the others are analogous).
+\begin{proof}
+    According to the lemma above $[e_{x}] = [e_{y}]$ gives us two options.
+    \begin{enumerate}
+        \item Either $e_{x} = e_{y}$
+        \item or there is a sequence $w = (a_1, \ldots, a_{n})$.
+    \end{enumerate}
+    In the first case the equality holds on the nose.
+    We then prove the second case by induction on the length of $w$.
+    \begin{itemize}
+        \item In the case $|w| = 2$, $e_{x} S e_{y}$ which means that there exists $z$ in $E_{Z}$ such that $f_{E};i_{1,E}(z) = e_{x}$ and $g_{E};i_{2,E}(z) = e_{y}$.
+              Recall that because $f$ and $g$ are homomorphisms, the following equalities hold
+              \[
+                s_{Z};f_{V}^{*};i_{1,V}^{*}(z) = f_{E};i_{1,E};s_{X+Y}(z) = s_{X+Y}(e_{x})
+              \]
+              \[
+                s_{Z};g_{V}^{*};i_{2,V}^{*}(z) = g_{E};i_{2,E};s_{X+Y}(z) = s_{X+Y}(e_{y})
+              \]
+              Then,
+              \[
+                \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}([e_{x}]) = s_{X+Y};s_{1}^{*}(e_{x}) = s_{Z};f_{V}^{*};i_{1,V}^{*};s_{1}^{*}(z)
+              \]
+              and
+              \[
+                \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}([e_{y}]) = s_{X+Y};s_{1}^{*}(e_{y}) = s_{Z};g_{V}^{*};i_{2,V}^{*};s_{1}^{*}(z)
+              \]
+              Then, by noting that $f_{V};i_{1,V}(z) = g_{V};i_{2,V}(z)$ for all $z$ in $V_{Z}$ (because $\sfrac{V_{X} + V_{Y}}{\sim R_{V}}$ is so defined) we have $s_{X+Y};s_1^{*}(e_{x}) = s_{X+Y};s_2^{*}(e_{y})$.
+        \item Suppose if $e_{x} S e_{y}$ by a chain of length $n$ or less then $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(e_{y})$.
+        \item Now suppose $e_{x} S e_{y}$ by a chain $w = (a_1, \ldots, a_{n+1})$ and a subchain $(a_1, \ldots, a_{n})$ satisfies the inductive hypothesis and hence $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(a_{n})$.
+              Note that $S_{E}, S_{V}$ relate only elements with the pre-image in $E_{X}$ (respectively, $V_{X}$) with elements with the pre-image in $E_{Y}$ (respectively, $V_{Y}$). 
+              Elements with the pre-image in the same set are not related.
+              Then we have $a_n S a_{n+1} = e_{y}$ and need to show that $s_{X+Y};s_1^{*}(a_{n+1}) = s_{X+Y};s_{1}^{*}(e_{y})$.
+              This follows by exactly the same argument as the base case as $a_{n+1}$ is necessarily in $E_{X}$.
+              And we finally have $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(a_{n+1}) = s_{X+Y};s_{1}^{*}(e_{y})$.
+    \end{itemize}
+    \update{Looks fine?}
+\end{proof}
+
+Next $\sfrac{i_{1,V}}{\sim R_{V}} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}} = i_{1,V};s'$ where $s' : V_{X}+V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}}$.
+and 
+$\sfrac{i_{1,E}}{\sim R_{E}} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim R_{E}} = i_{1,E};s'$, similarly for $\sfrac{i_{2,E}}{\sim R_{E}}$ and $\sfrac{i_{2,V}}{\sim R_{V}}$.
+We need to show
+\begin{enumerate}
+    \item $s_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = \sfrac{i_{1,E}}{\sim R_{E}};\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}(e)$
+    \item $t_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = \sfrac{i_{1,E}}{\sim R_{E}};\sfrac{t_{X+Y}}{\sim (R_{V},R_{E})}(e)$
+\end{enumerate}
+and likewise for $i_{2}$.
+Let's check the first one.
+$s_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = s_{X};i_{1,V}^{*};s'^{*}(e)$ which is a homomorphism because $i_{1,V}$ and \update{$s'$ are}.
+
+% Note that if $s : e_1 \mapsto \{v_1,v_2\}$ and $s : e_2 \mapsto \{v_3,v_4\}$  and $e_1 \sim e_2$ then $\sfrac{s}{\sim R} : [e_1,e_2] \to \{[v_1,v_3],[v_2,v_4]\}$ because $f$ and $g$ are homomorphisms.
+% Then, $\sfrac{i_1}{\sim R}$ maps $y \in Y$ to its equivalence class and likewise for $\sfrac{i_2}{\sim R}$.
+% we will denote the equivalence classes in $\sfrac{X + Y}{\sim R}$ as $[x]$ and $[y]$ depending on their pre-image being either in $X$ or $Y$.
+% Before we procced, we need to show that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are homomorphisms.
+% Let $e_{x}$ be an edge in $X$ and $v_x$ is its source. We need to check that $\sfrac{i_1}{\sim R}(v_x) = (\sfrac{i_1}{\sim R};\sfrac{s_{X}}{\sim R})(e_x)$.
+% $\sfrac{i_1}{\sim R}(v_x) = [v_x]$ and $(\sfrac{i_1}{\sim R};\sfrac{s_{X}}{\sim R})(e_x) = \sfrac{s_{X}}{\sim R}([e_x]) = [v_x]$ and hence $\sfrac{i_1}{\sim R}$ is a homomorphism, simialrly $\sfrac{i_2}{\sim R}$ is.
 Now, the square obviously commutes.
-To show that $\sfrac{X + Y}{\sim R}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1$ and $j_2$ such that the above square commutes there is a unique $u$ such that $j_1 = \sfrac{i_1}{\sim R} ; u$ and $j_2 = \sfrac{i_2}{\sim R} ; u$.
-We will define such $u$ as $u([x]) = j_1(x)$ and $u([y]) = j_2(y)$ and claim that is it a well-defined function by showing that if $t_1 = t_2$ then $u(t_1) = u(t_2)$.
-There are three options
-\begin{itemize}
-    \item $t_1 = [x]$ and $t_2 = [x]$, where $[x]$ is a singleton equivalence class.
-          The property then holds on the nose.
-    \item $t_1 = [y]$ and $t_2 = [y]$, where $[y]$ is a singleton equivalence class.
-    The property then holds on the nose.
-    \item $t_1 = [x]$ and $t_2 = [y]$ and there exists at least one $z$ in $Z$ such that $x = f(z)$ and $y = g(z)$.
-          By applying $j_1$ to the first we get $j_1(f(z)) = j_1(x)$ and similarly for $j_2$: $j_2(g(z)) = j_2(y)$ and $j_1(x) = j_2(y)$ by commutativity.
-          Hence, $u([x]) = u([y])$. 
-\end{itemize}
+To show that $\sfrac{X + Y}{\sim (R_{V},R_{E})}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1 = (j_{1,V}, j_{1,E})$ and $j_2 = (j_{2,V}, j_{2,E})$ such that the above square commutes there is a unique $u = (u_{V}, u_{E})$ such that $j_1 = \sfrac{i_{1}}{\sim (R_{V},R_{E})} ; u$ and $j_2 = \sfrac{i_2}{\sim (R_{V},R_{E})} ; u$.
+We will define such $u$ as $u_{E}([e_{x}]) = j_{1,E}(e_{x})$ and $u_{E}([e_{y}]) = j_{2,E}(e_y)$ (similarly for $u_{V}$) and claim that it is a well-defined function by showing that if $t_1 = t_2$ then $u_{E}(t_1) = u_{E}(t_2)$.
+There are three options analogoues to the proof above, so we will show the base case for $t_1 = [e_{x}]$ and $t_2 = [e_{y}]$.
+$t_1 = [e_x]$ and $t_2 = [e_y]$ and $t_1 = t_2$ implies that there exists at least one $z$ in $E_{Z}$ such that $e_{x} = f_{E}(z)$ and $[e_{x}] = \sfrac{i_{1,V}}{\sim R_{V}}(e_x)$, and $e_y = g_{E}(z)$ $[e_{y}] = \sfrac{i_{2,V}}{\sim R_{V}}(e_y)$.
+By applying $j_{1,E}$ to the first we get $j_{1,E};f_{E}(z) = j_{1,E}(e_x)$ and similarly for $j_{2,E}$: $j_{2,E};g_{E}(z) = j_{2,E}(e_y)$ and $j_{1,E}(e_{x}) = j_{2,E}(e_{y})$ by commutativity.
+Hence, $u_{E}([e_x]) = u_{E}([e_y])$.
+\update{Above $e_x$ and $e_y$ are different from $e_x$ and $e_y$ from the first proof. This is probablty not good.}
 
@@
     \[
-    \sfrac{(<_{c,X} + <_{c,Y})}{\sim R} \subseteq \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(E_{X} + E_{Y})}{\sim R} \to \sfrac{(E_{X} + E_{Y})}{\sim R}
-    \]
+    \sfrac{(<_{c,X+Y})}{\sim (R_{V},R_{E})} \subseteq \sfrac{(V_{X} + V_{Y})}{\sim R_{V}} + \sfrac{(E_{X} + E_{Y})}{\sim R_{E}} \times \sfrac{(E_{X} + E_{Y})}{\sim R_{E}}
+    \]
+    \update{We shall probably first define the quotient of the immediate parent because it is essentially a function and it will follow the same argument as the source function}
     \question{Below $e$ and $v$ do not necessarily mean edges and nodes, $v$ can be an edge as well.}
@@
 \[
-X_{ext} \xrightarrow{f_{ext}} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext}} Z_{ext}
-\]
-
-since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
+X_{ext} \xrightarrow{f_{ext};i_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};i_1} Z_{ext}
+\]
+where $i_1$ is a coproduct injection.
+Since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
 $h = a;p_1 + g|_{Y' \setminus Y_{ext}};p_2$, $k = b;p_2 + f|_{Y \setminus Y_{ext}};p_1$.
-\question{Is this a standard syntax for a narrowed function?}
 
@@
 % and then the disjoint union is just a coproduct. 
-Defining cospans up to isomorphism, associativity of coproducts and pushouts, and the presence of an id-cospan turns this construction into a valid category.
+Defining cospans up to isomorphism, associativity of coproducts and pushouts, and the presence of an id-cospan turns this construction into a proper category.
 
 Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC.
-More precisely, below are cospans for tensor product and symmetry. Tensor unit is given by the unit in $\catname{EHyp_{\Sigma}}$.
+More precisely, below are cospans for tensor product and symmetry.
+Tensor unit is given by the initial object in $\catname{EHyp_{\Sigma}}$.
 
@@
 	{X_{ext}} & X & A & Y & {Y_{ext}} \\
-	&& \otimes &&& {:=} & {X_{ext}+ext(X')} && {X+X'} && {A+B} && {Y+Y'} && {Y_{ext}+Y'_{ext}} \\
-	{ext(X')} & {X'} & B & {Y'} & {Y'_{ext}}
+	&& \otimes &&& {:=} & {X_{ext}+X'_{ext}} && {X+X'} && {A+B} && {Y+Y'} && {Y_{ext}+Y'_{ext}} \\
+	{X'_{ext}} & {X'} & B & {Y'} & {Y'_{ext}}
 	\arrow["{f_{ext}}", from=1-1, to=1-2]
@@
     New interfaces are shown in Figure~\ref{fig:e-hyp_composition_example_3}.
-    They are computed as shown above: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken.
+    They are computed as shown: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken.
     Similarly for the output interface.

### 2023-10-27 sections/combinatorial_semantics.tex
Puhsout in Hyp proof completed

@@
 
-Next $\sfrac{i_{1,V}}{\sim R_{V}} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}} = i_{1,V};s'$ where $s' : V_{X}+V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}}$.
+Next $\sfrac{i_{1,V}}{\sim R_{V}} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}} = i_{1,V};s_1$.
 and 
-$\sfrac{i_{1,E}}{\sim R_{E}} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim R_{E}} = i_{1,E};s'$, similarly for $\sfrac{i_{2,E}}{\sim R_{E}}$ and $\sfrac{i_{2,V}}{\sim R_{V}}$.
+$\sfrac{i_{1,E}}{\sim R_{E}} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim R_{E}} = i_{1,E};s_{E}'$, similarly for $\sfrac{i_{2,E}}{\sim R_{E}}$ and $\sfrac{i_{2,V}}{\sim R_{V}}$.
 We need to show
@@
 and likewise for $i_{2}$.
-Let's check the first one.
-$s_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = s_{X};i_{1,V}^{*};s'^{*}(e)$ which is a homomorphism because $i_{1,V}$ and \update{$s'$ are}.
+Let's check the first one for which we first need to check that $(s'_{E},s_{1})$ is a homomorphisms, i.e.
+\[
+s_{X+Y};s_1^{*}(e) = s'_{E};\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}(e) = \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}(s'_{E}(e)) = s_{X+Y};s_1^{*}(e)
+\]
+hence, $(\sfrac{i_{1,V}}{\sim R_{V}}, \sfrac{i_{1,E}}{\sim R_{V}})$ is a homomorphism because $(i_{1,V},i_{1,E})$ and $(s_1,s'_{E})$ are.
 

### 2023-10-27 sections/combinatorial_semantics.tex
Pushout for Hyp brushed up

--- 2023-10-26 sections/combinatorial_semantics.tex
@@
 \begin{lemma}
-    $x R y$ implies that either $x = y$ or there exists a sequence $(a_1,\ldots,a_{n-1},{a_n})$ such that $x = a_1$ and $y = a_{n}$ such that $a_k S a_{k+1}$ or $a_{k + 1} S a_{k}$ for $k < n$.
+    $x R y$ if and only if either $x = y$ or there exists a sequence $(a_1,\ldots,a_{n-1},{a_n})$ such that $x = a_1$ and $y = a_{n}$ such that $a_k S a_{k+1}$ or $a_{k + 1} S a_{k}$ for $k < n$.
 \end{lemma}
@@
 \]
-where $\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})} : \sfrac{E_{X} + E_{Y}}{\sim R_{E}} \to (\sfrac{V_{X} + V_{Y}}{\sim R_{V}})^{*}$.
-In particular, there is an obvious surjective function $s_1 : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim R_{V}})$ and $s_1^{*}$ is its extension to sequences. 
-Let $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim R_{V}})^{*} = s_{X+Y};s_1^{*}$ and then $\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}([e]) = s_{X+Y};s_1^{*}(e)$ where $[e]$ denotes the equivalence class of $e$ in $E_{X} + E_{Y}$.
+We will then refer to $\sim (R_{V},R_{E})$ just as $\sim$, i.e. writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and the exact relation will be clear from the context.
+We have 
+\[
+    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
+\]
+In particular, there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
+Let $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*} = s_{X+Y};[-]^{*}$ and then $\sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}$.
+There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
 We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has in image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-We need to check that if $t_1 = t_2$ then $\sfrac{s_{X+Y}}{\sim (R_{V,E})}(t_1) = \sfrac{s_{X+Y}}{\sim (R_{V,E})}(t_2)$. 
+We need to check that if $t_1 \sim t_2$ then $\sfrac{s_{X+Y}}{\sim}([t_1]) = \sfrac{s_{X+Y}}{\sim}([t_2])$. 
 There are three options
 \begin{enumerate}
-    \item $t_1 = [e_{y}^{1}]$ and $t_2 = [e_{y}^{2}]$
-    \item $t_1 = [e_{x}^{1}]$ and $t_2 = [e_{x}^{2}]$
-    \item $t_1 = [e_{x}]$ and $t_2 = [e_{y}]$ 
+    \item $t_1 = e_{y}^{1}$ and $t_2 = e_{y}^{2}$
+    \item $t_1 = e_{x}^{1}$ and $t_2 = e_{x}^{2}$
+    \item $t_1 = e_{x}$ and $t_2 = e_{y}$ 
 \end{enumerate}
@@
 \begin{proof}
-    According to the lemma above $[e_{x}] = [e_{y}]$ gives us two options.
+    According to the lemma above $e_{x} \sim e_{y}$ gives us two options.
     \begin{enumerate}
         \item Either $e_{x} = e_{y}$
-        \item or there is a sequence $w = (a_1, \ldots, a_{n})$.
+        \item or there is a sequence $w = (a_1, \ldots, a_{n})$ such that $a_1 = e_{x}$ and $a_{n} = e_{y}$.
     \end{enumerate}
@@
     \begin{itemize}
-        \item In the case $|w| = 2$, $e_{x} S e_{y}$ which means that there exists $z$ in $E_{Z}$ such that $f_{E};i_{1,E}(z) = e_{x}$ and $g_{E};i_{2,E}(z) = e_{y}$.
+        \item In the case $|w| = 2$, $e_{x} S e_{y}$ (or $e_{y} S e_{x}$) which means that there exists $z$ in $E_{Z}$ such that $f_{E};i_{1,E}(z) = e_{x}$ and $g_{E};i_{2,E}(z) = e_{y}$.
               Recall that because $f$ and $g$ are homomorphisms, the following equalities hold
@@
               \[
-                \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}([e_{x}]) = s_{X+Y};s_{1}^{*}(e_{x}) = s_{Z};f_{V}^{*};i_{1,V}^{*};s_{1}^{*}(z)
+                \sfrac{s_{X+Y}}{\sim}([e_{x}]) = [s_{X+Y}(e_{x})]^{*} = [s_{Z};f_{V}^{*};i_{1,V}^{*}(z)]^{*}
               \]
@@
               \[
-                \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}([e_{y}]) = s_{X+Y};s_{1}^{*}(e_{y}) = s_{Z};g_{V}^{*};i_{2,V}^{*};s_{1}^{*}(z)
-              \]
-              Then, by noting that $f_{V};i_{1,V}(z) = g_{V};i_{2,V}(z)$ for all $z$ in $V_{Z}$ (because $\sfrac{V_{X} + V_{Y}}{\sim R_{V}}$ is so defined) we have $s_{X+Y};s_1^{*}(e_{x}) = s_{X+Y};s_2^{*}(e_{y})$.
-        \item Suppose if $e_{x} S e_{y}$ by a chain of length $n$ or less then $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(e_{y})$.
+                \sfrac{s_{X+Y}}{\sim}([e_{y}]) = [s_{X+Y}(e_{y})]^{*} = [s_{Z};g_{V}^{*};i_{2,V}^{*}(z)]^{*}
+              \]
+              Then, by noting that $f_{V};i_{1,V}(z) = g_{V};i_{2,V}(z)$ for all $z$ in $V_{Z}$ (because $\sfrac{V_{X} + V_{Y}}{\sim}$ is so defined) we have $[s_{X+Y};(e_{x})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
+        \item Suppose if $e_{x} S e_{y}$ by a chain of length $n$ or less then $[s_{X+Y}(e_{x})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
         \item Now suppose $e_{x} S e_{y}$ by a chain $w = (a_1, \ldots, a_{n+1})$ and a subchain $(a_1, \ldots, a_{n})$ satisfies the inductive hypothesis and hence $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(a_{n})$.
@@
               Elements with the pre-image in the same set are not related.
-              Then we have $a_n S a_{n+1} = e_{y}$ and need to show that $s_{X+Y};s_1^{*}(a_{n+1}) = s_{X+Y};s_{1}^{*}(e_{y})$.
+              Then we have $a_n S a_{n+1} = e_{y}$ and need to show that $[s_{X+Y}(a_{n+1})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
               This follows by exactly the same argument as the base case as $a_{n+1}$ is necessarily in $E_{X}$.
-              And we finally have $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(a_{n+1}) = s_{X+Y};s_{1}^{*}(e_{y})$.
+              And we finally have $[s_{X+Y}(e_{x})]^{*} = [s_{X+Y}(a_{n+1})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
     \end{itemize}
@@
 
-Next $\sfrac{i_{1,V}}{\sim R_{V}} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}} = i_{1,V};s'$ where $s' : V_{X}+V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim R_{V}}$.
+Next 
+\[
+    \sfrac{i_{1,V}}{\sim} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim} = [i_{1,V}]
+\]
 and
-$\sfrac{i_{1,E}}{\sim R_{E}} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim R_{E}} = i_{1,E};s'$, similarly for $\sfrac{i_{2,E}}{\sim R_{E}}$ and $\sfrac{i_{2,V}}{\sim R_{V}}$.
+\[
+\sfrac{i_{1,E}}{\sim} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim} = [i_{1,E}]
+\]
+similarly for $\sfrac{i_{2,E}}{\sim}$ and $\sfrac{i_{2,V}}{\sim}$
 We need to show
 \begin{enumerate}
-    \item $s_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = \sfrac{i_{1,E}}{\sim R_{E}};\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}(e)$
-    \item $t_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = \sfrac{i_{1,E}}{\sim R_{E}};\sfrac{t_{X+Y}}{\sim (R_{V},R_{E})}(e)$
+    \item $s_{X};(\sfrac{i_{1,V}}{\sim})^{*}(e) = \sfrac{i_{1,E}}{\sim};\sfrac{s_{X+Y}}{\sim}(e)$
+    \item $t_{X};(\sfrac{i_{1,V}}{\sim})^{*}(e) = \sfrac{i_{1,E}}{\sim};\sfrac{t_{X+Y}}{\sim}(e)$
 \end{enumerate}
 and likewise for $i_{2}$.
-Let's check the first one.
-$s_{X};(\sfrac{i_{1,V}}{\sim R_{V}})^{*}(e) = s_{X};i_{1,V}^{*};s'^{*}(e)$ which is a homomorphism because $i_{1,V}$ and \update{$s'$ are}.
+Let's check the first one for which we first need to check that $([-]_{E},[-]_{V})$ is a homomorphisms, i.e.
+\[
+[s_{X+Y}(e)]^{*} = \sfrac{s_{X+Y}}{\sim}([e]) = [s_{X+Y}(e)]^{*}
+\]
+hence, $(\sfrac{i_{1,V}}{\sim}, \sfrac{i_{1,E}}{\sim})$ is a homomorphism because $(i_{1,V},i_{1,E})$ and $([-]_{E},[-]_{V})$ are.
 
@@
 Now, the square obviously commutes.
-To show that $\sfrac{X + Y}{\sim (R_{V},R_{E})}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1 = (j_{1,V}, j_{1,E})$ and $j_2 = (j_{2,V}, j_{2,E})$ such that the above square commutes there is a unique $u = (u_{V}, u_{E})$ such that $j_1 = \sfrac{i_{1}}{\sim (R_{V},R_{E})} ; u$ and $j_2 = \sfrac{i_2}{\sim (R_{V},R_{E})} ; u$.
-We will define such $u$ as $u_{E}([e_{x}]) = j_{1,E}(e_{x})$ and $u_{E}([e_{y}]) = j_{2,E}(e_y)$ (similarly for $u_{V}$) and claim that it is a well-defined function by showing that if $t_1 = t_2$ then $u_{E}(t_1) = u_{E}(t_2)$.
-There are three options analogoues to the proof above, so we will show the base case for $t_1 = [e_{x}]$ and $t_2 = [e_{y}]$.
-$t_1 = [e_x]$ and $t_2 = [e_y]$ and $t_1 = t_2$ implies that there exists at least one $z$ in $E_{Z}$ such that $e_{x} = f_{E}(z)$ and $[e_{x}] = \sfrac{i_{1,V}}{\sim R_{V}}(e_x)$, and $e_y = g_{E}(z)$ $[e_{y}] = \sfrac{i_{2,V}}{\sim R_{V}}(e_y)$.
-By applying $j_{1,E}$ to the first we get $j_{1,E};f_{E}(z) = j_{1,E}(e_x)$ and similarly for $j_{2,E}$: $j_{2,E};g_{E}(z) = j_{2,E}(e_y)$ and $j_{1,E}(e_{x}) = j_{2,E}(e_{y})$ by commutativity.
-Hence, $u_{E}([e_x]) = u_{E}([e_y])$.
-\update{Above $e_x$ and $e_y$ are different from $e_x$ and $e_y$ from the first proof. This is probablty not good.}
+To show that $\sfrac{X + Y}{\sim}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1 = (j_{1,V}, j_{1,E})$ and $j_2 = (j_{2,V}, j_{2,E})$ such that the above square commutes there is a unique $u = (u_{V}, u_{E})$ such that $j_1 = \sfrac{i_{1}}{\sim} ; u$ and $j_2 = \sfrac{i_2}{\sim} ; u$.
+We will define such $u$ as $u_{E}([i_{1,E}(e_{x})]) = j_{1,E}(e_{x})$ and $u_{E}([i_{2,E}(e_{y})]) = j_{2,E}(e_y)$ (similarly for $u_{V}$) and claim that it is a well-defined function by showing that if $t_1 \sim t_2$ then $u_{E}(t_1) = u_{E}(t_2)$.
+There are three options analogoues to the proof above, so we will show the base case for $t_1 = i_{1,E}(e_{x})$ and $t_2 = i_{2,E}(e_{y})$.
+$t_1 \sim t_2$ implies that there exists at least one $z$ in $E_{Z}$ such that $i_{1,E}(e_{x}) = f_{E};i_{1,E}(z)$ (or $e_{x} = f_{E}(z)$) and $[i_{1,E}(e_{x})] = \sfrac{i_{1,V}}{\sim}(e_x)$, and $e_y = g_{E}(z)$ $[in_{2,E}(e_{y})] = \sfrac{i_{2,V}}{\sim}(e_y)$.
+By applying $j_{1,E}$ to the first we get $j_{1,E}(f_{E}(z)) = j_{1,E}(e_x)$ and similarly for $j_{2,E}$: $j_{2,E}(g_{E}(z)) = j_{2,E}(e_y)$ and $j_{1,E}(e_{x}) = j_{2,E}(e_{y})$ by commutativity.
+Hence, $u_{E}([in_{1,E}(e_x)]) = u_{E}([in_{2,E}(e_y)])$.
+\update{Above $e_x$ and $e_y$ are different from $e_x$ and $e_y$ from the first proof, i.e. they are in $E_{X}$ and $E_{Y}$. This is probablty not good.}
 
@@
 Assume, that there exists $w : \sfrac{X+Y}{\sim R} \to Q$ for the same $j_1$ and $j_2$.
-In order for them to commute $w([x]) = j_1(x)$ and $w([y]) = j_2(y)$ which is the same as $u$.
+In order for them to commute $w_{E}([i_{1,E}(x)]) = j_1(x)$ and $w([in_{2,E}(y)]) = j_2(y)$ (similarly for $w_{V}$) which is the same as $u_{E}$.
 That is, $j_1$ and $j_2$ determine $u$ uniquely.

### 2023-10-27 sections/combinatorial_semantics.tex
Draft of pushout for EHyp

@@
     \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X} + s_{Y}, t_{X} + t_{Y}, \#_{X} + \#_{Y}, <_{c,X} + <_{c,Y} \}
-    \] 
-    once again omitting the labelling function.
-    And consider a relation $S = \{(x,y), (y,x) \in (X+Y) \times (X+Y) \; | \;\exists z \in Z \; . \; f(z) = x \text{ and } g(z) = y \}$ and let $R$ be its reflexive and transitive closure.
-    We then quotient $X + Y$ by $R$ and claim that it is a pushout.
-    \[
-        \sfrac{X + Y}{\sim R} = \{\sfrac{V_{X} + V_{Y}}{\sim R}, \sfrac{E_{X} + E_{Y}}{\sim R}, \sfrac{s_{X} + s_{Y}}{\sim R}, \sfrac{t_{X} + t_{Y}}{\sim R}, \sfrac{\#_{X} + \#_{Y}}{\sim R}, \sfrac{<_{c,X} + <_{c,Y}}{\sim R} \}
-    \]
-    where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$.
-    \[
-    \sfrac{(<_{c,X+Y})}{\sim (R_{V},R_{E})} \subseteq \sfrac{(V_{X} + V_{Y})}{\sim R_{V}} + \sfrac{(E_{X} + E_{Y})}{\sim R_{E}} \times \sfrac{(E_{X} + E_{Y})}{\sim R_{E}}
-    \]
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \#_{X+Y}, <_{c,X+Y} \}
+    \] 
+    once again omitting the labelling function and where source and target functions are as before.
+    And consider the same relations $S_{V},S_{E}$ and $R_{V}, R_{E}$.
+    We then quotient $X + Y$ by $R_{V},R_{E}$ and claim that it is a pushout.
+    \[
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X,Y}}{\sim ()}, \sfrac{\#_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{<_{c,X + Y}}{\sim (R_{V},R_{E})} \}
+    \]
+    where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$ and we will further omit writing the relations explicitly.
+    \[
+    \sfrac{(<_{c,X+Y})}{\sim} \subseteq \sfrac{(V_{X} + V_{Y})}{\sim R_{V}} + \sfrac{(E_{X} + E_{Y})}{\sim R_{E}} \times \sfrac{(E_{X} + E_{Y})}{\sim R_{E}}
+    \]
+    such that for each $(x_1, x_2) \in <_{c, X+Y}$ $\sfrac{<_{c,X+Y}}{\sim}$ contains $([x_1],[x_2])$.
+
     \update{We shall probably first define the quotient of the immediate parent because it is essentially a function and it will follow the same argument as the source function}

### 2023-10-31 sections/combinatorial_semantics.tex
Pushout attempt for EHyp

@@
 
-Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.
+\question{Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.}
 Finally, $u$ is unique.
@@
 
-    \item if $x_1$ is the immediate predecessor of $x_2$ in $\mathcal{F}$ then $\phi(x_1)$ is the immediate predecessor of $\phi(x_2)$ in $\mathcal{G}$
+    \item $\phi(<_{c,\mathcal{F}}^{\mu}(x)) = <_{c,\mathcal{G}}^{\mu}(\phi(x))$.
+          \update{We do not require such an equation when $[x) = \varnothing$}.
 
@@
     \]
+
     once again omitting the labelling function and where source and target functions are as before.
     And consider the same relations $S_{V},S_{E}$ and $R_{V}, R_{E}$.
-    We then quotient $X + Y$ by $R_{V},R_{E}$ and claim that it is a pushout.
-    \[
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X,Y}}{\sim ()}, \sfrac{\#_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{<_{c,X + Y}}{\sim (R_{V},R_{E})} \}
+
+    % \update{Now $<_{c,{X+Y}} : V_{X} + V_{Y} + E_{X} + E_{Y} \to E_{X} + E_{Y} + \bot_{X} + \bot_{Y}$
+    % And let's get back to the case when $<_{c,Z}^{\mu}(z) = \bot_{Z}$, $<_{c,X}^{\mu}(f_{V}(z)) = \bot_{X}$, $f_{E}(\bot_{Z}) = \bot_{X}$ and $g_{E}(\bot_{Z}) = <_{c,Y}^{\mu}(g_{V}(z))$
+    % We then want to show that $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]$
+    % Note that we have $[i_{1,E}(\bot_{X})] = [i_{2,E}(<_{c,Y}^{\mu}(g_{V}(z)))]$ and $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [i_{1,V}(\bot_{X})]$ because $i_{1,V}$ is a coproduct injection and $<_{c}^{\mu}(\bot) = \bot$.
+    % Now, because $g_{V};i_{2,V}$ is a homomorphism: $<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z)) = g_{V};i_{2,V}(<_{c,Z}(z)) = i_{2,V}(<_{c,Y}^{\mu}(g_{V}(z)))$
+    % }
+    We then quotient the sets of edges and nodes in $X + Y$ by $R_{V},R_{E}$
+    \[
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \#_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{c,\sfrac{X + Y}{\sim (R_{V},R_{E})}} \}
     \]
     where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$ and we will further omit writing the relations explicitly.
-    \[
-    \sfrac{(<_{c,X+Y})}{\sim} \subseteq \sfrac{(V_{X} + V_{Y})}{\sim R_{V}} + \sfrac{(E_{X} + E_{Y})}{\sim R_{E}} \times \sfrac{(E_{X} + E_{Y})}{\sim R_{E}}
-    \]
-    such that for each $(x_1, x_2) \in <_{c, X+Y}$ $\sfrac{<_{c,X+Y}}{\sim}$ contains $([x_1],[x_2])$.
-
-    \update{We shall probably first define the quotient of the immediate parent because it is essentially a function and it will follow the same argument as the source function}
-    \question{Below $e$ and $v$ do not necessarily mean edges and nodes, $v$ can be an edge as well.}
-    \question{I also kind of intuitively inventing the quotient of a relation, is there any accepted notion for this?}
-    and, whenever $e_{X} <_{c,X}^{\mu} v_{X}$ and $e_{Y} <_{c,Y}^{\mu} v_{Y}$ and $v_{x} R v_{y}$ then $e_{X} R e_{Y}$.
-    The latter condition means two things
+    Recall that $<_{c}$ is essentially a transitive clousure of $<_{c}^{\mu}$ and hence we can first define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]_{E})$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]_{V})$.
+    For this we will consider several cases where \update{$f$ and $g$ are mono. I require mono just to simplify things, we can get rid of this requirement if everything else is ok.}.
+    Assume that $u$ is an element of $E_{X+Y}$ (the case for a node is similar).
+    \begin{enumerate}
+        \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) \not = \varnothing$.
+              The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}(z))$.
+              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
+        \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) = \varnothing$ and $[f(z)) = \varnothing$ $[g(z)) \not = \varnothing$.
+              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(g(z))$.
+              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [g_{E};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]$.
+        \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) = \varnothing$ and $[f(z)) \not = \varnothing$ $[g(z)) = \varnothing$.
+              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(f(z))$.
+              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [f_{E};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]$.
+        \item $u$ has no pre-image in $Z$ and then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}(u)]$
+    \end{enumerate}
+    Clearly, all the cases above are disjoint.
+    Let's check that this is a well-defined function.
+    Suppose $t_1 \sim t_2$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_2])$.
+    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{E};i_{1,E}(z)$ and $t_2 = g_{E};i_{2,E}(z)$.
+    Then we need to consider the first three cases outlined above.
     \begin{itemize}
-        \label{cond:essential_for_pushout}
-        \item if $[z) = \varnothing$ then either $[f(z)) = \varnothing$ or $[g(z)) = \varnothing$.
-              \question{This point formally defines what it means to 'inherit' a relation.
-                        Let's say $v_1 = f(z)$ and $v_2 = g(z)$ and $e_1 <_{c,X}^{\mu} v_1$ and $[v_2) = \varnothing$,
-                        then, having $e_1 <_{c} [v_1,v_2]$ is fine as it does not break the condition of a predecessor being unique.}
-              \update{This point also says when we can compute a pushout: either non of X and Y should have a parent for $f(z)$ and $g(z)$ or only one of them should.}
-        \item if $[z) \not = \varnothing$ then there exists $z'$ such that $z' <_{c}^{\mu} z$ and $[g(z'), f(z')] <_{c}^{\mu} [g(z), f(z)]$
+        \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
+        \[
+            [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
+        \]
+        \[
+            [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
+        \]
+        and
+        \[ 
+            [f_{e};i_{1,E}(z')] = [g_{e};i_{2,E}(z')]
+        \]
+        by commutativity.
+        \item The other conditions satisfy functionality by definition.
     \end{itemize}
-    Then,
-    \[
-        \sfrac{\conflict_{X} + \conflict_{Y}}{\sim R} \subseteq \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R}
-    \]
-    meaning that whenever $e_1 \conflict e_2$ and $e_1' \conflict e_2'$ and $e_1 R e_1'$ then $[e_1, e_1'] \conflict e_2$ and $[e_1, e_1'] \conflict e_2'$.
-    $\sfrac{i_1}{\sim R}$ then maps $x \to [x]$ and $\sfrac{i_2}{\sim R}$ maps $y \to [y]$.
-    Before we proceed we need to check that $\sfrac{(X + Y)}{\sim R}$ is indeed an e-hypergraph.
-    Since $X + Y$ is clearly an e-hypergraph, we need to check whether identification breaks any of the requirements (1-7) above.
+
+    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms.
     \begin{enumerate}
-        \item Assume $[x_1,y_1] <_{c} [x_2,y_2]$ and $[x_1, y_1] \conflict [x_3]$ which means that $x_1 \conflict x_3$.
-              Hence $x_2 \conflict x_3$ (because $X$ is an e-hypergraph) and $[x_2,y_2] \conflict [x_3]$. Similarly, if $[x_1, y_1] \conflict [y_3]$ and $y_1 \conflict y_3$.
-        \item $\sfrac{\conflict}{\sim R}$ and $\sfrac{<_{c}}{\sim R}$ are disjoint because we quotient only inside the relations.
-        \item Again, we only quotient $\conflict$ on $x$ and $y$ for which original relation is defined.
-        \item This follows from $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ being homomorphisms.
-        \item We just perform transitive closure of the complement.
+        \item $s_{X+Y};[]_{V}^{*}(e) = []_{E};s_{\sfrac{X+Y}{\sim}}(e)$
+              This is the same as in $\catname{Hyp_{\Sigma}}$.
+        \item $[<_{c,X+Y}^{\mu}(e)]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E})$ for $e$ such that $[e) \not = \varnothing$.
+              Let's make case analysis on $[e]_{E}$.
+              \begin{itemize}
+                \item $e$ has a pre-image $z$ in $E_{Z}$.
+                    \begin{itemize}
+                        \item If $[z) \not = \varnothing$.
+                              Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E}) = [<_{c,X+Y}(e)]_{E}$ which is what we need.
+                        \item If $[z) = \varnothing$ and $[f_{E}(z)) = \varnothing$ and $[g_{E}(z)) \not = \varnothing$.
+                              Because $i_{2,E}$ is a homomorphism, we have $<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z)) = i_{2,E}(<_{c,Y}(g_{E}(z))) = i_{2,E}(e_{y})$.
+                              We next have either $e = f_{E};i_{1,E}(z)$ in which case $[e) = \varnothing$ because $i_{1,E}$ is a coproduct injection and \update{we do not require homomorphism-ness for such $e$}.
+                              Or, $e = g_{E};i_{2,E}(z)$ and 
+                              \[
+                              <_{c,\sfrac{X+Y}{\sim}}([e]) = [g_{E};<_{c,Y}{\mu};i_{2,E}(z)]_{E} = [i_{2,E}(e_{y})]
+                              \]
+                              and
+                              \begin{align*}
+                                [<_{c,X+Y}(e)] = [<_{c,X+Y}(g_{E};i_{2,E}(z))] &= \\
+                                &\text{because $[g_{E}(z)) \not = \varnothing$ and $i_{2,E}$ is homomorphism}\\
+                                 = [i_{2,E}(<_{c,Y}(g_{E}(z)))] = [i_{2,E}(e_{y})]
+                              \end{align*}
+                              so we get
+                              \[
+                              <_{c,\sfrac{X+Y}{\sim}}([e]) = [<_{c,X+Y}(e)]
+                              \]
+                        \item The remaining case is analogous.
+                    \end{itemize}
+                \item If $e$ has no pre-image in $E_{Z}$ then homomorphism-ness of $[]_{E}$ is similar to the first case above.
+              \end{itemize}
+    \end{enumerate}
+    The proof for when $e$ is a node is analogous.
+
+    Let's now prove the universal property.
+    $u_{E}([i_{1,E}(x)]) = j_{1,E}(x)$ and $u_{E}([i_{2,E}(y)]) = j_2(y)$.
+    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,E}$ and $j_{2,E}$.
+
+    Let's say that $t_1 \sim t_2$ then $u_{E}([t_1]) = u_{E}([t_2])$.
+    $u_{E}([t_1]) = u_{E}(f_{E};i_{1,E}(z)) = j_1{f_{E}(z)}$ and $u_{E}([t_2]) = u_{E}(g_{E};i_{2,E}(z)) = j_{2}(g_{E}(z))$ which are equal by commutativity.
+
+    \update{TODO: $\conflict$}
+    % We can think of relation $<_{c,X+Y}$ as of a function $<_{c} : V_{X+Y} + E_{X+Y} \to {E_{X+Y}}^{*}$, i.e. a function that for each node or edge returns the sequence of their predecessors.
+    % Then we can define the quotienting as before
+    % \[
+    % \sfrac{(<_{c,X+Y})}{\sim}([u]) = [<_{c,X+Y}(u)]_{E}^{*}\\
+    % \]
+    % or, if we consider a function that returns the immediate predecessor
+    % \[
+    % \sfrac{(<_{c,X+Y}^{\mu})}{\sim}([u]) = [<_{c,X+Y}^{\mu}(u)]_{E}
+    % \]
+    % To check that the function is well-defined we need to check that if $t_1 \sim t_2$ then $\sfrac{<_{c,X+Y}^{\mu}}{\sim}([t_1]) = \sfrac{<_{c,X+Y}^{\mu}}{\sim}([t_2])$ or $[<_{c,X+Y}^{\mu}(t_1)] = [<_{c,X+Y}^{\mu}(t_2)]$.
+    % Once again, $t_1 \sim t_2$ (suppose these are nodes) implies there exists $z$ such that $t_1 = f_{V}(z);i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$.
+    % Hence we have
+    % \[
+    %     [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))]
+    % \]
+    % \[
+    %     [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
+    % \]
+    % and two cases
+    % \begin{itemize}
+    %     \item $[z) \not = \varnothing$ in which case we can use the fact that $f_{V};i_{1,V}$ and $g_{V};i_{2,V}$ are homomorphisms.
+    %           This gives us 
+    %             \[
+    %                 f_{V};i_{1,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))
+    %             \] 
+    %             and
+    %             \[
+    %                 g_{V};i_{2,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))
+    %             \]
+    %             and finally,
+    %             \[
+    %                 [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
+    %             \] 
+    %             as they share the same pre-image.
+    %     \item Another case is when $[z) = \varnothing$. 
+    %           Assume $[f(z)) = \varnothing$, in which case $<_{c,Z}^{\mu}(z)$ and $<_{c,X}^{\mu}(f_{V}(z))$ are undefined.
+    %           This takes us to the situation where the lefthand-side of what we want to prove is undefined while the right-hand side is defined.
+    %           \question{Can I somehow require that in such case I take the right-hand side as a definition of the lefthand side? 
+    %           Otherwise the relation we have is too weak because it tells us only to identify things with the pre-image in $Z$ but eventually we also want to identify things that do not have such a pre-image.
+    %           }
+    %           What I want is to define 
+    %           \[
+    %             \sfrac{<_{c,X+Y}^{\mu}}{\sim}([f_{V};i_{1,V}(z)]) = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))] \text{ if } [f_{V};i_{1,V}(z)) = \varnothing
+    %           \]
+    % \end{itemize}
+
+    % \update{Approach 2}
+
+    % We can extend the set of edges in an e-hypergraph with a special edge $\bot$.
+    % And allow that it can be mapped to any other edge under a homomorphisms, i.e. we will not require that $s;\phi^{*}(\bot) = \phi;s;(\bot)$.
+    % \question{but this might break quotienting of sources and targets}
+    % This also means that set of edges now is $E_{\mathcal{G}} + \bot$.
+    % \question{What to do with $\bot$ after a coproduct?}.
+    % This gives us the ability to map this $\bot$ using homomorphisms.
+
+    % The pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
+    % First, a coproduct of $X+Y$ is computed which is
+
+    % \[
+    %     X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y} + \bot_{X} + \bot_{Y}, s_{X+Y}, t_{X+Y}, \#_{X+Y}, <_{c,X+Y} \}
+    % \]
+
+    % \[
+    %     R_{E} = \{ (x_i, y_j) \in (E_{X} + E_{Y} + \bot_{X} + \bot_{Y}) \times (E_{X} + E_{Y} + \bot_{X} + \bot_{Y} | \exists z \in E_{Z} + \bot_{Z} \; .
+\; x_i = f_{E};i_{1,E}(z) \text{ and } y_j = g_{E};i_{2,E}(z))\}
+    % \]
+
+    % Consider $<_{c,X+Y}^{\mu}$ as a function that returns the immediate predecessor, i.e. it has a type $V_{X} + V_{Y} + E_{X} + E_{Y} \to E_{X} + E_{Y} + \bot_{X} + \bot_{Y}$.
+    % And define its quotient
+    % \[
+    % \sfrac{<_{c,X+Y}^{\mu}}{\sim}([u]) = [<_{c,X+Y}^{\mu}(u)]_{E}
+    % \]
+
+    % We then need to check that whenever $t_1 \sim t_2$ $[<_{c,X+Y}^{\mu}(t_1)] = [<_{c,X+Y}^{\mu}(t_2)]$.
+    % $t_1 \sim t_2$ implies that there exists at least one $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$.
+    % Hence we have
+    % \[
+    %     [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))]
+    % \]
+    % \[
+    %     [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
+    % \]
+    % and two cases
+    % We then can use the fact that $f_{V};i_{1,V}$ and $g_{V};i_{2,V}$ are homomorphisms.
+    % This gives us 
+    % \[
+    %     f_{V};i_{1,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))
+    % \] 
+    % and
+    % \[
+    %     g_{V};i_{2,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))
+    % \]
+    % and finally,
+    % \[
+    %     [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
+    % \] 
+    % as they share the same pre-image ($<_{c,Z}^{\mu}(z)$).
+
+    % Now let's check that $\sfrac{s_{X+Y}}{\sim}$ is well-defined.
+    % We will consider a case where $f_{E}(\bot_{Z}) = \bot_{X}$ and $g_{E}(\bot_{Z}) \not = \bot$.
+    % We then need to check that $[s_{X+Y}(f_{E};i_{1,E}(z))]^{*} = [s_{X+Y}(g_{E};i_{2,E}(z))]$
+    % \begin{itemize}
+    %     \item $z = \bot_{Z}$ then $[{}] = \ldots$
+    % \end{itemize}
+
+    % \update{end of Approach 2}
+
+    % \update{We shall probably first define the quotient of the immediate parent because it is essentially a function and it will follow the same argument as the source function}
+    % \question{Below $e$ and $v$ do not necessarily mean edges and nodes, $v$ can be an edge as well.}
+    % \question{I also kind of intuitively inventing the quotient of a relation, is there any accepted notion for this?}
+    % and, whenever $e_{X} <_{c,X}^{\mu} v_{X}$ and $e_{Y} <_{c,Y}^{\mu} v_{Y}$ and $v_{x} R v_{y}$ then $e_{X} R e_{Y}$.
+    % The latter condition means two things
+    % \begin{itemize}
+    %     \label{cond:essential_for_pushout}
+    %     \item if $[z) = \varnothing$ then either $[f(z)) = \varnothing$ or $[g(z)) = \varnothing$.
+    %           \question{This point formally defines what it means to 'inherit' a relation.
+    %                     Let's say $v_1 = f(z)$ and $v_2 = g(z)$ and $e_1 <_{c,X}^{\mu} v_1$ and $[v_2) = \varnothing$,
+    %                     then, having $e_1 <_{c} [v_1,v_2]$ is fine as it does not break the condition of a predecessor being unique.}
+    %           \update{This point also says when we can compute a pushout: either non of X and Y should have a parent for $f(z)$ and $g(z)$ or only one of them should.}
+    %     \item if $[z) \not = \varnothing$ then there exists $z'$ such that $z' <_{c}^{\mu} z$ and $[g(z'), f(z')] <_{c}^{\mu} [g(z), f(z)]$
+    % \end{itemize}
+    % Then,
+    % \[
+    %     \sfrac{\conflict_{X} + \conflict_{Y}}{\sim R} \subseteq \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R}
+    % \]
+    % meaning that whenever $e_1 \conflict e_2$ and $e_1' \conflict e_2'$ and $e_1 R e_1'$ then $[e_1, e_1'] \conflict e_2$ and $[e_1, e_1'] \conflict e_2'$.
+    % $\sfrac{i_1}{\sim R}$ then maps $x \to [x]$ and $\sfrac{i_2}{\sim R}$ maps $y \to [y]$.
+    % Before we proceed we need to check that $\sfrac{(X + Y)}{\sim R}$ is indeed an e-hypergraph.
+    % Since $X + Y$ is clearly an e-hypergraph, we need to check whether identification breaks any of the requirements (1-7) above.
+    % \begin{enumerate}
+    %     \item Assume $[x_1,y_1] <_{c} [x_2,y_2]$ and $[x_1, y_1] \conflict [x_3]$ which means that $x_1 \conflict x_3$.
+    %           Hence $x_2 \conflict x_3$ (because $X$ is an e-hypergraph) and $[x_2,y_2] \conflict [x_3]$. Similarly, if $[x_1, y_1] \conflict [y_3]$ and $y_1 \conflict y_3$.
+    %     \item $\sfrac{\conflict}{\sim R}$ and $\sfrac{<_{c}}{\sim R}$ are disjoint because we quotient only inside the relations.
+    %     \item Again, we only quotient $\conflict$ on $x$ and $y$ for which original relation is defined.
+    %     \item This follows from $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ being homomorphisms.
+    %     \item We just perform transitive closure of the complement.
 \update{Maybe we can get rid of this point and instead later just take transitive closure of the complement when talking about interfaces}
-        \item Assume $[x_1, y_1] \conflict [x_2,y_2]$ and $[x_2,y_2] \conflict [x_3,y_3]$ and all of them are distinct.
-              Then $[x_1,y_1] \conflict [x_3,y_3]$ because either $x_1 \conflict x_3$ or $y_1 \conflict y_3$ (otherwise they would not have been merged into the same equivalence class).
-        \item We never merge $x_1$ and $x_2$ such that $x_1 <_{c} x_2$ into the same equivalence class because morphisms should preserve the order.
-    \end{enumerate}
-
-    Then we need to check that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are indeed homomorphisms.
-    For these we need to check that conditions (2-3) in~\ref{def:e-homo} are satisfied.
-    Let's check for $\sfrac{i_1}{\sim R}$.
-    We identify elements from $X$ with elements from $Y$ which does not affect (2) and the same reasoning applies to (3).
-    (1) was proved earlier for $\catname{Hyp_{\Sigma}}$.
-
-    Finally, by following essentially the same argument as for pushout in $\catname{Hyp_{\Sigma}}$ we can conclude that the structure defined above is indeed a pushout in $\catname{EHyp_{\Sigma}}$ and we know precise conditions under which this concrete pushout exists.
+    %     \item Assume $[x_1, y_1] \conflict [x_2,y_2]$ and $[x_2,y_2] \conflict [x_3,y_3]$ and all of them are distinct.
+    %           Then $[x_1,y_1] \conflict [x_3,y_3]$ because either $x_1 \conflict x_3$ or $y_1 \conflict y_3$ (otherwise they would not have been merged into the same equivalence class).
+    %     \item We never merge $x_1$ and $x_2$ such that $x_1 <_{c} x_2$ into the same equivalence class because morphisms should preserve the order.
+    % \end{enumerate}
+
+    % Then we need to check that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are indeed homomorphisms.
+    % For these we need to check that conditions (2-3) in~\ref{def:e-homo} are satisfied.
+    % Let's check for $\sfrac{i_1}{\sim R}$.
+    % We identify elements from $X$ with elements from $Y$ which does not affect (2) and the same reasoning applies to (3).
+    % (1) was proved earlier for $\catname{Hyp_{\Sigma}}$.
+
+    % Finally, by following essentially the same argument as for pushout in $\catname{Hyp_{\Sigma}}$ we can conclude that the structure defined above is indeed a pushout in $\catname{EHyp_{\Sigma}}$ and we know precise conditions under which this concrete pushout exists.
 \end{remark}

### 2023-10-31 sections/combinatorial_semantics.tex
Minor fixes

@@
         \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) \not = \varnothing$.
-              The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}(z))$.
+              The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
               We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
@@
               We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [f_{E};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]$.
-        \item $u$ has no pre-image in $Z$ and then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}(u)]$
+        \item $u$ has no pre-image in $Z$ and then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
     \end{enumerate}
@@
         \[ 
-            [f_{e};i_{1,E}(z')] = [g_{e};i_{2,E}(z')]
+            [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
         \]
@@
                         \item If $[z) \not = \varnothing$.
-                              Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E}) = [<_{c,X+Y}(e)]_{E}$ which is what we need.
+                              Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E}) = [<_{c,X+Y}^{\mu}(e)]_{E}$ which is what we need.
                         \item If $[z) = \varnothing$ and $[f_{E}(z)) = \varnothing$ and $[g_{E}(z)) \not = \varnothing$.
-                              Because $i_{2,E}$ is a homomorphism, we have $<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z)) = i_{2,E}(<_{c,Y}(g_{E}(z))) = i_{2,E}(e_{y})$.
+                              Because $i_{2,E}$ is a homomorphism, we have $<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z)) = i_{2,E}(<_{c,Y}^{\mu}(g_{E}(z))) = i_{2,E}(e_{y})$.
                               We next have either $e = f_{E};i_{1,E}(z)$ in which case $[e) = \varnothing$ because $i_{1,E}$ is a coproduct injection and \update{we do not require homomorphism-ness for such $e$}.
@@
                               \[
-                              <_{c,\sfrac{X+Y}{\sim}}([e]) = [g_{E};<_{c,Y}{\mu};i_{2,E}(z)]_{E} = [i_{2,E}(e_{y})]
+                              <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]) = [g_{E};<_{c,Y}^{\mu};i_{2,E}(z)]_{E} = [i_{2,E}(e_{y})]
                               \]
@@
                               \begin{align*}
-                                [<_{c,X+Y}(e)] = [<_{c,X+Y}(g_{E};i_{2,E}(z))] &= \\
+                                [<_{c,X+Y}^{\mu}(e)] = [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] &= \\
                                 &\text{because $[g_{E}(z)) \not = \varnothing$ and $i_{2,E}$ is homomorphism}\\
@@
                               \[
-                              <_{c,\sfrac{X+Y}{\sim}}([e]) = [<_{c,X+Y}(e)]
+                              <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]) = [<_{c,X+Y}^{\mu}(e)]
                               \]

### 2023-11-08 sections/introduction.tex
remove e-rewriting

@@
 As a corollary, we also formalize the rewriting theory of standard e-graphs in this way.
-Finally, we touch on the question of efficient implementation of rewriting for e-hypergraphs.
+%Finally, we touch on the question of efficient implementation of rewriting for e-hypergraphs.
 
@@
 
-\subsection{E-matching and E-rewriting}
-
-Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
+%\subsection{E-matching and E-rewriting}
+%
+%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
 In general, this involves unsharing the e-box structure before a redex becomes available.
@@
 Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-\begin{figure}
-\[
-	\tikzfig{combinatorial_semantics/example-rewriting}
-\]
-\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
+%\begin{figure}
+%\[
+%	\tikzfig{combinatorial_semantics/example-rewriting}
+%\]
+%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
 REDO THIS}
-\end{figure}
+%\end{figure}
 

### 2023-11-08 sections/combinatorial_semantics.tex
Working draft of pushouts in EHyp

@@
     The labelling $l$ is modified to include an extra value $\bot \;$ $l :E_{\mathcal{G}} \to \Sigma + 1$. 
-With an abuse of notation, we will confuse $\mathcal{G}$ with $V_{\mathcal{G}} \cup E_{\mathcal{G}}$, when treating it as a set.
-
-    $<_c \subset \mathcal{G} \times \mathcal{G}$ is a child relation imposing a partial order on $\mathcal{G}$.
-    We require that for a given $x \in \mathcal{G}$ a set of all predecessors of $x\;$ $[x) = \{x' \in \mathcal{G} \;||\; x' <_c x\; \}$ is finite and consists of edges $e \in E_{\mathcal{G}}$ such that $l(e) = \bot$ only.
+    % With an abuse of notation, we will confuse $\mathcal{G}$ with $V_{\mathcal{G}} \cup E_{\mathcal{G}}$, when treating it as a set.
+    \question{Shall I use $\cup$ instead of $+$?}
+    $<_c \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is a child relation imposing a strict partial order on $(V_\mathcal{G} + E_\mathcal{G})$.
+    We require that for a given $x \in (V_\mathcal{G} + E_\mathcal{G})$ a set of all predecessors of $x\;$ $[x) = \{x' \in E_\mathcal{G} \;||\; x' <_c x\; \}$ is finite and consists of edges $e \in E_{\mathcal{G}}$ such that $l(e) = \bot$ only.
     Next, each $x$ has at most one immediate predecessor.
@@
     Such order was called a \textbf{well-founded arborescent} strict partial order in~\cite{orders_and_events}.
-    Similarly, a set of all successors is $(x] = \{x' \in \mathcal{G} \; ||\; x < x'\}$ where $x'$ can be a node or an edge. 
+    Similarly, a set of all successors is $(x] = \{x' \in (V_\mathcal{G} + E_\mathcal{G}) \; ||\; x <_{c} x'\}$ where $x'$ can be a node or an edge. 
     % An edge $e \in E_{\mathcal{G}}$ is unlabelled, i.e. $l(e) = \bot$ iff $(e] \not = \varnothing$.
-    Finally, we require that $<_c$ respects connectivity, i.e. if $v \in s(e)$ then $e' < e$ iff $e' <_c v$.
-
-    $\langle \mathcal{G}, <_c, \# \rangle$ is an event-structure where $\# \subset \mathcal{G} \times \mathcal{G}$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
+    Finally, we require that $<_c^{\mu}$ respects connectivity, i.e. if $v \in s(e)$ then $e' <_c^{\mu} e$ iff $e' <_c^{\mu} v$.
+
+    $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
     \begin{enumerate}
+        \label{def:e-hypergraph-req}
         \item 
         \[
-    \forall x_1, x_2, x_3 \in \mathcal{G} \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
+    \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
     \] 
@@
     \item $<_c$ and $\#$ are mutually exclusive
-    \item $\#$ is defined \textit{only} for $x \in \mathcal{G}$ such that $[x) \not \eq \varnothing$
+    \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
     \item $\#$ should respect connectivity
-    \item \update{Complement of $\#$ should be transitive}
-    \item for all distinct $x_1, x_2, x_3 \in V \cup E$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
-    \item for all $x, x' \in \mathcal{G}$ $x \#_{\mu} x'$ implies $[x) = [x')$
+    \item \update{Complement of $\conflict_{\mu}$ should be transitive}
+    \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
+    \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
         \end{enumerate}
@@
 Intuitively, the child and conflict relation together make it possible to model e-classes in our e-hypergraphs: an unlabelled edge $e$ represents an e-class where the children of $e$ that are in immediate conflict correspond to nodes and edges in different e-class instances.
-Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$. Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.
+Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$.
+Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.
 
@@
     % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
-    \question{
+
         \item
-    if $x_1 \#_{\mathcal{F}} x_2$ then $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
-    }
+    if $x_1 \#_{\mathcal{F}} x_2$ then $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
+    Or, written functionally, when $x \in E_{\mathcal{F}}$
+    \[
+        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x)))
+        \subseteq
+        \conflict_{\mathcal{G}}(\phi_{E};i_{E_{\mathcal{G}}}(x))
+    \]
+    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+    And, when $x \in V_{\mathcal{F}}$
+    \[
+        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\conflict_{\mathcal{F}}(i_{V_{\mathcal{F}}}(x)))
+        \subseteq
+        \conflict_{\mathcal{G}}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+    \]
+    Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
 \end{enumerate}
 
-\update{Let's denote the fact of $x_1$ being the immediate predecessor of $x_2$ by $x_1 <_{\mu} x_2$.
-Also, note that preserving the immediate predecessor implies preserving order.
-Homomorphism definition seems to be fine: by not requiring the \textit{only if} direction in bullet 2 we basically allow either of two things
-
-\begin{itemize}
-    \item $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 <_{\mu} x_2$
-    \item $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1$ is unrelated to $x_2$
-\end{itemize}
-
-We can not have $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 < x_2$, i.e.
+% \update{Let's denote the fact of $x_1$ being the immediate predecessor of $x_2$ by $x_1 <_{\mu} x_2$.
+%  Also, note that preserving the immediate predecessor implies preserving order.
+% Homomorphism definition seems to be fine: by not requiring the \textit{only if} direction in bullet 2 we basically allow either of two things
+
+% \begin{itemize}
+%     \item $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 <_{\mu} x_2$
+%     \item $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1$ is unrelated to $x_2$
+% \end{itemize}
+
+% We can not have $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 < x_2$, i.e.
 the case where $x_1$ is not the immediate predecessor of $x_2$ because then it would mean that there exists at least $x_3$ such that $x_1 <_{\mu} x_3 <_{\mu} x_2$ and because $\phi$ should preserve them it would mean that $\phi(x_1) <_{\mu} \phi(x_3) <_{\mu} \phi(x_2)$ implying that $\phi(x_2)$ has two immediate predecessors: $\phi(x_1)$ and $\phi(x_3)$.
-Of course, if $\phi$ does not identify $x_1$ and $x_3$.
-The first bullet is fine because it means that $\phi$ reflects the immediate predecessor.
-The second bullet leaves us with two options depicted above.
-
-}
-
-\begin{figure}
-    \centering
-\begin{subfigure}{0.4\linewidth}
-
-\scalebox{0.5}{
-\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=node, label={above:$x_1$}] (19) at (-3.5, -4.5) {};
-		\node [style=node, label={above:$x_2$}] (20) at (-4.5, -6.5) {};
-		\node [style=node, label={above:$x_3$}] (21) at (-2.5, -6.5) {};
-		\node [style=node, label={above:$x_4$}] (22) at (-3.5, -8.5) {};
-		\node [style=node, label={above:$x_5$}] (23) at (-1.5, -8.5) {};
-		\node [style=node, label={right:$x_2$}] (24) at (4.5, -5.5) {};
-		\node [style=node, label={above:$x_1$}] (25) at (4.5, -2.5) {};
-		\node [style=node, label={below:$x_5$}] (26) at (4.5, -11.5) {};
-		\node [style=none] (27) at (4.5, -7.5) {};
-		\node [style=none] (28) at (4.5, -9.5) {};
-		\node [style=none] (29) at (4.5, -8.5) {$\vdots$};
-		\node [style=none] (30) at (0.5, -6.5) {};
-		\node [style=none] (31) at (2.5, -6.5) {};
-		\node [style=none] (32) at (1.5, -5.5) {$\phi$};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw (19) to (20);
-		\draw (19) to (21);
-		\draw (21) to (22);
-		\draw (21) to (23);
-		\draw (24) to (27.center);
-		\draw (26) to (28.center);
-		\draw (25) to (24);
-		\draw [style=diredge] (30.center) to (31.center);
-	\end{pgfonlayer}
-\end{tikzpicture}
-
-}
-
-\caption{Case 1}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\linewidth}
-
-\scalebox{0.5}{
-\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=node, label={above:$x_1$}] (19) at (-6.5, -4.5) {};
-		\node [style=node, label={above:$x_2$}] (20) at (-7.5, -6.5) {};
-		\node [style=node, label={above:$x_3$}] (21) at (-5.5, -6.5) {};
-		\node [style=node, label={above:$x_4$}] (30) at (-3, -4.5) {};
-		\node [style=node, label={above:$x_1$}] (31) at (2.5, -4.5) {};
-		\node [style=node, label={above:$x_2$}] (32) at (1.5, -6.5) {};
-		\node [style=node, label={above:$x_3$}] (33) at (2.75, -6.5) {};
-		\node [style=node, label={above:$x_4$}] (34) at (4, -6.5) {};
-		\node [style=none] (35) at (-1.5, -5.5) {};
-		\node [style=none] (36) at (-0.5, -5.5) {};
-		\node [style=none] (37) at (-1, -5) {$\phi$};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw (19) to (20);
-		\draw (19) to (21);
-		\draw (31) to (32);
-		\draw (31) to (33);
-		\draw (31) to (34);
-		\draw [style=diredge] (35.center) to (36.center);
-	\end{pgfonlayer}
-\end{tikzpicture}
-
-}
-
-\caption{Case 2}
-
-\end{subfigure}
-\end{figure}
-
-\update{Case (a) above is impossible on two occasions: for the image of $x_5$ to be a successor of the image of $x_2$, $x_5$ itself should be hierarchical.
+% Of course, if $\phi$ does not identify $x_1$ and $x_3$.
+% The first bullet is fine because it means that $\phi$ reflects the immediate predecessor.
+% The second bullet leaves us with two options depicted above.
+
+% }
+
+% \begin{figure}
+%     \centering
+% \begin{subfigure}{0.4\linewidth}
+
+% \scalebox{0.5}{
+% \begin{tikzpicture}
+% 	\begin{pgfonlayer}{nodelayer}
+% 		\node [style=node, label={above:$x_1$}] (19) at (-3.5, -4.5) {};
+% 		\node [style=node, label={above:$x_2$}] (20) at (-4.5, -6.5) {};
+% 		\node [style=node, label={above:$x_3$}] (21) at (-2.5, -6.5) {};
+% 		\node [style=node, label={above:$x_4$}] (22) at (-3.5, -8.5) {};
+% 		\node [style=node, label={above:$x_5$}] (23) at (-1.5, -8.5) {};
+% 		\node [style=node, label={right:$x_2$}] (24) at (4.5, -5.5) {};
+% 		\node [style=node, label={above:$x_1$}] (25) at (4.5, -2.5) {};
+% 		\node [style=node, label={below:$x_5$}] (26) at (4.5, -11.5) {};
+% 		\node [style=none] (27) at (4.5, -7.5) {};
+% 		\node [style=none] (28) at (4.5, -9.5) {};
+% 		\node [style=none] (29) at (4.5, -8.5) {$\vdots$};
+% 		\node [style=none] (30) at (0.5, -6.5) {};
+% 		\node [style=none] (31) at (2.5, -6.5) {};
+% 		\node [style=none] (32) at (1.5, -5.5) {$\phi$};
+% 	\end{pgfonlayer}
+% 	\begin{pgfonlayer}{edgelayer}
+% 		\draw (19) to (20);
+% 		\draw (19) to (21);
+% 		\draw (21) to (22);
+% 		\draw (21) to (23);
+% 		\draw (24) to (27.center);
+% 		\draw (26) to (28.center);
+% 		\draw (25) to (24);
+% 		\draw [style=diredge] (30.center) to (31.center);
+% 	\end{pgfonlayer}
+% \end{tikzpicture}
+
+% }
+
+% \caption{Case 1}
+% \end{subfigure}
+% \hfill
+% \begin{subfigure}{0.4\linewidth}
+
+% \scalebox{0.5}{
+% \begin{tikzpicture}
+% 	\begin{pgfonlayer}{nodelayer}
+% 		\node [style=node, label={above:$x_1$}] (19) at (-6.5, -4.5) {};
+% 		\node [style=node, label={above:$x_2$}] (20) at (-7.5, -6.5) {};
+% 		\node [style=node, label={above:$x_3$}] (21) at (-5.5, -6.5) {};
+% 		\node [style=node, label={above:$x_4$}] (30) at (-3, -4.5) {};
+% 		\node [style=node, label={above:$x_1$}] (31) at (2.5, -4.5) {};
+% 		\node [style=node, label={above:$x_2$}] (32) at (1.5, -6.5) {};
+% 		\node [style=node, label={above:$x_3$}] (33) at (2.75, -6.5) {};
+% 		\node [style=node, label={above:$x_4$}] (34) at (4, -6.5) {};
+% 		\node [style=none] (35) at (-1.5, -5.5) {};
+% 		\node [style=none] (36) at (-0.5, -5.5) {};
+% 		\node [style=none] (37) at (-1, -5) {$\phi$};
+% 	\end{pgfonlayer}
+% 	\begin{pgfonlayer}{edgelayer}
+% 		\draw (19) to (20);
+% 		\draw (19) to (21);
+% 		\draw (31) to (32);
+% 		\draw (31) to (33);
+% 		\draw (31) to (34);
+% 		\draw [style=diredge] (35.center) to (36.center);
+% 	\end{pgfonlayer}
+% \end{tikzpicture}
+
+% }
+
+% \caption{Case 2}
+
+% \end{subfigure}
+% \end{figure}
+
+% \update{Case (a) above is impossible on two occasions: for the image of $x_5$ to be a successor of the image of $x_2$, $x_5$ itself should be hierarchical.
 Then, because $\phi$ should preserve the immediate predecessor, the image of $x_3$ should be the immediate predecessor of the image of $x_5$ and the image of $x_1$ should be the immediate predecessor of the image of $x_3$ and of the image of $x_2$.
@@
 
-The case (b) above typechecks (for typed e-hypergraphs) only if $x_4$ is of type $0 \to 0$.
+% The case (b) above typechecks (for typed e-hypergraphs) only if $x_4$ is of type $0 \to 0$.
 This case is fine because we always can append, e.g.
@@
 This also assumes that the image is down-closed.
-}
-
-\question{There is an issue with requiring that $\phi$ should reflect the conflict though.
+% }
+
+% \question{There is an issue with requiring that $\phi$ should reflect the conflict though.
 When we consider e-cospans and the maps from the interfaces to the carrier we map some nodes from the interface to conflicted nodes in the carrier.
@@
 The same was done for down-closed-ness, for example.
-}
+% }
 
@@
 
-    We will compute pushouts in $\catname{EHyp_{\Sigma}}$ similarly to $\catname{EHyp_{\Sigma}}$.
+    We will compute pushouts in $\catname{EHyp_{\Sigma}}$ similarly to $\catname{Hyp_{\Sigma}}$ but additionally will make a few assumptions.
     Recall the diagram below.
@@
 
-    The pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
+    First, we assume that $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
+
+    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
     First, a coproduct of $X+Y$ is computed which is 
@@
     For this we will consider several cases where \update{$f$ and $g$ are mono. I require mono just to simplify things, we can get rid of this requirement if everything else is ok.}.
-    Assume that $u$ is an element of $E_{X+Y}$ (the case for a node is similar).
+    Assume that $u$ is an element of $V_{X+Y}$.
     \begin{enumerate}
-        \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) \not = \varnothing$.
-              The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
-              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) = \varnothing$ and $[f(z)) = \varnothing$ $[g(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(g(z))$.
-              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [g_{E};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]$.
-        \item $u$ has a pre-image $z$ in $E_{Z}$, $[z) = \varnothing$ and $[f(z)) \not = \varnothing$ $[g(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(f(z))$.
-              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [f_{E};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]$.
+        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
+        %       The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
+        %       We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
+        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$ $[g_{V}(z)) \not = \varnothing$.
+              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(g_{V}(z))$.
+              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [g_{V};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]$.
+        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
+              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(f_{V}(z))$.
+              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [f_{V};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]$.
         \item $u$ has no pre-image in $Z$ and then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
     \end{enumerate}
+    \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
+    Now, if $u$ is an element of $E_{X+Y}$.
+    \begin{enumerate}
+        \item We have the only case of $u$ having no pre-image in $E_{Z}$ because its empty.
+              Hence, we define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$.
+    \end{enumerate}
+    \question{I decided to keep Z discrete, otherwise the number of cases to consider within the proofs is too large.
+    This also means that we now can identify only nodes.
+    I will make changes to account for this later.
+    }
+    \update{If $f$ and $g$ are not mono, the definitions stay the same as $i_{2,E}$ and $i_{1,E}$ are mono.}
     Clearly, all the cases above are disjoint.
     Let's check that this is a well-defined function.
-    Suppose $t_1 \sim t_2$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{E};i_{1,E}(z)$ and $t_2 = g_{E};i_{2,E}(z)$.
-    Then we need to consider the first three cases outlined above.
-    \begin{itemize}
-        \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
-        \[
-            [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
-        \]
-        \[
-            [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
+    Suppose $t_1, t_2$ are nodes and $t_1 \sim t_2$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_2])$.
+    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$ and either $[f_{V}(t_1)) = \varnothing$ or $[g_{V}(z)) = \varnothing$.
+    Assume the latter, then, by definition.
+    \[
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_1]) = [f_{V};<_{c,X}^{\mu};i_{1,E}(z)]    
     \]
@@
     \[
-            [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
-        \]
-        by commutativity.
-        \item The other conditions satisfy functionality by definition.
-    \end{itemize}
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_2]) = [f_{V};<_{c,X}^{\mu};i_{1,E}(z)]    
+    \]
+    so we have well-defined-ness on the nose.
+
+    % \begin{itemize}
+        % \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
+        % \[
+        %     [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
+        % \]
+        % \[
+        %     [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
+        % \]
+        % and
+        % \[ 
+        %     [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
+        % \]
+        % by commutativity.
+        % \item The other conditions satisfy functionality by definition.
+
+    % \end{itemize}
 
@@
         \item $[<_{c,X+Y}^{\mu}(e)]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E})$ for $e$ such that $[e) \not = \varnothing$.
-              Let's make case analysis on $[e]_{E}$.
+              This is homomorphic by definition. 
+        \item $[<_{c,X+Y}^{\mu}(v)]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([v]_{V})$.
+              Let's make case analysis on $[v]_{V}$.
               \begin{itemize}
-                \item $e$ has a pre-image $z$ in $E_{Z}$.
+                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};i_{1,V}(z)$.
                     \begin{itemize}
-                        \item If $[z) \not = \varnothing$.
-                              Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E}) = [<_{c,X+Y}^{\mu}(e)]_{E}$ which is what we need.
-                        \item If $[z) = \varnothing$ and $[f_{E}(z)) = \varnothing$ and $[g_{E}(z)) \not = \varnothing$.
-                              Because $i_{2,E}$ is a homomorphism, we have $<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z)) = i_{2,E}(<_{c,Y}^{\mu}(g_{E}(z))) = i_{2,E}(e_{y})$.
-                              We next have either $e = f_{E};i_{1,E}(z)$ in which case $[e) = \varnothing$ because $i_{1,E}$ is a coproduct injection and \update{we do not require homomorphism-ness for such $e$}.
-                              Or, $e = g_{E};i_{2,E}(z)$ and 
-                              \[
-                              <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]) = [g_{E};<_{c,Y}^{\mu};i_{2,E}(z)]_{E} = [i_{2,E}(e_{y})]
-                              \]
-                              and
-                              \begin{align*}
-                                [<_{c,X+Y}^{\mu}(e)] = [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] &= \\
-                                &\text{because $[g_{E}(z)) \not = \varnothing$ and $i_{2,E}$ is homomorphism}\\
-                                 = [i_{2,E}(<_{c,Y}(g_{E}(z)))] = [i_{2,E}(e_{y})]
-                              \end{align*}
-                              so we get
-                              \[
-                              <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]) = [<_{c,X+Y}^{\mu}(e)]
-                              \]
-                        \item The remaining case is analogous.
+                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $i_{1,V}$ is a coproduct injection, it must be the case that $[i_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
+                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
+                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([v]) = [f_{V};<_{c,X};i_{1,E}(z)] = [i_{1,E}(<_{c,Y}^{\mu}(f_{V}(z)))] = [i_{1,E}(e_y)]$.
+                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [i_{1,E}(<_{c,Y}^{\mu}(f_{V}(z)))] = [i_{1,E}(e_y)]$.
                     \end{itemize}
-                \item If $e$ has no pre-image in $E_{Z}$ then homomorphism-ness of $[]_{E}$ is similar to the first case above.
+                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{V}$ above.
               \end{itemize}
     \end{enumerate}
-    The proof for when $e$ is a node is analogous.
-
+
+    Then $<_{c,\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
+    \question{I probably should've added injections for $<$ as well as it is defined on a coproduct (like I do below for $\conflict$) }
+
+    \update{Check that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive and assymetric and that it respects connectivity.}
+    \begin{itemize}
+        \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}^{\mu}}([e]) = [<_{c,X+Y}^{\mu}(e)]$ and $e = <_{c,X+Y}^{\mu}(e)$ because edges are not identified.
+              This is not possible as $<_{c,X+Y}^{\mu}$ is irreflexive. We need not to check this requirement for nodes because a node can not be a predecessor.
+        \item Let $e_1,e_2$ be edges. 
+              Then, if $[e_1] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_2])$ it must not be the case that $[e_2] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1])$.
+              We have $[e_1] = [<_{c,X+Y}^{\mu}(e_2)]$ and if it was the case that $[e_2] <_{c,\sfrac{X+Y}{\sim}}([e_1])$ then $[e_2] = <_{c,X+Y}^{\mu}(e_1)$ which would mean that $<_{c,X+Y}$ is not assymetric.
+              Once again, we do not need to check for nodes as a node can not be on the lefthand side of the relation.
+        \item We need to show that if $[v] \in s([e_1])$ and $[e_2] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1])$ then $[e_2] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([v])$.
+              We have $e_2 = <_{c,X+Y}^{\mu}(e_1)$ and $[v] \in [s(e_1)]^{*}$ meaning 
+              that there exists $v' \in s(e_1)$ such that $[v] = [v']$.
+              \begin{itemize}
+                        \item $v' = v$.
+Then, given $e_2 = <_{c,X+Y}^{\mu}(e_1)$ it must be the case that $e_2 = <_{c,X+Y}^{\mu}(v)$ and $[e_2] = [<_{c,X+Y}^{\mu}(v)] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([v])$.
+                        \item $v' \sim v$ and there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$ and because $e_2 <_{c,X+Y}^{\mu}(e_1)$ it must be the case that $e_2 <_{c,X+Y}^{\mu}(v')$.
+This also means that necessarily $[g_{V}(z)) \not = \varnothing$ and $[f_{V}(z)) = \varnothing$.
+                              If $[f_{V}(z)) = \varnothing$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([v]) = [g_{V};<_{c,Y};i_{2,E}(z)] = [<_{c,Y}^{\mu};i_{2,E}(g_{V}(z))] = [<_{c,X+Y}^{\mu}(i_{2,V}(g_{B}(z)))] = [<_{c,X+Y}^{\mu}(v')] = [e_2]$.
+                        \item Case when $v' = f_{V};i_{1,V}(z)$ is symmetric.
+               \end{itemize}
+            \end{itemize}
+    Let's now define $\conflict_{\sfrac{X+Y}{\sim}}$.
+    We can consider the original conflict relation as a function $\conflict_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
+    Quotienting the range of the function gives us $\conflict_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote it as $\langle []_{V}^{\conflict} []_{E}^{\conflict} \rangle$) applied to the return value of $\conflict_{X+Y}$.
+    We then define the new relation as
+
+    \[
+        \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}([x]_{E})) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x')))^{\Rightarrow}
+    \]
+    \[
+        \conflict_{\sfrac{X+Y}{\sim}}(i_{V_{X} + V_{Y}}([x]_{V})) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} i_{V_{X} + V_{Y}}(x')))^{\Rightarrow}
+    \]
+        where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$ and $\Rightarrow$ is a closure by (1) from~\ref{def:e-hypergraph-req}.
+        That is, the relation above is the smallest relation that contains $\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x'))$ and such that if $[x_1] <_{c,\sfrac{X+Y}{\sim}}([x_2])$ and $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$ then $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_2] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$.
+        And similarly for nodes.
+
+    Again, we need to check that this is a well-defined function and that our initial $([]_{V},[]_{E})$ respects the conflict.
+
+    \begin{itemize}
+        \item Assume that $t_1, t_2 \in E_{X+Y}$ and $t_1 \sim t_2$ then $\conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}[t_1]) = \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}[t_2])$.
+              $t_1 \sim t_2$ implies there exists $z \in E_{Z}$ such that $t_1 = f_{E};i_{1,E}(z)$ and $t_2 = g_{E};i_{2,E}(z)$
+              \begin{align*}
+                \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}([t_1])) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{t_i \in \{t_1, t_2\}} i_{E_{X} + E_{Y}}(t_i)))^{\Rightarrow}
+                %      \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
+                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
+                %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
+              \end{align*}
+              Similarly,
+              \[
+                \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}([t_2])) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{t_i \in \{t_1, t_2\}} i_{E_{X} + E_{Y}}(t_i)))^{\Rightarrow}
+              \]
+              and we get well-definedness on the nose.
+            %   \begin{align*}
+            %     \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}[t_2]) = \conflict_{X+Y};\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(i_{E_{X} + E_{Y}}(t_2)) = \\
+            %          \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_2))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(g_{E};i_{2,E};i_{E_{X} + E_{Y}}(z)) = \\
+            %          \text{By homomorphism-ness of $g_{E};i_{2,E}$} \\
+            %          = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle g_{V};i_{2,V};i_{V_{X} + V_{Y}} ,g_{E};i_{2,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}(z)))
+            %   \end{align*}
+            %   Then,
+            %   $[f_{E};i_{1,E}(z')]_{E} = [g_{E};i_{2,E}(z')]_{E}$ for all $z' \in \{z'' \in \conflict_{Z}(z) \text{ s.t. } z'' \in E_{Z}\}$
+            %   and
+            %   $[f_{V};i_{1,V}(z')]_{V} = [g_{V};i_{2,V}(z')]_{V}$ for all $z' \in \{z'' \in \conflict_{Z}(z) \text{ s.t. } z'' \in V_{Z}\}$
+            %   hence
+            %   \begin{align*}
+            %     \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle g_{V};i_{2,V};i_{V_{X} + V_{Y}} ,g_{E};i_{2,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z))) = \\
+            %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
+            %   \end{align*}
+        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
+            \[
+             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\conflict_{X+Y}(i_{E_{X+Y}}(x)))
+             \subseteq
+             \conflict_{\mathcal{\sfrac{X+Y}{\sim}}}([]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}(x))
+             \]
+            By definition, 
+            \[
+                \conflict_{\mathcal{\sfrac{X+Y}{\sim}}}([]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}(x)) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle (\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x))))^{\Rightarrow}
+            \]
+            and by recalling that $[]_V^{*} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{*}$ we get the inclusion.
+    \end{itemize}
+    All is left is to check that $\sfrac{X + Y}{\sim}$ is indeed an e-hypergraph.
+    For this we need to check conditions (1) - (7) in~\ref{def:e-hypergraph-req}.
+    % \begin{enumerate}
+    %     \item Assume all $x_1, x_2, x_3$ are in $E_{X+Y}$.
+    %      Given $[x_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([x_2])$ 
+    %      and 
+    %      $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_1]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))$ then it must be the case that
+    %     $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))$.
+
+    %     \[
+    %         i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_1]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))
+    %     \]
+    %     means that there exist $x',x'' \in E_{X} + E_{Y}\; | \;[x'] = [x_3] \text{ and } x'' \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x')) \text{ and } [x''] = [x_1]$.
+    %     We then begin by considering cases for $<$ and $[x_1], [x_2], [x_3], [x']$.
+    %     \begin{itemize}
+    %         \item $[x_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([x_2]) = [<_{c,X+Y}^{\mu}(x_2)]$.
+    %         We have two options.  
+    %             \begin{itemize}
+    %                 \item $x_2$ has a pre-image $z_{x_2}$ in $E_{Z}$ such that $[z_{x_2}) \not = \varnothing$.
+    %                 \begin{itemize}
+    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' = x_3$, $x'' = x_1$ hence $i_{E_{X} + E_{Y}}(x_1) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
+    %                       Then, because $X+Y$ is an e-hypergraph it must be the case that $i_{E_{X} + E_{Y}}(x_2) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
+    %                       And, by homomorphism-ness of $[]$ we get $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]_{E}) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]_{E}))$. 
+    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' \sim x_3$, $x'' = x_1$ hence there exists $z \in E_{Z}$ such that $x' = f_{E};i_{1,E}(z)$ and $x_3 = g_{E};i_{2,E}(z)$.
+    %                       Hence $i_{E_{X} + E_{Y}}(x_1) \in \conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z))$.
+    %                       Then, because $X+Y$ is an e-hypergraph it must be the case that $i_{E_{X} + E_{Y}}(x_2) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z)))$.
+    %                       And, by homomorphism-ness of $[]$ we get $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]_{E}) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([f_{E};i_{1,E}(z)]_{E})) = \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]_{E}))$.
+    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' = x_3$, $x'' \sim x_1$ hence there exists $z \in E_{Z}$ such that $x'' = f_{E};i_{1,E}(z)$ and $x_1 = g_{E};i_{2,E}(z)$.
+    %                       Hence $i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z)) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
+    %                       By definition $x_2 = g_{E};i_{2,E}(z_{x_2})$ (because $x_1$ and $x_2$ should have pre-image in either $X$ or $Y$, not both).
+    %                       This means there exists $z_2 <_{c,Z}^{\mu}(z_{x_2})$ and $g_{E};i_{2,E}(z_2) <_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z_{x_2})) = <_{c,X+Y}^{\mu}(x_2)$ and because there can be only one imemdiate predecessor we have $x_1 = g_{E};i_{2,E}(z_2)$ and $z = z_2$.
+    %                       \question{What next?}
+    %                       Because $X+Y$ is an e-hypergraph, we have $i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z')) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
+    %                       Then, by homomorphism-ness we get $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([f_{E};i_{1,E}(z')]) = i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))$.
+    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' \sim x_3$, $x'' \sim x_1$ hence there exist $z_1 \in E_{Z}$ such that $x'' = f_{E};i_{1,E}(z_1)$ and $x_1 = g_{E};i_{2,E}(z_1)$ and $z_2 \in E_{Z}$ such that $x' = f_{E};i_{1,E}(z_2)$ and $x_3 = g_{E};i_{2,E}(z_2)$.
+    %                       Hence, $i_{E_{X} + E_{Y}}(x'') \in \conflict_{X+Y}(i_{E_{X} + E_{Y}})(x')$ and $i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z_1)) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}})(f_{E};i_{1,E}(z_2))$.
+    %                       There also should exist $z_3$ such that $z_1 <_{c,Z}^{\mu}(z_3)$ and $g_{E};i_{2,E}(z_1) <_{c,Z}^{\mu}(g_{E};i_{2,E}(z_3))$
+    %             \end{itemize}
+    %             $x_2$ has no pre-image in $E_{Z}$.
+    %         \end{itemize}
+    %         \end{itemize}
+    % \end{enumerate}
+
+    \begin{enumerate}
+        \item (1) is given by construction.
+        \item There are separate cases for nodes and edges.
+              \begin{itemize}
+                \item Being not mutually exclusive means it is possible that $[e]_{E} <_{c,\sfrac{X+Y}{\sim}}([v]_{V})$ and $i_{\sfrac{V_{X}+V_{Y}}{\sim}}[v] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}[e])$.
+                \item Or, $[e_1]_{E} <_{c,\sfrac{X+Y}{\sim}}([e_2]_{E})$ and $i_{\sfrac{E_{X}+E_{Y}}{\sim}}[e_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}[e_2])$.
+               \end{itemize}
+              The first means that either
+              \begin{itemize}
+                \item $[e] = [<_{c,X+Y}(v)]$ if $v$ has no pre-image in $V_{Z}$ and there exist $v'$ and $e'$  such that $[v'] = [v]$, $[e'] = [e]$ and $i_{V_{X} + V_{Y}}(v') \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e'))$.
+                      As there is no pre-image in $V_{Z}$, $v = v'$. Then we have the following case.
+                    \begin{itemize}
+                        \item $e = <_{c,X+Y}(v)$ and $e' = e$.
+                              As we have $i_{V_{X} + V_{Y}}(v) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e))$ and this is not possible as $X+Y$ is an e-hypergraph.
+                    \end{itemize}
+                \item $v$ has a pre-image $z \in V_{Z}$, let's say, $f_{V};i_{1,V}(z) = v$, and $[f_{V}(z)) = \varnothing$.
+                      Then, either $v' = v$ and $i_{V_{X} + V_{Y}}(v) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e))$, which is not possible as it requires that $v$ has a predecessor in $X$.
+                      Or, $v' \sim v$ (meaning that $v' = g_{V};i_{2,V}(z)$) and $i_{V_{X} + V_{Y}}(v') \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e))$.
+                      Then, $<_{c,\sfrac{X+Y}{\sim}}([v]) = [g_{V};<_{c,Y};i_{2,E}(z)] = [e]$. Once again, it means that $e = g_{V};<_{c,Y};i_{2,E}(z)$.
+                      Overall, it means that $g_{V}(z) \in \conflict_{Y}(i_{E_{Y}}(g_{V};<_{c,Y}(z)))$ which is not possible as $Y$ is an e-hypergraph.
+                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+              \end{itemize}
+    \end{enumerate}
     Let's now prove the universal property.

### 2023-11-19 sections/combinatorial_semantics.tex
Parent relation for pushout is now proven correct

@@
 
+\begin{remark}
+    Both relations can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, i.e. on the coproduct of nodes and edges.
+    To make things well-typed, we will use corresponding coproduct injections $i_{V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $i_{E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either node or edge into these functions.
+    For example, an immediate successor of a node $x$ can be written functionally as $<_{c,\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x))$.
+\end{remark}
+
 \begin{definition}{E-hypergraphs homomorphism}
@@
 
-    \item $\phi(<_{c,\mathcal{F}}^{\mu}(x)) = <_{c,\mathcal{G}}^{\mu}(\phi(x))$.
+    \item \[
+        \phi_{E}(<_{c,\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x))) = <_{c,\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+        \] when $x$ is a node,
+        and
+        \[
+        \phi_{E}(<_{c,\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x))) = <_{c,\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))  
+        \] when $x$ is an edge.
           \update{We do not require such an equation when $[x) = \varnothing$}.
@@
 
-    First, we assume that $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
+    \begin{itemize}
+    \item First, we assume that $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
+    \item Second, we assume that $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$.
+\update{This essentially means that $f,g$ do not map nodes to levels of different hierarchy.}
+    \label{pushout:assumptions}
+    \end{itemize}
 
@@
     where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$ and we will further omit writing the relations explicitly.
-    Recall that $<_{c}$ is essentially a transitive clousure of $<_{c}^{\mu}$ and hence we can first define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]_{E})$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]_{V})$.
+    Recall that $<_{c}$ is essentially a transitive clousure of $<_{c}^{\mu}$ and hence we can first define 
+    \[<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
+    \] and 
+    \[
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
+    \] where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
     For this we will consider several cases where \update{$f$ and $g$ are mono. I require mono just to simplify things, we can get rid of this requirement if everything else is ok.}.
-    Assume that $u$ is an element of $V_{X+Y}$.
+    Assume that $u$ is an element of $V_{X+Y}$ and
     \begin{enumerate}
@@
         \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$ $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(g_{V}(z))$.
-              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [g_{V};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]$.
+              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z)))$.
+              We then define 
+              \[
+                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]
+              \]
         \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(f_{V}(z))$.
-              We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [f_{V};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]$.
-        \item $u$ has no pre-image in $Z$ and then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
+              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(i_{V_{X}}(f_{V}(z)))$.
+              We then define 
+              \[
+                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]
+              \]
+        \item $u$ has no pre-image in $Z$ and then 
+        \[
+            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))]
+        \]
     \end{enumerate}
@@
         \item We have the only case of $u$ having no pre-image in $E_{Z}$ because its empty.
-              Hence, we define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$.
+              Consider two cases.
+              \begin{itemize}
+                \item  $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u)) = u_1$.
+                        Then we define
+                    \[
+                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u))]
+                    \]
+                \item $[u) = \varnothing$ and for all $v_i \in s(u)$ and $v_j \in t(u)$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
+                      Then we define 
+                    \[
+                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                    \] where $v$ is either in $s([u])$ or $t([u])$.
+                     The second assumption in~\ref{pushout:assumptions} guarantees that all candidates for $[v]$ share the same predecessor, i.e. the definition does not depend on a particular $[v]$.
+              \end{itemize}
     \end{enumerate}
-    \question{I decided to keep Z discrete, otherwise the number of cases to consider within the proofs is too large.
-    This also means that we now can identify only nodes.
-    I will make changes to account for this later.
-    }
+
+    \begin{remark}
+        Essentially, what we build above is that $[e] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ when there exists $v' \in V_{X} + V_{Y}$ such that $[v'] = [v]$ and $e = <_{c,X+Y}^{\mu}(v')$.
+    \end{remark}
     \update{If $f$ and $g$ are not mono, the definitions stay the same as $i_{2,E}$ and $i_{1,E}$ are mono.}
     Clearly, all the cases above are disjoint.
+
     Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are nodes and $t_1 \sim t_2$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$ and either $[f_{V}(t_1)) = \varnothing$ or $[g_{V}(z)) = \varnothing$.
-    Assume the latter, then, by definition.
-    \[
-        <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_1]) = [f_{V};<_{c,X}^{\mu};i_{1,E}(z)]    
+    Suppose $t_1, t_2$ are nodes and $t_1 \sim t_2$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
+    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$ and two cases
+    \begin{itemize}
+        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(t_2)) = \varnothing$
+        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(t_2)) \not = \varnothing$
+    \end{itemize} 
+    Consider the first case, then, by definition.
+    \[
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
     \]
@@
     \[
-        <_{c,\sfrac{X+Y}{\sim}}^{\mu}([t_2]) = [f_{V};<_{c,X}^{\mu};i_{1,E}(z)]    
-    \]
-    so we have well-defined-ness on the nose.
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
+    \]
+    so we have well-defined-ness on the nose. We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
 
@@
               This is the same as in $\catname{Hyp_{\Sigma}}$.
-        \item $[<_{c,X+Y}^{\mu}(e)]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e]_{E})$ for $e$ such that $[e) \not = \varnothing$.
+        \item $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
+              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
               This is homomorphic by definition. 
-        \item $[<_{c,X+Y}^{\mu}(v)]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([v]_{V})$.
+        \item $[<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
               Let's make case analysis on $[v]_{V}$.
@@
                         \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([v]) = [f_{V};<_{c,X};i_{1,E}(z)] = [i_{1,E}(<_{c,Y}^{\mu}(f_{V}(z)))] = [i_{1,E}(e_y)]$.
-                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [i_{1,E}(<_{c,Y}^{\mu}(f_{V}(z)))] = [i_{1,E}(e_y)]$.
+                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}[v]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
+                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V};i_{V_{X} + V_{Y}}(z))] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
                     \end{itemize}
                 \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{V}$ above.
+                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
               \end{itemize}
@@
     \begin{itemize}
-        \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}^{\mu}}([e]) = [<_{c,X+Y}^{\mu}(e)]$ and $e = <_{c,X+Y}^{\mu}(e)$ because edges are not identified.
-              This is not possible as $<_{c,X+Y}^{\mu}$ is irreflexive. We need not to check this requirement for nodes because a node can not be a predecessor.
+        \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}^{\mu}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
+              Then we have two cases.
+              \begin{itemize}
+                \item Assume $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{c,\sfrac{X+Y}{\sim}^{\mu}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
+                      This would mean that $e_1 = e$ which would mean that the relation $<_{c,X+Y}^{\mu}$ is not irreflexive.
+                \item Assume $[e) = \varnothing$ and for all $v_{i} \in s(e)$ and $v_{j} \in t(e)$ $[[v_i]) \not = \varnothing$ $[[v_j]) \not = \varnothing$.
+                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                      \begin{itemize}
+                        \item If $v$ has no pre-image in $V_{Z}$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e]$.
+                              This would mean that $e$ is a predecessor of $v$ in $X+Y$, but $v$ is either in the sources or targets of $e$ which violates respectful connectivity condition.
+                        \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ such that $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$.
+                              Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{1,E}(e_x)] = [e]$.
+                              This means that $e$ has a pre-image in $E_{X}$, but because $v$ also has a pre-image in $V_{X}$ and its set of predecessors is not empty it would mean that $e$ has a predecessor which it does not.
+                        \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ such that $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$.
+                              Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)] = [e]$.
+                              It means that $e$ has a pre-image in $E_{Y}$, but $v$ is its source or target and has a pre-image in $V_{Y}$ which is not possible.
+                        \item The remaining two cases for $v = g_{V};i_{2,V}(z)$ are symmetric.
+                      \end{itemize}
+              \end{itemize}$[<_{c,X+Y}^{\mu}(e)]$ and $e = <_{c,X+Y}^{\mu}(e)$ because edges are not identified.
+                We do not need to check this requirement for nodes because a node can not be a predecessor.
         \item Let $e_1,e_2$ be edges. 
-              Then, if $[e_1] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_2])$ it must not be the case that $[e_2] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1])$.
-              We have $[e_1] = [<_{c,X+Y}^{\mu}(e_2)]$ and if it was the case that $[e_2] <_{c,\sfrac{X+Y}{\sim}}([e_1])$ then $[e_2] = <_{c,X+Y}^{\mu}(e_1)$ which would mean that $<_{c,X+Y}$ is not assymetric.
-              Once again, we do not need to check for nodes as a node can not be on the lefthand side of the relation.
-        \item We need to show that if $[v] \in s([e_1])$ and $[e_2] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1])$ then $[e_2] <_{c,\sfrac{X+Y}{\sim}}^{\mu}([v])$.
-              We have $e_2 = <_{c,X+Y}^{\mu}(e_1)$ and $[v] \in [s(e_1)]^{*}$ meaning 
-              that there exists $v' \in s(e_1)$ such that $[v] = [v']$.
+              Then, if $[e_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+              For $[e_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
               \begin{itemize}
-                        \item $v' = v$.
-Then, given $e_2 = <_{c,X+Y}^{\mu}(e_1)$ it must be the case that $e_2 = <_{c,X+Y}^{\mu}(v)$ and $[e_2] = [<_{c,X+Y}^{\mu}(v)] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([v])$.
-                        \item $v' \sim v$ and there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$ and because $e_2 <_{c,X+Y}^{\mu}(e_1)$ it must be the case that $e_2 <_{c,X+Y}^{\mu}(v')$.
-This also means that necessarily $[g_{V}(z)) \not = \varnothing$ and $[f_{V}(z)) = \varnothing$.
-                              If $[f_{V}(z)) = \varnothing$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([v]) = [g_{V};<_{c,Y};i_{2,E}(z)] = [<_{c,Y}^{\mu};i_{2,E}(g_{V}(z))] = [<_{c,X+Y}^{\mu}(i_{2,V}(g_{B}(z)))] = [<_{c,X+Y}^{\mu}(v')] = [e_2]$.
-                        \item Case when $v' = f_{V};i_{1,V}(z)$ is symmetric.
+                \item If $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
+                      Having at the same time $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
+                \begin{itemize}
+                    \item $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
+                     This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
+                    \item $[e_1) = \varnothing$ and for all $v_i \in s(e_1)$ and $v_j \in t(e_1)$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
+                          Then, $[e_2] = [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
+                          \begin{itemize}
+                            \item If $v$ has no pre-image in $V_{Z}$, then 
+                            \[
+                                [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e_2]
+                            \]
+                                  It would mean that $e_2$ is a predecessor of $v$ and, by connectivity, a predecessor of $e_1$.
+                                  However, $[e_1) = \varnothing$ and we get a contradiction.
+                            \item If $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
+                                  Then,
+                                  \[
+                                    [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [i_{2,E}(e_y)] = [e_2]
+                                  \]
+                                  This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is a predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
+                            \item Case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric to the above.
                           \end{itemize}
                     \end{itemize}
+                    \item $[e_2) = \varnothing$ and for all $v_i \in s(e_2)$ and $v_j \in t(e_2)$ $[[v_i]) \not = \varnothing$ $[[v_j]) \not = \varnothing$.
+                          Then, 
+                          \[
+                            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e_1]
+                          \]
+                          The cases when $[e_1) \not = \varnothing$ are similar to the above, so we will just jump to the interesting case when 
+                          $[e_1) = \varnothing$ and
+                          \[
+                            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]
+                          \]
+                          The cases we have are the following.
+                          \begin{itemize}
+                            \item $v$ has no pre-image in $V_{Z}$. 
+                                  Then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e_1]$
+                                  This implies that $e_1$ is the predecessor of $e_2$ because it is the predecessor of $v$ and $v$ is $e_2$'s sources of targets.
+                                  \begin{itemize}
+                                    \item $v'$ has no pre-image in $V_{Z}$.
+                                    Then $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [e_2]$.
+                                    This implies that $e_2$ is the predecessor of $v'$ and hence of $e_1$. 
+                                    This contradicts the assymetry of $<_{c,X+Y}^{\mu}$.
+                                    \item $v'$ has a pre-image $f_{V};i_{1,E}(z) = v'$ in $V_{Z}$ (and so should $e_1$ do) and therefore $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$ because $[e_1) = \varnothing$.
+                                          Then, $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [i_{2,E}(e_y)] = [e_2]$.
+                                          This means that $e_2$ has a pre-image in $E_{Y}$ but $e_1$ is the predecessor of $e_2$ and hence they should both have pre-image in $E_{X}$ (as $e_1$ does so).
+                                    \item Case when $v' = g_{V};i_{2,V}(z)$ is symmetric.
+                                    \end{itemize}
+                            \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ (and so should $e_2$ do) in $Z_{V}$ and therefore $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$.
+                                  Then, $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_{y})] = [e_1]$.
+                                  This implies that $e_1$ has a pre-image in $E_{Y}$.
+                                  \begin{itemize}
+                                    \item $v'$ has no pre-image in $V_{Z}$.
+                                          Then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [e_2]$.
+                                          This implies that $e_2$ is the predecessor of $v'$ and hence of $e_1$.
+                                          But $[e_1) = \varnothing$ and hence we have a contradiction.
+                                    \item $v'$ has a pre-image $f_{V};i_{1,V}(z') = v'$ in $V_{Z}$ (and so should $e_1$ do) and therefore $[f_{V}(z')) = \varnothing$ and $[g_{V}(z')) \not = \varnothing$.
+                                          This also means that $[g_{V}(z)) = [g_{V}(z'))$.
+                                          By definition we have that $e_1$ is the image of the predecessor of $g_{V}(z')$ and $e_2$ is the image of the predecessor of $g_{V}(z)$ and because their predecessors are equal it implies $e_1 = e_2$ but the relation is irreflexive.
+                                    \item $v'$ has a pre-image $g_{V};i_{2,V}(z') = v'$ in $V_{Z}$ and therefore $[f_{V}(z')) \not = \varnothing$ and $[g_{V}(z')) = \varnothing$.
+                                          This is impossible as it violates the requirement that $[f_{V}(z)) = [f_{V}(z'))$ for all $z,z' \in V_{Z}$.
+                                \end{itemize}
+                            \item The case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric.
+                            \end{itemize}
+              \end{itemize} 
+              Once again, we do not need to check for nodes as a node can not be on the lefthand side of the relation.
+        \item We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
+              $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us two cases.
+              \begin{itemize}
+                \item If $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
+                      $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
+                      \begin{itemize}
+                        \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
+                            That is, $e_2 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                        \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$.
+                              Once again, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$ and 
+                              \[
+                                [e_2] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+                              \]
+                        \end{itemize}
+                \item $[e_1) = \varnothing$ and for all $v_i \in s(e_1), v_j \in t(e_1)$ $[[v_i]) \not = \varnothing,\; [[v_j]) \not = \varnothing$. 
+                Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e_2]$ and this gives us the needed property on the nose.
+              \end{itemize}
+        \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+        Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $[v'] = [v]$.
+        \begin{itemize}
+            \item $v' = v$ and hence $v \in s(e_1)$.
+            \begin{itemize}
+                \item If $v$ has no pre-image in $V_{Z}$, then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]$.
+                      This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ which is what we want.
+                \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
+                \[ 
+                    [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{1,E}(e_x)]
+                \]
+                Because $e_{x}$ is the predecessor of $f_{V}(z)$ it should also be the predecessor of the pre-image of $e_1$ in $E_{X}$ as $v \in s(e_1)$.
+                Denote the pre-image of $e_1$ in $E_{X}$ as $e_{1}^{x}$. 
+                Then, $e_{x} = <_{c,X}^{\mu}(i_{E_{X}}(e_{1}^{x}))$ and
+
+                \begin{align*}
+                    [i_{1,E}(e_{x})] &= [i_{1,E}(<_{c,X}^{\mu}(i_{E_{X}}(e_{1}^{x})))] = \\
+                    &= [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(i_{1,E}(e_{1}^{x})))] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = \\
+                    &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [e_2]
+                \end{align*}
+                \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
+                \[ 
+                    [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]
+                \]
+                This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ necessarily has a pre-image in $E_{X}$.
+                Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [i_{2,E}(e_y)] = [e_2]$.
+                \item Cases when $v$ has a pre-image $g_{V};i_{2,V}(z)$ are symmetric.
+            \end{itemize}
+            \item $v \sim v'$ and hence there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$.
+                  This implies that $e_1$ has a pre-image in $E_{Y}$ and
+                  \[
+                    [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+                  \]
+                  \begin{itemize}
+                    \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$ which implies $[g_{V};i_{2,V}(z)) = [v') \not = \varnothing$.
+                    That is, there exists $e_3 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$ and by e-hypergraphness of $X+Y$ $e_3 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$.
+                    Then, 
+                    \begin{align*}
+                    [e_3] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = \\
+                    &= [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = \\
+                    &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V{Y}}{\sim}}([v]))
+                \end{align*}
+                \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$ which implies $[g_{V};i_{2,V}(z)) = [v') = \varnothing$ and hence $[e_1) = \varnothing$.
+                      By definition,
+                      \begin{align*}
+                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v'])) = \\
+                        &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v])) = [e_2]
+                    \end{align*}
+                \end{itemize}
+        \item Cases when $v = g_{V};i_{2,V}(z)$ are symmetric.
+        \end{itemize}    
+    \end{itemize}
+    \update{This will be changed to consistency relation, but the definition of quotiening will stay the same.}
     Let's now define $\conflict_{\sfrac{X+Y}{\sim}}$.

### 2023-11-28 sections/combinatorial_semantics.tex
Refined parent and consistency relations WIP

@@
     \update{We will denote the fact that $e$ is an immediate predecessor of $v$ as $e <_c^{\mu} v$.}
-    % , which essentially makes $<_{c}$ into a function that for any given $x$ returns its immediate predecessor.
-    Such order was called a \textbf{well-founded arborescent} strict partial order in~\cite{orders_and_events}.
     Similarly, a set of all successors is $(x] = \{x' \in (V_\mathcal{G} + E_\mathcal{G}) \; ||\; x <_{c} x'\}$ where $x'$ can be a node or an edge. 
@@
 
-    $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
+    Then, consider a relation $\consistency^{\mu} \subseteq (V_{\mathcal{G}} + E_{\mathcal{G}}) \times (V_{\mathcal{G}} + E_{\mathcal{G}})$ which we will call a \textit{consistency} relation and which has the following properties.
     \begin{enumerate}
-        \label{def:e-hypergraph-req}
-        \item 
-        \[
-    \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
-    \] 
-    if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
-    \item $<_c$ and $\#$ are mutually exclusive
-    \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
-    \item $\#$ should respect connectivity
-    \item \update{Complement of $\conflict_{\mu}$ should be transitive}
-    \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
-    \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
+        \label{def:consistency_properties}
+        \item For each $v \in V_{\mathcal{G}}$ $v \consistency^{\mu} v$
+        \item For each $e \in E_{\mathcal{G}}$ $e \consistency^{\mu} e$
+        \item If $v \in s(e)$ then $v \consistency^{\mu} e$
+        \item If $v \in t(e)$ then $v \consistency^{\mu} e$
+        \item The relation is symmetric and transitive
+        \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
     \end{enumerate}
 
-The last two requirements are known as \textit{confusion freeness}~\cite{ConfusionFreeEvents}.
-\end{definition}
-
-\Aleksei{Switch to the complent in the definition of the relation and call it a \textit{consistency} relation}
-
-Intuitively, the child and conflict relation together make it possible to model e-classes in our e-hypergraphs: an unlabelled edge $e$ represents an e-class where the children of $e$ that are in immediate conflict correspond to nodes and edges in different e-class instances.
-Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$.
-Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.
-
-\begin{lemma}[Complement of $\#$ is equivalence relation]
-\begin{proof}
-As $\#$ is irreflexive then its complement $\bar{\#}$ is reflexive because $\forall x \; .
+%     $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
+%     \begin{enumerate}
+%         \label{def:e-hypergraph-req}
+%         \item 
+%         \[
+%     \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
+%     \] 
+%     if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
+%     \item $<_c$ and $\#$ are mutually exclusive
+%     \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
+%     \item $\#$ should respect connectivity
+%     \item \update{Complement of $\conflict_{\mu}$ should be transitive}
+%     \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
+%     \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
+%         \end{enumerate}
+
+% The last two requirements are known as \textit{confusion freeness}~\cite{ConfusionFreeEvents}.
+\end{definition}
+
+% Intuitively, the child and conflict relation together make it possible to model e-classes in our e-hypergraphs: an unlabelled edge $e$ represents an e-class where the children of $e$ that are in immediate conflict correspond to nodes and edges in different e-class instances.
+% Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$.
+% Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.
+
+% \begin{lemma}[Complement of $\#$ is equivalence relation]
+% \begin{proof}
+% As $\#$ is irreflexive then its complement $\bar{\#}$ is reflexive because $\forall x \; .
 \; (x,x) \not \in \#$.
 Because $\#$ is symmetric so is $\bar{\#}$, otherwise if $(x_1, x_2) \in \bar{\#}$ and $(x_2,x_1) \not \in \bar{\#}$ then $(x_2,x_1) \in {\#}$ which is symmetric and hence $(x_1,x_2)$ should be in ${\#}$.
-Finally, the complement is transitive because of (5).
-
-\end{proof}
-\end{lemma}
+% Finally, the complement is transitive because of (5).
+
+% \end{proof}
+% \end{lemma}
 
 \begin{remark}
-    We will further call edges $e$ such that $l(e) = \bot$ \textit{hierarchical}. We will also refer to edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ as top-level.
+    We will further call edges $e$ such that $l(e) = \bot$ \textit{hierarchical}.
+    We will also refer to edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ as top-level.
 \end{remark}
 
+\begin{remark}
+    Because $\consistency^{\mu}$ respects connectivity and is reflexive, symmetric and transitive, it is essentially a relation between connected components that have the same predecessor.
+    This notion will be helpful in the following example.
+\end{remark}
 % We associate an e-class with a hierarchical edge, i.e. an edge with children. 
@@
 \begin{example}
+
+A few examples of e-hypergraphs (and their corresponding graphical interpretation) can be seen in Figure~\ref{fig:e-hypergraph-example}.
+In example~\ref{fig:example_a} the relations $<_{c}$ and $\consistency^{\mu}$ are the following
+\[
+    <_{c} = \{(v_3,e_1), (e_2, e_1), (v_4, e_1) \}
+\] and 
+\begin{align*}
+    \consistency^{\mu} = \{ (v_1,e_1),
+                            (v_2, e_1), 
+                            (v_3,e_2), (v_4,e_2)\}^{c}
+\end{align*}
+where $c$ denotes the reflexive, syymetric and transitive closure.
+In other words, $e_1$ is the predecessor of the connected component that contains edge $e_2$ and this connected component is consistent with itself only.
+The connected component that contains $e_1$ is also consistent with itself only.
+
+In the next example~\ref{fig:example_b} $e_1$ is the predecessor of two connected components but these components are consistent only with themselves.
+We denote this fact by using a vertical line. 
+Practically, hierarchical edges such as $e_1$ represent e-classes and connected components which are pairwise not consistent belong to different instances of the same e-class.
+
+In the last example~\ref{fig:example_c} two connected components that contain $e_2$ and $e_3$ respectively are consistent with themselves and with each other.
+\end{example}
+
+\begin{figure}[t]
+    \begin{subfigure}{0.3\linewidth}
+        \[
+            \scalebox{0.6}{
+                \tikzfig{combinatorial_semantics/e-hypergraph-example-1}
+            }
+            \]
+        \caption{}
+        \label{fig:example_a}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.3\linewidth}
+        \[
+          \scalebox{0.6}{
+            \tikzfig{combinatorial_semantics/e-hypergraph-example-2}
+          }  
+        \]
+        \caption{}
+        \label{fig:example_b}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.3\linewidth}
+        \[
+          \scalebox{0.6}{
+            \tikzfig{combinatorial_semantics/e-hypergraph-example-3}
+          }  
+        \]
+        \caption{}
+        \label{fig:example_c}
+    \end{subfigure}
+    \caption{E-hypergraph examples}
+    \label{fig:e-hypergraph-example}
+\end{figure}
+
+\begin{example}
 An example of an e-hypergraph can be found in Figure~\ref{fig:e-cospan}.
@@
         \item
-    if $x_1 \#_{\mathcal{F}} x_2$ then $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
+    if $x_1 \consistency_{\mathcal{F}}^{\mu} x_2$ then $\phi(x_1) \consistency_{\mathcal{G}}^{\mu} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
     Or, written functionally, when $x \in E_{\mathcal{F}}$
     \[
-        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x)))
+        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
         \subseteq
-        \conflict_{\mathcal{G}}(\phi_{E};i_{E_{\mathcal{G}}}(x))
+        \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
     \]
@@
     \[
-        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\conflict_{\mathcal{F}}(i_{V_{\mathcal{F}}}(x)))
+        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
         \subseteq
-        \conflict_{\mathcal{G}}(\phi_{V};i_{V_{\mathcal{G}}}(x))
-    \]
-    Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
+        \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+    \]
+    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
 \end{enumerate}
@@
 
-\begin{remark}
+\begin{definition}[Pushout in $\catname{EHyp_{\Sigma}}$]
 
@@
     \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \#_{X+Y}, <_{c,X+Y} \}
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{c,X+Y} \}
     \]
@@
               \]
-        \item $u$ has no pre-image in $Z$ and then 
-        \[
-            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))]
-        \]
+        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u)) = e'$
+        \[
+            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))] = [e']
+        \]
+        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))$ is undefined (i.e.
+$[u) = \varnothing$), but there exists $[e]$ such that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e']$ and $[u] \in t([e])$ or $[u] \in s([e])$, then
+              \[ 
+                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
+              \]
+              \update{Assumptions above guarantee that the predecessor of such node does not depend on the choice of $e$ if there are several of them.}
     \end{enumerate}
@@
     Now, if $u$ is an element of $E_{X+Y}$.
+    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
+              Consider two cases.
               \begin{enumerate}
-        \item We have the only case of $u$ having no pre-image in $E_{Z}$ because its empty.
-              Consider two cases.
-              \begin{itemize}
                 \item  $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u)) = u_1$.
@@
                     \]
-                \item $[u) = \varnothing$ and for all $v_i \in s(u)$ and $v_j \in t(u)$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
+                \item $[u) = \varnothing$ and for some $[v_i] \in s([u])$ and $[v_j] \in t([u])$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
                       Then we define 
@@
                         <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                    \] where $v$ is either in $s([u])$ or $t([u])$.
+                    \] where $[v]$ is either in $s([u])$ or $t([u])$.
                      The second assumption in~\ref{pushout:assumptions} guarantees that all candidates for $[v]$ share the same predecessor, i.e. the definition does not depend on a particular $[v]$.
-              \end{itemize}
               \end{enumerate}
+    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then steps (1) - (2) for edges and finally, step (4) for nodes.}
 
     \begin{remark}
-        Essentially, what we build above is that $[e] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ when there exists $v' \in V_{X} + V_{Y}$ such that $[v'] = [v]$ and $e = <_{c,X+Y}^{\mu}(v')$.
+        Essentially, what we build above is that if the relation is undefined, it is inherited from either incident nodes or edges.
     \end{remark}
@@
     \end{enumerate}
-
     Then $<_{c,\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
-    \question{I probably should've added injections for $<$ as well as it is defined on a coproduct (like I do below for $\conflict$) }
-
-    \update{Check that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive and assymetric and that it respects connectivity.}
+
+    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$.
+    We can consider the original consistency relation as a function $\consistency_{X+Y}^{\mu} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
+    Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
+    We then define the new relation for edges as
+    \begin{enumerate}
+        \item If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
+                \begin{align*}
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) =
+                    (\langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))))^{c}
+                \end{align*}
+        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V}))
+        \]
+    \end{enumerate}
+
+    and for nodes
+    \begin{enumerate}
+
+    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
+    \]
+    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E}))
+    \]
+    \end{enumerate}
+        where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$ and $c$ is a reflexive, symmetric and transitive closure.
+        % That is, the relation above is the smallest relation that contains $\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x'))$ and such that if $[x_1] <_{c,\sfrac{X+Y}{\sim}}([x_2])$ and $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$ then $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_2] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$.
+        % And similarly for nodes.
+
+    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency^{\mu}$.
+
+    \begin{itemize}
+        \item 
+        \begin{itemize}
+
+            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_2])$.
+              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};i_{1,V}(z)$ and $v_2 = g_{V};i_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
+              \begin{align*}
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+                %      \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
+                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
+                %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
+              \end{align*}
+              Similarly,
+              \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+              \]
+              and we get well-definedness on the nose.
+              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
+            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$ and there exists $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ such that $[v_1] \in s([e])$ (and automatically $[v_2] \in s([e])$).
+                Then,
+                  \[
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                  \]
+            \end{itemize}
+        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
+            \[
+             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
+             \subseteq
+             \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))
+             \]
+            By definition, 
+            \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x))))^{c}
+            \]
+            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion.
+            If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness is true because empty set is a subset of any set.
+            Similarly for edges.
+    \end{itemize}
+
+    Let's now prove the universal property.
+    Let $u_{E}([i_{1,E}(x)]) = j_{1,E}(x)$ and $u_{E}([i_{2,E}(y)]) = j_2(y)$.
+    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,E}$ and $j_{2,E}$.
+
+    Let's say that $t_1 \sim t_2$ then $u_{E}([t_1]) = u_{E}([t_2])$.
+    $u_{E}([t_1]) = u_{E}(f_{E};i_{1,E}(z)) = j_1{f_{E}(z)}$ and $u_{E}([t_2]) = u_{E}(g_{E};i_{2,E}(z)) = j_{2}(g_{E}(z))$ which are equal by commutativity.
+
+\end{definition}
+
+\begin{lemma}
+    Suppose $[e) = \varnothing$ and there exists $[v_1] \in s([e])$ and $[v_2] \in s([e])$ such that $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2]))$.
+    \begin{proof}
+        Because we build the relation in steps, $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$ mean that was achieved after running steps (1)-(3) for nodes.
+        From $[v_1] \in s([e])$ we get that there exists $v'_1$ such that $v_1 \sim v'_1$ and $v'_1 \in s(e)$ and similarly $v'_2 \sim v_2$.
+        \begin{itemize}
+            \item $v_1 = v'_1$ and $v_1 \in s(e)$, and $v_2 = v'_2$ $v_2 \in s(e)$.
+            \begin{itemize}
+                \item $v_1$ has a pre-image $f_{V};i_{1,V}(z)$ and hence $[f_{V}(z)) = \varnothing$ (because $[e) = \varnothing$) and $v_2$ has a pre-image $f_{V};i_{1,V}(z')$.
+                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{2,E}(e_{y})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e'_{y})]$.
+                      Because $e_{y} = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z)))$ and $e'_{y} = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z')))$ we have $e_{y} = e'_{y}$ because under our assumptions nodes in the image of $g_{V}$ should share the same predecessor.
+                      Hence, $[i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})]$.
+                \item Case for $v_1 = g_{V};i_{2,V}(z)$ is symmetric.
+
+                It can not be the case that the node has no pre-image as $[[v_i]) \not = \varnothing$ and $[e) = \varnothing$ implies that step (4) is needed but we only completed steps (1)-(3).
+            \end{itemize}
+            \item $v_1 \not = v'_1$ and $v'_2 \not = v_2$ and hence $v'_1 \in s([e])$ and $v'_2 \in s(e)$.
+            \begin{itemize}
+                \item $v'_1$ has a pre-image $f_{V};i_{1,V}(z)$ and hence $[f_{V}(z)) = \varnothing$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) = [i_{2,E}(e_{y})]$.
+                      $v'_2$ has a pre-image $f_{V};i_{1,V}(z')$ and for the same reason 
+                      \[
+                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) = [i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                      \]
+            \end{itemize}
+            \item The remaining cases are similar.
+        \end{itemize}
+    \end{proof}
+\end{lemma}
+
+\begin{lemma}
+    Suppose $[v) = \varnothing$ and there exist $e_1$ and $e_2$ such that $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ and $[v] \in s([e_1])$, $[v] \in s([e_2])$.
+    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$
+    \begin{proof}
+        Once again, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.
+        We then have the following cases.
+        \begin{itemize}
+            \item $v \in s(e_1)$ and $v \in s(e_2)$.
+                  The only possible case for $e_1$ and $e_2$ is that $[e_1) = \varnothing$ and $[e_2) = \varnothing$ (because $[v) = \varnothing$).
+                  Then, 
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
+                  \], where $[v_1] \in s([e_1])$ and $[[v_1]) \not = \varnothing$.
+                  And 
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                  \], where $[v_2] \in s([e_2])$ and $[[v_2]) \not = \varnothing$.
+                  The cases are following.
+                  \begin{itemize}
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and 
+                          $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{1,E}(e'_{x})]$. 
+                          Then, $e'_{x} = e_{x}$ because the nodes in the image of $f$ should have the same predecessor.
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and 
+                          $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_{y})]$
+                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$. The only possible case then is that $v_2 \in s(e_2)$ and $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$.
+                          Then $[v_1') = \varnothing$ and $[v_2) = \varnothing$. Show it is not possible.
+                          \begin{itemize}
+                            \item $v_1 = f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$.
+                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
+                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it cant be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}([v_2])}) = [i_{2,E}(e_y)]$.
+                                  If $v'_2 \in s(e_2)$ and $v_2 \sim v'_2$ then $v'_2$ has a pre-image in $V_{Y}$ and $v_2$ has a pre-image in $V_{X}$, i.e. $v_2 = f_{V};i_{1,V}(z_2)$. 
+                                  The equation above implies that $[g_{V}(z_2)) \not = \varnothing$ which is not the case as $[v'_2) = \varnothing$.
+                            \item $v_1 = g_{V};i_{2,V}(z_1)$ and $[g_{V}(z_1)) = \varnothing$. 
+                                  Then, necessarily, $v_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = f_{V};i_{1,V}(z_1)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
+                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$, i.e.
+$v_2 = g_{V};i_{2,V}(z_2)$ and then it should be $[g_{V}(z_2)) \not = \varnothing$ which is not the case.
+                                  If $v'_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$, i.e.
+$v'_2 = g_{V};i_{2,V}(z_2)$ and then it should be $[g_{V}(z_2)) \not = \varnothing$ which is not the case.
+                          \end{itemize}
+                    \item Other cases are not possible, for example, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$
+                          would mean that $e_2$ has a predecessor because $v_2$ does.
+                  \end{itemize}
+            \item $v' \in s(e_1)$ and $v \in s(e_2)$ and $v' \sim v$
+                  Suppose $[e_1) \not = \varnothing$. Then,
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))]
+                  \] and $[e_2) = \varnothing$ and
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                  \],
+                  where $[v_2] \in s([e_2])$ and $[[v_2]) \not = \varnothing$.
+                  \begin{itemize}
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{1,E}(e_x)]$.
+                          If $v_2 = f_{V};i_{1,V}(z_2)$, then $[f_{V}(z_2)) \not = \varnothing$ and hence $v'_2 \in s(e_2)$ and $v'_2 \sim v_2$.
+                          It should be the case that $v = g_{V};i_{2,V}(z)$ (if it had a pre-image in $V_{X}$ it would share the predecessor with the pre-image of $v_2$, but $[v) = \varnothing$) and $v' = f_{V};i_{1,V}(z)$, then $e_1$ has a pre-image in $E_{X}$ and $e_2$ has a pre-image in $E_{Y}$.
+                          Then $v'$ and $v_2$ both have pre-image in $V_{X}$ and because of the assumption $[f_{V}(z)) = [f_{V}(z_2))$ and $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [i_{1,E}(e_x)] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$.
+                    \item \update{TODO}
+                  \end{itemize}
+            \item $v \in s(e_1)$ and $v' \in s(e_2)$ and $v' \sim v$
+            \item $v' \in s(e_1)$ and $v'' \in s(e_2)$ and $v' \sim v$, $v'' \sim v$. 
+                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class.
+        \end{itemize}
+    \end{proof}
+\end{lemma}
+
+\begin{lemma}
+    $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive, assymetric and respects connectivity.
+    \begin{proof}
+        We will separately consider each property.
         \begin{itemize}
@@
                             \end{itemize}
-                \item $[e_1) = \varnothing$ and for all $v_i \in s(e_1), v_j \in t(e_1)$ $[[v_i]) \not = \varnothing,\; [[v_j]) \not = \varnothing$. 
-                Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e_2]$ and this gives us the needed property on the nose.
+                    \item $[e_1) = \varnothing$ and for some $[v_i] \in s([e_1]), [v_j] \in t([e_1])$ $[[v_i]) \not = \varnothing,\; [[v_j]) \not = \varnothing$. 
+                    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$.
+                    We now need to check that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                    \begin{itemize}
+                        \item If $v$ has no pre-image in $V_{Z}$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$, then, by definition
+                              \[
+                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+                              \]
+                    \end{itemize}
+                    We have that $[v] \in s([e_1])$ and $[v'] \in s([e_1])$ (or $[v'] \in t([e_1])$). 
+                    This means there exists $v_1$ such that $[v] = [v_1]$ and $v_1 \in s(e_1)$ and there exists $v_2$ such that $[v_2] = v'$ and $v_2 \in s(e_1)$.
+                    \begin{itemize}
+                        \item $v = v_1$ and $v \in s(e_1)$ and $v' = v_2$ and $v' \in s(e_1)$. 
+                              Because $[e_1) = \varnothing$, then $[v) = \varnothing$ and $[v') = \varnothing$ because the relation respects connectivity.
+                              \begin{itemize}
+                                \item If $v = f_{V};i_{1,V}(z)$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]$.
+                                      If $v' = f_{V};i_{1,V}(z')$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e'_y)]$.
+                                      We then have $e'_y = e_y$ because by assumption all nodes from $V_{Z}$ have the same immediate predecessor under $g$.
+                                \item If $v = f_{V};i_{1,V}(z)$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]$ and $v'$ has no pre-image in $V_{Z}$ then there necessarily exists an incident edge $[e']$ such that ...
+                                \item                                        
+                              \end{itemize} 
+                    \end{itemize}
                   \end{itemize}
@@
         \end{itemize}
-    \update{This will be changed to consistency relation, but the definition of quotiening will stay the same.}
-    Let's now define $\conflict_{\sfrac{X+Y}{\sim}}$.
-    We can consider the original conflict relation as a function $\conflict_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
-    Quotienting the range of the function gives us $\conflict_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote it as $\langle []_{V}^{\conflict} []_{E}^{\conflict} \rangle$) applied to the return value of $\conflict_{X+Y}$.
-    We then define the new relation as
-
-    \[
-        \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}([x]_{E})) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x')))^{\Rightarrow}
-    \]
-    \[
-        \conflict_{\sfrac{X+Y}{\sim}}(i_{V_{X} + V_{Y}}([x]_{V})) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} i_{V_{X} + V_{Y}}(x')))^{\Rightarrow}
-    \]
-        where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$ and $\Rightarrow$ is a closure by (1) from~\ref{def:e-hypergraph-req}.
-        That is, the relation above is the smallest relation that contains $\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x'))$ and such that if $[x_1] <_{c,\sfrac{X+Y}{\sim}}([x_2])$ and $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$ then $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_2] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$.
-        And similarly for nodes.
-
-    Again, we need to check that this is a well-defined function and that our initial $([]_{V},[]_{E})$ respects the conflict.
-
-    \begin{itemize}
-        \item Assume that $t_1, t_2 \in E_{X+Y}$ and $t_1 \sim t_2$ then $\conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}[t_1]) = \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}[t_2])$.
-              $t_1 \sim t_2$ implies there exists $z \in E_{Z}$ such that $t_1 = f_{E};i_{1,E}(z)$ and $t_2 = g_{E};i_{2,E}(z)$
-              \begin{align*}
-                \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}([t_1])) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{t_i \in \{t_1, t_2\}} i_{E_{X} + E_{Y}}(t_i)))^{\Rightarrow}
-                %      \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
-                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
-                %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
-              \end{align*}
-              Similarly,
-              \[
-                \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}([t_2])) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{t_i \in \{t_1, t_2\}} i_{E_{X} + E_{Y}}(t_i)))^{\Rightarrow}
-              \]
-              and we get well-definedness on the nose.
-            %   \begin{align*}
-            %     \conflict_{\sfrac{X+Y}{\sim}}(i_{E_{X} + E_{Y}}[t_2]) = \conflict_{X+Y};\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(i_{E_{X} + E_{Y}}(t_2)) = \\
-            %          \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_2))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(g_{E};i_{2,E};i_{E_{X} + E_{Y}}(z)) = \\
-            %          \text{By homomorphism-ness of $g_{E};i_{2,E}$} \\
-            %          = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle g_{V};i_{2,V};i_{V_{X} + V_{Y}} ,g_{E};i_{2,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}(z)))
-            %   \end{align*}
-            %   Then,
-            %   $[f_{E};i_{1,E}(z')]_{E} = [g_{E};i_{2,E}(z')]_{E}$ for all $z' \in \{z'' \in \conflict_{Z}(z) \text{ s.t. } z'' \in E_{Z}\}$
-            %   and
-            %   $[f_{V};i_{1,V}(z')]_{V} = [g_{V};i_{2,V}(z')]_{V}$ for all $z' \in \{z'' \in \conflict_{Z}(z) \text{ s.t. } z'' \in V_{Z}\}$
-            %   hence
-            %   \begin{align*}
-            %     \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle g_{V};i_{2,V};i_{V_{X} + V_{Y}} ,g_{E};i_{2,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z))) = \\
-            %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
-            %   \end{align*}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \[
-             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\conflict_{X+Y}(i_{E_{X+Y}}(x)))
-             \subseteq
-             \conflict_{\mathcal{\sfrac{X+Y}{\sim}}}([]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}(x))
-             \]
+    \end{proof}
+\end{lemma}
+
+\begin{lemma}
+    $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$ satisfies properties~\ref{def:consistency_properties}.
+    The properties (1),(2),(5) are satisfied by construction.
+    Hence we need to check wether properties (3),(4) and (6) are satisfied.
+    \begin{proof}
+        \begin{itemize}
+            \item Suppose $[v] \in s([e])$, then it must be the case that $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                  $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $[v] = [v']$.
+                  We then have two cases.
+                  \begin{itemize}
+                    \item $v = v'$ and then $v \in s(e)$. 
+                    This also means that $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(i_{V_{X} + V_{Y}}(v))$.
+                    By applying $([]_{V},[]_{E})$ we get 
+                    \[
+                       i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                    \]
+                        \item $v' \sim v'$ and $v' \in s(e)$ and hence $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$.
+                        By applying $([]_{V},[]_{E})$ we get 
+                        \[
+                        i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v'])
+                        \]      
+                        By functionality, 
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                        \]
+                        and hence
+                        \[
+                            i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                        \]
+                  \end{itemize}
+                  \item The case when $[v] \in t([e])$ is symmetric.
+                  \item If $i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then $[[v_1]) = [[v_2])$.
+                        The latter means that either $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$ or both functions are undefined.
                         By definition,
                         \[
-                \conflict_{\mathcal{\sfrac{X+Y}{\sim}}}([]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}(x)) = (\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle (\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x))))^{\Rightarrow}
-            \]
-            and by recalling that $[]_V^{*} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{*}$ we get the inclusion.
-    \end{itemize}
-    All is left is to check that $\sfrac{X + Y}{\sim}$ is indeed an e-hypergraph.
-    For this we need to check conditions (1) - (7) in~\ref{def:e-hypergraph-req}.
-    % \begin{enumerate}
-    %     \item Assume all $x_1, x_2, x_3$ are in $E_{X+Y}$.
-    %      Given $[x_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([x_2])$ 
-    %      and 
-    %      $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_1]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))$ then it must be the case that
-    %     $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))$.
-
-    %     \[
-    %         i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_1]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))
-    %     \]
-    %     means that there exist $x',x'' \in E_{X} + E_{Y}\; | \;[x'] = [x_3] \text{ and } x'' \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x')) \text{ and } [x''] = [x_1]$.
-    %     We then begin by considering cases for $<$ and $[x_1], [x_2], [x_3], [x']$.
-    %     \begin{itemize}
-    %         \item $[x_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}([x_2]) = [<_{c,X+Y}^{\mu}(x_2)]$.
-    %         We have two options.  
-    %             \begin{itemize}
-    %                 \item $x_2$ has a pre-image $z_{x_2}$ in $E_{Z}$ such that $[z_{x_2}) \not = \varnothing$.
-    %                 \begin{itemize}
-    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' = x_3$, $x'' = x_1$ hence $i_{E_{X} + E_{Y}}(x_1) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
-    %                       Then, because $X+Y$ is an e-hypergraph it must be the case that $i_{E_{X} + E_{Y}}(x_2) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
-    %                       And, by homomorphism-ness of $[]$ we get $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]_{E}) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]_{E}))$. 
-    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' \sim x_3$, $x'' = x_1$ hence there exists $z \in E_{Z}$ such that $x' = f_{E};i_{1,E}(z)$ and $x_3 = g_{E};i_{2,E}(z)$.
-    %                       Hence $i_{E_{X} + E_{Y}}(x_1) \in \conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z))$.
-    %                       Then, because $X+Y$ is an e-hypergraph it must be the case that $i_{E_{X} + E_{Y}}(x_2) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z)))$.
-    %                       And, by homomorphism-ness of $[]$ we get $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]_{E}) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([f_{E};i_{1,E}(z)]_{E})) = \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]_{E}))$.
-    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' = x_3$, $x'' \sim x_1$ hence there exists $z \in E_{Z}$ such that $x'' = f_{E};i_{1,E}(z)$ and $x_1 = g_{E};i_{2,E}(z)$.
-    %                       Hence $i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z)) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
-    %                       By definition $x_2 = g_{E};i_{2,E}(z_{x_2})$ (because $x_1$ and $x_2$ should have pre-image in either $X$ or $Y$, not both).
-    %                       This means there exists $z_2 <_{c,Z}^{\mu}(z_{x_2})$ and $g_{E};i_{2,E}(z_2) <_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z_{x_2})) = <_{c,X+Y}^{\mu}(x_2)$ and because there can be only one imemdiate predecessor we have $x_1 = g_{E};i_{2,E}(z_2)$ and $z = z_2$.
-    %                       \question{What next?}
-    %                       Because $X+Y$ is an e-hypergraph, we have $i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z')) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(x_3))$.
-    %                       Then, by homomorphism-ness we get $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([f_{E};i_{1,E}(z')]) = i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_2]) \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([x_3]))$.
-    %                 \item Suppose $x_1 = <_{c,X+Y}^{\mu}(x_2)$, $x' \sim x_3$, $x'' \sim x_1$ hence there exist $z_1 \in E_{Z}$ such that $x'' = f_{E};i_{1,E}(z_1)$ and $x_1 = g_{E};i_{2,E}(z_1)$ and $z_2 \in E_{Z}$ such that $x' = f_{E};i_{1,E}(z_2)$ and $x_3 = g_{E};i_{2,E}(z_2)$.
-    %                       Hence, $i_{E_{X} + E_{Y}}(x'') \in \conflict_{X+Y}(i_{E_{X} + E_{Y}})(x')$ and $i_{E_{X} + E_{Y}}(f_{E};i_{1,E}(z_1)) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}})(f_{E};i_{1,E}(z_2))$.
-    %                       There also should exist $z_3$ such that $z_1 <_{c,Z}^{\mu}(z_3)$ and $g_{E};i_{2,E}(z_1) <_{c,Z}^{\mu}(g_{E};i_{2,E}(z_3))$
-    %             \end{itemize}
-    %             $x_2$ has no pre-image in $E_{Z}$.
-    %         \end{itemize}
-    %         \end{itemize}
-    % \end{enumerate}
-
-    \begin{enumerate}
-        \item (1) is given by construction.
-        \item There are separate cases for nodes and edges.
-              \begin{itemize}
-                \item Being not mutually exclusive means it is possible that $[e]_{E} <_{c,\sfrac{X+Y}{\sim}}([v]_{V})$ and $i_{\sfrac{V_{X}+V_{Y}}{\sim}}[v] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}[e])$.
-                \item Or, $[e_1]_{E} <_{c,\sfrac{X+Y}{\sim}}([e_2]_{E})$ and $i_{\sfrac{E_{X}+E_{Y}}{\sim}}[e_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}[e_2])$.
-               \end{itemize}
-              The first means that either
-              \begin{itemize}
-                \item $[e] = [<_{c,X+Y}(v)]$ if $v$ has no pre-image in $V_{Z}$ and there exist $v'$ and $e'$  such that $[v'] = [v]$, $[e'] = [e]$ and $i_{V_{X} + V_{Y}}(v') \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e'))$.
-                      As there is no pre-image in $V_{Z}$, $v = v'$. Then we have the following case.
-                    \begin{itemize}
-                        \item $e = <_{c,X+Y}(v)$ and $e' = e$.
-                              As we have $i_{V_{X} + V_{Y}}(v) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e))$ and this is not possible as $X+Y$ is an e-hypergraph.
-                    \end{itemize}
-                \item $v$ has a pre-image $z \in V_{Z}$, let's say, $f_{V};i_{1,V}(z) = v$, and $[f_{V}(z)) = \varnothing$.
-                      Then, either $v' = v$ and $i_{V_{X} + V_{Y}}(v) \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e))$, which is not possible as it requires that $v$ has a predecessor in $X$.
-                      Or, $v' \sim v$ (meaning that $v' = g_{V};i_{2,V}(z)$) and $i_{V_{X} + V_{Y}}(v') \in \conflict_{X+Y}(i_{E_{X} + E_{Y}}(e))$.
-                      Then, $<_{c,\sfrac{X+Y}{\sim}}([v]) = [g_{V};<_{c,Y};i_{2,E}(z)] = [e]$. Once again, it means that $e = g_{V};<_{c,Y};i_{2,E}(z)$.
-                      Overall, it means that $g_{V}(z) \in \conflict_{Y}(i_{E_{Y}}(g_{V};<_{c,Y}(z)))$ which is not possible as $Y$ is an e-hypergraph.
-                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
-              \end{itemize}
-    \end{enumerate}
-    Let's now prove the universal property.
-    $u_{E}([i_{1,E}(x)]) = j_{1,E}(x)$ and $u_{E}([i_{2,E}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,E}$ and $j_{2,E}$.
-
-    Let's say that $t_1 \sim t_2$ then $u_{E}([t_1]) = u_{E}([t_2])$.
-    $u_{E}([t_1]) = u_{E}(f_{E};i_{1,E}(z)) = j_1{f_{E}(z)}$ and $u_{E}([t_2]) = u_{E}(g_{E};i_{2,E}(z)) = j_{2}(g_{E}(z))$ which are equal by commutativity.
-
-    \update{TODO: $\conflict$}
-    % We can think of relation $<_{c,X+Y}$ as of a function $<_{c} : V_{X+Y} + E_{X+Y} \to {E_{X+Y}}^{*}$, i.e. a function that for each node or edge returns the sequence of their predecessors.
-    % Then we can define the quotienting as before
-    % \[
-    % \sfrac{(<_{c,X+Y})}{\sim}([u]) = [<_{c,X+Y}(u)]_{E}^{*}\\
-    % \]
-    % or, if we consider a function that returns the immediate predecessor
-    % \[
-    % \sfrac{(<_{c,X+Y}^{\mu})}{\sim}([u]) = [<_{c,X+Y}^{\mu}(u)]_{E}
-    % \]
-    % To check that the function is well-defined we need to check that if $t_1 \sim t_2$ then $\sfrac{<_{c,X+Y}^{\mu}}{\sim}([t_1]) = \sfrac{<_{c,X+Y}^{\mu}}{\sim}([t_2])$ or $[<_{c,X+Y}^{\mu}(t_1)] = [<_{c,X+Y}^{\mu}(t_2)]$.
-    % Once again, $t_1 \sim t_2$ (suppose these are nodes) implies there exists $z$ such that $t_1 = f_{V}(z);i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$.
-    % Hence we have
-    % \[
-    %     [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))]
-    % \]
-    % \[
-    %     [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
-    % \]
-    % and two cases
-    % \begin{itemize}
-    %     \item $[z) \not = \varnothing$ in which case we can use the fact that $f_{V};i_{1,V}$ and $g_{V};i_{2,V}$ are homomorphisms.
-    %           This gives us 
-    %             \[
-    %                 f_{V};i_{1,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))
-    %             \] 
-    %             and
-    %             \[
-    %                 g_{V};i_{2,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))
-    %             \]
-    %             and finally,
-    %             \[
-    %                 [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
-    %             \] 
-    %             as they share the same pre-image.
-    %     \item Another case is when $[z) = \varnothing$. 
-    %           Assume $[f(z)) = \varnothing$, in which case $<_{c,Z}^{\mu}(z)$ and $<_{c,X}^{\mu}(f_{V}(z))$ are undefined.
-    %           This takes us to the situation where the lefthand-side of what we want to prove is undefined while the right-hand side is defined.
-    %           \question{Can I somehow require that in such case I take the right-hand side as a definition of the lefthand side? 
-    %           Otherwise the relation we have is too weak because it tells us only to identify things with the pre-image in $Z$ but eventually we also want to identify things that do not have such a pre-image.
-    %           }
-    %           What I want is to define 
-    %           \[
-    %             \sfrac{<_{c,X+Y}^{\mu}}{\sim}([f_{V};i_{1,V}(z)]) = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))] \text{ if } [f_{V};i_{1,V}(z)) = \varnothing
-    %           \]
-    % \end{itemize}
-
-    % \update{Approach 2}
-
-    % We can extend the set of edges in an e-hypergraph with a special edge $\bot$.
-    % And allow that it can be mapped to any other edge under a homomorphisms, i.e. we will not require that $s;\phi^{*}(\bot) = \phi;s;(\bot)$.
-    % \question{but this might break quotienting of sources and targets}
-    % This also means that set of edges now is $E_{\mathcal{G}} + \bot$.
-    % \question{What to do with $\bot$ after a coproduct?}.
-    % This gives us the ability to map this $\bot$ using homomorphisms.
-
-    % The pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    % First, a coproduct of $X+Y$ is computed which is
-
-    % \[
-    %     X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y} + \bot_{X} + \bot_{Y}, s_{X+Y}, t_{X+Y}, \#_{X+Y}, <_{c,X+Y} \}
-    % \]
-
-    % \[
-    %     R_{E} = \{ (x_i, y_j) \in (E_{X} + E_{Y} + \bot_{X} + \bot_{Y}) \times (E_{X} + E_{Y} + \bot_{X} + \bot_{Y} | \exists z \in E_{Z} + \bot_{Z} \; .
-\; x_i = f_{E};i_{1,E}(z) \text{ and } y_j = g_{E};i_{2,E}(z))\}
-    % \]
-
-    % Consider $<_{c,X+Y}^{\mu}$ as a function that returns the immediate predecessor, i.e. it has a type $V_{X} + V_{Y} + E_{X} + E_{Y} \to E_{X} + E_{Y} + \bot_{X} + \bot_{Y}$.
-    % And define its quotient
-    % \[
-    % \sfrac{<_{c,X+Y}^{\mu}}{\sim}([u]) = [<_{c,X+Y}^{\mu}(u)]_{E}
-    % \]
-
-    % We then need to check that whenever $t_1 \sim t_2$ $[<_{c,X+Y}^{\mu}(t_1)] = [<_{c,X+Y}^{\mu}(t_2)]$.
-    % $t_1 \sim t_2$ implies that there exists at least one $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$.
-    % Hence we have
-    % \[
-    %     [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))]
-    % \]
-    % \[
-    %     [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
-    % \]
-    % and two cases
-    % We then can use the fact that $f_{V};i_{1,V}$ and $g_{V};i_{2,V}$ are homomorphisms.
-    % This gives us 
-    % \[
-    %     f_{V};i_{1,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))
-    % \] 
-    % and
-    % \[
-    %     g_{V};i_{2,V}(<_{c,Z}^{\mu}(z)) = <_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))
-    % \]
-    % and finally,
-    % \[
-    %     [<_{c,X+Y}^{\mu}(f_{V};i_{1,V}(z))] = [<_{c,X+Y}^{\mu}(g_{V};i_{2,V}(z))]
-    % \] 
-    % as they share the same pre-image ($<_{c,Z}^{\mu}(z)$).
-
-    % Now let's check that $\sfrac{s_{X+Y}}{\sim}$ is well-defined.
-    % We will consider a case where $f_{E}(\bot_{Z}) = \bot_{X}$ and $g_{E}(\bot_{Z}) \not = \bot$.
-    % We then need to check that $[s_{X+Y}(f_{E};i_{1,E}(z))]^{*} = [s_{X+Y}(g_{E};i_{2,E}(z))]$
-    % \begin{itemize}
-    %     \item $z = \bot_{Z}$ then $[{}] = \ldots$
-    % \end{itemize}
-
-    % \update{end of Approach 2}
-
-    % \update{We shall probably first define the quotient of the immediate parent because it is essentially a function and it will follow the same argument as the source function}
-    % \question{Below $e$ and $v$ do not necessarily mean edges and nodes, $v$ can be an edge as well.}
-    % \question{I also kind of intuitively inventing the quotient of a relation, is there any accepted notion for this?}
-    % and, whenever $e_{X} <_{c,X}^{\mu} v_{X}$ and $e_{Y} <_{c,Y}^{\mu} v_{Y}$ and $v_{x} R v_{y}$ then $e_{X} R e_{Y}$.
-    % The latter condition means two things
-    % \begin{itemize}
-    %     \label{cond:essential_for_pushout}
-    %     \item if $[z) = \varnothing$ then either $[f(z)) = \varnothing$ or $[g(z)) = \varnothing$.
-    %           \question{This point formally defines what it means to 'inherit' a relation.
-    %                     Let's say $v_1 = f(z)$ and $v_2 = g(z)$ and $e_1 <_{c,X}^{\mu} v_1$ and $[v_2) = \varnothing$,
-    %                     then, having $e_1 <_{c} [v_1,v_2]$ is fine as it does not break the condition of a predecessor being unique.}
-    %           \update{This point also says when we can compute a pushout: either non of X and Y should have a parent for $f(z)$ and $g(z)$ or only one of them should.}
-    %     \item if $[z) \not = \varnothing$ then there exists $z'$ such that $z' <_{c}^{\mu} z$ and $[g(z'), f(z')] <_{c}^{\mu} [g(z), f(z)]$
-    % \end{itemize}
-    % Then,
-    % \[
-    %     \sfrac{\conflict_{X} + \conflict_{Y}}{\sim R} \subseteq \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R} \times \sfrac{(V_{X} + V_{Y} + E_{X} + E_{Y})}{\sim R}
-    % \]
-    % meaning that whenever $e_1 \conflict e_2$ and $e_1' \conflict e_2'$ and $e_1 R e_1'$ then $[e_1, e_1'] \conflict e_2$ and $[e_1, e_1'] \conflict e_2'$.
-    % $\sfrac{i_1}{\sim R}$ then maps $x \to [x]$ and $\sfrac{i_2}{\sim R}$ maps $y \to [y]$.
-    % Before we proceed we need to check that $\sfrac{(X + Y)}{\sim R}$ is indeed an e-hypergraph.
-    % Since $X + Y$ is clearly an e-hypergraph, we need to check whether identification breaks any of the requirements (1-7) above.
-    % \begin{enumerate}
-    %     \item Assume $[x_1,y_1] <_{c} [x_2,y_2]$ and $[x_1, y_1] \conflict [x_3]$ which means that $x_1 \conflict x_3$.
-    %           Hence $x_2 \conflict x_3$ (because $X$ is an e-hypergraph) and $[x_2,y_2] \conflict [x_3]$. Similarly, if $[x_1, y_1] \conflict [y_3]$ and $y_1 \conflict y_3$.
-    %     \item $\sfrac{\conflict}{\sim R}$ and $\sfrac{<_{c}}{\sim R}$ are disjoint because we quotient only inside the relations.
-    %     \item Again, we only quotient $\conflict$ on $x$ and $y$ for which original relation is defined.
-    %     \item This follows from $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ being homomorphisms.
-    %     \item We just perform transitive closure of the complement.
-\update{Maybe we can get rid of this point and instead later just take transitive closure of the complement when talking about interfaces}
-    %     \item Assume $[x_1, y_1] \conflict [x_2,y_2]$ and $[x_2,y_2] \conflict [x_3,y_3]$ and all of them are distinct.
-    %           Then $[x_1,y_1] \conflict [x_3,y_3]$ because either $x_1 \conflict x_3$ or $y_1 \conflict y_3$ (otherwise they would not have been merged into the same equivalence class).
-    %     \item We never merge $x_1$ and $x_2$ such that $x_1 <_{c} x_2$ into the same equivalence class because morphisms should preserve the order.
-    % \end{enumerate}
-
-    % Then we need to check that $\sfrac{i_1}{\sim R}$ and $\sfrac{i_2}{\sim R}$ are indeed homomorphisms.
-    % For these we need to check that conditions (2-3) in~\ref{def:e-homo} are satisfied.
-    % Let's check for $\sfrac{i_1}{\sim R}$.
-    % We identify elements from $X$ with elements from $Y$ which does not affect (2) and the same reasoning applies to (3).
-    % (1) was proved earlier for $\catname{Hyp_{\Sigma}}$.
-
-    % Finally, by following essentially the same argument as for pushout in $\catname{Hyp_{\Sigma}}$ we can conclude that the structure defined above is indeed a pushout in $\catname{EHyp_{\Sigma}}$ and we know precise conditions under which this concrete pushout exists.
-\end{remark}
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
+                        \]
+                        which means that either $v_1 \consistency_{X+Y}^{\mu} v_2$ or $v'_1 \consistency_{X+Y}^{\mu} v_2$ and $v'_1 \sim v_1$.
+                        \begin{itemize}
+                            \item $v_1 \consistency_{X+Y}^{\mu} v_2$ implies $[v_1) = [v_2)$.
+                            \begin{itemize}
+                                \item If $v_1$ does not have a pre-image in $V_{Z}$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1))]$
+                                \begin{itemize}
+                                    \item If $v_2$ does not have a pre-image in $V_{Z}$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))]$ and by assumption $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))$ and hence their images are also equal.
+                                    \item If $v_2$ has an image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and hence $[v_2) = \varnothing$ and $[v_1) = \varnothing$.
+                                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_{y})]$.
+                                \end{itemize}
+                            \end{itemize}
+                        \end{itemize}
+        \end{itemize}
+    \end{proof}
+\end{lemma}
+
+\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
+For this we need to check that the defined relations satisfy all the needed properties.
+
+\end{proposition}
 

### 2023-12-07 sections/combinatorial_semantics.tex
Pushout for EHyp with consistency and proofs

@@
         \label{def:consistency_properties}
-        \item For each $v \in V_{\mathcal{G}}$ $v \consistency^{\mu} v$
-        \item For each $e \in E_{\mathcal{G}}$ $e \consistency^{\mu} e$
+        \item The relation is an equivalence relation
         \item If $v \in s(e)$ then $v \consistency^{\mu} e$
         \item If $v \in t(e)$ then $v \consistency^{\mu} e$
-        \item The relation is symmetric and transitive
         \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
@@
 %     \item $\#$ should respect connectivity
-%     \item \update{Complement of $\conflict_{\mu}$ should be transitive}
+%     \item \update{Complement of $\hashtag_{\mu}$ should be transitive}
 %     \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
@@
     \]
-    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
+    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \hashtag_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \hashtag_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
 \end{enumerate}
@@
 
-    \begin{itemize}
-    \item First, we assume that $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
-    \item Second, we assume that $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$.
-\update{This essentially means that $f,g$ do not map nodes to levels of different hierarchy.}
+    \begin{enumerate}
+    \item $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
+    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$. \update{This essentially means that $f,g$ do not map nodes to levels of different hierarchy.}
+    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and vice versa.
+    \item $\consistency^{\mu}(f(v_i)) = \consistency^{\mu}(f(v_j))$ and $\consistency^{\mu}(f(g_i)) = \consistency^{\mu}(g(v_j))$.
+    \item $f$ and $g$ are mono.
     \label{pushout:assumptions}
-    \end{itemize}
+    \end{enumerate}
 
@@
     \] where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    For this we will consider several cases where \update{$f$ and $g$ are mono. I require mono just to simplify things, we can get rid of this requirement if everything else is ok.}.
+    We will consider several cases.
     Assume that $u$ is an element of $V_{X+Y}$ and
@@
               \end{enumerate}
-    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then steps (1) - (2) for edges and finally, step (4) for nodes.}
-
-    \begin{remark}
-        Essentially, what we build above is that if the relation is undefined, it is inherited from either incident nodes or edges.
-    \end{remark}
-    \update{If $f$ and $g$ are not mono, the definitions stay the same as $i_{2,E}$ and $i_{1,E}$ are mono.}
+    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for nodes until fixed point.
+            Essentially, this interleaving means that nodes inherit the relation from incident edges and vice versa.
+    }
     Clearly, all the cases above are disjoint.
@@
     Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
-    We then define the new relation for edges as
+    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
     \begin{enumerate}
@@
                 \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) =
-                    (\langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))))^{c}
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
+                    \langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)))
                 \end{align*}
-        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V}))
+        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{V_{X} + V_{Y}}([v]_{V}))
         \]
@@
     \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
-    \]
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E}))
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')))
+    \]
+    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{E_{X} + E_{Y}}([e]_{E}))
     \]
     \end{enumerate}
-        where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$ and $c$ is a reflexive, symmetric and transitive closure.
-        % That is, the relation above is the smallest relation that contains $\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x'))$ and such that if $[x_1] <_{c,\sfrac{X+Y}{\sim}}([x_2])$ and $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$ then $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_2] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$.
-        % And similarly for nodes.
+    where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
+
+    We then define
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
+    \]
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
+    \]
+    where $c$ is a reflexive, symmetric and transitive closure.
 
@@
     \begin{itemize}
-        \item 
+        \item Let us show well-definedness.
         \begin{itemize}
@@
               \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
-                %      \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+                %      \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
                 %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
-                %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
+                %      = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\hashtag_{Z}(i_{E_{Z}}(z)))
               \end{align*}
@@
               \[
-                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
               \]
@@
               We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$ and there exists $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ such that $[v_1] \in s([e])$ (and automatically $[v_2] \in s([e])$).
-                Then,
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$.
+                  This would mean that for all $v_i \in V_{Z}$ $\consistency^{\mu}(f_{V}(v)) = \varnothing$ and $\consistency^{\mu}(g_{V}(z)) = \varnothing$ and automatically
+                  \[
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
                   \]
@@
             \[
-                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x))))^{c}
-            \]
-            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion.
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
+            \]
+            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion $\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ because it is a closure.
             If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness is true because empty set is a subset of any set.
@@
 \begin{lemma}
+    \label{lemma:choice_of_nodes_agnostic}
     Suppose $[e) = \varnothing$ and there exists $[v_1] \in s([e])$ and $[v_2] \in s([e])$ such that $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2]))$.
     \begin{proof}
-        Because we build the relation in steps, $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$ mean that was achieved after running steps (1)-(3) for nodes.
+        We will show the statement is true for the base case when $[v_1]$ and $[v_2]$ do not inherit their relations.
+        In other words, $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$ mean that was achieved after running steps (1)-(3) for nodes.
         From $[v_1] \in s([e])$ we get that there exists $v'_1$ such that $v_1 \sim v'_1$ and $v'_1 \in s(e)$ and similarly $v'_2 \sim v_2$.
         \begin{itemize}
-            \item $v_1 = v'_1$ and $v_1 \in s(e)$, and $v_2 = v'_2$ $v_2 \in s(e)$.
+            \item $v_1 = v'_1$, $v_1 \in s(e)$, and $v_2 = v'_2$, $v_2 \in s(e)$.
             \begin{itemize}
@@
                 \item Case for $v_1 = g_{V};i_{2,V}(z)$ is symmetric.
-
-                It can not be the case that the node has no pre-image as $[[v_i]) \not = \varnothing$ and $[e) = \varnothing$ implies that step (4) is needed but we only completed steps (1)-(3).
-            \end{itemize}
-            \item $v_1 \not = v'_1$ and $v'_2 \not = v_2$ and hence $v'_1 \in s([e])$ and $v'_2 \in s(e)$.
+                \item Case when $v_1$ has no pre-image in $V_Z$ is not possible, because it would mean that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1))]$ (according to step (3)) which is not possible as it would mean that $[e) \not = \varnothing$ (because $[v_1) \not = \varnothing$).
+            \end{itemize}
+            \item $v_1 \not = v'_1$ and $v'_2 \not = v_2$ and hence $v'_1 \in s(e)$ and $v'_2 \in s(e)$.
             \begin{itemize}
@@
                       $v'_2$ has a pre-image $f_{V};i_{1,V}(z')$ and for the same reason 
-                      \[
-                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) = [i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                      \]
-            \end{itemize}
-            \item The remaining cases are similar.
-        \end{itemize}
+                      \begin{align*}
+                        &<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) =\\ 
+                        &[i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})] =\\
+                        &<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                    \end{align*}
+                \item Case when $v'_1 = g_{V};i_{2,V}(z)$ is similar.
+            \end{itemize}
+            \item $v_1 = v'_1$, $v'_2 \sim v_2$ and $v_1 \in s(e)$, $v'_2 \in s(e)$.
+            \begin{itemize}
+                \item $v_1 = f_{V};i_{1,V}(z)$ and hence $v'_2 = f_{V};i_{1,V}(z')$ and $[f_{V}(z)) = \varnothing$ as well as $[f_{V}(z')) = \varnothing$.
+                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{2,E}(e_{y})]$ and similarly $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e'_{y})]$.
+                      And by the following the same reasoning as above we get the equality.
+                \item Case when $v_1 = g_{V};i_{2,V}(z)$ is similar.
+            \end{itemize}
+            \item The remaining case is similar.
+        \end{itemize}
+        The proofs for the cases when $[v_1] \in t([e])$ and $[v_2] \in t([e])$ and for the combinations of $t,s$ are similar.
     \end{proof}
@@
 \begin{lemma}
+    \label{lemma:choice_of_edges_agnostic}
     Suppose $[v) = \varnothing$ and there exist $e_1$ and $e_2$ such that $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ and $[v] \in s([e_1])$, $[v] \in s([e_2])$.
@@
     \begin{proof}
-        Once again, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.
+        We will again show the base case when $[e_1]$ and $[e_2]$ inherit the relation atmost at 1 step.
+        That is, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.
         We then have the following cases.
@@
                           $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_{y})]$
-                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$. The only possible case then is that $v_2 \in s(e_2)$ and $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$.
-                          Then $[v_1') = \varnothing$ and $[v_2) = \varnothing$. Show it is not possible.
+                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$ (because they are from the same connected component).
                           \begin{itemize}
                             \item $v_1 = f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$.
-                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
-                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it cant be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}([v_2])}) = [i_{2,E}(e_y)]$.
+                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}(z)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
+                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it can not be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_y)]$.
                                   If $v'_2 \in s(e_2)$ and $v_2 \sim v'_2$ then $v'_2$ has a pre-image in $V_{Y}$ and $v_2$ has a pre-image in $V_{X}$, i.e. $v_2 = f_{V};i_{1,V}(z_2)$. 
@@
                     \item Other cases are not possible, for example, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$
-                          would mean that $e_2$ has a predecessor because $v_2$ does.
+                          would mean that $e_2$ has a predecessor because $v_2$ does and $v_2 \in s(e_2)$ because $v_2$ has no pre-image in $V_{Z}$.
+$[v_2]$ can not inherit the relation because we consider inheritance only upto 1 step: inheritance for $v_2$ would mean that $v$ inherits the relation in more than 1 step.
                   \end{itemize}
@@
                           Then $v'$ and $v_2$ both have pre-image in $V_{X}$ and because of the assumption $[f_{V}(z)) = [f_{V}(z_2))$ and $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [i_{1,E}(e_x)] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$.
-                    \item \update{TODO}
-                  \end{itemize}
+                          The case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$ is not possible as it would mean that $v_2 \in s(e_2)$, but $[e_2) = \varnothing$.
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e']))$ is also not possible as it would mean that $e_2$ inherits the relation in more than 1 step.
+                \end{itemize}
+                Cases $[e_1) = \varnothing$ and $[e_2) = \varnothing$ and $[e_1) = \varnothing$, $[e_2) \not = \varnothing$ are not possible.
             \item $v \in s(e_1)$ and $v' \in s(e_2)$ and $v' \sim v$
+                  This is symmetric to the above.
             \item $v' \in s(e_1)$ and $v'' \in s(e_2)$ and $v' \sim v$, $v'' \sim v$. 
-                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class.
-        \end{itemize}
+                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class so it reduces to the cases above.
+        \end{itemize}
+        The proofs for the cases when $[v] \in t([e_1])$ and $[v] \in t([e_2])$ and for the combinations of $t,s$ are similar.
     \end{proof}
 \end{lemma}
+
+\begin{remark}
+    To get the result for arbitrary number of steps we take the above lemmas as base case and procceed by induction on the number of steps.
+\end{remark}
+
+\begin{corollary}
+    \label{corollary:nodes}
+    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
+    \begin{align*}
+         [v') \not = \varnothing,\\
+         <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \end{align*}
+    In other words, there exists the node which initiates the inheritance for nodes that had no predecessor.
+\end{corollary}
+\begin{corollary}
+    \label{corollary:edges}
+    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $[v]$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
+    \begin{align*}
+        [v') \not = \varnothing,\\
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \end{align*}  
+    In other words, there exists the node which initiates the inheritance for edges that had no predecessor.
+\end{corollary}
 
@@
         \begin{itemize}
-            \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}^{\mu}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
+            \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
                   Then we have two cases.
                   \begin{itemize}
-                    \item Assume $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{c,\sfrac{X+Y}{\sim}^{\mu}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-                          This would mean that $e_1 = e$ which would mean that the relation $<_{c,X+Y}^{\mu}$ is not irreflexive.
-                    \item Assume $[e) = \varnothing$ and for all $v_{i} \in s(e)$ and $v_{j} \in t(e)$ $[[v_i]) \not = \varnothing$ $[[v_j]) \not = \varnothing$.
-                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                          \begin{itemize}
-                            \item If $v$ has no pre-image in $V_{Z}$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e]$.
-                                  This would mean that $e$ is a predecessor of $v$ in $X+Y$, but $v$ is either in the sources or targets of $e$ which violates respectful connectivity condition.
-                            \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ such that $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$.
-                                  Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{1,E}(e_x)] = [e]$.
-                                  This means that $e$ has a pre-image in $E_{X}$, but because $v$ also has a pre-image in $V_{X}$ and its set of predecessors is not empty it would mean that $e$ has a predecessor which it does not.
-                            \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ such that $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$.
-                                  Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)] = [e]$.
-                                  It means that $e$ has a pre-image in $E_{Y}$, but $v$ is its source or target and has a pre-image in $V_{Y}$ which is not possible.
-                            \item The remaining two cases for $v = g_{V};i_{2,V}(z)$ are symmetric.
-                          \end{itemize}
-                  \end{itemize}$[<_{c,X+Y}^{\mu}(e)]$ and $e = <_{c,X+Y}^{\mu}(e)$ because edges are not identified.
+                    \item Assume $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
+                          This would mean that $e_1 = e$ which would mean that the relation $<_{c,X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
+                    \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v]))$.
+                          \begin{itemize}
+                            \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};i_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
+                                  Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [e]$.
+                                  This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
+                                  Contradiction.
+                            \item Case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+                          \end{itemize}
+                  \end{itemize}                  
                     We do not need to check this requirement for nodes because a node can not be a predecessor.
@@
                          This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-                        \item $[e_1) = \varnothing$ and for all $v_i \in s(e_1)$ and $v_j \in t(e_1)$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
+                        \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
                               Then, $[e_2] = [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
                               \begin{itemize}
-                                \item If $v$ has no pre-image in $V_{Z}$, then 
-                                \[
-                                    [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e_2]
-                                \]
-                                      It would mean that $e_2$ is a predecessor of $v$ and, by connectivity, a predecessor of $e_1$.
-                                      However, $[e_1) = \varnothing$ and we get a contradiction.
-                                \item If $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
+                                % \item If $v$ has no pre-image in $V_{Z}$, then 
+                                % \[
+                                %     [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e_2]
+                                % \]
+                                %       It would mean that $e_2$ is a predecessor of $v$ and, by connectivity, a predecessor of $e_1$.
+                                %       However, $[e_1) = \varnothing$ and we get a contradiction.
+                                \item If $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
                                       Then,
@@
                                       \]
-                                      This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is a predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
+                                      This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
                                 \item Case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric to the above.
@@
                         \end{itemize}
-                        \item $[e_2) = \varnothing$ and for all $v_i \in s(e_2)$ and $v_j \in t(e_2)$ $[[v_i]) \not = \varnothing$ $[[v_j]) \not = \varnothing$.
+                        \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exits $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
                               Then, 
                               \[
-                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e_1]
-                              \]
-                              The cases when $[e_1) \not = \varnothing$ are similar to the above, so we will just jump to the interesting case when 
+                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
+                              \]
+                              The case when $[e_1) \not = \varnothing$ are similar to the above, so we will just jump to the interesting case when 
                               $[e_1) = \varnothing$ and
                               \[
-                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]
+                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
                               \]
@@
                               \begin{itemize}
-                                \item $v$ has no pre-image in $V_{Z}$. 
-                                      Then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e_1]$
-                                      This implies that $e_1$ is the predecessor of $e_2$ because it is the predecessor of $v$ and $v$ is $e_2$'s sources of targets.
-                                      \begin{itemize}
-                                        \item $v'$ has no pre-image in $V_{Z}$.
-                                        Then $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [e_2]$.
-                                        This implies that $e_2$ is the predecessor of $v'$ and hence of $e_1$. 
-                                        This contradicts the assymetry of $<_{c,X+Y}^{\mu}$.
-                                        \item $v'$ has a pre-image $f_{V};i_{1,E}(z) = v'$ in $V_{Z}$ (and so should $e_1$ do) and therefore $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$ because $[e_1) = \varnothing$.
-                                              Then, $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [i_{2,E}(e_y)] = [e_2]$.
-                                              This means that $e_2$ has a pre-image in $E_{Y}$ but $e_1$ is the predecessor of $e_2$ and hence they should both have pre-image in $E_{X}$ (as $e_1$ does so).
-                                        \item Case when $v' = g_{V};i_{2,V}(z)$ is symmetric.
-                                        \end{itemize}
-                                \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ (and so should $e_2$ do) in $Z_{V}$ and therefore $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$.
-                                      Then, $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_{y})] = [e_1]$.
-                                      This implies that $e_1$ has a pre-image in $E_{Y}$.
-                                      \begin{itemize}
-                                        \item $v'$ has no pre-image in $V_{Z}$.
-                                              Then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [e_2]$.
-                                              This implies that $e_2$ is the predecessor of $v'$ and hence of $e_1$.
-                                              But $[e_1) = \varnothing$ and hence we have a contradiction.
-                                        \item $v'$ has a pre-image $f_{V};i_{1,V}(z') = v'$ in $V_{Z}$ (and so should $e_1$ do) and therefore $[f_{V}(z')) = \varnothing$ and $[g_{V}(z')) \not = \varnothing$.
-                                              This also means that $[g_{V}(z)) = [g_{V}(z'))$.
-                                              By definition we have that $e_1$ is the image of the predecessor of $g_{V}(z')$ and $e_2$ is the image of the predecessor of $g_{V}(z)$ and because their predecessors are equal it implies $e_1 = e_2$ but the relation is irreflexive.
-                                        \item $v'$ has a pre-image $g_{V};i_{2,V}(z') = v'$ in $V_{Z}$ and therefore $[f_{V}(z')) \not = \varnothing$ and $[g_{V}(z')) = \varnothing$.
-                                              This is impossible as it violates the requirement that $[f_{V}(z)) = [f_{V}(z'))$ for all $z,z' \in V_{Z}$.
-                                    \end{itemize}
-                                \item The case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric.
-                                \end{itemize}
+                                \item $v_2 = f_{V};i_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};i_{2,V}(z_2)$.
+                                Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z_2)] = [e_1]$ which also implies that $e_1$ has a pre-image in $E_{Y}$.
+                                \begin{itemize}
+                                    \item $v_1 = g_{V};i_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};i_{1,V}(z_1)$.
+                                    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z_1)] = [e_2]$  
+                                    However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates our assumptions.
+                                \end{itemize}
+                                \item Case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
+                              \end{itemize}
+                            %   \begin{itemize}
+                            %     \item $v$ has no pre-image in $V_{Z}$. 
+                            %           Then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = [e_1]$
+                            %           This implies that $e_1$ is the predecessor of $e_2$ because it is the predecessor of $v$ and $v$ is $e_2$'s sources of targets.
+                            %           \begin{itemize}
+                            %             \item $v'$ has no pre-image in $V_{Z}$.
+                            %             Then $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [e_2]$.
+                            %             This implies that $e_2$ is the predecessor of $v'$ and hence of $e_1$. 
+                            %             This contradicts the assymetry of $<_{c,X+Y}^{\mu}$.
+                            %             \item $v'$ has a pre-image $f_{V};i_{1,E}(z) = v'$ in $V_{Z}$ (and so should $e_1$ do) and therefore $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$ because $[e_1) = \varnothing$.
+                            %                   Then, $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [i_{2,E}(e_y)] = [e_2]$.
+                            %                   This means that $e_2$ has a pre-image in $E_{Y}$ but $e_1$ is the predecessor of $e_2$ and hence they should both have pre-image in $E_{X}$ (as $e_1$ does so).
+                            %             \item Case when $v' = g_{V};i_{2,V}(z)$ is symmetric.
+                            %             \end{itemize}
+                            %     \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ (and so should $e_2$ do) in $Z_{V}$ and therefore $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$.
+                            %           Then, $<_{c,\sfrac{X+Y}{\sim}}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_{y})] = [e_1]$.
+                            %           This implies that $e_1$ has a pre-image in $E_{Y}$.
+                            %           \begin{itemize}
+                            %             \item $v'$ has no pre-image in $V_{Z}$.
+                            %                   Then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [e_2]$.
+                            %                   This implies that $e_2$ is the predecessor of $v'$ and hence of $e_1$.
+                            %                   But $[e_1) = \varnothing$ and hence we have a contradiction.
+                            %             \item $v'$ has a pre-image $f_{V};i_{1,V}(z') = v'$ in $V_{Z}$ (and so should $e_1$ do) and therefore $[f_{V}(z')) = \varnothing$ and $[g_{V}(z')) \not = \varnothing$.
+                            %                   This also means that $[g_{V}(z)) = [g_{V}(z'))$.
+                            %                   By definition we have that $e_1$ is the image of the predecessor of $g_{V}(z')$ and $e_2$ is the image of the predecessor of $g_{V}(z)$ and because their predecessors are equal it implies $e_1 = e_2$ but the relation is irreflexive.
+                            %             \item $v'$ has a pre-image $g_{V};i_{2,V}(z') = v'$ in $V_{Z}$ and therefore $[f_{V}(z')) \not = \varnothing$ and $[g_{V}(z')) = \varnothing$.
+                            %                   This is impossible as it violates the requirement that $[f_{V}(z)) = [f_{V}(z'))$ for all $z,z' \in V_{Z}$.
+                            %         \end{itemize}
+                            %     \item The case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric.
+                            %     \end{itemize}
                   \end{itemize} 
@@
                             \end{itemize}
-                    \item $[e_1) = \varnothing$ and for some $[v_i] \in s([e_1]), [v_j] \in t([e_1])$ $[[v_i]) \not = \varnothing,\; [[v_j]) \not = \varnothing$. 
-                    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$.
-                    We now need to check that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                    \begin{itemize}
-                        \item If $v$ has no pre-image in $V_{Z}$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$, then, by definition
-                              \[
-                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                              \]
-                    \end{itemize}
-                    We have that $[v] \in s([e_1])$ and $[v'] \in s([e_1])$ (or $[v'] \in t([e_1])$). 
-                    This means there exists $v_1$ such that $[v] = [v_1]$ and $v_1 \in s(e_1)$ and there exists $v_2$ such that $[v_2] = v'$ and $v_2 \in s(e_1)$.
-                    \begin{itemize}
-                        \item $v = v_1$ and $v \in s(e_1)$ and $v' = v_2$ and $v' \in s(e_1)$. 
-                              Because $[e_1) = \varnothing$, then $[v) = \varnothing$ and $[v') = \varnothing$ because the relation respects connectivity.
-                              \begin{itemize}
-                                \item If $v = f_{V};i_{1,V}(z)$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]$.
-                                      If $v' = f_{V};i_{1,V}(z')$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e'_y)]$.
-                                      We then have $e'_y = e_y$ because by assumption all nodes from $V_{Z}$ have the same immediate predecessor under $g$.
-                                \item If $v = f_{V};i_{1,V}(z)$ then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]$ and $v'$ has no pre-image in $V_{Z}$ then there necessarily exists an incident edge $[e']$ such that ...
-                                \item                                        
-                              \end{itemize} 
+                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
+                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
+                          Suppose $v_1 = f_{V};i_{1,V}(z)$ and then $v'_1 = g_{V};i_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [e_2]$ and hence $e_2$ has a pre-image in $E_{Y}$.
+                          \begin{itemize}
+                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
+                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
+                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
+                                  By the same argument, 
+                                  \[
+                                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
+                                  \]
                           \end{itemize}                    
@@
             \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $[v'] = [v]$.
+            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
             \begin{itemize}
@@
                           This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ which is what we want.
+                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
+                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
+                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$.
+                          Because $v \in s(e_1)$ there should also be a path from $e_1$ to $v'$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$
                     \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
                     \[ 
-                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{1,E}(e_x)]
-                    \]
-                    Because $e_{x}$ is the predecessor of $f_{V}(z)$ it should also be the predecessor of the pre-image of $e_1$ in $E_{X}$ as $v \in s(e_1)$.
-                    Denote the pre-image of $e_1$ in $E_{X}$ as $e_{1}^{x}$. 
-                    Then, $e_{x} = <_{c,X}^{\mu}(i_{E_{X}}(e_{1}^{x}))$ and
-
-                    \begin{align*}
-                        [i_{1,E}(e_{x})] &= [i_{1,E}(<_{c,X}^{\mu}(i_{E_{X}}(e_{1}^{x})))] = \\
-                        &= [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(i_{1,E}(e_{1}^{x})))] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = \\
-                        &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [e_2]
-                    \end{align*}
+                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
+                    \]
+                    This means that the pre-image of $e_2$ is the predecessor of $f_{V};(z)$ and hence $e_2$ should be the predecessor of $f_{V};i_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
+                    That is, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
                     \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
@@
                 \end{itemize}
-                \item $v \sim v'$ and hence there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$.
-                      This implies that $e_1$ has a pre-image in $E_{Y}$ and
-                      \[
-                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-                      \]
-                      \begin{itemize}
-                        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$ which implies $[g_{V};i_{2,V}(z)) = [v') \not = \varnothing$.
-                        That is, there exists $e_3 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$ and by e-hypergraphness of $X+Y$ $e_3 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$.
-                        Then, 
-                        \begin{align*}
-                        [e_3] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = \\
-                        &= [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = \\
-                        &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V{Y}}{\sim}}([v]))
-                    \end{align*}
-                    \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$ which implies $[g_{V};i_{2,V}(z)) = [v') = \varnothing$ and hence $[e_1) = \varnothing$.
-                          By definition,
-                          \begin{align*}
-                            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v'])) = \\
-                            &= <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v])) = [e_2]
-                        \end{align*}
-                    \end{itemize}
-            \item Cases when $v = g_{V};i_{2,V}(z)$ are symmetric.
+                \item $v \sim v'$.
+                    Follows by the same argument as above by recalling that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
             \end{itemize}    
@@
 \begin{lemma}
-    $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$ satisfies properties~\ref{def:consistency_properties}.
-    The properties (1),(2),(5) are satisfied by construction.
-    Hence we need to check wether properties (3),(4) and (6) are satisfied.
+    If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exist $e_1 \in E_{X+Y}$ such that either $[v] \in s([e_1])$ or $[v] \in t([e_1])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) \not = \varnothing$ and $e_2 \in E_{X+Y}$ such that either $[v] \in s([e_2])$ or $[v] \in t([e_2])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) \not = \varnothing$ then
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))
+    \]
     \begin{proof}
-        \begin{itemize}
-            \item Suppose $[v] \in s([e])$, then it must be the case that $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                  $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $[v] = [v']$.
-                  We then have two cases.
-                  \begin{itemize}
-                    \item $v = v'$ and then $v \in s(e)$. 
-                    This also means that $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(i_{V_{X} + V_{Y}}(v))$.
-                    By applying $([]_{V},[]_{E})$ we get 
-                    \[
-                       i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                    \]
-                        \item $v' \sim v'$ and $v' \in s(e)$ and hence $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$.
-                        By applying $([]_{V},[]_{E})$ we get 
-                        \[
-                        i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v'])
-                        \]      
-                        By functionality, 
-                        \[
-                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                        \]
-                        and hence
-                        \[
-                            i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                        \]
-                  \end{itemize}
-                  \item The case when $[v] \in t([e])$ is symmetric.
-                  \item If $i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then $[[v_1]) = [[v_2])$.
-                        The latter means that either $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$ or both functions are undefined.
-                        By definition,
-                        \[
-                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
-                        \]
-                        which means that either $v_1 \consistency_{X+Y}^{\mu} v_2$ or $v'_1 \consistency_{X+Y}^{\mu} v_2$ and $v'_1 \sim v_1$.
-                        \begin{itemize}
-                            \item $v_1 \consistency_{X+Y}^{\mu} v_2$ implies $[v_1) = [v_2)$.
-                            \begin{itemize}
-                                \item If $v_1$ does not have a pre-image in $V_{Z}$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1))]$
-                                \begin{itemize}
-                                    \item If $v_2$ does not have a pre-image in $V_{Z}$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))]$ and by assumption $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))$ and hence their images are also equal.
-                                    \item If $v_2$ has an image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and hence $[v_2) = \varnothing$ and $[v_1) = \varnothing$.
-                                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_{y})]$.
-                                \end{itemize}
-                            \end{itemize}
-                        \end{itemize}
-        \end{itemize}
+        The proof is analogous to the proof of an analogous lemma for $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
     \end{proof}
@@
 
+\begin{lemma}
+    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exist $v_1$ such that either $[v_1] \in s([e])$ or $[v_1] \in t([e])$ and $v_2$ such that either $[v_2] \in s([e])$ or $[v_2] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) \not = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) \not = \varnothing$ then
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
+    \]
+    \begin{proof}
+        The proof is analogous to the proof of the analogous lemma for $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
+    \end{proof}
+\end{lemma}
+
+\begin{corollary}
+    \label{corollary:consistency_nodes}
+    If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
+    \[
+        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \]
+\end{corollary}
+
+\begin{corollary}
+    \label{corollary:consistency_edges}
+    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
+    \[
+        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \]
+\end{corollary}
+
+\begin{lemma}
+    $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$ satisfies properties~\ref{def:consistency_properties}.
+    The property (1) is satisfied by construction because we explicitly perform a closure.
+    Hence we need to check whether properties (2),(3) and (4) are satisfied.
+    \begin{proof}
+        \begin{itemize}
+            \item Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                  $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
+                  We then have two cases.
+                  \begin{itemize}
+                    \item $v = v'$ and then $v \in s(e)$.
+                    \begin{itemize}
+                        \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))$.
+                                By applying $([]_{V},[]_{E})$ we get 
+                                \[
+                                i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                                \]
+                        \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
+                              It is then also the case that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
+                              Then, by definition, $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))$ and by reflexivity $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$
+                    \end{itemize}
+                    \item $v \sim v'$ follows by the same argument by using functionality, 
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                        \]
+                   \end{itemize}
+                  \item The case when $[v] \in t([e])$ is symmetric.
+                  \item If $i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
+                        $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
+                        By definition, either
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
+                        \]
+                        which means that either $v_1 \consistency_{X+Y}^{\mu} v_2$ or $v'_1 \consistency_{X+Y}^{\mu} v_2$ and $v'_1 \sim v_1$, or $v_1 \consistency_{X+Y}^{\mu} v_2'$ and $v_2' \sim v_2$, or $v_1' \consistency_{X+Y}^{\mu}(v_2')$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
+                        Suppose, $v_1' \consistency_{X+Y}^{\mu} v_2$ and $v_1' \sim v_1$.
+                        It implies that $[v_1') = [v_2)$ and that 
+                        \[
+                            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1'))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                        \]
+                        The rest cases are analogous.
+                        Or there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}^{\mu}(v_3') \not = \varnothing$ and
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
+                        \]
+                        Then the argument is the same but with using $v_3$ in place of $v_2$.
+                    \item Case for edges is analogous.
+        \end{itemize}
+    \end{proof}
+\end{lemma}
+
 \begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
 For this we need to check that the defined relations satisfy all the needed properties.
-
+\begin{proof}
+    This proposition then follows by taking two lemmas above.
+\end{proof}
 \end{proposition}
@@
 \begin{example}
-    Consider an example span below in Figure~\ref{fig:pushout_fail}.
+    Consider an example span in Figure~\ref{fig:pushout_fail}.
 \begin{figure}[h!]
@@
 
-    The pushout defined above does not exist for this span because neither of the set of predecessors of the images of $f$ are $\varnothing$ which violates~\ref{cond:essential_for_pushout}.
-    Luckily, we do not need arbitrary pushouts and pushouts that we actually need do exist.
+    The pushout defined above does not exist for this span because it does not satisfy the assumptions on two occasions.
+    First, the carrier e-hypergraph $Z$ is not descrete.
+    Second, it is not true that if $[f_V(z)) = \varnothing$ then $[g_{V}(z)) = \varnothing$.
 \end{example}
+
+\begin{example}
+    Consider an example span in Figure~\ref{fig:pushout_success_example}.
+    The span satisfies the assumptions since $Z$ is discrete, $[f_{V}(1)) = [f_{V}(2)) = \varnothing$ and $\consistency^{\mu}(g_{V}(1)) = \consistency^{\mu}(g_{V}(2))$.
+    The pushout can be seen in the lower left corner.
+    Note how $\consistency^{\mu}$ and $<_{c}^{\mu}$ get inherited from the right foot of the span.
+\end{example}
+
+\begin{figure}[h!]
+    \centering
+    \[
+    \scalebox{0.5}{
+        \tikzfig{combinatorial_semantics/pushout_success_example}
+    }
+    \]
+    \caption{Pushout example}
+    \label{fig:pushout_success_example}
+\end{figure}
 
@@
     % The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
-    \update{Note that pushout for such composition always exists as the interface in the middle is descrete (meaning nodes in it have no predecessors) and nodes in the images of the middle interface are at the outermost level as well.
-    Thus the span satisfies~\ref{cond:essential_for_pushout}.
+    \update{Note that pushout for such composition always exists as the interface in the middle is descrete and nodes in the images of the middle interface are at the outermost level as well.
+    Thus the span satisfies our assumptions.
     }

### 2023-12-14 sections/combinatorial_semantics.tex
Pushout diagrams changes and proof for complement

@@
 
-Naturality is implied by $+$ being a coproduct~\cite{Frobenius}.
+\update{Naturality and SMC laws are implied by $+$ being a coproduct[reference?]}.
 
@@
 Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a \textit{full} sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
+\update{Assuming that $D$ is a functor from PROP to $\catname{Hyp_{\Sigma}}$, $\catname{Csp_{D}(Hyp_{\Sigma})}$ is also a PROP and hence an SMC (Theorem 3.6~\cite{Frobenius}).}
 \end{definition}
@@
 \[
-X_{ext} \xrightarrow{f_{ext};i_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};i_1} Z_{ext}
+X_{ext} \xrightarrow{a_{ext};i_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} A +_{f_{ext};f,g_{ext};g} B \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};i_1} Z_{ext}
 \]
@@
 Since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
-$h = a;p_1 + g|_{Y' \setminus Y_{ext}};p_2$, $k = b;p_2 + f|_{Y \setminus Y_{ext}};p_1$.
-
+\[
+    h = \langle a;p_1, g|_{Y' \setminus Y_{ext}};p_2 \rangle
+\]
+
+\[
+    k = \langle b;p_2, f|_{Y \setminus Y_{ext}};p_1 \rangle
+\]
+\update{I've put copairings above}
 % \[
@@
 
-Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC.
+\update{Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC if $D$ is a functor from PROP to $\catname{EHyp_{\Sigma}}$.}
 More precisely, below are cospans for tensor product and symmetry.
@@
 
+\question{Or shall we say that symmetry is inherited from $\catname{EHyp_{\Sigma}}$ because coproduct is commutative and associative?}
 \end{definition}
@@
     \update{Note that pushout for such composition always exists as the interface in the middle is descrete and nodes in the images of the middle interface are at the outermost level as well.
-    Thus the span satisfies our assumptions.
+    Furthermore, $f_{ext};f$ and $g_{ext};g$ are monos and hence the span satisfies our assumptions.
     }
@@
     $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
-    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v) = \varnothing$.
+    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now consistent with each other and similarly for edges and nodes from $B$.
+    There is also a new hierarchical edge in $E_{C}$ that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ from $A$ and $B$ such that $[e) = \varnothing$ and $[v) = \varnothing$.
     And such that its sources and targets are top-level nodes.
@@
     This also does not increase their degrees to more than 1.
-    $f_{ext}$ (respectively, $g_{ext}$) is mono because its image $X$ (respectively, $Z$) is unchanged.
-    The morphism $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
-    Also note that $p_2$ narrowed to $B \setminus Y_{ext}$ is exactly $i_2$ and similarly for $p_1$ and $i_1$.
-    Pushout is then obtained by identifying the nodes in $A + B$ that share their pre-image in $Y_{ext}$.
-    These nodes are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
-    $g';i_2$ is equal to $g';p_2$ because the image of $g'$ does not contain any nodes with pre-image in $Y_{ext}$.
-    Next, $a$ is mono and $a;p_1$ is mono, because $p_1$ only identifies nodes with pre-image in $Y_{ext}$.
-    $h = a;p_1 + g';p_2$ and hence $h$ is mono.
+    $a_{axt};i_{1}$ is mono because $a_{ext}$ is mono and $i_{1}$ is a coproduct injection.
+    $p_1$ is mono because it only identifies $v_{1}, v_{2}$ in $V_{A}$ if there exist $y_1, y_2 \in V_{Y_{ext}}$ such that $v_{1} = f_{ext};f(y_1)$, $v_{2} = f_{ext};f(y_2)$ and there exists $v_3 \in V_{B}$ such that $v_3 = g_{ext};g(y_1)$ and $v_3 = g_{ext};g(y_2)$, but $g_{ext};g$ is mono.
+    Therefore, $a;p_1$ is mono.
+    $g|Y'\setminus Y_{ext}$ is mono because $g$ is and hence $g|Y'\setminus Y_{ext};p_2$ is mono and eventually $h$ is mono.
     Likewise, $k$ is mono.
@@
           Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
-          Hence, everything is preserved component-wise by copairing and coproduct.
+          Hence, everything is preserved component-wise by coproduct.
     \item Symmetry~\ref{morphism:symmetry} is an mda-cospan.
@@
 
-Consider an mda-cospan~\ref{mda-cospan} and an edge $e$ in $E_{A}$ such that $l(e) = \bot$.
+Consider an mda-cospan
+\[
+    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} C \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
+\]
+and an edge $e$ in $E_{C}$ such that $l(e) = \bot$.
 Then we take all nodes that are immediate successors of $e$.
@@
 
-% \begin{definition}[Set of inner subgraphs of an edge $e \in E_\mathcal{G}$]
-
-% Consider an edge $e \in \mathcal{G}$ such that $l(e) = \bot$.
-We denote with $\mathcal{H}_{e}$ a set of inner subgraphs of the edge $e$, that is $\mathcal{H}_e \triangleq \{e' \in E_\mathcal{G} \text{ and } v' \in V_{\mathcal{G}} \; | \; \exists n,m > 0 \; .
-\; p^n(e') = e \text{ and } p^m(v') = e\}$.
-\update{And such that $\mathcal{H}_{e}$ inherits all the relations from $\mathcal{H}$ for the corresponding nodes and edges.
-What happens to the parent of e.g.
-$v$ for which $p_{V}(v) = e$?}
-% \end{definition}
-
-% add sketch of a proof for saying that a complement is an equivalence, define partition
-% \begin{definition}[Instance subgraph of an edge $e \in E_\mathcal{G}$]
-
-% Consider a partition of $\mathcal{H}_e$ according to a complement of the relation $[\#_V,\#_E]$ denoted as $[\bar{\#}_{V},\bar{\#}_{E}]$.
-% Note that these relations are equivalence relations as the original relation is irreflexive.
-We call each $i^{th}$ subset of this partition as $i^{th}$ instance subgraph of $e$, that is, every $\mathcal{H}^i_e \subseteq \mathcal{H}_{e,\bar{\#}}$ is an instance subgraph of $e$.
-
-% \end{definition}
-
 \begin{definition}[Well-typed cospan]
 
-    Consider a hierarchical edge $e$ in $E_{A}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively.
-    Then consider partitions $V_{e, in}^{\bar{\#}}$ and $V_{e,out}^{\bar{\#}}$  of the nodes according to $\bar{\#}$.
-    We call $e$ well-typed if for all $v \in V_{e,in}^{\bar{\#}}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\bar{\#}}$ $|v| = |t(e)|$.
+    Consider a hierarchical edge $e$ in $E_{C}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively.
+    Then consider partitions $V_{e, in}^{\consistency}$ and $V_{e,out}^{\consistency}$  of the nodes according to $\consistency^{\mu}$ relation.
+    We call $e$ well-typed if for all $v \in V_{e,in}^{\consistency}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\consistency}$ $|v| = |t(e)|$.
     We will call an mda-cospan in $\MdaEcospans$ well-typed if all hierarchical edges in its carrier are well-typed.
@@
 
-For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\MdaCospans$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\# = \varnothing$, and $<_{c} = \varnothing$.
+For any mda-cospan $X \xrightarrow{} A \xleftarrow{} Y$ in $\MdaCospans$ there is a corresponding well-typed mda-cospan in $\WellTypedMdaEcospans$: $X \xrightarrow{} X \xrightarrow{} B \xleftarrow{} Y \xleftarrow{} Y$, where $B$ is an e-hypergraph lifting of $A$ where  $\consistency^{\mu} = \varnothing$, and $<_{c} = \varnothing$.
 It is well-typed on the nose as it has no hierarchical edges.
 
-% \update{Two cospans $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$, $p \xrightarrow{} \mathcal{G} \xleftarrow{} q$ in $\MdaCospans$ are composable when $out^{-1}_{out}(\mathcal{G})$ and $in^{-1}_{out}(\mathcal{H})$ are isomorphic.
-% The result of such composition is depicted below
-% }
+\subsection{DPO rewriting}
+
+\Aleksei{Should this be a separate section rather than a subsection?}
+
+% \begin{definition}[\cite{Frobenius}]
+% A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e.
+pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities.
+% Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as in Figure~\ref{fig:prop_rewriting}, for some rule $\langle l,r \rangle \in \mathcal{R}$
+% \end{definition}
+
+% \begin{figure}[h!]
+
 %     \[
-% \adjustbox{scale=0.75,center}{
-% \begin{tikzcd}
-%             &                         &                                   & k \arrow[ld] \arrow[rd] \arrow[lld] \arrow[rrd] &                                  &                         &             \\
-% n \arrow[r] & \mathcal{G} \arrow[rrd] & m \arrow[l]                       &                                                 & p \arrow[r]                      & \mathcal{H} \arrow[lld] & q \arrow[l] \\
-%             &                         & n \cup (p \setminus k)  \arrow[r] & \mathcal{G};\mathcal{H}                         & q \cup (m \setminus k) \arrow[l] &                         &            
-% \end{tikzcd}
+%     \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
+%     \]
+%     \caption{Rewriting in $\textsf{PROP}(\Sigma)$}
+%     \label{fig:prop_rewriting}
+% \end{figure}
+
+% \begin{definition}
+% Given a rewriting system $\mathcal{R}$ for a $\textsf{PROP}(\Sigma)$ we define 
+%  a rewriting system $\mathcal{R}^{+}$ in a $\textsf{PROP}^{+}(\Sigma)$  as a set of pairs $\langle l , r \rangle$ for each pair $\langle l, r \rangle$ in $\mathcal{R}$.
+% However, the rewriting happens modulo semilattice enriched laws rather than SMC laws.
+% We define what $a \Rightarrow_{\mathcal{R}^+} b$ means for some rule $\langle l , r\rangle $ based on the structure of $a$ and $b$.
+
+% If they are decomposable as in Figure~\ref{fig:prop_rewriting} then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
+% Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above ($\Rightarrow_{\mathcal{R}})$)) and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as in Figure~\ref{fig:enriched_prop_rewriting}.
+
+% \begin{figure}
+
+%     \[
+%     \scalebox{0.6}{
+%     \tikzfig{combinatorial_semantics/prop_plus_rewriting}
 %     }
 %     \]
-
-% where $k = out^{-1}_{out}(\mathcal{G})$ and $\mathcal{G;H}$ is a pushout of $\mathcal{G} \xleftarrow{} k \xrightarrow{} \mathcal{H}$.
-New input interface for the composition is obtained by adding the input interface of $\mathcal{H}$ to $n$ and removing $p$ (and likewise for the output interface).
-$p \setminus k$ can be defined as a following pushout complement (and likewise for $m \setminus k$):
-
-% \[
-% \adjustbox{scale=0.75,center}{
-% \begin{tikzcd}
-% k \arrow[d] & 0 \arrow[l] \arrow[d]   \\
-% p           & p \setminus k \arrow[l]
-% \end{tikzcd}
-% }
-% \]
-
-% then $n \cup (p \setminus k)$ is just a coproduct, since $n$ and $(p \setminus k)$ are disjoint.
-
-% We will also use the notion of a boundary complement to preserve monogomousness when rewriting.
-
-% \begin{definition}[Boundary complement]
-%     For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
-% and mono $f : L \to G$, a pushout complement as depicted below
-
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.3\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-%     \label{fig:boundary_complement}
+%     \caption{Rewriting in $\textsf{PROP}^{+}(\Sigma)$}
+%     \label{fig:enriched_prop_rewriting}
 % \end{figure}
 
-% is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to L^{\bot}$ and $d_2 : m \to L^{\bot}$ making the above triangle commute and such that
-% \[
-%     n + j \xrightarrow{[d_1,c_2]} L^{\bot} \xleftarrow{[d_2,c_1]} m + i
-% \]
-% is a monogamous cospan
 % \end{definition}
-
-\subsection{Syntactic rewriting of PROPs}
-
-\Aleksei{Should this be a separate section rather than a subsection?}
-
-\begin{definition}[\cite{Frobenius}]
-A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e.
-pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities.
-Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as in Figure~\ref{fig:prop_rewriting}, for some rule $\langle l,r \rangle \in \mathcal{R}$
-\end{definition}
-
-\begin{figure}[h!]
-
-    \[
-    \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
-    \]
-    \caption{Rewriting in $\textsf{PROP}(\Sigma)$}
-    \label{fig:prop_rewriting}
-\end{figure}
-
-\begin{definition}
-Given a rewriting system $\mathcal{R}$ for a $\textsf{PROP}(\Sigma)$ we define 
- a rewriting system $\mathcal{R}^{+}$ in a $\textsf{PROP}^{+}(\Sigma)$  as a set of pairs $\langle l , r \rangle$ for each pair $\langle l, r \rangle$ in $\mathcal{R}$.
-However, the rewriting happens modulo semilattice enriched laws rather than SMC laws.
-We define what $a \Rightarrow_{\mathcal{R}^+} b$ means for some rule $\langle l , r\rangle $ based on the structure of $a$ and $b$.
-
-If they are decomposable as in Figure~\ref{fig:prop_rewriting} then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
-Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above ($\Rightarrow_{\mathcal{R}})$)) and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as in Figure~\ref{fig:enriched_prop_rewriting}.
-
-\begin{figure}
-
-    \[
-    \scalebox{0.6}{
-    \tikzfig{combinatorial_semantics/prop_plus_rewriting}
-    }
-    \]
-    \caption{Rewriting in $\textsf{PROP}^{+}(\Sigma)$}
-    \label{fig:enriched_prop_rewriting}
-\end{figure}
-
-\end{definition}
-
-% \update{
-
-% \begin{definition}
-%     Functor $\llbracket \cdot \rrbracket : \text{PROP}^{+} \to \catname{MACsp_{D}(EHyp_{\Sigma})}$.
-
-%     We will build this functor inductively.
-As noted above, $\MdaCospans$ is a full subcategory of $\catname{MACsp_{D}(EHyp_{\Sigma})}$ and the acting of this functor on generators can be found in~\cite{Frobenius}.
-We need to define what $+$ from $\text{PROP}^{+}$ in $\catname{MACsp_{D}(EHyp_{\Sigma})}$ is.
-The definition could be found in Figure~\ref{fig:f+g} where well-typed mda-cospans for images of arbitrary morphisms $f, g : m \to n$ are depicted on the left and the image of $f + g$ is depicted on the right.
-A new edge is introduced which is the immediate predecessor of arbitrary e-hypergraphs $f$ and all edges and nodes of $f$ and $g$ are in conflict.
-The nodes are added to the interfaces and to the new edge to preserve well-typedness.
-
-% \end{definition}
-
-% }
 
@@
 
-\begin{definition}[$\ulcorner \cdot \urcorner$]
+\begin{definition}[\;$\ulcorner \cdot \urcorner$]
 
@@
 
-\begin{proposition}
-
-    $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$ is mda-cospan in $\MdaCospans$ iff $0 \xrightarrow{} \mathcal{G} \xleftarrow{} n + m$ is an mda-cospan.
-\end{proposition}
-
-In particular, the above proposition means that by fullness of $\llbracket \cdot \rrbracket$ there exists a corresponding SMC morphism for some $0 \xrightarrow{} \mathcal{G}\xleftarrow{} n + m$.
+% \begin{proposition}
+%     $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$ is mda-cospan in $\MdaCospans$ iff $0 \xrightarrow{} \mathcal{G} \xleftarrow{} n + m$ is an mda-cospan.
+% \end{proposition}
+
+% In particular, the above proposition means that by fullness of $\llbracket \cdot \rrbracket$ there exists a corresponding SMC morphism for some $0 \xrightarrow{} \mathcal{G}\xleftarrow{} n + m$.
 Theorem above essentially makes two routes below. In forward direction:
@@
 Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-\begin{figure}
-    \centering
-    \[
-    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-    \label{fig:interface_change_example}
-\end{figure}
 
@@
 % https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
-\[\begin{tikzcd}
-	{l_i+l_o} && L && {x_{ext} + y_{ext}} && R && {r_i+r_o} \\
-	\\
-	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
-	\\
-	&&&& {n_{ext}+m_{ext}}
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
-	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
-	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
-	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=3-1, to=3-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-7, to=3-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-7]
-	\arrow[from=1-9, to=1-7]
-	\arrow[from=3-9, to=3-7]
-\end{tikzcd}\]
-
-The original DPOI square is shown in red.
+\[
+ \scalebox{0.75}{
+    \tikzfig{combinatorial_semantics/DPOI_square_coloured}
+ }
+\]
+
+The parts of the original DPOI square is shown in blue.
 The extra bits simply account for more complicated interfaces.
@@
 The bottom half puts the same restrictions on $G$ and $H$.
-% The extra squares on both sides say that there should exist a matching of the interfaces together with a matching of e-hypergraphs.
-This requirement was implicit in the case of $\catname{MACsp_{D}(Hyp_{\Sigma}}$ because the whole interfaces and the outermost interfaces coincided.
-Because now the interfaces can change, we need to keep track of which parts of the interfaces are subjected to changes.
+For the pushout to exist we impose our assumptions~\ref{pushout:assumptions} on the correspoding arrow.
 
@@
 % https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-\[\begin{tikzcd}
-	{l_i+l_o} && L && {x_{ext} + y_{ext}} \\
-	\\
-	{g_i+g_o} && {G^{\urcorner}} && C \\
-	\\
-	&&&& {n_{ext}+m_{ext}}
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
-	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
-	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
-	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=3-1, to=3-3]
-\end{tikzcd}\]
+\[
+\scalebox{0.75}{
+    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
+}
+\]
 
@@
 
-% Intuitively, these pushout squares mean that when we delete the occurrence of $L$ from $G$ we also remove the occurrence of $L$'s interface within $G$'s interface preserving the outermost interfaces $i + j$.
-The highlighted arrows form the pushout square for the original boundary complement.
-
-% We call morphisms $i + j \to C \to G$ a boundary complement if
-
-% \begin{enumerate}
-%     \item The square above commutes and $L \xrightarrow{} G \xleftarrow{} C$ is a pushout
-%     \item There exists an mda cospan $g_i \setminus (l_i \setminus i) + j \xrightarrow{} C \xleftarrow{} g_o \setminus (l_o \setminus j) + i$.
-\question{I am unsure about categorical notation for this set difference.
-Maybe it can be paraphrased in the following way?}
-%     \item There exists an mda cospan $c_i + j \xrightarrow{} C \xleftarrow{} c_o + i$ where $c_i$ and $c_o$ are the following pushout complements in the category of discrete hypergraphs:
-% \end{enumerate}
+\begin{figure}
+    \centering
+    \[
+    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
+    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
+    \label{fig:interface_change_example}
+\end{figure}
 
@@
 First, let's check that this notion generalises the ordinary boundary complement from~\ref{def:boundary_original}.
-For this, we assume that child and conflict relations are empty in which case $x_{ext} + y_{ext}$ and $l_i + l_o$ coincide as well as $g_i + g_o$ and $n_{ext} + m_{ext}$.
+For this, we assume that $<_{c}$ and $\consistency^{\mu}$ relations are empty in which case $x_{ext} + y_{ext}$ and $l_i + l_o$ coincide as well as $g_i + g_o$ and $n_{ext} + m_{ext}$.
 Hence, $g_i \setminus (l_i \setminus x_{ext}) + y_{ext} = g_i + y_{ext} = n_{ext} + y_{ext}$ and the same for the output interfaces.
@@
     \begin{proof}
-        \update{TODO}
+        \update{I will try to reinvent a similar proof from~\cite{Frobenius2} as the cited proof does not make sense to me}
+        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of nodes and hyperedges.
+        Hence, pushout complement should yield pushout complements for the sets of nodes and hyperedges.
+        In particular,
+          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
+    \begin{figure}[h!]
+        \begin{subfigure}{0.45\textwidth}
+            \[\begin{tikzcd}
+                {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
+                \\
+                {V_{G}^{\urcorner}} &&&& {V_{C}}
+                \arrow["{b_{V}}", from=1-5, to=1-3]
+                \arrow["{a_{V}}", from=1-3, to=1-1]
+                \arrow["{f_{V}}", from=1-1, to=3-1]
+                \arrow["{g_{V}}"', from=3-5, to=3-1]
+                \arrow["c"', from=1-5, to=3-5]
+            \end{tikzcd}\]
+        \end{subfigure}
+        \hfill
+        \begin{subfigure}{0.45\textwidth}
+           % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
+        \[
+        \begin{tikzcd}
+            {E_{L}} && 0 && 0 \\
+            \\
+            {E_{G}^{\urcorner}} &&&& {E_{C}}
+            \arrow["{b_{E}}", from=1-5, to=1-3]
+            \arrow["{a_{E}}", from=1-3, to=1-1]
+            \arrow["{f_{E}}", from=1-1, to=3-1]
+            \arrow["{g_{E}}"', from=3-5, to=3-1]
+            \arrow["c"', from=1-5, to=3-5]
+        \end{tikzcd}
+        \]
+        \end{subfigure}
+    \end{figure}
+    $V$s and $E$s are sets of nodes and hyperedges respectively, marked squares are pushouts and $C$s are pushout complements.
+    Because $i+j$ is discrete, $E_{G}^{\urcorner}$ is the disjoint union of $E_{L}$ and $E_{C}$ and hence $E_{C} = E_{G}^{\urcorner} \setminus E_{L}$.
+    Then, since all the arrows are monos, the left square can be rewritten as follows.
+    \[
+      % https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3IFxcY29uZyBpICsgaiArIHoiXSxbMiwxLCJiX3tWfSJdLFsxLDAsImFfe1Z9Il0sWzAsMywiZl97Vn0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
+    \begin{tikzcd}
+        {i + j + x + y} && {i+j + x} && {i+j} \\
+        \\
+        {i + j + x + y + z} &&&& {i + j + w \cong i + j + z}
+        \arrow["{b_{V}}", from=1-5, to=1-3]
+        \arrow["{a_{V}}", from=1-3, to=1-1]
+        \arrow["{f_{V}}", from=1-1, to=3-1]
+        \arrow["{g_{E}}"', from=3-5, to=3-1]
+        \arrow["c"', from=1-5, to=3-5]
+    \end{tikzcd}
+    \]
+    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for nodes is $i + j + z$ up to isomorphism.
+    To show that pushout complement is unique we next need to show that source and target maps are unique.
+    Suppose that there exist
+    \[
+        C_{1} = (V_{C}, V_{E}, s_{1},t)
+    \]
+    and
+    \[
+        C_{2} = (V_{C},V_{E},s_{2},t)
+    \]
+    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(h)) = s_{G}^{*}(g_{E}(h)) = g_{V}^{*}(s_{2}(h))$.
+    Because $g$ is mono, $g_{V}^{*}(s_{1}(h)) = g_{V}^{*}(s_{2}(h))$ implies $s_{1}(h) = s_{2}(h)$.
+    Similarly for targets.
+    We also need to show that $<_{c}$ and $\consistency^{\mu}$ are unique.
+    Suppose that there exist
+    \[
+        C_{1} = (V_{C}, V_{E}, s,t, <_{1})
+    \]
+    and
+    \[
+        C_{2} = (V_{C},V_{E},s,t, <_{2})
+    \]
+    Because $g$ is homomorphism, it must be the case that $g_{E}(<_{1}(i_{V_{C}}(v))) = <_{G}(i_{V_{G}}(g_{V}(v))) = g_{E}(<_{2}(i_{V_{C}}(v)))$ which implies $<_{2}(i_{V_{C}}(v)) = <_{1}(i_{V_{C}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
+    Similarly for edges.
+    Now suppose that $<_{1}(i_{V_{C}}(v)) = e$ and $<_{2}(i_{V_{C}}(v))$ is undefined.
+    Then, $g_{E}(e) = <_{G}(i_{V_{G}}(g_{V}(v)))$ and there exists $v' \in V_{L}$ and $v'' \in V_{C_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
+    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $C_{1}$, but then it violates our assumptions as there exists some node $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = b_{V};a_{V}(u)$ and $[v') \not = \varnothing$.
+    Therefore we get a contradiction. 
+    Similarly for edges.
+    Finally, let's suppose there exist
+    \[
+        C_{1} = (V_{C}, V_{E}, s,t, <, \consistency_{1}^{\mu})
+    \]
+    and
+    \[
+        C_{2} = (V_{C},V_{E},s,t,<, \consistency_{2}^{\mu})
+    \]
+    Because $g$ is a homomorphism, we have
+    \[
+        \langle g_{V};i_{V_{G}}, g_{E};i_{E_{G}} \rangle^{*}(\consistency_{C_1}^{\mu}(i_{V_{C}}(v))) \subseteq \consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v))
+    \]
+    \[
+        \langle g_{V};i_{V_{G}}, g_{E};i_{E_{G}} \rangle^{*}(\consistency_{C_2}^{\mu}(i_{V_{C}}(v))) \subseteq \consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v))  
+    \]
+    \begin{itemize}
+        \item If $\consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v)) = \varnothing$ then both lefthand sides should be $\varnothing$ and the relations are equal.
+        \item Suppose that $\consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v)) \not = \varnothing$ which further means that $[g_{V};i_{V_{G}}(v)) \not = \varnothing$.
+        \begin{itemize}
+            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \subseteq \consistency_{C_2}^{\mu}(i_{V_{C}}(v))$.
+                  This means that there exists $v'$ such that $v' \consistency_{C_{2}}^{\mu} v$ and $v' \not \consistency_{C_{1}}^{\mu} v$.
+                  Because $g_{V}(v) \consistency^{\mu} v'$ it means that there exists $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{L}$ such that $w \consistency_{L}^{\mu} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
+                  This again contradicts our assumptions as $u \consistency_{C_{2}}^{\mu} u'$ and $w \consistency_{L}^{\mu} w'$.
+            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \cap \consistency_{C_2}^{\mu}(i_{V_{C}}(v)) = \varnothing$
+                  Similar to the above.
+            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \cap \consistency_{C_2}^{\mu}(i_{V_{C}}(v)) \not = \varnothing$
+                  Similar to the above.
+        \end{itemize}
+    \end{itemize}
     \end{proof}
+    \update{Can we check the proof? I haven't used all the conditions from monogamous-ness above.
+    Although, if the morphism $i+j \to L$ is mono, the complement is unique for not-monogamous hypergraphs as well.}
 \end{proposition}
@@
 \question{
-\begin{definition}[Convex down-closed conflict reflecting match] 
-    We call $m : L \to G$ in $\catname{EHyp_{\Sigma}}$ a convex down-closed conflict reflecting match if
+\begin{definition}[Convex down-closed match] 
+    We call $m : L \to G$ in $\catname{EHyp_{\Sigma}}$ a convex down-closed match if
 
@@
         \item $m(L)$ is a convex down-closed subgraph of $G$
-        \item if $m(x_1) \# m(x_2)$ then $x_1 \# x_2$
+        \item \update{I've removed the requirement that consistency should be reflected because well typedness prohibits otherwise formed matches.}
     \end{enumerate}
@@
     Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n_{ext} + m_{ext}$ 
-    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\WellTypedMdaEcospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $l_i + l_o \xrightarrow{} L \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} R \xleftarrow{} r_i + r_o$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
+    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $l_i + l_o \xrightarrow{} L \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} R \xleftarrow{} r_i + r_o$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
 
 % https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
-\[\begin{tikzcd}
-	{l_i+l_o} && L && {x_{ext} + y_{ext}} && R && {r_i+r_o} \\
-	\\
-	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
-	\\
-	&&&& {n_{ext}+m_{ext}}
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-3]
-	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=1-3, to=3-3]
-	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=3-5]
-	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-5]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-3]
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=3-1, to=3-3]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=5-5, to=3-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=3-5, to=3-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-7, to=3-7]
-	\arrow[draw={rgb,255:red,214;green,92;blue,92}, from=1-5, to=1-7]
-	\arrow[from=1-9, to=1-7]
-	\arrow[from=3-9, to=3-7]
-\end{tikzcd}\]
+\[\scalebox{0.75}{
+    \tikzfig{combinatorial_semantics/DPOI_square}
+    }
+\]
     and the following conditions hold
@@
     \begin{enumerate}
-        \item $m : L \to G$ is a convex down-closed conflict reflecting match;
+        \item $m : L \to G$ is a convex down-closed match;
         \item $x_{ext} + y_{ext} \to C \to G$ is a boundary complement in the leftmost pushout.
@@
 
-% using the two pushout squares below.
-% Intuitively, these squares combine the interfaces of $R$ and $C$ and identify those bits that share the same pre-image in the outermost interface $i + j$.
-
-% \[\begin{tikzcd}
-% i \arrow[r] \arrow[d] & r_i \arrow[d] & j \arrow[r] \arrow[d] & r_o \arrow[d] \\
-% c_i \arrow[r]         & h_i           & c_o \arrow[r]         & h_o          
-% \end{tikzcd}\]
-
-% Going back to the example from the figure~\ref{fig:dpoir}, $h_i = \{0'',0,0',2',2\}$, $r_i = \{0,0',2'\}$, $r_o = \{1',3',1\}$, $h_o = \{1',3',3,1,1''\}$, $c_i = \{0'',2,0\}$, $c_o = \{1'',1,3\}$, $i = \{0\}$, $j = \{1\}$, where $h_i$ is exactly $c_i \coprod_{\equiv_{i}} r_i = \{0'',2,0\} \coprod_{\equiv_{i}} \{0,0',2'\} = \{0'',0,0',2',2\}$ and the same for $h_o$.
+We then say that an mda-cospan 
+\[
+    n_{ext} \xrightarrow{a_1} g_{i} \xrightarrow{a_2} G \xleftarrow{b_2} g_{o} \xleftarrow{b_1} m_{ext}
+\]
+rewrites convexly into another mda-cospan
+
+\[
+    n_{ext} \xrightarrow{a_1} h_{i} \xrightarrow{a_2} H \xleftarrow{b_2} h_{o} \xleftarrow{b_1} m_{ext}
+\]
+
+under a rewrite rule 
+\[
+    \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
+    x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
+\]
+
+if 
+\[
+    G \xleftarrow{\langle a_2,b_2 \rangle} g_i + g_{o} \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext} \Rrightarrow H \xleftarrow{\langle a_2,b_2 \rangle} h_i + h_{o} \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext}
+\]
+under a rewrite rule
+\[
+    \langle l_{i} + l_{o} \xrightarrow{\langle a_2, b_2 \rangle} L \xleftarrow{\langle a_1, b_1 \rangle} x_{ext} + y_{ext},
+    r_{i} + r_{o} \xrightarrow{\langle a_2, b_2 \rangle} R \xleftarrow{\langle a_1, b_1 \rangle} x_{ext} + y_{ext} 
+    \rangle
+\]
 
@@
 Well-typedness is restored once the right-hand side of the rewrite rule is glued into $C$.
-\question{Uniqueness of this pushout complement? Is it a presheaf? As a presheaf category is adhesive, $\catname{Csp_{D}(EHyp)}$ can not be a presheaf as it is not adhesive.}
 \end{proof}
@@
 \Aleksei{Check the proof}
-
-\begin{lemma}(Decomposition).
-Let $m \xrightarrow{} G \xleftarrow{} n$ be a monogamous acyclic cospan and $L$ a convex and down-closed sub-hypergraph of $G$.
-Then there exists $k \in \mathcal{N}$ and a unique cospan $i \xrightarrow{} L \xleftarrow{} j$ such that G factors as $\ldots$
-
-\update{TODO: this lemma is needed to prove soundness of rewriting}
-\end{lemma}
 

### 2023-12-14 sections/combinatorial_semantics.tex
Pushout existence for structural rule

@@
     After we identify a convex down-closed image of the left-hand side of this rule within a graph $G$ we remove the image by keeping only the outermost nodes.
+    Note that because the image of all $a$'s and $b$'s are outermost nodes, we are allowed to pick any monomorphism for our pushout complement.
+    And for the same reason the pushout for the right square exists.
     Then we simply glue the right-hand side of the rule into the hole.
-    Since we glue along the outermost nodes this does not introduce any conflict with regard to child relation.
     Such equations are not absorbed by e-hypergraph representation unlike SMC equations because the right-hand side contains more edges, i.e. graphs $G$ and $H$ are not isomorphic.
     Similar argument applies to other rules as well.
-
 \end{proof}

### 2023-12-19 sections/combinatorial_semantics.tex
TODOs added

@@
 
-Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we add them as structural rewrite rules.
-Recall that the category $\catname{EHyp_{\Sigma}}$ does not have all pushouts and hence we need to explicitly show that the pushouts that we need exist.
+Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we will need to turn the latter into a semilattice-enriced category.
+We already have a $+$ in $\WellTypedMdaEcospans$ so we need to require that it satisfies a particular set of equations, for example, that $h \otimes (f + g) = h \otimes f + h \otimes g$.
+\update{How to formally state this? Because $h \otimes (f+g)$ is an SMC term from $\WellTypedMdaEcospans$ which is just an SMC so far.}
+We will add such equations as DPOI rewrite rules and show that for each such rule the corresponding pushout complement and the pushout exist.
 
@@
 \end{proposition}
+
+\update{How do we state the correspondence between term equations and their e-hypergraph counterpart without a functor?}
 

### 2023-12-19 sections/soundness_and_completeness.tex
TODOs added

--- 2023-10-17 sections/soundness_and_completeness.tex
@@
 \[
-[\cdot] : \textsf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans
+[\cdot] : \textsf{PROP}^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}
 \]
-
-by freeness, it suffices to define how the functor acts on generators.
-As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \MdaCospans$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}.
+where $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is $\WellTypedMdaEcospans$ quotiented by structural rewrite rules which makes this category into a semilattice-enrhiched one.
+By freeness, it suffices to define how the functor acts on generators \update{Identities and generators from the signature?}.
+% As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \MdaCospans$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}.
 Everything else follows by defining
@@
     \item $[ f + g ] = [ f ] + [ g ]$  for $f,g \in \Sigma$;
+    \item \update{Symmetry?}
 \end{enumerate}
 
-this also makes the functor into a semilattice-enriched SMC one on the nose.
+this also makes the functor into a symmetric monoidal semilattice-enriched SMC one on the nose.
 % \update{This essentially means that $[\cdot] = \llbracket \cdot \rrbracket ; \mathcal{I}$ plus three equations above.}
@@
             }
+        \question{Can we use normal form here as well? $f = g$ means there exists $h$ such that $f =_{SMC^{+}} h$ and $h =_{SMC^{+}}$ and $h$ is of the form $\phi_1 + \ldots + \phi_n$?
+                  Then we need to show that $[f] \Rrightarrow_{\mathcal{S}} [h]$ and $[h] \Rrightarrow_{\mathcal{S}} [g]$.
+        }
     \end{proof}
@@
     \begin{proof}
-        Consider the commutative diagram below.
+        Consider the commutative diagram below \update{Is it obvious that the diagram commutes?}.
 
@@
 Essentially we need to show that if $\llbracket f \rrbracket = \llbracket g \rrbracket$ then $f = g$ which is equivalent to if $f \not = g$ then $\llbracket f \rrbracket \not = \llbracket g \rrbracket$.
-First, we note that we have a \question{weak (modulo commutativity)} normal form for any term in $PROP^{+}(\Sigma)$ which is achieved by applying distributivity for tensor and composition as well as unpacking equations, that is, every term is representable as in Figure~\ref{fig:normal_form_string} where $\phi_1,\;\ldots\;,\phi_n$ are pure SMC terms.
+First, we note that we have a \question{weak (modulo commutativity)} normal form for any term \update{string diagram?} in $PROP^{+}(\Sigma)$ which is achieved by applying distributivity for tensor and composition as well as unpacking equations, that is, every term is representable as in Figure~\ref{fig:normal_form_string} where $\phi_1,\;\ldots\;,\phi_n$ are pure SMC terms.
 \update{It is not actaully a `normal' form because we can still apply SMC equations, so it should be normal modulo SMC and commutativity}.
@@
 We will write such forms as $\phi_1 + \ldots \phi_n$ and $A_1 + \ldots A_n$ respectively.
-Then, $\llbracket f \rrbracket \not = \llbracket g \rrbracket$ leads to two options
-
-% \begin{itemize}
-%     \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_m$ where $n \not = m$.
-By functoriality and weak normal form in $\textsf{PROP}^{+}(\Sigma)$, $f = a_1 + \ldots + a_n$, $g = b_1 + \ldots b_m$ and $\llbracket f \rrbracket = \llbracket a_1 \rrbracket + \ldots + \llbracket a_n \rrbracket$, $\llbracket g \rrbracket = \llbracket b_1 \rrbracket + \ldots \llbracket b_m \rrbracket$.
-This essentially means that $f \not = g$ as it has different number of summands.
-%     \item $\llbracket f \rrbracket$ = $A_1 + \ldots A_n$ and $\llbracket g \rrbracket = B_1 + \ldots B_n$.
-Then there exists a permutation $\sigma$ and an index $i$ such that $ A_i  \not =  B_{\sigma(i)}$.
-By functoriality there exists $a_i$ such that $\llbracket a_i \rrbracket = A_i$ and $b_j$ such that $\llbracket b_j \rrbracket = B_{\sigma(i)}$.
-As $A_i$ does not have any hierarchical edges it effectively lives in $\mathcal{I}(\catname{MACsp_{D}(Hyp_{\Sigma})})$.
-By faithfulness of $\llbracket \cdot \rrbracket;\mathcal{I}$ it follows that $a_i \not = b_j$ and $f \not = g$.
-\update{Prove that $\llbracket \cdot \rrbracket; \mathcal{I}$ is faithful?}
-% \end{itemize}
-
 Now, assume that $f = \phi_1 + \ldots + \phi_n$ and $g = \psi_1 + \ldots + \psi_m$.
-Then, by functoriality, $[f] = [A_1] + \ldots + [A_n]$ and $g = [B_1] + \ldots + [B_m]$. 
-$f \not g$ gives us two options.
+Then, by functoriality, $[f] = [\phi_1] + \ldots + [\phi_n] = A_1 + \ldots + A_n$ and $g = [\psi_1] + \ldots + [\psi_m] = B_{1} + \ldots + B_{m}$. 
+$f \not = g$ gives us two options.
 
@@
     \item $n \not = m$. Then $[f] \not = [g]$ as they are simply not isomorphic cospans of e-hypegraphs as they have different number of summands.
-    \item $n = m$.
-Then there exists a permutation $\sigma$ and index $i$ such that $\phi_i \not = \psi_{\sigma(i)}$.
+    \item $n = m$. Then there exists a permutation $\sigma$ and index $i$ such that $\phi_i \not = \psi_{\sigma(i)}$. 
     Remember that these are just SMC terms and by faithfulness of $\llbracket \cdot \rrbracket;\mathcal{I}$ we get $A_i \not = B_{\sigma(i)}$ and $[f] \not = [g]$.
@@
 
+% \begin{theorem}
+
+%     Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$ and $\langle l, r \rangle \in \mathcal{R}$.
+%     $a \Rightarrow{}^{+}_{\langle l, r \rangle} b$ iff $[a] \Rrightarrow{}^{+}_{\langle [l], [r] \rangle} [b]$.
+%     \begin{proof}
+%         $a \Rrightarrow^{+}_{\langle l, r \rangle} b$ means that either
+%         \begin{itemize}
+%             \item $a$ and $b$ are decomposable as
+%                   \[
+%                     \scalebox{0.75}{
+%                         \tikzfig{combinatorial_semantics/prop_rewriting}
+%                     }
+%                   \]
+%                   \question{Here we can probably just reuse the proof going to $\ulcorner a \urcorner$ and then the result follows by functoriality of $[\cdot]$ (or $\llangle \cdot \rrangle$)}
+%             \item or
+%                  \[
+%                     \scalebox{0.75}{
+%                         \tikzfig{combinatorial_semantics/prop_plus_rewriting}
+%                     }   
+%                  \]
+%                  and $a' \Rightarrow_{\langle l,r \rangle} b'$. This gives us the following decomposition in cospans
+%                  % https://q.uiver.app/#q=WzAsMjcsWzAsNCwibSJdLFsxLDQsIkFfMSJdLFsyLDQsImkgKyBrIl0sWzMsNCwiOyJdLFs0LDAsImkiXSxbNSwwLCJYXzEiXSxbNiwwLCJqIl0sWzUsMiwiXFx2ZG90cyJdLFs1LDQsIkEnIl0sWzQsNCwiaSJdLFs2LDQsImoiXSxbNSwxLCIrIl0sWzUsMywiKyJdLFs1LDUsIisiXSxbNSw2LCJcXHZkb3RzIl0sWzUsNywiKyJdLFs0LDgsImkiXSxbNSw4LCJYX3ciXSxbNiw4LCJqIl0sWzcsNCwiXFxvdGltZXMiXSxbOCw0LCJrIl0sWzksNCwiayJdLFsxMCw0LCJrIl0sWzExLDQsIjsiXSxbMTIsNCwiaiArIGsiXSxbMTMsNCwiQV8yIl0sWzE0LDQsIm4iXSxbMCwxXSxbMiwxXSxbNCw1XSxbNiw1XSxbOSw4XSxbMTAsOF0sWzE4LDE3XSxbMTYsMTddLFsyMCwyMV0sWzIyLDIxXSxbMjYsMjVdLFsyNCwyNV1d
+%                 \[
+%                     \adjustbox{scale=0.5, center}{
+%                 % \scalebox{0.75}{
+%                     \begin{tikzcd}
+%                     &&&& i & {X_1} & j \\
+%                     &&&&& {+} \\
+%                     &&&&& \vdots \\
+%                     &&&&& {+} \\
+%                     m & {A_1} & {i + k} & {;} & i & {A'} & j & \otimes & k & k & k & {;} & {j + k} & {A_2} & n \\
+%                     &&&&& {+} \\
+%                     &&&&& \vdots \\
+%                     &&&&& {+} \\
+%                     &&&& i & {X_w} & j
+%                     \arrow[from=5-1, to=5-2]
+%                     \arrow[from=5-3, to=5-2]
+%                     \arrow[from=1-5, to=1-6]
+%                     \arrow[from=1-7, to=1-6]
+%                     \arrow[from=5-5, to=5-6]
+%                     \arrow[from=5-7, to=5-6]
+%                     \arrow[from=9-7, to=9-6]
+%                     \arrow[from=9-5, to=9-6]
+%                     \arrow[from=5-9, to=5-10]
+%                     \arrow[from=5-11, to=5-10]
+%                     \arrow[from=5-15, to=5-14]
+%                     \arrow[from=5-13, to=5-14]
+%                 \end{tikzcd}}
+%                 \]
+%                 and there is a convex match of $L$ in $A'$.
+%                 It is then also down-closed and conflict reflecting as $A'$ is not hierarchical.
+%                 Need to show that arising pushout complement is a boundary complement.
+%                 \question{A problem is that the above terms does not decompose into a form of $A';C$ where $C$ is a pushout complement.
+%                 This happens because we take composition along the external interfaces and this $A'$ is within a box, so it is not a part of an external interface.
+%                 }
+%                 \Aleksei{discuss this. We can simplify PROP+ rewriting to just a box, but still have a problem with context C}
+%         \end{itemize}
+%         \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.
+%             But we only have this result for $\llangle \cdot \rrangle$ for terms with bent wires which is a functor to $PROP + Frob$.
+%         }
+%     \end{proof}
+% \end{theorem}
+
+We will next consider functors
+\[
+    \textsf{PROP}(\Sigma, \mathcal{E}) \xrightarrow{\mathcal{J}} \textsf{PROP}^{+}(\Sigma, \mathcal{E}) \xrightarrow{\llangle - \rrangle} \sfrac{\WellTypedMdaEcospans}{\mathcal{S},\mathcal{E}}
+\]
+where $\mathcal{E}$ is a set of equations of the corresponding SMT.
+\begin{theorem}[$\llangle - \rrangle$ is a functor]
+\end{theorem}
+
+\begin{theorem}[$\llangle - \rrangle$ is faithful]
+\end{theorem}
+
+\update{Or we can define rewriting for string diagrams in $\textsf{PROP}^{+}(\Sigma)$ based on rewriting in $\textsf{PROP}(\Sigma)$ and then say that $f \xrightarrow{}_{\mathcal{R}} g$ iff $[f] [g]$}
+
+\begin{definition}[Rewriting in $\textsf{PROP}^{+}(\Sigma)$]
+
+Recall the definition of rewriting in some $\textsf{PROP}(\Sigma) = \mathbb{A}$.
+We say that $a \in \mathbb{A}$ rewrites into $b \in \mathbb{A}$ under a rewrite rule $\langle l, r \rangle$ where $l,r \in \mathbb{A}$ if $a$ and $b$ are decomposable as below.
+\[
+\scalebox{0.75}{
+\tikzfig{combinatorial_semantics/prop_rewriting}
+}
+\]
+and we will use notation $a \Rightarrow_{\langle l, r \rangle} b$.
+
+Now consider an enriched prop $\mathbb{A}^{+}$ over the signature $\Sigma$ such that there is a functor $\mathcal{J} : \mathbb{A} \to \mathbb{A}^{+}$.
+We than say that $a \in \mathbb{A}^{+}$ rewrites into $b \in \mathbb{A}^{+}$ under a rewrite rule $\langle l, r \rangle$ where $l,r \in \mathcal{J}(\mathbb{A})$ if $a$ and $b$ are decomposable as below.
+
+\[
+\scalebox{0.75}{
+\tikzfig{combinatorial_semantics/prop_plus_rewriting}
+}
+\]
+
+and $a' \Rightarrow_{\langle l, r \rangle} b'$.
+We will use notation $a \Rightarrow_{\langle l, r \rangle}^{+} b$.
+\end{definition}
+
 \begin{theorem}
-
-    Let $\mathcal{R}$ be a rewriting system on $\textsf{PROP}(\Sigma)$.
-    $f \Rightarrow{}_{\mathcal{R}^{+}}$ iff $[f] \Rrightarrow{}_{[\mathcal{R}^{+}]} [g]$.
-    \begin{proof}
-        \update{It is tempting to reuse a similar result for $\llbracket \cdot \rrbracket$ because according to our definition $f \Rrightarrow_{\mathcal{R}^{+}}$ means that there is an SMC piece $f'$ inside some box that rewrites to $g'$.
-            But we only have this result for $\llangle \cdot \rrangle$ for terms with bent wires which is a functor to $PROP + Frob$.
-        }
-    \end{proof}
+    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$.
 \end{theorem}

### 2023-12-20 sections/combinatorial_semantics.tex
Soundness and fullness for PROP(\Sigma)+

--- 2023-12-14 sections/combinatorial_semantics.tex
@@
 
-\update{Naturality and SMC laws are implied by $+$ being a coproduct[reference?]}.
+This construction turns this data into a symmetric monoidal category~\cite{MonoidalCoproduct}.
 
@@
 Category $\catname{Csp_{D}(Hyp_{\Sigma})}$ is a \textit{full} sub-category of $\catname{Csp(Hyp_{\Sigma})}$ which has discrete hypergraphs as objects.
-\update{Assuming that $D$ is a functor from PROP to $\catname{Hyp_{\Sigma}}$, $\catname{Csp_{D}(Hyp_{\Sigma})}$ is also a PROP and hence an SMC (Theorem 3.6~\cite{Frobenius}).}
+% \update{Assuming that $D$ is a functor from PROP to $\catname{Hyp_{\Sigma}}$, $\catname{Csp_{D}(Hyp_{\Sigma})}$ is also a PROP and hence an SMC (Theorem 3.6~\cite{Frobenius}).}
 \end{definition}
@@
     \end{itemize}
-    \update{Looks fine?}
 \end{proof}
@@
 Hence, $u_{E}([in_{1,E}(e_x)]) = u_{E}([in_{2,E}(e_y)])$.
-\update{Above $e_x$ and $e_y$ are different from $e_x$ and $e_y$ from the first proof, i.e. they are in $E_{X}$ and $E_{Y}$. This is probablty not good.}
-
-\question{Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.}
+% \update{Above $e_x$ and $e_y$ are different from $e_x$ and $e_y$ from the first proof, i.e. they are in $E_{X}$ and $E_{Y}$. This is probablty not good.}
+
+Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.
 Finally, $u$ is unique.
@@
     % With an abuse of notation, we will confuse $\mathcal{G}$ with $V_{\mathcal{G}} \cup E_{\mathcal{G}}$, when treating it as a set.    
-    \question{Shall I use $\cup$ instead of $+$?}
     $<_c \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is a child relation imposing a strict partial order on $(V_\mathcal{G} + E_\mathcal{G})$.
@@
     Next, each $x$ has at most one immediate predecessor.
-    \update{We will denote the fact that $e$ is an immediate predecessor of $v$ as $e <_c^{\mu} v$.}
+    We will denote the fact that $e$ is an immediate predecessor of $v$ as $e <_c^{\mu} v$.
     Similarly, a set of all successors is $(x] = \{x' \in (V_\mathcal{G} + E_\mathcal{G}) \; ||\; x <_{c} x'\}$ where $x'$ can be a node or an edge. 
@@
         \] when $x$ is an edge.
-          \update{We do not require such an equation when $[x) = \varnothing$}.
+        We do not require such an equation when $[x) = \varnothing$.
 
@@
     \item $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$. \update{This essentially means that $f,g$ do not map nodes to levels of different hierarchy.}
+    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$.
+    This essentially means that $f,g$ do not map nodes to levels of different hierarchy.
     \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and vice versa.
@@
               \]
-              \update{Assumptions above guarantee that the predecessor of such node does not depend on the choice of $e$ if there are several of them.}
+              Assumptions above guarantee that the predecessor of such node does not depend on the choice of $e$ if there are several of them.
     \end{enumerate}
-    \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
+    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
     Now, if $u$ is an element of $E_{X+Y}$.
@@
               \end{enumerate}
-    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for nodes until fixed point.
+    We will also build this relation in steps, first we do steps for nodes (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for nodes until fixed point.
     Essentially, this interleaving means that nodes inherit the relation from incident edges and vice versa.
-    }
     Clearly, all the cases above are disjoint.
@@
         \item $W, X, Y, Z$ are discrete and $f,g,h,k$ are monos
-        \item \update{The image of $f;g$ (respectively, $k;h$) constitutes of nodes $v$ such that $[v) = \varnothing$}
+        \item The image of $f;g$ (respectively, $k;h$) consists of nodes $v$ such that $[v) = \varnothing$
     \end{enumerate}
@@
 \]
-\update{I've put copairings above}
 % \[
@@
 
-\update{Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC if $D$ is a functor from PROP to $\catname{EHyp_{\Sigma}}$.}
+Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC.
 More precisely, below are cospans for tensor product and symmetry.
@@
 \]\label{morphism:symmetry}
-
-\question{Or shall we say that symmetry is inherited from $\catname{EHyp_{\Sigma}}$ because coproduct is commutative and associative?}
 \end{definition}
@@
     % The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
-    \update{Note that pushout for such composition always exists as the interface in the middle is descrete and nodes in the images of the middle interface are at the outermost level as well.
+    Note that pushout for such composition always exists as the interface in the middle is descrete and nodes in the images of the middle interface are at the outermost level as well.
     Furthermore, $f_{ext};f$ and $g_{ext};g$ are monos and hence the span satisfies our assumptions.
-    }
 \end{remark}
@@
 
-\update{below definition depends on whether we require the homomorphism to reflect the conflicts}
+% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
 
@@
     \[
-    \scalebox{0.75}{
+    \scalebox{0.6}{
     \tikzfig{combinatorial_semantics/f_plus_g_new}
@@
 
-\update{We can probably omit the proofs below because they are obvious. But keeping them here serves as a sanity check.}
+% \update{We can probably omit the proofs below because they are obvious. But keeping them here serves as a sanity check.}
 
@@
 
-    This category is also symmetric monoidal and has $+$ by the lemma above.
+    This category is also symmetric monoidal and has $+$.
 \end{definition}
@@
 \end{definition}
-
-\update{I have run out of possible names for a category}
 
@@
     \begin{proof}
-        \update{I will try to reinvent a similar proof from~\cite{Frobenius2} as the cited proof does not make sense to me}
         The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of nodes and hyperedges.
@@
     \end{proof}
-    \update{Can we check the proof? I haven't used all the conditions from monogamous-ness above.
-    Although, if the morphism $i+j \to L$ is mono, the complement is unique for not-monogamous hypergraphs as well.}
+    % \update{Can we check the proof? I haven't used all the conditions from monogamous-ness above.
+    % Although, if the morphism $i+j \to L$ is mono, the complement is unique for not-monogamous hypergraphs as well.}
 \end{proposition}
 
-\question{
 \begin{definition}[Convex down-closed match] 
@@
         \item $m(L)$ is a convex down-closed subgraph of $G$
-        \item \update{I've removed the requirement that consistency should be reflected because well typedness prohibits otherwise formed matches.}
+        % \item \update{I've removed the requirement that consistency should be reflected because well typedness prohibits otherwise formed matches.}
     \end{enumerate}
 \end{definition}
-}
 
@@
 
-\Aleksei{Do we need to introduce substitution? This is needed if our rewrite rules are going to have variables as labels.}
+% \Aleksei{Do we need to introduce substitution? This is needed if our rewrite rules are going to have variables as labels.}
 
@@
 
-\Aleksei{Check the proof}
-
-\Aleksei{Can we reuse anything from~\cite{Frobenius}?}
-
-Consider an example below that illustrates the introduced constraints.
-
-\update{I've switched to $[ \cdot ]$ for our functor to be able to differenciate between $\llbracket \cdot \rrbracket$, $\llangle \cdot \rrangle$}
-\begin{remark}
-    For now, we will informally denote a combinatorial interpretation of a morphism $f$ in $\textsf{PROP}^{+}(\Sigma)$ as $[ f ]$.
-\end{remark}
-
-\begin{example}
-
-A DPOI square for rewriting in $\WellTypedMdaEcospans$ is depicted in Figure~\ref{fig:refined_dpoi}.
-A projection rule of $\llbracket f + g \rrbracket \to \llbracket f \rrbracket$ is shown in the top half of the diagram.
-The image of the matching $m$ is shown in red and its image is a convex down-closed e-hypergraph.
-Pushout complement is then computed by removing the image of $m$ from the target e-hypergraph by keeping intact the nodes that have a pre-image in the carrier of the rewriting span.
+% Pushout example
+
+% Consider an example below that illustrates the introduced constraints.
+
+% \begin{remark}
+%     For now, we will informally denote a combinatorial interpretation of a morphism $f$ in $\textsf{PROP}^{+}(\Sigma)$ as $[ f ]$.
+% \end{remark}
+
+% \begin{example}
+
+% A DPOI square for rewriting in $\WellTypedMdaEcospans$ is depicted in Figure~\ref{fig:refined_dpoi}.
+% A projection rule of $\llbracket f + g \rrbracket \to \llbracket f \rrbracket$ is shown in the top half of the diagram.
+% The image of the matching $m$ is shown in red and its image is a convex down-closed e-hypergraph.
+% Pushout complement is then computed by removing the image of $m$ from the target e-hypergraph by keeping intact the nodes that have a pre-image in the carrier of the rewriting span.
 The corresponding mda-cospan is depicted at the bottom: $c_i = g_i \setminus (l_i \setminus i) = \{18, 19, 1, 2, 13, 14, 5, 6, 7, 8\} \setminus (\{1,2,5,6,7,8\} \setminus \{1,2\}) = \{18, 19, 1, 2, 13, 14, 5, 6, 7, 8\} \setminus \{5,6,7,8\} = \{18,19,1,2,13,14\}$ and $c_i + j = \{18,19, 1, 2, 13, 14, 3, 4\}$ which is the input interface for $C$.
@@
 
-\end{example}
-
-\Aleksei{This figure yeilds "Dimension too large". Need to fix it.}
-
-Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we add them as structural rewrite rules.
-Recall that the category $\catname{EHyp_{\Sigma}}$ does not have all pushouts and hence we need to explicitly show that the pushouts that we need exist.
-
-\begin{proposition}
-Pushouts for all semilattice enrichment equations exist.
-
-\begin{proof}
-    Let us give a combinatorial interpretation for an instance of the first equation (distributivity) in~\ref{fig:string-equations}.
-    It can be found in Figure~\ref{fig:distributivity_interpretation}.
-    After we identify a convex down-closed image of the left-hand side of this rule within a graph $G$ we remove the image by keeping only the outermost nodes.
-    Then we simply glue the right-hand side of the rule into the hole.
-    Since we glue along the outermost nodes this does not introduce any conflict with regard to child relation.
-    Such equations are not absorbed by e-hypergraph representation unlike SMC equations because the right-hand side contains more edges, i.e. graphs $G$ and $H$ are not isomorphic.
-    Similar argument applies to other rules as well.
-
-\end{proof}
-\end{proposition}
-
-\update{I think I haven't paid any attention to this fact previously, but it seems that in~\cite{Frobenius} the functor $\llbracket - \rrbracket$ is between $PROP_{\Sigma_{Frob}}$ and $Csp_{D}(Hyp_{\Sigma})$, i.e.
+% \end{example}
+
+% \Aleksei{This figure yeilds "Dimension too large". Need to fix it.}
+
+% This is now in soundness and completeness section
+
+% Unlike SMC equations the equations for semilattice enriched categories are not absorbed by e-hypergraph representation and hence to further build a correspondence between $\textsf{PROP}^{+}(\Sigma)$ and $\WellTypedMdaEcospans$ we will need to turn the latter into a semilattice-enriced category.
+% We already have a $+$ in $\WellTypedMdaEcospans$ so we need to require that it satisfies a particular set of equations, for example, that $h \otimes (f + g) = h \otimes f + h \otimes g$.
+% \update{How to formally state this? Because $h \otimes (f+g)$ is an SMC term from $\WellTypedMdaEcospans$ which is just an SMC so far.}
+% We will add such equations as DPOI rewrite rules and show that for each such rule the corresponding pushout complement and the pushout exist.
+
+% \begin{proposition}
+% Pushouts for all semilattice enrichment equations exist.
+
+% \begin{proof}
+%     Let us give a combinatorial interpretation for an instance of the first equation (distributivity) in~\ref{fig:string-equations}.
+%     It can be found in Figure~\ref{fig:distributivity_interpretation}.
+%     After we identify a convex down-closed image of the left-hand side of this rule within a graph $G$ we remove the image by keeping only the outermost nodes.
+%     Note that because the image of all $a$'s and $b$'s are outermost nodes, we are allowed to pick any monomorphism for our pushout complement.
+%     And for the same reason the pushout for the right square exists.
+%     Then we simply glue the right-hand side of the rule into the hole.
+%     Such equations are not absorbed by e-hypergraph representation unlike SMC equations because the right-hand side contains more edges, i.e. graphs $G$ and $H$ are not isomorphic.
+%     Similar argument applies to other rules as well.
+% \end{proof}
+% \end{proposition}
+
+% \update{How do we state the correspondence between term equations and their e-hypergraph counterpart without a functor?}
+
+% \update{I think I haven't paid any attention to this fact previously, but it seems that in~\cite{Frobenius} the functor $\llbracket - \rrbracket$ is between $PROP_{\Sigma_{Frob}}$ and $Csp_{D}(Hyp_{\Sigma})$, i.e.
 PROP is quotiented by Frobenius equations, but they are absorbed in hypergraphs domain.
@@
 we will quotient PROP$^{+}$ by semilattice equations and quotient e-hypergraphs by the interpretation of these equations as they are not absorbed.
-The quotienting on the left is always implicit?}
-
-\begin{figure}
-    \centering
-    \[
-    \scalebox{0.45}{
-    \tikzfig{combinatorial_semantics/structural_rule_pushout}}\]  
-    \caption{Distributivity rewrite rule interpretation}
-    \label{fig:distributivity_interpretation}
-\end{figure}
-
-\begin{definition}
-    We will denote with $\mathcal{S}$ the interpretation of the equations in~\ref{fig:string-equations} as rewrite rules in $\WellTypedMdaEcospans$.
-    In particular, for all $l = r$ in~\ref{fig:string-equations}, $\mathcal{S}$ will contain $\langle [l], [r]\rangle$ and $\langle [r], [l] \rangle$.
-    \update{Technically we add those for all possible instances of $l$ and $r$ in~\ref{fig:string-equations}.}
-\end{definition}
+% The quotienting on the left is always implicit?}
+
+% \begin{figure}
+%     \centering
+%     \[
+%     \scalebox{0.45}{
+%     \tikzfig{combinatorial_semantics/structural_rule_pushout}}\]  
+%     \caption{Distributivity rewrite rule interpretation}
+%     \label{fig:distributivity_interpretation}
+% \end{figure}
+
+% \begin{definition}
+%     We will denote with $\mathcal{S}$ the interpretation of the equations in~\ref{fig:string-equations} as rewrite rules in $\WellTypedMdaEcospans$.
+%     In particular, for all $l = r$ in~\ref{fig:string-equations}, $\mathcal{S}$ will contain $\langle [l], [r]\rangle$ and $\langle [r], [l] \rangle$.
+%     \update{Technically we add those for all possible instances of $l$ and $r$ in~\ref{fig:string-equations}.}
+% \end{definition}
 
@@
     It is unable to express rewriting for rules that contain an edge with no inputs and outputs on their right-hand side.
-This is because there is no way to impose both conflict and child relation of such an edge as it has no incident nodes in its context.
+    This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
     An example of this could be seen in Figure~\ref{fig:non_expressible}.

### 2023-12-20 sections/soundness_and_completeness.tex
Soundness and fullness for PROP(\Sigma)+

@@
 
-\update{Below I kind of use $\llbracket - \rrbracket$ to define itself when using $\llbracket \mathcal{E} \rrbracket$.}
 In this section, we will build a correspondence between terms in a semilattice enriched $\textsf{PROP}^{+}(\Sigma)$ and morphisms in $\WellTypedMdaEcospans$.
-Similarly to $\textsf{PROP}^{+}(\Sigma)$ where terms are defined up to semilattice enrichment and SMC equations we will say that two morphisms $G$ and $F$ in $\WellTypedMdaEcospans$ are equal when $G \Rrightarrow{}_{\mathcal{S}} F$.
-As SMC equations are absorbed by e-hypergraph representation we only quotient by $\mathcal{S}$. 
-We are now ready to build a semilattice-enriched SMC functor
+% Similarly to $\textsf{PROP}^{+}(\Sigma)$ where terms are defined up to semilattice enrichment and SMC equations we will say that two morphisms $G$ and $F$ in $\WellTypedMdaEcospans$ are equal when $G \Rrightarrow{}_{\mathcal{S}} F$.
+% As SMC equations are absorbed by e-hypergraph representation we only quotient by $\mathcal{S}$. 
+We will build a semilattice-enriched SMC functor
 
 \[
-[\cdot] : \textsf{PROP}^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}
+[-] : \textsf{PROP}^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}
 \]
-where $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is $\WellTypedMdaEcospans$ quotiented by structural rewrite rules which makes this category into a semilattice-enrhiched one.
-By freeness, it suffices to define how the functor acts on generators \update{Identities and generators from the signature?}.
-% As $\textsf{PROP}(\Sigma)$ and $\textsf{PROP}^{+}(\Sigma)$ essentially have the same set of generators we can reuse the functor $\llbracket \cdot \rrbracket: \textsf{PROP}(\Sigma) \to \MdaCospans$ from~\cite{Frobenius2} and then follow the inclusion $\mathcal{I}$ defined in~\ref{functor:I}.
-Everything else follows by defining
+where $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is $\WellTypedMdaEcospans$ quotiented by structural rewrite rules $\mathcal{S} = \{\langle [l], [r] \rangle, \langle [r],[l] \rangle\}$ for all $l = r$ in~\ref{fig:string-equations}.
+By freeness, it suffices to define how the functor acts on generators.
+For every $f : n \to m \in \Sigma$, $[f]$ is defined as below.
 
-\begin{enumerate}
-    \item $[ f;g ] = [ f ] ; [ g ]$  for $f,g \in \Sigma$;
-    \item $[ f \otimes g ] = [ f ] \otimes [ g ]$  for $f,g \in \Sigma$;
-    \item $[ f + g ] = [ f ] + [ g ]$  for $f,g \in \Sigma$;
-    \item \update{Symmetry?}
-\end{enumerate}
-
-this also makes the functor into a symmetric monoidal semilattice-enriched SMC one on the nose.
-% \update{This essentially means that $[\cdot] = \llbracket \cdot \rrbracket ; \mathcal{I}$ plus three equations above.}
-
-\update{Anything about soundness?
-I.e.
-if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$?
-Because $f = g$ is modulo semilattice-enriched SMC equations we have to show that $\llbracket f \rrbracket \Rrightarrow \llbracket g \rrbracket$.
-Doesn't it require the soundness of rewriting?
+\[
+\scalebox{0.75}{
+\tikzfig{soundness_and_completeness/generator_translation}
 }
+\]
 
@@
     \begin{proof}
-        \update{
-            By induction on the number of applied equations to turn $f$ into $g$?
-            E.g. $f = g$ in one step essentially means that $f = l$ for some $l$ in~\ref{fig:string-equations} and $g = r$ (or the other way around).
-            Then we need to show that $[f] \Rrightarrow_{\langle l, r \rangle \in \mathcal{S}} [g']$ and $[g'] = [g]$.
+        $f = g$ means that $f$ as a term can be turned into $g$ by applying SMC equations and the equations from~\ref{fig:string-equations}.
+        Since $f_1 =_{\text{SMC}} f_2$ means that $[f_1] = [f_2]$ we will only consider the transformations induced by the equations from~\ref{fig:string-equations}.
+        We will also note that every subterm of $f$ corresponds to a convex down-closed sub e-hypergraph of $[f]$.
+        Then, proving the above statement boils down to showing that for every rewrite rule in $\mathcal{S}$ the corresponding pushout complement and pushout exist.
+        Suppose, $f =_{r} g$ where $r$ is the top left equation from~\ref{fig:string-equations}.
+        $[r]$ is then given below.
+        \[
+            \scalebox{0.45}{
+                \tikzfig{combinatorial_semantics/structural_rule_pushout}
             }
-        \question{Can we use normal form here as well? $f = g$ means there exists $h$ such that $f =_{SMC^{+}} h$ and $h =_{SMC^{+}}$ and $h$ is of the form $\phi_1 + \ldots + \phi_n$?
-                  Then we need to show that $[f] \Rrightarrow_{\mathcal{S}} [h]$ and $[h] \Rrightarrow_{\mathcal{S}} [g]$.
-        }
+        \]
+        After we identify a convex down-closed image of the left-hand side of this rule within a graph $[f]$ we remove the image by keeping only the outermost nodes.
+        Note that because the image of all $a$'s and $b$'s are outermost nodes, we are allowed to pick any monomorphism for our pushout complement.
+        And for the same reason the pushout for the right square exists.
+        Then we simply glue the right-hand side of the rule into the hole.
+        Similar argument applies to other rules as well.
+        The result for any number of applied equations follows by induction.
     \end{proof}
@@
 \begin{corollary}
-    $[\cdot] : PROP^{+}(\Sigma) \to \sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is sound in a sense that
-    if $f = g$ in $PROP^{+}(\Sigma)$ then $[f] = [g]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$
+    $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ is a semilattice-enriched SMC category and $[-]$ is a semilattice-enriched SMC functor.
 \end{corollary}
@@
     \begin{proof}
-        Consider the commutative diagram below \update{Is it obvious that the diagram commutes?}.
+        Consider the commutative diagram below.
 
@@
             \arrow["{\llbracket \cdot \rrbracket}"', from=1-1, to=3-1]
-            \arrow["{[\cdot]}", from=1-3, to=3-3]
+            \arrow["{[-]}", from=1-3, to=3-3]
             \arrow["{\mathcal{I}}"', from=3-1, to=3-3]
@@
 
-\begin{theorem}[$\llbracket - \rrbracket$ is faithful]
+\begin{theorem}
+$[-]$ is faithful.
 \begin{proof}
-Essentially we need to show that if $\llbracket f \rrbracket = \llbracket g \rrbracket$ then $f = g$ which is equivalent to if $f \not = g$ then $\llbracket f \rrbracket \not = \llbracket g \rrbracket$.
+Essentially we need to show that if $[f] = [g]$ then $f = g$ which is equivalent to if $f \not = g$ then $[f] \not = [g]$.
 First, we note that we have a \question{weak (modulo commutativity)} normal form for any term \update{string diagram?} in $PROP^{+}(\Sigma)$ which is achieved by applying distributivity for tensor and composition as well as unpacking equations, that is, every term is representable as in Figure~\ref{fig:normal_form_string} where $\phi_1,\;\ldots\;,\phi_n$ are pure SMC terms.
-\update{It is not actaully a `normal' form because we can still apply SMC equations, so it should be normal modulo SMC and commutativity}.
 
@@
 
+\end{proof}
+\end{theorem}
+
+\begin{theorem}
+$[-]$ is full.
+\begin{proof}
+As we quotient by $\mathcal{S}$ instead of arbitrary cospans we can only consider their normal forms $A_1 + \ldots A_n$ where all of $A_i$ do not contain any hierarchical edges.
+Because each of $A_i$ is a plain hypergraph, they have a pre-image in $\textsf{PROP}(\Sigma)$, i.e. $A_i = \llbracket \cdot \rrbracket;\mathcal{I}(a_i)$.
+And then $[\mathcal{J}(a_1) + \ldots + \mathcal{J}(a_n)] = [\mathcal{J}(a_1)] + \ldots + [\mathcal{J}(a_n)] = A_1 + \ldots + A_n$ and hence $\mathcal{J}(a_1) + \ldots + \mathcal{J}(a_n)$ is the pre-image of $A_1 + \ldots + A_n$.
 \end{proof}
@@
 
-We will next consider functors
-\[
-    \textsf{PROP}(\Sigma, \mathcal{E}) \xrightarrow{\mathcal{J}} \textsf{PROP}^{+}(\Sigma, \mathcal{E}) \xrightarrow{\llangle - \rrangle} \sfrac{\WellTypedMdaEcospans}{\mathcal{S},\mathcal{E}}
-\]
-where $\mathcal{E}$ is a set of equations of the corresponding SMT.
-\begin{theorem}[$\llangle - \rrangle$ is a functor]
-\end{theorem}
+% We will next consider functors
+% \[
+%     \textsf{PROP}(\Sigma, \mathcal{E}) \xrightarrow{\mathcal{J}} \textsf{PROP}^{+}(\Sigma, \mathcal{E}) \xrightarrow{\llangle - \rrangle} \sfrac{\WellTypedMdaEcospans}{\mathcal{S},\mathcal{E}}
+% \]
+% where $\mathcal{E}$ is a set of equations of the corresponding SMT.
+% \begin{theorem}[$\llangle - \rrangle$ is a functor]
+% \end{theorem}
 
-\begin{theorem}[$\llangle - \rrangle$ is faithful]
-\end{theorem}
+% \begin{theorem}[$\llangle - \rrangle$ is faithful]
+% \end{theorem}
 
-\update{Or we can define rewriting for string diagrams in $\textsf{PROP}^{+}(\Sigma)$ based on rewriting in $\textsf{PROP}(\Sigma)$ and then say that $f \xrightarrow{}_{\mathcal{R}} g$ iff $[f] [g]$}
+% \update{Or we can define rewriting for string diagrams in $\textsf{PROP}^{+}(\Sigma)$ based on rewriting in $\textsf{PROP}(\Sigma)$ and then say that $f \xrightarrow{}_{\mathcal{R}} g$ iff $[f] [g]$}
 
@@
 \begin{theorem}
-    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$.
+    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
 \end{theorem}

### 2023-12-21 sections/combinatorial_semantics.tex
Soundness and completeness for rewriting

@@
 
-\Aleksei{Actually, in~\cite{Frobenius} a hypergraph is defined as an object of the corresponding category and labelled hypergraphs are defined as a slice category.
+% \Aleksei{Actually, in~\cite{Frobenius} a hypergraph is defined as an object of the corresponding category and labelled hypergraphs are defined as a slice category.
 This is needed to show adhesiveness.
@@
 
-\update{I will build below $\Ecospans$ (E for extended) instead of $ECsp(\mathcal{C})$ for arbitrary $\mathcal{C}$ with colimits to avoid talking about pushouts in general below, because we use set difference.}
+% \update{I will build below $\Ecospans$ (E for extended) instead of $ECsp(\mathcal{C})$ for arbitrary $\mathcal{C}$ with colimits to avoid talking about pushouts in general below, because we use set difference.}
+
 \begin{definition}[Category $\Ecospans$]
@@
 
-\textbf{Update: 4 Oct}
-
-\begin{theorem}[Theorem 35~\cite{Frobenius2}]
-\label{theorem:rewriting_no_frob}
-    Let $R$ be any rewriting system on $\textsf{PROP}(\Sigma)$. Then,
-    \[
-    g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llangle ^{\ulcorner}g^{\urcorner} \rrangle \Rrightarrow_{\llangle \mathcal{R} \rrangle} \llangle ^{\ulcorner}h^{\urcorner} \rrangle 
-    \]
-
-    where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + Frob \to \catname{Csp_{D}(Hyp_{\Sigma})}$ is an isomorphism of props.
-    \update{
-        Last time I confused functors $\llangle \cdot \rrangle$ and $\llbracket \cdot \rrbracket$.
-        Note that the result above is a weaker version of the same result for rewriting in $\textsf{PROP}(\Sigma) + Frob$.
-        $Frob$ is an SMT with the signature $\Sigma_{Frob}$ which constitutes of \textit{only} Frobenius generators and terms are quotiented by both SMC laws and Frobenius equations.
-        So, essentially, $Frob = \textsf{PROP}(\Sigma_{Frob}, \mathcal{E}_{Frob})$. 
-        \question{Is it true that by quotienting $\textsf{PROP}(\Sigma)$ by SMC laws and quotienting $Frob$ by SMC and Frob laws we automatically quotient $\textsf{PROP}(\Sigma) + Frob$ by SMC and Frob laws?}
-    }
-\end{theorem}
-
-\begin{definition}[\;$\ulcorner \cdot \urcorner$]
-
-    This is a syntactic operation which is defined for terms in $\textsf{PROP}(\Sigma) + Frob$ and is given below
-    \[
-        \tikzfig{combinatorial_semantics/bending_wires}
-    \]
-
-\end{definition}
-
-\begin{remark}
-    We need the operation above because we defined the rewriting for cospans of hypergraphs only for cospans of the form $0 \to G \xleftarrow{} n + m$ (for $\catname{Csp_{D}(Hyp_{\Sigma})},\; \MdaCospans,\; \MdaEcospans$), but the functors take arbitrary terms (string diagrams) as input.
-    This operation turns an arbitrary term into a term with no inputs.
-    Now the left-hand and right-hand sides are equivalent in the sense that $f \Rightarrow_{\langle l, r \rangle} g$ if and only if $\ulcorner f \urcorner \Rrightarrow_{\langle \ulcorner l \urcorner, \ulcorner r \urcorner \rangle}
- \ulcorner g \urcorner$ and the latter string diagram can be interpreted as $0 \xrightarrow{} G \xleftarrow{} n + m$ for which we defined rewriting above.
-
-\end{remark}
-
-% \begin{proposition}
-%     $m \xrightarrow{} \mathcal{G} \xleftarrow{} n$ is mda-cospan in $\MdaCospans$ iff $0 \xrightarrow{} \mathcal{G} \xleftarrow{} n + m$ is an mda-cospan.
-% \end{proposition}
-
-% In particular, the above proposition means that by fullness of $\llbracket \cdot \rrbracket$ there exists a corresponding SMC morphism for some $0 \xrightarrow{} \mathcal{G}\xleftarrow{} n + m$.
-Theorem above essentially makes two routes below. In forward direction:
-
-\[
-\textsf{PROP}(\Sigma) \xrightarrow{i_1} \textsf{PROP}(\Sigma) + Frob \xrightarrow{\ulcorner \cdot \urcorner} \textsf{PROP}(\Sigma) + Frob \xrightarrow{\llangle \cdot \rrangle} \catname{Csp_{D}(Hyp_{\Sigma})}
-\]
-
-The proof of the theorem above in the forward direction then relies on the proof for rewriting in $\textsf{PROP}(\Sigma) + Frob$ and functoriality of $\llangle \cdot \rrangle$ and checking that a match is convex.
-
-In backwards direction
-
-\[
-  \catname{Csp_{D}(Hyp_{\Sigma})} \xrightarrow{\llbracket \cdot \rrbracket}^{-1} \textsf{PROP}(\Sigma)  
-\]
-
-relies on decomposition of a hypergraph given a convex match and on fullness of
-
-\[
-    \llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \MdaCospans
-\]
-
-\question{What options we have?
-\begin{enumerate}
-    \item Define $PROP(\Sigma)^{+}$ and $PROP(\Sigma)^{+} + Frob$
-    \item Say that $n \xrightarrow{} \mathcal{G} \xleftarrow{} m$ rewrites into $n \xrightarrow{} \mathcal{F} \xleftarrow{} m$ iff $0 \xrightarrow{} \mathcal{G} \xrightarrow{} n + m$ rewrites into $0 \xrightarrow{} \mathcal{F} \xrightarrow{} n + m$.
-          This gives the following route
-          \[
-            \textsf{PROP}(\Sigma)^{+} \xrightarrow{\llbracket \cdot \rrbracket} \MdaEcospans \xrightarrow{\ulcorner \cdot \urcorner} \MdaEcospans
-          \]
-
-          and the same in the other direction with flipped arrows.
-\end{enumerate}
-}
-\Aleksei{TODO: discuss this}
-
-If we quotient $\MdaCospans$ by $\llangle \mathcal{R} \rrangle$ by the theorem~\ref{theorem:rewriting_no_frob} and faithfulness and fullness of $\llbracket \cdot \rrbracket$ we get the following result.
-$f = g$ in $\textsf{PROP}(\Sigma,\mathcal{E})$ iff $\llangle ^{\ulcorner} f ^{\urcorner} \rrangle = \llangle ^{\ulcorner} f ^{\urcorner} \rrangle$ in $\MdaCospans$.
-
-\question{Do we want the same result for $\textsf{PROP}(\Sigma, \mathcal{E})^{+}$ and $\WellTypedMdaEcospans$?}
-
-\textbf{End of update}
+% \textbf{Update: 4 Oct}
+
+% \begin{theorem}[Theorem 35~\cite{Frobenius2}]
+% \label{theorem:rewriting_no_frob}
+%     Let $R$ be any rewriting system on $\textsf{PROP}(\Sigma)$. Then,
+%     \[
+%     g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llangle ^{\ulcorner}g^{\urcorner} \rrangle \Rrightarrow_{\llangle \mathcal{R} \rrangle} \llangle ^{\ulcorner}h^{\urcorner} \rrangle 
+%     \]
+
+%     where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + Frob \to \catname{Csp_{D}(Hyp_{\Sigma})}$ is an isomorphism of props.
+%     \update{
+%         Last time I confused functors $\llangle \cdot \rrangle$ and $\llbracket \cdot \rrbracket$.
+%         Note that the result above is a weaker version of the same result for rewriting in $\textsf{PROP}(\Sigma) + Frob$.
+%         $Frob$ is an SMT with the signature $\Sigma_{Frob}$ which constitutes of \textit{only} Frobenius generators and terms are quotiented by both SMC laws and Frobenius equations.
+%         So, essentially, $Frob = \textsf{PROP}(\Sigma_{Frob}, \mathcal{E}_{Frob})$. 
+%         \question{Is it true that by quotienting $\textsf{PROP}(\Sigma)$ by SMC laws and quotienting $Frob$ by SMC and Frob laws we automatically quotient $\textsf{PROP}(\Sigma) + Frob$ by SMC and Frob laws?}
+%     }
+% \end{theorem}
+
+% \begin{definition}[\;$\ulcorner \cdot \urcorner$]
+
+%     This is a syntactic operation which is defined for terms in $\textsf{PROP}(\Sigma) + Frob$ and is given below
+%     \[
+%         \tikzfig{combinatorial_semantics/bending_wires}
+%     \]
+
+% \end{definition}
+
+% \begin{remark}
+%     We need the operation above because we defined the rewriting for cospans of hypergraphs only for cospans of the form $0 \to G \xleftarrow{} n + m$ (for $\catname{Csp_{D}(Hyp_{\Sigma})},\; \MdaCospans,\; \MdaEcospans$), but the functors take arbitrary terms (string diagrams) as input.
+%     This operation turns an arbitrary term into a term with no inputs.
+%     Now the left-hand and right-hand sides are equivalent in the sense that $f \Rightarrow_{\langle l, r \rangle} g$ if and only if $\ulcorner f \urcorner \Rrightarrow_{\langle \ulcorner l \urcorner, \ulcorner r \urcorner \rangle}
+%  \ulcorner g \urcorner$ and the latter string diagram can be interpreted as $0 \xrightarrow{} G \xleftarrow{} n + m$ for which we defined rewriting above.
+
+% \end{remark}
+
+% Theorem above essentially makes two routes below. In forward direction:
+
+% \[
+% \textsf{PROP}(\Sigma) \xrightarrow{i_1} \textsf{PROP}(\Sigma) + Frob \xrightarrow{\ulcorner \cdot \urcorner} \textsf{PROP}(\Sigma) + Frob \xrightarrow{\llangle \cdot \rrangle} \catname{Csp_{D}(Hyp_{\Sigma})}
+% \]
+
+% The proof of the theorem above in the forward direction then relies on the proof for rewriting in $\textsf{PROP}(\Sigma) + Frob$ and functoriality of $\llangle \cdot \rrangle$ and checking that a match is convex.
+
+% In backwards direction
+
+% \[
+%   \catname{Csp_{D}(Hyp_{\Sigma})} \xrightarrow{\llbracket \cdot \rrbracket}^{-1} \textsf{PROP}(\Sigma)  
+% \]
+
+% relies on decomposition of a hypergraph given a convex match and on fullness of
+
+% \[
+%     \llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \MdaCospans
+% \]
+
+% \question{What options we have?
+% \begin{enumerate}
+%     \item Define $PROP(\Sigma)^{+}$ and $PROP(\Sigma)^{+} + Frob$
+%     \item Say that $n \xrightarrow{} \mathcal{G} \xleftarrow{} m$ rewrites into $n \xrightarrow{} \mathcal{F} \xleftarrow{} m$ iff $0 \xrightarrow{} \mathcal{G} \xrightarrow{} n + m$ rewrites into $0 \xrightarrow{} \mathcal{F} \xrightarrow{} n + m$.
+%           This gives the following route
+%           \[
+%             \textsf{PROP}(\Sigma)^{+} \xrightarrow{\llbracket \cdot \rrbracket} \MdaEcospans \xrightarrow{\ulcorner \cdot \urcorner} \MdaEcospans
+%           \]
+
+%           and the same in the other direction with flipped arrows.
+% \end{enumerate}
+% }
+% \Aleksei{TODO: discuss this}
+
+% If we quotient $\MdaCospans$ by $\llangle \mathcal{R} \rrangle$ by the theorem~\ref{theorem:rewriting_no_frob} and faithfulness and fullness of $\llbracket \cdot \rrbracket$ we get the following result.
+% $f = g$ in $\textsf{PROP}(\Sigma,\mathcal{E})$ iff $\llangle ^{\ulcorner} f ^{\urcorner} \rrangle = \llangle ^{\ulcorner} f ^{\urcorner} \rrangle$ in $\MdaCospans$.
+
+% \question{Do we want the same result for $\textsf{PROP}(\Sigma, \mathcal{E})^{+}$ and $\WellTypedMdaEcospans$?}
+
+% \textbf{End of update}
 
@@
     Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n_{ext} + m_{ext}$ 
-    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $l_i + l_o \xrightarrow{} L \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} R \xleftarrow{} r_i + r_o$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
+    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}}^{+} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $L \xleftarrow{} l_i + l_o \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} r_i + r_o \xrightarrow{} R$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
 
@@
     \begin{enumerate}
+        \label{dpoi:assumptions}
         \item $m : L \to G$ is a convex down-closed match;
+        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $x_{ext} + y_{ext}$
+        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $x_{ext} + y_{ext}$
         \item $x_{ext} + y_{ext} \to C \to G$ is a boundary complement in the leftmost pushout.
@@
 \end{definition}
+
+\begin{remark}
+The conditions 2 and 3 are needed to guarantee the existence of the pushout complement.
+\end{remark}
 

### 2023-12-30 main.tex
remove e-rewriting
Refined parent and consistency relations WIP
Pushout for EHyp with consistency and proofs
TODOs added
introduction revision

--- 2023-10-01 main.tex
@@
 % quiver is for quiver figures
+% including quiver redefines \Bbbk which leads to an error
 % \usepackage{quiver}
+
 \usepackage{tikz}
@@
 
+% conflict
+\newcommand{\hashtag}{{\#}}
+\newcommand{\consistency}{{\smile}}
 % There is already a remark in the preamble, but it does
@@
 
-\title{Equivalence Hypergraphs:\\ Semantics of E-Graphs and their Generalization to 
-Monoidal Theories}
-\author{Chris Barrett, Dan Ghica, Aleksei Tiurin}
+\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
+\author{Chris Barrett}
+\author{Dan R. Ghica}
+\author{Aleksei Tiurin}
 
@@
 \begin{abstract}
-Abstract.
+The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
+We give a categorical semantics for \emph{e-graphs} over algebraic theories, a generalisation of the structures which are the subject of this technique, and to their rewriting. 
+The proposed framework consists of Cartesian categories enriched over a semilattice,  further generalised to symmetric monoidal categories, to account for monoidal theories.
+The aim is to permit the optimisation of a larger class of programs, based on string diagrams.  
+Furthermore, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs,  which we call \emph{e-hypergraphs}, which have the usual advantage that many structural equations are absorbed into a general notion of isomorphism.
 \end{abstract}
@@
 \input{sections/soundness_and_completeness}
-\input{sections/e-graph-translation}
 \input{sections/conclusion}

### 2024-01-01 E-Hypergraphs-LICS24.tex
LICS24-WIP

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%

\documentclass[sigconf]{acmart}

\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage[
  colorinlistoftodos,
  prependcaption,
  % disable,
]{todonotes}
\usepackage{fontenc}

%\usepackage{minted}

\usepackage{subcaption}

\usepackage{hyperref}

\input{preamble}
\input{macros}
\input{sample.tikzstyles}
\input{hypergraph.tikzstyles}
\input{hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

% \usepackage{mathtools,mathrsfs}
\usepackage{tikz-cd}
\usepackage{tikzit}

% for quotienting
% \usepackage{faktor}
\usepackage{xfrac}

\usepackage{import}

% category names
%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}

\newcommand\HypI[1]{\textbf{HypI}(#1)}
\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{WMEHypI(\Sigma)}}}
% conflict
\newcommand{\hashtag}{{\#}}
\newcommand{\consistency}{{\smile}}
% There is already a remark in the preamble, but it does
% not work for some reason
\newtheorem{remark}{Remark}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
  2024}{Tallinn, Estonia}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\begin{document}

\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}

\author{Anonymous}
%\email{}
%\orcid{}
%\affiliation{
%  \institution{}
%  \streetaddress{}
%  \city{}
%  \state{}
%  \country{
%  \postcode{}
%}

%\renewcommand{\shortauthors}{}

\begin{abstract}
The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
We give a categorical semantics for \emph{e-graphs} over algebraic theories, a generalisation of the structures which are the subject of this technique, and to their rewriting. 
The proposed framework consists of Cartesian categories enriched over a semilattice,  further generalised to symmetric monoidal categories, to account for monoidal theories.
The aim is to permit the optimisation of a larger class of programs, based on string diagrams.  
Furthermore, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs,  which we call \emph{e-hypergraphs}, which have the usual advantage that many structural equations are absorbed into a general notion of isomorphism. 
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10010124.10010131.10010137</concept_id>
       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003809.10010031</concept_id>
       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10002950.10003624.10003633.10003637</concept_id>
       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10003798</concept_id>
       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[300]{Theory of computation~Categorical semantics}
\ccsdesc[300]{Theory of computation~Data structures design and analysis}
\ccsdesc[300]{Mathematics of computing~Hypergraphs}
\ccsdesc[300]{Theory of computation~Equational logic and rewriting}

\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

\newcommand\id{\textsf{id}}
\newcommand\sym{\textsf{sym}}

\maketitle

%-------------------- INTRODUCTION

\section{Introduction}

Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.  This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites. Typical approaches to this problem use heuristics to determine an ordering.

A recently proposed,  alternative solution is \textit{equality saturation} 
\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.  Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.

Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories. 
The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
The starting innovation of our approach is the semilattice-enrichment. 
The \textit{join} operator lifts the formalism so that it can now operate with \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial represetation.

Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques. 
This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.

A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.  We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework 
\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.  As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.

\subsection{E-graphs}

\begin{figure*}\label{fig:e-graph-example}
\[
    \scalebox{0.65}{
    \tikzfig{categorical-semantics/egraph-translation}
    }
\]
\caption{Example translation of e-graphs into string diagrams for semilattice enriched (traced) Cartesian categories. DAN: (a) SHOULD USE MATH FONT.}
\end{figure*}

E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.

Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.

Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
Initially each equivalence class has a single node. 
The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$. 
This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself. 
Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.

The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.

\subsection{Semantics via Semilattice Enriched Categories}

\begin{figure}\label{fig:egraph-strings}
\[
	\scalebox{0.65}{
	\tikzfig{categorical-semantics/egraph-strings}
	}
\]
\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
\end{figure}

Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{noauthor_09083347_nodate,joyal_geometry_1991, mellies_functorial_2006}. In Figure~\ref{fig:egraph-strings} we extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator  
\[
\infer{\phi_1 + \ldots + \phi_n: A \to B}{\{\phi_i: A \to B\}_{i \in \{1,\ldots,n\}}}
\]
The ability to take formal sums, which in concrete categories will usually be set union, is used to model the equivalence class structure of e-graphs.

DAN: TOP TO BOTTOM OR BOTTOM TO TOP? UNCLEAR.\\
String diagrams are read from top-to-bottom; we consider additional generators for the duplication and deletion transformations of a Cartesian category.  Note, further, that in the Cartesian case we can restrict to generators $c$ with a single output.\\
DAN: WHY? WITH OR WITHOUT LOSS OF GENERALITY? WHAT DOES THIS MEAN?

The translation from e-graphs to enriched string diagrams is illustrated informally below,
noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
\[
    \tikzfig{categorical-semantics/egraph-translation-1}
\]
DAN: IT WOULD BE CLEARER IMHO IF THE ARGUMENTS OF C1 AND C2 REMAIN CENTERED AND THE DISCARDED NODES ARE OFF CENTER.

Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}. In particular, note that the translation of the \textit{cyclic} e-graph (e)
requires the use of a categorical \textit{trace}, generating a cycle. In this paper, we focus on \textit{acyclic} e-graphs and their corresponding catgeories; but the above example illustrates that the extension to the cyclic case seems to be routine. As it would involve needless additional presentational complexity we leave it as further work.

Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal sums. \\
DAN: SUMS OR JOINS? CONSISTENT?\\
Thus we generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we outline in the conclusion of this paper.   
\\ DAN: THIS IS A BIT UNCLEAR. WHY IS THIS THE NATURAL LEVEL THE GENERALITY? IT'S NOT QUITE OBVIOUS.

\subsection{Combinatorial Representation of Enriched String Diagrams}

In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.  String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC. For example, we have the following equivalences between diagrams. 
%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/interchange}
	}
\]
This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.  Here, the generators $c_i$ become vertices, connected via hyper-edges.  Thus the expected quotient becomes simply hypergraph isomorphism. With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some $c_i$.

We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(e) of Figure \ref{fig:e-graph-example}. These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.  Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ in any hypergraph $G$ and replace it with (a sub-hypergraph corresponding to) $r$.
The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{???} is still applicable.
However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.

\subsection{Soundness and Completeness}
The main technical result of the paper is a proof soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs. While the structural equalities of SMCs are factored out in the representation,  important structural equalities arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be. They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.  Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).

Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched SMCs.  That is, we exhibit a translation functor $\llbracket-\rrbracket$ from the free semilattice enriched SMC over a symmetric monoidal theory to our (quotiented) category of e-hypergraphs such that 
morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the monoidal theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow_{DPO} \llbracket g \rrbracket$.

DAN: THIS NEEDS EXPANDING. EXPLAINING IN PLAIN WORDS SOME OF THE THEORETICAL PROPERTIES OF E-GRAPHS AND THE CATEGORICAL STATEMENT.

%\subsection{E-matching and E-rewriting}
%
%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph. In general, this involves unsharing the e-box structure before a redex becomes available.  Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure. In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$. Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.  
%\begin{figure}
%\[
%	\tikzfig{combinatorial_semantics/example-rewriting}
%\]
%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.  REDO THIS}
%\end{figure}

\subsection{Related Work}

Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.  This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.  We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.  Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.

Our use of string diagrams is also central to the intuition behind our work.  The string diagrams we work with are left essentially as informal reasoning devices, however there exists a breadth of work on similar diagrammatic formalisms for similar classes of categories.  We mention here work on hierarchical string diagrams \cite{ghica_hierarchical_2023},  functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.

DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.

\section{Categorical Semantics of E-Graphs}
CHRIS: INTRO TO SECTION: CLARIFY THAT WE GIVE E-GRAPHS AS CARTESIAN CASE,  BUT ALSO GENERALIZE TO MONOIDAL CASE.

First, we fix the following notation.  Given a catgeory $\mathbb{C}$ and object $A,B \in \mathbb{C}$ ,  we denote $\mathbb{C}(A,B)$ the corresponding hom-set.  We write the identity morphism on $A$ as $\id_A$,  or simply $A$.  We commonly write $f;g$ for composition in diagram order.  Composition in the usual order is written $g \circ f$.  We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $1$ and its symmetry natural transformation as $\sym$ \cite{maclane}.  We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.  We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.  We denote by $\textsf{SLatt}$ the category of semilattices.

\subsection{Symmetric Monoidal Theories and PROPs. }

A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations} -- formally,  pairs of terms freely generated over the signature.  We are interested here in symmetric monoidal theories, which are generalizations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.  First,  we generalize the notion of a signature. 
\begin{definition}[Monoidal signature]
A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $f: n \to m$,  with \textit{arity} $n$ and \textit{coarity} $m$.  %A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements. 
\end{definition}

A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
\begin{definition}[Products and permutations category]
A \textit{permutations and products category (PROP)} is a symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$. In this case, we will enote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$. A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects. 
We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ is $\textsf{PROP}(\Sigma)$.  
\end{definition}
The free category $\textsf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.

We can now specify an equation associated with a monoidal signature as a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
\begin{definition}[Symmetric Monoidal Theory]
A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textsf{PROP}(\Sigma)$ 
A \textit{symmetric monoidal theory} $\textsf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textsf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
\end{definition}

It will often be convenient to give presentations of SMTs in terms of string diagrams.
\begin{example}
The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
\[
	\scalebox{0.6}{
  	 \tikzfig{categorical-semantics/Cartesian-equipment}
	}
\]
together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-theory}	
	}
\]
The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-naturality}
	}
\]
By Fox's Theoem [cite],  $\textsf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
\end{example}

\subsection{Free Semilattice Enrichment: Formal Joins}

We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$. \footnote{Note, in particular,  that semilattices do not require a unit for $+$. This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct -- a degeneracy we wish to avoid.} More precisely our hom-sets will be given the structure of a semilattice.  
Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
\[
	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
\]
is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.  Technically, this amounts to asking that the tensor product is an enriched functor. 
We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{SLatt}$.

\begin{definition}[Semilattice-enriched PROP]
A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms. 
\begin{align*}
f ; (g+h) &= f;g + f;h &
(f+g) ; h &= f;h + g;h \\
f \otimes (g+h) &= f \otimes g + f \otimes h & 
(f+g) \otimes h &= f \otimes h + g \otimes h,
%\item A set $ob(\mathbb{C})$ of objects;
%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
\end{align*}
Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.  We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textsf{PROP}^+(\Sigma)$ ($\textsf{SMT}^+(\Sigma, \mathcal{E})$,  respectively). 
\end{definition}

The free category $\textsf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.

To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.  It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.

Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. Indeed, the diagrammatic language itself is not truly novel and is similar to \textit{e.g.}, \cite{duncan_generalised_2009}.

\begin{figure*}
\[  
    \scalebox{0.75}{
	\tikzfig{categorical-semantics/egraph-strings-equations}
    }
\]
\caption{Equations for a  semilattice enriched symmetric monoidal category}
\label{fig:string-equations}
\end{figure*}

\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}

In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the equivalence between hypergraphs of a particular form and terms in $\textsf{PROP}(\Sigma)$ established in~\cite{Frobenius, Frobenius2},  and how one can implement string diagram rewriting using DPO rewriting on hypergraphs.   
CHRIS: INTRO HYPERGRAPHS WITH INTERFACES,  RESTRICTIONS TO SMCS.  Q: INTRO DPO REWRITING HERE TOO? NOTE,   REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ morphisms!

First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$. 
We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.

\subsection{Hypergraphs}

Hypergraphs generalize graphs by allowing edges to have multiple sources and targets. 
\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of nodes, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.  The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$. 
A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold. 
\begin{enumerate}
    \item $(s_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;s_{\mathcal{G}})(e)$
    \item $(t_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;t_{\mathcal{G}})(e)$
    \item $l_{\mathcal{F}}(e) = \phi_E;l_{\mathcal{G}}(e)$
\end{enumerate}
We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$.
\end{definition}
Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
The initial object is given by the empty hypergraph.  
CHRIS: DESCRIPTION OF PUSHOUT HERE?

When we interpret string diagrams as hypergraphs,  we will interpret morphisms as edges,  and wires by vertices.  However,  we will additionally need a way to specify which vertices are to be considered as \textit{input} and \textit{output}. 
For this purpose,  will use the following construction to equip our hypergraphs with \textit{interfaces}. 
\begin{definition}[Symmetric monoidal category of cospans]
Let $\mathbb{C}$ be a category with all finite colimits.  A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.  This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
%\[
%\begin{tikzcd}
%                                               & A \arrow[dd, "\alpha"] &                                                \\
%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
%                                               & B                      &                                               
%\end{tikzcd}
%\]
%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.  Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows: 
\begin{align*}
    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) &\otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} &+ \mathcal{G'} \xleftarrow{g+g'} B + B' ~ , 
\end{align*}
and its unit by the initial object of $\mathbb{C}$.
Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
\end{definition}

Following~\cite{Frobenius2} morphisms of $\textsf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
\begin{definition}[Hypergraphs with interfaces]
\label{def:cspd}
Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
\end{definition}

In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.  The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.   
CHRIS: NOTE THIS IS OK FOR FROBENIUS,  NOT FOR SMC.

\begin{definition}[Degree,  path,  acyclic] 
The \emph{in-degree} of a node $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in s(e)$.  Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in t(e)$.
A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some node $v \in t(e_i)$ is also in $s(e_{i+1})$.
A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
\end{definition}

In order to represent morphisms of $\textsf{PROP}(\Sigma)$,  it suffices to restrict our attention to MDA-cospans,  as defined below.
\begin{definition}[Category of MDA Hypergraphs with Interfaces \cite{Frobenius2}]
\label{def:monogamy_hyp}
We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
\begin{enumerate}
    \item $\mathcal{G}$ is a directed acyclic hypergraph;
    \item $f$ and $g$ are monomorphisms;
    \item the in-degree and out-degree of every vertex is at most $1$;
    \item vertices with in-degree $0$ are precisely the image of $f$; and
    \item vertices with out-degree $0$ are precisely the image of $g$.
\end{enumerate}
We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
\end{definition}
Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  In fact,  due to 
\cite{Frobenius2},  we have the following theorem which says the category $\MdaCospans$ in fact \textit{characterizes} $\textsf{PROP}(\Sigma)$. 
\begin{theorem}[Corollary 26~\cite{Frobenius2}]
    $\textsf{PROP}(\Sigma) \cong \MdaCospans$
\end{theorem}
We continue to show a similar theorem relating our \textit{enriched} category $\textsf{PROP}^+(\Sigma)$ to a combinatorial category of e-hypergraphs with interfaces,  introduced in the next section.

CHRIS: EXPLAIN ABOUT MONOIDAL THEORIES AND EQUATIONS.

\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
CHRIS: PUT BELOW WITH OTHER PRELIMINARIES???

In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs. 
DPO rewriting is a standard technique for formalising graph rewriting categorically.

The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $L \rightsquigarrow R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $L$ within $\mathcal{G}$,  remove it,  and then plug $R$ into the hole.  In DPO rewriting,  such a rewrite rule is formalized as a span $L \xleftarrow{} K \xrightarrow{} R$ in some category $\mathcal{C}$ with pushouts,  where $K$ is some invariant that should be maintained during rewriting.
CHRIS: "WITHIN G",  BUT WE ARE IN AN ARBITRARY CATEGORY: NOT NECESSARILY GRAPHS?

A typical DPO square looks as follows,  where the two marked squares are pushouts. 
\[
\begin{tikzcd}
L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H
\end{tikzcd}
\]
Morphisms $K \to C \to G$ are called a \textit{pushout complement} and morphism $m$ is called a \textit{match},  defined when the corresponding pushout complement exists. When $f,g$ are expected to be monomorphisms.   Constructing $C$ can be intuitively understood as finding an occurrence of $L$ in $G$ and removing from this occurrence everything that has no pre-image in $K$ and then inserting in there everything from $R$ that has no pre-image in $K$ -- hence the two pushouts: one for deletion and one for insertion.  In this case,  we say that $G$ is rewritten into $H$ via the rewrite rule $L \xleftarrow{} K \xrightarrow{} R$.

However, this notion of DPO does not account for hypergraphs with interfaces, i.e. those from $\MdaCospans$ where $l,r$ from $\langle l, r \rangle$ are cospans. To support this we first assume that we only need to maintain the interfaces of $i \xrightarrow{} L \xleftarrow{} j$ and $i \xrightarrow{} R \xleftarrow{} j$ during rewriting, then we can take $K \cong i + j$, i.e. take $K$ to be the coproduct of $i,j$ (remark 3.14~\cite{Frobenius}). What remains is to account for the interfaces in $G$, $C$, and $H$, which brings us to the following modified DPOI square. $J \cong n + m$ where $n$ and $m$ are the input and output interfaces of $G$, $C$, and $H$ respectively. The only difference from DPO square above is that we now explicitly require that the initial object $G$, intermediate object $C$ and final object $H$ all share the same interfaces.

CHRIS: DEFINE DPOI FORMALLY,  IN DEFINITION ENVIRONMENT? PERHAPS COMBINE WITH BOUNDARY COMPLEMENT BELOW,  TWO FIGURES ON ONE LINE? 
\[
\begin{tikzcd}
L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H \\
                  & J \arrow[lu] \arrow[u] \arrow[ru]          &              
\end{tikzcd}
\]

To mimic rewrites in $\textsf{PROP}(\Sigma)$ we need to put some restrictions on pushout complements as not all of them lead to mda-cospans. CHRIS: "LEAD TO"? BE PRECISE! :)

\begin{definition}[Boundary complement for $\catname{MACsp_{D}(Hyp_{\Sigma}))}$]
\label{def:boundary_original}
For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
and mono $f : L \to G$, a pushout complement as depicted below

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
\end{figure}

is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to L^{\bot}$ and $d_2 : m \to L^{\bot}$ making the above triangle commute and such that
\[
    n + j \xrightarrow{[d_1,c_2]} L^{\bot} \xleftarrow{[d_2,c_1]} m + i
\]
is a monogamous cospan. 
\end{definition}

Intuitively, such a pushout complement requires that inputs get only glued to outputs and vice-verse in the two pushout squares.

\begin{definition}[Convex subgraphs]
We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ \emph{convex} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
\end{definition}
\begin{definition}[Convex match]

We call a match $m : L \to G$ convex if it is mono and its image is a convex subgraph of $G$.

\end{definition}

\begin{definition}[DPOI Rewriting]
Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ in $\catname{HypI(\Sigma)}$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ in $\catname{Hyp_{\Sigma}}$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold

\begin{itemize}
    \item $m : L \to G$ is a convex match;
    \item $i + j \to C \to G$ is a boundary complement in the leftmost pushout.
\end{itemize}

\end{definition}

\section{E-Hypergraphs}

CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.  WHAT DOES (4) FOR CONSISTENCY MEAN?

We follow analogous steps to those used to achieve a combinatorial representation of $\textsf{PROP}(\Sigma)$ as a particular subset of plain hypergraphs with interfaces.  First we define an e-hypergraph,  which supports a hierarchical notion of "e-box" with distinct components.  Then,  we extend this notion with interfaces via a generalization of the cospan construction.  Finally,  we restrict the morphisms under consideration to those satisfying a certain directed acyclic \textit{monogamy} condition.

In the following,  we elide the obvious inclusions of $V$ and $E$ into $V+E$. 
In other words, homomorphism should preserve immediate predecessors and the conflict relation.  CHRIS: CHECK I DIDNT INTRODUCE MISTAKES ABOVE.  SHOULD WE USE $< ; \phi$ NOTATION TO MATCH OTHER HYPERGRAPH DEFINITION?

\begin{definition}[Category of E-hypergraphs]
\label{def:e-homo}    
An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}; and an equivalence relation $\consistency$ on $V + E$ called the \textit{consistency relation}.  We define the following additional notation.  For $x \in V + E$,  the set of \emph{predecessors} and \emph{successors} of $x$,  respectively: 
\[
	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
\]
We will denote the unique immediate predecessor $e$ of $x$ by writing $e <_c^{\mu} x$.
We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
We require that the child relation satisfies the following conditions
\begin{enumerate}
	\item $[x)$ contains exclusively edges such that $l(e) = \bot$.
	\item Each $x$ has at most one immediate predecessor.
	\item if $v \in s(e)$ then $e' <_c^{\mu} e$ iff $e' <_c^{\mu} v$.
\end{enumerate}
The consistency relation must satisfy the following properties. 
    \begin{enumerate}
        \label{def:consistency_properties}
        \item The relation is an equivalence relation
        \item If $v \in s(e)$ then $v \consistency^{\mu} e$
        \item If $v \in t(e)$ then $v \consistency^{\mu} e$
        \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
    \end{enumerate}

A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
\begin{enumerate}
    \item $\phi$ is a hypergraph homomorphism
    \item for $v \in V_{\mathcal{F}}$,  if $[v)$ is not empty then 
	\[
        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
        \] 
        and for $e \in E_{\mathcal{F}}$,  if $[e)$ is not empty then 
        \[
        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
        \] 
    % \item
    % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 \consistency_{\mathcal{F}}^{\mu} x_2$ implies $\phi(x_1) \consistency_{\mathcal{G}}^{\mu} \phi(x_2)$. 
%    Or, written functionally, when $x \in E_{\mathcal{F}}$
%    \[
%        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
%        \subseteq
%        \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
%    \]
%    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$%, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
%    And, when $x \in V_{\mathcal{F}}$
%    \[
%        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
%        \subseteq
%        \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
%    \]
    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \hashtag_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \hashtag_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
\end{enumerate}
If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as arrows.  
\end{definition}
Condition $(3)$ on the child relations says that $<_c^{\mu}$ respects connectivity.

The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  A concrete description of the pushout of two morphisms in this category is given in the appendix.  Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.

\begin{remark}
When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate predecessor of all $x$ whose immediate predecessor is in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
\end{remark}

When modelling our enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.  Thus,  we introduce an \textit{extended} cospan construction.  Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extened cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist. 
\begin{definition}[Category of e-hypergraphs with interfaces]
    We will build the category of \textit{e-hypergraphs with interfaces},  denoted $\Ecospans$ as follows.  Its objects are discrete e-hypergraphs,  and the set of morphisms $\Ecospans(n,m)$ are isomorphism classes of \textit{extended cospans},  which are defined as follows:  
    \[
    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
    \]
    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices.

This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.  We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.  We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.  
We will sometimes confuse $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and we will sometimes confuse $n, n', m$ and $m'$ with their images in $\mathcal{G}$. 
Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.

    Two cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
    \[
    \scalebox{0.75}{
        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
    }
    \]
Composition of two morphisms $f : X_{ext} \to Y_{ext}$ and $g : Y_{ext} \to Z_{ext}$ is computed in two stages
First, $\mathcal{F} +_{f_{ext}, g_{ext}} \mathcal{G}$ is computed,  as shown below.

% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=

\[
\scalebox{0.9}
{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
\]
Then, new feet for $\mathcal{H} = A +_{f_{ext};f,g_{ext};g} B$ are computed as below
\[
X_{ext} \xrightarrow{a_{ext};\iota_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} \mathcal{H} \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};\iota_1} Z_{ext}
\]
where
\[
    h = \langle a;p_1, g|_{Y' \setminus Y_{ext}};p_2 \rangle
\]

\[
    k = \langle b;p_2, f|_{Y \setminus Y_{ext}};p_1 \rangle
\]
% \[
% \begin{tikzcd}
% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
% Y'          & Y' \setminus Y_{ext} \arrow[l]
% \end{tikzcd}
% \]
% and then the disjoint union is just a coproduct. 
The identity of composition is given by the obvious extended cospan of identities. 
$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp_{\Sigma}}$,  analogously to Definition~\ref{def:cspd}.  
\end{definition}
CHRIS: UPDATE DIAGRAMS TO USE MY NEW NOTATION,  TAKE LESS VERTICAL SPACE.  FIND CLEANER WAY TO WRITE NEW FEET.  USE $\iota$ FOR INJECTIONS,  WHAT IS $\langle \rangle$? Include example e-hypergraph-with-interfaces vs strings here?

As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism). 
% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=

\begin{definition}[Category of MDA e-hypergraphs with interfaces]
We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
\[
n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
\]
\begin{enumerate}
        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
        \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
        \item The in-degree and out-degree of every vertex is at most 1; 
        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
        \item Vertices in the strictly internal interface are not top-level. 
        % the item below is not assumed by the definition of a cospan of e-hypergraphs
        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
\end{enumerate}
\label{def:monogamy_ehyp} 
We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of MDA-extended-cospans. 
\end{definition}
CHRIS: COULD PUT CONDITION (5) IN THE PLAIN DEFINITION OF EXTENDED COSPANS? WE ALREADY ASK FOR EXTERNAL INTERFACES TO BE TOP-LEVEL.  ALSO,  (2) IS ALREADY COVERED IN THAT DEFINITION.

\begin{definition}[Well-typedness]
Consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
We 
\end{definition}
CHRIS: CATEGORY OF SUCH COSPANS? COMBINE WITH DEFINITION ABOVE? $\WellTypedMdaEcospans$

\section{DPOI-Rewriting for E-Hypergraphs}

We have a more general notion of interface and hence we need to adjust DPOI rewriting a little bit.

We have a more general notion of interface and hence we need to adjust DPOI rewriting a little bit.
The first issue is that whole interfaces are not preserved during rewriting, consider an example in Figure~\ref{fig:example_rewrite_ecospans} which corresponds to a rewrite rule of $\langle f, f+g \rangle$.

%\begin{figure}[t]
%    \centering
%    \[
%    \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_to_f_plus_g_example}}
%    \]
%    \caption{Example rewrite rule in $\WellTypedMdaEcospans$}
%    \label{fig:example_rewrite_ecospans}
%\end{figure}

The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
Note that the outermost interfaces are preserved.

Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.

The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
The entry in the middle of the span depicts the outermost interfaces which are preserved.
The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it. 
Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.

\update{$i,j$ are replaced}

To account for such changes in the interfaces we put some additional constraints on DPOI rewriting.
Consider a DPOI diagram with extra constraints below which encodes
 the notion that cospan 
 $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$
 rewrites into a cospan 
 $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$ via a rewrite rule $\langle l, r \rangle$,
 where
 $l := x_{ext} \xrightarrow{} l_i \xrightarrow{} L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
 and
 $r := x_{ext} \xrightarrow{} r_i \xrightarrow{} L \xleftarrow{} r_o \xleftarrow{} y_{ext}$.
 Note there are arrows from $x_{ext} + y_{ext} \to l_i + l_o$ (respectively, $r_i + r_o$) from the corresponding cospans which are not shown in the diagram.
 Similarly for the arrows from $n_{ext} + m_{ext}$.

% \[\begin{tikzcd}
% 	&&&& {i + j} \\
% 	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
% 	\\
% 	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
% 	\\
% 	&&&& {n+m}
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
% 	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
% 	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
% 	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
% 	\arrow[from=2-1, to=2-3]
% 	\arrow["{[f_i,f_o]}", from=2-1, to=4-1]
% 	\arrow[from=4-1, to=4-3]
% 	\arrow[from=2-9, to=2-7]
% 	\arrow[from=4-9, to=4-7]
% 	\arrow["{[h_i,h_o]}"{description}, from=2-9, to=4-9]
% 	\arrow[from=6-5, to=4-9]
% 	\arrow[from=6-5, to=4-1]
% 	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
% 	\arrow["{[i_r,j_r]}"{description}, from=1-5, to=2-9]
% \end{tikzcd}\]

% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
\[
 \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square_coloured}
 }
\]

The parts of the original DPOI square is shown in blue.
The extra bits simply account for more complicated interfaces.
For example, the left-hand side $L$ of a rewrite rule
 $\langle L, R \rangle$ which originally was some cospan
 $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
 now is a cospan
 $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
The top half of this diagram says that the outermost interfaces
 should coincide for $L$ and $R$.
The bottom half puts the same restrictions on $G$ and $H$.
For the pushout to exist we impose our assumptions~\ref{pushout:assumptions} on the correspoding arrow.

We also adjust the definition of a boundary complement.
\begin{definition}[Down-closed subgraph]
We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ \emph{down-closed} if for all $e \in E_{\mathcal{H}}$,   every element of $(e]$ is also in $\mathcal{H}$.
\end{definition}

\begin{definition}[Boundary complement in $\WellTypedMdaEcospans$]
\label{def:boundary_new}

% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
\[
\scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
}
\]

Consider the diagram above where the marked square is a pushout
 and $x_{ext} + y_{ext}$ and $n_{ext} + m_{ext}$ are 
 the \textit{outermost} interfaces for the corresponding 
 cospans.
We say that $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ is 
 a \textit{boundary} complement if $[i_c, i_j]$ is mono 
 and there exist $n_c : n_{ext} \to C$ and $m_c : m_{ext} \to C$ 
 making the above triangle commute and such that
there exists a \textit{not necessarily well-typed} mda-cospan

\[
n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}
\]

where $c_i$ and $c_o$ are computed as
 $c_i = g_i \setminus (l_i \setminus x_{ext})$ and
 $c_o = g_o \setminus (l_o \setminus y_{ext})$.
\end{definition}
%
%\begin{figure}
%    \centering
%    \[
%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
%    \label{fig:interface_change_example}
%\end{figure}

% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$. \question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}

\begin{remark}
Note that cospan $n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}$ is not well-typed.
The cospan's input interface contains nodes that previously were a part of the output interface.
\end{remark}

\begin{proposition}
    The boundary complement in~\ref{def:boundary_new} when exists is unique
\end{proposition}

\begin{definition}[Convex down-closed match] 
    We call $m : L \to G$ in $\catname{EHyp_{\Sigma}}$ a convex down-closed match if

    \begin{enumerate}
        \item $m$ is monomorphism
        \item $m(L)$ is a convex down-closed subgraph of $G$
        % \item \update{I've removed the requirement that consistency should be reflected because well typedness prohibits otherwise formed matches.}
    \end{enumerate}
\end{definition}
\

\begin{definition}[Rewriting in $\WellTypedMdaEcospans$]
    Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n_{ext} + m_{ext}$ 
    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}}^{+} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $L \xleftarrow{} l_i + l_o \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} r_i + r_o \xrightarrow{} R$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts

% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
\[\scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square}
    }
\]
    and the following conditions hold

    \begin{enumerate}
        \label{dpoi:assumptions}
        \item $m : L \to G$ is a convex down-closed match;
        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $x_{ext} + y_{ext}$
        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $x_{ext} + y_{ext}$
        \item $x_{ext} + y_{ext} \to C \to G$ is a boundary complement in the leftmost pushout.
    \end{enumerate}
\end{definition}

\begin{remark}
The conditions 2 and 3 are needed to guarantee the existence of the pushout complement.
\end{remark}

The whole interfaces of the resulting e-hypergraph $H$ from the cospan
 $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$
 are constructed as $h_i = (r_i \setminus x_{ext}) + c_i$ 
 and
 $h_o = (r_o \setminus y_{ext}) + c_o$.
 $h_i \to H$ is defined as a copairing of 
 $f_1 : r_i \setminus x_{ext} \to R \to H$ 
 and 
 $f_2 : c_i \to C \to H$. 
Since both $f_1$ and $f_2$ are mono, their copairing is also a mono. 
Similarly for $h_o$.

We then say that an mda-cospan 
\[
    n_{ext} \xrightarrow{a_1} g_{i} \xrightarrow{a_2} G \xleftarrow{b_2} g_{o} \xleftarrow{b_1} m_{ext}
\]
rewrites convexly into another mda-cospan

\[
    n_{ext} \xrightarrow{a_1} h_{i} \xrightarrow{a_2} H \xleftarrow{b_2} h_{o} \xleftarrow{b_1} m_{ext}
\]

under a rewrite rule 
\[
    \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
    x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
\]

if 
\[
    G \xleftarrow{\langle a_2,b_2 \rangle} g_i + g_{o} \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext} \Rrightarrow H \xleftarrow{\langle a_2,b_2 \rangle} h_i + h_{o} \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext}
\]
under a rewrite rule
\[
    \langle l_{i} + l_{o} \xrightarrow{\langle a_2, b_2 \rangle} L \xleftarrow{\langle a_1, b_1 \rangle} x_{ext} + y_{ext},
    r_{i} + r_{o} \xrightarrow{\langle a_2, b_2 \rangle} R \xleftarrow{\langle a_1, b_1 \rangle} x_{ext} + y_{ext} 
    \rangle
\]

\begin{proposition}
$n_{ext} \xrightarrow{n_h} h_i \xrightarrow{f} H \xleftarrow{g} h_o \xleftarrow{m_h} m_{ext}$ is a well-typed mda-cospan.
\end{proposition}

\section{Soundness and Completeness}\label{sec:soundness-and-completeness}

% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}

\begin{figure}
    \[
    \scalebox{0.6}{
    \tikzfig{combinatorial_semantics/f_plus_g_new}
    }
    \]
    \caption{$+$ of two morphisms in $\Ecospans$}
    \label{fig:A+B}
\end{figure}

Note that the necessary pushouts for $\mathcal{S}$ exist. 
\begin{definition}
We denote by $\WellTypedMdaEcospans/\mathcal{S}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
\[
	\phi \sim \psi \quad \text{if} \quad \phi \Rightarrow^*_{[\mathcal{S}]} \psi
\]
where $\mathcal{S}$ denotes set of semi-lattice equations the free functor $\llbracket- \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans$ induced by the interpretation functor sending $\Sigma$ to itself.  CHRIS: NEEDS WORK. 
\end{definition}

\begin{proposition}
The category $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ a semilattice-enriched PROP. 
\end{proposition}
\begin{proof}
The symmetric monoidal structure is inherited from coproduct in $\textbf{EHyp}(\Sigma)$.  It suffices to check that the symmetry and the tensor product of two well-typed e-hypergraphs is again a well-typed e-hypergraph. 
The formal join $+$ of two morphisms of $\WellTypedMdaEcospans$ is defined in the Figure~\ref{fig:A+B}. 
Etc.
\end{proof}
Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by rewrites,  for which we've checked the pushout exists.

\begin{proposition}
    For each cospan $A$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ there is a weak normal form such that 
\[
	A =_{\mathcal{S}} A_1 + \ldots + A_n
\]
 such that each $A_i$ contains no hierarchical edges.
    \begin{proof}
    By applying the rules 1-6 in $\mathcal{S}$ (the ones from Figure~\ref{fig:string-equations}) we either remove 1 hierarchical edge (rules 5-6) or do not change the number of them, but expand the edge onto other edges.
    We will have a redex for such a rule until there is only one hierarchical edge at the outermost level (and only one outermost edge) with no other hierarchical edges within it.
    \end{proof}
\end{proposition}

\begin{theorem}
We have the following equivalence of categories
\[
	PROP^{+}(\Sigma) \cong \WellTypedMdaEcospans~.
\]
\end{theorem}
\begin{proof}
Given by the free functor induced by the interpretation of $\Sigma$ as itself. 
Normal form argument. 
\end{proof}

Unfolding the definitions,  we can spell out the previous theorem in terms of DPO-rewriting. 
\begin{corollary}
For any $\phi, \psi \in \textbf{PROP}^{+}(\Sigma, \mathcal{E})$,  we have
\[
	\phi = \psi \quad \iff \quad \llbracket\phi \rrbracket \Rrightarrow_{\mathcal{S}, \mathcal{E}} \llbracket\psi\rrbracket~ . 
\]
 \end{corollary}
CHRIS: INCORPORATING EXTRA EQUATIONS OUGHT TO BE IMMEDIATE? 
%
%\begin{theorem}[Theorem 35~\cite{Frobenius2}]
%
%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
%\end{theorem}
%
%\begin{proposition}
%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
%\end{proposition}
%
%\begin{theorem}
%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
%\end{theorem}

\section{Conclusion}

We have described a categorical semantics for e-graphs over algebraic theories in terms of Cartesian categories enriched over a semilattice, and defined a natural generalization to account for monoidal theories in terms of SMCs,  similarly enriched.  Having described a translation of e-graphs into our formalism, we developed a combinatorial representation of morphisms in our free category which factor out the structural SMC equations, but remain sensitive to the complexity-relevant equations induced by the enrichment.  The design of such a representation and its corresponding  rewrite theory proved delicate, and -- together with its proof of completeness -- constitute the main technical contributions of the paper.  Finally, we considered an analogue of e-rewriting in our setting, pushing towards the practical considerations necessary for an implementation of our ideas.  Although we believe our work forms a solid basis for a formal mathematical theory of (acyclic) e-graphs,  there are a number of extensions which we intend to investigate in the future.

\subsection{Future Work and Applications}

We have already alluded to the fact we deal with only \textit{acyclic} e-graphs in this paper.  Nevertheless,  within this setting we present a generalization from Cartesian to (symmetric) monoidal theories,  giving rise to a host of potential new domains of application.  Further,  the string diagrammatic and graph rewriting framework we set up is ameenable to extension, including lifting the acyclicity restriction.

\begin{figure}\label{fig:applications}
\[
	\scalebox{0.75}{
	\tikzfig{conclusion/example-e-graphs}
	}
\]
\caption{Example applications of e-hypergraphs.  Left, the ZX-calculus.  Right, the $\lambda$-calculus as represented by string diagrams for Cartesian closed categories.  }
\end{figure}

Symmetric monoidal theories have received significant attention over recent years: the ZX-calculus can be used to characterize quantum processes, with applications to optimization of quantum circuits \cite{coecke_interacting_2011,ZX} ; similar theories also characterize signal flow diagrams 
\cite{baldan_categorical_2014}, digital circuits 
\cite{ghica_operational_2021}, as well as having application to computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  In future work, we hope to apply our generalization of e-graphs to optimization problems in these domains (e.g. Figure \ref{fig:applications})

The theories described above often have more structure than simply being symmetric monoidal: for example, they may contain comonoids, Frobenius algebras or traced structure (as in the case of e-graphs with cycles),  and this structure can be exploited to give more natural (in the sense of factoring out more equations) combinatorial representations.  In the above cases, such representations and their DPO-rewriting theory is well-studied 
\cite{ghica_rewriting_2023} % traced comonoid
[cite, cite, cite].  We consider that our development of hierarchical "e-box" structure within string diagrams and their associated (e-)hypergraphs ( categorically,  semilattice enrichment) is essentially orthogonal to these developments, and so ought to be easily combined to give notions of e-hypergraphs specifically suited to the optimization of these particular monoidal theories.

A particularly interesting extension that we intend to consider is the combination of our "e-structure" with (Cartesian) \textit{closed} categories, allowing the natural combination of e-graph optimization techniques with the $\lambda$-calculus.  The corresponding combinatorial representation would build on the recent work \cite{ghica_hierarchical_2023, alvarez-picallo_rewriting_2022, alvarez-picallo_functorial_2021}. % hierarchical

\bibliographystyle{acm}
\bibliography{bibliography}

\appendix

\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}

\subsection{Proof of the existence of pushouts}

\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}

\end{document}

### 2024-01-04 E-Hypergraphs-LICS24.tex
Small fixes to intro

@@
 \[
+	\hspace{2cm}
     \scalebox{0.65}{
@@
 \]
-\caption{Example translation of e-graphs into string diagrams for semilattice enriched (traced) Cartesian categories. DAN: (a) SHOULD USE MATH FONT.}
+\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
 \end{figure*}
@@
 \]
-The ability to take formal sums, which in concrete categories will usually be set union, is used to model the equivalence class structure of e-graphs.
-
-DAN: TOP TO BOTTOM OR BOTTOM TO TOP? UNCLEAR.\\
-String diagrams are read from top-to-bottom; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generators $c$ with a single output.\\
-DAN: WHY? WITH OR WITHOUT LOSS OF GENERALITY? WHAT DOES THIS MEAN?
+The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
+
+String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
+Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
+This means the (informal) translation given below is fully general for acyclic e-graphs.
 
@@
 \]
-DAN: IT WOULD BE CLEARER IMHO IF THE ARGUMENTS OF C1 AND C2 REMAIN CENTERED AND THE DISCARDED NODES ARE OFF CENTER.
-
 Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}. In particular, note that the translation of the \textit{cyclic} e-graph (e)
@@
 
-Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal sums.
-\\
-DAN: SUMS OR JOINS? CONSISTENT?\\
+Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal joins.
 Thus we generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we outline in the conclusion of this paper.   
@@
 CHRIS: INTRO TO SECTION: CLARIFY THAT WE GIVE E-GRAPHS AS CARTESIAN CASE,  BUT ALSO GENERALIZE TO MONOIDAL CASE. 
+CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX.
 
@@
 
-CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.  WHAT DOES (4) FOR CONSISTENCY MEAN?
+CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.
 
@@
 SHOULD WE USE $< ; \phi$ NOTATION TO MATCH OTHER HYPERGRAPH DEFINITION?
+WHAT DOES (4) FOR CONSISTENCY MEAN?
 
@@
 
-\begin{figure}\label{fig:applications}
+\begin{figure}
 \[
@@
 \]
-\caption{Example applications of e-hypergraphs.  Left, the ZX-calculus.  Right, the $\lambda$-calculus as represented by string diagrams for Cartesian closed categories.  }
-\end{figure}
+\[
+	\scalebox{0.75}{
+	\tikzfig{conclusion/example-traced}
+	}
+\]
+\caption{Example applications of e-hypergraphs.
+Top left, the ZX-calculus.
+Top right, the $\lambda$-calculus as represented by string diagrams for Cartesian closed categories.
+Below,  an extension of Figure \ref{fig:} by another rewrite step, $1 * x \to x$,  introducing a cycle modelled by traced structure.
+}
+\end{figure}\label{fig:applications}
 
@@
 We consider that our development of hierarchical "e-box" structure within string diagrams and their associated (e-)hypergraphs ( categorically,  semilattice enrichment) is essentially orthogonal to these developments, and so ought to be easily combined to give notions of e-hypergraphs specifically suited to the optimization of these particular monoidal theories.
-
 A particularly interesting extension that we intend to consider is the combination of our "e-structure" with (Cartesian) \textit{closed} categories, allowing the natural combination of e-graph optimization techniques with the $\lambda$-calculus.
 The corresponding combinatorial representation would build on the recent work \cite{ghica_hierarchical_2023, alvarez-picallo_rewriting_2022, alvarez-picallo_functorial_2021}.
-% hierarchical
+String diagrams for these various examples are included in Figure \ref{fig:applications}% hierarchical
 

### 2024-01-04 E-Hypergraphs-LICS24.tex
Section 3.2, moved theorems to the appendix

--- 2024-01-01 E-Hypergraphs-LICS24.tex
@@
 Morphisms $K \to C \to G$ are called a \textit{pushout complement} and morphism $m$ is called a \textit{match},  defined when the corresponding pushout complement exists.
-When $f,g$ are expected to be monomorphisms.
-Constructing $C$ can be intuitively understood as finding an occurrence of $L$ in $G$ and removing from this occurrence everything that has no pre-image in $K$ and then inserting in there everything from $R$ that has no pre-image in $K$ -- hence the two pushouts: one for deletion and one for insertion.
+When $f,g$ are expected to be monomorphisms, constructing $C$ can be intuitively understood as finding an occurrence of $L$ in $G$ and removing from this occurrence everything that has no pre-image in $K$ and then inserting in there everything from $R$ that has no pre-image in $K$ -- hence the two pushouts: one for deletion and one for insertion.
 In this case,  we say that $G$ is rewritten into $H$ via the rewrite rule $L \xleftarrow{} K \xrightarrow{} R$.
 
-However, this notion of DPO does not account for hypergraphs with interfaces, i.e.
+To express the rewriting of (hyper-)graphs with interfaces, e.g.
+the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout with \textit{interfaces} rewriting (DPOI).
+
+\begin{definition}[DPOI rewriting]
+Fix a category $\mathbb{C}$ with pushouts.
+Given morphisms $G \xleftarrow{} J$ and $H \xleftarrow{} J$ in $\mathbb{C}$, $G$ rewrites into $H$ with interface $J$ if there exists a rewrite rule $L \xleftarrow{} K \xrightarrow{} R$ and object $C$ with suitable morphisms in $\mathbb{C}$ such that the diagram below commutes and two marked squares are pushouts.
+\[
+\begin{tikzcd}
+L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
+G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H \\
+                  & J \arrow[lu] \arrow[u] \arrow[ru]          &              
+\end{tikzcd}
+\]
+
+\end{definition}
+When $\mathbb{C} = \catname{Hyp_{\Sigma}}$, the rewrite rules become cospans of hypergraphs with matching interfaces: $i \xrightarrow{} L \xleftarrow{} j$, $i \xrightarrow{} R \xleftarrow{} j$.
+By noting that a cospan $0 \xrightarrow{} L \xleftarrow{[f_{i},f_{j}]} i + j$ encodes the same data as a cospan $i \xrightarrow{f_{i}} L \xleftarrow{f_{j}} j$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $L \xleftarrow{} K \xrightarrow{} R$ where $K \cong i + j$.
+Likewise, for the target cospan $n \xrightarrow{} G \xleftarrow{} m$  $J \cong n + m$.
+% However, this notion of DPO does not account for hypergraphs with interfaces, i.e.
 those from $\MdaCospans$ where $l,r$ from $\langle l, r \rangle$ are cospans.
@@
 CHRIS: DEFINE DPOI FORMALLY,  IN DEFINITION ENVIRONMENT? PERHAPS COMBINE WITH BOUNDARY COMPLEMENT BELOW,  TWO FIGURES ON ONE LINE?
-\[
-\begin{tikzcd}
-L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
-G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H \\
-                  & J \arrow[lu] \arrow[u] \arrow[ru]          &              
-\end{tikzcd}
-\]
-
-To mimic rewrites in $\textsf{PROP}(\Sigma)$ we need to put some restrictions on pushout complements as not all of them lead to mda-cospans. CHRIS: "LEAD TO"? BE PRECISE! :)
-
-\begin{definition}[Boundary complement for $\catname{MACsp_{D}(Hyp_{\Sigma}))}$]
+
+To mimic rewrites in $\textsf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textsf{PROP}(\Sigma)$.
+CHRIS: "LEAD TO"?
+BE PRECISE!
+:)
+
+\begin{definition}[Boundary complement]
 \label{def:boundary_original}
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-\end{figure}
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
+% \end{figure}
+
+% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
+\[\begin{tikzcd}
+	L && {i+j} \\
+	\\
+	{G^{\urcorner}} && {L^{\bot}} \\
+	\\
+	&& {n+m}
+	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
+	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
+	\arrow["f"', from=1-1, to=3-1]
+	\arrow["{a = [a_1,a_2]}"', from=1-3, to=1-1]
+	\arrow["{c = [c_1,c_2]}", from=1-3, to=3-3]
+	\arrow["g", from=3-3, to=3-1]
+\end{tikzcd}\]
 
@@
 \end{definition}
+
+\begin{remark}
+    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e.
+morphisms in $\MdaCospans{\Sigma}$, is a part of the above definition.
+\end{remark}
 
@@
 
-\begin{definition}[DPOI Rewriting]
-Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ in $\catname{HypI(\Sigma)}$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ in $\catname{Hyp_{\Sigma}}$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold
+\begin{definition}
+Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold
 
@@
 \end{definition}
+
+ALEKSEI: THE WORD `CONVEXLY' ABOVE MEANS THAT THE REWRITING IS FORMULATED FOR MOMOGAMOUS MORPHISMS. SHALL WE STICK TO THIS FORMULATION?
 
@@
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
+
+\begin{theorem}[Theorem~\ref{th:existence_of_pushouts} restatement]
+Under certain conditions pushouts in $\catname{EHyp_{\Sigma}}$ exist.   
+\end{theorem}
 
@@
 
-\begin{proposition}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
+\begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
+    The boundary complement in~\ref{def:boundary_new} when exists is unique.
 \end{proposition}
@@
 
+\begin{theorem}[Existence of pushouts in $\catname{Hyp_{\Sigma}}$]
+
+    Pushouts in $\catname{Hyp_{\Sigma}}$ always exist.
+\end{theorem}
+
+\begin{proof}
+    Consider a diagram below
+    % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAC0QBfU9TXfIRQBGclVqMWbABrdeIDNjwEiZYePrNWiEAE05fJYKKj11TVJ0AdK3ABmAJzoBjYNIDUursBvYAtgAEAEpcBgr8ykLIAEyk0RqS2iAAitziMFAA5vBEoI4QfkhkIDgQSKISWmx2YfmFiMWlSLGVliCZINQMdABGMAwAChHGOg5YmQAWOLUOBeXUTYgAzOaJbL6OLsBYAPrC3r5YgSEzc4gtiyutSRtOrrvRB7ZHwaE8ebP1F2WIACzUfTAUCQAFolsULEkAFY7aKdEDdPqDYYqUbjKaneoVRb-ECA4HLCFrHQw4SYpBXHGrKo6JhpLhAA
+
+\[
+        \begin{tikzcd}
+        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{i_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
+        Y \arrow[r, "\sfrac{i_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+                                                                        &                                                                  & Q
+        \end{tikzcd}
+\]
+
+Pushout of two hypergraphs $X = \{V_{X}, E_{X}, s_{X}, t_{X}\}$ and $Y = \{V_{Y}, E_{Y}, s_{Y}, t_{Y}\}$ (we will omit labelling functions here since they are irrelevant) along $Z$ is computed in two steps.
+First, a coproduct of $X$ and $Y$ is taken which is 
+\[
+    X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X + Y}, t_{X+Y}\}
+\]
+where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $\langle s'_{X}, s'_{Y} \rangle : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};i_{1,V}^{*}$ and $s'_{Y} = s_{Y};i_{2,V}^{*}$ where $i_1,i_2$ are corresponding coproduct injections, similarly for $t_{X+Y}$.
+Recall that each arrow in $\catname{Hyp_{\Sigma}}$ is a pair of arrows for edges and nodes and we will further use subscripts $V$ and $E$ when referring to these functions.
+Then, consider relations 
+\[
+    S_{V} = \{(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \; \exists z \in V_{Z} \; .
+\; x_i = f_{V};i_{1,V}(z) \text{ and } y_j = g_{V};i_{2,V}(z) \text{where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\}
+\]
+\[
+    S_{E} = \{(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; | \; \exists z \in E_{Z} \; .
+\; x_i = f_{E};i_{1,E}(z) \text{ and } y_j = g_{E};i_{2,E}(z) \text{where $x_i \in E_{X}$ and $y_j \in E_{Y}$ }\}
+\]
+and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.
+
+\begin{lemma}
+    $x R y$ if and only if either $x = y$ or there exists a sequence $(a_1,\ldots,a_{n-1},{a_n})$ such that $x = a_1$ and $y = a_{n}$ such that $a_k S a_{k+1}$ or $a_{k + 1} S a_{k}$ for $k < n$.
+\end{lemma}
+
+A quotient of $X+Y$ by these relation is 
+\[
+    \sfrac{X+Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X+Y}}{\sim (R_{V},R_{E})}\}
+\]
+We will then refer to $\sim (R_{V},R_{E})$ just as $\sim$, i.e. writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and the exact relation will be clear from the context.
+We have 
+\[
+    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
+\]
+In particular, there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
+Let $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*} = s_{X+Y};[-]^{*}$ and then $\sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}$.
+There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
+We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has in image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
+We need to check that if $t_1 \sim t_2$ then $\sfrac{s_{X+Y}}{\sim}([t_1]) = \sfrac{s_{X+Y}}{\sim}([t_2])$. 
+There are three options
+\begin{enumerate}
+    \item $t_1 = e_{y}^{1}$ and $t_2 = e_{y}^{2}$
+    \item $t_1 = e_{x}^{1}$ and $t_2 = e_{x}^{2}$
+    \item $t_1 = e_{x}$ and $t_2 = e_{y}$ 
+\end{enumerate}
+We will prove the last point (all the others are analogous).
+
+According to the lemma above $e_{x} \sim e_{y}$ gives us two options.
+\begin{enumerate}
+    \item Either $e_{x} = e_{y}$
+    \item or there is a sequence $w = (a_1, \ldots, a_{n})$ such that $a_1 = e_{x}$ and $a_{n} = e_{y}$.
+\end{enumerate}
+In the first case the equality holds on the nose.
+We then prove the second case by induction on the length of $w$.
+\begin{itemize}
+    \item In the case $|w| = 2$, $e_{x} S e_{y}$ (or $e_{y} S e_{x}$) which means that there exists $z$ in $E_{Z}$ such that $f_{E};i_{1,E}(z) = e_{x}$ and $g_{E};i_{2,E}(z) = e_{y}$.
+            Recall that because $f$ and $g$ are homomorphisms, the following equalities hold
+            \[
+            s_{Z};f_{V}^{*};i_{1,V}^{*}(z) = f_{E};i_{1,E};s_{X+Y}(z) = s_{X+Y}(e_{x})
+            \]
+            \[
+            s_{Z};g_{V}^{*};i_{2,V}^{*}(z) = g_{E};i_{2,E};s_{X+Y}(z) = s_{X+Y}(e_{y})
+            \]
+            Then,
+            \[
+            \sfrac{s_{X+Y}}{\sim}([e_{x}]) = [s_{X+Y}(e_{x})]^{*} = [s_{Z};f_{V}^{*};i_{1,V}^{*}(z)]^{*}
+            \]
+            and
+            \[
+            \sfrac{s_{X+Y}}{\sim}([e_{y}]) = [s_{X+Y}(e_{y})]^{*} = [s_{Z};g_{V}^{*};i_{2,V}^{*}(z)]^{*}
+            \]
+            Then, by noting that $f_{V};i_{1,V}(z) = g_{V};i_{2,V}(z)$ for all $z$ in $V_{Z}$ (because $\sfrac{V_{X} + V_{Y}}{\sim}$ is so defined) we have $[s_{X+Y};(e_{x})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
+    \item Suppose if $e_{x} S e_{y}$ by a chain of length $n$ or less then $[s_{X+Y}(e_{x})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
+    \item Now suppose $e_{x} S e_{y}$ by a chain $w = (a_1, \ldots, a_{n+1})$ and a subchain $(a_1, \ldots, a_{n})$ satisfies the inductive hypothesis and hence $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(a_{n})$.
+            Note that $S_{E}, S_{V}$ relate only elements with the pre-image in $E_{X}$ (respectively, $V_{X}$) with elements with the pre-image in $E_{Y}$ (respectively, $V_{Y}$). 
+            Elements with the pre-image in the same set are not related.
+            Then we have $a_n S a_{n+1} = e_{y}$ and need to show that $[s_{X+Y}(a_{n+1})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
+            This follows by exactly the same argument as the base case as $a_{n+1}$ is necessarily in $E_{X}$.
+            And we finally have $[s_{X+Y}(e_{x})]^{*} = [s_{X+Y}(a_{n+1})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
+\end{itemize}
+
+Next 
+\[
+    \sfrac{i_{1,V}}{\sim} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim} = [i_{1,V}]
+\]
+and
+\[
+\sfrac{i_{1,E}}{\sim} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim} = [i_{1,E}]
+\]
+similarly for $\sfrac{i_{2,E}}{\sim}$ and $\sfrac{i_{2,V}}{\sim}$
+We need to show
+\begin{enumerate}
+    \item $s_{X};(\sfrac{i_{1,V}}{\sim})^{*}(e) = \sfrac{i_{1,E}}{\sim};\sfrac{s_{X+Y}}{\sim}(e)$
+    \item $t_{X};(\sfrac{i_{1,V}}{\sim})^{*}(e) = \sfrac{i_{1,E}}{\sim};\sfrac{t_{X+Y}}{\sim}(e)$
+\end{enumerate}
+and likewise for $i_{2}$.
+Let's check the first one for which we first need to check that $([-]_{E},[-]_{V})$ is a homomorphisms, i.e.
+\[
+[s_{X+Y}(e)]^{*} = \sfrac{s_{X+Y}}{\sim}([e]) = [s_{X+Y}(e)]^{*}
+\]
+hence, $(\sfrac{i_{1,V}}{\sim}, \sfrac{i_{1,E}}{\sim})$ is a homomorphism because $(i_{1,V},i_{1,E})$ and $([-]_{E},[-]_{V})$ are.
+
+Now, the square obviously commutes.
+To show that $\sfrac{X + Y}{\sim}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1 = (j_{1,V}, j_{1,E})$ and $j_2 = (j_{2,V}, j_{2,E})$ such that the above square commutes there is a unique $u = (u_{V}, u_{E})$ such that $j_1 = \sfrac{i_{1}}{\sim} ; u$ and $j_2 = \sfrac{i_2}{\sim} ; u$.
+We will define such $u$ as $u_{E}([i_{1,E}(e_{x})]) = j_{1,E}(e_{x})$ and $u_{E}([i_{2,E}(e_{y})]) = j_{2,E}(e_y)$ (similarly for $u_{V}$) and claim that it is a well-defined function by showing that if $t_1 \sim t_2$ then $u_{E}(t_1) = u_{E}(t_2)$.
+There are three options analogoues to the proof above, so we will show the base case for $t_1 = i_{1,E}(e_{x})$ and $t_2 = i_{2,E}(e_{y})$.
+$t_1 \sim t_2$ implies that there exists at least one $z$ in $E_{Z}$ such that $i_{1,E}(e_{x}) = f_{E};i_{1,E}(z)$ (or $e_{x} = f_{E}(z)$) and $[i_{1,E}(e_{x})] = \sfrac{i_{1,V}}{\sim}(e_x)$, and $e_y = g_{E}(z)$ $[in_{2,E}(e_{y})] = \sfrac{i_{2,V}}{\sim}(e_y)$.
+By applying $j_{1,E}$ to the first we get $j_{1,E}(f_{E}(z)) = j_{1,E}(e_x)$ and similarly for $j_{2,E}$: $j_{2,E}(g_{E}(z)) = j_{2,E}(e_y)$ and $j_{1,E}(e_{x}) = j_{2,E}(e_{y})$ by commutativity.
+Hence, $u_{E}([in_{1,E}(e_x)]) = u_{E}([in_{2,E}(e_y)])$.
+
+Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.
+Finally, $u$ is unique.
+Assume, that there exists $w : \sfrac{X+Y}{\sim R} \to Q$ for the same $j_1$ and $j_2$.
+In order for them to commute $w_{E}([i_{1,E}(x)]) = j_1(x)$ and $w([in_{2,E}(y)]) = j_2(y)$ (similarly for $w_{V}$) which is the same as $u_{E}$.
+That is, $j_1$ and $j_2$ determine $u$ uniquely.
+\end{proof}
+
+\begin{remark}
+    Both relations can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, i.e. on the coproduct of nodes and edges.
+    To make things well-typed, we will use corresponding coproduct injections $i_{V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $i_{E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either node or edge into these functions.
+    For example, an immediate successor of a node $x$ can be written functionally as $<_{c,\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x))$.
+\end{remark}
+
+\begin{remark}
+    Using functional notation we can reformulate the notion of homomorphism of two e-hypergraphs in the following way.
+    \begin{definition}{E-hypergraphs homomorphism}
+        \label{def:e-homo-2}    
+        A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+
+        \begin{enumerate}
+            \item $\phi$ is hypergraph homomorphism
+
+            \item \[
+                \phi_{E}(<_{c,\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x))) = <_{c,\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+                \] when $x$ is a node,
+                and
+                \[
+                \phi_{E}(<_{c,\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x))) = <_{c,\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))  
+                \] when $x$ is an edge.
+                We do not require such an equation when $[x) = \varnothing$.
+
+            % \item
+            % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
+
+                \item
+            When $x \in E_{\mathcal{F}}$
+            \[
+                \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
+                \subseteq
+                \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
+            \]
+            , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+            And, when $x \in V_{\mathcal{F}}$
+            \[
+                \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
+                \subseteq
+                \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+            \]
+            \end{enumerate}
+    \end{definition}
+\end{remark}
+
+\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
+\label{th:existence_of_pushouts}
+Consider the following span in $\catname{EHyp_{\Sigma}}$
+% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
+\[\begin{tikzcd}
+	Z && X \\
+	\\
+	Y
+	\arrow["f", from=1-1, to=1-3]
+	\arrow["g"', from=1-1, to=3-1]
+\end{tikzcd}\]
+such that
+\begin{enumerate}
+    \item $Z$ is a \textit{descrete} e-hypergraph.
+    \item $f,g$ are monos.
+    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
+    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
+    \item $\consistency^{\mu}(f(v_i)) = \consistency^{\mu}(f(v_j))$ and $\consistency^{\mu}(g(v_i)) = \consistency^{\mu}(g(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
+\end{enumerate}    
+then the pushout $X +_{f,g} Y$ exists.
+\end{theorem}
+\begin{proof}
+    Recall the diagram below.
+    \[
+        \begin{tikzcd}
+        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{i_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
+        Y \arrow[r, "\sfrac{i_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+                                                                        &                                                                  & Q
+        \end{tikzcd}
+    \]
+    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
+    First, a coproduct of $X+Y$ is computed which is 
+    \[
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{c,X+Y} \}
+    \]
+
+    once again omitting the labelling function and where source and target functions are as before.
+    And consider the same relations $S_{V},S_{E}$ and $R_{V}, R_{E}$.
+
+    We then quotient the sets of edges and nodes in $X + Y$ by $R_{V},R_{E}$
+    \[
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \#_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{c,\sfrac{X + Y}{\sim (R_{V},R_{E})}} \}
+    \]
+    where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$ and we will further omit writing the relations explicitly.
+    Recall that $<_{c}$ is essentially a transitive clousure of $<_{c}^{\mu}$ and hence we can first define 
+    \[<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
+    \] and 
+    \[
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
+    \] where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    We will consider several cases.
+    Assume that $u$ is an element of $V_{X+Y}$ and
+    \begin{enumerate}
+        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
+        %       The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
+        %       We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
+        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$ $[g_{V}(z)) \not = \varnothing$.
+              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z)))$.
+              We then define 
+              \[
+                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]
+              \]
+        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
+              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(i_{V_{X}}(f_{V}(z)))$.
+              We then define 
+              \[
+                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]
+              \]
+        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u)) = e'$
+        \[
+            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))] = [e']
+        \]
+        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))$ is undefined (i.e.
+$[u) = \varnothing$), but there exists $[e]$ such that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e']$ and $[u] \in t([e])$ or $[u] \in s([e])$, then
+              \[ 
+                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
+              \]
+              Assumptions above guarantee that the predecessor of such node does not depend on the choice of $e$ if there are several of them.
+    \end{enumerate}
+    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
+    Now, if $u$ is an element of $E_{X+Y}$.
+    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
+              Consider two cases.
+              \begin{enumerate}
+                \item  $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u)) = u_1$.
+                        Then we define
+                    \[
+                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u))]
+                    \]
+                \item $[u) = \varnothing$ and for some $[v_i] \in s([u])$ and $[v_j] \in t([u])$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
+                      Then we define 
+                    \[
+                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                    \] where $[v]$ is either in $s([u])$ or $t([u])$.
+                     The second assumption in~\ref{pushout:assumptions} guarantees that all candidates for $[v]$ share the same predecessor, i.e. the definition does not depend on a particular $[v]$.
+              \end{enumerate}
+    We will also build this relation in steps, first we do steps for nodes (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for nodes until fixed point.
+    Essentially, this interleaving means that nodes inherit the relation from incident edges and vice versa.
+    Clearly, all the cases above are disjoint.
+
+    Let's check that this is a well-defined function.
+    Suppose $t_1, t_2$ are nodes and $t_1 \sim t_2$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
+    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$ and two cases
+    \begin{itemize}
+        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(t_2)) = \varnothing$
+        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(t_2)) \not = \varnothing$
+    \end{itemize} 
+    Consider the first case, then, by definition.
+    \[
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
+    \]
+    and
+    \[
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
+    \]
+    so we have well-defined-ness on the nose. We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
+
+    % \begin{itemize}
+        % \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
+        % \[
+        %     [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
+        % \]
+        % \[
+        %     [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
+        % \]
+        % and
+        % \[ 
+        %     [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
+        % \]
+        % by commutativity.
+        % \item The other conditions satisfy functionality by definition.
+
+    % \end{itemize}
+
+    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms.
+    \begin{enumerate}
+        \item $s_{X+Y};[]_{V}^{*}(e) = []_{E};s_{\sfrac{X+Y}{\sim}}(e)$
+              This is the same as in $\catname{Hyp_{\Sigma}}$.
+        \item $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
+              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
+              This is homomorphic by definition. 
+        \item $[<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
+              Let's make case analysis on $[v]_{V}$.
+              \begin{itemize}
+                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};i_{1,V}(z)$.
+                    \begin{itemize}
+                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $i_{1,V}$ is a coproduct injection, it must be the case that $[i_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
+                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
+                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}[v]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
+                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V};i_{V_{X} + V_{Y}}(z))] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
+                    \end{itemize}
+                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
+              \end{itemize}
+    \end{enumerate}
+    Then $<_{c,\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
+
+    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$.
+    We can consider the original consistency relation as a function $\consistency_{X+Y}^{\mu} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
+    Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
+    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
+    \begin{enumerate}
+        \item If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
+                \begin{align*}
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
+                    \langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)))
+                \end{align*}
+        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{V_{X} + V_{Y}}([v]_{V}))
+        \]
+    \end{enumerate}
+
+    and for nodes
+    \begin{enumerate}
+
+    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')))
+    \]
+    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{E_{X} + E_{Y}}([e]_{E}))
+    \]
+    \end{enumerate}
+    where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
+
+    We then define
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
+    \]
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
+    \]
+    where $c$ is a reflexive, symmetric and transitive closure.
+
+    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency^{\mu}$.
+
+    \begin{itemize}
+        \item Let us show well-definedness.
+        \begin{itemize}
+
+            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_2])$.
+              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};i_{1,V}(z)$ and $v_2 = g_{V};i_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
+              \begin{align*}
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+                %      \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
+                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
+                %      = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\hashtag_{Z}(i_{E_{Z}}(z)))
+              \end{align*}
+              Similarly,
+              \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+              \]
+              and we get well-definedness on the nose.
+              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
+            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$.
+                  This would mean that for all $v_i \in V_{Z}$ $\consistency^{\mu}(f_{V}(v)) = \varnothing$ and $\consistency^{\mu}(g_{V}(z)) = \varnothing$ and automatically
+                  \[
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                  \]
+            \end{itemize}
+        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
+            \[
+             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
+             \subseteq
+             \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))
+             \]
+            By definition, 
+            \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
+            \]
+            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion $\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ because it is a closure.
+            If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness is true because empty set is a subset of any set.
+            Similarly for edges.
+    \end{itemize}
+
+    Let's now prove the universal property.
+    Let $u_{E}([i_{1,E}(x)]) = j_{1,E}(x)$ and $u_{E}([i_{2,E}(y)]) = j_2(y)$.
+    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,E}$ and $j_{2,E}$.
+
+    Let's say that $t_1 \sim t_2$ then $u_{E}([t_1]) = u_{E}([t_2])$.
+    $u_{E}([t_1]) = u_{E}(f_{E};i_{1,E}(z)) = j_1{f_{E}(z)}$ and $u_{E}([t_2]) = u_{E}(g_{E};i_{2,E}(z)) = j_{2}(g_{E}(z))$ which are equal by commutativity.
+\end{proof}
+
+\begin{lemma}
+    \label{lemma:choice_of_nodes_agnostic}
+    Suppose $[e) = \varnothing$ and there exists $[v_1] \in s([e])$ and $[v_2] \in s([e])$ such that $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2]))$.
+    \begin{proof}
+        We will show the statement is true for the base case when $[v_1]$ and $[v_2]$ do not inherit their relations.
+        In other words, $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$ mean that was achieved after running steps (1)-(3) for nodes.
+        From $[v_1] \in s([e])$ we get that there exists $v'_1$ such that $v_1 \sim v'_1$ and $v'_1 \in s(e)$ and similarly $v'_2 \sim v_2$.
+        \begin{itemize}
+            \item $v_1 = v'_1$, $v_1 \in s(e)$, and $v_2 = v'_2$, $v_2 \in s(e)$.
+            \begin{itemize}
+                \item $v_1$ has a pre-image $f_{V};i_{1,V}(z)$ and hence $[f_{V}(z)) = \varnothing$ (because $[e) = \varnothing$) and $v_2$ has a pre-image $f_{V};i_{1,V}(z')$.
+                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{2,E}(e_{y})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e'_{y})]$.
+                      Because $e_{y} = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z)))$ and $e'_{y} = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z')))$ we have $e_{y} = e'_{y}$ because under our assumptions nodes in the image of $g_{V}$ should share the same predecessor.
+                      Hence, $[i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})]$.
+                \item Case for $v_1 = g_{V};i_{2,V}(z)$ is symmetric.
+                \item Case when $v_1$ has no pre-image in $V_Z$ is not possible, because it would mean that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1))]$ (according to step (3)) which is not possible as it would mean that $[e) \not = \varnothing$ (because $[v_1) \not = \varnothing$).
+            \end{itemize}
+            \item $v_1 \not = v'_1$ and $v'_2 \not = v_2$ and hence $v'_1 \in s(e)$ and $v'_2 \in s(e)$.
+            \begin{itemize}
+                \item $v'_1$ has a pre-image $f_{V};i_{1,V}(z)$ and hence $[f_{V}(z)) = \varnothing$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) = [i_{2,E}(e_{y})]$.
+                      $v'_2$ has a pre-image $f_{V};i_{1,V}(z')$ and for the same reason 
+                      \begin{align*}
+                        &<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) =\\ 
+                        &[i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})] =\\
+                        &<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                    \end{align*}
+                \item Case when $v'_1 = g_{V};i_{2,V}(z)$ is similar.
+            \end{itemize}
+            \item $v_1 = v'_1$, $v'_2 \sim v_2$ and $v_1 \in s(e)$, $v'_2 \in s(e)$.
+            \begin{itemize}
+                \item $v_1 = f_{V};i_{1,V}(z)$ and hence $v'_2 = f_{V};i_{1,V}(z')$ and $[f_{V}(z)) = \varnothing$ as well as $[f_{V}(z')) = \varnothing$.
+                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{2,E}(e_{y})]$ and similarly $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e'_{y})]$.
+                      And by the following the same reasoning as above we get the equality.
+                \item Case when $v_1 = g_{V};i_{2,V}(z)$ is similar.
+            \end{itemize}
+            \item The remaining case is similar.
+        \end{itemize}
+        The proofs for the cases when $[v_1] \in t([e])$ and $[v_2] \in t([e])$ and for the combinations of $t,s$ are similar.
+    \end{proof}
+\end{lemma}
+
+\begin{lemma}
+    \label{lemma:choice_of_edges_agnostic}
+    Suppose $[v) = \varnothing$ and there exist $e_1$ and $e_2$ such that $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ and $[v] \in s([e_1])$, $[v] \in s([e_2])$.
+    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$
+    \begin{proof}
+        We will again show the base case when $[e_1]$ and $[e_2]$ inherit the relation atmost at 1 step.
+        That is, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.
+        We then have the following cases.
+        \begin{itemize}
+            \item $v \in s(e_1)$ and $v \in s(e_2)$.
+                  The only possible case for $e_1$ and $e_2$ is that $[e_1) = \varnothing$ and $[e_2) = \varnothing$ (because $[v) = \varnothing$).
+                  Then, 
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
+                  \], where $[v_1] \in s([e_1])$ and $[[v_1]) \not = \varnothing$.
+                  And 
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                  \], where $[v_2] \in s([e_2])$ and $[[v_2]) \not = \varnothing$.
+                  The cases are following.
+                  \begin{itemize}
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and 
+                          $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{1,E}(e'_{x})]$. 
+                          Then, $e'_{x} = e_{x}$ because the nodes in the image of $f$ should have the same predecessor.
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and 
+                          $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_{y})]$
+                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$ (because they are from the same connected component).
+                          \begin{itemize}
+                            \item $v_1 = f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$.
+                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}(z)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
+                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it can not be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_y)]$.
+                                  If $v'_2 \in s(e_2)$ and $v_2 \sim v'_2$ then $v'_2$ has a pre-image in $V_{Y}$ and $v_2$ has a pre-image in $V_{X}$, i.e. $v_2 = f_{V};i_{1,V}(z_2)$. 
+                                  The equation above implies that $[g_{V}(z_2)) \not = \varnothing$ which is not the case as $[v'_2) = \varnothing$.
+                            \item $v_1 = g_{V};i_{2,V}(z_1)$ and $[g_{V}(z_1)) = \varnothing$. 
+                                  Then, necessarily, $v_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = f_{V};i_{1,V}(z_1)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
+                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$, i.e.
+$v_2 = g_{V};i_{2,V}(z_2)$ and then it should be $[g_{V}(z_2)) \not = \varnothing$ which is not the case.
+                                  If $v'_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$, i.e.
+$v'_2 = g_{V};i_{2,V}(z_2)$ and then it should be $[g_{V}(z_2)) \not = \varnothing$ which is not the case.
+                          \end{itemize}
+                    \item Other cases are not possible, for example, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$
+                          would mean that $e_2$ has a predecessor because $v_2$ does and $v_2 \in s(e_2)$ because $v_2$ has no pre-image in $V_{Z}$.
+$[v_2]$ can not inherit the relation because we consider inheritance only upto 1 step: inheritance for $v_2$ would mean that $v$ inherits the relation in more than 1 step.
+                  \end{itemize}
+            \item $v' \in s(e_1)$ and $v \in s(e_2)$ and $v' \sim v$
+                  Suppose $[e_1) \not = \varnothing$. Then,
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))]
+                  \] and $[e_2) = \varnothing$ and
+                  \[
+                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                  \],
+                  where $[v_2] \in s([e_2])$ and $[[v_2]) \not = \varnothing$.
+                  \begin{itemize}
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{1,E}(e_x)]$.
+                          If $v_2 = f_{V};i_{1,V}(z_2)$, then $[f_{V}(z_2)) \not = \varnothing$ and hence $v'_2 \in s(e_2)$ and $v'_2 \sim v_2$.
+                          It should be the case that $v = g_{V};i_{2,V}(z)$ (if it had a pre-image in $V_{X}$ it would share the predecessor with the pre-image of $v_2$, but $[v) = \varnothing$) and $v' = f_{V};i_{1,V}(z)$, then $e_1$ has a pre-image in $E_{X}$ and $e_2$ has a pre-image in $E_{Y}$.
+                          Then $v'$ and $v_2$ both have pre-image in $V_{X}$ and because of the assumption $[f_{V}(z)) = [f_{V}(z_2))$ and $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [i_{1,E}(e_x)] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$.
+                          The case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$ is not possible as it would mean that $v_2 \in s(e_2)$, but $[e_2) = \varnothing$.
+                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e']))$ is also not possible as it would mean that $e_2$ inherits the relation in more than 1 step.
+                \end{itemize}
+                Cases $[e_1) = \varnothing$ and $[e_2) = \varnothing$ and $[e_1) = \varnothing$, $[e_2) \not = \varnothing$ are not possible.
+            \item $v \in s(e_1)$ and $v' \in s(e_2)$ and $v' \sim v$
+                  This is symmetric to the above.
+            \item $v' \in s(e_1)$ and $v'' \in s(e_2)$ and $v' \sim v$, $v'' \sim v$. 
+                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class so it reduces to the cases above.
+        \end{itemize}
+        The proofs for the cases when $[v] \in t([e_1])$ and $[v] \in t([e_2])$ and for the combinations of $t,s$ are similar.
+    \end{proof}
+\end{lemma}
+
+\begin{remark}
+    To get the result for arbitrary number of steps we take the above lemmas as base case and procceed by induction on the number of steps.
+\end{remark}
+
+\begin{corollary}
+    \label{corollary:nodes}
+    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
+    \begin{align*}
+         [v') \not = \varnothing,\\
+         <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \end{align*}
+    In other words, there exists the node which initiates the inheritance for nodes that had no predecessor.
+\end{corollary}
+\begin{corollary}
+    \label{corollary:edges}
+    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $[v]$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
+    \begin{align*}
+        [v') \not = \varnothing,\\
+        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \end{align*}  
+    In other words, there exists the node which initiates the inheritance for edges that had no predecessor.
+\end{corollary}
+
+\begin{lemma}
+    $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive, assymetric and respects connectivity.
+    \begin{proof}
+        We will separately consider each property.
+        \begin{itemize}
+            \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
+                  Then we have two cases.
+                  \begin{itemize}
+                    \item Assume $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
+                          This would mean that $e_1 = e$ which would mean that the relation $<_{c,X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
+                    \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v]))$.
+                          \begin{itemize}
+                            \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};i_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
+                                  Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [e]$.
+                                  This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
+                                  Contradiction.
+                            \item Case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+                          \end{itemize}
+                  \end{itemize}                  
+                    We do not need to check this requirement for nodes because a node can not be a predecessor.
+            \item Let $e_1,e_2$ be edges. 
+                  Then, if $[e_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+                  For $[e_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
+                  \begin{itemize}
+                    \item If $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
+                          Having at the same time $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
+                    \begin{itemize}
+                        \item $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
+                         This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
+                        \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
+                              Then, $[e_2] = [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
+                              \begin{itemize}
+                                \item If $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
+                                      Then,
+                                      \[
+                                        [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [i_{2,E}(e_y)] = [e_2]
+                                      \]
+                                      This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
+                                \item Case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric to the above.
+                              \end{itemize}
+                        \end{itemize}
+                        \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exits $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
+                              Then, 
+                              \[
+                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
+                              \]
+                              The case when $[e_1) \not = \varnothing$ are similar to the above, so we will just jump to the interesting case when 
+                              $[e_1) = \varnothing$ and
+                              \[
+                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
+                              \]
+                              The cases we have are the following.
+                              \begin{itemize}
+                                \item $v_2 = f_{V};i_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};i_{2,V}(z_2)$.
+                                Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z_2)] = [e_1]$ which also implies that $e_1$ has a pre-image in $E_{Y}$.
+                                \begin{itemize}
+                                    \item $v_1 = g_{V};i_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};i_{1,V}(z_1)$.
+                                    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z_1)] = [e_2]$  
+                                    However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates our assumptions.
+                                \end{itemize}
+                                \item Case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
+                              \end{itemize}
+                  \end{itemize} 
+                  Once again, we do not need to check for nodes as a node can not be on the lefthand side of the relation.
+            \item We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
+                  $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us two cases.
+                  \begin{itemize}
+                    \item If $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
+                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
+                          \begin{itemize}
+                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
+                                That is, $e_2 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$.
+                                  Once again, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$ and 
+                                  \[
+                                    [e_2] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+                                  \]
+                            \end{itemize}
+                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
+                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
+                          Suppose $v_1 = f_{V};i_{1,V}(z)$ and then $v'_1 = g_{V};i_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [e_2]$ and hence $e_2$ has a pre-image in $E_{Y}$.
+                          \begin{itemize}
+                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
+                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
+                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
+                                  By the same argument, 
+                                  \[
+                                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
+                                  \]
+                          \end{itemize}                    
+                  \end{itemize}
+            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
+            \begin{itemize}
+                \item $v' = v$ and hence $v \in s(e_1)$.
+                \begin{itemize}
+                    \item If $v$ has no pre-image in $V_{Z}$, then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]$.
+                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ which is what we want.
+                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
+                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
+                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$.
+                          Because $v \in s(e_1)$ there should also be a path from $e_1$ to $v'$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$
+                    \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
+                    \[ 
+                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
+                    \]
+                    This means that the pre-image of $e_2$ is the predecessor of $f_{V};(z)$ and hence $e_2$ should be the predecessor of $f_{V};i_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
+                    That is, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+                    \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
+                    \[ 
+                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]
+                    \]
+                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ necessarily has a pre-image in $E_{X}$.
+                    Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [i_{2,E}(e_y)] = [e_2]$.
+                    \item Cases when $v$ has a pre-image $g_{V};i_{2,V}(z)$ are symmetric.
+                \end{itemize}
+                \item $v \sim v'$.
+                    Follows by the same argument as above by recalling that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+            \end{itemize}    
+        \end{itemize}
+    \end{proof}
+\end{lemma}
+
+\begin{lemma}
+    If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exist $e_1 \in E_{X+Y}$ such that either $[v] \in s([e_1])$ or $[v] \in t([e_1])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) \not = \varnothing$ and $e_2 \in E_{X+Y}$ such that either $[v] \in s([e_2])$ or $[v] \in t([e_2])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) \not = \varnothing$ then
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))
+    \]
+    \begin{proof}
+        The proof is analogous to the proof of an analogous lemma for $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
+    \end{proof}
+\end{lemma}
+
+\begin{lemma}
+    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exist $v_1$ such that either $[v_1] \in s([e])$ or $[v_1] \in t([e])$ and $v_2$ such that either $[v_2] \in s([e])$ or $[v_2] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) \not = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) \not = \varnothing$ then
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
+    \]
+    \begin{proof}
+        The proof is analogous to the proof of the analogous lemma for $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
+    \end{proof}
+\end{lemma}
+
+\begin{corollary}
+    \label{corollary:consistency_nodes}
+    If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
+    \[
+        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \]
+\end{corollary}
+
+\begin{corollary}
+    \label{corollary:consistency_edges}
+    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
+    \[
+        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \]
+\end{corollary}
+
+\begin{lemma}
+    $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$ satisfies properties~\ref{def:consistency_properties}.
+    The property (1) is satisfied by construction because we explicitly perform a closure.
+    Hence we need to check whether properties (2),(3) and (4) are satisfied.
+    \begin{proof}
+        \begin{itemize}
+            \item Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                  $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
+                  We then have two cases.
+                  \begin{itemize}
+                    \item $v = v'$ and then $v \in s(e)$.
+                    \begin{itemize}
+                        \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))$.
+                                By applying $([]_{V},[]_{E})$ we get 
+                                \[
+                                i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                                \]
+                        \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
+                              It is then also the case that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
+                              Then, by definition, $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))$ and by reflexivity $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$
+                    \end{itemize}
+                    \item $v \sim v'$ follows by the same argument by using functionality, 
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                        \]
+                   \end{itemize}
+                  \item The case when $[v] \in t([e])$ is symmetric.
+                  \item If $i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
+                        $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
+                        By definition, either
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
+                        \]
+                        which means that either $v_1 \consistency_{X+Y}^{\mu} v_2$ or $v'_1 \consistency_{X+Y}^{\mu} v_2$ and $v'_1 \sim v_1$, or $v_1 \consistency_{X+Y}^{\mu} v_2'$ and $v_2' \sim v_2$, or $v_1' \consistency_{X+Y}^{\mu}(v_2')$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
+                        Suppose, $v_1' \consistency_{X+Y}^{\mu} v_2$ and $v_1' \sim v_1$.
+                        It implies that $[v_1') = [v_2)$ and that 
+                        \[
+                            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1'))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+                        \]
+                        The rest cases are analogous.
+                        Or there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}^{\mu}(v_3') \not = \varnothing$ and
+                        \[
+                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
+                        \]
+                        Then the argument is the same but with using $v_3$ in place of $v_2$.
+                    \item Case for edges is analogous.
+        \end{itemize}
+    \end{proof}
+\end{lemma}
+
+\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
+For this we need to check that the defined relations satisfy all the needed properties.
+\begin{proof}
+    This proposition then follows by taking two lemmas above.
+\end{proof}
+\end{proposition}
+
+\begin{proposition}
+\label{prop:boundary_complement}
+    The boundary complement in~\ref{def:boundary_new} when exists is unique
+    \begin{proof}
+        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of nodes and hyperedges.
+        Hence, pushout complement should yield pushout complements for the sets of nodes and hyperedges.
+        In particular,
+          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
+    \begin{figure}[h!]
+        \begin{subfigure}{0.45\textwidth}
+            \[\begin{tikzcd}
+                {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
+                \\
+                {V_{G}^{\urcorner}} &&&& {V_{C}}
+                \arrow["{b_{V}}", from=1-5, to=1-3]
+                \arrow["{a_{V}}", from=1-3, to=1-1]
+                \arrow["{f_{V}}", from=1-1, to=3-1]
+                \arrow["{g_{V}}"', from=3-5, to=3-1]
+                \arrow["c"', from=1-5, to=3-5]
+            \end{tikzcd}\]
+        \end{subfigure}
+        \hfill
+        \begin{subfigure}{0.45\textwidth}
+           % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
+        \[
+        \begin{tikzcd}
+            {E_{L}} && 0 && 0 \\
+            \\
+            {E_{G}^{\urcorner}} &&&& {E_{C}}
+            \arrow["{b_{E}}", from=1-5, to=1-3]
+            \arrow["{a_{E}}", from=1-3, to=1-1]
+            \arrow["{f_{E}}", from=1-1, to=3-1]
+            \arrow["{g_{E}}"', from=3-5, to=3-1]
+            \arrow["c"', from=1-5, to=3-5]
+        \end{tikzcd}
+        \]
+        \end{subfigure}
+    \end{figure}
+    $V$s and $E$s are sets of nodes and hyperedges respectively, marked squares are pushouts and $C$s are pushout complements.
+    Because $i+j$ is discrete, $E_{G}^{\urcorner}$ is the disjoint union of $E_{L}$ and $E_{C}$ and hence $E_{C} = E_{G}^{\urcorner} \setminus E_{L}$.
+    Then, since all the arrows are monos, the left square can be rewritten as follows.
+    \[
+      % https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3IFxcY29uZyBpICsgaiArIHoiXSxbMiwxLCJiX3tWfSJdLFsxLDAsImFfe1Z9Il0sWzAsMywiZl97Vn0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
+    \begin{tikzcd}
+        {i + j + x + y} && {i+j + x} && {i+j} \\
+        \\
+        {i + j + x + y + z} &&&& {i + j + w \cong i + j + z}
+        \arrow["{b_{V}}", from=1-5, to=1-3]
+        \arrow["{a_{V}}", from=1-3, to=1-1]
+        \arrow["{f_{V}}", from=1-1, to=3-1]
+        \arrow["{g_{E}}"', from=3-5, to=3-1]
+        \arrow["c"', from=1-5, to=3-5]
+    \end{tikzcd}
+    \]
+    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for nodes is $i + j + z$ up to isomorphism.
+    To show that pushout complement is unique we next need to show that source and target maps are unique.
+    Suppose that there exist
+    \[
+        C_{1} = (V_{C}, V_{E}, s_{1},t)
+    \]
+    and
+    \[
+        C_{2} = (V_{C},V_{E},s_{2},t)
+    \]
+    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(h)) = s_{G}^{*}(g_{E}(h)) = g_{V}^{*}(s_{2}(h))$.
+    Because $g$ is mono, $g_{V}^{*}(s_{1}(h)) = g_{V}^{*}(s_{2}(h))$ implies $s_{1}(h) = s_{2}(h)$.
+    Similarly for targets.
+    We also need to show that $<_{c}$ and $\consistency^{\mu}$ are unique.
+    Suppose that there exist
+    \[
+        C_{1} = (V_{C}, V_{E}, s,t, <_{1})
+    \]
+    and
+    \[
+        C_{2} = (V_{C},V_{E},s,t, <_{2})
+    \]
+    Because $g$ is homomorphism, it must be the case that $g_{E}(<_{1}(i_{V_{C}}(v))) = <_{G}(i_{V_{G}}(g_{V}(v))) = g_{E}(<_{2}(i_{V_{C}}(v)))$ which implies $<_{2}(i_{V_{C}}(v)) = <_{1}(i_{V_{C}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
+    Similarly for edges.
+    Now suppose that $<_{1}(i_{V_{C}}(v)) = e$ and $<_{2}(i_{V_{C}}(v))$ is undefined.
+    Then, $g_{E}(e) = <_{G}(i_{V_{G}}(g_{V}(v)))$ and there exists $v' \in V_{L}$ and $v'' \in V_{C_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
+    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $C_{1}$, but then it violates our assumptions as there exists some node $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = b_{V};a_{V}(u)$ and $[v') \not = \varnothing$.
+    Therefore we get a contradiction. 
+    Similarly for edges.
+    Finally, let's suppose there exist
+    \[
+        C_{1} = (V_{C}, V_{E}, s,t, <, \consistency_{1}^{\mu})
+    \]
+    and
+    \[
+        C_{2} = (V_{C},V_{E},s,t,<, \consistency_{2}^{\mu})
+    \]
+    Because $g$ is a homomorphism, we have
+    \[
+        \langle g_{V};i_{V_{G}}, g_{E};i_{E_{G}} \rangle^{*}(\consistency_{C_1}^{\mu}(i_{V_{C}}(v))) \subseteq \consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v))
+    \]
+    \[
+        \langle g_{V};i_{V_{G}}, g_{E};i_{E_{G}} \rangle^{*}(\consistency_{C_2}^{\mu}(i_{V_{C}}(v))) \subseteq \consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v))  
+    \]
+    \begin{itemize}
+        \item If $\consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v)) = \varnothing$ then both lefthand sides should be $\varnothing$ and the relations are equal.
+        \item Suppose that $\consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v)) \not = \varnothing$ which further means that $[g_{V};i_{V_{G}}(v)) \not = \varnothing$.
+        \begin{itemize}
+            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \subseteq \consistency_{C_2}^{\mu}(i_{V_{C}}(v))$.
+                  This means that there exists $v'$ such that $v' \consistency_{C_{2}}^{\mu} v$ and $v' \not \consistency_{C_{1}}^{\mu} v$.
+                  Because $g_{V}(v) \consistency^{\mu} v'$ it means that there exists $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{L}$ such that $w \consistency_{L}^{\mu} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
+                  This again contradicts our assumptions as $u \consistency_{C_{2}}^{\mu} u'$ and $w \consistency_{L}^{\mu} w'$.
+            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \cap \consistency_{C_2}^{\mu}(i_{V_{C}}(v)) = \varnothing$
+                  Similar to the above.
+            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \cap \consistency_{C_2}^{\mu}(i_{V_{C}}(v)) \not = \varnothing$
+                  Similar to the above.
+        \end{itemize}
+    \end{itemize}
+    \end{proof}
+    % \update{Can we check the proof? I haven't used all the conditions from monogamous-ness above.
+    % Although, if the morphism $i+j \to L$ is mono, the complement is unique for not-monogamous hypergraphs as well.}
+\end{proposition}
+
 \section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}

### 2024-01-05 E-Hypergraphs-LICS24.tex
figures update

--- 2024-01-04 E-Hypergraphs-LICS24.tex
@@
 
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques.
+Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
 This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits

### 2024-01-05 E-Hypergraphs-LICS24.tex
intro

@@
 The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics for \emph{e-graphs} over algebraic theories, a generalisation of the structures which are the subject of this technique, and to their rewriting. 
-The proposed framework consists of Cartesian categories enriched over a semilattice,  further generalised to symmetric monoidal categories, to account for monoidal theories.
-The aim is to permit the optimisation of a larger class of programs, based on string diagrams.  
-Furthermore, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs,  which we call \emph{e-hypergraphs}, which have the usual advantage that many structural equations are absorbed into a general notion of isomorphism.
+We give a categorical semantics to these structures, called \emph{e-graphs}, in terms of Cartesian categories enriched over a semilattice. 
+We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
+Finally, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs (\emph{e-hypergraphs}).
+They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
 \end{abstract}
@@
 
-\subsection{Semantics via Semilattice Enriched Categories}
+\subsection{Semilattice Enriched Categories}
 
@@
 makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{noauthor_09083347_nodate,joyal_geometry_1991, mellies_functorial_2006}.
-In Figure~\ref{fig:egraph-strings} we extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
+We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
 \[
@@
 
-Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal joins.
-Thus we generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we outline in the conclusion of this paper.   
-\\ DAN: THIS IS A BIT UNCLEAR. WHY IS THIS THE NATURAL LEVEL THE GENERALITY? IT'S NOT QUITE OBVIOUS.
+Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
+However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
+Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
 
@@
 Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some $c_i$.
+With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
 
@@
 \subsection{Soundness and Completeness}
-The main technical result of the paper is a proof soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs.
+The main technical result of the paper is a proof of soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs.
 While the structural equalities of SMCs are factored out in the representation,  important structural equalities arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
@@
 Our use of string diagrams is also central to the intuition behind our work.  
+DAN: INFORMAL REASONING DEVICES DOESN'T SOUND GOOD. WHAT DO YOU MEAN? 
 The string diagrams we work with are left essentially as informal reasoning devices, however there exists a breadth of work on similar diagrammatic formalisms for similar classes of categories.
@@
 CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
+DAN: I SUGGEST TO CITE HASEGAWA IN FURTHER WORK, FOR GENERALISING THE THEORY TO TRACED MC.
 

### 2024-01-05 E-Hypergraphs-LICS24.tex
section 2

@@
 
-First, we fix the following notation.
-Given a catgeory $\mathbb{C}$ and object $A,B \in \mathbb{C}$ ,  we denote $\mathbb{C}(A,B)$ the corresponding hom-set.
+Given a $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
 We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
@@
 
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations} -- formally,  pairs of terms freely generated over the signature.
+A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
 We are interested here in symmetric monoidal theories, which are generalizations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
@@
 \begin{definition}[Products and permutations category]
-A \textit{permutations and products category (PROP)} is a symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$.
-In this case, we will enote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$.
+A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$.
+In this case, we will denote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$.
 A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ is $\textsf{PROP}(\Sigma)$.  
+We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textsf{PROP}(\Sigma)$.  
 \end{definition}
@@
 
-We can now specify an equation associated with a monoidal signature as a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
+An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
 \begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textsf{PROP}(\Sigma)$
+A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textsf{PROP}(\Sigma)$.
 A \textit{symmetric monoidal theory} $\textsf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textsf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
@@
 \]
-By Fox's Theoem [cite],  $\textsf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
+By Fox's Theorem~cite{cite},  $\textsf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
 \end{example}
@@
 
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-\footnote{Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct -- a degeneracy we wish to avoid.} More precisely our hom-sets will be given the structure of a semilattice.
+We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.\footnote{Note, in particular,  that semilattices do not require a unit for $+$.
+This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct---a degeneracy we wish to avoid.
+DAN: REFERENCE OR PROOF?} More precisely our hom-sets will be given the structure of a semilattice.
 Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
@@
 We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics.
-Indeed, the diagrammatic language itself is not truly novel and is similar to \textit{e.g.}, \cite{duncan_generalised_2009}.
+Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
 

### 2024-01-05 E-Hypergraphs-LICS24.tex
fix figures, section 3.1 blurb.

@@
 \[
-	\hspace{2cm}
+	\hspace{1.3cm}
     \scalebox{0.65}{
@@
 
-\begin{figure}\label{fig:egraph-strings}
+\begin{figure}
 \[
@@
 \caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\end{figure}
+\end{figure}\label{fig:egraph-strings}
 
@@
 Our use of string diagrams is also central to the intuition behind our work.
-DAN: INFORMAL REASONING DEVICES DOESN'T SOUND GOOD. WHAT DO YOU MEAN? 
-The string diagrams we work with are left essentially as informal reasoning devices, however there exists a breadth of work on similar diagrammatic formalisms for similar classes of categories.
-We mention here work on hierarchical string diagrams \cite{ghica_hierarchical_2023},  functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-
-DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.
-
-\section{Categorical Semantics of E-Graphs}
-CHRIS: INTRO TO SECTION: CLARIFY THAT WE GIVE E-GRAPHS AS CARTESIAN CASE,  BUT ALSO GENERALIZE TO MONOIDAL CASE. 
+There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
+
 CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
 DAN: I SUGGEST TO CITE HASEGAWA IN FURTHER WORK, FOR GENERALISING THE THEORY TO TRACED MC.
+DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.
+
+\section{Categorical Semantics of E-Graphs}
+In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
 
@@
 A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textsf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textsf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
+We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
+\end{definition}
+The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
 
@@
 \begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textsf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textsf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textsf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
+A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textbf{PROP}(\Sigma)$.
+A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
 \end{definition}
@@
 \[
-	\scalebox{0.6}{
+	\scalebox{0.8}{
   	 \tikzfig{categorical-semantics/Cartesian-equipment}
@@
 \]
-By Fox's Theorem~cite{cite},  $\textsf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
+By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
 \end{example}
@@
 We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.\footnote{Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must also be a categorical coproduct---a degeneracy we wish to avoid.
-DAN: REFERENCE OR PROOF?} More precisely our hom-sets will be given the structure of a semilattice.
+This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid.} More precisely our hom-sets will be given the structure of a semilattice.
 Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
@@
 A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textsf{PROP}^+(\Sigma)$ ($\textsf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textsf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.
+We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
+\end{definition}
+
+The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.
 
@@
 
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the equivalence between hypergraphs of a particular form and terms in $\textsf{PROP}(\Sigma)$ established in~\cite{Frobenius, Frobenius2},  and how one can implement string diagram rewriting using DPO rewriting on hypergraphs.
-CHRIS: INTRO HYPERGRAPHS WITH INTERFACES,  RESTRICTIONS TO SMCS.  Q: INTRO DPO REWRITING HERE TOO? NOTE,   REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ morphisms!
+In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{Frobenius, Frobenius2}.
 
@@
 
+In ~\cite{Frobenius},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
 Hypergraphs generalize graphs by allowing edges to have multiple sources and targets. 
+When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
+However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
+This gives rise to the definition of a hypergraph with interfaces.  
+Subsequent work \cite{Frobenius2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous directed acyclic}),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
+
 \begin{definition}[Category of hypergraphs]\label{def:hypergraph}
@@
 The initial object is given by the empty hypergraph.
-CHRIS: DESCRIPTION OF PUSHOUT HERE?
-
-When we interpret string diagrams as hypergraphs,  we will interpret morphisms as edges,  and wires by vertices.
-However,  we will additionally need a way to specify which vertices are to be considered as \textit{input} and \textit{output}.
-For this purpose,  will use the following construction to equip our hypergraphs with \textit{interfaces}. 
+
 \begin{definition}[Symmetric monoidal category of cospans]
@@
 
-Following~\cite{Frobenius2} morphisms of $\textsf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
+Following~\cite{Frobenius2} morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
 \begin{definition}[Hypergraphs with interfaces]
@@
 The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-CHRIS: NOTE THIS IS OK FOR FROBENIUS,  NOT FOR SMC.
 
@@
 
-In order to represent morphisms of $\textsf{PROP}(\Sigma)$,  it suffices to restrict our attention to MDA-cospans,  as defined below.
+In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
 \begin{definition}[Category of MDA Hypergraphs with Interfaces \cite{Frobenius2}]
@@
 Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  In fact,  due to 
-\cite{Frobenius2},  we have the following theorem which says the category $\MdaCospans$ in fact \textit{characterizes} $\textsf{PROP}(\Sigma)$. 
+\cite{Frobenius2},  we have the following theorem which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
 \begin{theorem}[Corollary 26~\cite{Frobenius2}]
-    $\textsf{PROP}(\Sigma) \cong \MdaCospans$
+    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
 \end{theorem}
-We continue to show a similar theorem relating our \textit{enriched} category $\textsf{PROP}^+(\Sigma)$ to a combinatorial category of e-hypergraphs with interfaces,  introduced in the next section.
-
-CHRIS: EXPLAIN ABOUT MONOIDAL THEORIES AND EQUATIONS.
 
 \subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-CHRIS: PUT BELOW WITH OTHER PRELIMINARIES???
-
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs. 
+
+The correspondence of the previous section further extends to include the equations of a monoidal theory: the equations of a theory are considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
 DPO rewriting is a standard technique for formalising graph rewriting categorically. 
+In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
 
@@
 
-To mimic rewrites in $\textsf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textsf{PROP}(\Sigma)$.
+To mimic rewrites in $\textbf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textbf{PROP}(\Sigma)$.
 CHRIS: "LEAD TO"?
@@
 CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.  
-
-We follow analogous steps to those used to achieve a combinatorial representation of $\textsf{PROP}(\Sigma)$ as a particular subset of plain hypergraphs with interfaces.
+REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ morphisms!
+
+We follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ as a particular subset of plain hypergraphs with interfaces.
 First we define an e-hypergraph,  which supports a hierarchical notion of "e-box" with distinct components.

### 2024-01-05 E-Hypergraphs-LICS24.tex
Sections 3.2 and 5

@@
 By noting that a cospan $0 \xrightarrow{} L \xleftarrow{[f_{i},f_{j}]} i + j$ encodes the same data as a cospan $i \xrightarrow{f_{i}} L \xleftarrow{f_{j}} j$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $L \xleftarrow{} K \xrightarrow{} R$ where $K \cong i + j$.
-Likewise, for the target cospan $n \xrightarrow{} G \xleftarrow{} m$  $J \cong n + m$.
+Likewise, for the target cospan $n \xrightarrow{} G \xleftarrow{} m$,  $J \cong n + m$.
 % However, this notion of DPO does not account for hypergraphs with interfaces, i.e.
@@
 \begin{definition}
-Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold
+Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in the set of rewrite rules $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold
 
@@
 \end{definition}
+
+By resorting to convex matching and boundary complement we virtually require that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
 
@@
 
-We have a more general notion of interface and hence we need to adjust DPOI rewriting a little bit.
-
-We have a more general notion of interface and hence we need to adjust DPOI rewriting a little bit.
-The first issue is that whole interfaces are not preserved during rewriting, consider an example in Figure~\ref{fig:example_rewrite_ecospans} which corresponds to a rewrite rule of $\langle f, f+g \rangle$.
-
-%\begin{figure}[t]
-%    \centering
-%    \[
-%    \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_to_f_plus_g_example}}
-%    \]
-%    \caption{Example rewrite rule in $\WellTypedMdaEcospans$}
-%    \label{fig:example_rewrite_ecospans}
-%\end{figure}
-
-The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-Note that the outermost interfaces are preserved.
-
-Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-The entry in the middle of the span depicts the outermost interfaces which are preserved.
-The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-\update{$i,j$ are replaced}
-
-To account for such changes in the interfaces we put some additional constraints on DPOI rewriting.
+As extended cospans have a more general notion of interface we need to make some adjustments to DPOI rewriting.
+The main issue is that internal interfaces may not be preserved during rewriting, for example, if the right-hand side of the rewrite rule has more hierarchy than the left-hand side: it affects both the eventual e-hypergraph and the possible pushout complement.
+
+% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
+% Note that the outermost interfaces are preserved.
+
+% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
+% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
+
+% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
+% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
+% The entry in the middle of the span depicts the outermost interfaces which are preserved.
+% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
+% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
+% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
+% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
+
+To account for such changes in the interfaces we put some additional constraints on DPOI rewriting for the case of $\catname{EHyp_{\Sigma}}$.
 Consider a DPOI diagram with extra constraints below which encodes
- the notion that cospan 
- $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$
- rewrites into a cospan 
- $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$ via a rewrite rule $\langle l, r \rangle$,
- where
- $l := x_{ext} \xrightarrow{} l_i \xrightarrow{} L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
- and
- $r := x_{ext} \xrightarrow{} r_i \xrightarrow{} L \xleftarrow{} r_o \xleftarrow{} y_{ext}$.
- Note there are arrows from $x_{ext} + y_{ext} \to l_i + l_o$ (respectively, $r_i + r_o$) from the corresponding cospans which are not shown in the diagram.
- Similarly for the arrows from $n_{ext} + m_{ext}$.
-
+the notion that e-hypergraph $G$ rewrites into $H$ with external interfaces $n_{ext}+m_{ext}$.
+Note the new interfaces and how we use the same trick to go from cospan $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$ to $G \xleftarrow{} g_{i} + g_{o} \xleftarrow{} n_{ext} + m_{ext}$.
+The original DPO square is depicted in blue.
 % \[\begin{tikzcd}
@@
 
-The parts of the original DPOI square is shown in blue.
-The extra bits simply account for more complicated interfaces.
-For example, the left-hand side $L$ of a rewrite rule
- $\langle L, R \rangle$ which originally was some cospan
- $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
- now is a cospan
- $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
-from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
-The top half of this diagram says that the outermost interfaces
- should coincide for $L$ and $R$.
-The bottom half puts the same restrictions on $G$ and $H$.
-For the pushout to exist we impose our assumptions~\ref{pushout:assumptions} on the correspoding arrow.
-
-We also adjust the definition of a boundary complement.
-\begin{definition}[Down-closed subgraph]
-We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ \emph{down-closed} if for all $e \in E_{\mathcal{H}}$,   every element of $(e]$ is also in $\mathcal{H}$.
-\end{definition}
+% The parts of the original DPOI square is shown in blue.
+% For example, the left-hand side $L$ of a rewrite rule
+%  $\langle L, R \rangle$ which originally was some cospan
+%  $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
+%  now is a cospan
+%  $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
+% from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
+% The top half of this diagram says that the outermost interfaces
+%  should coincide for $L$ and $R$.
+% The bottom half puts the same restrictions on $G$ and $H$.
+% For the pushout to exist we impose our assumptions~\ref{pushout:assumptions} on the correspoding arrow.
+
+We then also adjust the definition of a boundary complement.
 
@@
 % https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
+
+For monogamous extended cospans $x_{ext} \xrightarrow{} l_{i} L \xleftarrow{} l_{o} \xleftarrow{} y_{ext}$ and $n_{ext} \xrightarrow{} g_{i} \xrightarrow{} G \xleftarrow{} g_{o} \xleftarrow{} m_{ext}$ and mono $f : L \to G$, a pushout complement $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ as depicted below
 \[
 \scalebox{0.75}{
+
     \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
@@
 
-Consider the diagram above where the marked square is a pushout
- and $x_{ext} + y_{ext}$ and $n_{ext} + m_{ext}$ are 
- the \textit{outermost} interfaces for the corresponding 
- cospans.
-We say that $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ is 
- a \textit{boundary} complement if $[i_c, i_j]$ is mono 
+is a \textit{boundary} complement if $[i_c, i_j]$ is mono 
  and there exist $n_c : n_{ext} \to C$ and $m_c : m_{ext} \to C$ 
@@
  $c_o = g_o \setminus (l_o \setminus y_{ext})$.
+
+ ALEKSEI: ARE MORPHISMS $c_i + y_{ext} \to C$ AND THE SYMMETRIC ONE OBVIOUS?
 \end{definition}
@@
 \begin{remark}
-Note that cospan $n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}$ is not well-typed.
-The cospan's input interface contains nodes that previously were a part of the output interface.
+Note that cospan $n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}$ is not necessarily well-typed.
+The cospan's input internal interface may contain nodes that previously were a part of the output internal interface and similarly for output internal interface.
 \end{remark}
 
-\begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-    The boundary complement in~\ref{def:boundary_new} when exists is unique.
-\end{proposition}
+\begin{remark}
+    The uniqueness and existence properties of such pushout complement are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
+\end{remark}
+
+% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
+%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
+% \end{proposition}
+
+\begin{definition}[Down-closed subgraph]
+    We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ \emph{down-closed} if for all $e \in E_{\mathcal{H}}$,   every element of $(e]$ is also in $\mathcal{H}$.
+\end{definition}
 
@@
         \item $m(L)$ is a convex down-closed subgraph of $G$
-        % \item \update{I've removed the requirement that consistency should be reflected because well typedness prohibits otherwise formed matches.}
     \end{enumerate}
@@
     Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n_{ext} + m_{ext}$ 
-    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}}^{+} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $L \xleftarrow{} l_i + l_o \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} r_i + r_o \xrightarrow{} R$ with the outermost interfaces $x_{ext} + y_{ext}$ in $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ such that the diagram below commutes and its marked squares are pushouts
+    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}}^{+} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $L \xleftarrow{} l_i + l_o \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} r_i + r_o \xrightarrow{} R$ with the outermost interfaces $x_{ext} + y_{ext}$ in the set of rewrite rules $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ such that the diagram below commutes and its marked squares are pushouts
 
@@
         \item $m : L \to G$ is a convex down-closed match;
-        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $x_{ext} + y_{ext}$
-        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $x_{ext} + y_{ext}$
         \item $x_{ext} + y_{ext} \to C \to G$ is a boundary complement in the leftmost pushout.
@@
 
-\begin{remark}
-The conditions 2 and 3 are needed to guarantee the existence of the pushout complement.
-\end{remark}
-
-The whole interfaces of the resulting e-hypergraph $H$ from the cospan
+The internal interfaces of the resulting e-hypergraph $H$ from the cospan
  $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$
@@
 if 
-\[
-    G \xleftarrow{\langle a_2,b_2 \rangle} g_i + g_{o} \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext} \Rrightarrow H \xleftarrow{\langle a_2,b_2 \rangle} h_i + h_{o} \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext}
-\]
+\begin{align*}    
+    G \xleftarrow{\langle a_2,b_2 \rangle} g_i + g_{o}& \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext} \\ &\Rrightarrow^{+}\\ H \xleftarrow{\langle a_2,b_2 \rangle} h_i + h_{o}& \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext}
+\end{align*}
+
 under a rewrite rule
@@
 \begin{enumerate}
+\label{pushout:assumptions}
     \item $Z$ is a \textit{descrete} e-hypergraph.
@@
 \end{proposition}
+
+\begin{definition}[Definition 3.16~\cite{Frobenius}]
+Morphisms $K \xrightarrow{i} L \xrightarrow{m} G$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every node of its source and target is either (i) not in the image of $m$ or (ii) in the image of $i ; m$.
+They satisfy the no-identification condition if any two nodes merged by $m$ are in the image of $i$.
+\end{definition}
+
+\begin{proposition}
+\label{prop:complement_existence}
+    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
+    % https://q.uiver.app/#q=WzAsNSxbMCwwLCJYIl0sWzIsMCwiWiJdLFsyLDIsIkMiXSxbMCwyLCJQXntcXHVyY29ybmVyfSJdLFsyLDUsIlxcYnVsbGV0Il0sWzEsMCwiaSIsMl0sWzEsMl0sWzIsM10sWzAsMywibSIsMl1d
+\[\begin{tikzcd}
+	X && Z \\
+	\\
+	{P^{\urcorner}} && C \\
+	\\
+	\\
+	&& \bullet
+	\arrow["i"', from=1-3, to=1-1]
+	\arrow[from=1-3, to=3-3]
+	\arrow[from=3-3, to=3-1]
+	\arrow["m"', from=1-1, to=3-1]
+\end{tikzcd}\]
+    Such that $m,i$ are monos and $Z$ is discrete. Then, pushout complement $Z \to C \to G$ exists if
+    \begin{enumerate}
+        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
+        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
+        \item The no-dangling condition holds.
+    \end{enumerate}
+\end{proposition}
+\begin{proof}
+    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
+    Then, no-dangling and no-identification conditions are necessary for pushout complement to exists~\cite{dpo}.
+    Points (1)-(2) above guarantee that after removing from $G$ everything in $L$ that has no pre-image in $Z$ the arrow $Z \to C$ will satisfy the assumptions~\ref{pushout:assumptions}.
+\end{proof}
 

### 2024-01-05 E-Hypergraphs-LICS24.tex
Removed subscript c and superscript \mu for rels

@@
 \]
-We will denote the unique immediate predecessor $e$ of $x$ by writing $e <_c^{\mu} x$.
+We will denote the unique immediate predecessor $e$ of $x$ by writing $e <^{\mu} x$.
 We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
@@
 	\item Each $x$ has at most one immediate predecessor.
-	\item if $v \in s(e)$ then $e' <_c^{\mu} e$ iff $e' <_c^{\mu} v$.
+	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
 \end{enumerate}
@@
         \item The relation is an equivalence relation
-        \item If $v \in s(e)$ then $v \consistency^{\mu} e$
-        \item If $v \in t(e)$ then $v \consistency^{\mu} e$
+        \item If $v \in s(e)$ then $v \consistency e$
+        \item If $v \in t(e)$ then $v \consistency e$
         \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
@@
     % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 \consistency_{\mathcal{F}}^{\mu} x_2$ implies $\phi(x_1) \consistency_{\mathcal{G}}^{\mu} \phi(x_2)$. 
+        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 \consistency_{\mathcal{F}} x_2$ implies $\phi(x_1) \consistency_{\mathcal{G}} \phi(x_2)$. 
 %    Or, written functionally, when $x \in E_{\mathcal{F}}$
@@
 \end{definition}
-Condition $(3)$ on the child relations says that $<_c^{\mu}$ respects connectivity.
+Condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity.
 
@@
 
-\begin{theorem}[Existence of pushouts in $\catname{Hyp_{\Sigma}}$]
-
-    Pushouts in $\catname{Hyp_{\Sigma}}$ always exist.
-\end{theorem}
-
-\begin{proof}
-    Consider a diagram below
-    % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAC0QBfU9TXfIRQBGclVqMWbABrdeIDNjwEiZYePrNWiEAE05fJYKKj11TVJ0AdK3ABmAJzoBjYNIDUursBvYAtgAEAEpcBgr8ykLIAEyk0RqS2iAAitziMFAA5vBEoI4QfkhkIDgQSKISWmx2YfmFiMWlSLGVliCZINQMdABGMAwAChHGOg5YmQAWOLUOBeXUTYgAzOaJbL6OLsBYAPrC3r5YgSEzc4gtiyutSRtOrrvRB7ZHwaE8ebP1F2WIACzUfTAUCQAFolsULEkAFY7aKdEDdPqDYYqUbjKaneoVRb-ECA4HLCFrHQw4SYpBXHGrKo6JhpLhAA
-
-\[
-        \begin{tikzcd}
-        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{i_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-        Y \arrow[r, "\sfrac{i_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                        &                                                                  & Q
-        \end{tikzcd}
-\]
-
-Pushout of two hypergraphs $X = \{V_{X}, E_{X}, s_{X}, t_{X}\}$ and $Y = \{V_{Y}, E_{Y}, s_{Y}, t_{Y}\}$ (we will omit labelling functions here since they are irrelevant) along $Z$ is computed in two steps.
-First, a coproduct of $X$ and $Y$ is taken which is 
-\[
-    X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X + Y}, t_{X+Y}\}
-\]
-where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $\langle s'_{X}, s'_{Y} \rangle : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};i_{1,V}^{*}$ and $s'_{Y} = s_{Y};i_{2,V}^{*}$ where $i_1,i_2$ are corresponding coproduct injections, similarly for $t_{X+Y}$.
-Recall that each arrow in $\catname{Hyp_{\Sigma}}$ is a pair of arrows for edges and nodes and we will further use subscripts $V$ and $E$ when referring to these functions.
-Then, consider relations 
-\[
-    S_{V} = \{(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \; \exists z \in V_{Z} \; .
-\; x_i = f_{V};i_{1,V}(z) \text{ and } y_j = g_{V};i_{2,V}(z) \text{where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\}
-\]
-\[
-    S_{E} = \{(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; | \; \exists z \in E_{Z} \; .
-\; x_i = f_{E};i_{1,E}(z) \text{ and } y_j = g_{E};i_{2,E}(z) \text{where $x_i \in E_{X}$ and $y_j \in E_{Y}$ }\}
-\]
-and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-\begin{lemma}
-    $x R y$ if and only if either $x = y$ or there exists a sequence $(a_1,\ldots,a_{n-1},{a_n})$ such that $x = a_1$ and $y = a_{n}$ such that $a_k S a_{k+1}$ or $a_{k + 1} S a_{k}$ for $k < n$.
-\end{lemma}
-
-A quotient of $X+Y$ by these relation is 
-\[
-    \sfrac{X+Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X+Y}}{\sim (R_{V},R_{E})}\}
-\]
-We will then refer to $\sim (R_{V},R_{E})$ just as $\sim$, i.e. writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and the exact relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-In particular, there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-Let $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*} = s_{X+Y};[-]^{*}$ and then $\sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}$.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has in image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-We need to check that if $t_1 \sim t_2$ then $\sfrac{s_{X+Y}}{\sim}([t_1]) = \sfrac{s_{X+Y}}{\sim}([t_2])$. 
-There are three options
-\begin{enumerate}
-    \item $t_1 = e_{y}^{1}$ and $t_2 = e_{y}^{2}$
-    \item $t_1 = e_{x}^{1}$ and $t_2 = e_{x}^{2}$
-    \item $t_1 = e_{x}$ and $t_2 = e_{y}$ 
-\end{enumerate}
-We will prove the last point (all the others are analogous).
-
-According to the lemma above $e_{x} \sim e_{y}$ gives us two options.
-\begin{enumerate}
-    \item Either $e_{x} = e_{y}$
-    \item or there is a sequence $w = (a_1, \ldots, a_{n})$ such that $a_1 = e_{x}$ and $a_{n} = e_{y}$.
-\end{enumerate}
-In the first case the equality holds on the nose.
-We then prove the second case by induction on the length of $w$.
-\begin{itemize}
-    \item In the case $|w| = 2$, $e_{x} S e_{y}$ (or $e_{y} S e_{x}$) which means that there exists $z$ in $E_{Z}$ such that $f_{E};i_{1,E}(z) = e_{x}$ and $g_{E};i_{2,E}(z) = e_{y}$.
-            Recall that because $f$ and $g$ are homomorphisms, the following equalities hold
-            \[
-            s_{Z};f_{V}^{*};i_{1,V}^{*}(z) = f_{E};i_{1,E};s_{X+Y}(z) = s_{X+Y}(e_{x})
-            \]
-            \[
-            s_{Z};g_{V}^{*};i_{2,V}^{*}(z) = g_{E};i_{2,E};s_{X+Y}(z) = s_{X+Y}(e_{y})
-            \]
-            Then,
-            \[
-            \sfrac{s_{X+Y}}{\sim}([e_{x}]) = [s_{X+Y}(e_{x})]^{*} = [s_{Z};f_{V}^{*};i_{1,V}^{*}(z)]^{*}
-            \]
-            and
-            \[
-            \sfrac{s_{X+Y}}{\sim}([e_{y}]) = [s_{X+Y}(e_{y})]^{*} = [s_{Z};g_{V}^{*};i_{2,V}^{*}(z)]^{*}
-            \]
-            Then, by noting that $f_{V};i_{1,V}(z) = g_{V};i_{2,V}(z)$ for all $z$ in $V_{Z}$ (because $\sfrac{V_{X} + V_{Y}}{\sim}$ is so defined) we have $[s_{X+Y};(e_{x})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
-    \item Suppose if $e_{x} S e_{y}$ by a chain of length $n$ or less then $[s_{X+Y}(e_{x})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
-    \item Now suppose $e_{x} S e_{y}$ by a chain $w = (a_1, \ldots, a_{n+1})$ and a subchain $(a_1, \ldots, a_{n})$ satisfies the inductive hypothesis and hence $s_{X+Y};s_{1}^{*}(e_{x}) = s_{X+Y};s_{1}^{*}(a_{n})$.
-            Note that $S_{E}, S_{V}$ relate only elements with the pre-image in $E_{X}$ (respectively, $V_{X}$) with elements with the pre-image in $E_{Y}$ (respectively, $V_{Y}$). 
-            Elements with the pre-image in the same set are not related.
-            Then we have $a_n S a_{n+1} = e_{y}$ and need to show that $[s_{X+Y}(a_{n+1})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
-            This follows by exactly the same argument as the base case as $a_{n+1}$ is necessarily in $E_{X}$.
-            And we finally have $[s_{X+Y}(e_{x})]^{*} = [s_{X+Y}(a_{n+1})]^{*} = [s_{X+Y}(e_{y})]^{*}$.
-\end{itemize}
-
-Next 
-\[
-    \sfrac{i_{1,V}}{\sim} : V_{X} \to \sfrac{V_{X} + V_{Y}}{\sim} = [i_{1,V}]
-\]
-and
-\[
-\sfrac{i_{1,E}}{\sim} : E_{X} \to \sfrac{E_{X} + E_{Y}}{\sim} = [i_{1,E}]
-\]
-similarly for $\sfrac{i_{2,E}}{\sim}$ and $\sfrac{i_{2,V}}{\sim}$
-We need to show
-\begin{enumerate}
-    \item $s_{X};(\sfrac{i_{1,V}}{\sim})^{*}(e) = \sfrac{i_{1,E}}{\sim};\sfrac{s_{X+Y}}{\sim}(e)$
-    \item $t_{X};(\sfrac{i_{1,V}}{\sim})^{*}(e) = \sfrac{i_{1,E}}{\sim};\sfrac{t_{X+Y}}{\sim}(e)$
-\end{enumerate}
-and likewise for $i_{2}$.
-Let's check the first one for which we first need to check that $([-]_{E},[-]_{V})$ is a homomorphisms, i.e.
-\[
-[s_{X+Y}(e)]^{*} = \sfrac{s_{X+Y}}{\sim}([e]) = [s_{X+Y}(e)]^{*}
-\]
-hence, $(\sfrac{i_{1,V}}{\sim}, \sfrac{i_{1,E}}{\sim})$ is a homomorphism because $(i_{1,V},i_{1,E})$ and $([-]_{E},[-]_{V})$ are.
-
-Now, the square obviously commutes.
-To show that $\sfrac{X + Y}{\sim}$ is a pushout we need to argue that for each $Q$ and morphisms $j_1 = (j_{1,V}, j_{1,E})$ and $j_2 = (j_{2,V}, j_{2,E})$ such that the above square commutes there is a unique $u = (u_{V}, u_{E})$ such that $j_1 = \sfrac{i_{1}}{\sim} ; u$ and $j_2 = \sfrac{i_2}{\sim} ; u$.
-We will define such $u$ as $u_{E}([i_{1,E}(e_{x})]) = j_{1,E}(e_{x})$ and $u_{E}([i_{2,E}(e_{y})]) = j_{2,E}(e_y)$ (similarly for $u_{V}$) and claim that it is a well-defined function by showing that if $t_1 \sim t_2$ then $u_{E}(t_1) = u_{E}(t_2)$.
-There are three options analogoues to the proof above, so we will show the base case for $t_1 = i_{1,E}(e_{x})$ and $t_2 = i_{2,E}(e_{y})$.
-$t_1 \sim t_2$ implies that there exists at least one $z$ in $E_{Z}$ such that $i_{1,E}(e_{x}) = f_{E};i_{1,E}(z)$ (or $e_{x} = f_{E}(z)$) and $[i_{1,E}(e_{x})] = \sfrac{i_{1,V}}{\sim}(e_x)$, and $e_y = g_{E}(z)$ $[in_{2,E}(e_{y})] = \sfrac{i_{2,V}}{\sim}(e_y)$.
-By applying $j_{1,E}$ to the first we get $j_{1,E}(f_{E}(z)) = j_{1,E}(e_x)$ and similarly for $j_{2,E}$: $j_{2,E}(g_{E}(z)) = j_{2,E}(e_y)$ and $j_{1,E}(e_{x}) = j_{2,E}(e_{y})$ by commutativity.
-Hence, $u_{E}([in_{1,E}(e_x)]) = u_{E}([in_{2,E}(e_y)])$.
-
-Next, $u$ is a homomorphism because $j_1$ and $j_2$ are.
-Finally, $u$ is unique.
-Assume, that there exists $w : \sfrac{X+Y}{\sim R} \to Q$ for the same $j_1$ and $j_2$.
-In order for them to commute $w_{E}([i_{1,E}(x)]) = j_1(x)$ and $w([in_{2,E}(y)]) = j_2(y)$ (similarly for $w_{V}$) which is the same as $u_{E}$.
-That is, $j_1$ and $j_2$ determine $u$ uniquely.
-\end{proof}
-
 \begin{remark}
-    Both relations can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, i.e. on the coproduct of nodes and edges.
-    To make things well-typed, we will use corresponding coproduct injections $i_{V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $i_{E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either node or edge into these functions.
-    For example, an immediate successor of a node $x$ can be written functionally as $<_{c,\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x))$.
+    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, i.e. on the coproduct of nodes and edges.
+    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either node or edge into these functions.
+    For example, an immediate successor of a node $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
 \end{remark}
@@
 
-            \item \[
-                \phi_{E}(<_{c,\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x))) = <_{c,\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+            \item When $[x) \not = \varnothing$, \[
+                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
                 \] when $x$ is a node,
@@
                 \[
-                \phi_{E}(<_{c,\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x))) = <_{c,\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))  
+                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
                 \] when $x$ is an edge.          
-                We do not require such an equation when $[x) = \varnothing$.
-
-            % \item
-            % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
-
                 \item
@@
             \[
-                \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
+                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
                 \subseteq
-                \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
-            \]
-            , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
+            \]
+            , where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
             And, when $x \in V_{\mathcal{F}}$
             \[
-                \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
+                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
                 \subseteq
-                \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
+                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
             \]
@@
 \begin{proof}
-    Recall the diagram below.
+    Consider the diagram below.
     \[
         \begin{tikzcd}
-        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{i_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-        Y \arrow[r, "\sfrac{i_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
+        Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
                                                                         &                                                                  & Q
@@
     \]
-
-    once again omitting the labelling function and where source and target functions are as before.
-    And consider the same relations $S_{V},S_{E}$ and $R_{V}, R_{E}$.
+    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{c,X+Y}, \consistency_{X+Y}$.
+    We will omit labels as they are irrelevant to pushout construction.
+    Consider relations
+    \begin{align*}    
+    S_{V} &= \{\\
+          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z) \text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
+          &\}
+    \end{align*}
+    \begin{align*}
+    S_{E} &= \{\\
+    &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
+    &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z) \text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$ }
+    &\}
+    \end{align*}
+and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.
 
     We then quotient the sets of edges and nodes in $X + Y$ by $R_{V},R_{E}$
-    \[
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})}, \sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \#_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{c,\sfrac{X + Y}{\sim (R_{V},R_{E})}} \}
-    \]
-    where quotienting on source and target functions is taken as in $Hyp_{\Sigma}$ and we will further omit writing the relations explicitly.
+    \begin{align*}
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
+            &\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \#_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{c,\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+        & \}    
+    \end{align*}
+    We will then refer to $\sim (R_{V},R_{E})$ just as $\sim$, i.e. writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and the exact relation will be clear from the context.
+We have 
+\[
+    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
+\]
+and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
+We then define $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*} = s_{X+Y};[-]^{*}$ and then $\sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}$.
+There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
+We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
+Likewise, $t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}$.
+
+So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{Frobenius} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined an $[-]$ is a homomorphism with respect to sources and targets.
+
 Recall that $<_{c}$ is essentially a transitive clousure of $<_{c}^{\mu}$ and hence we can first define 

### 2024-01-05 E-Hypergraphs-LICS24.tex
Section 3 tidy up.

--- 2024-01-05 E-Hypergraphs-LICS24.tex
@@
 \begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category where every object is the n-fold tensor product of some distinguished object $A$.
-In this case, we will denote the n-fold tensor product of the distinguished object by simply $n$, so that $n \otimes m = n+m$.
+A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
 A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
@@
 This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{Frobenius2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous directed acyclic}),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
+Subsequent work \cite{Frobenius2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
 
@@
 \subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-
-The correspondence of the previous section further extends to include the equations of a monoidal theory: the equations of a theory are considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
+The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
+The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
 DPO rewriting is a standard technique for formalising graph rewriting categorically. 
@@
 
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $L \rightsquigarrow R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $L$ within $\mathcal{G}$,  remove it,  and then plug $R$ into the hole.
-In DPO rewriting,  such a rewrite rule is formalized as a span $L \xleftarrow{} K \xrightarrow{} R$ in some category $\mathcal{C}$ with pushouts,  where $K$ is some invariant that should be maintained during rewriting.
-CHRIS: "WITHIN G",  BUT WE ARE IN AN ARBITRARY CATEGORY: NOT NECESSARILY GRAPHS?
-
+The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
+In DPO rewriting,  such a rewrite rule is formalized as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
 A typical DPO square looks as follows,  where the two marked squares are pushouts. 
@@
 \begin{tikzcd}
-L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
-G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H
+\mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+\mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner} \mathcal H
 \end{tikzcd}
 \]
-Morphisms $K \to C \to G$ are called a \textit{pushout complement} and morphism $m$ is called a \textit{match},  defined when the corresponding pushout complement exists.
-When $f,g$ are expected to be monomorphisms, constructing $C$ can be intuitively understood as finding an occurrence of $L$ in $G$ and removing from this occurrence everything that has no pre-image in $K$ and then inserting in there everything from $R$ that has no pre-image in $K$ -- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $G$ is rewritten into $H$ via the rewrite rule $L \xleftarrow{} K \xrightarrow{} R$.
-
-To express the rewriting of (hyper-)graphs with interfaces, e.g.
-the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout with \textit{interfaces} rewriting (DPOI).
-
-\begin{definition}[DPOI rewriting]
-Fix a category $\mathbb{C}$ with pushouts.
-Given morphisms $G \xleftarrow{} J$ and $H \xleftarrow{} J$ in $\mathbb{C}$, $G$ rewrites into $H$ with interface $J$ if there exists a rewrite rule $L \xleftarrow{} K \xrightarrow{} R$ and object $C$ with suitable morphisms in $\mathbb{C}$ such that the diagram below commutes and two marked squares are pushouts.
+
+The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement} -- the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
+The final graph $\mathcal H$ is the recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ -- hence the two pushouts: one for deletion and one for insertion.
+In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
+
+To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI).
+In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
+By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
+Thus we reach the following variant of DPO rewriting.
+\begin{definition}[DPOI rewriting]\label{def:dpoi}
+Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
 \[
 \begin{tikzcd}
-L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
-G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H \\
-                  & J \arrow[lu] \arrow[u] \arrow[ru]          &              
+\mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+\mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
+                  & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
 \end{tikzcd}
 \]
-
-\end{definition}
-When $\mathbb{C} = \catname{Hyp_{\Sigma}}$, the rewrite rules become cospans of hypergraphs with matching interfaces: $i \xrightarrow{} L \xleftarrow{} j$, $i \xrightarrow{} R \xleftarrow{} j$.
-By noting that a cospan $0 \xrightarrow{} L \xleftarrow{[f_{i},f_{j}]} i + j$ encodes the same data as a cospan $i \xrightarrow{f_{i}} L \xleftarrow{f_{j}} j$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $L \xleftarrow{} K \xrightarrow{} R$ where $K \cong i + j$.
-Likewise, for the target cospan $n \xrightarrow{} G \xleftarrow{} m$  $J \cong n + m$.
-% However, this notion of DPO does not account for hypergraphs with interfaces, i.e.
-those from $\MdaCospans$ where $l,r$ from $\langle l, r \rangle$ are cospans.
-To support this we first assume that we only need to maintain the interfaces of $i \xrightarrow{} L \xleftarrow{} j$ and $i \xrightarrow{} R \xleftarrow{} j$ during rewriting, then we can take $K \cong i + j$, i.e.
-take $K$ to be the coproduct of $i,j$ (remark 3.14~\cite{Frobenius}).
-What remains is to account for the interfaces in $G$, $C$, and $H$, which brings us to the following modified DPOI square.
-$J \cong n + m$ where $n$ and $m$ are the input and output interfaces of $G$, $C$, and $H$ respectively.
-The only difference from DPO square above is that we now explicitly require that the initial object $G$, intermediate object $C$ and final object $H$ all share the same interfaces.
-
-CHRIS: DEFINE DPOI FORMALLY,  IN DEFINITION ENVIRONMENT? PERHAPS COMBINE WITH BOUNDARY COMPLEMENT BELOW,  TWO FIGURES ON ONE LINE?
-
-To mimic rewrites in $\textbf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textbf{PROP}(\Sigma)$.
-CHRIS: "LEAD TO"?
-BE PRECISE!
-:)
-
+\end{definition}
+We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acylcicity.  
 \begin{definition}[Boundary complement]
 \label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
-and mono $f : L \to G$, a pushout complement as depicted below
+For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
+and mono $f : \mathcal L \to \mathcal G$, a pushout complement as depicted in the square below
 
@@
 \[\begin{tikzcd}
-	L && {i+j} \\
+	\mathcal L && {i+j} \\
 	\\
-	{G^{\urcorner}} && {L^{\bot}} \\
+	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
 	\\
@@
 	\arrow["f"', from=1-1, to=3-1]
-	\arrow["{a = [a_1,a_2]}"', from=1-3, to=1-1]
-	\arrow["{c = [c_1,c_2]}", from=1-3, to=3-3]
+	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
+	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
 	\arrow["g", from=3-3, to=3-1]
 \end{tikzcd}\]
-
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to L^{\bot}$ and $d_2 : m \to L^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} L^{\bot} \xleftarrow{[d_2,c_1]} m + i
+is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
+\[
+    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
 \]
@@
 \end{definition}
-
-\begin{remark}
-    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e.
-morphisms in $\MdaCospans{\Sigma}$, is a part of the above definition.
-\end{remark}
-
-Intuitively, such a pushout complement requires that inputs get only glued to outputs and vice-verse in the two pushout squares.
-
-\begin{definition}[Convex subgraphs]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ \emph{convex} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-\end{definition}
+%\begin{remark}
+%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
+%\end{remark}
+Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
+
 \begin{definition}[Convex match]
-
-We call a match $m : L \to G$ convex if it is mono and its image is a convex subgraph of $G$.
-
-\end{definition}
-
-\begin{definition}
-Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold
-
-\begin{itemize}
-    \item $m : L \to G$ is a convex match;
-    \item $i + j \to C \to G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-
-\end{definition}
-
-ALEKSEI: THE WORD `CONVEXLY' ABOVE MEANS THAT THE REWRITING IS FORMULATED FOR MOMOGAMOUS MORPHISMS. SHALL WE STICK TO THIS FORMULATION?
+We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
+We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
+
+\end{definition}
+
+\begin{definition}[Convex DPOI rewriting]
+Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
+Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
+and the following conditions hold
+\begin{itemize}
+    \item $m : L \to \mathcal G$ is a convex match;
+    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
+\end{itemize}
+\end{definition}
+
+CHRIS: WANT TO STATE THE CORRESPONDENCE THEOREM WITH SMTS HERE? NEED TO DEFINE $\lceil-\rceil$ AND INTERPRETATION? Hmm
+\begin{theorem}[Theorem 35~\cite{Frobenius2}]
+For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
+\[
+	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
+\]
+where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ . 
+\end{theorem}
 
@@
 In other words, homomorphism should preserve immediate predecessors and the conflict relation.
-CHRIS: CHECK I DIDNT INTRODUCE MISTAKES ABOVE.
 SHOULD WE USE $< ; \phi$ NOTATION TO MATCH OTHER HYPERGRAPH DEFINITION?

### 2024-01-05 E-Hypergraphs-LICS24.tex
Update E-Hypergraphs-LICS24.tex

--- 2024-01-05 E-Hypergraphs-LICS24.tex
@@
 \begin{tikzcd}
-L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
-G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H
+\mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+\mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner} \mathcal H
 \end{tikzcd}
 \]
-Morphisms $K \to C \to G$ are called a \textit{pushout complement} and morphism $m$ is called a \textit{match},  defined when the corresponding pushout complement exists.
-When $f,g$ are expected to be monomorphisms, constructing $C$ can be intuitively understood as finding an occurrence of $L$ in $G$ and removing from this occurrence everything that has no pre-image in $K$ and then inserting in there everything from $R$ that has no pre-image in $K$ -- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $G$ is rewritten into $H$ via the rewrite rule $L \xleftarrow{} K \xrightarrow{} R$.
-
-To express the rewriting of (hyper-)graphs with interfaces, e.g.
-the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout with \textit{interfaces} rewriting (DPOI).
-
-\begin{definition}[DPOI rewriting]
-Fix a category $\mathbb{C}$ with pushouts.
-Given morphisms $G \xleftarrow{} J$ and $H \xleftarrow{} J$ in $\mathbb{C}$, $G$ rewrites into $H$ with interface $J$ if there exists a rewrite rule $L \xleftarrow{} K \xrightarrow{} R$ and object $C$ with suitable morphisms in $\mathbb{C}$ such that the diagram below commutes and two marked squares are pushouts.
+
+The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement} -- the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
+The final graph $\mathcal H$ is the recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ -- hence the two pushouts: one for deletion and one for insertion.
+In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
+
+To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI).
+In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
+By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
+Thus we reach the following variant of DPO rewriting.
+\begin{definition}[DPOI rewriting]\label{def:dpoi}
+Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
 \[
 \begin{tikzcd}
-L \arrow[d, "m"'] & K \arrow[d] \arrow[l, "f"'] \arrow[r, "g"] & R \arrow[d]   \\
-G^{\urcorner}     & C \arrow[l] \arrow[r]                      & ^{\ulcorner}H \\
-                  & J \arrow[lu] \arrow[u] \arrow[ru]          &              
+\mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+\mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
+                  & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
 \end{tikzcd}
 \]
-
-\end{definition}
-When $\mathbb{C} = \catname{Hyp_{\Sigma}}$, the rewrite rules become cospans of hypergraphs with matching interfaces: $i \xrightarrow{} L \xleftarrow{} j$, $i \xrightarrow{} R \xleftarrow{} j$.
-By noting that a cospan $0 \xrightarrow{} L \xleftarrow{[f_{i},f_{j}]} i + j$ encodes the same data as a cospan $i \xrightarrow{f_{i}} L \xleftarrow{f_{j}} j$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $L \xleftarrow{} K \xrightarrow{} R$ where $K \cong i + j$.
-Likewise, for the target cospan $n \xrightarrow{} G \xleftarrow{} m$,  $J \cong n + m$.
-% However, this notion of DPO does not account for hypergraphs with interfaces, i.e.
-those from $\MdaCospans$ where $l,r$ from $\langle l, r \rangle$ are cospans.
-To support this we first assume that we only need to maintain the interfaces of $i \xrightarrow{} L \xleftarrow{} j$ and $i \xrightarrow{} R \xleftarrow{} j$ during rewriting, then we can take $K \cong i + j$, i.e.
-take $K$ to be the coproduct of $i,j$ (remark 3.14~\cite{Frobenius}).
-What remains is to account for the interfaces in $G$, $C$, and $H$, which brings us to the following modified DPOI square.
-$J \cong n + m$ where $n$ and $m$ are the input and output interfaces of $G$, $C$, and $H$ respectively.
-The only difference from DPO square above is that we now explicitly require that the initial object $G$, intermediate object $C$ and final object $H$ all share the same interfaces.
-
-CHRIS: DEFINE DPOI FORMALLY,  IN DEFINITION ENVIRONMENT? PERHAPS COMBINE WITH BOUNDARY COMPLEMENT BELOW,  TWO FIGURES ON ONE LINE?
-
-To mimic rewrites in $\textbf{PROP}(\Sigma)$ with the help of DPOI rewriting in $\catname{Hyp_{\Sigma}}$, we need to put some restrictions both on pushout complement and on matching as not all of them correspond to valid rewrites in $\textbf{PROP}(\Sigma)$.
-CHRIS: "LEAD TO"?
-BE PRECISE!
-:)
+\end{definition}
+We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acylcicity.
 
@@
 \label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} G \xleftarrow{b_2} m$
-and mono $f : L \to G$, a pushout complement as depicted below
+For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
+and mono $f : \mathcal L \to \mathcal G$, a pushout complement as depicted in the square below
 
@@
 \end{definition}
-
-\begin{remark}
-    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e.
-morphisms in $\MdaCospans{\Sigma}$, is a part of the above definition.
-\end{remark}
-
-Intuitively, such a pushout complement requires that inputs get only glued to outputs and vice-verse in the two pushout squares.
-
-\begin{definition}[Convex subgraphs]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ \emph{convex} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-\end{definition}
+%\begin{remark}
+%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
+%\end{remark}
+Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
+
 \begin{definition}[Convex match]
-
-We call a match $m : L \to G$ convex if it is mono and its image is a convex subgraph of $G$.
-
-\end{definition}
-
-\begin{definition}
-Given $G \xleftarrow{} n+m$ and $H \xleftarrow{} n + m$ in $\catname{Hyp_{\Sigma}}$, $G$ rewrites convexly into $H$ with interface $n + m$ --- notation $(G \xleftarrow{} n + m ) \Rrightarrow_{\mathcal{R}}  (H \xleftarrow{} n + m )$ --- if there exist rule $L \xleftarrow{} i + j \xrightarrow{} R$ in the set of rewrite rules $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} C \xleftarrow{} n+m$ such that the DPOI diagram above commutes and its marked squares are pushouts and the following conditions hold
-
-\begin{itemize}
-    \item $m : L \to G$ is a convex match;
-    \item $i + j \to C \to G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-
-\end{definition}
-
-By resorting to convex matching and boundary complement we virtually require that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-
-ALEKSEI: THE WORD `CONVEXLY' ABOVE MEANS THAT THE REWRITING IS FORMULATED FOR MOMOGAMOUS MORPHISMS. SHALL WE STICK TO THIS FORMULATION?
-
+We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
+We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
+
+\end{definition}
+
+\begin{definition}[Convex DPOI rewriting]
+Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
+Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
+and the following conditions hold
+\begin{itemize}
+    \item $m : L \to \mathcal G$ is a convex match;
+    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
+\end{itemize}
+\end{definition}
+%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
+
+CHRIS: WANT TO STATE THE CORRESPONDENCE THEOREM WITH SMTS HERE? NEED TO DEFINE $\lceil-\rceil$ AND INTERPRETATION? Hmm
+\begin{theorem}[Theorem 35~\cite{Frobenius2}]
+For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
+\[
+	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
+\]
+where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
+\end{theorem}
 \section{E-Hypergraphs}

### 2024-01-05 E-Hypergraphs-LICS24.tex
e-hypergraphs section

@@
 \begin{enumerate}
-    \item $(s_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;s_{\mathcal{G}})(e)$
-    \item $(t_{\mathcal{F}}\, ;\phi_V^*)(e) = (\phi_E\, ;t_{\mathcal{G}})(e)$
-    \item $l_{\mathcal{F}}(e) = \phi_E;l_{\mathcal{G}}(e)$
+    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
+    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
+    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
 \end{enumerate}
@@
 In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces \cite{Frobenius2}]
+\begin{definition}[Category of MDA Hypergraphs with Interfaces]
 \label{def:monogamy_hyp}
@@
 
-CHRIS: EXPLAIN THE CONDITIONS,  HOW WE WILL INTERPRET THE STRING DIAGRAMS.  
-REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ morphisms!
-
-We follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ as a particular subset of plain hypergraphs with interfaces.
-First we define an e-hypergraph,  which supports a hierarchical notion of "e-box" with distinct components.
+WHAT DOES (3) FOR CONSISTENCY MEAN? 
+In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in our semilattice enriched setting.
+First we define an \textit{e-hypergraph},  which supports a hierarchical notion of "e-box" with distinct components.
 Then,  we extend this notion with interfaces via a generalization of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying a certain directed acyclic \textit{monogamy} condition.
-
-In the following,  we elide the obvious inclusions of $V$ and $E$ into $V+E$. 
-In other words, homomorphism should preserve immediate predecessors and the conflict relation.
-SHOULD WE USE $< ; \phi$ NOTATION TO MATCH OTHER HYPERGRAPH DEFINITION?
-WHAT DOES (4) FOR CONSISTENCY MEAN?
-
+Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
+An example diagram representing our e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
+
+Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
+This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
+
+In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$. 
 \begin{definition}[Category of E-hypergraphs]
@@
 We define the following additional notation.
-For $x \in V + E$,  the set of \emph{predecessors} and \emph{successors} of $x$,  respectively:
+For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
 \[
@@
 \]
-We will denote the unique immediate predecessor $e$ of $x$ by writing $e <^{\mu} x$.
+We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
 We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions
+We require that the child relation satisfies the following conditions:
 \begin{enumerate}
-	\item $[x)$ contains exclusively edges such that $l(e) = \bot$.
-	\item Each $x$ has at most one immediate predecessor.
+	\item each parent set contains exclusively hierarchical edges;
+	\item Each $x$ has at most one immediate parent;
 	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
 \end{enumerate}
-The consistency relation must satisfy the following properties. 
+The consistency equivalence relation must satisfy the following conditions:
     \begin{enumerate}
         \label{def:consistency_properties}
-        \item The relation is an equivalence relation
-        \item If $v \in s(e)$ then $v \consistency e$
-        \item If $v \in t(e)$ then $v \consistency e$
-        \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
+        \item If $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$;
+        \item The relation is defined only for edges $x$ which are not top-level.
     \end{enumerate}
@@
 \end{definition}
-Condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity.
+Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of our string diagrams.  
+These edges can be designated as "parents" to sub-hypergraphs,  which interpret the various components of the e-boxes.
+These components are distinguished by the consistency relation, which partitions the sub-hypergraph below the hierarchical edge.
+Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
+The conditions on homomorphisms require to preserve immediate parents and the conflict relation.
 
@@
 
-\begin{theorem}[Theorem~\ref{th:existence_of_pushouts} restatement]
-Under certain conditions pushouts in $\catname{EHyp_{\Sigma}}$ exist.   
-\end{theorem}
-
 \begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate predecessor of all $x$ whose immediate predecessor is in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
+When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent is in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
 \end{remark}
@@
 
-\begin{definition}[Category of MDA e-hypergraphs with interfaces]
+\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
 We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
@@
 \label{def:monogamy_ehyp}
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of MDA-extended-cospans. 
-\end{definition}
-CHRIS: COULD PUT CONDITION (5) IN THE PLAIN DEFINITION OF EXTENDED COSPANS? WE ALREADY ASK FOR EXTERNAL INTERFACES TO BE TOP-LEVEL.  ALSO,  (2) IS ALREADY COVERED IN THAT DEFINITION.
-
-\begin{definition}[Well-typedness]
-Consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
+
+Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
 The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
@@
 We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-We 
-\end{definition}
-CHRIS: CATEGORY OF SUCH COSPANS? COMBINE WITH DEFINITION ABOVE? $\WellTypedMdaEcospans$
+
+We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
+\end{definition}
+CHRIS: COULD PUT CONDITION (5) IN THE PLAIN DEFINITION OF EXTENDED COSPANS? WE ALREADY ASK FOR EXTERNAL INTERFACES TO BE TOP-LEVEL.  ALSO,  (2) IS ALREADY COVERED IN THAT DEFINITION.
+
+Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
+This is beacuse the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
+Instead,  we will treat them via rewriting,  as we do with equations of the signature.
+This is because the semilattice equations involve (un)sharing,  and so we wish to maintain distinct e-hypergraphs expressing more or less shared representations of morphisms.
 
@@
 
+CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
 As extended cospans have a more general notion of interface we need to make some adjustments to DPOI rewriting.

### 2024-01-05 E-Hypergraphs-LICS24.tex
Section 5

@@
 As extended cospans have a more general notion of interface we need to make some adjustments to DPOI rewriting.
-The main issue is that internal interfaces may not be preserved during rewriting, for example, if the right-hand side of the rewrite rule has more hierarchy than the left-hand side: it affects both the eventual e-hypergraph and the possible pushout complement.
-
+The first thing is that we need to account for internal interfaces after all.
+The main issue is then that internal interfaces may not be preserved during rewriting.
+To illustrate this, consider the following rewrite rule
+\[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
+such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
+\[
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} m' \xleftarrow{} m
+\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$.
+When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, e.g.
+by computing the pushout complement, the internal interface of $\mathcal{G}$ should be modified as the verticies of $\mathcal{L}$ corresponding to strict internal interface of $\mathcal{L}$ (and hence to strict internal interface of $\mathcal{G}$) should be removed from $\mathcal{G}$ and therefore from $\mathcal{G}$'s internal interface.
+Then, when gluing $\mathcal{R}$ into the created context, we also introduce strict internal interfaces of $\mathcal{R}$ into this context.
 % The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
@@
 
-To account for such changes in the interfaces we put some additional constraints on DPOI rewriting for the case of $\catname{EHyp_{\Sigma}}$.
+To account for such changes in the interfaces we put some additional structure on DPOI rewriting for the case of $\catname{EHyp({\Sigma})}$.
 Consider a DPOI diagram with extra constraints below which encodes
 the notion that e-hypergraph $G$ rewrites into $H$ with external interfaces $n_{ext}+m_{ext}$.
-Note the new interfaces and how we use the same trick to go from cospan $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$ to $G \xleftarrow{} g_{i} + g_{o} \xleftarrow{} n_{ext} + m_{ext}$.
+Note the addition of internal interfaces and how we use the same trick to go from cospan $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$ to $G \xleftarrow{} g_{i} + g_{o} \xleftarrow{} n_{ext} + m_{ext}$.
 The original DPO square is depicted in blue.
@@
 
-% The parts of the original DPOI square is shown in blue.
-% For example, the left-hand side $L$ of a rewrite rule
-%  $\langle L, R \rangle$ which originally was some cospan
-%  $i \xrightarrow{} L \xleftarrow{} j$ in $\MdaCospans$ 
-%  now is a cospan
-%  $x_{ext} \xrightarrow{} l_i \to L \xleftarrow{} l_o \xleftarrow{} y_{ext}$
-% from $\WellTypedMdaEcospans$ where $l_i, l_o$ are whole input and output interfaces respectively.
-% The top half of this diagram says that the outermost interfaces
-%  should coincide for $L$ and $R$.
-% The bottom half puts the same restrictions on $G$ and $H$.
-% For the pushout to exist we impose our assumptions~\ref{pushout:assumptions} on the correspoding arrow.
-
-We then also adjust the definition of a boundary complement.
+Then, we explicitly show how the internal interfaces of the pushout complement are constructed.
 
@@
 
-For monogamous extended cospans $x_{ext} \xrightarrow{} l_{i} L \xleftarrow{} l_{o} \xleftarrow{} y_{ext}$ and $n_{ext} \xrightarrow{} g_{i} \xrightarrow{} G \xleftarrow{} g_{o} \xleftarrow{} m_{ext}$ and mono $f : L \to G$, a pushout complement $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ as depicted below
+For monogamous extended cospans $x_{ext} \xrightarrow{} l_{i} \xrightarrow{} L \xleftarrow{} l_{o} \xleftarrow{} y_{ext}$ and $n_{ext} \xrightarrow{} g_{i} \xrightarrow{} G \xleftarrow{} g_{o} \xleftarrow{} m_{ext}$ and mono $f : L \to G$, a pushout complement $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ as depicted below
 \[
@@
 \begin{definition}[Convex down-closed match] 
-    We call $m : L \to G$ in $\catname{EHyp_{\Sigma}}$ a convex down-closed match if
+    We call $m : L \to G$ in $\catname{EHyp(\Sigma)}$ a convex down-closed match if
 
@@
 \end{definition}
-\
+
+The whole process of rewriting including the construction of the internal interfaces of the resulting e-hypergraph is summarised in the following definition.
 
@@
 \begin{align*}    
-    G \xleftarrow{\langle a_2,b_2 \rangle} g_i + g_{o}& \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext} \\ &\Rrightarrow^{+}\\ H \xleftarrow{\langle a_2,b_2 \rangle} h_i + h_{o}& \xleftarrow{\langle a_1, b_1 \rangle} n_{ext} + m_{ext}
+    G \xleftarrow{[ a_2,b_2 ]} g_i + g_{o}& \xleftarrow{[a_1, b_1]} n_{ext} + m_{ext} \\ &\Rrightarrow^{+}\\ H \xleftarrow{[a_2,b_2]} h_i + h_{o}& \xleftarrow{[a_1, b_1]} n_{ext} + m_{ext}
 \end{align*}
@@
 \[
-    \langle l_{i} + l_{o} \xrightarrow{\langle a_2, b_2 \rangle} L \xleftarrow{\langle a_1, b_1 \rangle} x_{ext} + y_{ext},
-    r_{i} + r_{o} \xrightarrow{\langle a_2, b_2 \rangle} R \xleftarrow{\langle a_1, b_1 \rangle} x_{ext} + y_{ext} 
+    \langle l_{i} + l_{o} \xrightarrow{[a_2, b_2]} L \xleftarrow{[a_1, b_1]} x_{ext} + y_{ext},
+    r_{i} + r_{o} \xrightarrow{[a_2, b_2]} R \xleftarrow{[a_1, b_1]} x_{ext} + y_{ext} 
     \rangle
@@
 
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{Frobenius} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined an $[-]$ is a homomorphism with respect to sources and targets.
-
-Recall that $<_{c}$ is essentially a transitive clousure of $<_{c}^{\mu}$ and hence we can first define 
-    \[<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
+So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{Frobenius} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets.
+
+Recall that $<$ is essentially a transitive clousure of $<^{\mu}$ and hence we can first define 
+    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
     \] and 
     \[
-        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
+        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
     \] where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.

### 2024-01-06 E-Hypergraphs-LICS24.tex
changed arxiv in bib wherever possible

@@
 Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{noauthor_09083347_nodate,joyal_geometry_1991, mellies_functorial_2006}.
+makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
 We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
@@
 
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{Frobenius, Frobenius2}.
+In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
 
@@
 
-In ~\cite{Frobenius},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
+In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
 Hypergraphs generalize graphs by allowing edges to have multiple sources and targets. 
@@
 This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{Frobenius2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
+Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
 
@@
 
-Following~\cite{Frobenius2} morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
+Following~\cite{bonchi_string_2022-2} morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
 \begin{definition}[Hypergraphs with interfaces]
@@
 Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  In fact,  due to 
-\cite{Frobenius2},  we have the following theorem which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26~\cite{Frobenius2}]
+\cite{bonchi_string_2022-2},  we have the following theorem which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
+\begin{theorem}[Corollary 26~\cite{bonchi_string_2022-2}]
     $\textbf{PROP}(\Sigma) \cong \MdaCospans$
@@
 In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{Frobenius}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
+By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
 Thus we reach the following variant of DPO rewriting.
@@
 CHRIS: WANT TO STATE THE CORRESPONDENCE THEOREM WITH SMTS HERE? NEED TO DEFINE $\lceil-\rceil$ AND INTERPRETATION? Hmm
-\begin{theorem}[Theorem 35~\cite{Frobenius2}]
+\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
 For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
@@
 %
-%\begin{theorem}[Theorem 35~\cite{Frobenius2}]
+%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
 %
@@
 \begin{remark}
-    Using functional notation we can reformulate the notion of homomorphism of two e-hypergraphs in the following way.
-    \begin{definition}{E-hypergraphs homomorphism}
+    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
+    \begin{definition}
         \label{def:e-homo-2}    
-        A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
 
@@
           &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z) \text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
+          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
+          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
           &\}
@@
     &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z) \text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$ }
+    &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
+    &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
     &\}
@@
 
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{Frobenius} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets.
+So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets.
 
@@
 
-\begin{definition}[Definition 3.16~\cite{Frobenius}]
+\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
 Morphisms $K \xrightarrow{i} L \xrightarrow{m} G$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every node of its source and target is either (i) not in the image of $m$ or (ii) in the image of $i ; m$.

### 2024-01-08 E-Hypergraphs-LICS24.tex
intro

@@
 CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
-DAN: I SUGGEST TO CITE HASEGAWA IN FURTHER WORK, FOR GENERALISING THE THEORY TO TRACED MC.
 DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.
@@
 
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.\footnote{Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid.} More precisely our hom-sets will be given the structure of a semilattice.
+We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
+Note, in particular,  that semilattices do not require a unit for $+$.
+This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid.
+More precisely our hom-sets will be given the structure of a semilattice.
 Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).

### 2024-01-08 E-Hypergraphs-LICS24.tex
sections 1-4

@@
 Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid.
+This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
 More precisely our hom-sets will be given the structure of a semilattice.
@@
 Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}. 
+DAN: I THINK THE STRING DIAGRAMS EQUATIONS SHOULD BE ACCOMPANIED BY THE TERM EQUATIONS. THE NUMBER OF EQUATIONS IN THE DEFINITIONS AND IN THE FIGURE IS NOT THE SAME, WHICH IS A BIT CONCERNING.
 
@@
 \end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$.
+We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$. DAN: WHAT ARE THE MORPHISMS?
 \end{definition}
@@
 Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{align*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) &\otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} &+ \mathcal{G'} \xleftarrow{g+g'} B + B' ~ ,
-\end{align*}
+\begin{multline*}
+    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+\end{multline*}
 and its unit by the initial object of $\mathbb{C}$.
@@
 
-Following~\cite{bonchi_string_2022-2} morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces. 
+DAN: I AM NOT SURE 'DISCRETE HYPERGRAPHS' ARE DEFINED?
+Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
 \begin{definition}[Hypergraphs with interfaces]
@@
 
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement} -- the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
 When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is the recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ -- hence the two pushouts: one for deletion and one for insertion.
+The final graph $\mathcal H$ is the recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$--hence the two pushouts: one for deletion and one for insertion.
 In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
@@
 CHRIS: WANT TO STATE THE CORRESPONDENCE THEOREM WITH SMTS HERE? NEED TO DEFINE $\lceil-\rceil$ AND INTERPRETATION? Hmm
+DAN: I THINK THIS PRELIMINARY SECTION IS VERY LONG.
+UNLESS WE USE THIS THEOREM IN THE PAPER LATER CAN WE JUST STATE IT INFORMALLY, WITHOUT INTRODUCING THIS NOTATION?
+IF THE NOTATION IS USED, IT MUST BE DEFINED.
 \begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
@@
 WHAT DOES (3) FOR CONSISTENCY MEAN? 
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in our semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of "e-box" with distinct components.
+DAN: THERE IS LITERATURE ON HIERARCHICAL HYPERGRAPHS (DETLEF PLUMP AND UGO MONTANARI I THINK HAVE PAPERS) SOME OF IT IS REFERENCED FROM MY PREVIOUS PAPERS ON HIERARCHICAL HYPERGRAPHS.
+MAKE SURE YOU CITE THEM, THESE PEOPLE MAY GET ANNOYED.
+WE DID NOT INVENT HIERARCHICAL HYPERGRAPHS.
+ALSO MY PAPER WITH KOKO WHICH INTRODUCES THEM ONLY COMBINATORIALLY.
+In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
+First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
 Then,  we extend this notion with interfaces via a generalization of the cospan construction.
 Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing our e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
+An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
 
@@
 In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$. 
+DAND: THIS IS VERY LONG... SHOULD IT BE BROKEN DOWN INTO SEVERAL DEFINITIONS? 
 \begin{definition}[Category of E-hypergraphs]
@@
 \end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of our string diagrams.  
+Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
 These edges can be designated as "parents" to sub-hypergraphs,  which interpret the various components of the e-boxes.
@@
 
-When modelling our enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
+When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
 Thus,  we introduce an \textit{extended} cospan construction.
@@
 \begin{definition}[Category of e-hypergraphs with interfaces]
-    We will build the category of \textit{e-hypergraphs with interfaces},  denoted $\Ecospans$ as follows.
-Its objects are discrete e-hypergraphs,  and the set of morphisms $\Ecospans(n,m)$ are isomorphism classes of \textit{extended cospans},  which are defined as follows:
+    The category of \textit{e-hypergraphs with interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with homsets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
     \[
@@
 We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We will sometimes confuse $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and we will sometimes confuse $n, n', m$ and $m'$ with their images in $\mathcal{G}$. 
+We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n, n', m$ and $m'$ with their images in $\mathcal{G}$. 
 Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
@@
 Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve (un)sharing,  and so we wish to maintain distinct e-hypergraphs expressing more or less shared representations of morphisms.
+This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
+string diagrams for Cartesian categories.
 
@@
 Having described a translation of e-graphs into our formalism, we developed a combinatorial representation of morphisms in our free category which factor out the structural SMC equations, but remain sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding  rewrite theory proved delicate, and -- together with its proof of completeness -- constitute the main technical contributions of the paper.
+The design of such a representation and its corresponding  rewrite theory proved delicate, and---together with its proof of completeness---constitute the main technical contributions of the paper.
 Finally, we considered an analogue of e-rewriting in our setting, pushing towards the practical considerations necessary for an implementation of our ideas.

### 2024-01-08 E-Hypergraphs-LICS24.tex
fixes to S4

@@
 The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial represetation.
+The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial represetation.
 
@@
 
-\begin{figure*}\label{fig:e-graph-example}
+\begin{figure*}
 \[
@@
 \caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
+\label{fig:e-graph-example}
 \end{figure*}
@@
 
-\begin{definition}[Semilattice-enriched PROP]
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
+\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
+A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
 \begin{align*}
@@
 Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
+The first four equations are those displayed in Definition \ref{def:def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
+We overload the binary notation $+$ for our $n$-ary notation.
 We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
 Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-DAN: I THINK THE STRING DIAGRAMS EQUATIONS SHOULD BE ACCOMPANIED BY THE TERM EQUATIONS. THE NUMBER OF EQUATIONS IN THE DEFINITIONS AND IN THE FIGURE IS NOT THE SAME, WHICH IS A BIT CONCERNING.
 
@@
 The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
+We call a hypergraph \textit{discrete} if its set of edges is empty.
 A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
@@
 \end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$. DAN: WHAT ARE THE MORPHISMS?
+We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
 \end{definition}
@@
 
-DAN: I AM NOT SURE 'DISCRETE HYPERGRAPHS' ARE DEFINED?
 Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
@@
 
-WHAT DOES (3) FOR CONSISTENCY MEAN? 
-DAN: THERE IS LITERATURE ON HIERARCHICAL HYPERGRAPHS (DETLEF PLUMP AND UGO MONTANARI I THINK HAVE PAPERS) SOME OF IT IS REFERENCED FROM MY PREVIOUS PAPERS ON HIERARCHICAL HYPERGRAPHS.
-MAKE SURE YOU CITE THEM, THESE PEOPLE MAY GET ANNOYED.
-WE DID NOT INVENT HIERARCHICAL HYPERGRAPHS.
-ALSO MY PAPER WITH KOKO WHICH INTRODUCES THEM ONLY COMBINATORIALLY.
+Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
 In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
@@
 In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-DAND: THIS IS VERY LONG... SHOULD IT BE BROKEN DOWN INTO SEVERAL DEFINITIONS? 
-\begin{definition}[Category of E-hypergraphs]
+
+\begin{definition}[E-hypergraphs]
 \label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}; and an equivalence relation $\consistency$ on $V + E$ called the \textit{consistency relation}.
+An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
 We define the following additional notation.
@@
 \end{enumerate}
-The consistency equivalence relation must satisfy the following conditions:
-    \begin{enumerate}
+We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
+We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
+The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
 \label{def:consistency_properties}
-        \item If $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$;
-        \item The relation is defined only for edges $x$ which are not top-level.
-    \end{enumerate}
-
-A homomorphism $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+\end{definition}
+Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
+These edges can be designated as "parents" to sub-hypergraphs,  which interpret the various components of the e-boxes.
+These components are distinguished by the consistency relation, which partitions the sub-hypergraph below the hierarchical edge.
+Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
+\begin{definition}[E-hypergraph homomorphism]
+An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
 \begin{enumerate}
     \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $[v)$ is not empty then 
+    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
 	\[
@@
         \] 
-        and for $e \in E_{\mathcal{F}}$,  if $[e)$ is not empty then 
+        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
         \[
@@
 If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as arrows.  
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the sub-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-The conditions on homomorphisms require to preserve immediate parents and the conflict relation.
+\end{definition}
+The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
+\begin{definition}[Category of e-hypergraphs]
+The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
+\end{definition}
 
@@
     \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices.
+    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and that that vertices in the strictly internal interface (defined below) are not top-level.
 
@@
 \[
-    h = \langle a;p_1, g|_{Y' \setminus Y_{ext}};p_2 \rangle
-\]
-
-\[
-    k = \langle b;p_2, f|_{Y \setminus Y_{ext}};p_1 \rangle
+    h = [ a;p_1, ~g';p_2 ]
+\qquad
+    k = [ b;p_2, ~f';p_1 ]
 \]
@@
 % and then the disjoint union is just a coproduct. 
+where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
 The identity of composition is given by the obvious extended cospan of identities. 
@@
 \end{definition}
-CHRIS: UPDATE DIAGRAMS TO USE MY NEW NOTATION,  TAKE LESS VERTICAL SPACE.
-FIND CLEANER WAY TO WRITE NEW FEET.
-USE $\iota$ FOR INJECTIONS,  WHAT IS $\langle \rangle$?
-Include example e-hypergraph-with-interfaces vs strings here?
+CHRIS: UPDATE DIAGRAMS TO USE MY NEW NOTATION,  TAKE LESS VERTICAL SPACE.  FIND CLEANER WAY TO WRITE NEW FEET.  USE $\iota$ FOR INJECTIONS
 
 As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
+Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
+This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
+This also means that we omit the monomorphism condition from the following definition.
+We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
+
 % https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
@@
         \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-        \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
+       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
         \item The in-degree and out-degree of every vertex is at most 1; 
@@
         \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        \item Vertices in the strictly internal interface are not top-level. 
         % the item below is not assumed by the definition of a cospan of e-hypergraphs
@@
 \end{definition}
-CHRIS: COULD PUT CONDITION (5) IN THE PLAIN DEFINITION OF EXTENDED COSPANS? WE ALREADY ASK FOR EXTERNAL INTERFACES TO BE TOP-LEVEL.  ALSO,  (2) IS ALREADY COVERED IN THAT DEFINITION.
-
 Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
@@
 The corresponding combinatorial representation would build on the recent work \cite{ghica_hierarchical_2023, alvarez-picallo_rewriting_2022, alvarez-picallo_functorial_2021}.
-String diagrams for these various examples are included in Figure \ref{fig:applications}% hierarchical
+String diagrams for these various examples are included in Figure \ref{fig:applications}.
+There is also the question of developing a corresponding term calculus \cite{Heijltjes:FMCII}.
+% hierarchical
 

### 2024-01-08 E-Hypergraphs-LICS24.tex
small updates to S5

@@
 CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
-As extended cospans have a more general notion of interface we need to make some adjustments to DPOI rewriting.
-The first thing is that we need to account for internal interfaces after all.
-The main issue is then that internal interfaces may not be preserved during rewriting.
+Because extended cospans have a more general notion of interface,  we need to make some adjustments to DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics}.
+The first thing is that we need to account for is the existence of \textit{internal} interfaces,  and in particular that internal interfaces are not necessarily preserved during rewriting.
 To illustrate this, consider the following rewrite rule
@@
 When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, e.g.
-by computing the pushout complement, the internal interface of $\mathcal{G}$ should be modified as the verticies of $\mathcal{L}$ corresponding to strict internal interface of $\mathcal{L}$ (and hence to strict internal interface of $\mathcal{G}$) should be removed from $\mathcal{G}$ and therefore from $\mathcal{G}$'s internal interface.
+by computing the pushout complement, the internal interface of $\mathcal{G}$ should be modified.
+This is because the verticies of $\mathcal{L}$ corresponding to strict internal interface of $\mathcal{L}$ (and hence to strict internal interface of $\mathcal{G}$) should be removed from $\mathcal{G}$ and therefore from $\mathcal{G}$'s internal interface.
 Then, when gluing $\mathcal{R}$ into the created context, we also introduce strict internal interfaces of $\mathcal{R}$ into this context.
@@
 
-To account for such changes in the interfaces we put some additional structure on DPOI rewriting for the case of $\catname{EHyp({\Sigma})}$.
+To account for such changes in the interfaces we require additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$.
 Consider a DPOI diagram with extra constraints below which encodes
@@
 
-For monogamous extended cospans $x_{ext} \xrightarrow{} l_{i} \xrightarrow{} L \xleftarrow{} l_{o} \xleftarrow{} y_{ext}$ and $n_{ext} \xrightarrow{} g_{i} \xrightarrow{} G \xleftarrow{} g_{o} \xleftarrow{} m_{ext}$ and mono $f : L \to G$, a pushout complement $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ as depicted below
+For monogamous extended cospans $x_{ext} \xrightarrow{} l_{i} \xrightarrow{} L \xleftarrow{} l_{o} \xleftarrow{} y_{ext}$ and $n_{ext} \xrightarrow{} g_{i} \xrightarrow{} G \xleftarrow{} g_{o} \xleftarrow{} m_{ext}$ and mono $f : L \to G$, a pushout complement $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ as depicted in the square below
 \[

### 2024-01-08 E-Hypergraphs-LICS24.tex
update to S6.

@@
 \newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{WMEHypI(\Sigma)}}}
+\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
 % conflict
@@
 \end{figure}\label{fig:egraph-strings}
+CHRIS: DONT USE PHI/PSI FOR DIAGRAMS, USE F,G
 
@@
 CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
+Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}. 
 DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.
@@
 \begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $f: n \to m$,  with \textit{arity} $n$ and \textit{coarity} $m$.
+A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{coarity} $m$.
 %A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
@@
 \end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
+The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
 
@@
 \begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $\phi,\psi: n \to m$ of $\textbf{PROP}(\Sigma)$.
+A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
 A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
@@
 
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $f \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, as well as $f+g$, quotiented by the axioms of an enriched symmetric monoidal category.
+The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
 
@@
 \end{theorem}
-\section{E-Hypergraphs}
+\section{E-Hypergraphs}\label{sec:e-hypergraphs}
 
@@
 The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp_{\Sigma}}$,  analogously to Definition~\ref{def:cspd}.  
+$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
 \end{definition}
@@
 \end{definition}
+Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
+
 Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
@@
 
-\begin{remark}
 Note that cospan $n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}$ is not necessarily well-typed.
 The cospan's input internal interface may contain nodes that previously were a part of the output internal interface and similarly for output internal interface.
-\end{remark}
-
-\begin{remark}
-    The uniqueness and existence properties of such pushout complement are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-\end{remark}
+
+    The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
 
@@
 
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-
 \begin{figure}
@@
 
-Note that the necessary pushouts for $\mathcal{S}$ exist. 
-\begin{definition}
-We denote by $\WellTypedMdaEcospans/\mathcal{S}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	\phi \sim \psi \quad \text{if} \quad \phi \Rightarrow^*_{[\mathcal{S}]} \psi
-\]
-where $\mathcal{S}$ denotes set of semi-lattice equations the free functor $\llbracket- \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans$ induced by the interpretation functor sending $\Sigma$ to itself.
-CHRIS: NEEDS WORK.
-\end{definition}
-
-\begin{proposition}
-The category $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ a semilattice-enriched PROP. 
+Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
+Recall further that (typed) $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$, respectively.
+Write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
+We wish to specify a set of rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
+In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the action of the free functor from $\textbf{PROP}(\Sigma)$ into $\WellTypedMdaEcospans$ induced by the interpretation of $\Sigma$ as itself (detailed in \cite{bonchi_string_2022-2}).
+\begin{definition}[Interpretation function] 
+%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
+Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by (mutual) induction on $\Sigma$-terms by 
+\begin{align*}
+	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket \phi ; \psi \rrbracket_{n,m} &= \llbracket \phi \rrbracket^+_{n,m} ; \llbracket \psi \rrbracket_{n,m} \\
+	\llbracket 1 \rrbracket_{n,m} &= 1 & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
+	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
+	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
+\end{align*}
+where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.  We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context. 
+\end{definition}
+
+Let the function $\llbracket - \rrbracket^+$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms). 
+% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
+\begin{definition}[Quotient by rewrites]  
+Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
+\[
+	f \sim g \quad \text{if} \quad f \Rightarrow^*_{\llbracket \mathcal E \rrbracket^+} g ~ . 
+\]
+\end{definition}
+In particular,  let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations to be applied to $\Sigma^+$-terms to recover $\textbf{PROP}^+(\Sigma)$.
+This quotient in fact turns the interpretation function into a proper functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which can further be quotiented to account for extra equations arising in symmetric monoidal theories.
+Note that not all pushouts in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
+%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
+Thus we have the following result. 
+\begin{proposition}[Soundness]
+The category $\WellTypedMdaEcospans/{\mathcal{S}}$ a semilattice-enriched PROP. 
 \end{proposition}
-\begin{proof}
-The symmetric monoidal structure is inherited from coproduct in $\textbf{EHyp}(\Sigma)$.
-It suffices to check that the symmetry and the tensor product of two well-typed e-hypergraphs is again a well-typed e-hypergraph.
-The formal join $+$ of two morphisms of $\WellTypedMdaEcospans$ is defined in the Figure~\ref{fig:A+B}. 
-Etc.
-\end{proof}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by rewrites,  for which we've checked the pushout exists.
-
-\begin{proposition}
-    For each cospan $A$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ there is a weak normal form such that 
-\[
-	A =_{\mathcal{S}} A_1 + \ldots + A_n
-\]
- such that each $A_i$ contains no hierarchical edges.
-    \begin{proof}
-    By applying the rules 1-6 in $\mathcal{S}$ (the ones from Figure~\ref{fig:string-equations}) we either remove 1 hierarchical edge (rules 5-6) or do not change the number of them, but expand the edge onto other edges.
-    We will have a redex for such a rule until there is only one hierarchical edge at the outermost level (and only one outermost edge) with no other hierarchical edges within it.
-    \end{proof}
+Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
+
+The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system which results in a cospan satisfying the statement below.
+\begin{proposition}[Weak normal form]
+For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
+\[
+	f = f_1 + \ldots + f_n
+\]
+ such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
 \end{proposition}
 
-\begin{theorem}
+This normal form can be used to prove the following completeness theorem. 
+\begin{theorem}[Completeness]
 We have the following equivalence of categories
 \[
-	PROP^{+}(\Sigma) \cong \WellTypedMdaEcospans~.
+	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
 \]
@@
 \begin{proof}
-Given by the free functor induced by the interpretation of $\Sigma$ as itself. 
-Normal form argument. 
+The equivalence is given by the functor $\llbracket - \rrbracket$.
+First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
+In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA hypergraphs which is given by the restriction of $\llbracket - \rrbracket$ to $$\textbf{SMT}(\Sigma, \mathcal E)$ which is a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$.
+
+To see that the functor is faithful,  consider $f$ and $g$ of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ such that $\llbracket f \rrbracket = \llrbracket g \rrbracket$.
+We aim to show that $f = g$.
+First,  find weak normal forms of $\llbracket f \rrbracket$ and $\llrbracket g \rrbracket$,  namely $\phi_1 + \ldots + \phi_ni$ and $\psi_1 + \ldots + \psi_n$.
+We now need a further observation that if $\llbracket f \rrbracket \Rrightarrow_{{\llbracket \mathcal{S,E} \rrbracket}^+} \llrbracket g \rrbracket$,  then there exist rewrites $\phi_i \Rrightarrow_{\mathcal E} \psi_i$.
+Thus,  by $(*)$ we have that $\llbracket \phi_i \rrbracket^{-1} = \llbracket \psi_i \rrbracket^{-1}$.
+Noting that $f = \llbracket \phi_1 \rrbracket^{-1} + \ldots + \llbracket \phi_n \rrbracket^{-1}$ and $g = \llbracket \psi_1 \rrbracket^{-1} + \ldots + \llrbacket \psi_n \rrbracket^{-1}.
+CHRIS: THIS NEEDS FIXING.
+
+To see that the functor is full,  consider a cospan $f$ in $\WellTypedMdaEcospans$.
+We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ is a plain MDA hypergraph.
+Thus,  by $(*)$ we can find morphisms $g_i$ such that $\llbracket g_i \rrbracket = f_i$.
+Thus,  we have a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
 \end{proof}
@@
 Unfolding the definitions,  we can spell out the previous theorem in terms of DPO-rewriting. 
-\begin{corollary}
-For any $\phi, \psi \in \textbf{PROP}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	\phi = \psi \quad \iff \quad \llbracket\phi \rrbracket \Rrightarrow_{\mathcal{S}, \mathcal{E}} \llbracket\psi\rrbracket~ . 
+\begin{corollary}[Completeness of rewriting]
+For any $\phi, \psi \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
+\[
+	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^*_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
 \]
  \end{corollary}
-CHRIS: INCORPORATING EXTRA EQUATIONS OUGHT TO BE IMMEDIATE? 
 %

### 2024-01-08 E-Hypergraphs-LICS24.tex
fix

@@
 First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA hypergraphs which is given by the restriction of $\llbracket - \rrbracket$ to $$\textbf{SMT}(\Sigma, \mathcal E)$ which is a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$.
-
-To see that the functor is faithful,  consider $f$ and $g$ of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ such that $\llbracket f \rrbracket = \llrbracket g \rrbracket$.
+In each case,  the result follows by a normal form argument in combination with relying on the equivalence $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA hypergraphs which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ which is a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$.
+
+%To see that the functor is faithful,  consider $f$ and $g$ of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ such that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
 We aim to show that $f = g$.
-First,  find weak normal forms of $\llbracket f \rrbracket$ and $\llrbracket g \rrbracket$,  namely $\phi_1 + \ldots + \phi_ni$ and $\psi_1 + \ldots + \psi_n$.
-We now need a further observation that if $\llbracket f \rrbracket \Rrightarrow_{{\llbracket \mathcal{S,E} \rrbracket}^+} \llrbracket g \rrbracket$,  then there exist rewrites $\phi_i \Rrightarrow_{\mathcal E} \psi_i$.
+First,  find weak normal forms of $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$,  namely $\phi_1 + \ldots + \phi_ni$ and $\psi_1 + \ldots + \psi_n$.
+We now need a further observation that if $\llbracket f \rrbracket \Rrightarrow_{{\llbracket \mathcal{S,E} \rrbracket}^+} \llbracket g \rrbracket$,  then there exist rewrites $\phi_i \Rrightarrow_{\mathcal E} \psi_i$.
 Thus,  by $(*)$ we have that $\llbracket \phi_i \rrbracket^{-1} = \llbracket \psi_i \rrbracket^{-1}$.
 Noting that $f = \llbracket \phi_1 \rrbracket^{-1} + \ldots + \llbracket \phi_n \rrbracket^{-1}$ and $g = \llbracket \psi_1 \rrbracket^{-1} + \ldots + \llrbacket \psi_n \rrbracket^{-1}.
-CHRIS: THIS NEEDS FIXING.
-
+%CHRIS: THIS NEEDS FIXING. 
+%
 To see that the functor is full,  consider a cospan $f$ in $\WellTypedMdaEcospans$.

### 2024-01-09 E-Hypergraphs-LICS24.tex
S6 small update

@@
 
+CHRIS: INTRO.
+
 \begin{figure}
@@
 Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that (typed) $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$, respectively.
-Write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-We wish to specify a set of rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the action of the free functor from $\textbf{PROP}(\Sigma)$ into $\WellTypedMdaEcospans$ induced by the interpretation of $\Sigma$ as itself (detailed in \cite{bonchi_string_2022-2}).
+Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
+We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
+
+We wish to specify a set of DPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
+In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
 \begin{definition}[Interpretation function] 
 %Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by (mutual) induction on $\Sigma$-terms by 
+Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
 \begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket \phi ; \psi \rrbracket_{n,m} &= \llbracket \phi \rrbracket^+_{n,m} ; \llbracket \psi \rrbracket_{n,m} \\
+	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket \phi ; \psi \rrbracket_{n,m} &= \llbracket \phi \rrbracket_{n,m} ; \llbracket \psi \rrbracket_{n,m} \\
 	\llbracket 1 \rrbracket_{n,m} &= 1 & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
@@
 
-Let the function $\llbracket - \rrbracket^+$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms). 
+Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms). 
 % \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
+CHRIS: COMMENT THAT WE QUOTIENT TO PHRASE CATEGORICALLY OUR RESULTS. 
 \begin{definition}[Quotient by rewrites]  
@@
 \[
-	f \sim g \quad \text{if} \quad f \Rightarrow^*_{\llbracket \mathcal E \rrbracket^+} g ~ . 
-\]
-\end{definition}
-In particular,  let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations to be applied to $\Sigma^+$-terms to recover $\textbf{PROP}^+(\Sigma)$.
-This quotient in fact turns the interpretation function into a proper functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which can further be quotiented to account for extra equations arising in symmetric monoidal theories.
-Note that not all pushouts in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
+	f \sim g \quad \text{if} \quad f \Rightarrow^*_{\llbracket \mathcal E \rrbracket} g ~ . 
+\]
+\end{definition}
+In particular,  let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
+This quotient in fact turns the interpretation function into a proper functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which can further be quotiented to account for extra equations $\mathcal E$ arising in symmetric monoidal theories,  as below.
+\[
+	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
+\]
+
+Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
 %Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
@@
 \begin{proposition}[Soundness]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ a semilattice-enriched PROP. 
+The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
 \end{proposition}
+It immediately follows that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is a semilattice-enriched PROP,  and $\llbracket - \rrbracket_{\mathcal E}$ is the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
 Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
 
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system which results in a cospan satisfying the statement below.
+The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
 \begin{proposition}[Weak normal form]
@@
 
-This normal form can be used to prove the following completeness theorem. 
+This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
 \begin{theorem}[Completeness]
@@
 First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA hypergraphs which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ which is a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$.
+In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ as a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ (note that cospans embed in extended cospans when the internal interface is isomorphic to the external interface).
 
@@
 Noting that $f = \llbracket \phi_1 \rrbracket^{-1} + \ldots + \llbracket \phi_n \rrbracket^{-1}$ and $g = \llbracket \psi_1 \rrbracket^{-1} + \ldots + \llrbacket \psi_n \rrbracket^{-1}.
-%CHRIS: THIS NEEDS FIXING. 
+CHRIS: FAITHFULNESS. 
 %
 To see that the functor is full,  consider a cospan $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ is a plain MDA hypergraph.
+We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ is a plain MDA cospan.
 Thus,  by $(*)$ we can find morphisms $g_i$ such that $\llbracket g_i \rrbracket = f_i$.
@@
 \begin{corollary}[Completeness of rewriting]
-For any $\phi, \psi \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
+For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
 \[
@@
 %\end{theorem}
+CHRIS: CONCLUDING STATEMENT.
 

### 2024-01-09 E-Hypergraphs-LICS24.tex
Section 5 update

@@
 \begin{definition}[Convex DPOI rewriting]
+\label{def:convex_dpo}
 Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
@@
 The first thing is that we need to account for is the existence of \textit{internal} interfaces,  and in particular that internal interfaces are not necessarily preserved during rewriting.
-To illustrate this, consider the following rewrite rule
-\[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
-such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} m' \xleftarrow{} m
-\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$.
-When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, e.g.
-by computing the pushout complement, the internal interface of $\mathcal{G}$ should be modified.
-This is because the verticies of $\mathcal{L}$ corresponding to strict internal interface of $\mathcal{L}$ (and hence to strict internal interface of $\mathcal{G}$) should be removed from $\mathcal{G}$ and therefore from $\mathcal{G}$'s internal interface.
-Then, when gluing $\mathcal{R}$ into the created context, we also introduce strict internal interfaces of $\mathcal{R}$ into this context.
+The former means that rewrite rules become of the form
+\[
+\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
+\] where we use the fact that cospans 
+\[
+    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
+\]
+and
+\[
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
+\]
+encode the same data and we will further omit the arrows $0 \to 0 \to \mathcal{L}$.
+That is, a rewrite rule is a pair of cospans of e-hypergraphs with matching \textit{external} interfaces and arbitrary \textit{internal} interfaces.
+To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
+% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
+% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
+\[
+    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
+\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
+When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
+Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
+In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
+\[
+    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
+\]
+not necessarily coincide with the internal interfaces of the initial cospan.
 % The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
@@
 
-To account for such changes in the interfaces we require additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$.
-Consider a DPOI diagram with extra constraints below which encodes
-the notion that e-hypergraph $G$ rewrites into $H$ with external interfaces $n_{ext}+m_{ext}$.
-Note the addition of internal interfaces and how we use the same trick to go from cospan $n_{ext} \xrightarrow{} g_i \xrightarrow{} G \xleftarrow{} g_o \xleftarrow{} m_{ext}$ to $G \xleftarrow{} g_{i} + g_{o} \xleftarrow{} n_{ext} + m_{ext}$.
-The original DPO square is depicted in blue.
+To account for such changes in the interfaces we require additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$ which is summarised in the definition below.
+\begin{definition}[Extended DPOI (EDPOI) rewriting]
+\label{def:dpoi-e}
+Given a span of morphisms 
+\[
+    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
+\]
+in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites to} $\mathcal{H}$ \textit{with external interface} $n + k$ \textit{via a rewrite rule} 
+\[
+    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
+\]
+if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
 % \[\begin{tikzcd}
@@
  \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square_coloured}
+    \tikzfig{combinatorial_semantics/DPOI_square}
  }
 \]
-
-Then, we explicitly show how the internal interfaces of the pushout complement are constructed.
+\end{definition}
+To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph we also need to introduce a revised version of a boundary complement.
 
@@
 
-For monogamous extended cospans $x_{ext} \xrightarrow{} l_{i} \xrightarrow{} L \xleftarrow{} l_{o} \xleftarrow{} y_{ext}$ and $n_{ext} \xrightarrow{} g_{i} \xrightarrow{} G \xleftarrow{} g_{o} \xleftarrow{} m_{ext}$ and mono $f : L \to G$, a pushout complement $x_{ext} + y_{ext} \to C \to G^{\urcorner}$ as depicted in the square below
+For monogamous extended cospans 
+\[
+    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
+\] 
+and 
+\[
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
+\] and mono $f : \mathcal{L} \to \mathcal{G}$, a pushout complement 
+\[
+    i + j \to \mathcal{L}^{\bot} \to G^{\urcorner}
+\] 
+as depicted in the square below
 \[
@@
 
-is a \textit{boundary} complement if $[i_c, i_j]$ is mono 
- and there exist $n_c : n_{ext} \to C$ and $m_c : m_{ext} \to C$ 
+is a \textit{boundary} complement if $[c_1,c_2]$ is mono 
+ and there exist $d_1 : n \to C$ and $d_2 : k \to C$ 
  making the above triangle commute and such that
@@
 \[
-n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}
-\]
-
-where $c_i$ and $c_o$ are computed as
- $c_i = g_i \setminus (l_i \setminus x_{ext})$ and
- $c_o = g_o \setminus (l_o \setminus y_{ext})$.
-
- ALEKSEI: ARE MORPHISMS $c_i + y_{ext} \to C$ AND THE SYMMETRIC ONE OBVIOUS?
-\end{definition}
+n \xrightarrow{} l + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} r + i \xleftarrow{} k
+\]
+
+where $l$ and $r$ are
+\[
+    l = n' \setminus (i' \setminus i)
+\]
+and
+\[
+    r = k' \setminus (j' \setminus j)
+\]
+Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed apart from the part that has a pre-image in $i + j$ there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
+\[
+    D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
+\]
+Similarly for $r \to \mathcal{L}^{\bot}$ which we denote as $g_2$.
+Then there is a morphism $n \to l$ because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
+Similarly for $k \to r$.
+\end{definition}
+
+The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
+The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
 %
@@
 
-Note that cospan $n_{ext} \xrightarrow{} c_i + y_{ext} \xrightarrow{} C \xleftarrow{} c_o + x_{ext} \xleftarrow{} m_{ext}$ is not necessarily well-typed.
+\begin{remark}
+Note that cospan 
+\[
+    n \xrightarrow{} l + j \xrightarrow{} C \xleftarrow{} r + i \xleftarrow{} k
+\] is not necessarily well-typed.
 The cospan's input internal interface may contain nodes that previously were a part of the output internal interface and similarly for output internal interface.
@@
 
+\begin{definition}[Subgraph]
+
+A \textit{subgraph} of an e-hypergraph 
+\[
+    \mathcal{G} = \{V_{\mathcal{G}}, E_{\mathcal{G}}, s_{\mathcal{G}}, t_{\mathcal{G}}, l, <_{G},\consistency_{\mathcal{G}}\}
+\]
+is an e-hypergraph 
+\[
+    \mathcal{H} = \{V_{\mathcal{H}}, E_{\mathcal{H}}, s_{\mathcal{H}}, t_{\mathcal{H}}, l, <_{H},\consistency_{\mathcal{H}}\}
+\] such that $V_{\mathcal{H}} \subseteq V_{\mathcal{G}}$, $E_{\mathcal{H}} \subseteq E_{\mathcal{G}}$ and such that $<_{\mathcal{H}}$ and $\consistency_{\mathcal{H}}$ both satisfy the properties from~\ref{def:e-homo}.
+\end{definition}
+
 \begin{definition}[Down-closed subgraph]
-    We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ \emph{down-closed} if for all $e \in E_{\mathcal{H}}$,   every element of $(e]$ is also in $\mathcal{H}$.
+    We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ \emph{down-closed} if for all $e \in E_{\mathcal{H}}$,   all children of $r$ are also in $\mathcal{H}$.
 \end{definition}
@@
 \begin{definition}[Convex down-closed match] 
-    We call $m : L \to G$ in $\catname{EHyp(\Sigma)}$ a convex down-closed match if
+    We call $m : L \to G$ in $\catname{EHyp(\Sigma)}$ a \textit{convex down-closed match} if
 
     \begin{enumerate}
-        \item $m$ is monomorphism
+        \item $m$ is a monomorphism
         \item $m(L)$ is a convex down-closed subgraph of $G$
+        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
+        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
     \end{enumerate}
@@
 
-The whole process of rewriting including the construction of the internal interfaces of the resulting e-hypergraph is summarised in the following definition.
-
-\begin{definition}[Rewriting in $\WellTypedMdaEcospans$]
-    Given $G \xleftarrow{} g_i + g_o \xleftarrow{} n_{ext} + m_{ext}$ 
-    and $H \xleftarrow{} h_i + h_o \xleftarrow{} n_{ext} + m_{ext}$ in $\Ecospans$ with the outermost interfaces $n_{ext} + m_{ext}$, $G$ rewrites convexly into $H$ with the outermost interface $n_{ext} + m_{ext}$ --- notation $( G \xleftarrow{} n_{ext} + m_{ext} ) \Rrightarrow_{\mathcal{R}}^{+} (E \xleftarrow{} n_{ext} + m_{ext} )$ --- if there exists a rule $L \xleftarrow{} l_i + l_o \xleftarrow{} x_{ext} + y_{ext} \xrightarrow{} r_i + r_o \xrightarrow{} R$ with the outermost interfaces $x_{ext} + y_{ext}$ in the set of rewrite rules $\mathcal{R}$ and object $C$ and cospan arrows $x_{ext} + y_{ext} \xrightarrow{} C \xleftarrow{} n_{ext} + m_{ext}$ such that the diagram below commutes and its marked squares are pushouts
-
-% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
-\[\scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
-    }
-\]
-    and the following conditions hold
+The conditions (3)-(4), in particular, prohibit a match $\{f \mapsto f, g \mapsto g\}$ in $\mathcal{L} = f \otimes g$ and $\mathcal{G} = f + g$.
+
+We are now ready to define the \textit{convex} rewriting for e-hypergraphs with interfaces which we will later use to show the equivalence between $\textsf{PROP}(\Sigma)$ and $\WellTypedMdaEcospans$.
+It will be analogous to the one from Definition~\ref{def:convex_dpo} but again, to make sure the resulting cospan of e-hypergraph is monogamous we will have to construct internal interfaces separately.
+We will also limit matches to convex down-closed ones both to guarantee an mda result and so that the result is a well-defined e-hypergraph.
+
+\begin{definition}[Convex EDPOI rewriting]
+    Let $\mathfrak{R}$ be a set of EDPOI rewrite rules.
+    Then, given 
+    \[
+        \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
+    \] 
+    and 
+    \[
+        \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
+    \] in $\catname{EHyp(\Sigma)}$ with the external interfaces $n + k$, $\mathcal{G}$ \textit{rewrites convexly into} $\mathcal{H}$ \textit{with the outermost interface} $n + k$ --- notation $( \mathcal{G} \xleftarrow{} n + k) \Rrightarrow_{\mathfrak{R}}^{+} (\mathcal{H} \xleftarrow{} n + k)$ --- if there exists a rule
+    \[
+        \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
+    \] 
+    with the external interfaces $i + j$ in $\mathfrak{R}$ and object $\mathcal{C}$ and cospan arrows $i + j \xrightarrow{} \mathcal{C} \xleftarrow{} n + k$ 
+    such that the diagram of Definition~\ref{def:dpoi-e} commutes and its marked squares are pushouts, and the following conditions hold
 
     \begin{enumerate}
-        \label{dpoi:assumptions}
-        \item $m : L \to G$ is a convex down-closed match;
-        \item $x_{ext} + y_{ext} \to C \to G$ is a boundary complement in the leftmost pushout.
+        \label{dpoi-e:assumptions}
+        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex down-closed match;
+        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement in the leftmost pushout.
     \end{enumerate}
@@
 
-The internal interfaces of the resulting e-hypergraph $H$ from the cospan
- $n_{ext} \xrightarrow{} h_i \xrightarrow{} H \xleftarrow{} h_o \xleftarrow{} m_{ext}$
- are constructed as $h_i = (r_i \setminus x_{ext}) + c_i$ 
 and
- $h_o = (r_o \setminus y_{ext}) + c_o$.
- $h_i \to H$ is defined as a copairing of 
- $f_1 : r_i \setminus x_{ext} \to R \to H$ 
- and 
- $f_2 : c_i \to C \to H$. 
-Since both $f_1$ and $f_2$ are mono, their copairing is also a mono. 
-Similarly for $h_o$.
-
-We then say that an mda-cospan 
-\[
-    n_{ext} \xrightarrow{a_1} g_{i} \xrightarrow{a_2} G \xleftarrow{b_2} g_{o} \xleftarrow{b_1} m_{ext}
-\]
-rewrites convexly into another mda-cospan
-
-\[
-    n_{ext} \xrightarrow{a_1} h_{i} \xrightarrow{a_2} H \xleftarrow{b_2} h_{o} \xleftarrow{b_1} m_{ext}
+
+% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
+%  \[
+%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
+% \]
+\begin{align*}
+&n'' = (i'' \setminus i) + l\\
+&k'' = (j'' \setminus j) + r\\
+&n'' \to \mathcal{H} = [f_1,f_2] \text{ where }\\
+&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H}\\ 
+&f_2 : l \to \mathcal{C} \to \mathcal{H} 
+\end{align*}
+Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
+Similarly for $k'' \to \mathcal{H}$.
+
+\begin{proposition}
+    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
+\end{proposition}
+
+Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$ we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
+
+\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
+
+We say that an \textit{mda-cospan}
+\[
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
+\]
+\textit{rewrites convexly into another mda-cospan}
+
+\[
+    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
 \]
@@
 under a rewrite rule 
-\[
-    \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-    x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-\]
+% \[
+%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
+%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
+% \]
+
+\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
 
@@
 \begin{align*}    
-    G \xleftarrow{[ a_2,b_2 ]} g_i + g_{o}& \xleftarrow{[a_1, b_1]} n_{ext} + m_{ext} \\ &\Rrightarrow^{+}\\ H \xleftarrow{[a_2,b_2]} h_i + h_{o}& \xleftarrow{[a_1, b_1]} n_{ext} + m_{ext}
+    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
 \end{align*}
-
-under a rewrite rule
-\[
-    \langle l_{i} + l_{o} \xrightarrow{[a_2, b_2]} L \xleftarrow{[a_1, b_1]} x_{ext} + y_{ext},
-    r_{i} + r_{o} \xrightarrow{[a_2, b_2]} R \xleftarrow{[a_1, b_1]} x_{ext} + y_{ext} 
-    \rangle
-\]
-
-\begin{proposition}
-$n_{ext} \xrightarrow{n_h} h_i \xrightarrow{f} H \xleftarrow{g} h_o \xleftarrow{m_h} m_{ext}$ is a well-typed mda-cospan.
-\end{proposition}
+under the same rewrite rule.
+
+\end{definition}
+
+\begin{remark}
+There is a minor limitation for EDPOI rewriting above in that it is unable to express rewriting
+for rules that contain an edge with no inputs and outputs on their right-hand side (such edges correspond to morphisms of type $0 \to 0$).
+This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
+We claim that this is only a minor limitation because in practice such edges (morphisms) with no
+incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
+\end{remark}
 

### 2024-01-09 E-Hypergraphs-LICS24.tex
small addition Sec 5

@@
 The cospan's input internal interface may contain nodes that previously were a part of the output internal interface and similarly for output internal interface.
-
+\end{remark}
+
+\begin{remark}
 The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
+\end{remark}
 
@@
 under the same rewrite rule.
+We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
 

### 2024-01-09 E-Hypergraphs-LICS24.tex
faithfulness draft.

--- 2024-01-09 E-Hypergraphs-LICS24.tex
@@
 
+CHRIS: NOTE ALSO WEAK NORMAL FORM FOR $\Sigma^+$-TERMS?
 The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
@@
 First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ as a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ (note that cospans embed in extended cospans when the internal interface is isomorphic to the external interface).
-
-%To see that the functor is faithful,  consider $f$ and $g$ of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ such that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We aim to show that $f = g$.
-First,  find weak normal forms of $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$,  namely $\phi_1 + \ldots + \phi_ni$ and $\psi_1 + \ldots + \psi_n$.
-We now need a further observation that if $\llbracket f \rrbracket \Rrightarrow_{{\llbracket \mathcal{S,E} \rrbracket}^+} \llbracket g \rrbracket$,  then there exist rewrites $\phi_i \Rrightarrow_{\mathcal E} \psi_i$.
-Thus,  by $(*)$ we have that $\llbracket \phi_i \rrbracket^{-1} = \llbracket \psi_i \rrbracket^{-1}$.
-Noting that $f = \llbracket \phi_1 \rrbracket^{-1} + \ldots + \llbracket \phi_n \rrbracket^{-1}$ and $g = \llbracket \psi_1 \rrbracket^{-1} + \ldots + \llrbacket \psi_n \rrbracket^{-1}.
-CHRIS: FAITHFULNESS. 
-%
+In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ as a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ (note that cospans embed in extended cospans when the internal interface is isomorphic to the external interface); see Theorem 35 of ~\cite{bonchi_string_2022-2}.
+
+To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
+First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
+Applying the functor $\llbracket - \rrbracket$,  we see that
+\begin{align*}
+	\llbracket f \rrbracket &= \llbracket f_1 + \ldots + f_n \rrbracket &= \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \\
+	\llbracket g \rrbracket &= \llbracket g_1 + \ldots + g_m \rrbracket &= \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
+\end{align*}
+In particular,  note that $\llbracket - \rrbracket$ preserves weak normal forms.
+Therefore $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
+Thus if $n \neq m$ we have that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $.
+Otherwise,  if $n = m$,  we have that there for every permutation $\sigma$ there exists an index $i$ such that $f_i \neq g_{\sigma(i)}$,  and thus by $(*)$ that $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $.
+This implies that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $.
+
 To see that the functor is full,  consider a cospan $f$ in $\WellTypedMdaEcospans$.
@@
 %\end{theorem}
-CHRIS: CONCLUDING STATEMENT.
+CHRIS: CONCLUDING STATEMENT.  COMMENT ABOUT GENERALIZATION TO DIRECTED REWRITES.
 

### 2024-01-10 E-Hypergraphs-LICS24.tex
typoes sec 5 and 6

--- 2024-01-09 E-Hypergraphs-LICS24.tex
@@
 CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
-Because extended cospans have a more general notion of interface,  we need to make some adjustments to DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics}.
-The first thing is that we need to account for is the existence of \textit{internal} interfaces,  and in particular that internal interfaces are not necessarily preserved during rewriting.
+Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
+First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
 The former means that rewrite rules become of the form
@@
 \]
-not necessarily coincide with the internal interfaces of the initial cospan.
+not necessarily coincides with the internal interfaces of the initial cospan.
 % The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
@@
 
-To account for such changes in the interfaces we require additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$ which is summarised in the definition below.
+To account for such changes in the interfaces, consider additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$, summarised in the definition below.
 \begin{definition}[Extended DPOI (EDPOI) rewriting]
@@
 \end{definition}
-To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph we also need to introduce a revised version of a boundary complement.
+To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph boundary complement also requires a revised definition.
 
@@
     i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\] 
-and 
-\[
+\quad\text{ and }\quad
     n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
@@
  making the above triangle commute and such that
+ DAN: THIS (AND SUCH THAT) IS EXTREMELY UNCLEAR, PLEASE REWORD
 there exists a \textit{not necessarily well-typed} mda-cospan
@@
     l = n' \setminus (i' \setminus i)
-\]
-and
-\[
-    r = k' \setminus (j' \setminus j)
-\]
-Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed apart from the part that has a pre-image in $i + j$ there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
+\quad\text{ and }\quad
+    r = k' \setminus (j' \setminus j).
+\]
+Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
 \[
@@
 \]
-Similarly for $r \to \mathcal{L}^{\bot}$ which we denote as $g_2$.
-Then there is a morphism $n \to l$ because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
+Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
+Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
 Similarly for $k \to r$.
@@
 We are now ready to define the \textit{convex} rewriting for e-hypergraphs with interfaces which we will later use to show the equivalence between $\textsf{PROP}(\Sigma)$ and $\WellTypedMdaEcospans$.
-It will be analogous to the one from Definition~\ref{def:convex_dpo} but again, to make sure the resulting cospan of e-hypergraph is monogamous we will have to construct internal interfaces separately.
+It will be analogous to the one from Definition~\ref{def:convex_dpo} but again, to make sure the resulting cospan of e-hypergraphs is monogamous we will have to construct internal interfaces separately.
 We will also limit matches to convex down-closed ones both to guarantee an mda result and so that the result is a well-defined e-hypergraph.
@@
         \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
-    \] 
-    and 
-    \[
+    \quad 
+    \text{and} 
+    \quad
         \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
@@
 \end{definition}
-
 and
-
 % The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
@@
 
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
+The equations of Figure~\ref{fig:string-equations}---excepting the commutativity equation---can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
 \begin{proposition}[Weak normal form]

### 2024-01-10 E-Hypergraphs-LICS24.tex
faithfulness draft.

@@
 
-The equations of Figure~\ref{fig:string-equations}---excepting the commutativity equation---can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
+CHRIS: NOTE ALSO WEAK NORMAL FORM FOR $\Sigma^+$-TERMS?
+The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
 \begin{proposition}[Weak normal form]
@@
 First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ as a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ (note that cospans embed in extended cospans when the internal interface is isomorphic to the external interface).
-
-%To see that the functor is faithful,  consider $f$ and $g$ of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ such that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We aim to show that $f = g$.
-First,  find weak normal forms of $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$,  namely $\phi_1 + \ldots + \phi_ni$ and $\psi_1 + \ldots + \psi_n$.
-We now need a further observation that if $\llbracket f \rrbracket \Rrightarrow_{{\llbracket \mathcal{S,E} \rrbracket}^+} \llbracket g \rrbracket$,  then there exist rewrites $\phi_i \Rrightarrow_{\mathcal E} \psi_i$.
-Thus,  by $(*)$ we have that $\llbracket \phi_i \rrbracket^{-1} = \llbracket \psi_i \rrbracket^{-1}$.
-Noting that $f = \llbracket \phi_1 \rrbracket^{-1} + \ldots + \llbracket \phi_n \rrbracket^{-1}$ and $g = \llbracket \psi_1 \rrbracket^{-1} + \ldots + \llrbacket \psi_n \rrbracket^{-1}.
-CHRIS: FAITHFULNESS. 
-%
+In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ as a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ (note that cospans embed in extended cospans when the internal interface is isomorphic to the external interface); see Theorem 35 of ~\cite{bonchi_string_2022-2}.
+
+To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
+First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
+Applying the functor $\llbracket - \rrbracket$,  we see that
+\begin{align*}
+	\llbracket f \rrbracket &= \llbracket f_1 + \ldots + f_n \rrbracket &&= \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \\
+	\llbracket g \rrbracket &= \llbracket g_1 + \ldots + g_m \rrbracket &&= \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
+\end{align*}
+In particular,  note that $\llbracket - \rrbracket$ preserves weak normal forms.
+Therefore $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
+Thus if $n \neq m$ we have that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $.
+Otherwise,  if $n = m$,  we have that there for every permutation $\sigma$ there exists an index $i$ such that $f_i \neq g_{\sigma(i)}$,  and thus by $(*)$ that $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $.
+This implies that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $.
+
 To see that the functor is full,  consider a cospan $f$ in $\WellTypedMdaEcospans$.

### 2024-01-10 E-Hypergraphs-LICS24.tex
conclusion

@@
 
-We have described a categorical semantics for e-graphs over algebraic theories in terms of Cartesian categories enriched over a semilattice, and defined a natural generalization to account for monoidal theories in terms of SMCs,  similarly enriched.
-Having described a translation of e-graphs into our formalism, we developed a combinatorial representation of morphisms in our free category which factor out the structural SMC equations, but remain sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding  rewrite theory proved delicate, and---together with its proof of completeness---constitute the main technical contributions of the paper.
-Finally, we considered an analogue of e-rewriting in our setting, pushing towards the practical considerations necessary for an implementation of our ideas.
-Although we believe our work forms a solid basis for a formal mathematical theory of (acyclic) e-graphs,  there are a number of extensions which we intend to investigate in the future.
-
-\subsection{Future Work and Applications}
-
-We have already alluded to the fact we deal with only \textit{acyclic} e-graphs in this paper.
-Nevertheless,  within this setting we present a generalization from Cartesian to (symmetric) monoidal theories,  giving rise to a host of potential new domains of application.
-Further,  the string diagrammatic and graph rewriting framework we set up is ameenable to extension, including lifting the acyclicity restriction.
+We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
+In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
+Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
+As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
+The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
+Finally, we considered an analogue of e-rewriting in this setting, required in the developing potential implementations.
+In addition to the current mathematical treatment of (acyclic) e-graphs,  there are a number of extensions which we intend to investigate in the future.
+
+Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{GhicaZanasiTutorial}, as sketched in Fig.
+\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
+\ref{fig:app1}(b); or trace~\cite{Joyal, Hasegawa} which can be used to express feedback in categorical models of digital circuits~\cite{GhicaJung, GhicaKaye} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
+All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
 
@@
 \]
+\caption{E-graphs for functorial boxes and the ZX-calculus
+DAN: THE LINE BETWEEN THE APPLICATION NODE AND THE BOX IS TOO LONG}
+\label{fig:app1} 
+\end{figure}
+\begin{figure}
 \[
@@
 \]
-\caption{Example applications of e-hypergraphs.
-Top left, the ZX-calculus.
-Top right, the $\lambda$-calculus as represented by string diagrams for Cartesian closed categories.
-Below,  an extension of Figure \ref{fig:} by another rewrite step, $1 * x \to x$,  introducing a cycle modelled by traced structure.
-}
-\end{figure}\label{fig:applications}
-
-Symmetric monoidal theories have received significant attention over recent years: the ZX-calculus can be used to characterize quantum processes, with applications to optimization of quantum circuits \cite{coecke_interacting_2011,ZX} ; similar theories also characterize signal flow diagrams
-\cite{baldan_categorical_2014}, digital circuits 
-\cite{ghica_operational_2021}, as well as having application to computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.
-In future work, we hope to apply our generalization of e-graphs to optimization problems in these domains (e.g.
-Figure \ref{fig:applications})
+\label{fig:app2}
+\caption{E-graphs with cycles and e-hypergraphs with trace}
+\end{figure}
 
@@
 In the above cases, such representations and their DPO-rewriting theory is well-studied
-\cite{ghica_rewriting_2023} % traced comonoid
-[cite, cite, cite].
-We consider that our development of hierarchical "e-box" structure within string diagrams and their associated (e-)hypergraphs ( categorically,  semilattice enrichment) is essentially orthogonal to these developments, and so ought to be easily combined to give notions of e-hypergraphs specifically suited to the optimization of these particular monoidal theories.
-A particularly interesting extension that we intend to consider is the combination of our "e-structure" with (Cartesian) \textit{closed} categories, allowing the natural combination of e-graph optimization techniques with the $\lambda$-calculus.
-The corresponding combinatorial representation would build on the recent work \cite{ghica_hierarchical_2023, alvarez-picallo_rewriting_2022, alvarez-picallo_functorial_2021}.
-String diagrams for these various examples are included in Figure \ref{fig:applications}.
-There is also the question of developing a corresponding term calculus \cite{Heijltjes:FMCII}.
-% hierarchical
+\cite{ghica_rewriting_2023,cite,cite,cite}. % traced comonoid
+
+The hierarchical ``e-box'' structure within string diagrams and their associated (e-)hypergraphs (categorically, the semilattice enrichment) is essentially orthogonal to these developments.
+Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
+However, as we have seen in this paper, the technical details can be delicate.
+Finally, a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{https://doi.org/10.4204/EPTCS.266.4}.
+Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
 

### 2024-01-11 E-Hypergraphs-LICS24.tex
Figures update up to Section 5

@@
 \usepackage{tikzit}
+% to reduce tikz picture white space
+\usepackage{adjustbox}
 
@@
 A typical DPO square looks as follows,  where the two marked squares are pushouts. 
-\[
-\begin{tikzcd}
-\mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-\mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner} \mathcal H
-\end{tikzcd}
+% \[
+% \begin{tikzcd}
+% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
+% \end{tikzcd}
+% \]
+\[
+    \scalebox{0.8}{
+        \tikzfig{combinatorial_semantics/DPO_square}
+    }
 \]
@@
 Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
-\[
-\begin{tikzcd}
-\mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-\mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-                  & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-\end{tikzcd}
+% \[
+% \begin{tikzcd}
+% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
+%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
+% \end{tikzcd}
+% \]
+\[
+    \scalebox{0.8}{
+        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
+    }
 \]
@@
 % https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-\[\begin{tikzcd}
-	\mathcal L && {i+j} \\
-	\\
-	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-	\\
-	&& {n+m}
-	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-	\arrow["f"', from=1-1, to=3-1]
-	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-	\arrow["g", from=3-3, to=3-1]
-\end{tikzcd}\]
+% \[\begin{tikzcd}
+% 	\mathcal L && {i+j} \\
+% 	\\
+% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
+% 	\\
+% 	&& {n+m}
+% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
+% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
+% 	\arrow["f"', from=1-1, to=3-1]
+% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
+% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
+% 	\arrow[from=3-3, to=3-1]
+% \end{tikzcd}\]
+
+\[
+    \scalebox{0.8}{
+        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
+    }
+\]
+
 is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
@@
     \]
-Composition of two morphisms $f : X_{ext} \to Y_{ext}$ and $g : Y_{ext} \to Z_{ext}$ is computed in two stages
-First, $\mathcal{F} +_{f_{ext}, g_{ext}} \mathcal{G}$ is computed,  as shown below.
+Composition of two morphisms $h_1 : n \to m$ and $h_2 : m \to k$ is computed in two stages.
+First, $\mathcal{H} = \mathcal{F} +_{f_{ext}, g_{ext}} \mathcal{G}$ is computed,  as shown below.
 
@@
 \[
-\scalebox{0.9}
+\trimbox{0cm 0cm 0cm 1cm}{
+\adjustbox{scale=0.9,center}
 {\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-\]
-Then, new feet for $\mathcal{H} = A +_{f_{ext};f,g_{ext};g} B$ are computed as below
-\[
-X_{ext} \xrightarrow{a_{ext};\iota_1} X \uplus (Y' \setminus Y_{ext}) \xrightarrow{h} \mathcal{H} \xleftarrow{k} Z \uplus (Y \setminus Y_{ext}) \xleftarrow{g_{ext};\iota_1} Z_{ext}
+}
+\]
+Then, new feet for $\mathcal{H}$ are computed as below
+\[
+n \xrightarrow{f_{ext};\iota_1} n' \uplus (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' \uplus (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
 \]
@@
 \[
-    h = [ a;p_1, ~g';p_2 ]
+    h_1 = [ f_{int};p_1, ~g_{int}|_{m'' \setminus m};p_2 ]
 \qquad
-    k = [ b;p_2, ~f';p_1 ]
+    h_2 = [ g'_{int};p_2, ~f'_{int}|_{m' \setminus m};p_1 ]
 \]
@@
 % and then the disjoint union is just a coproduct. 
-where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
+% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
 The identity of composition is given by the obvious extended cospan of identities. 
@@
 First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
-The former means that rewrite rules become of the form
+The former means that we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching external interfaces.
+
+\[
+    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
+\]
+and
+\[
+    i \xrightarrow{} i'' \xrightarrow{} \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j
+\]
+
+We will next use the fact that cospans
+\[
+    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
+\]
+and
+\[
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
+\]
+endcode the same data and only consider cospans of the form
+
 \[
 \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] where we use the fact that cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\]
-and
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
-\]
-encode the same data and we will further omit the arrows $0 \to 0 \to \mathcal{L}$.
-That is, a rewrite rule is a pair of cospans of e-hypergraphs with matching \textit{external} interfaces and arbitrary \textit{internal} interfaces.
+\] to fit the DPO formalism.
+We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
 To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
@@
 \[
-    i + j \to \mathcal{L}^{\bot} \to G^{\urcorner}
+    i + j \to \mathcal{L}^{\bot} \to \mathcal{G}
 \] 
@@
 
-is a \textit{boundary} complement if $[c_1,c_2]$ is mono 
- and there exist $d_1 : n \to C$ and $d_2 : k \to C$ 
- making the above triangle commute and such that
- DAN: THIS (AND SUCH THAT) IS EXTREMELY UNCLEAR, PLEASE REWORD
-there exists a \textit{not necessarily well-typed} mda-cospan
-
+is a \textit{boundary} complement if the following holds
+\begin{enumerate}
+    \item $[c_1,c_2]$ is mono 
+    \item there exist $d_1 : n \to \mathcal{C}$ and $d_2 : k \to \mathcal{C}$ making the above triangle commute
+    \item there exists a \textit{not necessarily well-typed} mda-cospan
     \[
@@
     \]
+\end{enumerate}
 
@@
 Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
+That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
 Similarly for $k \to r$.
@@
 
-The conditions (3)-(4), in particular, prohibit a match $\{f \mapsto f, g \mapsto g\}$ in $\mathcal{L} = f \otimes g$ and $\mathcal{G} = f + g$.
+The conditions (3)-(4), in particular, prohibit a match $\{f \mapsto f, g \mapsto g\}$ in $\mathcal{L} =$ $\scalebox{0.4}{\tikzfig{combinatorial_semantics/f_times_g}}$ and $\mathcal{G} =$ $\scalebox{0.4}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}$.
 

### 2024-01-12 E-Hypergraphs-LICS24.tex
Sec 6 figure update

@@
 The former means that we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching external interfaces.
-
 \[
     i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\]
-and
-\[
+\qquad \text{and} \qquad
     i \xrightarrow{} i'' \xrightarrow{} \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j
 \]
-
 We will next use the fact that cospans
@@
     i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\]
-and
-\[
+\qquad \text{and} \qquad
     0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
@@
 endcode the same data and only consider cospans of the form
-
 \[
@@
     \[
-    \scalebox{0.6}{
+    \scalebox{0.55}{
     \tikzfig{combinatorial_semantics/f_plus_g_new}
@@
 \[
-	f \sim g \quad \text{if} \quad f \Rightarrow^*_{\llbracket \mathcal E \rrbracket} g ~ . 
+	f \sim g \quad \text{if} \quad f \Rrightarrow^{+,*}_{\llbracket \mathcal E \rrbracket} g ~ . 
 \]
@@
 \begin{proposition}[Soundness]
+\label{prop:soundness}
 The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
@@
 \begin{proposition}[Weak normal form]
+\label{prop:wnormal_form}
 For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
@@
 \[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^*_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
+	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
 \]
@@
 \label{pushout:assumptions}
-    \item $Z$ is a \textit{descrete} e-hypergraph.
+    \item $Z$ is a \textit{discrete} e-hypergraph.
     \item $f,g$ are monos.
@@
 
+\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
+The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
+\end{proposition}
+
+\begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
+
+For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
+\[
+    f = f_1 + \ldots + f_n
+\]
+such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
+\end{proposition}
+
 \end{document}

### 2024-01-12 E-Hypergraphs-LICS24.tex
Small fixes.

--- 2024-01-11 E-Hypergraphs-LICS24.tex
@@
 The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-Finally, we considered an analogue of e-rewriting in this setting, required in the developing potential implementations.
-In addition to the current mathematical treatment of (acyclic) e-graphs,  there are a number of extensions which we intend to investigate in the future.
-
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{GhicaZanasiTutorial}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{Joyal, Hasegawa} which can be used to express feedback in categorical models of digital circuits~\cite{GhicaJung, GhicaKaye} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
+
+In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
+Some of these are: support for `functorial boxes'~\cite{mellies} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{GhicaZanasiTutorial}, as sketched in Fig.
+\ref{fig:app1}~(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
+\ref{fig:app1}~(b); and  traces~\cite{Joyal, Hasegawa},  which can be used to express feedback in categorical models of digital circuits~\cite{GhicaJung, GhicaKaye},  and, indeed cycles in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
 All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
@@
 
-The theories described above often have more structure than simply being symmetric monoidal: for example, they may contain comonoids, Frobenius algebras or traced structure (as in the case of e-graphs with cycles),  and this structure can be exploited to give more natural (in the sense of factoring out more equations) combinatorial representations.
-In the above cases, such representations and their DPO-rewriting theory is well-studied
-\cite{ghica_rewriting_2023,cite,cite,cite}. % traced comonoid
-
-The hierarchical ``e-box'' structure within string diagrams and their associated (e-)hypergraphs (categorically, the semilattice enrichment) is essentially orthogonal to these developments.
+In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
+Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,cite,cite,cite}.
+% traced comonoid
+%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
 Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.

### 2024-01-13 E-Hypergraphs-LICS24.tex
Citations fixed, proofs to appendix added

--- 2024-01-12 E-Hypergraphs-LICS24.tex
@@
 
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{GhicaZanasiTutorial}, as sketched in Fig.
+Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
 \ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{Joyal, Hasegawa} which can be used to express feedback in categorical models of digital circuits~\cite{GhicaJung, GhicaKaye} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
+\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
 All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
@@
 \]
+\caption{E-graphs with cycles and e-hypergraphs with trace}
 \label{fig:app2}
-\caption{E-graphs with cycles and e-hypergraphs with trace}
 \end{figure}
@@
 However, as we have seen in this paper, the technical details can be delicate.
-Finally, a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{https://doi.org/10.4204/EPTCS.266.4}.
+Finally, a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal}.
 Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
@@
 \end{proposition}
+\begin{proof}
+    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $f = g$ in $\textsf{PROP}^{+}(\Sigma)$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket f \rrbracket \Rrightarrow_{\mathcal{S}}^{+,*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans$.
+    $f = g$ implies that there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
+    As $\textsf{PROP}^{+}(\Sigma)$ is quotiented by SMC equations we have the following contexts $\mathcal{C}$ for any term $f$ such that $f = \mathcal{C}[l]$ for some other term $l$.
+    \[
+        \mathcal{C} := [] ~|~ c_1 + \ldots + \mathcal{C} + \ldots + c_n ~|~ c_1 ; (id \otimes \mathcal{C}); c_2
+    \]
+    we will proceed by induction on the number of equation applied to make $f$ into $g$.
+    \begin{itemize}
+        \item Base case. $f = g$ implies that for some equation $l = r$ in Figure~\ref{fig:string-equations}, $f = \mathcal{C}[l]$ and $g = \mathcal{C}[r]$.
+              We then note that every subterm of $f$ corresponds to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence so does $l$.
+              By computing pushout complement we remove the image of $\llbracket l \rrbracket$ from $\llbracket f \rrbracket$ and replace it with $\llbracket r \rrbracket$ which yields $\llbracket g \rrbracket = \llbracket \mathcal{C} \rrbracket [\llbracket r \rrbracket] = \llbracket \mathcal{C}[r] \rrbracket$ and $\llbracket f \rrbracket \Rrightarrow^{+}_{l = r} \llbracket g \rrbracket$.
+        \item Suppose $f = g$ by applying less than $n$ equations implies $\llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}} \llbracket g \rrbracket$
+        \item Let $f = g$ by applying exactly $n$ equations and then $f = g'$ by applying first $n-1$ equations and $\llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}} \llbracket g' \rrbracket$ and $g' = g$ by applying one equation.
+              By reusing the base case we have that $\llbracket g' \rrbracket \Rrightarrow^{+}_{l = r} \llbracket g \rrbracket$ and hence $\llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}} \llbracket g \rrbracket$.
+    \end{itemize}
+    The only thing that is left to show that pushout complement and pushout exist for every equation in $\mathcal{S}$.
+    Suppose, $f = g$ by applying $l = r$ which is the top left equation from~\ref{fig:string-equations}.
+    $\llbracket r \rrbracket$ is then given in Figure~\ref{fig:structural_rule_pushout}.
+    \begin{figure*}
+        \centering
+        \[
+            \scalebox{0.45}{
+                \tikzfig{combinatorial_semantics/structural_rule_pushout}
+            }
+        \]
+        \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
+        \label{fig:structural_rule_pushout}
+    \end{figure*}
+
+    After we identify a convex down-closed image of the left-hand side of this rule within a graph $\llbracket f \rrbracket$ we remove the image by keeping only the top-level verticies.
+    Note that because the image of all $a$'s and $b$'s are top-level verticies, we are allowed to pick any monomorphism for our pushout complement.
+    And for the same reason the pushout for the right square exists.
+    Then we simply glue the right-hand side of the rule into the hole.
+    Similar argument applies to other rules as well.
+\end{proof}
 
@@
 \end{proposition}    
+\begin{proof}
+    First, we apply the interpretation of the rules (1)-(4) from Figure~\ref{fig:string-equations} which reduces the number of edges and vertices until there is only one top-level edge which is hierarchical.
+    There will be a redex for these rules as long as there are at least two connected edges at least one of which is hierarchical.
+    The above steps therefore produce a form like \[f_1 + \ldots + f_n\] where each $f_i$ is also of the form $g_1 + \ldots + g_m$ and so on.
+    Then we apply the interpretation of rules (5)-(6) and (8) which reduces the number of hierarchical edges at each step.
+    Once again, there is a redex for such a rule as long as there are at least two hierarchical edges.
+    This will yield a form $f_1 + \ldots + f_n$ such that each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$.
+    The latter is guaranteed by rule (8).
+\end{proof}
 

### 2024-01-14 E-Hypergraphs-LICS24.tex

@@
 The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-Finally, we considered an analogue of e-rewriting in this setting, required in the developing potential implementations.
-In addition to the current mathematical treatment of (acyclic) e-graphs,  there are a number of extensions which we intend to investigate in the future.
-
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
+
+In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
+Some of these are: support for `functorial boxes'~\cite{mellies} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{GhicaZanasiTutorial}, as sketched in Fig.
+\ref{fig:app1}~(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
+\ref{fig:app1}~(b); and  traces~\cite{Joyal, Hasegawa},  which can be used to express feedback in categorical models of digital circuits~\cite{GhicaJung, GhicaKaye},  and, indeed cycles in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
 All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.

### 2024-01-15 E-Hypergraphs-LICS24.tex
Polishing appendix (from the bottom)

@@
 \begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $K \xrightarrow{i} L \xrightarrow{m} G$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every node of its source and target is either (i) not in the image of $m$ or (ii) in the image of $i ; m$.
-They satisfy the no-identification condition if any two nodes merged by $m$ are in the image of $i$.
+Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every node of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
+They satisfy the no-identification condition if any two nodes merged by $m$ are in the image of $f$.
 \end{definition}
@@
     % https://q.uiver.app/#q=WzAsNSxbMCwwLCJYIl0sWzIsMCwiWiJdLFsyLDIsIkMiXSxbMCwyLCJQXntcXHVyY29ybmVyfSJdLFsyLDUsIlxcYnVsbGV0Il0sWzEsMCwiaSIsMl0sWzEsMl0sWzIsM10sWzAsMywibSIsMl1d
-\[\begin{tikzcd}
-	X && Z \\
-	\\
-	{P^{\urcorner}} && C \\
-	\\
-	\\
-	&& \bullet
-	\arrow["i"', from=1-3, to=1-1]
-	\arrow[from=1-3, to=3-3]
-	\arrow[from=3-3, to=3-1]
-	\arrow["m"', from=1-1, to=3-1]
-\end{tikzcd}\]
-    Such that $m,i$ are monos and $Z$ is discrete. Then, pushout complement $Z \to C \to G$ exists if
+% \[\begin{tikzcd}
+% 	X && Z \\
+% 	\\
+% 	{P^{\urcorner}} && C \\
+% 	\\
+% 	\\
+% 	&& \bullet
+% 	\arrow["i"', from=1-3, to=1-1]
+% 	\arrow[from=1-3, to=3-3]
+% 	\arrow[from=3-3, to=3-1]
+% 	\arrow["m"', from=1-1, to=3-1]
+% \end{tikzcd}\]
+\[
+    \tikzfig{appendix/pushout_square}
+\]
+    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
+    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
     \begin{enumerate}
-        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
-        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
+        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
+        \item $\consistency(m(v_i)) = \consistency(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
         \item The no-dangling condition holds.
@@
     Then, no-dangling and no-identification conditions are necessary for pushout complement to exists~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $G$ everything in $L$ that has no pre-image in $Z$ the arrow $Z \to C$ will satisfy the assumptions~\ref{pushout:assumptions}.
+    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
 \end{proof}
@@
     \begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of nodes and hyperedges.
-        Hence, pushout complement should yield pushout complements for the sets of nodes and hyperedges.
+        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of verticies and hyperedges.
+        Hence, pushout complement should yield pushout complements for the corresponding sets.
         In particular,
@@
         \begin{subfigure}{0.45\textwidth}
-            \[\begin{tikzcd}
-                {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-                \\
-                {V_{G}^{\urcorner}} &&&& {V_{C}}
-                \arrow["{b_{V}}", from=1-5, to=1-3]
-                \arrow["{a_{V}}", from=1-3, to=1-1]
-                \arrow["{f_{V}}", from=1-1, to=3-1]
-                \arrow["{g_{V}}"', from=3-5, to=3-1]
-                \arrow["c"', from=1-5, to=3-5]
-            \end{tikzcd}\]
+            % \[\begin{tikzcd}
+            %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
+            %     \\
+            %     {V_{G}^{\urcorner}} &&&& {V_{C}}
+            %     \arrow["{b_{V}}", from=1-5, to=1-3]
+            %     \arrow["{a_{V}}", from=1-3, to=1-1]
+            %     \arrow["{f_{V}}", from=1-1, to=3-1]
+            %     \arrow["{g_{V}}"', from=3-5, to=3-1]
+            %     \arrow["c"', from=1-5, to=3-5]
+            % \end{tikzcd}\]
+            \[
+                \tikzfig{appendix/pushout_verticies}
+            \]
         \end{subfigure}
@@
         \[
-        \begin{tikzcd}
-            {E_{L}} && 0 && 0 \\
-            \\
-            {E_{G}^{\urcorner}} &&&& {E_{C}}
-            \arrow["{b_{E}}", from=1-5, to=1-3]
-            \arrow["{a_{E}}", from=1-3, to=1-1]
-            \arrow["{f_{E}}", from=1-1, to=3-1]
-            \arrow["{g_{E}}"', from=3-5, to=3-1]
-            \arrow["c"', from=1-5, to=3-5]
-        \end{tikzcd}
+        % \begin{tikzcd}
+        %     {E_{L}} && 0 && 0 \\
+        %     \\
+        %     {E_{G}^{\urcorner}} &&&& {E_{C}}
+        %     \arrow["{b_{E}}", from=1-5, to=1-3]
+        %     \arrow["{a_{E}}", from=1-3, to=1-1]
+        %     \arrow["{f_{E}}", from=1-1, to=3-1]
+        %     \arrow["{g_{E}}"', from=3-5, to=3-1]
+        %     \arrow["c"', from=1-5, to=3-5]
+        % \end{tikzcd}
+        \tikzfig{appendix/pushout_edges}
         \]
@@
     \end{figure}
-    $V$s and $E$s are sets of nodes and hyperedges respectively, marked squares are pushouts and $C$s are pushout complements.
-    Because $i+j$ is discrete, $E_{G}^{\urcorner}$ is the disjoint union of $E_{L}$ and $E_{C}$ and hence $E_{C} = E_{G}^{\urcorner} \setminus E_{L}$.
-    Then, since all the arrows are monos, the left square can be rewritten as follows.
-    \[
-      % https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3IFxcY29uZyBpICsgaiArIHoiXSxbMiwxLCJiX3tWfSJdLFsxLDAsImFfe1Z9Il0sWzAsMywiZl97Vn0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
+    $V$s and $E$s are sets of verticies and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
+    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
+    Then, since all the arrows are monos, the first square can be rewritten as follows.
+    \[
+      % https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3IFxcY29uZyBpICsgaiArIHoiXSxbMiwxLCJoX3tWLGV4dH0iLDJdLFsxLDAsImhfe1YsaW50fSIsMl0sWzAsMywibV97Vn0iLDJdLFs0LDMsImdfe1Z9Il0sWzIsNCwiYyIsMl1d
     \begin{tikzcd}
@@
 	{i + j + x + y + z} &&&& {i + j + w \cong i + j + z}
-        \arrow["{b_{V}}", from=1-5, to=1-3]
-        \arrow["{a_{V}}", from=1-3, to=1-1]
-        \arrow["{f_{V}}", from=1-1, to=3-1]
-        \arrow["{g_{E}}"', from=3-5, to=3-1]
-        \arrow["c"', from=1-5, to=3-5]
+	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
+	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
+	\arrow["{m_{V}}"', from=1-1, to=3-1]
+	\arrow["{g_{V}}", from=3-5, to=3-1]
+	\arrow["c_{V}"', from=1-5, to=3-5]
 \end{tikzcd}
@@
     \[
-        C_{1} = (V_{C}, V_{E}, s_{1},t)
-    \]
-    and
-    \[
-        C_{2} = (V_{C},V_{E},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(h)) = s_{G}^{*}(g_{E}(h)) = g_{V}^{*}(s_{2}(h))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(h)) = g_{V}^{*}(s_{2}(h))$ implies $s_{1}(h) = s_{2}(h)$.
+        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
+    \qquad
+    \text{and}
+    \qquad
+        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
+    \]
+    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
+    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
     Similarly for targets.
-    We also need to show that $<_{c}$ and $\consistency^{\mu}$ are unique.
+    We also need to show that $<$ and $\consistency$ are unique.
     Suppose that there exist
     \[
-        C_{1} = (V_{C}, V_{E}, s,t, <_{1})
-    \]
-    and
-    \[
-        C_{2} = (V_{C},V_{E},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that $g_{E}(<_{1}(i_{V_{C}}(v))) = <_{G}(i_{V_{G}}(g_{V}(v))) = g_{E}(<_{2}(i_{V_{C}}(v)))$ which implies $<_{2}(i_{V_{C}}(v)) = <_{1}(i_{V_{C}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
+        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
+    \qquad
+    \text{and}
+    \qquad
+        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
+    \]
+    Because $g$ is homomorphism, it must be the case that 
+    \[
+        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
+    \]
+    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
     Similarly for edges.
-    Now suppose that $<_{1}(i_{V_{C}}(v)) = e$ and $<_{2}(i_{V_{C}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{G}(i_{V_{G}}(g_{V}(v)))$ and there exists $v' \in V_{L}$ and $v'' \in V_{C_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $C_{1}$, but then it violates our assumptions as there exists some node $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = b_{V};a_{V}(u)$ and $[v') \not = \varnothing$.
+    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
+    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
+    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
     Therefore we get a contradiction. 
@@
     \[
-        C_{1} = (V_{C}, V_{E}, s,t, <, \consistency_{1}^{\mu})
-    \]
-    and
-    \[
-        C_{2} = (V_{C},V_{E},s,t,<, \consistency_{2}^{\mu})
+        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
+    \quad
+    \text{and}
+    \quad
+        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
     \]
@@
     \[
-        \langle g_{V};i_{V_{G}}, g_{E};i_{E_{G}} \rangle^{*}(\consistency_{C_1}^{\mu}(i_{V_{C}}(v))) \subseteq \consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v))
-    \]
-    \[
-        \langle g_{V};i_{V_{G}}, g_{E};i_{E_{G}} \rangle^{*}(\consistency_{C_2}^{\mu}(i_{V_{C}}(v))) \subseteq \consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v))  
-    \]
-    \begin{itemize}
-        \item If $\consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v)) = \varnothing$ then both lefthand sides should be $\varnothing$ and the relations are equal.
-        \item Suppose that $\consistency_{G}^{\mu}(g_{V};i_{V_{G}}(v)) \not = \varnothing$ which further means that $[g_{V};i_{V_{G}}(v)) \not = \varnothing$.
-        \begin{itemize}
-            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \subseteq \consistency_{C_2}^{\mu}(i_{V_{C}}(v))$.
-                  This means that there exists $v'$ such that $v' \consistency_{C_{2}}^{\mu} v$ and $v' \not \consistency_{C_{1}}^{\mu} v$.
-                  Because $g_{V}(v) \consistency^{\mu} v'$ it means that there exists $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{L}$ such that $w \consistency_{L}^{\mu} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{C_{2}}^{\mu} u'$ and $w \consistency_{L}^{\mu} w'$.
-            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \cap \consistency_{C_2}^{\mu}(i_{V_{C}}(v)) = \varnothing$
+        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
+    \]
+    \[
+        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
+    \]
+    \begin{itemize}
+        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both lefthand sides should be $\varnothing$ and therefore the relations are equal.
+        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
+              Consider the following cases.
+        \begin{itemize}
+            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
+                  This means that there exists $v'$ such that 
+                  \[
+                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
+                    \quad
+                    \text{and}
+                    \quad
+                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
+                \]
+                  Because $g_{V}(v) \consistency_{\mathcal{G}} g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
+                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
+            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
                   Similar to the above.
-            \item $\consistency_{C_1}^{\mu}(i_{V_{C}}(v)) \cap \consistency_{C_2}^{\mu}(i_{V_{C}}(v)) \not = \varnothing$
+            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
                   Similar to the above.

### 2024-01-16 E-Hypergraphs-LICS24.tex
intro and S3 small fixes.

@@
 \subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  important structural equalities arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
+While the design of an appropriate combinatorial representation of semilattice enriched SMCs is delicate,  the main technical result of the paper is a proof of soundness and completeness of this representation.
+This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
+While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
 They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
@@
 
-Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched SMCs.
-That is, we exhibit a translation functor $\llbracket-\rrbracket$ from the free semilattice enriched SMC over a symmetric monoidal theory to our (quotiented) category of e-hypergraphs such that
-morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the monoidal theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow_{DPO} \llbracket g \rrbracket$.
-
-DAN: THIS NEEDS EXPANDING. EXPLAINING IN PLAIN WORDS SOME OF THE THEORETICAL PROPERTIES OF E-GRAPHS AND THE CATEGORICAL STATEMENT.
+Precisely,  our soundness and completeness result is the statement that,  for any morphisms $f$ and $g$ in an appropriate free semilattice-enriched SMC,  $f = g$  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
+In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
 
@@
 
-Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
+Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesizers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
 This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
@@
 There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-
-CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
 Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}. 
@@
 \cite{bonchi_string_2022-2},  we have the following theorem which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26~\cite{bonchi_string_2022-2}]
+\begin{theorem}[Corollary 26~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
     $\textbf{PROP}(\Sigma) \cong \MdaCospans$
@@
 %By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-
-CHRIS: WANT TO STATE THE CORRESPONDENCE THEOREM WITH SMTS HERE? NEED TO DEFINE $\lceil-\rceil$ AND INTERPRETATION? Hmm
-DAN: I THINK THIS PRELIMINARY SECTION IS VERY LONG.
-UNLESS WE USE THIS THEOREM IN THE PAPER LATER CAN WE JUST STATE IT INFORMALLY, WITHOUT INTRODUCING THIS NOTATION?
-IF THE NOTATION IS USED, IT MUST BE DEFINED.
-\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-\[
-	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-\]
-where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-\end{theorem}
+We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
+In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
+%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
+%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
+%\[
+%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
+%\]
+%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
+%\end{theorem}
 \section{E-Hypergraphs}\label{sec:e-hypergraphs}

### 2024-01-16 E-Hypergraphs-LICS24.tex
Update EDPOI section

@@
 We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
+We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
 We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
@@
 In DPO rewriting,  such a rewrite rule is formalized as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square looks as follows,  where the two marked squares are pushouts. 
+
 % \[
@@
 % \]
+CHRIS: WHY IS J DISPLAYING AND NOT R? 
 \[
@@
 By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
+We now make this identification throughout without further mention.
 Thus we reach the following variant of DPO rewriting.
@@
 For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement as depicted in the square below
+and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
 
@@
 Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathcal{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
+Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
 and the following conditions hold
@@
 Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the sub-hypergraph below the hierarchical edge.
+These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
+These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
 Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
@@
 Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extened cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
+
+We use the notation $n \setminus m$ to denote the discrete hypergraph with $n - m$ nodes,  in particular with removal of the nodes of $m$ removed from $n$ when nodes of $m$ are identified within $n$.
 \begin{definition}[Category of e-hypergraphs with interfaces]
@@
 
-    Two cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
+    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
     \[
@@
     \]
-Composition of two morphisms $h_1 : n \to m$ and $h_2 : m \to k$ is computed in two stages.
-First, $\mathcal{H} = \mathcal{F} +_{f_{ext}, g_{ext}} \mathcal{G}$ is computed,  as shown below.
-
+Composition of two morphisms 
+\begin{align*}
+	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
+	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
+\end{align*} is computed in two stages.
+First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
 % https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-
 \[
@@
 \]
-Then, new feet for $\mathcal{H}$ are computed as below
-\[
-n \xrightarrow{f_{ext};\iota_1} n' \uplus (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' \uplus (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where
-\[
-    h_1 = [ f_{int};p_1, ~g_{int}|_{m'' \setminus m};p_2 ]
+then,  the result of composition is defined as follows:
+\[
+n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
+\]
+where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
+\[
+    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
 \qquad
-    h_2 = [ g'_{int};p_2, ~f'_{int}|_{m' \setminus m};p_1 ]
+    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
 \]
@@
 \end{definition}
-CHRIS: UPDATE DIAGRAMS TO USE MY NEW NOTATION,  TAKE LESS VERTICAL SPACE.  FIND CLEANER WAY TO WRITE NEW FEET.  USE $\iota$ FOR INJECTIONS
 
@@
 
-CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
-Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
-The former means that we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching external interfaces.
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\qquad \text{and} \qquad
-    i \xrightarrow{} i'' \xrightarrow{} \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j
-\]
-We will next use the fact that cospans
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\qquad \text{and} \qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
-\]
-endcode the same data and only consider cospans of the form
-\[
-\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] to fit the DPO formalism.
-We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
-% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
-% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
-\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
-When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
-Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
-In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
-\[
-    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
-\]
-not necessarily coincides with the internal interfaces of the initial cospan.
+Because extended cospans have a more general notion of interface,  including \textit{internal} nodes,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
+This is the purpose of this section of the paper.
+
+We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
+Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessaril internal) interfaces,  as follows. 
+\[
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
+\qquad
+    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
+\] 
+Analogous to Section \ref{sec:combinatorial-semantics},  we observing that the following extended cospans express the same data
+\[
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
+\qquad
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
+\] allows us to encode rewrite rules as extended cospans of the following form
+\[
+\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
+\] which will fit into the DPO formalism. We make this identification throughout,  without further mention . 
+%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
+
+\begin{remark}
+There is a limitation of EDPOI rewriting as defined in this section: it is unable to express rewriting
+for rules whose target contains an edge with no incident vertices.  
+While this limitation can be overcome,  we find the technical development (relating to the child and consistency relations) to be significantly simpler if we simply avoid consideration of such edges.
+For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
+%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
+%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
+%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
+\end{remark}
+
 % The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
@@
 
-To account for such changes in the interfaces, consider additional structure on DPOI rewriting for $\catname{EHyp({\Sigma})}$, summarised in the definition below.
-\begin{definition}[Extended DPOI (EDPOI) rewriting]
-\label{def:dpoi-e}
-Given a span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites to} $\mathcal{H}$ \textit{with external interface} $n + k$ \textit{via a rewrite rule} 
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\]
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
-% \[\begin{tikzcd}
-% 	&&&& {i + j} \\
-% 	{l_i+l_o} && L &&&& R && {r_i+r_o} \\
-% 	\\
-% 	{g_i+g_o} && {G^{\urcorner}} && C && {^{\ulcorner}H} && {h_i+h_o} \\
-% 	\\
-% 	&&&& {n+m}
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-3]
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=2-7]
-% 	\arrow["m", color={rgb,255:red,214;green,92;blue,92}, from=2-3, to=4-3]
-% 	\arrow["{[i_c,j_c]}", color={rgb,255:red,214;green,92;blue,92}, from=1-5, to=4-5]
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=2-7, to=4-7]
-% 	\arrow["{[n_c,m_c]}"', color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-5]
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-7]
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=4-5, to=4-3]
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-3]
-% 	\arrow[color={rgb,255:red,214;green,92;blue,92}, from=6-5, to=4-7]
-% 	\arrow[from=2-1, to=2-3]
-% 	\arrow["{[f_i,f_o]}", from=2-1, to=4-1]
-% 	\arrow[from=4-1, to=4-3]
-% 	\arrow[from=2-9, to=2-7]
-% 	\arrow[from=4-9, to=4-7]
-% 	\arrow["{[h_i,h_o]}"{description}, from=2-9, to=4-9]
-% 	\arrow[from=6-5, to=4-9]
-% 	\arrow[from=6-5, to=4-1]
-% 	\arrow["{[i_l,j_l]}"{description}, from=1-5, to=2-1]
-% 	\arrow["{[i_r,j_r]}"{description}, from=1-5, to=2-9]
-% \end{tikzcd}\]
-
-% https://q.uiver.app/#q=WzAsMTEsWzIsMCwiTCJdLFs0LDAsInhfe2V4dH0gKyB5X3tleHR9Il0sWzIsMiwiR157XFx1cmNvcm5lcn0iXSxbNCwyLCJDIl0sWzQsNCwibl97ZXh0fSttX3tleHR9Il0sWzAsMCwibF9pK2xfbyJdLFswLDIsImdfaStnX28iXSxbNiwwLCJSIl0sWzYsMiwiXntcXHVsY29ybmVyfUgiXSxbOCwyLCJoX2kraF9vIl0sWzgsMCwicl9pK3JfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXSxbNCw4LCIiLDAseyJjb2xvdXIiOlswLDYwLDYwXX1dLFszLDgsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzcsOCwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbMSw3LCIiLDIseyJjb2xvdXIiOlswLDYwLDYwXX1dLFsxMCw3XSxbOSw4XV0=
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-\end{definition}
-To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph boundary complement also requires a revised definition.
-
-\begin{definition}[Boundary complement in $\WellTypedMdaEcospans$]
+To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph we must revise the definition of boundary complement.
+
+\begin{definition}[Extended boundary complement]
 \label{def:boundary_new}
@@
     n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$, a pushout complement 
-\[
-    i + j \to \mathcal{L}^{\bot} \to \mathcal{G}
-\] 
+\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
 as depicted in the square below
@@
 \]
-
-is a \textit{boundary} complement if the following holds
-\begin{enumerate}
-    \item $[c_1,c_2]$ is mono 
-    \item there exist $d_1 : n \to \mathcal{C}$ and $d_2 : k \to \mathcal{C}$ making the above triangle commute
-    \item there exists a \textit{not necessarily well-typed} mda-cospan
-    \[
-    n \xrightarrow{} l + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} r + i \xleftarrow{} k
-    \]
-\end{enumerate}
-
-where $l$ and $r$ are
-\[
-    l = n' \setminus (i' \setminus i)
-\quad\text{ and }\quad
-    r = k' \setminus (j' \setminus j).
-\]
+is a \textit{boundary} complement if $[c_1,c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute,  such that
+    \[
+    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
+    \]
+is a \textit{not necessarily well-typed} extended mda-cospan.
+
+CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
 Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
@@
 
-The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
+%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
+%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
+%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
 %
@@
 
-\begin{remark}
-Note that cospan 
-\[
-    n \xrightarrow{} l + j \xrightarrow{} C \xleftarrow{} r + i \xleftarrow{} k
-\] is not necessarily well-typed.
-The cospan's input internal interface may contain nodes that previously were a part of the output internal interface and similarly for output internal interface.
-\end{remark}
-
-\begin{remark}
-The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-\end{remark}
+Note that,  in the above definition,  the extended cospan 
+    \[
+    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
+    \]
+ is not necessarily well-typed.
+This is because its input internal interface may contain nodes that previously were a part of the output internal interface and vice versa. 
+%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
 
@@
 
-\begin{definition}[Subgraph]
-
-A \textit{subgraph} of an e-hypergraph 
-\[
-    \mathcal{G} = \{V_{\mathcal{G}}, E_{\mathcal{G}}, s_{\mathcal{G}}, t_{\mathcal{G}}, l, <_{G},\consistency_{\mathcal{G}}\}
-\]
-is an e-hypergraph 
-\[
-    \mathcal{H} = \{V_{\mathcal{H}}, E_{\mathcal{H}}, s_{\mathcal{H}}, t_{\mathcal{H}}, l, <_{H},\consistency_{\mathcal{H}}\}
-\] such that $V_{\mathcal{H}} \subseteq V_{\mathcal{G}}$, $E_{\mathcal{H}} \subseteq E_{\mathcal{G}}$ and such that $<_{\mathcal{H}}$ and $\consistency_{\mathcal{H}}$ both satisfy the properties from~\ref{def:e-homo}.
-\end{definition}
+The appropriate notion of \textit{match} for our hierarchical setting additionally requires that its image is a \textit{down-closed} graph,  as defined below. 
+This ensures the result of a rewrite is a well-typed MDA e-hypergraph.
 
 \begin{definition}[Down-closed subgraph]
-    We call a subgraph $\mathcal{H}$ of an e-hypergraph $\mathcal{G}$ \emph{down-closed} if for all $e \in E_{\mathcal{H}}$,   all children of $r$ are also in $\mathcal{H}$.
-\end{definition}
-
-\begin{definition}[Convex down-closed match] 
-    We call $m : L \to G$ in $\catname{EHyp(\Sigma)}$ a \textit{convex down-closed match} if
-
-    \begin{enumerate}
-        \item $m$ is a monomorphism
-        \item $m(L)$ is a convex down-closed subgraph of $G$
-        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
-        \item $\consistency^{\mu}(m(v_i)) = \consistency^{\mu}(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{Z}$.
-    \end{enumerate}
-\end{definition}
-
-The conditions (3)-(4), in particular, prohibit a match $\{f \mapsto f, g \mapsto g\}$ in $\mathcal{L} =$ $\scalebox{0.4}{\tikzfig{combinatorial_semantics/f_times_g}}$ and $\mathcal{G} =$ $\scalebox{0.4}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}$.
-
-We are now ready to define the \textit{convex} rewriting for e-hypergraphs with interfaces which we will later use to show the equivalence between $\textsf{PROP}(\Sigma)$ and $\WellTypedMdaEcospans$.
-It will be analogous to the one from Definition~\ref{def:convex_dpo} but again, to make sure the resulting cospan of e-hypergraphs is monogamous we will have to construct internal interfaces separately.
-We will also limit matches to convex down-closed ones both to guarantee an mda result and so that the result is a well-defined e-hypergraph.
-
+%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
+%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
+%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
+    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
+\end{definition}
+
+We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.
+for e-hypergraphs with interfaces,  which we will later use to show the equivalence between $\textsf{PROP}^+(\Sigma)$ and $\WellTypedMdaEcospans$.
+It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
+%To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
+% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
+% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
+%\[
+%    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
+%\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
+
+More precisely, when removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
+Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
+
+CHRIS: TODO: COMBINE TWO EDPOI DEFINITIONS
 \begin{definition}[Convex EDPOI rewriting]
+
+\label{def:dpoi-e}
+Given a span of morphisms 
+\[
+    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
+\]
+in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites to} $\mathcal{H}$ \textit{with external interface} $n + k$ \textit{via a rewrite rule} 
+\[
+    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
+\]
+if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
+\[
+ \scalebox{0.75}{
+    \tikzfig{combinatorial_semantics/DPOI_square}
+ }
+\]
+
     Let $\mathfrak{R}$ be a set of EDPOI rewrite rules.
@@
         \label{dpoi-e:assumptions}
-        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex down-closed match;
+        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
+        \item $m(L)$ is a down-closed e-hypergraph;
+        \item all $v,w$ that have a pre-image in $i+j$ share the same set of parents;
+        \item $m(v)  \consistency m(w)$ for all $v,w$ that have a pre-image in $i+j$.
         \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement in the leftmost pushout.
     \end{enumerate}
-\end{definition}
 and
@@
 Similarly for $k'' \to \mathcal{H}$.
-
-\begin{proposition}
-    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-\end{proposition}
-
-Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$ we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-
-\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-
-We say that an \textit{mda-cospan}
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\]
-\textit{rewrites convexly into another mda-cospan}
-
-\[
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-\]
-
-under a rewrite rule 
-% \[
-%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-% \]
-
-\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-
-if 
-\begin{align*}    
-    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-\end{align*}
-under the same rewrite rule.
-We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-
-\end{definition}
-
-\begin{remark}
-There is a minor limitation for EDPOI rewriting above in that it is unable to express rewriting
-for rules that contain an edge with no inputs and outputs on their right-hand side (such edges correspond to morphisms of type $0 \to 0$).
-This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
+\end{definition}
+The conditions (3) and (4) on a match, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+\[
+	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
+\]
+
+With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense. 
+%\begin{proposition}
+%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
+%\end{proposition}
+
+%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
+%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
+%
+%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
+%    
+%We say that an \textit{mda-cospan}
+%\[
+%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
+%\]
+%\textit{rewrites convexly into another mda-cospan}
+%
+%\[
+%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
+%\]
+%
+%under a rewrite rule 
+%% \[
+%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
+%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
+%% \]
+%
+%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
+%
+%if 
+%\begin{align*}    
+%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
+%\end{align*}
+%under the same rewrite rule.
+%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
+%
+%\end{definition}
+%
 
@@
 
-CHRIS: NOTE ALSO WEAK NORMAL FORM FOR $\Sigma^+$-TERMS?
 The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
@@
 \end{proposition}    
+Note that at similar argument can be used to rewrite $\Sigma^+$-terms into a similar form by using only the semilattice-enrichment equations.
 
@@
 \begin{align*}
-	\llbracket f \rrbracket &= \llbracket f_1 + \ldots + f_n \rrbracket &&= \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \\
-	\llbracket g \rrbracket &= \llbracket g_1 + \ldots + g_m \rrbracket &&= \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
+	\llbracket f \rrbracket = \llbracket f_1 + \ldots + f_n \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \\
+	\llbracket g \rrbracket = \llbracket g_1 + \ldots + g_m \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
 \end{align*}
@@
 \]
-\caption{E-graphs for functorial boxes and the ZX-calculus
-DAN: THE LINE BETWEEN THE APPLICATION NODE AND THE BOX IS TOO LONG}
+\caption{E-graphs for functorial boxes and the ZX-calculus}
 \label{fig:app1} 
@@
 However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal}.
+Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
 Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.

### 2024-01-17 E-Hypergraphs-LICS24.tex
S6 update

@@
 
-We use the notation $n \setminus m$ to denote the discrete hypergraph with $n - m$ nodes,  in particular with removal of the nodes of $m$ removed from $n$ when nodes of $m$ are identified within $n$.
+We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ nodes,  in particular with the nodes of $m$ removed from $n$ when nodes of $m$ is a sub-e-hypergraph of $n$.
 \begin{definition}[Category of e-hypergraphs with interfaces]
@@
 We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n, n', m$ and $m'$ with their images in $\mathcal{G}$. 
+We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
 Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
@@
 
-CHRIS: INTRO.
+This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
+We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
+As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
+Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
 
@@
 
-We wish to specify a set of DPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
+We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
+In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
 \begin{definition}[Interpretation function] 
@@
 
+In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
+This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
+In fact,  we prove that the functor is also full,  giving an equivalence of categories.
+
+We define the appropriate notion of quotient by EDPOI rewrites below.
 Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
 % \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-CHRIS: COMMENT THAT WE QUOTIENT TO PHRASE CATEGORICALLY OUR RESULTS. 
 \begin{definition}[Quotient by rewrites]  
@@
 \[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{+,*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-In particular,  let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-This quotient in fact turns the interpretation function into a proper functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which can further be quotiented to account for extra equations $\mathcal E$ arising in symmetric monoidal theories,  as below.
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
+	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
+\]
+\end{definition}
 
 Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
+Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
 %Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Thus we have the following result. 
+Then we have the following result. 
 \begin{proposition}[Soundness]
@@
 \end{proposition}
-It immediately follows that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is a semilattice-enriched PROP,  and $\llbracket - \rrbracket_{\mathcal E}$ is the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
 Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
+
+The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
+\[
+	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
+\]
+which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
+Unfolding definitions,  it follows that
+\[
+	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
+\]
+Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
 
@@
 \end{proposition}    
-Note that at similar argument can be used to rewrite $\Sigma^+$-terms into a similar form by using only the semilattice-enrichment equations.
+The proof follows the familar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
 
 This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Completeness]
+\begin{theorem}[Full completeness]
 We have the following equivalence of categories
@@
 
-Unfolding the definitions,  we can spell out the previous theorem in terms of DPO-rewriting. 
+Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
 \begin{corollary}[Completeness of rewriting]
@@
 \[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
+	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
 \]
@@
 %\end{theorem}
-CHRIS: CONCLUDING STATEMENT.  COMMENT ABOUT GENERALIZATION TO DIRECTED REWRITES.
+Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
+We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
 

### 2024-01-17 E-Hypergraphs-LICS24.tex
Minor fixes and S6 update

@@
 \end{figure}\label{fig:egraph-strings}
-CHRIS: DONT USE PHI/PSI FOR DIAGRAMS, USE F,G
 
@@
 \[
-\infer{\phi_1 + \ldots + \phi_n: A \to B}{\{\phi_i: A \to B\}_{i \in \{1,\ldots,n\}}}
+\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
 \]
@@
 \begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with homsets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
+    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with homsets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
     \[
@@
     where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and that that vertices in the strictly internal interface (defined below) are not top-level.
+We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
 
@@
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessaril internal) interfaces,  as follows. 
+Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
 \[
@@
 
+Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph.
+
 \begin{remark}
-There is a limitation of EDPOI rewriting as defined in this section: it is unable to express rewriting
+There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
 for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development (relating to the child and consistency relations) to be significantly simpler if we simply avoid consideration of such edges.
+While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
 For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
@@
 % Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-To guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph we must revise the definition of boundary complement.
 
@@
 We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-for e-hypergraphs with interfaces,  which we will later use to show the equivalence between $\textsf{PROP}^+(\Sigma)$ and $\WellTypedMdaEcospans$.
 It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
@@
 
-CHRIS: TODO: COMBINE TWO EDPOI DEFINITIONS
+CHRIS: FIX BELOW DEFINITION. 
 \begin{definition}[Convex EDPOI rewriting]
-
 \label{def:dpoi-e}
-Given a span of morphisms 
+Given an extended span of morphisms 
 \[
@@
 \]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites to} $\mathcal{H}$ \textit{with external interface} $n + k$ \textit{via a rewrite rule} 
+in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
 \[
@@
 \]
-
-    Let $\mathfrak{R}$ be a set of EDPOI rewrite rules.
-    Then, given 
-    \[
-        \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
-    \quad 
-    \text{and} 
-    \quad
-        \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-    \] in $\catname{EHyp(\Sigma)}$ with the external interfaces $n + k$, $\mathcal{G}$ \textit{rewrites convexly into} $\mathcal{H}$ \textit{with the outermost interface} $n + k$ --- notation $( \mathcal{G} \xleftarrow{} n + k) \Rrightarrow_{\mathfrak{R}}^{+} (\mathcal{H} \xleftarrow{} n + k)$ --- if there exists a rule
-    \[
-        \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-    \] 
-    with the external interfaces $i + j$ in $\mathfrak{R}$ and object $\mathcal{C}$ and cospan arrows $i + j \xrightarrow{} \mathcal{C} \xleftarrow{} n + k$ 
-    such that the diagram of Definition~\ref{def:dpoi-e} commutes and its marked squares are pushouts, and the following conditions hold
-
+ and the following conditions hold:
     \begin{enumerate}
@@
         \item $m(L)$ is a down-closed e-hypergraph;
-        \item all $v,w$ that have a pre-image in $i+j$ share the same set of parents;
-        \item $m(v)  \consistency m(w)$ for all $v,w$ that have a pre-image in $i+j$.
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement in the leftmost pushout.
-    \end{enumerate}
-and
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
+        \item all elements in the image of $i+j$ share the same set of parents;
+        \item $m(v)  \consistency m(w)$ for all $v,w$ in the image of $i+j$.
+        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
+	\item the internal interfaces are given by the following:
 		\begin{align*}
@@
 		\end{align*}
+    \end{enumerate}
+% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
+%  \[
+%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
+% \]
 Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
 Similarly for $k'' \to \mathcal{H}$.
+
+    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
 \end{definition}
@@
 With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
+
 %\begin{proposition}
@@
 \begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket \phi ; \psi \rrbracket_{n,m} &= \llbracket \phi \rrbracket_{n,m} ; \llbracket \psi \rrbracket_{n,m} \\
+	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
 	\llbracket 1 \rrbracket_{n,m} &= 1 & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\

### 2024-01-17 E-Hypergraphs-LICS24.tex
EDPOI rewrite defn

@@
 \] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
+if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
 \[
@@
 \]
- and the following conditions hold:
+such that the two marked squares are pushouts and the following conditions hold:
     \begin{enumerate}

### 2024-01-17 E-Hypergraphs-LICS24.tex
minor fixes

@@
 
-CHRIS: FIX BELOW DEFINITION. 
 \begin{definition}[Convex EDPOI rewriting]
@@
         \item $m(L)$ is a down-closed e-hypergraph;
-        \item all elements in the image of $i+j$ share the same set of parents;
-        \item $m(v)  \consistency m(w)$ for all $v,w$ in the image of $i+j$.
         \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
@@
 % \]
-Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
+Since both $f_1$ and $f_2$ are monos with disjoint image (?), their copairing is also a mono. 
 Similarly for $k'' \to \mathcal{H}$.
+CHRIS: FIX (4) ABOVE:
 
@@
 \end{definition}
-The conditions (3) and (4) on a match, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-
+%The conditions above, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+%\[
+%	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
+%\]
+%We use 
 With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
@@
 \end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.  We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context. 
+where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
+It uses the notation of dashed blue boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components seperated by another dashed line.
+We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
 \end{definition}

### 2024-01-17 E-Hypergraphs-LICS24.tex
completeness proof update.

@@
 These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ in any hypergraph $G$ and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{???} is still applicable.
+Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
+The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
 However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
@@
 
-Precisely,  our soundness and completeness result is the statement that,  for any morphisms $f$ and $g$ in an appropriate free semilattice-enriched SMC,  $f = g$  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
+Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
 In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
@@
 
+% E-graphs
 Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesizers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
@@
 
-Our use of string diagrams is also central to the intuition behind our work.
+% String diagrams
+Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
 There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}. 
-DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.
+% Hierarchical hypergraphs
+Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
+% String diagram rewrites
+Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
 
@@
 Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
+The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
 We overload the binary notation $+$ for our $n$-ary notation.
@@
 and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-
 % \begin{figure}[h!]
@@
 % \end{figure}
-
 % https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
@@
 % \end{tikzcd}\]
-
 \[
@@
 \]
-
 is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
@@
 The appropriate notion of \textit{match} for our hierarchical setting additionally requires that its image is a \textit{down-closed} graph,  as defined below. 
-This ensures the result of a rewrite is a well-typed MDA e-hypergraph.
-
 \begin{definition}[Down-closed subgraph]
@@
         \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces are given by the following:
+	\item the internal interfaces of $\mathcal H$ are such that:
+	\[
+		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
+	\]
 		\begin{align*}
-		&n'' = (i'' \setminus i) + l\\
-		&k'' = (j'' \setminus j) + r\\
-		&n'' \to \mathcal{H} = [f_1,f_2] \text{ where }\\
-		&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H}\\ 
-		&f_2 : l \to \mathcal{C} \to \mathcal{H} 
+		&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H} & \\ 
+		&f_2 : n' \setminus (i' \setminus i) \to \mathcal{C} \to \mathcal{H} & \\
 		\end{align*}
@@
 % \]
-Since both $f_1$ and $f_2$ are monos with disjoint image (?), their copairing is also a mono. 
-Similarly for $k'' \to \mathcal{H}$.
+
+$[f_1,f_2]: n'' \to \mathcal{H} $
+Since both $f_1$ and $f_2$ are monos with disjoint image (?), their copairing is also a mono.  Note $f_1$ is the obvious restriction,  $f_2$ is defined as $g_2$ in boundary complement. 
 CHRIS: FIX (4) ABOVE:
 
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-%The conditions above, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-%\[
-%	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-%\]
-%We use 
+    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
+\end{definition}
+The boundary complement conditions, in particular, prohibits finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+\[
+	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
+\]
+Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
+
 With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
@@
 First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$ as a subcategory of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$ (note that cospans embed in extended cospans when the internal interface is isomorphic to the external interface); see Theorem 35 of ~\cite{bonchi_string_2022-2}.
+In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
 
@@
 First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we see that
-\begin{align*}
-	\llbracket f \rrbracket = \llbracket f_1 + \ldots + f_n \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \\
-	\llbracket g \rrbracket = \llbracket g_1 + \ldots + g_m \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\end{align*}
-In particular,  note that $\llbracket - \rrbracket$ preserves weak normal forms.
-Therefore $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-Thus if $n \neq m$ we have that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $.
-Otherwise,  if $n = m$,  we have that there for every permutation $\sigma$ there exists an index $i$ such that $f_i \neq g_{\sigma(i)}$,  and thus by $(*)$ that $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $.
-This implies that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $.
-
-To see that the functor is full,  consider a cospan $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ is a plain MDA cospan.
-Thus,  by $(*)$ we can find morphisms $g_i$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we have a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
+Applying the functor $\llbracket - \rrbracket$,  we thus have that
+\[
+	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
+	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
+\]
+Observe  that therefore $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
+We conisder two cases: where $n=m$ and where $n\neq m$.
+In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since their weak normal forms cannot be equal.
+Otherwise,  if $n = m$,  and $\llbracket f \rrbracket = \llbracket g \rrbracket$,  we have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
+
+To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
+We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
+Thus,  by $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
+Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
 \end{proof}
@@
 In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,cite,cite,cite}.
+Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
 % traced comonoid

### 2024-01-17 E-Hypergraphs-LICS24.tex
Def 5.1 update

@@
 % \]
-CHRIS: WHY IS J DISPLAYING AND NOT R? 
+% CHRIS: WHY IS J DISPLAYING AND NOT R? 
 \[
@@
 \] 
-Analogous to Section \ref{sec:combinatorial-semantics},  we observing that the following extended cospans express the same data
+Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
 \[
@@
 \]
-is a \textit{boundary} complement if $[c_1,c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute,  such that
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
-is a \textit{not necessarily well-typed} extended mda-cospan.
-
-CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-\[
-    D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-\]
-Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-Similarly for $k \to r$.
-\end{definition}
+is a \textit{boundary} complement if
+\begin{enumerate} 
+    \item the image of $m$ is a convex down-closed e-hypergraph
+    \item $[c_1,c_2]$ is mono
+    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute
+    \item all $v,w$ in $\mathcal{G}$ with the pre-image in $i + j$ share the same set of parents
+    \item either for all $v,w$ in $\mathcal{G}$ with the pre-image in $i+j$, $v \consistency w$, or for all $v,w$ in $\mathcal{G}$ with the pre-image in $i+j$ $\consistency$ is undefined
+    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
+    \[
+    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+    \]
+    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
+    \[
+    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+    \]
+\end{enumerate}
+
+We first note that because the image of $m$ is a convex down-closed e-hypergraph, $i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono.
+Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
+
+\[
+    \tikzfig{combinatorial_semantics/pushout_difference}
+\]
+and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
+Similarly for $g_2$.
+$f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$ --- noting the image of $g_{ext}$ lies within $n' \setminus (i' \setminus i)$.
+Similarly for $f_2$.
+% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
+% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
+% \[
+%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
+% \]
+% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
+% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
+% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
+% Similarly for $k \to r$.
+\end{definition}
+
+Intuitively the above definition means that when the occurence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
+In both cases when removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
+The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
 
@@
 \begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, i.e. on the coproduct of nodes and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either node or edge into these functions.
-    For example, an immediate successor of a node $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
+    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of verticies and edges.
+    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
+    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
 \end{remark}
@@
 
-            \item When $[x) \not = \varnothing$, \[
+            \item When $x$ is not a top-level vertex, 
+                \[
                 \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \] when $x$ is a node,
+                \]
                 and
@@
                 \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is an edge.          
+                \] when $x$ is a not top-level edge.          
                 \item
@@
             \]
-            , where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
             And, when $x \in V_{\mathcal{F}}$
@@
     \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency^{\mu}(f(v_i)) = \consistency^{\mu}(f(v_j))$ and $\consistency^{\mu}(g(v_i)) = \consistency^{\mu}(g(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
+    \item $\consistency(f_{V};\iota_{V_{X}} (v_i)) = \consistency(f_{V};\iota_{V_{X}}(v_j))$ and $\consistency(g_{V};\iota_{V_{Y}}(v_i)) = \consistency(g_{V};\iota_{V_{Y}}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
 \end{enumerate}    
@@
     \[
+        \adjustbox{scale=1.25}{
             \begin{tikzcd}
@@
                                                                             &                                                                  & Q
-        \end{tikzcd}
+            \end{tikzcd}}
     \]
@@
     \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{c,X+Y} \}
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
     \]
@@
           &\}
+        \\
+        \\
+    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
     \end{align*}
-    \begin{align*}
-    S_{E} &= \{\\
-    &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    &\}
-    \end{align*}
-and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-    We then quotient the sets of edges and nodes in $X + Y$ by $R_{V},R_{E}$
+    % \begin{align*}
+    % S_{E} &= \{\\
+    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
+    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
+    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
+    % &\}
+    % \end{align*}
+and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
+
+We then quotient the set verticies and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
     \begin{align*}
         \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim R_{V}}, \sfrac{E_{X} + E_{Y}}{\sim R_{E}}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \#_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{c,\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
         & \}    
     \end{align*}
-    We will then refer to $\sim (R_{V},R_{E})$ just as $\sim$, i.e. writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and the exact relation will be clear from the context.
+We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
 We have 
@@
 and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-We then define $s_2 : E_{X} + E_{Y} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*} = s_{X+Y};[-]^{*}$ and then $\sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}$.
 There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
+We then define 
+\[
+    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
+\]
 We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, $t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}$.
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets.
+Likewise, 
+\[
+    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
+\]
+
+So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
 
@@
         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
     We will consider several cases.
-    Assume that $u$ is an element of $V_{X+Y}$ and
+    First, assume that $u$ is an element of $V_{X+Y}$ and
     \begin{enumerate}
@@
         %       We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$ $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z)))$.
+        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
+              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
               We then define 
               \[
-                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [i_{2,E}(e_y)]
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
               \]
         \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{c,X}^{\mu}(i_{V_{X}}(f_{V}(z)))$.
+              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
               We then define 
               \[
-                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))$ is undefined (i.e.
-$[u) = \varnothing$), but there exists $[e]$ such that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e']$ and $[u] \in t([e])$ or $[u] \in s([e])$, then
-              \[ 
-                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
-              \]
-              Assumptions above guarantee that the predecessor of such node does not depend on the choice of $e$ if there are several of them.
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
+              \]
+        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
+        \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
+        \]
+        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
+              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
+              \[ 
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[v])
+              \]
     \end{enumerate}
     % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, if $u$ is an element of $E_{X+Y}$.
+    Now, assume that $u$ is an element of $E_{X+Y}$.
     We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
@@
     \begin{enumerate}
-                \item  $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u)) = u_1$.
+    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
             Then we define
         \[
-                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(u))]
-                    \]
-                \item $[u) = \varnothing$ and for some $[v_i] \in s([u])$ and $[v_j] \in t([u])$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.
-                      Then we define 
-                    \[
-                        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                    \] where $[v]$ is either in $s([u])$ or $t([u])$.
-                     The second assumption in~\ref{pushout:assumptions} guarantees that all candidates for $[v]$ share the same predecessor, i.e. the definition does not depend on a particular $[v]$.
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
+        \]
+    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
+        \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+        \] 
     \end{enumerate}
-    We will also build this relation in steps, first we do steps for nodes (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for nodes until fixed point.
-    Essentially, this interleaving means that nodes inherit the relation from incident edges and vice versa.
+    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
+    % Essentially, this interleaving means that verticies inherit the relation from incident edges and vice versa.
     Clearly, all the cases above are disjoint.
 
+    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
+    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
+    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
+    Then consider sevaral cases.
+    \begin{itemize}
+        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
+        \begin{itemize}
+            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
+                  Hence, 
+                  \[
+                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
+                  \]
+                  and
+                  \[
+                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
+                  \]
+                  Then, $[e_1] \not = [e_2]$ implies
+                  \[
+                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
+                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
+                  \[
+                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
+                  \]
+                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
+                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
+                \end{itemize}
+                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
+                \begin{itemize}
+                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
+                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
+                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
+                \end{itemize}
+                \item The remaining two cases are symmetric.
+    \end{itemize}
+
     Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are nodes and $t_1 \sim t_2$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};i_{1,V}(z)$ and $t_2 = g_{V};i_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(t_2)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(t_2)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case, then, by definition.
-    \[
-        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
+    Suppose $t_1, t_2$ are verticies and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
+    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
+    \begin{itemize}
+        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
+        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
+    \end{itemize} 
+    Consider the first case. Then, by definition.
+    \[
+        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
     \]
@@
     \[
-        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
+        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
+    \]
+    so we have well-defined-ness on the nose. 
+    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
     % \begin{itemize}
@@
 
-    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms.
+    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
     \begin{enumerate}
-        \item $s_{X+Y};[]_{V}^{*}(e) = []_{E};s_{\sfrac{X+Y}{\sim}}(e)$
-              This is the same as in $\catname{Hyp_{\Sigma}}$.
-        \item $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
+         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
               Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
               This is homomorphic by definition. 
-        \item $[<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
+        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
               Let's make case analysis on $[v]_{V}$.
               \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};i_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $i_{1,V}$ is a coproduct injection, it must be the case that $[i_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
+                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
+                    \begin{itemize}
+                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
                         \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}[v]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
-                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V};i_{V_{X} + V_{Y}}(z))] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
-                    \end{itemize}
-                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+                              by definition, we have 
+                              \begin{align*}
+                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
+                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
+                                &= [\iota_{1,E}(e_x)] ~.
+                            \end{align*}
+                            Because $\iota_{1,E}$ is a homomorphism, we also have 
+                            \[
+                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
+                            \]
+                    \end{itemize}
+                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
                 \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
@@
     \end{enumerate}
-    Then $<_{c,\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$.
-    We can consider the original consistency relation as a function $\consistency_{X+Y}^{\mu} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
-    Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
+    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
+
+    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
+    We can consider the consistency relation from the coproduct as a function 
+    \[
+        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
+    \]
+    Quotienting the values of the function gives us 
+    \[
+        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
+     \]
+    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
     We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
     \begin{enumerate}
-        \item If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
+        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
                 \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    \langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)))
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
+                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
                 \end{align*}
-        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{V_{X} + V_{Y}}([v]_{V}))
+        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$, then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
         \]
@@
 
-    and for nodes
+    and for verticies
     \begin{enumerate}
 
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')))
-    \]
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{E_{X} + E_{Y}}([e]_{E}))
+    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v')))
+    \]
+    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
     \]
     \end{enumerate}
-    where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
+    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
 
@@
     \[
-      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
+      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
+    \]
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
     \]
@@
 
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency^{\mu}$.
+    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
 
@@
 
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_2])$.
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};i_{1,V}(z)$ and $v_2 = g_{V};i_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
+            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
+            \[
+                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
+            \]
+              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
               \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
                 %      \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
@@
               \[
-                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
+                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
               \]
@@
               We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$.
-                  This would mean that for all $v_i \in V_{Z}$ $\consistency^{\mu}(f_{V}(v)) = \varnothing$ and $\consistency^{\mu}(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+            \item Now suppose that both 
+            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
+            \quad 
+            \text{and} 
+            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
+            \]
+            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
+                  \[
+                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
                   \]
@@
         \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \[
-             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
-             \subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))
-             \]
+            \begin{align*}
+             [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(&\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\subseteq
+             &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
+            \end{align*}
             By definition, 
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
-            \]
-            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion $\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ because it is a closure.
-            If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness is true because empty set is a subset of any set.
+            \begin{align*}
+                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
+                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
+            \end{align*}
+            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
+            \[
+                []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
+            \] and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
+            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
             Similarly for edges.
@@
 
+    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
     Let's now prove the universal property.
-    Let $u_{E}([i_{1,E}(x)]) = j_{1,E}(x)$ and $u_{E}([i_{2,E}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,E}$ and $j_{2,E}$.
-
-    Let's say that $t_1 \sim t_2$ then $u_{E}([t_1]) = u_{E}([t_2])$.
-    $u_{E}([t_1]) = u_{E}(f_{E};i_{1,E}(z)) = j_1{f_{E}(z)}$ and $u_{E}([t_2]) = u_{E}(g_{E};i_{2,E}(z)) = j_{2}(g_{E}(z))$ which are equal by commutativity.
+    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
+    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
+
+    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
+    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
+    Then, we have
+    \[
+    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
+    \] and 
+    \[
+    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
+    \] which are equal by commutativity.
 \end{proof}
 
-\begin{lemma}
-    \label{lemma:choice_of_nodes_agnostic}
-    Suppose $[e) = \varnothing$ and there exists $[v_1] \in s([e])$ and $[v_2] \in s([e])$ such that $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2]))$.
-    \begin{proof}
-        We will show the statement is true for the base case when $[v_1]$ and $[v_2]$ do not inherit their relations.
-        In other words, $[[v_1]) \not = \varnothing$ and $[[v_2]) \not = \varnothing$ mean that was achieved after running steps (1)-(3) for nodes.
-        From $[v_1] \in s([e])$ we get that there exists $v'_1$ such that $v_1 \sim v'_1$ and $v'_1 \in s(e)$ and similarly $v'_2 \sim v_2$.
-        \begin{itemize}
-            \item $v_1 = v'_1$, $v_1 \in s(e)$, and $v_2 = v'_2$, $v_2 \in s(e)$.
-            \begin{itemize}
-                \item $v_1$ has a pre-image $f_{V};i_{1,V}(z)$ and hence $[f_{V}(z)) = \varnothing$ (because $[e) = \varnothing$) and $v_2$ has a pre-image $f_{V};i_{1,V}(z')$.
-                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{2,E}(e_{y})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e'_{y})]$.
-                      Because $e_{y} = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z)))$ and $e'_{y} = <_{c,Y}^{\mu}(i_{V_{Y}}(g_{V}(z')))$ we have $e_{y} = e'_{y}$ because under our assumptions nodes in the image of $g_{V}$ should share the same predecessor.
-                      Hence, $[i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})]$.
-                \item Case for $v_1 = g_{V};i_{2,V}(z)$ is symmetric.
-                \item Case when $v_1$ has no pre-image in $V_Z$ is not possible, because it would mean that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1))]$ (according to step (3)) which is not possible as it would mean that $[e) \not = \varnothing$ (because $[v_1) \not = \varnothing$).
-            \end{itemize}
-            \item $v_1 \not = v'_1$ and $v'_2 \not = v_2$ and hence $v'_1 \in s(e)$ and $v'_2 \in s(e)$.
-            \begin{itemize}
-                \item $v'_1$ has a pre-image $f_{V};i_{1,V}(z)$ and hence $[f_{V}(z)) = \varnothing$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) = [i_{2,E}(e_{y})]$.
-                      $v'_2$ has a pre-image $f_{V};i_{1,V}(z')$ and for the same reason 
-                      \begin{align*}
-                        &<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_1])) =\\ 
-                        &[i_{2,E}(e_{y})] = [i_{2,E}(e'_{y})] =\\
-                        &<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                    \end{align*}
-                \item Case when $v'_1 = g_{V};i_{2,V}(z)$ is similar.
-            \end{itemize}
-            \item $v_1 = v'_1$, $v'_2 \sim v_2$ and $v_1 \in s(e)$, $v'_2 \in s(e)$.
-            \begin{itemize}
-                \item $v_1 = f_{V};i_{1,V}(z)$ and hence $v'_2 = f_{V};i_{1,V}(z')$ and $[f_{V}(z)) = \varnothing$ as well as $[f_{V}(z')) = \varnothing$.
-                      Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{2,E}(e_{y})]$ and similarly $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e'_{y})]$.
-                      And by the following the same reasoning as above we get the equality.
-                \item Case when $v_1 = g_{V};i_{2,V}(z)$ is similar.
-            \end{itemize}
-            \item The remaining case is similar.
-        \end{itemize}
-        The proofs for the cases when $[v_1] \in t([e])$ and $[v_2] \in t([e])$ and for the combinations of $t,s$ are similar.
-    \end{proof}
-\end{lemma}
-
-\begin{lemma}
-    \label{lemma:choice_of_edges_agnostic}
-    Suppose $[v) = \varnothing$ and there exist $e_1$ and $e_2$ such that $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ and $[v] \in s([e_1])$, $[v] \in s([e_2])$.
-    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$
-    \begin{proof}
-        We will again show the base case when $[e_1]$ and $[e_2]$ inherit the relation atmost at 1 step.
-        That is, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.
-        We then have the following cases.
-        \begin{itemize}
-            \item $v \in s(e_1)$ and $v \in s(e_2)$.
-                  The only possible case for $e_1$ and $e_2$ is that $[e_1) = \varnothing$ and $[e_2) = \varnothing$ (because $[v) = \varnothing$).
-                  Then, 
-                  \[
-                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
-                  \], where $[v_1] \in s([e_1])$ and $[[v_1]) \not = \varnothing$.
-                  And 
-                  \[
-                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \], where $[v_2] \in s([e_2])$ and $[[v_2]) \not = \varnothing$.
-                  The cases are following.
-                  \begin{itemize}
-                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and 
-                          $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{1,E}(e'_{x})]$. 
-                          Then, $e'_{x} = e_{x}$ because the nodes in the image of $f$ should have the same predecessor.
-                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and 
-                          $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_{y})]$
-                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$ (because they are from the same connected component).
-                          \begin{itemize}
-                            \item $v_1 = f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$.
-                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}(z)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
-                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it can not be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_y)]$.
-                                  If $v'_2 \in s(e_2)$ and $v_2 \sim v'_2$ then $v'_2$ has a pre-image in $V_{Y}$ and $v_2$ has a pre-image in $V_{X}$, i.e. $v_2 = f_{V};i_{1,V}(z_2)$. 
-                                  The equation above implies that $[g_{V}(z_2)) \not = \varnothing$ which is not the case as $[v'_2) = \varnothing$.
-                            \item $v_1 = g_{V};i_{2,V}(z_1)$ and $[g_{V}(z_1)) = \varnothing$. 
-                                  Then, necessarily, $v_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = f_{V};i_{1,V}(z_1)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
-                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$, i.e.
-$v_2 = g_{V};i_{2,V}(z_2)$ and then it should be $[g_{V}(z_2)) \not = \varnothing$ which is not the case.
-                                  If $v'_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$, i.e.
-$v'_2 = g_{V};i_{2,V}(z_2)$ and then it should be $[g_{V}(z_2)) \not = \varnothing$ which is not the case.
-                          \end{itemize}
-                    \item Other cases are not possible, for example, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = [i_{1,E}(e_{x})]$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$
-                          would mean that $e_2$ has a predecessor because $v_2$ does and $v_2 \in s(e_2)$ because $v_2$ has no pre-image in $V_{Z}$.
-$[v_2]$ can not inherit the relation because we consider inheritance only upto 1 step: inheritance for $v_2$ would mean that $v$ inherits the relation in more than 1 step.
-                  \end{itemize}
-            \item $v' \in s(e_1)$ and $v \in s(e_2)$ and $v' \sim v$
-                  Suppose $[e_1) \not = \varnothing$. Then,
-                  \[
-                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))]
-                  \] and $[e_2) = \varnothing$ and
-                  \[
-                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \],
-                  where $[v_2] \in s([e_2])$ and $[[v_2]) \not = \varnothing$.
-                  \begin{itemize}
-                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{1,E}(e_x)]$.
-                          If $v_2 = f_{V};i_{1,V}(z_2)$, then $[f_{V}(z_2)) \not = \varnothing$ and hence $v'_2 \in s(e_2)$ and $v'_2 \sim v_2$.
-                          It should be the case that $v = g_{V};i_{2,V}(z)$ (if it had a pre-image in $V_{X}$ it would share the predecessor with the pre-image of $v_2$, but $[v) = \varnothing$) and $v' = f_{V};i_{1,V}(z)$, then $e_1$ has a pre-image in $E_{X}$ and $e_2$ has a pre-image in $E_{Y}$.
-                          Then $v'$ and $v_2$ both have pre-image in $V_{X}$ and because of the assumption $[f_{V}(z)) = [f_{V}(z_2))$ and $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = [i_{1,E}(e_x)] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$.
-                          The case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
-                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$ is not possible as it would mean that $v_2 \in s(e_2)$, but $[e_2) = \varnothing$.
-                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e']))$ is also not possible as it would mean that $e_2$ inherits the relation in more than 1 step.
-                \end{itemize}
-                Cases $[e_1) = \varnothing$ and $[e_2) = \varnothing$ and $[e_1) = \varnothing$, $[e_2) \not = \varnothing$ are not possible.
-            \item $v \in s(e_1)$ and $v' \in s(e_2)$ and $v' \sim v$
-                  This is symmetric to the above.
-            \item $v' \in s(e_1)$ and $v'' \in s(e_2)$ and $v' \sim v$, $v'' \sim v$. 
-                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class so it reduces to the cases above.
-        \end{itemize}
-        The proofs for the cases when $[v] \in t([e_1])$ and $[v] \in t([e_2])$ and for the combinations of $t,s$ are similar.
-    \end{proof}
-\end{lemma}
-
-\begin{remark}
-    To get the result for arbitrary number of steps we take the above lemmas as base case and procceed by induction on the number of steps.
-\end{remark}
+By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$.
 
@@
          [v') \not = \varnothing,\\
-         <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
     \end{align*}
-    In other words, there exists the node which initiates the inheritance for nodes that had no predecessor.
 \end{corollary}
@@
     \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $[v]$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
+    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
     \begin{align*}
         [v') \not = \varnothing,\\
-        <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
     \end{align*}  
-    In other words, there exists the node which initiates the inheritance for edges that had no predecessor.
 \end{corollary}
 
+\begin{corollary}
+    \label{corollary:consistency_nodes}
+    If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
+    \[
+        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \]
+\end{corollary}
+
+\begin{corollary}
+    \label{corollary:consistency_edges}
+    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
+    \[
+        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    \]
+\end{corollary}
+
 \begin{lemma}
-    $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive, assymetric and respects connectivity.
+    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive, assymetric and respects connectivity.
     \begin{proof}
@@
 \end{lemma}
-
-\begin{lemma}
-    If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exist $e_1 \in E_{X+Y}$ such that either $[v] \in s([e_1])$ or $[v] \in t([e_1])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) \not = \varnothing$ and $e_2 \in E_{X+Y}$ such that either $[v] \in s([e_2])$ or $[v] \in t([e_2])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2])) \not = \varnothing$ then
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))
-    \]
-    \begin{proof}
-        The proof is analogous to the proof of an analogous lemma for $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
-    \end{proof}
-\end{lemma}
-
-\begin{lemma}
-    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exist $v_1$ such that either $[v_1] \in s([e])$ or $[v_1] \in t([e])$ and $v_2$ such that either $[v_2] \in s([e])$ or $[v_2] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) \not = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) \not = \varnothing$ then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
-    \]
-    \begin{proof}
-        The proof is analogous to the proof of the analogous lemma for $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
-    \end{proof}
-\end{lemma}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
 
@@
         \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
-        \item $\consistency(m(v_i)) = \consistency(m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
+        \item $\consistency(m_{V};\iota_{V_{\mathcal{G}}}(v_i)) = \consistency(m_{V};\iota_{V_{\mathcal{G}}}(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
         \item The no-dangling condition holds.

### 2024-01-18 E-Hypergraphs-LICS24.tex
finish S5

@@
 \]
-is a \textit{boundary} complement if
+is a \textit{boundary complement} if
 \begin{enumerate} 
-    \item the image of $m$ is a convex down-closed e-hypergraph
-    \item $[c_1,c_2]$ is mono
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute
-    \item all $v,w$ in $\mathcal{G}$ with the pre-image in $i + j$ share the same set of parents
-    \item either for all $v,w$ in $\mathcal{G}$ with the pre-image in $i+j$, $v \consistency w$, or for all $v,w$ in $\mathcal{G}$ with the pre-image in $i+j$ $\consistency$ is undefined
+    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
+    \item $[c_1,c_2]$ is mono;
+    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
+    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
     \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
@@
 \end{enumerate}
-
-We first note that because the image of $m$ is a convex down-closed e-hypergraph, $i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono.
-Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-
-\[
-    \tikzfig{combinatorial_semantics/pushout_difference}
-\]
-and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-Similarly for $g_2$.
-$f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$ --- noting the image of $g_{ext}$ lies within $n' \setminus (i' \setminus i)$.
-Similarly for $f_2$.
+where $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$ --- noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$ --- and 
+similarly for $f_2$. The map $g_1$ is derived from $g_{int}$ as follows. 
+ %because the image of $m$ is a convex down-closed e-hypergraph, 
+Note that the internal interface $i'$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$.
+This induces an identification of $i'$ in  $n'$ which restricts to an identification of $i'\setminus i$ in $n'$.
+This gives rise to a restriction of $g_{int}$ of type $n' \setminus (i' \setminus i) \to \mathcal G$.
+Since $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface of $\mathcal G$ minus $i' \setminus i$ --- this observation induces the map $g_1:  n' \setminus (i' \setminus i) \to \mathcal{L}^\bot$ from $g_{int}$.
+A similar argument defines $g_2$.
+%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
+%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
+%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
+%
+%\[
+%    \tikzfig{combinatorial_semantics/pushout_difference}
+%\]
+%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
+%Similarly for $g_2$.
 % CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
@@
         \label{dpoi-e:assumptions}
-        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        \item $m(L)$ is a down-closed e-hypergraph;
+       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
+        %\item $m(L)$ is a down-closed e-hypergraph;
         \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
@@
 	\]
-		\begin{align*}
-		&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H} & \\ 
-		&f_2 : n' \setminus (i' \setminus i) \to \mathcal{C} \to \mathcal{H} & \\
-		\end{align*}
+	\item the map $f = [f_1,f_2]: n'' \to \mathcal H$ in the diagram above is defined as in Definition \ref{def:boundary_new} of boundary complement,  and $f_2$ as the restiction of the composite $i'' \to \mathcal R \to \mathcal H$ and,  and similarly for the map $g: k'' \to \mathcal H$.
     \end{enumerate}
@@
 % \]
-
-$[f_1,f_2]: n'' \to \mathcal{H} $
-Since both $f_1$ and $f_2$ are monos with disjoint image (?), their copairing is also a mono.  Note $f_1$ is the obvious restriction,  $f_2$ is defined as $g_2$ in boundary complement. 
-CHRIS: FIX (4) ABOVE:
-
     Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.

### 2024-01-18 E-Hypergraphs-LICS24.tex
fixes to S5

@@
 \end{enumerate}
-where $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$ --- noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$ --- and 
-similarly for $f_2$. The map $g_1$ is derived from $g_{int}$ as follows. 
+where $f_i$ and $g_i$ are defined as follows.  
  %because the image of $m$ is a convex down-closed e-hypergraph, 
-Note that the internal interface $i'$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$.
-This induces an identification of $i'$ in  $n'$ which restricts to an identification of $i'\setminus i$ in $n'$.
-This gives rise to a restriction of $g_{int}$ of type $n' \setminus (i' \setminus i) \to \mathcal G$.
-Since $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface of $\mathcal G$ minus $i' \setminus i$ --- this observation induces the map $g_1:  n' \setminus (i' \setminus i) \to \mathcal{L}^\bot$ from $g_{int}$.
-A similar argument defines $g_2$.
+Observe that the strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$.
+This induces an identification of $i' \setminus i$ in  $n'$.
+Thus,  $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$ --- noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$.
+The map $g_1$ is then derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface of $\mathcal G$ minus $i' \setminus i$.
+A similar argument defines $f_2$ and $g_2$.
 %$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a

### 2024-01-18 E-Hypergraphs-LICS24.tex
fix S5
typos

@@
 A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become vertices, connected via hyper-edges.
+Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
 Thus the expected quotient becomes simply hypergraph isomorphism.
@@
 \subsection{Soundness and Completeness}
-While the design of an appropriate combinatorial representation of semilattice enriched SMCs is delicate,  the main technical result of the paper is a proof of soundness and completeness of this representation.
+The main technical result of the paper is a proof of soundness and completeness of this representation.
 This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
@@
 
-Given a $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
+Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
 We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
@@
 % Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
+
+The appropriate notion of \textit{match} for our hierarchical setting additionally requires that its image is a \textit{down-closed} graph,  as defined below. 
+\begin{definition}[Down-closed subgraph]
+%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
+%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
+%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
+    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
+\end{definition}
 
@@
 % \end{proposition}
-
-The appropriate notion of \textit{match} for our hierarchical setting additionally requires that its image is a \textit{down-closed} graph,  as defined below. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
 

### 2024-01-18 E-Hypergraphs-LICS24.tex
proofs for duality; pattern matching references

@@
 The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called \emph{e-graphs}, in terms of Cartesian categories enriched over a semilattice. 
+We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
 We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs (\emph{e-hypergraphs}).
+Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
 They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 

### 2024-01-18 E-Hypergraphs-LICS24.tex
Soundness proof update, citation in conclusion fix

@@
 	\]
-	\item the map $f = [f_1,f_2]: n'' \to \mathcal H$ in the diagram above is defined as in Definition \ref{def:boundary_new} of boundary complement,  and $f_2$ as the restiction of the composite $i'' \to \mathcal R \to \mathcal H$ and,  and similarly for the map $g: k'' \to \mathcal H$.
+	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restiction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
     \end{enumerate}
@@
 In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of these are: support for `functorial boxes'~\cite{mellies} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{GhicaZanasiTutorial}, as sketched in Fig.
-\ref{fig:app1}~(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}~(b); and  traces~\cite{Joyal, Hasegawa},  which can be used to express feedback in categorical models of digital circuits~\cite{GhicaJung, GhicaKaye},  and, indeed cycles in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
+Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
+\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
+\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
 All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
@@
     \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V};\iota_{V_{X}} (v_i)) = \consistency(f_{V};\iota_{V_{X}}(v_j))$ and $\consistency(g_{V};\iota_{V_{Y}}(v_i)) = \consistency(g_{V};\iota_{V_{Y}}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
+    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
 \end{enumerate}    
@@
     \begin{enumerate}
-        \item $[m(v_i)) = [m(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V};\iota_{V_{\mathcal{G}}}(v_i)) = \consistency(m_{V};\iota_{V_{\mathcal{G}}}(v_j))$ for all $v_i,v_j$ that have a pre-image in $V_{\mathcal{K}}$.
+        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
+        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
         \item The no-dangling condition holds.
@@
     Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
+    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
+    \begin{itemize}
+        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$.
+        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
+    \end{itemize}
+    This means that $\mathcal{G}$ is isomorphic to the pushout constructed accroding to Theorem~\ref{th:existence_of_pushouts}.
 \end{proof}
@@
 
+\begin{figure*}[t!]
+    \centering
+    \[
+        \scalebox{0.45}{
+            \tikzfig{combinatorial_semantics/structural_rule_pushout}
+        }
+    \]
+    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
+    \label{fig:structural_rule_pushout}
+\end{figure*}
+
 \section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
@@
 \begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $f = g$ in $\textsf{PROP}^{+}(\Sigma)$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket f \rrbracket \Rrightarrow_{\mathcal{S}}^{+,*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans$.
-    $f = g$ implies that there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
-    As $\textsf{PROP}^{+}(\Sigma)$ is quotiented by SMC equations we have the following contexts $\mathcal{C}$ for any term $f$ such that $f = \mathcal{C}[l]$ for some other term $l$.
-    \[
-        \mathcal{C} := [] ~|~ c_1 + \ldots + \mathcal{C} + \ldots + c_n ~|~ c_1 ; (id \otimes \mathcal{C}); c_2
-    \]
-    we will proceed by induction on the number of equation applied to make $f$ into $g$.
-    \begin{itemize}
-        \item Base case. $f = g$ implies that for some equation $l = r$ in Figure~\ref{fig:string-equations}, $f = \mathcal{C}[l]$ and $g = \mathcal{C}[r]$.
-              We then note that every subterm of $f$ corresponds to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence so does $l$.
-              By computing pushout complement we remove the image of $\llbracket l \rrbracket$ from $\llbracket f \rrbracket$ and replace it with $\llbracket r \rrbracket$ which yields $\llbracket g \rrbracket = \llbracket \mathcal{C} \rrbracket [\llbracket r \rrbracket] = \llbracket \mathcal{C}[r] \rrbracket$ and $\llbracket f \rrbracket \Rrightarrow^{+}_{l = r} \llbracket g \rrbracket$.
-        \item Suppose $f = g$ by applying less than $n$ equations implies $\llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}} \llbracket g \rrbracket$
-        \item Let $f = g$ by applying exactly $n$ equations and then $f = g'$ by applying first $n-1$ equations and $\llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}} \llbracket g' \rrbracket$ and $g' = g$ by applying one equation.
-              By reusing the base case we have that $\llbracket g' \rrbracket \Rrightarrow^{+}_{l = r} \llbracket g \rrbracket$ and hence $\llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}} \llbracket g \rrbracket$.
-    \end{itemize}
-    The only thing that is left to show that pushout complement and pushout exist for every equation in $\mathcal{S}$.
-    Suppose, $f = g$ by applying $l = r$ which is the top left equation from~\ref{fig:string-equations}.
-    $\llbracket r \rrbracket$ is then given in Figure~\ref{fig:structural_rule_pushout}.
-    \begin{figure*}
-        \centering
-        \[
-            \scalebox{0.45}{
-                \tikzfig{combinatorial_semantics/structural_rule_pushout}
+    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textsf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
+    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
+    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \llbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
+    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
+    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
+
+    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
+    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
+    \[
+        \scalebox{0.8}{
+        \tikzfig{combinatorial_semantics/DPOI_square}
         }
     \]
-        \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-        \label{fig:structural_rule_pushout}
-    \end{figure*}
-
-    After we identify a convex down-closed image of the left-hand side of this rule within a graph $\llbracket f \rrbracket$ we remove the image by keeping only the top-level verticies.
-    Note that because the image of all $a$'s and $b$'s are top-level verticies, we are allowed to pick any monomorphism for our pushout complement.
-    And for the same reason the pushout for the right square exists.
-    Then we simply glue the right-hand side of the rule into the hole.
-    Similar argument applies to other rules as well.
+    We then further note that if $l$ is a subterm of $g$ then the nodes in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent.
+    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
+    The no-dangling condition is satisfied because we consider only mda-cospans.
+    Therefore, there exists a pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
+    Then necessarily verticies in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do in $\mathcal{G}$ so.
+    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
+    Hence $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
+    The rest follows by induction on the number of applied equations.
+    Similar argument then applies to other rules in $\mathcal{S}$ as well.
 \end{proof}
 
-\begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-    f = f_1 + \ldots + f_n
-\]
-such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-\begin{proof}
-    First, we apply the interpretation of the rules (1)-(4) from Figure~\ref{fig:string-equations} which reduces the number of edges and vertices until there is only one top-level edge which is hierarchical.
-    There will be a redex for these rules as long as there are at least two connected edges at least one of which is hierarchical.
-    The above steps therefore produce a form like \[f_1 + \ldots + f_n\] where each $f_i$ is also of the form $g_1 + \ldots + g_m$ and so on.
-    Then we apply the interpretation of rules (5)-(6) and (8) which reduces the number of hierarchical edges at each step.
-    Once again, there is a redex for such a rule as long as there are at least two hierarchical edges.
-    This will yield a form $f_1 + \ldots + f_n$ such that each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$.
-    The latter is guaranteed by rule (8).
-\end{proof}
-
 \end{document}

### 2024-01-19 E-Hypergraphs-LICS24.tex
fixes

--- 2024-01-18 E-Hypergraphs-LICS24.tex
@@
 When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is the recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$--hence the two pushouts: one for deletion and one for insertion.
+The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
 In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.

### 2024-01-19 E-Hypergraphs-LICS24.tex
Figures, padding and typos

@@
 \usepackage{subcaption}
+\usepackage{caption}
 
@@
 \]
+\captionsetup{belowskip=-5ex}
 \caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\end{figure}\label{fig:egraph-strings}
+\label{fig:egraph-strings}
+\end{figure}
 
@@
 \begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of verticies and edges.
+    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
     To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
@@
 
-We then quotient the set verticies and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
+We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
     \begin{align*}
@@
     % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that verticies inherit the relation from incident edges and vice versa.
+    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
     Clearly, all the cases above are disjoint.
@@
     Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are verticies and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
+    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
     $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
@@
 
-    and for verticies
+    and for vertices
     \begin{enumerate}
@@
     \begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of verticies and hyperedges.
+        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
         Hence, pushout complement should yield pushout complements for the corresponding sets.
@@
     \end{figure}
-    $V$s and $E$s are sets of verticies and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
+    $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
     Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
@@
     $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \llbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
+    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
     We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
@@
     \]
-    We then further note that if $l$ is a subterm of $g$ then the nodes in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent.
+    We then further note that if $l$ is a subterm of $g$ then the nodes in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
     This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
@@
     Therefore, there exists a pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily verticies in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do in $\mathcal{G}$ so.
+    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
     The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.

### 2024-01-19 E-Hypergraphs-LICS24.tex
typos

--- 2024-01-19 E-Hypergraphs-LICS24.tex
@@
 \begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of nodes, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
+A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
 The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
@@
 \begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a node $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in s(e)$.
+The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in s(e)$.
 Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in t(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some node $v \in t(e_i)$ is also in $s(e_{i+1})$.
+A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
 A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
@@
 \end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  In fact,  due to 
-\cite{bonchi_string_2022-2},  we have the following theorem which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
+Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
+\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
+\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
     $\textbf{PROP}(\Sigma) \cong \MdaCospans$
@@
 We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and nodes $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
+We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
 We require that the child relation satisfies the following conditions:
@@
 The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the appendix.
+A concrete description of the pushout of two morphisms in this category is given in the Appendix.
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
@@
 
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ nodes,  in particular with the nodes of $m$ removed from $n$ when nodes of $m$ is a sub-e-hypergraph of $n$.
+We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
 \begin{definition}[Category of e-hypergraphs with interfaces]
@@
 \[
-\trimbox{0cm 0cm 0cm 1cm}{
+\trimbox{0cm 0cm 0cm 0.75cm}{
 \adjustbox{scale=0.9,center}
@@
 string diagrams for Cartesian categories.
+An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
+Presently,  we will treat these equations via rewriting as well.
 
@@
 
-Because extended cospans have a more general notion of interface,  including \textit{internal} nodes,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
+Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
 This is the purpose of this section of the paper.
@@
 %We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that the rewriting yields a monogamous directed acyclic e-hypergraph.
-
 \begin{remark}
@@
 
-The appropriate notion of \textit{match} for our hierarchical setting additionally requires that its image is a \textit{down-closed} graph,  as defined below. 
+Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
+First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
 \begin{definition}[Down-closed subgraph]
@@
  %because the image of $m$ is a convex down-closed e-hypergraph, 
-Observe that the strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$.
-This induces an identification of $i' \setminus i$ in  $n'$.
-Thus,  $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$ --- noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$.
-The map $g_1$ is then derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface of $\mathcal G$ minus $i' \setminus i$.
-A similar argument defines $f_2$ and $g_2$.
+The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
+Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
+The maps $f_2$ and $g_2$ are defined similarly.
 %$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
@@
  is not necessarily well-typed.
-This is because its input internal interface may contain nodes that previously were a part of the output internal interface and vice versa. 
+This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
 %The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
@@
 \end{figure}
-
 Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  

### 2024-01-19 E-Hypergraphs-LICS24.tex
minor fixes to completeness proof

@@
 \]
-Observe  that therefore $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We conisder two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since their weak normal forms cannot be equal.
-Otherwise,  if $n = m$,  and $\llbracket f \rrbracket = \llbracket g \rrbracket$,  we have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
+Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
+We consider two cases: where $n=m$ and where $n\neq m$.
+In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
+Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
+We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
+Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
 
@@
 We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-Thus,  by $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
+By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
 Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.

### 2024-01-22 E-Hypergraphs-LICS24.tex
Spacing fixed, proofs aligned

@@
 
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf,9pt]{acmart}
 
@@
 In DPO rewriting,  such a rewrite rule is formalized as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
+A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
 
@@
 % CHRIS: WHY IS J DISPLAYING AND NOT R? 
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPO_square}
-    }
-\]
+% \[
+%     \scalebox{0.8}{
+%         \tikzfig{combinatorial_semantics/DPO_square}
+%     }
+% \]
 
@@
 
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI).
+To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
 In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
@@
 % \]
+\begin{figure}[t!]
+    \begin{subfigure}{0.45\linewidth}
+    \[
+            \scalebox{0.8}{
+                \tikzfig{combinatorial_semantics/DPO_square}
+            }
+    \] 
+    % \subcaption{DPO square}
+    % \label{fig:dpo}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\linewidth}
         \[
@@
 \]
+    % \subcaption{DPOI square}
+    % \label{fig:dpoi}
+    \end{subfigure}
+    \captionsetup{belowskip=-3ex}
+    \caption{DPO and DPOI squares}
+    \label{fig:dpo_dpoi}
+\end{figure}
+
 \end{definition}
@@
         \] 
-    % \item
-    % $x_1 \#_{\mathcal{F}} x_2$ iff $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$
         \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 \consistency_{\mathcal{F}} x_2$ implies $\phi(x_1) \consistency_{\mathcal{G}} \phi(x_2)$. 
-%    Or, written functionally, when $x \in E_{\mathcal{F}}$
-%    \[
-%        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
-%        \subseteq
-%        \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
-%    \]
-%    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$%, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-%    And, when $x \in V_{\mathcal{F}}$
-%    \[
-%        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
-%        \subseteq
-%        \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
-%    \]
-    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \hashtag_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \hashtag_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
 \end{enumerate}
@@
     \]
+    \captionsetup{belowskip=-4ex}
     \caption{$+$ of two morphisms in $\Ecospans$}
@@
             \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
             And, when $x \in V_{\mathcal{F}}$
@@
     \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{c,X+Y}, \consistency_{X+Y}$.
+    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
     We will omit labels as they are irrelevant to pushout construction.
@@
         % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
-        %       We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
+        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
+        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
         \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
@@
         \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[v])
+              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
+              \[ 
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
               \]
@@
     We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-    % \begin{itemize}
-        % \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
-        % \[
-        %     [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
-        % \]
-        % \[
-        %     [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
-        % \]
-        % and
-        % \[ 
-        %     [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
-        % \]
-        % by commutativity.
-        % \item The other conditions satisfy functionality by definition.
-
-    % \end{itemize}
 
@@
     \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v')))
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
     \]
@@
             \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
+              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
               \begin{align*}
                 \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-                %      \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
-                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
-                %      = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\hashtag_{Z}(i_{E_{Z}}(z)))
               \end{align*}
@@
             \begin{align*}
-             [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(&\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\subseteq
-             &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
+             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
+             &\subseteq
+             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
             \end{align*}
@@
             \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \[
-                []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \] and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
+            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
+            \begin{align*}
+                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
+                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
+            \end{align*}
+             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
             If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
@@
 
-By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$.
+So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
+The proof of the latter will follow below and the following corollaries will come in handy.
+
+By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
 
@@
     \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
+    \[
+        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
     \]
@@
     \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
+    \[
+        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
+    \]
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
     \]
@@
 \begin{lemma}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive, assymetric and respects connectivity.
+\label{lemma:child_irreflexive}
+    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
+\end{lemma}
 \begin{proof}
-        We will separately consider each property.
-        \begin{itemize}
-            \item Suppose there exist $[e]$ such that $[e] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
+    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
     Then we have two cases.
     \begin{itemize}
-                    \item Assume $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-                          This would mean that $e_1 = e$ which would mean that the relation $<_{c,X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-                    \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([v]))$.
-                          \begin{itemize}
-                            \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};i_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                                  Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [e]$.
+      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
+            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
+      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+            \begin{itemize}
+              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
+                    Then, 
+                    \[
+                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
+                    \]
                     This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
                     Contradiction.
-                            \item Case when $v = g_{V};i_{2,V}(z)$ is symmetric.
+              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
             \end{itemize}
@@
       We do not need to check this requirement for nodes because a node can not be a predecessor.
-            \item Let $e_1,e_2$ be edges. 
-                  Then, if $[e_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-                  For $[e_1] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-                  \begin{itemize}
-                    \item If $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-                          Having at the same time $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-                    \begin{itemize}
-                        \item $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
+\end{proof}
+
+\begin{lemma}
+\label{lemma:child_assymetric}
+    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is assymetric.
+\end{lemma}
+\begin{proof}
+    Let $e_1,e_2$ be edges. 
+    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
+    \begin{itemize}
+      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
+            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
+      \begin{itemize}
+          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
            This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
           \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                              Then, $[e_2] = [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                              \begin{itemize}
-                                \item If $v$ has a pre-image $f_{V};i_{1,V}(z)$ in $V_{Z}$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
+                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
+                \begin{itemize}
+                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
                         Then,
                         \[
-                                        [<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [i_{2,E}(e_y)] = [e_2]
+                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
                         \]
                         This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                                \item Case when $v$ has a pre-image $g_{V};i_{2,V}(z)$ is symmetric to the above.
+                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
                 \end{itemize}
@@
                 \[
-                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                              \]
-                              The case when $[e_1) \not = \varnothing$ are similar to the above, so we will just jump to the interesting case when 
+                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
+                \]
+                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
                 $[e_1) = \varnothing$ and
                 \[
-                                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
+                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
                 \]
@@
                 \begin{itemize}
-                                \item $v_2 = f_{V};i_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};i_{2,V}(z_2)$.
-                                Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z_2)] = [e_1]$ which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                                \begin{itemize}
-                                    \item $v_1 = g_{V};i_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};i_{1,V}(z_1)$.
-                                    Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z_1)] = [e_2]$  
-                                    However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates our assumptions.
-                                \end{itemize}
-                                \item Case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
+                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
+                  Then, 
+                  \[
+                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
+                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
+                  \begin{itemize}
+                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
+                      Then,
+                      \[
+                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
+                      \]  
+                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
+                  \end{itemize}
+                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
                 \end{itemize}
@@
     Once again, we do not need to check for nodes as a node can not be on the lefthand side of the relation.
-            \item We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-                  $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us two cases.
-                  \begin{itemize}
-                    \item If $<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
+\end{proof}
+
+\begin{lemma}
+\label{lemma:child_respect}
+    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
+\end{lemma}
+\begin{proof}
+        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
+        We have two cases.
+                  \begin{itemize}
+                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
                           $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
@@
                             \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};i_{1,V}(z)$ and $v' = g_{V};i_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))$ and 
-                                  \[
-                                    [e_2] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                  \]
+                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
+                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
+                                  \begin{align*}
+                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
+                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
+                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+                                \end{align*}
                             \end{itemize}
                     \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};i_{1,V}(z)$ and then $v'_1 = g_{V};i_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};i_{V_{Y}};<_{c,Y}^{\mu};i_{2,E}(z)] = [e_2]$ and hence $e_2$ has a pre-image in $E_{Y}$.
+                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
+                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
+                          \[
+                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
+                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
                           \begin{itemize}
                             \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
+                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
                             \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
                                   By the same argument, 
-                                  \[
-                                    <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                                  \]
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+                                  \begin{align*}
+                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
+                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
+                                                                                                          &= [e_2]
+                                   \end{align*}
+                          \end{itemize}                    
+                  \end{itemize}
+            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
             Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
@@
                 \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$, then $[e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]$.
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ which is what we want.
+                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
+                    \begin{align*}
+                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
+                    \end{align*}
+                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
+                          \begin{align*}
+                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
+                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
+                          \end{align*}
+                           which is what we want.
                     \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
                           By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$.
-                          Because $v \in s(e_1)$ there should also be a path from $e_1$ to $v'$ and $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = [e_2]$
-                    \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \[ 
-                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)]
-                    \]
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V};(z)$ and hence $e_2$ should be the predecessor of $f_{V};i_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))$ and $[e_2] = [<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e_1))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-                    \item $v$ has a pre-image $f_{V};i_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \[ 
-                        [e_2] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [i_{2,E}(e_y)]
-                    \]
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ necessarily has a pre-image in $E_{X}$.
-                    Then, by definition, $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([e_1]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [i_{2,E}(e_y)] = [e_2]$.
-                    \item Cases when $v$ has a pre-image $g_{V};i_{2,V}(z)$ are symmetric.
+                          Then, 
+                          \begin{align*}
+                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
+                             &= [e_2]
+                          \end{align*}
+                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
+                          \begin{align*}
+                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
+                             &= [e_2]
+                          \end{align*}
+                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
+                    \begin{align*} 
+                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
+                    \end{align*}
+                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
+                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
+                    \begin{align*}
+                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
+                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
+                    \end{align*}
+                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
+                    \begin{align*}
+                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
+                    \end{align*} 
+                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
+                    Then, by definition, 
+                    \begin{align*}
+                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
+                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
+                                                           &= [e_2]
+                    \end{align*}
+                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
                 \end{itemize}
                 \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
+                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
             \end{itemize}    
 \end{proof}
+
+\begin{lemma}
+\label{lemma:consistency_respect}
+$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
 \end{lemma}
-
-\begin{lemma}
-    $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$ satisfies properties~\ref{def:consistency_properties}.
-    The property (1) is satisfied by construction because we explicitly perform a closure.
-    Hence we need to check whether properties (2),(3) and (4) are satisfied.
 \begin{proof}
-        \begin{itemize}
-            \item Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
     $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
@@
       \begin{itemize}
-                        \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $i_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))$.
+          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
                   By applying $([]_{V},[]_{E})$ we get 
-                                \[
-                                i_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) \in \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}}) \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) \subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                                \]
+                  \begin{align*}
+                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
+                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
+                  \subseteq
+                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
+                  \subseteq
+                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
+                  \end{align*}
           \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                              It is then also the case that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                              Then, by definition, $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))$ and by reflexivity $i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$
+                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
+                Then, by definition, 
+                \begin{align*}
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
+                \end{align*}
+                and by reflexivity 
+                \[
+                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+                \]
       \end{itemize}
@@
           \[
-                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-                        \]
-                   \end{itemize}
-                  \item The case when $[v] \in t([e])$ is symmetric.
-                  \item If $i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-                        $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-                        By definition, either
-                        \[
-                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
-                        \]
-                        which means that either $v_1 \consistency_{X+Y}^{\mu} v_2$ or $v'_1 \consistency_{X+Y}^{\mu} v_2$ and $v'_1 \sim v_1$, or $v_1 \consistency_{X+Y}^{\mu} v_2'$ and $v_2' \sim v_2$, or $v_1' \consistency_{X+Y}^{\mu}(v_2')$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-                        Suppose, $v_1' \consistency_{X+Y}^{\mu} v_2$ and $v_1' \sim v_1$.
-                        It implies that $[v_1') = [v_2)$ and that 
-                        \[
-                            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1'))] = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2))] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+          \]
+     \end{itemize}
+The case when $[v] \in t([e])$ is symmetric.
+\end{proof}
+
+\begin{lemma}
+\label{lemma:consistency_child}
+    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for nodes and edges that share the same non-empty set of parents.
+\end{lemma}
+\begin{proof}
+  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
+        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
+        By definition, one case is when
+        \begin{align*}
+            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
+            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
+        \end{align*}
+        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
+        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
+        It implies that $[v_1') = [v_2) \not = \varnothing$.
+        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
+        \begin{align*}
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
+             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
+             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
+             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+        \end{align*}
+        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
+        Then,
+        \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
+        \]
+        and
+        \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
         \]
         The rest cases are analogous.
-                        Or there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}^{\mu}(v_3') \not = \varnothing$ and
-                        \[
-                            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
+        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
         \]
         Then the argument is the same but with using $v_3$ in place of $v_2$.
-                    \item Case for edges is analogous.
-        \end{itemize}
+  Case for edges is analogous.
 \end{proof}
-\end{lemma}
 
 \begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations satisfy all the needed properties.
+For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties.
+\end{proposition}
 \begin{proof}
-    This proposition then follows by taking two lemmas above.
+    This proposition then follows by applying the lemmas above.
+    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
 \end{proof}
-\end{proposition}
 
@@
     Then, since all the arrows are monos, the first square can be rewritten as follows.
-    \[
-      % https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3IFxcY29uZyBpICsgaiArIHoiXSxbMiwxLCJoX3tWLGV4dH0iLDJdLFsxLDAsImhfe1YsaW50fSIsMl0sWzAsMywibV97Vn0iLDJdLFs0LDMsImdfe1Z9Il0sWzIsNCwiYyIsMl1d
-    \begin{tikzcd}
+% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
+\[\begin{tikzcd}
 	{i + j + x + y} && {i+j + x} && {i+j} \\
 	\\
-	{i + j + x + y + z} &&&& {i + j + w \cong i + j + z}
+	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
 	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
@@
 	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c_{V}"', from=1-5, to=3-5]
-\end{tikzcd}
-    \]
+	\arrow["c"', from=1-5, to=3-5]
+\end{tikzcd}\]
     Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for nodes is $i + j + z$ up to isomorphism.
@@
                 \]
-                  Because $g_{V}(v) \consistency_{\mathcal{G}} g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
+                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
                   This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
@@
 \begin{figure*}[t!]
-    \centering
+    \begin{minipage}{0.1\textwidth}
+        $\;$
+    \end{minipage}
+    \hfill
+\begin{minipage}{0.7\textwidth}
     \[
@@
     \label{fig:structural_rule_pushout}
+\end{minipage}
+\hfill
+\begin{minipage}{0.1\textwidth}
+        $\;$
+\end{minipage}
 \end{figure*}

### 2024-01-23 E-Hypergraphs-LICS24.tex
Typos, alignment

@@
 The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial represetation.
+The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
 
@@
 requires the use of a categorical \textit{trace}, generating a cycle.
-In this paper, we focus on \textit{acyclic} e-graphs and their corresponding catgeories; but the above example illustrates that the extension to the cyclic case seems to be routine.
+In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
 As it would involve needless additional presentational complexity we leave it as further work.
@@
 
-\subsection{Symmetric Monoidal Theories and PROPs. }
+\subsection{Symmetric Monoidal Theories and PROPs.}
 
 A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalizations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalize the notion of a signature.
+We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
+First,  we generalise the notion of a signature.
 \begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{coarity} $m$.
+A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
 %A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
@@
 Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
+In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
 \[
@@
 \]
-is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
+is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
 In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
 Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{SLatt}$.
+We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
 
@@
 In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalize graphs by allowing edges to have multiple sources and targets. 
+Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
 When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
@@
 The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalized as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
+In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
 A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
@@
 \begin{figure}[t!]
-    \begin{subfigure}{0.45\linewidth}
+    \begin{subfigure}{0.4\linewidth}
     \[
@@
     \hfill
-    \begin{subfigure}{0.45\linewidth}
+    \begin{subfigure}{0.4\linewidth}
         \[
@@
 \end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acylcicity.
+We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
 
@@
 First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalization of the cospan construction.
+Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
 Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
@@
 Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extened cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
+Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
 
@@
 \begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with homsets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
+    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
     \[
@@
 Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is beacuse the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
+This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
 Instead,  we will treat them via rewriting,  as we do with equations of the signature.
@@
 \mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention . 
+\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
 %We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
@@
 
-Intuitively the above definition means that when the occurence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
+Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
+In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
 The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
@@
 
-More precisely, when removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
+More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
 Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
@@
 	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restiction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
+	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
     \end{enumerate}
@@
 where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed blue boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components seperated by another dashed line.
+It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
 We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
@@
 \end{proposition}    
-The proof follows the familar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
+The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
 
@@
 
+In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
+We will do the same for the corresponding pushout complement and show its uniqueness.
+These condition will play a crucial role in showing that for each equation in $\textsf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
+
 \subsection{Proof of the existence of pushouts}
 
+In constructing the pushout the functional versions of e-hypergraph relations will be useful.
 \begin{remark}
@@
 
-Recall that $<$ is essentially a transitive clousure of $<^{\mu}$ and hence we can first define 
+Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
     \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
@@
     First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider sevaral cases.
+    Then consider several cases.
     \begin{itemize}
@@
 
-    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
+    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
     \begin{enumerate}
@@
                 \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$, then
+        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
+        \]
+        then
         \[
@@
 
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \]
+    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
+    \begin{align*}
+        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
+        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
+    \end{align*}
     \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
@@
               \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
+                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
+                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
               \end{align*}
               Similarly,
-              \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \]
+              \begin{align*}
+                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
+                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
+            \end{align*}
               and we get well-definedness on the nose.
@@
 
-By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
+% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
 
@@
     \end{itemize}                  
-      We do not need to check this requirement for nodes because a node can not be a predecessor.
+      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
 \end{proof}
+
+We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
 
@@
 \label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is assymetric.
+    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
 \end{lemma}
@@
           \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exits $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
+          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
                 Then, 
@@
     \end{itemize} 
-    Once again, we do not need to check for nodes as a node can not be on the lefthand side of the relation.
+    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
 \end{proof}
@@
 \label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for nodes and edges that share the same non-empty set of parents.
+    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
 \end{lemma}
@@
 \begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties.
+For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
 \end{proposition}
@@
 
+We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
+
 \begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every node of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two nodes merged by $m$ are in the image of $f$.
+Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
+They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
 \end{definition}
@@
     Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-    % https://q.uiver.app/#q=WzAsNSxbMCwwLCJYIl0sWzIsMCwiWiJdLFsyLDIsIkMiXSxbMCwyLCJQXntcXHVyY29ybmVyfSJdLFsyLDUsIlxcYnVsbGV0Il0sWzEsMCwiaSIsMl0sWzEsMl0sWzIsM10sWzAsMywibSIsMl1d
-% \[\begin{tikzcd}
-% 	X && Z \\
-% 	\\
-% 	{P^{\urcorner}} && C \\
-% 	\\
-% 	\\
-% 	&& \bullet
-% 	\arrow["i"', from=1-3, to=1-1]
-% 	\arrow[from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% 	\arrow["m"', from=1-1, to=3-1]
-% \end{tikzcd}\]
 \[
@@
     By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exists~\cite{dpo}.
+    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
     Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
@@
     \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed accroding to Theorem~\ref{th:existence_of_pushouts}.
+    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
 \end{proof}
+
+\begin{figure*}[t!]
+    \begin{minipage}{0.1\textwidth}
+        $\;$
+    \end{minipage}
+    \hfill
+\begin{minipage}{0.7\textwidth}
+    \[
+        \scalebox{0.45}{
+            \tikzfig{combinatorial_semantics/structural_rule_pushout}
+        }
+    \]
+    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
+    \label{fig:structural_rule_pushout}
+\end{minipage}
+\hfill
+\begin{minipage}{0.1\textwidth}
+        $\;$
+\end{minipage}
+\end{figure*}
 
@@
           % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    \begin{figure}[h!]
-        \begin{subfigure}{0.45\textwidth}
-            % \[\begin{tikzcd}
-            %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-            %     \\
-            %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-            %     \arrow["{b_{V}}", from=1-5, to=1-3]
-            %     \arrow["{a_{V}}", from=1-3, to=1-1]
-            %     \arrow["{f_{V}}", from=1-1, to=3-1]
-            %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-            %     \arrow["c"', from=1-5, to=3-5]
-            % \end{tikzcd}\]
+    % \begin{figure}[h!]
+    %     \begin{subfigure}{0.45\textwidth}
+    %         % \[\begin{tikzcd}
+    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
+    %         %     \\
+    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
+    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
+    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
+    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
+    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
+    %         %     \arrow["c"', from=1-5, to=3-5]
+    %         % \end{tikzcd}\]
+    %         \[
+    %             \tikzfig{appendix/pushout_verticies}
+    %         \]
+    %     \end{subfigure}
+    %     \hfill
+    %     \begin{subfigure}{0.45\textwidth}
+    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
+    %     \[
+    %     % \begin{tikzcd}
+    %     %     {E_{L}} && 0 && 0 \\
+    %     %     \\
+    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
+    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
+    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
+    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
+    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
+    %     %     \arrow["c"', from=1-5, to=3-5]
+    %     % \end{tikzcd}
+    %     \tikzfig{appendix/pushout_edges}
+    %     \]
+    %     \end{subfigure}
+    % \end{figure}
     \[
@@
     \]
-        \end{subfigure}
-        \hfill
-        \begin{subfigure}{0.45\textwidth}
-           % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-        \[
-        % \begin{tikzcd}
-        %     {E_{L}} && 0 && 0 \\
-        %     \\
-        %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-        %     \arrow["{b_{E}}", from=1-5, to=1-3]
-        %     \arrow["{a_{E}}", from=1-3, to=1-1]
-        %     \arrow["{f_{E}}", from=1-1, to=3-1]
-        %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-        %     \arrow["c"', from=1-5, to=3-5]
-        % \end{tikzcd}
+    \[
         \tikzfig{appendix/pushout_edges}
     \]
-        \end{subfigure}
-    \end{figure}
-    $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
+    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
     Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
@@
 \end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for nodes is $i + j + z$ up to isomorphism.
+    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
     To show that pushout complement is unique we next need to show that source and target maps are unique.
@@
     \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both lefthand sides should be $\varnothing$ and therefore the relations are equal.
+        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
         \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
@@
 
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
+% \begin{figure*}[t!]
+%     \begin{minipage}{0.1\textwidth}
+%         $\;$
+%     \end{minipage}
+%     \hfill
+% \begin{minipage}{0.7\textwidth}
+%     \[
+%         \scalebox{0.45}{
+%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
+%         }
+%     \]
+%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
+%     \label{fig:structural_rule_pushout}
+% \end{minipage}
+% \hfill
+% \begin{minipage}{0.1\textwidth}
+%         $\;$
+% \end{minipage}
+% \end{figure*}
 
 \section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
+
+In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textsf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans$.
 
@@
     \]
-    We then further note that if $l$ is a subterm of $g$ then the nodes in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
+    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
     This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
     The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
+    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
     Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.

### 2024-01-25 E-Hypergraphs-LICS24.tex
Some fixes

@@
 \]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}. In particular, note that the translation of the \textit{cyclic} e-graph (e)
-requires the use of a categorical \textit{trace}, generating a cycle.
-In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
+Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
+% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
+% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
 As it would involve needless additional presentational complexity we leave it as further work.
@@
 
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(e) of Figure \ref{fig:e-graph-example}.
+We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
 These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
+Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
 The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
@@
     The boundary complement in~\ref{def:boundary_new} when exists is unique
+\end{proposition}
 \begin{proof}
@@
 \end{proof}
-    % \update{Can we check the proof? I haven't used all the conditions from monogamous-ness above.
-    % Although, if the morphism $i+j \to L$ is mono, the complement is unique for not-monogamous hypergraphs as well.}
-\end{proposition}
+
+Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
 
@@
     And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
+    \begin{figure}[h!]
     \[
@@
     \]
+    \end{figure}
     We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
@@
     The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
+    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
     The rest follows by induction on the number of applied equations.
@@
 
+% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
+%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
+%     \[
+%         f = f_1 + \ldots + f_n
+%     \]
+%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
+% \end{proposition}
+% \begin{proof}
+%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
+%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
+%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
+%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
+%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
+%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
+
+%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
+%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
+%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
+% \end{proof}
+
 \end{document}

### 2024-01-25 E-Hypergraphs-LICS24.tex
Final fixes

@@
 % E-graphs
-Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesizers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
+Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
 This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
@@
 Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $1$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
+We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
 We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
@@
 \begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in s(e)$.
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in t(e)$.
+The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
+Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
 A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
@@
 \begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
+Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
 % \[
@@
         \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 \consistency_{\mathcal{F}} x_2$ implies $\phi(x_1) \consistency_{\mathcal{G}} \phi(x_2)$. 
+        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
 \end{enumerate}
@@
 \begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent is in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
+When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
 \end{remark}
@@
     \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and that that vertices in the strictly internal interface (defined below) are not top-level.
+    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
 We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
@@
 It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-%To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
-% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
-% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
-%\[
-%    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
-%\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
-
 More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
@@
 \end{definition}
-The boundary complement conditions, in particular, prohibits finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
 \[
@@
 	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket 1 \rrbracket_{n,m} &= 1 & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
+	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
 	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
@@
 We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each equation in $\textsf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
+These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
 
@@
     Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
+\end{remark}
 \begin{definition}
@@
         \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism
+            \item $\phi$ is hypergraph homomorphism.
 
@@
             where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            And, when $x \in V_{\mathcal{F}}$
+            \item When $x \in V_{\mathcal{F}}$
             \[
@@
                 \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
+                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
             \]
@@
 \end{definition}
-\end{remark}
 
@@
     \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$.
+        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
         \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
@@
 
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textsf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans$.
+In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
 
@@
 \begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textsf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
+    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
     $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.

### 2024-03-19 Reviews-response.tex
reviews-response draft

%!LW recipe=pdflatex
\documentclass{article}

\usepackage{tikz-cd}
\usepackage{xargs}
\usepackage{amsmath}

\input{preamble}
\input{macros}
\input{sample.tikzstyles}
\input{hypergraph.tikzstyles}
\input{hypergraph.tikzdefs}

\newcommand\id{\textsf{id}}
\newcommand\sym{\textsf{sym}}

\begin{document}

\title{LICS24 Reviews Response}
\author{}

\maketitle

\section*{Review 1}
\paragraph{About theorem 6.5 (the central result of the paper: full completeness)}
\textit{I am doubtful about the statement because it seems to me that $[sym] = [id]$,
hence the functor $[-]$ is not faithful.
Here is a sketch of the argument: $[\sym]$ is a e-hypergraph with two vertices
where the inputs and the outputs are inverted. Now we embed it into a
hierarchical edge by applying the bottom left equality of Figure 3.
We get a cospan with 2 external inputs, 2 external ouputs, 2 strict internal
inputs, 2 strict internal outputs (the sets of strict internal inputs and
outputs are the same, but they are labelled differently by $f_{int}$ and $g_{int}$). We
do the same thing with $[\id]$ and notice that the two resulting cospans are
isomorphic. Thus they are equal in the category where cospans are quotiented by
isos.}

We assume that in $[\sym] = [\id]$ it was meant that $[\sym] = [\id \otimes \id]$, otherwise the equality does not typecheck.

By the definition $[\sym]$ is 
\[
    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
\]
represented as
\begin{figure}[h!]
    \[
    \scalebox{0.5}{
        \tikzfig{reviews/sym}
    }
    \]
\end{figure}

And $[\id \otimes \id]$ is \[
    n \xrightarrow{f'_{ext}} n'' \xrightarrow{f'_{int}} \mathcal{G} \xleftarrow{g'_{int}} m'' \xleftarrow{g'_{ext}} m
\]
represented as 
\begin{figure}[h!]
    \[
    \scalebox{0.5}{
        \tikzfig{reviews/id}
    }
    \]
\end{figure}

Clearly, it should be the case that $\{u_1 \mapsto u_1, \ldots u_4 \mapsto u_4\}$ and similarly for $w_i$. I think this does not follow from our definition of isomorphic cospans, so perhaps we should've required that $\alpha$ and $\gamma$ are identities (below).
\[
    \scalebox{0.75}{
        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
    }
\]
Then, the cospans above are not isomorphic. Consider the cases.
\begin{enumerate}
    \item ${v_1 \mapsto v_1}$ and ${v_2 \mapsto v_2}$ under $\beta$. Then $g_{int};\beta(w_4) \not = g'_{int}(w_4)$
    \item ${v_1 \mapsto v_2}$ and ${v_2 \mapsto v_1}$ under $\beta$. Then $f_{int};\beta(w_4) \not = f'_{int}(w_4)$
\end{enumerate}
This means that the corresponding diagram does not commute and therefore the cospans are not isomorphic.

\paragraph{In the proof}
In the first case, it is immediate that $[f] \not = [g]$, since equal terms
have weak normal forms with the same number of components.
\textit{This argument does not seem to be valid in presence of equations.
Consider an equation $a = a + b$ for example, where a and b are constants of the
signature. Then}
\[
    [a] = [a + b] = [a] + [b]
\]

We only consider equations of SMC terms, \textit{i.e.}, with no $+$ on either side of $=$.

\end{document}

### 2024-03-19 Reviews-response.tex
Added comments to the first review

@@
 \usepackage{amsmath}
+\usepackage{stmaryrd}
+\usepackage{amssymb}
 
@@
 \newcommand\sym{\textsf{sym}}
+\newcommand{\consistency}{{\smile}}
 
@@
 
+\paragraph{About Proposition A.14 (existence of the pushout complement)}
+\textit{Why doesn't the pushout complement trivially exist since you already assume
+given a pushout square?}
+
+It is indeed the case.
+The formulation should've been consider a morhisms $\mathcal{K} \to \mathcal{L} \to \mathcal{G}$.
+
+\paragraph{In the proof:}
+Then, no-dangling and no-identification
+conditions are necessary for pushout complement to exist [26].
+\textit{Are those conditions sufficient? Can you give a more precise reference? It seems
+to me that [26] (in particular Proposition 9 about the existence of pushout
+complements) focuses on the category of graphs only, so I don't see how it
+applies to the category of e-hypergraphs.}
+
+I do not know if those conditions are sufficient. 
+I am only sure the three conditions of A.14 are sufficient.
+We only proved that the pushout exists for a certain class of morphisms hence the additional conditions.
+We did not, however, prove that the pushout does not exist for other classes of morphisms.
+The no-dangling condition ensures that the sources or targets of the edge which is not a part of a match do not get deleted.
+If it was the case then there would be no way to obtain the original graph by computing a pushout as there would be not enough vertices.
+Similarly with no-identification condition, if $m$ was to identify two vertices of $\mathcal{L}$ while these were not identified by $f$ then there would be no way of getting $\mathcal{G}$ by only knowing $\mathcal{L}$, $\mathcal{L^{\bot}}$ and $\mathcal{K}$ as computing the union of $\mathcal{L}$ and $\mathcal{L^{\bot}}$ would produce more vertices than there was in $\mathcal{G}$.
+The proposition ensures that after we remove the image of $\mathcal{L}$ from $\mathcal{G}$ (excluding the image of $f;m$), all the morphisms in the pushout square are suitable for our computation of the pushout.
+There is a claim about the necessity and sufficiency of no-identification and no-dangling conditions for the pushout to exist in $\mathbf{Hyp}(\Sigma)$ in Proposition 3.17 of [3].
+
+\paragraph{About Proposition A.15 (uniqueness of the boundary complement)}
+The pushout in $\mathbf{EHyp}(\Sigma)$ should also be a pushout on the
+underlying sets of vertices and hyperedges.
+\textit{Why is that? Is it always the case for any pushout in $\mathbf{EHyp}(\Sigma)$?}
+
+This is by construction as we compute the disjoint union of vertices and edges and identify them along the corresponding morphisms.
+This construction yields the pushouts on the sets of vertices and edges.
+The construction of the corresponding consistency and child relations are orthogonal as these do not change the sets of edges and vertices.
+This is at least the case for a particular set of morphisms that we consider.
+
+\textit{Then, since all the arrows are monos}
+\textit{What if some vertex is both input and output in $\mathcal{L}$? Then $i' + j' \to L$ is not a
+monomorphism, is it?}
+
+Yes, then $i' + j' \to L$ is not a monomorphism.
+
+\paragraph{About remark 2 (restriction to signatures without any symbol of arity $0 \to 0$)}
+\textit{By incident vertices, do you mean input and output vertices? I think there is
+still a problem if you consider an hypergraph $L$ composing an edge of arity $0 \to
+1$ with an edge of arity $1 \to 0$: the boundary conditions prevent finding a match
+of $L$ in any $G$.}
+
+Yes, this is correct.
+
+\paragraph{About Def 4.1 (e-hypergraphs)}
+\textit{Why don't you require the condition (3) for t(e) as well?}
+
+That's a typo. It should've been similar to consistency relation.
+
+\textit{The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $e \consistency v$}
+\textit{What if 'e' is a top-level edge?}
+
+Should've been if $e$ is not a top-level edge.
+
+\paragraph{About def 4.2 (e-hypergraph homomorphism)}
+
+\textit{Condition (2): what if the image of 'v' or 'e' is top-level?}
+
+Then we do not require such an equality, \textit{i.e.}, it might hold or might not meaning that a top-level vertex of edge might end up either top-level or nested under a homomorphism but nested edges and vertices should preserve their predecessors.
+
+\paragraph{About def 4.5 (well-typed MDA hypergraphs)}
+\textit{Why aren't non top-level hierarchical edges required to be well-typed?}
+
+All hierarchical edges must be well-typed. The `apex' refers to the apex of a cospan, \textit{aka}, the carrier.
+
+\paragraph{About def 5.2 (extended boundary complement)}
+\textit{Isn't requiring that $i+j \to L^\bot$ is a mono too strong? For example if you
+want to account for an equality relating some term involving the identity
+morphism (for which all nodes are both inputs and outputs).}
+
+Yes, it is strong and was needed to make the proofs easier. 
+Dropping this condition is a future work.
+
+\textit{Are monomorphisms in the category of hypergraphs the morphisms that are
+injective on edges and vertices?}
+
+Yes.
+
+\paragraph{Minor remarks / typos}
+
+\textit{In the figure above, shouldn't there be $n$ (rather than just 1) pairs of black
+dots in each dashed subbox? (p2)}
+
+There is a dot for each unused input of the corresponding instance within a box.
+For example, $*$ requires $a$ and $2$ and therefore the wires from another copy of $2$ and from $1$ are terminated with dots.
+
+\textit{removing from this occurrence everything that has no pre-image in $\mathcal{K}$ (p6)}
+\textit{Isn't $C$ rather constructed by removing from $G$ everything from $L$ except for what
+comes from $K$?}
+
+Yes, this formulation is equivalent. We are removing $m(L)$ from $G$ with the exception of nodes and edges that have a pre-image in $K$, \textit{i.e.}, those that come from $K$.
+\textit{What is a monogamous cospan? p(7)}
+
+Apparently the definition of a monogamous cospan was overlooked.
+
+\textit{This is confusing: you are not defining any equivalence relation here. Rather, 
+this family of equivalence relations is provided by each hypergraph (if I
+understand correctly). Def 4.1 (p7)}
+
+Yes, it is indeed a family of relations.
+
+\textit{Shouldn't a homomorphism preserves labelling?}
+
+It preserves labelling as per the condition (1) on such a morphism: it should be a hypergraph-homomorphism, which, by definition, preserves labelling.
+
+\textit{Unfolding definitions, it follows that}
+$f = g \implies \llbracket f \rrbracket \Rrightarrow^{*} \llbracket g \rrbracket$.
+
+\textit{The literal meaning makes this statement trivial: if $f = g$, then $[f]$ rewrites to
+$[g]$ in 0 steps.
+If we replace $f = g$ with $p_i(f) = p_i(g)$, where $f$ and $g$ are unquotiented terms and
+$p_i$ is the canonical projection, this is wrong, because it would entail that the
+rewriting relation is symmetric, which is not the case, unless we consider} $\Rrightarrow^{*}$
+\textit{that each equation generates two rewrite rules, one for each direction of the
+equality (if is the case, this should be said explicitly). Corollary 6.6 suffers
+from the same defect, and the proof of Theorem 6.5 as well.}
+
+$\llbracket \mathcal{E} \rrbracket$ consists of $\langle l, r \rangle$ and $\langle l, r \rangle$ for each $l = r$ in $\mathcal{E}$.
+
+\textit{Aren't they partial functions, defined on any non-top level elements only?}
+Yes
+
+\textit{Condition (5) implicitly assumes that f(v) is not top-level, because the
+consistency relation is only defined for such elements.}
+
+It should be read as either all $f(v_i)$ are pair-wise consistent, or consistency for them is undefined, \textit{i.e.}, all of them are top-level.
+
+\textit{In which category is this pushout proven to exist? Clearly not in the category
+$\mathbf{EHyp}(\Sigma)$ of hypergraphs, since just after the proof you write that it was not
+proven to be a hypergraph.}
+
+It becomes a pushout in $\mathbf{EHyp}(\Sigma)$ right after it is proved that it is a valid e-hypergraph (Proposition A.12).
+Probably the proof should've been ordered the other way around.
+
+\textit{In the proof, you consider the identity relation on edges, then its
+reflexive-transitive-symmetric closure. Doesn't it coincide with the identity
+relation? Furthermore, doesn't quotienting a set X by the identity relation
+yields the same set X?
+Do you really need to take the transitive closure of the relation on vertices,
+since f and g are monomorphisms?}
+
+This was just to make the presentation uniform.
 \end{document}

### 2024-03-20 Reviews-response.tex
Another comments

@@
 This was just to make the presentation uniform.
+
+\section*{Review 2}
+\textit{The discussion about Figure 1 in Section (1.1) does not provide enough details to understand what the graph describes.
+In particular, in the e-hypergraph formalism, I am not sure I understand what the black nodes inside the equivalence classes mean.
+For instance, why are these nodes only connected to the bottom of the boxes and never to the top?}
+
+Black dots correspond to delete morphisms of the Cartesian category, \textit{i.e.}, they discard the unused input and hence are only conected to the bottom as inputs flow from bottom to top.
+
+\textit{Is the free monoid monad the list monad?}
+
+Yes
+
+\textit{Without further assumptions (e.g., some form of adhesivity), pushout complements may not be unique, assuming it exists.
+Is there any guarantee in the DPOI framework for hypergraphs with interfaces?
+How does the construction of boundary complement (Definition 3.8) relate to initial pushouts (see, for instance, Chapter 6 of Fundamentals of Algebraic Graph Transformation Ehrig et a.
+2006, page 125)?}
+
+The category of plain hypergraphs ($\mathbf{Hyp}$), as used in [3], is adhesive.
+We do not know if the category of e-hypergraphs ($\mathbf{EHyp}$) has any adhesive properties, however, in this work we are only interested in pushouts along monos and monomorphic matches and for such morphisms the pushout complement is unique as shown in A.15.
+The boundary complement definition is a generalisation of a similar concept from [4] which is needed to make sure we remain within needed class of morphisms when computing the complement and the pushout.
+We can not immediately relate initial pushouts to the boundary complement.
+
+\textit{One could argue that since the consistency relation is built from the child relation, it should not be part of the tuple defining the e-hypergraph.}
+
+?
+
+\textit{In Figure 5, why do $a_1$, ... $a_n$, and $a'_1$, ..., $a'_k$ appear as internal interfaces on the rightmost hypergraph?}
+
+Internal interface is external interface plus strict internal interface. Since both $a_i$'s and $a'_j$'s are external interfaces they should be present in internal interface as well.
+
+\section*{Review 3}
+\textit{However, I think the authors could improve the discussion of the 0-ary case and be more convincing about the fact that this could be handled without much difficulty.
+Otherwise, this would dramatically restrict the domain of application of those results.}
+
+One workaround is to add more context to the left-hand side of a rewrite rule. 
+Assume we have a rewrite rule $f \Rrightarrow g \otimes a$ where $a$ is a scalar.
+If the occurrence of $f$ within $\mathcal{G}$ is top-level, then we can proceed as usual.
+If, on the other hand, the occurrence is nested, then there is no way to specify that $a$ should have the same parent as $f$ had.
+The workaround can be seen in the following example.
+
+\[
+    \scalebox{0.5}{
+        \tikzfig{reviews/handling_scalars}
+    }
+\]
+
+$\mathcal{F}, \mathcal{G}_i$ are variable-labelled e-hypergraphs that can be unified with concrete e-hypergraphs.
+For example, $\mathcal{F}$ corresponds to $h$ and $\mathcal{G}_1$ corresponds to $k$.
+
+\section*{Review 4}
+
+Review 4 had no questions.
+
 \end{document}

### 2024-03-21 Reviews-response.tex
A little fix.
small update about iso

@@
 I think this does not follow from our definition of isomorphic cospans, so perhaps we should've required that $\alpha$ and $\gamma$ are identities (below).
+However, the interfaces are ordered according to a remark on page 21 in [3], so the only possible isomorphism between $n'$ and $n''$ (respectively, $m'$ and $m''$) is identity.
 \[
@@
 Internal interface is external interface plus strict internal interface. Since both $a_i$'s and $a'_j$'s are external interfaces they should be present in internal interface as well.
+The figure is essentially saying that we sum both internal interfaces and add extra copy of external interface (to account for the inputs and outputs of this new hierarchical edge).
 

### 2024-04-04 EGraphsExtendedAbstract.tex
Extended abstract

\documentclass[sigconf, 9pt, nonacm]{acmart}

\settopmatter{printacmref=false}

\usepackage{amsmath}
\usepackage{stmaryrd}
% \usepackage{amssymb}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage[
  colorinlistoftodos,
  prependcaption,
  % disable,
]{todonotes}
\usepackage{fontenc}

%\usepackage{minted}

\usepackage{subcaption}
\usepackage{caption}

\usepackage{hyperref}

\usepackage{docmute}
\usepackage{hyperref}
\usepackage{proof}

\ifthenelse{\isundefined{\ismain}}{
    \newcommand{\ismain}{0}
    \usepackage{xifthen}
}{}

\usepackage{enumitem}
\usepackage{amsthm}

\usepackage{aliascnt}
% \newaliascnt{definition}{thm}
% \newaliascnt{remark}{thm}
% \newaliascnt{example}{thm}
\newtheorem{thm}{Theorem}
\newtheorem{remark}[thm]{Remark}

\theoremstyle{definition}
\newtheorem{definition}[thm]{Definition}

\newtheorem{example}[thm]{Example}
\newtheorem{proposition}[thm]{Proposition}

\input{macros}
\input{sample.tikzstyles}
\input{hypergraph.tikzstyles}
\input{hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

% \usepackage{mathtools,mathrsfs}
\usepackage{tikz-cd}
\usepackage{tikzit}
% to reduce tikz picture white space
\usepackage{adjustbox}

% for quotienting
% \usepackage{faktor}
\usepackage{xfrac}

\usepackage{import}

% category names
%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}

\newcommand\HypI[1]{\textbf{HypI}(#1)}
\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
% conflict
\newcommand{\hashtag}{{\#}}
\newcommand{\consistency}{{\smile}}
% There is already a remark in the preamble, but it does
% not work for some reason

\newcommand\id{\textsf{id}}
\newcommand\sym{\textsf{sym}}

\usepackage{cleveref} % after hyperref!

\begin{document}

\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}

\author{Anonymous}

\maketitle

\section{Introduction}

E-graphs have been around for a while since being introduced in the 80s~\cite{nelson1980techniques} and yet there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
However, there is a fundamental limitation to the area of equational theories that e-graphs (as presented, \textit{e.g.}, in~\cite{EggPaper}) can be applied to.
Specifically, e-graphs support equality saturation for so-called \textit{algebraic} theories --- theories that consist of operators (or generators) that can have an arbitrary number of inputs but only a single output, and equations between them.
There are certain applications~\cite{zx, ghica_compositional_2023, probabilistic} where the operators in the corresponding theory can have multiple inputs and multiple outputs. 
Such theories are called \textit{symmetric monoidal theories} (SMTs), they consist of terms which are built from the operators by combining them with parallel and sequential composition, and they would certainly benefit from the equality saturation machinery offered by the e-graphs.
Even if we allowed terms in an e-graph to have multiple outputs, the rewriting for such terms using the current data structure that e-graphs are based on would be infeasible because rewriting in an SMT is performed modulo SMT axioms --- including such axioms into the set of rewrite rules would quickly explode the size of the hypothetical e-graph.
This motivates the need for generalised e-graphs and suggests that such generalisation should be developed with SMT axioms in mind.

This work in progress aims to propose a generalisation of e-graphs for arbitrary SMTs which is build by enriching the underlying categorical structure of a given SMT.
Then, it can be seen that original e-graphs is the case of our construction applied to Cartesian SMTs.
Apart from building the foundation for the application of e-graphs to a richer class of theories, this work also provides a new perspective on solving the notorious issues of alpha-equivalence and binding that arise when, for example, one wants to build e-graphs for functional languages~\cite{koehler2022sketchguided}.

\section{Categorical semantics of e-graphs}

In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.

Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.  We write the identity morphism on $A$ as $\id_A$,  or simply $A$.  We commonly write $f;g$ for composition in diagram order.  Composition in the usual order is written $g \circ f$.  We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.  We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.  We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.  We denote by $\textsf{SLatt}$ the category of semilattices.
We will also use string diagrams as a well-established two-dimensional syntax for the (equivalence classes of) terms in an SMC which shall be read from bottom-to-top.

\subsection{Symmetric Monoidal Theories and PROPs.}

A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.  We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.  First,  we generalise the notion of a signature. 
\begin{definition}[Monoidal signature]
A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.  %A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements. 
\end{definition}

A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
\begin{definition}[Products and permutations category]
A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
% A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects. 
We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
\end{definition}
The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.

An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
\begin{definition}[Symmetric Monoidal Theory]
A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
\end{definition}

It will often be convenient to give presentations of SMTs in terms of string diagrams.
\begin{example}
\label{example:csmt}
The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
\[
	\scalebox{0.8}{
  	 \tikzfig{categorical-semantics/Cartesian-equipment}
	}
\]
together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-theory}	
	}
\]
The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
\[
	\scalebox{0.6}{
	\tikzfig{categorical-semantics/Cartesian-naturality}
	}
\]
By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
\end{example}

\subsection{Free Semilattice Enrichment: Formal Joins}

To express the notion of e-classes of SMT terms we consider an SMT enriched over a category of semilattices as defined below.

\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms. 
\begin{align*}
f ; (g+h) &= f;g + f;h &
(f+g) ; h &= f;h + g;h \\
f \otimes (g+h) &= f \otimes g + f \otimes h & 
(f+g) \otimes h &= f \otimes h + g \otimes h,
%\item A set $ob(\mathbb{C})$ of objects;
%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
\end{align*}
Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
% A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.  
We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
\end{definition}

The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
Note that when we consider enriched SMTs ($\textbf{SMT}^{+}(\Sigma,\mathcal{E})$), $\mathcal{E}$ consists exclusively of $\Sigma$-terms, not $\Sigma^{+}$ terms.
Intuitively, the $+$ generator provides us with an opportunity to form equivalent classes of terms and the axioms above state the congruence over $+$.
For example, the first equation can be read as: if $g$ is equivalent to $h$ then $f;g$ is equivalent to $f;h$ for any $f$.

\begin{figure}
    \[
        \scalebox{0.65}{
        \tikzfig{categorical-semantics/egraph-strings}
        }
    \]
    \captionsetup{belowskip=-5ex}
    \caption{String diagrams for semilattice enriched symmetric monoidal categories.}
    \label{fig:egraph-strings}
    \end{figure}

To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms (and to denote the e-classes).
We denote this particular syntax as \textit{e-string diagrams}.
Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.  We overload the binary notation $+$ for our $n$-ary notation.  
We use string diagrams informally in this work, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.

\begin{figure*}
\[  
    \scalebox{0.75}{
	\tikzfig{categorical-semantics/egraph-strings-equations}
    }
\]
\caption{Equations for a  semilattice enriched symmetric monoidal category}
\label{fig:string-equations}
\end{figure*}

\section{E-graphs as terms of a Cartesian SMT}
We are now ready to present how (asyclic) e-graphs can be considered as terms of a Cartesian SMT.
Let $\Sigma$ be a monoidal signature in which every generator $c$ has a single output and $\mathcal{E}$ be the set of equations on $\Sigma$-terms. 
Then let $\textbf{CSMT}(\Sigma_{C},\mathcal{E}_{C})$ be the Cartesian SMT as introduced in Example~\ref{example:csmt}.
\begin{proposition}
    Acyclic e-graphs over a set of generators $\Sigma$ and equivalence classes of terms on $\textbf{CSMT}^{+}(\Sigma_{C},\mathcal{E}_{C})$ are in one-to-one correspondence.
\end{proposition}

\begin{figure}
    \[
        \scalebox{0.5}{
        \tikzfig{categorical-semantics/egraph-translation-2}
        }
    \]
    \caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
    \label{fig:e-graph-example}
\end{figure}

We will illustrate this proposition by an example.
\begin{example}
    Recall an excerpt of a simple e-graph example from~\cite{EggPaper} that we depict in Figure~\ref{fig:e-graph-example}.
    Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation. 
The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing ($\Delta$-generator) node.

Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
Initially each equivalence class has a single node. 
The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$. 
This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-string diagrams, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges (wires) connect to the equivalence class itself. 
Explicit discarding ($!$-generator) nodes indicate which of the class-level edges are connected to which nodes inside the class. 
Generally, the application of rewrite rules to e-graphs correspond to a sequence of equations applied to e-string diagrams.
A step-by-step application of such equations for the rewrite rule from $(a)$ to $(b)$ and $(b)$ to $(c)$ can be seen in Figure~\ref{fig:e-graph-example-a-b} and Figure~\ref{fig:e-graph-example-a-b} respectively.
The equations applied are the ones from Figure~\ref{fig:string-equations} and naturality of delete and copy.
One can note that the size of the eventual e-string diagram is comparable to the size of the corresponding e-graph.
\end{example}

We can also translate an arbitrary e-graph into an e-string diagram as shown belownoting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
\[
    \tikzfig{categorical-semantics/egraph-translation-1}
\]
Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.  This means the (informal) translation given above is fully general for acyclic e-graphs.
We believe that e-graphs with cycles can be represented as terms in Traced Cartesian SMTs and accounting for a trace above seems to be routine.

While we did not explicitly define how exactly equations are applied to e-string diagrams resorting rather to intuition, e-string diagrams can be formalised as concrete combinatorial structures by following an established approach of converting string diagrams for non-enriched SMCs to hypergraphs~\cite{bonchi_string_2022,bonchi_string_2022-1,bonchi_string_2022-2}.
We denote the corresponding combinatorial structure an e-hypergraph and claim that applying equations to e-string diagrams can be mimicked by performing a \textit{double pushout} rewriting on e-hypergraphs.
This allows us to express a complex e-graph rewriting, that includes creation of e-classes, merging, and computing congruence closure as a sequence of e-hypergraph rewriting steps which as we believe could facilitate reasoning about e-graphs.
The difference between e-string diagrams and e-hypergraphs is that the former are implicitly quotiented by the equations in Figure~\ref{fig:string-equations} while for the latter these equations become structural rewrite rules.
One can notice that we unfolded e-string diagrams in Figure~\ref{fig:e-graph-example-b-c} to identify a redex and then to factor out the nodes to achieve a better sharing.
This would also be the case when applying rewrite rule for e-hypergraph representation in a na\"ive way.
The work towards finding redexes and achieving maximal sharing without excessive unfolding is in progress.

\begin{figure*}
    \[
        % \hspace{1.3cm}
        \scalebox{0.5}{
        \tikzfig{categorical-semantics/egraph-translation-step-by-step-a-b}
        }
    \]
    \caption{Example translation from $(a)$ to $(b)$.}
    \label{fig:e-graph-example-a-b}
\end{figure*}

\begin{figure*}
    \[
        % \hspace{1.3cm}
        \scalebox{0.55}{
        \tikzfig{categorical-semantics/e-graph-substitution}
        }
    \]
    \caption{E-graph explicit substitution example.}
    \label{fig:e-graph-substitution}
\end{figure*}

\begin{figure*}
    \[
        % \hspace{1.3cm}
        \scalebox{0.5}{
        \tikzfig{categorical-semantics/e-string-substitution}
        }
    \]
    \caption{E-string diagrammatic substitution example.}
    \label{fig:e-string-substitution}
\end{figure*}

\section{E-string diagrams for lambda calculus}
In this section we will show how a particular kind of SMTs can provide a new perspective on solving the long-standing issues of $\alpha$-equivalence and binding when using e-graphs for languages like lambda calculus.
The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{De Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a subsittution of eqivalence classes that represent a lot of terms at once.
An example of a substitution can be seen in Figure~\ref{fig:e-graph-substitution}.
Notably, the terms that correspond to explicit substitution can be discarded at the end.

By recalling the correspondence between Cartesian Closed SMCs and lambda-calculus we can consider the corresponding semilattice enriched Cartesian Closed SMTs.
Apart from the generators in $\Sigma_{C}$ such theories also contain \textit{application} and $\textit{abstraction}$ maps that correspond to $app$ and $\lambda$ respectively.
A complete description of these maps and the corresponding equations that they obey can be found in~\cite{ghica-zanassi2023string}.
The corresponding e-string diagram can be seen in Figure~\ref{fig:e-string-substitution}.
The half-circle nodes correspond to application maps and the rounded rectangle with a wire attached to its left side is the abstraction map where the attached wire corresponds to the argument of the abstraction.
A distinguishable feature of these string diagrams is that they are automatically quotiented by $\alpha$-equivalence and that a substitution is merely a rewrite step that glues an argument of the application to the attached wire of the abstraction.

\begin{figure*}
    \[
        % \hspace{1.3cm}
        \scalebox{0.5}{
        \tikzfig{categorical-semantics/egraph-translation-step-by-step-b-c}
        }
    \]
    \caption{Example translation from $(b)$ to $(c)$.}
    \label{fig:e-graph-example-b-c}
\end{figure*}

\section{Conclusion}

In this work, we proposed a generalisation of e-graphs for SMTs which as a bonus provides a framework for reasoning about acyclic e-graphs.
One promising application of such a generalisation is performing equality saturation for lambda calculus as e-string diagrams ease the burden of dealing with $\alpha$-equivalent terms and binding.
Some other potential applications also include e-graphs for digital circuits~\cite{ghica_compositional_2023} and quantum computing~\cite{coecke_interacting_2011} where each domain arises from a certain kind of SMT.
We hope this work will foster discussion within e-graphs community, especially in designing a suitable data structure that would bring this theoretical generalisation into practice.

\bibliographystyle{acm}
\bibliography{bibliography}

\end{document}

### 2024-04-17 EGraphsExtendedAbstract.tex
the submitted version of extended abstract

@@
 
-\author{Anonymous}
+\author{Chris Barrett}
+\affiliation{
+  \institution{University of Oxford}
+  \country{United Kingdom}
+}
+
+\author{Aleksei Tiurin}
+\affiliation{
+  \institution{University of Birmingham}
+  \country{United Kingdom}
+}
+
+\author{Dan Ghica}
+\affiliation{
+  \institution{University of Birmingham}
+  \country{United Kingdom}
+}
 
@@
 
-This work in progress aims to propose a generalisation of e-graphs for arbitrary SMTs which is build by enriching the underlying categorical structure of a given SMT.
+Our work in progress aims to propose a generalisation of e-graphs for arbitrary SMTs which is build by enriching the underlying categorical structure of a given SMT.
 Then, it can be seen that original e-graphs is the case of our construction applied to Cartesian SMTs.
@@
 
-To express the notion of e-classes of SMT terms we consider an SMT enriched over a category of semilattices as defined below.
+To express the notion of e-classes of SMT terms we consider an SMT which is represented by an SMC enriched over a category of semilattices as defined below.
 
@@
 \section{E-graphs as terms of a Cartesian SMT}
-We are now ready to present how (asyclic) e-graphs can be considered as terms of a Cartesian SMT.
+We are now ready to present how (acyclic) e-graphs can be considered as terms of a Cartesian SMT.
 Let $\Sigma$ be a monoidal signature in which every generator $c$ has a single output and $\mathcal{E}$ be the set of equations on $\Sigma$-terms. 
@@
 \begin{proposition}
-    Acyclic e-graphs over a set of generators $\Sigma$ and equivalence classes of terms on $\textbf{CSMT}^{+}(\Sigma_{C},\mathcal{E}_{C})$ are in one-to-one correspondence.
+    For each acyclic e-graph over a set of generators $\Sigma$ and a set of equations $\mathcal{E}$ that have at most one root e-class, there is a corresponding equivalence class of terms in $\textbf{CSMT}^{+}(\Sigma_{C},\mathcal{E}_{C})$.
+    Moreover, if $e_1$ and $e_2$ are e-graphs such that $e_2$ is obtained from $e_1$ after applying some rewrite rule (\textit{i.e.}, $e_2$ is obtained after adding new terms into $e_1$ and performing the necessary merges and repairs), then $s_2$ can be obtained from $s_1$ by applying the equations from Figure~\ref{fig:string-equations} and $\mathcal{E}_{C}$ where $s_1$ and $s_2$ are the corresponding e-string diagrams for $e_1$ and $e_2$ respectively.
 \end{proposition}
@@
 A step-by-step application of such equations for the rewrite rule from $(a)$ to $(b)$ and $(b)$ to $(c)$ can be seen in Figure~\ref{fig:e-graph-example-a-b} and Figure~\ref{fig:e-graph-example-a-b} respectively.
-The equations applied are the ones from Figure~\ref{fig:string-equations} and naturality of delete and copy.
+The equations applied are the ones from Figure~\ref{fig:string-equations} and naturality of delete ($!$) and copy ($\Delta$).
 One can note that the size of the eventual e-string diagram is comparable to the size of the corresponding e-graph.
@@
 
-We can also translate an arbitrary e-graph into an e-string diagram as shown belownoting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
+We can then translate an arbitrary e-graph into an e-string diagram as shown below, noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
 \[
@@
 While we did not explicitly define how exactly equations are applied to e-string diagrams resorting rather to intuition, e-string diagrams can be formalised as concrete combinatorial structures by following an established approach of converting string diagrams for non-enriched SMCs to hypergraphs~\cite{bonchi_string_2022,bonchi_string_2022-1,bonchi_string_2022-2}.
-We denote the corresponding combinatorial structure an e-hypergraph and claim that applying equations to e-string diagrams can be mimicked by performing a \textit{double pushout} rewriting on e-hypergraphs.
+We denote the corresponding combinatorial structure an e-hypergraph and claim that applying equations to e-string diagrams can be mimicked by performing a \textit{double pushout} rewriting of e-hypergraphs.
 This allows us to express a complex e-graph rewriting, that includes creation of e-classes, merging, and computing congruence closure as a sequence of e-hypergraph rewriting steps which as we believe could facilitate reasoning about e-graphs.
@@
 One can notice that we unfolded e-string diagrams in Figure~\ref{fig:e-graph-example-b-c} to identify a redex and then to factor out the nodes to achieve a better sharing.
-This would also be the case when applying rewrite rule for e-hypergraph representation in a na\"ive way.
+This would also be the case when applying rewrite rules for e-hypergraph representation in a na\"ive way.
 The work towards finding redexes and achieving maximal sharing without excessive unfolding is in progress.
@@
         \scalebox{0.5}{
-        \tikzfig{categorical-semantics/e-string-substitution}
+        \tikzfig{categorical-semantics/e-string-substitution-2}
         }
@@
 
-\section{E-string diagrams for lambda calculus}
-In this section we will show how a particular kind of SMTs can provide a new perspective on solving the long-standing issues of $\alpha$-equivalence and binding when using e-graphs for languages like lambda calculus.
-The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
-If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{De Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
-The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a subsittution of eqivalence classes that represent a lot of terms at once.
-An example of a substitution can be seen in Figure~\ref{fig:e-graph-substitution}.
-Notably, the terms that correspond to explicit substitution can be discarded at the end.
-
-By recalling the correspondence between Cartesian Closed SMCs and lambda-calculus we can consider the corresponding semilattice enriched Cartesian Closed SMTs.
-Apart from the generators in $\Sigma_{C}$ such theories also contain \textit{application} and $\textit{abstraction}$ maps that correspond to $app$ and $\lambda$ respectively.
-A complete description of these maps and the corresponding equations that they obey can be found in~\cite{ghica-zanassi2023string}.
-The corresponding e-string diagram can be seen in Figure~\ref{fig:e-string-substitution}.
-The half-circle nodes correspond to application maps and the rounded rectangle with a wire attached to its left side is the abstraction map where the attached wire corresponds to the argument of the abstraction.
-A distinguishable feature of these string diagrams is that they are automatically quotiented by $\alpha$-equivalence and that a substitution is merely a rewrite step that glues an argument of the application to the attached wire of the abstraction.
-
-\begin{figure*}
+\begin{figure*}[t!]
+    \vspace{-3cm}
+    \centering
     \[
-        % \hspace{1.3cm}
-        \scalebox{0.5}{
+        \hspace{1.3cm}
+        \resizebox{0.8\textwidth}{!}{
         \tikzfig{categorical-semantics/egraph-translation-step-by-step-b-c}
@@
 \end{figure*}
+
+\section{E-string diagrams for lambda calculus}
+In this section we will show how a particular kind of SMTs can provide a new perspective on solving the long-standing issues of $\alpha$-equivalence and binding when using e-graphs for languages like lambda calculus.
+The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
+If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
+The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
+An example of a substitution can be seen in Figure~\ref{fig:e-graph-substitution}.
+An explicit substitution operator is the part of the syntax and the substitution of e-classes is defined inductively on the structure of all terms that these e-classes represent, \textit{i.e.}, computing a substitution of e-classes is achieved by computing a pair-wise substitution for all the instances of these e-classes.
+Notably, the terms that correspond to explicit substitution can be discarded at the end.
+
+By recalling the correspondence between Cartesian Closed SMCs and lambda-calculus we can consider the corresponding semilattice enriched Cartesian Closed SMTs.
+Apart from the generators in $\Sigma_{C}$ such theories also contain \textit{application} and $\textit{abstraction}$ maps that correspond to $app$ and $\lambda$ respectively.
+A complete description of these maps and the equations that they obey can be found in~\cite{ghica-zanassi2023string}.
+The e-string diagram version of the previous example can be seen in Figure~\ref{fig:e-string-substitution}.
+The half-circle nodes correspond to application maps and the rounded rectangles with a wire attached to their right sides are the abstraction maps where the attached wire corresponds to the argument of the abstraction.
+A distinguishable feature of these string diagrams is that they are automatically quotiented by $\alpha$-equivalence and that a $\beta$-reduction is merely a rewrite step that glues an argument of the application to the attached wire of the abstraction and the substitution is just string diagram composition.
+Another advantage of string diagrams is that every occurrence of the bound variable is essentially the same wire as opposed to de Brujin indices, so there is no need to modify any indices, the whole substitution is just a matter of reattaching the wires non-destructively.
+In particular, note how the substitution within $y + y$ was completely avoided.
+% The latter follows from the definition of the respective adjunction that defines these maps.
 

### 2024-05-01 E-Hypergraphs-LICS24.tex
halfway through amended definitions of \consistency and <

@@
     \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
+    \item \label{assumption:equal_predecessors} $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
+    \item \label{assumption:non_ambiguous_predecessors} If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
     \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
@@
     S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
+          &\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
+          &\}\\
+          &\cup\\
+          &\{\\
+          &\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
+          &\}\\
+          &\cup\\
+          &\{\\
+          &\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\\
+          &\}\\
         \\
@@
     % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
+and let relations $R_{V},R_{E}$ be their transitive closures respectively.
 
@@
     \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
+We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
+Where necessary we will refer to $S_{V}$ and $S_{R}$ as $\sim_{S}$.
 We have 
@@
 We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
+Similarly, we will write $v_{x}$ to refer to a vertex with a pre-image in $V_{X}$.
 Likewise, 
@@
 
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
+These definitions make $([]_{V},[]_{E})$ automatically a homomorphism with respect to source and target maps.
+These maps are also automatically well-defined functions since $[e_1] = [e_2]$ implies $e_1 = e_2$ since the mappings for edges are mono.
+
+% So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
 
 Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
+    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
     \] and 
     \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
+        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u])
     \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
@@
     \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
+        \item If there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ is defined and $u \sim v$
+              we let
+              \[
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
               \]
         \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
+              If there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
               \[ 
@@
     \end{enumerate}
+    There is no case when $u$ has a pre-image in $V_{Z}$ and yet $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e$ and there is a path from $[u]$ to $[v]$.
+    The existence of a path would mean that there is $u' \sim u$ and $v' \sim v$ such that there is a path from $u'$ to $v'$ and since $u$ has a pre-image in $V_{Z}$, $u = f_{V};\iota_{1,V}(z_1)$.
+    $v' \sim v$ means that both $v'$ and $v$ have a pre-image in $V_{Z}$ (unless $v' = v$ in which case $[u') \not = \varnothing$ as there is a path from $u'$ to $v'$) and since $u = f_{V};\iota_{1,V}(z_1)$ and $[u) = \varnothing$ $f_{V};\iota_{1,V}(z) = \varnothing$ for all $z$.
+    Since $[v) \not = \varnothing$ and $v' \sim v$, $v$ is necessarily in the image of $g_{V};\iota_{2,V}$ and for all $z$ $[g_{V};\iota_{2,V}(z)) \not = \varnothing$ as well as for some $z'$ such that $u'' = g_{V};\iota_{2,V}(z')$ and $u'' \sim_{S} u$ which would mean that the case (1) is applicable to $u$.
+    Similarly, when $u = g_{V};\iota_{2,V}(z)$.
+
+    Otherwise, we leave $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}{\sim}([u]))$ undefined.
     % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
+    Now, assume that $e$ is an element of $E_{X+Y}$.
     Consider two cases.
     \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
+    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e'$.
             Then we define
         \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]
+        \]
+    \item $[e) = \varnothing$ and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an undirected path from $[e]$ to $[v]$, we define 
+        \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
         \] 
     \end{enumerate}
+    Otherwise we leave $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{E_{X} + E_{Y}}{\sim}([e]))$ undefined.
     % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
@@
 
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
+    Let's show that the definition does not depend on a particular choice of $v$ in (1).
+    Suppose there exist $v_1$ and $v_2$ such that $v_1 \not = v_2$ and $[v_1) \not = \varnothing$ and $[v_2) \not = \varnothing$ and such that $u \sim v_1$ and $u \sim v_2$.
+    Then, it must be the case that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1)) = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))$.
+    Consider cases.
     \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
+        \item $u = v_1$ and $u \sim_{S} v_2$. Then there exists $z \in V_{Z}$ such that $u = f_{V};\iota_{1,V}(z) = v_1$ and $v_2 = g_{V};\iota_{2,V}(z)$.
+              Since $[v_1) \not = \varnothing$, it must be the case that $[v_2) = \varnothing$ according to assumption \ref{assumption:non_ambiguous_predecessors} and we get a contraction to $[v_2) \not = \varnothing$.
+        \item $u = v_1$ and $u \sim v_2$. The latter implies that $u$ has a pre-image in $V_{Z}$ and so does $v_2$.
+              For both $[v_1) \not = \varnothing$ and $[v_2) \not = \varnothing$ they should be both in the image of $f_{V};\iota_{1,V}$ or $g_{V};\iota_{2,V}$ and according to assumption \ref{assumption:equal_predecessors} their sets of predecessors must be equal.
+        \item $u \sim_{S} v_1$ and $u \sim_{S} v_2$. This implies $u = f_{V};\iota_{1,V}(z_1) = f_{V};\iota_{1,V}(z_2)$ and $v_{1} = g_{V};\iota_{2,V}(z_1)$ and $v_{2} = g_{V};\iota_{2,V}(z_2)$.
+              By assumption \ref{assumption:equal_predecessors} it must be the case that $[g_{V}(z_1)) = [g_{V}(z_2))$ which implies that $[v_1) = [v_2)$.
+              The case when $u = g_{V};\iota_{2,V}(z_1)$ is symmetric.
+        \item $u \sim v_1$ and $u \sim v_2$.
+Same as above, by assumption \ref{assumption:non_ambiguous_predecessors} both $v_1$ and $v_2$ should be in the image of $f_{V};\iota_{1,V}$ or $g_{V};\iota_{2,V}$ and by assumption \ref{assumption:equal_predecessors} $[v_1) = [v_2)$.
+    \end{itemize}
+
+    Let's show that the definition does not depend on a particular choice of $v$ in (2).
+    Suppose, $[u) = \varnothing$ and there exist $v_1$ and $v_2$ such that $<_{X+Y}^{\mu}(v_1) = e_1$ and $<_{X+Y}^{\mu}(v_2) = e_2$ and $e_1 \not = e_2$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
+    The existence of a path from $[u]$ to $[v_1]$ means that there is $v_1' \sim v_1$ such that there is a path from $u'$ to $v_1'$ and similarly for $v_2$ and $v_2'$.
+    We will proceed by case analysis.
     \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
+            \item If $v_1 = v_1'$ or $v_2 = v_2'$ it means that there is a path from $u$ to $v_1$ or from $u$ to $v_2$ and since $[v_1) \not = \varnothing$ as well as $[v_2) \not = \varnothing$ it must be the case that $[u) \not = \varnothing$ which contradicts the assumption that $[u) = \varnothing$.
+            \item Suppose $v_1 \sim_{S} v_1'$ which means there exists $z_1 \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_1' = g_{V};\iota_{2,V}(z_1)$ and suppose $v_2 \sim_{S} v_2'$ which further means there exists $z_2 \in V_{Z}$ such that $v_2 = f_{V};\iota_{1,V}(z_2)$ and $v_2' = g_{V};\iota_{2,V}(z_2)$.
+                  Then we have
+                  \begin{align*}
+                    <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1)) &= <_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z_1))\\
+                                                              &= \iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_1)))
+                \end{align*}
                   and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
+                  \begin{align*}
+                    <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2)) &= <_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z_2))\\
+                                                              &= \iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_2)))
+                \end{align*}
+                and $e_1 \not = e_2$ implies $<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_1)) \not = <_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_2))$ which contradicts the assumption \ref{assumption:equal_predecessors}.
+                The case if $v_2 = g_{V};\iota_{2,V}(z_2)$ and $v_1 = f_{V};\iota_{1,V}(z_1)$ ultimately contradicts the assumption \ref{assumption:non_ambiguous_predecessors} as it entails that both
+                \[
+                    <_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_1)) \not = \varnothing
+                \]
+                and
+                \[
+                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2)) \not = \varnothing
+                \]
+                The cases when $v_1 = g_{V};\iota_{2,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$, and $v_1 = g_{V};\iota_{2,V}(z_1)$ and $v_2 = g_{V};\iota_{2,V}(z_2)$ are analogous.
+            \item Suppose $v_1 \sim_{S} v_1'$ and $v_2 \sim v_2'$ via $w = (x_1, \ldots, x_n)$ such that there is a path from $u$ to $v_1'$ and from $u$ to $x_n = v_2'$.
+                  The mere fact that $v_2 \sim v_2'$ implies that $v_2$ should have a pre-image in $V_{Z}$ and the same holds for $v_1$.
+                  By the same reasoning as above it must be the case that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1)) = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))$.
+            \item The case $v_1 \sim v_1'$ via $w = (x_1, \ldots, x_n)$ and $v_2 \sim_{S} v_1$ is symmetric to the above.
+            \item The case when $v_1 \sim v_1'$ and $v_2 \sim v_2'$ is analogous.
     \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
+
+    Let's check that this is a well-defined function.
+    Suppose $v_1, v_2$ are vertices and $v_1 \sim v_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])$.
     \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
+        \item If $[v_2) \not = \varnothing$, then, by definition we have
+                \begin{align*}
+                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
+                     &= \text{ as $v_2 \sim v_2$ and $[v_2) \not = \varnothing$ }\\
+                     &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_2])
+                \end{align*}
+        \item The case when $[v_1) \not = \varnothing$ is symmetric.
     \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
+    % $v_1 \sim v_2$ implies that either $v_1 = v_2$ and the well-definedness is trivial, or there exists a sequence $w = (x_1, \ldots, x_n)$ such that $v_1 = x_1$, $v_2 = x_n$ and $x_i \sim_{S} x_{i+1}$ for $i < n$.
+    % We will proceed by induction on the length of $w$.
+    % \begin{itemize}
+    %     \item $|w|$ and then $v_1 \sim_{S} v_2$ which means there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_{2} = g_{V};\iota_{2,V}(z)$.
+    %           Consider the following cases.
+    %           \begin{itemize}
+    %             \item $[f_{V}(z)) \not = \varnothing$ and $g_{V}(z) = \varnothing$.
+    %                   Then, 
+    %                   \[
+    %                     <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
+    %                   \]
+    %                   and
+    %                   \[
+    %                     <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
+    %                   \]
+    %                   and we have well-definedness by definition.
+    %             \item $f_{V}(z) = \varnothing$ and $g_{V}(z) \not = \varnothing$ is symmetric to the above.
+    %           \end{itemize}
+    %     \item Assume that if $v_1 \sim v_2$ via $w = (x_1, \ldots, x_n)$ then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])$.
+    %     \item Now suppose $v_1 \sim v_2$ via $w = (x_1, \ldots, x_n, x_{n+1})$.
+    %           We know that 
+    %           \[
+    %             <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[x_n])
+    %           \]
+    %           and that
+    %           \[
+    %             <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[x_n]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])
+    %           \]
+    %           by the same argument as in the base case,
+    %           which implies
+    %           \[
+    %             <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])
+    %           \]
+    % \end{itemize}
     We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
 
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
+    We also need to make sure that $([]_{V},[]_{E})$ is homomorphic with respect to $<^{\mu}$.
+    That is, we need to check if $[v) \not = \varnothing$, then $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+    Since $v \sim v$ this is homomorphic by definition.
+    Similarly for edges.
+
     Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
@@
     which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
+    We then first define an auxiliary relation $\consistency^{\hashtag}$ similarly to how $<^{\mu}_{\sfrac{X+Y}{\sim}}$ was defined.
+We begin with defining $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ for vertices.
+
+    \begin{enumerate}
+        \item If there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $u \sim v$, we let
+              \begin{align*}
+                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
+                &=\\
+                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+            \end{align*}
+        \item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}^{\hashtag}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
+        Then we let
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+        \]
+    \end{enumerate}
+
+    Next we define $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}$ for edges.
+
     \begin{enumerate}
@@
                 \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
+        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[e]$ to $[v]$
         then
         \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]))
         \]
@@
 
-    and for vertices
+    The well-definedness of this construction follows by the same argument as the well-definedness of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
+    Then we define
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+    \]
+    and
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
+    \]
+
+    to be the reflexive, symmetric and transitive closures of $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$.
+
+    More precisely, 
+    \[
+      (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
+    \]
+    where $c$ denotes a closure, is the smallest set such that
+    \begin{itemize}
+        \item $[v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$
+        \item if $[u] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
+        \[
+            [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])))^{c}
+        \]
+        \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
+        \[
+            [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
+        \]
+        \item for any sequence $(x_1, \ldots, x_n)$ such that
+        \[
+            [x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]))
+        \] or 
+        \[
+            [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i}]))
+        \] for $i < n$ (respectively, when $[x_{i}] \in \sfrac{E_{X} + E_{Y}}{\sim}$) both
+        \[
+            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n}])))^{c}
+        \]
+        and
+        \[
+            [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{1}])))^{c}
+        \]
+    \end{itemize}
+    Similarly for
+    \[
+        (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
+    \]
+
+    Let's check that $([]_{V},[]_{E})$ is a homomorphism with respect to $\consistency$.
+    We need to check, if 
+    \begin{align*}
+        &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))\\
+        &\subseteq\\
+        &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+    \end{align*}
+
     \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
+        \item The first case is when $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$.
+              Then the property trivially holds as the empty set is a subset of any set.
+        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$.
+              Then, 
+              \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}})(v))
+              \]
+              since $v \sim v$,
+              and hence 
+              \begin{align*}
+                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}})(v))\\
+                &\subseteq\\
+                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                &\subseteq\\
+                &(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c} = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+              \end{align*}
+              by recalling that $[[]_{V}^{\consistency}, []_{E}^{\consistency}] = [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]$.
     \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
+
+    The cases for edges are analogous.
+
+    % and for vertices
+    % \begin{enumerate}
+
+    % \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
+    % \begin{align*}
+    %     &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
+    %     &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
+    % \end{align*}
+    % \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
+    % \[
+    %     \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
+    % \]
+    % \end{enumerate}
+    % where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    % % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
+
+    % We then define
+    % \[
+    %   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
+    % \]
+    % \[
+    %   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
+    % \]
+    % where $c$ is a reflexive, symmetric and transitive closure.
+
+    % Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
+
+    % \begin{itemize}
+    %     \item Let us show well-definedness.
+    %     \begin{itemize}
+
+    %         \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
+    %         \[
+    %             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
+    %         \]
+    %           $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
+    %           \begin{align*}
+    %             &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
+    %             & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
+    %           \end{align*}
+    %           Similarly,
+    %           \begin{align*}
+    %             &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
+    %             &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
+    %         \end{align*}
+    %           and we get well-definedness on the nose.
+    %           We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
+    %         \item Now suppose that both 
+    %         \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
+    %         \quad 
+    %         \text{and} 
+    %         \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
+    %         \]
+    %         This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
+    %               \[
+    %                 \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
+    %               \]
+    %         \end{itemize}
+    %     \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
+    %         \begin{align*}
+    %          &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
+    %          &\subseteq
+    %          \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
+    %         \end{align*}
+    %         By definition, 
+    %         \begin{align*}
+    %             &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
+    %              [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
+    %         \end{align*}
+    %         and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
+    %         \begin{align*}
+    %             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
+    %             &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
+    %         \end{align*}
+    %          and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
+    %         If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
+    %         Similarly for edges.
+    % \end{itemize}
 
@@
     Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
+    We will need to check that this a well-defined function as well as that it is indeed a homomorphism.
+
+    Let's say that $v_1 \sim v_2$ then we need to show that $u_{V}([v_1]) = u_{V}([v_2])$.
+    $v_1 \sim v_2$ means there is a sequence $w = (x_1, \ldots, x_n)$ such that $v_1 = x_1$, $v_2 = x_n$ and for all $i < n$ $x_i \sim_{S} x_{i+1}$. We will proceed by induction on the length of $w$.
+    \begin{itemize}
+        \item $|w| = 2$, then $v_1 \sim_{S} v_2$ and there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V})(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and we have
+        \[
+            u_{V}([v_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1({f_{V}(z)})
+        \] 
+        and 
+        \[
+            u_{V}([v_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
+        \] 
+        which are equal by commutativity.
+        \item Assume if $v_1 \sim v_2$ via $w : |w| \leq n$ then $u_{V}([v_1]) = u_{V}([v_2])$.
+        \item Suppose $v_1 \sim v_2$ via $w = (x_1, \ldots, x_n, x_{n+1})$.
+              By hypothesis, we know that $u_{V}([v_1]) = u_{V}([x_1]) = u_{V}([x_n])$ and that $u_{V}([x_n]) = u_{V}([x_{n+1}]) = u_{V}([v_2])$ and hence $u_{V}([v_1]) = u_{V}([v_2])$.
+    \end{itemize}
+    Because $v_1 \sim t_2$, it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
     Then, we have
     \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
+    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1({f_{V}(z)})
     \] and 
@@
     \] which are equal by commutativity.
+
+    Let's show that such $(u_{V},u_{E})$ is unique.
+    Assume that there is $u'_{V}$ such that $j_{1,V}(v) = u'_{V}([\iota_{1,V}(v)])$ and $j_{2,V}(v) = u'_{V}([\iota_{2,V}(v)])$.
+    Then we have $u_{V}[\iota_{1,V}(v)] = j_{1,V}(v) = u'_{V}([\iota_{1,V}(v)])$ for all $v \in V_{X}$ (similarly for $v \in V_{Y}$) and hence $u_{V} = u'_{V}$.
+    Similarly for $u_{E}$.
+
+    Let's now check if $u = (u_{V},u_{E})$ is a homomorphism.
+    \begin{enumerate}
+        \item The first condition to check is if $u_{V}^{*}(s_{\sfrac{X+Y}{\sim}}([e])) = s_{Q}(u_{E}([e]))$. 
+              By construction of $s_{\sfrac{X+Y}{\sim}}$ we have
+              \[
+                u_{V}^{*}(s_{\sfrac{X+Y}{\sim}}([e])) = u_{V}^{*}([s_{X+Y}(e)]^{*})
+              \]
+              Suppose $e = \iota_{E_{X}}(e')$ and then
+              \[
+                u_{V}^{*}([s_{X+Y}(e)]^{*}) = u_{V}^{*}([s_{X+Y}(\iota_{1,E}(e'))]^{*}) = u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*})
+              \]
+              Since $j_1$ and $j_2$ are homomorphisms, it must be the case that $s_{Q}(j_{1,E}(e')) = j_{1,V}^{*}(s_{X}(e'))$.
+              By definition, 
+              \begin{align*}                
+                u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*}) &= j_{1,V}^{*}((s_{X}(e')))\\ 
+                &= s_{Q}(j_{1,E}(e'))\\
+                &= s_{Q}(u_{E}([\iota_{1,E}(e')]))\\
+                &= s_{Q}(u_{E}([e]))
+            \end{align*}
+            Similarly for $e = \iota_{E_{Y}}(e')$.
+        \item The case for targets is analogous.
+        \item The next condition to check is if $[[v]) \not = \varnothing$ then 
+    \[
+        <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
+    \]
+    We need to consider a few cases.
+    \begin{itemize}
+        \item Suppose $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))]$ for $u \sim v$.
+              Then assume $u = \iota_{1,V}(u_{x})$ and then 
+              \begin{align*}
+                [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(\iota_{1,V}(u_{x})))]\\
+                &= [\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(u_{x})))]
+            \end{align*}
+            and
+            \begin{align*}
+                u_{E}([\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(u_{x})))]) &= j_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(u_{x})))\\
+                &= <_{Q}^{\mu}(\iota_{V_{Q}}(j_{1,V}(u_{x})))\\
+                &= <_{Q}^{\mu}(\iota_{V_{Q}}(u([\iota_{1,V}(u_{x})])))\\
+                &= <_{Q}^{\mu}(\iota_{V_{Q}}(u([v])))
+            \end{align*}
+            and we have
+            \[
+                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
+            \]
+            The case when $u = \iota_{2,V}(u_{y})$ is symmetric.
+        \item Suppose $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))$ such that there is a path from $[v]$ to $[v']$ and $v$ has no pre-image in $V_{Z}$.
+              By the argument above, we know that
+              \[
+                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v']))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']))
+              \]
+              and by definition
+              \[
+                u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v'])) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
+              \]
+              since $(u_{V},u_{E})$ preserves sources and targets and there is a path from $[v]$ to $[v']$, there is a path from $u_{V}([v])$ to $u_{V}([v'])$ and because $Q$ is an e-hypergraph
+              \[
+                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v']))) = <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v])))
+              \]
+              and finally
+              \[
+                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
+              \]
+        \item Suppose $v = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$. Then 
+        \[
+            <^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [g_{V};\iota_{V_{Y}};<^{\mu}_{Y};\iota_{2,E}(z)]
+        \]
+        Because $j_2$ is a homomorphism, it must be the case that
+        \[
+         j_{2,E}(g_{V};\iota_{V_{Y}};<^{\mu}_{Y}(z)) = <^{\mu}_{Q}(g_{V};j_{2,V};\iota_{V_{Q}}(z))
+        \]
+        By definition,
+        \begin{align*}
+            <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([\iota_{1,V}(f_{V}(z))])))\\
+            &= <^{\mu}_{Q}(\iota_{V_{Q}}(j_{1,V}(f_{V}(z))))\\
+            &\text{by commutativity}\\
+            &= <^{\mu}_{Q}(\iota_{V_{Q}}(j_{2,V}(g_{V}(z))))\\
+            &= j_{2,E}(g_{V};\iota_{V_{Y}};<^{\mu}_{Y}(z))
+        \end{align*}
+        And also
+        \begin{align*}
+            u_{E}([g_{V};\iota_{V_{Y}};<^{\mu}_{Y};\iota_{2,E}(z)]) &= j_{2,E}(g_{V};\iota_{V_{Y}};<^{\mu}_{Y}(z))\\
+            &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v])))
+        \end{align*}
+        \item Suppose $v = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$. Then 
+        \[
+            <^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [f_{V};\iota_{V_{X}};<^{\mu}_{X};\iota_{1,E}(z)]
+        \]
+        Because $j_1$ is a homomorphism, it must be the case that
+        \[
+         j_{1,E}(f_{V};\iota_{V_{X}};<^{\mu}_{X}(z)) = <^{\mu}_{Q}(f_{V};j_{1,V};\iota_{V_{Q}}(z))
+        \]
+        By definition,
+        \begin{align*}
+            <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([\iota_{1,V}(f_{V}(z))])))\\
+            &= <^{\mu}_{Q}(\iota_{V_{Q}}(j_{1,V}(f_{V}(z))))\\
+            &= j_{1,E}(f_{V};\iota_{V_{X}};<^{\mu}_{X}(z))
+        \end{align*}
+        And also
+        \begin{align*}
+            u_{E}([f_{V};\iota_{V_{X}};<^{\mu}_{X};\iota_{1,E}(z)]) &= j_{1,E}(f_{V};\iota_{V_{X}};<^{\mu}_{X}(z))\\
+            &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v])))
+        \end{align*}
+        \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
+        \item Suppose that $v$ has no pre-image in $V_{Z}$ and $v = \iota_{1,V}(v')$ such that $[v') \not = \varnothing$. Then,
+        \begin{align*}
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([\iota_{1,V}(v')]))\\
+            &= [<_{X+Y}^{\mu}(\iota_{V_{X}+V_{Y}})(\iota_{1,V}(v'))]\\
+            &= [\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v')))]
+        \end{align*}
+        By definition,
+        \begin{align*}
+            u_{E}([\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v')))]) &= j_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v')))\\
+            &= <_{Q}^{\mu}(\iota_{V_{Q}}(j_{1,V}(v')))\\
+            &= <_{Q}^{\mu}(\iota_{V_{Q}}(u_{V}([\iota_{1,V}(v')])))\\
+            &= u_{E}(<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))
+        \end{align*}
+        \item The last case to consider is when the predecessor of $[v]$ is defined by the predecessor of $[v']$ such that there is a path from $[v]$ to $[v']$. That is,
+        \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
+        \]
+        Because $u$ preserves sources and targets, there is a path between $u_{V}([v])$ and $u_{V}([v'])$.
+        Then
+        \begin{align*}            
+        &u_{E}(<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+        &= u_{E}(<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])))\\
+        &\text{(by reusing the results above for $v'$)}\\
+        &= <_{Q}^{\mu}(\iota_{V_{Q}}(u_{V}([v'])))\\
+        &\text{(because $Q$ is an e-hypergraph and $u_{V}(v)$ and $u_{V}(v')$ are connected)}\\
+        &= <_{Q}^{\mu}(\iota_{V_{Q}}(u_{V}([v])))
+        \end{align*}
+    \end{itemize}
+    The cases for edges are analogous.
+    \item The last condition to check is if
+    \[
+    [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+    \]
+    Recall that
+    \[
+      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
+    \]
+    and first we will check if
+    \[
+      [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+    \]
+    Consider cases
+    \begin{itemize}
+        \item If $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \varnothing$ then the inclusion is trivial and hence we assume that
+        $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$
+        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $v = \iota_{1,V}(v_{X})$, then
+        \begin{align*}
+            &\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+            &= [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))\\
+            &= [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(\iota_{1,V}(v_{X}))))\\
+            &= [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}([\iota_{1,V};\iota_{V_{X} + V_{Y}},\iota_{1,E};\iota_{E_{X} + E_{Y}}]^{*}(\consistency_{X}(\iota_{V_{X}}(v_{X}))))\\
+            &= [\iota_{1,V};[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},\iota_{1,E};[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X}(\iota_{V_{X}}(v_{X})))
+        \end{align*}
+        and
+        \begin{align*}
+            &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+            &=\\
+            &[\iota_{1,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{1,E};[]_{E};u_{E};\iota_{E_{Q}}]^{*}(\consistency_{X}(\iota_{V_{X}}(v_{X})))\\
+            &=\\
+            &[j_{1,V};\iota_{V_{Q}},j_{1,E};\iota_{E_{Q}}](\consistency_{X}(\iota_{V_{X}}(v_{X})))\\
+            &\subseteq\\
+            &\consistency_{Q}(j_{1,V};\iota_{V_{Q}}(v_{X}))\\
+            &=\\
+            &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+        \end{align*}
+        \item The case when $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $v = \iota_{2,V}(v_{Y})$ is symmetric.
+        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $v = f_{V};\iota_{1,V}(z)$, then
+              \begin{align*}
+                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                &=\\
+                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{2,V};\iota_{V_{X} + V_{Y}}(g_{V}(z))))\\
+                &=\\
+                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}([\iota_{2,V};\iota_{V_{X} + V_{Y}},\iota_{2,E};\iota_{E_{X}+E_{Y}}]^{*}(\consistency_{Y}(\iota_{V_{Y}}(g_{V}(z)))))
+            \end{align*}
+            and
+            \begin{align*}
+                &[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+                &=\\
+                &[\iota_{2,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{2,E};[]_{E};u_{E};\iota_{E_{Q}}]^{*}(\consistency_{Y}(\iota_{V_{Y}}(g_{V}(z))))\\
+                &=\\
+                &[j_{2,V};\iota_{V_{Q}},j_{2,E};\iota_{E_{Q}}]^{*}(\consistency_{Y}(\iota_{V_{Y}}(g_{V}(z))))\\
+                &\text{(by homomorphism-ness of $j_2$)}\\
+                &\subseteq\\
+                &\consistency_{Q}(j_{2,V};\iota_{V_{Q}}(g_{V}(z)))\\
+                &\consistency_{Q}(j_{1,V};\iota_{V_{Q}}(f_{V}(z)))\\
+                &=\\
+                &\consistency_{Q}(u_{V};\iota_{V_{Q}}([f_{V};\iota_{1,V};(z)]))\\
+                &=\\
+                &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+            \end{align*}
+        \item The remaining case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
+        \end{itemize}
+       So far we have shown that
+       \[
+        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+       \]
+       and
+       \[
+        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))) \subseteq \consistency_{Q}(u_{E};\iota_{E_{Q}}([e]))
+       \] follows by the same argument.
+       Then, to check if
+       \[
+        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+       \]
+       we need to verify if
+       \begin{itemize}
+        \item $[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\{[v]\}) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))$.
+              This holds because $\consistency_{Q}$ is reflexive.
+        \item $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
+        \[
+            [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\{[v]\}) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([u]))
+        \]
+              The first bit implies that $u_{V};\iota_{V_{Q}}([u]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))$ and since $\consistency_{Q}$ is symmetric it must be the case that $u_{V_{Q}};\iota_{V_{Q}}[v] \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([u]))$
+        \item There exists a sequence $w = (x_1, \ldots, x_n)$ such that 
+        \[
+            \iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_i]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]))
+        \]
+
+        or 
+        \[
+            \iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i}]))
+        \]
+        for $i < n$ (or similarly for edges), then $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
+        \begin{itemize}
+            \item Base case is when $w = (x_1,x_2)$.
+                  Then the statement is trivial as 
+                  \begin{align*}
+                  &[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_2])))\\
+                  &\subseteq\\
+                  &\consistency_{Q}(u_{V};\iota_{V_{Q}}([x_2]))
+                  \end{align*}
+            \item Suppose the statement is true for all $w$ such that $|w| \leq n$.
+            \item Consider a sequence $w = (x_1, \ldots, x_{n}, x_{n+1})$.
+                  By assumption, we know that $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
+                  The fact that $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n+1}]))$ (or its symmetric counterpart) implies that $u_{V};\iota_{V_{Q}}([x_{n}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))$ and by the transitivity and symmetry of $\consistency_{Q}$ it follows that
+                \[
+                    u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))
+                \]
+                and
+                \[
+                    u_{V};\iota_{V_{Q}}([x_{n+1}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{1}]))
+                \]
+        \end{itemize}
+       \end{itemize}
+       Therefore, we have
+       \[
+        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+       \]
+       and similarly for edges. Which concludes that $(u_{V},u_{E})$ is a homomorphism.
+    \end{enumerate}
 \end{proof}
+
+\begin{figure}
+    \[
+        \scalebox{0.5}{
+        \tikzfig{appendix/pushout_example}
+        }
+\]
+\caption{Pushout in $\catname{EHyp_{\Sigma}}$}
+\label{fig:pushout_example}
+\end{figure}
+
+\begin{example}
+    Consider a diagram in Figure~\ref{fig:pushout_example} where the subscripts define the corresponding morphisms.
+    E-hypergraph $Q$ is not a pushout as there is no homomorphism $u$ that would complete the diagram.
+    In particular, there is no $(u_{V},u_{E})$ such that 
+    \[
+        [u_{V};\iota_{P_{V}},u_{E};\iota_{P_{E}}]^{*}(\consistency_{Q}(\iota_{V_{Q}}(q_1))) \subseteq \consistency_{P}(u_{V};\iota_{P_{V}}(q_1))
+    \]
+    If we let $u_{V}(i_{1,V}(v)) = j_{1,V}(v)$ for all $v \in V_{X}$ and $u_{V}(i_{2,V}(v)) = j_{2,V}(v)$ and similarly for edges.
+    Then 
+    \begin{align*}
+        &[u_{V};\iota_{P_{V}}, u_{E};\iota_{P_{E}}]^{*}(\consistency_{Q}(\iota_{V_{Q}}(q_1)))\\
+        &=\\
+        &[u_{V};\iota_{P_{V}}, u_{E};\iota_{P_{E}}]^{*}(\{q_1, q_2, q_3, q_4, f, h, g\})\\
+        &=\\
+        &[\iota_{P_{V}}(j_{1,V}(x_1)), \iota_{P_{V}}(j_{1,V}(x_2)), \iota_{P_{V}}(j_{1,V}(x_3)),\\
+         &\; \iota_{P_{V}}(j_{1,V}(x_4)), \iota_{P_{E}}(j_{1,E}(f)), \iota_{P_{E}}(j_{1,E}(g)), \iota_{P_{V}}(j_{2,V}(y_1)),\\
+         &\; \iota_{P_{V}}(j_{2,V}(y_2)), \iota_{P_{E}}(j_{2,E}(h))]\\
+        &=\\
+        &[p_1, p_2, p_3, p_4, f, h, g]\\
+        &\not \subseteq\\
+        &\consistency_{P}(u_{V};\iota_{P_{V}}(q_1))\\
+        &=\\
+        &\consistency_{P}(j_{1,V};\iota_{P_{V}}(x_1))\\
+        &=\\
+        &\consistency_{P}(\iota_{P_{V}}(p_1))\\
+        &=\\
+        &[p_1, f, h, p_2]
+    \end{align*}
+\end{example}
 

### 2024-06-13 LICS/sections/appendix.tex
arxiv version added, repo restructured

--- 2024-05-01 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
-%-------------------- INTRODUCTION
-
-\section{Introduction}
-
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
-This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
-
-A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
-
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
-
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
-
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
-
-\subsection{E-graphs}
-
-\begin{figure*}
-\[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
-    }
-\]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
-
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
-
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
-
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
-
-\subsection{Semilattice Enriched Categories}
-
-\begin{figure}
-\[
-	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
-	}
-\]
-\captionsetup{belowskip=-5ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
-\end{figure}
-
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
-\[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
-\]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
-
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
-
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
-\[
-    \tikzfig{categorical-semantics/egraph-translation-1}
-\]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
-As it would involve needless additional presentational complexity we leave it as further work.
-
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
-
-\subsection{Combinatorial Representation of Enriched String Diagrams}
-
-In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
-For example, we have the following equivalences between diagrams.
-%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
-	}
-\]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
-Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
-
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
-These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
-
-\subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
-They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
-
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
-
-%\subsection{E-matching and E-rewriting}
-%
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
-In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
-In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
-Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-%\begin{figure}
-%\[
-%	\tikzfig{combinatorial_semantics/example-rewriting}
-%\]
-%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
-REDO THIS}
-%\end{figure}
-
-\subsection{Related Work}
-
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
-We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
-
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
-
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
-\end{figure}
-
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
+\section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
+
+In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
+We will also show the uniqueness of a boundary pushout complement.
+These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
+
+\begin{proposition}
+    Two cospans in $\catname{Hyp}(\Sigma)$ (\textit{i.e.}, morphisms in $\HypI{\Sigma}$) $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ and $n \xrightarrow{f'} \mathcal{G}' \xleftarrow{g'} m$ are equal if and only if
+    the following two cospans  $\catname{EHyp}(\Sigma)$ (\textit{i.e.} morphisms in $\Ecospans$)
+    \[
+    n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m
+    \]
+    \[
+        n \xrightarrow{f_{ext}'} n \xrightarrow{f_{int}'} \mathcal{G} \xleftarrow{g_{int}'} m \xleftarrow{g_{ext}'} m    
+    \]
+    such that $f = f_{ext};f_{int}, f'=f_{ext}';f_{int}'$ (respectively, $g = g_{ext};g_{int}, g' = g_{ext}';g_{int}'$) are equal.
 \end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
+This essentially means that $\catname{\HypI{\Sigma}}$ faithfully embeds into $\Ecospans$. 
 \begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
+    Recall that cospans are equal when they are isomorphic.
+    The proposition means that the following cospans are isomorphic (where $\alpha$ is an isomorphism)
+    \[\begin{tikzcd}
+            & \mathcal{G} \arrow[dd, "\alpha"] &                                                            \\
+    n \arrow[ru, "f", bend left] \arrow[rd, "f'"', bend right] &                                  & m \arrow[lu, "g"', bend right] \arrow[ld, "g'", bend left] \\
+            & \mathcal{G}'                     &                                                           
+    \end{tikzcd}
+    \]
+    if and only if the following cospans are isomorphic
+    \[
+        \begin{tikzcd}
+            & n \arrow[r, "f_{int}"] \arrow[dd, "\beta"] & \mathcal{G} \arrow[dd, "\alpha"] & m \arrow[l, "g_{int}"'] \arrow[dd, "\gamma"] &                                                                        \\
+n \arrow[ru, "f_{ext}", bend left] \arrow[rd, "f_{ext}'"', bend right] &                                            &                                  &                                              & m \arrow[lu, "g_{ext}"', bend right] \arrow[ld, "g_{ext}'", bend left] \\
+            & n \arrow[r, "f_{int}'"']                    & \mathcal{G}'                     & m \arrow[l, "g_{int}'"]                      &
+\end{tikzcd}    
+    \]
+    such that $f = f_{ext};f_{int}, f'=f_{ext}';f_{int}'$ (respectively, $g = g_{ext};g_{int}, g' = g_{ext}';g_{int}'$).
+    The direction from right to left is straightforward: since the bottom diagram commutes $f;\alpha = f_{int};f_{ext};\alpha = f_{int}';f_{ext}' = f'$ and similarly for $g$.
+
+    Let's show the direction from left to right.
+    Note that $f_{ext}$ and $f_{ext}'$ have inverses since they are injective and surjective (the injection is per the definition and surjection is because domain and codomain are equal).
+    Then we need $f_{ext};\beta = f_{ext}'$ and we define $\beta = f_{ext}^{-1};f_{ext}'$.
+    Similarly $\beta^{-1} = f_{ext}'^{-1};f_{ext}$. 
+    We can check that $\beta^{-1};\beta = f_{ext}'^{-1};f_{ext};f_{ext}^{-1};f_{ext}' = f_{ext}'^{-1};id;f_{ext}' = f_{ext}'^{-1};f_{ext}' = id$.
+    Then we have that $f_{ext};\beta;f_{int}' = f_{ext};f_{ext}^{-1};f_{ext}';f_{int}' = f_{ext}';f_{int}' = f' = f;\alpha = f_{ext};f_{int};\alpha$.
+    Since $f_{ext}$ is surjective, we have $\beta;f_{int}' = f_{int};\alpha$.
+    Analogously for $\beta^{-1}$ and the right-hand sides of the diagrams.
 \end{proof}
 
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
-
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
-\end{figure}
-
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
+\subsection{Existence of pushout}
 
@@
 \begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
+    Both $<$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
     To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
@@
     \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
     \item \label{assumption:equal_predecessors} $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
@@
     Consider relations
+    \ifdefined\ONECOLUMN
+    \begin{align*}
+            S_{V} &= \;\{
+            (x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |
+            \exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and }\\
+            &\qquad y_j = g_{V};\iota_{2,V}(z) \text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
+            \}\\
+            &\cup \;\{
+                (y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |
+          \exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and }\\
+          &\qquad y_j = g_{V};\iota_{2,V}(z) \text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
+        \}\\
+        &\cup \;\{(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\}\\
+        S_{E} &= \;\{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
+    \end{align*}
+    \else 
     \begin{align*}    
@@
     \end{align*}
+    \fi
     % \begin{align*}
@@
 We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
+\ifdefined \ONECOLUMN
+\begin{align*}
+    \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{
+        &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
+        &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\}    
+\end{align*}
+\else
     \begin{align*}
@@
     \end{align*}
+\fi
 We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
@@
               \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
+        \item \label{def:child_respects_connectivity} $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
               If there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
@@
     \end{enumerate}
+    Note that the existence of a path in the definition (2) implies that there exists $v' \sim v$ such that there is a path from $u$ to $v'$ and furthermore $v \not = v'$.
+    If $v' = v$ then it would mean that $[u) \not = \varnothing$ since $[v) \not = \varnothing$ and there is a path from $u$ to $v$.
     There is no case when $u$ has a pre-image in $V_{Z}$ and yet $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e$ and there is a path from $[u]$ to $[v]$.
@@
 
-    Otherwise, we leave $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}{\sim}([u]))$ undefined.
+    Otherwise, we leave $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))$ undefined.
     % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
@@
         \item The case when $[v_1) \not = \varnothing$ is symmetric.
+        \item If both $[v_2) = \varnothing$ and $[v_1) = \varnothing$ but there exists $[v_3) \not = \varnothing$ such that $v_1 \sim v_3$, then
+        \begin{align*}
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_3))]\\
+             &= \text{ as $v_3 \sim v_3$ and $[v_3) \not = \varnothing$ }\\
+             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_3])\\
+             &= \text{ as $v_2 \sim v_1 \sim v_3$}\\
+             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_2])
+        \end{align*}
     \end{itemize}
@@
         \item If there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $u \sim v$, we let
+              \ifdefined \ONECOLUMN
+              \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))
+                =
+                [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+              \]
+              \else
               \begin{align*}
@@
             \end{align*}
-        \item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}^{\hashtag}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
+            \fi
+        \item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
         Then we let
@@
     Then we define
+    \ifdefined \ONECOLUMN
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \qquad \text{and} \qquad \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
+    \]
+    \else
     \[
@@
     \]
-
-    to be the reflexive, symmetric and transitive closures of $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$.
-
-    More precisely, 
-    \[
-      (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    where $c$ denotes a closure, is the smallest set such that
-    \begin{itemize}
-        \item $[v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$
-        \item if $[u] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
-        \[
-            [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])))^{c}
-        \]
-        \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
-        \[
-            [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
-        \]
-        \item for any sequence $(x_1, \ldots, x_n)$ such that
-        \[
-            [x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]))
+    \fi
+
+    as closures of $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ as below.
+    \[
+      (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}
+    \]
+    where $c$ denotes a closure and $\iota_{\sfrac{*}{\sim}}$ is $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}$ depending on whether $[x_i]$ comes from $V_{X} + V_{Y}$ or $E_{X} + E_{Y}$, is the smallest set such that
+    \begin{itemize}
+        \item $[x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$
+        \item if $[y] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x]))$ then 
+        \[
+            [x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([y])))^{c}
+        \]
+        % \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
+        % \[
+        %     [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
+        % \]
+        \item for any sequence $([x_1], \ldots, [x_n])$ such that
+        \ifdefined \ONECOLUMN
+        \[[x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}])) \qquad \text{or} \qquad [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))\]
+        \else
+        \[
+            [x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}]))
         \] or 
         \[
-            [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i}]))
-        \] for $i < n$ (respectively, when $[x_{i}] \in \sfrac{E_{X} + E_{Y}}{\sim}$) both
-        \[
-            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n}])))^{c}
+            [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))
+        \]
+        \fi
+         for $i < n$ both
+        \ifdefined \ONECOLUMN
+        \[
+            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c} \qquad \text{and} \qquad [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
+        \]
+        \else
+         \[
+            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c}
         \]
@@
         \[
-            [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{1}])))^{c}
-        \]
-    \end{itemize}
-    Similarly for
-    \[
-        (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
-    \]
-
+            [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
+        \]
+        \fi
+    \end{itemize}
     Let's check that $([]_{V},[]_{E})$ is a homomorphism with respect to $\consistency$.
     We need to check, if
+    \ifdefined \ONECOLUMN
+    \[
+        [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+        \subseteq
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+    \]
+    \else
     \begin{align*}
@@
     \end{align*}
+    \fi
 
@@
               and hence
+              \ifdefined \ONECOLUMN
+              \begin{align*}
+                [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}})(v))
+                &\subseteq
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                &\subseteq
+                (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c} = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+              \end{align*}
+              \else
               \begin{align*}
@@
               \end{align*}
+              \fi
               by recalling that $[[]_{V}^{\consistency}, []_{E}^{\consistency}] = [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]$.
@@
     So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
+    By the Proposition~\ref{prop:pushout_is_e_hypergraph} the constructed $\sfrac{X+Y}{\sim}$ is an e-hypergraph.
+    To show that it is a pushout we need to check the universal property.
     Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
@@
               \]
-        \item Suppose $v = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$. Then 
-        \[
-            <^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [g_{V};\iota_{V_{Y}};<^{\mu}_{Y};\iota_{2,E}(z)]
-        \]
-        Because $j_2$ is a homomorphism, it must be the case that
-        \[
-         j_{2,E}(g_{V};\iota_{V_{Y}};<^{\mu}_{Y}(z)) = <^{\mu}_{Q}(g_{V};j_{2,V};\iota_{V_{Q}}(z))
-        \]
-        By definition,
-        \begin{align*}
-            <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([\iota_{1,V}(f_{V}(z))])))\\
-            &= <^{\mu}_{Q}(\iota_{V_{Q}}(j_{1,V}(f_{V}(z))))\\
-            &\text{by commutativity}\\
-            &= <^{\mu}_{Q}(\iota_{V_{Q}}(j_{2,V}(g_{V}(z))))\\
-            &= j_{2,E}(g_{V};\iota_{V_{Y}};<^{\mu}_{Y}(z))
-        \end{align*}
-        And also
-        \begin{align*}
-            u_{E}([g_{V};\iota_{V_{Y}};<^{\mu}_{Y};\iota_{2,E}(z)]) &= j_{2,E}(g_{V};\iota_{V_{Y}};<^{\mu}_{Y}(z))\\
-            &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v])))
-        \end{align*}
-        \item Suppose $v = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$. Then 
-        \[
-            <^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]) = [f_{V};\iota_{V_{X}};<^{\mu}_{X};\iota_{1,E}(z)]
-        \]
-        Because $j_1$ is a homomorphism, it must be the case that
-        \[
-         j_{1,E}(f_{V};\iota_{V_{X}};<^{\mu}_{X}(z)) = <^{\mu}_{Q}(f_{V};j_{1,V};\iota_{V_{Q}}(z))
-        \]
-        By definition,
-        \begin{align*}
-            <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([\iota_{1,V}(f_{V}(z))])))\\
-            &= <^{\mu}_{Q}(\iota_{V_{Q}}(j_{1,V}(f_{V}(z))))\\
-            &= j_{1,E}(f_{V};\iota_{V_{X}};<^{\mu}_{X}(z))
-        \end{align*}
-        And also
-        \begin{align*}
-            u_{E}([f_{V};\iota_{V_{X}};<^{\mu}_{X};\iota_{1,E}(z)]) &= j_{1,E}(f_{V};\iota_{V_{X}};<^{\mu}_{X}(z))\\
-            &= <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v])))
-        \end{align*}
-        \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-        \item Suppose that $v$ has no pre-image in $V_{Z}$ and $v = \iota_{1,V}(v')$ such that $[v') \not = \varnothing$. Then,
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([\iota_{1,V}(v')]))\\
-            &= [<_{X+Y}^{\mu}(\iota_{V_{X}+V_{Y}})(\iota_{1,V}(v'))]\\
-            &= [\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v')))]
-        \end{align*}
-        By definition,
-        \begin{align*}
-            u_{E}([\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v')))]) &= j_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v')))\\
-            &= <_{Q}^{\mu}(\iota_{V_{Q}}(j_{1,V}(v')))\\
-            &= <_{Q}^{\mu}(\iota_{V_{Q}}(u_{V}([\iota_{1,V}(v')])))\\
-            &= u_{E}(<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))
-        \end{align*}
-        \item The last case to consider is when the predecessor of $[v]$ is defined by the predecessor of $[v']$ such that there is a path from $[v]$ to $[v']$. That is,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-        \]
-        Because $u$ preserves sources and targets, there is a path between $u_{V}([v])$ and $u_{V}([v'])$.
-        Then
-        \begin{align*}            
-        &u_{E}(<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
-        &= u_{E}(<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])))\\
-        &\text{(by reusing the results above for $v'$)}\\
-        &= <_{Q}^{\mu}(\iota_{V_{Q}}(u_{V}([v'])))\\
-        &\text{(because $Q$ is an e-hypergraph and $u_{V}(v)$ and $u_{V}(v')$ are connected)}\\
-        &= <_{Q}^{\mu}(\iota_{V_{Q}}(u_{V}([v])))
-        \end{align*}
         \end{itemize}
@@
         \item If $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \varnothing$ then the inclusion is trivial and hence we assume that
-        $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$
-        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $v = \iota_{1,V}(v_{X})$, then
-        \begin{align*}
-            &\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-            &= [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))\\
-            &= [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(\iota_{1,V}(v_{X}))))\\
-            &= [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}([\iota_{1,V};\iota_{V_{X} + V_{Y}},\iota_{1,E};\iota_{E_{X} + E_{Y}}]^{*}(\consistency_{X}(\iota_{V_{X}}(v_{X}))))\\
-            &= [\iota_{1,V};[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},\iota_{1,E};[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X}(\iota_{V_{X}}(v_{X})))
-        \end{align*}
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
+        \]
+        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) \not = \varnothing$ and $v \sim u$ and then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(u)))
+        \]
+        Assume $u = \iota_{1,V}(u_{x})$, then
+        \ifdefined \ONECOLUMN
+        \begin{align*}
+            [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(u)))
+            &=
+            [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(\iota_{1,V}(u_{x}))))\\
+            &=
+            [\iota_{1,V};[]_{V}^{\consistency},\iota_{1,E};[]_{E}^{\consistency}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))
+        \end{align*}
+        \else
+        \begin{align*}
+            &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(u)))\\
+            &=\\
+            &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(\iota_{1,V}(u_{x}))))\\
+            &=\\
+            &[\iota_{1,V};[]_{V}^{\consistency},\iota_{1,E};[]_{E}^{\consistency}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))
+        \end{align*}
+        \fi
+
         and
+        \ifdefined \ONECOLUMN
+        \begin{align*}
+            [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))
+            &=
+            [\iota_{1,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{1,E};[]_{E};u_{E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
+            &=
+            [j_{1,V};\iota_{V_{Q}},j_{1,E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
+            &\subseteq
+            \consistency_{Q}(\iota_{V_{Q}}(j_{1,V}(u_{x})))\\
+            &=
+            \consistency_{Q}(\iota_{V_{Q}}(u_{V}[\iota_{1,V}(u_{x})]))\\
+            &=
+            \consistency_{Q}(\iota_{V_{Q}}(u_{V}[u]))\\
+            &=
+            \consistency_{Q}(\iota_{V_{Q}}(u_{V}[v]))
+        \end{align*}
+        \else
         \begin{align*}
@@
             &=\\
-            &[\iota_{1,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{1,E};[]_{E};u_{E};\iota_{E_{Q}}]^{*}(\consistency_{X}(\iota_{V_{X}}(v_{X})))\\
+            &[\iota_{1,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{1,E};[]_{E};u_{E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
             &=\\
-            &[j_{1,V};\iota_{V_{Q}},j_{1,E};\iota_{E_{Q}}](\consistency_{X}(\iota_{V_{X}}(v_{X})))\\
+            &[j_{1,V};\iota_{V_{Q}},j_{1,E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
             &\subseteq\\
-            &\consistency_{Q}(j_{1,V};\iota_{V_{Q}}(v_{X}))\\
+            &\consistency_{Q}(\iota_{V_{Q}}(j_{1,V}(u_{x})))\\
             &=\\
-            &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-        \end{align*}
-        \item The case when $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $v = \iota_{2,V}(v_{Y})$ is symmetric.
-        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $v = f_{V};\iota_{1,V}(z)$, then
+            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}[\iota_{1,V}(u_{x})]))\\
+            &=\\
+            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}[u]))\\
+            &=\\
+            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}[v]))
+        \end{align*}
+        \fi
+        The case when $u = \iota_{2,V}(u_{y})$ is symmetric.
+        \item Suppose that $v$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and there exists $u$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) \not = \varnothing$ and such that there is an undirected path from $[v]$ to $[u]$ and then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))    
+        \]
+        By the argument above it follows that
+        \[
+            [u_{V};\iota_{Q_{V}},u_{E};\iota_{Q_{E}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))) \subseteq \consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))
+        \]
+        \end{itemize}
+       Since $(u_{V},u_{E})$ preserves paths, there is a path from $u_{V}([u])$ to $u_{V}([v])$ and since $Q$ is an e-hypergraph
+       \ifdefined \ONECOLUMN
+       \begin{align*}
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+        &=
+        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
+        &\subseteq
+        \consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))\\
+        &=
+        &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([v])))
+       \end{align*}
+       \else
        \begin{align*}
@@
         &=\\
-                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{2,V};\iota_{V_{X} + V_{Y}}(g_{V}(z))))\\
+        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
+        &\subseteq\\
+        &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))\\
         &=\\
-                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}([\iota_{2,V};\iota_{V_{X} + V_{Y}},\iota_{2,E};\iota_{E_{X}+E_{Y}}]^{*}(\consistency_{Y}(\iota_{V_{Y}}(g_{V}(z)))))
-            \end{align*}
-            and
-            \begin{align*}
-                &[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
-                &=\\
-                &[\iota_{2,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{2,E};[]_{E};u_{E};\iota_{E_{Q}}]^{*}(\consistency_{Y}(\iota_{V_{Y}}(g_{V}(z))))\\
-                &=\\
-                &[j_{2,V};\iota_{V_{Q}},j_{2,E};\iota_{E_{Q}}]^{*}(\consistency_{Y}(\iota_{V_{Y}}(g_{V}(z))))\\
-                &\text{(by homomorphism-ness of $j_2$)}\\
-                &\subseteq\\
-                &\consistency_{Q}(j_{2,V};\iota_{V_{Q}}(g_{V}(z)))\\
-                &\consistency_{Q}(j_{1,V};\iota_{V_{Q}}(f_{V}(z)))\\
-                &=\\
-                &\consistency_{Q}(u_{V};\iota_{V_{Q}}([f_{V};\iota_{1,V};(z)]))\\
-                &=\\
-                &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-            \end{align*}
-        \item The remaining case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-        \end{itemize}
+        &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([v])))
+    \end{align*}
+    \fi
        So far we have shown that
@@
         \scalebox{0.5}{
-        \tikzfig{appendix/pushout_example}
+        \tikzfig{../figures/appendix/pushout_example}
         }
 \]
+\captionsetup{belowskip=-3ex}
 \caption{Pushout in $\catname{EHyp_{\Sigma}}$}
@@
 
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
+% So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
+% The proof of the latter will follow below and the following corollaries will come in handy.
+% We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
 
 % By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
 
@@
     \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
+      \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))] = [e]$
+      This would mean that $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e$ is the predecessor of $e$.
+      \item Suppose $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ is defined and there is a path from $[e]$ to $[v]$.
+            This implies
+            \[
+                [e] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
+            \]
+            and
+            \[
+               e = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))
+            \]
+            The existence of a path from $[e]$ to $[v]$ means there is a path from $e$ to $v'$ such that $v' \sim v$.
+            \begin{itemize}
+                \item If $v' = v$, then there is a path from $e$ to $v$ which contradicts $e = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$
+                \item Consider the case when $v' \sim v$.
+                      $e = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ implies that both $e$ and $v$ are in the image of $(\iota_{1,V},\iota_{1,E})$ or $(\iota_{2,V},\iota_{2,E})$.
+                      Suppose $e = \iota_{1,E}(e_{x})$ and $v = \iota_{1,V}(v_{x})$.
+                      Since $v' \sim v$, $v_{x} = f_{V}(z_1)$. 
+                      And since there is a path from $v'$ to $e$, $v' = \iota_{1,V}(v'_{x})$ and $v'_{x} = f_{V}(z_2)$.
+                      This further implies
+                      \begin{align*}
+                        <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{x})))\\
+                                                                &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_1))))\\
+                                                                &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_2))))\\
+                                                                &= \iota_{1,E}(e_{x})
+                    \end{align*}
+                      and
+                      \[
+                        e_{x} = <_{X}(\iota_{V_{X}}(f_{V}(z_2))) = <_{X}(\iota_{V_{X}}(v_{x}'))
+                      \]
+                      which contradicts that there is a path from $e_{x}$ to $v_{x}'$.
+                      The case when $e = \iota_{2,E}(e_{y})$ and $v = \iota_{2,V}(v_{y})$ is symmetric.
             \end{itemize}
@@
 \end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
 
@@
     \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
+      \item 
+      \[
+        [e_1] = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2))]
+      \]
+      which implies $e_1 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2))$.
       Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
       \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
+        \item $[e_2] = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]$ and $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ which implies that $e_1$ is the predecessor of $e_2$ and vice versa which contradicts e-hypergraphness of $X+Y$.
+        \item
+            \[
+                [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
+            \]
+            such that $[e_1) = \varnothing$ and there is a path from $[e_1]$ to $[v]$, i.e. a path from $e_1$ to $v'$ such that $v' \sim v$.
+            \begin{itemize}
+                \item $v' \sim v$ and there is a path from $e_1$ to $v'$. 
+                Suppose $e_2 = \iota_{1,E}(e_{2,x})$ and $v = \iota_{1,V}(v_{x})$.
+                Since $e_1$ is the predecessor of $e_2$, the former should also be in the image of $\iota_{1,E}$, i.e.
+$e_1 = \iota_{1,E}(e_{1,x})$, and since there is a path from $e_1$ to $v'$, $v' = \iota_{1,V}(v'_{x})$.
+                $v' \sim v$ implies $v_{x} = f_{V}(z_1)$ and $v'_{x} = f_{V}(z_2)$.
                 Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
+                \begin{align*}
+                    <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{x})))\\
+                                                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_1))))\\
+                                                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_2))))\\
+                                                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{x}')))\\
+                                                            &= <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))\\
+                                                            &= <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))
+                \end{align*}
+                which contradicts $e_1 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2))$.
+                The case when $e_2 = \iota_{2,E}(e_{2,y})$ and $v = \iota_{2,V}(v_{y})$ is symmetric.
+            \end{itemize}
+        \end{itemize}
+            \item The other case is when          
+            \[
+                [e_1] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))]
+            \]
+            such that there is a path from $e_2$ to $v'_1$ and $v_1 \sim v'_1$ and $[e_2) = \varnothing$.
+            \begin{itemize}
+                \item The case when $[e_2] = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]$ is analogous to the last case above.
+                \item Suppose 
+                \[
+                    [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]
+                \] such that there is a path from $e_1$ to $v'_2$ and $v_2 \sim v'_2$ and $[e_1) = \varnothing$.
+                    Suppose $e_1 = \iota_{1,E}(e_{1,x})$ and $v_1 = \iota_{1,V}(v_{1,x})$ which implies $v'_2 = \iota_{1,V}(v'_{2,x})$.
+                    Since $v_1 \sim v_1'$, $v_{1,x} = f_{V}(z_1)$ and similarly $v'_{2,x} = f_{V}(z_2)$.
+                    \begin{align*}
+                        e_1 &= <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))\\
+                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{1,x})))\\
+                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_1))))\\
+                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_2))))\\
+                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v'_{2,x})))\\
+                            &= <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'_2))\\
+                            &= <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))
+                    \end{align*}
+                    And we get a contradiction to irreflexivity of $<_{X+Y}^{\mu}$.
             \end{itemize}
@@
 \begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
+        We need to show that if $[v] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
         We have two cases.
         \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
+            \item \[
+            [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(e_1))]
+            \]
+            By definition, $s_{\sfrac{X+Y}{\sim}}([e_1]) = [s_{X+Y}(e_1)]^{*}$ and $[v] \in [s_{X+Y}(e_1)]^{*}$ means there exists $v' \sim v$ such that $v' \in s_{X+Y}(e_1)$.
+            \begin{itemize}
+                \item $v' = v$ and then $v \in s_{X+Y}(e_1)$ and therefore $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]$.
+                \item $v' \sim v$ and then 
                 \begin{align*}
@@
             \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
+            \item \[
+                [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))]    
+            \]
+            such that there is $v_1' \sim v_1$ and there is a path from $e_1$ to $v_1'$ and $[e_1) = \varnothing$.
+            $[v] \in [s_{X+Y}(e_1)]^{*}$ means there exists $v' \sim v$ such that $v' \in s_{X+Y}(e_1)$.
+            \begin{itemize}
+                \item $v' = v$ and then $v \in s_{X+Y}(e_1)$. Since there is a path from $e_1$ to $v_1'$, there is also a path from $v'$ to $v_1'$ and by definition
+                \[
+                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
+                \]
+                \item $v' \sim v$ and then there is a path from $v'_1$ to $v'$. 
+                Suppose $v_1 = f_{V};\iota_{1,V}(z_1)$ and since $v_1' \in s_{X+Y}(e_1)$, $[e_1) = \varnothing$, both $v_1'$ and $v'$ should be in the image of $g_{V};\iota_{2,V}$.
+                Since $[v_1) \not = \varnothing$ it must be the case that for all $z$ $[f_{V}(z)) \not = \varnothing$ as well as for some $v'' \sim v'$ such that $v'' = f_{V};\iota_{1,V}(z'')$.
+                Clearly such $v''$ exists.
+                It is either $v$, if $v = f_{V};\iota_{1,V}(z'')$, or some $x_i$ in $(x_1, \ldots, x_n)$ such that $x_1 = v'$ and $x_n = v$.
+                \begin{align*}
+                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))]\\
+                          &= [\iota_{1,E}(<_{X}(\iota_{V_{X}}(f_{V}(z_1))))]\\
+                          &= [\iota_{1,E}(<_{X}(\iota_{V_{X}}(f_{V}(z''))))]\\
+                          &= [<_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z'')))]\\
+                          &= [<_{X+Y}(\iota_{V_{X} + V_{Y}}(v''))]\\
+                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
+                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+                \end{align*}
+                The case when $v_1 = g_{V};\iota_{2,V}(z_1)$ is symmetric.
+            \end{itemize}
+            \end{itemize}
+            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
+            Once again, $[v] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ means that there exists $v' \in s_{X+Y}(e_1)$ such that $v' \sim v$.
+            \begin{itemize}
+                \item \[
+                [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v''))]
+                \] such that $v'' \sim v$ and $[v'') \not = \varnothing$.
+                \begin{itemize}
+                    \item If $v' = v$ and $v'' = v$ then $v \in s_{X+Y}(e_1)$ and since $[v'') = [v) \not = \varnothing$ and $v \in s_{X+Y}(e_1)$, it must be the case that
+                    $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(v))$ and further
+                    \begin{align*}
+                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
+                                                                                            &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(v))]\\
+                                                                                            &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
                                                                                             &= [e_2]
                     \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
+                    \item If $v' = v $ and $v'' \sim v$ then $v \in s_{X+Y}(e_1)$.
+                          Suppose $v'' = f_{V};\iota_{1,V}(z'')$.
+                          Consider cases.
+                          \begin{itemize}
+                            \item $v = f_{V};\iota_{1,V}(z)$, then $f_{V}(z) = f_{V}(z'') \not = \varnothing$ and $[e_1) = [v) = [v'')$
+                            \item $v = g_{V};\iota_{2,V}(z)$, then $g_{V}(z) = \varnothing$ and $[e_1) = \varnothing$.
+                                  Since there is a path from $v$ to $e_1$ and $[v'') \not = \varnothing$ and $v'' \sim v$, by definition,
+                                  \begin{align*}
+                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
+                                  \end{align*}
+                          \end{itemize}
+                          Case when $v'' = g_{V};\iota_{2,V}(z'')$ is symmetric.
+                    \item If $v' \sim v$ and $v'' = v$ then $v' \in s_{X+Y}(e_1)$. This implies that $[v) \not = \varnothing$
+                          \begin{itemize}
+                            \item If $v = f_{V};\iota_{1,V}(z)$ and $v' = f_{V};\iota_{1,V}(z')$ then $[e_1) = [v') = [v) \not = \varnothing $ since $[f_{V}(z')) = [f_{V}(z))$.
+                                  That is,
+                                  \begin{align*}
+                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}})(e_1)]\\
+                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v)]\\
+                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v'')]\\
                                                                                                     &= [e_2]
                                   \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
+                            \item $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z')$ and then $[g_{V}(z')) = \varnothing$ and $[e_1) = \varnothing$.
+                                  By definition,
+                                  \begin{align*}
+                                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
+                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
                                                                                                           &= [e_2]
                                   \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
+                          \end{itemize}
+                          The case when $v = g_{V};\iota_{1,V}(z)$ is symmetric.
+                    \item If $v' \sim v$ and $v'' \sim v$ then $v' \in s_{X+Y}{e_1}$, $[v'') \not = \varnothing$ and $v'' \sim v'$.
+                    \begin{itemize}
+                        \item If $v'' = f_{V};\iota_{1,V}(z'')$ and $v' = f_{V};\iota_{2,V}(z')$ then $[e_1) = [v') = [v'')$ and
+                            \begin{align*}
+                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}})(e_1)]\\
+                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v')]\\
+                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v'')]\\
                                                                                                     &= [e_2]
                                   \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
+                        \item If $v'' = f_{V};\iota_{1,V}(z'')$ and $v' = g_{V};\iota_{2,V}(z')$ then $[g_{V}(z')) = \varnothing$ and $[e_1) = \varnothing$. 
+                        Then
+                        \begin{align*}
+                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
+                        \end{align*}
+                        The case when $v'' = g_{V};\iota_{2,V}(z'')$ is symmetric.
+                    \end{itemize}
+                \end{itemize}
+                \item Now suppose that 
+                \[
+                [e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
+                \]
+                such that $[v'') \not = \varnothing$, $[v) = \varnothing$ and there is a path from $[v]$ to $[v'']$, i.e. there is a path from $v$ to $v'$ and $v' \sim v''$.
+                In its turn $[v] \in s_{\sfrac{X + Y}{\sim}}([e_1])$ implies $v_1 \in s_{X+Y}(e_1)$ and $v \sim v_1$.
+                As per the comment below~\ref{def:child_respects_connectivity}, necessarily $v = v_1$ as $v$ should not have a pre-image in $V_{Z}$.
+                This implies that $v \in s_{X+Y}(e_1)$ and that $[e_1) = \varnothing$.
+                As there is a path from $v$ to $v'$, there is a path from $e_1$ to $v'$ and by definition, since $v'' \sim v'$ and $[v'') \not = \varnothing$,
+                \[
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
+                \]
+            \end{itemize}    
+\end{proof}
+
+\begin{lemma}
+\label{lemma:closures_of_equal_sets_are_equal}
+If
+\[
+\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
+\]
+where $\iota_{\sfrac{*}{\sim}}$ is either $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}$ depending on whether $[x_i]$ is in $V_{X} + V_{Y}$ or $E_{X} + E_{Y}$.
+Then
+\[
+\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))
+\]
+That is, closures of both sides are equal.
+\end{lemma}
+
+\begin{proof}
+    To check that the sets after closures are equal, we need to check if every element of the set on the left is also in the set on the right and vice versa.
+    Apart from $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))$ the set on the left also contains
+    \begin{itemize}
+        \item $[x_1]$ as per the reflexivity.
+        So we need to check if $[x_1] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
+              Because 
+              \[
+              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
+              \]
+              there exists some $[y]$ such that 
+              \[
+              [y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))
+              \]
+               and
+            \[
+                [y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
+            \]
+            that is, we have a sequence $([x_1],[y],[x_2])$ and by transitivity
+               \[
+               [x_1] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))
+               \]
+                and
+            \[
+            [x_2] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_1]))
+            \]
+        \item $[y]$ for all $[y]$ such that $[x_1] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([y]))$.
+              Let's check if $[y] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
+              As above, there exists $[z]$ such that
+              \[
+              [z] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))
+              \]
+               and
+              \[
+                [z] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
+              \]
+              By a sequence $([y],[x_1],[z],[x_2])$ we get that
+              \[
+                [y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
+              \]
+              Similarly for $[y]$ such that $[x_2] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([y]))$.
+        \item For any sequence $([x_1],\ldots,[x_n])$ such that $[x_i] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_{i+1}]))$ or $[x_{i+1}] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))$ for $i < n$ the set on the left contains $[x_n]$.
+              Let's check if $[x_n] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
+              Since there exists $[y]$ such that $[y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))$ and $[y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))$ we can extend the sequence to $([x_2],[y],[x_1],\ldots,[x_n])$ and hence $[x_n] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
+              Similarly, given a sequence $([x_2],\ldots, [x_n])$ and extending it to $([x_1],[y],[x_2],\ldots,[x_n])$ we get $[x_n] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_1]))$.
     \end{itemize}
@@
 \begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
+    Suppose $[v_1] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ and $[[v_1]) = [[e_1]) \not = \varnothing$, then it must be the case that 
+    \[
+    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
+    \]
+    By the lemma above, to prove this, it is sufficient to show that
+
+    \[
+    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
+    \]
+
+    $[v_1] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ means there exists $v_1' \in s_{X+Y}(e_1)$ such that $v_1 \sim v_1'$.
     We then have two cases.
     \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
+      \item $v_1 = v_1'$ and then $v_1 \in s_{X+Y}(e_1)$.
+      \begin{itemize}
+         \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1''))]$ such that $v_1'' \sim v_1$ and $[v_1'') \not = \varnothing$.
+            \begin{itemize}
+                \item $v_1'' = v_1$ and then $[v_1) \not = \varnothing$ and $[e_1) \not = \varnothing$.
+                      Furthermore, necessarily $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)) = \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1))$.
                       By applying $([]_{V},[]_{E})$ we get
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
+                      \ifdefined \ONECOLUMN
+                      \begin{align*}
+                        \text{by functionality}\\
+                          [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1))) &=
+                          [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))\\
+                          \text{by unfolding the definition}\\
+                          \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
+                          \end{align*} 
+                      \else
+                      \begin{align*}
+                      \text{by functionality}\\
+                        [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1))) &=\\
+                        [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1))) &\\
+                        \text{by unfolding the definition}\\
+                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
+                        \end{align*} 
+                      \fi
+                \item $v_1'' \sim v_1$. If $v_1'' = f_{V};\iota_{1,V}(z_1'')$ and $v_1 = f_{V};\iota_{1,V}(z_1)$ we get
+                    \begin{align*}
+                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')))\\
+                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z_1''))))\\
+                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z_1))))\\
+                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
+                                                                                                                &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
+                    \end{align*}
+                    If $v_1'' = f_{V};\iota_{1,V}(z_1'')$ and $v_1 = g_{V};\iota_{1,V}(z_1)$ then $[v_1) = [e_1) = \varnothing$ and 
+                    \[
+                        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)) = \varnothing
+                    \]
+                    \begin{align*}
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\
+                                                                                                            &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
+                    \end{align*}
+                \end{itemize}
+                \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ such that $[v_2) \not = \varnothing$ and there is a path from $[v_1]$ to $[v_2]$.
+                By definition, this implies $[v_1) = \varnothing$ and $[e_1) = \varnothing$.
+                Again, by definition, we have
+                \begin{align*}
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
+                                                                                                    &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
+                \end{align*}
+                \end{itemize}
+                \item $v_1 \sim v_1'$ and then $v_1' \in s_{X+Y}(e_1)$.
+                \begin{itemize}
+                    \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1''))]$ such that $v_1 \sim v_1''$ and $[v_1'') \not = \varnothing$.
+                    \begin{itemize}
+                        \item $v_1 = v_1''$ and then $[v_1) \not = \varnothing$.
+                              Suppose $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_1' = f_{V};\iota_{1,V}(z_1')$. Then $[v_1') = [e_1) \not = \varnothing$ and 
+                            \begin{align*}
+                              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')))\\
+                                                                                                                     &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
+                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))\\
+                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1])
+                            \end{align*}
+                            Suppose $v_1' = g_{V};\iota_{2,V}(z_1')$ and then $[v_1') = [e_1) = \varnothing$.
+                            By definition,
+                            \[
+                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
+                            \]
+                            and since there is a path from $e_1$ to $v_1'$ and $v_1' \sim v_1$
+                            \[
+                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1]) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
+                            \]
+                            The cases when $v_1 = g_{V};\iota_{2,V}(z_1)$ is symmetric.
+                        \item $v_1 \sim v_1''$ and then $v_1' \sim v_1''$. Suppose $v_1'' = f_{V};\iota_{1,V}(z_1'')$.
+                              If $v_1'$ is in the image of $f_{V};\iota_{1,V}$, then
+                              \begin{align*}
+                                \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')) &= \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1''))\\ 
+                                                                               &= \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(e_1'))
+                              \end{align*}
+                              and
+                              \begin{align*}
+                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1]) &=
+                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1'])\\ &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1])
+                              \end{align*}
+                              If $v_1'$ is in the image of $g_{V};\iota_{2,V}$, then $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')) = \consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)) = \varnothing$ and by definition as there is a path from $[e_1]$ to $[v_1'']$
+                              \begin{align*}
+                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\
+                                                                                                                        &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
+                              \end{align*}
+                              The cases when $v_1'' = g_{V};\iota_{2,V}(z_1'')$ are symmetric.
+                    \end{itemize}
+                \end{itemize}
      \end{itemize}
@@
 \label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
+    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that are not top-level.
 \end{lemma}
 \begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
+  $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$ if and only if $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{*}{\sim}}([x]))$ is defined.
+  Let's first show from left to right.
+$\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$ implies $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$.
+  \begin{itemize}
+    \item $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) = [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\iota_{*}{\sim}(x'))$ such that $x \sim x'$.
+          By e-hypergraph-ness of $X+Y$ this implies that $[x') \not = \varnothing$ and hence $<_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x'])) = <_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) = [<_{X+Y}^{\mu}(\iota_{*}(x))]$
+    \item $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ such that there is a path from $[x]$ to $[v]$ and
+          $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$. This implies that $[v) \not = \varnothing$ and by definition
+          \[
+            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{*}{\sim}}([x])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{*}{\sim}}([v]))  
+          \]
+  \end{itemize}
+  The right-to-left direction is analogous.
 \end{proof}
@@
 \begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
+\label{prop:pushout_is_e_hypergraph}
 For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
@@
 
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
+We will then discuss the uniqueness of the corresponding pushout complement.
+
+% \begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
+% Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
+% They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
+% \end{definition}
+% Intuitively, the no-dangling condition guarantees that sources and targets of an edge are not deleted if the edge is not deleted.
+% The no-identification condition requires that if $m$ identifies two vertices, it must not be the case that one vertex should be removed and the other vertex should be preserved according to $\mathcal{K}$.
 
@@
 %         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
+%             \tikzfig{../figures/combinatorial_semantics/structural_rule_pushout}
 %         }
@@
 
+\begin{proposition}
+\label{prop:boundary_complement}
+    The boundary complement in~\ref{def:boundary_new} when exists is unique.
+\end{proposition}
+\begin{proof}
+        The pushout in~\ref{def:boundary_new} satisfies the assumptions in Definition~\ref{pushout:assumptions} (note that since $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j$ is an MDA cospan the image of $i + j$ in $\mathcal{L}$ consists of top-level vertices only) and therefore is isomorphic to the pushout constructed in~\ref{pushout:assumptions} and by construction should also be a pushout on the underlying sets of vertices and hyperedges.
+    % I think the above sufficies, but it can also be shown formally that every pushout in $\catname{EHyp_{\Sigma}}$ is a pushout on the underlying sets of edges and vertices by constructing an e-hyperghraph from the set of vertices and same for the set of edges and checking if the morphisms are well-defined.
+    %     Consider diagrams below,
+    %     \[
+    %         \begin{tikzcd}
+    %             & \mathcal{L} \arrow[d] \arrow[ldd, bend right] & i+j \arrow[l, "h"'] \arrow[d, "c"]           &                 & V_{\mathcal{L}} \arrow[d] \arrow[ldd, bend right] & i+j \arrow[l, "h_{V}"] \arrow[d, "c_{V}"]      \\
+    %             & \mathcal{G} \arrow[ld, "u"']                  & \mathcal{C} \arrow[l] \arrow[lld, bend left] &                 & V_{\mathcal{G}} \arrow[ld, "u_{V}"]               & V_\mathcal{C} \arrow[l] \arrow[lld, bend left] \\
+    % \mathcal{Q} &                                               &                                              & V_{\mathcal{Q}} &                                                   &
+    % \end{tikzcd}  
+    %     \]
+    %     If $\mathcal{L} \xrightarrow{} \mathcal{G} \xleftarrow{} C$ is a pushout in $\catname{EHyp}_{\Sigma}$ then $V_{\mathcal{L}} \xrightarrow{} V_{\mathcal{G}} \xleftarrow{} V_{\mathcal{C}}$ is a pushout in $\catname{Set}$.
+    %     The small square on the right commutes because the square on the left commutes, so we only need to check the universal property.
+    %     Suppose there exist $i+j \to V_{\mathcal{L}} \to V_{Q} = i+j \to V_{\mathcal{C}} \to V_{Q}$ and we need to check that there exists a unique $u_{V}$ such that these morphisms factors out through $u_{V}$.
+    %     % Hence, pushout complement should yield pushout complements for the corresponding sets.
+    %     Construct an e-hypergraph $\mathcal{Q} = (V,E,s,t,<,\consistency)$ such that
+    %     \begin{itemize}
+    %         \item $V = V_{Q}$
+    %         \item $E = \sfrac{E_{\mathcal{L}} + E_{\mathcal{C}}}{\mathcal{R}}$ where $e_1 \mathcal{R} e_2$ if there exists $v \in V_{Q}$ such that $v = i_{V}(v_1)$ and $v = i_{V}(v_2)$ and $e_1 = <^{\mu}_{\mathcal{L}}(\iota_{V_{\mathcal{L}}}(v_1)), e_2 = <^{\mu}_{\mathcal{C}}(\iota_{V_{\mathcal{C}}}(v_2))$
+    %     \end{itemize}
+        In particular,
+    \[
+        \tikzfig{../figures/appendix/pushout_verticies}
+    \]
+    \[
+        \tikzfig{../figures/appendix/pushout_edges}
+    \]
+    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
+    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
+    Then, since $m$ and $[c_1,c_2]$ are monos, the first square can be rewritten as follows.
+% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
+\[\begin{tikzcd}
+	{a + x + y} && {i + j + x} && {i+j} \\
+	\\
+	{a + x + y + z} &&&& {\begin{matrix} i + j + w \\ \cong \\ i + j + z \end{matrix}}
+	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
+	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
+	\arrow["{m_{V}}"', from=1-1, to=3-1]
+	\arrow["{g_{V}}", from=3-5, to=3-1]
+	\arrow["c"', from=1-5, to=3-5]
+\end{tikzcd}\]
+    Because $a + x + y + w$ also yields a pushout it must be the case that $a+x+y+w \cong a + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
+    One can further observe that $g_{V} = h_{V,ext};h_{V,int} + id_{n}$ and that $h_{V,ext}$ is mono and $h_{V,int} = [h_1,h_2]$ where $h_1,h_2$ are mono.
+    The above shows that edges and vertices in the pushout complement are uniquely determined.
+    To show that pushout complement is unique we next need to show that source and target maps are unique.
+    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{\mathcal{L^{\bot}}}(e)) = s_{\mathcal{G}}(g_{E}(e))$ and $s_{\mathcal{L^{\bot}}}(e) \subseteq (g_{V}^{-1})^{*}(s_{\mathcal{G}}(g_{E}(e)))$.
+    Since $g_{V} = h_{V,ext};[h_1,h_2] + id_{n}$ the inverse image of $g_{V}$ contains at most two elements: one with the pre-image in $i$ and another with the pre-image in $j$.
+    If $g_{V}$ identified more than two vertices it would mean that $[h_1,h_2]$ identified more than two vertices which would mean that two of them come both either from $i$ or $j$ which would imply that either $h_1$ or $h_2$ is not mono.
+    If the inverse image contains exactly one vertex then the $s_{\mathcal{L^{\bot}}}(e)$ is uniquely fixed, otherwise suppose it contains $\{v_1,v_2\}$ such that $v_1 \in i$ and $v_2 \in j$.
+    However, $v_1$ can not be in $s_{\mathcal{L^{\bot}}}(e)$ as per the (6) and (7) conditions of the boundary complement~\ref{def:boundary_new} the image of $j$ in $\mathcal{L}^{\bot}$ should consist of vertices of in-degree 0.
+
+    Similarly for $t_{\mathcal{L}^{\bot}}$.
+    % Suppose that there exist
+    % \[
+    %     \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
+    % \qquad
+    % \text{and}
+    % \qquad
+    %     \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
+    % \]
+    % Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
+    % Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
+    % Similarly for targets.
+    We also need to show that $<$ and $\consistency$ are unique.
+    Suppose that there exist
+    \[
+        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
+    \qquad
+    \text{and}
+    \qquad
+        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
+    \]
+    Because $g$ is homomorphism, it must be the case that 
+    \[
+        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
+    \]
+    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ is mono.
+    Similarly for edges.
+    We now show that if $[v) = \varnothing$ for $v \in V_{\mathcal{L}^{\bot}}$ then necessarily $[g_{V}(v)) = \varnothing$.
+    Suppose $[v) = \varnothing$ and $<_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = e$ and there is a connected component that contains $v$ and such that for all $v'$ and $e'$ in this component $[v') = \varnothing = [e')$.
+    Consider $\mathcal{G'}$ which is obtained from $\mathcal{G}$ by making $[g_{V}(v')) = [g_{E}(e')) = \varnothing$ for $v', e'$ above.
+    Clearly there is a morphism from $\mathcal{L}^{\bot}$ to $\mathcal{G}'$, i.e.
+the morphism $g' = (g_{V},g_{E})$ that has the same action on edges and vertices as the morphism $g$ does, because $v'$ such that $[v') \not = \varnothing$ can be mapped to $g_{V}(v')$ such that $[g_{V}(v')) \not = \varnothing$.
+    The only case when such a morphism may not exist is when $[v'') \not = \varnothing$, $[v') = \varnothing$ and $g_{V}(v'') = g_{V}(v')$.
+    This implies that both $v''$ and $v'$ have a pre-image in $i + j$ and according to the definition of a boundary complement it must be $[[c_1,c_2](z_1)) = [[c_1,c_2](z_2))$ for all $z_1, z_2 \in V_{i + j}$.
+    For the same reason there is also a morphism from $\mathcal{L}$ to $\mathcal{G}'$.
+    The only case when such a morphism may not exist is when $v'' \in V_{\mathcal{L}}$ such that $[v'') \not = \varnothing$ is mapped to $m_{V}(v'') = g_{V}(v')$ such that $[v') = \varnothing$ which implies $v'$ and $v''$ share a pre-image in $i + j$.
+    This leads to a contradiction as the image of $i + j$ in $\mathcal{L}$ should consist of top-level vertices exclusively.
+    This construction results in a commutative diagram below
+    \[
+        \begin{tikzcd}
+            & \mathcal{L} \arrow[d, "m"'] \arrow[ldd, "m'"', bend right=49] & i'+j' \arrow[l] & i+j \arrow[d] \arrow[l]                                           \\
+            & \mathcal{G}                                                   &                 & \mathcal{L}^{\bot} \arrow[ll, "g"'] \arrow[llld, "g'", bend left] \\
+\mathcal{G}' &                                                               &                 &                                                                  
+\end{tikzcd}
+    \]
+    that further results in a contradiction since there is no morphism from $\mathcal{G}$ to $\mathcal{G'}$ because for any such morphism $u$ it must be the case that $[u_{V}(g_{V}(v'))) \not = \varnothing = [g'_{V}(v')) = \varnothing$ which would mean that $\mathcal{G}$ is not a pushout.
+    This means that either both $<_{1}(\iota_{*}(x))$ and $<_{2}(\iota_{*}(x))$ are undefined for a given $x$ or $<_{1}(\iota_{*}(x)) = <_{2}(\iota_{*}(x))$ which concludes that the relations are equal and $\mathcal{L}_{1}^{\bot} = \mathcal{L}_{2}^{\bot}$.
+
+    Finally, let's suppose there exist
+    \[
+        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
+    \quad
+    \text{and}
+    \quad
+        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
+    \]
+    Because $g$ is a homomorphism, we have
+    \[
+        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
+    \]
+    \[
+        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
+    \]
+    \begin{itemize}
+        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
+        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
+              $\consistency_{1} \not = \consistency_{2}$ implies there exist $v, v' \in \mathcal{L}^{\bot}$ such that $v \consistency_{1} v'$ and $v \not \consistency_{2} v'$ for $[v) = [v') \not = \varnothing$.
+              Then there are connected components $C_1, C_2$ in $\mathcal{L}_{2}^{\bot}$ such that for all $x \in C_1$ and $y \in C_2$ $x \not \consistency_{2} y$ and $v \in C_1$, $v' \in C_2$.
+              Consider $g_2 = (g_{V},g_{E}) : \mathcal{L}_2^{\bot} \to \mathcal{G}$ and construct $\mathcal{G'}$ that is obtained from $\mathcal{G}$ by making $g(x) \not \consistency g(y)$ for $x \in C_1$ and $y \in C_2$.
+              There is a morphism $g_2' = g_2 : \mathcal{L}_2^{\bot} \to \mathcal{G}'$.
+Like above, the only case when such morphism may not exist is when for $v' \not \consistency v$ $g_{V}(v) = g_{V}(v')$.
+              This implies $v, v'$ have a pre-image in $i + j$ and by definition it must be $v' \consistency v$.
+              Similarly, there is a morphism $m' : \mathcal{L} \to \mathcal{G}'$.
+The only case when it may not exist is when for $v_1, v_1' \in \mathcal{L}$ such that $v_1 \consistency v_1'$, $m_{V}(v_1) = g_{V}(v_2)$ and $m_{V}(v_1') = g_{V}(v_2')$ where $v_2 \in C_1$ and $v_2' \in C_2$.
+              It implies $v_1, v_1'$ have pre-image in $i + j$ and by definition of a boundary-complement the image of $i + j$ should consist of top-level vertices exclusively.
+              This results in the diagram below there is no morphism $u : G \to G'$, as the existence of such a morphism implies $u(g(x)) = u(g'(x))$, $u(g(y)) = u(g'(y))$ for $x \in C_1$ and $y \in C_2$ and $u(g(x)) \consistency u(g(y))$ and $u(g'(x)) \not \consistency u(g'(y))$.
+              This contradicts the fact that $\mathcal{G}$ is a pushout.
+              Hence, $\consistency_{1} = \consistency_{2}$ and $\mathcal{L}_{2}^{\bot} = \mathcal{L}_{1}^{\bot}$.
+    \end{itemize}
+    Ultimately, we have shown that $\mathcal{L}^{\bot}$ is unique up to isomorphism.
+\end{proof}
+
+Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
+
+% \begin{remark}
+%     Consider morphisms $i + j \xrightarrow{[f_{ext},g_{ext}];[f_{int},g_{int}]} \mathcal{L} \xrightarrow{m} \mathcal{G}$.
+%     If both of these morphisms are mono, then the boundary complement exists. 
+%     Intuitively, it exists for the same reason the corresponding boundary complement for plain hypergraphs exists: it is constructed by removing from $\mathcal{G}$ everything that has no pre-image in $i + j$.
+% \end{remark}
+
+% \begin{figure*}[t!]
+%     \begin{minipage}{0.1\textwidth}
+%         $\;$
+%     \end{minipage}
+%     \hfill
+% \begin{minipage}{0.7\textwidth}
+%     \[
+%         \scalebox{0.45}{
+%             \tikzfig{../figures/combinatorial_semantics/structural_rule_pushout}
+%         }
+%     \]
+%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
+%     \label{fig:structural_rule_pushout}
+% \end{minipage}
+% \hfill
+% \begin{minipage}{0.1\textwidth}
+%         $\;$
+% \end{minipage}
+% \end{figure*}
+
 \section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
@@
 
+\begin{figure}
+    \[
+    \scalebox{0.75}{
+        \tikzfig{../figures/appendix/contexts}
+    }
+    \]
+    \caption{String diagrammatic contexts}
+    \label{fig:string_contexts}
+\end{figure}
+
+\begin{remark}
+    The construction in Figure~\ref{fig:A+B} naturally extends to $k$ operands.
+    Given a join of $k$ cospans of e-hypegraphs
+    \begin{align*}
+        n_1 \xrightarrow{} n_1' \xrightarrow{} &\mathcal{F}_1 \xleftarrow{} m_1' \xleftarrow{} m_1\\
+        &\;+\\
+        &\vdotswithin{+}\\
+        &\;+\\
+        n_k \xrightarrow{} n_k' \xrightarrow{} &\mathcal{F}_k \xleftarrow{} m_k' \xleftarrow{} m_k
+    \end{align*}
+    we will denote the carrier of the resulting cospan as
+    \[
+    \mathcal{F}_1 \; \hat{+}\; \ldots\; \hat{+}\; \mathcal{F}_{k}
+    \]
+    to distinguish it from the coproduct.
+\end{remark}
+
 \begin{proposition}[Proposition~\ref{prop:soundness} restatement]
+    \label{proof:appendix:soundness}
 The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
@@
 \begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
+    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $f = g$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket f\rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket f \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans$.
+    $f = g$ implies that either
+    \begin{itemize}
+        \item $f = \mathcal{C}[l] = g$ for some morphism $l \in \textbf{PROP}^{+}(\Sigma)$.
+        \item $f = \mathcal{C}[l]$ and $g = \mathcal{C}[r]$ where $l = r$ or $r = l$ is from $\mathcal{S}$.
+        \item There is a sequence $w = (x_1, \ldots, x_n)$ such that $x_1 = f$ and $x_n = g$ and $x_i = x_{i+1}$ or $x_{i+1} = x_{i}$ in the sense above for $i < n$.
+    \end{itemize}
+    where $\mathcal{C}$ is a context defined as per the Figure~\ref{fig:string_contexts} and $\mathcal{C}[l]$ is defined by replacing the occurrence of
+    \scalebox{0.5}{
+    \begin{tikzpicture}[tikzfig]
+        \begin{pgfonlayer}{nodelayer}
+            \node [style=empty diag] (0) at (-0.5, 0.5) {};
+        \end{pgfonlayer}
+    \end{tikzpicture}
     }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
+    in $\mathcal{C}$ with a string diagram $l$.
+    Let's prove the statement by induction on $\mathcal{C}$ when $|w| = 2$.
+    \begin{itemize}
+        \item If $\mathcal{C}$ is empty, then either $f = g$ syntactically and $\llbracket f \rrbracket = \llbracket g \rrbracket$ or $f = l = r = g$ for some $l = r$ or $r = l$ in $\mathcal{S}$.
+              The latter trivially yields the E-DPO diagram below
+              \[
+                \begin{tikzcd}
+                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                    \\
+                    {\mathcal{L}} && {i+j} && {\mathcal{R}} \\
+                    & {i'+j'} && {i''+j''} \\
+                    && {i+j}
+                    \arrow[from=1-1, to=3-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-3, to=1-2]
+                    \arrow[from=1-3, to=1-4]
+                    \arrow[from=1-3, to=3-3]
+                    \arrow[from=1-4, to=1-5]
+                    \arrow[from=1-5, to=3-5]
+                    \arrow["\lrcorner"{pos=0.05, rotate=90, description}, phantom, from=3-1, to=1-2]
+                    \arrow[from=3-3, to=3-1]
+                    \arrow[from=3-3, to=3-5]
+                    \arrow["\lrcorner"{pos=0.05, rotate=180, description}, phantom, from=3-5, to=1-4]
+                    \arrow[from=4-2, to=3-1]
+                    \arrow[from=4-4, to=3-5]
+                    \arrow[from=5-3, to=3-3]
+                    \arrow[from=5-3, to=4-2]
+                    \arrow[from=5-3, to=4-4]
+                \end{tikzcd}
+              \]
+              where $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j = \llbracket l \rrbracket = \llbracket f \rrbracket$ and $i \to i'' \to \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j = \llbracket r \rrbracket = \llbracket g \rrbracket$ and $i + j \to i + j \to \mathcal{L}$ is a boundary complement as per~\ref{def:boundary_new}.
+              Hence, $\llbracket f \rrbracket \Rrightarrow{}_{\langle \mathcal{L}, \mathcal{R} \rangle} \llbracket g \rrbracket$.
+              \item Suppose 
+              \[
+                \scalebox{0.75}{\tikzfig{../figures/appendix/contexts_2}}
+              \]
+              and by inductive hypothesis $\llbracket \mathcal{C}[l] \rrbracket \Rrightarrow{} \llbracket \mathcal{C}[r] \rrbracket$ which results in the top half of the diagram below.
+              \[\begin{tikzcd}
+                {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                \\
+                {\llbracket \mathcal{C}[l]\rrbracket} && {\mathcal{L}_1^{\bot}} && {\llbracket \mathcal{C}[r]\rrbracket} \\
+                & {n_1'+m_1'} & {n_1+m_1} & {n_1''+m_1''} \\
+                {\llbracket f \rrbracket} && {\mathcal{L}_2^{\bot}} && {\llbracket g \rrbracket} \\
+                & {n_2' + m_2'} && {n_2''+m_2''} \\
+                && {n_2 + m_2}
+                \arrow[from=1-1, to=3-1]
+                \arrow[from=1-2, to=1-1]
+                \arrow[from=1-3, to=1-2]
+                \arrow[from=1-3, to=1-4]
+                \arrow[from=1-3, to=3-3]
+                \arrow[from=1-4, to=1-5]
+                \arrow[from=1-5, to=3-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=90, description}, draw=none, from=3-1, to=1-2]
+                \arrow[from=3-1, to=5-1]
+                \arrow[from=3-3, to=3-1]
+                \arrow[from=3-3, to=3-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=180, description}, draw=none, from=3-5, to=1-4]
+                \arrow[from=3-5, to=5-5]
+                \arrow[from=4-2, to=3-1]
+                \arrow[from=4-3, to=3-3]
+                \arrow[from=4-3, to=4-2]
+                \arrow[from=4-3, to=4-4]
+                \arrow[from=4-3, to=5-3]
+                \arrow[from=4-4, to=3-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=90, description}, draw=none, from=5-1, to=4-2]
+                \arrow[from=5-3, to=5-1]
+                \arrow[from=5-3, to=5-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=180, description}, draw=none, from=5-5, to=4-4]
+                \arrow[from=6-2, to=5-1]
+                \arrow[from=6-4, to=5-5]
+                \arrow[from=7-3, to=5-3]
+                \arrow[from=7-3, to=6-2]
+                \arrow[from=7-3, to=6-4]
+            \end{tikzcd}\]
+            Clearly the occurrence of $\llbracket \mathcal{C}[l] \rrbracket$ in $\llbracket f \rrbracket$ is convex and down-closed and by following the argument in Theorem 35~\cite{bonchi_string_2022-2}, there exists $n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2$ such that
+            \ifdefined \ONECOLUMN
+            \begin{align*}
+                \llbracket f \rrbracket &= \;
+                (0 \to 0 \to \llbracket C[l] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)
+                ;
+                (n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)\\
+                \llbracket g \rrbracket &= \;
+                (0 \to 0 \to \llbracket C[r] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)
+                ;
+                (n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)
+                \end{align*}
+            \else
+            \begin{align*}
+            \llbracket f \rrbracket = \;
+            &(0 \to 0 \to \llbracket C[l] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)\\
+            &;\\
+            &(n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)\\
+            \llbracket g \rrbracket = \;
+            &(0 \to 0 \to \llbracket C[r] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)\\
+            &;\\
+            &(n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)
+            \end{align*}
+            \fi
+            and
+            % $n'_2 + m'_2 = n_3 + m_3 + (n_1' + m_1' \setminus (n_1 + m_1))$
+            \[
+                n_2' = n_3 + (n_1' \setminus n_1) \qquad m_2' = m_3 + (m_1' \setminus m_1)
+            \]
+            and
+            \[
+            n_2 + m_1 \xrightarrow{} n_2' \setminus (n_1' \setminus n_1) + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot} \xleftarrow{} m'_2 \setminus (m_1' \setminus m_1) + n_1 \xleftarrow{} m_2 + n_1    
+            \] is an MDA e-hypergraph.
+            By denoting the pushout of $\mathcal{L}_1^{\bot} \xleftarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot}$ with $\mathcal{L}_3^{\bot}$,
+            one can observe that $\llbracket f \rrbracket$ is the pushout object of $\mathcal{L} \xleftarrow{} i+j \xrightarrow{} \mathcal{L}$ as per the diagram below
+            \[\begin{tikzcd}
+                && {n_1+m_1} && {\mathcal{L}^{\bot}_2} \\
+                \\
+                {i+j} && {\mathcal{L}^{\bot}_1} && {\mathcal{L}^{\bot}_3} \\
+                \\
+                {\mathcal{L}} && {\mathcal{C}[l]} && {\llbracket f \rrbracket}
+                \arrow[from=1-3, to=1-5]
+                \arrow[from=1-3, to=3-3]
+                \arrow[from=1-5, to=3-5]
+                \arrow[from=3-1, to=3-3]
+                \arrow[from=3-1, to=5-1]
+                \arrow[from=3-3, to=3-5]
+                \arrow[from=3-3, to=5-3]
+                \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-3]
+                \arrow[from=3-5, to=5-5]
+                \arrow[from=5-1, to=5-3]
+                \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=5-3, to=3-1]
+                \arrow[from=5-3, to=5-5]
+                \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=5-5, to=3-3]
+            \end{tikzcd}
+            \]
+            The corestriction of $n_1 + m_1 \xrightarrow{[f_1,f_2]} \mathcal{L}_1^{\bot}$ to the image of $i + j \xrightarrow{[g_1,g_2]} \mathcal{L}_{1}^{\bot}$ is mono: suppose it is not mono, then there exist $z_1$, $z_2$, $z_3$ such that $f_1(z_1) = f_2(z_2) = g_1(z_3)$, then $[f_2,g_1]$ would not be mono, similarly if $f_1(z_1) = f_2(z_2) = g_2(z_3)$.\
+            By construction of the pushout the restriction of $\mathcal{L}_{1}^{\bot} \to \mathcal{L}_{3}^{\bot}$ is mono and hence the arrow $i + j \to \mathcal{L}_{3}^{\bot}$ is mono.
+            This implies that $\llbracket f \rrbracket \Rrightarrow_{\langle \mathcal{L},\mathcal{R} \rangle} \llbracket g \rrbracket$.
+            \item Consider the case when
+            \[
+            \scalebox{0.75}{\tikzfig{../figures/appendix/contexts_3}}    
+            \]
+            By hypothesis, there exists a diagram
+                \[
+                \begin{tikzcd}
+                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                    \\
+                    {\llbracket\mathcal{C}[l]\rrbracket} && {\mathcal{L}_1^{\bot}} && {\llbracket\mathcal{C}[r]\rrbracket} \\
+                    & {n'+m'} && {n''+m''} \\
+                    && {n+m}
+                    \arrow[from=1-1, to=3-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-3, to=1-2]
+                    \arrow[from=1-3, to=1-4]
+                    \arrow[from=1-3, to=3-3]
+                    \arrow[from=1-4, to=1-5]
+                    \arrow[from=1-5, to=3-5]
+                    \arrow[from=3-3, to=3-1]
+                    \arrow[from=3-3, to=3-5]
+                    \arrow[from=4-2, to=3-1]
+                    \arrow[from=4-4, to=3-5]
+                    \arrow[from=5-3, to=3-3]
+                    \arrow[from=5-3, to=4-2]
+                    \arrow[from=5-3, to=4-4]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
+                \end{tikzcd}
+                \]
+                which implies the existence of the following diagram where both squares are pushouts by construction and the matching is convex and down-closed.
+                \ifdefined \ONECOLUMN
+                \[\adjustbox{width=0.95\linewidth}{
+                    \begin{tikzcd}
+                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                    \\
+                    {\llbracket f_1 \rrbracket \; \hat{+}\; \ldots \hat{+} \; \llbracket\mathcal{C}[l]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \mathcal{L}_{1}^{\bot} \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket\mathcal{C}[r]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} \\
+                    & {n_1' + m_1'} && {n_1'' + m_1''} \\
+                    && {n + m}
+                    \arrow[from=1-1, to=3-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-3, to=1-2]
+                    \arrow[from=1-3, to=1-4]
+                    \arrow[from=1-3, to=3-3]
+                    \arrow[from=1-4, to=1-5]
+                    \arrow[from=1-5, to=3-5]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
+                    \arrow[from=3-3, to=3-1]
+                    \arrow[from=3-3, to=3-5]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
+                    \arrow[from=4-2, to=3-1]
+                    \arrow[from=4-4, to=3-5]
+                    \arrow[from=5-3, to=3-3]
+                    \arrow[from=5-3, to=4-2]
+                    \arrow[from=5-3, to=4-4]
+                \end{tikzcd}
+                }
+                \]
+                \else
+                \begin{figure*}[hbt!]
+                \[
+                    \begin{tikzcd}
+                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                    \\
+                    {\llbracket f_1 \rrbracket \; \hat{+}\; \ldots \hat{+} \; \llbracket\mathcal{C}[l]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \mathcal{L}_{1}^{\bot} \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket\mathcal{C}[r]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} \\
+                    & {n_1' + m_1'} && {n_1'' + m_1''} \\
+                    && {n + m}
+                    \arrow[from=1-1, to=3-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-3, to=1-2]
+                    \arrow[from=1-3, to=1-4]
+                    \arrow[from=1-3, to=3-3]
+                    \arrow[from=1-4, to=1-5]
+                    \arrow[from=1-5, to=3-5]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
+                    \arrow[from=3-3, to=3-1]
+                    \arrow[from=3-3, to=3-5]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
+                    \arrow[from=4-2, to=3-1]
+                    \arrow[from=4-4, to=3-5]
+                    \arrow[from=5-3, to=3-3]
+                    \arrow[from=5-3, to=4-2]
+                    \arrow[from=5-3, to=4-4]
+                \end{tikzcd}
+                \]
+                \caption{$\llbracket f \rrbracket \Rrightarrow{}_{\langle \mathcal{L},\mathcal{R} \rangle} \llbracket g \rrbracket$}
+                \label{fig:f_rewrites_to_g_under_plus}
+            \end{figure*}
+            \fi
+                and hence we have $\llbracket f \rrbracket \Rrightarrow_{\langle \mathcal{L}, \mathcal{R} \rangle} \llbracket g \rrbracket$
+    \end{itemize}
+    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans(\Sigma)/\mathcal{S}$.
+    % there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
+    % As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
+    % We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
+    % The mapping of external interfaces into $\llbracket l \rrbracket$ is mono and therefore the corresponding pushout complement exists.
+    % We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
+
+    % Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
+    % And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
+    % \begin{figure}[h!]
+    % \[
+    %     \scalebox{0.8}{
+    %     \tikzfig{../figures/combinatorial_semantics/DPOI_square}
+    %     }
+    % \]
+    % \end{figure}
+    % We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
+    % This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
+    % The no-dangling condition is satisfied because we consider only mda-cospans.
+    % Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
+    % Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
+    % The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
+    % Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
+    % The rest follows by induction on the number of applied equations.
+    % Similar argument then applies to other rules in $\mathcal{S}$ as well.
 \end{proof}
@@
 
-\end{document}
+\section{Additional figures for Section~\ref{sec:introduction}:Introduction}
+
+This section contains the figures with more elaborate transformation of e-string diagrams from the Figure~\ref{fig:e-graph-example}.
+
+\ifdefined\ONECOLUMN
+\begin{figure}[htb!]
+    \vspace{-3cm}
+    \centering
+    \[
+        \hspace{1.3cm}
+        \resizebox{0.8\textwidth}{!}{
+        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-b-c}
+        }
+    \]
+    \caption{Example translation from $(b)$ to $(c)$.}
+    \label{fig:e-graph-example-b-c}
+\end{figure}
+\else
+\begin{figure*}[htb!]
+    \vspace{-3cm}
+    \centering
+    \[
+        \hspace{1.3cm}
+        \resizebox{0.8\textwidth}{!}{
+        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-b-c}
+        }
+    \]
+    \caption{Example translation from $(b)$ to $(c)$.}
+    \label{fig:e-graph-example-b-c}
+\end{figure*}
+\fi
+
+\ifdefined\ONECOLUMN
+\begin{figure}[htb!]
+    \[
+        % \hspace{1.3cm}
+        \scalebox{0.4}{
+        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
+        }
+    \]
+    \caption{Example translation from $(a)$ to $(b)$.}
+    \label{fig:e-graph-example-a-b}
+\end{figure}
+\else
+\begin{figure*}[htb!]
+    \[
+        % \hspace{1.3cm}
+        \scalebox{0.5}{
+        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
+        }
+    \]
+    \caption{Example translation from $(a)$ to $(b)$.}
+    \label{fig:e-graph-example-a-b}
+\end{figure*}
+\fi
+
+\section{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
+\begin{example}
+    Consider cospans from Figure~\ref{fig:appendix:non-isomorphic-cospans}(a) where the $f_{int}, f_{ext}$ (respectively, $g_{int},g_{ext}$) maps are shown with red arrows.
+    The left cospan represents a symmetry placed inside a hierarchical edge while the right one depicts a tensor product of two identities put inside the hierarchical edge.
+    The colours of the vertices show the induced partition of the interfaces.
+    We will try to build an isomorphism step-by-step.
+    To make the diagram commute, clearly, it must be the case that $u_0 \mapsto u_0$ and $u_1 \mapsto u_1$.
+    Then, because the isomorphism should preserve the order within the subset of green verices, it must be the case that $u_2 \mapsto u_2$ and $u_3 \mapsto u_3$.
+    Similarly for the output interfaces.
+    To further make the diagram commute, we need to map $v_2 \mapsto v_3$ and $v_3 \mapsto v_2$.
+    This, however, contradicts the requirement that $g_{int};\gamma(w_2) = g_{int}(w_2)$.
+    Hence, the two cospans are not isomorphic.
+\end{example}
+
+\begin{example}
+    Now consider the cospans from Figure~\ref{fig:appendix:non-isomorphic-cospans}(b).
+    Notice how the position of external input interface relative to strict internal interface is different in the second cospan.
+    The isomorphism would map $u_0 \mapsto u_2, u_1 \mapsto u_1, u_2 \mapsto u_0, u_3 \mapsto u_3$.
+    The other mappings are identities. 
+    One can check that the corresponding diagram commutes.
+\end{example}
+
+\begin{figure*}
+    \begin{subfigure}[T]{0.48\textwidth}
+    \begin{subfigure}[T]{0.45\textwidth}
+        \[
+    \scalebox{0.6}{
+        \tikzfig{../figures/appendix/sym_cospan}
+    }    
+    \]
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[T]{0.45\textwidth}
+        \[
+            \scalebox{0.6}{
+                \tikzfig{../figures/appendix/id_x_id_cospan}
+            }    
+            \]
+    \end{subfigure}
+    \subcaption{\;}
+\end{subfigure}
+    \hfill
+\begin{subfigure}[T]{0.48\textwidth}
+    \begin{subfigure}[T]{0.45\textwidth}
+        \[
+            \scalebox{0.6}{
+                \tikzfig{../figures/appendix/id_x_id_cospan}
+            }    
+            \]
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[T]{0.45\textwidth}
+        \[
+            \scalebox{0.6}{
+                \tikzfig{../figures/appendix/id_x_id_cospan_2}
+            }    
+            \]
+    \end{subfigure}
+    \subcaption{\;}
+\end{subfigure}
+    \caption{Non-isomorphic cospans (a) and isomorphic cospans (b)}
+    \label{fig:appendix:non-isomorphic-cospans}
+\end{figure*}

### 2024-06-13 LICS/sections/combinatorial_semantics.tex
arxiv version added, repo restructured

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
-%-------------------- INTRODUCTION
-
-\section{Introduction}
-
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
-This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
-
-A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
-
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
-
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
-
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
-
-\subsection{E-graphs}
-
-\begin{figure*}
-\[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
-    }
-\]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
-
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
-
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
-
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
-
-\subsection{Semilattice Enriched Categories}
-
-\begin{figure}
-\[
-	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
-	}
-\]
-\captionsetup{belowskip=-5ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
-\end{figure}
-
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
-\[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
-\]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
-
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
-
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
-\[
-    \tikzfig{categorical-semantics/egraph-translation-1}
-\]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
-As it would involve needless additional presentational complexity we leave it as further work.
-
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
-
-\subsection{Combinatorial Representation of Enriched String Diagrams}
-
-In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
-For example, we have the following equivalences between diagrams.
-%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
-	}
-\]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
-Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
-
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
-These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
-
-\subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
-They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
-
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
-
-%\subsection{E-matching and E-rewriting}
-%
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
-In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
-In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
-Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-%\begin{figure}
-%\[
-%	\tikzfig{combinatorial_semantics/example-rewriting}
-%\]
-%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
-REDO THIS}
-%\end{figure}
-
-\subsection{Related Work}
-
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
-We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
-
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
-
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
 \section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
@@
 Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
+\ifdefined \ONECOLUMN
+\[
+    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+\]
+\else
 \begin{multline*}
@@
 \end{multline*}
+\fi
 and its unit by the initial object of $\mathbb{C}$.
@@
 
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
+In particular, this means that any morphism in $\HypI{\Sigma}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
 where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
@@
 \end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
+We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA cospans. 
 \end{definition}
@@
 
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
+The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal \mathcal{L}^{\bot} \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal{L}^{\bot}$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
 The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
@@
 \begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
+    \begin{subfigure}[T]{0.4\linewidth}
     \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
+                % \tikzfig{../figures/combinatorial_semantics/DPO_square}
+                \begin{tikzcd}
+                    {\mathcal{L}} & {\mathcal{K}} & {\mathcal{R}} \\
+                    {\mathcal{G}} & {\mathcal{L}^{\bot}} & {\mathcal{H}}
+                    \arrow["f"', from=1-1, to=2-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-2, to=1-3]
+                    \arrow[from=1-2, to=2-2]
+                    \arrow[from=1-3, to=2-3]
+                    \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=90}, draw=none, from=2-1, to=1-2]
+                    \arrow[from=2-2, to=2-1]
+                    \arrow[from=2-2, to=2-3]
+                    \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=180}, draw=none, from=2-3, to=1-2]
+                \end{tikzcd}
     \]
@@
     \hfill
-    \begin{subfigure}{0.4\linewidth}
+    \begin{subfigure}[T]{0.4\linewidth}
         \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
+    % \scalebox{0.8}{
+        % \tikzfig{../figures/combinatorial_semantics/DPOI_square_Hyp}
+    % }
+    \begin{tikzcd}
+        {\mathcal{L}} & {i+j} & {\mathcal{R}} \\
+        {\mathcal{G}} & {\mathcal{L}^{\bot}} & {\mathcal{H}} \\
+        & {n+m}
+        \arrow["f"', from=1-1, to=2-1]
+        \arrow[from=1-2, to=1-1]
+        \arrow[from=1-2, to=1-3]
+        \arrow[from=1-2, to=2-2]
+        \arrow[from=1-3, to=2-3]
+        \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=90}, draw=none, from=2-1, to=1-2]
+        \arrow[from=2-2, to=2-1]
+        \arrow[from=2-2, to=2-3]
+        \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=180}, draw=none, from=2-3, to=1-2]
+        \arrow[from=3-2, to=2-1]
+        \arrow[from=3-2, to=2-2]
+        \arrow[from=3-2, to=2-3]
+    \end{tikzcd}
 \]
@@
 \label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
+For MDA cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
 and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
@@
     \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
+        \tikzfig{../figures/combinatorial_semantics/DPOI_boundary_complement}
     }
@@
 \]
-is a monogamous cospan. 
+is an MDA cospan. 
 \end{definition}
@@
 We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
+We call a match $f : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
 
@@
 Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
+Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $\mathcal{L}^{\bot}$ and cospan arrows $i+j \xrightarrow{} \mathcal{L}^{\bot} \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
 and the following conditions hold
 \begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
+    \item $f : L \to \mathcal G$ is a convex match;
+    \item $i + j \to \mathcal{L}^{\bot} \to \mathcal G$ is a boundary complement in the leftmost pushout.
 \end{itemize}
@@
 %\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
-
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
-\end{figure}
-
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\end{document}

### 2024-06-13 LICS/sections/conclusion.tex
arxiv version added, repo restructured

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
-%-------------------- INTRODUCTION
-
-\section{Introduction}
-
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
-This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
-
-A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
-
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
-
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
-
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
-
-\subsection{E-graphs}
-
-\begin{figure*}
-\[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
-    }
-\]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
-
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
-
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
-
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
-
-\subsection{Semilattice Enriched Categories}
-
-\begin{figure}
-\[
-	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
-	}
-\]
-\captionsetup{belowskip=-5ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
-\end{figure}
-
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
-\[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
-\]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
-
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
-
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
-\[
-    \tikzfig{categorical-semantics/egraph-translation-1}
-\]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
-As it would involve needless additional presentational complexity we leave it as further work.
-
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
-
-\subsection{Combinatorial Representation of Enriched String Diagrams}
-
-In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
-For example, we have the following equivalences between diagrams.
-%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
-	}
-\]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
-Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
-
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
-These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
-
-\subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
-They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
-
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
-
-%\subsection{E-matching and E-rewriting}
-%
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
-In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
-In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
-Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-%\begin{figure}
-%\[
-%	\tikzfig{combinatorial_semantics/example-rewriting}
-%\]
-%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
-REDO THIS}
-%\end{figure}
-
-\subsection{Related Work}
-
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
-We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
-
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
-
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
-\end{figure}
-
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
 \section{Conclusion}
 
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
+We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over the category of semilattices. 
 In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
@@
 \begin{figure}
+\begin{subfigure}{0.45\linewidth}
 \[
 	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
+	\tikzfig{../figures/conclusion/example-e-graphs_1}
 	}
 \]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
+\subcaption{\;} 
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\linewidth}
+\[
+	\scalebox{0.75}{
+	\tikzfig{../figures/conclusion/example-e-graphs_2}
+	}
+\]
+\subcaption{\;}
+\end{subfigure}
+\caption{E-(hyper)graphs for functorial boxes and the ZX-calculus}
 \label{fig:app1}
 \end{figure}
+
 \begin{figure}
 	\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
+	\scalebox{0.65}{
+	\tikzfig{../figures/conclusion/example-traced}
 	}
@@
 Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\end{document}

### 2024-06-13 LICS/sections/e-hypergraphs.tex
arxiv version added, repo restructured

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
-%-------------------- INTRODUCTION
-
-\section{Introduction}
-
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
-This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
-
-A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
-
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
-
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
-
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
-
-\subsection{E-graphs}
-
-\begin{figure*}
-\[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
-    }
-\]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
-
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
-
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
-
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
-
-\subsection{Semilattice Enriched Categories}
-
-\begin{figure}
-\[
-	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
-	}
-\]
-\captionsetup{belowskip=-5ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
-\end{figure}
-
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
-\[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
-\]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
-
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
-
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
-\[
-    \tikzfig{categorical-semantics/egraph-translation-1}
-\]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
-As it would involve needless additional presentational complexity we leave it as further work.
-
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
-
-\subsection{Combinatorial Representation of Enriched String Diagrams}
-
-In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
-For example, we have the following equivalences between diagrams.
-%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
-	}
-\]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
-Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
-
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
-These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
-
-\subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
-They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
-
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
-
-%\subsection{E-matching and E-rewriting}
-%
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
-In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
-In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
-Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-%\begin{figure}
-%\[
-%	\tikzfig{combinatorial_semantics/example-rewriting}
-%\]
-%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
-REDO THIS}
-%\end{figure}
-
-\subsection{Related Work}
-
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
-We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
-
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
-
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
-\end{figure}
-
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
 \section{E-Hypergraphs}\label{sec:e-hypergraphs}
@@
 
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
+Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$  such that for all $c_1,c_2 \in \Sigma$, if arity (respectively, co-arity) of $c_1$ is 0, then co-arity (respectively, arity) of $c_2$ must be at least 1.
+This is to ensure we do not have terms of type $0 \to 0$.
 This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
@@
 We define the following additional notation.  
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
+For $x \in V + E$,  the set of \emph{parents} of $x$, is denoted as $[x) = \{x' ~|~ x' < x\; \}$.
+% \[
+% 	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
+% \]
+We will denote the immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
 We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
@@
 	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
+	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$ and similarly if $v \in t(e)$.
 \end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
+We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu x\}$ of elements which share the same parent.  
+We collect the union of all these equivalence relations into one relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
+Note that this relation is symmetric and transitive if considered on $V + E$.
+The consistency relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ and there exists $e'$ such that $e' <^{\mu} v$ and $e' <^{\mu} e$ then $v \consistency e$.
 \label{def:consistency_properties}
@@
 These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
+Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for the consistency relation. 
 \begin{definition}[E-hypergraph homomorphism]
@@
     \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
+    \item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
 	    \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
+        \text{if } e <_{\mathcal{F}}^{\mu} v \text{ then } \phi(e) <_{\mathcal{G}}^{\mu} \phi(v)
         \]
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
+        and for $e_1, e_2 \in E_{\mathcal{F}}$
         \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
+        \text{if } e_1 <_{\mathcal{F}}^{\mu} e_2 \text{ then } \phi(e_1) <_{\mathcal{G}}^{\mu} \phi(e_2)
         \]
@@
 \end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
+% If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
 \end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
+The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the consistency relation.  
 \begin{definition}[Category of e-hypergraphs]
@@
 A concrete description of the pushout of two morphisms in this category is given in the Appendix.  
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
+Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied, including the requirement that the shared domain of the morphisms that form the span for the pushout is a discrete e-hypergraph.
 
 \begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
+When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we let the immediate parent to be undefined for all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
 \end{remark}
@@
 \begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
+    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete \textit{ordered} e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
     \[
@@
     \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
+    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete ordered e-hypergraphs,  $f_{ext},g_{ext}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
 We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
 
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
+We will call $n$  \textit{external input interfaces} and $n'$ \textit{internal input interfaces}.
 We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.  
@@
 
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
+Consider a relation $R = \{x R y \text{ if } f_{int}(x) \consistency f_{int}(y) \text{ for } x,y \in n'\}$ and let $S$ be its reflexive closure.
+The latter partitions $n'$ into non-empty subsets $\{p_{j}\}_{j=1}^{k}$. We get an analogous partition for $m'$.
+An example of an extended cospan can be seen in Figure~\ref{fig:e-cospan-example}.
+The cospan represents a morphism that encodes symmetry put into a hierarchical edge.
+The partition of the interfaces is given with colours.
+The labels suggest how $f_{int}, f_{ext}$ (respectively, $g_{int}, g_{ext}$) act on vertices.
+The ordering of vertices in the interfaces is from left to right.
+
+\begin{figure}
+    \[
+    \scalebox{0.6}{
+        \tikzfig{../figures/combinatorial_semantics/extended_cospan_example}
+    }    
+    \]
+    \captionsetup{belowskip=-3ex}
+    \caption{Extended cospan example}
+    \label{fig:e-cospan-example}
+\end{figure}
+
+Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute and such that $\alpha$ and $\gamma$ preserve order within each $p_i$.
 \[
 \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
+    \tikzfig{../figures/combinatorial_semantics/isomorphic_e_cospans}
 }
 \]
+Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{Frobenius}.
+Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix.
+
+\begin{remark}\label{remark:embedding_functor}
+There is also a faithful identity-on-objects functor $E: \HypI{\Sigma} \to \Ecospans$ which maps a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ to an extended cospan $n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m$ such that $f_{ext};f_{int} = f$ and $g_{ext};g_{int} = g$.
+\end{remark}
+
 Composition of two morphisms 
@@
 \adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
+{\tikzfig{../figures/combinatorial_semantics/pushout_e_cospans}}
 }
@@
 As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
+% Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.  
+% This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.  
+% This also means that we omit the monomorphism condition from the following definition.  
 We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
@@
         \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
+        \item $f_{int}$, $g_{int}$ are monomorphisms; 
         \item The in-degree and out-degree of every vertex is at most 1; 
@@
 The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
+We call $e$ \emph{well-typed} if the size of each subset in the partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
+We will call an MDA extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
 
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
+We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA extended-cospans. 
 \end{definition}
@@
 Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.  
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
+This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.
 Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
+This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
+% , cf. string diagrams for Cartesian categories.  
+An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this.
 Presently,  we will treat these equations via rewriting as well.
@@
 There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
+for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.  
+While we believe this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such sub-hypergraphs.  
+For this reason,  in the remainder of the paper we consider signatures which can not result in construction of terms with type $0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges.
 %This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
@@
 
-For monogamous extended cospans 
+For MDA cospans 
 \[
@@
     n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
+\] and mono $m : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
 as depicted in the square below
@@
 
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
+    \tikzfig{../figures/combinatorial_semantics/DPOI_pushout_complement}
 }
@@
     \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
+    \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
     \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
@@
 This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
+\begin{remark}
+The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+\[
+	\mathcal{L} = \scalebox{0.8}{\tikzfig{../figures/combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{../figures/combinatorial_semantics/f_plus_g_inline}}
+\]
+\end{remark}
 %The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
@@
  \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
+    \tikzfig{../figures/combinatorial_semantics/DPOI_square}
  }
@@
 \end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
+% Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
 
@@
 %
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
-
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
-\end{figure}
-
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\end{document}

### 2024-06-13 LICS/sections/introduction.tex
arxiv version added, repo restructured

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
 %-------------------- INTRODUCTION
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
@@
 
+\ifdefined \ONECOLUMN
+\begin{figure}
+	\[
+		\scalebox{0.55}{
+		\tikzfig{../figures/categorical-semantics/egraph-translation}
+		}
+	\]
+	\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
+	\label{fig:e-graph-example}
+	\end{figure}
+\else
 \begin{figure*}
@@
     \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
+    \tikzfig{../figures/categorical-semantics/egraph-translation}
     }
@@
 \end{figure*}
+\fi
 
@@
 The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
+The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node
+\scalebox{0.3}{
+	\begin{tikzpicture}
+		\begin{pgfonlayer}{nodelayer}
+			\node [style=vertex] (0) at (1.5, 3.5) {};
+			\node [style=none] (1) at (1.5, 3) {};
+			\node [style=none] (2) at (1, 4) {};
+			\node [style=none] (3) at (2, 4) {};
+		\end{pgfonlayer}
+		\begin{pgfonlayer}{edgelayer}
+			\draw (0) to (1.center);
+			\draw [bend right=45] (0) to (3.center);
+			\draw [bend left=45] (0) to (2.center);
+		\end{pgfonlayer}
+	\end{tikzpicture}
+}
+.
 
@@
 This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
+Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes 
+\scalebox{0.4}{
+	\begin{tikzpicture}
+		\begin{pgfonlayer}{nodelayer}
+			\node [style=vertex] (0) at (1.5, 3.5) {};
+			\node [style=none] (1) at (1.5, 3) {};
+		\end{pgfonlayer}
+		\begin{pgfonlayer}{edgelayer}
+			\draw (0) to (1.center);
+		\end{pgfonlayer}
+	\end{tikzpicture}
+}, indicate which of the class-level edges are connected to which nodes inside the class.
 
 The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
+If we were to apply another rule, $x * 1 \to x$, we would identify the e-class of $a$ with the top-level e-class of the e-graph which would require a cycle to be introduced: the left argument of node $*$ would be the same e-class that $*$ belongs to.
+This would result into a possibility of extracting the optimal term consisting of only the node $a$, discarding everything else.
+To support such cycles we need a richer structure for our e-string diagrams.
+While equipping the e-string diagrams which such structure is straightforward, it is out of the scope of this work, and thus we limit ourselves to acyclic e-graphs and e-string diagrams for the purposes of presentation.
+Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the e-string diagrammatic version is guided by the equations of the underlying theory.
+A step-by-step application of such equations for the rewrite rule from $(a)$ to $(b)$ and $(b)$ to $(c)$ can be seen in Figure~\ref{fig:e-graph-example-a-b} and Figure~\ref{fig:e-graph-example-b-c} respectively.
+The equations applied are the ones from Figure~\ref{fig:string-equations} and naturality of delete and copy.
+One can note that the size of the eventual e-string diagram is comparable to the size of the corresponding e-graph.
+% Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
+% From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
 
@@
 	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
+	\tikzfig{../figures/categorical-semantics/egraph-strings}
 	}
 \]
-\captionsetup{belowskip=-5ex}
+\captionsetup{belowskip=-4ex}
 \caption{String diagrams for semilattice enriched symmetric monoidal categories.}
@@
 \[
-    \tikzfig{categorical-semantics/egraph-translation-1}
+    \tikzfig{../figures/categorical-semantics/egraph-translation-1}
 \]
+The figure on the right above is slightly simplified: if we denote the arity of each $c_i$ as $a_{c_{i}}$, then the dashed box on the right contains $\sum\limits_{i} a_{c_{i}}$ inputs and each component $c_j$ has $\sum\limits_{i \not = j} a_{c_{i}}$ weakening nodes inside.
 Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
@@
 In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
+String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of a \textit{symmetric monoidal category (SMC)}.
 For example, we have the following equivalences between diagrams. 
@@
 	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
+	\tikzfig{../figures/categorical-semantics/interchange}
 	}
@@
 Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
+We denote the tensor product of an SMC $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.  
 We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
 We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
+We denote by $\catname{SLatt}$ the category of semilattices.
 
@@
 
+% \begin{remark}
+%     Below we will assume that for all $c_1,c_2 \in \Sigma$, if arity (respectively, co-arity) of $c_1$ is 0, then co-arity (respectively, arity) of $c_2$ must be at least 1.
+% 	This is to ensure we do not have terms of type $0 \to 0$.
+% \end{remark}
+
 A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
@@
 \end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
+The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
 
@@
 	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
+  	 \tikzfig{../figures/categorical-semantics/Cartesian-equipment}
 	}
@@
 	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
+	\tikzfig{../figures/categorical-semantics/Cartesian-theory}	
 	}
 \]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
+The Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations of the SMT of commutative comonoids, and the following additional naturality equations, for every $c \in \Sigma_C$
 \[
 	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
+	\tikzfig{../figures/categorical-semantics/Cartesian-naturality}
 	}
@@
 
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
+\begin{definition}[Semilattice-enriched SMC]\label{def:enriched-prop}
+A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by an SMC $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
 \begin{align*}
@@
 Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
+A \textit{semilattice enriched strict SMC functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC functor on the underlying SMCs which additionally satisfies $F(f+g) = F(f)+F(g)$.
+All definitions extend to \textit{PROP} when the manipulated categories are PROPs rather than SMCs. 
 We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively). 
@@
 
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
+The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
 
@@
 
+\ifdefined \ONECOLUMN
+\begin{figure}
+	\[  
+		\scalebox{0.6}{
+		\tikzfig{../figures/categorical-semantics/egraph-strings-equations}
+		}
+	\]
+	\caption{Equations for a  semilattice enriched symmetric monoidal category}
+	\label{fig:string-equations}
+	\end{figure}
+\else
 \begin{figure*}
@@
     \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
+	\tikzfig{../figures/categorical-semantics/egraph-strings-equations}
     }
@@
 \end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
-\end{figure}
-
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
-
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
-\end{figure}
-
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\end{document}
+\fi

### 2024-06-13 draft/sections/introduction.tex
arxiv version added, repo restructured

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
+% !TEX root = ../main.tex
 
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
-%-------------------- INTRODUCTION
+\newcommand\mylet[2]{\textsf{let } #1 = #2 \textsf{ in }}
 
@@
 The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
+The \textit{join} operator lifts the formalism so that it can now operate with \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial represetation.
 
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
+Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques.
 This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
@@
 
-\begin{figure*}
+\begin{figure}\label{fig:e-graph-example}
 \[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
+    \scalebox{0.5}{
     \tikzfig{categorical-semantics/egraph-translation}
@@
 \]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
+\caption{Example translation of e-graphs into string diagrams for semilattice enriched (traced) Cartesian categories. DAN: (a) SHOULD USE MATH FONT.}
+\end{figure}
+
+\begin{figure}\label{fig:let-calculus}
+\begin{align*}
+    (a*2)/2\\
+    \mylet{y}{a*2} y/2 \\ 
+    \mylet{y}{\{a*2,a<\!\!<1\}} y/2 \\
+    \mylet{y}{\{a*2,a<\!\!<1\}} \{y/2, a*(2/2)\} \\
+    \mylet{x}{\{2/2, 1\}} \mylet{y}{\{a*2, a <\!\!< 1\}} \{y/2, a*x\} \\
+    \textsf{letrec } z = (\mylet{x}{\{2/2, 1\}} \mylet{y}{\{z*2, z <\!\!< 1\}} \{a, y/2, z*x\}) \textsf{ in } z
+\end{align*}
+\caption{Example term calculus corresponding to the string diagrams of Figure \ref{fig:e-graph-example} 
+DAN: THESE SEEM TO HAVE THE WRONG SHARING. DO WE NEED A TERM CALCULUS AT ALL? MAYBE MOVE TO APPENDIX.} 
+\end{figure}
 
@@
 Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
+The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2\footnote{In fact, a more accurate term representation of the DAG would be: $\mylet{t_0}{2}(a*t_0)/t_0$.}.
 The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
@@
 
-\subsection{Semilattice Enriched Categories}
+\subsection{Semantics via Semilattice Enriched Categories}
 
-\begin{figure}
+\begin{figure}\label{fig:egraph-strings}
 \[
-	\scalebox{0.65}{
     \tikzfig{categorical-semantics/egraph-strings}
-	}
 \]
-\captionsetup{belowskip=-5ex}
 \caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
 \end{figure}
@@
 Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
+makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{noauthor_09083347_nodate,joyal_geometry_1991, mellies_functorial_2006}.
+In Figure~\ref{fig:egraph-strings} we extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
 \[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
+\infer{\phi_1 + \ldots + \phi_n: A \to B}{\{\phi_i: A \to B\}_{i \in \{1,\ldots,n\}}}
 \]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
+The ability to take formal sums, which in concrete categories will usually be set union, is used to model the equivalence class structure of e-graphs.
 
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
+To aid understanding, in Figure \ref{fig:let-calculus} we sketch a representation of the string diagrams in Figure \ref{fig:e-graph-example} using a let-calculus, which is standard except for its augmentation with the typing rule below, corresponding to semilattice enrichment.
+The new construct forms a set of terms and satisfies the following additional congruence: $K\{\{t_1, \ldots, t_n\}\} = \{K\{t_1\}, \ldots, K\{t_n\}\}$.
+DAN: WHAT IS K?
+WHAT ARE THE NESTED CURLY BRACES?
+\[
+\infer{\Gamma \vdash \{t_1, \ldots, t_n\}: A}{\{\Gamma \vdash t_i: A\}_{i \in \{1, \ldots, n\}}}
+\]
+We leave the details of this calculus informal, as they are parenthetic to the main thrust of our work.
+
+DAN: TOP TO BOTTOM OR BOTTOM TO TOP? UNCLEAR.\\
+String diagrams are read from top-to-bottom; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
+Note, further, that in the Cartesian case we can restrict to generators $c$ with a single output.\\
+DAN: WHY? WITH OR WITHOUT LOSS OF GENERALITY? WHAT DOES THIS MEAN?
 
@@
 \]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
+DAN: IT WOULD BE CLEARER IMHO IF THE ARGUMENTS OF C1 AND C2 REMAIN CENTERED AND THE DISCARDED NODES ARE OFF CENTER.
+
+Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}. In particular, note that the translation of the \textit{cyclic} e-graph (e)
+requires the use of a categorical \textit{trace}, generating a cycle.
+In this paper, we focus on \textit{acyclic} e-graphs and their corresponding catgeories; but the above example illustrates that the extension to the cyclic case seems to be routine.
 As it would involve needless additional presentational complexity we leave it as further work.
 
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
+Note how the natural level of generality for our string diagrams is not Cartesian, but monoidal: the generators of a Cartesian category are required in order to embed e-graphs, but are orthogonal to the addition of formal sums.
+\\
+DAN: SUMS OR JOINS? CONSISTENT?\\
+Thus we generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we outline in the conclusion of this paper.   
+\\ DAN: THIS IS A BIT UNCLEAR. WHY IS THIS THE NATURAL LEVEL THE GENERALITY? IT'S NOT QUITE OBVIOUS.
 
@@
 \[
-	\scalebox{0.6}{
 	\tikzfig{categorical-semantics/interchange}
-	}
 \]
@@
 A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
+Here, the generators $c_i$ become vertices, connected via hyper-edges.
 Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
+With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some $c_i$.
 
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
+We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(e) of Figure \ref{fig:e-graph-example}.
 These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
+Because the generating equations can be applied in any context, we are lead to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ in any hypergraph $G$ and replace it with (a sub-hypergraph corresponding to) $r$.
+The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{???} is still applicable.
 However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
@@
 \subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
+The main technical result of the paper is a proof soundness and completeness of the combinatorial representation for semilattice enriched SMCs, extending the results of \cite{bonchi_string_2022-2} for plain SMCs.
+While the structural equalities of SMCs are factored out in the representation,  important structural equalities arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
 They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
@@
 
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
+Quotienting e-hypergraphs by the induced equivalence gives rise to a sound and complete model of semilattice enriched SMCs.
+That is, we exhibit a translation functor $\llbracket-\rrbracket$ from the free semilattice enriched SMC over a symmetric monoidal theory to our (quotiented) category of e-hypergraphs such that
+morphisms $f = g$ if and only if there exists a sequence of DPO-rewrites (each induced by a structural equality or the monoidal theory) between their translations: $\llbracket f \rrbracket \rightsquigarrow_{DPO} \llbracket g \rrbracket$.
+
+DAN: THIS NEEDS EXPANDING. EXPLAINING IN PLAIN WORDS SOME OF THE THEORETICAL PROPERTIES OF E-GRAPHS AND THE CATEGORICAL STATEMENT.
 
@@
 
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
+Although e-graphs were first developed in the 80's \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
 This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
@@
 
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
+Our use of string diagrams is also central to the intuition behind our work.
+The string diagrams we work with are left essentially as informal reasoning devices, however there exists a breadth of work on similar diagrammatic formalisms for similar classes of categories.
+We mention here work on hierarchical string diagrams \cite{ghica_hierarchical_2023},  functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
 
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
-\end{figure}
-
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
-
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
-\end{figure}
-
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\end{document}
+DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.

### 2024-06-13 draft/sections/soundness_and_completeness.tex
arxiv version added, repo restructured

--- 2023-12-20 sections/soundness_and_completeness.tex
@@
 \end{corollary}
+
+\begin{proposition}
+    For each cospan $A$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$ there is a week normal form such that $A =_{\mathcal{S}} A_1 + \ldots A_m$ where each of $A_i$ does not contain any hierarchical edges.
+    \begin{proof}
+    By applying the rules 1-6 in $\mathcal{S}$ (the ones from Figure~\ref{fig:string-equations}) we either remove 1 hierarchical edge (rules 5-6) or do not change the number of them, but expand the edge onto other edges.
+    We will have a redex for such a rule until there is only one hierarchical edge at the outermost level (and only one outermost edge) with no other hierarchical edges within it.
+    \end{proof}
+\end{proposition}
 
@@
 \begin{proof}
-As we quotient by $\mathcal{S}$ instead of arbitrary cospans we can only consider their normal forms $A_1 + \ldots A_n$ where all of $A_i$ do not contain any hierarchical edges.
-Because each of $A_i$ is a plain hypergraph, they have a pre-image in $\textsf{PROP}(\Sigma)$, i.e. $A_i = \llbracket \cdot \rrbracket;\mathcal{I}(a_i)$.
+As we quotient by $\mathcal{S}$ instead of arbitrary cospans we can only consider their normal forms $A_1 + \ldots A_n$ where all of $A_i$ do not contain any hierarchical edges as per the proposition above.
+Because each of $A_i$ is a plain hypergraph, they have a pre-image in $\textsf{PROP}(\Sigma)$, i.e.
+for all $i$ there exists $a_i$ in $\textsf{PROP}(\Sigma)$ such that $A_i = \llbracket a_i \rrbracket;\mathcal{I}$.
 And then $[\mathcal{J}(a_1) + \ldots + \mathcal{J}(a_n)] = [\mathcal{J}(a_1)] + \ldots + [\mathcal{J}(a_n)] = A_1 + \ldots + A_n$ and hence $\mathcal{J}(a_1) + \ldots + \mathcal{J}(a_n)$ is the pre-image of $A_1 + \ldots + A_n$.
@@
 \end{theorem}
-
-\question{Is proof ok?}
 
@@
 
+\begin{theorem}[Theorem 35~\cite{Frobenius2}]
+
+    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
+    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
+\end{theorem}
+
+\begin{remark}
+    If $\llbracket a \rrbracket = n \xrightarrow{} A \xleftarrow{} m$ then $\llangle ^{\ulcorner} A ^{\urcorner} \rrangle$ is isomorphic to $0 \xrightarrow{} A \xleftarrow{} n + m$.
+\end{remark}
+
+\begin{proposition}
+    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
+     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
+    \begin{proof}
+        The forward direction follows by the remark above and by recalling that $\Rrightarrow$ implies there exists a pushout square below
+        \[
+        \scalebox{1}{
+            \tikzfig{soundness_and_completeness/pushout_in_Hyp}
+        }    
+        \]
+        where the matching $m$ is convex, $i+j \to C \to G$ is a boundary complement~\cite{Frobenius2} and marked squares are pushouts
+        and by using the injection $\mathcal{K} : \catname{Hyp_{\Sigma}} \to \catname{EHyp_{\Sigma}}$ there exists the following pushout square
+        \[
+          \scalebox{0.75}{
+            \tikzfig{soundness_and_completeness/pushout_in_EHyp}
+          }  
+        \]
+        where $m$ is convex and down-closed (as there are no hierarchical edges) and $\mathcal{K}(i+j) \to C \to G$ is a boundary complement.
+        This by definition means that $\mathcal{K}(n) \xrightarrow{} \mathcal{K}(n) \xrightarrow{} G \xleftarrow{} \mathcal{K}(m) \xleftarrow{} \mathcal{K}(m) = [\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} = [\mathcal{J}(b)] = \mathcal{K}(n) \xrightarrow{} \mathcal{K}(n) \xrightarrow{} G \xleftarrow{} \mathcal{K}(m) \xleftarrow{} \mathcal{K}(m)$.
+        The other direction follows by applying the same argument backwards.
+    \end{proof}
+\end{proposition}
+
 \begin{theorem}
     $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
+    \begin{proof}
+        $a \Rightarrow_{\langle l, r \rangle}^{+} b$ implies that $a$ is decomposable as $a_1 + \ldots + a' + \ldots + a_k$ and $b$ is decomposable as $a_1 + \ldots + b' + \ldots + a_k$ and $a' \Rightarrow_{\langle l, r \rangle} b'$.
+        The latter implies by the Theorem 35~\cite{Frobenius2} that $\llangle \ulcorner a' \urcorner \rrangle \Rrightarrow_{\langle \llangle ^\ulcorner l ^\urcorner \rrangle, \llangle ^\ulcorner r ^\urcorner \rrangle \rangle} \llangle \ulcorner b' \urcorner \rrangle$ which means that $[a'] \Rrightarrow_{\langle [l], [r] \rangle}^{+} [b']$.
+        This means that there exists a convex matching of $[l]$ within $[a']$ and hence there is a convex match of $[l]$ in $[a_1] + \ldots + [a'] + \ldots + [a_k]$ and that the pushout complement is $[a_1] + \ldots + C + \ldots + [a_k]$ where $C$ is the pushout complement for $[a'] \Rrightarrow_{\langle [l], [r] \rangle} [b']$.
+        Similarly, the pushout is $[a_1] + \ldots [b'] + \ldots + [a_k] = [b]$.
+
+        For the converse direction, $[a] \Rrightarrow_{\langle [l],[r] \rangle}^{+} [b]$ implies that $[a] = A_{1} + \ldots A' + \ldots + A_{n}$ and $[b] = A_{1} + \ldots B' + \ldots + A_{n}$ and $A' \Rrightarrow_{\langle [l],[r] \rangle}^{+} B'$ because $[l]$ and $[r]$ have a pre-image in $\textsf{PROP}(\Sigma)$ and because we quotient by $\mathcal{S}$.
+        The latter means by the proposition above that $a' \Rightarrow_{\langle l, r \rangle} b'$ such that $A' = \llbracket a' \rrbracket; \mathcal{I}$ and $B' = \llbracket b' \rrbracket; \mathcal{I}$ and hence $a = a_1 + \ldots + \mathcal{I}(a') + \ldots + a_n \Rightarrow_{\langle l, r \rangle}^{+} a_1 + \ldots \mathcal{I}(b') + \ldots a_n = b$
+    \end{proof}
+    \update{It's again somewhat handwaving, do we need any supporing proofs for the above?}
 \end{theorem}

### 2024-06-13 extended_abstract/EGraphsExtendedAbstract.tex
arxiv version added, repo restructured

--- 2024-04-17 EGraphsExtendedAbstract.tex
@@
 
-\usepackage{amsmath}
 \usepackage{stmaryrd}
-% \usepackage{amssymb}
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\usepackage{docmute}
-\usepackage{hyperref}
-\usepackage{proof}
 
@@
 
-\usepackage{aliascnt}
-% \newaliascnt{definition}{thm}
-% \newaliascnt{remark}{thm}
-% \newaliascnt{example}{thm}
-\newtheorem{thm}{Theorem}
-\newtheorem{remark}[thm]{Remark}
+% \newtheorem{thm}{Theorem}
+% \newtheorem{remark}[thm]{Remark}
 
-\theoremstyle{definition}
-\newtheorem{definition}[thm]{Definition}
+% \theoremstyle{definition}
+% \newtheorem{definition}[thm]{Definition}
 
-\newtheorem{example}[thm]{Example}
-\newtheorem{proposition}[thm]{Proposition}
+% \newtheorem{example}[thm]{Example}
+% \newtheorem{proposition}[thm]{Proposition}
 
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
+\input{../macros}
+\input{../preamble}
+\input{../sample.tikzstyles}
+\input{../hypergraph.tikzstyles}
+\input{../hypergraph.tikzdefs}
 
@@
 
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-
 \newcommand\id{\textsf{id}}
 \newcommand\sym{\textsf{sym}}
-
-\usepackage{cleveref} % after hyperref!
 
@@
 	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
+	\tikzfig{../figures/categorical-semantics/Cartesian-theory}	
 	}
@@
 	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
+	\tikzfig{../figures/categorical-semantics/Cartesian-naturality}
 	}
@@
         \scalebox{0.65}{
-        \tikzfig{categorical-semantics/egraph-strings}
+        \tikzfig{../figures/categorical-semantics/egraph-strings}
         }
@@
         \scalebox{0.5}{
-        \tikzfig{categorical-semantics/egraph-translation-2}
+        \tikzfig{../figures/categorical-semantics/egraph-translation-2}
         }
@@
 \[
-    \tikzfig{categorical-semantics/egraph-translation-1}
+    \tikzfig{../figures/categorical-semantics/egraph-translation-1}
 \]
@@
         \scalebox{0.5}{
-        \tikzfig{categorical-semantics/egraph-translation-step-by-step-a-b}
+        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
         }
@@
         \scalebox{0.55}{
-        \tikzfig{categorical-semantics/e-graph-substitution}
+        \tikzfig{../figures/categorical-semantics/e-graph-substitution}
         }
@@
         \scalebox{0.5}{
-        \tikzfig{categorical-semantics/e-string-substitution-2}
+        \tikzfig{../figures/categorical-semantics/e-string-substitution-2}
         }
@@
         \resizebox{0.8\textwidth}{!}{
-        \tikzfig{categorical-semantics/egraph-translation-step-by-step-b-c}
+        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-b-c}
         }
@@
 \bibliographystyle{acm}
-\bibliography{bibliography}
+\bibliography{../bibliography}
 

### 2024-06-22 LICS/sections/e-hypergraphs.tex
slides and minor fixes

@@
 \]
-Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{Frobenius}.
+Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{bonchi_string_2022-1}.
 Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix.

### 2024-06-22 slides/slides.tex
slides and minor fixes

%!LW recipe=xelatex

\documentclass[aspectratio=169]{beamer}

\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{tikz-cd}
\usepackage{tikzit}
\usepackage{subcaption}
\usepackage{minted}

\input{../sample.tikzstyles}
\input{../hypergraph.tikzstyles}
\input{../hypergraph.tikzdefs}

\usetheme{metropolis}
% \usetheme{awesome}

% \setsansfont{Ubuntu}
% \setmonofont{Ubuntu Mono}

\setbeamertemplate{footline}
{
  \leavevmode
  \hbox{
  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}
    \usebeamerfont{author in head/foot}\insertshortauthor
  \end{beamercolorbox}

  \begin{beamercolorbox}[wd=.7\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}

  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}
    \insertframenumber{} / \inserttotalframenumber
  \end{beamercolorbox}
  }
}

% \setbeamertemplate{footline}{\insertshortauthor~(\insertshortinstitute)}
% \usepackage{custom_footer}

\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories} 
\date{June 24, 2024} % Use metropolis theme 
\author[Aleksei Tiurin]{$\text{Chris Barrett}^{\dagger}$, $\textbf{Aleksei Tiurin}^{\ddagger}$, $\text{Dan R. Ghica}^{\ddagger}$}
\institute{$^{\dagger}$University of Oxford, $^{\ddagger}$University of Birmingham} 
\begin{document}

\maketitle

\begin{frame}{Table of contents}
    \setbeamertemplate{section in toc}[sections numbered]
    \tableofcontents[hideallsubsections]
\end{frame}

\section{What and Why?}

\begin{frame}
    Traditionally, e-graphs are used for \alert{theories} over \alert{algebraic} signatures with a single sort of composition $()$

    \begin{example}[Commutative monoid]

    \begin{align*}
    \Sigma = \{&\mu : 2 \to 1, i : 0 \to 1\}\\
    \mathcal{E} = \{&\mu(x,i\;) \equiv x, \mu(i,x) \equiv x,
     \mu(\mu(x,y),z) \equiv \mu(x,\mu(y,z)),\\
    &\mu(x,y) \equiv \mu(y,x)\}    
    \end{align*}
    \end{example}

    However, there are more general theories that describe the interaction of processes, \textit{e.g.}, quantum
\end{frame}

\begin{frame}{}

Consider a \alert{co-algebraic} signature 
\[
\Sigma = \{id_n : n \to n,\sigma_{n,m} : n + m \to m + n,\mu : 1 \to 2, i : 1 \to 0\}
\]
and a family of terms $t$ such that every term is constructed from the generators above by combining them sequentially ($;$) and in parallel ($\otimes$).

Given $f : n \to m, g : m \to k$, $f\;;g : n \to k$

Given $f : n \to m$, $g : k \to l$, $f \otimes g : n + k \to m + l$

\begin{example}
    \[
    (\mu\;;\sigma_{1,1}\;;i \otimes id_{1}) \otimes (\mu\;;(id_{1}\otimes id_{1}))
    \]
\end{example}

\end{frame}

\begin{frame}{}

\small

\[
\Sigma = \{id_n : n \to n,\sigma_{n,m} : n + m \to m + n,\mu : 1 \to 2, i : 1 \to 0\}
\]

Such signatures also come with \alert{built-in} equations

\begin{minipage}[c][][c]{0.45\textwidth}
\begin{align*}    
(s\;;t)\;;u \equiv s\;;(t\;;u)\\
id_{n}\;;s \equiv s \equiv s\;;id_{n}\\
id_{0} \otimes s \equiv s \otimes id_{0}\\
(s \otimes t) \otimes u \equiv (s \otimes t) \otimes u\\
(s\;;u) \otimes (t\;;v) \equiv (s \otimes t)\;;(u \otimes v)\\
\end{align*}
\end{minipage}
\hfill
\begin{minipage}[c][][c]{0.45\textwidth}
\begin{align*}
(id_{m} \otimes id_{n}) \equiv id_{m + n}\\
(\sigma_{m,n} \otimes id_{o});(id_{n} \otimes \sigma_{m,o}) \equiv \sigma_{m,n+o}\\
\sigma_{m,n} ; \sigma_{n,m} \equiv id_{m+n}\\
(s \otimes id_{m})\;; \sigma_{n,m} = \sigma_{n,m}\;; (id_{m} \otimes s)
\end{align*}    
\end{minipage}

\vspace{-1.5em}

A \textit{symmetric monoidal theory (SMT)} over $\Sigma$ is a set of terms constructed out of the symbols from $\Sigma$ using $;$ and $\otimes$ \textit{quotiented} by the equations above and by some other equations $\mathcal{E}$

\begin{example}[Commutative comonoid]
    % \vspace{-1.5em}
    \[
        \mathcal{E} = \{ \mu \;; (\mu \otimes id_{1}) \equiv \mu \;; (id_{1} \otimes \mu), \mu \;; (i \otimes id_{1}) \equiv id_{1} \equiv \mu\;;(id_{1} \otimes i\;), \mu;\sigma_{1,1} \equiv \mu \}  
    \]
\end{example}

\end{frame}

\begin{frame}
    \vfill
    A given SMT is just a syntax and yields a corresponding free symmetric monoidal category (a \textbf{PROP}) so we can reason about the SMT categorically

    \vfill

    The equations above are the laws of a (strict) symmetric monoidal category
    \vfill

    One gets a semantics of an SMT by building a functor for the corresponding \textbf{PROP}, \textit{e.g.} a functor into a category of vector spaces \textbf{$\text{Vect}_{k}$}
\end{frame}

\begin{frame}
    SMTs crop up in various areas including
    \begin{itemize}
        \item Quantum computing
        \item Probabilistic programming
        \item Lambda calculus
        \item Digital circuits
        \item $\ldots$
    \end{itemize}
\end{frame}

\begin{frame}[containsverbatim]{}

It is possible to encode such theories using e-graphs, for example as

\begin{minted}{c++}
    enum CommutativeComonoids {
        "mu" = Mu,
        "i" = I,
        ";" = Compose([Id;2]),
        "tensor" = Tensor([Id;2]),
        "id" = Identity,
        "I" = Unit,
        "sym" = Sym    
    }
\end{minted}

\end{frame}

\begin{frame}
    However, the buit-in equations (the laws of monoidal category) lead to an explosion of the number of e-classes

    \begin{example}
        \vspace{1em}
        E-graph for
        \[
            (\mu\;;\sigma_{1,1}\;;i \otimes id_{1}) \otimes (\mu\;;(id_{1}\otimes id_{1}))
        \]

        \begin{figure}
            \includegraphics[width=\linewidth]{figures/dot_5.pdf}
        \end{figure}

        \end{example}

        \pause

        Can we somehow get rid of these equations?
\end{frame}

\begin{frame}{}

\small

SMTs come with a two-dimensional syntax called \textit{string diagrams} which represent equivalence classes of terms quotiented by built-in equations which are a convenient tool for graphical reasoning
\[
\scalebox{0.7}{
\tikzfig{figures/string_diagrams}
}
\]

If one slides boxes along the wires or cross wires even number of times, they get the same string diagram

\end{frame}

\begin{frame}{}

    \small

    \begin{minipage}{0.6\linewidth}

        In particular note how bi-functoriality law gets absorbed by the notation
    \end{minipage}
    \begin{minipage}{0.35\linewidth}
        \[
         \scalebox{0.7}{
            \tikzfig{figures/bifunctoriality}
         }   
        \]        
    \end{minipage}
    \begin{minipage}{0.6\linewidth}       
        Further structure on the monoidal theory is encoded in geometrical properties on these string diagrams
    \end{minipage}
    \begin{minipage}{0.35\linewidth}

    \end{minipage}
    % \begin{minipage}{0.6\linewidth}
    %     \begin{itemize}
    %         \item There are equations for which their term-equivalent is not immediately obvious (ZX-calculus Z-spider fusion)
    %     \end{itemize}        
    % \end{minipage}
    % \begin{minipage}{0.35\linewidth}
    %     \[
    %      \scalebox{0.7}{
    %         \tikzfig{figures/Z_spider_fusion}
    %      }   
    %     \]        
    % \end{minipage}

\end{frame}

\begin{frame}
    Another benefit of string diagrams is that they simplify the additional equations for which their term-equivalent form is not immediately obvious

    \[
         \scalebox{1}{
            \tikzfig{figures/Z_spider_fusion}
         }   
    \]
\end{frame}

\begin{frame}
\vfill
Notably, string diagrams can be represented combinatorially as \alert{hypergraphs with interfaces} where the laws of the symmetric monoidal category gets \alert{absorbed} by the notion of isomorphic hypergraphs
\vfill
Diagrammatic reasoning becomes graph rewriting
\vfill
Further structure is encoded as rewrites and/or by enforcing additional constraints on the graph structure
\vfill
\end{frame}

\begin{frame}
So \alert{yes}, we can get rid of these equations if we use \alert{hypergraphs}
\vfill
The work next formalises e-graphs for SMTs that would use string diagrams (hypergraphs) as a representation for terms
\end{frame}

% \begin{frame}{}

%     The applications of SMTs (and string diagrams) include
%     \begin{itemize}
%         \item Quantum computing
%         \item Probabilistic programming
%         \item Lambda calculus
%         \item Digital circuits
%         \item $\ldots$
%     \end{itemize}

%     Can we have e-graphs for such two-dimensional syntax (so that built-in equations are all absorbed rather than represented with e-classes)?

%     \pause

%     This work builds a foundation to answer the above question
% \end{frame}

\section{Underlying Theory}

\begin{frame}{}
    \vfill

    First we introduce the notion of an \textit{enriched} SMT where we can take formal joins of terms
    \[
    f_1 + \ldots +f_n, \text{\; where $f_{i} : m \to n$}    
    \]
    \vfill
    These formal joins will encode the equivalence classes, \textit{i.e.}, we will use them to express that $f_i \sim f_j$
    \vfill
    The joins are represented graphically as boxes
    \[
    \scalebox{0.8}{
        \tikzfig{figures/join_box}
    }    
    \]
\end{frame}

\begin{frame}{}
    \vfill
    Much like in e-graphs where $x \sim y$ implies $f(x) \sim f(y)$ for a functional symbol $f$ we have a notion of \textit{congruence} which is a bit more general since we have two sorts of composition
    \vfill
    These are encoded as string diagrams as well as the requirements that $+$ is an associative, commutative and idempotent operator
\end{frame}

\begin{frame}{}
    \[
        \scalebox{0.6}{
        \tikzfig{../figures/categorical-semantics/egraph-strings-equations}
        }
    \]
\end{frame}

\begin{frame}{}
\small
The non-destructive feature of rewriting with e-graphs is modelled by using the idempotence property of $+$
\vfill
Given a rewrite rule (in a string diagrammatic form) and a source string diagram
\begin{minipage}{0.4\linewidth}
    \[
    \scalebox{1}{
    \tikzfig{figures/rewrite_example_1}
    }
    \]
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
    \[
    \tikzfig{figures/rewrite_example_3}
    \]
\end{minipage}

\end{frame}
% \vfill
% These equations equip a set of terms (morphisms) with the same domain and co-domain with a structure of a semilattice and hence the models of such enriched SMTs are symmetric monoidal categories enriched over a category of semilattices

\begin{frame}
    We apply the equations (and the rewrite rule above) in the following way
    \[
    \tikzfig{figures/rewrite_example_2}
    \]
\end{frame}

\begin{frame}{}
    Such extended string diagrams have a concrete combinatorial representation which we call \textit{e-hypergraphs} with interfaces

        \begin{minipage}{0.45\linewidth}
            \[
            \tikzfig{figures/boxed_sym_string}    
            \]
        \end{minipage}
        \begin{minipage}{0.05\linewidth}
            \[
                \to
            \]
        \end{minipage}
        \begin{minipage}{0.45\linewidth}
            \[
            \scalebox{0.5}{
            \tikzfig{figures/extended_cospan_example}
            }
            \]
        \end{minipage}
\end{frame}

\begin{frame}{}
    \vfill
    Now equivalence classes represent equivalent subgraphs rather than subtrees
    \vfill
    Such combinatorial structures are shown to be \textit{sound} and \textit{complete} in terms of representing terms of enriched SMTs
    \vfill
    String diagrammatic reasoning becomes e-hypergraph rewriting
    \vfill
    The latter is defined as a \textit{double-pushout rewriting} which is a convenient framework to study properties of a rewriting system, e.g. confluence applicability of parallel rewrites etc.
\end{frame}

\section{Case Studies}

\begin{frame}{}
    If we consider an SMT with a copy and delete generators we get essentially an algebraic theory, the one for which e-graphs are designed

    That is, if we consider the generators (the inputs flow from bottom to top to follow the e-graphs presentation)
    \[
	\scalebox{0.7}{
  	 \tikzfig{../figures/categorical-semantics/Cartesian-equipment}
	}
    \]

    and equations

    \begin{minipage}{0.4\linewidth}
        \[
            \scalebox{0.55}{
            \tikzfig{../figures/categorical-semantics/Cartesian-theory}	
            }
            \]
    \end{minipage}
    \hfill
    \begin{minipage}{0.5\linewidth}
        \[
            \scalebox{0.55}{
            \tikzfig{../figures/categorical-semantics/Cartesian-naturality}
            }
        \]
    \end{minipage}

    We can show that our construction is indeed a generalisation of e-graphs

    The latter gives a categorical presentation of e-graphs and provides a framework for reasoning about them in terms of e-hypergraph rewriting
\end{frame}

\begin{frame}{}
    First, consider how an e-class gets translated into a string diagrammatic form

    \begin{example}
    \[
    \tikzfig{figures/e-graphs-translation-2}
    \]
    \end{example}

    A general translation can be drawn as follows
    \[
    \tikzfig{figures/e-graphs-translation}
    \]
\end{frame}

\begin{frame}{}
    \small
    This way applying equations to an e-graph can be seen as e-hypergraph rewriting

    \[
    \scalebox{0.5}{
    \tikzfig{figures/e-graphs-translation-3}
    }
    \]
    Can be rendered as
    \[
        % \hspace{1.3cm}
        \scalebox{0.38}{
        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
        }
    \]
\end{frame}

\begin{frame}{}
    \vfill
    Lambda calculus is a model for a Cartesian Closed SMT and the benefits of using string diagrams for lambda calculus is that they are automatically quotiented by $\alpha$-equivalence (variables just become wires) and by linear substitution
    \vfill
    Thus we can represent e-graphs for lambda calculus with string diagrams for an enriched Cartesian Closed SMT
\end{frame}

\begin{frame}
    \[
        % \hspace{1.3cm}
        \scalebox{0.4}{
        \tikzfig{../figures/categorical-semantics/e-graph-substitution}
        }
    \]
    \[
        % \hspace{1.3cm}
        \scalebox{0.38}{
        \tikzfig{../figures/categorical-semantics/e-string-substitution-2}
        }
    \]
\end{frame}

\section{Conclusion}

\begin{frame}{}
    \begin{minipage}{0.65\linewidth}
    \small
    \vfill
    \begin{itemize}
        \item We have presented a generalisation of e-graphs for SMTs which could broaden the area of application of e-graphs to
        \vfill
        \begin{itemize}
            \item Quantum processes
            \vfill
            \item Probabilistic programming
            \vfill
            \item Digital circuits
        \end{itemize}
        \vfill
        \item It also has benefits for algebraic theories like lambda calculus
        \vfill
        \item As a bonus we have a framework for reasoning about e-graphs in terms of graph rewriting
        \vfill
        \item We \textbf{did not} study the implementation of e-hypergraphs as a data structure and the corresponding equality saturation procedure
    \end{itemize}
\end{minipage}
\begin{minipage}{0.3\linewidth}
    \small
    \includegraphics[width=\linewidth]{figures/arxiv_qr}
    Link to a pre-print$\;\uparrow$\\

    Email for collaboration:\\
    \alert{\url{axt257@student.bham.ac.uk}}
\end{minipage}
\end{frame}

\end{document}

### 2024-06-23 slides/slides.tex
slides fixes
slides update

@@
 %!LW recipe=xelatex
-
 \documentclass[aspectratio=169]{beamer}
@@
 \usepackage{tikz}
+
 \usepackage{tikz-network}
@@
 \usepackage{minted}
+\usepackage{mathtools}
+
+% \hypersetup{
+%     colorlinks=true,
+%     linkcolor=gray,
+%     filecolor=magenta,      
+%     urlcolor=cyan,
+%     pdftitle={Overleaf Example},
+%     pdfpagemode=FullScreen,
+%     }
 
@@
 
-\usetheme{metropolis}
+\usetheme[subsectionpage=simple]{metropolis}
 % \usetheme{awesome}
@@
 
-% \setbeamertemplate{footline}{\insertshortauthor~(\insertshortinstitute)}
-% \usepackage{custom_footer}
+\makeatletter
+\metropolis@disablesubsectionpage
+\makeatother
+
+\newcommand{\bsubsection}[1]{\subsection{$\bullet$ #1}}
 
@@
 \begin{frame}{Table of contents}
+    \small
     \setbeamertemplate{section in toc}[sections numbered]
-    \tableofcontents[hideallsubsections]
+    \tableofcontents
 \end{frame}
@@
 \section{What and Why?}
+\bsubsection{Algebraic theories}
 
@@
     Traditionally, e-graphs are used for \alert{theories} over \alert{algebraic} signatures with a single sort of composition $()$
-
+    \pause
+    \vfill
     \begin{example}[Commutative monoid]
@@
 
-    However, there are more general theories that describe the interaction of processes, \textit{e.g.}, quantum
+    % However, there are more general theories that describe the interaction of processes, \textit{e.g.}, quantum
 \end{frame}
+
+\bsubsection{Monoidal theories}
 
@@
 
+\pause
 \begin{example}
     \[
-    (\mu\;;\sigma_{1,1}\;;i \otimes id_{1}) \otimes (\mu\;;(id_{1}\otimes id_{1}))
+    (\mu\;;(\sigma_{1,1}\;;i \otimes id_{1})) \otimes (\mu\;;(id_{1}\otimes id_{1}))
     \]
@@
 
-Such signatures also come with \alert{built-in} equations
+Such signatures also come with \alert{built-in} equations (the laws of a symmetric monoidal category)
 
@@
 
-\vspace{-1.5em}
+\end{frame}
+
+\begin{frame}
 
@@
 
-    The equations above are the laws of a (strict) symmetric monoidal category
+    One gets a semantics of an SMT by building a functor for the corresponding \textbf{PROP}, \textit{e.g.} a functor into a category of vector spaces \textbf{$\text{Vect}_{k}$}
+
     \vfill
 
-    One gets a semantics of an SMT by building a functor for the corresponding \textbf{PROP}, \textit{e.g.} a functor into a category of vector spaces \textbf{$\text{Vect}_{k}$}
+    If the functor has nice properties, we can reason about the target category (e.g. \textbf{$\text{Vect}_{k}$}) syntactically
 \end{frame}
@@
 
+\bsubsection{Why e-graphs struggle with monoidal theories}
+
 \begin{frame}[containsverbatim]{}
@@
         ";" = Compose([Id;2]),
-        "tensor" = Tensor([Id;2]),
+        "parallel" = Parallel([Id;2]),
         "id" = Identity,
@@
 \begin{frame}
-    However, the buit-in equations (the laws of monoidal category) lead to an explosion of the number of e-classes
+    However, the laws of monoidal category lead to an explosion of the number of e-classes
+\end{frame}
 
+\begin{frame}
     \begin{example}
@@
         \[
-            (\mu\;;\sigma_{1,1}\;;i \otimes id_{1}) \otimes (\mu\;;(id_{1}\otimes id_{1}))
+            (\mu\;;(\sigma_{1,1}\;;i \otimes id_{1})) \otimes (\mu\;;(id_{1}\otimes id_{1}))
         \]
 
+        Before saturation
+        \vspace{-2em}
+        \begin{figure}
+            \includegraphics[scale=0.4]{figures/egraph_before_saturation.pdf}
+        \end{figure}
+    \end{example}
+\end{frame}
+
+\begin{frame}
+    \begin{example}
+        \vspace{1em}
+        E-graph for
+        \[
+            (\mu\;;(\sigma_{1,1}\;;i \otimes id_{1})) \otimes (\mu\;;(id_{1}\otimes id_{1}))
+        \]
+        After saturation
         \begin{figure}
@@
         \end{example}
+\end{frame}
 
-        \pause
-
+\begin{frame}[standout]
     Can we somehow get rid of these equations?
 \end{frame}
+
+\bsubsection{String diagrams and hypergraphs}
 
@@
 \begin{frame}
-So \alert{yes}, we can get rid of these equations if we use \alert{hypergraphs}
+    \begin{example}
+        \begin{minipage}{0.45\linewidth}
+        Assuming $\mu \coloneqq $\raisebox{-1.5em}{\scalebox{0.6}{\begin{tikzpicture}
+            \begin{pgfonlayer}{nodelayer}
+                \node [style=round box] (0) at (0.5, 0.5) {$\mu$};
+                \node [style=none] (1) at (0.5, 1.5) {};
+                \node [style=none] (2) at (0, -0.5) {};
+                \node [style=none] (3) at (1, -0.5) {};
+                \node [style=none] (4) at (0.25, 0.25) {};
+                \node [style=none] (5) at (0.75, 0.25) {};
+            \end{pgfonlayer}
+            \begin{pgfonlayer}{edgelayer}
+                \draw (0) to (1.center);
+                \draw [bend right=15] (4.center) to (2.center);
+                \draw [bend left=15] (5.center) to (3.center);
+            \end{pgfonlayer}
+        \end{tikzpicture}}}
+        and $i \coloneqq \;\;$\raisebox{-0.2em}{\scalebox{0.6}{\begin{tikzpicture}
+            \begin{pgfonlayer}{nodelayer}
+                \node [style=none] (0) at (3.5, 1) {};
+                \node [style=vertex] (1) at (3.5, 0.5) {};
+            \end{pgfonlayer}
+            \begin{pgfonlayer}{edgelayer}
+                \draw (0.center) to (1);
+            \end{pgfonlayer}
+        \end{tikzpicture}        
+        }} a string diagrammatic version of
+        \[ 
+            (\mu\;;(\sigma_{1,1}\;;i \otimes id_{1})) \otimes (\mu\;;(id_{1}\otimes id_{1}))
+        \] looks like
+    \end{minipage}
+        \hfill
+    \begin{minipage}{0.45\linewidth}
+        \begin{tikzpicture}
+                \begin{pgfonlayer}{nodelayer}
+                    \node [style=round box] (0) at (0.5, 0.5) {$\mu$};
+                    \node [style=none] (1) at (0.5, 1.5) {};
+                    \node [style=none] (2) at (0, -0.25) {};
+                    \node [style=none] (3) at (1, -0.25) {};
+                    \node [style=none] (4) at (0.25, 0.25) {};
+                    \node [style=none] (5) at (0.75, 0.25) {};
+                    \node [style=vertex] (6) at (1, -2.25) {};
+                    \node [style=none] (7) at (1, -1.75) {};
+                    \node [style=round box] (8) at (3.5, 0.5) {$\mu$};
+                    \node [style=none] (9) at (3.5, 1.5) {};
+                    \node [style=none] (10) at (3, -0.25) {};
+                    \node [style=none] (11) at (4, -0.25) {};
+                    \node [style=none] (12) at (3.25, 0.25) {};
+                    \node [style=none] (13) at (3.75, 0.25) {};
+                    \node [style=none] (16) at (0.5, -1) {};
+                    \node [style=none] (17) at (0, -1.75) {};
+                    \node [style=none] (18) at (1, -1.75) {};
+                    \node [style=none] (19) at (0, -2.25) {};
+                    \node [style=none] (20) at (3, -2.25) {};
+                    \node [style=none] (21) at (4, -2.25) {};
+                \end{pgfonlayer}
+                \begin{pgfonlayer}{edgelayer}
+                    \draw (0) to (1.center);
+                    \draw [bend right=15] (4.center) to (2.center);
+                    \draw [bend left=15] (5.center) to (3.center);
+                    \draw (6) to (7.center);
+                    \draw (8) to (9.center);
+                    \draw [bend right=15] (12.center) to (10.center);
+                    \draw [bend left=15] (13.center) to (11.center);
+                    \draw [in=30, out=-90] (3.center) to (16.center);
+                    \draw [in=90, out=-135] (16.center) to (17.center);
+                    \draw [in=315, out=90] (18.center) to (16.center);
+                    \draw [in=-90, out=150] (16.center) to (2.center);
+                    \draw (17.center) to (19.center);
+                    \draw (10.center) to (20.center);
+                    \draw (11.center) to (21.center);
+                \end{pgfonlayer}
+            \end{tikzpicture}
+    \end{minipage}
     \vfill
-The work next formalises e-graphs for SMTs that would use string diagrams (hypergraphs) as a representation for terms
+    Which is already saturated with respect to the built-in equations
+    \end{example} 
+\end{frame}
+
+\begin{frame}[standout]
+So \alert{yes}, we can get rid of these equations if we use \alert{hypergraphs} (derived from string diagrams)
+\end{frame}
+
+\begin{frame}
+\vfill
+The work next formalises e-graphs for SMTs that would use string diagrams (hypergraphs) as a representation for terms in a bottom-up manner deriving this formalisation from an SMT itself
+\vfill
 \end{frame}
@@
     Now equivalence classes represent equivalent subgraphs rather than subtrees
+    \pause
     \vfill
-    Such combinatorial structures are shown to be \textit{sound} and \textit{complete} in terms of representing terms of enriched SMTs
+    \pause
+    Such combinatorial structures are shown to be \textit{sound} and \textit{complete} for representing terms of enriched SMTs
     \vfill
+    \pause
     String diagrammatic reasoning becomes e-hypergraph rewriting
     \vfill
-    The latter is defined as a \textit{double-pushout rewriting} which is a convenient framework to study properties of a rewriting system, e.g. confluence applicability of parallel rewrites etc.
+    \pause
+    The latter is defined as a \textit{double-pushout rewriting} which is a convenient framework to study properties of a rewriting system, e.g. confluence, applicability of parallel rewrites etc.
+    \vfill
+    \pause
+    Note that they absorb the laws of symmetric monoidal category but not the structural rules for $+$
 \end{frame}
@@
 \section{Case Studies}
+
+\bsubsection{(acyclic) E-graphs are e-hypergraphs for Cartesian monoidal theories}
 
@@
     We can show that our construction is indeed a generalisation of e-graphs
-
-    The latter gives a categorical presentation of e-graphs and provides a framework for reasoning about them in terms of e-hypergraph rewriting
 \end{frame}
@@
     }
-    \]
+    \]\footnote{[1]Willsey, M., Nandi, C., Remy Wang, Y., Flatt, O., Tatlock, Z., and Panchekha, P., “egg: Fast and Extensible Equality Saturation”, 2020.}
     Can be rendered as
@@
 
+\begin{frame}
+    \vfill
+    This serves as a sanity check for our construction, provides a categorical presentation of e-graphs and therefore a possible framework for reasoning about them in terms of e-hypergraph rewriting
+    \vfill
+\end{frame}
+
+\bsubsection{Support for binding}
+
 \begin{frame}{}
+    \small
     \vfill
@@
     \vfill
+    Compare the following
+
+    \begin{minipage}{0.3\linewidth}
+        \[
+        \scalebox{0.5}{
+        \tikzfig{figures/e-graphs-binding-example-plain}
+        }
+        \]
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.3\linewidth}
+        \[
+        \scalebox{0.5}{
+        \tikzfig{figures/e-graphs-binding-example-indices}
+        }
+        \]
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.3\linewidth}
+        \[
+        \scalebox{0.5}{
+        \tikzfig{figures/e-graphs-binding-string}
+        }
+        \]
+    \end{minipage}
+    \vfill
+
+\end{frame}
+
+\begin{frame}
+    \vfill
     Thus we can represent e-graphs for lambda calculus with string diagrams for an enriched Cartesian Closed SMT
+    \vfill
 \end{frame}
@@
     \]
+    \tiny{[2]\href{https://pldi23.sigplan.org/details/egraphs-2023-papers/12/Optimizing-Beta-Reduction-in-E-Graphs}{Optimizing-Beta-Reduction-in-E-Graphs}}
     \[
         % \hspace{1.3cm}
-        \scalebox{0.38}{
+        \scalebox{0.35}{
         \tikzfig{../figures/categorical-semantics/e-string-substitution-2}

### 2024-06-24 slides/slides.tex
slides update

@@
 \[
-\Sigma = \{id_n : n \to n,\sigma_{n,m} : n + m \to m + n,\mu : 1 \to 2, i : 1 \to 0\}
+\Sigma = \{id_0 : 0 \to 0, id_{1} : 1 \to 1, \sigma_{1,1} : 1 + 1 \to 1 + 1,\mu : 1 \to 2, i : 1 \to 0\}
 \]
@@
     enum CommutativeComonoids {
-        "mu" = Mu,
-        "i" = I,
+        "mu" = Mu([Id;1]),
+        "i" = I([Id;1]),
         ";" = Compose([Id;2]),
         "parallel" = Parallel([Id;2]),
-        "id" = Identity,
+        "id" = Identity([Id;1]),
         "I" = Unit,
-        "sym" = Sym    
+        "sym" = Sym([Id;2])   
     }
@@
 \begin{frame}
-    However, the laws of monoidal category lead to an explosion of the number of e-classes
+    However, because of the laws of monoidal category this does not scale very well
 \end{frame}
@@
 \begin{frame}
+    \small
     \begin{example}
@@
         Before saturation
-        \vspace{-2em}
+        \vspace{-3em}
         \begin{figure}
@@
         \begin{figure}
-            \includegraphics[width=\linewidth]{figures/dot_5.pdf}
+            \includegraphics[width=0.9\linewidth]{figures/dot_5.jpeg}
         \end{figure}
@@
     \vfill
-    Which is already saturated with respect to the built-in equations
+    Which is already `saturated' with respect to the built-in equations
     \end{example} 
+\end{frame}
+
+\begin{frame}
+
+A hypergraph form of the above example is
+
+\[
+\scalebox{0.6}{
+    \tikzfig{figures/hypergraph_with_interfaces}
+}    
+\]
+
 \end{frame}

### 2024-08-27 LICS/sections/abstract.tex
initial commit with closed monoidal

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-
-\documentclass[sigconf,9pt]{acmart}
-
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
-
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\author{Anonymous}
-%\email{}
-%\orcid{}
-%\affiliation{
-%  \institution{}
-%  \streetaddress{}
-%  \city{}
-%  \state{}
-%  \country{
-%  \postcode{}
-%}
-
-%\renewcommand{\shortauthors}{}
-
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
+The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for satisfiability modulo theory solvers.
+We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a category of semilattices.
 We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
@@
 They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism.
-\end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
-%\received{20 February 2007}
-%\received[revised]{12 March 2009}
-%\received[accepted]{5 June 2009}
-
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
-\maketitle
-
-%-------------------- INTRODUCTION
-
-\section{Introduction}
-
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
-This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
-
-A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
-
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
-
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
-
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
-
-\subsection{E-graphs}
-
-\begin{figure*}
-\[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
-    }
-\]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
-
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
-
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
-
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
-
-\subsection{Semilattice Enriched Categories}
-
-\begin{figure}
-\[
-	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
-	}
-\]
-\captionsetup{belowskip=-5ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
-\end{figure}
-
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
-\[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
-\]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
-
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
-
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
-\[
-    \tikzfig{categorical-semantics/egraph-translation-1}
-\]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
-As it would involve needless additional presentational complexity we leave it as further work.
-
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
-
-\subsection{Combinatorial Representation of Enriched String Diagrams}
-
-In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
-For example, we have the following equivalences between diagrams.
-%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
-	}
-\]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
-Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
-
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
-These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
-
-\subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
-They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
-
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
-
-%\subsection{E-matching and E-rewriting}
-%
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
-In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
-In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
-Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-%\begin{figure}
-%\[
-%	\tikzfig{combinatorial_semantics/example-rewriting}
-%\]
-%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
-REDO THIS}
-%\end{figure}
-
-\subsection{Related Work}
-
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
-We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
-
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
-
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
-\end{figure}
-
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
-
-\begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
-\end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
-
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
-\end{figure}
-
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\end{document}

### 2024-08-27 closed_monoidal/closed_monoidal.tex
initial commit with closed monoidal

%!LW recipe=pdflatex -> bibtex -> pdflatex * 2
\documentclass[acmsmall, screen, nonacm]{acmart}

\newcommand*{\ONECOLUMN}{}

% \input{../preamble}

\usepackage{hyperref}

\usepackage{amsthm}

\usepackage{mathtools}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage{fontenc}

\usepackage{subcaption}
\usepackage{caption}

\usepackage{tikz-cd}
\usepackage{tikzit}
% to reduce tikz picture white space
\usepackage{adjustbox}

% for quotienting
\usepackage{xfrac}

\usepackage{import}

\newcommand\HypI[1]{\textbf{HypI}(#1)}
\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
% conflict
\newcommand{\hashtag}{{\#}}
\newcommand{\consistency}{{\smile}}
% There is already a remark in the preamble, but it does
% not work for some reason
\newcommand\obj[1]{{\textbf{obj}(#1)}}

\usepackage{xifthen}
\ifthenelse{\isundefined{\ismain}}{
  \newcommand{\ismain}{0}
}{}

\usepackage{enumitem}

\newcommand{\ignora}[1]{ }

% \newcounter{thm}

% % \ignora{
% \usepackage{aliascnt}
% % alias-counters
% \newaliascnt{definition}{thm}
% \newaliascnt{proposition}{thm} 
% \newaliascnt{lemma}{thm}
% \newaliascnt{corollary}{thm}
% \newaliascnt{conjecture}{thm}
% \newaliascnt{remark}{thm}
% \newaliascnt{example}{thm}
% \newaliascnt{examples}{thm}
% \newaliascnt{assumption}{thm}
% \newaliascnt{construction}{thm}
% \newaliascnt{claim}{thm}

% % \theoremstyle{plain}
% \theoremstyle{definition}

% \newtheorem{theorem}[thm]{Theorem}
% \newtheorem{proposition}[proposition]{Proposition} 
% \aliascntresetthe{proposition}
% \newtheorem{lemma}[lemma]{Lemma}
% \aliascntresetthe{lemma}
% \newtheorem{corollary}[corollary]{Corollary}
% \aliascntresetthe{corollary}
% \newtheorem{conjecture}[conjecture]{Conjecture}
% \aliascntresetthe{conjecture}
% \newtheorem{remark}[remark]{Remark}
% \aliascntresetthe{remark}
% \newtheorem{assumption}[assumption]{Assumption}
% \aliascntresetthe{assumption}
% \newtheorem{construction}[construction]{Construction}
% \aliascntresetthe{construction}
% \newtheorem{claim}[claim]{Claim}
% \aliascntresetthe{claim}

% \theoremstyle{definition}
% \newtheorem{definition}[definition]{Definition}
% \aliascntresetthe{definition}
% \newtheorem{example}[example]{Example}
% \aliascntresetthe{example}
% \newtheorem{examples}[examples]{Examples}
% \aliascntresetthe{examples}
% }

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}

\usepackage{cleveref} % after hyperref!

\input{../macros}
\input{../sample.tikzstyles}
\input{../hypergraph.tikzstyles}
\input{../hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\newcommand\enriched[1]{{\overline{\mathcal{#1}}}}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
% \setcopyright{acmlicensed}
% \copyrightyear{2024}
% \acmYear{2024}
% \acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.

%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
% \acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\begin{abstract}
  These notes contain the foundations of categories enriched over a category of semilattices and the respective string diagrams for such categories.
  It introduces the required notions and carefully checks that all properties are satisfied.
\end{abstract}

\begin{document}

\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}

\maketitle

\begin{section}{Enriched Categories}

\begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a, \sigma)$ be a closed symmetric monoidal category where $l$ and $r$ are left and right unitors and $\sigma$ is a symmetry.
  We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
  \begin{enumerate}
    \item a class of objects $\obj{\enriched{C}}$
    \item for every pair of objects $A,B \in \obj{\enriched{C}}$ --- an (hom-)object $\obj{\enriched{C}}(A,B) \in \mathcal{V}$
    \item for every triple of objects $A,B,C \in \obj{\enriched{C}}$ --- a composition morphism in $\mathcal{V}$
    \[
    c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)
    \]
    \item for every object $A \in \obj{\enriched{C}}$ --- a unit morphism
    \[
      u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)
    \]
  \end{enumerate}
  such that the following diagrams commute (associativity and unitality)
  \[
  \begin{tikzcd}
    {(\enriched{C}(A,B) \otimes \enriched{C}(B,C)) \otimes \enriched{C}(C,D)} \arrow[rr, "{c_{A,B,C} \otimes 1}"] \arrow[d, "{a_{\enriched{C}(A,B),\enriched{C}(B,C),\enriched{C}(C,D)}}"']   &  & {\enriched{C}(A,C) \otimes \enriched{C}(C,D)} \arrow[dd, "{c_{A,C,D}}"]                                                                                     \\
    {\enriched{C}(A,B) \otimes (\enriched{C}(B,C) \otimes \enriched{C}(C,D))} \arrow[d, "{1\otimes c_{B,C,D}}"']                                             &  &                                                                                                                                       \\
    {\enriched{C}(A,B) \otimes \enriched{C}(B,D)} \arrow[rr, "{c_{A,B,D}}"]                                                                       &  & {\enriched{C}(A,D)}                                                                                                                             
  \end{tikzcd}
  \]

  \[
  \begin{tikzcd}
    {I_{\mathcal{V}} \otimes \enriched{C}(A,B)} \arrow[dd, "u_{A} \otimes 1"'] \arrow[r, "{l_{\enriched{C}(A,B)}}"]                                             & {\enriched{C}(A,B)} \arrow[dd, "{1_{\enriched{C}(A,B)}}"] & {\enriched{C}(A,B) \otimes I_{\mathcal{V}}} \arrow[l, "{r_{\enriched{C}(A,B)}}"'] \arrow[dd, "1 \otimes u_{B}"]                                                           \\
    &                                     &                                                                                                                                       \\
{\enriched{C}(A,A) \otimes \enriched{C}(A,B)} \arrow[r, "{c_{A,A,B}}"]                                                                        & {\enriched{C}(A,B)}                            & {\enriched{C}(A,B) \otimes \enriched{C}(B,B)} \arrow[l, "{c_{A,B,B}}"']                                                                                    
\end{tikzcd}
  \]
\end{definition}

\begin{remark}
  We will occasionally use $c_{A,B,C}$ as $;$, \textit{i.e.}, instead of $c_{A,B,C}(f,g)$ we will write $f ; g$ when it makes sense to talk about components of hom-objects.
\end{remark}

\begin{example}
When $\mathcal{V} = \catname{Set}$ we end up with the usual definition of a category.
$c_{A,B,C}$ becomes the usual morphism composition ($f;g$ rather than $g \circ f$) and $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$ picks out the identity morphism from the homset $\enriched{C}(A,A)$ by matching the only element of a singleton set $I \cong \{*\} \mapsto 1_{A}$ and component-wise the diagrams above commute.
\end{example}

TODO: another example

\begin{definition}[Semilattice]
  A \textit{semilattice} is a set equipped with an operation that we denote as $+$ which is associative, commutative and idempotent.
\end{definition}

Note that we do not require the existence of a unit for $+$. 
Semilattices that satisfy this extra requirement are sometimes called \textit{bounded}, i.e., they are idempotent commutative monoids.

\begin{definition}[Semilattice homomorphism]

A homomorphism between two semilattices $S_{1}$ and $S_{2}$ is a map $h$ that respects $+$.
That is, for all $s,s' \in S_{1}$, $h(s +_{S_{1}} s') = h(s) +_{S_{2}} h(s')$.
\end{definition}

\begin{definition}[Category of semilattices]

Semilattices with their respective homomorphisms form a category that we denote $\catname{SLat}$.
\end{definition}

\begin{proposition}
  $\catname{SLat}$ is a closed symmetric monoidal category.
\end{proposition}
\begin{proof}
  The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
  $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
  \begin{itemize}
    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \cong (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \cong (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
  \end{itemize}

  The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
  Clearly $S \otimes I \cong S$ by mapping $(s,*) \mapsto s$ for all $s \in S$ and vice versa.
  The symmetry, associators and unitors are then obvious morphisms.
  Finally, the category is closed since the set of homomorpisms between two semilattices is a semilattice by defining $(f + g)(x)$ as $f(x) + g(x)$.
  $f + g$ is a homomorphism since $(f + g)(x+y) = f(x+y) + g(x+y) = f(x) + f(y) + g(x) + g(y) = f(x) + g(x) + f(y) + g(y) = f(x+y) + g(x+y)$.
\end{proof}

This makes $\catname{SLat}$ a suitable base for enrichment and allows us to take $\mathcal{V} = \catname{SLat}$.
Before we proceed we need to define what an enriched functor is and define what it means to freely enrich a category over $\catname{SLat}$.

\begin{definition}[Isomorphism in $\enriched{C}$]

Let $\enriched{C}$ be a $\mathcal{V}$-enriched category.
Given two objects $A,B \in \enriched{C}$ and a morphism $f : I_{\mathcal{V}} \to \enriched{C}(A,B)$, we call $f$ an isomorphism if there is a $g : I_{\mathcal{V}} \to \enriched{C}(B,A)$ such that
\[
  I_{\mathcal{V}} \cong I_{\mathcal{V}} \otimes I_{\mathcal{V}} \xrightarrow{f \otimes g} \enriched{C}(A,B) \otimes \enriched{C}(B,A) \xrightarrow{c_{A,B,A}} \enriched{C}(A,A)
\]
equals $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$.
\end{definition}

TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.

\begin{definition}
  Let $\enriched{C}$ and $\enriched{D}$ be two $\mathcal{V}$-categories.
  A $\mathcal{V}$-enriched functor $F : \enriched{C} \to \enriched{D}$ is defined by the following data.
  \begin{enumerate}
    \item A mapping $F : \obj{\enriched{C}} \to \obj{\enriched{D}}$
    \item An object-indexed family of morphisms in $\mathcal{V}$ $F_{A,B} : \enriched{C}(A,B) \to \enriched{D}(FA,FB)$
  \end{enumerate}
  such that the following diagrams commute

  \[
  \begin{tikzcd}
	{\enriched{C}(A,A') \otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} \\
	\\
	{\enriched{D}(FA,FA') \otimes \enriched{D}(FA',FA'')} && {\enriched{D}(FA,FA'')}
	\arrow["{c_{A,A',A''}}"', from=1-1, to=1-3]
	\arrow["{F_{A,A'} \otimes F_{A',A''}}", from=1-1, to=3-1]
	\arrow["{F_{A,A''}}"', from=1-3, to=3-3]
	\arrow["{c_{FA,FA',FA''}}"', from=3-1, to=3-3]
\end{tikzcd}
\]

\[
\begin{tikzcd}
	I_{\mathcal{V}} && {\enriched{C}(A,A)} \\
	\\
	&& {\enriched{D}(FA,FA)}
	\arrow["{u_{A}}", from=1-1, to=1-3]
	\arrow["{u_{FA}}"', from=1-1, to=3-3]
	\arrow["{F_{A,A}}", from=1-3, to=3-3]
\end{tikzcd}
\]
\end{definition}

\begin{definition}
  Given two $\mathcal{V}$-functors $F,G : \enriched{C} \to \enriched{D}$ a $\mathcal{V}$-natural transformation is family of morphisms indexed by objects in $\enriched{C}$, $\alpha_{A} : I_{\mathcal{V}} \to \enriched{D}(FA,GA)$ in $\mathcal{V}$ such that the following diagram commutes
  \[
  \begin{tikzcd}
	& {\enriched{C}(A,A')\otimes I_{\mathcal{V}}} && {\enriched{D}(FA,FA') \otimes \enriched{D}(FA',GA')} \\
	{\enriched{C}(A,A')} &&&& {\enriched{D}(FA,GA')} \\
	& {I_{\mathcal{V}} \otimes \enriched{C}(A,A')} && {\enriched{D}(FA,GA) \otimes \enriched{D}(GA,GA')}
	\arrow["{G_{A,A'}\otimes \alpha_{A'}}"', from=1-2, to=1-4]
	\arrow["{c_{FA,FA',GA'}}"', from=1-4, to=2-5]
	\arrow["{r^{-1}}", from=2-1, to=1-2]
	\arrow["{l^{-1}}", from=2-1, to=3-2]
	\arrow["{\alpha_{A} \otimes F_{A,A'}}", from=3-2, to=3-4]
	\arrow["{c_{FA,GA,GA'}}", from=3-4, to=2-5]
\end{tikzcd}
\]
\end{definition}

\begin{remark}
  When $\mathcal{V}$ is $\catname{Set}$, a family of morphisms $\alpha_{A} : I_{\mathcal{V}} \to \enriched{C}(FA,GA)$ is defined by mapping $* \mapsto \alpha'_{A}$ where $\alpha'_{A} : FA \to GA$ is usual natural transformation.
  We can then trace the diagram to get

\[\begin{tikzcd}
	& {f \otimes *} && {Gf \otimes \alpha'_{A'}} \\
	f &&&& {Gf;\alpha'_{A} = \alpha'_{A};Ff} \\
	& {* \otimes f} && {\alpha'_{A} \otimes Ff}
	\arrow[from=1-2, to=1-4]
	\arrow[from=1-4, to=2-5]
	\arrow[from=2-1, to=1-2]
	\arrow[from=2-1, to=3-2]
	\arrow[from=3-2, to=3-4]
	\arrow[from=3-4, to=2-5]
\end{tikzcd}\]

which is the usual naturality condition.

\end{remark}

\begin{proposition}
  The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
\end{proposition}
\begin{proof}
  The functor is defined by letting $F(A) = \coprod_{A} I$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
  The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
  In particular, $F(I) = I_{\catname{SLat}}$ and therefore such functor is monoidal with $F(A \otimes B) \cong F(A) \otimes F(B)$.
\end{proof}

TODO: add a note that this follows from the definition of a copower.

\begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
  Let $\mathcal{C}$ be a small category.
  Freely enriched $\catname{SLat}$-category $\enriched{C}$ is defined as follows
  \begin{enumerate}
    \item $\obj{\enriched{C}} = \obj{\mathcal{C}}$;
    \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite
    \[
    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \cong F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
    \] 
    where the arrow is the image of the ordinary composition
    \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
    \[
    I_{\catname{SLat}} \cong F(I) \to F(\mathcal{C}(A,A))
    \]
    where the arrow is $* \mapsto 1_{A}$.
  \end{enumerate}
\end{definition}

We will then consider \textit{only} such freely-enriched categories.

\begin{remark}
  Suppose $\mathcal{C}(A,B) = \{f,g\}$ and $\mathcal{C}(B,C) = \{h\}$, then $f + g \in \enriched{C}(A,B)$ and in particular $(f+g);h = f;h + g;h$ since $c_{A,B,C}$ is a semilattice homomorphism.
\end{remark}

\begin{proposition}
\label{prop:functoriality_preserved}
  A functor $G : \mathcal{C} \to \mathcal{D}$ induces an enriched functor $\overline{G} : \enriched{C} \to \enriched{D}$.
\end{proposition}
\begin{proof}
  We define
  \begin{itemize}
    \item $\overline{G}(A) = G(A)$ for all $A \in \enriched{C}$
    \item $\overline{G}_{A,B}(\enriched{C}(A,B)) = F(G_{A,B}(\mathcal{C}(A,B)))$
  \end{itemize}
  The functoriality follows from the following diagrams
\[
\adjustbox{width=\linewidth}{
\begin{tikzcd}
	{\enriched{C}(A,A')\otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} && {(f, g)} && {f;g} \\
	\\
	{\enriched{D}(GA,GA') \otimes \enriched{D}(GA',GA'')} && {\enriched{D}(GA,GA'')} && {(Gf,Gg)} && {Gf;Gg = G(f;g)}
	\arrow[from=1-1, to=1-3]
	\arrow[from=1-1, to=3-1]
	\arrow[from=1-3, to=3-3]
	\arrow[from=1-5, to=1-7]
	\arrow[from=1-5, to=3-5]
	\arrow[from=1-7, to=3-7]
	\arrow[from=3-1, to=3-3]
	\arrow[from=3-5, to=3-7]
\end{tikzcd}}
\]
where the right-hand diagram is a component render of the left-hand diagram.
It commutes because $G$ is a functor. 
By freeness, the commutativity of more complex components follows from the commutativity for generators.
The following diagrams commute for the same reason.
\[\begin{tikzcd}
	I_{\catname{SLat}} && {\enriched{C}(A,A)} && {*} && {1_{A}} \\
	\\
	&& {\enriched{D}(GA,GA)} &&&& {G(1_{A}) = 1_{G(A)}}
	\arrow[from=1-1, to=1-3]
	\arrow[from=1-1, to=3-3]
	\arrow[from=1-3, to=3-3]
	\arrow[from=1-5, to=1-7]
	\arrow[from=1-5, to=3-7]
	\arrow[from=1-7, to=3-7]
\end{tikzcd}\]
\end{proof}

\begin{remark}
  Again, because $\overline{G}_{A,B}$ is a semilattice homomorphism, we have $\overline{G}(f+g) = \overline{G}(f) + \overline{G}(g)$ for any $\catname{SLat}$-enriched functor $\overline{G}$.
\end{remark}

\begin{example}
  Consider a monoidal bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
  It induces an enriched monoidal bifunctor $\overline{\otimes}$ on $\enriched{C}$.
  \begin{enumerate}
    \item $A \overline{\otimes} B \coloneq A \otimes B$;
    \item $\overline{l} : A \overline{\otimes} I \to A \coloneq I \to \enriched{C}(A \overline{\otimes} I, I)$, i.e. $\overline{l}$ picks out $l$ out of $\enriched{C}(A \overline{\otimes} I, I)$;
    \item similarly for $\overline{r}$, $\overline{\alpha}$.
  \end{enumerate}
  $\otimes_{A,B,C,D} : \mathcal{C}(A,B) \times \mathcal{C}(C,D) \to \mathcal{C}(A \otimes C, B \otimes D)$ becomes $\overline{\otimes}_{A,B,C,D} : \enriched{C}(A,B) \times \enriched{C}(C,D) \to \enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D)$.
  In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
\end{example}

\begin{proposition}[Proposition 6.2.8~\cite{Borceux_1994}]
  When $F,G$ are $\mathcal{V}$-valued $\mathcal{V}$-functors, the notion of $\mathcal{V}$-natural transformation reduces to
  $\alpha_{A} : FA \to GA$ and the following diagram
  \[\begin{tikzcd}
	{C(A,B)} && {[FA,FB]} \\
	\\
	{[GA,GB]} && {[FA,GB]}
	\arrow["{F_{A,B}}", from=1-1, to=1-3]
	\arrow["{G_{A,B}}"', from=1-1, to=3-1]
	\arrow["{[1,\alpha_{B}]}", from=1-3, to=3-3]
	\arrow["{[\alpha_{A},1]}"', from=3-1, to=3-3]
\end{tikzcd}\]
\end{proposition}

Note the flip of the domain and co-domain of the bottom arrow.
This is because the functor $[-,-] : \enriched{C}^{\text{op}} \otimes \enriched{C} \to \mathcal{V}$ is contravariant in its first argument.

\begin{definition}[$\mathcal{V}$-enriched adjunction]
  Let $\mathcal{V}$ be a closed symmetric monoidal category and $\enriched{C}$, $\enriched{D}$ two $\mathcal{V}$-categories.
  Two $\mathcal{V}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$ are $\mathcal{V}$-adjoint, $\overline{L}$ left adjoint to $\overline{R}$ and $\overline{R}$ right adjoint to $\overline{L}$, when there exist isomorphisms in $\mathcal{V}$
  \[
  \enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B) \;,
  \]
  which are $\mathcal{V}$-natural in $A$ and $B$.
\end{definition}

Equivalently, one can give a definition of an adjunction in terms of unit and co-unit natural transformations.

\begin{definition}[Section 1.11 in~\cite{Kelly2022BASICCO}]
  A $\mathcal{V}$-adjunction $\eta, \epsilon: \overline{L} \;\reflectbox{$\vdash$}\; \overline{R} : \enriched{C} \to \enriched{D}$ between functors $L : \enriched{D} \to \enriched{C}$ and $\enriched{R} : \enriched{C} \to \enriched{D}$ consists of $\mathcal{V}$-natural $\eta : 1 \to RL$ (the \textit{unit}) and $\epsilon : LR \to 1$ (the \textit{co-unit}) satisfying the triangular equations
  \[
  \overline{R}\epsilon . \eta_{\overline{L}} = 1 \qquad \epsilon_{\overline{L}} . \overline{R} \eta = 1
  \]
\end{definition}
Furthermore we have that the isomorphism $\alpha : \enriched{C}(\overline{L}A,B) \to \enriched{D}(A, \overline{R}B)$ has the form $\enriched{D}(\eta, 1)\overline{R}$:
\[
  \alpha = \enriched{C}(\overline{L}A, B) \xrightarrow{\overline{R}} \enriched{D}(\overline{RL}A, \overline{R}B) \xrightarrow{\enriched{D}(\eta, 1)} \enriched{D}(A, \enriched{R}B)
\]
and $\eta = u_{\overline{L}A};\alpha_{\overline{L}A} = I \xrightarrow{u_{\overline{L}A}}\enriched{C}(\overline{L}A, \overline{L}A) \xrightarrow{\alpha_{\overline{L}A}} \enriched{D}(A,\overline{RL}A)$.
Similarly, $\alpha^{-1} = \enriched{C}(1,\epsilon)\overline{L}$ and $\epsilon = u_{\overline{R}B};\alpha^{-1}_{\overline{R}B}$.

\begin{proposition}
\label{prop:naturality_preserved}
  For a natural transformation $\alpha : T \to S$ for $T,S : \mathcal{C} \to \mathcal{D}$ there is a $\catname{SLat}$-natural transformation $\overline{\alpha} : \overline{T} \to \overline{S}$ where $\overline{T},\overline{S} : \enriched{C} \to \enriched{D}$.
\end{proposition}
\begin{proof}
  The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e. a morphism that picks out the original $\alpha_{A}$.
  We need to check if so defined $\overline{\alpha}$ is $\catname{SLat}$-natural.
  This follows from the following commutativity diagrams
  \[
  \adjustbox{width=\linewidth}{
  \begin{tikzcd}
	&& {\enriched{C}(A,B) \otimes I_{\catname{SLat}}} && {\enriched{D}(\overline{T}A,\overline{T}B) \otimes \enriched{D}(\overline{T}B, \overline{S}B)} &&&&& {f \otimes *} && {\overline{T}f \otimes F\alpha_{B}} \\
	{\enriched{C}(A,B)} &&&&&& {\enriched{D}(\overline{T}A,\overline{S}B)} && f &&&& {F\alpha_{A};\overline{S}f = \overline{T}f ;F\alpha_{B}} \\
	&& {I_{\catname{SLat}} \otimes \enriched{C}(A,B)} && {\enriched{D}(\overline{T}A,\overline{S}A) \otimes \enriched{D}(\overline{S}A,\overline{S}B)} &&&&& {* \otimes f} && {F\alpha_{A} \otimes \overline{S}f}
	\arrow["{\overline{T}_{A,B} \otimes \overline{\alpha_{B}}}", from=1-3, to=1-5]
	\arrow["{c_{\overline{T}A,\overline{T}B,\overline{S}B}}", from=1-5, to=2-7]
	\arrow[from=1-10, to=1-12]
	\arrow[from=1-12, to=2-13]
	\arrow[from=2-1, to=1-3]
	\arrow[from=2-1, to=3-3]
	\arrow[from=2-9, to=1-10]
	\arrow[from=2-9, to=3-10]
	\arrow["{\overline{\alpha_{A}} \otimes \overline{S}(A,B)}", from=3-3, to=3-5]
	\arrow["{c_{\overline{T}A,\overline{S}A,\overline{T}B}}"', from=3-5, to=2-7]
	\arrow[from=3-10, to=3-12]
	\arrow[from=3-12, to=2-13]
\end{tikzcd}}
\]
The diagram on the right is the component-wise rendering of the diagram on the left.
It commutes because it is essentially a lifting of the following diagram (naturality of $\alpha$) using free enrichment functor $F$.

\[
\begin{tikzcd}
	TA && SA && {\overline{T}A} && {\overline{S}A} \\
	\\
	TB && SB && {\overline{T}B} && {\overline{S}B}
	\arrow["{\alpha_{A}}"', from=1-1, to=1-3]
	\arrow["Tf", from=1-1, to=3-1]
	\arrow["Sf"', from=1-3, to=3-3]
	\arrow["{F\alpha_{A}}"', from=1-5, to=1-7]
	\arrow["{\overline{T}f}", from=1-5, to=3-5]
	\arrow["{\overline{S}A}"', from=1-7, to=3-7]
	\arrow["{\alpha_{B}}", from=3-1, to=3-3]
	\arrow["{F\alpha_{B}}", from=3-5, to=3-7]
\end{tikzcd}
\]

The commutativity for more complex elements of a semilattice like $f + g$ follows component-wise.

\end{proof}

\begin{example}

Consider a symmetry natural transformation for a given symmetric monoidal category $\mathcal{C}$ $\sigma_{A,B} : A \otimes B \to B \otimes A$ and its freely $\catname{SLat}$ enriched version $\overline{\sigma_{A,B}} : I_{\catname{SLat}} \to \enriched{C}(A \overline{\otimes} B, B \overline{\otimes} A)$.
$\catname{SLat}$-naturality takes the following form

\[
\adjustbox{width=\linewidth}{
\begin{tikzcd}
	&& {(\enriched{C}(A,B) \times \enriched{C}(C,D)) \otimes I} && {\enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D) \otimes \enriched{C}(B \overline{\otimes} D, D \overline{\otimes} B)} \\
	{\enriched{C}(A,B) \times \enriched{C}(C,D)} &&&&&& {\enriched{C}(A \overline{\otimes} C, D \overline{\otimes} B)} \\
	&& {I \otimes (\enriched{C}(A,B) \times \enriched{C}(C,D))} && {\enriched{C}(A \overline{\otimes} C, C \overline{\otimes} A) \otimes \enriched{C}(C \overline{\otimes} A, D \overline{\otimes} B)} \\
	\\
	&& {(f,g) \otimes *} && {(f \overline{\otimes} g) \otimes \sigma} \\
	{(f,g)} &&&&&& {\sigma;(g \overline{\otimes}f) = (f\overline{\otimes}g);\sigma} \\
	&& {* \otimes (f,g)} && {\sigma \otimes (g \overline{\otimes} f)}
	\arrow["{\overline{\otimes} \otimes \overline{\sigma}_{B,D}}", from=1-3, to=1-5]
	\arrow[from=1-5, to=2-7]
	\arrow[from=2-1, to=1-3]
	\arrow[from=2-1, to=3-3]
	\arrow["{\overline{\sigma}_{C,A} \otimes \overline{\otimes}}"', from=3-3, to=3-5]
	\arrow[from=3-5, to=2-7]
	\arrow[from=5-3, to=5-5]
	\arrow[from=5-5, to=6-7]
	\arrow[from=6-1, to=5-3]
	\arrow[from=6-1, to=7-3]
	\arrow[from=7-3, to=7-5]
	\arrow[from=7-5, to=6-7]
\end{tikzcd}}\]

Given $(f_1 + f_2, g)$ we have $((f_1 + f_2) \;\overline{\otimes}\; g);\sigma = (f_1 \;\overline{\otimes}\; g) + (f_{2} \;\overline{\otimes}\; g);\sigma = (f_{1} \;\overline{\otimes}\; g);\sigma + (f_{2} \;\overline{\otimes}\; g);\sigma = \sigma;(g \;\overline{\otimes}\; f_{1}) + \sigma;(g \;\overline{\otimes}\; f_{2})$

\end{example}

\begin{proposition}

Functor $F: \mathcal{Cat} \to \catname{SLat}-\mathcal{Cat}$ is a 2-functor.
\end{proposition}
\begin{proof}
  This is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
\end{proof}

The above proposition means, in particular, that an ordinary adjunction gives rise to a corresponding $\catname{SLat}$-enriched adjuntion.
The latter, however, could be also shown Equivalently as below.

\begin{proposition}
  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$.
  Because $\overline{L}$ acts on objects as $L$ we will further omit the bar.
\end{proposition}
\begin{proof}
  The first adjunction implies the existence of natural isomorphisms (in $A$ and $B$) between $\mathcal{C}(L-,B)$ and $\mathcal{D}(-,RB)$ and $\mathcal{C}(LA,-)$ and $\mathcal{D}(A,R-)$ respectively.
  Naturality in $B$ is by definition the commutativity of the following diagram
  \begin{equation}
    \label{diag:naturality_plain}
  \begin{tikzcd}
	{\mathcal{C}(LA,B)} && {\mathcal{D}(A,RB)} \\
	\\
	{\mathcal{C}(LA,B')} && {\mathcal{D}(A,RB')}
	\arrow["{\alpha_{B}}"', from=1-1, to=1-3]
	\arrow["{\mathcal{C}(LA,f)}"', from=1-1, to=3-1]
	\arrow["{\mathcal{D}(A,Rf)}", from=1-3, to=3-3]
	\arrow["{\alpha_{B'}}", from=3-1, to=3-3]
  \end{tikzcd}
\end{equation}
  and similarly for the naturality in $A$.
  The claim is that $\enriched{C}(LA,B) \cong \enriched{D}(A,RB)$ is an $\catname{SLat}$-adjunction.
  Recall that $\enriched{C}(LA,B) = F(\mathcal{C}(LA,B))$ and similarly for another hom-set and clearly the isomorphism is preserved: given $\alpha$, an isomorphism from $\mathcal{C}(LA,B) \to \mathcal{D}(A,RB)$, the isomorphism from $\enriched{C}(LA,B) \to \enriched{D}(A,RB)$ is $\enriched{\alpha} = F(\alpha)$.
  We need to check that this $\enriched{\alpha}$ is natural.
  This is equivalent to checking if the following diagram commutes
  \[\begin{tikzcd}
	{\enriched{C}(B,B')} && {[\enriched{C}(LA,B),\enriched{C}(LA,B')]} \\
	\\
	{[\enriched{D}(A,RB),\enriched{D}(A,RB')]} && {[\enriched{C}(A,RB),\enriched{D}(A,RB')]}
	\arrow["{\enriched{C}(LA,-)}", from=1-1, to=1-3]
	\arrow["{\enriched{D}(A,R-)}"', from=1-1, to=3-1]
	\arrow["{[1,\overline{\alpha}_{B'}]}", from=1-3, to=3-3]
	\arrow["{[\overline{\alpha_{B}},1]}"', from=3-1, to=3-3]
\end{tikzcd}\]
which component-wise becomes

\[
\begin{tikzcd}
	f & {\enriched{C}(LA,f)} \\
	{\enriched{D}(A,Rf)} & {\overline{\alpha}_{B};\enriched{D}(A,Rf);1 = 1;\enriched{C}(LA,f);\overline{\alpha}_{B}'}
	\arrow[from=1-1, to=1-2]
	\arrow[from=1-1, to=2-1]
	\arrow[from=1-2, to=2-2]
	\arrow[from=2-1, to=2-2]
\end{tikzcd}
\]

The commutativity of the diagram above follows from applying $F$ to the diagram \ref{diag:naturality_plain}:
\[\begin{tikzcd}
	{\enriched{C}(LA,B)} && {\enriched{D}(A,RB)} \\
	\\
	{\enriched{C}(LA,B')} && {\enriched{D}(A,RB')}
	\arrow["{\overline{\alpha}_{B}}", from=1-1, to=1-3]
	\arrow["{\enriched{C}(LA,f)}"', from=1-1, to=3-1]
	\arrow["{\enriched{D}(A,Rf)}", from=1-3, to=3-3]
	\arrow["{\overline{\alpha}_{B'}}", from=3-1, to=3-3]
\end{tikzcd}
\]
The commutativity for more complex elements of $\enriched{C}(B,B')$, \textit{e.g.} $f + g$, follows component-wise.
Similarly for naturality in $A$.
Then, $\catname{Slat}$-naturality separately in $A$ and $B$ implies $\catname{SLat}$-naturality in both $A$ and $B$, a consequence of a more general statement about $\mathcal{V}$-naturality ((1.21) in~\cite{Kelly2022BASICCO})

\end{proof}

\end{section}

\begin{section}{String diagrams for freely $\catname{SLat}$-enriched symmetric monoidal categories}

String diagrams are a two-dimensional syntax for describing morphisms in symmetric monoidal categories which is typically given by the following grammar parameterised over a monoidal signature $\Sigma$ of symbols.
The labels on wires specify the types of the morphisms.

\[
\tikzfig{./figures/string_diagrams}
\]

The diagrams constructed by using the rules above are then subject to SMC equations.

By freely enriching home-sets to hom-objects in $\catname{SLat}$ we still can access the morphisms within these objects.
To support `new' morphisms we extend the grammar by adding a production for $f + g$

\[
\adjustbox{scale=0.8}{
\tikzfig{./figures/f_plus_g}
}
\]

Once we consider string diagrams for a $\catname{SLat}$-enriched category $\enriched{C}$, $;$ and $\otimes$ become $c_{A,B,C}$ and $\overline{\otimes}$ respectively.
Since the latter are semilattice homomorphisms, we have the following equations (with labels on wires omitted) for the newly introduced dashed boxes.

\[
\adjustbox{width=\linewidth}{
\tikzfig{./figures/egraph-strings-equations}
}
\]

The fact of $+$ being associative and commutative justifies the usage of $\ldots$ in the equations.
\end{section}

TODO: string diagrams for functorial boxes and natural transformations (unit-counit ones).

% \[
% \adjustbox{scale=0.95, lap={\width}{-12em}}{
%     \begin{tikzcd}
%         & {\enriched{C}(A_1,A_2)} \arrow[ld] \arrow[rd]                       &                                                                                                                                                                                                            \\
% {I \otimes \enriched{C}(A_1,A_2)} \arrow[dd, "{\alpha'_{A_1} \otimes \overline{\mathcal{D}}(\overline{L}A,-)}"]                                                                    &                                                                               & {\enriched{C}(A_1,A_2) \otimes I} \arrow[dd, "{\enriched{C}(\overline{L}A,-) \otimes \alpha'_{A_2}}"]                                                                                  \\
%         &                                                                               &                                                                                                                                                                                                            \\
% {\mathcal{V}(\enriched{C}(\overline{L}A,A_1),\overline{\mathcal{D}}(A,A_1)) \otimes \mathcal{V}(\overline{\mathcal{D}}(A,A_1),\overline{\mathcal{D}}(A,A_2)))} \arrow[rd, "\circ"] &                                                                               & {\mathcal{V}(\enriched{C}(\overline{L}A,A_1),\enriched{C}(\overline{L}A,A_2))) \otimes \mathcal{V}(\enriched{C}(LA,A_2),\overline{\mathcal{D}}(A,A_2))} \arrow[ld, "\circ"'] \\
%         & {\mathcal{V}(\enriched{C}(\overline{L}A,A_1), \overline{D}(A,A_2))} &                                                                                                                                                                                                           
% \end{tikzcd}
% }
% \]

% \[
%   \begin{tikzcd}
%     {\enriched{C}(B,B')} \arrow[rr, "{\enriched{C}(LA,-)}"] \arrow[d, "{\overline{\mathcal{D}}(A,R-)}"] &  & {[\enriched{C}(LA,B), \enriched{C}(LA,B')]} \arrow[d, "{[1,\overline{\alpha_{B'}}]}"] \\
%     {[\overline{\mathcal{D}}(A,RB),\overline{\mathcal{D}}(A,RB')]} \arrow[rr, "{[\overline{\alpha_{B}},1]}"]                &  & {[\enriched{C}(LA,B),\overline{\mathcal{D}}(A,RB')]}                                           
%     \end{tikzcd}
% \]

% \vspace{4em}

% \[
%   \begin{tikzcd}
%     {\enriched{C}(B,B')} \arrow[rr, "{\enriched{C}(LA,-)}"] \arrow[d, "{\overline{\mathcal{D}}(A,R-)}"] &  & {[\enriched{C}(LA,B), \enriched{C}(LA,B')]} \arrow[d, "{[1,\overline{\alpha_{B'}}]}"]                             \\
%     {[\overline{\mathcal{D}}(A,RB),\overline{\mathcal{D}}(A,RB')]} \arrow[rr, "{[\overline{\alpha_{B}},1]}"]                &  & {[\enriched{C}(LA,B),\overline{\mathcal{D}}(A,RB')]}                                                                        \\
%     f \arrow[rr] \arrow[d]                                                                                                  &  & {\enriched{C}(LA,f)} \arrow[d]                                                                                              \\
%     {\overline{\mathcal{D}}(A,Rf)} \arrow[rr]                                                                               &  & {1;\enriched{C}(LA,f);\overline{\alpha_{B'}} = \overline{\alpha_{B}};\overline{\mathcal{D}}(A,Rf);1}
%     \end{tikzcd}
% \]

\bibliographystyle{acm}
\bibliography{bibliography}

\end{document}

### 2024-08-27 LICS/sections/appendix.tex
added missing github action

@@
 
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
+\begin{theorem}[Existence of pushouts in $\catname{EHyp}(\Sigma)$]
 \label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
+Consider the following span in $\catname{EHyp}(\Sigma)$
 % https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
@@
 \]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
+and there is an obvious surjective function $[]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
+There is also $[] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
 We then define 
 \[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
+    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[]^{*}(e) = [s_{X+Y}(e)]^{*}
 \]
@@
 
-% So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
+% So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
 
@@
 
+\begin{remark}
+    If in a span $Y \xleftarrow{g} Z \xrightarrow{f} X$ of morphisms in $\catname{EHyp}(\Sigma)$ the morphisms $f$ and $g$ are monos, then the arrows $f'$ and $g'$ of the pushout $Y \xrightarrow{g'} X +_{f,g} Y \xleftarrow{f'} X$ are also monos.
+\end{remark}
+This is because $[]_{V}$ identifies two vertices $v_1$ and $v_2$ only if there exists a vertex $z$ in $V_{Z}$ such that $f_{V};\iota_{1,V}(z) = v_1$ and $g_{V};\iota_{2,V}(z) = v_2$ and therefore $f' = \iota_{1};[]$ and $g' = \iota_{2};[]$ are monos.
+
 \begin{figure}
@@
 \captionsetup{belowskip=-3ex}
-\caption{Pushout in $\catname{EHyp_{\Sigma}}$}
+\caption{Pushout in $\catname{EHyp}(\Sigma)$}
 \label{fig:pushout_example}
@@
 
-We will then discuss the uniqueness of the corresponding pushout complement.
+We will then discuss the uniqueness of the corresponding boundary pushout complement.
 
@@
 
-\begin{proposition}
+\begin{proposition}[Proposition~\ref{prop:boundary_unique} restatement]
 \label{prop:boundary_complement}
@@
 \begin{proof}
-        The pushout in~\ref{def:boundary_new} satisfies the assumptions in Definition~\ref{pushout:assumptions} (note that since $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j$ is an MDA cospan the image of $i + j$ in $\mathcal{L}$ consists of top-level vertices only) and therefore is isomorphic to the pushout constructed in~\ref{pushout:assumptions} and by construction should also be a pushout on the underlying sets of vertices and hyperedges.
+        The pushout in~\ref{def:boundary_new} satisfies the assumptions in Definition~\ref{pushout:assumptions} (note that since $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j$ is an MDA cospan, the image of $i + j$ in $\mathcal{L}$ consists of top-level vertices only) and therefore is isomorphic to the pushout constructed in~\ref{pushout:assumptions} and by construction should also be a pushout on the underlying sets of vertices and hyperedges.
     % I think the above sufficies, but it can also be shown formally that every pushout in $\catname{EHyp_{\Sigma}}$ is a pushout on the underlying sets of edges and vertices by constructing an e-hyperghraph from the set of vertices and same for the set of edges and checking if the morphisms are well-defined.
@@
     to distinguish it from the coproduct.
+\end{remark}
+
+\begin{remark}
+    Below we will also occasionally conflate the cospan resulting from $\llbracket - \rrbracket$ with its carrier which will be clear from the context.
+    For example, if $\llbracket f \rrbracket = n \to n' \to \mathcal{F} \xleftarrow{} m' \xleftarrow{} m$, then we will also use $\llbracket f \rrbracket$ to denote $\mathcal{F}$.
 \end{remark}
@@
             n_2 + m_1 \xrightarrow{} n_2' \setminus (n_1' \setminus n_1) + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot} \xleftarrow{} m'_2 \setminus (m_1' \setminus m_1) + n_1 \xleftarrow{} m_2 + n_1    
-            \] is an MDA e-hypergraph.
+            \] is an MDA cospan.
             By denoting the pushout of $\mathcal{L}_1^{\bot} \xleftarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot}$ with $\mathcal{L}_3^{\bot}$,
-            one can observe that $\llbracket f \rrbracket$ is the pushout object of $\mathcal{L} \xleftarrow{} i+j \xrightarrow{} \mathcal{L}$ as per the diagram below
+            one can observe that $\llbracket f \rrbracket$ is the pushout object of $\mathcal{L} \xleftarrow{} i+j \xrightarrow{} \mathcal{L}_{3}^{\bot}$ as per the diagram below
             \[\begin{tikzcd}
@@
     \end{itemize}
-    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans(\Sigma)/\mathcal{S}$.
+    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
     % there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
-    % As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
+    % As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[]$ is defined by computing the pushout along the common interfaces.
     % We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
@@
 
+\begin{lemma}
+\label{lemma:decomposition}
+    Let $n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m$ be an MDA extended cospan.
+    If $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $\mathcal{G}$ such that the immediate predecessors of top-level connected components of $\mathcal{L}$ in $\mathcal{G}$ are either undefined or coincide and pair-wise consistent then either
+    \begin{itemize}
+    \item there exists a cospan $i \to i' \to \mathcal{G}' \xleftarrow{} j' \xleftarrow{} j$ and $k \in \mathbb{N}$ such that $\mathcal{G}$ is decomposable as
+    \begin{align}
+        \label{fig:decomposition:1}
+    \scalebox{0.6}{
+        \tikzfig{../figures/appendix/G_decomposition_1}
+    }
+    \end{align}
+    \item or, $\mathcal{G}$ is decomposable as
+    \begin{align}
+        \label{fig:decomposition:2}
+    \scalebox{0.6}{
+        \tikzfig{../figures/appendix/G_decomposition_2}
+    }
+    \end{align}
+    \end{itemize}
+    such that $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $\mathcal{G}'$ and all the cospans are MDA.
+\end{lemma}
+\begin{proof}
+    Suppose the image of top-level components of $\mathcal{L}$ consists of top-level edges and vertices.
+    Then let $\mathcal{C}_1$ be the smallest e-hypergraph consisting of the input vertices of $\mathcal{G}$ and every hyperedge $h$ (and its successors) that is not in $\mathcal{L}$ but has a path to it.
+    Let $\mathcal{C}_{2}$ be the smallest e-hypergraph such that $\mathcal{G} = \mathcal{C}_{1} \cup \mathcal{L} \cup \mathcal{C}_{2}$.
+    These e-hypergraphs overlap only on vertices so we define
+    \begin{align*}
+        i = V_{\mathcal{C}_{1}} \cap V_{\mathcal{L}}\\
+        j = V_{\mathcal{C}_{2}} \cap V_{\mathcal{L}}\\
+        k = (V_{\mathcal{C}_{1}} \cap V_{\mathcal{C}_{2}}) \setminus V_{\mathcal{L}}
+    \end{align*}
+    and let $i'$ be the union of $i$ and the vertices of $\mathcal{L}$ in the image of $f'$ and $j'$ be the union of $j$ and the vertices of $\mathcal{L}$ in the image of $g'$.
+    Let $n_{1}'$ be pre-image of the co-restriction of $f'$ to $\mathcal{C}_{1}$ and $n_{2}'$ be the pre-image of the co-restriction of $f'$ to $\mathcal{C}_{2}$.
+    Similarly, for $m_{1}'$ and $m_{2}'$ with regard to $g'$.
+    Then there are cospans
+    \begin{align*}
+        n \to n_{1}' \to &\mathcal{C}_{1} \xleftarrow{} m_{1}' + k + i \xleftarrow{} k + i\\
+        i \to i' \to &\mathcal{L} \xleftarrow{} j' \xleftarrow{} j\\
+        k + j \to n_{2}' + k + j \to &\mathcal{C}_{2} \xleftarrow{} m_{2}' \xleftarrow{} m\\
+    \end{align*}
+    and $n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m$ is the co-limit of the following diagram
+    \[
+        n \to n_{1}' \to \mathcal{C}_{1} \xleftarrow{} m_{1}' + k \xleftarrow{} i + k \to i' + k \to \mathcal{L} \xleftarrow{} k + j' \xleftarrow{} k + j \xrightarrow{} n_{2}' + k + j \to \mathcal{C}_{2} \xleftarrow{} m_{2}' \xleftarrow{} m
+    \]
+    The two spans identify precisely those nodes from $\mathcal{G}$ that occur in more than one sub-hypergraph, so this amounts to simply taking the union
+    \[
+    n \to n_{1}' \cup n_{2}' \cup i' \to \mathcal{C}_{1} \cup \mathcal{C}_{2} \cup \mathcal{L} \xleftarrow{} m_{1}' \cup m_{2}' \cup j' \xleftarrow{} m  = n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m
+    \]
+    If we let $\mathcal{G}' = \mathcal{L}$ we get the decomposition as in~(\ref{fig:decomposition:1}).
+    This follows the construction of a similar lemma in~\cite{bonchi_string_2022-1} (Lemma 24). 
+    The cospans with $\mathcal{C}_{1}$ and $\mathcal{C}_{2}$ are MDAs because they are closed under successors and predecessors and the cospan with $\mathcal{L}$ is an MDA because it is convex.
+
+    Now suppose that the image of top-level components of $\mathcal{L}$ consist of non top-level edges and vertices $e$ and $v$.
+    Take a hyperedge $h$ such that $h \leq e$ and for all $e' \not = h$ such that $e' \leq e$, $h \leq e'$.
+    Let $\mathcal{G}''$ be $h$ including all its sources and targets and successors.
+    If $\mathcal{G} = \mathcal{G}''$ then it is decomposable as in~(\ref{fig:decomposition:2}) and let $\mathcal{G}'$ be the component that contains the image of $\mathcal{L}$ which exists by pair-wise consistency.
+    All $\mathcal{G}_{i}$ (and their respective cospans) are MDA by definitions.
+    Otherwise, if $\mathcal{G} \not = \mathcal{G}''$ then $\mathcal{G}''$ is a convex down-closed sub-e-hypergraph of $\mathcal{G}$ and 
+    by applying the same construction as above we get~(\ref{fig:decomposition:1}).
+\end{proof}
+
 % \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]

### 2024-08-27 LICS/sections/conclusion.tex
added missing github action

@@
 In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
+Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string}, as sketched in Fig.
 \ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.

### 2024-08-27 LICS/sections/e-hypergraphs.tex
added missing github action

@@
     \[
-    \scalebox{0.6}{
+    \scalebox{0.5}{
         \tikzfig{../figures/combinatorial_semantics/extended_cospan_example}
@@
     \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
+    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended MDA cospan
     \[
@@
     \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
+    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended MDA cospan
     \[
@@
 Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
+In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding MDA cospans above.
 The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
@@
 \end{remark}
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
 
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
+\begin{proposition}
+\label{prop:boundary_unique}
+    The boundary complement in~\ref{def:boundary_new} when exists is unique.
+\end{proposition}
 

### 2024-08-27 LICS/sections/introduction.tex
added missing github action

@@
 This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
+\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica-zanassi2023string}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
 However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
@@
 Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
+There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica-zanassi2023string} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
 % Hierarchical hypergraphs
@@
 We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
+It is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
 Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.

### 2024-09-09 closed_monoidal/closed_monoidal.tex
some updates

@@
 \begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
-  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a, \sigma)$ be a closed symmetric monoidal category where $l$ and $r$ are left and right unitors and $\sigma$ is a symmetry.
+  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a)$ be a monoidal category where $l$ and $r$ are left and right unitors.
   We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
@@
 
+\begin{remark}
+
+Requiring that a category $\mathcal{V}$ is monoidal is the most basic requirement for enriched category theory.
+To make the notion of enriched category more useful, in particular to be able to talk about a free enrichment, we will further require that $\mathcal{V}$ is a closed symmetric monoidal category.
+\end{remark}
+
 \begin{example}
@@
   \begin{itemize}
-    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \cong (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
-    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \cong (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
+    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
+    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
   \end{itemize}
@@
 
-\begin{definition}[Isomorphism in $\enriched{C}$]
+\begin{definition}[Isomorphism in $\enriched{C}$. Lemma 3.5.12~\cite{Riehl_2014}]
 
@@
 
-TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
-Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.
+% TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
+% Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.
 
@@
 
+\begin{definition}
+Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
+\end{definition}
+
+Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.
+
 \begin{proposition}
@@
 \begin{proof}
-  The functor is defined by letting $F(A) = \coprod_{A} I$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
+  The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
   The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
-  In particular, $F(I) = I_{\catname{SLat}}$ and therefore such functor is monoidal with $F(A \otimes B) \cong F(A) \otimes F(B)$.
+  The adjunction is then given by the following natural isomorphisms
+  \[
+  \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
+  \]
+  The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
+  Furthermore, we have, 
+  \[
+  F(I) \cong I_{\catname{SLat}}
+  \]
+  and 
+  \begin{align*}
+  F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
+        &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
+        &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
+        &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
+        &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
+        &\cong F(A \times B)
+  \end{align*}
+In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
 \end{proof}
 
-TODO: add a note that this follows from the definition of a copower.
+\begin{remark}
+  The above proposition in fact works for any symmetric monoidal closed category $\mathcal{V}$ picked as a base for enrichment (Proposition 6.4.6~\cite{Borceux_1994}) 
+\end{remark}
+
+\begin{remark}
+
+  Free semilattice on a set $S$ can be though of as a set of non-empty subsets of $S$.
+\end{remark}
 
 \begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
+\label{def:free_slat}
   Let $\mathcal{C}$ be a small category.
@@
     \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
-    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite
+    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite $\cong ; F(;)$ below
     \[
-    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \cong F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
+    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \xrightarrow{\cong} F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
     \] 
-    where the arrow is the image of the ordinary composition
+    where the rightmost arrow is the image of the ordinary composition;
     \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
     \[
-    I_{\catname{SLat}} \cong F(I) \to F(\mathcal{C}(A,A))
+    I_{\catname{SLat}} \xrightarrow{\cong} F(I) \to F(\mathcal{C}(A,A))
     \]
-    where the arrow is $* \mapsto 1_{A}$.
+    where the rightmost arrow is $* \mapsto 1_{A}$.
   \end{enumerate}
 \end{definition}
+
+This is again a specialisation of a more general definition of a free $\mathcal{V}$-category (Proposition 6.4.7~\cite{Borceux_1994}).
 
@@
 \begin{example}
-  Consider a monoidal bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
-  It induces an enriched monoidal bifunctor $\overline{\otimes}$ on $\enriched{C}$.
+  Let $\mathcal{C} = (\mathcal{C}_{0},I,\otimes, l, r, \alpha)$ be a monoidal category and consider the bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
+  It induces an enriched monoidal category $\enriched{C} = (\enriched{C}_{0}, I, \enriched{\otimes}, \overline{l}, \overline{r}, \enriched{\alpha})$ such that
   \begin{enumerate}
@@
 \begin{proof}
-  The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e.
+  Recall that an ordinary natural transformation is in fact a $\catname{Set}$-enriched natural transformation and has the form $\alpha_{A} : I \to \mathcal{D}(TA,SA)$.
+  We then obtain an $\catname{SLat}$-natural transformation $\overline{\alpha_{A}} = F(\alpha_{A}) : F(I) \to F(\mathcal{D}(TA,SA))$, or $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(TA,SA)$ given as $* \mapsto F\alpha_{A}$.
+  Again, we can use $TA$ or $\overline{T}A$ interchangeably as their action on objects is the same.
+  % The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e.
 a morphism that picks out the original $\alpha_{A}$.
@@
 \begin{proposition}
-
-Functor $F: \mathcal{Cat} \to \catname{SLat}-\mathcal{Cat}$ is a 2-functor.
+Functor $F : \catname{Set} \to \catname{SLat}$ induces a 2-functor $F^{2}: \catname{Cat} \to \catname{SLat}-\catname{Cat}$ from the category of small categories to the category of small categories enriched over a category of semilattices.
 \end{proposition}
 \begin{proof}
-  This is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
+  The functoriality of $F^{2}$ is given by Definition~\ref{def:free_slat} and 2-functoriality is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
 \end{proof}
@@
 The above proposition means, in particular, that an ordinary adjunction gives rise to a corresponding $\catname{SLat}$-enriched adjuntion.
-The latter, however, could be also shown Equivalently as below.
+The latter, however, could be also shown equivalently as below.
 
 \begin{proposition}
-  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$.
+  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors $L : \mathcal{D} \to \mathcal{C}$ and $R : \mathcal{C} \to \mathcal{D}$, we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$ of $\catname{SLat}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$.
   Because $\overline{L}$ acts on objects as $L$ we will further omit the bar.
@@
 \[
+\adjustbox{scale=0.7}{
 \tikzfig{./figures/string_diagrams}
+}
 \]
@@
 The fact of $+$ being associative and commutative justifies the usage of $\ldots$ in the equations.
+
+Then we can reuse the diagrams for functorial boxes~\cite{GhicaZanassiStringLambda}.
+Given a functor $F : \mathcal{C} \to \mathcal{D}$ and a morphism $f \in C(A,B)$ we can depict the application of $Ff$ as below (left) together with the functoriality equations in Figure~\ref{fig:functoriality_laws}
+
+\[
+  \adjustbox{scale=0.7}{
+  \tikzfig{./figures/functorial_string}
+  }
+\]
+
+\begin{figure}
+  \begin{subfigure}[c]{0.4\linewidth}
+    \[
+    \adjustbox{scale=0.7}{
+    \tikzfig{./figures/functoriality_eq_1_string}
+    }
+    \]
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.4\linewidth}
+    \[
+    \adjustbox{scale=0.7}{
+    \tikzfig{./figures/functoriality_eq_2_string}
+    }
+    \]
+  \end{subfigure}
+  \caption{String diagrammatic equations for functoriality}
+  \label{fig:functoriality_laws}
+\end{figure}
 \end{section}
 
-TODO: string diagrams for functorial boxes and natural transformations (unit-counit ones).
+A functor $\overline{F}$ induced by the functor $F$ has the same diagrammatic rendering with the addition of an extra equation.
+
+\[
+\adjustbox{scale=0.7}{
+\tikzfig{./figures/functoriality_enriched_eq}
+}
+\]
+
+Since a component of a natural transformation $\overline{\alpha}$ between two $\catname{SLat}$-enriched functors $\overline{F} : \enriched{C} \to \enriched{D}$ and $\overline{G}: \enriched{C} \to \enriched{D}$ has the form $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(FA, GA)$, it can be identified with the element of $\enriched{D}(FA,GA)$, i.e.
+we can use the syntax for plain morphism to depict components of natural transformations.
+This allows us to reuse the whole syntax machinery for Cartesian closed categories for their free $\catname{SLat}$-enriched version which induce an additional equation
+
+\[
+\adjustbox{scale=0.7}{
+  \tikzfig{./figures/lambda_enriched_eq}
+}
+\]
+
+We can similarly lift a trace from a traced monoidal category $\mathcal{C}$ to $\enriched{C}$.
+Recall that a trace is an operator
+\[
+\text{tr}^{X} : \mathcal{C}(A \otimes X, B \otimes X) \to \mathcal{C}(A,B)
+\]
+\textit{i.e.}, a mapping between hom-sets, such that it satisfies the equations
+\begin{enumerate}
+  \item \textit{Tightening} (naturality in $A,B$) For all $A, B, C, D, X \in \mathcal{C}$, $g : A \to B$, $f : B \otimes X \to C \otimes X$, $h : C \to D$
+  \[
+  \text{tr}^{X}((h \otimes id_{X});f;(g \otimes id_{X})) = h;\text{tr}^{X}(f);g
+  \]
+  \item \textit{Sliding} (naturality in $X$) For all $A,B,X,Y \in \mathcal{C}$, $f : A \otimes X \to B \otimes Y$, $g : Y \to X$
+  \[
+   \text{tr}^{X}(f;(id_{B} \otimes g)) = \text{tr}^{Y}((id_{A} \otimes g); f) 
+  \]
+  \item \textit{Vanishing} For all $A,B,X,Y \in \mathcal{C}$,
+  \[
+  \text{tr}^{I}(f) = f,\; \forall f : A \to B \qquad \text{and} \qquad \text{tr}^{X \otimes Y}(f) = \text{tr}^{X}(\text{tr}^{Y}(f)), \; \forall f : A \otimes X \otimes Y \to B \otimes X \otimes Y
+  \]
+  \item \textit{Strength} For all $A,B,C,D,X \in \mathcal{C}$, $f : C \otimes X \to D \otimes X$, $g : A \to B$
+    \[
+    \text{tr}^{X}(g \otimes f) = g \otimes \text{tr}^{X}(f)
+    \] 
+  \item \textit{Yanking} (when $\mathcal{C}$ is symmetric)
+    \[
+    \text{tr}^{X}(\sigma_{X,X}) = id_{X}
+    \]
+\end{enumerate}
+
+By freely enriching $\mathcal{C}$ we obtain an enriched trace
+\[
+\text{tr}^{X} : \enriched{C}(A \overline{\otimes} X, B \overline{\otimes} X) \to \enriched{C}(A,B)
+\]
+and all the equations are preserved with the addition of a new equation
+\[
+  \text{tr}^{X}(f + g) = \text{tr}^{X}(f) + \text{tr}^{X}(g)
+\]
+for $f,g : A \overline{\otimes} X \to B \overline{\otimes} X$
+TODO: trace?
 

### 2024-09-17 closed_monoidal/closed_monoidal.tex
updated slides and closed monoidal notes

@@
   In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
+  \question{I need to show that $\overline{\otimes}$ is a bifunctor.
+Functoriality in each argument does not entail bifunctoriality.
+This is the difference between monoidal and pre-monoidal categories.
+  }
 \end{example}
@@
 for $f,g : A \overline{\otimes} X \to B \overline{\otimes} X$
-TODO: trace?
+
+\section{E-Hypergraphs for $\catname{SLat}$-enriched symmetric closed monoidal categories}
+
+We can extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness.
+
+\begin{definition}
+Closed monoidal signature $\Sigma$ is a pair $(\Sigma_{O},\Sigma_{M})$ of signatures for objects and signatures for morphisms with sources and targets in $\text{obj}_{\Sigma_{O}}$ defined as
+\begin{itemize}
+  \item $I$ (monoidal unit) is in $\text{obj}_{\Sigma_{O}}$
+  \item For every $o \in \Sigma_{O}$, $o \in \text{obj}_{\Sigma_{O}}$;
+  \item For every $o_1$ and $o_2$ $ \in \text{obj}_{\Sigma_{O}}$, $o_1 \otimes o_2 \in \text{obj}{\Sigma_{O}}$ and $o_1 \multimap o_2 \in \text{obj}_{\Sigma_{O}}$
+\end{itemize}
+\end{definition}
+
+\begin{definition}
+ Closed symmetric monoidal theory ($\catname{CSMT}$) over a closed monoidal signature $\Sigma$ and equations $\mathcal{E}$ denoted as $\catname{CSMT}(\Sigma, \mathcal{E})$ is given by a set of terms $T$
+ \begin{align*}
+ T \coloneq\; &t : A \to B \in \Sigma_{M} \;|\; t_1 \otimes t_2 : A \otimes C \to B \otimes D \text{ for } t_{1} : A \to B, t_{2} : C \to D \in T\\
+            & \;|\; t_{1};t_{2} : A \to C \text{ for } t_{1} : A \to B, t_{2} : B \to C \in T \; | \;  \Lambda(t) : A \to X \multimap B \text{ for } t : A \otimes X \to B \\
+            &  \;|\; \text{eval}(t_1 \otimes t_2) : B \text{ for } t_1 : X \multimap B, t_2 : X \in T \\
+            &  \;|\; \text{sym}_{A,B} \;|\; id_{A}
+ \end{align*}
+ quotiented by the laws of a closed symmetric monoidal category.
+\end{definition}
+\question{Maybe types and terms combined is not a good idea}
+\begin{definition}
+
+We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{red}{<}, \textcolor{blue}{<},\consistency)$ where
+\begin{itemize}
+  \item $V$ is a set of vertices;
+  \item $E = \textcolor{gray}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that include \textit{plain} edges and two types of \textit{hierarchical} edges respectively;
+  \item $s,t : E \to V^{*}$ are source and target functions;
+  \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \text{obj}_{\Sigma_{O}}$ are \textit{label} functions;
+  \item $\textcolor{red}{<} : V + E \to \textcolor{red}{E}$ and $\textcolor{blue}{<} : V + E \to \textcolor{blue}{E}$ are partial \textit{child} functions that are disjoint: if $\textcolor{red}{<}(x)$ is defined then $\textcolor{blue}{<}(x)$ is undefined and vice-versa;
+  \item $\consistency = \bigcup_{p} \consistency_{p}$, where $\consistency_{p}$ is a \textit{consistency} relation defined on each set $\{x \in V + E \;|\; p \;\textcolor{red}{<^{\mu}}\; x\}$;
+\end{itemize} 
+\end{definition}
+
+$\textcolor{red}{E}$ will contain edges that encode the equivalence classes and $\textcolor{blue}{E}$ the ones that defined the lambda abstraction.
+All the above functions and relations must satisfy the obvious properties (TODO: add them).
+
+Similarly, we have a category of (extended) cospans formed by such e-hypergraphs where morphisms are of the form
+\[
+n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+\]
+
+where $n$ and $m$ are external input and output interfaces respectively and $n'$, $m'$ are input and output internal interfaces.
+$n,n',m,m'$ are discrete ordered e-hypergraphs.
+We further require that $f_{\text{ext}}$ and $g_{\text{ext}}$ are monos and all vertices in the image of $f_{\text{ext}};f_{\text{int}}$ (respectively, $g_{\text{ext}};g_{\text{int}}$) are top-level.
+
+To build a correspondence between terms of a closed SMC and e-hypergraphs with interfaces (above) we restrict the cospans of the latter to monogamous cospans.
+
+\begin{definition}
+  We call a cospan 
+  \[
+n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+\]
+\textit{weak} monogamous if
+\begin{itemize}
+  \item in- and out- degrees of every vertex is at most 1;
+  \item $f_{\text{int}}$ and $g_{\text{int}}$ are monos;
+  \item Vertices with in-degree (respectively, out-degree) of 0 are precisely the image of $f_{\text{int}}$ (respectively, $g_{\text{int}}$).
+\end{itemize}
+
+\end{definition}
+
+\begin{definition}[Isomorphic cospans]
+Consider a relation 
+\[
+R = \{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\}
+\]
+for $x, y \in n'$.
+And let $S$ be its reflexive closure.
+The latter partitions $n'$ into non-empty subsets $\{p_{j}\}^{k}_{j=1}$.
+We get an analogous partition for $m'$.
+
+Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute and such that $\alpha$ and $\gamma$ preserve order within each $p_j$.
+\[
+\scalebox{0.8}{
+    \tikzfig{../figures/combinatorial_semantics/isomorphic_e_cospans}
+}
+\]
+\end{definition}
+
+Consider an example of isomorphic and non-isomorphic cospans below and their respective string diagrams
+
+\begin{figure}
+  \begin{subfigure}[c]{0.4\linewidth}
+    \scalebox{0.6}{
+    \tikzfig{./figures/closed_iso_example_1}
+    }
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.4\linewidth}
+    \scalebox{0.6}{
+    \tikzfig{./figures/closed_iso_example_2}
+    }
+  \end{subfigure}
+\end{figure}
+
+\begin{definition}
+
+We define ordered sets of \textit{input} (respectively, \textit{output}) vertices of a hierarchical edge $e$ as the vertices $v_{i}$ such that $e < v_{i}$ which are in the image of $f_{\text{int}}$ (respectively, $g_{\text{ext}}$) with the induced ordering.
+\end{definition}
+
+\begin{definition}
+
+Given an ordered set $S$ of labelled vertices we can form words formed by concatenating the labels.
+We will denote such words as $w(S)$.
+\end{definition}
+
+\begin{definition}
+We call a monogamous cospan
+\[
+  n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+\]
+\textit{well-typed} if all hierarchical edges of $\mathcal{G}$ are well-typed in the sense below.
+\begin{itemize}
+  \item For each $e \in \textcolor{red}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
+        $e$ is well-typed if for each element $S$ of the partition of $I$ $w(S) = w(s(e))$ and similarly for $O$ and $t(e)$.
+  \item For each $e \in \textcolor{blue}{E}$ consider sets $I$ and $O$ of its input and output vertices.
+        $e$ is well-typed if there exists an object $B \in \text{obj}_{\Sigma_{O}}$ such that $[...s(e), B] = w(I)$ and $B \multimap w(O) = w(t(e))$
+\end{itemize}
+\end{definition}
 

### 2024-09-17 slides/slides.tex
slides update
updated slides and closed monoidal notes

@@
     enum CommutativeComonoids {
-        "mu" = Mu([Id;1]),
-        "i" = I([Id;1]),
+        "mu" = Mu, // mu : 1 -> 2
+        "i" = I, // i : 1 -> 0
         ";" = Compose([Id;2]),
         "parallel" = Parallel([Id;2]),
-        "id" = Identity([Id;1]),
+        "id" = Identity1,
+        "id_2" = Identity2,
         "I" = Unit,
-        "sym" = Sym([Id;2])   
+        "sym_1_1" = Sym1_1,
+        "sym_2_2" = Sym2_2,
+        "sym_2_1" = Sym2_1,
+        "sym_1_2" = Sym1_2,
+        Symbol(egg::Symbol),
     }
@@
         \begin{figure}
-            \includegraphics[scale=0.4]{figures/egraph_before_saturation.pdf}
+            \includegraphics[scale=0.4]{figures/egraph_before_saturation_fix.pdf}
         \end{figure}

### 2024-09-30 closed_monoidal/closed_monoidal.tex
small fix

@@
 we can use the syntax for plain morphism to depict components of natural transformations.
-This allows us to reuse the whole syntax machinery for Cartesian closed categories for their free $\catname{SLat}$-enriched version which induce an additional equation
+This means, in particular, that given a usual adjuntion $\mathcal{C}(A \otimes X, Y) \cong \mathcal{D}(A, X \multimap Y)$ with respetive natural transformations $\eta : I \to X \multimap (- \otimes X)$ and $\varepsilon : (X \multimap -) \otimes X \to I$ that define a Cartesian closed monoidal category, we can reuse the graphical syntax introduced in~\cite{GhicaZanassiStringLambda}
+for depicting morphisms in a freely $\catname{SLat}$-enriched Cartesian closed monoidal categories.
+Free enrichment induces an additional equation for \textit{abstraction} box below.
 

### 2024-11-20 LICS/sections/appendix.tex
wip

@@
 \begin{remark}
+    The first assumption in~\ref{pushout:assumptions} can be dropped with extra requirements that for all egdes $e_{i}, e_{j} \in E_{Z}$:
+    \begin{itemize}
+        \item $[f_{E}(e_{i})) = [f_{E}(e_{j}))$ and $[g_{E}(e_{i})) = [g_{E}(e_{j}))$ \textcolor{red}{This has to be dropped}
+        \item \textcolor{red}{Find appropriate conditions}
+    \end{itemize}
+\end{remark}
+
+\begin{remark}
     If in a span $Y \xleftarrow{g} Z \xrightarrow{f} X$ of morphisms in $\catname{EHyp}(\Sigma)$ the morphisms $f$ and $g$ are monos, then the arrows $f'$ and $g'$ of the pushout $Y \xrightarrow{g'} X +_{f,g} Y \xleftarrow{f'} X$ are also monos.
@@
 \[\begin{tikzcd}
-	{a + x + y} && {i + j + x} && {i+j} \\
+	{a + y} && {i + j + x} && {i+j} \\
 	\\
-	{a + x + y + z} &&&& {\begin{matrix} i + j + w \\ \cong \\ i + j + z \end{matrix}}
+	{a + y + z} &&&& {\begin{matrix} i + j + w \\ \cong \\ i + j + z \end{matrix}}
 	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
@@
 \end{tikzcd}\]
-    Because $a + x + y + w$ also yields a pushout it must be the case that $a+x+y+w \cong a + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
+    Where $y$ is the image of $i + j$. Clearly, $i + j + z$ is a pushout complement, as computing the pushout identifies $i + j$ with its image within $y$ and leaves $z$ as is.
+    Suppose there is another pushout complement $i + j + w$.
+    Because $a + y + w$ also yields a pushout it must be the case that $a + y + w \cong a + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
+    \textcolor{red}{The same reasoning applies to show $E_{\mathcal{L}^{\bot}}$} is unique.
     One can further observe that $g_{V} = h_{V,ext};h_{V,int} + id_{n}$ and that $h_{V,ext}$ is mono and $h_{V,int} = [h_1,h_2]$ where $h_1,h_2$ are mono.
@@
                 \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=5-5, to=3-3]
+                \arrow["\dagger", draw=none, from=5-5, to=3-3]
             \end{tikzcd}
             \]
+            where the square marked with $\dagger$ is a pushout because of pushout pasting law.
             The corestriction of $n_1 + m_1 \xrightarrow{[f_1,f_2]} \mathcal{L}_1^{\bot}$ to the image of $i + j \xrightarrow{[g_1,g_2]} \mathcal{L}_{1}^{\bot}$ is mono: suppose it is not mono, then there exist $z_1$, $z_2$, $z_3$ such that $f_1(z_1) = f_2(z_2) = g_1(z_3)$, then $[f_2,g_1]$ would not be mono, similarly if $f_1(z_1) = f_2(z_2) = g_2(z_3)$.\
@@
                 which implies the existence of the following diagram where both squares are pushouts by construction and the matching is convex and down-closed.
+                \textcolor{red}{Consider cases when $\mathcal{L}$ has no-input and no-output sub-hypergraph and when it does not.
+Basically, repeat the above reasoning here as well, but say that if $n+m$ is empty, then there is no pushout.}
                 \ifdefined \ONECOLUMN
@@
 \end{figure*}
+
+\section{DPO rewriting for arbitrary signatures}
+\label{sec:dpo-fix}
+Here we will discuss how one can formulate DPO rewrite rules where the left-hand of right-hand sides contain sub-hypergraphs with no inputs and outputs at the outermost level.
+First we will illustrate where the restriction in the original formulation of the DPO rewriting as per the Definition~\ref{def:convex_dpo} comes from via several examples: one for Cartesian equations (which are important to express e-graphs as e-hypergraphs) and two for the structural rules we need for $\catname{MEHypI}_{\Sigma} / \mathcal{S}$ to be a category enriched over $\catname{SLat}$.
+Consider a rewrite rule induced by the Cartesian structure $a;! = id_{I}$ for $a : 0 \to 1$ which is rendered string diagrammatically as 
+\[
+\tikzfig{../figures/appendix/cartesian_rewrite_string}
+\]
+which becomes the following extended cospan of e-hypergraphs
+\[
+\tikzfig{../figures/appendix/cartesian_rewrite_hypergraph}
+\]
+
+Assume we want to perform a rewrite within the e-hypergraph in Figure~\ref{fig:dpo-stuck} (left).
+
+\begin{figure}
+    \begin{subfigure}{0.4\linewidth}
+    \[
+    \adjustbox{scale=0.7}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_source_graph}
+    }
+    \]
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.55\linewidth}
+        \[
+        \adjustbox{scale=0.7}{
+        \tikzfig{../figures/appendix/cartesian_rewrite_dpo_stuck}
+        }
+        \]
+    \end{subfigure}
+    \caption{DPO rewriting stuck}
+    \label{fig:dpo-stuck}
+\end{figure}
+
+However, we are immediately stuck as there is no matching of the left-hand side of the rule within the graph as there is no pushout complement: the only way we can impose a constraint that an edge shall have a particular predecessor in the pushout is through its adjacent vertices, but the sub-hypergraph for $a;!$ does not have any as demonstrated in the right subfigure.
+
+Another problematic rewrite arises when we want to utilise the structural rule of the free enrichment
+
+\[
+\adjustbox{scale=0.75}{
+\tikzfig{../figures/appendix/structural_rewrite_problem}
+}
+\]
+
+if we consider the following DPO square
+
+\[
+\adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck}
+}
+\]
+
+% we can not express id_I + id_I in such a way
+% The last example is the rewrite that corresponds to the idempotence for the identity on the tensor unit $id_{I} = id_{I} + id_{I}$
+
+% \[
+% \adjustbox{scale=0.75}{
+%     \tikzfig{../figures/appendix/structural_id_unit_problem}
+% }
+% \]
+
+% with the DPO square as
+
+% \[
+% \adjustbox{scale=0.75}{
+% \tikzfig{../figures/appendix/structural_rewrite_id_unit_dpo_stuck}
+% }
+% \]
+
+\begin{definition}
+
+We will define the action of $+$ on two cospans where either of the carriers is the empty e-hypergraph as
+
+\begin{align*}
+n \xrightarrow{} n' \xrightarrow{} &\;\mathcal{F} \xleftarrow{} m' \xleftarrow{} m\\
+&\;+ \hspace{6em} \Coloneqq \hspace{2em} n  \xrightarrow{} n' \xrightarrow{} \;\mathcal{F} \xleftarrow{} m' \xleftarrow{} m\\
+0 \xrightarrow{} 0 \xrightarrow{} &\;\varnothing \xleftarrow{} 0 \xleftarrow{} 0.
+\end{align*}
+
+\end{definition}
+
+This is so that we can absorb some equations and to seamlessly introduce what follows.
+
+% \begin{remark}
+% By putting no restrictions on the signature we need to enforce some restrictions on the equations in $\mathcal{E}$.
+% Recall that $\llbracket id_{I} + id_{I}\rrbracket = \llbracket id_{I} \rrbracket$ as per the Remark~\ref{remark:f+g}.
+% This implies that for all $f : 0 \to 0$, $\llbracket f + id_{I} \rrbracket$
+% Consider a cospan of e-hypergraphs that corresponds to a morphism $f + g + (id_{I} + id_{I})$
+
+% \[
+% \adjustbox{scale=0.75}{
+%     \tikzfig{../figures/appendix/remark_on_equations}
+% }
+% \]
+
+% By using the idempotence equation for $id_{I} + id_{I}$ we can rewrite it to
+
+% \[
+%     \adjustbox{scale=0.75}{
+%         \tikzfig{../figures/appendix/remark_on_equations_2}
+%     }
+% \]
+
+% However, there is no rewrite rule to go back and equate these two cospans in $\MdaEcospans / \mathcal{S,E}$.
+% To fix that, we need to introduce the following rewrite rule schema
+% \[
+% \adjustbox{scale=0.75}{
+%     \tikzfig{../figures/appendix/remark_on_equations_3}
+% }
+% \]
+% By introducing this equation we automatically require that every morphism $0 \to 0$ is essentially $id_{I}$.
+
+% \end{remark}
+
+We will start completing all the above DPO squares by first considering the structural rules.
+We will include special unlabelled edges with no inputs and outputs to the interfaces to be able to impose the hierarchical relation on subgraphs with no inputs and outputs.
+That is, the previously mentioned structural rule will become
+
+\[
+\adjustbox{scale=0.75}{
+    \tikzfig{../figures/appendix/structural_rewrite_problem_fix}
+}
+\]
+
+where the labels show how interface edges are mapped to the edges of the corresponding e-hypergraphs.
+For example, the edge in the interface labelled $a_1$ is mapped to the outermost edge on the right.
+We can now compute the pushout complement and complete the DPO square below
+
+\[
+\adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
+}
+\]
+
+Handling deletion or introduction rewrite rules will require the introduction of auxiliary rules.
+In particular, two rules for introduction and elimination of these special edges
+
+\begin{figure}[h!]
+\begin{subfigure}{0.45\linewidth}
+    \[
+    \adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/idI_elimination_hypergraph}
+    }
+    \]
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\linewidth}
+    \[
+    \adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/idI_introduction_hypergraph}
+    }
+    \]
+\end{subfigure}
+\end{figure}
+
+and similar rules to introduce and eliminate in nested contexts
+
+\begin{figure}[h!]
+    \begin{subfigure}{\linewidth}
+        \[
+        \adjustbox{scale=0.55}{
+        \tikzfig{../figures/appendix/idI_elimination_nested_hypergraph}
+        }
+        \]
+    \end{subfigure}
+    \vspace{1em}
+    \begin{subfigure}{\linewidth}
+        \[
+        \adjustbox{scale=0.55}{
+        \tikzfig{../figures/appendix/idI_introduction_nested_hypergraph}
+        }
+        \]
+    \end{subfigure}
+\end{figure}
+
+and, finally, the rules to introduce such edges as new consistent components.
+Intuitively such edge plays the role of a tensor unit identity that can be inserted in any context.
+
+\begin{figure}[h!]
+    \begin{subfigure}{0.45\linewidth}
+        \[
+        \adjustbox{scale=0.55}{
+        \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
+        }
+        \]
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\linewidth}
+        \[
+        \adjustbox{scale=0.55}{
+        \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
+        }
+        \]
+    \end{subfigure}
+    \end{figure}
+
+% \[
+% \adjustbox{scale=0.75}{
+% \tikzfig{../figures/appendix/auxiliary_schema}
+% }
+% \]
+
+% This rule makes sense in any SMT where there is a unique morphism $I \to I$ which is $id_{I}$.
+Then the deletion rule is reconstructed as
+
+\[
+\adjustbox{scale=0.75}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_fix}
+}
+\]
+
+and the full DPO square becomes
+
+\[
+\adjustbox{scale=0.75}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix}
+}
+\]
+
+The target e-hypergraph is not exactly what we wanted, however, but we can use the auxiliary rule to remove this unlabelled edge.
+
+\[
+\adjustbox{scale=0.75}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix_2}
+}
+\]
+
+% and then utilise the idempotence rule for the identity of the tensor unit
+
+% \[
+%     \adjustbox{scale=0.75}{
+%         \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix_3}
+%     }
+% \]
+
+Generally, given an interpretation of a structural rule (where the right-hand side contains a single edge at the outermost level) $\llangle \llbracket l \rrbracket, \llbracket r \rrbracket \rrangle$, for each connected component $C_{i}$ at the outermost level with no inputs and outputs and for each hyperedge with no inputs \textit{or} outputs  in $C_{i}$  we introduce an empty hyperedge into the interfaces which are mapped to the single edge at the outermost level in $\llbracket r \rrbracket$.
+Symmetric rules are handled by performing the above in the other direction.
+
+Rules for deletion and introduction from Cartesian structure are interpreted similarly, by mapping all interface edges to a single empty hyperedge on the right (respectively, left) when deleting (respectively, introducing).
+For an arbitrary rule $\llangle \llbracket l \rrbracket, \llbracket r \rrbracket \rrangle$ induced by $\mathcal{E}$ consider connected components with no inputs and outputs $C^{l}_{i}$ and $C^{r}_{j}$ for left-hand and right-hand sides respectively.
+Let $c^{l}_{1,1}, \ldots, c^{l}_{k,n}$ and $c^{r}_{1,1}, \ldots, c^{r}_{p,m}$ where $c^{l}_{i,j}$ is $i$-th edge with no inputs or outputs in component $C^{l}_{j}$ and similarly for $c^{r}_{i,j}$ and let the interface contain the largest of the two sets.
+And let the morphism be any valid morphism from $\{^{l}_{1,1}, \ldots, c^{l}_{k,n}\} \to c^{r}_{1,1}, \ldots, c^{r}_{p,m}$ is the former is larger as a set, or the other way around if the latter is larger.
+Graphically, this is expressed as follows
+
+\[
+\adjustbox{scale=0.65}{
+    \tikzfig{../figures/appendix/rule_transformation_example}
+}
+\]
+
+assuming that $|c^{l}_{1,1}, \ldots, c^{l}_{k,n}| > |c^{r}_{1,1}, \ldots, c^{r}_{p,m}|$ and $f = \{ c^{l}_{1,1} \mapsto c^{l}_{1,1}, \ldots, c^{l}_{k,n} \mapsto c^{l}_{k,n} \}$ and $g$ is any valid mapping of $c^{l}_{1,1}, \ldots, c^{l}_{k,n}$ to $c^{r}_{1,1}, \ldots, c^{r}_{p,m}$.
+
+\section{Going from e-graph to e-hypergraph}
+
+\begin{figure}
+
+\begin{align*}
+    \text{functional symbols}& \hspace{2em} f,g\\
+    \text{e-class id}& \hspace{2em} a,b\\
+    \text{terms}& \hspace{2em} t \Coloneqq f \;|\; f (t_1, \ldots, t_m) \hspace{2em} &m \geq 1\\
+    \text{e-nodes}& \hspace{2em} n \Coloneqq f \;|\; f (a_1, \ldots, a_m) \hspace{2em} &m \geq 1\\
+    \text{e-classes}& \hspace{2em} c \Coloneqq \{n_1, \ldots, n_m\}   \hspace{2em} &m \geq 1
+\end{align*}
+\caption{E-graph components}
+\label{fig:e-graph-components}
+\end{figure}
+
+\begin{definition}[E-graph~\cite{EggPaper}]
+    E-graph is a data-structure that consists of
+    \begin{itemize}
+        \item components defined in a grammar in Figure~\ref{fig:e-graph-components}
+        \item a union find structure $U$ that stores equivalence relation on e-class ids
+        \item \textit{e-class} map $M$ that maps e-class ids to e-classes.
+All equivalent e-class ids map to the same e-class, \textit{i.e.}, if $a \equiv_{id} b$ then $M[a]$ and $M[b]$ are the same set.
+              An e-class id $a$ is said to refer to e-class $M[\text{find}(a)]$. NB: $\text{find}()$ returns the canonical representative for the equivalence, i.e. the last predecessor of $a$ in $U$.
+        \item The \textit{hashcons} map $H$ that maps e-nodes to e-class ids.
+    \end{itemize}
+\end{definition}
+
+\begin{definition}[Hashcons invariant]
+The hashcons $H$ must map all canonical e-nodes to their e-class ids:
+\[
+\text{e-node } n \in M[a] \text{ iff } H[\text{canonicalize}(n)] = \text{find}(a)
+\]
+
+\begin{algorithm}
+    \caption{Pattern Matching Pseudocode Example}
+    \begin{algorithmic}[1]
+    \ForAll{e-class-id in H}
+    \State edge $edge \gets$ mkEdge(e-class-id)
+        \ForAll {e-node in H \textbf{where} H[e-node] == e-class-id}
+            \State \textbf{match} e-node \textbf{with}
+                \State \quad \textbf{case} \texttt{f}:
+                %    \State \quad \quad edge.add(mk_edge(f))
+        \EndFor
+    \EndFor
+    % \Function{matchPattern}{shape}
+    %     \State \textbf{match} shape \textbf{with}
+    %     \State \quad \textbf{case} \texttt{Circle(radius)}:
+    %     \State \quad \quad \Return $\pi \cdot \texttt{radius}^2$
+    %     \State \quad \textbf{case} \texttt{Rectangle(width, height)}:
+    %     \State \quad \quad \Return \texttt{width} $\times$ \texttt{height}
+    %     \State \quad \textbf{case} \texttt{Triangle(base, height)}:
+    %     \State \quad \quad \Return $\frac{1}{2} \times \texttt{base} \times \texttt{height}$
+    %     \State \quad \textbf{case} \texttt{Square(side)}:
+    %     \State \quad \quad \Return \texttt{side}$^2$
+    %     \State \quad \textbf{case} \texttt{\_}:
+    %     \State \quad \quad \Return ``Unknown shape''
+    % \EndFunction
+    \end{algorithmic}
+    \end{algorithm}
+\end{definition}

### 2024-11-20 LICS/sections/e-hypergraphs.tex
wip

@@
 for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
-While we believe this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such sub-hypergraphs.  
-For this reason,  in the remainder of the paper we consider signatures which can not result in construction of terms with type $0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges.
+While this limitation can be overcome by specifically encoding such rewrite rules, we find the technical development to be simpler if we avoid consideration of such rules and therefore consider signatures which can not result in construction of terms with type $0 \to 0$.
+The discussion of a general case can be found in Section~\ref{sec:dpo-fix} of Appendix.
 %This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.

### 2024-12-24 LICS/sections/appendix.tex
makefile added

@@
 \begin{remark}
-    The first assumption in~\ref{pushout:assumptions} can be dropped with extra requirements that for all egdes $e_{i}, e_{j} \in E_{Z}$:
-    \begin{itemize}
-        \item $[f_{E}(e_{i})) = [f_{E}(e_{j}))$ and $[g_{E}(e_{i})) = [g_{E}(e_{j}))$ \textcolor{red}{This has to be dropped}
-        \item \textcolor{red}{Find appropriate conditions}
+    The first assumption in~\ref{pushout:assumptions} can be weakened by allowing $E_{Z}$ to contain unlabelled edges with no inputs and outputs such that:
+    \begin{itemize}
+        \item $[f_{E}(e_{i})) = [f_{E}(e_{j}))$ and $[g_{E}(e_{i})) = [g_{E}(e_{j}))$ for all $e_{i},e_{j} \in E_{Z}$ 
+        \item If $[f_{E}(e)) \not = \varnothing$ then $[g_{E}(e)) = \varnothing$ and if $[g_{E}(e)) \not = \varnothing$ then $[f_{E}(e)) = \varnothing$.
+        \item $\consistency(f_{E}(e_{i})) = \consistency(f_{E}(e_{j}))$ and $\consistency(g_{E}(e_i)) = \consistency(g_{E}(e_j))$ for all $e_i,e_j$ in $E_{Z}$.
     \end{itemize}
@@
 This is because $[]_{V}$ identifies two vertices $v_1$ and $v_2$ only if there exists a vertex $z$ in $V_{Z}$ such that $f_{V};\iota_{1,V}(z) = v_1$ and $g_{V};\iota_{2,V}(z) = v_2$ and therefore $f' = \iota_{1};[]$ and $g' = \iota_{2};[]$ are monos.
+Similarly for edges.
 
@@
     \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since $m$ and $[c_1,c_2]$ are monos, the first square can be rewritten as follows.
+    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements and where $E_{0}$ is either empty or contains edges with no inputs and outputs only.
+    % Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
+    Since $m$ and $[c_1,c_2]$ are monos, the first square can be rewritten as follows.
 % https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
@@
     Because $a + y + w$ also yields a pushout it must be the case that $a + y + w \cong a + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    \textcolor{red}{The same reasoning applies to show $E_{\mathcal{L}^{\bot}}$} is unique.
-    One can further observe that $g_{V} = h_{V,ext};h_{V,int} + id_{n}$ and that $h_{V,ext}$ is mono and $h_{V,int} = [h_1,h_2]$ where $h_1,h_2$ are mono.
-    The above shows that edges and vertices in the pushout complement are uniquely determined.
+    \textcolor{red}{The same reasoning applies to show that $E_{\mathcal{L}^{\bot}}$} is unique and hence the sets of edges and vertices are uniquely determined in the pushout complement.
+    One can observe that $g_{V} = h_{V,ext};h_{V,int} + id_{n}$ and that $h_{V,ext}$ is mono and $h_{V,int} = [h_1,h_2]$ where $h_1,h_2$ are mono.
     To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{\mathcal{L^{\bot}}}(e)) = s_{\mathcal{G}}(g_{E}(e))$ and $s_{\mathcal{L^{\bot}}}(e) \subseteq (g_{V}^{-1})^{*}(s_{\mathcal{G}}(g_{E}(e)))$.
-    Since $g_{V} = h_{V,ext};[h_1,h_2] + id_{n}$ the inverse image of $g_{V}$ contains at most two elements: one with the pre-image in $i$ and another with the pre-image in $j$.
-    If $g_{V}$ identified more than two vertices it would mean that $[h_1,h_2]$ identified more than two vertices which would mean that two of them come both either from $i$ or $j$ which would imply that either $h_1$ or $h_2$ is not mono.
-    If the inverse image contains exactly one vertex then the $s_{\mathcal{L^{\bot}}}(e)$ is uniquely fixed, otherwise suppose it contains $\{v_1,v_2\}$ such that $v_1 \in i$ and $v_2 \in j$.
-    However, $v_1$ can not be in $s_{\mathcal{L^{\bot}}}(e)$ as per the (6) and (7) conditions of the boundary complement~\ref{def:boundary_new} the image of $j$ in $\mathcal{L}^{\bot}$ should consist of vertices of in-degree 0.
-
+    Suppose, on contrary, that there exist
+    \[
+    \mathcal{L}_{1}^{\bot} = \langle V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1} \rangle \qquad \mathcal{L}_{2}^{\bot} = \langle V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{2} \rangle
+    \]
+
+    such that there is $e \in E_{\mathcal{L}^{\bot}}$ and $s_{1}(e) \not = s_{2}(e)$.
+    As $g$ is a homomorphism, it must be the case that $g^{*}_{V}(s_{1}(e)) = g^{*}_{V}(s_{2}(e)) = s_{\mathcal{G}}(g(e))$, or
+    $(h_{V,ext};[h_{1},h_{2}] + id_{n})(s_{1}(e)) = (h_{V,ext};[h_{1},h_{2}] + id_{n})(s_{2}(e))$.
+    For the latter to hold when $s_{1}(e) \not = s_{2}(e)$ it must be the case that $h_{V,ext};[h_{1},h_{2}]$ identifies $v_{1} \in s_{1}(e)$ with $v_{2} \in s_{2}(e)$ and necessarily $v_{1}$ and $v_{2}$ are in the image of $c$.
+    This can only happen if $v_1 \in i$ and $v_2 \in j$, \textit{i.e.} they belong to input and output interfaces respectively.
+    However, this contradicts the fact that $v_2 \in s_2(e)$ as a vertex which is a source of some edge can not be in the output interface as per the (6) and (7) conditions of the boundary complement~\ref{def:boundary_new} the image of $j$ in $\mathcal{L}^{\bot}$ should consist of vertices of in-degree 0.
     Similarly for $t_{\mathcal{L}^{\bot}}$.
+
+    % Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{\mathcal{L^{\bot}}}(e)) = s_{\mathcal{G}}(g_{E}(e))$ and $s_{\mathcal{L^{\bot}}}(e) \subseteq (g_{V}^{-1})^{*}(s_{\mathcal{G}}(g_{E}(e)))$.
+    % Since $g_{V} = h_{V,ext};[h_1,h_2] + id_{n}$ the inverse image of $g_{V}$ contains at most two elements: one with the pre-image in $i$ and another with the pre-image in $j$.
+    % If $g_{V}$ identified more than two vertices it would mean that $[h_1,h_2]$ identified more than two vertices which would mean that two of them come both either from $i$ or $j$ which would imply that either $h_1$ or $h_2$ is not mono.
+    % If the inverse image contains exactly one vertex then the $s_{\mathcal{L^{\bot}}}(e)$ is uniquely fixed, otherwise suppose it contains $\{v_1,v_2\}$ such that $v_1 \in i$ and $v_2 \in j$.
+    % However, $v_1$ can not be in $s_{\mathcal{L^{\bot}}}(e)$ as per the (6) and (7) conditions of the boundary complement~\ref{def:boundary_new} the image of $j$ in $\mathcal{L}^{\bot}$ should consist of vertices of in-degree 0.
+
     % Suppose that there exist
@@
     \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ is mono.
+    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ \textcolor{red}{only identifies edges with no successors}.
     Similarly for edges.
@@
     Clearly there is a morphism from $\mathcal{L}^{\bot}$ to $\mathcal{G}'$, i.e.
-the morphism $g' = (g_{V},g_{E})$ that has the same action on edges and vertices as the morphism $g$ does, because $v'$ such that $[v') \not = \varnothing$ can be mapped to $g_{V}(v')$ such that $[g_{V}(v')) \not = \varnothing$.
+the morphism $g' = (g_{V},g_{E})$ that has the same action on edges and vertices as the morphism $g$ does, because $v'$ such that $[v') = \varnothing$ can be mapped to $g_{V}(v')$ such that $[g_{V}(v')) \not = \varnothing$.
     The only case when such a morphism may not exist is when $[v'') \not = \varnothing$, $[v') = \varnothing$ and $g_{V}(v'') = g_{V}(v')$.
@@
     This means that either both $<_{1}(\iota_{*}(x))$ and $<_{2}(\iota_{*}(x))$ are undefined for a given $x$ or $<_{1}(\iota_{*}(x)) = <_{2}(\iota_{*}(x))$ which concludes that the relations are equal and $\mathcal{L}_{1}^{\bot} = \mathcal{L}_{2}^{\bot}$.
+    The same reasoning applies when we consider $<_{1}(\iota_{E_{\mathcal{L}^{\bot}}}(e))$ and $<_{2}(\iota_{E_{\mathcal{L}^{\bot}}}(e))$.
 
@@
     Ultimately, we have shown that $\mathcal{L}^{\bot}$ is unique up to isomorphism.
+    The same reasoning applies when we consider edges instead of vertices.
 \end{proof}
@@
 \end{definition}
+
+\begin{lemma} Two e-classes $e_1$ and $e_2$ are merged if and only if there are rewrite rules that turn every term representable by $e_1$ into $e_2$.
+
+\end{lemma}
+
+\begin{lemma}
+Every hierarchical edge can be decomposed as follows.
+
+\[
+\tikzfig{../figures/appendix/hyperedge_decomposition}
+\]
+\end{lemma}
+
+Then, given e-classes $e_1$ and $e_2$ for sub-hypergraphs $C$ and $D$ respectively, we introduce e-nodes into the e-graph $A(e_1)$ and $B(e_2)$ that reside in e-class $e_3$.

### 2024-12-24 LICS/sections/e-hypergraphs.tex
makefile added

@@
 This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
+\textcolor{red}{The definitions and proofs without this extra assumption are given in the Appendix.}
 
@@
 \label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.  
+An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}, for which we introduce the following additional notation.
 For $x \in V + E$,  the set of \emph{parents} of $x$, is denoted as $[x) = \{x' ~|~ x' < x\; \}$.
@@
 	\item Each $x$ has at most one immediate parent;
+	\item The maximal edges must be labelled;
 	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$ and similarly if $v \in t(e)$.
 \end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu x\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into one relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-Note that this relation is symmetric and transitive if considered on $V + E$.
-The consistency relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ and there exists $e'$ such that $e' <^{\mu} v$ and $e' <^{\mu} e$ then $v \consistency e$.
+Finally, $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p <^\mu x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p <^{\mu}(v)$ and $p <^{\mu}(e)$ then $v \consistency_{p} e$.
+We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p <^{\mu} v\}$ and $E_{p} = \{ e ~ | ~ p <^{\mu} e\}$.
+% We collect the union of all these equivalence relations into one relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
+% Note that this relation is symmetric and transitive if considered on $V + E$.
+% The consistency relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ and there exists $e'$ such that $e' <^{\mu} v$ and $e' <^{\mu} e$ then $v \consistency e$.
 \label{def:consistency_properties}
 \end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
+Intuitively,  the child relation equips a hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of string diagrams.  
 These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.  
 These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for the consistency relation. 
+% Technically,  condition $(4)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for the consistency relation. 
+An example of an e-hypergraph can be seen below.
+\[
+\adjustbox{scale=0.75}{
+\tikzfig{../figures/combinatorial_semantics/e-hypergraph-example-2}
+}
+\]
+
+This e-hypergraph corresponds to the following tuple
+\[
+(V = (v_1, v_2, v_3, v_4, v_5, v_6\}, E = \{e_1, e_2, e_3\}, s, t, l, <, \consistency)
+\]
+\begin{align*}
+s(e_1) = \{v_1\} & t(e_1) = \{v_2\} & l(e_1) = \bot \\
+s(e_2) = \{v_3\} & t(e_2) = \{v_4\} & l(e_2) = e_2 \\
+s(e_3) = \{v_5\} & t(e_3) = \{v_6\} & l(e_3) = e_3 \\
+\end{align*}
+
+and $[v_3) = [v_4) = [v_5) = [v_6) = [e_2) = [e_3) = \{e_1\}$ and $[v_1) = [v_2) = [e_1) = \varnothing$.
+$v_3 \consistency v_4 \consistency e_2$ and $v_5 \consistency e_3 \consistency v_6$ and the relation is undefined for the rest vertices and edges.
+Because $v_3 \not \consistency v_5$ and because the relation is closed under connectivity we delimit the connected components by a dashed vertical line.
+The condition $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$, in particular, does not allow hierarchical edges with no delimiting lines.
+$e_2$ and $e_3$ are maximal edges.
 \begin{definition}[E-hypergraph homomorphism]
@@
 
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we let the immediate parent to be undefined for all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
+% This remark is confusing
+% \begin{remark}
+% When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we let the immediate parent to be undefined for all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
+% \end{remark}
 
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
+When modelling enriched string diagrams using e-hypergraphs,  we require keeping track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
 Thus,  we introduce an \textit{extended} cospan construction.

### 2024-12-24 LICS/sections/introduction.tex
makefile added

@@
 Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
+A na\"ive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
 However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
@@
 The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
+\textcolor{red}{The \textit{join} operator lifts the formalism} so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
 

### 2024-12-24 closed_monoidal/closed_monoidal.tex
makefile added

@@
 
+\section{Motivation for e-graphs}
+
+a particular kind of SMTs can provide a new perspective on solving the long-standing issues of $\alpha$-equivalence and binding when using e-graphs for languages like lambda calculus.
+The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
+If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
+The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
+An example of a substitution can be seen in Figure~\ref{fig:e-graph-substitution}.
+An explicit substitution operator is the part of the syntax and the substitution of e-classes is defined inductively on the structure of all terms that these e-classes represent, \textit{i.e.}, computing a substitution of e-classes is achieved by computing a pair-wise substitution for all the instances of these e-classes.
+Notably, the terms that correspond to explicit substitution can be discarded at the end.
+
+By recalling the correspondence between Cartesian Closed SMCs and lambda-calculus we can consider the corresponding semilattice enriched Cartesian Closed SMTs.
+Apart from the generators in $\Sigma_{C}$ such theories also contain \textit{application} and $\textit{abstraction}$ maps that correspond to $app$ and $\lambda$ respectively.
+A complete description of these maps and the equations that they obey can be found in~\cite{GhicaZanassiStringLambda}.
+The e-string diagram version of the previous example can be seen in Figure~\ref{fig:e-string-substitution}.
+The half-circle nodes correspond to application maps and the rounded rectangles with a wire attached to their right sides are the abstraction maps where the attached wire corresponds to the argument of the abstraction.
+A distinguishable feature of these string diagrams is that they are automatically quotiented by $\alpha$-equivalence and that a $\beta$-reduction is merely a rewrite step that glues an argument of the application to the attached wire of the abstraction and the substitution is just string diagram composition.
+Another advantage of string diagrams is that every occurrence of the bound variable is essentially the same wire as opposed to de Brujin indices, so there is no need to modify any indices, the whole substitution is just a matter of reattaching the wires non-destructively.
+In particular, note how the substitution within $y + y$ was completely avoided.
+
+\textcolor{red}{Another motivating example is sharing after $\beta$-redution, add figure}
+
+\begin{figure*}
+  \[
+      % \hspace{1.3cm}
+      \adjustbox{width=\textwidth}{
+      \tikzfig{../figures/categorical-semantics/e-graph-substitution}
+      }
+  \]
+  \caption{E-graph explicit substitution example.}
+  \label{fig:e-graph-substitution}
+\end{figure*}
+
+\begin{figure*}
+  \[
+      % \hspace{1.3cm}
+      \adjustbox{width=\textwidth}{
+      \tikzfig{../figures/categorical-semantics/e-string-substitution-2}
+      }
+  \]
+  \caption{E-string diagrammatic substitution example.}
+  \label{fig:e-string-substitution}
+\end{figure*}
+
 \section{E-Hypergraphs for $\catname{SLat}$-enriched symmetric closed monoidal categories}

### 2024-12-29 LICS/sections/e-hypergraphs.tex
small edits

@@
 \begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
+The extended DPOI rewriting as defined in this section trades off simplicity for expressiveness: it is unable to express rewriting for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
 While this limitation can be overcome by specifically encoding such rewrite rules, we find the technical development to be simpler if we avoid consideration of such rules and therefore consider signatures which can not result in construction of terms with type $0 \to 0$.
-The discussion of a general case can be found in Section~\ref{sec:dpo-fix} of Appendix.
+This limitation has no practical impact in applications of interest. 
+However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of Appendix.
 %This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.

### 2024-12-29 LICS/sections/introduction.tex
title
small edits

@@
 Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the e-string diagrammatic version is guided by the equations of the underlying theory.
-A step-by-step application of such equations for the rewrite rule from $(a)$ to $(b)$ and $(b)$ to $(c)$ can be seen in Figure~\ref{fig:e-graph-example-a-b} and Figure~\ref{fig:e-graph-example-b-c} respectively.
-The equations applied are the ones from Figure~\ref{fig:string-equations} and naturality of delete and copy.
+A detailed step-by-step application of the equations for the rewrite rules from $(a)$ to $(b)$ and $(b)$ to $(c)$ is given in Figures~\ref{fig:e-graph-example-a-b} and~\ref{fig:e-graph-example-b-c}, respectively, in the Appendix.
+The equations applied are the ones from Figure~\ref{fig:string-equations} along with the naturality of delete and copy.
 One can note that the size of the eventual e-string diagram is comparable to the size of the corresponding e-graph.
@@
 Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
+We omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \textbf{SLatt}$.
+It remains to be seen if more general enrichments have interesting applications.
 
@@
 We overload the binary notation $+$ for our $n$-ary notation.
-It is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
+We will later prove the intuitive fact that these diagrams are sound and complete with respect to their intended categorical semantics, noting that similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
 

### 2025-01-06 LICS/sections/combinatorial_semantics.tex
some edits in intro and categorical semantics

@@
 
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
+In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of SMTs including a Frobenius algebra. 
 Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
@@
 This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
+Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of such hypergraphs.
 
@@
 
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
+Then, morphisms of $\catname{S}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
 \begin{definition}[Hypergraphs with interfaces]

### 2025-01-06 LICS/sections/introduction.tex
some edits in intro and categorical semantics

@@
 A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
+\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively}, \textit{i.e.}, forgetting the part of the program that is replaced by the rewrite, at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
 Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
@@
 We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
+Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple Cartesian categories enriched over a category of semilattices (\textit{semilattice-enriched}).
 The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
 The starting innovation of our approach is the semilattice-enrichment. 
-\textcolor{red}{The \textit{join} operator lifts the formalism} so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
+By using the binary operator of a given semilattice we can operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
 
@@
 
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
+Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
+Initially each equivalence class has a single node (subgraph).
 The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
@@
 
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
+Note that the motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing as used in e-graphs.
 However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
@@
 We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.  
-We denote by $\catname{SLatt}$ the category of semilattices.
+We denote by $\catname{SLatt}$ the closed monoidal category of semilattices which is defined in the Appendix (\textcolor{red}{add definition of $\catname{SLatt}$ in the Appendix}).
 
@@
 
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
+A \textit{symmetric monoidal theory} is then defined using $\Sigma$-terms.
+A set of such terms is the set obtained by taking well-typed combinations of $c \in \Sigma, id_{1}, \text{sym}_{1,1}, (;), \text{ and } \otimes$.
+A categorical presentation of a set of $\Sigma$-terms is given by a freely generated products and permutations category.
+
 \begin{definition}[Products and permutations category]
@@
 A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
+We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\catname{S}(\Sigma)$.
+Such PROPs and functors between them define the category $\catname{PROP}$ and, in particular, $\catname{S}(\Sigma) \in \catname{PROP}$.
 \end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
+The free category $\textbf{S}(\Sigma)$ can be syntactically constructed by quotienting the set of $\Sigma$-terms by the axioms of a symmetric monoidal category.
+Thus, $\textbf{S}$ stands for \textit{syntactic} as another way of saying free.
 
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
+An equation associated with a monoidal signature is a pair of parallel $\Sigma$-terms, leading to the following definition of a symmetric monoidal theory. 
+\begin{definition}[Symmetric Monoidal Theory (Definition 2.1~\cite{bonchi_string_2022-1})]
+A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel $\Sigma$-terms $f,g: n \to m$.
+A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by a set of $\Sigma$-terms quotiented by the least congruence including $\mathcal{E}$.
+\end{definition}
+
+\begin{definition}(Definition 2.4~\cite{bonchi_string_2022-1})
+Let $\catname{S}(\Sigma, \mathcal{E})$ a free PROP presented by $\catname{SMT}(\Sigma, \mathcal{E})$.
+That is, it is given by $\catname{S}(\Sigma)$ additionally quotiented by the least congruence including $\mathcal{E}$.
 \end{definition}
@@
 \]
-The Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations of the SMT of commutative comonoids, and the following additional naturality equations, for every $c \in \Sigma_C$
+A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations of the SMT of commutative comonoids, and the following additional naturality equations, for every $c \in \Sigma_C$
 \[
@@
 \]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
+By Fox's Theorem~\cite{fox},  $\textbf{S}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
 \end{example}
@@
 
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
+We will use enrichment to axiomatize the e-box (e-class) structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$. 
 More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
+% Note, in particular,  that we do not require semilattices to have a unit for $+$. 
+% This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
+Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  
 In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
@@
 \]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
+is defined as an $M$-morphism which satisfies certain axioms.
+Composition being an $M$-morphism implies it respects the additional structure on hom-sets.
+%  In other words,  composition must respect the additional structure on hom-sets.  
+% In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
 Technically, this amounts to asking that the tensor product is an enriched functor.
@@
 
-\begin{definition}[Semilattice-enriched SMC]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by an SMC $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC functor on the underlying SMCs which additionally satisfies $F(f+g) = F(f)+F(g)$.
-All definitions extend to \textit{PROP} when the manipulated categories are PROPs rather than SMCs. 
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively). 
+\begin{definition}[Semilattice-enriched category]
+A semilattice-enriched category $\mathbb{C}$ is defined by the following data
+\begin{itemize}
+	\item a set of objects $\mathbb{C}$
+	\item for every pair of objects $A,B \in \mathbb{C})$ --- a hom-object $\mathbb{C}(A,B \in \catname{SLatt}$
+	\item for every triple of objects $A,B,C \in \mathbb{C}$ --- a composition morphism
+	\[
+		\circ: \mathbb{C}(B,C) \otimes \mathbb{C}(A,B) \to \mathbb{C}(A,C)
+	\]
+	\item for every object $A \in C$ --- a unit morphism
+	\[
+	u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLatt}$}
+	\]
+\end{itemize}
+such that particular coherence diagrams commute~\cite{Borceux_1994}.
 \end{definition}
 
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
+Intuitively such enrichment equips every hom-set with a commutative, associative and idempotent operator $+$ that is respected by the composition in the following sense
+\[
+	f ; (g+h) = f;g + f;h \qquad (f+g) ; h = f;h + g;h
+\]
+for all appropriately typed morphisms. 
+We take $(;\!)$ to bind more tightly than $+$.
+
+Note, in particular,  that we do not require semilattices to have a unit for $+$, \textit{i.e.,} we work with \textit{unbounded} semilattices.
+We also will not utilize the ordering that is induced by the operator of a semilattice in any meaningful way, so a semilattice can be thought of as a commutative idempotent semigroup throughout the paper.
+We stick to the term semilattice as it is shorter.
+
+\begin{definition}[Semilattice-enriched SMC]\label{def:enriched-prop}
+A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by a semilattice-enriched category $\mathbb{C}$ as above that additionally has
+\begin{itemize}
+\item a unit object $I_{C} \in \mathbb{C}$
+\item for every pair of objects $A,B \in \mathbb{C}$ --- an object $A \otimes B \in \mathbb{C}$
+\item for all $A,B,C,D$ --- a tensor morphism $ - \otimes_{C} - : \mathbb{C}(A,C) \otimes \mathbb{C}(B,D) \to \mathbb{C}(A \otimes B, C \otimes D)$
+\end{itemize}
+such that particular coherence diagrams commute~\cite{enriched_monoidal}.
+\end{definition}
+The latter morphism being in $\catname{SLatt}$ implies the following equations that $\otimes$ satisfy
+\[
+f \otimes (g+h) = f \otimes g + f \otimes h \qquad (f+g) \otimes h = f \otimes h + g \otimes h
+\]
+
+Again, we take $\otimes$ to bind more tightly than $+$.
+
+\begin{proposition}(A specialized case of Proposition 6.4.7~\cite{Borceux_1994})
+There is a 2-adjunction 
+% https://q.uiver.app/#q=WzAsMixbMCwwLCJcXGNhdG5hbWV7U0xhdHR9LVxcY2F0bmFtZXtDYXR9Il0sWzIsMCwiXFxjYXRuYW1le0NhdH0iXSxbMSwwLCJGIiwyLHsiY3VydmUiOjR9XSxbMCwxLCJHIiwyLHsiY3VydmUiOjR9XSxbMiwzLCIiLDAseyJsZXZlbCI6MSwic3R5bGUiOnsibmFtZSI6ImFkanVuY3Rpb24ifX1dXQ==
+\[\begin{tikzcd}
+	\catname{SLatt}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
+	\end{tikzcd}
+\]
+that is induced by a usual free-forgetful adjunction
+
+\[\begin{tikzcd}
+	\catname{SLatt} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
+	\end{tikzcd}
+\]
+\end{proposition}
+
+In particular, 2-functor $F$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice enriched $\mathbb{C} \in \catname{SLatt}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
+% SMC $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
+% \begin{align*}
+% f ; (g+h) &= f;g + f;h &
+% (f+g) ; h &= f;h + g;h \\
+% f \otimes (g+h) &= f \otimes g + f \otimes h & 
+% (f+g) \otimes h &= f \otimes h + g \otimes h,
+% %\item A set $ob(\mathbb{C})$ of objects;
+% %\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
+% %\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
+% \end{align*}
+% Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
+% A \textit{semilattice enriched strict SMC functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC functor on the underlying SMCs which additionally satisfies $F(f+g) = F(f)+F(g)$.
+% All definitions extend to \textit{PROP} when the manipulated categories are PROPs rather than SMCs. 
+% We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively). 
+% \end{definition}
+\begin{definition}
+We define $\catname{S}^{+}(\Sigma)$ and $\catname{S}^{+}(\Sigma, \mathcal{E})$ to be semilattice-enriched SMCs obtained as $\mathcal{F}(\catname{S}(\Sigma))$ and $\mathcal{F}(\catname{S}(\Sigma, \mathcal{E}))$ respectively.
+Both of these categories are in $\catname{PROP}^{+}$ that is obtained from $\catname{PROP}$ via the same adjunction.
+\end{definition}
+Equivalently, they can be constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category (and the least congruence including $\mathcal{E}$ in the case of $\catname{S}^{+}(\Sigma, \mathcal{E})$).
+% The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
 

### 2025-01-08 LICS/sections/introduction.tex
edit on soundness and completeness and draft for closed

@@
 Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
+The initial term, $(a*2)/2$ (which corresponds to subfigure (a)), is already represented efficiently in the e-graph by sharing the node 2.
 The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node
@@
 Initially each equivalence class has a single node (subgraph).
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
+The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary (containing more than one node or subgraph) equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
 This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
@@
 \captionsetup{belowskip=-4ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
+\caption{String diagrams for semilattice-enriched symmetric monoidal categories.}
 \label{fig:egraph-strings}
@@
 
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
+Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category (see \textit{e.g.}~\cite{GhicaZanassiStringLambda})
 makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
@@
 A \textit{symmetric monoidal theory} is then defined using $\Sigma$-terms.
-A set of such terms is the set obtained by taking well-typed combinations of $c \in \Sigma, id_{1}, \text{sym}_{1,1}, (;), \text{ and } \otimes$.
+A set of such terms is the set obtained by taking well-typed combinations of $c \in \Sigma, \textsf{id}_{I}, \textsf{id}_{1}, \text{sym}_{1,1}, (;), \text{ and } \otimes$.
 A categorical presentation of a set of $\Sigma$-terms is given by a freely generated products and permutations category.
@@
 \begin{definition}
-We define $\catname{S}^{+}(\Sigma)$ and $\catname{S}^{+}(\Sigma, \mathcal{E})$ to be semilattice-enriched SMCs obtained as $\mathcal{F}(\catname{S}(\Sigma))$ and $\mathcal{F}(\catname{S}(\Sigma, \mathcal{E}))$ respectively.
+We define $\catname{S}(\Sigma)^{+}$ and $\catname{S}(\Sigma, \mathcal{E})^{+}$ to be semilattice-enriched SMCs obtained as $\mathcal{F}(\catname{S}(\Sigma))$ and $\mathcal{F}(\catname{S}(\Sigma, \mathcal{E}))$ respectively.
 Both of these categories are in $\catname{PROP}^{+}$ that is obtained from $\catname{PROP}$ via the same adjunction.
 \end{definition}
-Equivalently, they can be constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category (and the least congruence including $\mathcal{E}$ in the case of $\catname{S}^{+}(\Sigma, \mathcal{E})$).
+Equivalently, they can be constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category (and the least congruence including $\mathcal{E}$ in the case of $\catname{S}(\Sigma, \mathcal{E})^{+}$).
 % The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sym{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
@@
 
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
+Figure \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} apart from the last component which is our notation for $+$; the first component denotes an empty diagram.
+Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
 The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.

### 2025-01-08 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
edit on soundness and completeness and draft for closed

--- 2024-09-17 closed_monoidal/closed_monoidal.tex
@@
-%!LW recipe=pdflatex -> bibtex -> pdflatex * 2
-\documentclass[acmsmall, screen, nonacm]{acmart}
+\documentclass[peerreviewcls]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+%Template version as of 6/27/2024
 
-\newcommand*{\ONECOLUMN}{}
-
-% \input{../preamble}
-
-\usepackage{hyperref}
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
 
@@
 
-\usepackage{import}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{remark}[theorem]{Remark}
+
+% some macros that maybe should be moved elsewhere
 
@@
 \newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newcommand\obj[1]{{\textbf{obj}(#1)}}
 
-\usepackage{xifthen}
-\ifthenelse{\isundefined{\ismain}}{
-  \newcommand{\ismain}{0}
-}{}
+\input{../../macros}
+\input{../../sample.tikzstyles}
+\input{../../hypergraph.tikzstyles}
+\input{../../hypergraph.tikzdefs}
 
-\usepackage{enumitem}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\begin{document}
 
-\newcommand{\ignora}[1]{ }
+\title{E-graphs With Bindings
+}
 
-% \newcounter{thm}
-
-% % \ignora{
-% \usepackage{aliascnt}
-% % alias-counters
-% \newaliascnt{definition}{thm}
-% \newaliascnt{proposition}{thm} 
-% \newaliascnt{lemma}{thm}
-% \newaliascnt{corollary}{thm}
-% \newaliascnt{conjecture}{thm}
-% \newaliascnt{remark}{thm}
-% \newaliascnt{example}{thm}
-% \newaliascnt{examples}{thm}
-% \newaliascnt{assumption}{thm}
-% \newaliascnt{construction}{thm}
-% \newaliascnt{claim}{thm}
-
-% % \theoremstyle{plain}
-% \theoremstyle{definition}
-
-% \newtheorem{theorem}[thm]{Theorem}
-% \newtheorem{proposition}[proposition]{Proposition} 
-% \aliascntresetthe{proposition}
-% \newtheorem{lemma}[lemma]{Lemma}
-% \aliascntresetthe{lemma}
-% \newtheorem{corollary}[corollary]{Corollary}
-% \aliascntresetthe{corollary}
-% \newtheorem{conjecture}[conjecture]{Conjecture}
-% \aliascntresetthe{conjecture}
-% \newtheorem{remark}[remark]{Remark}
-% \aliascntresetthe{remark}
-% \newtheorem{assumption}[assumption]{Assumption}
-% \aliascntresetthe{assumption}
-% \newtheorem{construction}[construction]{Construction}
-% \aliascntresetthe{construction}
-% \newtheorem{claim}[claim]{Claim}
-% \aliascntresetthe{claim}
-
-% \theoremstyle{definition}
-% \newtheorem{definition}[definition]{Definition}
-% \aliascntresetthe{definition}
-% \newtheorem{example}[example]{Example}
-% \aliascntresetthe{example}
-% \newtheorem{examples}[examples]{Examples}
-% \aliascntresetthe{examples}
+% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% \and
 % }
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{remark}[theorem]{Remark}
-
-\usepackage{cleveref} % after hyperref!
-
-\input{../macros}
-\input{../sample.tikzstyles}
-\input{../hypergraph.tikzstyles}
-\input{../hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-\newcommand\enriched[1]{{\overline{\mathcal{#1}}}}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-% \setcopyright{acmlicensed}
-% \copyrightyear{2024}
-% \acmYear{2024}
-% \acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{abstract}
-  These notes contain the foundations of categories enriched over a category of semilattices and the respective string diagrams for such categories.
-  It introduces the required notions and carefully checks that all properties are satisfied.
-\end{abstract}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
+\author{Anonymous}
 
@@
 
-\begin{section}{Enriched Categories}
+\begin{abstract}
+This is abstract.
+\end{abstract}
 
-\begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
-  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a)$ be a monoidal category where $l$ and $r$ are left and right unitors.
-  We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
-  \begin{enumerate}
-    \item a class of objects $\obj{\enriched{C}}$
-    \item for every pair of objects $A,B \in \obj{\enriched{C}}$ --- an (hom-)object $\obj{\enriched{C}}(A,B) \in \mathcal{V}$
-    \item for every triple of objects $A,B,C \in \obj{\enriched{C}}$ --- a composition morphism in $\mathcal{V}$
-    \[
-    c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)
-    \]
-    \item for every object $A \in \obj{\enriched{C}}$ --- a unit morphism
-    \[
-      u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)
-    \]
-  \end{enumerate}
-  such that the following diagrams commute (associativity and unitality)
-  \[
-  \begin{tikzcd}
-    {(\enriched{C}(A,B) \otimes \enriched{C}(B,C)) \otimes \enriched{C}(C,D)} \arrow[rr, "{c_{A,B,C} \otimes 1}"] \arrow[d, "{a_{\enriched{C}(A,B),\enriched{C}(B,C),\enriched{C}(C,D)}}"']   &  & {\enriched{C}(A,C) \otimes \enriched{C}(C,D)} \arrow[dd, "{c_{A,C,D}}"]                                                                                     \\
-    {\enriched{C}(A,B) \otimes (\enriched{C}(B,C) \otimes \enriched{C}(C,D))} \arrow[d, "{1\otimes c_{B,C,D}}"']                                             &  &                                                                                                                                       \\
-    {\enriched{C}(A,B) \otimes \enriched{C}(B,D)} \arrow[rr, "{c_{A,B,D}}"]                                                                       &  & {\enriched{C}(A,D)}
-  \end{tikzcd}
-  \]
+\begin{IEEEkeywords}
+component, formatting, style, styling, insert.
+\end{IEEEkeywords}
 
-  \[
-  \begin{tikzcd}
-    {I_{\mathcal{V}} \otimes \enriched{C}(A,B)} \arrow[dd, "u_{A} \otimes 1"'] \arrow[r, "{l_{\enriched{C}(A,B)}}"]                                             & {\enriched{C}(A,B)} \arrow[dd, "{1_{\enriched{C}(A,B)}}"] & {\enriched{C}(A,B) \otimes I_{\mathcal{V}}} \arrow[l, "{r_{\enriched{C}(A,B)}}"'] \arrow[dd, "1 \otimes u_{B}"]                                                           \\
-    &                                     &                                                                                                                                       \\
-{\enriched{C}(A,A) \otimes \enriched{C}(A,B)} \arrow[r, "{c_{A,A,B}}"]                                                                        & {\enriched{C}(A,B)}                            & {\enriched{C}(A,B) \otimes \enriched{C}(B,B)} \arrow[l, "{c_{A,B,B}}"']
-\end{tikzcd}
-  \]
-\end{definition}
+\section{Introduction}
+E-graphs~\cite{EggPaper} are a data structure designed to facilitate equality reasoning in the context of term rewriting systems, \textit{e.g.}, for program optimisation.
+They aim to solve the phase-ordering problem that arises when performing term rewriting, that is, the problem of one rewrite blocking the application of another rewrite due to removing the redex from the initial term.
+They do this by making the process of applying rewrite rules non-destructive: an application of a rewrite rule $\langle l, r \rangle$ simply adds the right-hand side $r$ to the e-graph that has a term with a redex for $l$ such that the new e-graph contains the original term and the $r$ at the same time in a compact manner by utilising sharing as much as possible.
+This way the application of another rule $\langle l, r' \rangle$ would not be blocked as the residual e-graph still contains the redex with $l$ which would have been blocked should we simply replaced all the occurrences of $l$ in the initial term with $r$.
+E-graphs have been around for a while since being introduced in the 80s~\cite{nelson1980techniques} and yet there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
+This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
 
-\begin{remark}
-  We will occasionally use $c_{A,B,C}$ as $;$, \textit{i.e.}, instead of $c_{A,B,C}(f,g)$ we will write $f ; g$ when it makes sense to talk about components of hom-objects.
-\end{remark}
+There are domains, however, where the effective usage of e-graphs requires \textit{ad hoc} tweaks to achieve the desired performance, namely domains that include variable \textit{binding} and the closely related notion of \textit{substitution}, for example, lambda calculus with $\beta$-reduction.
+The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
+If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
+The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
+This work shows how these two problems can be uniformly solved with \textit{string diagrams} by building on top of the recent result of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} that provided a categorical semantics for e-graphs by showing a correspondence between e-graphs and string diagrams for Cartesian semilattice-enriched symmetric monoidal categories or concrete combinatorial structures called \textit{e-hypergraphs} that encode them.
+It proceeds in doing so by showing how e-graphs with bindings can be realized as string diagrams (or e-hypergraphs) for closed (Cartesian) semilattice-enriched symmetric monoidal categories.
 
-\begin{remark}
+\subsection{E-graphs}
 
-Requiring that a category $\mathcal{V}$ is monoidal is the most basic requirement for enriched category theory.
-To make the notion of enriched category more useful, in particular to be able to talk about a free enrichment, we will further require that $\mathcal{V}$ is a closed symmetric monoidal category.
-\end{remark}
+E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes.
+The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
+This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
 
-\begin{example}
-When $\mathcal{V} = \catname{Set}$ we end up with the usual definition of a category.
-$c_{A,B,C}$ becomes the usual morphism composition ($f;g$ rather than $g \circ f$) and $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$ picks out the identity morphism from the homset $\enriched{C}(A,A)$ by matching the only element of a singleton set $I \cong \{*\} \mapsto 1_{A}$ and component-wise the diagrams above commute.
-\end{example}
-
-TODO: another example
-
-\begin{definition}[Semilattice]
-  A \textit{semilattice} is a set equipped with an operation that we denote as $+$ which is associative, commutative and idempotent.
-\end{definition}
-
-Note that we do not require the existence of a unit for $+$. 
-Semilattices that satisfy this extra requirement are sometimes called \textit{bounded}, i.e., they are idempotent commutative monoids.
-
-\begin{definition}[Semilattice homomorphism]
-
-A homomorphism between two semilattices $S_{1}$ and $S_{2}$ is a map $h$ that respects $+$.
-That is, for all $s,s' \in S_{1}$, $h(s +_{S_{1}} s') = h(s) +_{S_{2}} h(s')$.
-\end{definition}
-
-\begin{definition}[Category of semilattices]
-
-Semilattices with their respective homomorphisms form a category that we denote $\catname{SLat}$.
-\end{definition}
-
-\begin{proposition}
-  $\catname{SLat}$ is a closed symmetric monoidal category.
-\end{proposition}
-\begin{proof}
-  The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
-  $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
-  \begin{itemize}
-    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
-    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
-  \end{itemize}
-
-  The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
-  Clearly $S \otimes I \cong S$ by mapping $(s,*) \mapsto s$ for all $s \in S$ and vice versa.
-  The symmetry, associators and unitors are then obvious morphisms.
-  Finally, the category is closed since the set of homomorpisms between two semilattices is a semilattice by defining $(f + g)(x)$ as $f(x) + g(x)$.
-  $f + g$ is a homomorphism since $(f + g)(x+y) = f(x+y) + g(x+y) = f(x) + f(y) + g(x) + g(y) = f(x) + g(x) + f(y) + g(y) = f(x+y) + g(x+y)$.
-\end{proof}
-
-This makes $\catname{SLat}$ a suitable base for enrichment and allows us to take $\mathcal{V} = \catname{SLat}$.
-Before we proceed we need to define what an enriched functor is and define what it means to freely enrich a category over $\catname{SLat}$.
-
-\begin{definition}[Isomorphism in $\enriched{C}$. Lemma 3.5.12~\cite{Riehl_2014}]
-
-Let $\enriched{C}$ be a $\mathcal{V}$-enriched category.
-Given two objects $A,B \in \enriched{C}$ and a morphism $f : I_{\mathcal{V}} \to \enriched{C}(A,B)$, we call $f$ an isomorphism if there is a $g : I_{\mathcal{V}} \to \enriched{C}(B,A)$ such that
-\[
-  I_{\mathcal{V}} \cong I_{\mathcal{V}} \otimes I_{\mathcal{V}} \xrightarrow{f \otimes g} \enriched{C}(A,B) \otimes \enriched{C}(B,A) \xrightarrow{c_{A,B,A}} \enriched{C}(A,A)
-\]
-equals $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$.
-\end{definition}
-
-% TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
-% Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.
-
-\begin{definition}
-  Let $\enriched{C}$ and $\enriched{D}$ be two $\mathcal{V}$-categories.
-  A $\mathcal{V}$-enriched functor $F : \enriched{C} \to \enriched{D}$ is defined by the following data.
-  \begin{enumerate}
-    \item A mapping $F : \obj{\enriched{C}} \to \obj{\enriched{D}}$
-    \item An object-indexed family of morphisms in $\mathcal{V}$ $F_{A,B} : \enriched{C}(A,B) \to \enriched{D}(FA,FB)$
-  \end{enumerate}
-  such that the following diagrams commute
-
-  \[
-  \begin{tikzcd}
-	{\enriched{C}(A,A') \otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} \\
-	\\
-	{\enriched{D}(FA,FA') \otimes \enriched{D}(FA',FA'')} && {\enriched{D}(FA,FA'')}
-	\arrow["{c_{A,A',A''}}"', from=1-1, to=1-3]
-	\arrow["{F_{A,A'} \otimes F_{A',A''}}", from=1-1, to=3-1]
-	\arrow["{F_{A,A''}}"', from=1-3, to=3-3]
-	\arrow["{c_{FA,FA',FA''}}"', from=3-1, to=3-3]
-\end{tikzcd}
-\]
-
-\[
-\begin{tikzcd}
-	I_{\mathcal{V}} && {\enriched{C}(A,A)} \\
-	\\
-	&& {\enriched{D}(FA,FA)}
-	\arrow["{u_{A}}", from=1-1, to=1-3]
-	\arrow["{u_{FA}}"', from=1-1, to=3-3]
-	\arrow["{F_{A,A}}", from=1-3, to=3-3]
-\end{tikzcd}
-\]
-\end{definition}
-
-\begin{definition}
-  Given two $\mathcal{V}$-functors $F,G : \enriched{C} \to \enriched{D}$ a $\mathcal{V}$-natural transformation is family of morphisms indexed by objects in $\enriched{C}$, $\alpha_{A} : I_{\mathcal{V}} \to \enriched{D}(FA,GA)$ in $\mathcal{V}$ such that the following diagram commutes
-  \[
-  \begin{tikzcd}
-	& {\enriched{C}(A,A')\otimes I_{\mathcal{V}}} && {\enriched{D}(FA,FA') \otimes \enriched{D}(FA',GA')} \\
-	{\enriched{C}(A,A')} &&&& {\enriched{D}(FA,GA')} \\
-	& {I_{\mathcal{V}} \otimes \enriched{C}(A,A')} && {\enriched{D}(FA,GA) \otimes \enriched{D}(GA,GA')}
-	\arrow["{G_{A,A'}\otimes \alpha_{A'}}"', from=1-2, to=1-4]
-	\arrow["{c_{FA,FA',GA'}}"', from=1-4, to=2-5]
-	\arrow["{r^{-1}}", from=2-1, to=1-2]
-	\arrow["{l^{-1}}", from=2-1, to=3-2]
-	\arrow["{\alpha_{A} \otimes F_{A,A'}}", from=3-2, to=3-4]
-	\arrow["{c_{FA,GA,GA'}}", from=3-4, to=2-5]
-\end{tikzcd}
-\]
-\end{definition}
-
-\begin{remark}
-  When $\mathcal{V}$ is $\catname{Set}$, a family of morphisms $\alpha_{A} : I_{\mathcal{V}} \to \enriched{C}(FA,GA)$ is defined by mapping $* \mapsto \alpha'_{A}$ where $\alpha'_{A} : FA \to GA$ is usual natural transformation.
-  We can then trace the diagram to get
-
-\[\begin{tikzcd}
-	& {f \otimes *} && {Gf \otimes \alpha'_{A'}} \\
-	f &&&& {Gf;\alpha'_{A} = \alpha'_{A};Ff} \\
-	& {* \otimes f} && {\alpha'_{A} \otimes Ff}
-	\arrow[from=1-2, to=1-4]
-	\arrow[from=1-4, to=2-5]
-	\arrow[from=2-1, to=1-2]
-	\arrow[from=2-1, to=3-2]
-	\arrow[from=3-2, to=3-4]
-	\arrow[from=3-4, to=2-5]
-\end{tikzcd}\]
-
-which is the usual naturality condition.
-
-\end{remark}
-
-\begin{definition}
-Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
-\end{definition}
-
-Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.
-
-\begin{proposition}
-  The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
-\end{proposition}
-\begin{proof}
-  The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
-  The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
-  The adjunction is then given by the following natural isomorphisms
-  \[
-  \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
-  \]
-  The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
-  Furthermore, we have, 
-  \[
-  F(I) \cong I_{\catname{SLat}}
-  \]
-  and 
-  \begin{align*}
-  F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
-        &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
-        &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
-        &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
-        &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
-        &\cong F(A \times B)
-  \end{align*}
-In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
-\end{proof}
-
-\begin{remark}
-  The above proposition in fact works for any symmetric monoidal closed category $\mathcal{V}$ picked as a base for enrichment (Proposition 6.4.6~\cite{Borceux_1994}) 
-\end{remark}
-
-\begin{remark}
-
-  Free semilattice on a set $S$ can be though of as a set of non-empty subsets of $S$.
-\end{remark}
-
-\begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
-\label{def:free_slat}
-  Let $\mathcal{C}$ be a small category.
-  Freely enriched $\catname{SLat}$-category $\enriched{C}$ is defined as follows
-  \begin{enumerate}
-    \item $\obj{\enriched{C}} = \obj{\mathcal{C}}$;
-    \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
-    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite $\cong ; F(;)$ below
-    \[
-    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \xrightarrow{\cong} F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
-    \] 
-    where the rightmost arrow is the image of the ordinary composition;
-    \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
-    \[
-    I_{\catname{SLat}} \xrightarrow{\cong} F(I) \to F(\mathcal{C}(A,A))
-    \]
-    where the rightmost arrow is $* \mapsto 1_{A}$.
-  \end{enumerate}
-\end{definition}
-
-This is again a specialisation of a more general definition of a free $\mathcal{V}$-category (Proposition 6.4.7~\cite{Borceux_1994}).
-
-We will then consider \textit{only} such freely-enriched categories.
-
-\begin{remark}
-  Suppose $\mathcal{C}(A,B) = \{f,g\}$ and $\mathcal{C}(B,C) = \{h\}$, then $f + g \in \enriched{C}(A,B)$ and in particular $(f+g);h = f;h + g;h$ since $c_{A,B,C}$ is a semilattice homomorphism.
-\end{remark}
-
-\begin{proposition}
-\label{prop:functoriality_preserved}
-  A functor $G : \mathcal{C} \to \mathcal{D}$ induces an enriched functor $\overline{G} : \enriched{C} \to \enriched{D}$.
-\end{proposition}
-\begin{proof}
-  We define
-  \begin{itemize}
-    \item $\overline{G}(A) = G(A)$ for all $A \in \enriched{C}$
-    \item $\overline{G}_{A,B}(\enriched{C}(A,B)) = F(G_{A,B}(\mathcal{C}(A,B)))$
-  \end{itemize}
-  The functoriality follows from the following diagrams
-\[
-\adjustbox{width=\linewidth}{
-\begin{tikzcd}
-	{\enriched{C}(A,A')\otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} && {(f, g)} && {f;g} \\
-	\\
-	{\enriched{D}(GA,GA') \otimes \enriched{D}(GA',GA'')} && {\enriched{D}(GA,GA'')} && {(Gf,Gg)} && {Gf;Gg = G(f;g)}
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=1-1, to=3-1]
-	\arrow[from=1-3, to=3-3]
-	\arrow[from=1-5, to=1-7]
-	\arrow[from=1-5, to=3-5]
-	\arrow[from=1-7, to=3-7]
-	\arrow[from=3-1, to=3-3]
-	\arrow[from=3-5, to=3-7]
-\end{tikzcd}}
-\]
-where the right-hand diagram is a component render of the left-hand diagram.
-It commutes because $G$ is a functor. 
-By freeness, the commutativity of more complex components follows from the commutativity for generators.
-The following diagrams commute for the same reason.
-\[\begin{tikzcd}
-	I_{\catname{SLat}} && {\enriched{C}(A,A)} && {*} && {1_{A}} \\
-	\\
-	&& {\enriched{D}(GA,GA)} &&&& {G(1_{A}) = 1_{G(A)}}
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=1-1, to=3-3]
-	\arrow[from=1-3, to=3-3]
-	\arrow[from=1-5, to=1-7]
-	\arrow[from=1-5, to=3-7]
-	\arrow[from=1-7, to=3-7]
-\end{tikzcd}\]
-\end{proof}
-
-\begin{remark}
-  Again, because $\overline{G}_{A,B}$ is a semilattice homomorphism, we have $\overline{G}(f+g) = \overline{G}(f) + \overline{G}(g)$ for any $\catname{SLat}$-enriched functor $\overline{G}$.
-\end{remark}
-
-\begin{example}
-  Let $\mathcal{C} = (\mathcal{C}_{0},I,\otimes, l, r, \alpha)$ be a monoidal category and consider the bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
-  It induces an enriched monoidal category $\enriched{C} = (\enriched{C}_{0}, I, \enriched{\otimes}, \overline{l}, \overline{r}, \enriched{\alpha})$ such that
-  \begin{enumerate}
-    \item $A \overline{\otimes} B \coloneq A \otimes B$;
-    \item $\overline{l} : A \overline{\otimes} I \to A \coloneq I \to \enriched{C}(A \overline{\otimes} I, I)$, i.e. $\overline{l}$ picks out $l$ out of $\enriched{C}(A \overline{\otimes} I, I)$;
-    \item similarly for $\overline{r}$, $\overline{\alpha}$.
-  \end{enumerate}
-  $\otimes_{A,B,C,D} : \mathcal{C}(A,B) \times \mathcal{C}(C,D) \to \mathcal{C}(A \otimes C, B \otimes D)$ becomes $\overline{\otimes}_{A,B,C,D} : \enriched{C}(A,B) \times \enriched{C}(C,D) \to \enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D)$.
-  In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
-  \question{I need to show that $\overline{\otimes}$ is a bifunctor.
-Functoriality in each argument does not entail bifunctoriality.
-This is the difference between monoidal and pre-monoidal categories.
-  }
-\end{example}
-
-\begin{proposition}[Proposition 6.2.8~\cite{Borceux_1994}]
-  When $F,G$ are $\mathcal{V}$-valued $\mathcal{V}$-functors, the notion of $\mathcal{V}$-natural transformation reduces to
-  $\alpha_{A} : FA \to GA$ and the following diagram
-  \[\begin{tikzcd}
-	{C(A,B)} && {[FA,FB]} \\
-	\\
-	{[GA,GB]} && {[FA,GB]}
-	\arrow["{F_{A,B}}", from=1-1, to=1-3]
-	\arrow["{G_{A,B}}"', from=1-1, to=3-1]
-	\arrow["{[1,\alpha_{B}]}", from=1-3, to=3-3]
-	\arrow["{[\alpha_{A},1]}"', from=3-1, to=3-3]
-\end{tikzcd}\]
-\end{proposition}
-
-Note the flip of the domain and co-domain of the bottom arrow.
-This is because the functor $[-,-] : \enriched{C}^{\text{op}} \otimes \enriched{C} \to \mathcal{V}$ is contravariant in its first argument.
-
-\begin{definition}[$\mathcal{V}$-enriched adjunction]
-  Let $\mathcal{V}$ be a closed symmetric monoidal category and $\enriched{C}$, $\enriched{D}$ two $\mathcal{V}$-categories.
-  Two $\mathcal{V}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$ are $\mathcal{V}$-adjoint, $\overline{L}$ left adjoint to $\overline{R}$ and $\overline{R}$ right adjoint to $\overline{L}$, when there exist isomorphisms in $\mathcal{V}$
-  \[
-  \enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B) \;,
-  \]
-  which are $\mathcal{V}$-natural in $A$ and $B$.
-\end{definition}
-
-Equivalently, one can give a definition of an adjunction in terms of unit and co-unit natural transformations.
-
-\begin{definition}[Section 1.11 in~\cite{Kelly2022BASICCO}]
-  A $\mathcal{V}$-adjunction $\eta, \epsilon: \overline{L} \;\reflectbox{$\vdash$}\; \overline{R} : \enriched{C} \to \enriched{D}$ between functors $L : \enriched{D} \to \enriched{C}$ and $\enriched{R} : \enriched{C} \to \enriched{D}$ consists of $\mathcal{V}$-natural $\eta : 1 \to RL$ (the \textit{unit}) and $\epsilon : LR \to 1$ (the \textit{co-unit}) satisfying the triangular equations
-  \[
-  \overline{R}\epsilon . \eta_{\overline{L}} = 1 \qquad \epsilon_{\overline{L}} . \overline{R} \eta = 1
-  \]
-\end{definition}
-Furthermore we have that the isomorphism $\alpha : \enriched{C}(\overline{L}A,B) \to \enriched{D}(A, \overline{R}B)$ has the form $\enriched{D}(\eta, 1)\overline{R}$:
-\[
-  \alpha = \enriched{C}(\overline{L}A, B) \xrightarrow{\overline{R}} \enriched{D}(\overline{RL}A, \overline{R}B) \xrightarrow{\enriched{D}(\eta, 1)} \enriched{D}(A, \enriched{R}B)
-\]
-and $\eta = u_{\overline{L}A};\alpha_{\overline{L}A} = I \xrightarrow{u_{\overline{L}A}}\enriched{C}(\overline{L}A, \overline{L}A) \xrightarrow{\alpha_{\overline{L}A}} \enriched{D}(A,\overline{RL}A)$.
-Similarly, $\alpha^{-1} = \enriched{C}(1,\epsilon)\overline{L}$ and $\epsilon = u_{\overline{R}B};\alpha^{-1}_{\overline{R}B}$.
-
-\begin{proposition}
-\label{prop:naturality_preserved}
-  For a natural transformation $\alpha : T \to S$ for $T,S : \mathcal{C} \to \mathcal{D}$ there is a $\catname{SLat}$-natural transformation $\overline{\alpha} : \overline{T} \to \overline{S}$ where $\overline{T},\overline{S} : \enriched{C} \to \enriched{D}$.
-\end{proposition}
-\begin{proof}
-  Recall that an ordinary natural transformation is in fact a $\catname{Set}$-enriched natural transformation and has the form $\alpha_{A} : I \to \mathcal{D}(TA,SA)$.
-  We then obtain an $\catname{SLat}$-natural transformation $\overline{\alpha_{A}} = F(\alpha_{A}) : F(I) \to F(\mathcal{D}(TA,SA))$, or $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(TA,SA)$ given as $* \mapsto F\alpha_{A}$.
-  Again, we can use $TA$ or $\overline{T}A$ interchangeably as their action on objects is the same.
-  % The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e.
-a morphism that picks out the original $\alpha_{A}$.
-  We need to check if so defined $\overline{\alpha}$ is $\catname{SLat}$-natural.
-  This follows from the following commutativity diagrams
-  \[
-  \adjustbox{width=\linewidth}{
-  \begin{tikzcd}
-	&& {\enriched{C}(A,B) \otimes I_{\catname{SLat}}} && {\enriched{D}(\overline{T}A,\overline{T}B) \otimes \enriched{D}(\overline{T}B, \overline{S}B)} &&&&& {f \otimes *} && {\overline{T}f \otimes F\alpha_{B}} \\
-	{\enriched{C}(A,B)} &&&&&& {\enriched{D}(\overline{T}A,\overline{S}B)} && f &&&& {F\alpha_{A};\overline{S}f = \overline{T}f ;F\alpha_{B}} \\
-	&& {I_{\catname{SLat}} \otimes \enriched{C}(A,B)} && {\enriched{D}(\overline{T}A,\overline{S}A) \otimes \enriched{D}(\overline{S}A,\overline{S}B)} &&&&& {* \otimes f} && {F\alpha_{A} \otimes \overline{S}f}
-	\arrow["{\overline{T}_{A,B} \otimes \overline{\alpha_{B}}}", from=1-3, to=1-5]
-	\arrow["{c_{\overline{T}A,\overline{T}B,\overline{S}B}}", from=1-5, to=2-7]
-	\arrow[from=1-10, to=1-12]
-	\arrow[from=1-12, to=2-13]
-	\arrow[from=2-1, to=1-3]
-	\arrow[from=2-1, to=3-3]
-	\arrow[from=2-9, to=1-10]
-	\arrow[from=2-9, to=3-10]
-	\arrow["{\overline{\alpha_{A}} \otimes \overline{S}(A,B)}", from=3-3, to=3-5]
-	\arrow["{c_{\overline{T}A,\overline{S}A,\overline{T}B}}"', from=3-5, to=2-7]
-	\arrow[from=3-10, to=3-12]
-	\arrow[from=3-12, to=2-13]
-\end{tikzcd}}
-\]
-The diagram on the right is the component-wise rendering of the diagram on the left.
-It commutes because it is essentially a lifting of the following diagram (naturality of $\alpha$) using free enrichment functor $F$.
-
-\[
-\begin{tikzcd}
-	TA && SA && {\overline{T}A} && {\overline{S}A} \\
-	\\
-	TB && SB && {\overline{T}B} && {\overline{S}B}
-	\arrow["{\alpha_{A}}"', from=1-1, to=1-3]
-	\arrow["Tf", from=1-1, to=3-1]
-	\arrow["Sf"', from=1-3, to=3-3]
-	\arrow["{F\alpha_{A}}"', from=1-5, to=1-7]
-	\arrow["{\overline{T}f}", from=1-5, to=3-5]
-	\arrow["{\overline{S}A}"', from=1-7, to=3-7]
-	\arrow["{\alpha_{B}}", from=3-1, to=3-3]
-	\arrow["{F\alpha_{B}}", from=3-5, to=3-7]
-\end{tikzcd}
-\]
-
-The commutativity for more complex elements of a semilattice like $f + g$ follows component-wise.
-
-\end{proof}
-
-\begin{example}
-
-Consider a symmetry natural transformation for a given symmetric monoidal category $\mathcal{C}$ $\sigma_{A,B} : A \otimes B \to B \otimes A$ and its freely $\catname{SLat}$ enriched version $\overline{\sigma_{A,B}} : I_{\catname{SLat}} \to \enriched{C}(A \overline{\otimes} B, B \overline{\otimes} A)$.
-$\catname{SLat}$-naturality takes the following form
-
-\[
-\adjustbox{width=\linewidth}{
-\begin{tikzcd}
-	&& {(\enriched{C}(A,B) \times \enriched{C}(C,D)) \otimes I} && {\enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D) \otimes \enriched{C}(B \overline{\otimes} D, D \overline{\otimes} B)} \\
-	{\enriched{C}(A,B) \times \enriched{C}(C,D)} &&&&&& {\enriched{C}(A \overline{\otimes} C, D \overline{\otimes} B)} \\
-	&& {I \otimes (\enriched{C}(A,B) \times \enriched{C}(C,D))} && {\enriched{C}(A \overline{\otimes} C, C \overline{\otimes} A) \otimes \enriched{C}(C \overline{\otimes} A, D \overline{\otimes} B)} \\
-	\\
-	&& {(f,g) \otimes *} && {(f \overline{\otimes} g) \otimes \sigma} \\
-	{(f,g)} &&&&&& {\sigma;(g \overline{\otimes}f) = (f\overline{\otimes}g);\sigma} \\
-	&& {* \otimes (f,g)} && {\sigma \otimes (g \overline{\otimes} f)}
-	\arrow["{\overline{\otimes} \otimes \overline{\sigma}_{B,D}}", from=1-3, to=1-5]
-	\arrow[from=1-5, to=2-7]
-	\arrow[from=2-1, to=1-3]
-	\arrow[from=2-1, to=3-3]
-	\arrow["{\overline{\sigma}_{C,A} \otimes \overline{\otimes}}"', from=3-3, to=3-5]
-	\arrow[from=3-5, to=2-7]
-	\arrow[from=5-3, to=5-5]
-	\arrow[from=5-5, to=6-7]
-	\arrow[from=6-1, to=5-3]
-	\arrow[from=6-1, to=7-3]
-	\arrow[from=7-3, to=7-5]
-	\arrow[from=7-5, to=6-7]
-\end{tikzcd}}\]
-
-Given $(f_1 + f_2, g)$ we have $((f_1 + f_2) \;\overline{\otimes}\; g);\sigma = (f_1 \;\overline{\otimes}\; g) + (f_{2} \;\overline{\otimes}\; g);\sigma = (f_{1} \;\overline{\otimes}\; g);\sigma + (f_{2} \;\overline{\otimes}\; g);\sigma = \sigma;(g \;\overline{\otimes}\; f_{1}) + \sigma;(g \;\overline{\otimes}\; f_{2})$
-
-\end{example}
-
-\begin{proposition}
-Functor $F : \catname{Set} \to \catname{SLat}$ induces a 2-functor $F^{2}: \catname{Cat} \to \catname{SLat}-\catname{Cat}$ from the category of small categories to the category of small categories enriched over a category of semilattices.
-\end{proposition}
-\begin{proof}
-  The functoriality of $F^{2}$ is given by Definition~\ref{def:free_slat} and 2-functoriality is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
-\end{proof}
-
-The above proposition means, in particular, that an ordinary adjunction gives rise to a corresponding $\catname{SLat}$-enriched adjuntion.
-The latter, however, could be also shown equivalently as below.
-
-\begin{proposition}
-  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors $L : \mathcal{D} \to \mathcal{C}$ and $R : \mathcal{C} \to \mathcal{D}$, we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$ of $\catname{SLat}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$.
-  Because $\overline{L}$ acts on objects as $L$ we will further omit the bar.
-\end{proposition}
-\begin{proof}
-  The first adjunction implies the existence of natural isomorphisms (in $A$ and $B$) between $\mathcal{C}(L-,B)$ and $\mathcal{D}(-,RB)$ and $\mathcal{C}(LA,-)$ and $\mathcal{D}(A,R-)$ respectively.
-  Naturality in $B$ is by definition the commutativity of the following diagram
-  \begin{equation}
-    \label{diag:naturality_plain}
-  \begin{tikzcd}
-	{\mathcal{C}(LA,B)} && {\mathcal{D}(A,RB)} \\
-	\\
-	{\mathcal{C}(LA,B')} && {\mathcal{D}(A,RB')}
-	\arrow["{\alpha_{B}}"', from=1-1, to=1-3]
-	\arrow["{\mathcal{C}(LA,f)}"', from=1-1, to=3-1]
-	\arrow["{\mathcal{D}(A,Rf)}", from=1-3, to=3-3]
-	\arrow["{\alpha_{B'}}", from=3-1, to=3-3]
-  \end{tikzcd}
-\end{equation}
-  and similarly for the naturality in $A$.
-  The claim is that $\enriched{C}(LA,B) \cong \enriched{D}(A,RB)$ is an $\catname{SLat}$-adjunction.
-  Recall that $\enriched{C}(LA,B) = F(\mathcal{C}(LA,B))$ and similarly for another hom-set and clearly the isomorphism is preserved: given $\alpha$, an isomorphism from $\mathcal{C}(LA,B) \to \mathcal{D}(A,RB)$, the isomorphism from $\enriched{C}(LA,B) \to \enriched{D}(A,RB)$ is $\enriched{\alpha} = F(\alpha)$.
-  We need to check that this $\enriched{\alpha}$ is natural.
-  This is equivalent to checking if the following diagram commutes
-  \[\begin{tikzcd}
-	{\enriched{C}(B,B')} && {[\enriched{C}(LA,B),\enriched{C}(LA,B')]} \\
-	\\
-	{[\enriched{D}(A,RB),\enriched{D}(A,RB')]} && {[\enriched{C}(A,RB),\enriched{D}(A,RB')]}
-	\arrow["{\enriched{C}(LA,-)}", from=1-1, to=1-3]
-	\arrow["{\enriched{D}(A,R-)}"', from=1-1, to=3-1]
-	\arrow["{[1,\overline{\alpha}_{B'}]}", from=1-3, to=3-3]
-	\arrow["{[\overline{\alpha_{B}},1]}"', from=3-1, to=3-3]
-\end{tikzcd}\]
-which component-wise becomes
-
-\[
-\begin{tikzcd}
-	f & {\enriched{C}(LA,f)} \\
-	{\enriched{D}(A,Rf)} & {\overline{\alpha}_{B};\enriched{D}(A,Rf);1 = 1;\enriched{C}(LA,f);\overline{\alpha}_{B}'}
-	\arrow[from=1-1, to=1-2]
-	\arrow[from=1-1, to=2-1]
-	\arrow[from=1-2, to=2-2]
-	\arrow[from=2-1, to=2-2]
-\end{tikzcd}
-\]
-
-The commutativity of the diagram above follows from applying $F$ to the diagram \ref{diag:naturality_plain}:
-\[\begin{tikzcd}
-	{\enriched{C}(LA,B)} && {\enriched{D}(A,RB)} \\
-	\\
-	{\enriched{C}(LA,B')} && {\enriched{D}(A,RB')}
-	\arrow["{\overline{\alpha}_{B}}", from=1-1, to=1-3]
-	\arrow["{\enriched{C}(LA,f)}"', from=1-1, to=3-1]
-	\arrow["{\enriched{D}(A,Rf)}", from=1-3, to=3-3]
-	\arrow["{\overline{\alpha}_{B'}}", from=3-1, to=3-3]
-\end{tikzcd}
-\]
-The commutativity for more complex elements of $\enriched{C}(B,B')$, \textit{e.g.} $f + g$, follows component-wise.
-Similarly for naturality in $A$.
-Then, $\catname{Slat}$-naturality separately in $A$ and $B$ implies $\catname{SLat}$-naturality in both $A$ and $B$, a consequence of a more general statement about $\mathcal{V}$-naturality ((1.21) in~\cite{Kelly2022BASICCO})
-
-\end{proof}
-
-\end{section}
-
-\begin{section}{String diagrams for freely $\catname{SLat}$-enriched symmetric monoidal categories}
-
-String diagrams are a two-dimensional syntax for describing morphisms in symmetric monoidal categories which is typically given by the following grammar parameterised over a monoidal signature $\Sigma$ of symbols.
-The labels on wires specify the types of the morphisms.
-
+\begin{figure*}
 \[
 \adjustbox{scale=0.7}{
-\tikzfig{./figures/string_diagrams}
+\tikzfig{figures/e-graph-example}
 }
 \]
+\caption{E-graph example (top) and its equivalent string diagram representation (bottom)}
+\label{fig:e-graph-example}
+\end{figure*}
 
-The diagrams constructed by using the rules above are then subject to SMC equations.
+Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph as developed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+The initial term, $(a*2)/2$ (which corresponds to subfigure (a)), is already represented efficiently in the e-graph by sharing the node 2.
+The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node
+\scalebox{0.3}{
+	\begin{tikzpicture}
+		\begin{pgfonlayer}{nodelayer}
+			\node [style=vertex] (0) at (1.5, 3.5) {};
+			\node [style=none] (1) at (1.5, 3) {};
+			\node [style=none] (2) at (1, 4) {};
+			\node [style=none] (3) at (2, 4) {};
+		\end{pgfonlayer}
+		\begin{pgfonlayer}{edgelayer}
+			\draw (0) to (1.center);
+			\draw [bend right=45] (0) to (3.center);
+			\draw [bend left=45] (0) to (2.center);
+		\end{pgfonlayer}
+	\end{tikzpicture}
+}
+.
 
-By freely enriching home-sets to hom-objects in $\catname{SLat}$ we still can access the morphisms within these objects.
-To support `new' morphisms we extend the grammar by adding a production for $f + g$
+Nodes of an e-graph are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
+Initially each equivalence class has a single node (subgraph).
+The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary (containing more than one node or subgraph) equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
+This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
+Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes 
+\scalebox{0.4}{
+	\begin{tikzpicture}
+		\begin{pgfonlayer}{nodelayer}
+			\node [style=vertex] (0) at (1.5, 3.5) {};
+			\node [style=none] (1) at (1.5, 3) {};
+		\end{pgfonlayer}
+		\begin{pgfonlayer}{edgelayer}
+			\draw (0) to (1.center);
+		\end{pgfonlayer}
+	\end{tikzpicture}
+}, indicate which of the class-level edges are connected to which nodes inside the class.
 
+The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.
+
+\textcolor{red}{Add example showing that using De Brujin indices with e-graphs breaks sharing}
+
+\textcolor{red}{Add example showing $\beta$-reduction in e-graphs compared to string diagrams}
+
+\begin{figure*}
     \[
-\adjustbox{scale=0.8}{
-\tikzfig{./figures/f_plus_g}
+        % \hspace{1.3cm}
+        \adjustbox{width=\textwidth}{
+        \tikzfig{../../figures/categorical-semantics/e-graph-substitution}
         }
     \]
+    \caption{E-graph explicit substitution example.}
+    \label{fig:e-graph-substitution}
+  \end{figure*}
 
-Once we consider string diagrams for a $\catname{SLat}$-enriched category $\enriched{C}$, $;$ and $\otimes$ become $c_{A,B,C}$ and $\overline{\otimes}$ respectively.
-Since the latter are semilattice homomorphisms, we have the following equations (with labels on wires omitted) for the newly introduced dashed boxes.
-
+  \begin{figure*}
     \[
-\adjustbox{width=\linewidth}{
-\tikzfig{./figures/egraph-strings-equations}
+        % \hspace{1.3cm}
+        \adjustbox{width=\textwidth}{
+        \tikzfig{../../figures/categorical-semantics/e-string-substitution-2}
         }
     \]
+    \caption{E-string diagrammatic substitution example.}
+    \label{fig:e-string-substitution}
+  \end{figure*}
 
-The fact of $+$ being associative and commutative justifies the usage of $\ldots$ in the equations.
+\section{Categorical semantics}
 
-Then we can reuse the diagrams for functorial boxes~\cite{GhicaZanassiStringLambda}.
-Given a functor $F : \mathcal{C} \to \mathcal{D}$ and a morphism $f \in C(A,B)$ we can depict the application of $Ff$ as below (left) together with the functoriality equations in Figure~\ref{fig:functoriality_laws}
+Free semilattice enrichment of a closed SMC.
 
-\[
-  \adjustbox{scale=0.7}{
-  \tikzfig{./figures/functorial_string}
-  }
-\]
-
-\begin{figure}
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.7}{
-    \tikzfig{./figures/functoriality_eq_1_string}
-    }
-    \]
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.7}{
-    \tikzfig{./figures/functoriality_eq_2_string}
-    }
-    \]
-  \end{subfigure}
-  \caption{String diagrammatic equations for functoriality}
-  \label{fig:functoriality_laws}
-\end{figure}
-\end{section}
-
-A functor $\overline{F}$ induced by the functor $F$ has the same diagrammatic rendering with the addition of an extra equation.
-
-\[
-\adjustbox{scale=0.7}{
-\tikzfig{./figures/functoriality_enriched_eq}
-}
-\]
-
-Since a component of a natural transformation $\overline{\alpha}$ between two $\catname{SLat}$-enriched functors $\overline{F} : \enriched{C} \to \enriched{D}$ and $\overline{G}: \enriched{C} \to \enriched{D}$ has the form $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(FA, GA)$, it can be identified with the element of $\enriched{D}(FA,GA)$, i.e.
-we can use the syntax for plain morphism to depict components of natural transformations.
-This allows us to reuse the whole syntax machinery for Cartesian closed categories for their free $\catname{SLat}$-enriched version which induce an additional equation
-
-\[
-\adjustbox{scale=0.7}{
-  \tikzfig{./figures/lambda_enriched_eq}
-}
-\]
-
-We can similarly lift a trace from a traced monoidal category $\mathcal{C}$ to $\enriched{C}$.
-Recall that a trace is an operator
-\[
-\text{tr}^{X} : \mathcal{C}(A \otimes X, B \otimes X) \to \mathcal{C}(A,B)
-\]
-\textit{i.e.}, a mapping between hom-sets, such that it satisfies the equations
-\begin{enumerate}
-  \item \textit{Tightening} (naturality in $A,B$) For all $A, B, C, D, X \in \mathcal{C}$, $g : A \to B$, $f : B \otimes X \to C \otimes X$, $h : C \to D$
-  \[
-  \text{tr}^{X}((h \otimes id_{X});f;(g \otimes id_{X})) = h;\text{tr}^{X}(f);g
-  \]
-  \item \textit{Sliding} (naturality in $X$) For all $A,B,X,Y \in \mathcal{C}$, $f : A \otimes X \to B \otimes Y$, $g : Y \to X$
-  \[
-   \text{tr}^{X}(f;(id_{B} \otimes g)) = \text{tr}^{Y}((id_{A} \otimes g); f) 
-  \]
-  \item \textit{Vanishing} For all $A,B,X,Y \in \mathcal{C}$,
-  \[
-  \text{tr}^{I}(f) = f,\; \forall f : A \to B \qquad \text{and} \qquad \text{tr}^{X \otimes Y}(f) = \text{tr}^{X}(\text{tr}^{Y}(f)), \; \forall f : A \otimes X \otimes Y \to B \otimes X \otimes Y
-  \]
-  \item \textit{Strength} For all $A,B,C,D,X \in \mathcal{C}$, $f : C \otimes X \to D \otimes X$, $g : A \to B$
-    \[
-    \text{tr}^{X}(g \otimes f) = g \otimes \text{tr}^{X}(f)
-    \] 
-  \item \textit{Yanking} (when $\mathcal{C}$ is symmetric)
-    \[
-    \text{tr}^{X}(\sigma_{X,X}) = id_{X}
-    \]
-\end{enumerate}
-
-By freely enriching $\mathcal{C}$ we obtain an enriched trace
-\[
-\text{tr}^{X} : \enriched{C}(A \overline{\otimes} X, B \overline{\otimes} X) \to \enriched{C}(A,B)
-\]
-and all the equations are preserved with the addition of a new equation
-\[
-  \text{tr}^{X}(f + g) = \text{tr}^{X}(f) + \text{tr}^{X}(g)
-\]
-for $f,g : A \overline{\otimes} X \to B \overline{\otimes} X$
-
-\section{E-Hypergraphs for $\catname{SLat}$-enriched symmetric closed monoidal categories}
+\section{Combinatorial semantics}
 
@@
 \end{definition}
-\question{Maybe types and terms combined is not a good idea}
+% \question{Maybe types and terms combined is not a good idea}
 \begin{definition}
@@
 \scalebox{0.8}{
-    \tikzfig{../figures/combinatorial_semantics/isomorphic_e_cospans}
+    \tikzfig{../../figures/combinatorial_semantics/isomorphic_e_cospans}
 }
@@
   \begin{subfigure}[c]{0.4\linewidth}
-    \scalebox{0.6}{
+    \[
+    \adjustbox{scale=0.6}{
     \tikzfig{./figures/closed_iso_example_1}
     }
+    \]
   \end{subfigure}
@@
   \begin{subfigure}[c]{0.4\linewidth}
-    \scalebox{0.6}{
+    \[
+    \adjustbox{scale=0.6}{
     \tikzfig{./figures/closed_iso_example_2}
     }
+    \]
   \end{subfigure}
@@
 
-% \[
-% \adjustbox{scale=0.95, lap={\width}{-12em}}{
-%     \begin{tikzcd}
-%         & {\enriched{C}(A_1,A_2)} \arrow[ld] \arrow[rd]                       &                                                                                                                                                                                                            \\
-% {I \otimes \enriched{C}(A_1,A_2)} \arrow[dd, "{\alpha'_{A_1} \otimes \overline{\mathcal{D}}(\overline{L}A,-)}"]                                                                    &                                                                               & {\enriched{C}(A_1,A_2) \otimes I} \arrow[dd, "{\enriched{C}(\overline{L}A,-) \otimes \alpha'_{A_2}}"]                                                                                  \\
-%         &                                                                               &                                                                                                                                                                                                            \\
-% {\mathcal{V}(\enriched{C}(\overline{L}A,A_1),\overline{\mathcal{D}}(A,A_1)) \otimes \mathcal{V}(\overline{\mathcal{D}}(A,A_1),\overline{\mathcal{D}}(A,A_2)))} \arrow[rd, "\circ"] &                                                                               & {\mathcal{V}(\enriched{C}(\overline{L}A,A_1),\enriched{C}(\overline{L}A,A_2))) \otimes \mathcal{V}(\enriched{C}(LA,A_2),\overline{\mathcal{D}}(A,A_2))} \arrow[ld, "\circ"'] \\
-%         & {\mathcal{V}(\enriched{C}(\overline{L}A,A_1), \overline{D}(A,A_2))} &
-% \end{tikzcd}
-% }
-% \]
-
-% \[
-%   \begin{tikzcd}
-%     {\enriched{C}(B,B')} \arrow[rr, "{\enriched{C}(LA,-)}"] \arrow[d, "{\overline{\mathcal{D}}(A,R-)}"] &  & {[\enriched{C}(LA,B), \enriched{C}(LA,B')]} \arrow[d, "{[1,\overline{\alpha_{B'}}]}"] \\
-%     {[\overline{\mathcal{D}}(A,RB),\overline{\mathcal{D}}(A,RB')]} \arrow[rr, "{[\overline{\alpha_{B}},1]}"]                &  & {[\enriched{C}(LA,B),\overline{\mathcal{D}}(A,RB')]}
-%     \end{tikzcd}
-% \]
-
-% \vspace{4em}
-
-% \[
-%   \begin{tikzcd}
-%     {\enriched{C}(B,B')} \arrow[rr, "{\enriched{C}(LA,-)}"] \arrow[d, "{\overline{\mathcal{D}}(A,R-)}"] &  & {[\enriched{C}(LA,B), \enriched{C}(LA,B')]} \arrow[d, "{[1,\overline{\alpha_{B'}}]}"]                             \\
-%     {[\overline{\mathcal{D}}(A,RB),\overline{\mathcal{D}}(A,RB')]} \arrow[rr, "{[\overline{\alpha_{B}},1]}"]                &  & {[\enriched{C}(LA,B),\overline{\mathcal{D}}(A,RB')]}                                                                        \\
-%     f \arrow[rr] \arrow[d]                                                                                                  &  & {\enriched{C}(LA,f)} \arrow[d]                                                                                              \\
-%     {\overline{\mathcal{D}}(A,Rf)} \arrow[rr]                                                                               &  & {1;\enriched{C}(LA,f);\overline{\alpha_{B'}} = \overline{\alpha_{B}};\overline{\mathcal{D}}(A,Rf);1}
-%     \end{tikzcd}
-% \]
-
 \bibliographystyle{acm}
-\bibliography{bibliography}
+\bibliography{../../bibliography}
 

### 2025-01-09 LICS/sections/appendix.tex
some updates

@@
 \begin{align*}
-n \xrightarrow{} n' \xrightarrow{} &\;\mathcal{F} \xleftarrow{} m' \xleftarrow{} m\\
-&\;+ \hspace{6em} \Coloneqq \hspace{2em} n  \xrightarrow{} n' \xrightarrow{} \;\mathcal{F} \xleftarrow{} m' \xleftarrow{} m\\
+0 \xrightarrow{} n \xrightarrow{} &\;\mathcal{F} \xleftarrow{} m \xleftarrow{} 0\\
+&\;+ \hspace{6em} \Coloneqq \hspace{2em} 0  \xrightarrow{} n \xrightarrow{} \;\mathcal{F} \xleftarrow{} m \xleftarrow{} 0\\
 0 \xrightarrow{} 0 \xrightarrow{} &\;\varnothing \xleftarrow{} 0 \xleftarrow{} 0.
@@
 \]
+\end{definition}
+
+The pseudocode below shows the translation from a \textit{canonical} (the one for which the above invariant holds).
+For the following e-graph with e-classes shown as dashed boxes it produces the following cospan of e-hypergraphs.
+\texttt{connect} and \texttt{mkEdge} are responsible for making things shared using $\triangle$.
+
+\begin{figure}
+    \begin{subfigure}{0.4\linewidth}
+\[
+\adjustbox{scale=0.6}{
+\tikzfig{../figures/appendix/e-graph-example_1}
+}
+\]
+\end{subfigure}
+\begin{subfigure}{0.4\linewidth}
+    \[
+    \adjustbox{scale=0.6}{
+    \tikzfig{../figures/appendix/e-hypergraph-example_1}
+    }
+    \]
+\end{subfigure}
+\end{figure}
 
 \begin{algorithm}
-    \caption{Pattern Matching Pseudocode Example}
+    \caption{E-graph to e-hypergraph translation}
     \begin{algorithmic}[1]
-    \ForAll{e-class-id in H}
-    \State edge $edge \gets$ mkEdge(e-class-id)
-        \ForAll {e-node in H \textbf{where} H[e-node] == e-class-id}
-            \State \textbf{match} e-node \textbf{with}
-                \State \quad \textbf{case} \texttt{f}:
-                %    \State \quad \quad edge.add(mk_edge(f))
+    \Function{e-class-to-HEdge}{e\_class\_id}
+    \State e\_nodes $\gets$ [e\_node$_{1}$ $\ldots$ e\_node$_{n}$] \textbf{in} H.keys() \textbf{where} H[e\_node$_{i}$] == M[e\_class\_id]
+    % \State \textbf{for} [$\text{e_node}_{1}, \ldots, \text{e_node}_{n}$] \textbf{in} H \textbf{where} H[$\text{e_node}_{i}$] == M[e-class-id] \texbf{do}
+    \ForAll {e\_node$_{i}$ in e\_nodes}
+    \State edge $\gets$ mkEdge(e\_class\_id)
+    \Comment {Create a hierarchical edge that represents an e-class}
+    \State \textbf{match} e\_node$_{i}$ \textbf{with}
+    \State \qquad \textbf{case} f($e_1 \ldots e_{n}$) :
+    \State \qquad \qquad \textbf{if} is\_singleton(e\_class\_id) \textbf{then}
+    % \If{is\_singleton(e\_class\_id)}
+        \State \qquad\qquad \hspace{2em} edge $\gets$ mkEdge(f)
+        \Comment {Create an edge that represents the generator f $: k \to 1$}
+        \State \qquad \qquad \hspace{2em} edges $\gets$ [\Call{e-class-to-HEdge}{$e_1$} $\ldots$ \Call{e-class-to-HEdge}{$e_n$}]
+        \State \qquad \qquad \hspace{2em} edge.connect(edges)
+        \State \qquad \qquad \hspace{2em} \Return edge
+    % \Else
+    \State \qquad \qquad \textbf{else}
+    \State \qquad \qquad \hspace{2em} graph $\gets$ discard([e\_node$_1$$\ldots$e\_node$_{i-1}$]) $\cup$ mkEdge(f) $\cup$ discard([e\_node$_{i+1}$ $\ldots$ e\_node$_{n}$])
+        \State \qquad \qquad \hspace{2em} edges $\gets$ [\Call{e-class-to-HEdge}{$e_1$} $\ldots$ \Call{e-class-to-HEdge}{$e_n$}]
+        \State \qquad \qquad \hspace{2em} edge.add(graph)
+        \Comment {Add \texttt{graph} as a consistent component into hierarchical edge \texttt{edge}}
+        \State \qquad \qquad \hspace{2em} edge.connect(edges)
+    % \EndIf
     \EndFor
+    \State \Return edge
+    % \Comment {There will be a single e-node in a singleton e-class}
+    % \ForAll {e-node in H \textbf{where} H[e-node] == M[e-class-id]}
+    %     \State \textbf{match} e-node \textbf{with}
+    %     \State \quad \textbf{case} \texttt{f}: \Return (mkEdge(f))
+    %     \State \quad \textbf{case} \texttt{f}($e_1$, $\ldots$, $e_{n}$) :
+    %     % this is nested
+    %         \State \qquad\qquad edge $\gets$ mkEdge(f)
+    %         \State \qquad\qquad edge.connect([e-class-to-HEdge(e) | e $\gets$ [$e_1$, \ldots, $e_{n}$]])
+    %         \State \qquad\qquad \Return edge
+    % \EndFor
+    % \Else
+    % \State edge $\gets$ mkEdge(M[e-class-id])
+    % \ForAll {e-node in H \textbf{where} H[e-node] == M[e-class-id]}
+    %     \State \textbf{match} e-node \textbf{with}
+    %     \State \quad \textbf{case} \texttt{f}: edge.add(f)
+    %     \State \quad \textbf{case} \texttt{f}($e_1$, $\ldots$, $e_{n}$): edge.add(f)
+    % \EndFor
+    % \EndIf
+    \EndFunction
+    \State
+    \State graph $\gets \varnothing$ 
+    \ForAll{e\_class\_id in H.values()}
+    \State graph $\gets$ graph $\cup$ \Call {e-class-to-HEdge}{e\_class\_id})
+
     \EndFor
@@
 \end{algorithm}
-\end{definition}
 

### 2025-01-09 LICS/sections/e-hypergraphs.tex
some updates

@@
 
+\begin{definition}
+\label{def:iso}
 Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute and such that $\alpha$ and $\gamma$ preserve order within each $p_i$.
@@
 \]
+\end{definition}
 Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{bonchi_string_2022-1}.

### 2025-01-09 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
some updates

@@
 
+\subsection{String diagrams}
+
 \subsection{E-graphs}
@@
 
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph as developed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent string diagram representation, an e-hypergraph as developed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
 The initial term, $(a*2)/2$ (which corresponds to subfigure (a)), is already represented efficiently in the e-graph by sharing the node 2.
@@
 
+A typical way on encoding $\beta$-reduction in lambda-calculus in e-graphs is by explicit substitution~\cite{EggPaper,koehler2022sketchguided}: by adding a special substitution nodes and introduce rewrite rules for them.
+After the reduction has been done these nodes serve no purpose and can be removed.
+We show that $\beta$-reduction comes naturally as a composition when using string diagrams for enriched closed monoidal categories.
+Consider an example in Figure~\ref{fig:e-graph-substitution} where successive iterations of an e-graph for a term $(\lambda x . (y + y) + x) 1$ are shown following the $\beta$-reduction.
+First the subterm for an explicit substitution is introduced which is deemed equivalent to the top-level application term.
+It is then rewritten using the rules for explicit substitution at the top of each column.
+At the end all explicit substitution nodes can be removed.
+
+Now consider an example from Figure~\ref{fig:e-string-substitution} which encodes the same $\beta$-reduction for the same term.
+The $\lambda$-abstraction is represented by a rounded box and the bound variable is the dangling wire.
+The application node is depicted as 
+\adjustbox{scale=0.015}{\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=none] (0) at (1, 6.75) {};
+		\node [style=none] (1) at (1.25, 7.75) {};
+		\node [style=none] (2) at (2, 8) {};
+		\node [style=none] (3) at (2.75, 7.75) {};
+		\node [style=none] (4) at (3, 6.75) {};
+		\node [style=none] (5) at (1.5, 6.75) {};
+		\node [style=none] (6) at (2.5, 6.75) {};
+		\node [style=none] (7) at (2, 9) {};
+		\node [style=none] (8) at (1.5, 5.75) {};
+		\node [style=none] (9) at (2.5, 5.75) {};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=new edge style 2] (3.center)
+			 to [in=60, out=-30, looseness=0.75] (4.center)
+			 to (0.center)
+			 to [in=-150, out=120, looseness=0.75] (1.center)
+			 to [in=-180, out=45, looseness=0.75] (2.center)
+			 to [in=135, out=0, looseness=0.75] cycle;
+		\draw (2.center) to (7.center);
+		\draw (5.center) to (8.center);
+		\draw (9.center) to (6.center);
+	\end{pgfonlayer}
+\end{tikzpicture}
+}. 
+The reduction is just the re-attachment of wires.
+Note how we avoided re-attachment for subterms as they did not use the bound variable wire, unlike in the case $[x \mapsto 1](y+y)$ before.
+
+A common solution to replace explicit variable names in e-graphs is by using De Brujin indices~\cite{koehler2022sketchguided}.
+The issue with this approach is that it interferes with sharing as equivalent subterms can be structurally different.
+Consider the example $\lambda f . f (\lambda x . f x)$ which is represented as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$ using De Brujin indices.
+Notice how two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$ which means they can not be shared in an e-graph.
+A string diagrammatic representation of the same term makes it possible for $f$ to be shared as demonstrated below.
+
+\[
+\adjustbox{scale=0.6}{
+    \tikzfig{figures/de-brujin-string}
+}
+\]
+
+Using De Brujin indices when performing $\beta$-reduction also introduces the complexity of representing the reduction as a rewrite rule.
+Special shifting operators need to be introduced on the right-hand side of a rewrite rule which makes it not merely syntactic~\cite{koehler2022sketchguided}.
+
 \textcolor{red}{Add example showing that using De Brujin indices with e-graphs breaks sharing}
@@
         \adjustbox{width=\textwidth}{
-        \tikzfig{../../figures/categorical-semantics/e-graph-substitution}
+        \tikzfig{figures/e-graph-substitution}
         }
@@
         \adjustbox{width=\textwidth}{
-        \tikzfig{../../figures/categorical-semantics/e-string-substitution-2}
+        \tikzfig{figures/e-string-substitution-2}
         }
     \]
-    \caption{E-string diagrammatic substitution example.}
+    \caption{String diagrammatic substitution example.}
     \label{fig:e-string-substitution}

### 2025-01-10 LICS/sections/conclusion.tex
some updates

@@
 In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
+Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string}
+% , as sketched in Fig. \ref{fig:app1}(a)
+; the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}
+% , as sketched in Fig. \ref{fig:app1}(b)
+; or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes
+% , as sketched in Fig.~\ref{fig:app2}. 
 All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
 
-\begin{figure}
-\begin{subfigure}{0.45\linewidth}
-\[
-	\scalebox{0.75}{
-	\tikzfig{../figures/conclusion/example-e-graphs_1}
-	}
-\]
-\subcaption{\;} 
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.45\linewidth}
-\[
-	\scalebox{0.75}{
-	\tikzfig{../figures/conclusion/example-e-graphs_2}
-	}
-\]
-\subcaption{\;}
-\end{subfigure}
-\caption{E-(hyper)graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1}
-\end{figure}
+% \begin{figure}
+% \begin{subfigure}{0.45\linewidth}
+% \[
+% 	\scalebox{0.75}{
+% 	\tikzfig{../figures/conclusion/example-e-graphs_1}
+% 	}
+% \]
+% \subcaption{\;} 
+% \end{subfigure}
+% \hfill
+% \begin{subfigure}{0.45\linewidth}
+% \[
+% 	\scalebox{0.75}{
+% 	\tikzfig{../figures/conclusion/example-e-graphs_2}
+% 	}
+% \]
+% \subcaption{\;}
+% \end{subfigure}
+% \caption{E-(hyper)graphs for functorial boxes and the ZX-calculus}
+% \label{fig:app1}
+% \end{figure}
 
-\begin{figure}
-	\[
-	\scalebox{0.65}{
-	\tikzfig{../figures/conclusion/example-traced}
-	}
-\]
-	\caption{E-graphs with cycles and e-hypergraphs with trace}
-	\label{fig:app2}
-\end{figure}
+% \begin{figure}
+% 	\[
+% 	\scalebox{0.65}{
+% 	\tikzfig{../figures/conclusion/example-traced}
+% 	}
+% \]
+% 	\caption{E-graphs with cycles and e-hypergraphs with trace}
+% 	\label{fig:app2}
+% \end{figure}
 

### 2025-01-14 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
some updates
motivation and contribution

@@
 \section{Introduction}
-E-graphs~\cite{EggPaper} are a data structure designed to facilitate equality reasoning in the context of term rewriting systems, \textit{e.g.}, for program optimisation.
-They aim to solve the phase-ordering problem that arises when performing term rewriting, that is, the problem of one rewrite blocking the application of another rewrite due to removing the redex from the initial term.
-They do this by making the process of applying rewrite rules non-destructive: an application of a rewrite rule $\langle l, r \rangle$ simply adds the right-hand side $r$ to the e-graph that has a term with a redex for $l$ such that the new e-graph contains the original term and the $r$ at the same time in a compact manner by utilising sharing as much as possible.
-This way the application of another rule $\langle l, r' \rangle$ would not be blocked as the residual e-graph still contains the redex with $l$ which would have been blocked should we simply replaced all the occurrences of $l$ in the initial term with $r$.
-E-graphs have been around for a while since being introduced in the 80s~\cite{nelson1980techniques} and yet there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
 
+\subsection{Motivation and contribution}
+
+E-graphs (short for \emph{equality graphs})~\cite{EggPaper} are a data structure incorporating equivalence representation and used in automated theorem proving, compiler optimisation, and program analysis.
+Using e-graphs mitigates the so-called \emph{phase-ordering problem} of term rewriting, that is the fact that the way rewrites are scheduled may enable or block further rewrites due to introduction or elimination or redexes.
+This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule instead of replacing the left-hand side by the right-hand side, adds the latter to the graph so that all pre-existing redexes are protected and new ones are introduced.
+Explosive growth of the graph is avoided by the clever use of sharing of vertices.
+
+E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest particularly for optimisation and synthesis using so called \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, the repeated application of rewrites to an e-graph until a fixpoint is reached.
+The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
+Of particular interest to us is using e-graphs in the context of functional programming languages, which requires a representation of the $\lambda$-calculus using it. 
+Pioneering work in this direction has been done by~\cite{koehler2022sketchguided}.
+The well-known twin challenges of binding and capture-avoiding substitution remain however somewhat problematic in this work because DAN: IT IS UNCLEAR WHY THIS IS PROBLEMATIC IN LOC.
+CIT.
+THE TEXT BELOW IS UNCONVINCING.
+CAN YOU ILLUSTRATE THE PROBLEM WITH SOME SIMPLE EXAMPLES?
+\begin{quotation}
 There are domains, however, where the effective usage of e-graphs requires \textit{ad hoc} tweaks to achieve the desired performance, namely domains that include variable \textit{binding} and the closely related notion of \textit{substitution}, for example, lambda calculus with $\beta$-reduction.
@@
 The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
-This work shows how these two problems can be uniformly solved with \textit{string diagrams} by building on top of the recent result of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} that provided a categorical semantics for e-graphs by showing a correspondence between e-graphs and string diagrams for Cartesian semilattice-enriched symmetric monoidal categories or concrete combinatorial structures called \textit{e-hypergraphs} that encode them.
-It proceeds in doing so by showing how e-graphs with bindings can be realized as string diagrams (or e-hypergraphs) for closed (Cartesian) semilattice-enriched symmetric monoidal categories.
+\end{quotation}
+
+In this paper we take what we consider to be a principled approach to e-graphs with bindings by lifting to monoidal closed categories prior work showing how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}
+The approach in \emph{loc.
+cit.} was originally motivated by the desire to support the machinery of categorial (DPO) graph rewriting for e-graphs, but by extending the semilattice-enrichment to monoidal-categories it paves the way to a natural derivation of e-graphs with binding.
 
@@
 The application node is depicted as 
-\adjustbox{scale=0.015}{\begin{tikzpicture}
+\adjustbox{scale=0.15}{\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 The issue with this approach is that it interferes with sharing as equivalent subterms can be structurally different.
-Consider the example $\lambda f . f (\lambda x . f x)$ which is represented as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$ using De Brujin indices.
+Consider the example $\lambda f . f ((\lambda x . f x) 2)$ which is represented as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$ using De Brujin indices.
 Notice how two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$ which means they can not be shared in an e-graph.

### 2025-01-15 LICS/sections/appendix.tex
wip

@@
 Here we will discuss how one can formulate DPO rewrite rules where the left-hand of right-hand sides contain sub-hypergraphs with no inputs and outputs at the outermost level.
-First we will illustrate where the restriction in the original formulation of the DPO rewriting as per the Definition~\ref{def:convex_dpo} comes from via several examples: one for Cartesian equations (which are important to express e-graphs as e-hypergraphs) and two for the structural rules we need for $\catname{MEHypI}_{\Sigma} / \mathcal{S}$ to be a category enriched over $\catname{SLat}$.
+First we will illustrate where the restriction in the original formulation of the DPO rewriting as per the Definition~\ref{def:convex_dpo} comes from via several examples: one for Cartesian equations (which are important to express e-graphs as e-hypergraphs) and two for the structural rules we need for $\WellTypedMdaEcospans / \mathcal{S}$ to be a category enriched over $\catname{SLat}$.
 Consider a rewrite rule induced by the Cartesian structure $a;! = id_{I}$ for $a : 0 \to 1$ which is rendered string diagrammatically as 
@@
 
-\begin{definition}[Hashcons invariant]
+\begin{definition}[Congruence invariant~\cite{EggPaper}]
+Two e-nodes $f(a_{1}, \ldots, a_{n})$ and $f_{b_{1}, \ldots, b_{n}}$ are \textit{congruent}, if $find(a_{i}) = find(b_{i})$ for all $i$.
+The equivalence relation on e-nodes must be closed under this congruence, that is, two congruent e-nodes must reside in the same e-class.
+\end{definition}
+
+\begin{definition}[Hashcons invariant~\cite{EggPaper}]
 The hashcons $H$ must map all canonical e-nodes to their e-class ids:
@@
 
-The pseudocode below shows the translation from a \textit{canonical} (the one for which the above invariant holds).
+\begin{definition}[~\cite{EggPaper}]
+A rewriting over an e-graph is given by composition of the following operations
+\begin{itemize}
+    \item \textit{add} takes an e-node $n$ and:
+    \begin{itemize}
+        \item if $n$ is present in $H$, return $H[n]$;
+        \item else create e-class id $a$ and set $H[n] = a$ such that $M[a] = \{n\}$.
+    \end{itemize}
+    \item \textit{merge} given two e-class ids $a$ and $b$ unions them in the union find $U$ such that $M[a] = M[b] = M[a] \cup M[b]$.
+\end{itemize}
+\end{definition}
+Then, given a rewrite rule $l \to r$, as, for example, $x * 2 \to x \ll 1$ in Figure~\ref{fig:e-graph-example}, applying it to an e-graph amounts several steps.
+First, finding a tuple $(\sigma, c)$ where e-class id $c$ contains $\sigma(l)$ where sigma is the mapping from variables to e-class ids, for example, $c = H[a * 2]$ and $\sigma(l) = (a * 2)$.
+Then, adding $\sigma(r)$ to the e-graph and merging $c$ with $add(\sigma(r))$.
+The latter step may require additional merges to maintain the invariant.
+If two terms $f(a,b)$ and $f(a,c)$ reside in different e-classes, merging e-classes for $a$ and $b$ would require merging e-classes for $f(a,b)$ and $f(a,c)$ as they become congruent as congruent e-nodes must reside in the same e-class.
+The process of maintaining the invariants by additional merging is called \textit{upward merging}~\cite{EggPaper}.
+We will denote the process of obtaining e-graph $e_2$ from e-graph $e_1$ after applying a rule $l \to r$ (including upward merging) as $e_1 \leadsto e_2$.
+
+\begin{definition}
+The pseudocode in Algorithm~\ref{alg:translation} defines the translation function $\llbracket \rrbracket : \catname{E}\text{-}\catname{graph} \to \WellTypedMdaEcospans$ that turns a \textit{canonical} (the one for which the above invariants hold) to a cospan of e-hypergraphs.
+\end{definition}
+
 For the following e-graph with e-classes shown as dashed boxes it produces the following cospan of e-hypergraphs.
-\texttt{connect} and \texttt{mkEdge} are responsible for making things shared using $\triangle$.
+\texttt{connect} and \texttt{mkEdge} are responsible for making things shared using $\triangle$ and discarding unused inputs via $!$.
 
@@
     \caption{E-graph to e-hypergraph translation}
+    \label{alg:translation}
     \begin{algorithmic}[1]
@@
     \State \Return edge
-    % \Comment {There will be a single e-node in a singleton e-class}
-    % \ForAll {e-node in H \textbf{where} H[e-node] == M[e-class-id]}
-    %     \State \textbf{match} e-node \textbf{with}
-    %     \State \quad \textbf{case} \texttt{f}: \Return (mkEdge(f))
-    %     \State \quad \textbf{case} \texttt{f}($e_1$, $\ldots$, $e_{n}$) :
-    %     % this is nested
-    %         \State \qquad\qquad edge $\gets$ mkEdge(f)
-    %         \State \qquad\qquad edge.connect([e-class-to-HEdge(e) | e $\gets$ [$e_1$, \ldots, $e_{n}$]])
-    %         \State \qquad\qquad \Return edge
-    % \EndFor
-    % \Else
-    % \State edge $\gets$ mkEdge(M[e-class-id])
-    % \ForAll {e-node in H \textbf{where} H[e-node] == M[e-class-id]}
-    %     \State \textbf{match} e-node \textbf{with}
-    %     \State \quad \textbf{case} \texttt{f}: edge.add(f)
-    %     \State \quad \textbf{case} \texttt{f}($e_1$, $\ldots$, $e_{n}$): edge.add(f)
-    % \EndFor
-    % \EndIf
     \EndFunction
@@
     \EndFor
-    % \Function{matchPattern}{shape}
-    %     \State \textbf{match} shape \textbf{with}
-    %     \State \quad \textbf{case} \texttt{Circle(radius)}:
-    %     \State \quad \quad \Return $\pi \cdot \texttt{radius}^2$
-    %     \State \quad \textbf{case} \texttt{Rectangle(width, height)}:
-    %     \State \quad \quad \Return \texttt{width} $\times$ \texttt{height}
-    %     \State \quad \textbf{case} \texttt{Triangle(base, height)}:
-    %     \State \quad \quad \Return $\frac{1}{2} \times \texttt{base} \times \texttt{height}$
-    %     \State \quad \textbf{case} \texttt{Square(side)}:
-    %     \State \quad \quad \Return \texttt{side}$^2$
-    %     \State \quad \textbf{case} \texttt{\_}:
-    %     \State \quad \quad \Return ``Unknown shape''
-    % \EndFunction
     \end{algorithmic}
@@
 
-\begin{lemma} Two e-classes $e_1$ and $e_2$ are merged if and only if there are rewrite rules that turn every term representable by $e_1$ into $e_2$.
-
+\begin{lemma} 
+    Two e-classes $e_1$ and $e_2$ are merged if and only if there are rewrite rules that turn every term representable by $e_1$ into $e_2$.
 \end{lemma}
@@
 \begin{lemma}
-Every hierarchical edge can be decomposed as follows.
-
-\[
-\tikzfig{../figures/appendix/hyperedge_decomposition}
-\]
+Given a set of equations $\mathcal{E}$ between $\Sigma$-terms that induces a set of rewrite rules $l \to r$ and $r \to l$ for every $l = r \in \mathcal{E}$, $e_1 \leadsto e_2$ if and only if $\llbracket e_1 \rrbracket \Rrightarrow{}^{*} \llbracket e_2 \llbracket$.
 \end{lemma}
-
-Then, given e-classes $e_1$ and $e_2$ for sub-hypergraphs $C$ and $D$ respectively, we introduce e-nodes into the e-graph $A(e_1)$ and $B(e_2)$ that reside in e-class $e_3$.
+\begin{proof}
+
+We will prove this by example showing how an addition of an e-node followed by a series of merges can be expressed as a series of DPO rewrites.
+Consider an example in Figure~\ref{fig:e-graph-rewrite}.
+It shows the rewrite $b \to c$ applied to e-graph that initially encodes a single term $f(a,b) + f(a,c)$.
+The first step is $\textit{merge}(b, \textit{add}(c))$ which unions e-classes for $b$ and $c$ by making the latter be the parent of the former, \textit{i.e.}, \text{find}$(b) = c$.
+Then it recanonicalizes the e-node for $f(a,b)$ so that it points to the canonical e-class id for $b$.
+This makes it possible to further merge two e-classes as they contain congruent e-nodes.
+The interpretation of these steps as DPO rewrites is given in Figure~\ref{fig:e-string-rewrite}.
+Subfigure $(a)$ is the direct translation of the initial e-graph by using $\llbracket \rrbracket$.
+Then we utilize the idempotence twice as $b = b + b$ and $c = c + c$ (subfigure (b)) and use the fact that our rewrite rules are symmetric as they are induced by the set of equations and apply $b \to c$ and $c \to b$ (subfigure (c)) and then use the rules for sharing from the Cartesian structure (subfigures (d)-(e)).
+The reverse direction follows by finding normal forms for $\llbracket e_{1} \rrbracket$ and $\llbracket e_{2} \rrbracket$.
+\end{proof}
+
+\begin{remark}
+
+Technically, rewrite rules formed by $\Sigma$-terms do not contain any variables as they are not a part of the signature while the rewrite rules for a given e-graph do contain variables as, for example, $x * 2 \to x \ll 1$.
+This is however not a limitation, as variables are merely interfaces and we can express the above-mentioned rule as a pair of $\Sigma$-terms $*;id \otimes 2 \to \ll ; id \otimes 1$.
+\end{remark}
+
+\begin{figure}
+
+\[
+\adjustbox{width=\linewidth}{
+    \tikzfig{../figures/appendix/e-graph-rewrite-example}
+}
+\]
+\caption{E-graph rewrite example.}
+\label{fig:e-graph-rewrite}
+\end{figure}
+
+\begin{figure}
+
+    \[
+        \adjustbox{width=\linewidth}{
+            \tikzfig{../figures/appendix/e-string-rewrite-example}
+        }
+        \]
+        \caption{String diagram rewrite example.}
+        \label{fig:e-string-rewrite}
+\end{figure}
+
+% \begin{lemma}
+% Every hierarchical edge can be decomposed as follows.
+
+% \[
+% \tikzfig{../figures/appendix/hyperedge_decomposition}
+% \]
+% \end{lemma}
+
+% Then, given e-classes $e_1$ and $e_2$ for sub-hypergraphs $C$ and $D$ respectively, we introduce e-nodes into the e-graph $A(e_1)$ and $B(e_2)$ that reside in e-class $e_3$.

### 2025-01-15 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
wip

\section{Categorical semantics of E-Graphs with Bindings}

In this section we introduce the required preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories, and the string diagrammatic formalism we will use to represent them.
We also show a specific case of semilattice-enriched closed monoidal categories and how closed structure interacts with the enrichment structure.

We will consider symmetric monoidal categories generated by \textit{monoidal signatures}.

\begin{definition}

A \textit{monoidal} \textit{signature} $\Sigma = (\Sigma_{O}, \Sigma_{M}, t)$ is given by a set of objects $\Sigma_{O}$, set of generators $\Sigma_{M}$ and a type assignment function $t : \Sigma_{M} \to \Sigma_{O}^{*} \times \Sigma_{O}^{*}$ that assigns lists of input and output types to a given generator.
\end{definition}

\begin{definition}

A set of $\Sigma$-terms generated by monoidal signature $\Sigma$ is given by the following grammar
\[
\tau \Coloneq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}
\],
where $\phi$ ranges over the set of generators $\Sigma_{M}$ and $A,B$ range over the set of objects $\Sigma_{O}$.
\end{definition}

Identities and symmetries of more complex types, \textit{e.g.}, $\id_{A \otimes B}$ and $\sym_{A \otimes B, C}$, are given by the composition and tensoring of $\id_{A}$ and $\sym_{A,B}$ as, $\textit{e.g.}$, $\id_{A \otimes B} = \id_{A} \otimes \id_{B}$ and $\sym_{A \otimes B, C} = (\id_{A} \otimes \sym_{B,C});(\sym_{A,C} \otimes \id_{B})$.

\begin{definition}
A free symmetric monoidal category $\catname{S}(\Sigma)$ is given by a set of $\Sigma$-terms quotiented by the axioms of a symmetric monoidal category.
The objects are lists of types from $\Sigma_{O}$ with tensoring given by list concatenation.
\end{definition}

\begin{definition}

A symmetric monoidal theory $\catname{SMT}(\Sigma, \mathcal{E})$ is given by a set of $\Sigma$-terms quotiented by a set of equations $\mathcal{E}$ consisting of pairs of appropriately typed $\Sigma$ terms $l = r$.
\end{definition}

\begin{definition}
A categorical presentation of an $\catname{SMT}(\Sigma, \mathcal{E})$ is a free symmetric monoidal category $\catname{S}(\Sigma, \mathcal{E})$ given by a set of $\Sigma$ terms quotiented by the laws of a symmetric monoidal category and equations of $\mathcal{E}$.
\end{definition}

We can similarly consider \textit{closed} terms that give rise to a free \textit{closed} symmetric monoidal category.

\begin{definition}
\label{def:closed}
A monoidal (right) closed category is a monoidal category $\catname{C}$ satisfying for
every pair of objects $B,C$ there is an object $B \xRightarrow{} C$ and a morphism $\textbf{ev}_{B,C} : (B \xRightarrow{} C) \otimes B \to
C$, and for every triple of objects $A,B,C$ there is an operation $\Lambda_{A,B,C} : \catname{C}(A \otimes B,C) \to
\catname{C}(A,B \xRightarrow{} C)$ satisfying the following equations for all $f : A \otimes B \to C$ and $g : Z \to A$:
\begin{itemize}
\item $f = \Lambda_{A,B,C}(f) \otimes \id_{B}; \textbf{ev}_{B,C}$
\item $\id_{B \xRightarrow{} C} = \Lambda_{B \xRightarrow{} C,B,C}(\textbf{ev}_{B,C})$
\item $\Lambda_{Z,B,C}(g \otimes \id_{B}; f) = g;\Lambda_{A,B,C}(f)$
\end{itemize}
\end{definition}

This definition is equivalent to the more usual one that defines a (right) closed category through a pair of adjoint functors $- \otimes B$ and $B \xRightarrow{} -$~\cite{ghica-zanassi2023string}.
If the tensor is symmetric, then the category is simply \textit{closed}.

\begin{definition}

\textit{Closed} $\Sigma$-terms are given by the following grammar
\[
    \tau \coloneq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2} \; | \; \textbf{ev}_{A,B} \; | \; \Lambda(\tau)
\]
, where the types are given as
\[
A,B \coloneq T \;|\; A \otimes B \; | \; A \xRightarrow{} B
\], where $T$ is in $\Sigma_{O}$.
\end{definition}

\begin{definition}
The free closed symmetric monoidal category $\catname{Sc}(\Sigma)$ is given by a set of closed $\Sigma$-terms quotiented by the laws of a symmetric monoidal category and the equations from Definition~\ref{def:closed}.
\end{definition}

To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with additional structure of a semilattice induced by an idempotent, commutative and associative operator $+$.
This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them.

\begin{definition}[Semilattice-enriched category]
    A semilattice-enriched category $\mathbb{C}$ is defined by the following data
    \begin{itemize}
        \item a set of objects $\mathbb{C}$
        \item for every pair of objects $A,B \in \mathbb{C})$ --- a hom-object $\mathbb{C}(A,B \in \catname{SLatt}$
        \item for every triple of objects $A,B,C \in \mathbb{C}$ --- a composition morphism
        \[
            \circ: \mathbb{C}(B,C) \otimes \mathbb{C}(A,B) \to \mathbb{C}(A,C)
        \]
        \item for every object $A \in C$ --- a unit morphism
        \[
        u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLatt}$}
        \]
    \end{itemize}
    such that particular coherence diagrams commute~\cite{Borceux_1994}.
\end{definition}

\begin{definition}[Semilattice-enriched SMC]\label{def:enriched-prop}
    A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by a semilattice-enriched category $\mathbb{C}$ as above that additionally has
    \begin{itemize}
    \item a unit object $I_{C} \in \mathbb{C}$
    \item for every pair of objects $A,B \in \mathbb{C}$ --- an object $A \otimes B \in \mathbb{C}$
    \item for all $A,B,C,D$ --- a tensor morphism $ - \otimes_{C} - : \mathbb{C}(A,C) \otimes \mathbb{C}(B,D) \to \mathbb{C}(A \otimes B, C \otimes D)$
    \end{itemize}
    such that particular coherence diagrams commute~\cite{enriched_monoidal}.
    \end{definition}
    The latter morphism being in $\catname{SLatt}$ implies the following equations that $\otimes$ satisfy
    \[
    f \otimes (g+h) = f \otimes g + f \otimes h \qquad (f+g) \otimes h = f \otimes h + g \otimes h
    \]

    Again, we take $\otimes$ to bind more tightly than $+$.

    \begin{proposition}(A specialized case of Proposition 6.4.7~\cite{Borceux_1994})
    There is a 2-adjunction 
    % https://q.uiver.app/#q=WzAsMixbMCwwLCJcXGNhdG5hbWV7U0xhdHR9LVxcY2F0bmFtZXtDYXR9Il0sWzIsMCwiXFxjYXRuYW1le0NhdH0iXSxbMSwwLCJGIiwyLHsiY3VydmUiOjR9XSxbMCwxLCJHIiwyLHsiY3VydmUiOjR9XSxbMiwzLCIiLDAseyJsZXZlbCI6MSwic3R5bGUiOnsibmFtZSI6ImFkanVuY3Rpb24ifX1dXQ==
    \[\begin{tikzcd}
        \catname{SLatt}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
        \end{tikzcd}
    \]
    that is induced by a usual free-forgetful adjunction

    \[\begin{tikzcd}
        \catname{SLatt} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
        \end{tikzcd}
    \]
    \end{proposition}

In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice enriched $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
Being a 2-functor it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{Slat}$-adjunction.
Thus, applying the functor to $\catname{Sc}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{Sc}(\Sigma)^{+}$ by lifting the adjunction $- \otimes B \dashv B \xRightarrow{} -$.

### 2025-01-15 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
wip
intro fix

@@
 \documentclass[peerreviewcls]{IEEEtran}
-\IEEEoverridecommandlockouts
+% \IEEEoverridecommandlockouts
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
@@
 Of particular interest to us is using e-graphs in the context of functional programming languages, which requires a representation of the $\lambda$-calculus using it. 
-Pioneering work in this direction has been done by~\cite{koehler2022sketchguided}.
-The well-known twin challenges of binding and capture-avoiding substitution remain however somewhat problematic in this work because DAN: IT IS UNCLEAR WHY THIS IS PROBLEMATIC IN LOC.
+\textcolor{red}{Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
+the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when doing equality saturation for $\lambda$-calculus with e-graphs.
+As e-graph endeavours to represent multiple terms at once with as much sharing as possible, using variable names for binding means that two $\alpha$-equivalent terms can not be shared as they are structurally different.
+Even though such simple encoding makes expressing $\beta$-reduction as a rewrite rule in terms of explicit substitution simpler, applying such rules requires computing extra free variable predicates and populates a given e-graph with extra nodes for explicit substitution.
+On the other hand, encoding binding using $\textit{de Brujin}$ while still fails to represent $\alpha$-equivalent terms as being structurally equivalent, as \textit{e.g.}, in $\lambda .
+\%0 ((\lambda .
+\%1 \%0) 42)$ the same binder is represented with $\%0$ and $\%1$, avoids explicit substitution when implementing $\beta$-reduction at the cost of including \textit{ad hoc} operators for index shifting into the rewrite rules.
+Generally, the incurred redundancy these challenges remain problematic because variable binding is not inherent to e-graphs as they do not provide an internal support for it.
+Other works suggested an approach to add internal support for binding which however made the formulation of e-graph considerably more complicated~\cite{slotted-egraphs}.
+}
+\textcolor{blue}{The reference above is a workshop unpublished work, I do not know if it is good to use such reference.} 
+% The well-known twin challenges of binding and capture-avoiding substitution remain however somewhat problematic in this work because DAN: IT IS UNCLEAR WHY THIS IS PROBLEMATIC IN LOC.
 CIT.
@@
 CAN YOU ILLUSTRATE THE PROBLEM WITH SOME SIMPLE EXAMPLES?
-\begin{quotation}
-There are domains, however, where the effective usage of e-graphs requires \textit{ad hoc} tweaks to achieve the desired performance, namely domains that include variable \textit{binding} and the closely related notion of \textit{substitution}, for example, lambda calculus with $\beta$-reduction.
-The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
-If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
-The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
-\end{quotation}
+% \begin{quotation}
+% There are domains, however, where the effective usage of e-graphs requires \textit{ad hoc} tweaks to achieve the desired performance, namely domains that include variable \textit{binding} and the closely related notion of \textit{substitution}, for example, lambda calculus with $\beta$-reduction.
+% The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
+% If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
+% The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
+% \end{quotation}
 
@@
 
-\section{Categorical semantics}
-
-Free semilattice enrichment of a closed SMC.
+\input{sections/categorical_semantics}
 

### 2025-01-15 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
intro fix

@@
 \[
-\tau \Coloneq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}
+\tau \coloneq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}
 \],
@@
 Being a 2-functor it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{Slat}$-adjunction.
-Thus, applying the functor to $\catname{Sc}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{Sc}(\Sigma)^{+}$ by lifting the adjunction $- \otimes B \dashv B \xRightarrow{} -$.
+Thus, applying the functor to $\catname{Sc}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{Sc}(\Sigma)^{+}$ by lifting the adjunction $(- \otimes B, B \xRightarrow{} -, \eta, \varepsilon)$.
+We can similarly derive $\catname{SLat}$-enriched versions of  $\Lambda_{A,B,C}$ and $\textbf{ev}_{A,B}$ morphisms using the lifted adjunction as $\textbf{ev}_{A,B} = \varepsilon_{A}$ and $\Lambda_{A,B,C}(f) = \eta; (B \xRightarrow{} f)$.
+The following equation can then be derived which will be useful in encoding equivalences inside abstractions.
+\begin{align*}
+\Lambda_{A,B,C}(f + g) &= \eta;(B \xRightarrow{} (f + g))\\
+                       &= \eta;(B \xRightarrow{} (f) + B \xRightarrow{} (g))\\
+                       &= \eta;(B \xRightarrow{} (f)) + \eta;(B \xRightarrow{} (g))\\
+                       &= \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g)
+\end{align*}

### 2025-01-18 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
string diagrams section

--- 2025-01-14 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
@@
 Of particular interest to us is using e-graphs in the context of functional programming languages, which requires a representation of the $\lambda$-calculus using it. 
-Pioneering work in this direction has been done by~\cite{koehler2022sketchguided}.
-The well-known twin challenges of binding and capture-avoiding substitution remain however somewhat problematic in this work because DAN: IT IS UNCLEAR WHY THIS IS PROBLEMATIC IN LOC.
-CIT.
-THE TEXT BELOW IS UNCONVINCING.
-CAN YOU ILLUSTRATE THE PROBLEM WITH SOME SIMPLE EXAMPLES?
-\begin{quotation}
-There are domains, however, where the effective usage of e-graphs requires \textit{ad hoc} tweaks to achieve the desired performance, namely domains that include variable \textit{binding} and the closely related notion of \textit{substitution}, for example, lambda calculus with $\beta$-reduction.
-The first issue appears due to the fact that lambda calculus terms are quotiented by renaming of bound variables.
-If such renaming is not accounted for by the language encoding, \textit{e.g.,} by using \textit{de Brujin} indices, populating an e-graph with $\alpha$-equivalent terms increases the size of an e-graph without adding any meaningful equivalences.
-The other issue is that having binding entails having a substitution and computing substitution in e-graphs is hard as it is not a mere substitution of terms but rather a substitution of equivalence classes that represent a lot of terms at once.
-\end{quotation}
+Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
+the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when doing equality saturation for $\lambda$-calculus with e-graphs.
+A key challenge is that as e-graphs represent multiple terms at once, with as much sharing as possible, using variable names for binding means that two $\alpha$-equivalent terms can not be shared as they are not identical.
+
+\begin{red}
+DRG: UNCLEAR.
+Even though such simple encoding makes expressing $\beta$-reduction as a rewrite rule in terms of explicit substitution simpler, applying such rules requires computing extra free variable predicates and populates a given e-graph with extra nodes for explicit substitution.
+The commonly deployed solution to quotienting terms by $\alpha$-equivalence, encoding binding using $\textit{de Brujin}$ indices, still fails to represent $\alpha$-equivalent terms as being structurally equivalent for the purpose of e-graphs.
+Consider the simple example term $\lambda .
+\%0 ((\lambda .
+\%1 \%0) 42)$ the same binder is represented with $\%0$ and $\%1$, avoids explicit substitution when implementing $\beta$-reduction at the cost of including \textit{ad hoc} operators for index shifting into the rewrite rules.
+Generally, the incurred redundancy these challenges remain problematic because variable binding is not inherent to e-graphs as they do not provide an internal support for it.
+There are also other inchoate approaches to adding internal support for binding which, however, make the formulation of e-graph considerably more complicated~\cite{slotted-egraphs}.
+\end{red}
 
@@
 \subsection{String diagrams}
+
+We build on the technology of string diagrams, which is by now well developed in a number of areas. 
+String diagrams have the advantage of being a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, all based on solid mathematical foundations.
+To solve the new problems discuss above, of dealing with binding and capture avoiding substitution, we just follow the categorical breadcrumb trail.
+
+The general background literature for string diagrams is summarised in~\cite{https://arxiv.org/abs/2305.08768}.
+The particular context of string diagrams with bindings is given in~\cite{https://arxiv.org/abs/2305.18945}, while string diagrams for e-graphs without bindings are defined in~\cite{https://arxiv.org/abs/2406.15882}.
+
+String diagrams provide a geometric or topological framework for understanding monoidal categories, representing objects as \emph{strings} and morphisms as \emph{nodes}, where input strings flow in and output strings flow out.
+Key structures in monoidal categories---such as the monoidal product, tensor unit, braiding, and dualities---are represented through operations like placing, weaving, or bending strings. 
+This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by symbolic notation. 
+Originating from notations by Hotz~\cite{??} (for automata) and Penrose~\cite{??} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, offering a dual perspective to commutative diagrams through Poincaré duality.
+
+As described above, string diagrams are commonly an informal yet rigorous two-dimensional syntax for the internal language of a category. 
+They can be formalised by representing them as graphs or hypergraphs~\cite{https://dl.acm.org/doi/full/10.1145/3502719}, which has two further benefits. 
+The first benefit is that the property of two string diagrams being equal if and only if they ``look the same'' can be spelled out as graph isomorphism, eliminating any possible ambiguity of interpretation.
+The second benefit is that equational reasoning can also be formalised as (categorical) graph rewriting, for instance double-pushout rewriting (DPO).
+
+Finally, to represent diagrammatically two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{https://link.springer.com/chapter/10.1007/11874683_1}.
+The diagrammatic boxes require the introduction of further structure to (hyper)graphs in the form of a hierarchical layering of the graphs, as studied in \emph{loc. cit.}.
+
+DRG TODO: ADD AN EXAMPLE OF A STRING DIAGRAM HERE BEFORE DISCUSSING IT IN THE NEXT SECTIONS.
+BEST WE HAVE A RUNNING EXAMPLE THAT IS RELATED TO THE DIFFICULTY OF REPRESENTING E-GRAPHS FOR LAMBDA CALCULUS.
 

### 2025-01-18 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
e-graphs

@@
 
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes.
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
+E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the DAG representation of terms to include equivalence classes.
+The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, a `Hello world!' of e-graphs~\cite{EggPaper}.
 
@@
 
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent string diagram representation, an e-hypergraph as developed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-The initial term, $(a*2)/2$ (which corresponds to subfigure (a)), is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node
-\scalebox{0.3}{
+Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node
+\scalebox{0.4}{
 	\begin{tikzpicture}
@@
 Nodes of an e-graph are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
-Initially each equivalence class has a single node (subgraph).
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary (containing more than one node or subgraph) equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
+A term graph is embedded as an e-graph by making each equivalence class have a single node.
+The first rewrite rule, replacing multiplication by \emph{shift-left} creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) nodes,  now sharing the node `$a$'.
+This second e-graph illustrates the other representational difference, besides sharing, between conventional e-graphs and e-hypergraphs.
+In the former, edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
 Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes 
@@
 
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.
+The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rewrite rules.
 
-A typical way on encoding $\beta$-reduction in lambda-calculus in e-graphs is by explicit substitution~\cite{EggPaper,koehler2022sketchguided}: by adding a special substitution nodes and introduce rewrite rules for them.
-After the reduction has been done these nodes serve no purpose and can be removed.
-We show that $\beta$-reduction comes naturally as a composition when using string diagrams for enriched closed monoidal categories.
-Consider an example in Figure~\ref{fig:e-graph-substitution} where successive iterations of an e-graph for a term $(\lambda x . (y + y) + x) 1$ are shown following the $\beta$-reduction.
-First the subterm for an explicit substitution is introduced which is deemed equivalent to the top-level application term.
-It is then rewritten using the rules for explicit substitution at the top of each column.
-At the end all explicit substitution nodes can be removed.
+A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
+The auxiliary nature of these nodes is revealed by the fact that after rewrites they must be removed from the graph. 
+We will see how in the categorically-directed representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.
 
-Now consider an example from Figure~\ref{fig:e-string-substitution} which encodes the same $\beta$-reduction for the same term.
+\subsection{Conventional vs. categorical e-graphs with binding}
+
+Before diving into the technical details let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
+In Figure~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following the $\beta$-reductions.
+A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term, to be rewritten using the rules for explicit substitution at the top of each column.
+At the end, all explicit substitution nodes can be removed.
+
+The same example is given in Figure~\ref{fig:e-string-substitution} using string diagrams. 
 The $\lambda$-abstraction is represented by a rounded box and the bound variable is the dangling wire.
@@
 }. 
-The reduction is just the re-attachment of wires.
-Note how we avoided re-attachment for subterms as they did not use the bound variable wire, unlike in the case $[x \mapsto 1](y+y)$ before.
+The reduction is now simply the re-attachment of wires, and the overall process is visibly simpler. 
+DRG UNCLEAR: Note how we avoided re-attachment for subterms as they did not use the bound variable wire, unlike in the case $[x \mapsto 1](y+y)$ before.
 
-A common solution to replace explicit variable names in e-graphs is by using De Brujin indices~\cite{koehler2022sketchguided}.
+DRG UNCLEAR: A common solution to replace explicit variable names in e-graphs is by using De Brujin indices~\cite{koehler2022sketchguided}.
 The issue with this approach is that it interferes with sharing as equivalent subterms can be structurally different.
@@
 
+DRG USE WRAP-BOX. 
 \[

### 2025-01-18 LICS/sections/introduction.tex
wip

@@
 
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
+Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, \emph{e.g.}, in terms of running time or memory.
+However, applying some rewrites may enable or disable the application of subsequent rewrites, so the choice of application order significantly impacts the quality of the resulting optimisation.  
 This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
+Typical approaches to this problem use heuristics to determine an optimal ordering.
 
-A recently proposed,  alternative solution is \textit{equality saturation} 
+A recently proposed alternative solution is \textit{equality saturation} 
 \cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively}, \textit{i.e.}, forgetting the part of the program that is replaced by the rewrite, at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
 Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A na\"ive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
+A na\"ive approach for this is unfeasible, since the size of this set grows exponentially with the number of rewrites. 
+\textit{Equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making this technique computationally tractable in practical applications.
 
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple Cartesian categories enriched over a category of semilattices (\textit{semilattice-enriched}).
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
+Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations and the foundations of e-graphs in particular.
+We propose to address this gap via a categorical axiomatisation for e-graphs and their rewriting.
+Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and morphisms of Cartesian categories enriched over a category of semilattices (\textit{semilattice-enriched}).
+The morphisms of the latter can also be depicted using \emph{string diagrams}~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}: a two-dimensional syntax that encodes equivalence classes of morphism terms.
+Enriched categories generalise plain categories by imposing more structure making the categorical language more expressive~\cite{Borceux_1994,Kelly2022BASICCO} which will be useful for modelling equivalence between terms as encountered in e-graphs.
+The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022,bonchi_string_2022-1,bonchi_string_2022-2,ghica_rewriting_2023,ghica_hierarchical_2023}.
+The key ingredient of our categorical approach is the semilattice-enrichment.
+By the latter we will mean an idempotent commutative semigroup given by a binary operator but will stick to the word semilattice as it is shorter.
 By using the binary operator of a given semilattice we can operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
 
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
+Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that are already phrased in categorical language with a defined notion of rewriting and thus these applications will benefit from the mentioned optimisation technique.
+% now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
 This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
 \cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica-zanassi2023string}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
+% However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
 
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}. 
+A final contribution of this work is giving a concrete combinatorial representation of string diagrams for semilattice-enriched categories in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
 We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework 
 \cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.  
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
+As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature~\cite{EggPaper}.
 
@@
 
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
+E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.
+They generalise the typical \emph{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (subgraphs).
 The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
+This example illustrates how an optimiser using e-graphs may non-destructively apply a set of rewrites to initial term $(a * 2) / 2$, presented first in~\cite{EggPaper}.
 
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
+Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent string diagram representation: an e-hypergraph. 
 The initial term, $(a*2)/2$ (which corresponds to subfigure (a)), is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node
+The string diagram version is slightly different to the conventional e-graph version, as it represents the sharing of the node 2 explicitly, using a sharing (contraction) node
 \scalebox{0.3}{
@@
 Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
-Initially each equivalence class has a single node (subgraph).
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary (containing more than one node or subgraph) equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
+Initially, each equivalence class has a single node (subgraph).
+The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator, creates the first non-unitary (containing more than one node or subgraph) equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing: that of node $a$.
+This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs: in the former, edges connect directly to nodes inside the equivalence class, whereas in the latter, edges connect to the equivalence class itself.
 Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes 
@@
 The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.
-If we were to apply another rule, $x * 1 \to x$, we would identify the e-class of $a$ with the top-level e-class of the e-graph which would require a cycle to be introduced: the left argument of node $*$ would be the same e-class that $*$ belongs to.
-This would result into a possibility of extracting the optimal term consisting of only the node $a$, discarding everything else.
-To support such cycles we need a richer structure for our e-string diagrams.
-While equipping the e-string diagrams which such structure is straightforward, it is out of the scope of this work, and thus we limit ourselves to acyclic e-graphs and e-string diagrams for the purposes of presentation.
-Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the e-string diagrammatic version is guided by the equations of the underlying theory.
+If we were to apply another rule, $x * 1 \to x$, we would identify the e-class (dashed box) containing $a$ with the e-class containing $a * 1$ (the topmost dashed box) which would require a cycle to be introduced: the left argument of node $*$ would be the same e-class that $*$ belongs to.
+Then, a deemed optimal (as its size as a graph is minimal) term $a$ could be extracted from the e-graph by traversing all the e-classes.
+To support such cycles we need a richer categorical structure for our string diagrams.
+Such categorical structure may be added by combining the results of this work with the results of~\cite{ghica_rewriting_2023} and for the purposes of presentation we will limit ourselves to the case of acyclic e-graphs.
+% While equipping the string diagrams which such structure is straightforward, it is out of the scope of this work, and thus we limit ourselves to acyclic e-graphs and e-string diagrams for the purposes of presentation.
+Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the string diagrammatic version is guided by the equations of the underlying theory.
 A detailed step-by-step application of the equations for the rewrite rules from $(a)$ to $(b)$ and $(b)$ to $(c)$ is given in Figures~\ref{fig:e-graph-example-a-b} and~\ref{fig:e-graph-example-b-c}, respectively, in the Appendix.
 The equations applied are the ones from Figure~\ref{fig:string-equations} along with the naturality of delete and copy.
-One can note that the size of the eventual e-string diagram is comparable to the size of the corresponding e-graph.
+One can note that the size of the eventual string diagram is comparable to the size of the corresponding e-graph.
 % Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
@@
 
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category (see \textit{e.g.}~\cite{GhicaZanassiStringLambda})
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
+Recall that there is a correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category (see \textit{e.g.}~\cite{ghica-zanassi2023string}).
+% Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category (see \textit{e.g.}~\cite{ghica-zanassi2023string})
+It makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a semilattice-enriched Cartesian category, taken as string diagrams.
 We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
@@
 \]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
+Taking formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
 
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
+String diagrams are read from bottom-to-top, and we implicitly consider additional generators for the duplication and deletion transformations of a (semilattice-enriched) Cartesian category.
+Note, further, that in the Cartesian case we can restrict generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
 This means the (informal) translation given below is fully general for acyclic e-graphs.
 
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
+The translation from e-graphs to enriched string diagrams is illustrated informally in Figure~\ref{fig:e-graph-to-string},
+noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding, in each component, all unnecessary inputs.
+
+\begin{figure}
 \[
@@
 \]
-The figure on the right above is slightly simplified: if we denote the arity of each $c_i$ as $a_{c_{i}}$, then the dashed box on the right contains $\sum\limits_{i} a_{c_{i}}$ inputs and each component $c_j$ has $\sum\limits_{i \not = j} a_{c_{i}}$ weakening nodes inside.
+\caption{E-graph e-class translation to string diagram.}
+\label{fig:e-graph-to-string}
+\end{figure}
+The subfigure on the right is slightly simplified: if we denote the arity of each $c_i$ as $a_{c_{i}}$, then the dashed box on the right should contain $\sum\limits_{i} a_{c_{i}}$ inputs and each component $c_j$ has $\sum\limits_{i \not = j} a_{c_{i}}$ weakening nodes inside.
 Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
@@
 
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing as used in e-graphs.
+The motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing as used in e-graphs.
 However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
+Thus, we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
 
@@
 String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of a \textit{symmetric monoidal category (SMC)}.
-For example, we have the following equivalences between diagrams. 
+For example, we have the following equivalences between diagrams:
 %\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
@@
 \]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
+The latter representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
 A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.

### 2025-01-18 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
wip

@@
-\documentclass[peerreviewcls]{IEEEtran}
+\documentclass[]{IEEEtran}
 % \IEEEoverridecommandlockouts
@@
 
-\title{E-graphs With Bindings
+\title{E-Graphs With Bindings
 }
@@
 \begin{abstract}
-This is abstract.
+    The technique of equality saturation for program optimization and reasoning has recently gained significant attention due to the renewed popularity of equality graphs (e-graphs).
+    E-graphs represent multiple terms that are equivalent modulo a given set of rewrite rules. 
+    This capability allows rewriting to be performed over a family of terms simultaneously.    
+    However, there are domains where e-graphs fail to achieve optimal performance.
+Notably, e-graphs struggle in domains involving variable binding, such as $\lambda$-calculus, because they lack native support for binding.
+    Building on recent work that provides a categorical interpretation of e-graphs as morphisms in a semilattice-enriched symmetric monoidal category, we extend this framework.
+Specifically, we show how e-graphs with bindings can be derived as morphisms in a closed symmetric monoidal category with similar enrichment.
+This extension provides a principled approach to equality saturation for $\lambda$-calculus.
+    We then define a concrete combinatorial representation of these morphisms using hierarchical hypergraphs.
+Additionally, we introduce a corresponding notion of double-pushout (DPO) rewriting.
+Finally, we demonstrate the equivalence between term rewriting and DPO rewriting, with the key property that DPO rewriting absorbs the equations of the symmetric monoidal category.
 \end{abstract}
@@
 the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when doing equality saturation for $\lambda$-calculus with e-graphs.
-A key challenge is that as e-graphs represent multiple terms at once, with as much sharing as possible, using variable names for binding means that two $\alpha$-equivalent terms can not be shared as they are not identical.
-
-\begin{red}
+\textcolor{blue}{
+The challenge is two-fold.
+On one hand, as an e-graph represent a family of terms at once a $\beta$-reduction should also operate with families of terms which makes it a very expensive rewrite.
+On the other hand, e-graphs should represent these families with as much sharing of subterms as possible, and a particular choice of binding encoding affects this requirement.
+In this work we will focus on the second part of this challenge.
+The most simple way of encoding binding in e-graph is by using explicit variable names.
+A major drawback of such encoding is that it fails to share $\alpha$-equivalent terms as, $\textit{e.g.}$, $\lambda x .
+x$ and $\lambda y .
+y$ are structurally different and hence will be represented as different subterms in an e-graph that uses such encoding.
+$\beta$-reduction is then implemented by introducing an explicit substitution operator into the term calculus.
+When applying a $\beta$-reduction rule, such explicit substitution nodes quickly populate an e-graph which results in performance issues.
+The rewrite rule itself also becomes more complicated as there is a need to compute free variable predicates in order to apply the rule.
+}
+\textcolor{red}{
 DRG: UNCLEAR.
@@
 There are also other inchoate approaches to adding internal support for binding which, however, make the formulation of e-graph considerably more complicated~\cite{slotted-egraphs}.
-\end{red}
+}
 
@@
 \input{sections/categorical_semantics}
-
-\section{Combinatorial semantics}
-
-We can extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness.
-
-\begin{definition}
-Closed monoidal signature $\Sigma$ is a pair $(\Sigma_{O},\Sigma_{M})$ of signatures for objects and signatures for morphisms with sources and targets in $\text{obj}_{\Sigma_{O}}$ defined as
-\begin{itemize}
-  \item $I$ (monoidal unit) is in $\text{obj}_{\Sigma_{O}}$
-  \item For every $o \in \Sigma_{O}$, $o \in \text{obj}_{\Sigma_{O}}$;
-  \item For every $o_1$ and $o_2$ $ \in \text{obj}_{\Sigma_{O}}$, $o_1 \otimes o_2 \in \text{obj}{\Sigma_{O}}$ and $o_1 \multimap o_2 \in \text{obj}_{\Sigma_{O}}$
-\end{itemize}
-\end{definition}
-
-\begin{definition}
- Closed symmetric monoidal theory ($\catname{CSMT}$) over a closed monoidal signature $\Sigma$ and equations $\mathcal{E}$ denoted as $\catname{CSMT}(\Sigma, \mathcal{E})$ is given by a set of terms $T$
- \begin{align*}
- T \coloneq\; &t : A \to B \in \Sigma_{M} \;|\; t_1 \otimes t_2 : A \otimes C \to B \otimes D \text{ for } t_{1} : A \to B, t_{2} : C \to D \in T\\
-            & \;|\; t_{1};t_{2} : A \to C \text{ for } t_{1} : A \to B, t_{2} : B \to C \in T \; | \;  \Lambda(t) : A \to X \multimap B \text{ for } t : A \otimes X \to B \\
-            &  \;|\; \text{eval}(t_1 \otimes t_2) : B \text{ for } t_1 : X \multimap B, t_2 : X \in T \\
-            &  \;|\; \text{sym}_{A,B} \;|\; id_{A}
- \end{align*}
- quotiented by the laws of a closed symmetric monoidal category.
-\end{definition}
-% \question{Maybe types and terms combined is not a good idea}
-\begin{definition}
-
-We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{red}{<}, \textcolor{blue}{<},\consistency)$ where
-\begin{itemize}
-  \item $V$ is a set of vertices;
-  \item $E = \textcolor{gray}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that include \textit{plain} edges and two types of \textit{hierarchical} edges respectively;
-  \item $s,t : E \to V^{*}$ are source and target functions;
-  \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \text{obj}_{\Sigma_{O}}$ are \textit{label} functions;
-  \item $\textcolor{red}{<} : V + E \to \textcolor{red}{E}$ and $\textcolor{blue}{<} : V + E \to \textcolor{blue}{E}$ are partial \textit{child} functions that are disjoint: if $\textcolor{red}{<}(x)$ is defined then $\textcolor{blue}{<}(x)$ is undefined and vice-versa;
-  \item $\consistency = \bigcup_{p} \consistency_{p}$, where $\consistency_{p}$ is a \textit{consistency} relation defined on each set $\{x \in V + E \;|\; p \;\textcolor{red}{<^{\mu}}\; x\}$;
-\end{itemize} 
-\end{definition}
-
-$\textcolor{red}{E}$ will contain edges that encode the equivalence classes and $\textcolor{blue}{E}$ the ones that defined the lambda abstraction.
-All the above functions and relations must satisfy the obvious properties (TODO: add them).
-
-Similarly, we have a category of (extended) cospans formed by such e-hypergraphs where morphisms are of the form
-\[
-n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
-\]
-
-where $n$ and $m$ are external input and output interfaces respectively and $n'$, $m'$ are input and output internal interfaces.
-$n,n',m,m'$ are discrete ordered e-hypergraphs.
-We further require that $f_{\text{ext}}$ and $g_{\text{ext}}$ are monos and all vertices in the image of $f_{\text{ext}};f_{\text{int}}$ (respectively, $g_{\text{ext}};g_{\text{int}}$) are top-level.
-
-To build a correspondence between terms of a closed SMC and e-hypergraphs with interfaces (above) we restrict the cospans of the latter to monogamous cospans.
-
-\begin{definition}
-  We call a cospan 
-  \[
-n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
-\]
-\textit{weak} monogamous if
-\begin{itemize}
-  \item in- and out- degrees of every vertex is at most 1;
-  \item $f_{\text{int}}$ and $g_{\text{int}}$ are monos;
-  \item Vertices with in-degree (respectively, out-degree) of 0 are precisely the image of $f_{\text{int}}$ (respectively, $g_{\text{int}}$).
-\end{itemize}
-
-\end{definition}
-
-\begin{definition}[Isomorphic cospans]
-Consider a relation 
-\[
-R = \{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\}
-\]
-for $x, y \in n'$.
-And let $S$ be its reflexive closure.
-The latter partitions $n'$ into non-empty subsets $\{p_{j}\}^{k}_{j=1}$.
-We get an analogous partition for $m'$.
-
-Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute and such that $\alpha$ and $\gamma$ preserve order within each $p_j$.
-\[
-\scalebox{0.8}{
-    \tikzfig{../../figures/combinatorial_semantics/isomorphic_e_cospans}
-}
-\]
-\end{definition}
-
-Consider an example of isomorphic and non-isomorphic cospans below and their respective string diagrams
-
-\begin{figure}
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.6}{
-    \tikzfig{./figures/closed_iso_example_1}
-    }
-    \]
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.6}{
-    \tikzfig{./figures/closed_iso_example_2}
-    }
-    \]
-  \end{subfigure}
-\end{figure}
-
-\begin{definition}
-
-We define ordered sets of \textit{input} (respectively, \textit{output}) vertices of a hierarchical edge $e$ as the vertices $v_{i}$ such that $e < v_{i}$ which are in the image of $f_{\text{int}}$ (respectively, $g_{\text{ext}}$) with the induced ordering.
-\end{definition}
-
-\begin{definition}
-
-Given an ordered set $S$ of labelled vertices we can form words formed by concatenating the labels.
-We will denote such words as $w(S)$.
-\end{definition}
-
-\begin{definition}
-We call a monogamous cospan
-\[
-  n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
-\]
-\textit{well-typed} if all hierarchical edges of $\mathcal{G}$ are well-typed in the sense below.
-\begin{itemize}
-  \item For each $e \in \textcolor{red}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
-        $e$ is well-typed if for each element $S$ of the partition of $I$ $w(S) = w(s(e))$ and similarly for $O$ and $t(e)$.
-  \item For each $e \in \textcolor{blue}{E}$ consider sets $I$ and $O$ of its input and output vertices.
-        $e$ is well-typed if there exists an object $B \in \text{obj}_{\Sigma_{O}}$ such that $[...s(e), B] = w(I)$ and $B \multimap w(O) = w(t(e))$
-\end{itemize}
-\end{definition}
+\input{sections/combinatorial_semantics}
 

### 2025-01-18 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
wip

--- 2024-09-17 closed_monoidal/closed_monoidal.tex
@@
-%!LW recipe=pdflatex -> bibtex -> pdflatex * 2
-\documentclass[acmsmall, screen, nonacm]{acmart}
-
-\newcommand*{\ONECOLUMN}{}
-
-% \input{../preamble}
-
-\usepackage{hyperref}
-
-\usepackage{amsthm}
-
-\usepackage{mathtools}
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage{fontenc}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-\usepackage{xfrac}
-
-\usepackage{import}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newcommand\obj[1]{{\textbf{obj}(#1)}}
-
-\usepackage{xifthen}
-\ifthenelse{\isundefined{\ismain}}{
-  \newcommand{\ismain}{0}
-}{}
-
-\usepackage{enumitem}
-
-\newcommand{\ignora}[1]{ }
-
-% \newcounter{thm}
-
-% % \ignora{
-% \usepackage{aliascnt}
-% % alias-counters
-% \newaliascnt{definition}{thm}
-% \newaliascnt{proposition}{thm} 
-% \newaliascnt{lemma}{thm}
-% \newaliascnt{corollary}{thm}
-% \newaliascnt{conjecture}{thm}
-% \newaliascnt{remark}{thm}
-% \newaliascnt{example}{thm}
-% \newaliascnt{examples}{thm}
-% \newaliascnt{assumption}{thm}
-% \newaliascnt{construction}{thm}
-% \newaliascnt{claim}{thm}
-
-% % \theoremstyle{plain}
-% \theoremstyle{definition}
-
-% \newtheorem{theorem}[thm]{Theorem}
-% \newtheorem{proposition}[proposition]{Proposition} 
-% \aliascntresetthe{proposition}
-% \newtheorem{lemma}[lemma]{Lemma}
-% \aliascntresetthe{lemma}
-% \newtheorem{corollary}[corollary]{Corollary}
-% \aliascntresetthe{corollary}
-% \newtheorem{conjecture}[conjecture]{Conjecture}
-% \aliascntresetthe{conjecture}
-% \newtheorem{remark}[remark]{Remark}
-% \aliascntresetthe{remark}
-% \newtheorem{assumption}[assumption]{Assumption}
-% \aliascntresetthe{assumption}
-% \newtheorem{construction}[construction]{Construction}
-% \aliascntresetthe{construction}
-% \newtheorem{claim}[claim]{Claim}
-% \aliascntresetthe{claim}
-
-% \theoremstyle{definition}
-% \newtheorem{definition}[definition]{Definition}
-% \aliascntresetthe{definition}
-% \newtheorem{example}[example]{Example}
-% \aliascntresetthe{example}
-% \newtheorem{examples}[examples]{Examples}
-% \aliascntresetthe{examples}
-% }
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{remark}[theorem]{Remark}
-
-\usepackage{cleveref} % after hyperref!
-
-\input{../macros}
-\input{../sample.tikzstyles}
-\input{../hypergraph.tikzstyles}
-\input{../hypergraph.tikzdefs}
-
-\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
-\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
-\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
-\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
-
-% These are for comments
-\newcommand\question[1]{{\color{azure}#1}}
-\newcommand\update[1]{{\color{americanrose}#1}}
-
-\newcommand\enriched[1]{{\overline{\mathcal{#1}}}}
-
-\makeatletter
-\newsavebox{\@brx}
-\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
-  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
-  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
-\makeatother
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-% \setcopyright{acmlicensed}
-% \copyrightyear{2024}
-% \acmYear{2024}
-% \acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-\begin{abstract}
-  These notes contain the foundations of categories enriched over a category of semilattices and the respective string diagrams for such categories.
-  It introduces the required notions and carefully checks that all properties are satisfied.
-\end{abstract}
-
-\begin{document}
-
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
-
-\maketitle
-
-\begin{section}{Enriched Categories}
-
-\begin{definition}[Definition 6.2.1~\cite{Borceux_1994}]
-  Let $\mathcal{V} = (\mathcal{V}_{0}, I_{\mathcal{V}}, \otimes, l, r, a)$ be a monoidal category where $l$ and $r$ are left and right unitors.
-  We define a $\mathcal{V}$-enriched category $\enriched{C}$ consisting of the following data
-  \begin{enumerate}
-    \item a class of objects $\obj{\enriched{C}}$
-    \item for every pair of objects $A,B \in \obj{\enriched{C}}$ --- an (hom-)object $\obj{\enriched{C}}(A,B) \in \mathcal{V}$
-    \item for every triple of objects $A,B,C \in \obj{\enriched{C}}$ --- a composition morphism in $\mathcal{V}$
-    \[
-    c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)
-    \]
-    \item for every object $A \in \obj{\enriched{C}}$ --- a unit morphism
-    \[
-      u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)
-    \]
-  \end{enumerate}
-  such that the following diagrams commute (associativity and unitality)
-  \[
-  \begin{tikzcd}
-    {(\enriched{C}(A,B) \otimes \enriched{C}(B,C)) \otimes \enriched{C}(C,D)} \arrow[rr, "{c_{A,B,C} \otimes 1}"] \arrow[d, "{a_{\enriched{C}(A,B),\enriched{C}(B,C),\enriched{C}(C,D)}}"']   &  & {\enriched{C}(A,C) \otimes \enriched{C}(C,D)} \arrow[dd, "{c_{A,C,D}}"]                                                                                     \\
-    {\enriched{C}(A,B) \otimes (\enriched{C}(B,C) \otimes \enriched{C}(C,D))} \arrow[d, "{1\otimes c_{B,C,D}}"']                                             &  &                                                                                                                                       \\
-    {\enriched{C}(A,B) \otimes \enriched{C}(B,D)} \arrow[rr, "{c_{A,B,D}}"]                                                                       &  & {\enriched{C}(A,D)}
-  \end{tikzcd}
-  \]
-
-  \[
-  \begin{tikzcd}
-    {I_{\mathcal{V}} \otimes \enriched{C}(A,B)} \arrow[dd, "u_{A} \otimes 1"'] \arrow[r, "{l_{\enriched{C}(A,B)}}"]                                             & {\enriched{C}(A,B)} \arrow[dd, "{1_{\enriched{C}(A,B)}}"] & {\enriched{C}(A,B) \otimes I_{\mathcal{V}}} \arrow[l, "{r_{\enriched{C}(A,B)}}"'] \arrow[dd, "1 \otimes u_{B}"]                                                           \\
-    &                                     &                                                                                                                                       \\
-{\enriched{C}(A,A) \otimes \enriched{C}(A,B)} \arrow[r, "{c_{A,A,B}}"]                                                                        & {\enriched{C}(A,B)}                            & {\enriched{C}(A,B) \otimes \enriched{C}(B,B)} \arrow[l, "{c_{A,B,B}}"']
-\end{tikzcd}
-  \]
-\end{definition}
-
-\begin{remark}
-  We will occasionally use $c_{A,B,C}$ as $;$, \textit{i.e.}, instead of $c_{A,B,C}(f,g)$ we will write $f ; g$ when it makes sense to talk about components of hom-objects.
-\end{remark}
-
-\begin{remark}
-
-Requiring that a category $\mathcal{V}$ is monoidal is the most basic requirement for enriched category theory.
-To make the notion of enriched category more useful, in particular to be able to talk about a free enrichment, we will further require that $\mathcal{V}$ is a closed symmetric monoidal category.
-\end{remark}
-
-\begin{example}
-When $\mathcal{V} = \catname{Set}$ we end up with the usual definition of a category.
-$c_{A,B,C}$ becomes the usual morphism composition ($f;g$ rather than $g \circ f$) and $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$ picks out the identity morphism from the homset $\enriched{C}(A,A)$ by matching the only element of a singleton set $I \cong \{*\} \mapsto 1_{A}$ and component-wise the diagrams above commute.
-\end{example}
-
-TODO: another example
-
-\begin{definition}[Semilattice]
-  A \textit{semilattice} is a set equipped with an operation that we denote as $+$ which is associative, commutative and idempotent.
-\end{definition}
-
-Note that we do not require the existence of a unit for $+$. 
-Semilattices that satisfy this extra requirement are sometimes called \textit{bounded}, i.e., they are idempotent commutative monoids.
-
-\begin{definition}[Semilattice homomorphism]
-
-A homomorphism between two semilattices $S_{1}$ and $S_{2}$ is a map $h$ that respects $+$.
-That is, for all $s,s' \in S_{1}$, $h(s +_{S_{1}} s') = h(s) +_{S_{2}} h(s')$.
-\end{definition}
-
-\begin{definition}[Category of semilattices]
-
-Semilattices with their respective homomorphisms form a category that we denote $\catname{SLat}$.
-\end{definition}
-
-\begin{proposition}
-  $\catname{SLat}$ is a closed symmetric monoidal category.
-\end{proposition}
-\begin{proof}
-  The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
-  $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
-  \begin{itemize}
-    \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
-    \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
-  \end{itemize}
-
-  The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
-  Clearly $S \otimes I \cong S$ by mapping $(s,*) \mapsto s$ for all $s \in S$ and vice versa.
-  The symmetry, associators and unitors are then obvious morphisms.
-  Finally, the category is closed since the set of homomorpisms between two semilattices is a semilattice by defining $(f + g)(x)$ as $f(x) + g(x)$.
-  $f + g$ is a homomorphism since $(f + g)(x+y) = f(x+y) + g(x+y) = f(x) + f(y) + g(x) + g(y) = f(x) + g(x) + f(y) + g(y) = f(x+y) + g(x+y)$.
-\end{proof}
-
-This makes $\catname{SLat}$ a suitable base for enrichment and allows us to take $\mathcal{V} = \catname{SLat}$.
-Before we proceed we need to define what an enriched functor is and define what it means to freely enrich a category over $\catname{SLat}$.
-
-\begin{definition}[Isomorphism in $\enriched{C}$. Lemma 3.5.12~\cite{Riehl_2014}]
-
-Let $\enriched{C}$ be a $\mathcal{V}$-enriched category.
-Given two objects $A,B \in \enriched{C}$ and a morphism $f : I_{\mathcal{V}} \to \enriched{C}(A,B)$, we call $f$ an isomorphism if there is a $g : I_{\mathcal{V}} \to \enriched{C}(B,A)$ such that
-\[
-  I_{\mathcal{V}} \cong I_{\mathcal{V}} \otimes I_{\mathcal{V}} \xrightarrow{f \otimes g} \enriched{C}(A,B) \otimes \enriched{C}(B,A) \xrightarrow{c_{A,B,A}} \enriched{C}(A,A)
-\]
-equals $u_{A} : I_{\mathcal{V}} \to \enriched{C}(A,A)$.
-\end{definition}
-
-% TODO: I took this definition from \url{https://mathoverflow.net/questions/326575/isomorphisms-in-enriched-categories#:~:text=equals%20id%3A1%E2%86%92,C(X%2CY).} 
-% Need a proper reference. Riehl, for example, only defines isomorphism of $A$ $B$ in terms of objects of the underlying category.
-
-\begin{definition}
-  Let $\enriched{C}$ and $\enriched{D}$ be two $\mathcal{V}$-categories.
-  A $\mathcal{V}$-enriched functor $F : \enriched{C} \to \enriched{D}$ is defined by the following data.
-  \begin{enumerate}
-    \item A mapping $F : \obj{\enriched{C}} \to \obj{\enriched{D}}$
-    \item An object-indexed family of morphisms in $\mathcal{V}$ $F_{A,B} : \enriched{C}(A,B) \to \enriched{D}(FA,FB)$
-  \end{enumerate}
-  such that the following diagrams commute
-
-  \[
-  \begin{tikzcd}
-	{\enriched{C}(A,A') \otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} \\
-	\\
-	{\enriched{D}(FA,FA') \otimes \enriched{D}(FA',FA'')} && {\enriched{D}(FA,FA'')}
-	\arrow["{c_{A,A',A''}}"', from=1-1, to=1-3]
-	\arrow["{F_{A,A'} \otimes F_{A',A''}}", from=1-1, to=3-1]
-	\arrow["{F_{A,A''}}"', from=1-3, to=3-3]
-	\arrow["{c_{FA,FA',FA''}}"', from=3-1, to=3-3]
-\end{tikzcd}
-\]
-
-\[
-\begin{tikzcd}
-	I_{\mathcal{V}} && {\enriched{C}(A,A)} \\
-	\\
-	&& {\enriched{D}(FA,FA)}
-	\arrow["{u_{A}}", from=1-1, to=1-3]
-	\arrow["{u_{FA}}"', from=1-1, to=3-3]
-	\arrow["{F_{A,A}}", from=1-3, to=3-3]
-\end{tikzcd}
-\]
-\end{definition}
-
-\begin{definition}
-  Given two $\mathcal{V}$-functors $F,G : \enriched{C} \to \enriched{D}$ a $\mathcal{V}$-natural transformation is family of morphisms indexed by objects in $\enriched{C}$, $\alpha_{A} : I_{\mathcal{V}} \to \enriched{D}(FA,GA)$ in $\mathcal{V}$ such that the following diagram commutes
-  \[
-  \begin{tikzcd}
-	& {\enriched{C}(A,A')\otimes I_{\mathcal{V}}} && {\enriched{D}(FA,FA') \otimes \enriched{D}(FA',GA')} \\
-	{\enriched{C}(A,A')} &&&& {\enriched{D}(FA,GA')} \\
-	& {I_{\mathcal{V}} \otimes \enriched{C}(A,A')} && {\enriched{D}(FA,GA) \otimes \enriched{D}(GA,GA')}
-	\arrow["{G_{A,A'}\otimes \alpha_{A'}}"', from=1-2, to=1-4]
-	\arrow["{c_{FA,FA',GA'}}"', from=1-4, to=2-5]
-	\arrow["{r^{-1}}", from=2-1, to=1-2]
-	\arrow["{l^{-1}}", from=2-1, to=3-2]
-	\arrow["{\alpha_{A} \otimes F_{A,A'}}", from=3-2, to=3-4]
-	\arrow["{c_{FA,GA,GA'}}", from=3-4, to=2-5]
-\end{tikzcd}
-\]
-\end{definition}
-
-\begin{remark}
-  When $\mathcal{V}$ is $\catname{Set}$, a family of morphisms $\alpha_{A} : I_{\mathcal{V}} \to \enriched{C}(FA,GA)$ is defined by mapping $* \mapsto \alpha'_{A}$ where $\alpha'_{A} : FA \to GA$ is usual natural transformation.
-  We can then trace the diagram to get
-
-\[\begin{tikzcd}
-	& {f \otimes *} && {Gf \otimes \alpha'_{A'}} \\
-	f &&&& {Gf;\alpha'_{A} = \alpha'_{A};Ff} \\
-	& {* \otimes f} && {\alpha'_{A} \otimes Ff}
-	\arrow[from=1-2, to=1-4]
-	\arrow[from=1-4, to=2-5]
-	\arrow[from=2-1, to=1-2]
-	\arrow[from=2-1, to=3-2]
-	\arrow[from=3-2, to=3-4]
-	\arrow[from=3-4, to=2-5]
-\end{tikzcd}\]
-
-which is the usual naturality condition.
-
-\end{remark}
-
-\begin{definition}
-Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
-\end{definition}
-
-Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.
-
-\begin{proposition}
-  The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
-\end{proposition}
-\begin{proof}
-  The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
-  The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
-  The adjunction is then given by the following natural isomorphisms
-  \[
-  \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
-  \]
-  The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
-  Furthermore, we have, 
-  \[
-  F(I) \cong I_{\catname{SLat}}
-  \]
-  and 
-  \begin{align*}
-  F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
-        &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
-        &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
-        &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
-        &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
-        &\cong F(A \times B)
-  \end{align*}
-In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
-\end{proof}
-
-\begin{remark}
-  The above proposition in fact works for any symmetric monoidal closed category $\mathcal{V}$ picked as a base for enrichment (Proposition 6.4.6~\cite{Borceux_1994}) 
-\end{remark}
-
-\begin{remark}
-
-  Free semilattice on a set $S$ can be though of as a set of non-empty subsets of $S$.
-\end{remark}
-
-\begin{definition}[Proposition 6.4.7~\cite{Borceux_1994}]
-\label{def:free_slat}
-  Let $\mathcal{C}$ be a small category.
-  Freely enriched $\catname{SLat}$-category $\enriched{C}$ is defined as follows
-  \begin{enumerate}
-    \item $\obj{\enriched{C}} = \obj{\mathcal{C}}$;
-    \item $\enriched{C}(A,B) = F(\mathcal{C}(A,B))$, where $F$ is the left adjoint to $U : \catname{SLat} \to \catname{Set}$;
-    \item $c_{A,B,C} : \enriched{C}(A,B) \otimes \enriched{C}(B,C) \to \enriched{C}(A,C)$ is the composite $\cong ; F(;)$ below
-    \[
-    F(\mathcal{C}(A,B)) \otimes F(\mathcal{C}(B,C)) \xrightarrow{\cong} F(\mathcal{C}(A,B) \otimes \mathcal{C}(B,C)) \xrightarrow{F(;)} F(\mathcal{C}(A,C))
-    \] 
-    where the rightmost arrow is the image of the ordinary composition;
-    \item $u_{A} : I_{\catname{SLat}} \to \enriched{C}(A,A)$ is the composite
-    \[
-    I_{\catname{SLat}} \xrightarrow{\cong} F(I) \to F(\mathcal{C}(A,A))
-    \]
-    where the rightmost arrow is $* \mapsto 1_{A}$.
-  \end{enumerate}
-\end{definition}
-
-This is again a specialisation of a more general definition of a free $\mathcal{V}$-category (Proposition 6.4.7~\cite{Borceux_1994}).
-
-We will then consider \textit{only} such freely-enriched categories.
-
-\begin{remark}
-  Suppose $\mathcal{C}(A,B) = \{f,g\}$ and $\mathcal{C}(B,C) = \{h\}$, then $f + g \in \enriched{C}(A,B)$ and in particular $(f+g);h = f;h + g;h$ since $c_{A,B,C}$ is a semilattice homomorphism.
-\end{remark}
-
-\begin{proposition}
-\label{prop:functoriality_preserved}
-  A functor $G : \mathcal{C} \to \mathcal{D}$ induces an enriched functor $\overline{G} : \enriched{C} \to \enriched{D}$.
-\end{proposition}
-\begin{proof}
-  We define
-  \begin{itemize}
-    \item $\overline{G}(A) = G(A)$ for all $A \in \enriched{C}$
-    \item $\overline{G}_{A,B}(\enriched{C}(A,B)) = F(G_{A,B}(\mathcal{C}(A,B)))$
-  \end{itemize}
-  The functoriality follows from the following diagrams
-\[
-\adjustbox{width=\linewidth}{
-\begin{tikzcd}
-	{\enriched{C}(A,A')\otimes \enriched{C}(A',A'')} && {\enriched{C}(A,A'')} && {(f, g)} && {f;g} \\
-	\\
-	{\enriched{D}(GA,GA') \otimes \enriched{D}(GA',GA'')} && {\enriched{D}(GA,GA'')} && {(Gf,Gg)} && {Gf;Gg = G(f;g)}
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=1-1, to=3-1]
-	\arrow[from=1-3, to=3-3]
-	\arrow[from=1-5, to=1-7]
-	\arrow[from=1-5, to=3-5]
-	\arrow[from=1-7, to=3-7]
-	\arrow[from=3-1, to=3-3]
-	\arrow[from=3-5, to=3-7]
-\end{tikzcd}}
-\]
-where the right-hand diagram is a component render of the left-hand diagram.
-It commutes because $G$ is a functor. 
-By freeness, the commutativity of more complex components follows from the commutativity for generators.
-The following diagrams commute for the same reason.
-\[\begin{tikzcd}
-	I_{\catname{SLat}} && {\enriched{C}(A,A)} && {*} && {1_{A}} \\
-	\\
-	&& {\enriched{D}(GA,GA)} &&&& {G(1_{A}) = 1_{G(A)}}
-	\arrow[from=1-1, to=1-3]
-	\arrow[from=1-1, to=3-3]
-	\arrow[from=1-3, to=3-3]
-	\arrow[from=1-5, to=1-7]
-	\arrow[from=1-5, to=3-7]
-	\arrow[from=1-7, to=3-7]
-\end{tikzcd}\]
-\end{proof}
-
-\begin{remark}
-  Again, because $\overline{G}_{A,B}$ is a semilattice homomorphism, we have $\overline{G}(f+g) = \overline{G}(f) + \overline{G}(g)$ for any $\catname{SLat}$-enriched functor $\overline{G}$.
-\end{remark}
-
-\begin{example}
-  Let $\mathcal{C} = (\mathcal{C}_{0},I,\otimes, l, r, \alpha)$ be a monoidal category and consider the bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to \mathcal{C}$ on $\mathcal{C}$.
-  It induces an enriched monoidal category $\enriched{C} = (\enriched{C}_{0}, I, \enriched{\otimes}, \overline{l}, \overline{r}, \enriched{\alpha})$ such that
-  \begin{enumerate}
-    \item $A \overline{\otimes} B \coloneq A \otimes B$;
-    \item $\overline{l} : A \overline{\otimes} I \to A \coloneq I \to \enriched{C}(A \overline{\otimes} I, I)$, i.e. $\overline{l}$ picks out $l$ out of $\enriched{C}(A \overline{\otimes} I, I)$;
-    \item similarly for $\overline{r}$, $\overline{\alpha}$.
-  \end{enumerate}
-  $\otimes_{A,B,C,D} : \mathcal{C}(A,B) \times \mathcal{C}(C,D) \to \mathcal{C}(A \otimes C, B \otimes D)$ becomes $\overline{\otimes}_{A,B,C,D} : \enriched{C}(A,B) \times \enriched{C}(C,D) \to \enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D)$.
-  In particular, $f + g \;\overline{\otimes}\; h + k = (f \;\overline{\otimes}\; h + k) + (g \;\overline{\otimes}\; h + k) = (f \;\overline{\otimes}\; k) + (f \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; h) + (g \;\overline{\otimes}\; k)$, where $+$ binds tighter than $\overline{\otimes}$.
-  \question{I need to show that $\overline{\otimes}$ is a bifunctor.
-Functoriality in each argument does not entail bifunctoriality.
-This is the difference between monoidal and pre-monoidal categories.
-  }
-\end{example}
-
-\begin{proposition}[Proposition 6.2.8~\cite{Borceux_1994}]
-  When $F,G$ are $\mathcal{V}$-valued $\mathcal{V}$-functors, the notion of $\mathcal{V}$-natural transformation reduces to
-  $\alpha_{A} : FA \to GA$ and the following diagram
-  \[\begin{tikzcd}
-	{C(A,B)} && {[FA,FB]} \\
-	\\
-	{[GA,GB]} && {[FA,GB]}
-	\arrow["{F_{A,B}}", from=1-1, to=1-3]
-	\arrow["{G_{A,B}}"', from=1-1, to=3-1]
-	\arrow["{[1,\alpha_{B}]}", from=1-3, to=3-3]
-	\arrow["{[\alpha_{A},1]}"', from=3-1, to=3-3]
-\end{tikzcd}\]
-\end{proposition}
-
-Note the flip of the domain and co-domain of the bottom arrow.
-This is because the functor $[-,-] : \enriched{C}^{\text{op}} \otimes \enriched{C} \to \mathcal{V}$ is contravariant in its first argument.
-
-\begin{definition}[$\mathcal{V}$-enriched adjunction]
-  Let $\mathcal{V}$ be a closed symmetric monoidal category and $\enriched{C}$, $\enriched{D}$ two $\mathcal{V}$-categories.
-  Two $\mathcal{V}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$ are $\mathcal{V}$-adjoint, $\overline{L}$ left adjoint to $\overline{R}$ and $\overline{R}$ right adjoint to $\overline{L}$, when there exist isomorphisms in $\mathcal{V}$
-  \[
-  \enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B) \;,
-  \]
-  which are $\mathcal{V}$-natural in $A$ and $B$.
-\end{definition}
-
-Equivalently, one can give a definition of an adjunction in terms of unit and co-unit natural transformations.
-
-\begin{definition}[Section 1.11 in~\cite{Kelly2022BASICCO}]
-  A $\mathcal{V}$-adjunction $\eta, \epsilon: \overline{L} \;\reflectbox{$\vdash$}\; \overline{R} : \enriched{C} \to \enriched{D}$ between functors $L : \enriched{D} \to \enriched{C}$ and $\enriched{R} : \enriched{C} \to \enriched{D}$ consists of $\mathcal{V}$-natural $\eta : 1 \to RL$ (the \textit{unit}) and $\epsilon : LR \to 1$ (the \textit{co-unit}) satisfying the triangular equations
-  \[
-  \overline{R}\epsilon . \eta_{\overline{L}} = 1 \qquad \epsilon_{\overline{L}} . \overline{R} \eta = 1
-  \]
-\end{definition}
-Furthermore we have that the isomorphism $\alpha : \enriched{C}(\overline{L}A,B) \to \enriched{D}(A, \overline{R}B)$ has the form $\enriched{D}(\eta, 1)\overline{R}$:
-\[
-  \alpha = \enriched{C}(\overline{L}A, B) \xrightarrow{\overline{R}} \enriched{D}(\overline{RL}A, \overline{R}B) \xrightarrow{\enriched{D}(\eta, 1)} \enriched{D}(A, \enriched{R}B)
-\]
-and $\eta = u_{\overline{L}A};\alpha_{\overline{L}A} = I \xrightarrow{u_{\overline{L}A}}\enriched{C}(\overline{L}A, \overline{L}A) \xrightarrow{\alpha_{\overline{L}A}} \enriched{D}(A,\overline{RL}A)$.
-Similarly, $\alpha^{-1} = \enriched{C}(1,\epsilon)\overline{L}$ and $\epsilon = u_{\overline{R}B};\alpha^{-1}_{\overline{R}B}$.
-
-\begin{proposition}
-\label{prop:naturality_preserved}
-  For a natural transformation $\alpha : T \to S$ for $T,S : \mathcal{C} \to \mathcal{D}$ there is a $\catname{SLat}$-natural transformation $\overline{\alpha} : \overline{T} \to \overline{S}$ where $\overline{T},\overline{S} : \enriched{C} \to \enriched{D}$.
-\end{proposition}
-\begin{proof}
-  Recall that an ordinary natural transformation is in fact a $\catname{Set}$-enriched natural transformation and has the form $\alpha_{A} : I \to \mathcal{D}(TA,SA)$.
-  We then obtain an $\catname{SLat}$-natural transformation $\overline{\alpha_{A}} = F(\alpha_{A}) : F(I) \to F(\mathcal{D}(TA,SA))$, or $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(TA,SA)$ given as $* \mapsto F\alpha_{A}$.
-  Again, we can use $TA$ or $\overline{T}A$ interchangeably as their action on objects is the same.
-  % The component of a natural transformation $\overline{\alpha}_{A}$ is given by $I_{\catname{SLat}} \to \enriched{D}(\overline{T}A, \overline{S}A)$, i.e.
-a morphism that picks out the original $\alpha_{A}$.
-  We need to check if so defined $\overline{\alpha}$ is $\catname{SLat}$-natural.
-  This follows from the following commutativity diagrams
-  \[
-  \adjustbox{width=\linewidth}{
-  \begin{tikzcd}
-	&& {\enriched{C}(A,B) \otimes I_{\catname{SLat}}} && {\enriched{D}(\overline{T}A,\overline{T}B) \otimes \enriched{D}(\overline{T}B, \overline{S}B)} &&&&& {f \otimes *} && {\overline{T}f \otimes F\alpha_{B}} \\
-	{\enriched{C}(A,B)} &&&&&& {\enriched{D}(\overline{T}A,\overline{S}B)} && f &&&& {F\alpha_{A};\overline{S}f = \overline{T}f ;F\alpha_{B}} \\
-	&& {I_{\catname{SLat}} \otimes \enriched{C}(A,B)} && {\enriched{D}(\overline{T}A,\overline{S}A) \otimes \enriched{D}(\overline{S}A,\overline{S}B)} &&&&& {* \otimes f} && {F\alpha_{A} \otimes \overline{S}f}
-	\arrow["{\overline{T}_{A,B} \otimes \overline{\alpha_{B}}}", from=1-3, to=1-5]
-	\arrow["{c_{\overline{T}A,\overline{T}B,\overline{S}B}}", from=1-5, to=2-7]
-	\arrow[from=1-10, to=1-12]
-	\arrow[from=1-12, to=2-13]
-	\arrow[from=2-1, to=1-3]
-	\arrow[from=2-1, to=3-3]
-	\arrow[from=2-9, to=1-10]
-	\arrow[from=2-9, to=3-10]
-	\arrow["{\overline{\alpha_{A}} \otimes \overline{S}(A,B)}", from=3-3, to=3-5]
-	\arrow["{c_{\overline{T}A,\overline{S}A,\overline{T}B}}"', from=3-5, to=2-7]
-	\arrow[from=3-10, to=3-12]
-	\arrow[from=3-12, to=2-13]
-\end{tikzcd}}
-\]
-The diagram on the right is the component-wise rendering of the diagram on the left.
-It commutes because it is essentially a lifting of the following diagram (naturality of $\alpha$) using free enrichment functor $F$.
-
-\[
-\begin{tikzcd}
-	TA && SA && {\overline{T}A} && {\overline{S}A} \\
-	\\
-	TB && SB && {\overline{T}B} && {\overline{S}B}
-	\arrow["{\alpha_{A}}"', from=1-1, to=1-3]
-	\arrow["Tf", from=1-1, to=3-1]
-	\arrow["Sf"', from=1-3, to=3-3]
-	\arrow["{F\alpha_{A}}"', from=1-5, to=1-7]
-	\arrow["{\overline{T}f}", from=1-5, to=3-5]
-	\arrow["{\overline{S}A}"', from=1-7, to=3-7]
-	\arrow["{\alpha_{B}}", from=3-1, to=3-3]
-	\arrow["{F\alpha_{B}}", from=3-5, to=3-7]
-\end{tikzcd}
-\]
-
-The commutativity for more complex elements of a semilattice like $f + g$ follows component-wise.
-
-\end{proof}
-
-\begin{example}
-
-Consider a symmetry natural transformation for a given symmetric monoidal category $\mathcal{C}$ $\sigma_{A,B} : A \otimes B \to B \otimes A$ and its freely $\catname{SLat}$ enriched version $\overline{\sigma_{A,B}} : I_{\catname{SLat}} \to \enriched{C}(A \overline{\otimes} B, B \overline{\otimes} A)$.
-$\catname{SLat}$-naturality takes the following form
-
-\[
-\adjustbox{width=\linewidth}{
-\begin{tikzcd}
-	&& {(\enriched{C}(A,B) \times \enriched{C}(C,D)) \otimes I} && {\enriched{C}(A \overline{\otimes} C, B \overline{\otimes} D) \otimes \enriched{C}(B \overline{\otimes} D, D \overline{\otimes} B)} \\
-	{\enriched{C}(A,B) \times \enriched{C}(C,D)} &&&&&& {\enriched{C}(A \overline{\otimes} C, D \overline{\otimes} B)} \\
-	&& {I \otimes (\enriched{C}(A,B) \times \enriched{C}(C,D))} && {\enriched{C}(A \overline{\otimes} C, C \overline{\otimes} A) \otimes \enriched{C}(C \overline{\otimes} A, D \overline{\otimes} B)} \\
-	\\
-	&& {(f,g) \otimes *} && {(f \overline{\otimes} g) \otimes \sigma} \\
-	{(f,g)} &&&&&& {\sigma;(g \overline{\otimes}f) = (f\overline{\otimes}g);\sigma} \\
-	&& {* \otimes (f,g)} && {\sigma \otimes (g \overline{\otimes} f)}
-	\arrow["{\overline{\otimes} \otimes \overline{\sigma}_{B,D}}", from=1-3, to=1-5]
-	\arrow[from=1-5, to=2-7]
-	\arrow[from=2-1, to=1-3]
-	\arrow[from=2-1, to=3-3]
-	\arrow["{\overline{\sigma}_{C,A} \otimes \overline{\otimes}}"', from=3-3, to=3-5]
-	\arrow[from=3-5, to=2-7]
-	\arrow[from=5-3, to=5-5]
-	\arrow[from=5-5, to=6-7]
-	\arrow[from=6-1, to=5-3]
-	\arrow[from=6-1, to=7-3]
-	\arrow[from=7-3, to=7-5]
-	\arrow[from=7-5, to=6-7]
-\end{tikzcd}}\]
-
-Given $(f_1 + f_2, g)$ we have $((f_1 + f_2) \;\overline{\otimes}\; g);\sigma = (f_1 \;\overline{\otimes}\; g) + (f_{2} \;\overline{\otimes}\; g);\sigma = (f_{1} \;\overline{\otimes}\; g);\sigma + (f_{2} \;\overline{\otimes}\; g);\sigma = \sigma;(g \;\overline{\otimes}\; f_{1}) + \sigma;(g \;\overline{\otimes}\; f_{2})$
-
-\end{example}
-
-\begin{proposition}
-Functor $F : \catname{Set} \to \catname{SLat}$ induces a 2-functor $F^{2}: \catname{Cat} \to \catname{SLat}-\catname{Cat}$ from the category of small categories to the category of small categories enriched over a category of semilattices.
-\end{proposition}
-\begin{proof}
-  The functoriality of $F^{2}$ is given by Definition~\ref{def:free_slat} and 2-functoriality is implied by Propositions~\ref{prop:functoriality_preserved} and \ref{prop:naturality_preserved}.
-\end{proof}
-
-The above proposition means, in particular, that an ordinary adjunction gives rise to a corresponding $\catname{SLat}$-enriched adjuntion.
-The latter, however, could be also shown equivalently as below.
-
-\begin{proposition}
-  Given an adjunction $\mathcal{C}(LA,B) \cong \mathcal{D}(A,RB)$ of plain functors $L : \mathcal{D} \to \mathcal{C}$ and $R : \mathcal{C} \to \mathcal{D}$, we also get an $\catname{SLat}$-enriched adjunction $\enriched{C}(\overline{L}A,B) \cong \enriched{D}(A, \overline{R}B)$ of $\catname{SLat}$-functors $\overline{L} : \enriched{D} \to \enriched{C}$ and $\overline{R} : \enriched{C} \to \enriched{D}$.
-  Because $\overline{L}$ acts on objects as $L$ we will further omit the bar.
-\end{proposition}
-\begin{proof}
-  The first adjunction implies the existence of natural isomorphisms (in $A$ and $B$) between $\mathcal{C}(L-,B)$ and $\mathcal{D}(-,RB)$ and $\mathcal{C}(LA,-)$ and $\mathcal{D}(A,R-)$ respectively.
-  Naturality in $B$ is by definition the commutativity of the following diagram
-  \begin{equation}
-    \label{diag:naturality_plain}
-  \begin{tikzcd}
-	{\mathcal{C}(LA,B)} && {\mathcal{D}(A,RB)} \\
-	\\
-	{\mathcal{C}(LA,B')} && {\mathcal{D}(A,RB')}
-	\arrow["{\alpha_{B}}"', from=1-1, to=1-3]
-	\arrow["{\mathcal{C}(LA,f)}"', from=1-1, to=3-1]
-	\arrow["{\mathcal{D}(A,Rf)}", from=1-3, to=3-3]
-	\arrow["{\alpha_{B'}}", from=3-1, to=3-3]
-  \end{tikzcd}
-\end{equation}
-  and similarly for the naturality in $A$.
-  The claim is that $\enriched{C}(LA,B) \cong \enriched{D}(A,RB)$ is an $\catname{SLat}$-adjunction.
-  Recall that $\enriched{C}(LA,B) = F(\mathcal{C}(LA,B))$ and similarly for another hom-set and clearly the isomorphism is preserved: given $\alpha$, an isomorphism from $\mathcal{C}(LA,B) \to \mathcal{D}(A,RB)$, the isomorphism from $\enriched{C}(LA,B) \to \enriched{D}(A,RB)$ is $\enriched{\alpha} = F(\alpha)$.
-  We need to check that this $\enriched{\alpha}$ is natural.
-  This is equivalent to checking if the following diagram commutes
-  \[\begin{tikzcd}
-	{\enriched{C}(B,B')} && {[\enriched{C}(LA,B),\enriched{C}(LA,B')]} \\
-	\\
-	{[\enriched{D}(A,RB),\enriched{D}(A,RB')]} && {[\enriched{C}(A,RB),\enriched{D}(A,RB')]}
-	\arrow["{\enriched{C}(LA,-)}", from=1-1, to=1-3]
-	\arrow["{\enriched{D}(A,R-)}"', from=1-1, to=3-1]
-	\arrow["{[1,\overline{\alpha}_{B'}]}", from=1-3, to=3-3]
-	\arrow["{[\overline{\alpha_{B}},1]}"', from=3-1, to=3-3]
-\end{tikzcd}\]
-which component-wise becomes
-
-\[
-\begin{tikzcd}
-	f & {\enriched{C}(LA,f)} \\
-	{\enriched{D}(A,Rf)} & {\overline{\alpha}_{B};\enriched{D}(A,Rf);1 = 1;\enriched{C}(LA,f);\overline{\alpha}_{B}'}
-	\arrow[from=1-1, to=1-2]
-	\arrow[from=1-1, to=2-1]
-	\arrow[from=1-2, to=2-2]
-	\arrow[from=2-1, to=2-2]
-\end{tikzcd}
-\]
-
-The commutativity of the diagram above follows from applying $F$ to the diagram \ref{diag:naturality_plain}:
-\[\begin{tikzcd}
-	{\enriched{C}(LA,B)} && {\enriched{D}(A,RB)} \\
-	\\
-	{\enriched{C}(LA,B')} && {\enriched{D}(A,RB')}
-	\arrow["{\overline{\alpha}_{B}}", from=1-1, to=1-3]
-	\arrow["{\enriched{C}(LA,f)}"', from=1-1, to=3-1]
-	\arrow["{\enriched{D}(A,Rf)}", from=1-3, to=3-3]
-	\arrow["{\overline{\alpha}_{B'}}", from=3-1, to=3-3]
-\end{tikzcd}
-\]
-The commutativity for more complex elements of $\enriched{C}(B,B')$, \textit{e.g.} $f + g$, follows component-wise.
-Similarly for naturality in $A$.
-Then, $\catname{Slat}$-naturality separately in $A$ and $B$ implies $\catname{SLat}$-naturality in both $A$ and $B$, a consequence of a more general statement about $\mathcal{V}$-naturality ((1.21) in~\cite{Kelly2022BASICCO})
-
-\end{proof}
-
-\end{section}
-
-\begin{section}{String diagrams for freely $\catname{SLat}$-enriched symmetric monoidal categories}
-
-String diagrams are a two-dimensional syntax for describing morphisms in symmetric monoidal categories which is typically given by the following grammar parameterised over a monoidal signature $\Sigma$ of symbols.
-The labels on wires specify the types of the morphisms.
-
-\[
-\adjustbox{scale=0.7}{
-\tikzfig{./figures/string_diagrams}
-}
-\]
-
-The diagrams constructed by using the rules above are then subject to SMC equations.
-
-By freely enriching home-sets to hom-objects in $\catname{SLat}$ we still can access the morphisms within these objects.
-To support `new' morphisms we extend the grammar by adding a production for $f + g$
-
-\[
-\adjustbox{scale=0.8}{
-\tikzfig{./figures/f_plus_g}
-}
-\]
-
-Once we consider string diagrams for a $\catname{SLat}$-enriched category $\enriched{C}$, $;$ and $\otimes$ become $c_{A,B,C}$ and $\overline{\otimes}$ respectively.
-Since the latter are semilattice homomorphisms, we have the following equations (with labels on wires omitted) for the newly introduced dashed boxes.
-
-\[
-\adjustbox{width=\linewidth}{
-\tikzfig{./figures/egraph-strings-equations}
-}
-\]
-
-The fact of $+$ being associative and commutative justifies the usage of $\ldots$ in the equations.
-
-Then we can reuse the diagrams for functorial boxes~\cite{GhicaZanassiStringLambda}.
-Given a functor $F : \mathcal{C} \to \mathcal{D}$ and a morphism $f \in C(A,B)$ we can depict the application of $Ff$ as below (left) together with the functoriality equations in Figure~\ref{fig:functoriality_laws}
-
-\[
-  \adjustbox{scale=0.7}{
-  \tikzfig{./figures/functorial_string}
-  }
-\]
-
-\begin{figure}
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.7}{
-    \tikzfig{./figures/functoriality_eq_1_string}
-    }
-    \]
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.7}{
-    \tikzfig{./figures/functoriality_eq_2_string}
-    }
-    \]
-  \end{subfigure}
-  \caption{String diagrammatic equations for functoriality}
-  \label{fig:functoriality_laws}
-\end{figure}
-\end{section}
-
-A functor $\overline{F}$ induced by the functor $F$ has the same diagrammatic rendering with the addition of an extra equation.
-
-\[
-\adjustbox{scale=0.7}{
-\tikzfig{./figures/functoriality_enriched_eq}
-}
-\]
-
-Since a component of a natural transformation $\overline{\alpha}$ between two $\catname{SLat}$-enriched functors $\overline{F} : \enriched{C} \to \enriched{D}$ and $\overline{G}: \enriched{C} \to \enriched{D}$ has the form $\overline{\alpha_{A}} : I_{\catname{SLat}} \to \enriched{D}(FA, GA)$, it can be identified with the element of $\enriched{D}(FA,GA)$, i.e.
-we can use the syntax for plain morphism to depict components of natural transformations.
-This allows us to reuse the whole syntax machinery for Cartesian closed categories for their free $\catname{SLat}$-enriched version which induce an additional equation
-
-\[
-\adjustbox{scale=0.7}{
-  \tikzfig{./figures/lambda_enriched_eq}
-}
-\]
-
-We can similarly lift a trace from a traced monoidal category $\mathcal{C}$ to $\enriched{C}$.
-Recall that a trace is an operator
-\[
-\text{tr}^{X} : \mathcal{C}(A \otimes X, B \otimes X) \to \mathcal{C}(A,B)
-\]
-\textit{i.e.}, a mapping between hom-sets, such that it satisfies the equations
-\begin{enumerate}
-  \item \textit{Tightening} (naturality in $A,B$) For all $A, B, C, D, X \in \mathcal{C}$, $g : A \to B$, $f : B \otimes X \to C \otimes X$, $h : C \to D$
-  \[
-  \text{tr}^{X}((h \otimes id_{X});f;(g \otimes id_{X})) = h;\text{tr}^{X}(f);g
-  \]
-  \item \textit{Sliding} (naturality in $X$) For all $A,B,X,Y \in \mathcal{C}$, $f : A \otimes X \to B \otimes Y$, $g : Y \to X$
-  \[
-   \text{tr}^{X}(f;(id_{B} \otimes g)) = \text{tr}^{Y}((id_{A} \otimes g); f) 
-  \]
-  \item \textit{Vanishing} For all $A,B,X,Y \in \mathcal{C}$,
-  \[
-  \text{tr}^{I}(f) = f,\; \forall f : A \to B \qquad \text{and} \qquad \text{tr}^{X \otimes Y}(f) = \text{tr}^{X}(\text{tr}^{Y}(f)), \; \forall f : A \otimes X \otimes Y \to B \otimes X \otimes Y
-  \]
-  \item \textit{Strength} For all $A,B,C,D,X \in \mathcal{C}$, $f : C \otimes X \to D \otimes X$, $g : A \to B$
-    \[
-    \text{tr}^{X}(g \otimes f) = g \otimes \text{tr}^{X}(f)
-    \] 
-  \item \textit{Yanking} (when $\mathcal{C}$ is symmetric)
-    \[
-    \text{tr}^{X}(\sigma_{X,X}) = id_{X}
-    \]
-\end{enumerate}
-
-By freely enriching $\mathcal{C}$ we obtain an enriched trace
-\[
-\text{tr}^{X} : \enriched{C}(A \overline{\otimes} X, B \overline{\otimes} X) \to \enriched{C}(A,B)
-\]
-and all the equations are preserved with the addition of a new equation
-\[
-  \text{tr}^{X}(f + g) = \text{tr}^{X}(f) + \text{tr}^{X}(g)
-\]
-for $f,g : A \overline{\otimes} X \to B \overline{\otimes} X$
-
-\section{E-Hypergraphs for $\catname{SLat}$-enriched symmetric closed monoidal categories}
+\section{Combinatorial semantics}
 
@@
 \end{definition}
-
-% \[
-% \adjustbox{scale=0.95, lap={\width}{-12em}}{
-%     \begin{tikzcd}
-%         & {\enriched{C}(A_1,A_2)} \arrow[ld] \arrow[rd]                       &                                                                                                                                                                                                            \\
-% {I \otimes \enriched{C}(A_1,A_2)} \arrow[dd, "{\alpha'_{A_1} \otimes \overline{\mathcal{D}}(\overline{L}A,-)}"]                                                                    &                                                                               & {\enriched{C}(A_1,A_2) \otimes I} \arrow[dd, "{\enriched{C}(\overline{L}A,-) \otimes \alpha'_{A_2}}"]                                                                                  \\
-%         &                                                                               &                                                                                                                                                                                                            \\
-% {\mathcal{V}(\enriched{C}(\overline{L}A,A_1),\overline{\mathcal{D}}(A,A_1)) \otimes \mathcal{V}(\overline{\mathcal{D}}(A,A_1),\overline{\mathcal{D}}(A,A_2)))} \arrow[rd, "\circ"] &                                                                               & {\mathcal{V}(\enriched{C}(\overline{L}A,A_1),\enriched{C}(\overline{L}A,A_2))) \otimes \mathcal{V}(\enriched{C}(LA,A_2),\overline{\mathcal{D}}(A,A_2))} \arrow[ld, "\circ"'] \\
-%         & {\mathcal{V}(\enriched{C}(\overline{L}A,A_1), \overline{D}(A,A_2))} &
-% \end{tikzcd}
-% }
-% \]
-
-% \[
-%   \begin{tikzcd}
-%     {\enriched{C}(B,B')} \arrow[rr, "{\enriched{C}(LA,-)}"] \arrow[d, "{\overline{\mathcal{D}}(A,R-)}"] &  & {[\enriched{C}(LA,B), \enriched{C}(LA,B')]} \arrow[d, "{[1,\overline{\alpha_{B'}}]}"] \\
-%     {[\overline{\mathcal{D}}(A,RB),\overline{\mathcal{D}}(A,RB')]} \arrow[rr, "{[\overline{\alpha_{B}},1]}"]                &  & {[\enriched{C}(LA,B),\overline{\mathcal{D}}(A,RB')]}
-%     \end{tikzcd}
-% \]
-
-% \vspace{4em}
-
-% \[
-%   \begin{tikzcd}
-%     {\enriched{C}(B,B')} \arrow[rr, "{\enriched{C}(LA,-)}"] \arrow[d, "{\overline{\mathcal{D}}(A,R-)}"] &  & {[\enriched{C}(LA,B), \enriched{C}(LA,B')]} \arrow[d, "{[1,\overline{\alpha_{B'}}]}"]                             \\
-%     {[\overline{\mathcal{D}}(A,RB),\overline{\mathcal{D}}(A,RB')]} \arrow[rr, "{[\overline{\alpha_{B}},1]}"]                &  & {[\enriched{C}(LA,B),\overline{\mathcal{D}}(A,RB')]}                                                                        \\
-%     f \arrow[rr] \arrow[d]                                                                                                  &  & {\enriched{C}(LA,f)} \arrow[d]                                                                                              \\
-%     {\overline{\mathcal{D}}(A,Rf)} \arrow[rr]                                                                               &  & {1;\enriched{C}(LA,f);\overline{\alpha_{B'}} = \overline{\alpha_{B}};\overline{\mathcal{D}}(A,Rf);1}
-%     \end{tikzcd}
-% \]
-
-\bibliographystyle{acm}
-\bibliography{bibliography}
-
-\end{document}

### 2025-01-18 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
introduction

@@
 Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
-the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when doing equality saturation for $\lambda$-calculus with e-graphs.
-\textcolor{blue}{
-The challenge is two-fold.
-On one hand, as an e-graph represent a family of terms at once a $\beta$-reduction should also operate with families of terms which makes it a very expensive rewrite.
-On the other hand, e-graphs should represent these families with as much sharing of subterms as possible, and a particular choice of binding encoding affects this requirement.
-In this work we will focus on the second part of this challenge.
-The most simple way of encoding binding in e-graph is by using explicit variable names.
-A major drawback of such encoding is that it fails to share $\alpha$-equivalent terms as, $\textit{e.g.}$, $\lambda x .
-x$ and $\lambda y .
-y$ are structurally different and hence will be represented as different subterms in an e-graph that uses such encoding.
-$\beta$-reduction is then implemented by introducing an explicit substitution operator into the term calculus.
-When applying a $\beta$-reduction rule, such explicit substitution nodes quickly populate an e-graph which results in performance issues.
-The rewrite rule itself also becomes more complicated as there is a need to compute free variable predicates in order to apply the rule.
-}
-\textcolor{red}{
-DRG: UNCLEAR.
-Even though such simple encoding makes expressing $\beta$-reduction as a rewrite rule in terms of explicit substitution simpler, applying such rules requires computing extra free variable predicates and populates a given e-graph with extra nodes for explicit substitution.
-The commonly deployed solution to quotienting terms by $\alpha$-equivalence, encoding binding using $\textit{de Brujin}$ indices, still fails to represent $\alpha$-equivalent terms as being structurally equivalent for the purpose of e-graphs.
-Consider the simple example term $\lambda .
-\%0 ((\lambda .
-\%1 \%0) 42)$ the same binder is represented with $\%0$ and $\%1$, avoids explicit substitution when implementing $\beta$-reduction at the cost of including \textit{ad hoc} operators for index shifting into the rewrite rules.
-Generally, the incurred redundancy these challenges remain problematic because variable binding is not inherent to e-graphs as they do not provide an internal support for it.
-There are also other inchoate approaches to adding internal support for binding which, however, make the formulation of e-graph considerably more complicated~\cite{slotted-egraphs}.
-}
+the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise applying equality saturation to $\lambda$-calculus e-graphs.
 
-In this paper we take what we consider to be a principled approach to e-graphs with bindings by lifting to monoidal closed categories prior work showing how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}
+In the textual representation of $\lambda$ terms, $\alpha$-equivalent terms are distinct and thus cannot be shared, which is not only problematic but also inefficient. 
+Using De Bruijin indices is also problematic because the same variable can be assigned different indices depending on under the scope of how many binders it lies. 
+This again interferes with sharing. 
+As a simple example, consider the term $\lambda f . f ((\lambda x . f x) 2)$ which is encoded as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$.
+Notice how two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$, which cannot be shared in an e-graph.
+
+Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
+However, calculi of explicit substitutions have many technical subtleties which also directly impact how equivalence is defined, for instance \emph{graph equivalence} instituted between terms which have non-interfering explicit substitutions applied in different orders~\cite{doi:10.1145/2535838.2535886}.
+These difficulties are not handled in \emph{loc. cit.}.
+
+All these issues can be resolved by the use of \emph{string diagrams}.
+Thus, in this paper we take what we consider to be a principled approach to e-graphs with bindings by lifting to monoidal closed categories prior work showing how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}
 The approach in \emph{loc.
@@
 
-DRG TODO: ADD AN EXAMPLE OF A STRING DIAGRAM HERE BEFORE DISCUSSING IT IN THE NEXT SECTIONS.
-BEST WE HAVE A RUNNING EXAMPLE THAT IS RELATED TO THE DIFFICULTY OF REPRESENTING E-GRAPHS FOR LAMBDA CALCULUS.
+Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, note its string diagrammatic representation below, in which the two occurrences of $f$ are now shared.
+$\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire attached to the frame,
+the application node a half-circle
+\adjustbox{scale=0.15}{\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=none] (0) at (1, 6.75) {};
+		\node [style=none] (1) at (1.25, 7.75) {};
+		\node [style=none] (2) at (2, 8) {};
+		\node [style=none] (3) at (2.75, 7.75) {};
+		\node [style=none] (4) at (3, 6.75) {};
+		\node [style=none] (5) at (1.5, 6.75) {};
+		\node [style=none] (6) at (2.5, 6.75) {};
+		\node [style=none] (7) at (2, 9) {};
+		\node [style=none] (8) at (1.5, 5.75) {};
+		\node [style=none] (9) at (2.5, 5.75) {};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=new edge style 2] (3.center)
+			 to [in=60, out=-30, looseness=0.75] (4.center)
+			 to (0.center)
+			 to [in=-150, out=120, looseness=0.75] (1.center)
+			 to [in=-180, out=45, looseness=0.75] (2.center)
+			 to [in=135, out=0, looseness=0.75] cycle;
+		\draw (2.center) to (7.center);
+		\draw (5.center) to (8.center);
+		\draw (9.center) to (6.center);
+	\end{pgfonlayer}
+\end{tikzpicture}
+}, 
+and sharing a \emph{contraction node} DRG: IT'S UPSIDE DOWN?
+\scalebox{0.3}{
+	\begin{tikzpicture}
+		\begin{pgfonlayer}{nodelayer}
+			\node [style=vertex] (0) at (1.5, 3.5) {};
+			\node [style=none] (1) at (1.5, 3) {};
+			\node [style=none] (2) at (1, 4) {};
+			\node [style=none] (3) at (2, 4) {};
+		\end{pgfonlayer}
+		\begin{pgfonlayer}{edgelayer}
+			\draw (0) to (1.center);
+			\draw [bend right=45] (0) to (3.center);
+			\draw [bend left=45] (0) to (2.center);
+		\end{pgfonlayer}
+	\end{tikzpicture}
+}
+.
+
+DRG USE WRAP-BOX. 
+\[
+\adjustbox{scale=0.6}{
+    \tikzfig{figures/de-brujin-string}
+}
+\]
 
@@
 Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node
-\scalebox{0.4}{
-	\begin{tikzpicture}
-		\begin{pgfonlayer}{nodelayer}
-			\node [style=vertex] (0) at (1.5, 3.5) {};
-			\node [style=none] (1) at (1.5, 3) {};
-			\node [style=none] (2) at (1, 4) {};
-			\node [style=none] (3) at (2, 4) {};
-		\end{pgfonlayer}
-		\begin{pgfonlayer}{edgelayer}
-			\draw (0) to (1.center);
-			\draw [bend right=45] (0) to (3.center);
-			\draw [bend left=45] (0) to (2.center);
-		\end{pgfonlayer}
-	\end{tikzpicture}
-}
-.
+Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using the contraction node.
 
@@
 At the end, all explicit substitution nodes can be removed.
+The same example is then given again in Figure~\ref{fig:e-string-substitution}, this time using string diagrams.
 
-The same example is given in Figure~\ref{fig:e-string-substitution} using string diagrams. 
-The $\lambda$-abstraction is represented by a rounded box and the bound variable is the dangling wire.
-The application node is depicted as 
-\adjustbox{scale=0.15}{\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=none] (0) at (1, 6.75) {};
-		\node [style=none] (1) at (1.25, 7.75) {};
-		\node [style=none] (2) at (2, 8) {};
-		\node [style=none] (3) at (2.75, 7.75) {};
-		\node [style=none] (4) at (3, 6.75) {};
-		\node [style=none] (5) at (1.5, 6.75) {};
-		\node [style=none] (6) at (2.5, 6.75) {};
-		\node [style=none] (7) at (2, 9) {};
-		\node [style=none] (8) at (1.5, 5.75) {};
-		\node [style=none] (9) at (2.5, 5.75) {};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw [style=new edge style 2] (3.center)
-			 to [in=60, out=-30, looseness=0.75] (4.center)
-			 to (0.center)
-			 to [in=-150, out=120, looseness=0.75] (1.center)
-			 to [in=-180, out=45, looseness=0.75] (2.center)
-			 to [in=135, out=0, looseness=0.75] cycle;
-		\draw (2.center) to (7.center);
-		\draw (5.center) to (8.center);
-		\draw (9.center) to (6.center);
-	\end{pgfonlayer}
-\end{tikzpicture}
-}. 
-The reduction is now simply the re-attachment of wires, and the overall process is visibly simpler. 
-DRG UNCLEAR: Note how we avoided re-attachment for subterms as they did not use the bound variable wire, unlike in the case $[x \mapsto 1](y+y)$ before.
-
-DRG UNCLEAR: A common solution to replace explicit variable names in e-graphs is by using De Brujin indices~\cite{koehler2022sketchguided}.
-The issue with this approach is that it interferes with sharing as equivalent subterms can be structurally different.
-Consider the example $\lambda f . f ((\lambda x . f x) 2)$ which is represented as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$ using De Brujin indices.
-Notice how two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$ which means they can not be shared in an e-graph.
-A string diagrammatic representation of the same term makes it possible for $f$ to be shared as demonstrated below.
-
-DRG USE WRAP-BOX. 
-\[
-\adjustbox{scale=0.6}{
-    \tikzfig{figures/de-brujin-string}
-}
-\]
-
-Using De Brujin indices when performing $\beta$-reduction also introduces the complexity of representing the reduction as a rewrite rule.
-Special shifting operators need to be introduced on the right-hand side of a rewrite rule which makes it not merely syntactic~\cite{koehler2022sketchguided}.
-
-\textcolor{red}{Add example showing that using De Brujin indices with e-graphs breaks sharing}
-
-\textcolor{red}{Add example showing $\beta$-reduction in e-graphs compared to string diagrams}
+Using string diagrams, $\beta$-reduction is simply the re-wiring of the diagram where an application node meets an abstraction, and explicit substitution simplifies into mere composition. 
+This is appears to be simpler just by comparing diagrams, but the technical advantages of using string diagrams run deep. 
+The explicit occurrences of variable names in the older approach raise obvious questions of scope, binding and explicit substitution, whereas the use of terms inside the explicit substitution nodes also raises many technical difficulties which makes proving various mathematical properties essentially intractable.
+Additionally, explicit substitutions need to be propagated recursively throughout the graph, which is also inefficient. 
+Using De Brujin indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
 
@@
 
+\subsection{Contributions}
+
+DRG: LIST THE MAIN DEFINITIONS AND THEOREMS AND WHAT THEY MEAN
+
 \input{sections/categorical_semantics}

### 2025-01-19 LICS/sections/appendix.tex
wip

@@
+\section{$\catname{SLat}$}
+\label{sec:appendix:1}
+
+In this section we will define the category of semilattices that we use as a base for enrichment throughout the paper.
+
+\begin{definition}[Semilattice]
+    A \textit{semilattice} is a set equipped with an operation that we denote as $+$ which is associative, commutative and idempotent.
+  \end{definition}
+
+  Note that we do not require the existence of a unit for $+$. 
+  Semilattices that satisfy this extra requirement are sometimes called \textit{bounded}, i.e., they are idempotent commutative monoids.
+
+  \begin{definition}[Semilattice homomorphism]
+
+  A homomorphism between two semilattices $S_{1}$ and $S_{2}$ is a map $h$ that respects $+$.
+  That is, for all $s,s' \in S_{1}$, $h(s +_{S_{1}} s') = h(s) +_{S_{2}} h(s')$.
+  \end{definition}
+
+  \begin{definition}[Category of semilattices]
+
+  Semilattices with their respective homomorphisms form a category that we denote $\catname{SLat}$.
+  \end{definition}
+
+  \begin{proposition}
+    $\catname{SLat}$ is a closed symmetric monoidal category.
+  \end{proposition}
+  \begin{proof}
+    The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
+    $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
+    \begin{itemize}
+      \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
+      \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
+    \end{itemize}
+
+    The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
+    Clearly $S \otimes I \cong S$ by mapping $(s,*) \mapsto s$ for all $s \in S$ and vice versa.
+    The symmetry, associators and unitors are then obvious morphisms.
+    Finally, the category is closed since the set of homomorpisms between two semilattices is a semilattice by defining $(f + g)(x)$ as $f(x) + g(x)$.
+    $f + g$ is a homomorphism since $(f + g)(x+y) = f(x+y) + g(x+y) = f(x) + f(y) + g(x) + g(y) = f(x) + g(x) + f(y) + g(y) = f(x+y) + g(x+y)$.
+  \end{proof}
+
+  This makes $\catname{SLat}$ a suitable base for enrichment and allows us to take $\mathcal{V} = \catname{SLat}$.
+
+  \begin{definition}
+    Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
+    \end{definition}
+
+    Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.
+
+    \begin{proposition}[Special case of Proposition 6.4.6~\cite{Borceux_1994}]
+      The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
+    \end{proposition}
+    \begin{proof}
+      The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
+      The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
+      The adjunction is then given by the following natural isomorphisms
+      \[
+      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
+      \]
+      The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
+      Furthermore, we have, 
+      \[
+      F(I) \cong I_{\catname{SLat}}
+      \]
+      and 
+      \begin{align*}
+      F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
+            &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
+            &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
+            &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
+            &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
+            &\cong F(A \times B)
+      \end{align*}
+    In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
+    \end{proof}
+
 \section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
+\label{sec:appendix:2}
 

### 2025-01-19 LICS/sections/combinatorial_semantics.tex
wip

@@
 First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
+We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $R$ to a relation on ordered sequences. 
 We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
@@
 We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
+We will also refer to $A,B$ as \textit{feet} of the cospan, and to $C$ as \textit{carrier}.
 
@@
 
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
+In order to represent only morphisms of $\catname{S}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
 \begin{definition}[Category of MDA Hypergraphs with Interfaces]
@@
 Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
+\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\catname{S}(\Sigma)$. 
 \begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
+    $\catname{S}(\Sigma) \cong \MdaCospans$
 \end{theorem}
@@
 \subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
+The correspondence of the previous section means that the symmetric monoidal equations of $\catname{S}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
 The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
@@
 Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
+Yet, this restriction is not enough to make DPOI rewriting complete for $\catname{S,\mathcal{E}}$: there might be cospans with carriers $\mathcal{G}$ and $\mathcal{H}$ such that one can be rewritten into another but their corresponding morphisms are not equal modulo $\mathcal{E}$~\cite{bonchi_string_2022-1}.
+To make the DPOI rewriting complete, we need to add a restriction on the image of the match.
 
@@
 %By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
+We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\catname{S}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
 In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.

### 2025-01-19 LICS/sections/e-hypergraphs.tex
wip

@@
 Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.  
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
+In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\catname{S}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice-enriched setting.  
 First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components. 
@@
 This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-\textcolor{red}{The definitions and proofs without this extra assumption are given in the Appendix.}
+A discussion of a general case can be found in section~\ref{sec:dpo-fix} of Appendix.
 
@@
 An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}, for which we introduce the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} of $x$, is denoted as $[x) = \{x' ~|~ x' < x\; \}$.
+For $x \in V + E$,  the set of \emph{parents} of $x$ is denoted as $[x) = \{x' ~|~ x' < x\; \}$.
 % \[
@@
 We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
+An edge $e$ is called maximal if $\{x' | e < x'\} = \varnothing$.
 We require that the child relation satisfies the following conditions:
@@
 	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item The maximal edges must be labelled;
+	\item each $x$ has at most one immediate parent;
+	\item the maximal edges must be labelled;
 	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$ and similarly if $v \in t(e)$.
@@
 \begin{align*}
-s(e_1) = \{v_1\} & t(e_1) = \{v_2\} & l(e_1) = \bot \\
-s(e_2) = \{v_3\} & t(e_2) = \{v_4\} & l(e_2) = e_2 \\
-s(e_3) = \{v_5\} & t(e_3) = \{v_6\} & l(e_3) = e_3 \\
+s(e_1) = \{v_1\} &\; t(e_1) = \{v_2\} &\; l(e_1) = \bot \\
+s(e_2) = \{v_3\} &\; t(e_2) = \{v_4\} &\; l(e_2) = e_2 \\
+s(e_3) = \{v_5\} &\; t(e_3) = \{v_6\} &\; l(e_3) = e_3 \\
 \end{align*}

### 2025-01-19 LICS/sections/introduction.tex
wip

@@
 	\]
-	\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
+	\caption{Example translation of acyclic e-graphs into string diagrams for semilattice-enriched Cartesian categories. }
 	\label{fig:e-graph-example}
@@
 \]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
+\caption{Example translation of acyclic e-graphs into string diagrams for semilattice-enriched Cartesian categories. }
 \label{fig:e-graph-example}
@@
 The latter representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient.
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
+\emph{Hypergraphs} are a combinatorial representation designed to address this issue~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
+% A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
 Here, the generators $c_i$ become hyper-edges,  and wires become vertices.  
-Thus the expected quotient becomes simply hypergraph isomorphism.
+Thus, the expected quotient becomes simply hypergraph isomorphism. 
 With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
@@
 Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
+The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022}, is still applicable.
+However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs} --- hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
 Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
@@
 \subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
+The main technical result of the paper is a proof of soundness and completeness of this representation with respect to morphisms of a semilattice-enriched symmetric monoidal categories and their respective rewriting.
 This extends the results of \cite{bonchi_string_2022-2} for plain SMCs. 
@@
 Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
+In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in Figure~\ref{fig:e-graph-to-string} justifying our claim of developing a mathematical theory of e-graphs.
 
@@
 %
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
+%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice-enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
 In general, this involves unsharing the e-box structure before a redex becomes available.
@@
 We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.  
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
+Some further avenues for investigation, including potential applications, are given in the \nameref{sec:conclusion}.
 
@@
 \section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
+In this section,  we introduce the required preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
 
@@
 We write the identity morphism on $A$ as $\id_A$,  or simply $A$.  
-We commonly write $f;g$ for composition in diagram order.
+We commonly write $f;g$ for composition in diagrammatic order.  
 Composition in the usual order is written $g \circ f$.  
-We denote the tensor product of an SMC $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.  
+We denote the tensor product of an SMC $\mathbb{C}$ by $\otimes$,  its unit by $I$, and its symmetry natural transformation as $\sym$ \cite{maclane}.  
 We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.  
 We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.  
-We denote by $\catname{SLatt}$ the closed monoidal category of semilattices which is defined in the Appendix (\textcolor{red}{add definition of $\catname{SLatt}$ in the Appendix}).
+We denote by $\catname{SLat}$ the closed monoidal category of semilattices which is defined in section~\ref{sec:appendix:1} of the Appendix.
 
@@
 \]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
+together with the following associativity, commutativity and unitality equations, given in terms of equations of string diagrams. 
 \[
@@
 \]
-A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations of the SMT of commutative comonoids, and the following additional naturality equations, for every $c \in \Sigma_C$
+A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations of the SMT of commutative comonoids, and the following additional naturality of copy-delete equations, for every $c \in \Sigma_C$
 \[
@@
 % This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  
+Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$ (\textit{e.g.},  $\catname{SLat}$).
 In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
@@
 Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \textbf{SLatt}$.
-It remains to be seen if more general enrichments have interesting applications.
+We omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \catname{SLat}$.
 
@@
 	\item a set of objects $\mathbb{C}$
-	\item for every pair of objects $A,B \in \mathbb{C})$ --- a hom-object $\mathbb{C}(A,B \in \catname{SLatt}$
+	\item for every pair of objects $A,B \in \mathbb{C}$ --- a hom-object (or, hom-semilattice) $\mathbb{C}(A,B) \in \catname{SLat}$
 	\item for every triple of objects $A,B,C \in \mathbb{C}$ --- a composition morphism
@@
 	\[
-	u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLatt}$}
+	u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLat}$}
 	\]
@@
 such that particular coherence diagrams commute~\cite{Borceux_1994}.
+We will also call such categories $\catname{SLat}$-categories.
 \end{definition}
@@
 Note, in particular,  that we do not require semilattices to have a unit for $+$, \textit{i.e.,} we work with \textit{unbounded} semilattices.
-We also will not utilize the ordering that is induced by the operator of a semilattice in any meaningful way, so a semilattice can be thought of as a commutative idempotent semigroup throughout the paper.
-We stick to the term semilattice as it is shorter.
+% We also will not utilize the ordering that is induced by the operator of a semilattice in any meaningful way, so a semilattice can be thought of as a commutative idempotent semigroup throughout the paper.
+% We stick to the term semilattice as it is shorter.
 
-\begin{definition}[Semilattice-enriched SMC]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by a semilattice-enriched category $\mathbb{C}$ as above that additionally has
+\begin{definition}[$\catname{SLat}$-functor]
+	Let $\mathbb{C}$ and $\mathbb{D}$ be two $\catname{SLat}$-categories.
+	An $\catname{SLat}$-functor $F : \mathbb{C} \to \mathbb{D}$ is defined by the following data.
+	\begin{enumerate}
+	  \item A mapping $F : \mathbb{C} \to \mathbb{D}$
+	  \item An object-indexed family of morphisms in $\catname{SLat}$ $F_{A,B} : \mathbb{C}(A,B) \to \mathbb{D}(FA,FB)$
+	\end{enumerate}
+	such that certain coherence diagrams commute~\cite{Borceux_1994}.
+	$F$ being a morphism in $\catname{SLat}$ has a property $F(f + g) = F(f) + F(g)$.
+	An $\catname{SLat}$-functor $F$ defines an equivalence between two $\catname{SLat}$-categories if it is \textit{full}, \textit{faithful}, and \textit{essentially surjective}~\cite{Kelly2022BASICCO}.
+\end{definition}
+
+\begin{definition}[$\catname{SLat}$-SMC]\label{def:enriched-prop}
+A \textit{semilattice-enriched strict SMC}  $\mathbb{C}$ is given by an $\catname{SLat}$-category $\mathbb{C}$ as above that additionally has
 \begin{itemize}
@@
 \end{definition}
-The latter morphism being in $\catname{SLatt}$ implies the following equations that $\otimes$ satisfy
+The latter morphism being in $\catname{SLat}$ implies the following equations that $\otimes$ satisfy
 \[
@@
 
-\begin{proposition}(A specialized case of Proposition 6.4.7~\cite{Borceux_1994})
+\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also Appendix section~\ref{sec:appendix:1})
 There is a 2-adjunction 
@@
 
-In particular, 2-functor $F$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice enriched $\mathbb{C} \in \catname{SLatt}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
+In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice-enriched $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
 % SMC $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
@@
 % Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-% A \textit{semilattice enriched strict SMC functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC functor on the underlying SMCs which additionally satisfies $F(f+g) = F(f)+F(g)$.
+% A \textit{semilattice-enriched strict SMC functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC functor on the underlying SMCs which additionally satisfies $F(f+g) = F(f)+F(g)$.
 % All definitions extend to \textit{PROP} when the manipulated categories are PROPs rather than SMCs. 
@@
 We define $\catname{S}(\Sigma)^{+}$ and $\catname{S}(\Sigma, \mathcal{E})^{+}$ to be semilattice-enriched SMCs obtained as $\mathcal{F}(\catname{S}(\Sigma))$ and $\mathcal{F}(\catname{S}(\Sigma, \mathcal{E}))$ respectively.
-Both of these categories are in $\catname{PROP}^{+}$ that is obtained from $\catname{PROP}$ via the same adjunction.
+Both of these $\catname{SLat}$-categories are in $\catname{PROP}^{+}$ which we define to be the image of $\catname{PROP}$ via the same adjunction.
 \end{definition}
@@
 
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
+To aid reasoning, we introduce a new language of string diagrams for $\catname{SLat}$-SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
 It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
@@
 	\]
-	\caption{Equations for a  semilattice enriched symmetric monoidal category}
+	\caption{Equations for a  semilattice-enriched symmetric monoidal category}
 	\label{fig:string-equations}
@@
 \]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
+\caption{Equations for a  semilattice-enriched symmetric monoidal category}
 \label{fig:string-equations}

### 2025-01-19 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
wip

@@
 	\begin{pgfonlayer}{edgelayer}
-		\draw [style=new edge style 2] (3.center)
+		\draw [style=lambda_unit_string] (3.center)
 			 to [in=60, out=-30, looseness=0.75] (4.center)
@@
 
+\appendix
+
+\input{sections/appendix}
+
 \end{document}

### 2025-01-19 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
wip

--- 2025-01-09 LICS/sections/appendix.tex
@@
-\section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
+\section{$\catname{SLat}$}
+\label{sec:appendix:1}
 
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will also show the uniqueness of a boundary pushout complement.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
+In this section we will define the category of semilattices that we use as a base for enrichment throughout the paper.
+
+\begin{definition}[Semilattice]
+    A \textit{semilattice} is a set equipped with an operation that we denote as $+$ which is associative, commutative and idempotent.
+  \end{definition}
+
+  Note that we do not require the existence of a unit for $+$. 
+  Semilattices that satisfy this extra requirement are sometimes called \textit{bounded}, i.e., they are idempotent commutative monoids.
+
+  \begin{definition}[Semilattice homomorphism]
+
+  A homomorphism between two semilattices $S_{1}$ and $S_{2}$ is a map $h$ that respects $+$.
+  That is, for all $s,s' \in S_{1}$, $h(s +_{S_{1}} s') = h(s) +_{S_{2}} h(s')$.
+  \end{definition}
+
+  \begin{definition}[Category of semilattices]
+
+  Semilattices with their respective homomorphisms form a category that we denote $\catname{SLat}$.
+  \end{definition}
 
   \begin{proposition}
-    Two cospans in $\catname{Hyp}(\Sigma)$ (\textit{i.e.}, morphisms in $\HypI{\Sigma}$) $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ and $n \xrightarrow{f'} \mathcal{G}' \xleftarrow{g'} m$ are equal if and only if
-    the following two cospans  $\catname{EHyp}(\Sigma)$ (\textit{i.e.} morphisms in $\Ecospans$)
-    \[
-    n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m
-    \]
-    \[
-        n \xrightarrow{f_{ext}'} n \xrightarrow{f_{int}'} \mathcal{G} \xleftarrow{g_{int}'} m \xleftarrow{g_{ext}'} m    
-    \]
-    such that $f = f_{ext};f_{int}, f'=f_{ext}';f_{int}'$ (respectively, $g = g_{ext};g_{int}, g' = g_{ext}';g_{int}'$) are equal.
+    $\catname{SLat}$ is a closed symmetric monoidal category.
   \end{proposition}
-This essentially means that $\catname{\HypI{\Sigma}}$ faithfully embeds into $\Ecospans$. 
   \begin{proof}
-    Recall that cospans are equal when they are isomorphic.
-    The proposition means that the following cospans are isomorphic (where $\alpha$ is an isomorphism)
-    \[\begin{tikzcd}
-            & \mathcal{G} \arrow[dd, "\alpha"] &                                                            \\
-    n \arrow[ru, "f", bend left] \arrow[rd, "f'"', bend right] &                                  & m \arrow[lu, "g"', bend right] \arrow[ld, "g'", bend left] \\
-            & \mathcal{G}'                     &                                                           
-    \end{tikzcd}
-    \]
-    if and only if the following cospans are isomorphic
-    \[
-        \begin{tikzcd}
-            & n \arrow[r, "f_{int}"] \arrow[dd, "\beta"] & \mathcal{G} \arrow[dd, "\alpha"] & m \arrow[l, "g_{int}"'] \arrow[dd, "\gamma"] &                                                                        \\
-n \arrow[ru, "f_{ext}", bend left] \arrow[rd, "f_{ext}'"', bend right] &                                            &                                  &                                              & m \arrow[lu, "g_{ext}"', bend right] \arrow[ld, "g_{ext}'", bend left] \\
-            & n \arrow[r, "f_{int}'"']                    & \mathcal{G}'                     & m \arrow[l, "g_{int}'"]                      &
-\end{tikzcd}    
-    \]
-    such that $f = f_{ext};f_{int}, f'=f_{ext}';f_{int}'$ (respectively, $g = g_{ext};g_{int}, g' = g_{ext}';g_{int}'$).
-    The direction from right to left is straightforward: since the bottom diagram commutes $f;\alpha = f_{int};f_{ext};\alpha = f_{int}';f_{ext}' = f'$ and similarly for $g$.
+    The tensor product of two semilattices $S_{1}$ and $S_{2}$ is defined as follows.
+    $S_{1} \otimes S_{2}$ consists of pairs $(s_1,s_2)$ $s_{1} \in S_{1}$, $s_{2} \in S_{2}$ quotiented by commutativity, idempotence and associativity and additionally by the following relations
+    \begin{itemize}
+      \item $(s_{1} +_{S_{1}} s_{1}',s_{2}) \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
+      \item $(s_{1}, s_{2} +_{S_{2}} s_{2}') \equiv (s_{1},s_{2}) +_{S_{1} \otimes S_{2}} (s_{1}',s_{2})$
+    \end{itemize}
 
-    Let's show the direction from left to right.
-    Note that $f_{ext}$ and $f_{ext}'$ have inverses since they are injective and surjective (the injection is per the definition and surjection is because domain and codomain are equal).
-    Then we need $f_{ext};\beta = f_{ext}'$ and we define $\beta = f_{ext}^{-1};f_{ext}'$.
-    Similarly $\beta^{-1} = f_{ext}'^{-1};f_{ext}$. 
-    We can check that $\beta^{-1};\beta = f_{ext}'^{-1};f_{ext};f_{ext}^{-1};f_{ext}' = f_{ext}'^{-1};id;f_{ext}' = f_{ext}'^{-1};f_{ext}' = id$.
-    Then we have that $f_{ext};\beta;f_{int}' = f_{ext};f_{ext}^{-1};f_{ext}';f_{int}' = f_{ext}';f_{int}' = f' = f;\alpha = f_{ext};f_{int};\alpha$.
-    Since $f_{ext}$ is surjective, we have $\beta;f_{int}' = f_{int};\alpha$.
-    Analogously for $\beta^{-1}$ and the right-hand sides of the diagrams.
+    The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
+    Clearly $S \otimes I \cong S$ by mapping $(s,*) \mapsto s$ for all $s \in S$ and vice versa.
+    The symmetry, associators and unitors are then obvious morphisms.
+    Finally, the category is closed since the set of homomorpisms between two semilattices is a semilattice by defining $(f + g)(x)$ as $f(x) + g(x)$.
+    $f + g$ is a homomorphism since $(f + g)(x+y) = f(x+y) + g(x+y) = f(x) + f(y) + g(x) + g(y) = f(x) + g(x) + f(y) + g(y) = f(x+y) + g(x+y)$.
   \end{proof}
 
-\subsection{Existence of pushout}
+  This makes $\catname{SLat}$ a suitable base for enrichment.
 
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
+  \begin{definition}
+    Forgetful functor $U : \catname{SLat} \to \catname{Set}$ is given by $\catname{SLat}(I_{\catname{SLat}, -}) : \catname{SLat} \to \catname{Set}$.
+    \end{definition}
 
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+    Intuitively, the above functor returns the underlying set of a given semilattice $S$ as each morphism from $\{*\} \to S$ picks out an element of $S$.
 
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
+    \begin{proposition}[Special case of Proposition 6.4.6~\cite{Borceux_1994}]
+      The forgetful functor $U : \catname{SLat} \to \catname{Set}$ has a left adjoint free functor $F : \catname{Set} \to \catname{SLat}$.
+    \end{proposition}
+    \begin{proof}
+      The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
+      The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
+      The adjunction is then given by the following natural isomorphisms
+      \begin{align*}
+      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) &\cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B))\\
+                                                         &\cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B))\\
+                                                         &\cong \catname{Set}(A, U(B))
+      \end{align*}
+      The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
+      Furthermore, we have, 
       \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
+      F(I) \cong I_{\catname{SLat}}
       \]
       and 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp}(\Sigma)$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp}(\Sigma)$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item \label{assumption:equal_predecessors} $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item \label{assumption:non_ambiguous_predecessors} If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \ifdefined\ONECOLUMN
       \begin{align*}
-            S_{V} &= \;\{
-            (x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |
-            \exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and }\\
-            &\qquad y_j = g_{V};\iota_{2,V}(z) \text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
-            \}\\
-            &\cup \;\{
-                (y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |
-          \exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and }\\
-          &\qquad y_j = g_{V};\iota_{2,V}(z) \text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
-        \}\\
-        &\cup \;\{(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\}\\
-        S_{E} &= \;\{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
+      F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
+            &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
+            &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
+            &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
+            &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
+            &\cong F(A \times B)
       \end{align*}
-    \else 
-    \begin{align*}    
-    S_{V} &= \{\\
-          &\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}\\
-          &\cup\\
-          &\{\\
-          &\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}\\
-          &\cup\\
-          &\{\\
-          &\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\\
-          &\}\\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    \fi
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-\ifdefined \ONECOLUMN
-\begin{align*}
-    \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{
-        &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-        &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\}    
-\end{align*}
-\else
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-\fi
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-Where necessary we will refer to $S_{V}$ and $S_{R}$ as $\sim_{S}$.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
-There is also $[] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Similarly, we will write $v_{x}$ to refer to a vertex with a pre-image in $V_{X}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-These definitions make $([]_{V},[]_{E})$ automatically a homomorphism with respect to source and target maps.
-These maps are also automatically well-defined functions since $[e_1] = [e_2]$ implies $e_1 = e_2$ since the mappings for edges are mono.
-
-% So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u])
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        \item If there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ is defined and $u \sim v$
-              we let
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-              \]
-        \item \label{def:child_respects_connectivity} $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
-              If there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    Note that the existence of a path in the definition (2) implies that there exists $v' \sim v$ such that there is a path from $u$ to $v'$ and furthermore $v \not = v'$.
-    If $v' = v$ then it would mean that $[u) \not = \varnothing$ since $[v) \not = \varnothing$ and there is a path from $u$ to $v$.
-    There is no case when $u$ has a pre-image in $V_{Z}$ and yet $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e$ and there is a path from $[u]$ to $[v]$.
-    The existence of a path would mean that there is $u' \sim u$ and $v' \sim v$ such that there is a path from $u'$ to $v'$ and since $u$ has a pre-image in $V_{Z}$, $u = f_{V};\iota_{1,V}(z_1)$.
-    $v' \sim v$ means that both $v'$ and $v$ have a pre-image in $V_{Z}$ (unless $v' = v$ in which case $[u') \not = \varnothing$ as there is a path from $u'$ to $v'$) and since $u = f_{V};\iota_{1,V}(z_1)$ and $[u) = \varnothing$ $f_{V};\iota_{1,V}(z) = \varnothing$ for all $z$.
-    Since $[v) \not = \varnothing$ and $v' \sim v$, $v$ is necessarily in the image of $g_{V};\iota_{2,V}$ and for all $z$ $[g_{V};\iota_{2,V}(z)) \not = \varnothing$ as well as for some $z'$ such that $u'' = g_{V};\iota_{2,V}(z')$ and $u'' \sim_{S} u$ which would mean that the case (1) is applicable to $u$.
-    Similarly, when $u = g_{V};\iota_{2,V}(z)$.
-
-    Otherwise, we leave $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))$ undefined.
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $e$ is an element of $E_{X+Y}$.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e'$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]
-        \]
-    \item $[e) = \varnothing$ and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an undirected path from $[e]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    Otherwise we leave $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{E_{X} + E_{Y}}{\sim}([e]))$ undefined.
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $v$ in (1).
-    Suppose there exist $v_1$ and $v_2$ such that $v_1 \not = v_2$ and $[v_1) \not = \varnothing$ and $[v_2) \not = \varnothing$ and such that $u \sim v_1$ and $u \sim v_2$.
-    Then, it must be the case that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1)) = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))$.
-    Consider cases.
-    \begin{itemize}
-        \item $u = v_1$ and $u \sim_{S} v_2$. Then there exists $z \in V_{Z}$ such that $u = f_{V};\iota_{1,V}(z) = v_1$ and $v_2 = g_{V};\iota_{2,V}(z)$.
-              Since $[v_1) \not = \varnothing$, it must be the case that $[v_2) = \varnothing$ according to assumption \ref{assumption:non_ambiguous_predecessors} and we get a contraction to $[v_2) \not = \varnothing$.
-        \item $u = v_1$ and $u \sim v_2$. The latter implies that $u$ has a pre-image in $V_{Z}$ and so does $v_2$.
-              For both $[v_1) \not = \varnothing$ and $[v_2) \not = \varnothing$ they should be both in the image of $f_{V};\iota_{1,V}$ or $g_{V};\iota_{2,V}$ and according to assumption \ref{assumption:equal_predecessors} their sets of predecessors must be equal.
-        \item $u \sim_{S} v_1$ and $u \sim_{S} v_2$. This implies $u = f_{V};\iota_{1,V}(z_1) = f_{V};\iota_{1,V}(z_2)$ and $v_{1} = g_{V};\iota_{2,V}(z_1)$ and $v_{2} = g_{V};\iota_{2,V}(z_2)$.
-              By assumption \ref{assumption:equal_predecessors} it must be the case that $[g_{V}(z_1)) = [g_{V}(z_2))$ which implies that $[v_1) = [v_2)$.
-              The case when $u = g_{V};\iota_{2,V}(z_1)$ is symmetric.
-        \item $u \sim v_1$ and $u \sim v_2$.
-Same as above, by assumption \ref{assumption:non_ambiguous_predecessors} both $v_1$ and $v_2$ should be in the image of $f_{V};\iota_{1,V}$ or $g_{V};\iota_{2,V}$ and by assumption \ref{assumption:equal_predecessors} $[v_1) = [v_2)$.
-    \end{itemize}
-
-    Let's show that the definition does not depend on a particular choice of $v$ in (2).
-    Suppose, $[u) = \varnothing$ and there exist $v_1$ and $v_2$ such that $<_{X+Y}^{\mu}(v_1) = e_1$ and $<_{X+Y}^{\mu}(v_2) = e_2$ and $e_1 \not = e_2$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    The existence of a path from $[u]$ to $[v_1]$ means that there is $v_1' \sim v_1$ such that there is a path from $u'$ to $v_1'$ and similarly for $v_2$ and $v_2'$.
-    We will proceed by case analysis.
-    \begin{itemize}
-            \item If $v_1 = v_1'$ or $v_2 = v_2'$ it means that there is a path from $u$ to $v_1$ or from $u$ to $v_2$ and since $[v_1) \not = \varnothing$ as well as $[v_2) \not = \varnothing$ it must be the case that $[u) \not = \varnothing$ which contradicts the assumption that $[u) = \varnothing$.
-            \item Suppose $v_1 \sim_{S} v_1'$ which means there exists $z_1 \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_1' = g_{V};\iota_{2,V}(z_1)$ and suppose $v_2 \sim_{S} v_2'$ which further means there exists $z_2 \in V_{Z}$ such that $v_2 = f_{V};\iota_{1,V}(z_2)$ and $v_2' = g_{V};\iota_{2,V}(z_2)$.
-                  Then we have
-                  \begin{align*}
-                    <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1)) &= <_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z_1))\\
-                                                              &= \iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_1)))
-                \end{align*}
-                  and
-                  \begin{align*}
-                    <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2)) &= <_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z_2))\\
-                                                              &= \iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_2)))
-                \end{align*}
-                and $e_1 \not = e_2$ implies $<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_1)) \not = <_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_2))$ which contradicts the assumption \ref{assumption:equal_predecessors}.
-                The case if $v_2 = g_{V};\iota_{2,V}(z_2)$ and $v_1 = f_{V};\iota_{1,V}(z_1)$ ultimately contradicts the assumption \ref{assumption:non_ambiguous_predecessors} as it entails that both
-                \[
-                    <_{X}^{\mu}(f_{V};\iota_{V_{X}}(z_1)) \not = \varnothing
-                \]
-                and
-                \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2)) \not = \varnothing
-                \]
-                The cases when $v_1 = g_{V};\iota_{2,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$, and $v_1 = g_{V};\iota_{2,V}(z_1)$ and $v_2 = g_{V};\iota_{2,V}(z_2)$ are analogous.
-            \item Suppose $v_1 \sim_{S} v_1'$ and $v_2 \sim v_2'$ via $w = (x_1, \ldots, x_n)$ such that there is a path from $u$ to $v_1'$ and from $u$ to $x_n = v_2'$.
-                  The mere fact that $v_2 \sim v_2'$ implies that $v_2$ should have a pre-image in $V_{Z}$ and the same holds for $v_1$.
-                  By the same reasoning as above it must be the case that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1)) = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))$.
-            \item The case $v_1 \sim v_1'$ via $w = (x_1, \ldots, x_n)$ and $v_2 \sim_{S} v_1$ is symmetric to the above.
-            \item The case when $v_1 \sim v_1'$ and $v_2 \sim v_2'$ is analogous.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $v_1, v_2$ are vertices and $v_1 \sim v_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])$.
-    \begin{itemize}
-        \item If $[v_2) \not = \varnothing$, then, by definition we have
-                \begin{align*}
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-                     &= \text{ as $v_2 \sim v_2$ and $[v_2) \not = \varnothing$ }\\
-                     &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_2])
-                \end{align*}
-        \item The case when $[v_1) \not = \varnothing$ is symmetric.
-        \item If both $[v_2) = \varnothing$ and $[v_1) = \varnothing$ but there exists $[v_3) \not = \varnothing$ such that $v_1 \sim v_3$, then
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_3))]\\
-             &= \text{ as $v_3 \sim v_3$ and $[v_3) \not = \varnothing$ }\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_3])\\
-             &= \text{ as $v_2 \sim v_1 \sim v_3$}\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_2])
-        \end{align*}
-    \end{itemize}
-    % $v_1 \sim v_2$ implies that either $v_1 = v_2$ and the well-definedness is trivial, or there exists a sequence $w = (x_1, \ldots, x_n)$ such that $v_1 = x_1$, $v_2 = x_n$ and $x_i \sim_{S} x_{i+1}$ for $i < n$.
-    % We will proceed by induction on the length of $w$.
-    % \begin{itemize}
-    %     \item $|w|$ and then $v_1 \sim_{S} v_2$ which means there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_{2} = g_{V};\iota_{2,V}(z)$.
-    %           Consider the following cases.
-    %           \begin{itemize}
-    %             \item $[f_{V}(z)) \not = \varnothing$ and $g_{V}(z) = \varnothing$.
-    %                   Then, 
-    %                   \[
-    %                     <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    %                   \]
-    %                   and
-    %                   \[
-    %                     <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    %                   \]
-    %                   and we have well-definedness by definition.
-    %             \item $f_{V}(z) = \varnothing$ and $g_{V}(z) \not = \varnothing$ is symmetric to the above.
-    %           \end{itemize}
-    %     \item Assume that if $v_1 \sim v_2$ via $w = (x_1, \ldots, x_n)$ then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])$.
-    %     \item Now suppose $v_1 \sim v_2$ via $w = (x_1, \ldots, x_n, x_{n+1})$.
-    %           We know that 
-    %           \[
-    %             <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[x_n])
-    %           \]
-    %           and that
-    %           \[
-    %             <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[x_n]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])
-    %           \]
-    %           by the same argument as in the base case,
-    %           which implies
-    %           \[
-    %             <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[v_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_2])
-    %           \]
-    % \end{itemize}
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $([]_{V},[]_{E})$ is homomorphic with respect to $<^{\mu}$.
-    That is, we need to check if $[v) \not = \varnothing$, then $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    Since $v \sim v$ this is homomorphic by definition.
-    Similarly for edges.
-
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ similarly to how $<^{\mu}_{\sfrac{X+Y}{\sim}}$ was defined.
-We begin with defining $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ for vertices.
-
-    \begin{enumerate}
-        \item If there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $u \sim v$, we let
-              \ifdefined \ONECOLUMN
-              \[
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))
-                =
-                [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
-              \]
-              \else
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
-                &=\\
-                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
-            \end{align*}
-            \fi
-        \item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
-        Then we let
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-        \]
-    \end{enumerate}
-
-    Next we define $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}$ for edges.
-
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[e]$ to $[v]$
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]))
-        \]
-    \end{enumerate}
-
-    The well-definedness of this construction follows by the same argument as the well-definedness of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-    Then we define
-    \ifdefined \ONECOLUMN
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \qquad \text{and} \qquad \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
-    \]
-    \else
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-    \]
-    and
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
-    \]
-    \fi
-
-    as closures of $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ as below.
-    \[
-      (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}
-    \]
-    where $c$ denotes a closure and $\iota_{\sfrac{*}{\sim}}$ is $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}$ depending on whether $[x_i]$ comes from $V_{X} + V_{Y}$ or $E_{X} + E_{Y}$, is the smallest set such that
-    \begin{itemize}
-        \item $[x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$
-        \item if $[y] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x]))$ then 
-        \[
-            [x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([y])))^{c}
-        \]
-        % \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
-        % \[
-        %     [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
-        % \]
-        \item for any sequence $([x_1], \ldots, [x_n])$ such that
-        \ifdefined \ONECOLUMN
-        \[[x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}])) \qquad \text{or} \qquad [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))\]
-        \else
-        \[
-            [x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}]))
-        \] or 
-        \[
-            [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))
-        \]
-        \fi
-         for $i < n$ both
-        \ifdefined \ONECOLUMN
-        \[
-            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c} \qquad \text{and} \qquad [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
-        \]
-        \else
-         \[
-            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c}
-        \]
-        and
-        \[
-            [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
-        \]
-        \fi
-    \end{itemize}
-    Let's check that $([]_{V},[]_{E})$ is a homomorphism with respect to $\consistency$.
-    We need to check, if
-    \ifdefined \ONECOLUMN
-    \[
-        [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
-        \subseteq
-        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-    \]
-    \else
-    \begin{align*}
-        &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))\\
-        &\subseteq\\
-        &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-    \end{align*}
-    \fi
-
-    \begin{enumerate}
-        \item The first case is when $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$.
-              Then the property trivially holds as the empty set is a subset of any set.
-        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$.
-              Then, 
-              \[
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}})(v))
-              \]
-              since $v \sim v$,
-              and hence
-              \ifdefined \ONECOLUMN
-              \begin{align*}
-                [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}})(v))
-                &\subseteq
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                &\subseteq
-                (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c} = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-              \end{align*}
-              \else
-              \begin{align*}
-                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}})(v))\\
-                &\subseteq\\
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                &\subseteq\\
-                &(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c} = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-              \end{align*}
-              \fi
-              by recalling that $[[]_{V}^{\consistency}, []_{E}^{\consistency}] = [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]$.
-    \end{enumerate}
-
-    The cases for edges are analogous.
-
-    % and for vertices
-    % \begin{enumerate}
-
-    % \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    % \begin{align*}
-    %     &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-    %     &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    % \end{align*}
-    % \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    % \[
-    %     \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    % \]
-    % \end{enumerate}
-    % where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    % We then define
-    % \[
-    %   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    % \]
-    % \[
-    %   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    % \]
-    % where $c$ is a reflexive, symmetric and transitive closure.
-
-    % Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    % \begin{itemize}
-    %     \item Let us show well-definedness.
-    %     \begin{itemize}
-
-    %         \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-    %         \[
-    %             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-    %         \]
-    %           $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-    %           \begin{align*}
-    %             &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-    %             & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-    %           \end{align*}
-    %           Similarly,
-    %           \begin{align*}
-    %             &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-    %             &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-    %         \end{align*}
-    %           and we get well-definedness on the nose.
-    %           We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-    %         \item Now suppose that both 
-    %         \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-    %         \quad 
-    %         \text{and} 
-    %         \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-    %         \]
-    %         This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-    %               \[
-    %                 \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-    %               \]
-    %         \end{itemize}
-    %     \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-    %         \begin{align*}
-    %          &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-    %          &\subseteq
-    %          \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-    %         \end{align*}
-    %         By definition, 
-    %         \begin{align*}
-    %             &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-    %              [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-    %         \end{align*}
-    %         and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-    %         \begin{align*}
-    %             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-    %             &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-    %         \end{align*}
-    %          and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-    %         If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-    %         Similarly for edges.
-    % \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    By the Proposition~\ref{prop:pushout_is_e_hypergraph} the constructed $\sfrac{X+Y}{\sim}$ is an e-hypergraph.
-    To show that it is a pushout we need to check the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function as well as that it is indeed a homomorphism.
-
-    Let's say that $v_1 \sim v_2$ then we need to show that $u_{V}([v_1]) = u_{V}([v_2])$.
-    $v_1 \sim v_2$ means there is a sequence $w = (x_1, \ldots, x_n)$ such that $v_1 = x_1$, $v_2 = x_n$ and for all $i < n$ $x_i \sim_{S} x_{i+1}$. We will proceed by induction on the length of $w$.
-    \begin{itemize}
-        \item $|w| = 2$, then $v_1 \sim_{S} v_2$ and there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V})(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and we have
-        \[
-            u_{V}([v_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1({f_{V}(z)})
-        \] 
-        and 
-        \[
-            u_{V}([v_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-        \] 
-        which are equal by commutativity.
-        \item Assume if $v_1 \sim v_2$ via $w : |w| \leq n$ then $u_{V}([v_1]) = u_{V}([v_2])$.
-        \item Suppose $v_1 \sim v_2$ via $w = (x_1, \ldots, x_n, x_{n+1})$.
-              By hypothesis, we know that $u_{V}([v_1]) = u_{V}([x_1]) = u_{V}([x_n])$ and that $u_{V}([x_n]) = u_{V}([x_{n+1}]) = u_{V}([v_2])$ and hence $u_{V}([v_1]) = u_{V}([v_2])$.
-    \end{itemize}
-    Because $v_1 \sim t_2$, it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1({f_{V}(z)})
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-
-    Let's show that such $(u_{V},u_{E})$ is unique.
-    Assume that there is $u'_{V}$ such that $j_{1,V}(v) = u'_{V}([\iota_{1,V}(v)])$ and $j_{2,V}(v) = u'_{V}([\iota_{2,V}(v)])$.
-    Then we have $u_{V}[\iota_{1,V}(v)] = j_{1,V}(v) = u'_{V}([\iota_{1,V}(v)])$ for all $v \in V_{X}$ (similarly for $v \in V_{Y}$) and hence $u_{V} = u'_{V}$.
-    Similarly for $u_{E}$.
-
-    Let's now check if $u = (u_{V},u_{E})$ is a homomorphism.
-    \begin{enumerate}
-        \item The first condition to check is if $u_{V}^{*}(s_{\sfrac{X+Y}{\sim}}([e])) = s_{Q}(u_{E}([e]))$. 
-              By construction of $s_{\sfrac{X+Y}{\sim}}$ we have
-              \[
-                u_{V}^{*}(s_{\sfrac{X+Y}{\sim}}([e])) = u_{V}^{*}([s_{X+Y}(e)]^{*})
-              \]
-              Suppose $e = \iota_{E_{X}}(e')$ and then
-              \[
-                u_{V}^{*}([s_{X+Y}(e)]^{*}) = u_{V}^{*}([s_{X+Y}(\iota_{1,E}(e'))]^{*}) = u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*})
-              \]
-              Since $j_1$ and $j_2$ are homomorphisms, it must be the case that $s_{Q}(j_{1,E}(e')) = j_{1,V}^{*}(s_{X}(e'))$.
-              By definition, 
-              \begin{align*}                
-                u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*}) &= j_{1,V}^{*}((s_{X}(e')))\\ 
-                &= s_{Q}(j_{1,E}(e'))\\
-                &= s_{Q}(u_{E}([\iota_{1,E}(e')]))\\
-                &= s_{Q}(u_{E}([e]))
-            \end{align*}
-            Similarly for $e = \iota_{E_{Y}}(e')$.
-        \item The case for targets is analogous.
-        \item The next condition to check is if $[[v]) \not = \varnothing$ then 
-    \[
-        <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
-    \]
-    We need to consider a few cases.
-    \begin{itemize}
-        \item Suppose $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))]$ for $u \sim v$.
-              Then assume $u = \iota_{1,V}(u_{x})$ and then 
-              \begin{align*}
-                [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(\iota_{1,V}(u_{x})))]\\
-                &= [\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(u_{x})))]
-            \end{align*}
-            and
-            \begin{align*}
-                u_{E}([\iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(u_{x})))]) &= j_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(u_{x})))\\
-                &= <_{Q}^{\mu}(\iota_{V_{Q}}(j_{1,V}(u_{x})))\\
-                &= <_{Q}^{\mu}(\iota_{V_{Q}}(u([\iota_{1,V}(u_{x})])))\\
-                &= <_{Q}^{\mu}(\iota_{V_{Q}}(u([v])))
-            \end{align*}
-            and we have
-            \[
-                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
-            \]
-            The case when $u = \iota_{2,V}(u_{y})$ is symmetric.
-        \item Suppose $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))$ such that there is a path from $[v]$ to $[v']$ and $v$ has no pre-image in $V_{Z}$.
-              By the argument above, we know that
-              \[
-                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v']))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']))
-              \]
-              and by definition
-              \[
-                u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v'])) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
-              \]
-              since $(u_{V},u_{E})$ preserves sources and targets and there is a path from $[v]$ to $[v']$, there is a path from $u_{V}([v])$ to $u_{V}([v'])$ and because $Q$ is an e-hypergraph
-              \[
-                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v']))) = <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v])))
-              \]
-              and finally
-              \[
-                <^{\mu}_{Q}(\iota_{V_{Q}}(u_{V}([v]))) = u_{E}(<^{\mu}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v]))
-              \]
-        \end{itemize}
-    The cases for edges are analogous.
-    \item The last condition to check is if
-    \[
-    [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-    \]
-    Recall that
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    and first we will check if
-    \[
-      [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-    \]
-    Consider cases
-    \begin{itemize}
-        \item If $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \varnothing$ then the inclusion is trivial and hence we assume that
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        \item Suppose $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) \not = \varnothing$ and $v \sim u$ and then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(u)))
-        \]
-        Assume $u = \iota_{1,V}(u_{x})$, then
-        \ifdefined \ONECOLUMN
-        \begin{align*}
-            [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(u)))
-            &=
-            [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(\iota_{1,V}(u_{x}))))\\
-            &=
-            [\iota_{1,V};[]_{V}^{\consistency},\iota_{1,E};[]_{E}^{\consistency}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))
-        \end{align*}
-        \else
-        \begin{align*}
-            &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(u)))\\
-            &=\\
-            &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X}+V_{Y}}(\iota_{1,V}(u_{x}))))\\
-            &=\\
-            &[\iota_{1,V};[]_{V}^{\consistency},\iota_{1,E};[]_{E}^{\consistency}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))
-        \end{align*}
-        \fi
-
-        and
-        \ifdefined \ONECOLUMN
-        \begin{align*}
-            [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))
-            &=
-            [\iota_{1,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{1,E};[]_{E};u_{E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
-            &=
-            [j_{1,V};\iota_{V_{Q}},j_{1,E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
-            &\subseteq
-            \consistency_{Q}(\iota_{V_{Q}}(j_{1,V}(u_{x})))\\
-            &=
-            \consistency_{Q}(\iota_{V_{Q}}(u_{V}[\iota_{1,V}(u_{x})]))\\
-            &=
-            \consistency_{Q}(\iota_{V_{Q}}(u_{V}[u]))\\
-            &=
-            \consistency_{Q}(\iota_{V_{Q}}(u_{V}[v]))
-        \end{align*}
-        \else
-        \begin{align*}
-            &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
-            &=\\
-            &[\iota_{1,V};[]_{V};u_{V};\iota_{V_{Q}},\iota_{1,E};[]_{E};u_{E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
-            &=\\
-            &[j_{1,V};\iota_{V_{Q}},j_{1,E};\iota_{Q_{E}}]^{*}(\consistency_{X}(\iota_{V_{X}}(u_{x})))\\
-            &\subseteq\\
-            &\consistency_{Q}(\iota_{V_{Q}}(j_{1,V}(u_{x})))\\
-            &=\\
-            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}[\iota_{1,V}(u_{x})]))\\
-            &=\\
-            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}[u]))\\
-            &=\\
-            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}[v]))
-        \end{align*}
-        \fi
-        The case when $u = \iota_{2,V}(u_{y})$ is symmetric.
-        \item Suppose that $v$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and there exists $u$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) \not = \varnothing$ and such that there is an undirected path from $[v]$ to $[u]$ and then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))    
-        \]
-        By the argument above it follows that
-        \[
-            [u_{V};\iota_{Q_{V}},u_{E};\iota_{Q_{E}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))) \subseteq \consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))
-        \]
-        \end{itemize}
-       Since $(u_{V},u_{E})$ preserves paths, there is a path from $u_{V}([u])$ to $u_{V}([v])$ and since $Q$ is an e-hypergraph
-       \ifdefined \ONECOLUMN
-       \begin{align*}
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-        &=
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
-        &\subseteq
-        \consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))\\
-        &=
-        &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([v])))
-       \end{align*}
-       \else
-       \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-        &=\\
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
-        &\subseteq\\
-        &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))\\
-        &=\\
-        &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([v])))
-    \end{align*}
-    \fi
-       So far we have shown that
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-       \]
-       and
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))) \subseteq \consistency_{Q}(u_{E};\iota_{E_{Q}}([e]))
-       \] follows by the same argument.
-       Then, to check if
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-       \]
-       we need to verify if
-       \begin{itemize}
-        \item $[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\{[v]\}) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))$.
-              This holds because $\consistency_{Q}$ is reflexive.
-        \item $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
-        \[
-            [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\{[v]\}) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([u]))
-        \]
-              The first bit implies that $u_{V};\iota_{V_{Q}}([u]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))$ and since $\consistency_{Q}$ is symmetric it must be the case that $u_{V_{Q}};\iota_{V_{Q}}[v] \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([u]))$
-        \item There exists a sequence $w = (x_1, \ldots, x_n)$ such that 
-        \[
-            \iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_i]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]))
-        \]
-
-        or 
-        \[
-            \iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i+1}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{i}]))
-        \]
-        for $i < n$ (or similarly for edges), then $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
-        \begin{itemize}
-            \item Base case is when $w = (x_1,x_2)$.
-                  Then the statement is trivial as 
-                  \begin{align*}
-                  &[u_{V};\iota_{V_{Q}},u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_2])))\\
-                  &\subseteq\\
-                  &\consistency_{Q}(u_{V};\iota_{V_{Q}}([x_2]))
-                  \end{align*}
-            \item Suppose the statement is true for all $w$ such that $|w| \leq n$.
-            \item Consider a sequence $w = (x_1, \ldots, x_{n}, x_{n+1})$.
-                  By assumption, we know that $u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_n]))$.
-                  The fact that $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n}]) \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x_{n+1}]))$ (or its symmetric counterpart) implies that $u_{V};\iota_{V_{Q}}([x_{n}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))$ and by the transitivity and symmetry of $\consistency_{Q}$ it follows that
-                \[
-                    u_{V};\iota_{V_{Q}}([x_1]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{n+1}]))
-                \]
-                and
-                \[
-                    u_{V};\iota_{V_{Q}}([x_{n+1}]) \in \consistency_{Q}(u_{V};\iota_{V_{Q}}([x_{1}]))
-                \]
-        \end{itemize}
-       \end{itemize}
-       Therefore, we have
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-       \]
-       and similarly for edges. Which concludes that $(u_{V},u_{E})$ is a homomorphism.
-    \end{enumerate}
+    In the above we used the fact that functors $- \otimes X$ and $X \otimes -$ preserve colimits as they are both left-adjoint by symmetry and monoidal closedness of $\catname{SLat}$.
     \end{proof}
-
-\begin{remark}
-    The first assumption in~\ref{pushout:assumptions} can be weakened by allowing $E_{Z}$ to contain unlabelled edges with no inputs and outputs such that:
-    \begin{itemize}
-        \item $[f_{E}(e_{i})) = [f_{E}(e_{j}))$ and $[g_{E}(e_{i})) = [g_{E}(e_{j}))$ for all $e_{i},e_{j} \in E_{Z}$ 
-        \item If $[f_{E}(e)) \not = \varnothing$ then $[g_{E}(e)) = \varnothing$ and if $[g_{E}(e)) \not = \varnothing$ then $[f_{E}(e)) = \varnothing$.
-        \item $\consistency(f_{E}(e_{i})) = \consistency(f_{E}(e_{j}))$ and $\consistency(g_{E}(e_i)) = \consistency(g_{E}(e_j))$ for all $e_i,e_j$ in $E_{Z}$.
-    \end{itemize}
-\end{remark}
-
-\begin{remark}
-    If in a span $Y \xleftarrow{g} Z \xrightarrow{f} X$ of morphisms in $\catname{EHyp}(\Sigma)$ the morphisms $f$ and $g$ are monos, then the arrows $f'$ and $g'$ of the pushout $Y \xrightarrow{g'} X +_{f,g} Y \xleftarrow{f'} X$ are also monos.
-\end{remark}
-This is because $[]_{V}$ identifies two vertices $v_1$ and $v_2$ only if there exists a vertex $z$ in $V_{Z}$ such that $f_{V};\iota_{1,V}(z) = v_1$ and $g_{V};\iota_{2,V}(z) = v_2$ and therefore $f' = \iota_{1};[]$ and $g' = \iota_{2};[]$ are monos.
-Similarly for edges.
-
-\begin{figure}
-    \[
-        \scalebox{0.5}{
-        \tikzfig{../figures/appendix/pushout_example}
-        }
-\]
-\captionsetup{belowskip=-3ex}
-\caption{Pushout in $\catname{EHyp}(\Sigma)$}
-\label{fig:pushout_example}
-\end{figure}
-
-\begin{example}
-    Consider a diagram in Figure~\ref{fig:pushout_example} where the subscripts define the corresponding morphisms.
-    E-hypergraph $Q$ is not a pushout as there is no homomorphism $u$ that would complete the diagram.
-    In particular, there is no $(u_{V},u_{E})$ such that 
-    \[
-        [u_{V};\iota_{P_{V}},u_{E};\iota_{P_{E}}]^{*}(\consistency_{Q}(\iota_{V_{Q}}(q_1))) \subseteq \consistency_{P}(u_{V};\iota_{P_{V}}(q_1))
-    \]
-    If we let $u_{V}(i_{1,V}(v)) = j_{1,V}(v)$ for all $v \in V_{X}$ and $u_{V}(i_{2,V}(v)) = j_{2,V}(v)$ and similarly for edges.
-    Then 
-    \begin{align*}
-        &[u_{V};\iota_{P_{V}}, u_{E};\iota_{P_{E}}]^{*}(\consistency_{Q}(\iota_{V_{Q}}(q_1)))\\
-        &=\\
-        &[u_{V};\iota_{P_{V}}, u_{E};\iota_{P_{E}}]^{*}(\{q_1, q_2, q_3, q_4, f, h, g\})\\
-        &=\\
-        &[\iota_{P_{V}}(j_{1,V}(x_1)), \iota_{P_{V}}(j_{1,V}(x_2)), \iota_{P_{V}}(j_{1,V}(x_3)),\\
-         &\; \iota_{P_{V}}(j_{1,V}(x_4)), \iota_{P_{E}}(j_{1,E}(f)), \iota_{P_{E}}(j_{1,E}(g)), \iota_{P_{V}}(j_{2,V}(y_1)),\\
-         &\; \iota_{P_{V}}(j_{2,V}(y_2)), \iota_{P_{E}}(j_{2,E}(h))]\\
-        &=\\
-        &[p_1, p_2, p_3, p_4, f, h, g]\\
-        &\not \subseteq\\
-        &\consistency_{P}(u_{V};\iota_{P_{V}}(q_1))\\
-        &=\\
-        &\consistency_{P}(j_{1,V};\iota_{P_{V}}(x_1))\\
-        &=\\
-        &\consistency_{P}(\iota_{P_{V}}(p_1))\\
-        &=\\
-        &[p_1, f, h, p_2]
-    \end{align*}
-\end{example}
-
-% So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-% The proof of the latter will follow below and the following corollaries will come in handy.
-% We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))] = [e]$
-      This would mean that $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e$ is the predecessor of $e$.
-      \item Suppose $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ is defined and there is a path from $[e]$ to $[v]$.
-            This implies
-            \[
-                [e] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-            \]
-            and
-            \[
-               e = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))
-            \]
-            The existence of a path from $[e]$ to $[v]$ means there is a path from $e$ to $v'$ such that $v' \sim v$.
-            \begin{itemize}
-                \item If $v' = v$, then there is a path from $e$ to $v$ which contradicts $e = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$
-                \item Consider the case when $v' \sim v$.
-                      $e = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ implies that both $e$ and $v$ are in the image of $(\iota_{1,V},\iota_{1,E})$ or $(\iota_{2,V},\iota_{2,E})$.
-                      Suppose $e = \iota_{1,E}(e_{x})$ and $v = \iota_{1,V}(v_{x})$.
-                      Since $v' \sim v$, $v_{x} = f_{V}(z_1)$. 
-                      And since there is a path from $v'$ to $e$, $v' = \iota_{1,V}(v'_{x})$ and $v'_{x} = f_{V}(z_2)$.
-                      This further implies
-                      \begin{align*}
-                        <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{x})))\\
-                                                                &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_1))))\\
-                                                                &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_2))))\\
-                                                                &= \iota_{1,E}(e_{x})
-                    \end{align*}
-                      and
-                      \[
-                        e_{x} = <_{X}(\iota_{V_{X}}(f_{V}(z_2))) = <_{X}(\iota_{V_{X}}(v_{x}'))
-                      \]
-                      which contradicts that there is a path from $e_{x}$ to $v_{x}'$.
-                      The case when $e = \iota_{2,E}(e_{y})$ and $v = \iota_{2,V}(v_{y})$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item 
-      \[
-        [e_1] = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2))]
-      \]
-      which implies $e_1 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2))$.
-      Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-        \item $[e_2] = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]$ and $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ which implies that $e_1$ is the predecessor of $e_2$ and vice versa which contradicts e-hypergraphness of $X+Y$.
-        \item
-            \[
-                [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-            \]
-            such that $[e_1) = \varnothing$ and there is a path from $[e_1]$ to $[v]$, i.e. a path from $e_1$ to $v'$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' \sim v$ and there is a path from $e_1$ to $v'$. 
-                Suppose $e_2 = \iota_{1,E}(e_{2,x})$ and $v = \iota_{1,V}(v_{x})$.
-                Since $e_1$ is the predecessor of $e_2$, the former should also be in the image of $\iota_{1,E}$, i.e.
-$e_1 = \iota_{1,E}(e_{1,x})$, and since there is a path from $e_1$ to $v'$, $v' = \iota_{1,V}(v'_{x})$.
-                $v' \sim v$ implies $v_{x} = f_{V}(z_1)$ and $v'_{x} = f_{V}(z_2)$.
-                Then,
-                \begin{align*}
-                    <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{x})))\\
-                                                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_1))))\\
-                                                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_2))))\\
-                                                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{x}')))\\
-                                                            &= <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))\\
-                                                            &= <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))
-                \end{align*}
-                which contradicts $e_1 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2))$.
-                The case when $e_2 = \iota_{2,E}(e_{2,y})$ and $v = \iota_{2,V}(v_{y})$ is symmetric.
-            \end{itemize}
-        \end{itemize}
-            \item The other case is when          
-            \[
-                [e_1] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))]
-            \]
-            such that there is a path from $e_2$ to $v'_1$ and $v_1 \sim v'_1$ and $[e_2) = \varnothing$.
-            \begin{itemize}
-                \item The case when $[e_2] = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]$ is analogous to the last case above.
-                \item Suppose 
-                \[
-                    [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]
-                \] such that there is a path from $e_1$ to $v'_2$ and $v_2 \sim v'_2$ and $[e_1) = \varnothing$.
-                    Suppose $e_1 = \iota_{1,E}(e_{1,x})$ and $v_1 = \iota_{1,V}(v_{1,x})$ which implies $v'_2 = \iota_{1,V}(v'_{2,x})$.
-                    Since $v_1 \sim v_1'$, $v_{1,x} = f_{V}(z_1)$ and similarly $v'_{2,x} = f_{V}(z_2)$.
-                    \begin{align*}
-                        e_1 &= <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))\\
-                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v_{1,x})))\\
-                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_1))))\\
-                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z_2))))\\
-                            &= \iota_{1,E}(<_{X}^{\mu}(\iota_{V_{X}}(v'_{2,x})))\\
-                            &= <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'_2))\\
-                            &= <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))
-                    \end{align*}
-                    And we get a contradiction to irreflexivity of $<_{X+Y}^{\mu}$.
-            \end{itemize}
-      \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-        \begin{itemize}
-            \item \[
-            [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(e_1))]
-            \]
-            By definition, $s_{\sfrac{X+Y}{\sim}}([e_1]) = [s_{X+Y}(e_1)]^{*}$ and $[v] \in [s_{X+Y}(e_1)]^{*}$ means there exists $v' \sim v$ such that $v' \in s_{X+Y}(e_1)$.
-            \begin{itemize}
-                \item $v' = v$ and then $v \in s_{X+Y}(e_1)$ and therefore $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]$.
-                \item $v' \sim v$ and then 
-                \begin{align*}
-                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \end{align*}
-            \end{itemize}
-            \item \[
-                [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))]    
-            \]
-            such that there is $v_1' \sim v_1$ and there is a path from $e_1$ to $v_1'$ and $[e_1) = \varnothing$.
-            $[v] \in [s_{X+Y}(e_1)]^{*}$ means there exists $v' \sim v$ such that $v' \in s_{X+Y}(e_1)$.
-            \begin{itemize}
-                \item $v' = v$ and then $v \in s_{X+Y}(e_1)$. Since there is a path from $e_1$ to $v_1'$, there is also a path from $v'$ to $v_1'$ and by definition
-                \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                \item $v' \sim v$ and then there is a path from $v'_1$ to $v'$. 
-                Suppose $v_1 = f_{V};\iota_{1,V}(z_1)$ and since $v_1' \in s_{X+Y}(e_1)$, $[e_1) = \varnothing$, both $v_1'$ and $v'$ should be in the image of $g_{V};\iota_{2,V}$.
-                Since $[v_1) \not = \varnothing$ it must be the case that for all $z$ $[f_{V}(z)) \not = \varnothing$ as well as for some $v'' \sim v'$ such that $v'' = f_{V};\iota_{1,V}(z'')$.
-                Clearly such $v''$ exists.
-                It is either $v$, if $v = f_{V};\iota_{1,V}(z'')$, or some $x_i$ in $(x_1, \ldots, x_n)$ such that $x_1 = v'$ and $x_n = v$.
-                \begin{align*}
-                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1))]\\
-                          &= [\iota_{1,E}(<_{X}(\iota_{V_{X}}(f_{V}(z_1))))]\\
-                          &= [\iota_{1,E}(<_{X}(\iota_{V_{X}}(f_{V}(z''))))]\\
-                          &= [<_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z'')))]\\
-                          &= [<_{X+Y}(\iota_{V_{X} + V_{Y}}(v''))]\\
-                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
-                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \end{align*}
-                The case when $v_1 = g_{V};\iota_{2,V}(z_1)$ is symmetric.
-            \end{itemize}
-            \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ means that there exists $v' \in s_{X+Y}(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item \[
-                [e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v''))]
-                \] such that $v'' \sim v$ and $[v'') \not = \varnothing$.
-                \begin{itemize}
-                    \item If $v' = v$ and $v'' = v$ then $v \in s_{X+Y}(e_1)$ and since $[v'') = [v) \not = \varnothing$ and $v \in s_{X+Y}(e_1)$, it must be the case that
-                    $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(v))$ and further
-                    \begin{align*}
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                                                                                            &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(v))]\\
-                                                                                            &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                                                                                            &= [e_2]
-                    \end{align*}
-                    \item If $v' = v $ and $v'' \sim v$ then $v \in s_{X+Y}(e_1)$.
-                          Suppose $v'' = f_{V};\iota_{1,V}(z'')$.
-                          Consider cases.
-                          \begin{itemize}
-                            \item $v = f_{V};\iota_{1,V}(z)$, then $f_{V}(z) = f_{V}(z'') \not = \varnothing$ and $[e_1) = [v) = [v'')$
-                            \item $v = g_{V};\iota_{2,V}(z)$, then $g_{V}(z) = \varnothing$ and $[e_1) = \varnothing$.
-                                  Since there is a path from $v$ to $e_1$ and $[v'') \not = \varnothing$ and $v'' \sim v$, by definition,
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
-                                  \end{align*}
-                          \end{itemize}
-                          Case when $v'' = g_{V};\iota_{2,V}(z'')$ is symmetric.
-                    \item If $v' \sim v$ and $v'' = v$ then $v' \in s_{X+Y}(e_1)$. This implies that $[v) \not = \varnothing$
-                          \begin{itemize}
-                            \item If $v = f_{V};\iota_{1,V}(z)$ and $v' = f_{V};\iota_{1,V}(z')$ then $[e_1) = [v') = [v) \not = \varnothing $ since $[f_{V}(z')) = [f_{V}(z))$.
-                                  That is,
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}})(e_1)]\\
-                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v)]\\
-                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v'')]\\
-                                                                                                    &= [e_2]
-                                  \end{align*}
-                            \item $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z')$ and then $[g_{V}(z')) = \varnothing$ and $[e_1) = \varnothing$.
-                                  By definition,
-                                  \begin{align*}
-                                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
-                                                                                                          &= [e_2]
-                                  \end{align*}
-                          \end{itemize}
-                          The case when $v = g_{V};\iota_{1,V}(z)$ is symmetric.
-                    \item If $v' \sim v$ and $v'' \sim v$ then $v' \in s_{X+Y}{e_1}$, $[v'') \not = \varnothing$ and $v'' \sim v'$.
-                    \begin{itemize}
-                        \item If $v'' = f_{V};\iota_{1,V}(z'')$ and $v' = f_{V};\iota_{2,V}(z')$ then $[e_1) = [v') = [v'')$ and
-                            \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}})(e_1)]\\
-                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v')]\\
-                                                                                                    &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}})(v'')]\\
-                                                                                                    &= [e_2]
-                                  \end{align*}
-                        \item If $v'' = f_{V};\iota_{1,V}(z'')$ and $v' = g_{V};\iota_{2,V}(z')$ then $[g_{V}(z')) = \varnothing$ and $[e_1) = \varnothing$. 
-                        Then
-                        \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
-                        \end{align*}
-                        The case when $v'' = g_{V};\iota_{2,V}(z'')$ is symmetric.
-                    \end{itemize}
-                \end{itemize}
-                \item Now suppose that 
-                \[
-                [e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \]
-                such that $[v'') \not = \varnothing$, $[v) = \varnothing$ and there is a path from $[v]$ to $[v'']$, i.e. there is a path from $v$ to $v'$ and $v' \sim v''$.
-                In its turn $[v] \in s_{\sfrac{X + Y}{\sim}}([e_1])$ implies $v_1 \in s_{X+Y}(e_1)$ and $v \sim v_1$.
-                As per the comment below~\ref{def:child_respects_connectivity}, necessarily $v = v_1$ as $v$ should not have a pre-image in $V_{Z}$.
-                This implies that $v \in s_{X+Y}(e_1)$ and that $[e_1) = \varnothing$.
-                As there is a path from $v$ to $v'$, there is a path from $e_1$ to $v'$ and by definition, since $v'' \sim v'$ and $[v'') \not = \varnothing$,
-                \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
-                \]
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:closures_of_equal_sets_are_equal}
-If
-\[
-\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
-\]
-where $\iota_{\sfrac{*}{\sim}}$ is either $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}$ depending on whether $[x_i]$ is in $V_{X} + V_{Y}$ or $E_{X} + E_{Y}$.
-Then
-\[
-\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))
-\]
-That is, closures of both sides are equal.
-\end{lemma}
-
-\begin{proof}
-    To check that the sets after closures are equal, we need to check if every element of the set on the left is also in the set on the right and vice versa.
-    Apart from $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))$ the set on the left also contains
-    \begin{itemize}
-        \item $[x_1]$ as per the reflexivity.
-        So we need to check if $[x_1] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
-              Because 
-              \[
-              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
-              \]
-              there exists some $[y]$ such that 
-              \[
-              [y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))
-              \]
-               and
-            \[
-                [y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
-            \]
-            that is, we have a sequence $([x_1],[y],[x_2])$ and by transitivity
-               \[
-               [x_1] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))
-               \]
-                and
-            \[
-            [x_2] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_1]))
-            \]
-        \item $[y]$ for all $[y]$ such that $[x_1] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([y]))$.
-              Let's check if $[y] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
-              As above, there exists $[z]$ such that
-              \[
-              [z] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))
-              \]
-               and
-              \[
-                [z] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
-              \]
-              By a sequence $([y],[x_1],[z],[x_2])$ we get that
-              \[
-                [y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))
-              \]
-              Similarly for $[y]$ such that $[x_2] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([y]))$.
-        \item For any sequence $([x_1],\ldots,[x_n])$ such that $[x_i] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_{i+1}]))$ or $[x_{i+1}] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))$ for $i < n$ the set on the left contains $[x_n]$.
-              Let's check if $[x_n] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
-              Since there exists $[y]$ such that $[y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_1]))$ and $[y] \in \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x_2]))$ we can extend the sequence to $([x_2],[y],[x_1],\ldots,[x_n])$ and hence $[x_n] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_2]))$.
-              Similarly, given a sequence $([x_2],\ldots, [x_n])$ and extending it to $([x_1],[y],[x_2],\ldots,[x_n])$ we get $[x_n] \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_1]))$.
-    \end{itemize}
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v_1] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ and $[[v_1]) = [[e_1]) \not = \varnothing$, then it must be the case that 
-    \[
-    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-    \]
-    By the lemma above, to prove this, it is sufficient to show that
-
-    \[
-    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-    \]
-
-    $[v_1] \in s_{\sfrac{X+Y}{\sim}}([e_1])$ means there exists $v_1' \in s_{X+Y}(e_1)$ such that $v_1 \sim v_1'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v_1 = v_1'$ and then $v_1 \in s_{X+Y}(e_1)$.
-      \begin{itemize}
-         \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1''))]$ such that $v_1'' \sim v_1$ and $[v_1'') \not = \varnothing$.
-            \begin{itemize}
-                \item $v_1'' = v_1$ and then $[v_1) \not = \varnothing$ and $[e_1) \not = \varnothing$.
-                      Furthermore, necessarily $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)) = \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1))$.
-                      By applying $([]_{V},[]_{E})$ we get
-                      \ifdefined \ONECOLUMN
-                      \begin{align*}
-                        \text{by functionality}\\
-                          [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1))) &=
-                          [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))\\
-                          \text{by unfolding the definition}\\
-                          \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                          \end{align*} 
-                      \else
-                      \begin{align*}
-                      \text{by functionality}\\
-                        [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1))) &=\\
-                        [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1))) &\\
-                        \text{by unfolding the definition}\\
-                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                        \end{align*} 
-                      \fi
-                \item $v_1'' \sim v_1$. If $v_1'' = f_{V};\iota_{1,V}(z_1'')$ and $v_1 = f_{V};\iota_{1,V}(z_1)$ we get
-                    \begin{align*}
-                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')))\\
-                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z_1''))))\\
-                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z_1))))\\
-                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
-                                                                                                                &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    If $v_1'' = f_{V};\iota_{1,V}(z_1'')$ and $v_1 = g_{V};\iota_{1,V}(z_1)$ then $[v_1) = [e_1) = \varnothing$ and 
-                    \[
-                        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)) = \varnothing
-                    \]
-                    \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\
-                                                                                                            &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
-                    \end{align*}
-                \end{itemize}
-                \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ such that $[v_2) \not = \varnothing$ and there is a path from $[v_1]$ to $[v_2]$.
-                By definition, this implies $[v_1) = \varnothing$ and $[e_1) = \varnothing$.
-                Again, by definition, we have
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                                                                                                    &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                \end{align*}
-                \end{itemize}
-                \item $v_1 \sim v_1'$ and then $v_1' \in s_{X+Y}(e_1)$.
-                \begin{itemize}
-                    \item $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1''))]$ such that $v_1 \sim v_1''$ and $[v_1'') \not = \varnothing$.
-                    \begin{itemize}
-                        \item $v_1 = v_1''$ and then $[v_1) \not = \varnothing$.
-                              Suppose $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_1' = f_{V};\iota_{1,V}(z_1')$. Then $[v_1') = [e_1) \not = \varnothing$ and 
-                            \begin{align*}
-                              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')))\\
-                                                                                                                     &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
-                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))\\
-                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1])
-                            \end{align*}
-                            Suppose $v_1' = g_{V};\iota_{2,V}(z_1')$ and then $[v_1') = [e_1) = \varnothing$.
-                            By definition,
-                            \[
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
-                            \]
-                            and since there is a path from $e_1$ to $v_1'$ and $v_1' \sim v_1$
-                            \[
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1]) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
-                            \]
-                            The cases when $v_1 = g_{V};\iota_{2,V}(z_1)$ is symmetric.
-                        \item $v_1 \sim v_1''$ and then $v_1' \sim v_1''$. Suppose $v_1'' = f_{V};\iota_{1,V}(z_1'')$.
-                              If $v_1'$ is in the image of $f_{V};\iota_{1,V}$, then
-                              \begin{align*}
-                                \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')) &= \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1''))\\ 
-                                                                               &= \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(e_1'))
-                              \end{align*}
-                              and
-                              \begin{align*}
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1]) &=
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1'])\\ &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1])
-                              \end{align*}
-                              If $v_1'$ is in the image of $g_{V};\iota_{2,V}$, then $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')) = \consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)) = \varnothing$ and by definition as there is a path from $[e_1]$ to $[v_1'']$
-                              \begin{align*}
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\
-                                                                                                                        &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
-                              \end{align*}
-                              The cases when $v_1'' = g_{V};\iota_{2,V}(z_1'')$ are symmetric.
-                    \end{itemize}
-                \end{itemize}
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that are not top-level.
-\end{lemma}
-\begin{proof}
-  $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$ if and only if $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{*}{\sim}}([x]))$ is defined.
-  Let's first show from left to right.
-$\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$ implies $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$.
-  \begin{itemize}
-    \item $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) = [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\iota_{*}{\sim}(x'))$ such that $x \sim x'$.
-          By e-hypergraph-ness of $X+Y$ this implies that $[x') \not = \varnothing$ and hence $<_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x'])) = <_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) = [<_{X+Y}^{\mu}(\iota_{*}(x))]$
-    \item $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{*}{\sim}}([x])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ such that there is a path from $[x]$ to $[v]$ and
-          $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$. This implies that $[v) \not = \varnothing$ and by definition
-          \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{*}{\sim}}([x])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{*}{\sim}}([v]))  
-          \]
-  \end{itemize}
-  The right-to-left direction is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-\label{prop:pushout_is_e_hypergraph}
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the uniqueness of the corresponding boundary pushout complement.
-
-% \begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-% Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-% They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-% \end{definition}
-% Intuitively, the no-dangling condition guarantees that sources and targets of an edge are not deleted if the edge is not deleted.
-% The no-identification condition requires that if $m$ identifies two vertices, it must not be the case that one vertex should be removed and the other vertex should be preserved according to $\mathcal{K}$.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{../figures/combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\begin{proposition}[Proposition~\ref{prop:boundary_unique} restatement]
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique.
-\end{proposition}
-\begin{proof}
-        The pushout in~\ref{def:boundary_new} satisfies the assumptions in Definition~\ref{pushout:assumptions} (note that since $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j$ is an MDA cospan, the image of $i + j$ in $\mathcal{L}$ consists of top-level vertices only) and therefore is isomorphic to the pushout constructed in~\ref{pushout:assumptions} and by construction should also be a pushout on the underlying sets of vertices and hyperedges.
-    % I think the above sufficies, but it can also be shown formally that every pushout in $\catname{EHyp_{\Sigma}}$ is a pushout on the underlying sets of edges and vertices by constructing an e-hyperghraph from the set of vertices and same for the set of edges and checking if the morphisms are well-defined.
-    %     Consider diagrams below,
-    %     \[
-    %         \begin{tikzcd}
-    %             & \mathcal{L} \arrow[d] \arrow[ldd, bend right] & i+j \arrow[l, "h"'] \arrow[d, "c"]           &                 & V_{\mathcal{L}} \arrow[d] \arrow[ldd, bend right] & i+j \arrow[l, "h_{V}"] \arrow[d, "c_{V}"]      \\
-    %             & \mathcal{G} \arrow[ld, "u"']                  & \mathcal{C} \arrow[l] \arrow[lld, bend left] &                 & V_{\mathcal{G}} \arrow[ld, "u_{V}"]               & V_\mathcal{C} \arrow[l] \arrow[lld, bend left] \\
-    % \mathcal{Q} &                                               &                                              & V_{\mathcal{Q}} &                                                   &
-    % \end{tikzcd}  
-    %     \]
-    %     If $\mathcal{L} \xrightarrow{} \mathcal{G} \xleftarrow{} C$ is a pushout in $\catname{EHyp}_{\Sigma}$ then $V_{\mathcal{L}} \xrightarrow{} V_{\mathcal{G}} \xleftarrow{} V_{\mathcal{C}}$ is a pushout in $\catname{Set}$.
-    %     The small square on the right commutes because the square on the left commutes, so we only need to check the universal property.
-    %     Suppose there exist $i+j \to V_{\mathcal{L}} \to V_{Q} = i+j \to V_{\mathcal{C}} \to V_{Q}$ and we need to check that there exists a unique $u_{V}$ such that these morphisms factors out through $u_{V}$.
-    %     % Hence, pushout complement should yield pushout complements for the corresponding sets.
-    %     Construct an e-hypergraph $\mathcal{Q} = (V,E,s,t,<,\consistency)$ such that
-    %     \begin{itemize}
-    %         \item $V = V_{Q}$
-    %         \item $E = \sfrac{E_{\mathcal{L}} + E_{\mathcal{C}}}{\mathcal{R}}$ where $e_1 \mathcal{R} e_2$ if there exists $v \in V_{Q}$ such that $v = i_{V}(v_1)$ and $v = i_{V}(v_2)$ and $e_1 = <^{\mu}_{\mathcal{L}}(\iota_{V_{\mathcal{L}}}(v_1)), e_2 = <^{\mu}_{\mathcal{C}}(\iota_{V_{\mathcal{C}}}(v_2))$
-    %     \end{itemize}
-        In particular,
-    \[
-        \tikzfig{../figures/appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{../figures/appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements and where $E_{0}$ is either empty or contains edges with no inputs and outputs only.
-    % Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Since $m$ and $[c_1,c_2]$ are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{a + y} && {i + j + x} && {i+j} \\
-	\\
-	{a + y + z} &&&& {\begin{matrix} i + j + w \\ \cong \\ i + j + z \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Where $y$ is the image of $i + j$. Clearly, $i + j + z$ is a pushout complement, as computing the pushout identifies $i + j$ with its image within $y$ and leaves $z$ as is.
-    Suppose there is another pushout complement $i + j + w$.
-    Because $a + y + w$ also yields a pushout it must be the case that $a + y + w \cong a + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    \textcolor{red}{The same reasoning applies to show that $E_{\mathcal{L}^{\bot}}$} is unique and hence the sets of edges and vertices are uniquely determined in the pushout complement.
-    One can observe that $g_{V} = h_{V,ext};h_{V,int} + id_{n}$ and that $h_{V,ext}$ is mono and $h_{V,int} = [h_1,h_2]$ where $h_1,h_2$ are mono.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose, on contrary, that there exist
-    \[
-    \mathcal{L}_{1}^{\bot} = \langle V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1} \rangle \qquad \mathcal{L}_{2}^{\bot} = \langle V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{2} \rangle
-    \]
-
-    such that there is $e \in E_{\mathcal{L}^{\bot}}$ and $s_{1}(e) \not = s_{2}(e)$.
-    As $g$ is a homomorphism, it must be the case that $g^{*}_{V}(s_{1}(e)) = g^{*}_{V}(s_{2}(e)) = s_{\mathcal{G}}(g(e))$, or
-    $(h_{V,ext};[h_{1},h_{2}] + id_{n})(s_{1}(e)) = (h_{V,ext};[h_{1},h_{2}] + id_{n})(s_{2}(e))$.
-    For the latter to hold when $s_{1}(e) \not = s_{2}(e)$ it must be the case that $h_{V,ext};[h_{1},h_{2}]$ identifies $v_{1} \in s_{1}(e)$ with $v_{2} \in s_{2}(e)$ and necessarily $v_{1}$ and $v_{2}$ are in the image of $c$.
-    This can only happen if $v_1 \in i$ and $v_2 \in j$, \textit{i.e.} they belong to input and output interfaces respectively.
-    However, this contradicts the fact that $v_2 \in s_2(e)$ as a vertex which is a source of some edge can not be in the output interface as per the (6) and (7) conditions of the boundary complement~\ref{def:boundary_new} the image of $j$ in $\mathcal{L}^{\bot}$ should consist of vertices of in-degree 0.
-    Similarly for $t_{\mathcal{L}^{\bot}}$.
-
-    % Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{\mathcal{L^{\bot}}}(e)) = s_{\mathcal{G}}(g_{E}(e))$ and $s_{\mathcal{L^{\bot}}}(e) \subseteq (g_{V}^{-1})^{*}(s_{\mathcal{G}}(g_{E}(e)))$.
-    % Since $g_{V} = h_{V,ext};[h_1,h_2] + id_{n}$ the inverse image of $g_{V}$ contains at most two elements: one with the pre-image in $i$ and another with the pre-image in $j$.
-    % If $g_{V}$ identified more than two vertices it would mean that $[h_1,h_2]$ identified more than two vertices which would mean that two of them come both either from $i$ or $j$ which would imply that either $h_1$ or $h_2$ is not mono.
-    % If the inverse image contains exactly one vertex then the $s_{\mathcal{L^{\bot}}}(e)$ is uniquely fixed, otherwise suppose it contains $\{v_1,v_2\}$ such that $v_1 \in i$ and $v_2 \in j$.
-    % However, $v_1$ can not be in $s_{\mathcal{L^{\bot}}}(e)$ as per the (6) and (7) conditions of the boundary complement~\ref{def:boundary_new} the image of $j$ in $\mathcal{L}^{\bot}$ should consist of vertices of in-degree 0.
-
-    % Suppose that there exist
-    % \[
-    %     \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    % \qquad
-    % \text{and}
-    % \qquad
-    %     \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    % \]
-    % Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    % Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    % Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ \textcolor{red}{only identifies edges with no successors}.
-    Similarly for edges.
-    We now show that if $[v) = \varnothing$ for $v \in V_{\mathcal{L}^{\bot}}$ then necessarily $[g_{V}(v)) = \varnothing$.
-    Suppose $[v) = \varnothing$ and $<_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = e$ and there is a connected component that contains $v$ and such that for all $v'$ and $e'$ in this component $[v') = \varnothing = [e')$.
-    Consider $\mathcal{G'}$ which is obtained from $\mathcal{G}$ by making $[g_{V}(v')) = [g_{E}(e')) = \varnothing$ for $v', e'$ above.
-    Clearly there is a morphism from $\mathcal{L}^{\bot}$ to $\mathcal{G}'$, i.e.
-the morphism $g' = (g_{V},g_{E})$ that has the same action on edges and vertices as the morphism $g$ does, because $v'$ such that $[v') = \varnothing$ can be mapped to $g_{V}(v')$ such that $[g_{V}(v')) \not = \varnothing$.
-    The only case when such a morphism may not exist is when $[v'') \not = \varnothing$, $[v') = \varnothing$ and $g_{V}(v'') = g_{V}(v')$.
-    This implies that both $v''$ and $v'$ have a pre-image in $i + j$ and according to the definition of a boundary complement it must be $[[c_1,c_2](z_1)) = [[c_1,c_2](z_2))$ for all $z_1, z_2 \in V_{i + j}$.
-    For the same reason there is also a morphism from $\mathcal{L}$ to $\mathcal{G}'$.
-    The only case when such a morphism may not exist is when $v'' \in V_{\mathcal{L}}$ such that $[v'') \not = \varnothing$ is mapped to $m_{V}(v'') = g_{V}(v')$ such that $[v') = \varnothing$ which implies $v'$ and $v''$ share a pre-image in $i + j$.
-    This leads to a contradiction as the image of $i + j$ in $\mathcal{L}$ should consist of top-level vertices exclusively.
-    This construction results in a commutative diagram below
-    \[
-        \begin{tikzcd}
-            & \mathcal{L} \arrow[d, "m"'] \arrow[ldd, "m'"', bend right=49] & i'+j' \arrow[l] & i+j \arrow[d] \arrow[l]                                           \\
-            & \mathcal{G}                                                   &                 & \mathcal{L}^{\bot} \arrow[ll, "g"'] \arrow[llld, "g'", bend left] \\
-\mathcal{G}' &                                                               &                 &                                                                  
-\end{tikzcd}
-    \]
-    that further results in a contradiction since there is no morphism from $\mathcal{G}$ to $\mathcal{G'}$ because for any such morphism $u$ it must be the case that $[u_{V}(g_{V}(v'))) \not = \varnothing = [g'_{V}(v')) = \varnothing$ which would mean that $\mathcal{G}$ is not a pushout.
-    This means that either both $<_{1}(\iota_{*}(x))$ and $<_{2}(\iota_{*}(x))$ are undefined for a given $x$ or $<_{1}(\iota_{*}(x)) = <_{2}(\iota_{*}(x))$ which concludes that the relations are equal and $\mathcal{L}_{1}^{\bot} = \mathcal{L}_{2}^{\bot}$.
-    The same reasoning applies when we consider $<_{1}(\iota_{E_{\mathcal{L}^{\bot}}}(e))$ and $<_{2}(\iota_{E_{\mathcal{L}^{\bot}}}(e))$.
-
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              $\consistency_{1} \not = \consistency_{2}$ implies there exist $v, v' \in \mathcal{L}^{\bot}$ such that $v \consistency_{1} v'$ and $v \not \consistency_{2} v'$ for $[v) = [v') \not = \varnothing$.
-              Then there are connected components $C_1, C_2$ in $\mathcal{L}_{2}^{\bot}$ such that for all $x \in C_1$ and $y \in C_2$ $x \not \consistency_{2} y$ and $v \in C_1$, $v' \in C_2$.
-              Consider $g_2 = (g_{V},g_{E}) : \mathcal{L}_2^{\bot} \to \mathcal{G}$ and construct $\mathcal{G'}$ that is obtained from $\mathcal{G}$ by making $g(x) \not \consistency g(y)$ for $x \in C_1$ and $y \in C_2$.
-              There is a morphism $g_2' = g_2 : \mathcal{L}_2^{\bot} \to \mathcal{G}'$.
-Like above, the only case when such morphism may not exist is when for $v' \not \consistency v$ $g_{V}(v) = g_{V}(v')$.
-              This implies $v, v'$ have a pre-image in $i + j$ and by definition it must be $v' \consistency v$.
-              Similarly, there is a morphism $m' : \mathcal{L} \to \mathcal{G}'$.
-The only case when it may not exist is when for $v_1, v_1' \in \mathcal{L}$ such that $v_1 \consistency v_1'$, $m_{V}(v_1) = g_{V}(v_2)$ and $m_{V}(v_1') = g_{V}(v_2')$ where $v_2 \in C_1$ and $v_2' \in C_2$.
-              It implies $v_1, v_1'$ have pre-image in $i + j$ and by definition of a boundary-complement the image of $i + j$ should consist of top-level vertices exclusively.
-              This results in the diagram below there is no morphism $u : G \to G'$, as the existence of such a morphism implies $u(g(x)) = u(g'(x))$, $u(g(y)) = u(g'(y))$ for $x \in C_1$ and $y \in C_2$ and $u(g(x)) \consistency u(g(y))$ and $u(g'(x)) \not \consistency u(g'(y))$.
-              This contradicts the fact that $\mathcal{G}$ is a pushout.
-              Hence, $\consistency_{1} = \consistency_{2}$ and $\mathcal{L}_{2}^{\bot} = \mathcal{L}_{1}^{\bot}$.
-    \end{itemize}
-    Ultimately, we have shown that $\mathcal{L}^{\bot}$ is unique up to isomorphism.
-    The same reasoning applies when we consider edges instead of vertices.
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{remark}
-%     Consider morphisms $i + j \xrightarrow{[f_{ext},g_{ext}];[f_{int},g_{int}]} \mathcal{L} \xrightarrow{m} \mathcal{G}$.
-%     If both of these morphisms are mono, then the boundary complement exists. 
-%     Intuitively, it exists for the same reason the corresponding boundary complement for plain hypergraphs exists: it is constructed by removing from $\mathcal{G}$ everything that has no pre-image in $i + j$.
-% \end{remark}
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{../figures/combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{figure}
-    \[
-    \scalebox{0.75}{
-        \tikzfig{../figures/appendix/contexts}
-    }
-    \]
-    \caption{String diagrammatic contexts}
-    \label{fig:string_contexts}
-\end{figure}
-
-\begin{remark}
-    The construction in Figure~\ref{fig:A+B} naturally extends to $k$ operands.
-    Given a join of $k$ cospans of e-hypegraphs
-    \begin{align*}
-        n_1 \xrightarrow{} n_1' \xrightarrow{} &\mathcal{F}_1 \xleftarrow{} m_1' \xleftarrow{} m_1\\
-        &\;+\\
-        &\vdotswithin{+}\\
-        &\;+\\
-        n_k \xrightarrow{} n_k' \xrightarrow{} &\mathcal{F}_k \xleftarrow{} m_k' \xleftarrow{} m_k
-    \end{align*}
-    we will denote the carrier of the resulting cospan as
-    \[
-    \mathcal{F}_1 \; \hat{+}\; \ldots\; \hat{+}\; \mathcal{F}_{k}
-    \]
-    to distinguish it from the coproduct.
-\end{remark}
-
-\begin{remark}
-    Below we will also occasionally conflate the cospan resulting from $\llbracket - \rrbracket$ with its carrier which will be clear from the context.
-    For example, if $\llbracket f \rrbracket = n \to n' \to \mathcal{F} \xleftarrow{} m' \xleftarrow{} m$, then we will also use $\llbracket f \rrbracket$ to denote $\mathcal{F}$.
-\end{remark}
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-    \label{proof:appendix:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $f = g$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket f\rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket f \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans$.
-    $f = g$ implies that either
-    \begin{itemize}
-        \item $f = \mathcal{C}[l] = g$ for some morphism $l \in \textbf{PROP}^{+}(\Sigma)$.
-        \item $f = \mathcal{C}[l]$ and $g = \mathcal{C}[r]$ where $l = r$ or $r = l$ is from $\mathcal{S}$.
-        \item There is a sequence $w = (x_1, \ldots, x_n)$ such that $x_1 = f$ and $x_n = g$ and $x_i = x_{i+1}$ or $x_{i+1} = x_{i}$ in the sense above for $i < n$.
-    \end{itemize}
-    where $\mathcal{C}$ is a context defined as per the Figure~\ref{fig:string_contexts} and $\mathcal{C}[l]$ is defined by replacing the occurrence of
-    \scalebox{0.5}{
-    \begin{tikzpicture}[tikzfig]
-        \begin{pgfonlayer}{nodelayer}
-            \node [style=empty diag] (0) at (-0.5, 0.5) {};
-        \end{pgfonlayer}
-    \end{tikzpicture}
-    }
-    in $\mathcal{C}$ with a string diagram $l$.
-    Let's prove the statement by induction on $\mathcal{C}$ when $|w| = 2$.
-    \begin{itemize}
-        \item If $\mathcal{C}$ is empty, then either $f = g$ syntactically and $\llbracket f \rrbracket = \llbracket g \rrbracket$ or $f = l = r = g$ for some $l = r$ or $r = l$ in $\mathcal{S}$.
-              The latter trivially yields the E-DPO diagram below
-              \[
-                \begin{tikzcd}
-                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
-                    \\
-                    {\mathcal{L}} && {i+j} && {\mathcal{R}} \\
-                    & {i'+j'} && {i''+j''} \\
-                    && {i+j}
-                    \arrow[from=1-1, to=3-1]
-                    \arrow[from=1-2, to=1-1]
-                    \arrow[from=1-3, to=1-2]
-                    \arrow[from=1-3, to=1-4]
-                    \arrow[from=1-3, to=3-3]
-                    \arrow[from=1-4, to=1-5]
-                    \arrow[from=1-5, to=3-5]
-                    \arrow["\lrcorner"{pos=0.05, rotate=90, description}, phantom, from=3-1, to=1-2]
-                    \arrow[from=3-3, to=3-1]
-                    \arrow[from=3-3, to=3-5]
-                    \arrow["\lrcorner"{pos=0.05, rotate=180, description}, phantom, from=3-5, to=1-4]
-                    \arrow[from=4-2, to=3-1]
-                    \arrow[from=4-4, to=3-5]
-                    \arrow[from=5-3, to=3-3]
-                    \arrow[from=5-3, to=4-2]
-                    \arrow[from=5-3, to=4-4]
-                \end{tikzcd}
-              \]
-              where $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j = \llbracket l \rrbracket = \llbracket f \rrbracket$ and $i \to i'' \to \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j = \llbracket r \rrbracket = \llbracket g \rrbracket$ and $i + j \to i + j \to \mathcal{L}$ is a boundary complement as per~\ref{def:boundary_new}.
-              Hence, $\llbracket f \rrbracket \Rrightarrow{}_{\langle \mathcal{L}, \mathcal{R} \rangle} \llbracket g \rrbracket$.
-              \item Suppose 
-              \[
-                \scalebox{0.75}{\tikzfig{../figures/appendix/contexts_2}}
-              \]
-              and by inductive hypothesis $\llbracket \mathcal{C}[l] \rrbracket \Rrightarrow{} \llbracket \mathcal{C}[r] \rrbracket$ which results in the top half of the diagram below.
-              \[\begin{tikzcd}
-                {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
-                \\
-                {\llbracket \mathcal{C}[l]\rrbracket} && {\mathcal{L}_1^{\bot}} && {\llbracket \mathcal{C}[r]\rrbracket} \\
-                & {n_1'+m_1'} & {n_1+m_1} & {n_1''+m_1''} \\
-                {\llbracket f \rrbracket} && {\mathcal{L}_2^{\bot}} && {\llbracket g \rrbracket} \\
-                & {n_2' + m_2'} && {n_2''+m_2''} \\
-                && {n_2 + m_2}
-                \arrow[from=1-1, to=3-1]
-                \arrow[from=1-2, to=1-1]
-                \arrow[from=1-3, to=1-2]
-                \arrow[from=1-3, to=1-4]
-                \arrow[from=1-3, to=3-3]
-                \arrow[from=1-4, to=1-5]
-                \arrow[from=1-5, to=3-5]
-                \arrow["\lrcorner"{pos=0.025, rotate=90, description}, draw=none, from=3-1, to=1-2]
-                \arrow[from=3-1, to=5-1]
-                \arrow[from=3-3, to=3-1]
-                \arrow[from=3-3, to=3-5]
-                \arrow["\lrcorner"{pos=0.025, rotate=180, description}, draw=none, from=3-5, to=1-4]
-                \arrow[from=3-5, to=5-5]
-                \arrow[from=4-2, to=3-1]
-                \arrow[from=4-3, to=3-3]
-                \arrow[from=4-3, to=4-2]
-                \arrow[from=4-3, to=4-4]
-                \arrow[from=4-3, to=5-3]
-                \arrow[from=4-4, to=3-5]
-                \arrow["\lrcorner"{pos=0.025, rotate=90, description}, draw=none, from=5-1, to=4-2]
-                \arrow[from=5-3, to=5-1]
-                \arrow[from=5-3, to=5-5]
-                \arrow["\lrcorner"{pos=0.025, rotate=180, description}, draw=none, from=5-5, to=4-4]
-                \arrow[from=6-2, to=5-1]
-                \arrow[from=6-4, to=5-5]
-                \arrow[from=7-3, to=5-3]
-                \arrow[from=7-3, to=6-2]
-                \arrow[from=7-3, to=6-4]
-            \end{tikzcd}\]
-            Clearly the occurrence of $\llbracket \mathcal{C}[l] \rrbracket$ in $\llbracket f \rrbracket$ is convex and down-closed and by following the argument in Theorem 35~\cite{bonchi_string_2022-2}, there exists $n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2$ such that
-            \ifdefined \ONECOLUMN
-            \begin{align*}
-                \llbracket f \rrbracket &= \;
-                (0 \to 0 \to \llbracket C[l] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)
-                ;
-                (n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)\\
-                \llbracket g \rrbracket &= \;
-                (0 \to 0 \to \llbracket C[r] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)
-                ;
-                (n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)
-                \end{align*}
-            \else
-            \begin{align*}
-            \llbracket f \rrbracket = \;
-            &(0 \to 0 \to \llbracket C[l] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)\\
-            &;\\
-            &(n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)\\
-            \llbracket g \rrbracket = \;
-            &(0 \to 0 \to \llbracket C[r] \rrbracket \xleftarrow{} n_1' + m_1' \xleftarrow{} n_1 + m_1)\\
-            &;\\
-            &(n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2)
-            \end{align*}
-            \fi
-            and
-            % $n'_2 + m'_2 = n_3 + m_3 + (n_1' + m_1' \setminus (n_1 + m_1))$
-            \[
-                n_2' = n_3 + (n_1' \setminus n_1) \qquad m_2' = m_3 + (m_1' \setminus m_1)
-            \]
-            and
-            \[
-            n_2 + m_1 \xrightarrow{} n_2' \setminus (n_1' \setminus n_1) + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot} \xleftarrow{} m'_2 \setminus (m_1' \setminus m_1) + n_1 \xleftarrow{} m_2 + n_1    
-            \] is an MDA cospan.
-            By denoting the pushout of $\mathcal{L}_1^{\bot} \xleftarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot}$ with $\mathcal{L}_3^{\bot}$,
-            one can observe that $\llbracket f \rrbracket$ is the pushout object of $\mathcal{L} \xleftarrow{} i+j \xrightarrow{} \mathcal{L}_{3}^{\bot}$ as per the diagram below
-            \[\begin{tikzcd}
-                && {n_1+m_1} && {\mathcal{L}^{\bot}_2} \\
-                \\
-                {i+j} && {\mathcal{L}^{\bot}_1} && {\mathcal{L}^{\bot}_3} \\
-                \\
-                {\mathcal{L}} && {\mathcal{C}[l]} && {\llbracket f \rrbracket}
-                \arrow[from=1-3, to=1-5]
-                \arrow[from=1-3, to=3-3]
-                \arrow[from=1-5, to=3-5]
-                \arrow[from=3-1, to=3-3]
-                \arrow[from=3-1, to=5-1]
-                \arrow[from=3-3, to=3-5]
-                \arrow[from=3-3, to=5-3]
-                \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-3]
-                \arrow[from=3-5, to=5-5]
-                \arrow[from=5-1, to=5-3]
-                \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=5-3, to=3-1]
-                \arrow[from=5-3, to=5-5]
-                \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=5-5, to=3-3]
-                \arrow["\dagger", draw=none, from=5-5, to=3-3]
-            \end{tikzcd}
-            \]
-            where the square marked with $\dagger$ is a pushout because of pushout pasting law.
-            The corestriction of $n_1 + m_1 \xrightarrow{[f_1,f_2]} \mathcal{L}_1^{\bot}$ to the image of $i + j \xrightarrow{[g_1,g_2]} \mathcal{L}_{1}^{\bot}$ is mono: suppose it is not mono, then there exist $z_1$, $z_2$, $z_3$ such that $f_1(z_1) = f_2(z_2) = g_1(z_3)$, then $[f_2,g_1]$ would not be mono, similarly if $f_1(z_1) = f_2(z_2) = g_2(z_3)$.\
-            By construction of the pushout the restriction of $\mathcal{L}_{1}^{\bot} \to \mathcal{L}_{3}^{\bot}$ is mono and hence the arrow $i + j \to \mathcal{L}_{3}^{\bot}$ is mono.
-            This implies that $\llbracket f \rrbracket \Rrightarrow_{\langle \mathcal{L},\mathcal{R} \rangle} \llbracket g \rrbracket$.
-            \item Consider the case when
-            \[
-            \scalebox{0.75}{\tikzfig{../figures/appendix/contexts_3}}    
-            \]
-            By hypothesis, there exists a diagram
-                \[
-                \begin{tikzcd}
-                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
-                    \\
-                    {\llbracket\mathcal{C}[l]\rrbracket} && {\mathcal{L}_1^{\bot}} && {\llbracket\mathcal{C}[r]\rrbracket} \\
-                    & {n'+m'} && {n''+m''} \\
-                    && {n+m}
-                    \arrow[from=1-1, to=3-1]
-                    \arrow[from=1-2, to=1-1]
-                    \arrow[from=1-3, to=1-2]
-                    \arrow[from=1-3, to=1-4]
-                    \arrow[from=1-3, to=3-3]
-                    \arrow[from=1-4, to=1-5]
-                    \arrow[from=1-5, to=3-5]
-                    \arrow[from=3-3, to=3-1]
-                    \arrow[from=3-3, to=3-5]
-                    \arrow[from=4-2, to=3-1]
-                    \arrow[from=4-4, to=3-5]
-                    \arrow[from=5-3, to=3-3]
-                    \arrow[from=5-3, to=4-2]
-                    \arrow[from=5-3, to=4-4]
-                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
-                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
-                \end{tikzcd}
-                \]
-                which implies the existence of the following diagram where both squares are pushouts by construction and the matching is convex and down-closed.
-                \textcolor{red}{Consider cases when $\mathcal{L}$ has no-input and no-output sub-hypergraph and when it does not.
-Basically, repeat the above reasoning here as well, but say that if $n+m$ is empty, then there is no pushout.}
-                \ifdefined \ONECOLUMN
-                \[\adjustbox{width=0.95\linewidth}{
-                    \begin{tikzcd}
-                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
-                    \\
-                    {\llbracket f_1 \rrbracket \; \hat{+}\; \ldots \hat{+} \; \llbracket\mathcal{C}[l]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \mathcal{L}_{1}^{\bot} \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket\mathcal{C}[r]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} \\
-                    & {n_1' + m_1'} && {n_1'' + m_1''} \\
-                    && {n + m}
-                    \arrow[from=1-1, to=3-1]
-                    \arrow[from=1-2, to=1-1]
-                    \arrow[from=1-3, to=1-2]
-                    \arrow[from=1-3, to=1-4]
-                    \arrow[from=1-3, to=3-3]
-                    \arrow[from=1-4, to=1-5]
-                    \arrow[from=1-5, to=3-5]
-                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
-                    \arrow[from=3-3, to=3-1]
-                    \arrow[from=3-3, to=3-5]
-                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
-                    \arrow[from=4-2, to=3-1]
-                    \arrow[from=4-4, to=3-5]
-                    \arrow[from=5-3, to=3-3]
-                    \arrow[from=5-3, to=4-2]
-                    \arrow[from=5-3, to=4-4]
-                \end{tikzcd}
-                }
-                \]
-                \else
-                \begin{figure*}[hbt!]
-                \[
-                    \begin{tikzcd}
-                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
-                    \\
-                    {\llbracket f_1 \rrbracket \; \hat{+}\; \ldots \hat{+} \; \llbracket\mathcal{C}[l]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \mathcal{L}_{1}^{\bot} \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} && {\llbracket f_1 \rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket\mathcal{C}[r]\rrbracket \; \hat{+} \; \ldots \; \hat{+} \; \llbracket f_n \rrbracket} \\
-                    & {n_1' + m_1'} && {n_1'' + m_1''} \\
-                    && {n + m}
-                    \arrow[from=1-1, to=3-1]
-                    \arrow[from=1-2, to=1-1]
-                    \arrow[from=1-3, to=1-2]
-                    \arrow[from=1-3, to=1-4]
-                    \arrow[from=1-3, to=3-3]
-                    \arrow[from=1-4, to=1-5]
-                    \arrow[from=1-5, to=3-5]
-                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
-                    \arrow[from=3-3, to=3-1]
-                    \arrow[from=3-3, to=3-5]
-                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
-                    \arrow[from=4-2, to=3-1]
-                    \arrow[from=4-4, to=3-5]
-                    \arrow[from=5-3, to=3-3]
-                    \arrow[from=5-3, to=4-2]
-                    \arrow[from=5-3, to=4-4]
-                \end{tikzcd}
-                \]
-                \caption{$\llbracket f \rrbracket \Rrightarrow{}_{\langle \mathcal{L},\mathcal{R} \rangle} \llbracket g \rrbracket$}
-                \label{fig:f_rewrites_to_g_under_plus}
-            \end{figure*}
-            \fi
-                and hence we have $\llbracket f \rrbracket \Rrightarrow_{\langle \mathcal{L}, \mathcal{R} \rangle} \llbracket g \rrbracket$
-    \end{itemize}
-    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
-    % there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
-    % As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[]$ is defined by computing the pushout along the common interfaces.
-    % We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    % The mapping of external interfaces into $\llbracket l \rrbracket$ is mono and therefore the corresponding pushout complement exists.
-    % We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    % Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    % And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    % \begin{figure}[h!]
-    % \[
-    %     \scalebox{0.8}{
-    %     \tikzfig{../figures/combinatorial_semantics/DPOI_square}
-    %     }
-    % \]
-    % \end{figure}
-    % We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    % This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    % The no-dangling condition is satisfied because we consider only mda-cospans.
-    % Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    % Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    % The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    % Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    % The rest follows by induction on the number of applied equations.
-    % Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:decomposition}
-    Let $n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m$ be an MDA extended cospan.
-    If $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $\mathcal{G}$ such that the immediate predecessors of top-level connected components of $\mathcal{L}$ in $\mathcal{G}$ are either undefined or coincide and pair-wise consistent then either
-    \begin{itemize}
-    \item there exists a cospan $i \to i' \to \mathcal{G}' \xleftarrow{} j' \xleftarrow{} j$ and $k \in \mathbb{N}$ such that $\mathcal{G}$ is decomposable as
-    \begin{align}
-        \label{fig:decomposition:1}
-    \scalebox{0.6}{
-        \tikzfig{../figures/appendix/G_decomposition_1}
-    }
-    \end{align}
-    \item or, $\mathcal{G}$ is decomposable as
-    \begin{align}
-        \label{fig:decomposition:2}
-    \scalebox{0.6}{
-        \tikzfig{../figures/appendix/G_decomposition_2}
-    }
-    \end{align}
-    \end{itemize}
-    such that $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $\mathcal{G}'$ and all the cospans are MDA.
-\end{lemma}
-\begin{proof}
-    Suppose the image of top-level components of $\mathcal{L}$ consists of top-level edges and vertices.
-    Then let $\mathcal{C}_1$ be the smallest e-hypergraph consisting of the input vertices of $\mathcal{G}$ and every hyperedge $h$ (and its successors) that is not in $\mathcal{L}$ but has a path to it.
-    Let $\mathcal{C}_{2}$ be the smallest e-hypergraph such that $\mathcal{G} = \mathcal{C}_{1} \cup \mathcal{L} \cup \mathcal{C}_{2}$.
-    These e-hypergraphs overlap only on vertices so we define
-    \begin{align*}
-        i = V_{\mathcal{C}_{1}} \cap V_{\mathcal{L}}\\
-        j = V_{\mathcal{C}_{2}} \cap V_{\mathcal{L}}\\
-        k = (V_{\mathcal{C}_{1}} \cap V_{\mathcal{C}_{2}}) \setminus V_{\mathcal{L}}
-    \end{align*}
-    and let $i'$ be the union of $i$ and the vertices of $\mathcal{L}$ in the image of $f'$ and $j'$ be the union of $j$ and the vertices of $\mathcal{L}$ in the image of $g'$.
-    Let $n_{1}'$ be pre-image of the co-restriction of $f'$ to $\mathcal{C}_{1}$ and $n_{2}'$ be the pre-image of the co-restriction of $f'$ to $\mathcal{C}_{2}$.
-    Similarly, for $m_{1}'$ and $m_{2}'$ with regard to $g'$.
-    Then there are cospans
-    \begin{align*}
-        n \to n_{1}' \to &\mathcal{C}_{1} \xleftarrow{} m_{1}' + k + i \xleftarrow{} k + i\\
-        i \to i' \to &\mathcal{L} \xleftarrow{} j' \xleftarrow{} j\\
-        k + j \to n_{2}' + k + j \to &\mathcal{C}_{2} \xleftarrow{} m_{2}' \xleftarrow{} m\\
-    \end{align*}
-    and $n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m$ is the co-limit of the following diagram
-    \[
-        n \to n_{1}' \to \mathcal{C}_{1} \xleftarrow{} m_{1}' + k \xleftarrow{} i + k \to i' + k \to \mathcal{L} \xleftarrow{} k + j' \xleftarrow{} k + j \xrightarrow{} n_{2}' + k + j \to \mathcal{C}_{2} \xleftarrow{} m_{2}' \xleftarrow{} m
-    \]
-    The two spans identify precisely those nodes from $\mathcal{G}$ that occur in more than one sub-hypergraph, so this amounts to simply taking the union
-    \[
-    n \to n_{1}' \cup n_{2}' \cup i' \to \mathcal{C}_{1} \cup \mathcal{C}_{2} \cup \mathcal{L} \xleftarrow{} m_{1}' \cup m_{2}' \cup j' \xleftarrow{} m  = n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m
-    \]
-    If we let $\mathcal{G}' = \mathcal{L}$ we get the decomposition as in~(\ref{fig:decomposition:1}).
-    This follows the construction of a similar lemma in~\cite{bonchi_string_2022-1} (Lemma 24). 
-    The cospans with $\mathcal{C}_{1}$ and $\mathcal{C}_{2}$ are MDAs because they are closed under successors and predecessors and the cospan with $\mathcal{L}$ is an MDA because it is convex.
-
-    Now suppose that the image of top-level components of $\mathcal{L}$ consist of non top-level edges and vertices $e$ and $v$.
-    Take a hyperedge $h$ such that $h \leq e$ and for all $e' \not = h$ such that $e' \leq e$, $h \leq e'$.
-    Let $\mathcal{G}''$ be $h$ including all its sources and targets and successors.
-    If $\mathcal{G} = \mathcal{G}''$ then it is decomposable as in~(\ref{fig:decomposition:2}) and let $\mathcal{G}'$ be the component that contains the image of $\mathcal{L}$ which exists by pair-wise consistency.
-    All $\mathcal{G}_{i}$ (and their respective cospans) are MDA by definitions.
-    Otherwise, if $\mathcal{G} \not = \mathcal{G}''$ then $\mathcal{G}''$ is a convex down-closed sub-e-hypergraph of $\mathcal{G}$ and 
-    by applying the same construction as above we get~(\ref{fig:decomposition:1}).
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
-
-\section{Additional figures for Section~\ref{sec:introduction}:Introduction}
-
-This section contains the figures with more elaborate transformation of e-string diagrams from the Figure~\ref{fig:e-graph-example}.
-
-\ifdefined\ONECOLUMN
-\begin{figure}[htb!]
-    \vspace{-3cm}
-    \centering
-    \[
-        \hspace{1.3cm}
-        \resizebox{0.8\textwidth}{!}{
-        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-b-c}
-        }
-    \]
-    \caption{Example translation from $(b)$ to $(c)$.}
-    \label{fig:e-graph-example-b-c}
-\end{figure}
-\else
-\begin{figure*}[htb!]
-    \vspace{-3cm}
-    \centering
-    \[
-        \hspace{1.3cm}
-        \resizebox{0.8\textwidth}{!}{
-        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-b-c}
-        }
-    \]
-    \caption{Example translation from $(b)$ to $(c)$.}
-    \label{fig:e-graph-example-b-c}
-\end{figure*}
-\fi
-
-\ifdefined\ONECOLUMN
-\begin{figure}[htb!]
-    \[
-        % \hspace{1.3cm}
-        \scalebox{0.4}{
-        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
-        }
-    \]
-    \caption{Example translation from $(a)$ to $(b)$.}
-    \label{fig:e-graph-example-a-b}
-\end{figure}
-\else
-\begin{figure*}[htb!]
-    \[
-        % \hspace{1.3cm}
-        \scalebox{0.5}{
-        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
-        }
-    \]
-    \caption{Example translation from $(a)$ to $(b)$.}
-    \label{fig:e-graph-example-a-b}
-\end{figure*}
-\fi
-
-\section{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
-\begin{example}
-    Consider cospans from Figure~\ref{fig:appendix:non-isomorphic-cospans}(a) where the $f_{int}, f_{ext}$ (respectively, $g_{int},g_{ext}$) maps are shown with red arrows.
-    The left cospan represents a symmetry placed inside a hierarchical edge while the right one depicts a tensor product of two identities put inside the hierarchical edge.
-    The colours of the vertices show the induced partition of the interfaces.
-    We will try to build an isomorphism step-by-step.
-    To make the diagram commute, clearly, it must be the case that $u_0 \mapsto u_0$ and $u_1 \mapsto u_1$.
-    Then, because the isomorphism should preserve the order within the subset of green verices, it must be the case that $u_2 \mapsto u_2$ and $u_3 \mapsto u_3$.
-    Similarly for the output interfaces.
-    To further make the diagram commute, we need to map $v_2 \mapsto v_3$ and $v_3 \mapsto v_2$.
-    This, however, contradicts the requirement that $g_{int};\gamma(w_2) = g_{int}(w_2)$.
-    Hence, the two cospans are not isomorphic.
-\end{example}
-
-\begin{example}
-    Now consider the cospans from Figure~\ref{fig:appendix:non-isomorphic-cospans}(b).
-    Notice how the position of external input interface relative to strict internal interface is different in the second cospan.
-    The isomorphism would map $u_0 \mapsto u_2, u_1 \mapsto u_1, u_2 \mapsto u_0, u_3 \mapsto u_3$.
-    The other mappings are identities. 
-    One can check that the corresponding diagram commutes.
-\end{example}
-
-\begin{figure*}
-    \begin{subfigure}[T]{0.48\textwidth}
-    \begin{subfigure}[T]{0.45\textwidth}
-        \[
-    \scalebox{0.6}{
-        \tikzfig{../figures/appendix/sym_cospan}
-    }    
-    \]
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[T]{0.45\textwidth}
-        \[
-            \scalebox{0.6}{
-                \tikzfig{../figures/appendix/id_x_id_cospan}
-            }    
-            \]
-    \end{subfigure}
-    \subcaption{\;}
-\end{subfigure}
-    \hfill
-\begin{subfigure}[T]{0.48\textwidth}
-    \begin{subfigure}[T]{0.45\textwidth}
-        \[
-            \scalebox{0.6}{
-                \tikzfig{../figures/appendix/id_x_id_cospan}
-            }    
-            \]
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[T]{0.45\textwidth}
-        \[
-            \scalebox{0.6}{
-                \tikzfig{../figures/appendix/id_x_id_cospan_2}
-            }    
-            \]
-    \end{subfigure}
-    \subcaption{\;}
-\end{subfigure}
-    \caption{Non-isomorphic cospans (a) and isomorphic cospans (b)}
-    \label{fig:appendix:non-isomorphic-cospans}
-\end{figure*}
-
-\section{DPO rewriting for arbitrary signatures}
-\label{sec:dpo-fix}
-Here we will discuss how one can formulate DPO rewrite rules where the left-hand of right-hand sides contain sub-hypergraphs with no inputs and outputs at the outermost level.
-First we will illustrate where the restriction in the original formulation of the DPO rewriting as per the Definition~\ref{def:convex_dpo} comes from via several examples: one for Cartesian equations (which are important to express e-graphs as e-hypergraphs) and two for the structural rules we need for $\catname{MEHypI}_{\Sigma} / \mathcal{S}$ to be a category enriched over $\catname{SLat}$.
-Consider a rewrite rule induced by the Cartesian structure $a;! = id_{I}$ for $a : 0 \to 1$ which is rendered string diagrammatically as 
-\[
-\tikzfig{../figures/appendix/cartesian_rewrite_string}
-\]
-which becomes the following extended cospan of e-hypergraphs
-\[
-\tikzfig{../figures/appendix/cartesian_rewrite_hypergraph}
-\]
-
-Assume we want to perform a rewrite within the e-hypergraph in Figure~\ref{fig:dpo-stuck} (left).
-
-\begin{figure}
-    \begin{subfigure}{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.7}{
-    \tikzfig{../figures/appendix/cartesian_rewrite_source_graph}
-    }
-    \]
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.55\linewidth}
-        \[
-        \adjustbox{scale=0.7}{
-        \tikzfig{../figures/appendix/cartesian_rewrite_dpo_stuck}
-        }
-        \]
-    \end{subfigure}
-    \caption{DPO rewriting stuck}
-    \label{fig:dpo-stuck}
-\end{figure}
-
-However, we are immediately stuck as there is no matching of the left-hand side of the rule within the graph as there is no pushout complement: the only way we can impose a constraint that an edge shall have a particular predecessor in the pushout is through its adjacent vertices, but the sub-hypergraph for $a;!$ does not have any as demonstrated in the right subfigure.
-
-Another problematic rewrite arises when we want to utilise the structural rule of the free enrichment
-
-\[
-\adjustbox{scale=0.75}{
-\tikzfig{../figures/appendix/structural_rewrite_problem}
-}
-\]
-
-if we consider the following DPO square
-
-\[
-\adjustbox{scale=0.55}{
-    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck}
-}
-\]
-
-% we can not express id_I + id_I in such a way
-% The last example is the rewrite that corresponds to the idempotence for the identity on the tensor unit $id_{I} = id_{I} + id_{I}$
-
-% \[
-% \adjustbox{scale=0.75}{
-%     \tikzfig{../figures/appendix/structural_id_unit_problem}
-% }
-% \]
-
-% with the DPO square as
-
-% \[
-% \adjustbox{scale=0.75}{
-% \tikzfig{../figures/appendix/structural_rewrite_id_unit_dpo_stuck}
-% }
-% \]
-
-\begin{definition}
-
-We will define the action of $+$ on two cospans where either of the carriers is the empty e-hypergraph as
-
-\begin{align*}
-0 \xrightarrow{} n \xrightarrow{} &\;\mathcal{F} \xleftarrow{} m \xleftarrow{} 0\\
-&\;+ \hspace{6em} \Coloneqq \hspace{2em} 0  \xrightarrow{} n \xrightarrow{} \;\mathcal{F} \xleftarrow{} m \xleftarrow{} 0\\
-0 \xrightarrow{} 0 \xrightarrow{} &\;\varnothing \xleftarrow{} 0 \xleftarrow{} 0.
-\end{align*}
-
-\end{definition}
-
-This is so that we can absorb some equations and to seamlessly introduce what follows.
-
-% \begin{remark}
-% By putting no restrictions on the signature we need to enforce some restrictions on the equations in $\mathcal{E}$.
-% Recall that $\llbracket id_{I} + id_{I}\rrbracket = \llbracket id_{I} \rrbracket$ as per the Remark~\ref{remark:f+g}.
-% This implies that for all $f : 0 \to 0$, $\llbracket f + id_{I} \rrbracket$
-% Consider a cospan of e-hypergraphs that corresponds to a morphism $f + g + (id_{I} + id_{I})$
-
-% \[
-% \adjustbox{scale=0.75}{
-%     \tikzfig{../figures/appendix/remark_on_equations}
-% }
-% \]
-
-% By using the idempotence equation for $id_{I} + id_{I}$ we can rewrite it to
-
-% \[
-%     \adjustbox{scale=0.75}{
-%         \tikzfig{../figures/appendix/remark_on_equations_2}
-%     }
-% \]
-
-% However, there is no rewrite rule to go back and equate these two cospans in $\MdaEcospans / \mathcal{S,E}$.
-% To fix that, we need to introduce the following rewrite rule schema
-% \[
-% \adjustbox{scale=0.75}{
-%     \tikzfig{../figures/appendix/remark_on_equations_3}
-% }
-% \]
-% By introducing this equation we automatically require that every morphism $0 \to 0$ is essentially $id_{I}$.
-
-% \end{remark}
-
-We will start completing all the above DPO squares by first considering the structural rules.
-We will include special unlabelled edges with no inputs and outputs to the interfaces to be able to impose the hierarchical relation on subgraphs with no inputs and outputs.
-That is, the previously mentioned structural rule will become
-
-\[
-\adjustbox{scale=0.75}{
-    \tikzfig{../figures/appendix/structural_rewrite_problem_fix}
-}
-\]
-
-where the labels show how interface edges are mapped to the edges of the corresponding e-hypergraphs.
-For example, the edge in the interface labelled $a_1$ is mapped to the outermost edge on the right.
-We can now compute the pushout complement and complete the DPO square below
-
-\[
-\adjustbox{scale=0.55}{
-    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
-}
-\]
-
-Handling deletion or introduction rewrite rules will require the introduction of auxiliary rules.
-In particular, two rules for introduction and elimination of these special edges
-
-\begin{figure}[h!]
-\begin{subfigure}{0.45\linewidth}
-    \[
-    \adjustbox{scale=0.55}{
-    \tikzfig{../figures/appendix/idI_elimination_hypergraph}
-    }
-    \]
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.45\linewidth}
-    \[
-    \adjustbox{scale=0.55}{
-    \tikzfig{../figures/appendix/idI_introduction_hypergraph}
-    }
-    \]
-\end{subfigure}
-\end{figure}
-
-and similar rules to introduce and eliminate in nested contexts
-
-\begin{figure}[h!]
-    \begin{subfigure}{\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
-        \tikzfig{../figures/appendix/idI_elimination_nested_hypergraph}
-        }
-        \]
-    \end{subfigure}
-    \vspace{1em}
-    \begin{subfigure}{\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
-        \tikzfig{../figures/appendix/idI_introduction_nested_hypergraph}
-        }
-        \]
-    \end{subfigure}
-\end{figure}
-
-and, finally, the rules to introduce such edges as new consistent components.
-Intuitively such edge plays the role of a tensor unit identity that can be inserted in any context.
-
-\begin{figure}[h!]
-    \begin{subfigure}{0.45\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
-        \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
-        }
-        \]
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.45\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
-        \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
-        }
-        \]
-    \end{subfigure}
-    \end{figure}
-
-% \[
-% \adjustbox{scale=0.75}{
-% \tikzfig{../figures/appendix/auxiliary_schema}
-% }
-% \]
-
-% This rule makes sense in any SMT where there is a unique morphism $I \to I$ which is $id_{I}$.
-Then the deletion rule is reconstructed as
-
-\[
-\adjustbox{scale=0.75}{
-    \tikzfig{../figures/appendix/cartesian_rewrite_fix}
-}
-\]
-
-and the full DPO square becomes
-
-\[
-\adjustbox{scale=0.75}{
-    \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix}
-}
-\]
-
-The target e-hypergraph is not exactly what we wanted, however, but we can use the auxiliary rule to remove this unlabelled edge.
-
-\[
-\adjustbox{scale=0.75}{
-    \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix_2}
-}
-\]
-
-% and then utilise the idempotence rule for the identity of the tensor unit
-
-% \[
-%     \adjustbox{scale=0.75}{
-%         \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix_3}
-%     }
-% \]
-
-Generally, given an interpretation of a structural rule (where the right-hand side contains a single edge at the outermost level) $\llangle \llbracket l \rrbracket, \llbracket r \rrbracket \rrangle$, for each connected component $C_{i}$ at the outermost level with no inputs and outputs and for each hyperedge with no inputs \textit{or} outputs  in $C_{i}$  we introduce an empty hyperedge into the interfaces which are mapped to the single edge at the outermost level in $\llbracket r \rrbracket$.
-Symmetric rules are handled by performing the above in the other direction.
-
-Rules for deletion and introduction from Cartesian structure are interpreted similarly, by mapping all interface edges to a single empty hyperedge on the right (respectively, left) when deleting (respectively, introducing).
-For an arbitrary rule $\llangle \llbracket l \rrbracket, \llbracket r \rrbracket \rrangle$ induced by $\mathcal{E}$ consider connected components with no inputs and outputs $C^{l}_{i}$ and $C^{r}_{j}$ for left-hand and right-hand sides respectively.
-Let $c^{l}_{1,1}, \ldots, c^{l}_{k,n}$ and $c^{r}_{1,1}, \ldots, c^{r}_{p,m}$ where $c^{l}_{i,j}$ is $i$-th edge with no inputs or outputs in component $C^{l}_{j}$ and similarly for $c^{r}_{i,j}$ and let the interface contain the largest of the two sets.
-And let the morphism be any valid morphism from $\{^{l}_{1,1}, \ldots, c^{l}_{k,n}\} \to c^{r}_{1,1}, \ldots, c^{r}_{p,m}$ is the former is larger as a set, or the other way around if the latter is larger.
-Graphically, this is expressed as follows
-
-\[
-\adjustbox{scale=0.65}{
-    \tikzfig{../figures/appendix/rule_transformation_example}
-}
-\]
-
-assuming that $|c^{l}_{1,1}, \ldots, c^{l}_{k,n}| > |c^{r}_{1,1}, \ldots, c^{r}_{p,m}|$ and $f = \{ c^{l}_{1,1} \mapsto c^{l}_{1,1}, \ldots, c^{l}_{k,n} \mapsto c^{l}_{k,n} \}$ and $g$ is any valid mapping of $c^{l}_{1,1}, \ldots, c^{l}_{k,n}$ to $c^{r}_{1,1}, \ldots, c^{r}_{p,m}$.
-
-\section{Going from e-graph to e-hypergraph}
-
-\begin{figure}
-
-\begin{align*}
-    \text{functional symbols}& \hspace{2em} f,g\\
-    \text{e-class id}& \hspace{2em} a,b\\
-    \text{terms}& \hspace{2em} t \Coloneqq f \;|\; f (t_1, \ldots, t_m) \hspace{2em} &m \geq 1\\
-    \text{e-nodes}& \hspace{2em} n \Coloneqq f \;|\; f (a_1, \ldots, a_m) \hspace{2em} &m \geq 1\\
-    \text{e-classes}& \hspace{2em} c \Coloneqq \{n_1, \ldots, n_m\}   \hspace{2em} &m \geq 1
-\end{align*}
-\caption{E-graph components}
-\label{fig:e-graph-components}
-\end{figure}
-
-\begin{definition}[E-graph~\cite{EggPaper}]
-    E-graph is a data-structure that consists of
-    \begin{itemize}
-        \item components defined in a grammar in Figure~\ref{fig:e-graph-components}
-        \item a union find structure $U$ that stores equivalence relation on e-class ids
-        \item \textit{e-class} map $M$ that maps e-class ids to e-classes.
-All equivalent e-class ids map to the same e-class, \textit{i.e.}, if $a \equiv_{id} b$ then $M[a]$ and $M[b]$ are the same set.
-              An e-class id $a$ is said to refer to e-class $M[\text{find}(a)]$. NB: $\text{find}()$ returns the canonical representative for the equivalence, i.e. the last predecessor of $a$ in $U$.
-        \item The \textit{hashcons} map $H$ that maps e-nodes to e-class ids.
-    \end{itemize}
-\end{definition}
-
-\begin{definition}[Hashcons invariant]
-The hashcons $H$ must map all canonical e-nodes to their e-class ids:
-\[
-\text{e-node } n \in M[a] \text{ iff } H[\text{canonicalize}(n)] = \text{find}(a)
-\]
-\end{definition}
-
-The pseudocode below shows the translation from a \textit{canonical} (the one for which the above invariant holds).
-For the following e-graph with e-classes shown as dashed boxes it produces the following cospan of e-hypergraphs.
-\texttt{connect} and \texttt{mkEdge} are responsible for making things shared using $\triangle$.
-
-\begin{figure}
-    \begin{subfigure}{0.4\linewidth}
-\[
-\adjustbox{scale=0.6}{
-\tikzfig{../figures/appendix/e-graph-example_1}
-}
-\]
-\end{subfigure}
-\begin{subfigure}{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.6}{
-    \tikzfig{../figures/appendix/e-hypergraph-example_1}
-    }
-    \]
-\end{subfigure}
-\end{figure}
-
-\begin{algorithm}
-    \caption{E-graph to e-hypergraph translation}
-    \begin{algorithmic}[1]
-    \Function{e-class-to-HEdge}{e\_class\_id}
-    \State e\_nodes $\gets$ [e\_node$_{1}$ $\ldots$ e\_node$_{n}$] \textbf{in} H.keys() \textbf{where} H[e\_node$_{i}$] == M[e\_class\_id]
-    % \State \textbf{for} [$\text{e_node}_{1}, \ldots, \text{e_node}_{n}$] \textbf{in} H \textbf{where} H[$\text{e_node}_{i}$] == M[e-class-id] \texbf{do}
-    \ForAll {e\_node$_{i}$ in e\_nodes}
-    \State edge $\gets$ mkEdge(e\_class\_id)
-    \Comment {Create a hierarchical edge that represents an e-class}
-    \State \textbf{match} e\_node$_{i}$ \textbf{with}
-    \State \qquad \textbf{case} f($e_1 \ldots e_{n}$) :
-    \State \qquad \qquad \textbf{if} is\_singleton(e\_class\_id) \textbf{then}
-    % \If{is\_singleton(e\_class\_id)}
-        \State \qquad\qquad \hspace{2em} edge $\gets$ mkEdge(f)
-        \Comment {Create an edge that represents the generator f $: k \to 1$}
-        \State \qquad \qquad \hspace{2em} edges $\gets$ [\Call{e-class-to-HEdge}{$e_1$} $\ldots$ \Call{e-class-to-HEdge}{$e_n$}]
-        \State \qquad \qquad \hspace{2em} edge.connect(edges)
-        \State \qquad \qquad \hspace{2em} \Return edge
-    % \Else
-    \State \qquad \qquad \textbf{else}
-    \State \qquad \qquad \hspace{2em} graph $\gets$ discard([e\_node$_1$$\ldots$e\_node$_{i-1}$]) $\cup$ mkEdge(f) $\cup$ discard([e\_node$_{i+1}$ $\ldots$ e\_node$_{n}$])
-        \State \qquad \qquad \hspace{2em} edges $\gets$ [\Call{e-class-to-HEdge}{$e_1$} $\ldots$ \Call{e-class-to-HEdge}{$e_n$}]
-        \State \qquad \qquad \hspace{2em} edge.add(graph)
-        \Comment {Add \texttt{graph} as a consistent component into hierarchical edge \texttt{edge}}
-        \State \qquad \qquad \hspace{2em} edge.connect(edges)
-    % \EndIf
-    \EndFor
-    \State \Return edge
-    % \Comment {There will be a single e-node in a singleton e-class}
-    % \ForAll {e-node in H \textbf{where} H[e-node] == M[e-class-id]}
-    %     \State \textbf{match} e-node \textbf{with}
-    %     \State \quad \textbf{case} \texttt{f}: \Return (mkEdge(f))
-    %     \State \quad \textbf{case} \texttt{f}($e_1$, $\ldots$, $e_{n}$) :
-    %     % this is nested
-    %         \State \qquad\qquad edge $\gets$ mkEdge(f)
-    %         \State \qquad\qquad edge.connect([e-class-to-HEdge(e) | e $\gets$ [$e_1$, \ldots, $e_{n}$]])
-    %         \State \qquad\qquad \Return edge
-    % \EndFor
-    % \Else
-    % \State edge $\gets$ mkEdge(M[e-class-id])
-    % \ForAll {e-node in H \textbf{where} H[e-node] == M[e-class-id]}
-    %     \State \textbf{match} e-node \textbf{with}
-    %     \State \quad \textbf{case} \texttt{f}: edge.add(f)
-    %     \State \quad \textbf{case} \texttt{f}($e_1$, $\ldots$, $e_{n}$): edge.add(f)
-    % \EndFor
-    % \EndIf
-    \EndFunction
-    \State
-    \State graph $\gets \varnothing$ 
-    \ForAll{e\_class\_id in H.values()}
-    \State graph $\gets$ graph $\cup$ \Call {e-class-to-HEdge}{e\_class\_id})
-
-    \EndFor
-    % \Function{matchPattern}{shape}
-    %     \State \textbf{match} shape \textbf{with}
-    %     \State \quad \textbf{case} \texttt{Circle(radius)}:
-    %     \State \quad \quad \Return $\pi \cdot \texttt{radius}^2$
-    %     \State \quad \textbf{case} \texttt{Rectangle(width, height)}:
-    %     \State \quad \quad \Return \texttt{width} $\times$ \texttt{height}
-    %     \State \quad \textbf{case} \texttt{Triangle(base, height)}:
-    %     \State \quad \quad \Return $\frac{1}{2} \times \texttt{base} \times \texttt{height}$
-    %     \State \quad \textbf{case} \texttt{Square(side)}:
-    %     \State \quad \quad \Return \texttt{side}$^2$
-    %     \State \quad \textbf{case} \texttt{\_}:
-    %     \State \quad \quad \Return ``Unknown shape''
-    % \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{lemma} Two e-classes $e_1$ and $e_2$ are merged if and only if there are rewrite rules that turn every term representable by $e_1$ into $e_2$.
-
-\end{lemma}
-
-\begin{lemma}
-Every hierarchical edge can be decomposed as follows.
-
-\[
-\tikzfig{../figures/appendix/hyperedge_decomposition}
-\]
-\end{lemma}
-
-Then, given e-classes $e_1$ and $e_2$ for sub-hypergraphs $C$ and $D$ respectively, we introduce e-nodes into the e-graph $A(e_1)$ and $B(e_2)$ that reside in e-class $e_3$.

### 2025-01-19 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
wip

--- 2025-01-15 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
@@
 \[
-\tau \Coloneq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}
+\tau \coloneqq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}
 \],
@@
 A monoidal (right) closed category is a monoidal category $\catname{C}$ satisfying for
-every pair of objects $B,C$ there is an object $B \xRightarrow{} C$ and a morphism $\textbf{ev}_{B,C} : (B \xRightarrow{} C) \otimes B \to
+every pair of objects $B,C$ there is an object $B \multimap C$ and a morphism $\textbf{ev}_{B,C} : (B \multimap C) \otimes B \to
 C$, and for every triple of objects $A,B,C$ there is an operation $\Lambda_{A,B,C} : \catname{C}(A \otimes B,C) \to
-\catname{C}(A,B \xRightarrow{} C)$ satisfying the following equations for all $f : A \otimes B \to C$ and $g : Z \to A$:
+\catname{C}(A,B \multimap{} C)$ satisfying the following equations for all $f : A \otimes B \to C$ and $g : Z \to A$:
 \begin{itemize}
 \item $f = \Lambda_{A,B,C}(f) \otimes \id_{B}; \textbf{ev}_{B,C}$
-\item $\id_{B \xRightarrow{} C} = \Lambda_{B \xRightarrow{} C,B,C}(\textbf{ev}_{B,C})$
+\item $\id_{B \multimap{} C} = \Lambda_{B \multimap{} C,B,C}(\textbf{ev}_{B,C})$
 \item $\Lambda_{Z,B,C}(g \otimes \id_{B}; f) = g;\Lambda_{A,B,C}(f)$
@@
 
-This definition is equivalent to the more usual one that defines a (right) closed category through a pair of adjoint functors $- \otimes B$ and $B \xRightarrow{} -$~\cite{ghica-zanassi2023string}.
+This definition is equivalent to the more usual one that defines a (right) closed category through a pair of adjoint functors $- \otimes B$ and $B \multimap{} -$~\cite{ghica-zanassi2023string}.
 If the tensor is symmetric, then the category is simply \textit{closed}.
@@
 
-\textit{Closed} $\Sigma$-terms are given by the following grammar
-\[
-    \tau \coloneq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2} \; | \; \textbf{ev}_{A,B} \; | \; \Lambda(\tau)
-\]
-, where the types are given as
-\[
-A,B \coloneq T \;|\; A \otimes B \; | \; A \xRightarrow{} B
-\], where $T$ is in $\Sigma_{O}$.
+\textit{Closed} $\Sigma$-terms are defined inductively by first defining a family of objects $obj_{\Sigma_{O}}$.
+\begin{itemize}
+\item A designated $I$ is in $obj_{\Sigma_{O}}$.
+\item If $A \in obj_{\Sigma_{O}}$ and $B \in obj_{\Sigma_{O}}$, then $A \otimes B \in obj_{\Sigma_{O}}$ and $A \multimap B \in obj_{\Sigma_{O}}$.
+\end{itemize}
+A family of closed $\Sigma$ terms is then defined as
+\begin{itemize}
+    \item $\phi : A \to B \in \Sigma_{M}$ is a closed $\Sigma$-term.
+    \item If $A$ and $B$ are $obj_{\Sigma_{O}}$, then $\id_{A}$ and $\sym_{A,B}$ are closed $\Sigma$-terms.
+    \item If $f : A \to B$ and $g : B \to C$ are closed $\Sigma$-terms, then so is $f;g : A \to C$. 
+    \item If $f : A \to B$ and $g : C \to D$ are closed $\Sigma$-terms, then so is $f \otimes g : A \otimes C \to B \otimes D$.
+    \item If $A$ and $B$ are $obj_{\Sigma_{O}}$, then $\textbf{ev}_{A,B} : A \multimap B \otimes A \to B$ is a closed $\Sigma$-term.
+    \item If $h : X \otimes A \to B$ is a closed $\Sigma$-term, then $\Lambda_{X,A,B}(h) : X \to (A \multimap B)$ is a closed $\Sigma$-term.
+\end{itemize}
 \end{definition}
@@
 \begin{definition}
-The free closed symmetric monoidal category $\catname{Sc}(\Sigma)$ is given by a set of closed $\Sigma$-terms quotiented by the laws of a symmetric monoidal category and the equations from Definition~\ref{def:closed}.
+The free closed symmetric monoidal category $\catname{CS}(\Sigma)$ is given by a set of closed $\Sigma$-terms quotiented by the laws of a symmetric monoidal category and the equations from Definition~\ref{def:closed}.
 \end{definition}
@@
 To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with additional structure of a semilattice induced by an idempotent, commutative and associative operator $+$.
-This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them.
+This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix section~\ref{sec:appendix:1}).
 
@@
         \item a set of objects $\mathbb{C}$
-        \item for every pair of objects $A,B \in \mathbb{C})$ --- a hom-object $\mathbb{C}(A,B \in \catname{SLatt}$
+        \item for every pair of objects $A,B \in \mathbb{C}$ --- a hom-object (or, hom-semilattice) $\mathbb{C}(A,B) \in \catname{SLat}$
         \item for every triple of objects $A,B,C \in \mathbb{C}$ --- a composition morphism
@@
         \[
-        u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLatt}$}
+        u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLat}$}
         \]
@@
     such that particular coherence diagrams commute~\cite{Borceux_1994}.
+\end{definition}
+
+\begin{definition}[$\catname{SLat}$-functor]
+	Let $\mathbb{C}$ and $\mathbb{D}$ be two $\catname{SLat}$-categories.
+	An $\catname{SLat}$-functor $F : \mathbb{C} \to \mathbb{D}$ is defined by the following data.
+	\begin{itemize}
+	  \item A mapping $F : \mathbb{C} \to \mathbb{D}$
+	  \item An object-indexed family of morphisms in $\catname{SLat}$ $F_{A,B} : \mathbb{C}(A,B) \to \mathbb{D}(FA,FB)$
+	\end{itemize}
+	such that certain coherence diagrams commute~\cite{Borceux_1994}.
+	$F$ being a morphism in $\catname{SLat}$ has a property $F(f + g) = F(f) + F(g)$.
 \end{definition}
@@
     \end{definition}
-    The latter morphism being in $\catname{SLatt}$ implies the following equations that $\otimes$ satisfy
+    The latter morphism being in $\catname{SLat}$ implies the following equations that $\otimes$ satisfy
     \[
@@
     \[\begin{tikzcd}
-        \catname{SLatt}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
+        \catname{SLat}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
         \end{tikzcd}
@@
     \[\begin{tikzcd}
-        \catname{SLatt} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
+        \catname{SLat} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
         \end{tikzcd}
@@
 Being a 2-functor it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{Slat}$-adjunction.
-Thus, applying the functor to $\catname{Sc}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{Sc}(\Sigma)^{+}$ by lifting the adjunction $- \otimes B \dashv B \xRightarrow{} -$.
+Thus, applying the functor to $\catname{CS}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{CS}(\Sigma)^{+}$ by lifting the adjunction $(- \otimes B, B \multimap{} -, \eta, \varepsilon)$.
+We can similarly derive $\catname{SLat}$-enriched versions of  $\Lambda_{A,B,C}$ and $\textbf{ev}_{A,B}$ morphisms using the lifted adjunction as $\textbf{ev}_{A,B} = \varepsilon_{A}$ and $\Lambda_{A,B,C}(f) = \eta; (B \multimap{} f)$.
+The following equation can then be derived which will be useful in encoding equivalences inside abstractions.
+\begin{align*}
+\Lambda_{A,B,C}(f + g) &= \eta;(B \multimap (f + g))\\
+                       &= \eta;(B \multimap (f) + B \multimap (g))\\
+                       &= \eta;(B \multimap (f)) + \eta;(B \multimap (g))\\
+                       &= \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g)
+\end{align*}
+
+Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of $\Sigma^{+}$-terms.

### 2025-01-19 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
wip

@@
 \section{Combinatorial semantics}
+
+\textcolor{red}{Add intro to hypergraphs}
 
@@
 \begin{definition}
-Closed monoidal signature $\Sigma$ is a pair $(\Sigma_{O},\Sigma_{M})$ of signatures for objects and signatures for morphisms with sources and targets in $\text{obj}_{\Sigma_{O}}$ defined as
+We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{red}{<}, \textcolor{blue}{<},\consistency)$ where
 \begin{itemize}
-  \item $I$ (monoidal unit) is in $\text{obj}_{\Sigma_{O}}$
-  \item For every $o \in \Sigma_{O}$, $o \in \text{obj}_{\Sigma_{O}}$;
-  \item For every $o_1$ and $o_2$ $ \in \text{obj}_{\Sigma_{O}}$, $o_1 \otimes o_2 \in \text{obj}{\Sigma_{O}}$ and $o_1 \multimap o_2 \in \text{obj}_{\Sigma_{O}}$
+  \item $V$ is a set of vertices.
+  \item $E = \textcolor{gray}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
+  \item $s,t : E \to V^{*}$ are source and target functions.
+  \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are \textit{label} functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
+  \item $\textcolor{red}{<} \subseteq \textcolor{red}{E} \times V + E$ and $\textcolor{blue}{<} \subseteq \textcolor{blue}{E} \times V + E $ with $\textcolor{red}{E}, \textcolor{blue} \subset E$ are disjoint partial orders.
+        For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{red}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{blue}{<^{\mu}} y_{i+1}\; \}$.
+        We will denote the immediate predecessor $y$ of $x$ as $y \textcolor{red}{<^{\mu}} x$ and $y \textcolor{blue}{<^{\mu}} x$.
+        We will call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
+        Each of these relations should satisfy the following
+        \begin{enumerate}
+          \item each set of predecessors consists exclusively of hierarchical edges;
+          \item each $x$ has at most one immediate predecessor;
+          \item edges $e$ such that there is no $x$ such that $e \textcolor{red}{<^{\mu}} x$ or $e \textcolor{red}{<^{\mu}} x$ are labelled;
+          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{red}{<^{\mu}} e$ iff $e' \textcolor{red}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{blue}{<}$.
+        \end{enumerate}
+  \item $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p \textcolor{red}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{red}{<^{\mu}}(v)$ and $p \textcolor{red}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
+  We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{red}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{red}{<^{\mu}} e\}$.
 \end{itemize} 
+$\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations can consist of exclusively hierarchical edges, then we have $\textcolor{gray}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
 \end{definition}
 
-\begin{definition}
- Closed symmetric monoidal theory ($\catname{CSMT}$) over a closed monoidal signature $\Sigma$ and equations $\mathcal{E}$ denoted as $\catname{CSMT}(\Sigma, \mathcal{E})$ is given by a set of terms $T$
- \begin{align*}
- T \coloneq\; &t : A \to B \in \Sigma_{M} \;|\; t_1 \otimes t_2 : A \otimes C \to B \otimes D \text{ for } t_{1} : A \to B, t_{2} : C \to D \in T\\
-            & \;|\; t_{1};t_{2} : A \to C \text{ for } t_{1} : A \to B, t_{2} : B \to C \in T \; | \;  \Lambda(t) : A \to X \multimap B \text{ for } t : A \otimes X \to B \\
-            &  \;|\; \text{eval}(t_1 \otimes t_2) : B \text{ for } t_1 : X \multimap B, t_2 : X \in T \\
-            &  \;|\; \text{sym}_{A,B} \;|\; id_{A}
- \end{align*}
- quotiented by the laws of a closed symmetric monoidal category.
-\end{definition}
-% \question{Maybe types and terms combined is not a good idea}
-\begin{definition}
+Intuitively, edges from $\textcolor{red}{E}$ will encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ will encode lambda-abstraction boxes.
 
-We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{red}{<}, \textcolor{blue}{<},\consistency)$ where
-\begin{itemize}
-  \item $V$ is a set of vertices;
-  \item $E = \textcolor{gray}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that include \textit{plain} edges and two types of \textit{hierarchical} edges respectively;
-  \item $s,t : E \to V^{*}$ are source and target functions;
-  \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \text{obj}_{\Sigma_{O}}$ are \textit{label} functions;
-  \item $\textcolor{red}{<} : V + E \to \textcolor{red}{E}$ and $\textcolor{blue}{<} : V + E \to \textcolor{blue}{E}$ are partial \textit{child} functions that are disjoint: if $\textcolor{red}{<}(x)$ is defined then $\textcolor{blue}{<}(x)$ is undefined and vice-versa;
-  \item $\consistency = \bigcup_{p} \consistency_{p}$, where $\consistency_{p}$ is a \textit{consistency} relation defined on each set $\{x \in V + E \;|\; p \;\textcolor{red}{<^{\mu}}\; x\}$;
-\end{itemize} 
+Consider an example in the middle part of Figure~\ref{fig:e-cospan-example}.
+The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
+\[
+\begin{array}{cccc}
+  v_{4} \textcolor{red}{<^{\mu}} e_{1} & v_{7} \textcolor{blue}{<^{\mu}} e_{2} & v_{4} \consistency e_{3} & v_{4} \not \consistency v_{10}\\
+  e_{3} \textcolor{red}{<^{\mu}} e_{1} & e_{5} \textcolor{blue}{<^{\mu}} e_{2} &  e_{3} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
+  e_{2} \textcolor{red}{<^{\mu}} e_{1} & w_{7} \textcolor{blue}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
+  \ldots & \ldots & \ldots & \ldots \\
+  v_{12} \textcolor{red}{<^{\mu}} e_{1} & w_{9} \textcolor{blue}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
+\end{array}
+\]
+$\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
+Edges in $\textcolor{red}{E}$ are depicted with a dashed box, while edges in $\textcolor{blue}{E}$ --- with a round solid box.
+
+\begin{figure}
+
+\[
+\adjustbox{scale=0.6}{
+\tikzfig{./figures/closed_iso_example_2}
+}
+\]
+\caption{Cospan of e-hypergraphs example}
+\label{fig:e-cospan-example}
+\end{figure}
+
+\begin{definition}[E-hypergraph homomorphism]
+TODO
 \end{definition}
 
-$\textcolor{red}{E}$ will contain edges that encode the equivalence classes and $\textcolor{blue}{E}$ the ones that defined the lambda abstraction.
-All the above functions and relations must satisfy the obvious properties (TODO: add them).
+E-hypergraphs and their homomorphisms define a category $\catname{EHyp}({\Sigma})$.
 
@@
 
-Consider an example of isomorphic and non-isomorphic cospans below and their respective string diagrams
+% Consider an example of isomorphic and non-isomorphic cospans below and their respective string diagrams
 
-\begin{figure}
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.6}{
-    \tikzfig{./figures/closed_iso_example_1}
-    }
-    \]
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[c]{0.4\linewidth}
-    \[
-    \adjustbox{scale=0.6}{
-    \tikzfig{./figures/closed_iso_example_2}
-    }
-    \]
-  \end{subfigure}
-\end{figure}
+% \begin{figure}
+%   \begin{subfigure}[c]{0.4\linewidth}
+%     \[
+%     \adjustbox{scale=0.6}{
+%     \tikzfig{./figures/closed_iso_example_1}
+%     }
+%     \]
+%   \end{subfigure}
+%   \hfill
+%   \begin{subfigure}[c]{0.4\linewidth}
+%     \[
+%     \adjustbox{scale=0.6}{
+%     \tikzfig{./figures/closed_iso_example_2}
+%     }
+%     \]
+%   \end{subfigure}
+% \end{figure}
 
@@
 \end{definition}
+
+\textcolor{red}{TODO: dpo rewriting}

### 2025-01-21 LICS/sections/appendix.tex
wip

@@
 \section{$\catname{SLat}$}
-\label{sec:appendix:1}
+\label{sec:appendix:slat}
 
@@
 \section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
-\label{sec:appendix:2}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
+\label{sec:appendix:pushout}
+
+In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we will explicitly construct such a pushout.
 We will also show the uniqueness of a boundary pushout complement.
@@
     Then we need $f_{ext};\beta = f_{ext}'$ and we define $\beta = f_{ext}^{-1};f_{ext}'$.
-    Similarly $\beta^{-1} = f_{ext}'^{-1};f_{ext}$. 
+    Similarly, $\beta^{-1} = f_{ext}'^{-1};f_{ext}$. 
     We can check that $\beta^{-1};\beta = f_{ext}'^{-1};f_{ext};f_{ext}^{-1};f_{ext}' = f_{ext}'^{-1};id;f_{ext}' = f_{ext}'^{-1};f_{ext}' = id$.
@@
 \section{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
+\label{sec:appendix:iso}
 \begin{example}

### 2025-01-21 LICS/sections/combinatorial_semantics.tex
wip

@@
 This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of such hypergraphs.
+Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
 
@@
 \label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
+Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete \textit{ordered} hypergraphs as objects.
 \end{definition}
@@
 The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
+The ordering of the interface hypergraphs is the way to differentiate between the interpretations of morphisms $f \otimes g$ and $g \otimes f$.
 
@@
 
-In order to represent only morphisms of $\catname{S}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
+In order to make the interpretation complete with respect to $\catname{S}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
+\begin{definition}[Category of MDA Hypergraphs with Interfaces~\cite{bonchi_string_2022-2}]
 \label{def:monogamy_hyp}
@@
 Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-Yet, this restriction is not enough to make DPOI rewriting complete for $\catname{S,\mathcal{E}}$: there might be cospans with carriers $\mathcal{G}$ and $\mathcal{H}$ such that one can be rewritten into another but their corresponding morphisms are not equal modulo $\mathcal{E}$~\cite{bonchi_string_2022-1}.
+Yet, this restriction is not enough to make DPOI rewriting complete for $\catname{S}(\Sigma,\mathcal{E})$: there might be cospans with carriers $\mathcal{G}$ and $\mathcal{H}$ such that one can be rewritten into another but their corresponding morphisms are not equal modulo $\mathcal{E}$~\cite{bonchi_string_2022-2}.
 To make the DPOI rewriting complete, we need to add a restriction on the image of the match.
@@
 %By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\catname{S}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
+We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\catname{S}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as (cospans of) hypergraphs.
 In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.

### 2025-01-21 LICS/sections/e-hypergraphs.tex
wip

@@
 Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.  
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
+% An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
 
@@
 We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p <^{\mu} v\}$ and $E_{p} = \{ e ~ | ~ p <^{\mu} e\}$.
+Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $<$ and $\consistency$.
 % We collect the union of all these equivalence relations into one relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
@@
 % Technically,  condition $(4)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for the consistency relation. 
-An example of an e-hypergraph can be seen below.
+An example of an e-hypergraph can be seen below in Figure~\ref{fig:e-cospan-example} (within a grey shaded region).
+The e-hypergraph is given by $E = \{e_1, e_2\}, V = \{v_1, \ldots, w_4 \}$, $s$,$t$, $l = \{e_1 \mapsto \bot, e_2 \mapsto f\}$ and
 \[
-\adjustbox{scale=0.75}{
-\tikzfig{../figures/combinatorial_semantics/e-hypergraph-example-2}
-}
+\begin{array}{ccc}
+    v_3 <^{\mu} e_1 & v_3 \consistency v_4 & v_3 \not \consistency v_5\\
+    \ldots & \ldots & \ldots\\
+    w_4 <^{\mu} e_1 & w_3 \consistency w_4 & v_4 \not \consistency w_4
+\end{array}
 \]
-
-This e-hypergraph corresponds to the following tuple
-\[
-(V = (v_1, v_2, v_3, v_4, v_5, v_6\}, E = \{e_1, e_2, e_3\}, s, t, l, <, \consistency)
-\]
-\begin{align*}
-s(e_1) = \{v_1\} &\; t(e_1) = \{v_2\} &\; l(e_1) = \bot \\
-s(e_2) = \{v_3\} &\; t(e_2) = \{v_4\} &\; l(e_2) = e_2 \\
-s(e_3) = \{v_5\} &\; t(e_3) = \{v_6\} &\; l(e_3) = e_3 \\
-\end{align*}
-
-and $[v_3) = [v_4) = [v_5) = [v_6) = [e_2) = [e_3) = \{e_1\}$ and $[v_1) = [v_2) = [e_1) = \varnothing$.
-$v_3 \consistency v_4 \consistency e_2$ and $v_5 \consistency e_3 \consistency v_6$ and the relation is undefined for the rest vertices and edges.
-Because $v_3 \not \consistency v_5$ and because the relation is closed under connectivity we delimit the connected components by a dashed vertical line.
+A hierarchical edge is depicted with a dashed box.
+The partition induced by $\consistency_{e_1}$ is depicted with a dashed vertical line.
 The condition $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$, in particular, does not allow hierarchical edges with no delimiting lines.
-$e_2$ and $e_3$ are maximal edges.
+$e_1$ is the maximal edge and hence labelled.
 \begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ between e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
 \begin{enumerate}
@@
 The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.  
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied, including the requirement that the shared domain of the morphisms that form the span for the pushout is a discrete e-hypergraph.
+A concrete description of the pushout of two morphisms in this category is given in the Appendix section~\ref{sec:appendix:pushout}.
+Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does in all scenarios that we are interested in, most importantly the pushout for the composition of two cospans with discrete feet exists.
 
@@
 
+\end{definition}
 We will call $n$  \textit{external input interfaces} and $n'$ \textit{internal input interfaces}.
@@
 Consider a relation $R = \{x R y \text{ if } f_{int}(x) \consistency f_{int}(y) \text{ for } x,y \in n'\}$ and let $S$ be its reflexive closure.
-The latter partitions $n'$ into non-empty subsets $\{p_{j}\}_{j=1}^{k}$. We get an analogous partition for $m'$.
+The latter partitions $n'$ into non-empty subsets $\{p_{j}\}_{j=1}^{k}$. 
+We get an analogous partition for $m'$.
 An example of an extended cospan can be seen in Figure~\ref{fig:e-cospan-example}.
-The cospan represents a morphism that encodes symmetry put into a hierarchical edge.
 The partition of the interfaces is given with colours.
@@
 Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{bonchi_string_2022-1}.
-Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix.
+Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix section~\ref{sec:appendix:iso}.
 
@@
 $\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.
-\end{definition}
 
@@
         \item $f_{int}$, $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
+        \item in-degree and out-degree of every vertex is at most 1; 
+        \item vertices with in-degree $0$ are precisely the image of $f_{int}$; 
+        \item vertices with out-degree $0$ are precisely the image of $g_{int}$; 
         % the item below is not assumed by the definition of a cospan of e-hypergraphs
@@
 We call $e$ \emph{well-typed} if the size of each subset in the partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
+We will call an MDA extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its carrier are well-typed.
 We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA extended-cospans. 
@@
 
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.  
+Unlike in the previous section,  in order to give an equivalence with $\catname{S}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
 This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.

### 2025-01-21 LICS/sections/introduction.tex
wip

@@
 We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.  
-We denote by $\catname{SLat}$ the closed monoidal category of semilattices which is defined in section~\ref{sec:appendix:1} of the Appendix.
+We denote by $\catname{SLat}$ the closed monoidal category of semilattices which is defined in section~\ref{sec:appendix:slat} of the Appendix.
 
@@
 
-\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also Appendix section~\ref{sec:appendix:1})
+\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also Appendix section~\ref{sec:appendix:slat})
 There is a 2-adjunction 

### 2025-01-21 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
wip

@@
 To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with additional structure of a semilattice induced by an idempotent, commutative and associative operator $+$.
-This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix section~\ref{sec:appendix:1}).
+This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix section~\ref{sec:appendix:slat}).
 
@@
 
-    \begin{proposition}(A specialized case of Proposition 6.4.7~\cite{Borceux_1994})
+    \begin{proposition}(A specialised case of Proposition 6.4.7~\cite{Borceux_1994})
     There is a 2-adjunction 
@@
 
-Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of $\Sigma^{+}$-terms.
+Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of $\Sigma^{+}$-terms, the latter being defined as follows.
+Every $\Sigma$-term is also a $\Sigma^{+}$-term. 
+Given two $\Sigma^{+}$-terms $f : A \to B$ and $g : A \to B$, $f + g : A \to B$ is also a $\Sigma^{+}$-term.
+
+To aid reasoning, we introduce a new language of string diagrams for $\catname{SLat}$-SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
+It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
+Figure \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} apart from the last component which is our notation for $+$; the first component denotes an empty diagram.
+Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
+The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
+We overload the binary notation $+$ for our $n$-ary notation.
+We will later prove the intuitive fact that these diagrams are sound and complete with respect to their intended categorical semantics, noting that similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.

### 2025-01-21 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
wip

@@
   \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are \textit{label} functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
-  \item $\textcolor{red}{<} \subseteq \textcolor{red}{E} \times V + E$ and $\textcolor{blue}{<} \subseteq \textcolor{blue}{E} \times V + E $ with $\textcolor{red}{E}, \textcolor{blue} \subset E$ are disjoint partial orders.
+  \item $\textcolor{red}{<} \subseteq \textcolor{red}{E} \times V + E$ and $\textcolor{blue}{<} \subseteq \textcolor{blue}{E} \times V + E $ with $\textcolor{red}{E}, \textcolor{blue} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{red}{<^{\mu}}$ and $\textcolor{blue}{<^{\mu}}$.
         For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{red}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{blue}{<^{\mu}} y_{i+1}\; \}$.
-        We will denote the immediate predecessor $y$ of $x$ as $y \textcolor{red}{<^{\mu}} x$ and $y \textcolor{blue}{<^{\mu}} x$.
         We will call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
@@
 $\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations can consist of exclusively hierarchical edges, then we have $\textcolor{gray}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
+Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $<$ and $\consistency$.
 \end{definition}
@@
 Intuitively, edges from $\textcolor{red}{E}$ will encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ will encode lambda-abstraction boxes.
+Also note the typing on the $l_{V}$ function.
+We work with freely generated closed semilattice-enriched categoeries and to avoid unnecessary complications we limit the labelling of vertices to such a type.
+In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
 
 Consider an example in the middle part of Figure~\ref{fig:e-cospan-example}.
-The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
+The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{red}{E} = \{e_{1}\}$, $\textcolor{blue}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
 \[
 \begin{array}{cccc}
-  v_{4} \textcolor{red}{<^{\mu}} e_{1} & v_{7} \textcolor{blue}{<^{\mu}} e_{2} & v_{4} \consistency e_{3} & v_{4} \not \consistency v_{10}\\
-  e_{3} \textcolor{red}{<^{\mu}} e_{1} & e_{5} \textcolor{blue}{<^{\mu}} e_{2} &  e_{3} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
-  e_{2} \textcolor{red}{<^{\mu}} e_{1} & w_{7} \textcolor{blue}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
+  v_{4} \textcolor{red}{<^{\mu}} e_{1} & v_{7} \textcolor{blue}{<^{\mu}} e_{2} & v_{4} \consistency e_{2} & v_{4} \not \consistency v_{10}\\
+  e_{3} \textcolor{red}{<^{\mu}} e_{1} & v_{8} \textcolor{blue}{<^{\mu}} e_{2} &  e_{2} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
+  e_{2} \textcolor{red}{<^{\mu}} e_{1} & e_{5} \textcolor{blue}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
   \ldots & \ldots & \ldots & \ldots \\
@@
 \]
+and, in particular, we have $[e_{5}) = \{e_2, e_1\}$.
 $\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
@@
 \begin{definition}[E-hypergraph homomorphism]
-TODO
+  An \textit{e-hypergraph homomorphism} $\phi:\mathcal{F}\to\mathcal{G}$ between e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+  \begin{enumerate}
+      \item $\phi$ is a hypergraph homomorphism;
+      \item $\phi_{E}(\textcolor{red}{E_{\mathcal{F}}}) \subseteq \textcolor{red}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{blue}{E_{\mathcal{F}}}) \subseteq \textcolor{blue}{E_{\mathcal{G}}}$;
+      \item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
+        \begin{align*}
+          &\text{if } e \textcolor{red}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(v)\\
+          &\text{and}\\
+          &\text{if } e \textcolor{blue}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(v)
+        \end{align*}
+          and for $e_1, e_2 \in E_{\mathcal{F}}$
+          \begin{align*}
+          &\text{if } e_1 \textcolor{red}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(e_2)\\
+          &\text{and}\\
+          &\text{if } e_1 \textcolor{blue}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+          \end{align*}
+          \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
+  \end{enumerate}
+  % If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
+  \end{definition}
+  The conditions on e-hypergraph homomorphisms require preserving immediate predecessors and the consistency relation.
+  \begin{definition}[Category of e-hypergraphs]
+  The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
   \end{definition}
 
-E-hypergraphs and their homomorphisms define a category $\catname{EHyp}({\Sigma})$.
+The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  
+A concrete description of the pushout of two morphisms in this category is given in the Appendix section~\ref{sec:appendix:pushout}.
+Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied, most importantly a pushout for the composition of two cospans with discrete feet exists as described below.
 
-Similarly, we have a category of (extended) cospans formed by such e-hypergraphs where morphisms are of the form
+As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, we need a notion of an interface for our e-hypergraphs for them to model string diagrams as the latter have dedicated inputs and outputs.
+Because string diagrams for closed semilattice-enriched symmetric monoidal categories are nested: they contain sub string diagrams inside e-boxes and lambda abstraction boxes we need to account for the interfaces in these nested contexts as well.
+It is worth noting that these nested inputs and outputs of a string diagram do not participate in composition of string diagrams and hence they need to be put aside from the ones that participate in the composition.
+We thus arrive to a notion of an \textit{extended} cospans of discrete e-hypergraphs and will use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n \setminus m$ vertices, in particular
+with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
+
+\begin{definition}[Category of e-hypergraphs with interfaces]
+  The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete \textit{ordered} e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes (defined below) of \textit{extended cospans}, defined as follows:
   \[
-n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+  n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
   \]
-
-where $n$ and $m$ are external input and output interfaces respectively and $n'$, $m'$ are input and output internal interfaces.
-$n,n',m,m'$ are discrete ordered e-hypergraphs.
-We further require that $f_{\text{ext}}$ and $g_{\text{ext}}$ are monos and all vertices in the image of $f_{\text{ext}};f_{\text{int}}$ (respectively, $g_{\text{ext}};g_{\text{int}}$) are top-level.
-
-To build a correspondence between terms of a closed SMC and e-hypergraphs with interfaces (above) we restrict the cospans of the latter to monogamous cospans.
-
-\begin{definition}
-  We call a cospan 
-  \[
-n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
-\]
-\textit{weak} monogamous if
-\begin{itemize}
-  \item in- and out- degrees of every vertex is at most 1;
-  \item $f_{\text{int}}$ and $g_{\text{int}}$ are monos;
-  \item Vertices with in-degree (respectively, out-degree) of 0 are precisely the image of $f_{\text{int}}$ (respectively, $g_{\text{int}}$).
-\end{itemize}
+  where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete ordered e-hypergraphs,  $f_{ext},g_{ext}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
+  We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
 
 \end{definition}
+We will call $n$  \textit{external input interfaces} and $n'$ \textit{internal input interfaces}.
+We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.  
+We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.  
+We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
+Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
+Graphically we will depict such extended cospans as in Figure~\ref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
+All non-black vertices form strictly internal interfaces.
 
@@
 
-% Consider an example of isomorphic and non-isomorphic cospans below and their respective string diagrams
+In Figure~\ref{fig:e-cospan-example} the partitions for input and output internal interfaces are given with colours.
 
-% \begin{figure}
-%   \begin{subfigure}[c]{0.4\linewidth}
+Composition of two morphisms 
+\begin{align*}
+	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
+	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
+\end{align*} is computed in two stages.
+First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
+% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
+\[
+\trimbox{0cm 0cm 0cm 0.75cm}{
+\adjustbox{scale=0.9,center}
+{\tikzfig{./figures/pushout_e_cospans}}
+}
+\]
+then,  the result of composition is defined as follows:
+\[
+n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
+\]
+where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
+\[
+    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
+\quad
+    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
+\]
 % \[
-%     \adjustbox{scale=0.6}{
-%     \tikzfig{./figures/closed_iso_example_1}
-%     }
+% \begin{tikzcd}
+% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
+% Y'          & Y' \setminus Y_{ext} \arrow[l]
+% \end{tikzcd}
 % \]
-%   \end{subfigure}
-%   \hfill
-%   \begin{subfigure}[c]{0.4\linewidth}
-%     \[
-%     \adjustbox{scale=0.6}{
-%     \tikzfig{./figures/closed_iso_example_2}
-%     }
-%     \]
-%   \end{subfigure}
-% \end{figure}
+% and then the disjoint union is just a coproduct. 
+% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
+The identity of composition is given by the obvious extended cospan of identities. 
+$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.
+As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
 
+For the same reasons as in the case of $\catname{S}(\Sigma)$ we will restrict cospans to MDA cospans.
+We also need to enforce typing constraints on hierarchical edges so that they represent well-typed closed $\Sigma^{+}$-terms.
 \begin{definition}
+  We call a cospan 
+  \[
+n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+\]
+monogamous directed acyclic if
+\begin{itemize}
+  \item underlying hypergraph of $\mathcal{G}$ is directed acyclic;
+  \item in- and out- degrees of every vertex is at most 1;
+  \item $f_{\text{int}}$ and $g_{\text{int}}$ are monos;
+  \item vertices with in-degree (respectively, out-degree) of 0 are precisely the image of $f_{\text{int}}$ (respectively, $g_{\text{int}}$).
+\end{itemize}
 
-We define ordered sets of \textit{input} (respectively, \textit{output}) vertices of a hierarchical edge $e$ as the vertices $v_{i}$ such that $e < v_{i}$ which are in the image of $f_{\text{int}}$ (respectively, $g_{\text{ext}}$) with the induced ordering.
-\end{definition}
-
-\begin{definition}
-
-Given an ordered set $S$ of labelled vertices we can form words formed by concatenating the labels.
-We will denote such words as $w(S)$.
+MDA cospans and discrete e-hypergraphs form a category $\MdaEcospans$.
 \end{definition}
@@
     \item For each $e \in \textcolor{blue}{E}$ consider sets $I$ and $O$ of its input and output vertices.
-        $e$ is well-typed if there exists an object $B \in \text{obj}_{\Sigma_{O}}$ such that $[...s(e), B] = w(I)$ and $B \multimap w(O) = w(t(e))$
+          $e$ is well-typed if there exists an object $B \in \text{obj}_{\Sigma_{O}}$ such that $[w(s(e)), B] = w(I)$ and $B \multimap w(O) = w(t(e))$.
   \end{itemize}
+
+  We will denote the category of well typed MDA cospans as $\WellTypedMdaEcospans$.
+  It is a proper category as identities, symmetries, and tensor unit are well-typed and composition and tensor product of two well-typed MDA cospans is again a well typed MDA cospan.
   \end{definition}
 
-\textcolor{red}{TODO: dpo rewriting}
+Consider, again, Figure ~\ref{fig:e-cospan-example}.
+To make edge $e_{1}$ well-typed, it is necessary that $w([v_{4},v_{5},v_{6}]) = w([v_{10},v_{11},v_{12}]) = w([v_1,v_2,v_3])$ and $w([w_{4},w_{5},w_{6}]) = w([w_{10},v_{12},v_{11}]) = w([w_1,w_2,w_3])$ (note the change of ordering in the output of $e_{1}$).
+To make edge $e_{2}$ well-typed, it is necessary that $w([v_4,v_5,v_9]) = w([v_7,v_8,v_9])$ and $w([u_1]) = w([v_9]) \multimap w([v_8,v_7,w_9])$.
+
+Unlike in the previous section,  in order to give an equivalence with $\catname{S}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
+This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.
+Instead,  we will treat them via rewriting,  as we do with equations of the signature.
+This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
+% , cf. string diagrams for Cartesian categories.  
+An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this.
+Presently,  we will treat these equations via rewriting as well.
+
+\section{DPOI-Rewriting for E-Hypergraphs}
+
+Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
+This is the purpose of this section of the paper.
+
+We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
+Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
+\[
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
+\qquad
+    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
+\] 
+Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
+\[
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
+\qquad
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
+\] allows us to encode rewrite rules as extended cospans of the following form
+\[
+\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
+\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
+%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
+\begin{remark}
+The extended DPOI rewriting as defined in this section trades off simplicity for expressiveness: it is unable to express rewriting for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
+While this limitation can be overcome by specifically encoding such rewrite rules, we find the technical development to be simpler if we avoid consideration of such rules and therefore consider signatures which can not result in construction of terms with type $0 \to 0$.
+This limitation has no practical impact in applications of interest. 
+However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of Appendix.
+%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
+%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
+%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
+\end{remark}
+
+% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
+% Note that the outermost interfaces are preserved.
+
+% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
+% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
+
+% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
+% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
+% The entry in the middle of the span depicts the outermost interfaces which are preserved.
+% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
+% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
+% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
+% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
+
+Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
+First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
+\begin{definition}[Down-closed subgraph]
+%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
+%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
+%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
+    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
+\end{definition}
+
+\begin{definition}[Extended boundary complement]
+\label{def:boundary_new}
+
+% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
+
+For MDA cospans 
+\[
+    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
+\quad\text{ and }\quad
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
+\] and mono $m : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
+as depicted in the square below
+\[
+\scalebox{0.75}{
+
+    \tikzfig{./figures/DPOI_pushout_complement}
+}
+\]
+is a \textit{boundary complement} if
+\begin{enumerate} 
+    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
+    \item $[c_1,c_2]$ is mono;
+    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
+    \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
+    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
+    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended MDA cospan
+    \[
+    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+    \]
+    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended MDA cospan
+    \[
+    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+    \]
+\end{enumerate}
+where $f_i$ and $g_i$ are defined as follows.  
+ %because the image of $m$ is a convex down-closed e-hypergraph, 
+The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
+Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
+The maps $f_2$ and $g_2$ are defined similarly.
+%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
+%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
+%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
+%
+%\[
+%    \tikzfig{combinatorial_semantics/pushout_difference}
+%\]
+%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
+%Similarly for $g_2$.
+% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
+% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
+% \[
+%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
+% \]
+% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
+% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
+% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
+% Similarly for $k \to r$.
+\end{definition}
+
+Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
+In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding MDA cospans above.
+The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
+
+%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
+%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
+%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
+%
+%\begin{figure}
+%    \centering
+%    \[
+%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
+%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
+%    \label{fig:interface_change_example}
+%\end{figure}
+
+% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
+\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
+
+Note that,  in the above definition,  the extended cospan 
+    \[
+    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
+    \]
+ is not necessarily well-typed.
+This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
+\begin{remark}
+The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+\[
+	\mathcal{L} = \scalebox{0.8}{\tikzfig{./figures/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{./figures/f_plus_g_inline}}
+\]
+\end{remark}
+
+\begin{proposition}
+\label{prop:boundary_unique}
+    The boundary complement in~\ref{def:boundary_new} when exists is unique.
+\end{proposition}
+
+We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
+It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
+More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
+Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
+
+\begin{definition}[Convex EDPOI rewriting]
+\label{def:dpoi-e}
+Given an extended span of morphisms 
+\[
+    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
+\]
+in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
+\[
+    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
+\] 
+if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
+\[
+ \scalebox{0.75}{
+    \tikzfig{./figures/DPOI_square}
+ }
+\]
+such that the two marked squares are pushouts and the following conditions hold:
+    \begin{enumerate}
+        \label{dpoi-e:assumptions}
+       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
+        %\item $m(L)$ is a down-closed e-hypergraph;
+        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
+	\item the internal interfaces of $\mathcal H$ are such that:
+	\[
+		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
+	\]
+	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
+    \end{enumerate}
+% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
+%  \[
+%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
+% \]
+    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
+\end{definition}
+% Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
+
+With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.

### 2025-01-22 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
fixed some references and figs

@@
 \usepackage{xfrac} 
+\usepackage{wrapfig}
 
@@
 Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
-However, calculi of explicit substitutions have many technical subtleties which also directly impact how equivalence is defined, for instance \emph{graph equivalence} instituted between terms which have non-interfering explicit substitutions applied in different orders~\cite{doi:10.1145/2535838.2535886}.
+However, calculi of explicit substitutions have many technical subtleties which also directly impact how equivalence is defined, for instance \emph{graph equivalence} instituted between terms which have non-interfering explicit substitutions applied in different orders~\cite{accattoli2014nonstandard}.
 These difficulties are not handled in \emph{loc. cit.}.
@@
 
-The general background literature for string diagrams is summarised in~\cite{https://arxiv.org/abs/2305.08768}.
-The particular context of string diagrams with bindings is given in~\cite{https://arxiv.org/abs/2305.18945}, while string diagrams for e-graphs without bindings are defined in~\cite{https://arxiv.org/abs/2406.15882}.
+The general background literature for string diagrams is summarised in~\cite{piedeleu2023introductionstringdiagramscomputer}.
+The particular context of string diagrams with bindings is given in~\cite{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
 
@@
 This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by symbolic notation. 
-Originating from notations by Hotz~\cite{??} (for automata) and Penrose~\cite{??} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, offering a dual perspective to commutative diagrams through Poincaré duality.
+Originating from notations by Hotz~\cite{hotzsd} (for automata) and Penrose~\cite{penrose1984spinors} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, offering a dual perspective to commutative diagrams through Poincaré duality.
 
 As described above, string diagrams are commonly an informal yet rigorous two-dimensional syntax for the internal language of a category. 
-They can be formalised by representing them as graphs or hypergraphs~\cite{https://dl.acm.org/doi/full/10.1145/3502719}, which has two further benefits. 
+They can be formalised by representing them as graphs or hypergraphs~\cite{1011453502719}, which has two further benefits. 
 The first benefit is that the property of two string diagrams being equal if and only if they ``look the same'' can be spelled out as graph isomorphism, eliminating any possible ambiguity of interpretation.
@@
 
-Finally, to represent diagrammatically two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{https://link.springer.com/chapter/10.1007/11874683_1}.
+Finally, to represent diagrammatically two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
 The diagrammatic boxes require the introduction of further structure to (hyper)graphs in the form of a hierarchical layering of the graphs, as studied in \emph{loc. cit.}.
+
+\begin{wrapfigure}{L}{0.2\textwidth}
+\[
+\adjustbox{scale=0.6}{
+    \tikzfig{figures/de-brujin-string}
+}
+\]
+\end{wrapfigure}
 
@@
 }, 
-and sharing a \emph{contraction node} DRG: IT'S UPSIDE DOWN?
+and sharing a \emph{contraction node} 
 \scalebox{0.3}{
@@
 .
-
-DRG USE WRAP-BOX. 
-\[
-\adjustbox{scale=0.6}{
-    \tikzfig{figures/de-brujin-string}
-}
-\]
 
@@
 Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes
+
 \scalebox{0.4}{

### 2025-01-22 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
section 2 edits

@@
 
-In this section we introduce the required preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories, and the string diagrammatic formalism we will use to represent them.
-We also show a specific case of semilattice-enriched closed monoidal categories and how closed structure interacts with the enrichment structure.
+This section introduces preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories, and the string diagram formalism used to represent them.
+We also show a specific case of semilattice-enriched closed monoidal categories and how the closed structure interacts with the enrichment structure.
 
-We will consider symmetric monoidal categories generated by \textit{monoidal signatures}.
-
+We recall some basic definitions.
 \begin{definition}
-
-A \textit{monoidal} \textit{signature} $\Sigma = (\Sigma_{O}, \Sigma_{M}, t)$ is given by a set of objects $\Sigma_{O}$, set of generators $\Sigma_{M}$ and a type assignment function $t : \Sigma_{M} \to \Sigma_{O}^{*} \times \Sigma_{O}^{*}$ that assigns lists of input and output types to a given generator.
+A \textit{monoidal} \textit{signature} $\Sigma = (\Sigma_{O}, \Sigma_{M}, t)$ is given by a set of objects (called types) $\Sigma_{O}$, a set of generators (operations) $\Sigma_{M}$ and a type assignment function $t : \Sigma_{M} \to \Sigma_{O}^{*} \times \Sigma_{O}^{*}$ assigning lists of input and output types to a given operation.
 \end{definition}
@@
 \begin{definition}
-
 A set of $\Sigma$-terms generated by monoidal signature $\Sigma$ is given by the following grammar
@@
 \],
-where $\phi$ ranges over the set of generators $\Sigma_{M}$ and $A,B$ range over the set of objects $\Sigma_{O}$.
+where $\phi$ ranges over the set of operators $\Sigma_{M}$ and $A,B$ range over the set of objects $\Sigma_{O}$.
 \end{definition}
 
-Identities and symmetries of more complex types, \textit{e.g.}, $\id_{A \otimes B}$ and $\sym_{A \otimes B, C}$, are given by the composition and tensoring of $\id_{A}$ and $\sym_{A,B}$ as, $\textit{e.g.}$, $\id_{A \otimes B} = \id_{A} \otimes \id_{B}$ and $\sym_{A \otimes B, C} = (\id_{A} \otimes \sym_{B,C});(\sym_{A,C} \otimes \id_{B})$.
+Identities and symmetries of complex types, \textit{e.g.}, $\id_{A \otimes B}$ and $\sym_{A \otimes B, C}$, are given by the composition and tensoring of $\id_{A}$ and $\sym_{A,B}$ as, $\textit{e.g.}$, $\id_{A \otimes B} = \id_{A} \otimes \id_{B}$ and $\sym_{A \otimes B, C} = (\id_{A} \otimes \sym_{B,C});(\sym_{A,C} \otimes \id_{B})$.
 
@@
 \begin{definition}
-
 A symmetric monoidal theory $\catname{SMT}(\Sigma, \mathcal{E})$ is given by a set of $\Sigma$-terms quotiented by a set of equations $\mathcal{E}$ consisting of pairs of appropriately typed $\Sigma$ terms $l = r$.
@@
 \begin{definition}
-A categorical presentation of an $\catname{SMT}(\Sigma, \mathcal{E})$ is a free symmetric monoidal category $\catname{S}(\Sigma, \mathcal{E})$ given by a set of $\Sigma$ terms quotiented by the laws of a symmetric monoidal category and equations of $\mathcal{E}$.
+A categorical presentation of  $\catname{SMT}(\Sigma, \mathcal{E})$ is a free symmetric monoidal category $\catname{S}(\Sigma, \mathcal{E})$ given by a set of $\Sigma$ terms quotiented by the laws of a symmetric monoidal category and equations of $\mathcal{E}$.
 \end{definition}
@@
 \label{def:closed}
-A monoidal (right) closed category is a monoidal category $\catname{C}$ satisfying for
+A monoidal (right) closed category is a monoidal category $\catname{C}$ satisfying that for
 every pair of objects $B,C$ there is an object $B \multimap C$ and a morphism $\textbf{ev}_{B,C} : (B \multimap C) \otimes B \to
 C$, and for every triple of objects $A,B,C$ there is an operation $\Lambda_{A,B,C} : \catname{C}(A \otimes B,C) \to
-\catname{C}(A,B \multimap{} C)$ satisfying the following equations for all $f : A \otimes B \to C$ and $g : Z \to A$:
+\catname{C}(A,B \multimap{} C)$ such that for all $f : A \otimes B \to C$ and $g : Z \to A$ the following equations hold:
 \begin{itemize}
@@
 
-This definition is equivalent to the more usual one that defines a (right) closed category through a pair of adjoint functors $- \otimes B$ and $B \multimap{} -$~\cite{ghica-zanassi2023string}.
-If the tensor is symmetric, then the category is simply \textit{closed}.
+This definition is equivalent to defining a (right) closed category through a pair of adjoint functors $- \otimes B$ and $B \multimap{} -$.
+If the tensor is symmetric, then the category is said to be simply \textit{closed}.
 
 \begin{definition}
-
 \textit{Closed} $\Sigma$-terms are defined inductively by first defining a family of objects $obj_{\Sigma_{O}}$.
@@
 
-To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with additional structure of a semilattice induced by an idempotent, commutative and associative operator $+$.
+To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator $+$.
 This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix section~\ref{sec:appendix:slat}).
 
-\begin{definition}[Semilattice-enriched category]
-    A semilattice-enriched category $\mathbb{C}$ is defined by the following data
+\begin{definition}[Semilattice-enriched category~\cite{Borceux_1994}]
+    A semilattice-enriched category $\mathbb{C}$ is defined by the following data:
     \begin{itemize}
@@
     \end{itemize}
-    such that particular coherence diagrams commute~\cite{Borceux_1994}.
+    such that particular coherence diagrams commute.
 \end{definition}
@@
 
-\begin{definition}[Semilattice-enriched SMC]\label{def:enriched-prop}
+\begin{definition}[Semilattice-enriched SMC~\cite{enriched_monoidal}]\label{def:enriched-prop}
     A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by a semilattice-enriched category $\mathbb{C}$ as above that additionally has
@@
     \end{itemize}
-    such that particular coherence diagrams commute~\cite{enriched_monoidal}.
+    such that particular coherence diagrams commute.
     \end{definition}
-    The latter morphism being in $\catname{SLat}$ implies the following equations that $\otimes$ satisfy
+    The latter morphism being in $\catname{SLat}$ implies the following equations for $\otimes$:
     \[
@@
 We can similarly derive $\catname{SLat}$-enriched versions of  $\Lambda_{A,B,C}$ and $\textbf{ev}_{A,B}$ morphisms using the lifted adjunction as $\textbf{ev}_{A,B} = \varepsilon_{A}$ and $\Lambda_{A,B,C}(f) = \eta; (B \multimap{} f)$.
-The following equation can then be derived which will be useful in encoding equivalences inside abstractions.
+The following distributivity law can then be derived, forming the basis of encoding equivalences inside abstractions.
 \begin{align*}

### 2025-01-22 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
section 3 edits

@@
 
-We can extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness.
+We extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness.
 
@@
         For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{red}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{blue}{<^{\mu}} y_{i+1}\; \}$.
-        We will call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
+        We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
         Each of these relations should satisfy the following
@@
 $\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations can consist of exclusively hierarchical edges, then we have $\textcolor{gray}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
-Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $<$ and $\consistency$.
+Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $<$ and~$\consistency$.
 \end{definition}
 
-Intuitively, edges from $\textcolor{red}{E}$ will encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ will encode lambda-abstraction boxes.
+Intuitively, edges from $\textcolor{red}{E}$ encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ encode lambda-abstraction boxes.
 Also note the typing on the $l_{V}$ function.
-We work with freely generated closed semilattice-enriched categoeries and to avoid unnecessary complications we limit the labelling of vertices to such a type.
+We work with freely generated closed semilattice-enriched categories and to avoid unnecessary complications we limit the labelling of vertices to such a type.
 In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
@@
 A concrete description of the pushout of two morphisms in this category is given in the Appendix section~\ref{sec:appendix:pushout}.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied, most importantly a pushout for the composition of two cospans with discrete feet exists as described below.
+Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied. 
+Most importantly a pushout for the composition of two cospans with discrete feet exists, as described below.
 
-As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, we need a notion of an interface for our e-hypergraphs for them to model string diagrams as the latter have dedicated inputs and outputs.
-Because string diagrams for closed semilattice-enriched symmetric monoidal categories are nested: they contain sub string diagrams inside e-boxes and lambda abstraction boxes we need to account for the interfaces in these nested contexts as well.
-It is worth noting that these nested inputs and outputs of a string diagram do not participate in composition of string diagrams and hence they need to be put aside from the ones that participate in the composition.
-We thus arrive to a notion of an \textit{extended} cospans of discrete e-hypergraphs and will use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n \setminus m$ vertices, in particular
+As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, e-hypergraphs requires interfaces in order to model string diagrams with dedicated inputs and outputs.
+Because string diagrams for closed semilattice-enriched symmetric monoidal categories are nested, containing substring diagrams inside e-boxes and lambda abstraction boxes, we need to account for the interfaces in these nested contexts as well.
+Nested inputs and outputs of a string diagram do not participate in composition of string diagrams, hence they need to be distinguished from the ones that participate in the composition.
+We thus propose \textit{extended} cospans of discrete e-hypergraphs and use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n \setminus m$ vertices, in particular
 with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
@@
   where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete ordered e-hypergraphs,  $f_{ext},g_{ext}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-  We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
+  We sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
 
 \end{definition}
-We will call $n$  \textit{external input interfaces} and $n'$ \textit{internal input interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.  
+We call $n$  \textit{external input interfaces}, $n'$ \textit{internal input interfaces},
+and $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.  
 We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.  
@@
 Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-Graphically we will depict such extended cospans as in Figure~\ref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
+Graphically, we depict such extended cospans as in Figure~\ref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
 All non-black vertices form strictly internal interfaces.
@@
 Consider a relation 
-\[
-R = \{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\}
-\]
+\begin{multline*}
+R = \{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\}
+\end{multline*}
 for $x, y \in n'$.
@@
 where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
+\begin{align*}
     h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\quad
+\\
     h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
+\end{align*}
 % \[
@@
 
-For the same reasons as in the case of $\catname{S}(\Sigma)$ we will restrict cospans to MDA cospans.
+For the same reasons as in the case of $\catname{S}(\Sigma)$ we restrict cospans to MDA cospans.
 We also need to enforce typing constraints on hierarchical edges so that they represent well-typed closed $\Sigma^{+}$-terms.
@@
 
-  We will denote the category of well typed MDA cospans as $\WellTypedMdaEcospans$.
-  It is a proper category as identities, symmetries, and tensor unit are well-typed and composition and tensor product of two well-typed MDA cospans is again a well typed MDA cospan.
+  We denote the category of well typed MDA cospans as $\WellTypedMdaEcospans$.
   \end{definition}
+This is a proper category as identities, symmetries, and tensor unit are well-typed and composition and tensor product of two well-typed MDA cospans is again a well typed MDA cospan.
 
@@
 This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
+Instead,  we treat them via rewriting,  as we do with equations of the signaturem because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
 % , cf. string diagrams for Cartesian categories.  
-An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
+An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this,  so we also handle these equations via rewriting.
 
@@
 
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
+Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments, defined in this section.
 
@@
 \mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
+\] 
+which fits into the DPO formalism. We make this identification throughout,  without further mention. 
 %We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
@@
 This limitation has no practical impact in applications of interest. 
-However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of Appendix.
+However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of the Appendix.
 %This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
@@
 
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
+Before introducing \textit{extended} DPOI rewriting,  the definition of boundary complement must guarantee that rewriting yields a monogamous directed acyclic e-hypergraph. 
+First,  we introduce the  ancillary notion of a \textit{down-closed} graph:
 \begin{definition}[Down-closed subgraph]
@@
 
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
+Intuitively speaking, the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
 In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding MDA cospans above.
@@
 
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
+The enhanced definition of extended DPOI ensures that the combinatorial representation of semilattice-enriched SMCs is sound and complete, in an appropriate sense.

### 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
fixed grey color

@@
   \item $V$ is a set of vertices.
-  \item $E = \textcolor{gray}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
+  \item $E = \colorbox{yellow}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
   \item $s,t : E \to V^{*}$ are source and target functions.
@@
 \end{itemize} 
-$\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations can consist of exclusively hierarchical edges, then we have $\textcolor{gray}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
+$\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations can consist of exclusively hierarchical edges, then we have $\colorbox{yellow}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
 Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $<$ and~$\consistency$.

### 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
wip

@@
 
-\textcolor{red}{Add intro to hypergraphs}
+\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
+
+In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
+
+First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
+We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $R$ to a relation on ordered sequences. 
+We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
+We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
+We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
+We will also refer to $A,B$ as \textit{feet} of the cospan, and to $C$ as \textit{carrier}.
+
+\subsection{Hypergraphs}
+
+In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of SMTs including a Frobenius algebra. 
+Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
+When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
+However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
+This gives rise to the definition of a hypergraph with interfaces.  
+Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
+
+\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
+A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
+The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
+We call a hypergraph \textit{discrete} if its set of edges is empty.
+A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
+\begin{enumerate}
+    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
+    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
+    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
+\end{enumerate}
+We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
+\end{definition}
+Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
+The initial object is given by the empty hypergraph.
+
+\begin{definition}[Symmetric monoidal category of cospans]
+Let $\mathbb{C}$ be a category with all finite colimits.
+A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
+This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
+%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
+%\[
+%\begin{tikzcd}
+%                                               & A \arrow[dd, "\alpha"] &                                                \\
+%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
+%                                               & B                      &                                               
+%\end{tikzcd}
+%\]
+%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
+Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
+Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
+\ifdefined \ONECOLUMN
+\[
+    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+\]
+\else
+\begin{multline*}
+    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+\end{multline*}
+\fi
+and its unit by the initial object of $\mathbb{C}$.
+Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
+\end{definition}
+
+Then, morphisms of $\catname{S}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
+\begin{definition}[Hypergraphs with interfaces]
+\label{def:cspd}
+Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete \textit{ordered} hypergraphs as objects.
+\end{definition}
+
+In particular, this means that any morphism in $\HypI{\Sigma}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
+where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
+The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
+The ordering of the interface hypergraphs is the way to differentiate between the interpretations of morphisms $f \otimes g$ and $g \otimes f$.
+
+\begin{definition}[Degree,  path,  acyclic] 
+The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
+Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
+A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
+A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
+\end{definition}
+
+In order to make the interpretation complete with respect to $\catname{S}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
+\begin{definition}[Category of MDA Hypergraphs with Interfaces~\cite{bonchi_string_2022-2}]
+\label{def:monogamy_hyp}
+We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
+\begin{enumerate}
+    \item $\mathcal{G}$ is a directed acyclic hypergraph;
+    \item $f$ and $g$ are monomorphisms;
+    \item the in-degree and out-degree of every vertex is at most $1$;
+    \item vertices with in-degree $0$ are precisely the image of $f$; and
+    \item vertices with out-degree $0$ are precisely the image of $g$.
+\end{enumerate}
+We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA cospans.
+\end{definition}
+Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
+\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\catname{S}(\Sigma)$. 
+\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
+    $\catname{S}(\Sigma) \cong \MdaCospans$
+\end{theorem}
+
+\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
+The correspondence of the previous section means that the symmetric monoidal equations of $\catname{S}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
+The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
+DPO rewriting is a standard technique for formalising graph rewriting categorically. 
+In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
+
+The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
+In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
+A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
+
+% \[
+% \begin{tikzcd}
+% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
+% \end{tikzcd}
+% \]
+% CHRIS: WHY IS J DISPLAYING AND NOT R? 
+% \[
+%     \scalebox{0.8}{
+%         \tikzfig{combinatorial_semantics/DPO_square}
+%     }
+% \]
+
+The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal \mathcal{L}^{\bot} \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal{L}^{\bot}$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
+The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
+In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
+
+To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
+In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
+By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
+We now make this identification throughout without further mention.  
+Thus we reach the following variant of DPO rewriting.
+\begin{definition}[DPOI rewriting]\label{def:dpoi}
+Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
+% \[
+% \begin{tikzcd}
+% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
+% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
+%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
+% \end{tikzcd}
+% \]
+\begin{figure}[t!]
+    \begin{subfigure}[T]{0.4\linewidth}
+    \[
+                % \tikzfig{../figures/combinatorial_semantics/DPO_square}
+                \begin{tikzcd}
+                    {\mathcal{L}} & {\mathcal{K}} & {\mathcal{R}} \\
+                    {\mathcal{G}} & {\mathcal{L}^{\bot}} & {\mathcal{H}}
+                    \arrow["f"', from=1-1, to=2-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-2, to=1-3]
+                    \arrow[from=1-2, to=2-2]
+                    \arrow[from=1-3, to=2-3]
+                    \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=90}, draw=none, from=2-1, to=1-2]
+                    \arrow[from=2-2, to=2-1]
+                    \arrow[from=2-2, to=2-3]
+                    \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=180}, draw=none, from=2-3, to=1-2]
+                \end{tikzcd}
+    \]
+    % \subcaption{DPO square}
+    % \label{fig:dpo}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[T]{0.4\linewidth}
+        \[
+    % \scalebox{0.8}{
+        % \tikzfig{../figures/combinatorial_semantics/DPOI_square_Hyp}
+    % }
+    \begin{tikzcd}
+        {\mathcal{L}} & {i+j} & {\mathcal{R}} \\
+        {\mathcal{G}} & {\mathcal{L}^{\bot}} & {\mathcal{H}} \\
+        & {n+m}
+        \arrow["f"', from=1-1, to=2-1]
+        \arrow[from=1-2, to=1-1]
+        \arrow[from=1-2, to=1-3]
+        \arrow[from=1-2, to=2-2]
+        \arrow[from=1-3, to=2-3]
+        \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=90}, draw=none, from=2-1, to=1-2]
+        \arrow[from=2-2, to=2-1]
+        \arrow[from=2-2, to=2-3]
+        \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=180}, draw=none, from=2-3, to=1-2]
+        \arrow[from=3-2, to=2-1]
+        \arrow[from=3-2, to=2-2]
+        \arrow[from=3-2, to=2-3]
+    \end{tikzcd}
+\]
+    % \subcaption{DPOI square}
+    % \label{fig:dpoi}
+    \end{subfigure}
+    \captionsetup{belowskip=-3ex}
+    \caption{DPO and DPOI squares}
+    \label{fig:dpo_dpoi}
+\end{figure}
+
+\end{definition}
+We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
+
+\begin{definition}[Boundary complement]
+\label{def:boundary_original}
+For MDA cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
+and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
+% \end{figure}
+% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
+% \[\begin{tikzcd}
+% 	\mathcal L && {i+j} \\
+% 	\\
+% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
+% 	\\
+% 	&& {n+m}
+% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
+% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
+% 	\arrow["f"', from=1-1, to=3-1]
+% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
+% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
+% 	\arrow[from=3-3, to=3-1]
+% \end{tikzcd}\]
+\[
+    \scalebox{0.8}{
+        \tikzfig{../figures/combinatorial_semantics/DPOI_boundary_complement}
+    }
+\]
+is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
+\[
+    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
+\]
+is an MDA cospan. 
+\end{definition}
+
+Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
+Yet, this restriction is not enough to make DPOI rewriting complete for $\catname{S}(\Sigma,\mathcal{E})$: there might be cospans with carriers $\mathcal{G}$ and $\mathcal{H}$ such that one can be rewritten into another but their corresponding morphisms are not equal modulo $\mathcal{E}$~\cite{bonchi_string_2022-2}.
+To make the DPOI rewriting complete, we need to add a restriction on the image of the match.
+
+\begin{definition}[Convex match]
+We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
+We call a match $f : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
+
+\end{definition}
+
+\begin{definition}[Convex DPOI rewriting]
+\label{def:convex_dpo}
+Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
+Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $\mathcal{L}^{\bot}$ and cospan arrows $i+j \xrightarrow{} \mathcal{L}^{\bot} \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
+and the following conditions hold
+\begin{itemize}
+    \item $f : L \to \mathcal G$ is a convex match;
+    \item $i + j \to \mathcal{L}^{\bot} \to \mathcal G$ is a boundary complement in the leftmost pushout.
+\end{itemize}
+\end{definition}
+
+We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\catname{S}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as (cospans of) hypergraphs.
+In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
 

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/abstract.tex
wip

--- 2024-08-27 LICS/sections/abstract.tex
@@
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for satisfiability modulo theory solvers.
+% The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for satisfiability modulo theory solvers.
+The technique of equality saturation, which equips graphs with an equivalence relation, has proven effective for program optimisation.
 We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a category of semilattices.
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
+This approach generalises to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
 Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
 They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
+The advantage of this new principled approach to e-graphs is that for the first time it is possible to support double-pushout (DPO) rewriting for these structures which constitutes the main contribution of this paper.

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/appendix.tex
wip

--- 2025-01-21 LICS/sections/appendix.tex
@@
-\section{$\catname{SLat}$}
+\subsection{$\catname{SLat}$}
 \label{sec:appendix:slat}
 
-In this section we will define the category of semilattices that we use as a base for enrichment throughout the paper.
+In this section we define the category of semilattices that we use as a base for enrichment throughout the paper.
 
@@
       The adjunction is then given by the following natural isomorphisms
-      \[
-      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \cong \catname{Set}(A, U(B))
-      \]
+      \begin{align*}
+      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) &\cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B))\\
+                                                         &\cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B))\\
+                                                         &\cong \catname{Set}(A, U(B))
+      \end{align*}
       The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
@@
 
-\section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
+    \begin{definition}
+        Given two $\catname{SLat}$-functors $F,G : \mathbb{C} \to \mathbb{D}$ a $\catname{SLat}$-natural transformation is family of morphisms indexed by objects in $\mathbb{C}$, $\alpha_{A} : I_{\catname{SLat}} \to \mathbb{D}(FA,GA)$ in $\catname{SLat}$ such that the following diagram commutes
+        \[
+        \adjustbox{width=\linewidth}{
+        \begin{tikzcd}
+          & {\mathbb{C}(A,A')\otimes I_{\mathcal{V}}} && {\mathbb{D}(FA,FA') \otimes \mathbb{D}(FA',GA')} \\
+          {\mathbb{C}(A,A')} &&&& {\mathbb{D}(FA,GA')} \\
+          & {I_{\mathcal{V}} \otimes \mathbb{C}(A,A')} && {\mathbb{D}(FA,GA) \otimes \mathbb{D}(GA,GA')}
+          \arrow["{G_{A,A'}\otimes \alpha_{A'}}"', from=1-2, to=1-4]
+          \arrow["{c_{FA,FA',GA'}}"', from=1-4, to=2-5]
+          \arrow["{r^{-1}}", from=2-1, to=1-2]
+          \arrow["{l^{-1}}", from=2-1, to=3-2]
+          \arrow["{\alpha_{A} \otimes F_{A,A'}}", from=3-2, to=3-4]
+          \arrow["{c_{FA,GA,GA'}}", from=3-4, to=2-5]
+      \end{tikzcd}}
+      \]
+      \end{definition}
+
+      Intuitively, in the case of free enrichment $\alpha_{A}$ picks a usual natural transformation from $\mathbb{D}(FA,GA)$.
+
+      \begin{example}
+        Consider a copy natural transformation $\triangle_{A} : A \to A \times A$ of $\catname{S}(\Sigma_{C},\mathcal{E}_{C})$.
+        The later Cartesian PROP is lifted to a Cartesian $\catname{SLat}$-PROP where the naturality condition can be traced component-wise where ${*}$ is a one element semilattice:
+
+      \[\adjustbox{width=\linewidth}{
+      \begin{tikzcd}  
+          & {f + g \otimes *} && {\triangle_{A} \otimes ((f+g) \times (f+g))} \\
+          f + g &&&& {\triangle_{A'};((f + g) \times (f+g)) = (f + g);\triangle_{A'};} \\
+          & {* \otimes f + g} && {f + g \otimes \triangle_{A'}}
+          \arrow[from=1-2, to=1-4]
+          \arrow[from=1-4, to=2-5]
+          \arrow[from=2-1, to=1-2]
+          \arrow[from=2-1, to=3-2]
+          \arrow[from=3-2, to=3-4]
+          \arrow[from=3-4, to=2-5]
+      \end{tikzcd}}
+      \]
+
+      which is the usual naturality condition.
+
+      \end{example}
+
+\subsection{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
 \label{sec:appendix:pushout}
 
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we will explicitly construct such a pushout.
-We will also show the uniqueness of a boundary pushout complement.
+In this section we define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we explicitly construct such a pushout.
+We also show the uniqueness of a boundary pushout complement.
 These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
@@
     \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = (&\\
+            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})},\\
+            &\consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},\\
+            &<_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+        &)    
     \end{align*}
@@
               Suppose $e = \iota_{E_{X}}(e')$ and then
-              \[
-                u_{V}^{*}([s_{X+Y}(e)]^{*}) = u_{V}^{*}([s_{X+Y}(\iota_{1,E}(e'))]^{*}) = u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*})
-              \]
+              \begin{align*}
+                u_{V}^{*}([s_{X+Y}(e)]^{*}) &= u_{V}^{*}([s_{X+Y}(\iota_{1,E}(e'))]^{*})\\
+                                            &= u_{V}^{*}([(\iota_{1,V}^{*}(s_{X}(e')))]^{*})
+              \end{align*}
               Since $j_1$ and $j_2$ are homomorphisms, it must be the case that $s_{Q}(j_{1,E}(e')) = j_{1,V}^{*}(s_{X}(e'))$.
@@
     \item The last condition to check is if
-    \[
-    [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-    \]
+    \begin{align*}
+    &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+    &\;\subseteq\\
+    &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+    \end{align*}
     Recall that
@@
     and first we will check if
-    \[
-      [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-    \]
+    \begin{align*}
+      &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+      &\;\subseteq\\
+      &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+    \end{align*}
     Consider cases
@@
         By the argument above it follows that
-        \[
-            [u_{V};\iota_{Q_{V}},u_{E};\iota_{Q_{E}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))) \subseteq \consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))
-        \]
+        \begin{align*}
+            &[u_{V};\iota_{Q_{V}},u_{E};\iota_{Q_{E}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])))\\
+            &\;\subseteq\\
+            &\consistency_{Q}(\iota_{V_{Q}}(u_{V}([u])))
+        \end{align*}
         \end{itemize}
@@
        So far we have shown that
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-       \]
+       \begin{align*}
+        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+        &\;\subseteq\\
+        & \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+       \end{align*}
        and
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))) \subseteq \consistency_{Q}(u_{E};\iota_{E_{Q}}([e]))
-       \] follows by the same argument.
+       \begin{align*}
+        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))\\
+        &\;\subseteq\\
+        &\consistency_{Q}(u_{E};\iota_{E_{Q}}([e]))
+       \end{align*} follows by the same argument.
        Then, to check if
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-       \]
+       \begin{align*}
+        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+        &\; \subseteq\\
+        &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+       \end{align*}
        we need to verify if
@@
        Therefore, we have
-       \[
-        [u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))) \subseteq \consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
-       \]
+       \begin{align*}
+        &[u_{V};\iota_{V_{Q}}, u_{E};\iota_{E_{Q}}]^{*}(\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))\\
+        &\;\subseteq\\
+        &\consistency_{Q}(u_{V};\iota_{V_{Q}}([v]))
+       \end{align*}
        and similarly for edges. Which concludes that $(u_{V},u_{E})$ is a homomorphism.
@@
                                   \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
+                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
+                                    &= [e_2]
                                   \end{align*}
@@
                         \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
+                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
+                                                                                                    &= [e_2]
                         \end{align*}
@@
                 As there is a path from $v$ to $v'$, there is a path from $e_1$ to $v'$ and by definition, since $v'' \sim v'$ and $[v'') \not = \varnothing$,
-                \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v''])) = [e_2]
-                \]
+                \begin{align*}
+                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))\\
+                 &\;= [e_2]
+                \end{align*}
             \end{itemize}    
@@
                         \text{by unfolding the definition}\\
-                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
+                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &=\\
+                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))&
                         \end{align*} 
@@
                     \begin{align*}
-                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')))\\
+                        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}&(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) =\\
+                                                                                                                &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')))\\
                                                                                                                 &= [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(f_{V};\iota_{1,V}(z_1''))))\\
@@
                             \begin{align*}
-                              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')))\\
-                                                                                                                     &= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
-                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))\\
-                                                                                                                     &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1])
+                              \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}&(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) =\\
+                              &\;= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1')))\\
+                                                                                                                     &\;= [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e_1)))\\
+                                                                                                                     &\;= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))\\
+                                                                                                                     &\;= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1])
                             \end{align*}
@@
                             By definition,
-                            \[
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
-                            \]
+                            \begin{align*}
+                                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) =\\
+                                &\;[[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
+                            \end{align*}
                             and since there is a path from $e_1$ to $v_1'$ and $v_1' \sim v_1$
-                            \[
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1]) = [[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
-                            \]
+                            \begin{align*}
+                                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e_1]) =\\
+                                &\;[[]_{V}^{\hashtag},[]_{E}^{\hashtag}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)))
+                            \end{align*}
                             The cases when $v_1 = g_{V};\iota_{2,V}(z_1)$ is symmetric.
@@
                               \begin{align*}
-                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\
+                                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}&(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) =\\
+                                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1'']))\\ 
                                                                                                                         &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))
@@
     Because $a + y + w$ also yields a pushout it must be the case that $a + y + w \cong a + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    \textcolor{red}{The same reasoning applies to show that $E_{\mathcal{L}^{\bot}}$} is unique and hence the sets of edges and vertices are uniquely determined in the pushout complement.
+    The same reasoning applies to show that $E_{\mathcal{L}^{\bot}}$ is unique and hence the sets of edges and vertices are uniquely determined in the pushout complement.
     One can observe that $g_{V} = h_{V,ext};h_{V,int} + id_{n}$ and that $h_{V,ext}$ is mono and $h_{V,int} = [h_1,h_2]$ where $h_1,h_2$ are mono.
@@
     \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ \textcolor{red}{only identifies edges with no successors}.
+    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$ since $g_{E}$ only identifies edges with no successors.
     Similarly for edges.
@@
 
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
+\subsection{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
+\label{sec:appendix:detailed_proofs}
+In this section we will elaborate on soundness and completeness of our translation $\llbracket - \rrbracket : \catname{S}^{+}(\Sigma, \mathcal{E}) \to \WellTypedMdaEcospans / \mathcal{S}, \mathcal{E}$.
+First we recite the interpretation function $[]$ on $\Sigma$-terms as defined in~\cite{bonchi_string_2022-2} that induces a functor $\catname{S}(\Sigma) \to \MdaCospans$ and eventually a functor $\catname{S}^{+}(\Sigma) \to \MdaCospans^{+}$.
+The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Figure~\ref{fig:base_cases}.
+Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.
+
+\begin{remark}\label{remark:embedding_functor}
+    There is a faithful identity-on-objects functor $E: \HypI{\Sigma} \to \Ecospans$ which maps a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ to an extended cospan $n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m$ such that $f_{ext};f_{int} = f$ and $g_{ext};g_{int} = g$.
+\end{remark}
+
+\begin{lemma}
+\label{lemma:normal_form}
+For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{normal form} such that 
+\[
+	f = f_1 + \ldots + f_n
+\]
+	such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
+\end{lemma}
+
+\begin{figure*}
+\[
+\adjustbox{width=\textwidth}{
+\tikzfig{../figures/appendix/interpretation}
+}
+\]
+\caption{Base cases for $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$}
+\label{fig:base_cases}
+\end{figure*}
+
+\begin{lemma}[Lemma~\ref{lemma:cospans_plus_equiv} restatement]
+
+    We have the following equivalence of $\catname{SLat}$-categories
+    \[
+    \MdaCospans^{+} \cong \MdaEcospans/\mathcal{S}~.
+    \]
+    \end{lemma}
+    \begin{proof}
+    We will denote the semilattice-enriched functor giving this equivalence as $M$.
+    Observe that in $\MdaCospans^{+}$ we can consider only canonical morphisms of the form $f_{1} + \ldots + f_{n}$ where each of $f_{i}$ is an MDA cospan (in the sense of Definition~\ref{def:monogamy_hyp}) and for $i \not = j$, $f_{i} \not = f_{j}$.
+    Then, $M$ is defined as follows.
+    \begin{itemize}
+    \item It maps each discrete hypergraph in $\MdaCospans^{+}$ to a corresponding discrete e-hypergraph in $\MdaEcospans/\mathcal{S}$.
+    \item Every morphism $f_{1} + \ldots + f_{n}$ is mapped to $\text{E}(f_{1}) + \ldots + \text{E}(f_{n})$ where $\text{E}$ is the faithful functor induced by Remark~\ref{remark:embedding_functor}.
+          This makes $M$ into $\catname{SLat}$-functor by construction.
+    \end{itemize}
+    To show the functor $M$ is faithful, suppose $f = f_{1} + \ldots + f_{n} \not = g_{1} + \ldots + g_{m} = g$.
+    Then it must be that $Mf \not = Mg$.
+    If $n \not = m$ then, immediately $Mf \not = Mg$ as $Mf$ and $Mg$ contain different number of non-isomorphic components.
+    If $n = m$ then assume for a contradiction that $Mf = Mg$. 
+    There exists a permutation $\sigma$, such that for every index $i$, $Mf_{i} = Mg_{\sigma(i)}$.
+    Thus, by $\text{E}$ we have that $f_{i} = g_{\sigma(i)}$, implying $f = g$ which is a contradiction.
+    Therefore, $Mf \not = Mg$ as required.
+
+    To show that the functor is full, consider a morphism $f = f_{1} + \ldots + f_{n}$ in $\WellTypedMdaEcospans / \mathcal{S}$ (by Lemma~\ref{lemma:normal_form}). 
+    As each of $f_{i}$ does not contain any hierarchical edges, they are in the image of $\text{E}$, which for each $f_{i}$ gives us $g_{i}$, such that $f_{i} = Mg_{i}$.
+    This gives us a morphism $g = g_{1} + \ldots + {g_{n}}$ in $\MdaCospans^{+}$ such that $Mg = f$.
+    This concludes that the functor is full.
+    As it is also surjective on objects, it makes it into an equivalence.
+\end{proof}
 
@@
 
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
+\begin{theorem}[Theorem~\ref{thm:full-completeness} restatement]
     \label{proof:appendix:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
+    Let $\mathcal{E}$ be a set of equations $l = r$ of $\Sigma$-terms, then we have the following equivalence of $\catname{SLat}$-categories
+    \[
+        \catname{S}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
+    \]
+    where $\mathcal{E}$ on the right is overloaded as a set of EDPOI rewrite rules containing $\langle \llbracket l \rrbracket, \llbracket r \rrbracket  \rangle$ and $\langle \llbracket r \rrbracket, \llbracket l \rrbracket  \rangle$ for every equation $l = r$.
+\end{theorem}
 \begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $f = g$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket f\rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket f \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans$.
+    The statement boils down to showing that the $\catname{SLat}$ functor defining the equivalence is well-defined, full and faithful as it is obviously surjective.
+    Proof of fullness follows the same argument as in the case of $\llbracket - \rrbracket: \catname{S}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$, so we proceed with showing well-definedness and faithfulness.
+    The former is equivalent to showing that if $f = g$ in $\catname{S}^{+}(\Sigma)$ modulo $\mathcal{E}$ then $\llbracket f\rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S,E}~.$ or, equivalently, $\llbracket f \rrbracket \Rrightarrow_{\mathcal{E}}^{*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
     $f = g$ implies that either
     \begin{itemize}
-        \item $f = \mathcal{C}[l] = g$ for some morphism $l \in \textbf{PROP}^{+}(\Sigma)$.
-        \item $f = \mathcal{C}[l]$ and $g = \mathcal{C}[r]$ where $l = r$ or $r = l$ is from $\mathcal{S}$.
-        \item There is a sequence $w = (x_1, \ldots, x_n)$ such that $x_1 = f$ and $x_n = g$ and $x_i = x_{i+1}$ or $x_{i+1} = x_{i}$ in the sense above for $i < n$.
+        \item $f = \mathcal{C}[l] = g$ for some morphism $l \in \catname{S}^{+}(\Sigma)$;
+        \item $f = \mathcal{C}[l]$ and $g = \mathcal{C}[r]$ for an equation $l = r$ or $r = l$ from $\mathcal{E}$;
+        \item there is a sequence $w = (x_1, \ldots, x_n)$ such that $x_1 = f$ and $x_n = g$ and $x_i = x_{i+1}$ or $x_{i+1} = x_{i}$ in the sense above for $i < n$.
     \end{itemize}
@@
     \begin{itemize}
-        \item If $\mathcal{C}$ is empty, then either $f = g$ syntactically and $\llbracket f \rrbracket = \llbracket g \rrbracket$ or $f = l = r = g$ for some $l = r$ or $r = l$ in $\mathcal{S}$.
-              The latter trivially yields the E-DPO diagram below
+        \item If $\mathcal{C}$ is empty, then either $f = g$ syntactically and $\llbracket f \rrbracket = \llbracket g \rrbracket$ or $f = l = r = g$ for some $l = r$ or $r = l$.
+              The latter trivially yields the EDPO diagram below
               \[
@@
               and by inductive hypothesis $\llbracket \mathcal{C}[l] \rrbracket \Rrightarrow{} \llbracket \mathcal{C}[r] \rrbracket$ which results in the top half of the diagram below.
-              \[\begin{tikzcd}
+
+              \[
+              \adjustbox{width=\linewidth}{
+              \begin{tikzcd}
                 {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
@@
                 \arrow[from=7-3, to=6-4]
-            \end{tikzcd}\]
+            \end{tikzcd}}
+            \]
             Clearly the occurrence of $\llbracket \mathcal{C}[l] \rrbracket$ in $\llbracket f \rrbracket$ is convex and down-closed and by following the argument in Theorem 35~\cite{bonchi_string_2022-2}, there exists $n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2$ such that
@@
                 \[
+                \adjustbox{width=\linewidth}{
                 \begin{tikzcd}
@@
                     \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
-                \end{tikzcd}
+                \end{tikzcd}}
                 \]
                 which implies the existence of the following diagram where both squares are pushouts by construction and the matching is convex and down-closed.
-                \textcolor{red}{Consider cases when $\mathcal{L}$ has no-input and no-output sub-hypergraph and when it does not.
+                % \textcolor{red}{Consider cases when $\mathcal{L}$ has no-input and no-output sub-hypergraph and when it does not.
 Basically, repeat the above reasoning here as well, but say that if $n+m$ is empty, then there is no pushout.}
@@
                 \begin{figure*}[hbt!]
-                \[
+                \[\adjustbox{width=\linewidth}{
                     \begin{tikzcd}
@@
                     \arrow[from=5-3, to=4-4]
-                \end{tikzcd}
+                \end{tikzcd}}
                 \]
@@
     \end{itemize}
-    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
+    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ modulo $\mathcal{E}$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$  in $\WellTypedMdaEcospans / \mathcal{S}, \mathcal{E}$.
+
+    Then, faithfulness amounts to showing that if $\llbracket f \rrbracket \Rrightarrow_{\llbracket \mathcal{E} \rrbracket}^{*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ then $f = g$ modulo $\mathcal{E}$ in $\catname{S}^{+}(\Sigma)$.
+% By the theorem~\ref{thm:completeness_simple} if $\llbracket f \rrbracket = \llbracket g \rrbracket$ via rules from $\mathcal{S}$ then $f = g$, so we will show that the same holds when $\llbracket f \rrbracket = \llbracket g \rrbracket$ via rules from $\mathcal{E}$.
+Consider a rewrite rule $\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle = \langle \mathcal{L}, \mathcal{R} \rangle$ such that $\llbracket f \rrbracket \Rrightarrow_{\langle \mathcal{L}, \mathcal{R} \rangle} \llbracket g \rrbracket$ and suppose
+\[
+	\llbracket f \rrbracket = n \xrightarrow{} n' \xrightarrow{} \mathcal{F} \xleftarrow{} m' \xleftarrow{} m ~.
+\]
+This implies the existence of a diagram as in~\ref{def:dpoi-e}.
+We will proceed by induction on the level of the occurrence of $\mathcal{L}$ within $\mathcal{F}$.
+\begin{itemize}
+	\item Suppose the image of $\mathcal{L}$ in $\mathcal{F}$ is top-level.
+		  Then, by lemma~\ref{lemma:decomposition}, there exists a decomposition of $\llbracket f \rrbracket$ in terms of MDA cospans
+		\begin{align*}
+			(k \xrightarrow{id_{k}} k \xrightarrow{id_{k}} &k \xleftarrow{id_{k}} k \xleftarrow{id_{k}} k)\\
+			n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k; \qquad\qquad\qquad &\otimes \qquad\qquad\qquad ; j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m\\
+			(i \xrightarrow{} i' \xrightarrow{} &\mathcal{L} \xleftarrow{} j' \xleftarrow{} j)
+		\end{align*}
+		By fullness of $\llbracket - \rrbracket$ (from Theorem~\ref{thm:completeness_simple}) there exist $c_{1}$ and $c_{2}$ such that 
+		\begin{align*}
+		&\llbracket c_{1} \rrbracket = n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k\\
+        &\llbracket c_{2} \rrbracket = j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m
+        \end{align*}
+		By functoriality, $\llbracket f \rrbracket = \llbracket c_{1};(id \otimes l);c_{2} \rrbracket$ and by faithfulness we have that $f = c_{1};(id \otimes l);c_{2}$ to which we can apply $\langle l, r \rangle$ and get $g' = c_{1};(id \otimes r);c_{2}$.
+		We then have $\llbracket g' \rrbracket = \llbracket c_{1};(id \otimes r);c_{2} \rrbracket =  \llbracket g \rrbracket$ since the boundary complement is unique (Proposition ~\ref{prop:boundary_unique}) and $g' = g$ by faithfulness.
+
+		\item If the image of $\mathcal{L}$ is not top-level, then either $\llbracket f \rrbracket$ is decomposable as
+		\begin{align*}
+			(k \xrightarrow{id_{k}} k \xrightarrow{id_{k}} &k \xleftarrow{id_{k}} k \xleftarrow{id_{k}} k)\\
+			n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k; \qquad\qquad\qquad &\otimes \qquad\qquad\qquad ; j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m\\
+			(i \xrightarrow{} i' \xrightarrow{} &\mathcal{F'} \xleftarrow{} j' \xleftarrow{} j)
+		\end{align*}
+		such that $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $F'$.
+		Since the image of $\mathcal{L}$ lies within $\mathcal{F}'$, there is an EDPO rewrite step that turns
+		\[
+			(i \xrightarrow{} i' \xrightarrow{} \mathcal{F'} \xleftarrow{} j' \xleftarrow{} j)	
+		\]
+		into
+		\[
+			(i \xrightarrow{} i'' \xrightarrow{} \mathcal{G'} \xleftarrow{} j'' \xleftarrow{} j)	
+		\]
+		via $\langle \mathcal{L}, \mathcal{R} \rangle$.
+		By fullness and inductive hypothesis, there exist $f'$ and $g'$ such that 
+		\begin{align*}
+			&\llbracket f' \rrbracket = i \xrightarrow{} i' \xrightarrow{} \mathcal{F'} \xleftarrow{} j' \xleftarrow{} j\\
+            &\llbracket g' \rrbracket = (i \xrightarrow{} i'' \xrightarrow{} \mathcal{G'} \xleftarrow{} j'' \xleftarrow{} j)
+        \end{align*}
+		and that $f' = g'$.
+		By decomposition and fullness we have $\llbracket f \rrbracket = \llbracket c_{1};(id \otimes f');c_{2} \rrbracket$ and $f = c_{1};(id \otimes f');c_{2}$ by faithfulness.
+		By applying $f' = g'$ we get $f = g'' = c_{1};(id \otimes g');c_{2}$.
+		Finally, $\llbracket g'' \rrbracket = \llbracket c_{1};(id \otimes g');c_{2} \rrbracket = \llbracket g \rrbracket$ (by the uniqueness of the boundary complement) and $g'' = g = f$ by faithfulness.
+
+		\item or, $\llbracket f \rrbracket$ is decomposable as
+		\begin{align*}
+			n_1 \xrightarrow{} n_1' \xrightarrow{} &\mathcal{F}_1 \xleftarrow{} m_1' \xleftarrow{} m_1\\
+			&\;+\\
+			&\vdotswithin{+}\\
+			&\;+\\
+			n_p \xrightarrow{} n_p' \xrightarrow{} &\mathcal{F}_p \xleftarrow{} m_p' \xleftarrow{} m_p\\
+			&\;+\\
+			&\vdotswithin{+}\\
+			&\;+\\
+			n_k \xrightarrow{} n_k' \xrightarrow{} &\mathcal{F}_k \xleftarrow{} m_k' \xleftarrow{} m_k
+		\end{align*}
+		such that the image of $\mathcal{L}$ lies within $\mathcal{F}_{p}$.
+		By inductive hypothesis we have that $\llbracket f_{p} \rrbracket = n_p \xrightarrow{} n_{p}' \xrightarrow{} \mathcal{F}_{p} \xleftarrow{} m_{p}' \xleftarrow{} m_{p}$ rewrites into $\llbracket g_{p} \rrbracket = n_p \xrightarrow{} n_{p}'' \xrightarrow{} \mathcal{G}_{p} \xleftarrow{} m_{p}'' \xleftarrow{} m_{p}$ such that $f_{p} = g_{p}$.
+		By fullness and the fact that $\llbracket - \rrbracket$ is enriched, there exist $\llbracket f_1 + \ldots + f_{p} + \ldots + f_{k} \rrbracket = \llbracket f_{1} \rrbracket + \ldots + \llbracket f_{p} \rrbracket + \ldots + \llbracket f_{k} \rrbracket = \llbracket f \rrbracket$.
+		By applying $f_{p} = g_{p}$ we get $f = g''$ and $\llbracket g'' \rrbracket = \llbracket f_{1} + \ldots + g_{p} + \ldots f_{k} \rrbracket = \llbracket g \rrbracket$ and $f = g'' = g$.
+\end{itemize}
     % there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
@@
     and $n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m$ is the co-limit of the following diagram
-    \[
-        n \to n_{1}' \to \mathcal{C}_{1} \xleftarrow{} m_{1}' + k \xleftarrow{} i + k \to i' + k \to \mathcal{L} \xleftarrow{} k + j' \xleftarrow{} k + j \xrightarrow{} n_{2}' + k + j \to \mathcal{C}_{2} \xleftarrow{} m_{2}' \xleftarrow{} m
+    % \[
+    %     n \to n_{1}' \to \mathcal{C}_{1} \xleftarrow{} m_{1}' + k \xleftarrow{} i + k \to i' + k \to \mathcal{L} \xleftarrow{} k + j' \xleftarrow{} k + j \xrightarrow{} n_{2}' + k + j \to \mathcal{C}_{2} \xleftarrow{} m_{2}' \xleftarrow{} m
+    % \]
+    \[
+    \adjustbox{width=\linewidth}{
+    \begin{tikzcd}
+        n \arrow[r] & n'_{1} \arrow[r] & \mathcal{C}_{1} & m'_{1} + k \arrow[l]     & i+k \arrow[l] \arrow[r] & i'+k \arrow[rd] &             \\
+                    &                  &                 &                          &                         &                 & \mathcal{L} \\
+        m \arrow[r] & m'_{2} \arrow[r] & \mathcal{C}_{2} & n'_{2} + k + j \arrow[l] & k+j \arrow[r] \arrow[l] & k+j' \arrow[ru] &            
+    \end{tikzcd}
+    }
     \]
     The two spans identify precisely those nodes from $\mathcal{G}$ that occur in more than one sub-hypergraph, so this amounts to simply taking the union
-    \[
-    n \to n_{1}' \cup n_{2}' \cup i' \to \mathcal{C}_{1} \cup \mathcal{C}_{2} \cup \mathcal{L} \xleftarrow{} m_{1}' \cup m_{2}' \cup j' \xleftarrow{} m  = n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m
-    \]
+    \begin{align*}
+    &n \to n_{1}' \cup n_{2}' \cup i' \to \mathcal{C}_{1} \cup \mathcal{C}_{2} \cup \mathcal{L} \xleftarrow{} m_{1}' \cup m_{2}' \cup j' \xleftarrow{} m  =\\
+    &\;= n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m
+    \end{align*}
     If we let $\mathcal{G}' = \mathcal{L}$ we get the decomposition as in~(\ref{fig:decomposition:1}).
@@
 
-\section{Additional figures for Section~\ref{sec:introduction}:Introduction}
+\subsection{Additional figures for Section~\ref{sec:introduction}:Introduction}
 
@@
 
-\section{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
+\subsection{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
 \label{sec:appendix:iso}
+
+First we will explore the notion of isomorphism of two extended cospans in more detail.
+Our example will be Figure~\ref{fig:appendix:non-isomorphic-cospans}.
+Technically, we do not allow hierarchical edges with no delimiting dashed line, but here, for the purposes of illustration we make an exception to make the example compact.
 \begin{example}
@@
     To make the diagram commute, clearly, it must be the case that $u_0 \mapsto u_0$ and $u_1 \mapsto u_1$.
-    Then, because the isomorphism should preserve the order within the subset of green verices, it must be the case that $u_2 \mapsto u_2$ and $u_3 \mapsto u_3$.
+    Then, because the isomorphism should preserve the order within the subset of green vertices, it must be the case that $u_2 \mapsto u_2$ and $u_3 \mapsto u_3$.
     Similarly for the output interfaces.
@@
 
+Then we will elaborate on the notion of a boundary complement.
+
+\begin{example}
+
+Consider a pushout square in Figure~\ref{fig:not_boundary_complement} (a) where the labels suggest how arrows map vertices both from the interfaces and between the e-hypergraphs.
+While this is a valid pushout square, such pushout complement does not correspond to a rewrite of $\Sigma^{+}$-terms --- as there is no redex for $f \otimes g$ within $(f + g);g$ even modulo SMC equations.
+This pushout fails to be a boundary complement by violating conditions (3) and (4) of Definition~\ref{def:boundary_new}.
+Now consider the square in subfigure (b): it fails to be a boundary complement by violating conditions (4) and (7).
+\end{example}
+
+\begin{example}
+
+Now consider a full DPO square in Figure~\ref{fig:not_boundary_complement} (c).
+First note that the pushout complement is a boundary complement as it satisfies all the conditions in Definition~\ref{def:boundary_new}.
+In particular, observe that in order to make an extended MDA cospan with the boundary complement as a carrier, strict internal interfaces should be empty.
+Indeed, strict internal interfaces of the left-hand side coincide with the ones of the source e-hypergraph.
+Then, for the resulting e-hypergraph to form a valid well-typed MDA extended cospan we need to introduce the strict internal interfaces of the right-hand side of the rewrite rule.
+\end{example}
+
 \begin{figure*}
@@
         \[
-    \scalebox{0.6}{
+    \scalebox{0.55}{
         \tikzfig{../figures/appendix/sym_cospan}
@@
         \[
-            \scalebox{0.6}{
+            \scalebox{0.55}{
                 \tikzfig{../figures/appendix/id_x_id_cospan}
@@
         \[
-            \scalebox{0.6}{
+            \scalebox{0.55}{
                 \tikzfig{../figures/appendix/id_x_id_cospan_2}
@@
 
-\section{DPO rewriting for arbitrary signatures}
+\begin{figure*}
+    \begin{subfigure}{0.48\textwidth}
+\[
+\adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/boundary_complement_non_example}
+}
+\]
+\subcaption{Boundary complement non-example}
+    \end{subfigure}
+\hfill
+    \begin{subfigure}{0.48\textwidth}
+        \[
+        \adjustbox{scale=0.55}{
+            \tikzfig{../figures/appendix/boundary_complement_non_example_2}
+        }
+        \]
+            \subcaption{Boundary complement non-example}
+            \end{subfigure}
+\begin{subfigure}{\textwidth}
+    \[
+    \adjustbox{scale=0.55}{
+        \tikzfig{../figures/appendix/boundary_complement_example}
+    }
+    \]
+        \subcaption{Complete DPO square}
+        \end{subfigure}
+\caption{DPO examples}
+\label{fig:not_boundary_complement}
+\end{figure*}
+
+\subsection{DPO rewriting for arbitrary signatures}
 \label{sec:dpo-fix}
@@
 \[
+\adjustbox{scale=0.6}{
 \tikzfig{../figures/appendix/cartesian_rewrite_string}
+}
 \]
@@
 \[
+\adjustbox{scale=0.6}{
 \tikzfig{../figures/appendix/cartesian_rewrite_hypergraph}
+}
 \]
@@
 Assume we want to perform a rewrite within the e-hypergraph in Figure~\ref{fig:dpo-stuck} (left).
-
-\begin{figure}
+However, we are immediately stuck as there is no matching of the left-hand side of the rule within the graph as there is no pushout complement: the only way we can impose a constraint that an edge shall have a particular predecessor in the pushout is through its adjacent vertices, but the sub-hypergraph for $a;!$ does not have any as demonstrated in the right subfigure.
+
+\begin{figure*}
     \begin{subfigure}{0.4\linewidth}
@@
     \label{fig:dpo-stuck}
-\end{figure}
-
-However, we are immediately stuck as there is no matching of the left-hand side of the rule within the graph as there is no pushout complement: the only way we can impose a constraint that an edge shall have a particular predecessor in the pushout is through its adjacent vertices, but the sub-hypergraph for $a;!$ does not have any as demonstrated in the right subfigure.
-
+\end{figure*}
 Another problematic rewrite arises when we want to utilise the structural rule of the free enrichment
-
-\[
-\adjustbox{scale=0.75}{
+\[
+\adjustbox{width=\linewidth}{
 \tikzfig{../figures/appendix/structural_rewrite_problem}
@@
 \]
-
 if we consider the following DPO square
-
-\[
-\adjustbox{scale=0.55}{
+\[
+\adjustbox{width=\linewidth}{
     \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck}
@@
 We will start completing all the above DPO squares by first considering the structural rules.
-We will include special unlabelled edges with no inputs and outputs to the interfaces to be able to impose the hierarchical relation on subgraphs with no inputs and outputs.
+We will include special unlabelled edges with no inputs and outputs to the interfaces to be able to impose the hierarchical relation on subgraphs with no inputs and outputs and will not impose any restrictions on labelling preservation when mapping these special edges.
 That is, the previously mentioned structural rule will become
@@
 \[
-\adjustbox{scale=0.75}{
+\adjustbox{width=\linewidth}{
     \tikzfig{../figures/appendix/structural_rewrite_problem_fix}
@@
 \[
-\adjustbox{scale=0.55}{
+\adjustbox{width=\linewidth}{
     \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
@@
 
-and similar rules to introduce and eliminate in nested contexts
+and similar rules to introduce and eliminate in nested contexts (Figure~\ref{fig:nested_elimination}).
 
@@
     \end{subfigure}
+    \caption{Special edge nested elimination}
+    \label{fig:nested_elimination}
 \end{figure}
 
-and, finally, the rules to introduce such edges as new consistent components.
+and, finally, the rules to introduce such edges as new consistent components (Figure~\ref{fig:consistent_introduction}).
 Intuitively such edge plays the role of a tensor unit identity that can be inserted in any context.
@@
 \begin{figure}[h!]
-    \begin{subfigure}{0.45\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
+    \begin{subfigure}{\linewidth}
+        \[
+        \adjustbox{width=\linewidth}{
         \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
@@
     \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.45\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
+    \vspace{1em}
+    \begin{subfigure}{\linewidth}
+        \[
+        \adjustbox{width=\linewidth}{
         \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
@@
     \end{subfigure}
+    \caption{New consistent component introduction rule}
+    \label{fig:consistent_introduction}
 \end{figure}
@@
 \[
-\adjustbox{scale=0.75}{
+\adjustbox{width=0.8\linewidth}{
     \tikzfig{../figures/appendix/cartesian_rewrite_fix}
@@
 \[
-\adjustbox{scale=0.75}{
+\adjustbox{width=\linewidth}{
     \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix}
@@
 \[
-\adjustbox{scale=0.75}{
+\adjustbox{width=\linewidth}{
     \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix_2}
@@
 \[
-\adjustbox{scale=0.65}{
+\adjustbox{width=\linewidth}{
     \tikzfig{../figures/appendix/rule_transformation_example}
@@
 
-\section{Going from e-graph to e-hypergraph}
+The Definition~\ref{def:boundary_new} will then need a few additional constraints.
+\begin{itemize}
+\item For all top-level edges $e_{i}, e_{j}$ of type $0 \to 0$ in $\mathcal{L}$ it must be either $[m(e_{i})) = [m(e_{j})) = \varnothing$ or $m(e_{i}) \consistency m(e_{j})$.
+\item For all edges $e_{i}, e_{j}$ of type $0 \to 0$ in $i + j$ it must be either $[[c_1,c_2](e_{i})) = [[c_1,c_2](e_{j})) = \varnothing$ or $[c_1,c_2](e_{i}) \consistency [c_1,c_2](e_{j})$.
+\item For all edges $e$ of type $0 \to 0$ in $i + j$ it must be that $l([c_1,c_2](e)) = l(e) = \bot$ and $[c_1,c_2](e)$ must be maximal.
+\end{itemize}
+Then, when constructing the corresponding extended cospans, such auxiliary edges must be ignored in the interfaces.
+
+\subsection{E-graph interpretation in $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$}
+\label{sec:appendix:e-graph-translation}
+
+In this section we will elaborate on the interpretation of e-graphs as morphisms of $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$.
+We will first recite the formal definition of an e-graph.
 
@@
 \begin{definition}[Congruence invariant~\cite{EggPaper}]
-Two e-nodes $f(a_{1}, \ldots, a_{n})$ and $f_{b_{1}, \ldots, b_{n}}$ are \textit{congruent}, if $find(a_{i}) = find(b_{i})$ for all $i$.
+Two e-nodes $f(a_{1}, \ldots, a_{n})$ and $f_{b_{1}, \ldots, b_{n}}$ are \textit{congruent}, if $\text{find}(a_{i}) = \text{find}(b_{i})$ for all $i$.
 The equivalence relation on e-nodes must be closed under this congruence, that is, two congruent e-nodes must reside in the same e-class.
@@
 \begin{definition}
-The pseudocode in Algorithm~\ref{alg:translation} defines the translation function $\llbracket \rrbracket : \catname{E}\text{-}\catname{graph} \to \WellTypedMdaEcospans$ that turns a \textit{canonical} (the one for which the above invariants hold) to a cospan of e-hypergraphs.
+The pseudocode in Algorithm~\ref{alg:translation} defines the translation function $\llbracket \rrbracket : \catname{E}\text{-}\catname{graph} \to \catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$ that turns a \textit{canonical} (the one for which the above invariants hold) \textit{acyclic} \textit{connected} e-graph into a cospan of e-hypergraphs.
 \end{definition}
@@
 \begin{figure}
-    \begin{subfigure}{0.4\linewidth}
+    \begin{subfigure}[C]{0.35\linewidth}
 \[
@@
 \]
+\caption{\;}
 \end{subfigure}
-\begin{subfigure}{0.4\linewidth}
+\hfill
+\begin{subfigure}[C]{0.6\linewidth}
     \[
@@
     \]
+\caption{\;}
 \end{subfigure}
+\hfill
+\caption{E-graph (a) and a corresponding morphism of $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$ (b)}
+\label{fig:e-graph-to-cospan}
 \end{figure}
 
-\begin{algorithm}
+\begin{algorithm*}
     \caption{E-graph to e-hypergraph translation}
@@
     \end{algorithmic}
-\end{algorithm}
+\end{algorithm*}
 
@@
 
+\begin{remark}
+Technically, $\mathcal{E}_{C}$ is not induced by equations between $\Sigma$-terms, but rather between $\Sigma^{+}$-terms as we can apply naturality of copy and delete to boxes as well.
+This does not pose any issues as the proof of Theorem~\ref{thm:full-completeness} works for any $\mathcal{E}$ as long as the conditions of Definition~\ref{def:boundary_new} are satisfied.
+\end{remark}
+
 \begin{figure}

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/combinatorial_semantics.tex
wip

--- 2025-01-21 LICS/sections/combinatorial_semantics.tex
@@
 
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
+% In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of SMTs including a Frobenius algebra.
+% Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
+% When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
+% However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
+% This gives rise to the definition of a hypergraph with interfaces.  
+% Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
+
+In ~\cite{bonchi_string_2022-1}, hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of SMTs, including Frobenius algebras.
+Hypergraphs extend graphs by allowing edges with multiple sources and targets. 
+Interpreting string diagrams as hypergraphs maps morphisms to edges and wires to vertices, while the dangling wires at the diagram's top and bottom give rise to \textit{input} and \textit{output} interfaces for a hypergraph.
+Subsequent work \cite{bonchi_string_2022-2} refines this to \textit{monogamous} directed acyclic hypergraphs, establishing a correspondence between symmetric monoidal string diagrams (without Frobenius algebra) and the category of (cospans of) such hypergraphs.
 
@@
 \begin{enumerate}
+    \vspace{-1.25mm}
     \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
@@
     \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
+    \vspace{-0.5mm}
 \end{enumerate}
+% \[
+% \begin{array}{ll}
+%     \phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e)) & \phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))\\
+%     l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e)) & \;
+% \end{array}
+% \]
 We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
@@
 \else
+\vspace{-2mm}
 \begin{multline*}
     (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+    \vspace{-1.25cm}
 \end{multline*}
@@
 In particular, this means that any morphism in $\HypI{\Sigma}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
+where $n,\;m$ are discrete ordered hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
 The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
@@
 \subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\catname{S}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
+% The correspondence of the previous section means that the symmetric monoidal equations of $\catname{S}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
 The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
+% DPO rewriting is a standard technique for formalising graph rewriting categorically. 
+% In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
+The correspondence from the previous section absorbs the symmetric monoidal equations of $\catname{S}(\Sigma, \mathcal{E})$ into the hypergraph representation. 
+The equations $\mathcal{E}$ become rewrite rules on string diagrams, modelled as DPO rewrites on hypergraphs with interfaces.
+This section reviews standard DPO rewriting for hypergraphs and the \textit{convex} rewriting, which is sound and complete for arbitrary SMTs.
 
+DPO rewriting is a standard categorical approach to graph transformations~\cite{dpo}.
 The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
@@
 
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal \mathcal{L}^{\bot} \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal{L}^{\bot}$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
+% The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal \mathcal{L}^{\bot} \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+% When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal{L}^{\bot}$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
+% The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
+% In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
+
+The morphism $\mathcal L \to \mathcal G$ is called a \textit{match}, defined when a corresponding \textit{pushout complement} exists, comprising morphisms $\mathcal K \to \mathcal{L}^\bot \to \mathcal G$ completing the top half of the pushout square $\mathcal K \to \mathcal L \to \mathcal G$.
+For monomorphic rewrite rules, constructing $\mathcal{L}^\bot$ involves identifying $\mathcal L$ in $\mathcal G$ and removing elements with no pre-image in $\mathcal K$. 
+The final graph $\mathcal H$ is then obtained by inserting into the resulting ``hole'' the elements of $\mathcal R$ without pre-image in $\mathcal K$, resulting in two pushouts: one for deletion and one for insertion.
+Thus, we say that $\mathcal G$ is rewritten into $\mathcal H$ via $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
 
 To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
+In the case of $ \catname{Hyp}(\Sigma)$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
 By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
 We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
+Thus, we reach the following variant of DPO rewriting.
 \begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
+Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal{L}^{\bot}$ and morphisms which complete the right commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
 % \[
@@
 \begin{figure}[t!]
-    \begin{subfigure}[T]{0.4\linewidth}
+    \vspace{-0.5cm}
+    \begin{subfigure}[T]{0.35\linewidth}
     \[
@@
     \hfill
-    \begin{subfigure}[T]{0.4\linewidth}
+    \begin{subfigure}[T]{0.35\linewidth}
         \[
@@
     \end{subfigure}
-    \captionsetup{belowskip=-3ex}
+    \hfill
+    \captionsetup{skip=0pt, belowskip=0pt}
     \caption{DPO and DPOI squares}
     \label{fig:dpo_dpoi}
+    \vspace{-7mm}
 \end{figure}
@@
 \end{definition}
+\vspace{-1ex}
 We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.  
-
+\vspace{-1ex}
 \begin{definition}[Boundary complement]
@@
 %By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\catname{S}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as (cospans of) hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
+It is then proved in~\cite{bonchi_string_2022-2} that this notion of DPOI rewriting is sound and complete for $\catname{S}(\Sigma, \mathcal{E})$: for any morphisms $f$ and $g$ in $\catname{S}(\Sigma)$,  we have that $f = g$ modulo $\mathcal{E}$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as (cospans of) hypergraphs.
+As SMC equations get absorbed by the representation, it makes this approach computationally appealing.
+% In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
 %\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/conclusion.tex
wip

--- 2025-01-10 LICS/sections/conclusion.tex
@@
 \section{Conclusion}
+\label{sec:conclusion}
 
@@
 In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string}
+Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string};
 % , as sketched in Fig. \ref{fig:app1}(a)
-; the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}
+the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX};
 % , as sketched in Fig. \ref{fig:app1}(b)
-; or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes
+or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes.
 % , as sketched in Fig.~\ref{fig:app2}. 
 All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
+In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure as the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
+ % traced comonoid
 
@@
 
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
+% In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
 Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
@@
 However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
+% Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
 Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/e-hypergraphs.tex
wip

--- 2025-01-21 LICS/sections/e-hypergraphs.tex
@@
 
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$  such that for all $c_1,c_2 \in \Sigma$, if arity (respectively, co-arity) of $c_1$ is 0, then co-arity (respectively, arity) of $c_2$ must be at least 1.
-This is to ensure we do not have terms of type $0 \to 0$.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
+% Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$  such that for all $c_1,c_2 \in \Sigma$, if arity (respectively, co-arity) of $c_1$ is 0, then co-arity (respectively, arity) of $c_2$ must be at least 1.
+% This is to ensure we do not have terms of type $0 \to 0$.
+% This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
+\begin{remark}
+Here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ for which the corresponding family of $\Sigma$-terms does not contain terms of type $0 \to 0$ (apart from the monoidal unit).
+This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section.
 A discussion of a general case can be found in section~\ref{sec:dpo-fix} of Appendix.
+\end{remark}
 
@@
 Intuitively,  the child relation equips a hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.  
+These edges can be designated as ``parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.  
 These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
@@
 Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
+
+% To save space
+% Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
 
@@
 
-Consider a relation $R = \{x R y \text{ if } f_{int}(x) \consistency f_{int}(y) \text{ for } x,y \in n'\}$ and let $S$ be its reflexive closure.
-The latter partitions $n'$ into non-empty subsets $\{p_{j}\}_{j=1}^{k}$. 
-We get an analogous partition for $m'$.
-An example of an extended cospan can be seen in Figure~\ref{fig:e-cospan-example}.
-The partition of the interfaces is given with colours.
-The labels suggest how $f_{int}, f_{ext}$ (respectively, $g_{int}, g_{ext}$) act on vertices.
-The ordering of vertices in the interfaces is from left to right.
+% Consider a relation $R = \{x R y \text{ if } f_{int}(x) \consistency f_{int}(y) \text{ for } x,y \in n'\}$ and let $S$ be its reflexive closure.
+% The latter partitions $n'$ into non-empty subsets $\{p_{j}\}_{j=1}^{k}$. 
+% We get an analogous partition for $m'$.
 
@@
     \]
-    \captionsetup{belowskip=-3ex}
+    \captionsetup{belowskip=-3ex, skip=0pt}
     \caption{Extended cospan example}
@@
 \label{def:iso}
-Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute and such that $\alpha$ and $\gamma$ preserve order within each $p_i$.
+
+Consider two extended cospans below 
 \[
@@
 \]
+and consider a relation 
+\[
+R = \{x R y \text{ if } f_{int}(x) \consistency f_{int}(y) \text{ for } x,y \in n'\}
+\]
+and let $S$ be its reflexive closure.
+The latter partitions $n'$ into non-empty subsets $\{p_{j}\}_{j=1}^{k}$. 
+We get an analogous partition for $m'$, $n''$, $m''$.
+Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the above diagram commute and such that $\alpha$ and $\gamma$ preserve order within each $p_i$.
 \end{definition}
 Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{bonchi_string_2022-1}.
+An example of an extended cospan can be seen in Figure~\ref{fig:e-cospan-example}.
+The partition of the interfaces is given with colours.
+The labels suggest how $f_{int}, f_{ext}$ (respectively, $g_{int}, g_{ext}$) act on vertices.
+The ordering of vertices in the interfaces is from left to right.
 Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix section~\ref{sec:appendix:iso}.
 
-\begin{remark}\label{remark:embedding_functor}
-There is also a faithful identity-on-objects functor $E: \HypI{\Sigma} \to \Ecospans$ which maps a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ to an extended cospan $n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m$ such that $f_{ext};f_{int} = f$ and $g_{ext};g_{int} = g$.
-\end{remark}
+% \begin{remark}\label{remark:embedding_functor}
+% There is also a faithful identity-on-objects functor $E: \HypI{\Sigma} \to \Ecospans$ which maps a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ to an extended cospan $n \xrightarrow{f_{ext}} n \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m \xleftarrow{g_{ext}} m$ such that $f_{ext};f_{int} = f$ and $g_{ext};g_{int} = g$.
+% \end{remark}
 
@@
     h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
+\;
     h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
@@
 \end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
+
+% To save space
+% Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
 
 Unlike in the previous section,  in order to give an equivalence with $\catname{S}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.
+This is because the equations for semilattice enrichment are not subsumeable by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.
+% , as semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
 Instead,  we will treat them via rewriting,  as we do with equations of the signature.  
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
+% This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
 % , cf. string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
+
+% To save space
+% An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this.
+% Presently,  we will treat these equations via rewriting as well.
 
@@
     0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
+\;
     0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
@@
 %We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-The extended DPOI rewriting as defined in this section trades off simplicity for expressiveness: it is unable to express rewriting for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
-While this limitation can be overcome by specifically encoding such rewrite rules, we find the technical development to be simpler if we avoid consideration of such rules and therefore consider signatures which can not result in construction of terms with type $0 \to 0$.
-This limitation has no practical impact in applications of interest. 
-However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of Appendix.
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
+
+% \begin{remark}
+% The extended DPOI rewriting as defined in this section trades off simplicity for expressiveness: it is unable to express rewriting for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
+% While this limitation can be overcome by specifically encoding such rewrite rules, we find the technical development to be simpler if we avoid consideration of such rules and therefore consider signatures which can not result in construction of terms with type $0 \to 0$.
+% This limitation has no practical impact in applications of interest. 
+% However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of Appendix.
+% %This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
+% %We claim that this is only a minor limitation because in practice such edges (morphisms) with no
+% %incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
+% \end{remark}
 
@@
     \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+    \hspace{-2em} n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
     \]
@@
     \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+    \hspace{-2em} n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
     \]
@@
 
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
+First, note that conditions (6) and (7) are disjoint.
+Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level (premise for condition (6)) then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces (as in the case of Definition~\ref{def:dpoi}) and when the occurrence is nested (premise of condition (7)), those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces: one can think about the latter as applying Definition~\ref{def:dpoi} in a nested context.
 In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding MDA cospans above.
@@
     \[
+    \vspace{-1mm}
     n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
+    \vspace{-0.5mm}
     \]
@@
 This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
-\begin{remark}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{../figures/combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{../figures/combinatorial_semantics/f_plus_g_inline}}
-\]
-\end{remark}
+Examples and non-examples of a boundary complement can be found in Appendix section~\ref{sec:appendix:iso}.
+% \begin{remark}
+% The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+% \[
+% 	\mathcal{L} = \scalebox{0.8}{\tikzfig{../figures/combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{../figures/combinatorial_semantics/f_plus_g_inline}}
+% \]
+% \end{remark}
 
@@
 % \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
 \end{definition}
+Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ for reflexive, symmetric, and transitive closure of $\Rrightarrow{}$.
+An example of rewriting of extended cospans can be found in Appendix section~\ref{sec:appendix:iso}.
 % Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
 
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
+With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of $\catname{SLat}$-SMCs is sound and complete in an appropriate sense.
 

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/introduction.tex
wip

--- 2025-01-21 LICS/sections/introduction.tex
@@
 Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and morphisms of Cartesian categories enriched over a category of semilattices (\textit{semilattice-enriched}).
-The morphisms of the latter can also be depicted using \emph{string diagrams}~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}: a two-dimensional syntax that encodes equivalence classes of morphism terms.
-Enriched categories generalise plain categories by imposing more structure making the categorical language more expressive~\cite{Borceux_1994,Kelly2022BASICCO} which will be useful for modelling equivalence between terms as encountered in e-graphs.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022,bonchi_string_2022-1,bonchi_string_2022-2,ghica_rewriting_2023,ghica_hierarchical_2023}.
+% The morphisms of the latter can also be depicted using \emph{string diagrams}~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}: a two-dimensional syntax that encodes equivalence classes of morphism terms.
+% Enriched categories~\cite{Borceux_1994,Kelly2022BASICCO} generalise plain categories by imposing more structure making the categorical language more expressive~\cite{Borceux_1994,Kelly2022BASICCO} which will be useful for modelling equivalence between terms as encountered in e-graphs.
+The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3,fscd,ghica-zanassi2023string}.
 The key ingredient of our categorical approach is the semilattice-enrichment.
@@
 
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that are already phrased in categorical language with a defined notion of rewriting and thus these applications will benefit from the mentioned optimisation technique.
+Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that are induced by monoidal rather than algebraic theories with a defined notion of rewriting and thus these applications will benefit from the mentioned optimisation technique.
 % now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
@@
 
-A final contribution of this work is giving a concrete combinatorial representation of string diagrams for semilattice-enriched categories in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
+A final contribution of this work is giving a concrete combinatorial representation of string diagrams for semilattice-enriched categories in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs} or monoidal e-graphs, as the former are a generalisation of e-graphs for monoidal theories.
 We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework 
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.  
+\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3},  proving the representation \textit{sound and complete} with respect to our categorical semantics.  
 As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature~\cite{EggPaper}.
@@
 \]
+\captionsetup{skip=0pt, belowskip=-5mm}
 \caption{Example translation of acyclic e-graphs into string diagrams for semilattice-enriched Cartesian categories. }
@@
 This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs: in the former, edges connect directly to nodes inside the equivalence class, whereas in the latter, edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes 
+Explicit discarding (delete) nodes, depicted with black dots with a single wire inside the dashed boxes 
 \scalebox{0.4}{
@@
 \]
-\captionsetup{belowskip=-4ex}
+\captionsetup{skip=0pt, belowskip=-4ex}
 \caption{String diagrams for semilattice-enriched symmetric monoidal categories.}
@@
 Note, further, that in the Cartesian case we can restrict generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
+This means the (informal) translation given next is fully general for acyclic e-graphs.
 
@@
 noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding, in each component, all unnecessary inputs.
+Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
 
@@
 \]
+\captionsetup{skip=0pt}
 \caption{E-graph e-class translation to string diagram.}
 \label{fig:e-graph-to-string}
+\vspace{-8mm}
 \end{figure}
-The subfigure on the right is slightly simplified: if we denote the arity of each $c_i$ as $a_{c_{i}}$, then the dashed box on the right should contain $\sum\limits_{i} a_{c_{i}}$ inputs and each component $c_j$ has $\sum\limits_{i \not = j} a_{c_{i}}$ weakening nodes inside.
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
+% The subfigure on the right is slightly simplified: if we denote the arity of each $c_i$ as $a_{c_{i}}$, then the dashed box on the right should contain $\sum\limits_{i} a_{c_{i}}$ inputs and each component $c_j$ has $\sum\limits_{i \not = j} a_{c_{i}}$ weakening nodes inside.
+% Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
 % In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
@@
 
-The motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing as used in e-graphs.
+The motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing as could be seen in Figure~\ref{fig:e-graph-example}.
 However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
@@
 The latter representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient.
-\emph{Hypergraphs} are a combinatorial representation designed to address this issue~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
+\emph{Hypergraphs} are a combinatorial representation designed to address this issue~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3}.
 % A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
@@
 Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022}, is still applicable.
+The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022-1}, is still applicable.
 However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs} --- hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
@@
 % String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
+Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022-1,bonchi_string_2022-2, bonchi_string_2022-3}.
 
@@
 Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.  
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.  
+We write the identity morphism on $A$ as $\id_A$.  
 We commonly write $f;g$ for composition in diagrammatic order.  
@@
 We denote by $\catname{SLat}$ the closed monoidal category of semilattices which is defined in section~\ref{sec:appendix:slat} of the Appendix.
+Given two adjoint functors $F : \mathbb{A} \to \mathbb{B}$ and $G : \mathbb{B} \to \mathbb{A}$ with $F$ being left-adjoint, we write $F \dashv G : \mathbb{B} \to \mathbb{A}$ for this adjunction.
 
@@
 \]
-A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations of the SMT of commutative comonoids, and the following additional naturality of copy-delete equations, for every $c \in \Sigma_C$
+A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations $\mathcal{E}$ that include the SMT of commutative comonoids, and the following additional naturality of copy-delete equations, for every $c \in \Sigma_C$
 \[
@@
 % This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$ (\textit{e.g.},  $\catname{SLat}$).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism which satisfies certain axioms.
-Composition being an $M$-morphism implies it respects the additional structure on hom-sets.
+Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$, but we will omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \catname{SLat}$.
+%  (\textit{e.g.},  $\catname{SLat}$).
+% In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
+% \[
+% 	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
+% \]
+% is defined as an $M$-morphism which satisfies certain axioms.
+% Composition being an $M$-morphism implies it respects the additional structure on hom-sets.
 %  In other words,  composition must respect the additional structure on hom-sets.  
@@
 Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \catname{SLat}$.
+% We omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \catname{SLat}$.
 
@@
 % https://q.uiver.app/#q=WzAsMixbMCwwLCJcXGNhdG5hbWV7U0xhdHR9LVxcY2F0bmFtZXtDYXR9Il0sWzIsMCwiXFxjYXRuYW1le0NhdH0iXSxbMSwwLCJGIiwyLHsiY3VydmUiOjR9XSxbMCwxLCJHIiwyLHsiY3VydmUiOjR9XSxbMiwzLCIiLDAseyJsZXZlbCI6MSwic3R5bGUiOnsibmFtZSI6ImFkanVuY3Rpb24ifX1dXQ==
-\[\begin{tikzcd}
-	\catname{SLatt}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
-	\end{tikzcd}
+% \[\begin{tikzcd}
+% 	\catname{SLatt}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
+% 	\end{tikzcd}
+% \]
+\vspace{-1mm}
+\[
+\mathcal{F} \dashv \mathcal{U} : \catname{SLat}\text{--}\catname{Cat} \to \catname{Cat}
+\vspace{-1mm}
 \]
@@
 
-\[\begin{tikzcd}
-	\catname{SLatt} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
-	\end{tikzcd}
+% \[\begin{tikzcd}
+% 	\catname{SLatt} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
+% 	\end{tikzcd}
+% \]
+\vspace{-1mm}
+\[
+F \dashv U : \catname{SLat} \to \catname{Set}
+\vspace{-1mm}
 \]
 \end{proposition}
-
-In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice-enriched $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
+In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free $\catname{SLat}$-category $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
 % SMC $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
@@
 \begin{definition}
-We define $\catname{S}(\Sigma)^{+}$ and $\catname{S}(\Sigma, \mathcal{E})^{+}$ to be semilattice-enriched SMCs obtained as $\mathcal{F}(\catname{S}(\Sigma))$ and $\mathcal{F}(\catname{S}(\Sigma, \mathcal{E}))$ respectively.
+We define $\catname{S}(\Sigma)^{+}$ and $\catname{S}(\Sigma, \mathcal{E})^{+}$ to be $\catname{SLat}$-SMCs obtained as $\mathcal{F}(\catname{S}(\Sigma))$ and $\mathcal{F}(\catname{S}(\Sigma, \mathcal{E}))$ respectively.
 Both of these $\catname{SLat}$-categories are in $\catname{PROP}^{+}$ which we define to be the image of $\catname{PROP}$ via the same adjunction.
@@
 We overload the binary notation $+$ for our $n$-ary notation.
-We will later prove the intuitive fact that these diagrams are sound and complete with respect to their intended categorical semantics, noting that similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
+% We will later prove the intuitive fact that these diagrams are sound and complete with respect to their intended categorical semantics, noting that similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
+Notably, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
 
@@
 \]
+\captionsetup{skip=0pt}
 \caption{Equations for a  semilattice-enriched symmetric monoidal category}
 \label{fig:string-equations}
+\vspace{-8mm}
 \end{figure*}

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/combinatorial_semantics.tex
cleanup prose, make fit in 12 pages

@@
 
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
+In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  recapitulating the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
 
@@
 The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-The ordering of the interface hypergraphs is the way to differentiate between the interpretations of morphisms $f \otimes g$ and $g \otimes f$.
+The ordering of the interface associated to a hypergraph is used to differentiate between the interpretations of morphisms $f \otimes g$ and $g \otimes f$.
 
@@
 It is then proved in~\cite{bonchi_string_2022-2} that this notion of DPOI rewriting is sound and complete for $\catname{S}(\Sigma, \mathcal{E})$: for any morphisms $f$ and $g$ in $\catname{S}(\Sigma)$,  we have that $f = g$ modulo $\mathcal{E}$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as (cospans of) hypergraphs.
-As SMC equations get absorbed by the representation, it makes this approach computationally appealing.
+As SMC equations get absorbed by the representation, it makes this approach computationally appealing as they do not need to be represented explicitly.
 % In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/e-hypergraphs.tex
cleanup prose, make fit in 12 pages

@@
 The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  
-A concrete description of the pushout of two morphisms in this category is given in the Appendix section~\ref{sec:appendix:pushout}.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does in all scenarios that we are interested in, most importantly the pushout for the composition of two cospans with discrete feet exists.
+A concrete description of the pushout of two morphisms in this category is given in Appendix~\ref{sec:appendix:pushout}.
+Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does in all scenarios that we are interested in; most importantly, the pushout for the composition of two cospans with discrete feet always exists.
 
@@
 
-When modelling enriched string diagrams using e-hypergraphs,  we require keeping track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
+When modelling enriched string diagrams using e-hypergraphs,  we must keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which also must be tracked.
 Thus,  we introduce an \textit{extended} cospan construction.
@@
 The partition of the interfaces is given with colours.
-The labels suggest how $f_{int}, f_{ext}$ (respectively, $g_{int}, g_{ext}$) act on vertices.
-The ordering of vertices in the interfaces is from left to right.
-Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix section~\ref{sec:appendix:iso}.
+The labels suggest how $f_{int}, f_{ext}$ (respectively, $g_{int}, g_{ext}$) act on vertices, with the ordering of vertices in the interfaces given left to right.
+Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix~\ref{sec:appendix:iso}.
 
@@
 \end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
+First, $\mathcal{H}$ is computed as the result of the pushout square shown below: 
 % https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
@@
 Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
 
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
+Thus,  we wish for a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows: 
 \[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
+    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m,
 \qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
+    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m.
 \] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
+Analogously to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data:
 \[
@@
     0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
+\] allows us to encode rewrite rules as extended cospans of the following form:
 \[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
+\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R},
 \] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
@@
 First, note that conditions (6) and (7) are disjoint.
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level (premise for condition (6)) then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces (as in the case of Definition~\ref{def:dpoi}) and when the occurrence is nested (premise of condition (7)), those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces: one can think about the latter as applying Definition~\ref{def:dpoi} in a nested context.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding MDA cospans above.
+Intuitively, the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level (premise for condition (6)) then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces (as in the case of Definition~\ref{def:dpoi}), and when the occurrence is nested (premise of condition (7)), those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces: one can think about the latter as applying Definition~\ref{def:dpoi} in a nested context.
+In both cases, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding MDA cospans above.
 The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
@@
 This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
-Examples and non-examples of a boundary complement can be found in Appendix section~\ref{sec:appendix:iso}.
+Examples and non-examples of a boundary complement can be found in Appendix~\ref{sec:appendix:iso}.
 % \begin{remark}
@@
 \label{prop:boundary_unique}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique.
+    The boundary complement in~\ref{def:boundary_new} when it exists is unique.
 \end{proposition}
@@
 Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ for reflexive, symmetric, and transitive closure of $\Rrightarrow{}$.
-An example of rewriting of extended cospans can be found in Appendix section~\ref{sec:appendix:iso}.
+An example of rewriting of extended cospans can be found in Appendix~\ref{sec:appendix:iso}.
 % Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.

### 2025-01-23 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
fixed grey color
last section left for closed

@@
 \usepackage{textcomp}
-\usepackage{xcolor}
+\usepackage{xcolor,soul}
 
 \usepackage{amsthm}
+\usepackage{hyperref}
+\usepackage{stmaryrd}
 
@@
 \begin{abstract}
-    The technique of equality saturation for program optimization and reasoning has recently gained significant attention due to the renewed popularity of equality graphs (e-graphs).
+    The technique of equality saturation for program optimisation and reasoning has recently gained significant attention due to the renewed popularity of equality graphs (e-graphs).
     E-graphs represent multiple terms that are equivalent modulo a given set of rewrite rules. 
@@
 	Notably, e-graphs struggle in domains involving variable binding, such as $\lambda$-calculus, because they lack native support for binding.
-    Building on recent work that provides a categorical interpretation of e-graphs as morphisms in a semilattice-enriched symmetric monoidal category, we extend this framework.
+    Building on recent work that provided a categorical interpretation of e-graphs as morphisms in a semilattice-enriched symmetric monoidal category, we extend this framework. 
 	Specifically, we show how e-graphs with bindings can be derived as morphisms in a closed symmetric monoidal category with similar enrichment. 
@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 All these issues can be resolved by the use of \emph{string diagrams}.
-Thus, in this paper we take what we consider to be a principled approach to e-graphs with bindings by lifting to monoidal closed categories prior work showing how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}
+Thus, in this paper we take what we consider to be a principled approach to e-graphs with bindings by lifting to monoidal closed categories prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} showing how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
 The approach in \emph{loc.
@@
 String diagrams have the advantage of being a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, all based on solid mathematical foundations.
-To solve the new problems discuss above, of dealing with binding and capture avoiding substitution, we just follow the categorical breadcrumb trail.
+To solve the new problems discusses above, of dealing with binding and capture avoiding substitution, we just follow the categorical breadcrumb trail.
 
@@
 
-E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the DAG representation of terms to include equivalence classes.
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, a `Hello world!' of e-graphs~\cite{EggPaper}.
+E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
+The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph as first presented in ~\cite{EggPaper}.
 
@@
 In Figure~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following the $\beta$-reductions.
-A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term, to be rewritten using the rules for explicit substitution at the top of each column.
+A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box), to be rewritten using the rules for explicit substitution at the top of each column.
 At the end, all explicit substitution nodes can be removed.
@@
 \input{sections/combinatorial_semantics}
+\input{sections/soundness}
 

### 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
last section left for closed

@@
 
-We recall some basic definitions.
+Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.  
+We write the identity morphism on $A$ as $\id_A$.  
+We commonly write $f;g$ for composition in diagrammatic order.  
+Composition in the usual order is written $g \circ f$.  
+We denote the tensor product of an SMC $\mathbb{C}$ by $\otimes$,  its unit by $I$, and its symmetry natural transformation as $\sym$ \cite{maclane}.  
+We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.  
+We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
+
+First, recall some basic definitions.
 \begin{definition}
@@
 The free closed symmetric monoidal category $\catname{CS}(\Sigma)$ is given by a set of closed $\Sigma$-terms quotiented by the laws of a symmetric monoidal category and the equations from Definition~\ref{def:closed}.
+We can similarly consider a free closed symmetric monoidal category that is induced by an $\catname{SMT}(\Sigma, \mathcal{E})$ over closed $\Sigma$-terms which we will denote with $\catname{CS}(\Sigma, \mathcal{E})$.
 \end{definition}
 
-To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator $+$.
+To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator \textit{join} $+$ as presented in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
 This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix section~\ref{sec:appendix:slat}).
@@
 The following distributivity law can then be derived, forming the basis of encoding equivalences inside abstractions.
-\begin{align*}
+\begin{equation}
+\begin{aligned}
 \Lambda_{A,B,C}(f + g) &= \eta;(B \multimap (f + g))\\
@@
                        &= \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g)
-\end{align*}
+\end{aligned}
+\end{equation}
+\label{law:distributivity}
 
@@
 Given two $\Sigma^{+}$-terms $f : A \to B$ and $g : A \to B$, $f + g : A \to B$ is also a $\Sigma^{+}$-term.
+Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed $\catname{SLat}$-SMC, including the distributivity law~\ref{law:distributivity}.
 
-To aid reasoning, we introduce a new language of string diagrams for $\catname{SLat}$-SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-Figure \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} apart from the last component which is our notation for $+$; the first component denotes an empty diagram.
-Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
+To aid reasoning, we introduce a new language of string diagrams for closed $\catname{SLat}$-SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms as well as another hierarchical ``box'' for lambda abstraction.
+It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting and in interpreting $\lambda$-terms.
+Figure \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
+Note the orientation of string diagrams are now from top to bottom as opposed to what was used in~\nameref{sec:introduction}: the latter was mimicking the orientation of e-graphs, \textit{i.e.}, traditional orientation for expression trees.
+% Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
+% The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
 We overload the binary notation $+$ for our $n$-ary notation.
-We will later prove the intuitive fact that these diagrams are sound and complete with respect to their intended categorical semantics, noting that similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
+% We will later prove the intuitive fact that these diagrams are sound and complete with respect to their intended categorical semantics, noting that similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
+
+In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over some signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
+We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
+That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
+We call these string diagrams, or more precisely, the combinatorial representation of that will follow, e-graphs with bindings.
+
+\begin{figure*}
+\[
+\adjustbox{scale=0.75}{
+\tikzfig{./figures/egraph-strings}
+}
+\]
+\caption{String diagrams for closed semilattice-enriched symmetric monoidal categories.}
+\label{fig:egraph-strings}
+\end{figure*}

### 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
last section left for closed

@@
 We will also refer to $A,B$ as \textit{feet} of the cospan, and to $C$ as \textit{carrier}.
+Given an ordered sequence of labels from some alphabet $\Sigma$, $[l_1, \ldots, l_n]$ we will denote with $w([l_1, \ldots, l_n])$ the word concatenated from these labels using $\otimes$, \textit{e.g.}, $w([A,B]) = A \otimes B$.
 
@@
 Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
+We then evoke the definition of hypergraphs and corresponding cospans as introduced in~\cite{bonchi_string_2022-1,bonchi_string_2022-2} over not uni-typed signatures.
 
 \begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
+A \emph{hypergraph $\mathcal{G}$ over monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l = (l_{V} : V \to \Sigma_{O},l_{E} : E \to \Sigma_{M})$ is a pair of labelling functions that assigns each edge and vertex an operator and a type from monoidal signature $\Sigma$ respectively.
+The labelling function must respect the typing of the operator: for all edges $e$, such that $l(e) = c : A \to B$,  $l_{V}^{*}(s(e)) = A$ and $l_{V}^{*}(t(e)) = B$.  
 We call a hypergraph \textit{discrete} if its set of edges is empty.   
@@
     \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
+    \item $l_{V,\mathcal{F}}(v) = l_{V,\mathcal{G}}(\phi_V(v))$
+    \item $l_{E,\mathcal{F}}(e) = l_{E,\mathcal{G}}(\phi_E(e))$
 \end{enumerate}
+That is, a homomorphism should preserve sources and targets of edges and preserve the labels.
 We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms.
 \end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
+$\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
 The initial object is given by the empty hypergraph.
 
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
+To properly interpret morphisms of a symmetric monoidal category we need to account for the dangling wires at the diagram's top and bottom that represent inputs and outputs of a morphism.
+Such dangling wires are represented as interfaces for a given hypergraph giving rise to a category of cospans of hypergraphs as defined next.
+
+% \begin{definition}[Symmetric monoidal category of cospans]
+% Let $\mathbb{C}$ be a category with all finite colimits.  
+% A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
 This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
+
+% Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
 Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\ifdefined \ONECOLUMN
-\[
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\]
-\else
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-\fi
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
+% \begin{multline*}
+%     (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+% \end{multline*}
+% and its unit by the initial object of $\mathbb{C}$.
+% Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
+% \end{definition}
 
-Then, morphisms of $\catname{S}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
+% Then, morphisms of $\catname{S}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
 \begin{definition}[Hypergraphs with interfaces]
 \label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete \textit{ordered} hypergraphs as objects.
+A hypergraph with interfaces is a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\catname{Hyp}(\Sigma)$ where $n$, $m$ are discrete ordered hypergraphs that denote interfaces.
+Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the category with discrete \textit{ordered} hypergraphs as objects and isomorphism classes of hypergraphs with interfaces as morphisms.
 \end{definition}
+$\HypI{\Sigma}$ is a symmetric monoidal category with the tensor product given by coproduct in $\catname{Hyp}(\Sigma)$:
 
-In particular, this means that any morphism in $\HypI{\Sigma}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-The ordering of the interface hypergraphs is the way to differentiate between the interpretations of morphisms $f \otimes g$ and $g \otimes f$.
+\begin{align*}
+n \xrightarrow{f} &\mathcal{F} \xleftarrow{g} m\\
+&\times \qquad \quad \Coloneqq n + n' \xrightarrow{f + f'} \mathcal{F} + \mathcal{G} \xleftarrow{g + g'} m + m'\\
+n' \xrightarrow{f'} &\mathcal{G} \xleftarrow{g'} m'
+\end{align*}
+Unit is given by the initial object of $\catname{Hyp}(\Sigma)$.
+Symmetry is inherited from coproduct in $\catname{Hyp}(\Sigma)$~\cite{MonoidalCoproduct}.
+
+Composition of two cospans $n \xrightarrow{f} \mathcal{F} \xleftarrow{g} m;m \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} k$ is given by the pushout
+\[
+n \xrightarrow{f} \mathcal{F} +_{g,f'} \mathcal{G} \xleftarrow{g'} k
+\]
+
+that is computed by identifying vertices in $\mathcal{F} + \mathcal{G}$ that share pre-image in $m$.
+As pushout is only defined up to isomorphism, it justifies the morphisms in this category being isomorphism classes of cospans.
+
+To make such cospans complete with respect to representing morphism of $\catname{S}(\Sigma)$, we need to impose some restriction in their form.
 
@@
 
-In order to make the interpretation complete with respect to $\catname{S}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
 \begin{definition}[Category of MDA Hypergraphs with Interfaces~\cite{bonchi_string_2022-2}]
@@
 
+Practical meaning of the above theorem is that two $\Sigma$-terms $f$ and $g$ are equal modulo SMC equations if and only if they are represented as the same hypergraph with interfaces.
+The theorem defines a full and faithful functor that is induced by interpretation function of morphism terms that renders the latter as cospans of hypergraphs and is given in Appendix~\ref{sec:appendix:interpretation}.
+Term rewriting and hence equality modulo $\mathcal{E}$ is then implemented as DPO rewrites over such hypergraphs with interfaces that we recall next.
+
 \subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\catname{S}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
+DPO rewriting is a standard technique for formalising graph transformations categorically.
+% In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
 The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
@@
 A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
 
@@
 Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
+
 \begin{figure}[t!]
@@
 \end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
+To make DPOI rewriting sound and complete with respect to rewriting of $\Sigma$-terms modulo SMC laws and equations $\mathcal{E}$ the notion of both pushout complement and match must be restricted: otherwise there might be DPO rewrites that are not derivable from SMC laws and $\mathcal{E}$-equations.
+This gives rise to the following notion of a \textit{boundary complement} and \textit{convex} DPOI rewriting.
 
@@
 and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
 \[
-    \scalebox{0.8}{
-        \tikzfig{../figures/combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
+    \adjustbox{scale=0.8}{
+        \tikzfig{./figures/DPOI_boundary_complement}
+    }\]
 is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
@@
 Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-Yet, this restriction is not enough to make DPOI rewriting complete for $\catname{S}(\Sigma,\mathcal{E})$: there might be cospans with carriers $\mathcal{G}$ and $\mathcal{H}$ such that one can be rewritten into another but their corresponding morphisms are not equal modulo $\mathcal{E}$~\cite{bonchi_string_2022-2}.
-To make the DPOI rewriting complete, we need to add a restriction on the image of the match.
+% Yet, this restriction is not enough to make DPOI rewriting complete for $\catname{S}(\Sigma,\mathcal{E})$: there might be cospans with carriers $\mathcal{G}$ and $\mathcal{H}$ such that one can be rewritten into another but their corresponding morphisms are not equal modulo $\mathcal{E}$~\cite{bonchi_string_2022-2}.
+% To make the DPOI rewriting complete, we need to add a restriction on the image of the match.
 
@@
 
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\catname{S}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as (cospans of) hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
+Then, the authors of~\cite{bonchi_string_2022-2} showed that this enhanced notion of DPOI rewriting is sound and complete in the following sense.
+Any two $\Sigma$-terms $f$ and $g$ are equal modulo SMC laws and equations $\mathcal{E}$ if and only if $\llbracket f \rrbracket \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} \llbracket g \rrbracket$ where $\llbracket - \rrbracket$ is the functor giving the equivalence in Theorem~\ref{thm:prop-equiv} and $\llbracket E \rrbracket = \{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle, \langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle \text{ for } l=r \in \mathcal{E} \}$.
+We then proceed with building a combinatorial representation for closed $\Sigma^{+}$-terms to obtain a notion of DPOI rewriting that would also subsume equality modulo SMC laws.
 
-We extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness.
+\section{Closed E-hypergraphs}
+Prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
+The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established area of hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
+The latter work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
+Notably, the laws of a semilattice enrichment in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
+
+We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness and when evoking a term e-hypergraph we will mean the notion introduced in what follows rather than what was defined in \textit{loc.
+cit.}.
 
@@
 \end{itemize} 
-$\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations can consist of exclusively hierarchical edges, then we have $\colorbox{yellow}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
-Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $<$ and~$\consistency$.
+$\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges; then we have $\colorbox{yellow}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
+Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{red}{<}, \textcolor{blue}{<}$ and~$\consistency$.
 \end{definition}
@@
 In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
+$\textcolor{blue}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{red}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
 
@@
 
+Intuitively, such notion of isomorphism makes it possible to freely rearrange interfaces for disjoint components.
+For example, in the input interface of cospan in Figure~\ref{fig:e-cospan-example}, we can swap 
+\begin{tikzpicture}
+\begin{pgfonlayer}{nodelayer}
+		\node [style=red node, label={above:$v_{4}$}] (0) at (-0.5, 3.5) {};
+		\node [style=red node, label={above:$v_5$}] (1) at (0, 3.5) {};
+		\node [style=red node, label={above:$v_6$}] (2) at (0.5, 3.5) {};
+	\end{pgfonlayer}
+\end{tikzpicture}
+with 
+\begin{tikzpicture}
+\begin{pgfonlayer}{nodelayer}
+    \node [style=blue node, label={above:$v_{7}$}] (0) at (-0.5, 3.5) {};
+    \node [style=blue node, label={above:$v_8$}] (1) at (0, 3.5) {};
+    \node [style=blue node, label={above:$v_9$}] (2) at (0.5, 3.5) {};
+\end{pgfonlayer}
+\end{tikzpicture}
+but we can not swap
+\begin{tikzpicture}
+    \begin{pgfonlayer}{nodelayer}
+            \node [style=red node, label={above:$v_{4}$}] (0) at (-0.5, 3.5) {};
+            \node [style=red node, label={above:$v_5$}] (1) at (0, 3.5) {};
+        \end{pgfonlayer}
+\end{tikzpicture}
+with each other as it would result in a non-isomorphic cospan.
+
 Composition of two morphisms 
@@
 Consider, again, Figure ~\ref{fig:e-cospan-example}.
-To make edge $e_{1}$ well-typed, it is necessary that $w([v_{4},v_{5},v_{6}]) = w([v_{10},v_{11},v_{12}]) = w([v_1,v_2,v_3])$ and $w([w_{4},w_{5},w_{6}]) = w([w_{10},v_{12},v_{11}]) = w([w_1,w_2,w_3])$ (note the change of ordering in the output of $e_{1}$).
-To make edge $e_{2}$ well-typed, it is necessary that $w([v_4,v_5,v_9]) = w([v_7,v_8,v_9])$ and $w([u_1]) = w([v_9]) \multimap w([v_8,v_7,w_9])$.
+To make edge $e_{1}$ well-typed, it is necessary that $l_{V}^{*}([v_{4},v_{5},v_{6}]) = l_{V}^{*}([v_{10},v_{11},v_{12}]) = l_{V}^{*}([v_1,v_2,v_3])$ and $l_{V}^{*}([w_{4},w_{5},w_{6}]) = l_{V}^{*}([w_{10},v_{12},v_{11}]) = l_{V}^{*}([w_1,w_2,w_3])$ (note the change of ordering in the output of $e_{1}$).
+To make edge $e_{2}$ well-typed, it is necessary that $l_{V}^{*}([v_4,v_5,v_9]) = l_{V}^{*}([v_7,v_8,v_9])$ and $l_{V}([u_1]) = l_{V}([v_9]) \multimap l_{V}^{*}([v_8,v_7,w_9])$.
 
-Unlike in the previous section,  in order to give an equivalence with $\catname{S}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation, apart from the commutativity equation which is absorbed by the definition of isomorphic cospans.
-Instead,  we treat them via rewriting,  as we do with equations of the signaturem because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away.
-% , cf. string diagrams for Cartesian categories.  
-An ideal combinatorial representation would factor out the two semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity and idempotence) --- however, a limitation of the current work is that we do not achieve this,  so we also handle these equations via rewriting.
+We then proceed with definition a notion of DPOI rewriting for such e-hypergraphs.
 
@@
 Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments, defined in this section.
+The construction in this section closely follows the one introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} with some adjustments to account for $\textcolor{blue}{<}$.
 
@@
 %We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-The extended DPOI rewriting as defined in this section trades off simplicity for expressiveness: it is unable to express rewriting for rules whose source or target contain a maximal connected sub-hypergraph with no input and output vertices.
-While this limitation can be overcome by specifically encoding such rewrite rules, we find the technical development to be simpler if we avoid consideration of such rules and therefore consider signatures which can not result in construction of terms with type $0 \to 0$.
-This limitation has no practical impact in applications of interest. 
-However, the general case without this limitation can be found in Section~\ref{sec:dpo-fix} of the Appendix.
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
 
@@
     \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
+    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{blue}{E}$ s.t.
+$v,w \textcolor{blue}{<}^{\mu} e$;
+    \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{blue}{E}$ s.t.
+$v,w \textcolor{blue}{<}^{\mu} e$;
     \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
@@
 
+The conditions (3) and (4) require that top-level connected components in $\mathcal{L}$ do not end up inside edges of different types within $\mathcal{G}$ and $\mathcal{L}^{\bot}$ under $m$ and $[c_1,c_2]$.
+
 %The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
@@
 
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
-\begin{remark}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{./figures/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{./figures/f_plus_g_inline}}
-\]
-\end{remark}
+% Note that,  in the above definition,  the extended cospan 
+%     \[
+%     n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
+%     \]
+%  is not necessarily well-typed.
+% This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
+% \begin{remark}
+% The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
+% \[
+% 	\mathcal{L} = \scalebox{0.8}{\tikzfig{./figures/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{./figures/f_plus_g_inline}}
+% \]
+% \end{remark}
 
@@
 
-The enhanced definition of extended DPOI ensures that the combinatorial representation of semilattice-enriched SMCs is sound and complete, in an appropriate sense.
+The enhanced definition of extended DPOI ensures that the combinatorial representation of closed semilattice-enriched SMCs is sound and complete for rewriting of closed $\Sigma^{+}$-terms.

### 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
last section left for closed

\section{Soundness and completeness}

In this section we explore soundness and completeness of DPOI rewriting introduced in previous section with respect to rewriting of closed $\Sigma^{+}$-terms.
First we recall the rewriting of $\Sigma$-terms modulo SMC laws.

\begin{definition}
    We say that a $\Sigma$-term $f$ rewrites ($\leadsto$) to a $\Sigma$-term $g$ modulo SMC-laws via a rewrite rule $\langle l, r \rangle$ if they are representable as
    \[
    f = c_{1};(id_{k} \otimes l);c_{2} \qquad g = c_{2};(id_{k} \otimes r);c_{2}~.
    \]
    These representations are so-called SMC normal forms.
\end{definition}

This notion generalises to rewriting of closed $\Sigma$-terms by introducing contexts (in string diagrammatic notation).

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/appendix.tex
last section left for closed

@@
 \]
-This implies the existence of a diagram as in~\ref{def:dpoi-e}.
-We will proceed by induction on the level of the occurrence of $\mathcal{L}$ within $\mathcal{F}$.
-\begin{itemize}
-	\item Suppose the image of $\mathcal{L}$ in $\mathcal{F}$ is top-level.
-		  Then, by lemma~\ref{lemma:decomposition}, there exists a decomposition of $\llbracket f \rrbracket$ in terms of MDA cospans
+
+\begin{figure*}
+    \begin{subfigure}{\linewidth}
 \begin{align*}
@@
 \end{align*}
-		By fullness of $\llbracket - \rrbracket$ (from Theorem~\ref{thm:completeness_simple}) there exist $c_{1}$ and $c_{2}$ such that 
-		\begin{align*}
-		&\llbracket c_{1} \rrbracket = n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k\\
-        &\llbracket c_{2} \rrbracket = j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m
-        \end{align*}
-		By functoriality, $\llbracket f \rrbracket = \llbracket c_{1};(id \otimes l);c_{2} \rrbracket$ and by faithfulness we have that $f = c_{1};(id \otimes l);c_{2}$ to which we can apply $\langle l, r \rangle$ and get $g' = c_{1};(id \otimes r);c_{2}$.
-		We then have $\llbracket g' \rrbracket = \llbracket c_{1};(id \otimes r);c_{2} \rrbracket =  \llbracket g \rrbracket$ since the boundary complement is unique (Proposition ~\ref{prop:boundary_unique}) and $g' = g$ by faithfulness.
-
-		\item If the image of $\mathcal{L}$ is not top-level, then either $\llbracket f \rrbracket$ is decomposable as
+\subcaption{\;}
+\end{subfigure}
+\vspace{1ex}
+\begin{subfigure}{\linewidth}
     \begin{align*}
@@
     \end{align*}
+\subcaption{\;}
+\end{subfigure}
+\caption{Decomposition}
+\label{fig:decomposition_1}
+\end{figure*}
+
+This implies the existence of a diagram as in~\ref{def:dpoi-e}.
+We will proceed by induction on the level of the occurrence of $\mathcal{L}$ within $\mathcal{F}$.
+\begin{itemize}
+	\item Suppose the image of $\mathcal{L}$ in $\mathcal{F}$ is top-level.
+		  Then, by lemma~\ref{lemma:decomposition}, there exists a decomposition of $\llbracket f \rrbracket$ in terms of MDA cospans as in Figure~\ref{fig:decomposition_1}.
+
+		% \begin{align*}
+		% 	(k \xrightarrow{id_{k}} k \xrightarrow{id_{k}} &k \xleftarrow{id_{k}} k \xleftarrow{id_{k}} k)\\
+		% 	n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k; \qquad\qquad\qquad &\otimes \qquad\qquad\qquad ; j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m\\
+		% 	(i \xrightarrow{} i' \xrightarrow{} &\mathcal{L} \xleftarrow{} j' \xleftarrow{} j)
+		% \end{align*}
+		By fullness of $\llbracket - \rrbracket$ (from Theorem~\ref{thm:completeness_simple}) there exist $c_{1}$ and $c_{2}$ such that 
+		\begin{align*}
+		&\llbracket c_{1} \rrbracket = n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k\\
+        &\llbracket c_{2} \rrbracket = j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m
+        \end{align*}
+		By functoriality, $\llbracket f \rrbracket = \llbracket c_{1};(id \otimes l);c_{2} \rrbracket$ and by faithfulness we have that $f = c_{1};(id \otimes l);c_{2}$ to which we can apply $\langle l, r \rangle$ and get $g' = c_{1};(id \otimes r);c_{2}$.
+		We then have $\llbracket g' \rrbracket = \llbracket c_{1};(id \otimes r);c_{2} \rrbracket =  \llbracket g \rrbracket$ since the boundary complement is unique (Proposition ~\ref{prop:boundary_unique}) and $g' = g$ by faithfulness.
+
+		\item If the image of $\mathcal{L}$ is not top-level, then either $\llbracket f \rrbracket$ is decomposable as in Figure~\ref{fig:decomposition_2}
+		% \begin{align*}
+		% 	(k \xrightarrow{id_{k}} k \xrightarrow{id_{k}} &k \xleftarrow{id_{k}} k \xleftarrow{id_{k}} k)\\
+		% 	n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k; \qquad\qquad\qquad &\otimes \qquad\qquad\qquad ; j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m\\
+		% 	(i \xrightarrow{} i' \xrightarrow{} &\mathcal{F'} \xleftarrow{} j' \xleftarrow{} j)
+		% \end{align*}
 		such that $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $F'$.
@@
 		By inductive hypothesis we have that $\llbracket f_{p} \rrbracket = n_p \xrightarrow{} n_{p}' \xrightarrow{} \mathcal{F}_{p} \xleftarrow{} m_{p}' \xleftarrow{} m_{p}$ rewrites into $\llbracket g_{p} \rrbracket = n_p \xrightarrow{} n_{p}'' \xrightarrow{} \mathcal{G}_{p} \xleftarrow{} m_{p}'' \xleftarrow{} m_{p}$ such that $f_{p} = g_{p}$.
-		By fullness and the fact that $\llbracket - \rrbracket$ is enriched, there exist $\llbracket f_1 + \ldots + f_{p} + \ldots + f_{k} \rrbracket = \llbracket f_{1} \rrbracket + \ldots + \llbracket f_{p} \rrbracket + \ldots + \llbracket f_{k} \rrbracket = \llbracket f \rrbracket$.
+		By fullness and the fact that $\llbracket - \rrbracket$ is $\catname{SLat}$-enriched, there exist $\llbracket f_1 + \ldots + f_{p} + \ldots + f_{k} \rrbracket = \llbracket f_{1} \rrbracket + \ldots + \llbracket f_{p} \rrbracket + \ldots + \llbracket f_{k} \rrbracket = \llbracket f \rrbracket$.
 		By applying $f_{p} = g_{p}$ we get $f = g''$ and $\llbracket g'' \rrbracket = \llbracket f_{1} + \ldots + g_{p} + \ldots f_{k} \rrbracket = \llbracket g \rrbracket$ and $f = g'' = g$.
@@
     \begin{tikzcd}
-        n \arrow[r] & n'_{1} \arrow[r] & \mathcal{C}_{1} & m'_{1} + k \arrow[l]     & i+k \arrow[l] \arrow[r] & i'+k \arrow[rd] &             \\
+        n \arrow[r] & n'_{1} \arrow[r] & \mathcal{C}_{1} & m'_{1} + i + k \arrow[l]     & i+k \arrow[l] \arrow[r] & i'+k \arrow[rd] &             \\
                     &                  &                 &                          &                         &                 & \mathcal{L} \\

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
wip

@@
-\section{$\catname{SLat}$}
+\subsection{$\catname{SLat}$}
 \label{sec:appendix:1}
@@
     \end{proof}
+
+\subsection{Pushout in $\catname{EHyp}(\Sigma)$}
+\label{sec:appendix:pushout}
+
+In this section we will describe the construction of the pushout in $\catname{EHyp}(\Sigma)$.
+TODO
+
+\subsection{Interpretation of $\Sigma$-terms as cospans of hypergraphs}
+\label{sec:appendix:interpretation}
+\begin{figure*}
+  \[
+  \adjustbox{width=\textwidth}{
+  \tikzfig{./figures/interpretation}
+  }
+  \]
+  \caption{Base cases for $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$}
+  \label{fig:base_cases}
+\end{figure*}
+
+The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Figure~\ref{fig:base_cases}.
+Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/soundness.tex

@@
 \begin{definition}
+\label{def:rewrite}
     We say that a $\Sigma$-term $f$ rewrites ($\leadsto$) to a $\Sigma$-term $g$ modulo SMC-laws via a rewrite rule $\langle l, r \rangle$ if they are representable as
@@
 
-This notion generalises to rewriting of closed $\Sigma$-terms by introducing contexts (in string diagrammatic notation).
+This notion generalises to rewriting of closed $\Sigma$-terms
+
+\begin{definition}
+    We say that a closed-$\Sigma$-term $f$ rewrites ($\leadsto$) to a closed-$\Sigma$-term $g$ modulo SMC-laws via a rewrite rule $\langle l, r \rangle$ if they are representable as either in Definition~\ref{def:rewrite}, or as
+    \[
+    f = c_{1};(id_{k} \otimes d);c_{2} \qquad g = c_{2};(id_{k} \otimes e);c_{2}
+    \]
+    and $d \leadsto_{\langle l, r \rangle} e$
+\end{definition}
+
+\begin{lemma}
+
+    Every closed $\Sigma^{+}$ term $f$ can be equivalently represented as
+    \[
+    f_{1} + \ldots + f_{n}
+    \] where none of $f_{i}$ contain the join operator.
+\end{lemma}
+\begin{proof}
+Such forms are given by orienting $\catname{SLat}$-equations such that $+$ gets propagated to the top, as, for example $f \otimes (g + h) = f \otimes g + g \otimes h$.
+Most notably, we can also escape lambda abstractions by using distributivity law~\ref{law:distributivity}.
+\end{proof}
+
+This takes us to the following definition of rewriting for closed $\Sigma^{+}$-terms.
+\begin{definition}
+    We say that a closed-$\Sigma^{+}$-term $f$ rewrites ($\leadsto$) to a closed-$\Sigma^{+}$-term $g$ modulo SMC-laws via a rewrite rule $\langle l, r \rangle$ if they are representable as
+    \[
+    f = f_{1} + \ldots + f_{n} \qquad g = f_{1} + \ldots + f_{n}
+    \]
+    such that there is a permutation $\sigma$ such that there exist indices $i,j$ and $\sigma(f_{i}) \leadsto f_{j}$.
+\end{definition}
+
+To formulate the correspondence between rewriting systems we first define interpretation of closed $\Sigma^{+}$ terms in $\WellTypedMdaEcospans$.
+It follows by first defining the interpretation of generators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in Appendix~\ref{sec:appendix:interpretation}.
+The only missing cases are $[\textsf{ev}_{A,B}]$ and $[\Lambda_{A,B,C}]$.
+
+\begin{figure}
+    \begin{subfigure}{0.45\linewidth}
+\[
+\adjustbox{scale=0.5}{
+    \tikzfig{./figures/ev_interpretation}
+}
+\]
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\linewidth}
+        \[
+        \adjustbox{scale=0.5}{
+            \tikzfig{./figures/lambda_interpretation}
+        }
+        \]
+    \end{subfigure}
+
+\end{figure}
+
+Then we can make $\WellTypedMdaEcospans$ into closed $\catname{SLat}$ category by defining a join of two cospans, $\Lambda$ of a cospan, and a designated $\textsf{ev}$ cospan and introducing DPOI rewrite rules to make these construction satisfy the laws of closed $\catname{SLat}$ symmetric monoidal category.
+The join of two cospans is defined as in Figure~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules one for each axiom.
+For example, the distributivity rule for $\otimes$ is given by the following schema rule
+\[
+\adjustbox{width=\linewidth}{
+\tikzfig{./figures/semilattice_rule_1}~.
+}
+\]
+
+We collect all such rewrite schemas into a set $\mathcal{S}$ and then a category $\WellTypedMdaEcospans / \mathcal{S}$ is a closed $\catname{SLat}$ SMC.
+
+\begin{proposition}
+
+For two closed $\Sigma^{+}$-terms $f$ and $g$ and a $\mathcal{E}$-equation $l = r$  $f \leadsto g$ if and only if $\llbracket f \rrbracket \Rrightarrow{} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
+\end{proposition}
+\begin{proof}
+
+\end{proof}
+
+\begin{figure}
+\[
+        \adjustbox{scale=0.5}{
+            \tikzfig{./figures/f_plus_g_new}
+        }
+\]
+\caption{$+$ of two morphisms in $\WellTypedMdaEcospans$}
+\label{fig:f+g}
+\end{figure}

### 2025-01-24 LICS_2025/monoidal_e_graphs/sections/abstract.tex

@@
 % The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for satisfiability modulo theory solvers.
-The technique of equality saturation, which equips graphs with an equivalence relation, has proven effective for program optimisation.
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a category of semilattices.
+The technique of \emph{equality saturation}, which equips graphs with an equivalence relation, has proven effective for program optimisation.
+We give a categorical semantics to these structures, called \emph{e-graphs}, in terms of Cartesian categories enriched over the category of semilattices.
 This approach generalises to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
+Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call \emph{e-hypergraphs}.
 They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism.
-The advantage of this new principled approach to e-graphs is that for the first time it is possible to support double-pushout (DPO) rewriting for these structures which constitutes the main contribution of this paper.
+This new principled approach to e-graphs enables double-pushout (DPO) rewriting for these structures, which constitutes the main contribution of this paper.

### 2025-01-24 LICS_2025/monoidal_e_graphs/sections/conclusion.tex

@@
 In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
+Having described a translation of e-graphs into this categorical framework, we also provide a combinatorial representation of morphisms in the free category.
 As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.  
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
+We then showed our theory to be sound and complete with respect to our categorical semantics.
 

### 2025-01-24 LICS_2025/monoidal_e_graphs/sections/e-hypergraphs.tex

@@
     \captionsetup{belowskip=-3ex, skip=0pt}
-    \caption{Extended cospan example}
+    \caption{Morphism of $\Ecospans$}
     \label{fig:e-cospan-example}
@@
 \end{definition}
-Intuitively, such construction suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter; and that when considering individual components of hierarchical edges with their respective portions of strict internal input and output interfaces as plain cospans of hypergraphs (by only considering top-level vertices and edges of these components) they should be isomorphic cospans of hypergraphs in the sense from~\cite{bonchi_string_2022-1}.
 An example of an extended cospan can be seen in Figure~\ref{fig:e-cospan-example}.
@@
 The labels suggest how $f_{int}, f_{ext}$ (respectively, $g_{int}, g_{ext}$) act on vertices, with the ordering of vertices in the interfaces given left to right.
-Examples of isomorphic and non-isomorphic extended cospans can be found in Appendix~\ref{sec:appendix:iso}.
+Intuitively, such notion of isomorphism suggests that the position of the image of $n$ within $n'$ relative to the strict internal interface does not matter, for example, vertices
+\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=red node, label={above:$v_{3}$}] (0) at (-0.5, 3.5) {};
+		\node [style=red node, label={above:$v_4$}] (1) at (0, 3.5) {};
+	\end{pgfonlayer}
+\end{tikzpicture}
+can be swapped with vertices
+\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=blue node, label={above:$v_{5}$}] (0) at (-0.5, 3.5) {};
+		\node [style=blue node, label={above:$v_6$}] (1) at (0, 3.5) {};
+	\end{pgfonlayer}
+\end{tikzpicture}
+in the input interface yielding an isomorphic cospan.
+However, we may not swap
+\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=red node, label={above:$v_{3}$}] (0) at (-0.5, 3.5) {};
+		\node [style=red node, label={above:$v_4$}] (1) at (0, 3.5) {};
+	\end{pgfonlayer}
+\end{tikzpicture}
+with each other --- this would yield a non-isomorphic cospan.
+Further examples of isomorphic and non-isomorphic extended cospans can be found in Appendix~\ref{sec:appendix:iso}.
 

### 2025-01-24 LICS_2025/monoidal_e_graphs/sections/introduction.tex

@@
 This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an optimal ordering.
+The optimal solution to this problem is often intractably hard to compute, so typical approaches in practice use heuristics to determine a pseudooptimal ordering instead.
 
 A recently proposed alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively}, \textit{i.e.}, forgetting the part of the program that is replaced by the rewrite, at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
+\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} (\textit{i.e.}, at each step, forgetting the part of the program that is replaced by the rewrite), a \textit{set} of equivalent programs is maintained instead, where each rewrite step \textit{non-destructively} grows the set.
 Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
@@
 Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations and the foundations of e-graphs in particular.
-We propose to address this gap via a categorical axiomatisation for e-graphs and their rewriting.
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and morphisms of Cartesian categories enriched over a category of semilattices (\textit{semilattice-enriched}).
+We address this by giving a categorical axiomatisation for e-graphs and their rewriting.
+Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and morphisms of Cartesian categories enriched over the category of semilattices (\textit{semilattice-enriched}).
 % The morphisms of the latter can also be depicted using \emph{string diagrams}~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}: a two-dimensional syntax that encodes equivalence classes of morphism terms.
 % Enriched categories~\cite{Borceux_1994,Kelly2022BASICCO} generalise plain categories by imposing more structure making the categorical language more expressive~\cite{Borceux_1994,Kelly2022BASICCO} which will be useful for modelling equivalence between terms as encountered in e-graphs.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3,fscd,ghica-zanassi2023string}.
-The key ingredient of our categorical approach is the semilattice-enrichment.
-By the latter we will mean an idempotent commutative semigroup given by a binary operator but will stick to the word semilattice as it is shorter.
+Our approach is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3,fscd,ghica-zanassi2023string}.
+The key ingredient of our categorical approach is the semilattice-enrichment\footnote{We could alternatively use the language of idempotent commutative semigroups, which form an equivalent category, but stick to the \enquote{semilattice} terminology for brevity.}.
 By using the binary operator of a given semilattice we can operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
 
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that are induced by monoidal rather than algebraic theories with a defined notion of rewriting and thus these applications will benefit from the mentioned optimisation technique.
-% now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica-zanassi2023string}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.
-% However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
+\emph{Monoidal} category theory has seen applications in digital~\cite{ghica_compositional_2023} and quantum circuits~\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica-zanassi2023string}, and computational linguistics~\cite{wazni_quantum_2022,coecke_lambek_2013}.
+These topics all have \emph{computation} at their core, where a typical goal is to transform one expression into a \enquote{better} version of itself, with respect to some sound theory of rewriting, analogously to program optimisation.
+Our work does not presuppose any particular monoidal theory, but rather provides a general categorical framework, which we hope will serve as a foundation for a principled approach to efficient rewriting theories in these settings.
 
-A final contribution of this work is giving a concrete combinatorial representation of string diagrams for semilattice-enriched categories in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs} or monoidal e-graphs, as the former are a generalisation of e-graphs for monoidal theories.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework 
+Finally, we give a concrete combinatorial representation of string diagrams for semilattice-enriched categories in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs} or monoidal e-graphs, generalising e-graphs for monoidal theories.
+Additionally, we also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework 
 \cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3},  proving the representation \textit{sound and complete} with respect to our categorical semantics.  
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature~\cite{EggPaper}.
+As a corollary, this also provides a formalisation of the rewriting theory of e-graphs~\cite{EggPaper}.
 
@@
 
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.
+E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory simultaneously.
 They generalise the typical \emph{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (subgraphs).
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example illustrates how an optimiser using e-graphs may non-destructively apply a set of rewrites to initial term $(a * 2) / 2$, presented first in~\cite{EggPaper}.
+The key concepts of e-graphs can be intuitively grasped from some simple examples, as shown in Figure~\ref{fig:e-graph-example}.
+This example, adapted from~\cite{EggPaper}, illustrates how an optimiser using e-graphs may non-destructively apply a set of rewrites to initial term $(a * 2) / 2$.
 
@@
 If we were to apply another rule, $x * 1 \to x$, we would identify the e-class (dashed box) containing $a$ with the e-class containing $a * 1$ (the topmost dashed box) which would require a cycle to be introduced: the left argument of node $*$ would be the same e-class that $*$ belongs to.
-Then, a deemed optimal (as its size as a graph is minimal) term $a$ could be extracted from the e-graph by traversing all the e-classes.
+Then, an optimal term $a$ could be extracted from the e-graph by traversing all the e-classes, where its optimality is justified by its minimal size as a graph.
 To support such cycles we need a richer categorical structure for our string diagrams.
-Such categorical structure may be added by combining the results of this work with the results of~\cite{ghica_rewriting_2023} and for the purposes of presentation we will limit ourselves to the case of acyclic e-graphs.
+This is obtained by combining this work with the results of~\cite{ghica_rewriting_2023}, and for simplicity we will limit ourselves to the case of acyclic e-graphs, although this is not a limitation of the general theory.
 % While equipping the string diagrams which such structure is straightforward, it is out of the scope of this work, and thus we limit ourselves to acyclic e-graphs and e-string diagrams for the purposes of presentation.
-Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the string diagrammatic version is guided by the equations of the underlying theory.
-A detailed step-by-step application of the equations for the rewrite rules from $(a)$ to $(b)$ and $(b)$ to $(c)$ is given in Figures~\ref{fig:e-graph-example-a-b} and~\ref{fig:e-graph-example-b-c}, respectively, in the Appendix.
+Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the string diagrammatic version are guided by the equations of the underlying theory.
+A detailed step-by-step application of the equations for the rewrite rules from $(a)$ to $(b)$ and $(b)$ to $(c)$ is given in Figures~\ref{fig:e-graph-example-a-b} and~\ref{fig:e-graph-example-b-c}, respectively.
 The equations applied are the ones from Figure~\ref{fig:string-equations} along with the naturality of delete and copy.
-One can note that the size of the eventual string diagram is comparable to the size of the corresponding e-graph.
+Note that the size of the eventual string diagram is comparable to the size of the corresponding e-graph.
 % Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
@@
 It makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a semilattice-enriched Cartesian category, taken as string diagrams.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
+We extend the usual vocabulary of string diagrams for symmetric monoidal categories with an additional generator:
 \[
@@
 \]
-Taking formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
+Taking formal (non-empty) joins of morphisms in this way is used to model the equivalence class structure of e-graphs.
 
-String diagrams are read from bottom-to-top, and we implicitly consider additional generators for the duplication and deletion transformations of a (semilattice-enriched) Cartesian category.
-Note, further, that in the Cartesian case we can restrict generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
+String diagrams are read from bottom-to-top, and we implicitly consider additional generators for the duplication and deletion transformations of a (semilattice-enriched) Cartesian category, and we call the string diagrams for this enriched setting \emph{enriched string diagrams}.
+Note that in the Cartesian case we can restrict generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.  
 This means the (informal) translation given next is fully general for acyclic e-graphs.
@@
 
-The motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing as could be seen in Figure~\ref{fig:e-graph-example}.
+The motivating example of our work, e-graphs, requires the Cartesian structure to express arbitrary sharing (see Figure~\ref{fig:e-graph-example}).
 However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
@@
 
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
+We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations, such as the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
 These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.  
@@
 However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs} --- hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
+The technical aspects of finding the correct definitions are complex, and constitutes the main body of this paper.
 
 \subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation with respect to morphisms of a semilattice-enriched symmetric monoidal categories and their respective rewriting.
+The main technical result of the paper is a proof of soundness and completeness of this representation with respect to morphisms of a semilattice-enriched symmetric monoidal categories and their respective rewriting theory.
 This extends the results of \cite{bonchi_string_2022-2} for plain SMCs. 
@@
 We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.  
-Some further avenues for investigation, including potential applications, are given in the \nameref{sec:conclusion}.
+We elaborate on some potential applications in the \nameref{sec:conclusion}.
 
@@
 \section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
+In this section,  we introduce some preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
 
@@
 
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
+A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature which are identified.
 We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
@@
 \]
-together with the following associativity, commutativity and unitality equations, given in terms of equations of string diagrams. 
+together with the following associativity, commutativity and unitality equations, given in terms of equations of string diagrams: 
 \[
@@
 \]
-A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations $\mathcal{E}$ that include the SMT of commutative comonoids, and the following additional naturality of copy-delete equations, for every $c \in \Sigma_C$
+A Cartesian SMT is given by a set of generators $\Sigma_C$ which additionally contains the generators of commutative comonoids, together with the equations $\mathcal{E}$ that include the SMT of commutative comonoids, and the following additional naturality of copy-delete equations; for every $c \in \Sigma_C$:
 \[
@@
 % This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$, but we will omit further details of the general case of enrichment here, and instead work with the concrete definition required for our application, where $M = \catname{SLat}$.
+Generally,  we may define a category \textit{enriched} in any other (closed) monoidal category $M$, but we restrict ourselves to the concrete definition required for our application, where $M = \catname{SLat}$.
 %  (\textit{e.g.},  $\catname{SLat}$).
@@
 
-Note, in particular,  that we do not require semilattices to have a unit for $+$, \textit{i.e.,} we work with \textit{unbounded} semilattices.
+In particular,  that we do not require semilattices to have a unit for $+$, \textit{i.e.,} we work with \textit{unbounded} semilattices.
 % We also will not utilize the ordering that is induced by the operator of a semilattice in any meaningful way, so a semilattice can be thought of as a commutative idempotent semigroup throughout the paper.
@@
 
-\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also Appendix section~\ref{sec:appendix:slat})
+\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also Appendix~\ref{sec:appendix:slat})
 There is a 2-adjunction 
@@
 To aid reasoning, we introduce a new language of string diagrams for $\catname{SLat}$-SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
+This is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
 

### 2025-01-24 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
wip
sections done

@@
 \newtheorem{remark}[theorem]{Remark}
+\newtheorem{lemma}[theorem]{Lemma}
 
@@
 \begin{abstract}
-    The technique of equality saturation for program optimisation and reasoning has recently gained significant attention due to the renewed popularity of equality graphs (e-graphs).
-    E-graphs represent multiple terms that are equivalent modulo a given set of rewrite rules. 
-    This capability allows rewriting to be performed over a family of terms simultaneously.    
-    However, there are domains where e-graphs fail to achieve optimal performance. 
-	Notably, e-graphs struggle in domains involving variable binding, such as $\lambda$-calculus, because they lack native support for binding.
-    Building on recent work that provided a categorical interpretation of e-graphs as morphisms in a semilattice-enriched symmetric monoidal category, we extend this framework. 
-	Specifically, we show how e-graphs with bindings can be derived as morphisms in a closed symmetric monoidal category with similar enrichment. 
-	This extension provides a principled approach to equality saturation for $\lambda$-calculus.
-    We then define a concrete combinatorial representation of these morphisms using hierarchical hypergraphs. 
-	Additionally, we introduce a corresponding notion of double-pushout (DPO) rewriting. 
-	Finally, we demonstrate the equivalence between term rewriting and DPO rewriting, with the key property that DPO rewriting absorbs the equations of the symmetric monoidal category.
+    Equality saturation, a technique for program optimisation and reasoning, has gained attention due to the resurgence of equality graphs (e-graphs). 
+	E-graphs represent equivalence classes of terms under rewrite rules, enabling simultaneous rewriting across a family of terms. 
+	However, they struggle in domains like $\lambda$-calculus that involve variable binding, due to a lack of native support for bindings.
+	Building on recent work interpreting e-graphs categorically as morphisms in semilattice-enriched symmetric monoidal categories, we extend this framework to closed symmetric monoidal categories to handle bindings.
+	We provide a concrete combinatorial representation using hierarchical hypergraphs and introduce a corresponding double-pushout (DPO) rewriting mechanism. 
+	Finally, we establish the equivalence of term rewriting and DPO rewriting, with the key property that the combinatorial representation absorbs the equations of the symmetric monoidal category.
 \end{abstract}
-
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert.
-\end{IEEEkeywords}
 
@@
 As described above, string diagrams are commonly an informal yet rigorous two-dimensional syntax for the internal language of a category. 
-They can be formalised by representing them as graphs or hypergraphs~\cite{1011453502719}, which has two further benefits. 
+They can be formalised by representing them as graphs or hypergraphs~\cite{bonchi_string_2022-1}, which has two further benefits. 
 The first benefit is that the property of two string diagrams being equal if and only if they ``look the same'' can be spelled out as graph isomorphism, eliminating any possible ambiguity of interpretation.
@@
 E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph as first presented in ~\cite{EggPaper}.
+The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph as first presented in~\cite{EggPaper}.
 
-\begin{figure*}
+\begin{figure*}[h!]
 \[
-\adjustbox{scale=0.7}{
+\adjustbox{scale=0.6}{
 \tikzfig{figures/e-graph-example}
@@
 \begin{figure*}
+	\begin{subfigure}{\linewidth}
     \[
@@
     \]
-    \caption{E-graph explicit substitution example.}
+	% \captionsetup{skip=0pt}
+    \subcaption{E-graph explicit substitution example.}
     \label{fig:e-graph-substitution}
-  \end{figure*}
-
-  \begin{figure*}
+	\end{subfigure}
+	\begin{subfigure}{\linewidth}
 	\[
@@
     \]
-    \caption{String diagrammatic substitution example.}
-    \label{fig:e-string-substitution}
+	\subcaption{String diagrammatic substitution example.}
+	\end{subfigure}
 \end{figure*}
+  %   \begin{figure*}
+%     \[
+%         % \hspace{1.3cm}
+%         \adjustbox{width=\textwidth}{
+%         \tikzfig{figures/e-string-substitution-2}
+%         }
+%     \]
+% 	\captionsetup{skip=0pt}
+%     \caption{String diagrammatic substitution example.}
+%     \label{fig:e-string-substitution}
+%   \end{figure*}
 
@@
 
-DRG: LIST THE MAIN DEFINITIONS AND THEOREMS AND WHAT THEY MEAN
+In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over some signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
+We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
+That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
+We call these string diagrams, or more precisely, the combinatorial representation of them, e-graphs with bindings.
+
+We then define a combinatorial representation for such string diagrams in terms of hierarchical hypergraphs with a suitable notion of DPO rewriting.
+Finally, we show that this notion of rewriting is sound and complete with respect to rewriting of morphism terms.
+
+% DRG: LIST THE MAIN DEFINITIONS AND THEOREMS AND WHAT THEY MEAN
 

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
sections done

@@
 \subsection{$\catname{SLat}$}
-\label{sec:appendix:1}
+\label{sec:appendix:slat}
 
@@
 
-In this section we will describe the construction of the pushout in $\catname{EHyp}(\Sigma)$.
-TODO
+In constructing the pushout the functional versions of e-hypergraph relations will be useful.
+\begin{remark}
+    $\textcolor{red}{<}$, $\textcolor{blue}{<}$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
+    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
+    For example, an immediate successor of a vertex $x$ can be written functionally as $\textcolor{red}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))$.
+\end{remark}
+
+\begin{remark}
+    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
+\end{remark}
+\begin{definition}
+        \label{def:e-homo-2}    
+        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+
+        \begin{enumerate}
+            \item $\phi$ is hypergraph homomorphism.
+
+            \item When $x$ is not a top-level vertex such that $\textcolor{red}{<}(\iota_{V}(x))$ is defined, 
+                \[
+                \phi_{E}(\textcolor{red}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))) = \textcolor{red}{<_{\mathcal{G}}^{\mu}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
+                \]
+                and
+                \[
+                \phi_{E}(\textcolor{red}{<_{\mathcal{F}}^{\mu}}(\iota_{E_{\mathcal{F}}}(x))) = \textcolor{red}{<_{\mathcal{G}}^{\mu}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
+                \] when $x$ is a not top-level edge.
+                \item When $x$ is not a top-level vertex such that $\textcolor{blue}{<}(\iota_{V}(x))$ is defined, 
+                \[
+                \phi_{E}(\textcolor{blue}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))) = \textcolor{blue}{<_{\mathcal{G}}^{\mu}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
+                \]
+                and
+                \[
+                \phi_{E}(\textcolor{blue}{<_{\mathcal{F}}^{\mu}}(\iota_{E_{\mathcal{F}}}(x))) = \textcolor{blue}{<_{\mathcal{G}}^{\mu}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
+                \] when $x$ is a not top-level edge.
+                \item
+            When $x \in E_{\mathcal{F}}$
+            \[
+                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
+                \subseteq
+                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
+            \]
+            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+            \item When $x \in V_{\mathcal{F}}$
+            \[
+                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
+                \subseteq
+                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
+            \]
+            \end{enumerate}
+\end{definition}
+
+\begin{theorem}[Existence of pushouts in $\catname{EHyp}(\Sigma)$]
+\label{th:existence_of_pushouts}
+Consider the following span in $\catname{EHyp}(\Sigma)$
+% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
+\[\begin{tikzcd}
+	Z && X \\
+	\\
+	Y
+	\arrow["f", from=1-1, to=1-3]
+	\arrow["g"', from=1-1, to=3-1]
+\end{tikzcd}\]
+such that
+\begin{enumerate}
+\label{pushout:assumptions}
+    \item $Z$ is a \textit{discrete} e-hypergraph.
+    \item \label{assumption:equal_predecessors} $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
+    \item \label{assumption:non_ambiguous_predecessors} If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
+    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
+\end{enumerate}    
+then the pushout $X +_{f,g} Y$ exists.
+\end{theorem}
+\begin{proof}
+
+We next explicitly construct a pushout.
+Consider the diagram below.
+    \[
+        \adjustbox{scale=1.25}{
+            \begin{tikzcd}
+            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
+            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+                                                                            &                                                                  & Q
+            \end{tikzcd}}
+    \]
+    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
+    First, a coproduct of $X+Y$ is computed which is 
+    \[
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, \textcolor{red}{<_{X+Y}}, \textcolor{blue}{<_{X+Y}}\}
+    \]
+    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, \textcolor{red}{<_{X+Y}}, \textcolor{blue}{<_{X+Y}}, \consistency_{X+Y}$.
+    We will omit labels as they are irrelevant to pushout construction.
+    Consider relations
+    \begin{align*}    
+    S_{V} &= \{\\
+          &\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
+          &\}\\
+          &\cup\\
+          &\{\\
+          &\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
+          &\}\\
+          &\cup\\
+          &\{\\
+          &\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\\
+          &\}\\
+        \\
+    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
+    \end{align*}
+and let relations $R_{V},R_{E}$ be their transitive closures respectively.
+
+We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
+
+\begin{align*}
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = (&\\
+            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
+            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})},\\
+            &\consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},\\
+            &\textcolor{red}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+            &\textcolor{blue}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+        &)    
+\end{align*}
+
+We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
+Where necessary we will refer to $S_{V}$ and $S_{R}$ as $\sim_{S}$.
+We have 
+\[
+    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
+\]
+and there is an obvious surjective function $[]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
+There is also $[] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
+We then define 
+\[
+    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[]^{*}(e) = [s_{X+Y}(e)]^{*}
+\]
+We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
+Similarly, we will write $v_{x}$ to refer to a vertex with a pre-image in $V_{X}$.
+Likewise, 
+\[
+    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
+\]
+
+These definitions make $([]_{V},[]_{E})$ automatically a homomorphism with respect to source and target maps.
+These maps are also automatically well-defined functions since $[e_1] = [e_2]$ implies $e_1 = e_2$ since the mappings for edges are mono.
+
+Recall that $\textcolor{red}{<}$ is essentially a transitive closure of $\textcolor{red}{<^{\mu}}$ and hence we can first define 
+    \[\textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
+    \] and 
+    \[
+        \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u])
+    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+    We will consider several cases.
+    First, assume that $u$ is an element of $V_{X+Y}$ and
+    \begin{enumerate}
+        \item If there exists $v$ such that $\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))$ is defined and $u \sim v$
+              we let
+              \[
+                \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = [\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))]
+              \]
+        \item \label{def:child_respects_connectivity} $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
+              If there exists $v$ such that $\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
+              \[ 
+                \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+              \]
+    \end{enumerate}
+    % Note that the existence of a path in the definition (2) implies that there exists $v' \sim v$ such that there is a path from $u$ to $v'$ and furthermore $v \not = v'$.
+    % If $v' = v$ then it would mean that $[u) \not = \varnothing$ since $[v) \not = \varnothing$ and there is a path from $u$ to $v$.
+    % There is no case when $u$ has a pre-image in $V_{Z}$ and yet $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e$ and there is a path from $[u]$ to $[v]$.
+    % The existence of a path would mean that there is $u' \sim u$ and $v' \sim v$ such that there is a path from $u'$ to $v'$ and since $u$ has a pre-image in $V_{Z}$, $u = f_{V};\iota_{1,V}(z_1)$.
+    % $v' \sim v$ means that both $v'$ and $v$ have a pre-image in $V_{Z}$ (unless $v' = v$ in which case $[u') \not = \varnothing$ as there is a path from $u'$ to $v'$) and since $u = f_{V};\iota_{1,V}(z_1)$ and $[u) = \varnothing$ $f_{V};\iota_{1,V}(z) = \varnothing$ for all $z$.
+    % Since $[v) \not = \varnothing$ and $v' \sim v$, $v$ is necessarily in the image of $g_{V};\iota_{2,V}$ and for all $z$ $[g_{V};\iota_{2,V}(z)) \not = \varnothing$ as well as for some $z'$ such that $u'' = g_{V};\iota_{2,V}(z')$ and $u'' \sim_{S} u$ which would mean that the case (1) is applicable to $u$.
+    % Similarly, when $u = g_{V};\iota_{2,V}(z)$.
+
+    Otherwise, we leave $\textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))$ undefined.
+    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
+    Now, assume that $e$ is an element of $E_{X+Y}$.
+    Consider two cases.
+    \begin{enumerate}
+    \item  $\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e)) = e'$.
+            Then we define
+        \[
+            \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = [\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e))]
+        \]
+    \item $[e) = \varnothing$ and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an undirected path from $[e]$ to $[v]$, we define 
+        \[
+            \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+        \] 
+    \end{enumerate}
+    Otherwise we leave $\textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{E_{X} + E_{Y}}{\sim}([e]))$ undefined.
+    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
+    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
+    Clearly, all the cases above are disjoint.
+    We similarly define $\textcolor{blue}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
+
+    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
+    We can consider the consistency relation from the coproduct as a function 
+    \[
+        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
+    \]
+    Quotienting the values of the function gives us 
+    \[
+        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
+     \]
+    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
+    We then first define an auxiliary relation $\consistency^{\hashtag}$ similarly to how $<^{\mu}_{\sfrac{X+Y}{\sim}}$ was defined.
+We begin with defining $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ for vertices.
+
+    \begin{enumerate}
+        \item If there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $u \sim v$, we let
+              \ifdefined \ONECOLUMN
+              \[
+                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))
+                =
+                [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+              \]
+              \else
+              \begin{align*}
+                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
+                &=\\
+                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+            \end{align*}
+            \fi
+        \item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
+        Then we let
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+        \]
+    \end{enumerate}
+
+    Next we define $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}$ for edges.
+
+    \begin{enumerate}
+        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
+                \begin{align*}
+                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
+                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
+                \end{align*}
+        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[e]$ to $[v]$
+        then
+        \[
+            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]))
+        \]
+    \end{enumerate}
+
+    The well-definedness of this construction follows by the same argument as the well-definedness of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
+    Then we define
+    \ifdefined \ONECOLUMN
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \qquad \text{and} \qquad \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
+    \]
+    \else
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+    \]
+    and
+    \[
+        \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
+    \]
+    \fi
+
+    as closures of $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ as below.
+    \[
+      (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}
+    \]
+    where $c$ denotes a closure and $\iota_{\sfrac{*}{\sim}}$ is $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}$ depending on whether $[x_i]$ comes from $V_{X} + V_{Y}$ or $E_{X} + E_{Y}$, is the smallest set such that
+    \begin{itemize}
+        \item $[x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$
+        \item if $[y] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x]))$ then 
+        \[
+            [x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([y])))^{c}
+        \]
+        % \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
+        % \[
+        %     [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
+        % \]
+        \item for any sequence $([x_1], \ldots, [x_n])$ such that
+        \ifdefined \ONECOLUMN
+        \[[x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}])) \qquad \text{or} \qquad [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))\]
+        \else
+        \[
+            [x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}]))
+        \] or 
+        \[
+            [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))
+        \]
+        \fi
+         for $i < n$ both
+        \ifdefined \ONECOLUMN
+        \[
+            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c} \qquad \text{and} \qquad [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
+        \]
+        \else
+         \[
+            [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c}
+        \]
+        and
+        \[
+            [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
+        \]
+        \fi
+    \end{itemize}
+
+    This construction is analogous to how a pushout was constructed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and we refer the reader for more detailed proofs there.
+    The only addition is $\textcolor{blue}{<_{\sfrac{X+Y}{\sim}}}^{\mu}$ that is constructed analogously to $\textcolor{red}{<_{\sfrac{X+Y}{\sim}}}^{\mu}$.
+\end{proof}
+The construction can be illustrated with the example in Figure~\ref{fig:pushout_example}.
+Vertices and edges that are being identified as they have a common preimage inherit the relations from the vertex or edge that has it defined either directly or via a path.
+In particular, a coproduct of two feet contains\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node, label={above:$v_1$}] (0) at (-5.5, 3.5) {};
+		\node [style=node, label={above:$v_2$}] (1) at (-4.5, 3.5) {};
+		\node [style=node, label={above:$v_3$}] (2) at (-3.5, 3.5) {};
+		\node [style=node, label={above:$v'_1$}] (3) at (-2.5, 3.5) {};
+		\node [style=node, label={above:$v'_2$}] (4) at (-1.5, 3.5) {};
+		\node [style=node, label={above:$v'_3$}] (5) at (-0.5, 3.5) {};
+	\end{pgfonlayer}
+\end{tikzpicture}
+such that $\textcolor{blue}{<_{X+Y}}$ is only defined for $v_1,v_2,v_3$.
+Therefore, we assign $\textcolor{blue}{<_{\sfrac{X+Y}{\sim}}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [\textcolor{blue}{<_{X+Y}}(\iota_{V_{X} + V_{Y}}(v_1))]$ and so on.
+This motivates the requirement for conditions (2), (3), and (4) of Theorem~\ref{th:existence_of_pushouts}, otherwise such inheritance would be ambiguous.
+
+\begin{figure}[t]
+\[
+\adjustbox{scale=0.6}{
+  \tikzfig{./figures/pushout_example}
+}
+\]
+\caption{Pushout in $\catname{EHyp}(\Sigma)$}
+\label{fig:pushout_example}
+\end{figure}
 
@@
 Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.
+
+\subsection{DPOI rewriting}
+
+Consider an example of DPOI rewriting in Figure~\ref{fig:dpoi-example} as defined in Definition~\ref{def:dpoi-e}.
+The rule in the top half of the diagram is an instance of a $\beta$-reduction rule as defined for string diagrams~\cite{ghica2024stringdiagramslambdacalculifunctional}.
+Note that the match satisfies all the conditions of the boundary complement as all vertices from the interface have the same parent.
+First, to compute the boundary complement, we remove from the source e-hypergraph everything that has no pre-image in the interface above.
+The mda-cospan for the complement that will have .
+
+\begin{figure}
+\[
+\adjustbox{width=\linewidth}{
+\tikzfig{./figures/dpo_example}
+}
+\]
+\caption{$\beta$-rule DPOI example.}
+\label{fig:dpoi-example}
+\end{figure}

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
sections done

@@
 We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
+Given two adjoint functors $F : \mathbb{A} \to \mathbb{B}$ and $G : \mathbb{B} \to \mathbb{A}$ with $F$ being left-adjoint, we write $F \dashv G : \mathbb{B} \to \mathbb{A}$ for this adjunction.
 
@@
 To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator \textit{join} $+$ as presented in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix section~\ref{sec:appendix:slat}).
+This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix~\ref{sec:appendix:slat}).
 
 \begin{definition}[Semilattice-enriched category~\cite{Borceux_1994}]
+\label{def:slat-cat}
     A semilattice-enriched category $\mathbb{C}$ is defined by the following data:
@@
 \end{definition}
+As the composition is a semilattice homomorphism (morphism of $\catname{SLat}$), we have the following equations 
+\[
+(f + g); h = f;h + g;h \quad f;(g+h) = f;g + f;h
+\]
 
@@
 
+    Then, the axioms of $\catname{SLat}$-SMC consist of equations in Definitions~\ref{def:enriched-prop}~\ref{def:slat-cat} and associativity, commutativity, and idempotence equations for $+$.
+
     \begin{proposition}(A specialised case of Proposition 6.4.7~\cite{Borceux_1994})
@@
         % https://q.uiver.app/#q=WzAsMixbMCwwLCJcXGNhdG5hbWV7U0xhdHR9LVxcY2F0bmFtZXtDYXR9Il0sWzIsMCwiXFxjYXRuYW1le0NhdH0iXSxbMSwwLCJGIiwyLHsiY3VydmUiOjR9XSxbMCwxLCJHIiwyLHsiY3VydmUiOjR9XSxbMiwzLCIiLDAseyJsZXZlbCI6MSwic3R5bGUiOnsibmFtZSI6ImFkanVuY3Rpb24ifX1dXQ==
-    \[\begin{tikzcd}
-        \catname{SLat}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
-        \end{tikzcd}
+        % \[\begin{tikzcd}
+        % 	\catname{SLatt}\text{--}\catname{Cat}\arrow[rr, "\mathcal{U}", bend left] & \hspace{-1em}\top & \catname{Cat} \arrow[ll, "\mathcal{F}", bend left]
+        % 	\end{tikzcd}
+        % \]
+        \vspace{-1mm}
+        \[
+        \mathcal{F} \dashv \mathcal{U} : \catname{SLat}\text{--}\catname{Cat} \to \catname{Cat}
+        \vspace{-1mm}
         \]
@@
 
-    \[\begin{tikzcd}
-        \catname{SLat} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
-        \end{tikzcd}
+        % \[\begin{tikzcd}
+        % 	\catname{SLatt} \arrow[rr, "U", bend left] & \top & \catname{Set} \arrow[ll, "F", bend left]
+        % 	\end{tikzcd}
+        % \]
+        \vspace{-1mm}
+        \[
+        F \dashv U : \catname{SLat} \to \catname{Set}
+        \vspace{-1mm}
         \]        
     \end{proposition}
-
 In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice enriched $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
-Being a 2-functor it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{Slat}$-adjunction.
+Being a 2-functor, it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{Slat}$-adjunction.
 Thus, applying the functor to $\catname{CS}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{CS}(\Sigma)^{+}$ by lifting the adjunction $(- \otimes B, B \multimap{} -, \eta, \varepsilon)$.
@@
 
-Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of $\Sigma^{+}$-terms, the latter being defined as follows.
-Every $\Sigma$-term is also a $\Sigma^{+}$-term. 
-Given two $\Sigma^{+}$-terms $f : A \to B$ and $g : A \to B$, $f + g : A \to B$ is also a $\Sigma^{+}$-term.
+Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of closed $\Sigma^{+}$-terms, the latter being defined as follows.
+Every closed $\Sigma$-term is also a $\Sigma^{+}$-term. 
+Given two closed $\Sigma^{+}$-terms $f : A \to B$ and $g : A \to B$, $f + g : A \to B$ is also a closed $\Sigma^{+}$-term.
 Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed $\catname{SLat}$-SMC, including the distributivity law~\ref{law:distributivity}.
@@
 
-In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over some signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
-We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
-That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
-We call these string diagrams, or more precisely, the combinatorial representation of that will follow, e-graphs with bindings.
+% In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over some signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
+% We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
+% That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
+% We call these string diagrams, or more precisely, the combinatorial representation of that will follow, e-graphs with bindings.
 
@@
 \[
-\adjustbox{scale=0.75}{
+\adjustbox{scale=0.65}{
 \tikzfig{./figures/egraph-strings}
@@
 \]
+\captionsetup{skip=0pt, belowskip=-2ex}
 \caption{String diagrams for closed semilattice-enriched symmetric monoidal categories.}

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
sections done

@@
 We will also refer to $A,B$ as \textit{feet} of the cospan, and to $C$ as \textit{carrier}.
-Given an ordered sequence of labels from some alphabet $\Sigma$, $[l_1, \ldots, l_n]$ we will denote with $w([l_1, \ldots, l_n])$ the word concatenated from these labels using $\otimes$, \textit{e.g.}, $w([A,B]) = A \otimes B$.
+% Given an ordered sequence of labels from some alphabet $\Sigma$, $[l_1, \ldots, l_n]$ we will denote with $w([l_1, \ldots, l_n])$ the word concatenated from these labels using $\otimes$, \textit{e.g.}, $w([A,B]) = A \otimes B$.
 
@@
 
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of SMTs including a Frobenius algebra. 
+In ~\cite{bonchi_string_2022-2}, certain class of hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of arbitrary SMTs.
 Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
@@
 This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
-We then evoke the definition of hypergraphs and corresponding cospans as introduced in~\cite{bonchi_string_2022-1,bonchi_string_2022-2} over not uni-typed signatures.
+% Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
+We then evoke the definition of hypergraphs and corresponding cospans as introduced in~\cite{bonchi_string_2022-1,bonchi_string_2022-2}.
 
@@
 \begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{V,\mathcal{F}}(v) = l_{V,\mathcal{G}}(\phi_V(v))$
-    \item $l_{E,\mathcal{F}}(e) = l_{E,\mathcal{G}}(\phi_E(e))$
+    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e)) \qquad \phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
+    \item $l_{V,\mathcal{F}}(v) = l_{V,\mathcal{G}}(\phi_V(v)) \qquad\;\; l_{E,\mathcal{F}}(e) = l_{E,\mathcal{G}}(\phi_E(e))$
 \end{enumerate}
@@
 $\HypI{\Sigma}$ is a symmetric monoidal category with the tensor product given by coproduct in $\catname{Hyp}(\Sigma)$:
-
 \begin{align*}
@@
 \]
-
 that is computed by identifying vertices in $\mathcal{F} + \mathcal{G}$ that share pre-image in $m$.
@@
 
-To make such cospans complete with respect to representing morphism of $\catname{S}(\Sigma)$, we need to impose some restriction in their form.
-
+To make such cospans complete with respect to representing morphism of $\catname{S}(\Sigma)$, we need to impose some restriction on their form.
 \begin{definition}[Degree,  path,  acyclic] 
@@
 \begin{figure}[t!]
-    \begin{subfigure}[T]{0.4\linewidth}
+    \begin{subfigure}[T]{0.35\linewidth}
     \[
@@
     \end{subfigure}
+    \hfill
     \captionsetup{belowskip=-3ex}
@@
 Then, the authors of~\cite{bonchi_string_2022-2} showed that this enhanced notion of DPOI rewriting is sound and complete in the following sense.
-Any two $\Sigma$-terms $f$ and $g$ are equal modulo SMC laws and equations $\mathcal{E}$ if and only if $\llbracket f \rrbracket \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} \llbracket g \rrbracket$ where $\llbracket - \rrbracket$ is the functor giving the equivalence in Theorem~\ref{thm:prop-equiv} and $\llbracket E \rrbracket = \{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle, \langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle \text{ for } l=r \in \mathcal{E} \}$.
+Any two $\Sigma$-terms $f$ and $g$ are equal modulo SMC laws and equations $\mathcal{E}$ if and only if $\llbracket f \rrbracket \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} \llbracket g \rrbracket$ where $\llbracket - \rrbracket$ is the functor giving the equivalence in Theorem~\ref{thm:prop-equiv} and $\llbracket \mathcal{E} \rrbracket = \{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle, \langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle \text{ for } l=r \in \mathcal{E} \}$.
 We then proceed with building a combinatorial representation for closed $\Sigma^{+}$-terms to obtain a notion of DPOI rewriting that would also subsume equality modulo SMC laws.
@@
 The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established area of hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
-The latter work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
+The first work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
 Notably, the laws of a semilattice enrichment in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
@@
 \[
+\vspace{-1ex}
 \begin{array}{cccc}
@@
 Edges in $\textcolor{red}{E}$ are depicted with a dashed box, while edges in $\textcolor{blue}{E}$ --- with a round solid box.
-
 \begin{figure}
-
 \[
@@
 \]
+\captionsetup{skip=0pt, belowskip=-1ex}
 \caption{Cospan of e-hypergraphs example}
@@
 \end{figure}
-
 \begin{definition}[E-hypergraph homomorphism]
@@
         \begin{align*}
-          &\text{if } e \textcolor{red}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(v)\\
-          &\text{and}\\
-          &\text{if } e \textcolor{blue}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(v)
+          \text{if } e \textcolor{red}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(v)
+          \qquad
+          \text{if } e \textcolor{blue}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(v)
         \end{align*}
@@
           \begin{align*}
-          &\text{if } e_1 \textcolor{red}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(e_2)\\
-          &\text{and}\\
-          &\text{if } e_1 \textcolor{blue}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+          \text{if } e_1 \textcolor{red}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+          \qquad
+          \text{if } e_1 \textcolor{blue}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
           \end{align*}
@@
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied. 
-Most importantly a pushout for the composition of two cospans with discrete feet exists, as described below.
+Most importantly a pushout for the composition of two cospans with discrete feet always exists, as described below.
 
-As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, e-hypergraphs requires interfaces in order to model string diagrams with dedicated inputs and outputs.
+As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, e-hypergraphs require interfaces in order to model string diagrams with dedicated inputs and outputs.
 Because string diagrams for closed semilattice-enriched symmetric monoidal categories are nested, containing substring diagrams inside e-boxes and lambda abstraction boxes, we need to account for the interfaces in these nested contexts as well.
@@
 \begin{align*}
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\\
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
+    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ] \; h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
+% \\
+    % h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
 \end{align*}
@@
   \end{definition}
-This is a proper category as identities, symmetries, and tensor unit are well-typed and composition and tensor product of two well-typed MDA cospans is again a well typed MDA cospan.
+% This is a proper category as identities, symmetries, and tensor unit are well-typed and composition and tensor product of two well-typed MDA cospans is again a well typed MDA cospan.
 
@@
 Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
+\begin{align*}
+    &0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m\\
+    &0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
+\end{align*} 
+allows us to encode rewrite rules as extended cospans of the following form
 \[
@@
     \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+    \hspace{-2em}n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
     \]
@@
 % \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
+    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ for its reflexive, transitive and symmetric closure.
 \end{definition}
@@
 The enhanced definition of extended DPOI ensures that the combinatorial representation of closed semilattice-enriched SMCs is sound and complete for rewriting of closed $\Sigma^{+}$-terms.
+We use the term DPOI to mean EDPOI when it is clear from the context.

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
sections done

--- 2025-01-23 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
@@
 \begin{definition}
-    We say that a $\Sigma$-term $f$ rewrites ($\leadsto$) to a $\Sigma$-term $g$ modulo SMC-laws via a rewrite rule $\langle l, r \rangle$ if they are representable as
+\label{def:rewrite}
+    We say that a $\Sigma$-term $f$ rewrites to a $\Sigma$-term $g$ modulo SMC laws via a rewrite rule $\langle l, r \rangle$ (notation, $\leadsto_{\langle l, r \rangle}$) if they are representable as
     \[
-    f = c_{1};(id_{k} \otimes l);c_{2} \qquad g = c_{2};(id_{k} \otimes r);c_{2}~.
+    f = c_{1};(id_{k} \otimes l);c_{2} \qquad g = c_{1};(id_{k} \otimes r);c_{2}~.
     \]
-    These representations are so-called SMC normal forms.
+\end{definition}
+This notion generalises to rewriting of closed $\Sigma$-terms
+\begin{definition}
+    For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
+    \[
+    f = c_{1};(id_{k} \otimes d);c_{2} \qquad g = c_{1};(id_{k} \otimes e);c_{2}
+    \]
+    and $d \leadsto_{\langle l, r \rangle} e$
+\end{definition}
+\begin{lemma}
+\label{lemma:normal_form}
+    Every closed $\Sigma^{+}$ term $f$ can be equivalently represented as
+    \[
+    f_{1} + \ldots + f_{n}
+    \] modulo SMC laws and $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} where none of $f_{i}$ contain the join operator.
+\end{lemma}
+% \begin{proof}
+% Such forms are given by orienting $\catname{SLat}$-equations such that $+$ gets propagated to the top, as, for example $f \otimes (g + h) = f \otimes g + g \otimes h$.
+% Most notably, we can also escape lambda abstractions by using distributivity law~\ref{law:distributivity}.
+% \end{proof}
+This takes us to the following definition of rewriting for closed $\Sigma^{+}$-terms.
+\begin{definition}
+    For closed $\Sigma^{+}$-term $f,g$, $f \leadsto_{\langle l, r \rangle}$ SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
+    \[
+    f = f_{1} + \ldots + f_{i} + \ldots + f_{n} \qquad g = f_{1} + \ldots + f_{j} + \ldots + f_{n}
+    \]
+    such that there is a permutation $\sigma$ such that there exist indices $i,j$ and $\sigma(f_{i}) \leadsto f_{j}$.
+\end{definition}
+Next, to formulate the correspondence between rewriting systems we first define interpretation $\llbracket - \rrbracket$ of closed $\Sigma^{+}$ terms in $\WellTypedMdaEcospans$.
+It follows by first defining the interpretation of operators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in Appendix~\ref{sec:appendix:interpretation}.
+The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in Figure~\ref{fig:ev_and_lambda}.
+
+\begin{figure}
+    \begin{subfigure}{0.45\linewidth}
+\[
+\adjustbox{scale=0.6}{
+    \tikzfig{./figures/ev_interpretation}
+}
+\]
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\linewidth}
+        \[
+        \adjustbox{scale=0.6}{
+            \tikzfig{./figures/lambda_interpretation}
+        }
+        \]
+    \end{subfigure}
+\caption{$\llbracket \textsf{ev} \rrbracket$ (left) and $\llbracket \Lambda(f) \rrbracket$ (right)}
+\label{fig:ev_and_lambda}
+\end{figure}
+
+Then we can make $\WellTypedMdaEcospans$ into $\catname{SLat}$-category by defining a join of two cospans and introducing DPOI rewrite rules to satisfy the laws of $\catname{SLat}$ symmetric monoidal category.
+We also add a structural rewrite schema rule corresponding to distributivity law~\ref{law:distributivity}.
+The join of two cospans is defined as in Figure~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules, one for each axiom.
+For example, the distributivity rule for $\otimes$ is given by the following schema rule
+\[
+\adjustbox{width=\linewidth}{
+\tikzfig{./figures/semilattice_rule_1}~.
+}
+\]
+
+We collect all such rewrite schemas into a set $\mathcal{S}$ and then a category $\WellTypedMdaEcospans / \mathcal{S}$ is a closed $\catname{SLat}$ SMC.
+
+\begin{definition}[Quotient by rewrites]  
+    Given a set of DPOI rewrite rules $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation.
+    \[
+        f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\mathcal{E}} g ~ . 
+    \]
 \end{definition}
 
-This notion generalises to rewriting of closed $\Sigma$-terms by introducing contexts (in string diagrammatic notation).
+This explicit quotienting allows us to only consider cospans of the form $f_1 + \ldots + f_{n}$, similar to Lemma~\ref{lemma:normal_form}, such that carrier of $f_{i}$ does not contain any hierarchical edges from $\textcolor{red}{E}$.
+
+$\llbracket - \rrbracket$ is then induced by its definition on operators by induction
+\[
+\llbracket f;g \rrbracket = \llbracket f \rrbracket ; \llbracket g \rrbracket \quad
+\llbracket f \otimes g \rrbracket = \llbracket f \rrbracket \otimes \llbracket g \rrbracket \quad
+\llbracket f + g \rrbracket = \llbracket f \rrbracket + \llbracket g \rrbracket
+\]
+
+\begin{proposition}[Proposition 26 and Proposition 27~\cite{fscd}]
+\label{prop:fscd}
+For two closed $\Sigma$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{blue}{E}$.
+\end{proposition}
+
+\begin{proposition}
+    For two closed $\Sigma^{+}$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
+\end{proposition}
+\begin{proof}
+By definition, we have 
+\[
+    f = f_{1} + \ldots + f_{n} \qquad g = f_{1} + \ldots + f_{n}
+\], such that $f_{i} \leadsto f_{j}$ for some $i, j$.
+This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and by noting that every $f_{i}$ is a $\Sigma$-term and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{red}{E}$ we can apply Proposition~\ref{prop:fscd}.
+\end{proof}
+
+\begin{figure}
+\[
+        \adjustbox{scale=0.5}{
+            \tikzfig{./figures/f_plus_g_new}
+        }
+\]
+\caption{$+$ of two morphisms in $\WellTypedMdaEcospans$}
+\label{fig:f+g}
+\end{figure}
+
+\section{Conclusion}
+In this work we introduced e-graphs with bindings formalised as morphisms of a closed $\catname{SLat}$-SMC.
+We then provided a combinatorial representation for such morphisms in terms of hierarchical hypergraphs (e-hypergraphs) with a suitable notion of sound and complete DPO rewriting, making it possible for an e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
+An important future work is exploring whether such combinatorial structures may lead to better algorithmics for doing equality saturation for $\lambda$-calculus.

### 2025-01-24 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
fixes

@@
 
-In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over some signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
+In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
 We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
fixes

@@
 \subsection{DPOI rewriting}
+\label{sec:appendix:dpoi}
 
@@
 First, to compute the boundary complement, we remove from the source e-hypergraph everything that has no pre-image in the interface above.
-The mda-cospan for the complement that will have .
+The mda-cospan for the complement then will be
+\[
+[u_1] \to [u_1, v_1, v_2] + [v_3] \to \mathcal{L}^{\bot} \xleftarrow{} [v_3] + [v_1,v_2] \xleftarrow{} [u_2] ~.
+\]
+Note how we removed strict internal interfaces of $\mathcal{L}$ from internal interfaces of $\mathcal{G}$, yielding $[u_1,v_1,v_2]$ and similarly for output interfaces.
+After computing the pushout for the second half of the diagram we add strict internal interfaces of $\mathcal{R}$ but since there are none, we are done as per condition (2) of Definition~\ref{def:dpoi-e}.
 

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
fixes

@@
-\section{Combinatorial semantics}
-
 \section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
@@
 The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  
-A concrete description of the pushout of two morphisms in this category is given in the Appendix section~\ref{sec:appendix:pushout}.
+A concrete description of the pushout of two morphisms in this category is given in the Appendix~\ref{sec:appendix:pushout}.
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied. 
@@
 Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
+\begin{definition}[Convex EDPOI rewriting]\label{def:dpoi-e}
 Given an extended span of morphisms 
+\vspace{-5mm}
 \[
@@
 in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
+\vspace{-3mm}
 \[
@@
 
+An example of EDPOI rewrite can be seen in Appendix~\ref{sec:appendix:dpoi}.
 The enhanced definition of extended DPOI ensures that the combinatorial representation of closed semilattice-enriched SMCs is sound and complete for rewriting of closed $\Sigma^{+}$-terms.

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
fit 12 pages

@@
 To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
+In the case of $ \catname{Hyp(\Sigma)}$,  we wish for a rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
 By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
@@
 \] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
+Analogously to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data:
 \begin{align*}
@@
 \end{align*} 
-allows us to encode rewrite rules as extended cospans of the following form
+allows us to encode rewrite rules as extended cospans of the following form:
 \[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
+\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R},
 \] 
@@
 Given an extended span of morphisms 
-\vspace{-5mm}
 \[
@@
 in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\vspace{-3mm}
 \[

### 2025-01-24 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
fit 12 pages
proofreading pass, clean up prose

@@
 \usepackage{wrapfig}
+
+\usepackage[final]{microtype}
 
@@
 Using e-graphs mitigates the so-called \emph{phase-ordering problem} of term rewriting, that is the fact that the way rewrites are scheduled may enable or block further rewrites due to introduction or elimination or redexes.
-This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule instead of replacing the left-hand side by the right-hand side, adds the latter to the graph so that all pre-existing redexes are protected and new ones are introduced.
+This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule, instead of replacing the left-hand side by the right-hand side, adds the latter to the graph so that all pre-existing redexes are protected and new ones are introduced.
 Explosive growth of the graph is avoided by the clever use of sharing of vertices.
 
-E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest particularly for optimisation and synthesis using so called \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, the repeated application of rewrites to an e-graph until a fixpoint is reached.
+E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using so-called technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint is reached.
 The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
-Of particular interest to us is using e-graphs in the context of functional programming languages, which requires a representation of the $\lambda$-calculus using it. 
+Of particular interest to us is using e-graphs in the context of functional programming languages, which builds upon an internal representation of the $\lambda$-calculus. 
 Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
@@
 
-In the textual representation of $\lambda$ terms, $\alpha$-equivalent terms are distinct and thus cannot be shared, which is not only problematic but also inefficient. 
-Using De Bruijin indices is also problematic because the same variable can be assigned different indices depending on under the scope of how many binders it lies. 
+In the textual representation of $\lambda$-terms, $\alpha$-equivalent terms are distinct and thus cannot be shared, which adds a layer of bookkeeping and complexity.
+The traditional approach to handle this is by using de Bruijn indices~\cite{de1972lambda}, for which $\alpha$-equivalent terms are syntactically identified; however, this is also problematic because the same variable can be assigned different indices depending on its scope with respect to that instance of the variable.
 This again interferes with sharing. 
-As a simple example, consider the term $\lambda f . f ((\lambda x . f x) 2)$ which is encoded as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$.
-Notice how two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$, which cannot be shared in an e-graph.
+As a simple example, consider the term $\lambda f . f ((\lambda x . f x) 2)$ which, with de Bruijn indices,  is encoded as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$.
+The two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$, which cannot be shared in an e-graph.
 
 Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
-However, calculi of explicit substitutions have many technical subtleties which also directly impact how equivalence is defined, for instance \emph{graph equivalence} instituted between terms which have non-interfering explicit substitutions applied in different orders~\cite{accattoli2014nonstandard}.
-These difficulties are not handled in \emph{loc. cit.}.
+The implementations of $\beta$-reduction in e-graphs do not go into much detail regarding explicit substitution, and in particular do not include rules like $f[x/1][y/2] = f[y/2][x/1]$ for non-interfering substitutions, while such rules are absorbed by graph equivalence~\cite{accattoli2014nonstandard}.
 
-All these issues can be resolved by the use of \emph{string diagrams}.
-Thus, in this paper we take what we consider to be a principled approach to e-graphs with bindings by lifting to monoidal closed categories prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} showing how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
-The approach in \emph{loc.
-cit.} was originally motivated by the desire to support the machinery of categorial (DPO) graph rewriting for e-graphs, but by extending the semilattice-enrichment to monoidal-categories it paves the way to a natural derivation of e-graphs with binding.
+All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
+\cite{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
+In this paper, we extend this technique to e-graphs with bindings, lifting to monoidal \emph{closed} categories.
 
@@
 
-We build on the technology of string diagrams, which is by now well developed in a number of areas. 
-String diagrams have the advantage of being a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, all based on solid mathematical foundations.
-To solve the new problems discusses above, of dealing with binding and capture avoiding substitution, we just follow the categorical breadcrumb trail.
+We build on the technology of string diagrams, which is by now well-developed in a number of areas. 
+String diagrams have the advantage of being a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
+These categorical techniques provide insight in how to solve the problems of binding and capture-avoiding substitution in e-graphs, as we develop in this paper.
 
@@
 
-String diagrams provide a geometric or topological framework for understanding monoidal categories, representing objects as \emph{strings} and morphisms as \emph{nodes}, where input strings flow in and output strings flow out.
-Key structures in monoidal categories---such as the monoidal product, tensor unit, braiding, and dualities---are represented through operations like placing, weaving, or bending strings. 
-This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by symbolic notation. 
-Originating from notations by Hotz~\cite{hotzsd} (for automata) and Penrose~\cite{penrose1984spinors} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, offering a dual perspective to commutative diagrams through Poincaré duality.
+String diagrams provide a topological framework for understanding monoidal categories, representing objects as \emph{strings} and morphisms as \emph{nodes}, where input strings flow in and output strings flow out.
+Key structures in monoidal categories---such as the monoidal product, tensor unit, braiding, and dualities---are represented graphically, through operations like juxtaposing, weaving, or bending strings.
+This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by traditional symbolic notation. 
+Originating from notations by Hotz~\cite{hotzsd} (for automata) and Penrose~\cite{penrose1984spinors} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, and provide a mathematically rigorous diagrammatic calculus~\cite{joyal_geometry_1991}.
 
-As described above, string diagrams are commonly an informal yet rigorous two-dimensional syntax for the internal language of a category. 
-They can be formalised by representing them as graphs or hypergraphs~\cite{bonchi_string_2022-1}, which has two further benefits. 
-The first benefit is that the property of two string diagrams being equal if and only if they ``look the same'' can be spelled out as graph isomorphism, eliminating any possible ambiguity of interpretation.
-The second benefit is that equational reasoning can also be formalised as (categorical) graph rewriting, for instance double-pushout rewriting (DPO).
+Commonly, string diagrams are used informally to provide a two-dimensional syntax for the internal language of a category, but for our purposes we must be more formal. 
+Specifically, we require a concrete encoding of string diagrams, analogous to the de Bruijn index representation of $\lambda$-terms.
+For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics for this purpose.
+For instance, hypergraphs admit the property that two string diagrams are equal if and only if they ``look the same'' (\emph{i.e.} they admit a graph isomorphism); secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.
 
-Finally, to represent diagrammatically two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
+Finally, for the diagrammatical representation of two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
 The diagrammatic boxes require the introduction of further structure to (hyper)graphs in the form of a hierarchical layering of the graphs, as studied in \emph{loc. cit.}.
 
-\begin{wrapfigure}{L}{0.2\textwidth}
-\[
+\begin{figure}
+  \centering
   \adjustbox{scale=0.6}{
@@
   }
-\]
-\end{wrapfigure}
+\caption{String-diagrammatic representation of $\lambda f . f ((\lambda x . f x) 2)$.}\label{fig:de-brujin-string}
+\end{figure}
 
-Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, note its string diagrammatic representation below, in which the two occurrences of $f$ are now shared.
+Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, in Figure~\ref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
 $\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire attached to the frame,
@@
 E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph as first presented in~\cite{EggPaper}.
+The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
 
@@
 Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using the contraction node.
+Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node.
 
@@
 In the former, edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes, depicted with black dots with a single wire inside the dashed boxes
-
+Explicit discarding (weakening) nodes are depicted with black dots with a single wire
 \scalebox{0.4}{
@@
 	\end{tikzpicture}
-}, indicate which of the class-level edges are connected to which nodes inside the class.
+}
+inside the dashed boxes, indicating which of the class-level edges are connected to which nodes inside the class.
 
@@
 A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
-The auxiliary nature of these nodes is revealed by the fact that after rewrites they must be removed from the graph. 
-We will see how in the categorically-directed representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.
+The auxiliary nature of these nodes is revealed by the fact that, after rewriting, they must be removed from the graph. 
+We will see how in the categorical representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.
 
@@
 
-Before diving into the technical details let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
-In Figure~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following the $\beta$-reductions.
+First, let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
+In Figure~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
 A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box), to be rewritten using the rules for explicit substitution at the top of each column.
@@
 The explicit occurrences of variable names in the older approach raise obvious questions of scope, binding and explicit substitution, whereas the use of terms inside the explicit substitution nodes also raises many technical difficulties which makes proving various mathematical properties essentially intractable.
-Additionally, explicit substitutions need to be propagated recursively throughout the graph, which is also inefficient. 
-Using De Brujin indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
+Additionally, explicit substitutions need to be propagated recursively throughout the graph, which adds another layer of technical and conceptual overhead. 
+Using de Bruijn indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
 
@@
     \]
-	\subcaption{String diagrammatic substitution example.}
+	\subcaption{String diagrammatic substitution example.}\label{fig:e-string-substitution}
 	\end{subfigure}

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
proofreading pass, clean up prose

@@
 Prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
-The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established area of hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
+The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established body of literature on hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
 The first work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
@@
 We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness and when evoking a term e-hypergraph we will mean the notion introduced in what follows rather than what was defined in \textit{loc.
-cit.}.
+We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness, and when evoking a term e-hypergraph modify the original definitions as follows.
 
@@
 Also note the typing on the $l_{V}$ function.
-We work with freely generated closed semilattice-enriched categories and to avoid unnecessary complications we limit the labelling of vertices to such a type.
+We work with freely generated closed semilattice-enriched categories, and to avoid unnecessary complications we limit the labelling of vertices to such a type.
 In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
@@
 The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  
-A concrete description of the pushout of two morphisms in this category is given in the Appendix~\ref{sec:appendix:pushout}.
+A concrete description of the pushout of two morphisms in this category is given in Appendix~\ref{sec:appendix:pushout}.
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied. 

### 2025-01-24 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
fixes
proofreading pass, clean up prose

@@
 \[
-    f = f_{1} + \ldots + f_{n} \qquad g = f_{1} + \ldots + f_{n}
-\], such that $f_{i} \leadsto f_{j}$ for some $i, j$.
-This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and by noting that every $f_{i}$ is a $\Sigma$-term and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{red}{E}$ we can apply Proposition~\ref{prop:fscd}.
+  f = f_{1} + \ldots + f_{n} \quad \text{and} \quad g = f_{1} + \ldots + f_{n},
+\] such that $f_{i} \leadsto f_{j}$ for some $i, j$.
+This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and, by noting that every $f_{i}$ is a $\Sigma$-term, and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{red}{E}$, we can apply Proposition~\ref{prop:fscd}.
 \end{proof}
@@
 \]
+\captionsetup{belowskip=-1ex}
 \caption{$+$ of two morphisms in $\WellTypedMdaEcospans$}

### 2025-01-26 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
made formatting more compact by inlining some maths

@@
 
-We build on the technology of string diagrams, which is by now well-developed in a number of areas. 
-String diagrams have the advantage of being a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
+We build on the technology of string diagrams, by now well-developed in a number of areas. 
+String diagrams are a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
 These categorical techniques provide insight in how to solve the problems of binding and capture-avoiding substitution in e-graphs, as we develop in this paper.
@@
 
-Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, in Figure~\ref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
+Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, in Fig.~\ref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
 $\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire attached to the frame,
@@
 E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
+The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Fig.~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
 
@@
 First, let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
-In Figure~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
+In Fig.~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
 A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box), to be rewritten using the rules for explicit substitution at the top of each column.
 At the end, all explicit substitution nodes can be removed.
-The same example is then given again in Figure~\ref{fig:e-string-substitution}, this time using string diagrams.
+The same example is then given again in Fig.~\ref{fig:e-string-substitution}, this time using string diagrams.
 

### 2025-01-26 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
fixes
made formatting more compact by inlining some maths

@@
 \begin{definition}
-A set of $\Sigma$-terms generated by monoidal signature $\Sigma$ is given by the following grammar
+A set of $\Sigma$-terms generated by monoidal signature $\Sigma$ is given by the following grammar:
 \[
-\tau \coloneqq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}
-\],
+\tau \coloneqq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2},
+\]
 where $\phi$ ranges over the set of operators $\Sigma_{M}$ and $A,B$ range over the set of objects $\Sigma_{O}$.
@@
         \item for every triple of objects $A,B,C \in \mathbb{C}$ --- a composition morphism 
-        \[
-            \circ: \mathbb{C}(B,C) \otimes \mathbb{C}(A,B) \to \mathbb{C}(A,C)
-        \]
+        $\circ: \mathbb{C}(B,C) \otimes \mathbb{C}(A,B) \to \mathbb{C}(A,C)$
         \item for every object $A \in C$ --- a unit morphism 
-        \[
-        u_{A} : I \to \mathbb{C}(A,A), \text{ where $I$ is the monoidal unit for $\catname{SLat}$}
-        \]
+        $u_{A} : I \to \mathbb{C}(A,A)$, where $I$ is the monoidal unit for $\catname{SLat}$.
     \end{itemize}
@@
 \end{definition}
-As the composition is a semilattice homomorphism (morphism of $\catname{SLat}$), we have the following equations 
-\[
-(f + g); h = f;h + g;h \quad f;(g+h) = f;g + f;h
-\]
+As the composition is a semilattice homomorphism (morphism of $\catname{SLat}$), we require that 
+$
+(f + g); h = f;h + g;h$ and $f;(g+h) = f;g + f;h
+$.
 
@@
     \end{definition}
-    The latter morphism being in $\catname{SLat}$ implies the following equations for $\otimes$:
-    \[
-    f \otimes (g+h) = f \otimes g + f \otimes h \qquad (f+g) \otimes h = f \otimes h + g \otimes h
-    \]
-
-    Again, we take $\otimes$ to bind more tightly than $+$.
+    The latter morphism being in $\catname{SLat}$ implies that for $\otimes$ we have 
+    $
+    f \otimes (g+h) = f \otimes g + f \otimes h $ and $ (f+g) \otimes h = f \otimes h + g \otimes h
+    $.
+    As before, we take $\otimes$ to bind more tightly than $+$.
 
@@
         \]
-        that is induced by a usual free-forgetful adjunction
+        that is induced by the usual free-forgetful adjunction
 

### 2025-01-26 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
made formatting more compact by inlining some maths

@@
 
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  recapitulating the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
+We recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  recapitulating the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
 
@@
 \end{enumerate}
-That is, a homomorphism should preserve sources and targets of edges and preserve the labels.
+\end{definition}
 We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms.
-\end{definition}
 $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
@@
 Composition of two cospans $n \xrightarrow{f} \mathcal{F} \xleftarrow{g} m;m \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} k$ is given by the pushout
-\[
+$
 n \xrightarrow{f} \mathcal{F} +_{g,f'} \mathcal{G} \xleftarrow{g'} k
-\]
-that is computed by identifying vertices in $\mathcal{F} + \mathcal{G}$ that share pre-image in $m$.
+$,
+computed by identifying vertices in $\mathcal{F} + \mathcal{G}$ that share pre-image in $m$.
 As pushout is only defined up to isomorphism, it justifies the morphisms in this category being isomorphism classes of cospans.
-
 To make such cospans complete with respect to representing morphism of $\catname{S}(\Sigma)$, we need to impose some restriction on their form.
@@
     \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
+    \item the in-degree and out-degree of every vertex is at most~$1$;
+    \item vertices with in-degree $0$ are precisely the image of $f$; 
     \item vertices with out-degree $0$ are precisely the image of $g$.
@@
 is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
+$
     n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
+$
 is an MDA cospan. 
@@
 \begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $f : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
+A subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ is a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
+A match $f : \mathcal L \to \mathcal G$ is a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
 
@@
 \begin{definition}
-We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{red}{<}, \textcolor{blue}{<},\consistency)$ where
+We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{orange}{<}, \textcolor{blue}{<},\consistency)$ where
 \begin{itemize}
   \item $V$ is a set of vertices.
-  \item $E = \colorbox{yellow}{E} \cup \textcolor{red}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
+  \item $E = \colorbox{yellow}{E} \cup \textcolor{orange}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
   \item $s,t : E \to V^{*}$ are source and target functions.
   \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are \textit{label} functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
-  \item $\textcolor{red}{<} \subseteq \textcolor{red}{E} \times V + E$ and $\textcolor{blue}{<} \subseteq \textcolor{blue}{E} \times V + E $ with $\textcolor{red}{E}, \textcolor{blue} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{red}{<^{\mu}}$ and $\textcolor{blue}{<^{\mu}}$.
-        For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{red}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{blue}{<^{\mu}} y_{i+1}\; \}$.
+  \item $\textcolor{orange}{<} \subseteq \textcolor{orange}{E} \times V + E$ and $\textcolor{blue}{<} \subseteq \textcolor{blue}{E} \times V + E $ with $\textcolor{orange}{E}, \textcolor{blue} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{orange}{<^{\mu}}$ and $\textcolor{blue}{<^{\mu}}$.
+        For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{orange}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{blue}{<^{\mu}} y_{i+1}\; \}$.
         We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
@@
           \item each $x$ has at most one immediate predecessor;
-          \item edges $e$ such that there is no $x$ such that $e \textcolor{red}{<^{\mu}} x$ or $e \textcolor{red}{<^{\mu}} x$ are labelled;
-          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{red}{<^{\mu}} e$ iff $e' \textcolor{red}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{blue}{<}$.
+          \item edges $e$ such that there is no $x$ such that $e \textcolor{orange}{<^{\mu}} x$ or $e \textcolor{orange}{<^{\mu}} x$ are labelled;
+          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{orange}{<^{\mu}} e$ iff $e' \textcolor{orange}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{blue}{<}$.
         \end{enumerate}
-  \item $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p \textcolor{red}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{red}{<^{\mu}}(v)$ and $p \textcolor{red}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
-  We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{red}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{red}{<^{\mu}} e\}$.
+  \item $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p \textcolor{orange}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{orange}{<^{\mu}}(v)$ and $p \textcolor{orange}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
+  We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{orange}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{orange}{<^{\mu}} e\}$.
 \end{itemize} 
-$\textcolor{red}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges; then we have $\colorbox{yellow}{E} = E \setminus \textcolor{red}{E} \cup \textcolor{blue}{E}$.
-Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{red}{<}, \textcolor{blue}{<}$ and~$\consistency$.
+$\textcolor{orange}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{orange}{E} \cup \textcolor{blue}{E}$.
+Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{orange}{<}, \textcolor{blue}{<}$ and~$\consistency$.
 \end{definition}
 
-Intuitively, edges from $\textcolor{red}{E}$ encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ encode lambda-abstraction boxes.
+Intuitively, edges from $\textcolor{orange}{E}$ encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ encode lambda-abstraction boxes.
 Also note the typing on the $l_{V}$ function.
@@
 In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
-$\textcolor{blue}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{red}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+$\textcolor{blue}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{orange}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
 
 Consider an example in the middle part of Figure~\ref{fig:e-cospan-example}.
-The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{red}{E} = \{e_{1}\}$, $\textcolor{blue}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
+The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{orange}{E} = \{e_{1}\}$, $\textcolor{blue}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
 \[
@@
 \begin{array}{cccc}
-  v_{4} \textcolor{red}{<^{\mu}} e_{1} & v_{7} \textcolor{blue}{<^{\mu}} e_{2} & v_{4} \consistency e_{2} & v_{4} \not \consistency v_{10}\\
-  e_{3} \textcolor{red}{<^{\mu}} e_{1} & v_{8} \textcolor{blue}{<^{\mu}} e_{2} &  e_{2} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
-  e_{2} \textcolor{red}{<^{\mu}} e_{1} & e_{5} \textcolor{blue}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
+  v_{4} \textcolor{orange}{<^{\mu}} e_{1} & v_{7} \textcolor{blue}{<^{\mu}} e_{2} & v_{4} \consistency e_{2} & v_{4} \not \consistency v_{10}\\
+  e_{3} \textcolor{orange}{<^{\mu}} e_{1} & v_{8} \textcolor{blue}{<^{\mu}} e_{2} &  e_{2} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
+  e_{2} \textcolor{orange}{<^{\mu}} e_{1} & e_{5} \textcolor{blue}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
   \ldots & \ldots & \ldots & \ldots \\
-  v_{12} \textcolor{red}{<^{\mu}} e_{1} & w_{9} \textcolor{blue}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
+  v_{12} \textcolor{orange}{<^{\mu}} e_{1} & w_{9} \textcolor{blue}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
 \end{array}
@@
 $\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
-Edges in $\textcolor{red}{E}$ are depicted with a dashed box, while edges in $\textcolor{blue}{E}$ --- with a round solid box.
+Edges in $\textcolor{orange}{E}$ are depicted with a dashed box, while edges in $\textcolor{blue}{E}$ --- with a round solid box.
 \begin{figure}
@@
       \item $\phi$ is a hypergraph homomorphism;
-      \item $\phi_{E}(\textcolor{red}{E_{\mathcal{F}}}) \subseteq \textcolor{red}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{blue}{E_{\mathcal{F}}}) \subseteq \textcolor{blue}{E_{\mathcal{G}}}$;
+      \item $\phi_{E}(\textcolor{orange}{E_{\mathcal{F}}}) \subseteq \textcolor{orange}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{blue}{E_{\mathcal{F}}}) \subseteq \textcolor{blue}{E_{\mathcal{G}}}$;
       \item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
         \begin{align*}
-          \text{if } e \textcolor{red}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(v)
+          \text{if } e \textcolor{orange}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{orange}{<_{\mathcal{G}}^{\mu}} \phi(v)
           \qquad
@@
           \begin{align*}
-          \text{if } e_1 \textcolor{red}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{red}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+          \text{if } e_1 \textcolor{orange}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{orange}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
           \qquad
@@
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied. 
-Most importantly a pushout for the composition of two cospans with discrete feet always exists, as described below.
+However, a pushout for the composition of two cospans with discrete feet always exists, as described below.
 
@@
 \begin{multline*}
-R = \{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\}
+R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\bigr\}
 \end{multline*}
@@
 Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
+$
+	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m
+	$ and $
+	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} \mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
+$ is computed in two stages.
 First, $\mathcal{H}$ is computed as the result of the pushout square shown below: 
@@
   We call a cospan 
-  \[
+  $
 n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
-\]
+$
 monogamous directed acyclic if
@@
   \begin{itemize}
-    \item For each $e \in \textcolor{red}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
+    \item For each $e \in \textcolor{orange}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
           $e$ is well-typed if for each element $S$ of the partition of $I$ $w(S) = w(s(e))$ and similarly for $O$ and $t(e)$.
@@
 The construction in this section closely follows the one introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} with some adjustments to account for $\textcolor{blue}{<}$.
-
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
@@
 \] 
-Analogously to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data:
-\begin{align*}
-    &0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m\\
-    &0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\end{align*} 
-allows us to encode rewrite rules as extended cospans of the following form:
-\[
+Analogously to Section \ref{sec:combinatorial-semantics}, observing that  cospan
+$
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m$ expresses the same data as $
+    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
+$
+allows us to encode rewrite rules as extended cospans of the form
+$
 \mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R},
-\] 
+$ 
 which fits into the DPO formalism. We make this identification throughout,  without further mention. 
@@
 For MDA cospans 
-\[
+$
     i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
+$ and $
     n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $m : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
+$ and mono $m : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
 as depicted in the square below
@@
     \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended MDA cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
+    \begin{multline*}
+    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \\ \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+    \end{multline*}
 \end{enumerate}
@@
 Given an extended span of morphisms 
-\[
+$
 \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
+$
 in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
+$
 \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\]
+$
 if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
@@
 	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
+	$
+		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) $ and $ k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)
+	$
 	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.

### 2025-01-26 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
made formatting more compact by inlining some maths

@@
     We say that a $\Sigma$-term $f$ rewrites to a $\Sigma$-term $g$ modulo SMC laws via a rewrite rule $\langle l, r \rangle$ (notation, $\leadsto_{\langle l, r \rangle}$) if they are representable as
-    \[
-    f = c_{1};(id_{k} \otimes l);c_{2} \qquad g = c_{1};(id_{k} \otimes r);c_{2}~.
-    \]
+    $
+    f = c_{1};(id_{k} \otimes l);c_{2} $ and $ g = c_{1};(id_{k} \otimes r);c_{2}
+    $.
 \end{definition}
@@
     For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
-    \[
-    f = c_{1};(id_{k} \otimes d);c_{2} \qquad g = c_{1};(id_{k} \otimes e);c_{2}
-    \]
+    $
+    f = c_{1};(id_{k} \otimes d);c_{2} $ and $ g = c_{1};(id_{k} \otimes e);c_{2}
+    $
     and $d \leadsto_{\langle l, r \rangle} e$
@@
     Every closed $\Sigma^{+}$ term $f$ can be equivalently represented as
-    \[
+    $
     f_{1} + \ldots + f_{n}
-    \] modulo SMC laws and $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} where none of $f_{i}$ contain the join operator.
+    $ modulo SMC laws and $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} where none of $f_{i}$ contain the join operator.
 \end{lemma}
@@
     For closed $\Sigma^{+}$-term $f,g$, $f \leadsto_{\langle l, r \rangle}$ SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
-    \[
-    f = f_{1} + \ldots + f_{i} + \ldots + f_{n} \qquad g = f_{1} + \ldots + f_{j} + \ldots + f_{n}
-    \]
+    $
+    f = f_{1} + \ldots + f_{i} + \ldots + f_{n} $ and $ g = f_{1} + \ldots + f_{j} + \ldots + f_{n}
+    $
     such that there is a permutation $\sigma$ such that there exist indices $i,j$ and $\sigma(f_{i}) \leadsto f_{j}$.
@@
 \begin{definition}[Quotient by rewrites]  
-    Given a set of DPOI rewrite rules $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation.
-    \[
-        f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\mathcal{E}} g ~ . 
-    \]
+    Given a set of DPOI rewrite rules $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the  relation
+    $
+        f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\mathcal{E}} g 
+    $.
 \end{definition}
@@
 By definition, we have 
-\[
-  f = f_{1} + \ldots + f_{n} \quad \text{and} \quad g = f_{1} + \ldots + f_{n},
-\] such that $f_{i} \leadsto f_{j}$ for some $i, j$.
+$
+  f = f_{1} + \ldots + f_{n} $ and $  g = f_{1} + \ldots + f_{n}$,
+such that $f_{i} \leadsto f_{j}$ for some $i, j$.
 This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and, by noting that every $f_{i}$ is a $\Sigma$-term, and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{red}{E}$, we can apply Proposition~\ref{prop:fscd}.

### 2025-01-27 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
clean-up

@@
 \newtheorem{lemma}[theorem]{Lemma}
+
+% change red to orange
+\definecolor{e-color}{named}{orange}
+\definecolor{closed-color}{named}{blue}
 
@@
 Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
-The implementations of $\beta$-reduction in e-graphs do not go into much detail regarding explicit substitution, and in particular do not include rules like $f[x/1][y/2] = f[y/2][x/1]$ for non-interfering substitutions, while such rules are absorbed by graph equivalence~\cite{accattoli2014nonstandard}.
+The implementations of $\beta$-reduction in e-graphs do not go into much detail regarding explicit substitution, and in particular do not include rules like $[x\!\mapsto\!1][y\!\mapsto\!2]f\!=\![y\!\mapsto\!2][x\!\mapsto\!1]f$ for non-interfering substitutions, while such rules are absorbed by graph equivalence~\cite{accattoli2014nonstandard}.
 

### 2025-01-27 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
clean-up

@@
 \begin{remark}
-    $\textcolor{red}{<}$, $\textcolor{blue}{<}$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
+    $\textcolor{e-color}{<}$, $\textcolor{closed-color}{<}$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
     To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $\textcolor{red}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))$.
+    For example, an immediate successor of a vertex $x$ can be written functionally as $\textcolor{e-color}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))$.
 \end{remark}
@@
 
-            \item When $x$ is not a top-level vertex such that $\textcolor{red}{<}(\iota_{V}(x))$ is defined, 
+            \item When $x$ is not a top-level vertex such that $\textcolor{e-color}{<}(\iota_{V}(x))$ is defined, 
                 \[
-                \phi_{E}(\textcolor{red}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))) = \textcolor{red}{<_{\mathcal{G}}^{\mu}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
+                \phi_{E}(\textcolor{e-color}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))) = \textcolor{e-color}{<_{\mathcal{G}}^{\mu}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
                 \]
@@
                 \[
-                \phi_{E}(\textcolor{red}{<_{\mathcal{F}}^{\mu}}(\iota_{E_{\mathcal{F}}}(x))) = \textcolor{red}{<_{\mathcal{G}}^{\mu}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
+                \phi_{E}(\textcolor{e-color}{<_{\mathcal{F}}^{\mu}}(\iota_{E_{\mathcal{F}}}(x))) = \textcolor{e-color}{<_{\mathcal{G}}^{\mu}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
                 \] when $x$ is a not top-level edge.
-                \item When $x$ is not a top-level vertex such that $\textcolor{blue}{<}(\iota_{V}(x))$ is defined, 
+                \item When $x$ is not a top-level vertex such that $\textcolor{closed-color}{<}(\iota_{V}(x))$ is defined, 
                 \[
-                \phi_{E}(\textcolor{blue}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))) = \textcolor{blue}{<_{\mathcal{G}}^{\mu}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
+                \phi_{E}(\textcolor{closed-color}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))) = \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
                 \]
@@
                 \[
-                \phi_{E}(\textcolor{blue}{<_{\mathcal{F}}^{\mu}}(\iota_{E_{\mathcal{F}}}(x))) = \textcolor{blue}{<_{\mathcal{G}}^{\mu}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
+                \phi_{E}(\textcolor{closed-color}{<_{\mathcal{F}}^{\mu}}(\iota_{E_{\mathcal{F}}}(x))) = \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
                 \] when $x$ is a not top-level edge.
@@
     \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, \textcolor{red}{<_{X+Y}}, \textcolor{blue}{<_{X+Y}}\}
+        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, \textcolor{e-color}{<_{X+Y}}, \textcolor{closed-color}{<_{X+Y}}\}
     \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, \textcolor{red}{<_{X+Y}}, \textcolor{blue}{<_{X+Y}}, \consistency_{X+Y}$.
+    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, \textcolor{e-color}{<_{X+Y}}, \textcolor{closed-color}{<_{X+Y}}, \consistency_{X+Y}$.
     We will omit labels as they are irrelevant to pushout construction.
@@
             &\consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},\\
-            &\textcolor{red}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-            &\textcolor{blue}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+            &\textcolor{e-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
+            &\textcolor{closed-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
         &)    
@@
 
-Recall that $\textcolor{red}{<}$ is essentially a transitive closure of $\textcolor{red}{<^{\mu}}$ and hence we can first define 
-    \[\textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
+Recall that $\textcolor{e-color}{<}$ is essentially a transitive closure of $\textcolor{e-color}{<^{\mu}}$ and hence we can first define 
+    \[\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
     \] and 
     \[
-        \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u])
+        \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u])
     \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
@@
     \begin{enumerate}
-        \item If there exists $v$ such that $\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))$ is defined and $u \sim v$
+        \item If there exists $v$ such that $\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))$ is defined and $u \sim v$
               we let
               \[
-                \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = [\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))]
+                \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = [\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))]
               \]
         \item \label{def:child_respects_connectivity} $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
-              If there exists $v$ such that $\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
+              If there exists $v$ such that $\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
               \[ 
-                \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+                \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
               \]
@@
 
-    Otherwise, we leave $\textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))$ undefined.
+    Otherwise, we leave $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))$ undefined.
     % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
@@
     \begin{enumerate}
-    \item  $\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e)) = e'$.
+    \item  $\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e)) = e'$.
             Then we define
         \[
-            \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = [\textcolor{red}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e))]
+            \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = [\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e))]
         \]
@@
         \[
-            \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = \textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+            \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
         \] 
     \end{enumerate}
-    Otherwise we leave $\textcolor{red}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{E_{X} + E_{Y}}{\sim}([e]))$ undefined.
+    Otherwise we leave $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{E_{X} + E_{Y}}{\sim}([e]))$ undefined.
     % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
@@
     Clearly, all the cases above are disjoint.
-    We similarly define $\textcolor{blue}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
+    We similarly define $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
 
@@
     This construction is analogous to how a pushout was constructed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and we refer the reader for more detailed proofs there.
-    The only addition is $\textcolor{blue}{<_{\sfrac{X+Y}{\sim}}}^{\mu}$ that is constructed analogously to $\textcolor{red}{<_{\sfrac{X+Y}{\sim}}}^{\mu}$.
+    The only addition is $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$ that is constructed analogously to $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
 \end{proof}
-The construction can be illustrated with the example in Figure~\ref{fig:pushout_example}.
+The construction can be illustrated with the example in Fig.~\ref{fig:pushout_example}.
 Vertices and edges that are being identified as they have a common preimage inherit the relations from the vertex or edge that has it defined either directly or via a path.
@@
 \end{tikzpicture}
-such that $\textcolor{blue}{<_{X+Y}}$ is only defined for $v_1,v_2,v_3$.
-Therefore, we assign $\textcolor{blue}{<_{\sfrac{X+Y}{\sim}}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [\textcolor{blue}{<_{X+Y}}(\iota_{V_{X} + V_{Y}}(v_1))]$ and so on.
+such that $\textcolor{closed-color}{<_{X+Y}}$ is only defined for $v_1,v_2,v_3$.
+Therefore, we assign $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [\textcolor{closed-color}{<_{X+Y}}(\iota_{V_{X} + V_{Y}}(v_1))]$ and so on.
 This motivates the requirement for conditions (2), (3), and (4) of Theorem~\ref{th:existence_of_pushouts}, otherwise such inheritance would be ambiguous.
@@
 
-The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Figure~\ref{fig:base_cases}.
+The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Fig.~\ref{fig:base_cases}.
 Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.
@@
 
-Consider an example of DPOI rewriting in Figure~\ref{fig:dpoi-example} as defined in Definition~\ref{def:dpoi-e}.
+Consider an example of DPOI rewriting in Fig.~\ref{fig:dpoi-example} as defined in Definition~\ref{def:dpoi-e}.
 The rule in the top half of the diagram is an instance of a $\beta$-reduction rule as defined for string diagrams~\cite{ghica2024stringdiagramslambdacalculifunctional}.

### 2025-01-27 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
clean-up

@@
 In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
+A typical DPO square is depicted in Fig.~\ref{fig:dpo_dpoi}~(left).
 
@@
 
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
+To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Fig.~\ref{fig:dpo_dpoi}~(right).
 In the case of $ \catname{Hyp(\Sigma)}$,  we wish for a rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
@@
 \begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
+Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Fig.~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
 
@@
 \begin{definition}
-We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{orange}{<}, \textcolor{blue}{<},\consistency)$ where
+We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{e-color}{<}, \textcolor{closed-color}{<},\consistency)$ where
 \begin{itemize}
   \item $V$ is a set of vertices.
-  \item $E = \colorbox{yellow}{E} \cup \textcolor{orange}{E} \cup \textcolor{blue}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
+  \item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
   \item $s,t : E \to V^{*}$ are source and target functions.
   \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are \textit{label} functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
-  \item $\textcolor{orange}{<} \subseteq \textcolor{orange}{E} \times V + E$ and $\textcolor{blue}{<} \subseteq \textcolor{blue}{E} \times V + E $ with $\textcolor{orange}{E}, \textcolor{blue} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{orange}{<^{\mu}}$ and $\textcolor{blue}{<^{\mu}}$.
-        For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{orange}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{blue}{<^{\mu}} y_{i+1}\; \}$.
+  \item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times V + E$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times V + E $ with $\textcolor{e-color}{E}, \textcolor{closed-color} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
+        For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{<^{\mu}} y_{i+1}\; \}$.
         We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
@@
           \item each $x$ has at most one immediate predecessor;
-          \item edges $e$ such that there is no $x$ such that $e \textcolor{orange}{<^{\mu}} x$ or $e \textcolor{orange}{<^{\mu}} x$ are labelled;
-          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{orange}{<^{\mu}} e$ iff $e' \textcolor{orange}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{blue}{<}$.
+          \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{<^{\mu}} x$ or $e \textcolor{e-color}{<^{\mu}} x$ are labelled;
+          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{e-color}{<^{\mu}} e$ iff $e' \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<}$.
         \end{enumerate}
-  \item $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p \textcolor{orange}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{orange}{<^{\mu}}(v)$ and $p \textcolor{orange}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
-  We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{orange}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{orange}{<^{\mu}} e\}$.
+  \item $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{<^{\mu}}(v)$ and $p \textcolor{e-color}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
+  We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{<^{\mu}} e\}$.
 \end{itemize} 
-$\textcolor{orange}{E}$ and $\textcolor{blue}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{orange}{E} \cup \textcolor{blue}{E}$.
-Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{orange}{<}, \textcolor{blue}{<}$ and~$\consistency$.
+$\textcolor{e-color}{E}$ and $\textcolor{closed-color}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$.
+Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
 \end{definition}
 
-Intuitively, edges from $\textcolor{orange}{E}$ encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{blue}{E}$ encode lambda-abstraction boxes.
+Intuitively, edges from $\textcolor{e-color}{E}$ encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{closed-color}{E}$ encode lambda-abstraction boxes.
 Also note the typing on the $l_{V}$ function.
@@
 In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
-$\textcolor{blue}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{orange}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+$\textcolor{closed-color}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{e-color}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
 
-Consider an example in the middle part of Figure~\ref{fig:e-cospan-example}.
-The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{orange}{E} = \{e_{1}\}$, $\textcolor{blue}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
+Consider an example in the middle part of Fig.~\ref{fig:e-cospan-example}.
+The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{e-color}{E} = \{e_{1}\}$, $\textcolor{closed-color}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
 \[
@@
 \begin{array}{cccc}
-  v_{4} \textcolor{orange}{<^{\mu}} e_{1} & v_{7} \textcolor{blue}{<^{\mu}} e_{2} & v_{4} \consistency e_{2} & v_{4} \not \consistency v_{10}\\
-  e_{3} \textcolor{orange}{<^{\mu}} e_{1} & v_{8} \textcolor{blue}{<^{\mu}} e_{2} &  e_{2} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
-  e_{2} \textcolor{orange}{<^{\mu}} e_{1} & e_{5} \textcolor{blue}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
+  v_{4} \textcolor{e-color}{<^{\mu}} e_{1} & v_{7} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{4} \consistency e_{2} & v_{4} \not \consistency v_{10}\\
+  e_{3} \textcolor{e-color}{<^{\mu}} e_{1} & v_{8} \textcolor{closed-color}{<^{\mu}} e_{2} &  e_{2} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
+  e_{2} \textcolor{e-color}{<^{\mu}} e_{1} & e_{5} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
   \ldots & \ldots & \ldots & \ldots \\
-  v_{12} \textcolor{orange}{<^{\mu}} e_{1} & w_{9} \textcolor{blue}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
+  v_{12} \textcolor{e-color}{<^{\mu}} e_{1} & w_{9} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
 \end{array}
@@
 $\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
-Edges in $\textcolor{orange}{E}$ are depicted with a dashed box, while edges in $\textcolor{blue}{E}$ --- with a round solid box.
+Edges in $\textcolor{e-color}{E}$ are depicted with a dashed box, while edges in $\textcolor{closed-color}{E}$ --- with a round solid box.
 \begin{figure}
@@
       \item $\phi$ is a hypergraph homomorphism;
-      \item $\phi_{E}(\textcolor{orange}{E_{\mathcal{F}}}) \subseteq \textcolor{orange}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{blue}{E_{\mathcal{F}}}) \subseteq \textcolor{blue}{E_{\mathcal{G}}}$;
+      \item $\phi_{E}(\textcolor{e-color}{E_{\mathcal{F}}}) \subseteq \textcolor{e-color}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{closed-color}{E_{\mathcal{F}}}) \subseteq \textcolor{closed-color}{E_{\mathcal{G}}}$;
       \item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
         \begin{align*}
-          \text{if } e \textcolor{orange}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{orange}{<_{\mathcal{G}}^{\mu}} \phi(v)
+          \text{if } e \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(v)
           \qquad
-          \text{if } e \textcolor{blue}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(v)
+          \text{if } e \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(v)
         \end{align*}
@@
           \begin{align*}
-          \text{if } e_1 \textcolor{orange}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{orange}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+          \text{if } e_1 \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
           \qquad
-          \text{if } e_1 \textcolor{blue}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{blue}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+          \text{if } e_1 \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
           \end{align*}
@@
 Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-Graphically, we depict such extended cospans as in Figure~\ref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
+Graphically, we depict such extended cospans as in Fig.~\ref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
 All non-black vertices form strictly internal interfaces.
@@
 \begin{multline*}
-R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{blue}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{blue}{<}^{\mu} f_{\text{int}}(y)\bigr\}
+R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<}^{\mu} f_{\text{int}}(y)\bigr\}
 \end{multline*}
@@
 
-In Figure~\ref{fig:e-cospan-example} the partitions for input and output internal interfaces are given with colours.
+In Fig.~\ref{fig:e-cospan-example} the partitions for input and output internal interfaces are given with colours.
 
 Intuitively, such notion of isomorphism makes it possible to freely rearrange interfaces for disjoint components.
-For example, in the input interface of cospan in Figure~\ref{fig:e-cospan-example}, we can swap 
+For example, in the input interface of cospan in Fig.~\ref{fig:e-cospan-example}, we can swap 
 \begin{tikzpicture}
@@
   \begin{itemize}
-    \item For each $e \in \textcolor{orange}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
+    \item For each $e \in \textcolor{e-color}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
           $e$ is well-typed if for each element $S$ of the partition of $I$ $w(S) = w(s(e))$ and similarly for $O$ and $t(e)$.
-    \item For each $e \in \textcolor{blue}{E}$ consider sets $I$ and $O$ of its input and output vertices.
+    \item For each $e \in \textcolor{closed-color}{E}$ consider sets $I$ and $O$ of its input and output vertices.
           $e$ is well-typed if there exists an object $B \in \text{obj}_{\Sigma_{O}}$ such that $[w(s(e)), B] = w(I)$ and $B \multimap w(O) = w(t(e))$.
@@
 
-Consider, again, Figure ~\ref{fig:e-cospan-example}.
+Consider, again, Fig. ~\ref{fig:e-cospan-example}.
 To make edge $e_{1}$ well-typed, it is necessary that $l_{V}^{*}([v_{4},v_{5},v_{6}]) = l_{V}^{*}([v_{10},v_{11},v_{12}]) = l_{V}^{*}([v_1,v_2,v_3])$ and $l_{V}^{*}([w_{4},w_{5},w_{6}]) = l_{V}^{*}([w_{10},v_{12},v_{11}]) = l_{V}^{*}([w_1,w_2,w_3])$ (note the change of ordering in the output of $e_{1}$).
@@
 Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments, defined in this section.
-The construction in this section closely follows the one introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} with some adjustments to account for $\textcolor{blue}{<}$.
+The construction in this section closely follows the one introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} with some adjustments to account for $\textcolor{closed-color}{<}$.
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
@@
 % Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
+% Consider a hypothetical DPO square below in Fig.~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
 
@@
     \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{blue}{E}$ s.t.
-$v,w \textcolor{blue}{<}^{\mu} e$;
-    \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{blue}{E}$ s.t.
-$v,w \textcolor{blue}{<}^{\mu} e$;
+    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
+$v,w \textcolor{closed-color}{<}^{\mu} e$;
+    \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
+$v,w \textcolor{closed-color}{<}^{\mu} e$;
     \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and

### 2025-01-27 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
clean-up

@@
 It follows by first defining the interpretation of operators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in Appendix~\ref{sec:appendix:interpretation}.
-The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in Figure~\ref{fig:ev_and_lambda}.
+The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in Fig.~\ref{fig:ev_and_lambda}.
 
@@
 We also add a structural rewrite schema rule corresponding to distributivity law~\ref{law:distributivity}.
-The join of two cospans is defined as in Figure~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules, one for each axiom.
+The join of two cospans is defined as in Fig.~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules, one for each axiom.
 For example, the distributivity rule for $\otimes$ is given by the following schema rule
@@
 \label{prop:fscd}
-For two closed $\Sigma$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{blue}{E}$.
+For two closed $\Sigma$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
 \end{proposition}

### 2025-05-19 LICS/sections/appendix.tex
camera-ready and arxiv script

@@
 
-In this section we will define the category of semilattices that we use as a base for enrichment throughout the paper.
+In this section we define the category of semilattices that we use as a base for enrichment throughout the paper.
 
@@
 
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we will explicitly construct such a pushout.
-We will also show the uniqueness of a boundary pushout complement.
+In this section we define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists, and we explicitly construct such a pushout.
+We also show the uniqueness of a boundary pushout complement.
 These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
@@
 
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
+In this section we will elaborate on soundness and completeness of our translation $\llbracket - \rrbracket : \catname{S}^{+}(\Sigma, \mathcal{E}) \to \WellTypedMdaEcospans / \mathcal{S}, \mathcal{E}$.
+First we recite the interpretation function $[]$ on $\Sigma$-terms as defined in~\cite{bonchi_string_2022-2} that induces a functor $\catname{S}(\Sigma) \to \MdaCospans$ and eventually a functor $\catname{S}^{+}(\Sigma) \to \MdaCospans^{+}$.
+The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Figure~\ref{fig:base_cases}.
+Then it is defined inductively by letting $[f \otimes g] \Coloneq [f] \otimes [g]$ and $[f;g] \Coloneq [f];[g]$ and by freeness induces the corresponding functor.
+
+\begin{figure}
+\[
+\adjustbox{width=\textwidth}{
+\tikzfig{../figures/appendix/interpretation}
+}
+\]
+\caption{Base cases for $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$}
+\label{fig:base_cases}
+\end{figure}
+
+\begin{lemma}[Lemma~\ref{lemma:cospans_plus_equiv} restatement]
+
+    We have the following equivalence of $\catname{SLat}$-categories
+    \[
+    \MdaCospans^{+} \cong \MdaEcospans/\mathcal{S}~.
+    \]
+    \end{lemma}
+    \begin{proof}
+    We will denote the semilattice-enriched functor giving this equivalence as $M$.
+    Observe that in $\MdaCospans^{+}$ we can consider only canonical morphisms of the form $f_{1} + \ldots + f_{n}$ where each of $f_{i}$ is an MDA cospan (in the sense of Definition~\ref{def:monogamy_hyp}) and for $i \not = j$, $f_{i} \not = f_{j}$.
+    Then, $M$ is defined as follows.
+    \begin{itemize}
+    \item It maps each discrete hypergraph in $\MdaCospans^{+}$ to a corresponding discrete e-hypergraph in $\MdaEcospans/\mathcal{S}$.
+    \item Every morphism $f_{1} + \ldots + f_{n}$ is mapped to $\text{In}(f_{1}) + \ldots + \text{In}(f_{n})$ where $\text{In}$ is the faithful functor induced by Remark~\ref{remark:embedding_functor}.
+          This makes $M$ into $\catname{SLat}$-functor by construction.
+    \end{itemize}
+    To show the functor $M$ is faithful, suppose $f = f_{1} + \ldots + f_{n} \not = g_{1} + \ldots + g_{m} = g$.
+    Then it must be that $Mf \not = Mg$.
+    If $n \not = m$ then, immediately $Mf \not = Mg$ as $Mf$ and $Mg$ contain different number of non-isomorphic components.
+    If $n = m$ then assume for a contradiction that $Mf = Mg$. 
+    There exists a permutation $\sigma$, such that for every index $i$, $Mf_{i} = Mg_{\sigma(i)}$.
+    Thus, by $\text{In}$ we have that $f_{i} = g_{\sigma(i)}$, implying $f = g$ which is a contradiction.
+    Therefore, $Mf \not = Mg$ as required.
+
+    To show that the functor is full, consider a morphism $f = f_{1} + \ldots + f_{n}$ in $\WellTypedMdaEcospans / \mathcal{S}$ (by Lemma~\ref{lemma:normal_form}). 
+    As each of $f_{i}$ does not contain any hierarchical edges, they are in the image of $\text{In}$, which for each $f_{i}$ gives us $g_{i}$, such that $f_{i} = Mg_{i}$.
+    This gives us a morphism $g = g_{1} + \ldots + {g_{n}}$ in $\MdaCospans^{+}$ such that $Mg = f$.
+    This concludes that the functor is full.
+    As it is also surjective on objects, it makes it into an equivalence.
+\end{proof}
 
@@
                 which implies the existence of the following diagram where both squares are pushouts by construction and the matching is convex and down-closed.
-                \textcolor{red}{Consider cases when $\mathcal{L}$ has no-input and no-output sub-hypergraph and when it does not.
+                % \textcolor{red}{Consider cases when $\mathcal{L}$ has no-input and no-output sub-hypergraph and when it does not.
 Basically, repeat the above reasoning here as well, but say that if $n+m$ is empty, then there is no pushout.}
@@
     \end{itemize}
-    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
+    The rest of the argument follows by induction on the length of $w$ and ultimately we have shown that if $f = g$ modulo $\mathcal{E}$ then $\llbracket f \rrbracket = \llbracket g \rrbracket$  in $\WellTypedMdaEcospans / \mathcal{S}, \mathcal{E}$.
+
+    Then, faithfulness amounts to showing that if $\llbracket f \rrbracket \Rrightarrow_{\llbracket \mathcal{E} \rrbracket}^{*} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ then $f = g$ modulo $\mathcal{E}$ in $\catname{S}^{+}(\Sigma)$.
+% By the theorem~\ref{thm:completeness_simple} if $\llbracket f \rrbracket = \llbracket g \rrbracket$ via rules from $\mathcal{S}$ then $f = g$, so we will show that the same holds when $\llbracket f \rrbracket = \llbracket g \rrbracket$ via rules from $\mathcal{E}$.
+Consider a rewrite rule $\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle = \langle \mathcal{L}, \mathcal{R} \rangle$ such that $\llbracket f \rrbracket \Rrightarrow_{\langle \mathcal{L}, \mathcal{R} \rangle} \llbracket g \rrbracket$ and suppose
+\[
+	\llbracket f \rrbracket = n \xrightarrow{} n' \xrightarrow{} \mathcal{F} \xleftarrow{} m' \xleftarrow{} m ~.
+\]
+This implies the existence of a diagram as in~\ref{def:dpoi-e}.
+We will proceed by induction on the level of the occurrence of $\mathcal{L}$ within $\mathcal{F}$.
+\begin{itemize}
+	\item Suppose the image of $\mathcal{L}$ in $\mathcal{F}$ is top-level.
+		  Then, by lemma~\ref{lemma:decomposition}, there exists a decomposition of $\llbracket f \rrbracket$ in terms of MDA cospans
+		\begin{align*}
+			(k \xrightarrow{id_{k}} k \xrightarrow{id_{k}} &k \xleftarrow{id_{k}} k \xleftarrow{id_{k}} k)\\
+			n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k; \qquad\qquad\qquad &\otimes \qquad\qquad\qquad ; j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m\\
+			(i \xrightarrow{} i' \xrightarrow{} &\mathcal{L} \xleftarrow{} j' \xleftarrow{} j)
+		\end{align*}
+		By fullness of $\llbracket - \rrbracket$ (from Theorem~\ref{thm:completeness_simple}) there exist $c_{1}$ and $c_{2}$ such that 
+		\[
+		\llbracket c_{1} \rrbracket = n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k \qquad \llbracket c_{2} \rrbracket = j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m
+		\]
+		By functoriality, $\llbracket f \rrbracket = \llbracket c_{1};(id \otimes l);c_{2} \rrbracket$ and by faithfulness we have that $f = c_{1};(id \otimes l);c_{2}$ to which we can apply $\langle l, r \rangle$ and get $g' = c_{1};(id \otimes r);c_{2}$.
+		We then have $\llbracket g' \rrbracket = \llbracket c_{1};(id \otimes r);c_{2} \rrbracket =  \llbracket g \rrbracket$ since the boundary complement is unique (Proposition ~\ref{prop:boundary_unique}) and $g' = g$ by faithfulness.
+
+		\item If the image of $\mathcal{L}$ is not top-level, then either $\llbracket f \rrbracket$ is decomposable as
+		\begin{align*}
+			(k \xrightarrow{id_{k}} k \xrightarrow{id_{k}} &k \xleftarrow{id_{k}} k \xleftarrow{id_{k}} k)\\
+			n \xrightarrow{} n_{1}' \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m_{1}' + i + k \xleftarrow{} i + k; \qquad\qquad\qquad &\otimes \qquad\qquad\qquad ; j + k \xrightarrow{} j + k + n'_{2} \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_{2} \xleftarrow{} m\\
+			(i \xrightarrow{} i' \xrightarrow{} &\mathcal{F'} \xleftarrow{} j' \xleftarrow{} j)
+		\end{align*}
+		such that $\mathcal{L}$ is a convex down-closed sub-e-hypergraph of $F'$.
+		Since the image of $\mathcal{L}$ lies within $\mathcal{F}'$, there is an EDPO rewrite step that turns
+		\[
+			(i \xrightarrow{} i' \xrightarrow{} \mathcal{F'} \xleftarrow{} j' \xleftarrow{} j)	
+		\]
+		into
+		\[
+			(i \xrightarrow{} i'' \xrightarrow{} \mathcal{G'} \xleftarrow{} j'' \xleftarrow{} j)	
+		\]
+		via $\langle \mathcal{L}, \mathcal{R} \rangle$.
+		By fullness and inductive hypothesis, there exist $f'$ and $g'$ such that 
+		\[
+			\llbracket f' \rrbracket = i \xrightarrow{} i' \xrightarrow{} \mathcal{F'} \xleftarrow{} j' \xleftarrow{} j \qquad \llbracket g' \rrbracket = (i \xrightarrow{} i'' \xrightarrow{} \mathcal{G'} \xleftarrow{} j'' \xleftarrow{} j)
+		\]
+		and that $f' = g'$.
+		By decomposition and fullness we have $\llbracket f \rrbracket = \llbracket c_{1};(id \otimes f');c_{2} \rrbracket$ and $f = c_{1};(id \otimes f');c_{2}$ by faithfulness.
+		By applying $f' = g'$ we get $f = g'' = c_{1};(id \otimes g');c_{2}$.
+		Finally, $\llbracket g'' \rrbracket = \llbracket c_{1};(id \otimes g');c_{2} \rrbracket = \llbracket g \rrbracket$ (by the uniqueness of the boundary complement) and $g'' = g = f$ by faithfulness.
+
+		\item or, $\llbracket f \rrbracket$ is decomposable as
+		\begin{align*}
+			n_1 \xrightarrow{} n_1' \xrightarrow{} &\mathcal{F}_1 \xleftarrow{} m_1' \xleftarrow{} m_1\\
+			&\;+\\
+			&\vdotswithin{+}\\
+			&\;+\\
+			n_p \xrightarrow{} n_p' \xrightarrow{} &\mathcal{F}_p \xleftarrow{} m_p' \xleftarrow{} m_p\\
+			&\;+\\
+			&\vdotswithin{+}\\
+			&\;+\\
+			n_k \xrightarrow{} n_k' \xrightarrow{} &\mathcal{F}_k \xleftarrow{} m_k' \xleftarrow{} m_k
+		\end{align*}
+		such that the image of $\mathcal{L}$ lies within $\mathcal{F}_{p}$.
+		By inductive hypothesis we have that $\llbracket f_{p} \rrbracket = n_p \xrightarrow{} n_{p}' \xrightarrow{} \mathcal{F}_{p} \xleftarrow{} m_{p}' \xleftarrow{} m_{p}$ rewrites into $\llbracket g_{p} \rrbracket = n_p \xrightarrow{} n_{p}'' \xrightarrow{} \mathcal{G}_{p} \xleftarrow{} m_{p}'' \xleftarrow{} m_{p}$ such that $f_{p} = g_{p}$.
+		By fullness and the fact that $\llbracket - \rrbracket$ is enriched, there exist $\llbracket f_1 + \ldots + f_{p} + \ldots + f_{k} \rrbracket = \llbracket f_{1} \rrbracket + \ldots + \llbracket f_{p} \rrbracket + \ldots + \llbracket f_{k} \rrbracket = \llbracket f \rrbracket$.
+		By applying $f_{p} = g_{p}$ we get $f = g''$ and $\llbracket g'' \rrbracket = \llbracket f_{1} + \ldots + g_{p} + \ldots f_{k} \rrbracket = \llbracket g \rrbracket$ and $f = g'' = g$.
+\end{itemize}
     % there is a sequence of SMC and enrichment equations such that when applied to $f$ it is turned into $g$.
@@
 \label{sec:appendix:iso}
+
+First we will explore the notion of isomorphism of two extended cospans in more detail.
+Our example will be Figure~\ref{fig:appendix:non-isomorphic-cospans}.
+Technically, we do not allow hierarchical edges with no delimiting dashed line, but here, for the purposes of illustration we make an exception to make the example compact.
 \begin{example}
@@
 
+\begin{figure}
+    \begin{subfigure}{0.48\textwidth}
+\[
+\adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/boundary_complement_non_example}
+}
+\]
+\subcaption{Boundary complement non-example}
+    \end{subfigure}
+\hfill
+    \begin{subfigure}{0.48\textwidth}
+        \[
+        \adjustbox{scale=0.55}{
+            \tikzfig{../figures/appendix/boundary_complement_non_example_2}
+        }
+        \]
+            \subcaption{Boundary complement non-example}
+            \end{subfigure}
+\begin{subfigure}{\textwidth}
+    \[
+    \adjustbox{scale=0.55}{
+        \tikzfig{../figures/appendix/boundary_complement_example}
+    }
+    \]
+        \subcaption{Complete DPO square}
+        \end{subfigure}
+\caption{DPO examples}
+\label{fig:not_boundary_complement}
+\end{figure}
+
 \section{DPO rewriting for arbitrary signatures}
@@
 Assume we want to perform a rewrite within the e-hypergraph in Figure~\ref{fig:dpo-stuck} (left).
+However, we are immediately stuck as there is no matching of the left-hand side of the rule within the graph as there is no pushout complement: the only way we can impose a constraint that an edge shall have a particular predecessor in the pushout is through its adjacent vertices, but the sub-hypergraph for $a;!$ does not have any as demonstrated in the right subfigure.
 
@@
 \end{figure}
-
-However, we are immediately stuck as there is no matching of the left-hand side of the rule within the graph as there is no pushout complement: the only way we can impose a constraint that an edge shall have a particular predecessor in the pushout is through its adjacent vertices, but the sub-hypergraph for $a;!$ does not have any as demonstrated in the right subfigure.
 
@@
 
-\section{Going from e-graph to e-hypergraph}
+The Definition~\ref{def:boundary_new} will then need a few additional constraints.
+\begin{itemize}
+\item For all top-level edges $e_{i}, e_{j}$ of type $0 \to 0$ in $\mathcal{L}$ it must be either $[m(e_{i})) = [m(e_{j})) = \varnothing$ or $m(e_{i}) \consistency m(e_{j})$.
+\item For all edges $e_{i}, e_{j}$ of type $0 \to 0$ in $i + j$ it must be either $[[c_1,c_2](e_{i})) = [[c_1,c_2](e_{j})) = \varnothing$ or $[c_1,c_2](e_{i}) \consistency [c_1,c_2](e_{j})$.
+\item For all edges $e$ of type $0 \to 0$ in $i + j$ it must be that $l([c_1,c_2](e)) = l(e) = \bot$ and $[c_1,c_2](e)$ must be maximal.
+\end{itemize}
+Then, when constructing the corresponding extended cospans, such auxiliary edges must be ignored in the interfaces.
+
+\section{E-graph interpretation in $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$}
+\label{sec:appendix:e-graph-translation}
+
+In this section we will elaborate on the interpretation of e-graphs as morphisms of $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$.
+We will first recite the formal definition of an e-graph.
 
@@
 \begin{figure}
-    \begin{subfigure}{0.4\linewidth}
+    \begin{subfigure}[C]{0.4\linewidth}
 \[
@@
 \]
+\caption{\;}
 \end{subfigure}
-\begin{subfigure}{0.4\linewidth}
+\hfill
+\begin{subfigure}[C]{0.4\linewidth}
     \[
@@
     \]
+\caption{\;}
 \end{subfigure}
+\caption{E-graph (a) and a corresponding morphism of $\catname{MEHypI}(\Sigma \cup \Sigma_{C})$ (b)}
+\label{fig:e-graph-to-cospan}
 \end{figure}
@@
 % Then, given e-classes $e_1$ and $e_2$ for sub-hypergraphs $C$ and $D$ respectively, we introduce e-nodes into the e-graph $A(e_1)$ and $B(e_2)$ that reside in e-class $e_3$.
+
+\section{Additional figures}
+
+Additional translation steps
+
+\begin{figure}
+    \[
+    \adjustbox{width=\linewidth}{
+    \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-c-e}
+    }
+    \]
+    \caption{Translation (c)-(d)}
+    \end{figure}
+\begin{figure}
+
+\[
+\adjustbox{width=\linewidth}{
+\tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-d-e}
+}
+\]
+\caption{Translation (d)-(e)}
+\end{figure}

### 2025-05-19 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
camera-ready and arxiv script

@@
 
-\section{Introduction}
-\label{sec:introduction}
-
-\subsection{Motivation and contribution}
-
-E-graphs (short for \emph{equality graphs})~\cite{EggPaper} are a data structure incorporating equivalence representation and used in automated theorem proving, compiler optimisation, and program analysis.
-Using e-graphs mitigates the so-called \emph{phase-ordering problem} of term rewriting, that is the fact that the way rewrites are scheduled may enable or block further rewrites due to introduction or elimination or redexes.
-This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule, instead of replacing the left-hand side by the right-hand side, adds the latter to the graph so that all pre-existing redexes are protected and new ones are introduced.
-Explosive growth of the graph is avoided by the clever use of sharing of vertices.
-
-E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using so-called technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint is reached.
-The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
-Of particular interest to us is using e-graphs in the context of functional programming languages, which builds upon an internal representation of the $\lambda$-calculus. 
-Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
-the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise applying equality saturation to $\lambda$-calculus e-graphs.
-
-In the textual representation of $\lambda$-terms, $\alpha$-equivalent terms are distinct and thus cannot be shared, which adds a layer of bookkeeping and complexity.
-The traditional approach to handle this is by using de Bruijn indices~\cite{de1972lambda}, for which $\alpha$-equivalent terms are syntactically identified; however, this is also problematic because the same variable can be assigned different indices depending on its scope with respect to that instance of the variable.
-This again interferes with sharing. 
-As a simple example, consider the term $\lambda f . f ((\lambda x . f x) 2)$ which, with de Bruijn indices,  is encoded as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$.
-The two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$, which cannot be shared in an e-graph.
-
-Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
-The implementations of $\beta$-reduction in e-graphs do not go into much detail regarding explicit substitution, and in particular do not include rules like $[x\!\mapsto\!1][y\!\mapsto\!2]f\!=\![y\!\mapsto\!2][x\!\mapsto\!1]f$ for non-interfering substitutions, while such rules are absorbed by graph equivalence~\cite{accattoli2014nonstandard}.
-
-All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
-\cite{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
-In this paper, we extend this technique to e-graphs with bindings, lifting to monoidal \emph{closed} categories.
-
-\subsection{String diagrams}
-
-We build on the technology of string diagrams, by now well-developed in a number of areas. 
-String diagrams are a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
-These categorical techniques provide insight in how to solve the problems of binding and capture-avoiding substitution in e-graphs, as we develop in this paper.
-
-The general background literature for string diagrams is summarised in~\cite{piedeleu2023introductionstringdiagramscomputer}.
-The particular context of string diagrams with bindings is given in~\cite{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-
-String diagrams provide a topological framework for understanding monoidal categories, representing objects as \emph{strings} and morphisms as \emph{nodes}, where input strings flow in and output strings flow out.
-Key structures in monoidal categories---such as the monoidal product, tensor unit, braiding, and dualities---are represented graphically, through operations like juxtaposing, weaving, or bending strings.
-This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by traditional symbolic notation. 
-Originating from notations by Hotz~\cite{hotzsd} (for automata) and Penrose~\cite{penrose1984spinors} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, and provide a mathematically rigorous diagrammatic calculus~\cite{joyal_geometry_1991}.
-
-Commonly, string diagrams are used informally to provide a two-dimensional syntax for the internal language of a category, but for our purposes we must be more formal. 
-Specifically, we require a concrete encoding of string diagrams, analogous to the de Bruijn index representation of $\lambda$-terms.
-For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics for this purpose.
-For instance, hypergraphs admit the property that two string diagrams are equal if and only if they ``look the same'' (\emph{i.e.} they admit a graph isomorphism); secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.
-
-Finally, for the diagrammatical representation of two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
-The diagrammatic boxes require the introduction of further structure to (hyper)graphs in the form of a hierarchical layering of the graphs, as studied in \emph{loc. cit.}.
-
-\begin{figure}
-  \centering
-  \adjustbox{scale=0.6}{
-      \tikzfig{figures/de-brujin-string}
-  }
-\caption{String-diagrammatic representation of $\lambda f . f ((\lambda x . f x) 2)$.}\label{fig:de-brujin-string}
-\end{figure}
-
-Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, in Fig.~\ref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
-$\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire attached to the frame,
-the application node a half-circle
-\adjustbox{scale=0.15}{\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=none] (0) at (1, 6.75) {};
-		\node [style=none] (1) at (1.25, 7.75) {};
-		\node [style=none] (2) at (2, 8) {};
-		\node [style=none] (3) at (2.75, 7.75) {};
-		\node [style=none] (4) at (3, 6.75) {};
-		\node [style=none] (5) at (1.5, 6.75) {};
-		\node [style=none] (6) at (2.5, 6.75) {};
-		\node [style=none] (7) at (2, 9) {};
-		\node [style=none] (8) at (1.5, 5.75) {};
-		\node [style=none] (9) at (2.5, 5.75) {};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw [style=lambda_unit_string] (3.center)
-			 to [in=60, out=-30, looseness=0.75] (4.center)
-			 to (0.center)
-			 to [in=-150, out=120, looseness=0.75] (1.center)
-			 to [in=-180, out=45, looseness=0.75] (2.center)
-			 to [in=135, out=0, looseness=0.75] cycle;
-		\draw (2.center) to (7.center);
-		\draw (5.center) to (8.center);
-		\draw (9.center) to (6.center);
-	\end{pgfonlayer}
-\end{tikzpicture}
-}, 
-and sharing a \emph{contraction node} 
-\scalebox{0.3}{
-	\begin{tikzpicture}
-		\begin{pgfonlayer}{nodelayer}
-			\node [style=vertex] (0) at (1.5, 3.5) {};
-			\node [style=none] (1) at (1.5, 3) {};
-			\node [style=none] (2) at (1, 4) {};
-			\node [style=none] (3) at (2, 4) {};
-		\end{pgfonlayer}
-		\begin{pgfonlayer}{edgelayer}
-			\draw (0) to (1.center);
-			\draw [bend right=45] (0) to (3.center);
-			\draw [bend left=45] (0) to (2.center);
-		\end{pgfonlayer}
-	\end{tikzpicture}
-}
-.
-
-\subsection{E-graphs}
-
-E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Fig.~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
-
-\begin{figure*}[h!]
-\[
-\adjustbox{scale=0.6}{
-\tikzfig{figures/e-graph-example}
-}
-\]
-\caption{E-graph example (top) and its equivalent string diagram representation (bottom)}
-\label{fig:e-graph-example}
-\end{figure*}
-
-Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node.
-
-Nodes of an e-graph are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
-A term graph is embedded as an e-graph by making each equivalence class have a single node.
-The first rewrite rule, replacing multiplication by \emph{shift-left} creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) nodes,  now sharing the node `$a$'.
-This second e-graph illustrates the other representational difference, besides sharing, between conventional e-graphs and e-hypergraphs.
-In the former, edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes are depicted with black dots with a single wire
-\scalebox{0.4}{
-	\begin{tikzpicture}
-		\begin{pgfonlayer}{nodelayer}
-			\node [style=vertex] (0) at (1.5, 3.5) {};
-			\node [style=none] (1) at (1.5, 3) {};
-		\end{pgfonlayer}
-		\begin{pgfonlayer}{edgelayer}
-			\draw (0) to (1.center);
-		\end{pgfonlayer}
-	\end{tikzpicture}
-}
-inside the dashed boxes, indicating which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rewrite rules.
-
-A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
-The auxiliary nature of these nodes is revealed by the fact that, after rewriting, they must be removed from the graph. 
-We will see how in the categorical representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.
-
-\subsection{Conventional vs. categorical e-graphs with binding}
-
-First, let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
-In Fig.~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
-A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box), to be rewritten using the rules for explicit substitution at the top of each column.
-At the end, all explicit substitution nodes can be removed.
-The same example is then given again in Fig.~\ref{fig:e-string-substitution}, this time using string diagrams.
-
-Using string diagrams, $\beta$-reduction is simply the re-wiring of the diagram where an application node meets an abstraction, and explicit substitution simplifies into mere composition. 
-This is appears to be simpler just by comparing diagrams, but the technical advantages of using string diagrams run deep. 
-The explicit occurrences of variable names in the older approach raise obvious questions of scope, binding and explicit substitution, whereas the use of terms inside the explicit substitution nodes also raises many technical difficulties which makes proving various mathematical properties essentially intractable.
-Additionally, explicit substitutions need to be propagated recursively throughout the graph, which adds another layer of technical and conceptual overhead. 
-Using de Bruijn indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
-
-\begin{figure*}
-	\begin{subfigure}{\linewidth}
-    \[
-        % \hspace{1.3cm}
-        \adjustbox{width=\textwidth}{
-        \tikzfig{figures/e-graph-substitution}
-        }
-    \]
-	% \captionsetup{skip=0pt}
-    \subcaption{E-graph explicit substitution example.}
-    \label{fig:e-graph-substitution}
-	\end{subfigure}
-	\begin{subfigure}{\linewidth}
-	\[
-        % \hspace{1.3cm}
-        \adjustbox{width=\textwidth}{
-        \tikzfig{figures/e-string-substitution-2}
-        }
-    \]
-	\subcaption{String diagrammatic substitution example.}\label{fig:e-string-substitution}
-	\end{subfigure}
-\end{figure*}
-  %   \begin{figure*}
-%     \[
-%         % \hspace{1.3cm}
-%         \adjustbox{width=\textwidth}{
-%         \tikzfig{figures/e-string-substitution-2}
-%         }
-%     \]
-% 	\captionsetup{skip=0pt}
-%     \caption{String diagrammatic substitution example.}
-%     \label{fig:e-string-substitution}
-%   \end{figure*}
-
-\subsection{Contributions}
-
-In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
-We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
-That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
-We call these string diagrams, or more precisely, the combinatorial representation of them, e-graphs with bindings.
-
-We then define a combinatorial representation for such string diagrams in terms of hierarchical hypergraphs with a suitable notion of DPO rewriting.
-Finally, we show that this notion of rewriting is sound and complete with respect to rewriting of morphism terms.
-
-% DRG: LIST THE MAIN DEFINITIONS AND THEOREMS AND WHAT THEY MEAN
-
+\input{sections/introduction}
 \input{sections/categorical_semantics}

### 2025-05-19 LICS_2025/monoidal_e_graphs/sections/appendix.tex
wip
camera-ready and arxiv script

@@
+\ifdefined\ONECOLUMN
+% \section{}
+\section{$\catname{SLat}$}
+\else
 \subsection{$\catname{SLat}$}
+\fi
 \label{sec:appendix:slat}
@@
 
+    Note that this construction does not identify $(s_1,s_2) +_{S_1 \otimes S_2} (s_1',s_2')$ and $(s_1 +_{S_1} s_1', s_2 +_{S_2} s_2')$.
     The unit for this tensor product is $I = \{*\}$ --- a one-element semilattice.
@@
 
-  This makes $\catname{SLat}$ a suitable base for enrichment and allows us to take $\mathcal{V} = \catname{SLat}$.
+  This makes $\catname{SLat}$ a suitable base for enrichment.
+%    and allows us to take $\mathcal{V} = \catname{SLat}$.
 
@@
       The functor $F$ is defined by letting $F(A) = \coprod_{A} I_{\catname{SLat}}$ where the latter is a coproduct which is a `free' product of semilattices defined as follows.
-      The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$.
+      The elements of $S_{1} \coprod S_{2}$ are sequences $s_{1} + s_{2} + \ldots + s_{n}$ where each $s_{i}$ is either from $S_{1}$ or $S_{2}$ quotiented by all relations in $S_{1}$ and $S_{2}$ and by the necessary equations to turn it into a semilattice.
       The adjunction is then given by the following natural isomorphisms
@@
       \end{align*}
-      The fist isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
+      The first isomorphism is given by the fact that hom-functor makes limits into colimits in its first argument, the second being given by the property that $|\catname{Set}(A,B)| = |B|^{|A|} = |\underbrace{B \times \ldots \times B}_{|A|}|$, and the last isomorphism is given by the definition of $U$.
       Furthermore, we have, 
@@
       \begin{tikzcd}  
-          & {f + g \otimes *} && {\triangle_{A} \otimes ((f+g) \times (f+g))} \\
-          f + g &&&& {\triangle_{A'};((f + g) \times (f+g)) = (f + g);\triangle_{A'};} \\
+          & {f + g \otimes *} && {\triangle_{A} \otimes ((f+g) \times (f+g))}\\
+          f + g &&&& {\triangle_{A'};((f + g) \times (f+g)) = (f + g);\triangle_{A'}}\\
           & {* \otimes f + g} && {f + g \otimes \triangle_{A'}}
@@
 
+\ifdefined\ONECOLUMN
+\section{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
+\else
 \subsection{Proofs for Section \ref{sec:e-hypergraphs}: E-hypergraphs}
+\fi
 \label{sec:appendix:pushout}
@@
 
+\ifdefined\ONECOLUMN
+\section{Existence of pushout}
+\else
 \subsection{Existence of pushout}
+\fi
 
@@
     \begin{align*}    
-    S_{V} &= \{\\
-          &\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+    S_{V} = \{&
+          \;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
           &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}\\
-          &\cup\\
-          &\{\\
-          &\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\}\\
+          \cup&\\
+          \{&
+          \;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
           &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}\\
-          &\cup\\
-          &\{\\
-          &\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\\
-          &\}\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\}\\
+          \cup&\\
+          \{
+          &\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\}\\
         \\
@@
     \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = (&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = (&
+            \sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
             &\sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},\\
@@
             &\consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},\\
-            &<_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        &)    
+            &<_{\sfrac{X + Y}{\sim (R_{V},R_{E})}})    
     \end{align*}
@@
               \fi
-              by recalling that $[[]_{V}^{\consistency}, []_{E}^{\consistency}] = [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]$.
+              by recalling that 
+              \[
+              [[]_{V}^{\consistency}, []_{E}^{\consistency}] = [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]
+              \].
     \end{enumerate}
@@
 
+\begin{lemma}
+$\consistency_{\sfrac{X+Y}{\sim}}^{p}$ is not total.
+\end{lemma}
+\begin{proof}
+The property holds as we require that the arrows of the span for the pushout do not map vertices to different components and therefore they can not be merged into one component.
+\end{proof}
+
 \begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
@@
 
+\ifdefined\ONECOLUMN
+\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
+\else
 \subsection{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
+\fi
 \label{sec:appendix:detailed_proofs}
@@
               The latter trivially yields the EDPO diagram below
-              \[
+              \ifdefined\ONECOLUMN
+              \[
+              \adjustbox{scale=0.8}{
                 \begin{tikzcd}
@@
                 \end{tikzcd}
-              \]
+              }
+              \]
+              \else
+              \[
+                \begin{tikzcd}
+                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                    \\
+                    {\mathcal{L}} && {i+j} && {\mathcal{R}} \\
+                    & {i'+j'} && {i''+j''} \\
+                    && {i+j}
+                    \arrow[from=1-1, to=3-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-3, to=1-2]
+                    \arrow[from=1-3, to=1-4]
+                    \arrow[from=1-3, to=3-3]
+                    \arrow[from=1-4, to=1-5]
+                    \arrow[from=1-5, to=3-5]
+                    \arrow["\lrcorner"{pos=0.05, rotate=90, description}, phantom, from=3-1, to=1-2]
+                    \arrow[from=3-3, to=3-1]
+                    \arrow[from=3-3, to=3-5]
+                    \arrow["\lrcorner"{pos=0.05, rotate=180, description}, phantom, from=3-5, to=1-4]
+                    \arrow[from=4-2, to=3-1]
+                    \arrow[from=4-4, to=3-5]
+                    \arrow[from=5-3, to=3-3]
+                    \arrow[from=5-3, to=4-2]
+                    \arrow[from=5-3, to=4-4]
+                \end{tikzcd}
+              \]
+              \fi
               where $i \to i' \to \mathcal{L} \xleftarrow{} j' \xleftarrow{} j = \llbracket l \rrbracket = \llbracket f \rrbracket$ and $i \to i'' \to \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j = \llbracket r \rrbracket = \llbracket g \rrbracket$ and $i + j \to i + j \to \mathcal{L}$ is a boundary complement as per~\ref{def:boundary_new}.
@@
 
+              \ifdefined\ONECOLUMN
+              \[
+              \adjustbox{width=0.7\linewidth}{
+              \begin{tikzcd}
+                {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                \\
+                {\llbracket \mathcal{C}[l]\rrbracket} && {\mathcal{L}_1^{\bot}} && {\llbracket \mathcal{C}[r]\rrbracket} \\
+                & {n_1'+m_1'} & {n_1+m_1} & {n_1''+m_1''} \\
+                {\llbracket f \rrbracket} && {\mathcal{L}_2^{\bot}} && {\llbracket g \rrbracket} \\
+                & {n_2' + m_2'} && {n_2''+m_2''} \\
+                && {n_2 + m_2}
+                \arrow[from=1-1, to=3-1]
+                \arrow[from=1-2, to=1-1]
+                \arrow[from=1-3, to=1-2]
+                \arrow[from=1-3, to=1-4]
+                \arrow[from=1-3, to=3-3]
+                \arrow[from=1-4, to=1-5]
+                \arrow[from=1-5, to=3-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=90, description}, draw=none, from=3-1, to=1-2]
+                \arrow[from=3-1, to=5-1]
+                \arrow[from=3-3, to=3-1]
+                \arrow[from=3-3, to=3-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=180, description}, draw=none, from=3-5, to=1-4]
+                \arrow[from=3-5, to=5-5]
+                \arrow[from=4-2, to=3-1]
+                \arrow[from=4-3, to=3-3]
+                \arrow[from=4-3, to=4-2]
+                \arrow[from=4-3, to=4-4]
+                \arrow[from=4-3, to=5-3]
+                \arrow[from=4-4, to=3-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=90, description}, draw=none, from=5-1, to=4-2]
+                \arrow[from=5-3, to=5-1]
+                \arrow[from=5-3, to=5-5]
+                \arrow["\lrcorner"{pos=0.025, rotate=180, description}, draw=none, from=5-5, to=4-4]
+                \arrow[from=6-2, to=5-1]
+                \arrow[from=6-4, to=5-5]
+                \arrow[from=7-3, to=5-3]
+                \arrow[from=7-3, to=6-2]
+                \arrow[from=7-3, to=6-4]
+            \end{tikzcd}}
+            \]
+              \else
               \[
@@
             \]
+            \fi
             Clearly the occurrence of $\llbracket \mathcal{C}[l] \rrbracket$ in $\llbracket f \rrbracket$ is convex and down-closed and by following the argument in Theorem 35~\cite{bonchi_string_2022-2}, there exists $n_1 + m_1 \xrightarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_2^{\bot} \xleftarrow{} n_3 + m_3 \xleftarrow{} n_2 + m_2$ such that
@@
             and
-            \[
-            n_2 + m_1 \xrightarrow{} n_2' \setminus (n_1' \setminus n_1) + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot} \xleftarrow{} m'_2 \setminus (m_1' \setminus m_1) + n_1 \xleftarrow{} m_2 + n_1    
-            \] is an MDA cospan.
+            % fix overflowing
+            % \[
+            % n_2 + m_1 \xrightarrow{} n_2' \setminus (n_1' \setminus n_1) + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot} \xleftarrow{} m'_2 \setminus (m_1' \setminus m_1) + n_1 \xleftarrow{} m_2 + n_1
+            % \]
+            \tikzcdset{row sep/mysizerow/.initial=0.1ex}
+            \tikzcdset{column sep/mysizecol/.initial=0.1ex}
+            \[\
+\begin{tikzcd}[cramped, row sep=mysizerow, column sep = mysizecol]
+                n_2 + m_1 &&&&&&&& m_2 + n_1\\
+                & \hspace{-2em} n_2' \setminus (n_1' \setminus n_1) + m_1 &\; & & \mathcal{L}_{2}^{\bot} &\; && m'_2 \setminus (m_1' \setminus m_1) + n_1\hspace{-2em} &\\
+                \arrow[from=1-1, to=2-2]
+                \arrow[from=1-9, to=2-8]
+                \arrow[from=2-2, to=2-5]
+                \arrow[from=2-8, to=2-5]
+            \end{tikzcd}
+\]
+            is an MDA cospan.
             By denoting the pushout of $\mathcal{L}_1^{\bot} \xleftarrow{} n_1 + m_1 \xrightarrow{} \mathcal{L}_{2}^{\bot}$ with $\mathcal{L}_3^{\bot}$,
@@
             By hypothesis, there exists a diagram
+            \ifdefined\ONECOLUMN
+            \[
+                \adjustbox{width=0.7\linewidth}{
+                \begin{tikzcd}
+                    {\mathcal{L}} & {i'+j'} & {i+j} & {i''+j''} & {\mathcal{R}} \\
+                    \\
+                    {\llbracket\mathcal{C}[l]\rrbracket} && {\mathcal{L}_1^{\bot}} && {\llbracket\mathcal{C}[r]\rrbracket} \\
+                    & {n'+m'} && {n''+m''} \\
+                    && {n+m}
+                    \arrow[from=1-1, to=3-1]
+                    \arrow[from=1-2, to=1-1]
+                    \arrow[from=1-3, to=1-2]
+                    \arrow[from=1-3, to=1-4]
+                    \arrow[from=1-3, to=3-3]
+                    \arrow[from=1-4, to=1-5]
+                    \arrow[from=1-5, to=3-5]
+                    \arrow[from=3-3, to=3-1]
+                    \arrow[from=3-3, to=3-5]
+                    \arrow[from=4-2, to=3-1]
+                    \arrow[from=4-4, to=3-5]
+                    \arrow[from=5-3, to=3-3]
+                    \arrow[from=5-3, to=4-2]
+                    \arrow[from=5-3, to=4-4]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=90}, draw=none, from=3-1, to=1-2]
+                    \arrow["\lrcorner"{description, pos=0.025, rotate=180}, draw=none, from=3-5, to=1-4]
+                \end{tikzcd}}
+                \]
+            \else
                 \[
@@
                 \]
+                \fi
                 which implies the existence of the following diagram where both squares are pushouts by construction and the matching is convex and down-closed.
@@
     % \]
+    \ifdefined\ONECOLUMN
+    \[
+    n \xrightarrow{} n'_{1} \xrightarrow{} \mathcal{C}_{1} \xleftarrow{} m'_{1} + i + k \xleftarrow{} i+k \xrightarrow{} i' + k \xrightarrow{} \mathcal{L} \xleftarrow{} k + j' \xleftarrow{} k + j \xrightarrow{} n'_2 + k + j \xrightarrow{} \mathcal{C}_{2} \xleftarrow{} m'_2 \xleftarrow{} m
+    \]
+    \else
     \[
@@
     \]
+    \fi
     The two spans identify precisely those nodes from $\mathcal{G}$ that occur in more than one sub-hypergraph, so this amounts to simply taking the union
+    \ifdefined\ONECOLUMN
+    \begin{align*}
+        n \to n_{1}' \cup n_{2}' \cup i' \to \mathcal{C}_{1} \cup \mathcal{C}_{2} \cup \mathcal{L} \xleftarrow{} m_{1}' \cup m_{2}' \cup j' \xleftarrow{} m  = n \xrightarrow{f} n' \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} m' \xleftarrow{g} m
+    \end{align*}
+    \else
     \begin{align*}
@@
     \end{align*}
+    \fi
     If we let $\mathcal{G}' = \mathcal{L}$ we get the decomposition as in~(\ref{fig:decomposition:1}).
@@
 
+\ifdefined\ONECOLUMN
+\section{Additional figures for Section~\ref{sec:introduction}:Introduction}
+\else
 \subsection{Additional figures for Section~\ref{sec:introduction}:Introduction}
+\fi
 
@@
 
+\ifdefined\ONECOLUMN
+\section{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
+\else
 \subsection{Additional figures for Section~\ref{sec:e-hypergraphs}: E-hypergraphs}
+\fi
 \label{sec:appendix:iso}
@@
 
+\ifdefined\ONECOLUMN
+\section{DPO rewriting for arbitrary signatures}
+\else
 \subsection{DPO rewriting for arbitrary signatures}
+\fi
 \label{sec:dpo-fix}
@@
 Another problematic rewrite arises when we want to utilise the structural rule of the free enrichment
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.6\linewidth}{
+\tikzfig{../figures/appendix/structural_rewrite_problem}
+}
+\]
+\else
 \[
@@
 \]
+\fi
 if we consider the following DPO square
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.8\linewidth}{
+    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck}
+}
+\]
+\else
 \[
@@
 \]
+\fi
 
@@
 We will define the action of $+$ on two cospans where either of the carriers is the empty e-hypergraph as
-
 \begin{align*}
@@
 \end{align*}
-
 \end{definition}
@@
 
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.7\linewidth}{
+    \tikzfig{../figures/appendix/structural_rewrite_problem_fix}
+}
+\]
+\else
 \[
@@
 
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.8\linewidth}{
+    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
+}
+\]
+\else
 \[
@@
 
+\ifdefined\ONECOLUMN
 \begin{figure}[h!]
@@
         \[
-    \adjustbox{scale=0.55}{
+        \adjustbox{scale=0.7}{
         \tikzfig{../figures/appendix/idI_elimination_hypergraph}
@@
         \[
+        \adjustbox{scale=0.7}{
+        \tikzfig{../figures/appendix/idI_introduction_hypergraph}
+        }
+        \]
+    \end{subfigure}
+    \end{figure}
+\else
+\begin{figure}[h!]
+\begin{subfigure}{0.45\linewidth}
+    \[
+    \adjustbox{scale=0.55}{
+    \tikzfig{../figures/appendix/idI_elimination_hypergraph}
+    }
+    \]
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\linewidth}
+    \[
     \adjustbox{scale=0.55}{
@@
 \end{figure}
-
-and similar rules to introduce and eliminate in nested contexts (Figure~\ref{fig:nested_elimination}).
+\fi
+and similar rules to introduce and eliminate in nested contexts (Figure~\ref{fig:nested_elimination});
 
@@
 
+\ifdefined\ONECOLUMN
+\begin{figure}[h!]
+    \begin{subfigure}{0.6\linewidth}
+        \[
+        \adjustbox{width=\linewidth}{
+        \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
+        }
+        \]
+        \subcaption{}
+    \end{subfigure}
+    \begin{subfigure}{0.6\linewidth}
+        \[
+        \adjustbox{width=\linewidth}{
+        \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
+        }
+        \]
+        \subcaption{}
+    \end{subfigure}
+    \caption{New consistent component introduction (a) and elimination (b) rule}
+    \label{fig:consistent_introduction}
+\end{figure}
+\else
 \begin{figure}[h!]
@@
         \end{subfigure}
-    \caption{New consistent component introduction rule}
+    \caption{New consistent component introduction (top) and elimination (bottom) rule}
     \label{fig:consistent_introduction}
 \end{figure}
+\fi
 
@@
 
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.6\linewidth}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_fix}
+}
+\]
+\else
 \[
@@
 \]
-
+\fi
 and the full DPO square becomes
 
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.7\linewidth}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix}
+}
+\]
+\else
 \[
@@
 \]
-
+\fi
 The target e-hypergraph is not exactly what we wanted, however, but we can use the auxiliary rule to remove this unlabelled edge.
 
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.7\linewidth}{
+    \tikzfig{../figures/appendix/cartesian_rewrite_dpo_fix_2}
+}
+\]
+\else
 \[
@@
 \]
-
+\fi
 % and then utilise the idempotence rule for the identity of the tensor unit
@@
 
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.7\linewidth}{
+    \tikzfig{../figures/appendix/rule_transformation_example}
+}
+\]
+\else
 \[
@@
 \]
-
+\fi
 assuming that $|c^{l}_{1,1}, \ldots, c^{l}_{k,n}| > |c^{r}_{1,1}, \ldots, c^{r}_{p,m}|$ and $f = \{ c^{l}_{1,1} \mapsto c^{l}_{1,1}, \ldots, c^{l}_{k,n} \mapsto c^{l}_{k,n} \}$ and $g$ is any valid mapping of $c^{l}_{1,1}, \ldots, c^{l}_{k,n}$ to $c^{r}_{1,1}, \ldots, c^{r}_{p,m}$.
@@
 
+\ifdefined\ONECOLUMN
+\section{E-graph interpretation in $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$}
+\else
 \subsection{E-graph interpretation in $\catname{MEHypI}(\Sigma_{C}) / \mathcal{S}, \mathcal{E}_{C}$}
+\fi
 \label{sec:appendix:e-graph-translation}

### 2025-05-19 LICS_2025/monoidal_e_graphs/sections/combinatorial_semantics.tex
camera-ready and arxiv script

@@
 
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
+First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over $\catname{Set}$.
 We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $R$ to a relation on ordered sequences. 
@@
 \begin{figure}[t!]
-    \vspace{-0.5cm}
-    \begin{subfigure}[T]{0.35\linewidth}
+    \begin{subfigure}[T]{0.35\columnwidth}
     \[
@@
                     \arrow[from=1-3, to=2-3]
-                    \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=90}, draw=none, from=2-1, to=1-2]
+                    \arrow["{\lrcorner}"{rotate=90, xshift=-0.25em, yshift=0.9em,font = \Large}, draw=none, from=2-1, to=1-2]
                     \arrow[from=2-2, to=2-1]
                     \arrow[from=2-2, to=2-3]
-                    \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=180}, draw=none, from=2-3, to=1-2]
+                    \arrow["{\lrcorner}"{rotate=180,xshift=-0.5em, yshift=2.35em,font = \Large}, draw=none, from=2-3, to=1-2]
 \end{tikzcd}
@@
     \end{subfigure}
-    \hfill
-    \begin{subfigure}[T]{0.35\linewidth}
+    % \hfill
+    \hspace{0.15\columnwidth}
+    \begin{subfigure}[T]{0.35\columnwidth}
         \[
@@
         \arrow[from=1-3, to=2-3]
-        \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=90}, draw=none, from=2-1, to=1-2]
+        \arrow["{\lrcorner}"{rotate=90,xshift=-0.25em, yshift=1em,font = \Large}, draw=none, from=2-1, to=1-2]
         \arrow[from=2-2, to=2-1]
         \arrow[from=2-2, to=2-3]
-        \arrow["\lrcorner"{anchor=center, pos=0.125, rotate=180}, draw=none, from=2-3, to=1-2]
+        \arrow["{\lrcorner}"{rotate=180,xshift=-0.9em, yshift=2.25em,font = \Large}, draw=none, from=2-3, to=1-2]
         \arrow[from=3-2, to=2-1]
@@
     \end{subfigure}
-    \hfill
-    \captionsetup{skip=0pt, belowskip=0pt}
     \caption{DPO and DPOI squares}
     \label{fig:dpo_dpoi}
-    \vspace{-7mm}
 \end{figure}

### 2025-05-19 LICS_2025/monoidal_e_graphs/sections/e-hypergraphs.tex
camera-ready and arxiv script

@@
 An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}, for which we introduce the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} of $x$ is denoted as $[x) = \{x' ~|~ x' < x\; \}$.
+For $x \in V + E$,  the set of \emph{parents} (or, \emph{predecessors}) of $x$ is denoted as $[x) = \{x' ~|~ x' < x\; \}$.
 % \[
@@
 % \]
-We will denote the immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
+We will denote the immediate parent (predecessor) $e$ of $x$ by writing $e <^{\mu} x$, and likewise $x$ is then called an immediate \emph{child} (or, \emph{successor}).
 We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
@@
 	\item each $x$ has at most one immediate parent;
-	\item the maximal edges must be labelled;
+	\item for all $e$ such that $e$ is maximal, $l(e) \not = \bot$;
 	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$ and similarly if $v \in t(e)$.
@@
 \begin{array}{ccc}
-    v_3 <^{\mu} e_1 & v_3 \consistency v_4 & v_3 \not \consistency v_5\\
+    e_1 <^{\mu} v_3 & v_3 \consistency v_4 & v_3 \not \consistency v_5\\
     \ldots & \ldots & \ldots\\
-    w_4 <^{\mu} e_1 & w_3 \consistency w_4 & v_4 \not \consistency w_4
+    e_1 <^{\mu} w_4 & w_3 \consistency w_4 & v_4 \not \consistency w_4
 \end{array}
@@
 The condition $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$, in particular, does not allow hierarchical edges with no delimiting lines.
-$e_1$ is the maximal edge and hence labelled.
+Finally, $e_1$ is the maximal edge and hence labelled.
 \begin{definition}[E-hypergraph homomorphism]
@@
     0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\;
+\]
+\[
     0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
@@
     \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended MDA cospan
+    % \[
+    % \hspace{-2em} n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+    % \]
+\tikzcdset{row sep/mysize/.initial=0.25ex}.
 \[
-    \hspace{-2em} n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+\hspace{-2em}
+\begin{tikzcd}[cramped, row sep=mysize, column sep = tiny]
+	{n + j} &&&&&& {k + i} \\
+	& {n' \setminus (i' \setminus i) + j} & \vspace{16em} & {\mathcal{L}^{\bot}} & \vspace{16em} & {k' \setminus (j' \setminus j) + i}
+    \arrow["{f_1 + id_{j}}"{xshift=-0.7ex}, from=1-1, to=2-2]
+	\arrow["{f_2 + id_{i}}"'{xshift=1ex}, from=1-7, to=2-6]
+	\arrow["{[g_1,c_2]}"{yshift=0.5ex}, from=2-2, to=2-4]
+	\arrow["{[g_2,c_1]}"'{yshift=0.5ex}, from=2-6, to=2-4]
+\end{tikzcd}
 \]

### 2025-05-19 LICS_2025/monoidal_e_graphs/sections/introduction.tex
camera-ready and arxiv script

@@
 Our approach is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3,fscd,ghica-zanassi2023string}.
-The key ingredient of our categorical approach is the semilattice-enrichment\footnote{We could alternatively use the language of idempotent commutative semigroups, which form an equivalent category, but stick to the \enquote{semilattice} terminology for brevity.}.
+The key ingredient of our categorical approach is the semilattice-enrichment.
+% \footnote{We could alternatively use the language of idempotent commutative semigroups, which form an equivalent category, but stick to the \enquote{semilattice} terminology for brevity.}
 By using the binary operator of a given semilattice we can operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
@@
 
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.
+The other steps, (c) to (d), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules.
 If we were to apply another rule, $x * 1 \to x$, we would identify the e-class (dashed box) containing $a$ with the e-class containing $a * 1$ (the topmost dashed box) which would require a cycle to be introduced: the left argument of node $*$ would be the same e-class that $*$ belongs to.
@@
 The equations applied are the ones from Figure~\ref{fig:string-equations} along with the naturality of delete and copy.
-Note that the size of the eventual string diagram is comparable to the size of the corresponding e-graph.
+Note that the degree of sharing in the eventual string diagram is comparable to the degree of sharing in the corresponding e-graph.
 % Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
@@
 \label{fig:e-graph-to-string}
-\vspace{-8mm}
 \end{figure}
@@
 	such that certain coherence diagrams commute~\cite{Borceux_1994}.
-	$F$ being a morphism in $\catname{SLat}$ has a property $F(f + g) = F(f) + F(g)$.
+	$F_{A,B}$ being a morphism in $\catname{SLat}$ has a property $F_{A,B}(f + g) = F_{A,B}(f) + F_{A,B}(g)$.
 	An $\catname{SLat}$-functor $F$ defines an equivalence between two $\catname{SLat}$-categories if it is \textit{full}, \textit{faithful}, and \textit{essentially surjective}~\cite{Kelly2022BASICCO}.

### 2025-05-19 arxiv.py

@@ -1,0 +1,118 @@
+import shutil
+from pathlib import Path
+from pylatexenc.latexwalker import LatexWalker, LatexEnvironmentNode, LatexMacroNode
+from pylatexenc.latex2text import LatexNodes2Text
+import sys
+
+FIGURE_EXTENSIONS = ['.pdf', '.png', '.jpg', '.jpeg', '.eps', '.tikz']
+FIGURE_MACROS = {'includegraphics', 'tikzfig'}
+INPUT_MACROS = {'input', 'include'}
+BIB_MACROS = {'bibliography'}
+
+def find_macros_recursively(latex_code):
+    walker = LatexWalker(latex_code)
+    nodelist, pos, len_ = walker.get_latex_nodes()
+    return extract_macros(nodelist)
+
+def extract_macros(nodes):
+    macros = []
+    nodes_iter = iter(nodes)
+    while True:
+        try:
+            node = next(nodes_iter)
+            if isinstance(node, LatexMacroNode):
+                if node.macroname in FIGURE_MACROS | INPUT_MACROS | BIB_MACROS:
+                    if node.macroname == r'tikzfig':
+                        node = next(nodes_iter)
+                        macros.append((r'tikzfig', LatexNodes2Text().nodelist_to_text(node.nodelist)))
+                    else:                        
+                        arg = node.nodeargd.argnlist[0]
+                        if arg is not None:
+                            macros.append((node.macroname, LatexNodes2Text().nodelist_to_text(arg.nodelist)))
+            elif hasattr(node, 'nodelist'):
+                macros.extend(extract_macros(node.nodelist))
+            elif hasattr(node, 'nodeargd'):
+                for arg in node.nodeargd.argnlist:
+                    if hasattr(arg, 'nodelist'):
+                        macros.extend(extract_macros(arg.nodelist))
+        except StopIteration:
+            break
+    # for node in nodes:
+    #     if isinstance(node, LatexMacroNode):
+    #         if node.macroname in FIGURE_MACROS | INPUT_MACROS | BIB_MACROS:
+    #             arg = node.nodeargd.argnlist[0]
+    #             if arg is not None:
+    #                 macros.append((node.macroname, LatexNodes2Text().nodelist_to_text(arg.nodelist)))
+    #     elif hasattr(node, 'nodelist'):
+    #         macros.extend(extract_macros(node.nodelist))
+    #     elif hasattr(node, 'nodeargd'):
+    #         for arg in node.nodeargd.argnlist:
[…]

### 2025-06-12 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
notes for slotted e-graphs and arxiv script

@@
 As pushout is only defined up to isomorphism, it justifies the morphisms in this category being isomorphism classes of cospans.
-To make such cospans complete with respect to representing morphism of $\catname{S}(\Sigma)$, we need to impose some restriction on their form.
+To make such cospans complete with respect to representing morphisms of $\catname{S}(\Sigma)$, we need to impose some restrictions on their form.
 \begin{definition}[Degree,  path,  acyclic] 
@@
 
-Practical meaning of the above theorem is that two $\Sigma$-terms $f$ and $g$ are equal modulo SMC equations if and only if they are represented as the same hypergraph with interfaces.
-The theorem defines a full and faithful functor that is induced by interpretation function of morphism terms that renders the latter as cospans of hypergraphs and is given in Appendix~\ref{sec:appendix:interpretation}.
+The practical meaning of the above theorem is that two $\Sigma$-terms $f$ and $g$ are equal modulo SMC equations if and only if they are represented as the same hypergraph with interfaces.
+The theorem defines a full and faithful functor that is induced by the interpretation function of morphism terms that renders the latter as cospans of hypergraphs and is given in Appendix~\ref{sec:appendix:interpretation}.
 Term rewriting and hence equality modulo $\mathcal{E}$ is then implemented as DPO rewrites over such hypergraphs with interfaces that we recall next.
@@
 We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness, and when evoking a term e-hypergraph modify the original definitions as follows.
+We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness, and when evoking a term `e-hypergraph' we will mean the structure defined as follows.
 
@@
   \item $V$ is a set of vertices.
-  \item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that that is formed of three disjoint sets of hyperedges defined below.
+  \item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that is formed of three disjoint sets of hyperedges defined below.
   \item $s,t : E \to V^{*}$ are source and target functions.
@@
 Consider a relation
+\ifdefined\ONECOLUMN
+\[
+    R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(y)\bigr\}
+\]
+\else 
 \begin{multline*}
-R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<}^{\mu} f_{\text{int}}(y)\bigr\}
+R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(y)\bigr\}
 \end{multline*}
+\fi
 for $x, y \in n'$.
@@
     \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
-$v,w \textcolor{closed-color}{<}^{\mu} e$;
+$v,w \textcolor{closed-color}{<^{\mu}} e$;
     \item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
-$v,w \textcolor{closed-color}{<}^{\mu} e$;
+$v,w \textcolor{closed-color}{<^{\mu}} e$;
     \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
@@
     \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended MDA cospan
+    \ifdefined\ONECOLUMN
+    \[
+        n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+    \]
+    \else
     \begin{multline*}
@@
     \end{multline*}
+    \fi
 \end{enumerate}

### 2025-06-12 LICS_2025/e_graphs_with_bindings/sections/introduction.tex
notes for slotted e-graphs and arxiv script

\section{Introduction}
\label{sec:introduction}

\subsection{Motivation and contribution}

E-graphs (short for \emph{equality graphs})~\cite{EggPaper} are a data structure incorporating equivalence representation and used in automated theorem proving, compiler optimisation, and program analysis. 
Using e-graphs mitigates the so-called \emph{phase-ordering problem} of term rewriting, that is the fact that the way rewrites are scheduled may enable or block further rewrites due to introduction or elimination or redexes. 
This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule, instead of replacing the left-hand side by the right-hand side, adds the latter to the graph so that all pre-existing redexes are protected and new ones are introduced. 
Explosive growth of the graph is avoided by the clever use of sharing of vertices.

E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using so-called technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint is reached. 
The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
Of particular interest to us is using e-graphs in the context of functional programming languages, which builds upon an internal representation of the $\lambda$-calculus. 
Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise applying equality saturation to $\lambda$-calculus e-graphs.

In the textual representation of $\lambda$-terms, $\alpha$-equivalent terms are distinct and thus cannot be shared, which adds a layer of bookkeeping and complexity.
The traditional approach to handle this is by using de Bruijn indices~\cite{de1972lambda}, for which $\alpha$-equivalent terms are syntactically identified; however, this is also problematic because the same variable can be assigned different indices depending on its scope with respect to that instance of the variable.
This again interferes with sharing. 
As a simple example, consider the term $\lambda f . f ((\lambda x . f x) 2)$ which, with de Bruijn indices,  is encoded as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$.
The two different occurrences of the same bound variable $f$ are represented differently by $\%0$ and $\%1$, which cannot be shared in an e-graph.

Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
The implementations of $\beta$-reduction in e-graphs do not go into much detail regarding explicit substitution, and in particular do not include rules like $[x\!\mapsto\!1][y\!\mapsto\!2]f\!=\![y\!\mapsto\!2][x\!\mapsto\!1]f$ for non-interfering substitutions, while such rules are absorbed by graph equivalence~\cite{accattoli2014nonstandard}.

All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
\cite{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
In this paper, we extend this technique to e-graphs with bindings, lifting to monoidal \emph{closed} categories.

\subsection{String diagrams}

We build on the technology of string diagrams, by now well-developed in a number of areas. 
String diagrams are a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
These categorical techniques provide insight in how to solve the problems of binding and capture-avoiding substitution in e-graphs, as we develop in this paper.

The general background literature for string diagrams is summarised in~\cite{piedeleu2023introductionstringdiagramscomputer}.
The particular context of string diagrams with bindings is given in~\cite{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.

String diagrams provide a topological framework for understanding monoidal categories, representing objects as \emph{strings} and morphisms as \emph{nodes}, where input strings flow in and output strings flow out. 
Key structures in monoidal categories---such as the monoidal product, tensor unit, braiding, and dualities---are represented graphically, through operations like juxtaposing, weaving, or bending strings. 
This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by traditional symbolic notation. 
Originating from notations by Hotz~\cite{hotzsd} (for automata) and Penrose~\cite{penrose1984spinors} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, and provide a mathematically rigorous diagrammatic calculus~\cite{joyal_geometry_1991}.

Commonly, string diagrams are used informally to provide a two-dimensional syntax for the internal language of a category, but for our purposes we must be more formal. 
Specifically, we require a concrete encoding of string diagrams, analogous to the de Bruijn index representation of $\lambda$-terms.
For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics for this purpose.
For instance, hypergraphs admit the property that two string diagrams are equal if and only if they ``look the same'' (\emph{i.e.} they admit a graph isomorphism); secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.

Finally, for the diagrammatical representation of two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}. 
The diagrammatic boxes require the introduction of further structure to (hyper)graphs in the form of a hierarchical layering of the graphs, as studied in \emph{loc. cit.}.

\begin{figure}
  \centering
  \adjustbox{scale=0.6}{
      \tikzfig{figures/de-brujin-string}
  }
\caption{String-diagrammatic representation of $\lambda f . f ((\lambda x . f x) 2)$.}\label{fig:de-brujin-string}
\end{figure}

Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, in Fig.~\ref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
$\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire attached to the frame,
the application node a half-circle
\adjustbox{scale=0.15}{\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=none] (0) at (1, 6.75) {};
		\node [style=none] (1) at (1.25, 7.75) {};
		\node [style=none] (2) at (2, 8) {};
		\node [style=none] (3) at (2.75, 7.75) {};
		\node [style=none] (4) at (3, 6.75) {};
		\node [style=none] (5) at (1.5, 6.75) {};
		\node [style=none] (6) at (2.5, 6.75) {};
		\node [style=none] (7) at (2, 9) {};
		\node [style=none] (8) at (1.5, 5.75) {};
		\node [style=none] (9) at (2.5, 5.75) {};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=lambda_unit_string] (3.center)
			 to [in=60, out=-30, looseness=0.75] (4.center)
			 to (0.center)
			 to [in=-150, out=120, looseness=0.75] (1.center)
			 to [in=-180, out=45, looseness=0.75] (2.center)
			 to [in=135, out=0, looseness=0.75] cycle;
		\draw (2.center) to (7.center);
		\draw (5.center) to (8.center);
		\draw (9.center) to (6.center);
	\end{pgfonlayer}
\end{tikzpicture}
}, 
and sharing a \emph{contraction node} 
\scalebox{0.3}{
	\begin{tikzpicture}
		\begin{pgfonlayer}{nodelayer}
			\node [style=vertex] (0) at (1.5, 3.5) {};
			\node [style=none] (1) at (1.5, 3) {};
			\node [style=none] (2) at (1, 4) {};
			\node [style=none] (3) at (2, 4) {};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw (0) to (1.center);
			\draw [bend right=45] (0) to (3.center);
			\draw [bend left=45] (0) to (2.center);
		\end{pgfonlayer}
	\end{tikzpicture}
}
.

\subsection{E-graphs}

E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Fig.~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.

\begin{figure*}[h!]
\[
\adjustbox{scale=0.6}{
\tikzfig{figures/e-graph-example}
}
\]
\caption{E-graph example (top) and its equivalent string diagram representation (bottom)~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}}
\label{fig:e-graph-example}
\end{figure*}

Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}. 
Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node.

Nodes of an e-graph are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
A term graph is embedded as an e-graph by making each equivalence class have a single node.
The first rewrite rule, replacing multiplication by \emph{shift-left} creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) nodes,  now sharing the node `$a$'.
This second e-graph illustrates the other representational difference, besides sharing, between conventional e-graphs and e-hypergraphs.
In the former, edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
Explicit discarding (weakening) nodes are depicted with black dots with a single wire
\scalebox{0.4}{
	\begin{tikzpicture}
		\begin{pgfonlayer}{nodelayer}
			\node [style=vertex] (0) at (1.5, 3.5) {};
			\node [style=none] (1) at (1.5, 3) {};
		\end{pgfonlayer}
		\begin{pgfonlayer}{edgelayer}
			\draw (0) to (1.center);
		\end{pgfonlayer}
	\end{tikzpicture}
}
inside the dashed boxes, indicating which of the class-level edges are connected to which nodes inside the class.

The other steps, (c) to (d), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rewrite rules.

A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
The auxiliary nature of these nodes is revealed by the fact that, after rewriting, they must be removed from the graph. 
We will see how in the categorical representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.

\subsection{Conventional vs. categorical e-graphs with binding}

First, let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
In Fig.~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box), to be rewritten using the rules for explicit substitution at the top of each column.
At the end, all explicit substitution nodes can be removed.
The same example is then given again in Fig.~\ref{fig:e-string-substitution}, this time using string diagrams.

Using string diagrams, $\beta$-reduction is simply the re-wiring of the diagram where an application node meets an abstraction, and explicit substitution simplifies into mere composition. 
This is appears to be simpler just by comparing diagrams, but the technical advantages of using string diagrams run deep. 
The explicit occurrences of variable names in the older approach raise obvious questions of scope, binding and explicit substitution, whereas the use of terms inside the explicit substitution nodes also raises many technical difficulties which makes proving various mathematical properties essentially intractable. 
Additionally, explicit substitutions need to be propagated recursively throughout the graph, which adds another layer of technical and conceptual overhead. 
Using de Bruijn indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
Notably, a work of~\cite{slotted-egraphs} has emerged recently which introduced a formulation of e-graphs with native support for binding, \textit{i.e.,} where binding is a first-class operation along with a substitution.
It significantly complicates the formulation of an e-graph and corresponding algorithmics and based on \textit{nominal} techniques~\cite{Pitts_2013} unlike the approach presented in this work.
We believe the approach used in this work strikes with its simplicity although a more detailed comparison with the work of remains a future work.

\begin{figure*}
	\begin{subfigure}{\linewidth}
    \[
        % \hspace{1.3cm}
        \adjustbox{width=\textwidth}{
        \tikzfig{figures/e-graph-substitution}
        }
    \]
	% \captionsetup{skip=0pt}
    \subcaption{E-graph explicit substitution example.}
    \label{fig:e-graph-substitution}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\[
        % \hspace{1.3cm}
        \adjustbox{width=\textwidth}{
        \tikzfig{figures/e-string-substitution-2}
        }
    \]
	\subcaption{String diagrammatic substitution example.}\label{fig:e-string-substitution}
	\end{subfigure}
\end{figure*}
  %   \begin{figure*}
%     \[
%         % \hspace{1.3cm}
%         \adjustbox{width=\textwidth}{
%         \tikzfig{figures/e-string-substitution-2}
%         }
%     \]
% 	\captionsetup{skip=0pt}
%     \caption{String diagrammatic substitution example.}
%     \label{fig:e-string-substitution}
%   \end{figure*}

\subsection{Contributions}

In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
We call these string diagrams, or more precisely, the combinatorial representation of them, e-graphs with bindings.

We then define a combinatorial representation for such string diagrams in terms of hierarchical hypergraphs with a suitable notion of DPO rewriting.
Finally, we show that this notion of rewriting is sound and complete with respect to rewriting of morphism terms.

% DRG: LIST THE MAIN DEFINITIONS AND THEOREMS AND WHAT THEY MEAN

### 2025-06-12 LICS_2025/monoidal_e_graphs/sections/appendix.tex
notes for slotted e-graphs and arxiv script

@@
             % \]
-            \tikzcdset{row sep/mysizerow/.initial=0.1ex}
+            \tikzcdset{row sep/mysizerow/.initial=0.3ex}
             \tikzcdset{column sep/mysizecol/.initial=0.1ex}
@@
 \begin{tikzcd}[cramped, row sep=mysizerow, column sep = mysizecol]
-                n_2 + m_1 &&&&&&&& m_2 + n_1\\
+                n_2 + m_1\hspace{-2em} &&&&&&&& \hspace{-2em}m_2 + n_1\\
                 & \hspace{-2em} n_2' \setminus (n_1' \setminus n_1) + m_1 &\; & & \mathcal{L}_{2}^{\bot} &\; && m'_2 \setminus (m_1' \setminus m_1) + n_1\hspace{-2em} &\\
@@
             \[
-            \scalebox{0.75}{\tikzfig{../figures/appendix/contexts_3}}    
+            \hspace{-1em}\scalebox{0.75}{\tikzfig{../figures/appendix/contexts_3}}    
             \]
@@
 
-\begin{figure*}
+\begin{figure*}[t!]
     \begin{subfigure}{0.4\linewidth}
@@
 \end{figure*}
+
+\begin{figure*}[!t]
+    \centering
+    \begin{minipage}[t]{0.48\linewidth}
+        \centering
+            \begin{subfigure}{\linewidth}
+                \[
+                \adjustbox{width=0.95\linewidth}{
+                \tikzfig{../figures/appendix/idI_elimination_nested_hypergraph}
+                }
+                \]
+            \end{subfigure}
+            \vspace{1em}
+            \begin{subfigure}{\linewidth}
+                \[
+                \adjustbox{width=0.95\linewidth}{
+                \tikzfig{../figures/appendix/idI_introduction_nested_hypergraph}
+                }
+                \]
+            \end{subfigure}
+            \caption{Special edge nested introduction (top) and elimination (bottom) rules}
+            \label{fig:nested_elimination}
+    \end{minipage}%
+    \hfill
+    \begin{minipage}[t]{0.48\linewidth}
+        \centering
+            \begin{subfigure}{\linewidth}
+                \[
+                \adjustbox{width=0.95\linewidth}{
+                \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
+                }
+                \]
+            \end{subfigure}
+            \vspace{1em}
+                \begin{subfigure}{\linewidth}
+                    \[
+                    \adjustbox{width=0.95\linewidth}{
+                    \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
+                    }
+                    \]
+                \end{subfigure}
+            \caption{New consistent component introduction (top) and elimination (bottom) rules}
+            \label{fig:consistent_introduction}        
+    \end{minipage}
+\end{figure*}
+\begin{figure}[t!]
+\[
+\adjustbox{width=\linewidth}{
+    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
+}
+\]
+\caption{Complete DPO square for free enrichment structural rule}
+\label{fig:completed_dpo}
+\end{figure}
+
 Another problematic rewrite arises when we want to utilise the structural rule of the free enrichment
@@
 For example, the edge in the interface labelled $a_1$ is mapped to the outermost edge on the right.
-We can now compute the pushout complement and complete the DPO square below
+We can now compute the pushout complement and complete the DPO square \ifdefined\ONECOLUMN below \else in Figure~\ref{fig:completed_dpo}. \fi
 
@@
 \else
-\[
-\adjustbox{width=\linewidth}{
-    \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
-}
-\]
+% \[
+% \adjustbox{width=\linewidth}{
+%     \tikzfig{../figures/appendix/structural_rewrite_dpo_stuck_fix}
+% }
+% \]
 \fi
@@
 Handling deletion or introduction rewrite rules will require the introduction of auxiliary rules.
-In particular, two rules for introduction and elimination of these special edges
-
+In particular, two rules for introduction and elimination of these special edges (Figure~\ref{fig:idI_intro_elim}),
 \ifdefined\ONECOLUMN
@@
 \end{subfigure}
+\caption{Special edge flat introduction and elimination}
+\label{fig:idI_intro_elim}
 \end{figure}
 \fi
-and similar rules to introduce and eliminate in nested contexts (Figure~\ref{fig:nested_elimination});
-
-\begin{figure}[h!]
-    \begin{subfigure}{\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
-        \tikzfig{../figures/appendix/idI_elimination_nested_hypergraph}
-        }
-        \]
-    \end{subfigure}
-    \vspace{1em}
-    \begin{subfigure}{\linewidth}
-        \[
-        \adjustbox{scale=0.55}{
-        \tikzfig{../figures/appendix/idI_introduction_nested_hypergraph}
-        }
-        \]
-    \end{subfigure}
-    \caption{Special edge nested elimination}
-    \label{fig:nested_elimination}
-\end{figure}
-
+similar rules to introduce and eliminate in nested contexts (Figure~\ref{fig:nested_elimination});
+% \begin{figure}[t!]
+%     \begin{subfigure}{\linewidth}
+%         \[
+%         \adjustbox{scale=0.55}{
+%         \tikzfig{../figures/appendix/idI_elimination_nested_hypergraph}
+%         }
+%         \]
+%     \end{subfigure}
+%     \vspace{1em}
+%     \begin{subfigure}{\linewidth}
+%         \[
+%         \adjustbox{scale=0.55}{
+%         \tikzfig{../figures/appendix/idI_introduction_nested_hypergraph}
+%         }
+%         \]
+%     \end{subfigure}
+%     \caption{Special edge nested elimination}
+%     \label{fig:nested_elimination}
+% \end{figure}
 and, finally, the rules to introduce such edges as new consistent components (Figure~\ref{fig:consistent_introduction}).
@@
 \else
-\begin{figure}[h!]
-    \begin{subfigure}{\linewidth}
-        \[
-        \adjustbox{width=\linewidth}{
-        \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
-        }
-        \]
-    \end{subfigure}
-    \vspace{1em}
-        \begin{subfigure}{\linewidth}
-            \[
-            \adjustbox{width=\linewidth}{
-            \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
-            }
-            \]
-        \end{subfigure}
-    \caption{New consistent component introduction (top) and elimination (bottom) rule}
-    \label{fig:consistent_introduction}
-\end{figure}
+% \begin{figure}[t!]
+%     \begin{subfigure}{\linewidth}
+%         \[
+%         \adjustbox{width=\linewidth}{
+%         \tikzfig{../figures/appendix/idI_introduction_new_component_hypergraph}
+%         }
+%         \]
+%     \end{subfigure}
+%     \vspace{1em}
+%         \begin{subfigure}{\linewidth}
+%             \[
+%             \adjustbox{width=\linewidth}{
+%             \tikzfig{../figures/appendix/idI_elimination_new_component_hypergraph}
+%             }
+%             \]
+%         \end{subfigure}
+%     \caption{New consistent component introduction (top) and elimination (bottom) rule}
+%     \label{fig:consistent_introduction}
+% \end{figure}
 \fi
@@
 \else
+\begin{figure}[h!]
 \[
@@
 \]
+\caption{Complete DPO square for the deletion rule}
+\label{fig:cartesian_deletion_dpo}
+\end{figure}
 \fi
@@
 \else
+\begin{figure}
 \[
@@
 \]
+\caption{Complete DPO square for the deletion rule (after one more rule applied)}
+\label{fig:cartesian_dpo_fix}
+\end{figure}
 \fi

### 2025-06-12 LICS_2025/monoidal_e_graphs/sections/e-hypergraphs.tex
notes for slotted e-graphs and arxiv script

@@
 This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section.
-A discussion of a general case can be found in section~\ref{sec:dpo-fix} of Appendix.
+A discussion of a general case can be found in \ifcameraready{section~\ref{sec:dpo-fix} of Appendix}{\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 \end{remark}
@@
 The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.  
-A concrete description of the pushout of two morphisms in this category is given in Appendix~\ref{sec:appendix:pushout}.
+A concrete description of the pushout of two morphisms in this category is given in \ifcameraready{Appendix~\ref{sec:appendix:pushout}}{\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does in all scenarios that we are interested in; most importantly, the pushout for the composition of two cospans with discrete feet always exists.
@@
 with each other --- this would yield a non-isomorphic cospan.
-Further examples of isomorphic and non-isomorphic extended cospans can be found in Appendix~\ref{sec:appendix:iso}.
+Further examples of isomorphic and non-isomorphic extended cospans can be found in \ifcameraready{Appendix~\ref{sec:appendix:iso}}{\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 
@@
 This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa.
-Examples and non-examples of a boundary complement can be found in Appendix~\ref{sec:appendix:iso}.
+Examples and non-examples of a boundary complement can be found in \ifcameraready{Appendix~\ref{sec:appendix:iso}}{\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 % \begin{remark}
@@
 Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ for reflexive, symmetric, and transitive closure of $\Rrightarrow{}$.
-An example of rewriting of extended cospans can be found in Appendix~\ref{sec:appendix:iso}.
+An example of rewriting of extended cospans can be found in \ifcameraready{Appendix~\ref{sec:appendix:iso}}{\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 % Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.

### 2025-06-12 LICS_2025/monoidal_e_graphs/sections/introduction.tex
notes for slotted e-graphs and arxiv script

@@
 Note that while the transformations of the e-graph are guided by the algorithm, the transformations of the string diagrammatic version are guided by the equations of the underlying theory.
-A detailed step-by-step application of the equations for the rewrite rules from $(a)$ to $(b)$ and $(b)$ to $(c)$ is given in Figures~\ref{fig:e-graph-example-a-b} and~\ref{fig:e-graph-example-b-c}, respectively.
+A detailed step-by-step application of the equations for the rewrite rules from $(a)$ to $(b)$ and $(b)$ to $(c)$ \ifcameraready{is given in Figures~\ref{fig:e-graph-example-a-b} and~\ref{fig:e-graph-example-b-c}, respectively}{can be found in~\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 The equations applied are the ones from Figure~\ref{fig:string-equations} along with the naturality of delete and copy.
@@
 We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.  
-We denote by $\catname{SLat}$ the closed monoidal category of semilattices which is defined in section~\ref{sec:appendix:slat} of the Appendix.
+We denote by $\catname{SLat}$ the closed monoidal category of semilattices which is defined in \ifcameraready{section~\ref{sec:appendix:slat} of the Appendix}{\cite{tiurin2025equivalencehypergraphsdporewriting}}.
 Given two adjoint functors $F : \mathbb{A} \to \mathbb{B}$ and $G : \mathbb{B} \to \mathbb{A}$ with $F$ being left-adjoint, we write $F \dashv G : \mathbb{B} \to \mathbb{A}$ for this adjunction.
@@
 
-\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also Appendix~\ref{sec:appendix:slat})
+\begin{proposition}(A specialized case of Proposition 6.4.3~\cite{Borceux_1994}. See also \ifcameraready{Appendix~\ref{sec:appendix:slat}}{\cite{tiurin2025equivalencehypergraphsdporewriting}})
 There is a 2-adjunction 

### 2025-06-12 POPL2026/SlottedNotes.tex
notes for slotted e-graphs and arxiv script

--- 2024-01-25 E-Hypergraphs-LICS24.tex
@@
 
-\documentclass[sigconf,9pt]{acmart}
+\documentclass[acmsmall,screen, nonacm, anonymous]{acmart}
 
-\usepackage{tikz}
-\usepackage{tikz-network}
-\usepackage{xspace}
-\usepackage{xargs}
-\usepackage[
-  colorinlistoftodos,
-  prependcaption,
-  % disable,
-]{todonotes}
-\usepackage{fontenc}
+\newcommand*{\ONECOLUMN}{}
 
-%\usepackage{minted}
-
-\usepackage{subcaption}
-\usepackage{caption}
-
-\usepackage{hyperref}
-
-\input{preamble}
-\input{macros}
-\input{sample.tikzstyles}
-\input{hypergraph.tikzstyles}
-\input{hypergraph.tikzdefs}
+\input{../preamble}
+\input{../macros}
+\input{../sample.tikzstyles}
+\input{../hypergraph.tikzstyles}
+\input{../hypergraph.tikzdefs}
 
@@
 \newcommand\update[1]{{\color{americanrose}#1}}
-
-% \usepackage{mathtools,mathrsfs}
-\usepackage{tikz-cd}
-\usepackage{tikzit}
-% to reduce tikz picture white space
-\usepackage{adjustbox}
-
-% for quotienting
-% \usepackage{faktor}
-\usepackage{xfrac}
-
-\usepackage{import}
-
-% category names
-%\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}
-
-\newcommand\HypI[1]{\textbf{HypI}(#1)}
-\newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-% conflict
-\newcommand{\hashtag}{{\#}}
-\newcommand{\consistency}{{\smile}}
-% There is already a remark in the preamble, but it does
-% not work for some reason
-\newtheorem{remark}{Remark}
 
@@
 %% complete the rights form.
-\setcopyright{acmlicensed}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{XXXXXXX.XXXXXXX}
+% \setcopyright{acmlicensed}
+% \copyrightyear{2024}
+% \acmYear{2024}
+% \acmDOI{XXXXXXX.XXXXXXX}
 
 %% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[LICS '24]{Logic in Computer Science}{July 8--12,
-  2024}{Tallinn, Estonia}
+
 %%
@@
 %%  June 03--05, 2018, Woodstock, NY}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+% \acmISBN{978-1-4503-XXXX-X/18/06}
 
@@
 
+\settopmatter{printacmref=false}
+\authorsaddresses{}
+
 \begin{document}
 
-\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
+\title{Notes on slotted e-graphs}
 
 \author{Anonymous}
+
 %\email{}
@@
 
-\begin{abstract}
-The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
-We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
-We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
-Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
-They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
-\end{abstract}
-
 %%
@@
 
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10003752.10010124.10010131.10010137</concept_id>
-       <concept_desc>Theory of computation~Categorical semantics</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003809.10010031</concept_id>
-       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002950.10003624.10003633.10003637</concept_id>
-       <concept_desc>Mathematics of computing~Hypergraphs</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10003752.10003790.10003798</concept_id>
-       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Theory of computation~Categorical semantics}
-\ccsdesc[300]{Theory of computation~Data structures design and analysis}
-\ccsdesc[300]{Mathematics of computing~Hypergraphs}
-\ccsdesc[300]{Theory of computation~Equational logic and rewriting}
-
-\keywords{e-graphs, categorical semantics, enriched category theory,  hypergraph rewriting,  DPO rewriting.}
-
 %\received{20 February 2007}
@@
 
-\newcommand\id{\textsf{id}}
-\newcommand\sym{\textsf{sym}}
-
 \maketitle
 
-%-------------------- INTRODUCTION
+\section*{Section}
 
-\section{Introduction}
+Slotted e-graphs~\cite{slotted-egraphs} are an extension of e-graphs with two main objectives (apart from representing sets of equivalent terms)
+\begin{itemize}
+  \item Uniquely represent terms that are different up to a `renaming' of free variables, i.e., terms $a + b$ and $c + d$ where $a,b,c,d$ are distinct variables are represented as the same e-node.
+  \item Uniquely represent terms that are different up to renaming of bound variables, i.e., terms $\text{bind}\;\$s_1\; t_1$ and $\text{bind}\; \$s_2\;t_2$ are represented by the same e-node if $(s_1 \; s') \text{bind}\; \$s_1\;t_1 = (s_2 \; s') \text{bind}\;\$s_2\;t_2$ where $s'$ is a fresh name.
+\end{itemize}
 
-Rewrite-driven program optimisation consists of the application of a sequence of semantics-preserving rewrites which may improve the cost of execution, such as time or space (memory). 
-Because the application of some rewrites can either enable or block the subsequent application of others, the choice of application order significantly impacts the quality of the resulting optimisation.
-This long-standing issue is known as the \textit{phase-ordering problem}, with a \textit{phase} referring to a particular set of rewrites.
-Typical approaches to this problem use heuristics to determine an ordering.
+The first is achieved by parameterising each e-class by the ordered set of free variables (\textit{slots}) the terms of the e-class contain.
+For example, a term $a + b$ contains two free variables $a$ and $b$ and therefore the e-class that contains this term is parameterised by two variables.
+Non duplication of $a + b$ and $c + d$ is achieved by computing normal forms (\textit{shapes}) of such terms.
+In this particular example the shape is $\$0 + \$1$.
+One of the components of a slotted e-graph is a hash-map that maps such shapes to e-class id's that contain equivalent terms.
+Since two terms evaluate to the same shape, they correspond to a singe entry in an e-graph.
 
-A recently proposed,  alternative solution is \textit{equality saturation} 
-\cite{10.1145/1594834.1480915}: instead of maintaining a \textit{single},  putatively optimised program which is rewritten \textit{destructively} at each step, a \textit{set} of equivalent programs is maintained instead, whereas each rewrite step \textit{non-destructively} grows the set.
-Upon reaching a fixed point (\textit{saturation}),  a \textit{globally} optimal program can then be extracted from this set.
-A naive approach is clearly unfeasible, since the size of this set grows exponentially with the number of rewrites. 
-However, \textit{equality graphs (e-graphs)} \cite{EggPaper} are a data structure that can represent this set compactly, thus making the set tractable in practical applications.
+The other one is achieved by both computing the shape and generating a fresh slot each time a term with a binder is added to an e-graph.
+This makes it impossible to write a capturing $\beta$-reduction such as $(\lambda x . \lambda y . x) y$.
 
-Although equality saturation is already a state-of-the-art algorithmic optimisation technique, there is an opportunity to better understand its mathematical foundations. 
-We propose to address this issue via a categorical axiomatisation for e-graphs and their rewriting. 
-Considering programs as represented by terms of an arbitrary algebraic theory, we demonstrate a correspondence between e-graphs and (string diagrams for) simple \textit{semilattice-enriched} Cartesian categories.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, as encountered in a large body of research. 
-The starting innovation of our approach is the semilattice-enrichment. 
-The \textit{join} operator lifts the formalism so that it can now operate with (non-empty) \textit{sets} of terms (string diagrams, morphisms); the rest of the paper is about working out the mathematical implications of introducing this enrichment to string diagrams and their concrete combinatorial representation.
+Both of the above objectives can be achieved by using string diagrams for Cartesian closed SMCs by essentially performing a closure conversion of every functional symbol.
 
-Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}.
-This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
-\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
-However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.
-
-A final contribution of this work is giving a concrete combinatorial representation of enriched string diagrams in terms of (hierarchical)  hypergraphs, which we call \textit{e-hypergraphs}.
-We also give a specification of e-hypergraph rewriting via a suitable extension of the double pushout (DPO) rewriting framework
-\cite{dpo, bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022},  proving the representation \textit{sound and complete} with respect to our categorical semantics.
-As a corollary, this also provides a formalisation of the rewriting theory of e-graphs as encountered in the literature.
-
-\subsection{E-graphs}
-
-\begin{figure*}
-\[
-	\hspace{1.3cm}
-    \scalebox{0.65}{
-    \tikzfig{categorical-semantics/egraph-translation}
-    }
-\]
-\caption{Example translation of acyclic e-graphs into string diagrams for semilattice enriched Cartesian categories. }
-\label{fig:e-graph-example}
-\end{figure*}
-
-E-graphs are a data structure which can efficiently represent many equivalent terms of an algebraic theory.  They generalise the typical DAG representation of terms to include equivalence classes. 
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Figure~\ref{fig:e-graph-example}.
-This example is a `Hello world!' of e-graphs, presented first in~\cite{EggPaper}.
-
-Each column represents a term (or term rewrite rule), the conventional e-graph representation, and the equivalent e-string diagram representation, an e-hypergraph. 
-The initial term, $(a*2)/2$, is already represented efficiently in the e-graph by sharing the node 2.
-The string diagram version is slightly different than conventional e-graphs by representing the sharing of the node 2 explicitly, using a sharing (contraction) node.
-
-Nodes are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs. 
-Initially each equivalence class has a single node. 
-The first rewrite rule, replacing multiplication by the more efficient \emph{shift-left} operator creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) operator nodes, with a new sharing, that of node $a$.
-This second e-graph illustrates the other difference, besides sharing, between conventional e-graphs and e-hypergraphs, namely the fact that in the former edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes indicate which of the class-level edges are connected to which nodes inside the class.
-
-The other steps, (c) to (e), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rules. 
-Note that in the final e-graph, (e), a cycle is introduced, making it the representation of infinitely many terms: $a, a*1, (a*1)*1, \ldots$. 
-From the saturated graph the optimal term consisting of only the node $a$ can be then extracted, discarding everything else.
-
-\subsection{Semilattice Enriched Categories}
+Consider the example in Figure~\ref{fig:slotted_a_b_c_d} where subfigure (c) represents a term $(a + b) + (c + d)$ and note how subterms for $(a+b)$ and $(c + d)$ are shared.
+The headers of e-classes specify which free variables parameterise the e-class.
+Subfigure (d) represents an e-graph after an equivalent term $a + (b + (c + d))$ was added.
 
 \begin{figure}
-\[
-	\scalebox{0.65}{
-	\tikzfig{categorical-semantics/egraph-strings}
-	}
-\]
-\captionsetup{belowskip=-5ex}
-\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
-\label{fig:egraph-strings}
+  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_example}
+  \caption{Slotted e-graph example~\cite{slotted-egraphs}}
+  \label{fig:slotted_a_b_c_d}
 \end{figure}
 
-Recalling the correspondence between DAG representations of algebraic terms and  string-diagrams for morphisms of a suitable Cartesian category
-makes perspicuous the correspondence between (acyclic) e-graphs and morphisms of a suitably enriched Cartesian category, taken as string diagrams~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}.
-We extend the usual vocabulary of string diagrams for symmetric monoidal categories with the additional generator
-\[
-\infer{f_1 + \ldots + f_n: A \to B}{\{f_i: A \to B\}_{i \in \{1,\ldots,n\}}}
-\]
-The ability to take formal (non-empty) joins of morphisms is used to model the equivalence class structure of e-graphs.
+We can express the same data using string diagrams as in Figure~\ref{fig:slotted_a_b_c_d_string_1}.
+Note that there is no 1-to-1 corresponding between the slotted e-graph and the resulting string diagram (unlike in Figure~\ref{fig:slotted_a_b_c_d_string_2} which can not be obtained by applying rewrite rules to the initial string diagram).
+However, the e-box in red roughly corresponds to the invocations $(a)\;(b,c,d)$ and $(a,b)\;(c,d)$ of the slotted e-graph.
+We can decompose a string diagram in such a way because of the copy-delete rule.
+Also note that we can technically achieve more sharing by not only doing closure conversion on subtrees but on subgraphs as well, i.e., we could share the $\lambda x y . x + y$ 3 times instead of 2.
 
-String diagrams are read from bottom-to-top; we consider additional generators for the duplication and deletion transformations of a Cartesian category.
-Note, further, that in the Cartesian case we can restrict to generating morphisms $c$ to have a single output,  without loss of generality,  by applying the universal properties of the Cartesian product.
-This means the (informal) translation given below is fully general for acyclic e-graphs.
+In the presence of $\beta$-reduction rule we need a richer structure for our string diagrams though.
+Consider a term $(\lambda\; x\; .\; f\;x)\; y$ which evaluates to $f \; y$ under $\beta$-reduction.
+The corresponding slotted e-graph (after a $\beta$-reduction has been applied) can be seen in Figure~\ref{fig:f_x_y_slotted}.
+Note that it contains a cycle (the rightmost \texttt{app} e-node references \texttt{bind} and the latter references the e-class containing \texttt{app} again).
+This e-graph contains an infinite family of terms: $(\lambda\; x\; .\; f\;x)\; y$, $f \; y$, $(\lambda\; x\; .\; (\lambda z\; .\; f\;z) x\; y$ and so on.
+To fully support this, we need a trace.
+An example could be seen in Figure~\ref{fig:f_x_y_string_traced}.
 
-The translation from e-graphs to enriched string diagrams is illustrated informally below,
-noting how the typing constraints of the $+$ constructor are satisfied in the image of the translation by discarding in each component all unnecessary inputs.
-\[
-    \tikzfig{categorical-semantics/egraph-translation-1}
-\]
-Examples of this translation are given in the second row of Figure \ref{fig:e-graph-example}.
-% In particular, note that the translation of the \textit{cyclic} e-graph (e) requires the use of a categorical \textit{trace}, generating a cycle. 
-% In this paper, we focus on \textit{acyclic} e-graphs and their corresponding categories; but the above example illustrates that the extension to the cyclic case seems to be routine.
-As it would involve needless additional presentational complexity we leave it as further work.
-
-Note that the motivating example of our work, e-graphs, requires the Cartesian structure. 
-However, we will see how all the core of the theory of e-hypergraph rewriting only requires the monoidal structure. 
-Thus we can readily generalise e-graphs from algebraic to monoidal theories, giving rise to a  host of new potential applications, which we briefly outline in the conclusion of this paper.
-
-\subsection{Combinatorial Representation of Enriched String Diagrams}
-
-In order to \textit{implement} generalised e-graphs, rewriting must be performed on concrete representations of the corresponding string diagrams.
-String diagrams can be considered equivalently as either topological objects (\textit{i.e.}, taken modulo ``connectivity'') or as a 2-dimensional syntax quotiented by the equations of an SMC.
-For example, we have the following equivalences between diagrams.
-%\[(f_1 \otimes f_2) ; (g_1 \otimes g_2) = (f_1 ; g_1) \otimes (f_2 ; g_2)\] 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/interchange}
-	}
-\]
-This representation makes string diagrams unamenable to efficient implementation due to the difficulty of calculating the quotient. 
-A large body of research work shows how this issue can be solved by taking a different approach, representing string diagrams as hypergraphs~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022}.
-Here, the generators $c_i$ become hyper-edges,  and wires become vertices.
-Thus the expected quotient becomes simply hypergraph isomorphism.
-With appropriate restrictions on the form these hypergraphs can take, this approach can be used to \textit{characterise} the free symmetric monoidal category generated by some~$c_i$.
-
-We are interested not only in the free SMC over a set of generators, but also in \textit{(symmetric) monoidal theories} with extra equations such as, for instance, the rewrite rules (b)--(d) of Figure \ref{fig:e-graph-example}.
-These can be seen as equations between string diagrams, and thus rewrites of their hypergraph representations.
-Because the generating equations can be applied in any context, we are led to the notion of \textit{(hyper)graph rewriting}: given an equation $l=r$, we require some way to identify (a sub-hypergraph corresponding to) $l$ within a hypergraph and replace it with (a sub-hypergraph corresponding to) $r$.
-The standard methodology to giving specifications of such graph rewriting, known as \textit{double pushout (DPO) rewriting} \cite{dpo, bonchi_string_2022} is still applicable.
-However, these concepts must be now generalised from hypergraphs to \textit{e-hypergraphs}, hypergraphs with two additional relations denoting the hierarchical structure introduced by so-called \textit{e-boxes}: the generator for semilattice enrichment, and the separation of the components of each e-box.
-Giving the appropriate definitions turns out to be a technical matter of some complexity, and constitutes the main body of the paper.
-
-\subsection{Soundness and Completeness}
-The main technical result of the paper is a proof of soundness and completeness of this representation.
-This extends the results of \cite{bonchi_string_2022-2} for plain SMCs.
-While the structural equalities of SMCs are factored out in the representation,  those arising from the enrichment (see the equations of Figure \ref{fig:string-equations}) are not, and should not be.
-They represent the un/sharing of subdiagrams with respect to the e-box structure,  which is precisely what allows for the compact representation of equivalence classes.
-Instead, we consider DPO-rewrites implementing both the structural equalities for enrichment (which involve the e-box structure) and the equalities arising from the generating monoidal theory (which do not).
-
-Precisely,  our soundness and completeness result is the statement that morphisms in an appropriate free semilattice-enriched SMC are equal  if and only if there exists a sequence of DPO-rewrites --- each induced by a structural equality or the monoidal theory --- between their combinatorial representations.
-In the particular case of a monoidal signature including natural copy and delete maps,  we thus recover a sound and complete representation of semilattice-enriched Cartesian categories,  with the translation described earlier in the introduction justifying our claim of developing a mathematical theory of e-graphs.
-
-%\subsection{E-matching and E-rewriting}
-%
-%Having defined e-hypergraphs and shown them to correctly model the rewriting of string diagrams for semilattice enriched SMCs, we note that the naive implementation of DPO-rewriting is inefficient: to find a redex within an e-hypergraph involves finding a structurally equivalent e-hypergraph which contains the redex as a subgraph.
-In general, this involves unsharing the e-box structure before a redex becomes available.
-Thus, we define a notion of \textit{e-matching} for e-hypergraphs: that is, to find redexes working modulo the e-box structure.
-In particular, we wish to locate the smallest subgraph $G' \subseteq G$ "containing" (in an appropriate sense) the redex $L$.
-Given this,  \textit{e-rewriting} $L \to R$, intended to achieve equality saturation, is simple to define due to its non-destructive nature: we rewrite $G' \to G' + R$ in $G$.
-%\begin{figure}
-%\[
-%	\tikzfig{combinatorial_semantics/example-rewriting}
-%\]
-%\caption{Example application of e-rewriting for rewrite $(x*y)/z \to x*(y/z)$: find an appropriate minimal sub-e-hypergraph containing the redex $(x*y)/z$ and non-destructively add the reduct $x*(y/z)$.
-REDO THIS}
-%\end{figure}
-
-\subsection{Related Work}
-
-% E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
-This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.
-We are hopeful that our novel perspective on e-graphs can be extended to give uniform foundations to these investigations.
-Some further avenues for investigation, including potential applications, are given in the conclusion of this paper.
-
-% String diagrams
-Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.
-There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}.
-% Hierarchical hypergraphs
-Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-% String diagram rewrites
-Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.
-
-\section{Categorical Semantics of E-Graphs}
-In this section,  we introduce the required preliminaries on semilattice enriched symmetric monoidal categories generated by monoidal theories,  and the string diagrammatic formalism we will use to represent them.
-
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
-We write the identity morphism on $A$ as $\id_A$,  or simply $A$.
-We commonly write $f;g$ for composition in diagram order.
-Composition in the usual order is written $g \circ f$.
-We denote the tensor product of a \textit{symmetric monoidal category (SMC)} $\mathbb{C}$ by $\otimes$,  its unit by $I$ and its symmetry natural transformation as $\sym$ \cite{maclane}.
-We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
-We elide all associativity and unit isomorphisms associated with monoidal categories,  and often omit subscripts on identities and natural transformations where it can be inferred.
-We denote by $\textsf{SLatt}$ the category of semilattices.
-
-\subsection{Symmetric Monoidal Theories and PROPs.}
-
-A presentation of an algebraic theory is traditionally given by a \textit{signature} of $n$-ary operations and a set of \textit{equations}---formally,  pairs of terms freely generated over the signature.
-We are interested here in symmetric monoidal theories, which are generalisations of algebraic theories where the operations of the signature may have arbitrary \textit{co-arities}.
-First,  we generalise the notion of a signature.
-\begin{definition}[Monoidal signature]
-A \textit{(monoidal) signature} $\Sigma$ is given by a set of \textit{generators} $c: n \to m$,  with \textit{arity} $n$ and \textit{co-arity} $m$.
-%A \textit{signature homomorphism} $h: \Sigma \to \Gamma$ is a function from $\Sigma$ to $\Gamma$ which preserves the (co-)arities of elements.
-\end{definition}
-
-A categorical presentation of a monoidal signature is given by a freely generated products and permutations category.  
-\begin{definition}[Products and permutations category]
-A \textit{products and permutations category (PROP)} is a strict symmetric monoidal category with natural numbers as objects,  and such that $n \otimes m = n+m$.
-A \textit{PROP functor} is a strict SMC-functor which is additionally identity-on-objects.
-We denote the free \textit{products and permutations category} over a monoidal signature $\Sigma$ by $\textbf{PROP}(\Sigma)$.  
-\end{definition}
-The free category $\textbf{PROP}(\Sigma)$ can be syntactically constructed from taking (typed) categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$, quotiented by the axioms of a symmetric monoidal category.
-
-An equation associated with a monoidal signature is a pair of parallel morphisms of the form above, leading to the following definition of a symmetric monoidal theory. 
-\begin{definition}[Symmetric Monoidal Theory]
-A \textit{presentation of a symmetric monoidal theory} $(\Sigma, \mathcal{E})$ is given by a pair of a monoidal signature $\Sigma$ and a set $\mathcal{E}$ of pairs of parallel morphisms $f,g: n \to m$ of $\textbf{PROP}(\Sigma)$.
-A \textit{symmetric monoidal theory} $\textbf{SMT}(\Sigma,\mathcal{E})$ generated by a presentation $(\Sigma, \mathcal{E})$ is given by $\textbf{PROP}(\Sigma)$ quotiented by the least congruence including $\mathcal{E}$.
-\end{definition}
-
-It will often be convenient to give presentations of SMTs in terms of string diagrams.
-\begin{example}
-The SMT of \textit{commutative comonoids} is given by the following generators, ${\Delta, !}$, depicted in string diagrammatic form:
-\[
-	\scalebox{0.8}{
-  	 \tikzfig{categorical-semantics/Cartesian-equipment}
-	}
-\]
-together with the following associativity, commutativity and unitality equations, again given in terms of string diagrams. 
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-theory}	
-	}
-\]
-The Cartesian SMT is given by a set of generators $\Sigma_C$, together with the generators and equations of the SMT of commutative monoids, and the following additional naturality equations, for every $c \in \Sigma_C$
-\[
-	\scalebox{0.6}{
-	\tikzfig{categorical-semantics/Cartesian-naturality}
-	}
-\]
-By Fox's Theorem~\cite{fox},  $\textbf{SMT}(\Sigma_C, \mathcal{E}_C)$ is equivalent to the free Cartesian category over $\Sigma_C$. 
-\end{example}
-
-\subsection{Free Semilattice Enrichment: Formal Joins}
-
-We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.
-Note, in particular,  that semilattices do not require a unit for $+$.
-This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane}---a degeneracy we wish to avoid.
-More precisely our hom-sets will be given the structure of a semilattice.
-Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).
-In this case,  hom-sets are generalised to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition
-\[
-	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
-\]
-is defined as an $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
-In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.
-Technically, this amounts to asking that the tensor product is an enriched functor.
-We omit further details of the general case of enrichment here, and instead work with the following more concrete definition, where $M = \textbf{SLatt}$.
-
-\begin{definition}[Semilattice-enriched PROP]\label{def:enriched-prop}
-A \textit{semilattice enriched strict SMC (PROP)}  $\mathbb{C}$ is given by an SMC (PROP) $(\mathbb{C}, \otimes, 1,+)$ where every hom-set additionally has the structure of a semilattice $(\mathbb{C}(A,B), +)$ -- that is,  a set equipped with an associative,  commutative,  and idempotent operation -- which respects the composition and tensor product in the following ways,  for all appropriately typed morphisms.
-\begin{align*}
-f ; (g+h) &= f;g + f;h &
-(f+g) ; h &= f;h + g;h \\
-f \otimes (g+h) &= f \otimes g + f \otimes h & 
-(f+g) \otimes h &= f \otimes h + g \otimes h,
-%\item A set $ob(\mathbb{C})$ of objects;
-%\item For every pair $A,B \in ob(\mathbb{C})$, a hom-semigroup $Hom(A,B)$;
-%\item An element $\textsf{id}_{A,B} \in Hom(A,B)$; 
-\end{align*}
-Note,  we take $(;\!)$ and $\otimes$ to bind more tightly than $+$.
-A \textit{semilattice enriched strict SMC (PROP) functor} $F: (\mathbb{C}, \otimes, 1,+) \to (\mathbb{D}, \otimes, 1,+)$ is given by a strict SMC (PROP) functor on the underlying SMCs (PROPs) which additionally satisfies $F(f+g) = F(f)+F(g)$.
-We denote the \textit{freely enriched PROP (SMT)} over a monoidal signature $\Sigma$ as $\textbf{PROP}^+(\Sigma)$ ($\textbf{SMT}^+(\Sigma, \mathcal{E})$,  respectively).
-\end{definition}
-
-The free category $\textbf{PROP}^+(\Sigma)$ can be syntactically constructed from taking (typed) $\Sigma^+$-terms  -- namely, those freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$ (and $f+g$, respectively) -- quotiented by the axioms of an enriched symmetric monoidal category.
-
-To aid reasoning, we introduce a new language of string diagrams for semilattice enriched SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting.
-
-Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations.
-The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
-We overload the binary notation $+$ for our $n$-ary notation.
-We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
-Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
-
-\begin{figure*}
-\[  
-    \scalebox{0.75}{
-	\tikzfig{categorical-semantics/egraph-strings-equations}
-    }
-\]
-\caption{Equations for a  semilattice enriched symmetric monoidal category}
-\label{fig:string-equations}
-\end{figure*}
-
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
-
-In this section,  we recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  building to the recapitulation of the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
-
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.  
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $\psi$ to a relation on ordered sequences. 
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$. 
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-
-\subsection{Hypergraphs}
-
-In ~\cite{bonchi_string_2022-1},  hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for SMTs including a Frobenius algebra. 
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets. 
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.  
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram. 
-This gives rise to the definition of a hypergraph with interfaces.  
-Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams and the category of such hypergraphs.
-
-\begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-A \emph{hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l : E \to \Sigma$ is a labelling function that assigns each edge a generator from monoidal signature $\Sigma$.
-The labelling function must respect the typing of the generator: for all edges $e$,  $|s(e)| = m$ and $|t(e)| = n$,  where $l(e) : m \to n$.
-We call a hypergraph \textit{discrete} if its set of edges is empty.
-A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-\begin{enumerate}
-    \item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e))$
-    \item $\phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-    \item $l_{\mathcal{F}}(e) = l_{\mathcal{G}}(\phi_E(e))$
-\end{enumerate}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms. 
-\end{definition}
-Note that $\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
-
-\begin{definition}[Symmetric monoidal category of cospans]
-Let $\mathbb{C}$ be a category with all finite colimits.
-A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
-%Two cospans $X \xrightarrow{} A \xleftarrow{} Y$ and $X \xrightarrow{} B \xleftarrow{} Y$ are isomorphic if the following diagram commutes and $\alpha$ is iso.
-%\[
-%\begin{tikzcd}
-%                                               & A \arrow[dd, "\alpha"] &                                                \\
-%X \arrow[ru, bend left] \arrow[rd, bend right] &                        & Y \arrow[ld, bend left] \arrow[lu, bend right] \\
-%                                               & B                      &                                               
-%\end{tikzcd}
-%\]
-%For $X \in \mathbb{C}$ an $id$-cospan is $X \xrightarrow{id_X} X \xleftarrow{id_X} X$.
-Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
-Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
-\begin{multline*}
-    (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
-\end{multline*}
-and its unit by the initial object of $\mathbb{C}$.
-Symmetry is inherited from $\mathbb{C}$ as the coproduct~\cite{MonoidalCoproduct}.
-\end{definition}
-
-Then, morphisms of $\textbf{PROP}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
-\label{def:cspd}
-Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the full sub-category of $\catname{Csp(Hyp(\Sigma))}$ with discrete hypergraphs as objects.
-\end{definition}
-
-In particular, this means that any morphism in $\catname{HypI(\Sigma)}$ is of the form $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$,
-where $n,\;m$ are discrete hypergraphs and $\mathcal{G}$ is an arbitrary hypergraph.
-The morphisms $f$ and $g$ into $\mathcal{G}$ will specify which vertices are to be considered as inputs and outputs,  respectively.
-
-\begin{definition}[Degree,  path,  acyclic] 
-The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.  
-Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
-\end{definition}
-
-In order to represent only morphisms of $\textbf{PROP}(\Sigma)$,  it is necessary to restrict our attention to the following subset of hypergraphs with interfaces. 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces]
-\label{def:monogamy_hyp}
-We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
-\begin{enumerate}
-    \item $\mathcal{G}$ is a directed acyclic hypergraph;
-    \item $f$ and $g$ are monomorphisms;
-    \item the in-degree and out-degree of every vertex is at most $1$;
-    \item vertices with in-degree $0$ are precisely the image of $f$; and
-    \item vertices with out-degree $0$ are precisely the image of $g$.
-\end{enumerate}
-We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA-cospans. 
-\end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\textbf{PROP}(\Sigma)$. 
-\begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
-    $\textbf{PROP}(\Sigma) \cong \MdaCospans$
-\end{theorem}
-
-\subsection{DPOI-Rewriting for Hypergraphs with Interfaces}
-The correspondence of the previous section means that the symmetric monoidal equations of $\textbf{SMT}(\Sigma, \mathcal{E})$ are absorbed in the hypergraph representation.
-The equations $\mathcal{E}$ are then to be considered as rewrite rules on string diagrams,  which are then modelled as DPO-rewrites on hypergraphs with interfaces.
-DPO rewriting is a standard technique for formalising graph rewriting categorically. 
-In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Figure~\ref{fig:dpo_dpoi}~(left).
-
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & \mathcal K \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal G^{\urcorner}     & \mathcal C \arrow[l] \arrow[r] & ^{\ulcorner} \mathcal H
-% \end{tikzcd}
-% \]
-% CHRIS: WHY IS J DISPLAYING AND NOT R? 
-% \[
-%     \scalebox{0.8}{
-%         \tikzfig{combinatorial_semantics/DPO_square}
-%     }
-% \]
-
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal C \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal C$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
-
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Figure~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
-We now make this identification throughout without further mention.  
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Figure~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
-% \[
-% \begin{tikzcd}
-% \mathcal L \arrow[d, "m"'] & i+j \arrow[d] \arrow[l] \arrow[r] & \mathcal R \arrow[d]   \\
-% \mathcal{G}^{\urcorner}     & \mathcal C \arrow[l] \arrow[r]                      & ^{\ulcorner}\mathcal H \\
-%                   & n+m \arrow[lu] \arrow[u] \arrow[ru]          &              
-% \end{tikzcd}
-% \]
-\begin{figure}[t!]
-    \begin{subfigure}{0.4\linewidth}
-    \[
-            \scalebox{0.8}{
-                \tikzfig{combinatorial_semantics/DPO_square}
-            }
-    \] 
-    % \subcaption{DPO square}
-    % \label{fig:dpo}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-        \[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square_Hyp}
-    }
-\]
-    % \subcaption{DPOI square}
-    % \label{fig:dpoi}
-    \end{subfigure}
-    \captionsetup{belowskip=-3ex}
-    \caption{DPO and DPOI squares}
-    \label{fig:dpo_dpoi}
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_string_1_2}
+  \caption{String diagram example}
+  \label{fig:slotted_a_b_c_d_string_1}
 \end{figure}
 
-\end{definition}
-We need to restrict the notions of pushout complement and of match in order to preserve monogamy and directed acyclicity.
-
-\begin{definition}[Boundary complement]
-\label{def:boundary_original}
-For monogamous cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
-% \begin{figure}[h!]
-%     \centering
-%     \includegraphics[width=0.4\linewidth]{figures/combinatorial_semantics/boundary_complement.png}
-% \end{figure}
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJMIl0sWzIsMCwiaStqIl0sWzAsMiwiR157XFx1cmNvcm5lcn0iXSxbMiw0LCJuK20iXSxbMiwyLCJMXntcXGJvdH0iXSxbMyw0LCJbZF8xLGRfMl0iLDJdLFszLDIsIltiXzEsYl8yXSJdLFswLDIsImYiLDJdLFsxLDAsImEgPSBbYV8xLGFfMl0iLDJdLFsxLDQsImMgPSBbY18xLGNfMl0iXSxbNCwyLCJnIl1d
-% \[\begin{tikzcd}
-% 	\mathcal L && {i+j} \\
-% 	\\
-% 	{\mathcal G^{\urcorner}} && {\mathcal{L}^{\bot}} \\
-% 	\\
-% 	&& {n+m}
-% 	\arrow["{[d_1,d_2]}"', from=5-3, to=3-3]
-% 	\arrow["{[b_1,b_2]}", from=5-3, to=3-1]
-% 	\arrow["f"', from=1-1, to=3-1]
-% 	\arrow["{[a_1,a_2]}"', from=1-3, to=1-1]
-% 	\arrow["{[c_1,c_2]}", from=1-3, to=3-3]
-% 	\arrow[from=3-3, to=3-1]
-% \end{tikzcd}\]
-\[
-    \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_boundary_complement}
-    }
-\]
-is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
-\[
-    n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
-\]
-is a monogamous cospan. 
-\end{definition}
-%\begin{remark}
-%    Note how the requirement for $i \xrightarrow{} L \xleftarrow{} j$ and $n \xrightarrow{} G \xleftarrow{} m$ being monogamous, i.e. morphisms in $\MdaCospans$, is a part of the above definition.
-%\end{remark}
-Intuitively,  requiring boundary complements ensures that inputs are glued exclusively to outputs and vice-versa in the two pushout squares.
-
-\begin{definition}[Convex match]
-We call a subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-We call a match $m : \mathcal L \to \mathcal G$ a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
-\end{definition}
-
-\begin{definition}[Convex DPOI rewriting]
-\label{def:convex_dpo}
-Let $\mathfrak{R}$ be a set of DPOI rewrite rules. 
-Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $C$ and cospan arrows $i+j \xrightarrow{} \mathcal C \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-and the following conditions hold
-\begin{itemize}
-    \item $m : L \to \mathcal G$ is a convex match;
-    \item $i + j \to \mathcal C \to \mathcal G$ is a boundary complement in the leftmost pushout.
-\end{itemize}
-\end{definition}
-%By requiring convex matching and boundary complements we ensure that the original cospans above are monogamous, e.g., $n \xrightarrow{} G \xleftarrow{} m$.
-We do not formally recite the main result of \cite{bonchi_string_2022-2} here,  but the work proves that for any morphisms $f$ and $g$ in $\mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that $f = g$ if and only if there exist a sequence of DPO rewrites (induced by $\mathcal{E}$) between the representations of $f$ and $g$ as hypergraphs.
-In particular,  Theorem \ref{thm:prop-equiv} means that the structural SMC equations are factored out in the representation of $f$ and $g$,  and the DPO rewrites necessary are only those induced by $\mathcal{E}$.
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%For any $f,g \in \mathbf{SMT}(\Sigma, \mathcal{E})$,  we have that 
-%\[
-%	f = g \iff \llangle \lceil f \rceil \rrangle \Rrightarrow^*_{\mathfrak{R}} \llangle \lceil g \rceil \rrangle~, 
-%\]
-%where $\mathfrak{R} = \llangle \lceil \mathcal{E} \rceil \rrangle$ and $ \Rrightarrow^*_{\mathfrak{R}}$ denotes the existence of a sequence of convex rewrites. 
-%\end{theorem}
-\section{E-Hypergraphs}\label{sec:e-hypergraphs}
-
-Hierarchical (hyper-)graphs are a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}.
-In this section,  we follow analogous steps to those used to achieve a combinatorial representation of $\textbf{PROP}(\Sigma)$ by $\mathbf{MHypI}(\Sigma)$,  but in the semilattice enriched setting.
-First we define an \textit{e-hypergraph},  which supports a hierarchical notion of ``e-box'' with distinct components.
-Then,  we extend this notion with interfaces via a generalisation of the cospan construction.
-Finally,  we restrict the morphisms under consideration to those satisfying the MDA condition,  as in the previous section.
-An example diagram representing e-hypergraphs with interfaces can be found in Figure \ref{fig:A+B}.
-
-Note that here,  and in the rest of the paper,  we will restrict to dealing with signatures $\Sigma$ containing only morphisms $f: n \to m$ where either $n$ or $m$ are non-zero.
-This is a restriction we make in order to avoid complications to the technical development of DPOI rewriting for e-hypergraphs given in the subsequent section,  and we explain it further there.
-
-In the following,  we elide the obvious injections of $V$ and $E$ into $V+E$.
-
-\begin{definition}[E-hypergraphs]
-\label{def:e-homo}    
-An \emph{e-hypergraph $\mathcal{G}$ over a signature $\Sigma$} is a tuple $(V,E,s,t,l,<,\consistency)$,  where $(V,E,s,t)$ is an unlabelled hypergraph;  $l :E \to \Sigma + 1$ is a labelling function modified to include an extra value $\bot$ in its codomain; $<$ is a strict partial order on $V + E$ called the \textit{child relation}.
-We define the following additional notation.
-For $x \in V + E$,  the set of \emph{parents} and \emph{children} of $x$,  respectively:
-\[
-	[x) = \{x' ~|~ x' <_c x\; \} \qquad (x] = \{x' ~|~  x <_{c} x'\}
-\]
-We will denote the unique immediate parent $e$ of $x$ by writing $e <^{\mu} x$.
-We call edges $e$ such that $l(e) = \bot$ \textit{hierarchical},  and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \emph{top-level}.
-We require that the child relation satisfies the following conditions:
-\begin{enumerate}
-	\item each parent set contains exclusively hierarchical edges;
-	\item Each $x$ has at most one immediate parent;
-	\item if $v \in s(e)$ then $e' <^{\mu} e$ iff $e' <^{\mu} v$.
-\end{enumerate}
-We additionally define an equivalence relation $\consistency_p$,  called the \textit{consistency relation},  on each set $\{x \in V + E ~|~ p <^\mu c\}$ of elements which share the same parent.  
-We collect the union of all these equivalence relations into on relation $\consistency = \bigcup_{p} \consistency_p$,  which is defined on a subset of $V+E$. 
-The consistency equivalence relation must additionally satisfy that if $v \in s(e)$ or $v \in t(e)$ then $v \consistency e$. 
-\label{def:consistency_properties}
-\end{definition}
-Intuitively,  the child relation equips the hypergraph with a hierarchical structure: certain edges,  labelled with $\bot$,  are meant to interpret the dashed e-boxes of  string diagrams.  
-These edges can be designated as "parents" to sub-e-hypergraphs,  which interpret the various components of the e-boxes.
-These components are distinguished by the consistency relation, which partitions the maximal sub-e-hypergraph below the hierarchical edge.
-Technically,  condition $(3)$ on the child relations says that $<^{\mu}$ respects connectivity of edges and vertices,  and similarly for conditions $(1)$ of the consistency relation. 
-\begin{definition}[E-hypergraph homomorphism]
-An \textit{e-hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-\begin{enumerate}
-    \item $\phi$ is a hypergraph homomorphism
-    \item for $v \in V_{\mathcal{F}}$,  if $v$ is not top-level then 
-	\[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(v)) = <_{\mathcal{G}}^{\mu}(\phi_{V}(v))
-        \] 
-        and for $e \in E_{\mathcal{F}}$,  if $e$ is not top-level then 
-        \[
-        \phi_{E}(<_{\mathcal{F}}^{\mu}(e)) = <_{\mathcal{G}}^{\mu}(\phi_{E}(e))  
-        \] 
-        \item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$. 
-\end{enumerate}
-If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
-\end{definition}
-The conditions on e-hypergraph homomorphisms require to preserve immediate parents and the conflict relation.  
-\begin{definition}[Category of e-hypergraphs]
-The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.  
-\end{definition}
-
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in the Appendix.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but we prove that it does when certain technical conditions are satisfied,  including that the two morphisms involved are monos and their shared domain is a discrete hypergraph.
-
-\begin{remark}
-When we wish to consider $\phi(\mathcal{F}) \subseteq \mathcal{G}$ as an e-hypergraph in its own right,  we assign $\bot$ to be the immediate parent of all $x$ whose immediate parent in $\mathcal{G}$ is outside $\phi(\mathcal{F})$.
-\end{remark}
-
-When modelling enriched string diagrams using e-hypergraphs,  we require to keep track of more than simply the ultimate external inputs and outputs of the diagram: each internal e-box has its own inputs and outputs,  which we must also keep track of.
-Thus,  we introduce an \textit{extended} cospan construction.
-Since the $\catname{EHyp(\Sigma)}$ does not have pushouts in general,  we skip straight to considering a category of extended cospans of discrete e-hypergraphs (\textit{i.e.},  discrete hypergraphs considered as e-hypergraphs in the obvious way),  for which pushouts do exist.
-
-We use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n - m$ vertices,  in particular with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
-\begin{definition}[Category of e-hypergraphs with interfaces]
-    The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes of \textit{extended cospans}, defined as follows:
-    \[
-    n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-    \]
-    where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete e-hypergraphs,  $f_{ext},f_{int},g_{ext},g_{int}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
-We will sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
-
-This justifies calling $n$  \textit{external input interfaces} and $n'$ \textit{internal input (output) interfaces}.
-We call $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-
-    Two extended cospans are isomorphic if there exist isomorphisms $\alpha$, $\beta$ and $\gamma$ making the following diagram commute.
-    \[
-    \scalebox{0.75}{
-        \tikzfig{combinatorial_semantics/isomorphic_e_cospans}
-    }
-    \]
-Composition of two morphisms 
-\begin{align*}
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} &\mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m\\
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} &\mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
-\end{align*} is computed in two stages.
-First, $\mathcal{H}$ is computed as the result of the pushout square shown below; 
-% https://q.uiver.app/#q=WzAsMTAsWzAsMCwiZXh0KFgpIl0sWzEsMSwiWCJdLFsyLDIsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzUsMCwiWSciXSxbNiwyLCJCIl0sWzcsMSwiWiJdLFs4LDAsImV4dChaKSJdLFs0LDMsIl57XFx1bGNvcm5lcn1BK197Zl97ZXh0fTtmLGdfe2V4dH07Z31CIl0sWzAsMV0sWzEsMl0sWzMsMiwiZiJdLFs0LDMsImZfe2V4dH0iXSxbNCw1LCJnX3tleHR9IiwyXSxbNSw2LCJnIiwyXSxbNyw2XSxbOCw3XSxbMiw5XSxbNiw5XV0=
-\[
-\trimbox{0cm 0cm 0cm 0.75cm}{
-\adjustbox{scale=0.9,center}
-{\tikzfig{combinatorial_semantics/pushout_e_cospans}}
-}
-\]
-then,  the result of composition is defined as follows:
-\[
-n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
-\]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph. 
-\[
-    h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ]
-\qquad
-    h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
-\]
-% \[
-% \begin{tikzcd}
-% Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
-% \end{tikzcd}
-% \]
-% and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
-The identity of composition is given by the obvious extended cospan of identities. 
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.  
-\end{definition}
-
-As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
-Unlike in the previous section,  we require (extended) cospans to consist of monomorphisms.
-This is to ensure the existence of the pushouts needed for composition,  since pushouts do not necessarily exist in $\textbf{EHyp}(\Sigma)$.
-This also means that we omit the monomorphism condition from the following definition.
-We additionally introduce a notion of \textit{well-typing} of e-hypergraphs,  which ensures each component of an e-box has the same number of inputs (outputs) as every other component.
-
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
-
-\begin{definition}[Category of well-typed MDA e-hypergraphs with interfaces]
-We call an extended cospan in $\Ecospans$,  as below,   \textit{monogamous directed acyclic (MDA)} if
-\[
-n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
-\]
-\begin{enumerate}
-        \item underlying hypergraph of $\mathcal{G}$ is directed acyclic; 
-       % \item $f_{ext}$, $f_{int}$, $g_{ext}$ and $g_{int}$ are monomorphisms; 
-        \item The in-degree and out-degree of every vertex is at most 1; 
-        \item Vertices with in-degree $0$ are precisely the image of $f_{int}$; 
-        \item Vertices with out-degree $0$ are precisely the image of $g_{int}$; 
-        % the item below is not assumed by the definition of a cospan of e-hypergraphs
-        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
-\end{enumerate}
-\label{def:monogamy_ehyp}
-
-Further,  consider the sets of input and output vertices,  $I$ and $O$ respectively,  of a hierarchical edge $e$ of an e-hypergraph $\mathcal{G}$. 
-The consistency relation of $\mathcal{G}$ partitions $I$ and $O$.  
-We call $e$ \emph{well-typed} if the size of each partition of $I$ is $|s(e)|$ and the size of each partition of $O$ is $|t(e)|$.
-We will call an MDA-extended-cospan in $\MdaEcospans$ \emph{well-typed} if all hierarchical edges in its apex are well-typed.
-
-We denote by $\MdaEcospans$ the subcategory of $\Ecospans$ consisting of well-typed MDA-extended-cospans. 
-\end{definition}
-Note,  $\MdaEcospans$ is indeed a category,  and in fact inherits the symmetric monoidal structure of $\Ecospans$:  it is easy to check that identities,  the monoidal unit,  symmetry,  and the composition and tensor product of any two well-typed cospans are all (again) well-typed.
-
-Unlike in the previous section,  in order to give an equivalence with $\textbf{PROP}^+(\Sigma)$,  we must first develop the theory of DPOI rewriting for e-hypergraphs.
-This is because the equations for semilattice enrichment are not intended to be subsumed by the e-hypergraph representation.
-Instead,  we will treat them via rewriting,  as we do with equations of the signature.
-This is because the semilattice equations involve sharing and copying, operations on string diagrams which are not usually quotiented away, cf.
-string diagrams for Cartesian categories.
-An ideal combinatorial representation would factor out the three semilattice equations which \textit{do not} involve sharing and copying (namely,  associativity, commutativity, and idempotence) --- however, a limitation of the current work is that we do not achieve this.
-Presently,  we will treat these equations via rewriting as well.
-
-\section{DPOI-Rewriting for E-Hypergraphs}
-
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
-This is the purpose of this section of the paper.
-
-We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows. 
-\[
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
-\qquad
-    n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
-\] 
-Analogous to Section \ref{sec:combinatorial-semantics}, observing that the following extended cospans express the same data
-\[
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m
-\qquad
-    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
-\] allows us to encode rewrite rules as extended cospans of the following form
-\[
-\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R}
-\] which will fit into the DPO formalism. We make this identification throughout,  without further mention. 
-%We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
-\begin{remark}
-There is a limitation of extended DPOI rewriting as defined in this section: it is unable to express rewriting
-for rules whose target contains an edge with no incident vertices.  
-While this limitation can be overcome,  we find the technical development to be significantly simpler if we simply avoid consideration of such edges.  
-For this reason,  in the remainder of the paper we consider signatures which do not contain $c: 0 \to 0$,  thus restricting possible e-hypergraphs to those without such problematic edges. 
-%This is because there is no way to impose both consistency and child relation of such an edge as it has no incident nodes in its context.
-%We claim that this is only a minor limitation because in practice such edges (morphisms) with no
-%incident nodes correspond to ‘scalars’, if we consider for example a category of vector spaces, and scalars are usually removed by being absorbed by vectors rather than introduced during rewrites.
-\end{remark}
-
-% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
-% Note that the outermost interfaces are preserved.
-
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Figure~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
-
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
-% Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
-% The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
-
-Before introducing our definition of \textit{extended} DPOI rewriting,  we must modify the definition of boundary complement to guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  notion of a \textit{down-closed} graph,  which will be necessary for the subsequent development. 
-\begin{definition}[Down-closed subgraph]
-%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
-%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-    We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
-\end{definition}
-
-\begin{definition}[Extended boundary complement]
-\label{def:boundary_new}
-
-% https://q.uiver.app/#q=WzAsNyxbMiwwLCJMIl0sWzQsMCwieF97ZXh0fSArIHlfe2V4dH0iXSxbMiwyLCJHXntcXHVyY29ybmVyfSJdLFs0LDIsIkMiXSxbNCw0LCJuX3tleHR9K21fe2V4dH0iXSxbMCwwLCJsX2krbF9vIl0sWzAsMiwiZ19pK2dfbyJdLFsxLDAsIiIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzAsMiwibSIsMCx7ImNvbG91ciI6WzAsNjAsNjBdfSxbMCw2MCw2MCwxXV0sWzEsMywiW2lfYyxqX2NdIiwwLHsiY29sb3VyIjpbMCw2MCw2MF19LFswLDYwLDYwLDFdXSxbNCwzLCJbbl9jLG1fY10iLDIseyJjb2xvdXIiOlswLDYwLDYwXX0sWzAsNjAsNjAsMV1dLFszLDIsIiIsMix7ImNvbG91ciI6WzAsNjAsNjBdfV0sWzQsMiwiIiwyLHsiY29sb3VyIjpbMCw2MCw2MF19XSxbNSwwXSxbNiwyXV0=
-
-For monogamous extended cospans 
-\[
-    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
-\quad\text{ and }\quad
-    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-\] and mono $f : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
-as depicted in the square below
-\[
-\scalebox{0.75}{
-
-    \tikzfig{combinatorial_semantics/DPOI_pushout_complement}
-}
-\]
-is a \textit{boundary complement} if
-\begin{enumerate} 
-    \item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-    \item $[c_1,c_2]$ is mono;
-    \item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same set of parents, and either $v,w$ are top-level or else $v \consistency w$;
-    \item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended mda-cospan
-    \[
-    n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
-    \]
-    \item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended mda-cospan
-    \[
-    n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
-    \]
-\end{enumerate}
-where $f_i$ and $g_i$ are defined as follows.  
- %because the image of $m$ is a convex down-closed e-hypergraph, 
-The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
-The maps $f_2$ and $g_2$ are defined similarly.
-%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
-%
-%\[
-%    \tikzfig{combinatorial_semantics/pushout_difference}
-%\]
-%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
-%Similarly for $g_2$.
-% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
-% \[
-%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
-% \]
-% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
-% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
-% Similarly for $k \to r$.
-\end{definition}
-
-Intuitively the above definition means that when the occurrence of $\mathcal{L}$ within $\mathcal{G}$ is top-level then $\mathcal{L}$'s external interfaces become the part of $\mathcal{L}^{\bot}$'s external interfaces and when the occurrence is nested, those interfaces become the part of $\mathcal{L}^{\bot}$'s strict internal interfaces.
-In both cases when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, $\mathcal{L}$'s strict internal interfaces are removed from strict internal interfaces of $\mathcal{G}$ to form internal interfaces of the corresponding mda-cospans above.
-The morphisms are then constructed correspondingly to require that inputs (respectively, outputs) of $\mathcal{R}$ are glued to the outputs (respectively, inputs) of $\mathcal{L}^{\bot}$.
-
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
-%CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
-%The only difference is that we account for internal interfaces to require monogamous-ness and these interfaces must be computed explicitly to define the monogamous cospan.
-%
-%\begin{figure}
-%    \centering
-%    \[
-%    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
-%    \label{fig:interface_change_example}
-%\end{figure}
-
-% The second point above means that when cutting the mono-occurrence of $L$ out from $G$ we also need to remove the inner interfaces of $L$ from the whole interface of $G$.
-\question{According to the pushout squares above, $c_i$ is constructed by removing from $g_i$ everything from $l_i$ which does not have a pre-image in $i$, i.e., everything except the outermost interfaces (same holds for $c_o$)}
-
-Note that,  in the above definition,  the extended cospan 
-    \[
-    n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
-    \]
- is not necessarily well-typed.
-This is because its input internal interface may contain vertices that previously were a part of the output internal interface and vice versa. 
-%The uniqueness and existence properties of boundary complements as defined above are discussed in the appendix (Proposition~\ref{prop:boundary_complement} and Proposition~\ref{prop:complement_existence}).
-
-% \begin{proposition}[Proposition~\ref{prop:boundary_complement} restatement]
-%     The boundary complement in~\ref{def:boundary_new} when exists is unique.
-% \end{proposition}
-
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.  
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
-
-\begin{definition}[Convex EDPOI rewriting]
-\label{def:dpoi-e}
-Given an extended span of morphisms 
-\[
-    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
-\]
-in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
-\[
-    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
-\] 
-if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
-\[
- \scalebox{0.75}{
-    \tikzfig{combinatorial_semantics/DPOI_square}
- }
-\]
-such that the two marked squares are pushouts and the following conditions hold:
-    \begin{enumerate}
-        \label{dpoi-e:assumptions}
-       % \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-        %\item $m(L)$ is a down-closed e-hypergraph;
-        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
-	\item the internal interfaces of $\mathcal H$ are such that:
-	\[
-		n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) \qquad k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)\\
-	\]
-	\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
-    \end{enumerate}
-% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
-%  \[
-%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
-% \]
-    Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ if there exists a sequence of such rewrites.
-\end{definition}
-The boundary complement conditions, in particular, prohibit finding a match for $\mathcal L$ in $\mathcal G$,  below.  
-\[
-	\mathcal{L} = \scalebox{0.8}{\tikzfig{combinatorial_semantics/f_times_g}} \qquad \mathcal{G} = \scalebox{0.6}{\tikzfig{combinatorial_semantics/f_plus_g_inline}}
-\]
-Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
-
-With our modified definition of extended DPOI rewriting in hand,  we proceed to prove that our combinatorial representation of semilattice-enriched SMCs is sound and complete in an appropriate sense.
-
-%\begin{proposition}
-%    $n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k$ is a well-typed mda-cospan.
-%\end{proposition}
-
-%Above we defined rewriting for cospans with empty input interfaces, when we will further build a correspondence between $\textsf{PROP}(\Sigma)^{+}$ and $\WellTypedMdaEcospans$.
-%we will operate with cospans with non-empty input interfaces as well, so below we formulate what it means to rewrite cospans with non-empty input interfaces.
-%
-%\begin{definition}[Convex EDPOI in $\WellTypedMdaEcospans$]
-%    
-%We say that an \textit{mda-cospan}
-%\[
-%    n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-%\]
-%\textit{rewrites convexly into another mda-cospan}
-%
-%\[
-%    n \xrightarrow{} n'' \xrightarrow{} \mathcal{H} \xleftarrow{} k'' \xleftarrow{} k
-%\]
-%
-%under a rewrite rule 
-%% \[
-%%     \langle x_{ext} \xrightarrow{a_1} l_{i} \xrightarrow{a_2} L \xleftarrow{b_2} l_{o} \xleftarrow{b_1} y_{ext}, 
-%%     x_{ext} \xrightarrow{a_1} r_{i} \xrightarrow{a_2} R \xleftarrow{b_2} r_{o} \xleftarrow{b_1} y_{ext} \rangle
-%% \]
-%
-%\[\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}\]
-%
-%if 
-%\begin{align*}    
-%    \mathcal{G} \xleftarrow{} n + k \Rrightarrow^{+} \mathcal{H} \xleftarrow{} n + k
-%\end{align*}
-%under the same rewrite rule.
-%We will use the same $\Rrightarrow^{+}$ and the concrete definition will be clear from the context.
-%
-%\end{definition}
-%
-
-\section{Soundness and Completeness}\label{sec:soundness-and-completeness}
-
-This section contains the main technical results of the paper,  building on those of \cite{bonchi_string_2022-2}. 
-We first construct an interpretation of terms generating $\textbf{PROP}^+(\Sigma)$ in $\WellTypedMdaEcospans$.  
-As expected,  the SMC equations are factored out in our representation of terms in $\WellTypedMdaEcospans$,  but the semilattice equations are not.  
-Instead,  we implement the semilattice equations via EDPOI rewrites,  leading to the following soundness and completeness result: for any morphisms $f$ and $g$ in $\textbf{SMT}(\Sigma, \mathcal E)$,  $f = g$  if and only if there exists a sequence of EDPOI-rewrites --- each induced by either a structural semilattice equation or $\mathcal E$ --- between their combinatorial representations.
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{figures/slotted_a_b_c_d_string_2}
+  \caption{String diagram example}
+  \label{fig:slotted_a_b_c_d_string_2}
+\end{figure}
 
 \begin{figure}
-    \[
-    \scalebox{0.55}{
-    \tikzfig{combinatorial_semantics/f_plus_g_new}
-    }
-    \]
-    \captionsetup{belowskip=-4ex}
-    \caption{$+$ of two morphisms in $\Ecospans$}
-    \label{fig:A+B}
+  \includegraphics[width=0.5\linewidth]{figures/f_x_y_slotted}
+  \caption{Slotted e-graph for $\lambda x\; .\; f\;x$}
+  \label{fig:f_x_y_slotted}
 \end{figure}
-Recall from Section \ref{sec:e-hypergraphs} that $ \WellTypedMdaEcospans$ has a symmetric monoidal structure inherited from the coproduct structure of $\textbf{EHypI}(\Sigma)$.  
-Recall further that typed $\Sigma^+$-terms are those categorical combinator terms freely constructed from generators $c \in \Sigma$, $\textsf{id}_1$, $\sigma_{1,1}$, $(;\!)$ and $\otimes$,  and $f+g$.
-We will write $\Sigma^+(n,m)$ for the set of $\Sigma^+$-terms of type $n \to m$.
-
-We wish to specify a set of EDPOI rewrite rules corresponding to those in Figure \ref{fig:string-equations}.
-In order to do this,  we specify a function which maps  uninterpreted $\Sigma^+$-terms into $\WellTypedMdaEcospans$ morphisms,  and which also extends the expected interpretation of $\Sigma$-terms in $\WellTypedMdaEcospans$ (as given by the free functor from $\textbf{PROP}(\Sigma)$ induced by the interpretation of $\Sigma$ as itself,  detailed in \cite{bonchi_string_2022-2}).
-\begin{definition}[Interpretation function] 
-%Let $\llbracket - \rrbracket: \textbf{PROP}(\Sigma) \to \WellTypedMdaEcospans$ be the unique SMC-functor induced by the interpretation sending $\Sigma$ to itself.  
-Let $\llbracket - \rrbracket_{n,m}: \Sigma^+(n,m) \to \WellTypedMdaEcospans(n,m)$ be the function defined by induction on the type derivation of $\Sigma$-terms by 
-\begin{align*}
-	\llbracket \id \rrbracket_{n,m} &= \id  &	\llbracket f ; g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} ; \llbracket g \rrbracket_{n,m} \\
-	\llbracket I \rrbracket_{n,m} &= I & \llbracket f \otimes g \rrbracket_{n,m} &= \llbracket f \rrbracket^+_{n,m} \otimes \llbracket g \rrbracket_{n,m} \\
-	\llbracket \sym \rrbracket_{n,m} &= \sym  & \llbracket c \rrbracket_{n,m} &= c   \\ 
-	\llbracket f+g \rrbracket_{n,m} &= \llbracket f \rrbracket_{n,m} + \llbracket g \rrbracket^+_{n,m}
-\end{align*}
-where $c \in \Sigma$ and $+$ in $\WellTypedMdaEcospans$ is defined by Figure \ref{fig:A+B}.
-It uses the notation of dashed boxes for hierarchical edges, with its children drawn inside,  and maximally consistent components separated by another dashed line.
-We will omit the subscripts in $\llbracket - \rrbracket_{n,m}$ where it is clear from context.
-\end{definition}
-
-In order to phrase our soundness and completeness results categorically,  we define the following quotient of $\WellTypedMdaEcospans$ by EDPOI rewrites.
-This means that soundness is expressed as the extension of $\llbracket - \rrbracket$ to a semilattice-enriched SMC functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$,  and completeness is expressed as the faithfulness of this functor.
-In fact,  we prove that the functor is also full,  giving an equivalence of categories.
-
-We define the appropriate notion of quotient by EDPOI rewrites below.
-Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms).
-% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
-\begin{definition}[Quotient by rewrites]  
-Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
-\[
-	f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\llbracket \mathcal E \rrbracket} g ~ . 
-\]
-\end{definition}
-
-Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.
-Let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
-%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
-Then we have the following result. 
-\begin{proposition}[Soundness]
-\label{prop:soundness}
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.
-
-The reason for calling the previous result \textit{soundness} is that the interpretation function on $\Sigma^+$-terms induces a semilattice-enriched SMC functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which further induces the following semilattice-enriched SMC functor,  for a set of equations $\mathcal E$:
-\[
-	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
-\]
-which by uniqueness must be the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself.
-Unfolding definitions,  it follows that
-\[
-	f = g \quad \Rightarrow \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
-Note that it is immediate from the previous result that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is indeed a semilattice-enriched PROP.
-
-The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}.
-\begin{proposition}[Weak normal form]
-\label{prop:wnormal_form}
-For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-\[
-	f = f_1 + \ldots + f_n
-\]
- such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-\end{proposition}    
-The proof follows the familiar argument (\textit{i.e.},  termination of distributivity rewrites) that $\Sigma^+$-terms can be rewritten via the semilattice-enrichment equations into a similar normal form.
-
-This normal form can be used to prove the main theorem of the paper: the (full) completeness of our combinatorial representation of $\textbf{SMT}^{+}(\Sigma, \mathcal {E} )$. 
-\begin{theorem}[Full completeness]
-We have the following equivalence of categories
-\[
-	\textbf{SMT}^{+}(\Sigma, \mathcal {E} ) \cong \WellTypedMdaEcospans / \mathcal{S,E}~.
-\]
-\end{theorem}
-\begin{proof}
-The equivalence is given by the functor $\llbracket - \rrbracket$.
-First, note that the functor is surjective on objects,  so it suffices to prove the functor is both full and faithful.
-In each case,  the result follows by a normal form argument in combination with relying on the equivalence $(*)$ $\textbf{SMT}(\Sigma, \mathcal E) \cong \textbf{MHypI}(\Sigma)/\mathcal{E}$ of symmetric monoidal theories and (appropriately quotiented) plain MDA cospans which is given by the obvious restriction of $\llbracket - \rrbracket$ to $\textbf{SMT}(\Sigma, \mathcal E)$; see Theorem 35 of ~\cite{bonchi_string_2022-2}.
-
-To see that the functor is faithful,  we prove that $f \neq g$ in $\textbf{SMT}^+(\Sigma, \mathcal E)$ implies $\llbracket f \rrbracket \neq \llbracket g \rrbracket$.
-First,  calculate weak normal form $\Sigma^+$-terms of $f$ and $g$ as $f_1 + \ldots + f_n$ and $g_1 + \ldots + g_m$,  respectively.
-Applying the functor $\llbracket - \rrbracket$,  we thus have that
-\[
-	\llbracket f \rrbracket = \llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket \qquad
-	\llbracket g \rrbracket = \llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket ~ . 
-\]
-Observe that $\llbracket f_1 \rrbracket + \ldots + \llbracket f_n \rrbracket$ and $\llbracket g_1 \rrbracket + \ldots + \llbracket g_m \rrbracket$ are weak normal forms.
-We consider two cases: where $n=m$ and where $n\neq m$.
-In the first case,  it is immediate that $\llbracket f \rrbracket \neq \llbracket g \rrbracket $,  since equal terms have weak normal forms with the same number of components.
-Otherwise,  if $n = m$,  assume for contradiction that $\llbracket f \rrbracket = \llbracket g \rrbracket$.
-We have that there exists a permutation $\sigma$ such that for every index $i$ we have $\llbracket f_i \rrbracket \neq \llbracket g_{\sigma(i)} \rrbracket $,  and thus by  $(*)$ that $f_i = g_{\sigma(i)}$,  implying that $f = g$ which is a contradiction.
-Therefore $\llbracket f \rrbracket \neq \llbracket g \rrbracket$ as required.
-
-To see that the functor is full,  consider a morphism $f$ in $\WellTypedMdaEcospans$.
-We can find an equivalent weak normal form $f_1 + \ldots + f_n$,  where each $f_i$ contains no hierarchical edges.
-By $(*)$ we can find morphisms $g_i$ in $\textbf{SMT}(\Sigma, \mathcal E)$ such that $\llbracket g_i \rrbracket = f_i$.
-Thus,  we can construct a morphism $g_1 + \ldots + g_n$ such that $\llbracket g_1 + \ldots + g_n \rrbracket = \llbracket g_1 \rrbracket + \ldots \llbracket g_n \rrbracket = f_1 + \ldots + f_n$,  as required.
-\end{proof}
-
-Unfolding the definitions,  we can spell out the previous theorem in terms of EDPOI-rewriting. 
-\begin{corollary}[Completeness of rewriting]
-For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
-\[
-	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
-\]
- \end{corollary}
-%
-%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
-%
-%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
-%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
-%\end{theorem}
-%
-%\begin{proposition}
-%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
-%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
-%\end{proposition}
-%
-%\begin{theorem}
-%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
-%\end{theorem}
-Note that in string diagram rewrite theory as developed in \cite{bonchi_string_2022-2},  some extra steps are taken to prove the claim that $f \rightsquigarrow g$ if and only if $\llbracket f \rrbracket \Rrightarrow^{*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket$,  for an appropriate definition of $\Sigma$-term rewriting $\rightsquigarrow$.
-We omit this extra detail for brevity and clarity,  dealing with equational theories rather than rewrite theories,  but it is straightforward to extend our results to prove an analogous claim.
-
-\section{Conclusion}
-
-We have seen how e-graphs over algebraic theories can be expressed in terms of Cartesian categories enriched over a semilattice. 
-In fact, e-graphs are an instance of a more general construction that can account for monoidal theories in terms of SMCs,  similarly enriched.  
-Having described a translation of e-graphs into this categorical framework, a combinatorial representation of morphisms in the free category is also provided. 
-As expected, this representation factors out the structural SMC equations, while remaining sensitive to the complexity-relevant equations induced by the enrichment.
-The design of such a representation and its corresponding rewrite theory proved to be delicate, and, together with its proof of completeness, constitute the main technical contributions of the paper.
-
-In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
-Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica_hierarchical_2023}, as sketched in Fig.
-\ref{fig:app1}(a); the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}, as sketched in Fig.
-\ref{fig:app1}(b); or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes, as sketched in Fig.~\ref{fig:app2}.
-All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
 
 \begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-e-graphs}
-	}
-\]
-\caption{E-graphs for functorial boxes and the ZX-calculus}
-\label{fig:app1} 
-\end{figure}
-\begin{figure}
-\[
-	\scalebox{0.75}{
-	\tikzfig{conclusion/example-traced}
-	}
-\]
-\caption{E-graphs with cycles and e-hypergraphs with trace}
-\label{fig:app2}
+  \includegraphics[width=0.5\linewidth]{figures/f_x_y_string_traced}
+  \caption{String diagram example}
+  \label{fig:f_x_y_string_traced}
 \end{figure}
 
-In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.
-Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
-% traced comonoid
-%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments.
-Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.
-However, as we have seen in this paper, the technical details can be delicate.
-Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}.
-Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
-
 \bibliographystyle{acm}
-\bibliography{bibliography}
-
-\appendix
-
-\section{Proofs for Section \ref{sec:combinatorial-semantics}: Combinatorial Semantics}
-
-In this section we will define the conditions under which the pushout in $\catname{EHyp}(\Sigma)$ exists and we will explicitly construct such a pushout.
-We will do the same for the corresponding pushout complement and show its uniqueness.
-These condition will play a crucial role in showing that for each enrichment equation in $\textbf{PROP}^{+}(\Sigma)$ there is a sequence of rewrites in $\WellTypedMdaEcospans$.
-
-\subsection{Proof of the existence of pushouts}
-
-In constructing the pushout the functional versions of e-hypergraph relations will be useful.
-\begin{remark}
-    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
-\end{remark}
-
-\begin{remark}
-    Using functional notation we can reformulate the notion of a homomorphism between two e-hypergraphs in the following way.
-\end{remark}
-\begin{definition}
-        \label{def:e-homo-2}    
-        A \emph{homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ of e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
-
-        \begin{enumerate}
-            \item $\phi$ is hypergraph homomorphism.
-
-            \item When $x$ is not a top-level vertex, 
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{V};\iota_{V_{\mathcal{G}}}(x))
-                \]
-                and
-                \[
-                \phi_{E}(<_{\mathcal{F}}^{\mu}(\iota_{E_{\mathcal{F}}}(x))) = <_{\mathcal{G}}^{\mu}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))  
-                \] when $x$ is a not top-level edge.
-                \item
-            When $x \in E_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}} ]^{*}(\consistency_{\mathcal{F}}(\iota_{E_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{E};\iota_{E_{\mathcal{G}}}(x))
-            \]
-            where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
-            \item When $x \in V_{\mathcal{F}}$
-            \[
-                [\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{F}}(\iota_{V_{\mathcal{F}}}(x)))
-                \subseteq
-                \consistency_{\mathcal{G}}(\phi_{V};\iota_{V_{\mathcal{G}}}(x)).
-            \]
-            \end{enumerate}
-\end{definition}
-
-\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
-\label{th:existence_of_pushouts}
-Consider the following span in $\catname{EHyp_{\Sigma}}$
-% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
-\[\begin{tikzcd}
-	Z && X \\
-	\\
-	Y
-	\arrow["f", from=1-1, to=1-3]
-	\arrow["g"', from=1-1, to=3-1]
-\end{tikzcd}\]
-such that
-\begin{enumerate}
-\label{pushout:assumptions}
-    \item $Z$ is a \textit{discrete} e-hypergraph.
-    \item $f,g$ are monos.
-    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
-    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
-    \item $\consistency(f_{V}(v_i)) = \consistency(f_{V}(v_j))$ and $\consistency(g_{V}(v_i)) = \consistency(g_{V}(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
-\end{enumerate}    
-then the pushout $X +_{f,g} Y$ exists.
-\end{theorem}
-\begin{proof}
-    Consider the diagram below.
-    \[
-        \adjustbox{scale=1.25}{
-            \begin{tikzcd}
-            Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-            Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
-                                                                            &                                                                  & Q
-            \end{tikzcd}}
-    \]
-    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-    First, a coproduct of $X+Y$ is computed which is 
-    \[
-        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{X+Y} \}
-    \]
-    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{X+Y}, \consistency_{X+Y}$.
-    We will omit labels as they are irrelevant to pushout construction.
-    Consider relations
-    \begin{align*}    
-    S_{V} &= \{\\
-          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
-          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}
-        \\
-        \\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
-    \end{align*}
-    % \begin{align*}
-    % S_{E} &= \{\\
-    % &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
-    % &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
-    % &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
-    % &\}
-    % \end{align*}
-and let relations $R_{V},R_{E}$ be their reflexive symmetric and transitive closures respectively.
-
-We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
-    \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        & \}    
-    \end{align*}
-We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
-We have 
-\[
-    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
-\]
-and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
-There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
-We then define 
-\[
-    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
-\]
-We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
-Likewise, 
-\[
-    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
-\]
-
-So far, the construction matches the construction of the pushout in $\catname{Hyp_{\Sigma}}$ which can be found in~\cite{bonchi_string_2022-1} and hence $s_{\sfrac{X+Y}{\sim}}$ and $t_{\sfrac{X+Y}{\sim}}$ are well-defined and $[-]$ is a homomorphism with respect to sources and targets and therefore $\sfrac{\iota_1}{\sim}$ and $\sfrac{\iota_{2}}{\sim}$ are.
-
-Recall that $<$ is essentially a transitive closure of $<^{\mu}$ and hence we can first define 
-    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
-    \] and 
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
-    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    We will consider several cases.
-    First, assume that $u$ is an element of $V_{X+Y}$ and
-    \begin{enumerate}
-        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
-        %       The last requirement means that there necessarily exist $x = <_{X}^{\mu}(f_{E}(z)) = f_{E}(<_{Z}^{\mu}(z))$ and $y = <_{Y}^{\mu}(g_{E}(z)) = g_{E}(<_{Z}^{\mu}(z))$.
-        %       We then define $<_{\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{X+Y}^{\mu}(u)]$
-        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
-              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
-              \]
-        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
-              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
-              We then define 
-              \[
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
-              \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
-        \]
-        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
-              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
-              \[ 
-                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-              \]
-    \end{enumerate}
-    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-    Now, assume that $u$ is an element of $E_{X+Y}$.
-    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
-    Consider two cases.
-    \begin{enumerate}
-    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
-            Then we define
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
-        \]
-    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-        \] 
-    \end{enumerate}
-    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
-    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
-    Clearly, all the cases above are disjoint.
-
-    Let's show that the definition does not depend on a particular choice of $[v]$ in (4).
-    Suppose, $[u) = \varnothing$ and there exist $[v_1]$ and $[v_2]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [e_1]$ and $<_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [e_2]$ and $[e_1] \not = [e_2]$ and such that there is a path from $[u]$ to $[v_1]$ and from $[u]$ to $[v_2]$.
-    First, both $v_1$ and $v_2$ should have a pre-image in $V_{Z}$, otherwise it would mean that there is a path from, \textit{e.g.},  $u$ to $v_1$ and that $[v_1) \not = \varnothing$ which would imply that $[u) \not = \varnothing$ and we get contradiction.
-    Then consider several cases.
-    \begin{itemize}
-        \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = f_{V};\iota_{1,V}(z_2)$.
-        \begin{itemize}
-            \item $u$ has a pre-image in $V_{X}$. Then $[f_{V}(z_1)) = \varnothing$ and $[g_{V}(z_1)) \not = \varnothing$, similarly for $z_2$.
-                  Hence, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_1]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] = [e_1]
-                  \]
-                  and
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}([v_2]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_2]
-                  \]
-                  Then, $[e_1] \not = [e_2]$ implies
-                  \[
-                    [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_1)] \not = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)]
-                  \] and because $[]_{E}$ and $\iota_{2,E}$ are injective, it further means that
-                  \[
-                    <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_1)) \not = <_{Y}^{\mu}(g_{V};\iota_{V_{Y}}(z_2))
-                  \]
-                  which contradicts assumption (3) from~\ref{pushout:assumptions}.
-                  \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item $v_1 = f_{V};\iota_{1,V}(z_1)$ and $v_2 = g_{V};\iota_{1,V}(z_2)$
-                \begin{itemize}
-                    \item $u$ has a pre-image in $V_{X}$. Then, $[f_{V}(z_1)) = \varnothing$, $[g_{V}(z_1)) \not = \varnothing$, $[f_{V}(z_2)) = \varnothing$, $[g_{V}(z_2)) \not = \varnothing$.
-                          This contradicts the assumption (4) from~\ref{pushout:assumptions}.
-                    \item Case when $u$ has a pre-image in $V_{Y}$ is symmetric.
-                \end{itemize}
-                \item The remaining two cases are symmetric.
-    \end{itemize}
-
-    Let's check that this is a well-defined function.
-    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
-    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
-    \begin{itemize}
-        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
-        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
-    \end{itemize} 
-    Consider the first case. Then, by definition.
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    and
-    \[
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-    \]
-    so we have well-defined-ness on the nose. 
-    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
-
-    We also need to make sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms with respect to $<^{\mu}$.
-    \begin{enumerate}
-         \item $[<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
-              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
-              This is homomorphic by definition. 
-        \item $[<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]_{E} = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
-              Let's make case analysis on $[v]_{V}$.
-              \begin{itemize}
-                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$.
-                    \begin{itemize}
-                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $\iota_{1,V}$ is a coproduct injection, it must be the case that $[\iota_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
-                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
-                              by definition, we have 
-                              \begin{align*}
-                                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{V_{X} + V_{Y}}[v]) &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]\\ 
-                                &= [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))]\\
-                                &= [\iota_{1,E}(e_x)] ~.
-                            \end{align*}
-                            Because $\iota_{1,E}$ is a homomorphism, we also have 
-                            \[
-                                [<_{X+Y}^{\mu}(f_{V};\iota_{1,V};\iota_{V_{X} + V_{Y}}(z))] = [\iota_{1,E}(<_{X}^{\mu}(f_{V};\iota_{V_{X}}(z)))] = [\iota_{1,E}(e_x)] ~.
-                            \]
-                    \end{itemize}
-                \item The case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
-              \end{itemize}
-    \end{enumerate}
-    Then $<_{\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
-
-    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
-    We can consider the consistency relation from the coproduct as a function 
-    \[
-        \consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
-    \]
-    Quotienting the values of the function gives us 
-    \[
-        \consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
-     \]
-    which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
-    \begin{enumerate}
-        \item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
-                \begin{align*}
-                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-                    [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
-                \end{align*}
-        \item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ such that there is an undirected path from $[e]$ to $[v]$ and such that 
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing
-        \]
-        then
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]_{V}))
-        \]
-    \end{enumerate}
-
-    and for vertices
-    \begin{enumerate}
-
-    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$, then
-    \begin{align*}
-        &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))\\
-        &= [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')))
-    \end{align*}
-    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $[u]$ such that there is an undirected path from $[v]$ to $[u]$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([u])) \not = \varnothing$, then
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{E_{X} + E_{Y}}([u]_{V}))
-    \]
-    \end{enumerate}
-    where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
-    % We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.
-
-    We then define
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
-    \]
-    \[
-      \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
-    \]
-    where $c$ is a reflexive, symmetric and transitive closure.
-
-    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency$.
-
-    \begin{itemize}
-        \item Let us show well-definedness.
-        \begin{itemize}
-
-            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then we need to show that
-            \[
-                \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{E_{X} + E_{Y}}[v_2])
-            \]
-              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};\iota_{1,V}(z)$ and $v_2 = g_{V};\iota_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\
-                & = ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-              \end{align*}
-              Similarly,
-              \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-                &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_i))))^{c}
-            \end{align*}
-              and we get well-definedness on the nose.
-              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
-            \item Now suppose that both 
-            \[\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_1)) = \varnothing
-            \quad 
-            \text{and} 
-            \quad\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v_2)) = \varnothing
-            \]
-            This would mean that for all $z \in V_{Z}$, $\consistency(f_{V}(z)) = \varnothing$ and $\consistency(g_{V}(z)) = \varnothing$ and automatically
-                  \[
-                    \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-                  \]
-            \end{itemize}
-        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
-            \begin{align*}
-             &[[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-             &\subseteq
-             \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-            By definition, 
-            \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x])) =\\
-                 [[]_{V}^{\consistency}, []_{E}^{\consistency} ]^{*} &(\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))
-            \end{align*}
-            and by recalling that $[]_V^{\consistency} = []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion 
-            \begin{align*}
-                &[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)))\\
-                &\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))
-            \end{align*}
-             and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([x]))$ because it is a closure.
-            If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness holds because empty set is a subset of any set.
-            Similarly for edges.
-    \end{itemize}
-
-    So far we have shown that relations $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ are well-defined and $([]_{V},[]_{E})$ is a homomorphism.
-    Let's now prove the universal property.
-    Let $u_{V}([\iota_{1,V}(x)]) = j_{1,V}(x)$ and $u_{V}([\iota_{2,V}(y)]) = j_2(y)$.
-    We will need to check that this a well-defined function, while homomorphism-ness follows from homomorphism-ness of $j_{1,V}$ and $j_{2,V}$.
-
-    Let's say that $t_1 \sim t_2$ then we need to show that $u_{V}([t_1]) = u_{V}([t_2])$.
-    Because $t_1 \sim t_2$ it means that there exists $z \in V_{Z}$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$.
-    Then, we have
-    \[
-    u_{V}([t_1]) = u_{V}([f_{V};\iota_{1,V}(z)]) = j_1{f_{V}(z)}
-    \] and 
-    \[
-    u_{V}([t_2]) = u_{V}([g_{V};\iota_{2,V}(z)]) = j_{2}(g_{V}(z))
-    \] which are equal by commutativity.
-\end{proof}
-
-So far we have shown that the constructed $\sfrac{X+Y}{\sim}$ is a pushout but we are yet to show that it is an e-hypergraph.
-The proof of the latter will follow below and the following corollaries will come in handy.
-
-% By construction we have the following corollaries for $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$.
-
-\begin{corollary}
-    \label{corollary:nodes}
-    If $[v) = \varnothing$ and $[[v]) \not = \varnothing$ then there exist $v''$ such that there exists an undirected path from $v$ to $v''$ and there exists $v' \sim v''$ such that
-    \begin{align*}
-         [v') \not = \varnothing,\\
-         <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}
-\end{corollary}
-\begin{corollary}
-    \label{corollary:edges}
-    If $[e) = \varnothing$ and $[[e]) \not = \varnothing$ then there exist $v$ such that there is an undirected path from $e$ to $v$ and such that there exists $v' \sim v$
-    \begin{align*}
-        [v') \not = \varnothing,\\
-        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \end{align*}  
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_nodes}
-    If $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \not = \varnothing$ then there exists $v'$ such that there is a path from $v$ to $v'$ and there exists $v'' \sim v'$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{corollary}
-    \label{corollary:consistency_edges}
-    If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ then there exists $v$ such that there is a path from $e$ to $v$ and there exists $v' \sim v$ such that
-    \[
-        \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v')) \not = \varnothing
-    \]
-    \[
-        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))
-    \]
-\end{corollary}
-
-\begin{lemma}
-\label{lemma:child_irreflexive}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive.
-\end{lemma}
-\begin{proof}
-    Suppose there exist $[e]$ such that $[e] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))$.
-    Then we have two cases.
-    \begin{itemize}
-      \item Assume $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e)) = e_1$, then  $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e_1]$. 
-            This would mean that $e_1 = e$ as $[]_{E}$ is injective, which would mean that the relation $<_{X+Y}^{\mu}$ is not irreflexive because $e_1$ is the predecessor of $e$.
-      \item Assume $[e) = \varnothing$, then, by Corollary~\ref{corollary:edges} there exists $v$ such there is a path from $e$ to $v$ and there exists $v' \sim v$ and $[v') \not = \varnothing$ and therefore $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \begin{itemize}
-              \item $v$ has a pre-image in $V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and since there is a path from $e$ to $v$ $[f_{V}(z)) = \varnothing$ and $v' = g_{V};\iota_{2,V}(z)$, $[g_{V}(z)) \not = \varnothing$.
-                    Then, 
-                    \[
-                      <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e]
-                    \]
-                    This means that $e$ has a pre-image in $E_{Y}$, but because there is a path from $e$ to $v$, $e$ should have a pre-image in $E_{X}$ (because $v$ has one in $V_{X}$). 
-                    Contradiction.
-              \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric.
-            \end{itemize}
-    \end{itemize}                  
-      We do not need to check this requirement for vertices because a vertex can not be a predecessor.
-\end{proof}
-
-We will then proceed by showing that both relations satisfy the properties as defined in Definition~\ref{def:e-homo}.
-
-\begin{lemma}
-\label{lemma:child_assymetric}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ is asymmetric.
-\end{lemma}
-\begin{proof}
-    Let $e_1,e_2$ be edges. 
-    Then, if $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ it must not be the case that $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-    For $[e_1] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_2]))$ we have two cases.
-    \begin{itemize}
-      \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_2)) = e_3$ then $[e_1] = [e_3]$ which yields $e_1 = e_3$.
-            Having at the same time $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ gives us also two cases.
-      \begin{itemize}
-          \item $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_4$ then $[e_2] = [e_4]$ and $e_2 = e_4$.
-           This would mean that $e_1$ is a predecessor of $e_2$ and vice versa which violates the e-hypergraphness of $X+Y$.
-          \item $[e_1) = \varnothing$ and then by Corollary~\ref{corollary:edges} there exists $v$ such that there is a path from $e_1$ to $v$ and there exists $v' \sim v$ such that $[v') \not = \varnothing$.
-                Then, $[e_2] = [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])]$.
-                \begin{itemize}
-                  \item If $v$ has a pre-image in $V_{Z}$ such that $f_{V};\iota_{1,V}(z)$ (and hence $e_1$ has a pre-image in $E_{X}$) then necessarily $[f_{V}(z)) = \varnothing$ (because $[e_1) = \varnothing$) and hence $[g_{V}(z)) \not = \varnothing$.
-                        Then,
-                        \[
-                          [<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])] = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                        \]
-                        This means that $e_2$ has a pre-image in $E_{Y}$, but because $e_1$ is the predecessor of $e_2$ they both should have pre-image in either $E_{X}$ or $E_{Y}$.
-                  \item Case when $v = g_{V};\iota_{2,V}(z)$ is symmetric to the above.
-                \end{itemize}
-          \end{itemize}
-          \item $[e_2) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_2$ such that there is a path from $e_2$ to $v_2$ and there exists $v'_2 \sim v_2$ such that $[v'_2) \not = \varnothing$.
-                Then, 
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_2])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [e_1]
-                \]
-                The case when $[e_1) \not = \varnothing$ is similar to the above, so we will just jump to the interesting case when 
-                $[e_1) = \varnothing$ and
-                \[
-                  <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]
-                \]
-                The cases we have are the following.
-                \begin{itemize}
-                  \item $v_2 = f_{V};\iota_{1,V}(z_2)$ and necessarily $[f_{V}(z_2)) = \varnothing$ and $v'_2 = g_{V};\iota_{2,V}(z_2)$.
-                  Then, 
-                  \[
-                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z_2)] = [e_1]
-                  \] which also implies that $e_1$ has a pre-image in $E_{Y}$.
-                  \begin{itemize}
-                      \item $v_1 = g_{V};\iota_{2,V}(z_1)$ (because there is a path from $e_1$ to $v_1$) and necessarily $[g_{V}(z_1)) = \varnothing$ (because $[e_1) = \varnothing$) and $v'_1 = f_{V};\iota_{1,V}(z_1)$.
-                      Then,
-                      \[
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z_1)] = [e_2]
-                      \]  
-                      However, it means that $[f_{V}(z_2)) = \varnothing$ and $[f_{V}(z_1)) \not = \varnothing$ which violates assumption (3) in Theorem~\ref{th:existence_of_pushouts}.
-                  \end{itemize}
-                  \item Case when $v_2 = g_{V};\iota_{2,V}(z_2)$ is symmetric.
-                \end{itemize}
-    \end{itemize} 
-    Once again, we do not need to check for vertices as a vertex can not be on the left-hand side of the relation.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:child_respect}
-    $<_{\sfrac{X+Y}{\sim}}^{\mu}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-        We need to show that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])$.
-        We have two cases.
-                  \begin{itemize}
-                    \item If $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1)) = e_3$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))] = [e_3] = [e_2]$ which means that $e_2$ is the predecessor of $e_1$.
-                          $[v] \in s([e_1])$ means that there exists $v'$ such that $[v] = [v']$ and $v' \in s(e_1)$.
-                          \begin{itemize}
-                            \item $v = v'$ then $v \in s(e_1)$ and because $e_2$ is the predecessor of $e_1$ it is also a predecessor of $v$ because $X+Y$ is an e-hypergraph.
-                                That is, $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))$ and we have $[e_2] = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-                            \item $v \sim v'$ which means there exists $z \in V_{Z}$ such that $v = f_{V};\iota_{1,V}(z)$ and $v' = g_{V};\iota_{2,V}(z)$ and $v' \in s(e_1)$.
-                                  Once again, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ implies $e_2 = <_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))$ and 
-                                  \begin{align*}
-                                    [e_2] &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v'))]\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                                \end{align*}
-                            \end{itemize}
-                    \item $[e_1) = \varnothing$ and by Corollary~\ref{corollary:edges} there exists $v_1$ such that there is a path from $e_1$ to $v_1$ and there exists $v'_1 \sim v_1$ such that $[v'_1) \not = \varnothing$.
-                          Then, $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))$.
-                          Suppose $v_1 = f_{V};\iota_{1,V}(z)$ and then $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$ and 
-                          \[
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [e_2]
-                          \] and hence $e_2$ has a pre-image in $E_{Y}$.
-                          \begin{itemize}
-                            \item $v = v'$ and then $v \in s(e_1)$ and $[v) = \varnothing$.
-                                  Because there is a path from $e_1$ to $v_1$ there is also a path from $v$ to $v_1$ and hence $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [e_2]$
-                            \item $v \sim v'$ and $v' \in s(e_1)$ and $[v') = \varnothing$.
-                                  By the same argument, 
-                                  \begin{align*}
-                                    <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                                                                                                          &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]))\\ 
-                                                                                                          &= [e_2]
-                                   \end{align*}
-                          \end{itemize}                    
-                  \end{itemize}
-            \item Finally, we need to check that if $[v] \in s([e_1])$ and $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then $[e_2] = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))$.
-            Once again, $[v] \in s([e_1])$ means that there exists $v' \in s(e_1)$ such that $v' \sim v$.
-            \begin{itemize}
-                \item $v' = v$ and hence $v \in s(e_1)$.
-                \begin{itemize}
-                    \item If $v$ has no pre-image in $V_{Z}$ and $[v) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v))]
-                    \end{align*}
-                          This essentially means that $e_2$ is the predecessor of $v$ and, by e-hypergraphness of $X+Y$, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                          \begin{align*}
-                            [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\ 
-                                  &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                          \end{align*}
-                           which is what we want.
-                    \item $v$ has no pre-image in $V$ and $[v) = \varnothing$.
-                          By Corollary~\ref{corollary:nodes} there exists $v'$ such that there is a path from $v$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                          Then, 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                          Because $v \in s(e_1)$, there should also be a path from $e_1$ to $v'$ and 
-                          \begin{align*}
-                            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v']))\\
-                             &= [e_2]
-                          \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then
-                    \begin{align*} 
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
-                    \end{align*}
-                    This means that the pre-image of $e_2$ is the predecessor of $f_{V}(z)$ and hence $e_2$ should be the predecessor of $f_{V};\iota_{1,V}(z)$ and hence of $e_1$ since $v_1 \in s(e_1)$ and $X+Y$ is an e-hypergraph.
-                    That is, $e_2 = <_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))$ and 
-                    \begin{align*}
-                        [e_2] &= [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(e_1))]\\
-                               &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e_1]))
-                    \end{align*}
-                    \item $v$ has a pre-image $f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then
-                    \begin{align*}
-                        [e_2] &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                              &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]
-                    \end{align*} 
-                    This means that $e_2$ has a pre-image in $E_{Y}$ and $[e_1) = \varnothing$ since $e_1$ has a path to $v$.
-                    Then, by definition, 
-                    \begin{align*}
-                        <_{\sfrac{X+Y}{\sim}}^{\mu}([e_1]) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])\\
-                                                           &= [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)]\\
-                                                           &= [e_2]
-                    \end{align*}
-                    \item Cases when $v$ has a pre-image $g_{V};\iota_{2,V}(z)$ are symmetric.
-                \end{itemize}
-                \item $v \sim v'$.
-                    Follows by the same argument as above by recalling that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-            \end{itemize}    
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_respect}
-$\consistency_{\sfrac{X+Y}{\sim}}$ respects connectivity.
-\end{lemma}
-\begin{proof}
-    Suppose $[v] \in s([e])$ and $[[v]) = [[e]) \not = \varnothing$, then it must be the case that $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$.
-    $[v] \in s([e])$ means there exists $v' \in s(e)$ such that $v \sim v'$.
-    We then have two cases.
-    \begin{itemize}
-      \item $v = v'$ and then $v \in s(e)$.
-      \begin{itemize}
-          \item $[v) \not = \varnothing$ and then $[e) \not = \varnothing$ and necessarily $\iota_{E_{X} + E_{Y}}(e) \in \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))$.
-                  By applying $([]_{V},[]_{E})$ we get 
-                  \begin{align*}
-                  \iota_{\sfrac{E_{X}+E_{Y}}{\sim}}([e]) &\in \\
-                  [[]_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}},[]_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}})]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v))) &\\
-                  \subseteq
-                  \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&\\
-                  \subseteq
-                   \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])&
-                  \end{align*}
-          \item $[v) = \varnothing$ and then $[e) = \varnothing$ then there exists $v'$ such that there is a path from $e$ to $v'$ and $v'' \sim v'$ such that $[v'') \not = \varnothing$.
-                It is then also the case that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) = \varnothing$ and $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'')) \not = \varnothing$.
-                Then, by definition, 
-                \begin{align*}
-                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))\\
-                 &= \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v'']))
-                \end{align*}
-                and by reflexivity 
-                \[
-                    \iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
-                \]
-      \end{itemize}
-      \item $v \sim v'$ follows by the same argument by using functionality, 
-          \[
-              \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v']) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
-          \]
-     \end{itemize}
-The case when $[v] \in t([e])$ is symmetric.
-\end{proof}
-
-\begin{lemma}
-\label{lemma:consistency_child}
-    $\consistency_{\sfrac{X+Y}{\sim}}$ is defined only for vertices and edges that share the same non-empty set of parents.
-\end{lemma}
-\begin{proof}
-  If $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1]) \in \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))$ then
-        $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_1])) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}([v_2]))$.
-        By definition, one case is when
-        \begin{align*}
-            &\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))\\
-            &= ([[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*} (\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v_2]_{V}} \consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v'))))^{c}
-        \end{align*}
-        which means that either $v_1~\consistency_{X+Y}~v_2$ or $v'_1 ~\consistency_{X+Y}~ v_2$ and $v'_1 \sim v_1$, or $v_1 ~\consistency_{X+Y}~ v_2'$ and $v_2' \sim v_2$, or $v_1' ~\consistency_{X+Y}~ v_2'$ and $v_1' \sim v_1$, $v_2' \sim v_2$.
-        Suppose, $v_1' ~\consistency_{X+Y}~ v_2$ and $v_1' \sim v_1$.
-        It implies that $[v_1') = [v_2) \not = \varnothing$.
-        If $v_1$ has no pre-image in $V_{Z}$, then $v_1 = v'_1$ and we have
-        \begin{align*}
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1']))\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_1'))]\\
-             &= [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v_2))]\\
-             &= <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
-        \end{align*}
-        If $v_1$ has a pre-image in $V_{Z}$ and $v_1 = f_{V};\iota_{1,V}(z)$ and $[f_{V}(z)) = \varnothing$ (because $[v'_1) = [v_2) \not = \varnothing$) and $v'_1 = g_{V};\iota_{2,V}(z)$ and $[g_{V}(z)) \not = \varnothing$.
-        Then,
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        and
-        \[
-            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{X+Y}^{\mu}(\iota_{V_{X} +V_{Y}}(v_2))] = [\iota_{2,E}(<_{Y}^{\mu}(\iota_{V_{Y}}(v_2)))]
-        \]
-        The rest cases are analogous.
-        Another case is when there exists $v_3$ such that there is a path from $v_2$ to $v_3$ and there is $v_3' \sim v_3$ such that $\consistency_{X+Y}(v_3') \not = \varnothing$ and
-        \[
-            \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_3]))
-        \]
-        Then the argument is the same but with using $v_3$ in place of $v_2$.
-  Case for edges is analogous.
-\end{proof}
-
-\begin{proposition}[$\sfrac{X+Y}{\sim}$ is an e-hypergraph]
-For this we need to check that the defined relations --- $<_{\sfrac{X+Y}{\sim}}$ and $\consistency_{\sfrac{X+Y}{\sim}}$ --- satisfy all the needed properties as per Definition~\ref{def:e-homo}.
-\end{proposition}
-\begin{proof}
-    This proposition then follows by applying the lemmas above.
-    % Lemmas~\ref{lemma:child_assymetric}~\ref{lemma:child_irreflexive}~\ref{lemma:child_respect}~\ref{lemma:consistency_child}~\ref{lemma:consistency_respect}.
-\end{proof}
-
-We will then discuss the existence and uniqueness of the corresponding pushout complement by first recalling the result for plain hypergraphs.
-
-\begin{definition}[Definition 3.16~\cite{bonchi_string_2022-1}]
-Morphisms $\mathcal{K} \xrightarrow{f} \mathcal{L} \xrightarrow{m} \mathcal{G}$ satisfy the \textit{no-dangling} condition if, for every hyperedge not in the image of $m$, every vertex of its source and target is either (i) not in the image of $m$ or (ii) in the image of $f ; m$.
-They satisfy the no-identification condition if any two vertices merged by $m$ are in the image of $f$.
-\end{definition}
-
-\begin{proposition}
-\label{prop:complement_existence}
-    Consider a pushout square in $\catname{EHyp_{\Sigma}}$ below
-\[
-    \tikzfig{appendix/pushout_square}
-\]
-    Such that $m,f$ are monos and $\mathcal{K}$ is discrete.
-    Then, pushout complement $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ exists if
-    \begin{enumerate}
-        \item $[m_{V}(f_{V}(v_i))) = [m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item $\consistency(m_{V}(f(v_i))) = \consistency(m_{V}(f_{V}(v_j)))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-        \item The no-dangling condition holds.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    By requiring that $m$ is mono makes the no-identification condition satisfied on the nose.
-    Then, no-dangling and no-identification conditions are necessary for pushout complement to exist~\cite{dpo}.
-    Points (1)-(2) above guarantee that after removing from $\mathcal{G}$ everything in $\mathcal{L}$ that has no pre-image in $\mathcal{K}$ the arrow $\mathcal{K} \to \mathcal{L}^{\bot}$ will satisfy the assumptions~\ref{pushout:assumptions}.
-    In particular, they mean that after removing the image of $\mathcal{L}$ from $\mathcal{K}$ we have the following
-    \begin{itemize}
-        \item $[c_{V}(v_{i})) = [c_{V}(v_{j}))$ for all $v_{i},v_{j}$ in $V_{\mathcal{K}}$;
-        \item $\consistency(c_{V}(v_i)) = \consistency(c_{V}(v_j))$ for all $v_i,v_j$ in $V_{\mathcal{K}}$.
-    \end{itemize}
-    This means that $\mathcal{G}$ is isomorphic to the pushout constructed according to Theorem~\ref{th:existence_of_pushouts}.
-\end{proof}
-
-\begin{figure*}[t!]
-    \begin{minipage}{0.1\textwidth}
-        $\;$
-    \end{minipage}
-    \hfill
-\begin{minipage}{0.7\textwidth}
-    \[
-        \scalebox{0.45}{
-            \tikzfig{combinatorial_semantics/structural_rule_pushout}
-        }
-    \]
-    \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-    \label{fig:structural_rule_pushout}
-\end{minipage}
-\hfill
-\begin{minipage}{0.1\textwidth}
-        $\;$
-\end{minipage}
-\end{figure*}
-
-\begin{proposition}
-\label{prop:boundary_complement}
-    The boundary complement in~\ref{def:boundary_new} when exists is unique
-\end{proposition}
-\begin{proof}
-        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.
-        Hence, pushout complement should yield pushout complements for the corresponding sets.
-        In particular,
-          % https://q.uiver.app/#q=WzAsNSxbMCwwLCJWX3tMfSJdLFsyLDAsImxfe2l9ICsgbF97b30iXSxbNCwwLCJpK2oiXSxbMCwyLCJWX3tHfSJdLFs0LDIsIlZfe0N9Il0sWzIsMSwiYl97Vn0iXSxbMSwwLCJhX3tWfSJdLFswLDMsImZfe1Z9Il0sWzQsMywiZ197Vn0iLDJdLFsyLDQsImMiLDJdXQ==
-    % \begin{figure}[h!]
-    %     \begin{subfigure}{0.45\textwidth}
-    %         % \[\begin{tikzcd}
-    %         %     {V_{L}} && {l_{i} + l_{o}} && {i+j} \\
-    %         %     \\
-    %         %     {V_{G}^{\urcorner}} &&&& {V_{C}}
-    %         %     \arrow["{b_{V}}", from=1-5, to=1-3]
-    %         %     \arrow["{a_{V}}", from=1-3, to=1-1]
-    %         %     \arrow["{f_{V}}", from=1-1, to=3-1]
-    %         %     \arrow["{g_{V}}"', from=3-5, to=3-1]
-    %         %     \arrow["c"', from=1-5, to=3-5]
-    %         % \end{tikzcd}\]
-    %         \[
-    %             \tikzfig{appendix/pushout_verticies}
-    %         \]
-    %     \end{subfigure}
-    %     \hfill
-    %     \begin{subfigure}{0.45\textwidth}
-    %        % https://q.uiver.app/#q=WzAsNSxbMCwwLCJFX3tMfSJdLFsyLDAsIjAiXSxbNCwwLCIwIl0sWzAsMiwiRV97R31ee1xcdXJjb3JuZXJ9Il0sWzQsMiwiRV97Q30iXSxbMiwxLCJiX3tFfSJdLFsxLDAsImFfe0V9Il0sWzAsMywiZl97RX0iXSxbNCwzLCJnX3tFfSIsMl0sWzIsNCwiYyIsMl1d
-    %     \[
-    %     % \begin{tikzcd}
-    %     %     {E_{L}} && 0 && 0 \\
-    %     %     \\
-    %     %     {E_{G}^{\urcorner}} &&&& {E_{C}}
-    %     %     \arrow["{b_{E}}", from=1-5, to=1-3]
-    %     %     \arrow["{a_{E}}", from=1-3, to=1-1]
-    %     %     \arrow["{f_{E}}", from=1-1, to=3-1]
-    %     %     \arrow["{g_{E}}"', from=3-5, to=3-1]
-    %     %     \arrow["c"', from=1-5, to=3-5]
-    %     % \end{tikzcd}
-    %     \tikzfig{appendix/pushout_edges}
-    %     \]
-    %     \end{subfigure}
-    % \end{figure}
-    \[
-        \tikzfig{appendix/pushout_verticies}
-    \]
-    \[
-        \tikzfig{appendix/pushout_edges}
-    \]
-    where $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.
-    Because $i+j$ is discrete, $E_{\mathcal{G}}$ is the disjoint union of $E_{\mathcal{L}}$ and $E_{\mathcal{L}^{\bot}}$ and hence $E_{\mathcal{L}^{\bot}} = E_{\mathcal{G}} \setminus E_{\mathcal{L}}$.
-    Then, since all the arrows are monos, the first square can be rewritten as follows.
-% https://q.uiver.app/#q=WzAsNSxbMCwwLCJpICsgaiArIHggKyB5Il0sWzIsMCwiaStqICsgeCJdLFs0LDAsImkraiJdLFswLDIsImkgKyBqICsgeCArIHkgKyB6Il0sWzQsMiwiaSArIGogKyB3XFxcXFxcY29uZ1xcXFwgaSArIGogKyB6Il0sWzIsMSwiaF97VixleHR9IiwyXSxbMSwwLCJoX3tWLGludH0iLDJdLFswLDMsIm1fe1Z9IiwyXSxbNCwzLCJnX3tWfSJdLFsyLDQsImMiLDJdXQ==
-\[\begin{tikzcd}
-	{i + j + x + y} && {i+j + x} && {i+j} \\
-	\\
-	{i + j + x + y + z} &&&& {\begin{matrix} i + j + z \\ \cong \\ i + j + w \end{matrix}}
-	\arrow["{h_{V,ext}}"', from=1-5, to=1-3]
-	\arrow["{h_{V,int}}"', from=1-3, to=1-1]
-	\arrow["{m_{V}}"', from=1-1, to=3-1]
-	\arrow["{g_{V}}", from=3-5, to=3-1]
-	\arrow["c"', from=1-5, to=3-5]
-\end{tikzcd}\]
-    Because $i + j + x + y + w$ also yields a pushout it must be the case that $i+j+x+y+w \cong i + j + x + y + z$ and therefore $w \cong z$ and pushout complement for vertices is $i + j + z$ up to isomorphism.
-    To show that pushout complement is unique we next need to show that source and target maps are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s_{1},t)
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s_{2},t)
-    \]
-    Because $g$ is a homomorphism, it must be the case that $g_{V}^{*}(s_{1}(e)) = s_{\mathcal{G}}^{*}(g_{E}(e)) = g_{V}^{*}(s_{2}(e))$.
-    Because $g$ is mono, $g_{V}^{*}(s_{1}(e)) = g_{V}^{*}(s_{2}(e))$ implies $s_{1}(e) = s_{2}(e)$.
-    Similarly for targets.
-    We also need to show that $<$ and $\consistency$ are unique.
-    Suppose that there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <_{1})
-    \qquad
-    \text{and}
-    \qquad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},E_{\mathcal{L}^{\bot}},s,t, <_{2})
-    \]
-    Because $g$ is homomorphism, it must be the case that 
-    \[
-        g_{E}(<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v))) = g_{E}(<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)))
-    \]
-    which implies $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = <_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ for all $v$ such that $[v) \not = \varnothing$.
-    Similarly for edges.
-    Now suppose that $<_{1}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = e$ and $<_{2}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$ is undefined.
-    Then, $g_{E}(e) = <_{\mathcal{G}}(\iota_{V_{\mathcal{G}}}(g_{V}(v)))$ and there exists $v' \in V_{\mathcal{L}}$ and $v'' \in V_{\mathcal{L}^{\bot}_{2}}$ such that $[v') \not = \varnothing$ and there is a path from $v$ to $v''$ and $v''$ and $v'$ share a pre-image in $i+j$.
-    However, because there is a path from $v$ to $v''$ it must be the case that $[v'') \not = \varnothing$ in $\mathcal{L}^{\bot}_{1}$, but then it violates our assumptions as there exists some vertex $u$ in $i+j$ such that $v'' = c_{V}(u)$ and $[v'') \not = \varnothing$ and $v' = h_{V,ext};h_{V,int}(u)$ and $[v') \not = \varnothing$.
-    Therefore we get a contradiction. 
-    Similarly for edges.
-    Finally, let's suppose there exist
-    \[
-        \mathcal{L}^{\bot}_{1} = (V_{\mathcal{L}^{\bot}}, E_{\mathcal{L}^{\bot}}, s,t, <, \consistency_{1})
-    \quad
-    \text{and}
-    \quad
-        \mathcal{L}^{\bot}_{2} = (V_{\mathcal{L}^{\bot}},V_{\mathcal{L}^{\bot}},s,t,<, \consistency_{2})
-    \]
-    Because $g$ is a homomorphism, we have
-    \[
-        [~g_{V};\iota_{V_{\mathcal{G}}}, g_{E};\iota_{E_{\mathcal{G}}}~]^{*}(\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \[
-        [~g_{V};\iota_{V_{G}}, g_{E};\iota_{E_{\mathcal{G}}}]^{*}(\consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))) \subseteq \consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v))
-    \]
-    \begin{itemize}
-        \item If $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) = \varnothing$ then both left-hand sides should be $\varnothing$ and therefore the relations are equal.
-        \item Suppose that $\consistency_{\mathcal{G}}(g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$ which further means that $[g_{V};\iota_{V_{\mathcal{G}}}(v)) \not = \varnothing$.
-              Consider the following cases.
-        \begin{itemize}
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \subseteq \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v))$.
-                  This means that there exists $v'$ such that 
-                  \[
-                    v' \consistency_{\mathcal{L}^{\bot}_{2}} v
-                    \quad
-                    \text{and}
-                    \quad
-                    v' \not \consistency_{\mathcal{L}^{\bot}_{1}} v
-                \]
-                  Because $g_{V}(v) ~\consistency_{\mathcal{G}}~ g_{V}(v')$ it means that there exist $u,u'$ such that there is a path from $v$ to $u$ and from $v'$ to $u'$ and there exist $w,w' \in V_{\mathcal{L}}$ such that $w \consistency_{\mathcal{L}} w'$ and $w,u$ share pre-image in $i+j$ as well as $w',u'$.
-                  This again contradicts our assumptions as $u \consistency_{\mathcal{L}^{\bot}_{2}} u'$ and $w \consistency_{\mathcal{L}} w'$.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) = \varnothing$
-                  Similar to the above.
-            \item $\consistency_{\mathcal{L}^{\bot}_{1}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \cap \consistency_{\mathcal{L}^{\bot}_{2}}(\iota_{V_{\mathcal{L}^{\bot}}}(v)) \not = \varnothing$
-                  Similar to the above.
-        \end{itemize}
-    \end{itemize}
-\end{proof}
-
-Practically, the above proposition means that for any given match $m$ there is only one possible result of rewriting.
-
-% \begin{figure*}[t!]
-%     \begin{minipage}{0.1\textwidth}
-%         $\;$
-%     \end{minipage}
-%     \hfill
-% \begin{minipage}{0.7\textwidth}
-%     \[
-%         \scalebox{0.45}{
-%             \tikzfig{combinatorial_semantics/structural_rule_pushout}
-%         }
-%     \]
-%     \caption{The image of the top left equation from Figure~\ref{fig:string-equations} under $\llbracket - \rrbracket$}
-%     \label{fig:structural_rule_pushout}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.1\textwidth}
-%         $\;$
-% \end{minipage}
-% \end{figure*}
-
-\section{Proofs for Section \ref{sec:soundness-and-completeness}: Soundness and Completeness}
-
-In this section we will elaborate on soundness of our translation $\llbracket - \rrbracket : \textbf{PROP}^{+}(\Sigma) \to \WellTypedMdaEcospans / \mathcal{S}$.
-
-\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
-The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
-\end{proposition}
-\begin{proof}
-    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textbf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
-    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
-    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
-    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
-    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.
-
-    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
-    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
-    \begin{figure}[h!]
-    \[
-        \scalebox{0.8}{
-        \tikzfig{combinatorial_semantics/DPOI_square}
-        }
-    \]
-    \end{figure}
-    We then further note that if $l$ is a subterm of $g$ then the vertices in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
-    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
-    The no-dangling condition is satisfied because we consider only mda-cospans.
-    Therefore, there exists a unique pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
-    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
-    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
-    Hence, $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
-    The rest follows by induction on the number of applied equations.
-    Similar argument then applies to other rules in $\mathcal{S}$ as well.
-\end{proof}
-
-% \begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]
-%     For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
-%     \[
-%         f = f_1 + \ldots + f_n
-%     \]
-%      such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
-% \end{proposition}
-% \begin{proof}
-%     We will define the depth of an edge $e$ as the number of parents it has, \textit{i.e.}, by $|[e)|$.
-%     Then we will build the required weak normal form in steps and first will build a form $f = f_1 + \ldots + f_n$ such that every of $f_i$ is also of the form $g_1 + \ldots + g_k$ and so on and we will assume that $f$ has at least one hierarchical edge (otherwise is is already a weak normal form).
-%     In order to achieve the latter we will apply rewrite rules that correspond to equations (1)-(4) from Figure~\ref{fig:string-equations}.
-%     At step 1 we search for matches among the edges with depth 0 and then note that each such rewrite reduces the number of such edges until there is only one edge of depth 0, and it is necessarily a hierarchical one: because we have a finite number of edges of each depth, this process terminates.
-%     At step $k$ we similarly search for matches among the edges with depth $k-1$ noting that this process terminates when we end up with $l$ groups of edges of depth $k-1$ such that edges within a group are pair-wise not consistent.
-%     Because we do not increase the maximum depth in $f$ we will eventually reach a maximum value of $k$ and stop because there will be no hierarchical edges of depth $max(k)$ (otherwise their children would have depth $max(k) + 1$).
-
-%     Then we will apply rules (5), from right to left, (6) and (8), from right to left.
-%     Note that each of these rules decreases the number of hierarchical edges, so we apply them until there is only one hierarchical edge of depth 0, \textit{i.e.}, the top-level one.
-%     This brings us to a weak normal form $f = f_1 + \ldots + f_n$ where each of $f_i$ does not contain any hierarchical edges and for $i \not = j$, $f_i \not = f_j$, the latter is guaranteed by rule (8).
-% \end{proof}
+\bibliography{../bibliography}
 

### 2025-06-12 arxiv.py

@@ -1,6 +1,6 @@
 import shutil
 from pathlib import Path
-from pylatexenc.latexwalker import LatexWalker, LatexEnvironmentNode, LatexMacroNode
+from pylatexenc.latexwalker import LatexWalker, LatexEnvironmentNode, LatexMacroNode, LatexSpecialsNode
 from pylatexenc.latex2text import LatexNodes2Text
 import sys
 
@@ -20,7 +20,9 @@ def extract_macros(nodes):
     while True:
         try:
             node = next(nodes_iter)
-            if isinstance(node, LatexMacroNode):
+            if isinstance(node, LatexSpecialsNode):
+                continue
+            elif isinstance(node, LatexMacroNode):
                 if node.macroname in FIGURE_MACROS | INPUT_MACROS | BIB_MACROS:
                     if node.macroname == r'tikzfig':
                         node = next(nodes_iter)

### 2025-06-19 POPL2026/SlottedNotes.tex
slides wip

@@
 \newcommand\update[1]{{\color{americanrose}#1}}
+
+\usepackage{minted}
 
@@
 
+\subsection{E-matching}
+
+E-matching is a function $match(p,m * a, \beta, m_p)$ that takes in a pattern $p$, an applied-id $m*a$ to match against, a subtitution from variables to applied-ids $\beta$ and a renaming $m_p$ that maps slots of an e-graph to slots of the pattern.
+
+\begin{listing}
+\begin{minted}[linenos, fontsize=\small, breaklines]{rust}
+  struct State {
+    // uses egraph slots.
+    partial_subst: Subst, // type Subst = HashMap<String, AppliedId>
+
+    // maps from the pattern slots to the e-graph slots.
+    partial_slotmap: SlotMap, //type SlotMap = HashMap<Slot,Slot>
+  }
+
+  pub fn ematch_all<L: Language, N: Analysis<L>>(
+    eg: &EGraph<L, N>,
+    pattern: &Pattern<L>,
+) -> Vec<Subst> {
+    let mut out = Vec::new();
+    for i in eg.ids() {
+        let i = eg.mk_sem_identity_applied_id(i); // applies identity renaming to an id to turn it into AppliedId
+        out.extend(
+            ematch_impl(pattern, State::default(), i, eg) // this is `match'
+                .into_iter()
+                .map(final_subst),
+        );
+    }
+    out
+}
+
+\end{minted}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[linenos, fontsize=\small, breaklines]{rust}
+fn ematch_impl<L: Language, N: Analysis<L>>(
+    pattern: &Pattern<L>,
+    st: State,
+    i: AppliedId,
+    eg: &EGraph<L, N>,
+) -> Vec<State> {
+    match &pattern {
+        Pattern::PVar(v) => {
+            let mut st = st;
+            if let Some(j) = st.partial_subst.get(v) {
+                if !eg.eq(&i, j) {
+                    return Vec::new();
+                }
+            } else {
+                st.partial_subst.insert(v.clone(), i);
+            }
+            vec![st]
+        }
+        Pattern::ENode(n, children) => {
+            let mut out = Vec::new();
+            for nn in eg.enodes_applied(&i) { //traverses e_nodes of the e-class
+                let d = std::mem::discriminant(n);
+                let dd = std::mem::discriminant(&nn);
+                if d != dd {
+                    continue;
+                };
+
+                //takes the e_node that matches 'n'.
+                ematch_node(&st, eg, &n, children, &mut out, &nn);
+            }
+            out
+        }
+        Pattern::Subst(..) => panic!(),
+    }
+}
+
+\end{minted}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[linenos, fontsize=\small, breaklines]{rust}
+fn ematch_node<L: Language, N: Analysis<L>>(
+    st: &State,
+    eg: &EGraph<L, N>,
+    n: &L, // from the pattern
+    children: &[Pattern<L>],
+    out: &mut Vec<State>,
+    nn: &L, //candidate e_node from e-class
+) {
+    'nodeloop: for n2 in eg.get_group_compatible_weak_variants(&nn) {
+        if CHECKS {
+            assert_eq!(&nullify_app_ids(n), n);
+        }
+
+        let clear_n2 = nullify_app_ids(&n2);
+        // We can use weak_shape here, as the inputs are nullified
+        // i.e. they only have id0() without slot args, so there are no permutations possible.
+        let (n_sh, _) = n.weak_shape();
+        let (clear_n2_sh, _) = clear_n2.weak_shape();
+        if n_sh != clear_n2_sh {
+            continue 'nodeloop;
+        } //checks if the shapes are equal, for example lambda x . f and lambda y . f have the same shape
+
+        let mut st = st.clone();
+
+        for (x, y) in clear_n2
+            .all_slot_occurrences()
+            .into_iter()
+            .zip(n.all_slot_occurrences().into_iter())
+        {
+          // checks if mapping m_p is not ambiguous
+            if !try_insert_compatible_slotmap_bij(x, y, &mut st.partial_slotmap) {
+                continue 'nodeloop;
+            }
+        }
+
+        let mut acc = vec![st];
+        // recursively matches children
+        for (sub_id, sub_pat) in n2.applied_id_occurrences().into_iter().zip(children.iter()) {
+            let mut next = Vec::new();
+            for a in acc {
+                next.extend(ematch_impl(sub_pat, a, sub_id.clone(), eg));
+            }
+            acc = next;
+        }
+
+        out.extend(acc);
+    }
+}
+
+\end{minted}
+\end{listing}
+
 \bibliographystyle{acm}

### 2025-06-19 slides/EGRAPHS_2024/slides.tex
slides wip

--- 2024-09-17 slides/slides.tex
@@
 
-\input{../sample.tikzstyles}
-\input{../hypergraph.tikzstyles}
-\input{../hypergraph.tikzdefs}
+\input{../../sample.tikzstyles}
+\input{../../hypergraph.tikzstyles}
+\input{../../hypergraph.tikzdefs}
 
@@
 
-% \setsansfont{Ubuntu}
-% \setmonofont{Ubuntu Mono}
+\setsansfont{Ubuntu}
+\setmonofont{Ubuntu Mono}
 
@@
 \date{June 24, 2024} % Use metropolis theme 
-\author[Aleksei Tiurin]{$\text{Chris Barrett}^{\dagger}$, $\textbf{Aleksei Tiurin}^{\ddagger}$, $\text{Dan R. Ghica}^{\ddagger}$}
-\institute{$^{\dagger}$University of Oxford, $^{\ddagger}$University of Birmingham} 
+\author[Aleksei Tiurin]{$\textbf{Aleksei Tiurin}^{\dagger}$,$\text{Chris Barrett}^{\dagger}$,$\text{Dan R. Ghica}^{\dager,\ddagger}$, $\text{Nick Hu}^{\dagger}$}
+\institute{$^{\dagger}$University of Birmingham, $^{\dagger}$ Huawei Central Software Institute}
 \begin{document}
@@
         \scalebox{0.6}{
-        \tikzfig{../figures/categorical-semantics/egraph-strings-equations}
+        \tikzfig{../../figures/categorical-semantics/egraph-strings-equations}
         }
@@
 	\scalebox{0.7}{
-  	 \tikzfig{../figures/categorical-semantics/Cartesian-equipment}
+  	 \tikzfig{../../figures/categorical-semantics/Cartesian-equipment}
 	}
@@
             \scalebox{0.55}{
-            \tikzfig{../figures/categorical-semantics/Cartesian-theory}	
+            \tikzfig{../../figures/categorical-semantics/Cartesian-theory}	
             }
@@
             \scalebox{0.55}{
-            \tikzfig{../figures/categorical-semantics/Cartesian-naturality}
+            \tikzfig{../../figures/categorical-semantics/Cartesian-naturality}
             }
@@
         \scalebox{0.38}{
-        \tikzfig{../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
+        \tikzfig{../../figures/categorical-semantics/egraph-translation-step-by-step-a-b}
         }
@@
         \scalebox{0.4}{
-        \tikzfig{../figures/categorical-semantics/e-graph-substitution}
+        \tikzfig{../../figures/categorical-semantics/e-graph-substitution}
         }
@@
         \scalebox{0.35}{
-        \tikzfig{../figures/categorical-semantics/e-string-substitution-2}
+        \tikzfig{../../figures/categorical-semantics/e-string-substitution-2}
         }

### 2025-07-04 POPL2026/SlottedNotes.tex
notes update

@@
 
+\section{Explicit substitution}
+
+Lambda calculus with explicit substitution (also $\lambda_{\sigma}$-calculus) made the substitution operator a first-class element of the calculus (rather than a meta-level operation) to bridge the gap between lambda calculus and its concrete implementations~\cite{Abadi_Cardelli_Curien_Lévy_1991}.
+However, the existence of unique normal forms turned out to be a delicate matter and to achieve uniqueness, lambda terms must be quotiented by \textit{graphical} equivalence in addition to $\alpha$-equivalence~\cite{Standardisation}.
+We claim that using e-hypergraphs over e-graphs (or slotted e-graphs) may be more beneficial for doing equational reasoning for lambda calculus with explicit substitution.
+
+The rules can be seen in a table below (plust a standard rule for $(\lambda x . u)v = u[x / v]$)
+
+\begin{tabular}{lcc}
+\texttt{subst\_same\_var} & $x[x / u] = u$ &\\
+\texttt{subst\_diff\_var} & $y[x / u] = y$ & if $x \not \in \mathcal{F}(u)$\\
+\texttt{subst\_application} & $(uu')[x / v] = u[x / v]u'[x / v]$ & \\
+\texttt{subst\_lambda} & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$ & if $x \not \in \mathcal{F}(v)$ \\
+\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$\\
+\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])[x / v[x / v']]$ & if ???\\
+\texttt{subst\_comm} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$ & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)$\\
+\end{tabular}
+
+while such rules are absorbed when using string diagrams (or hypergraphs) akin to proof nets, they are not absorbed by e-graph representation.
+Below one could see the number of e-nodes for different set of equations enabled for saturating an expression \texttt{plus 2 2} using Church encoding.
+
+\begin{tabular}{lcccc}
+  Set of equations & built-in subst & explicit subst & explicit subst & explicit subst\\
+  &&& + interchange&  + interchange\\
+  &&&&  + comm\\
+  \# of e-nodes & 40 & 161 & 289 & 356
+\end{tabular}
+
+String diagrams represent the same equivalence without loss of generality and without handling equations for substitutions explicitly.
+
 \bibliographystyle{acm}

### 2025-07-07 POPL2026/notes/SlottedNotes.tex
notes and popl template

--- 2025-07-04 POPL2026/SlottedNotes.tex
@@
 
-\input{../preamble}
-\input{../macros}
-\input{../sample.tikzstyles}
-\input{../hypergraph.tikzstyles}
-\input{../hypergraph.tikzdefs}
+\input{../../preamble}
+\input{../../macros}
+\input{../../sample.tikzstyles}
+\input{../../hypergraph.tikzstyles}
+\input{../../hypergraph.tikzdefs}
 
@@
 while such rules are absorbed when using string diagrams (or hypergraphs) akin to proof nets, they are not absorbed by e-graph representation.
-Below one could see the number of e-nodes for different set of equations enabled for saturating an expression \texttt{plus 2 2} using Church encoding.
+In Table~\ref{tbl:slotted} one could see the number of e-nodes for different set of equations enabled for saturating an expression \texttt{plus 2 2} using Church encoding.
 
+\begin{table}
 \begin{tabular}{lcccc}
@@
 \end{tabular}
+\caption{Slotted e-graphs benchmarks}
+\label{tbl:slotted}
+\end{table}
 
@@
 
+Recall that by the virtue of free enrichment we have the equality $\Lambda(f + g) = \Lambda(f) + \Lambda(g)$.
+The intuitive interpretation of this equality is that $f \sim g$ (due to rewrite rules being applied) implies $\Lambda(f) \sim \Lambda(g)$.
+This is exactly one of the equivalences that a slotted e-graph from~\cite{slotted-egraphs} maintains.
+
+\[
+\text{bind}\;\$s_1\; x \cong \text{bind}\;\$s_2\;y \;\; \text{if} (\$s_1 \$f) * x \cong (\$s_2 \$f) * y, \text{where $f$ is fresh}
+\]
+
 \bibliographystyle{acm}
-\bibliography{../bibliography}
+\bibliography{../../bibliography}
 

### 2025-07-07 POPL2026/paper/sections/categorical_semantics.tex
notes and popl template

--- 2025-01-26 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
@@
 It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting and in interpreting $\lambda$-terms.
-Figure \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
+Fig.
+\ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
 Note the orientation of string diagrams are now from top to bottom as opposed to what was used in~\nameref{sec:introduction}: the latter was mimicking the orientation of e-graphs, \textit{i.e.}, traditional orientation for expression trees.
-% Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
+% Fig. \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
 % The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.

### 2025-07-07 POPL2026/paper/sections/soundness.tex
notes and popl template

--- 2025-01-27 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
@@
 \begin{definition}
-    For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
+    For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle} g$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
     $
@@
 For example, the distributivity rule for $\otimes$ is given by the following schema rule
+\ifdefined\ONECOLUMN
+\[
+\adjustbox{width=0.7\linewidth}{
+\tikzfig{./figures/semilattice_rule_1}~.
+}
+\]
+\else
 \[
@@
 
-This explicit quotienting allows us to only consider cospans of the form $f_1 + \ldots + f_{n}$, similar to Lemma~\ref{lemma:normal_form}, such that carrier of $f_{i}$ does not contain any hierarchical edges from $\textcolor{red}{E}$.
+This explicit quotienting allows us to only consider cospans of the form $f_1 + \ldots + f_{n}$, similar to Lemma~\ref{lemma:normal_form}, such that carrier of $f_{i}$ does not contain any hierarchical edges from $\textcolor{e-color}{E}$.
 
@@
 \label{prop:fscd}
-For two closed $\Sigma$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
+For two closed $\Sigma$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$, where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
 \end{proposition}
@@
 \begin{proposition}
-    For two closed $\Sigma^{+}$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
+    For two closed $\Sigma^{+}$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
 \end{proposition}
@@
 such that $f_{i} \leadsto f_{j}$ for some $i, j$.
-This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and, by noting that every $f_{i}$ is a $\Sigma$-term, and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{red}{E}$, we can apply Proposition~\ref{prop:fscd}.
+This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and, by noting that every $f_{i}$ is a $\Sigma$-term, and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{e-color}{E}$, we can apply Proposition~\ref{prop:fscd}.
 \end{proof}

### 2025-07-08 POPL2026/paper/sections/introduction.tex
notes and popl template
comparison update

--- 2025-06-12 LICS_2025/e_graphs_with_bindings/sections/introduction.tex
@@
 A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
-The auxiliary nature of these nodes is revealed by the fact that, after rewriting, they must be removed from the graph. 
+The auxiliary nature of these nodes is revealed by the fact that, after rewriting, they may be removed from the graph. 
 We will see how in the categorical representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.
@@
 In Fig.~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
-A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box), to be rewritten using the rules for explicit substitution at the top of each column.
+A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box).
+The substitution is then propagated through child e-classes using additional rewrite rules (depicted at the top of each column).
 At the end, all explicit substitution nodes can be removed.
@@
 Using de Bruijn indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
-Notably, a work of~\cite{slotted-egraphs} has emerged recently which introduced a formulation of e-graphs with native support for binding, \textit{i.e.,} where binding is a first-class operation along with a substitution.
-It significantly complicates the formulation of an e-graph and corresponding algorithmics and based on \textit{nominal} techniques~\cite{Pitts_2013} unlike the approach presented in this work.
-We believe the approach used in this work strikes with its simplicity although a more detailed comparison with the work of remains a future work.
+A recent work of~\cite{slotted-egraphs} on so-called \textit{slotted} e-graphs has addressed the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by making variables into first-class citizens of an e-graph and made binding a built-in operation for an e-graph.
+An example of a slotted e-graph that represents a term $\lambda x . (a + b) + (x + c)$ can be seen in Figure~\ref{fig:slotted}(left).
+
+\begin{figure}[t!]
+	\begin{subfigure}[t]{0.45\linewidth}
+\[
+\adjustbox{scale=0.8}{
+\tikzfig{figures/slotted-e-graph-example}
+}
+\]
+	\end{subfigure}
+	\hfill
+\begin{subfigure}[t]{0.45\linewidth}
+\[
+\adjustbox{scale=0.6}{
+\tikzfig{figures/slotted-string-2}
+}
+\]
+
+\end{subfigure}
+	\caption{}
+	\label{fig:slotted}
+\end{figure}
+
+A notable feature of a slotted e-graph is that every e-class is parameterised by a set of free variables of the terms represented by this e-class.
+The references then additionally convey the information about how variables are instantiated.
+For example, an e-class containing $+$ that is parameterised by variable names $s_2$ and $s_3$ gets intantiated once with $a,b$ (via $s_4 \mapsto a, s_2 \mapsto s_4$ and $s_5 \mapsto b, s_3 \mapsto s_5$) and once  with $x,c$.
+The fact that the variable $x$ is bound is evident from the corresponding e-class being parameterised by only three variables.
+If we were to add another e-node such as $e + d$, there would be a reference added to an already existing e-class such that $s_1 \mapsto e$, $s_2 \mapsto d$.
+We can achieve the same degree of sharing (in the sense we have the same number of nodes for $+$) using string diagrams by doing a closure conversion of $+$ (Figure~\ref{fig:slotted}(right)).
+However, the addition of these parameters and the corresponding alternations in the algorithmics makes the deeper correspondence between string diagrams and slotted e-graphs obscure.
+Despite this, there are applications where the use of string diagrams may be more beneficial than that of (slotted) e-graphs, for example in the case of lambda calculus with explicit substitution that we will describe next.
+Also, the categorical model that we built can help to explain the congruences maintained by slotted e-graphs which is further elaborated on in Section~\ref{sec:categorical}.
+
+% Notably, a work of~\cite{slotted-egraphs} has emerged recently which introduced a formulation of e-graphs with native support for binding, \textit{i.e.,} where binding is a first-class operation along with a substitution.
+% It significantly complicates the formulation of an e-graph and corresponding algorithmics and based on \textit{nominal} techniques~\cite{Pitts_2013} unlike the approach presented in this work.
+% We believe the approach used in this work strikes with its simplicity although a more detailed comparison with the work of remains a future work.
+
+\subsection{Lambda calculus with explicit substitutions and graphical equivalence of terms}
 

### 2025-07-08 POPL2026/paper/sections/applications.tex
wip

\section{Application}

In this section we will elaborate on how using e-hypergraphs for lambda calculus with explicit substitution may be more beneficial than (slotted) e-graphs.

While the work of~\cite{slotted-egraphs} has made the variables and binders into first class citizens of an e-graph, substitution and $\beta$-reduction in the context of $\lambda$-calculus remain challenging.
The issue with substitution is that despite it acting on a small part of the term, it results in duplication of large portions of an e-graph which is practically undesirable.
While it is possible to define substitution as a built-in operation in a slotted e-graph, in practice, using explicit substitution is more appropriate as it allows for a finer grained control over the duplication.
Even more so, explicit substitution becomes mandatory if we want to encode lambda calculus with explicit substitution in an e-graph.
Below is the Table~\ref{tbl:slotted} that features the number of e-nodes when performing equality saturation for a $\beta$-reduction using a lambda term $plus 2 2$ using Church encoding of arithmetic.

\begin{table}
    \begin{tabular}{lcccc}
      Set of equations & built-in subst & explicit subst & explicit subst & explicit subst\\
      &&& + interchange&  + interchange\\
      &&&&  + comm\\
      \# of e-nodes & 40 & 161 & 289 & 356
    \end{tabular}
    \caption{Slotted e-graphs benchmarks}
    \label{tbl:slotted}
\end{table}

The first row shows which rewrite rules were enabled, and the second row features the number of e-nodes after the e-graph is saturated (the initial term contained 16 e-nodes).
The first column used the built-in implementation of substitution that does not use the explicit substitution.
As it could be seen from the table, saturating the e-graph with respect to explicit substitution equations introduces a huge overhead (as the number of e-nodes that actually contain the result of a substitution is 40 as per the second column).
These equations, however, are absorbed by string diagrammatic (e-hypergraph) representation. Therefore, it is theoretically possible to achieve the size efficiency of built-in substitution without sacrificing any equations, i.e. providing complete information of all the redexes contained in a lambda term with explicit substitutions.

\textcolor{red}{Put the string diagrammatic encoding of substitution here?}

### 2025-07-08 POPL2026/paper/sections/categorical_semantics.tex
comparison update
wip

@@
 \section{Categorical semantics of E-Graphs with Bindings}
+\label{sec:categorical}
 
@@
 
+\begin{remark}
+
+The above distributivity law closely correspond to one of the congruences maintained by a slotted e-graph, namely
+
+\[
+\text{bind}\;s_1\; x \cong \text{bind}\;s_2\;y \;\; \text{if}\; x \cong_{\alpha} y
+\]
+The congruence intuitively says that if $x$ and $y$ are congruent (up to $\alpha$-equivalence), then the enclosing binders are also congruent.
+While this is a \textit{natural} congruence to define when constructing an e-graph with support for binders, in our framework it follows from the free enrichment by a purely categorical argument justifying that it is the \textit{right} congruence to define.
+\end{remark}
+
 Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of closed $\Sigma^{+}$-terms, the latter being defined as follows.

### 2025-07-08 POPL2026/paper/sections/introduction.tex
wip

@@
 A recent work of~\cite{slotted-egraphs} on so-called \textit{slotted} e-graphs has addressed the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by making variables into first-class citizens of an e-graph and made binding a built-in operation for an e-graph.
+They also solved the problem of explicit variable names guaranteeing freshness inside binders.
 An example of a slotted e-graph that represents a term $\lambda x . (a + b) + (x + c)$ can be seen in Figure~\ref{fig:slotted}(left).
@@
 However, the addition of these parameters and the corresponding alternations in the algorithmics makes the deeper correspondence between string diagrams and slotted e-graphs obscure.
-Despite this, there are applications where the use of string diagrams may be more beneficial than that of (slotted) e-graphs, for example in the case of lambda calculus with explicit substitution that we will describe next.
+Despite this, there are applications where the use of string diagrams may be more beneficial than that of (slotted) e-graphs, for example in the case of lambda calculus with explicit substitution that we will describe next \textcolor{red}{Or in Applications section?}.
 Also, the categorical model that we built can help to explain the congruences maintained by slotted e-graphs which is further elaborated on in Section~\ref{sec:categorical}.

### 2025-07-08 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
notes and popl template
start section 7. lambda calculus w/ explicit substitution

--- 2025-05-19 LICS_2025/e_graphs_with_bindings/E-graphs-with-bindings.tex
@@
-\documentclass[]{IEEEtran}
-% \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-%Template version as of 6/27/2024
+\documentclass[acmsmall,screen, nonacm, review, anonymous, authoryear]{acmart}
 
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{amsmath,amsfonts}
 \usepackage{algorithmic}
@@
 \usepackage{fontenc}
+\usepackage{csquotes}
 
@@
 
-\usepackage[final]{microtype}
+\usepackage{microtype}
 
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\title{Categorical Foundations of E-Graphs With Binding}
+
 \begin{document}
-
-\title{E-Graphs With Bindings
-}
-
-% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
-% \and
-% }
-
-\author{Anonymous}
-
-\maketitle
 
@@
 
+\maketitle
+
 \input{sections/introduction}
@@
 \input{sections/soundness}
+\input{sections/lambda-calculus}
+
+\section{Conclusion}
+In this work we introduced e-graphs with bindings formalised as morphisms of a closed $\catname{SLat}$-SMC.
+We then provided a combinatorial representation for such morphisms in terms of hierarchical hypergraphs (e-hypergraphs) with a suitable notion of sound and complete DPO rewriting, making it possible for an e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
+An important future work is exploring whether such combinatorial structures may lead to better algorithmics for doing equality saturation for $\lambda$-calculus.
 
@@
 
-\input{sections/appendix}
+% \input{sections/appendix}
 

### 2025-07-08 POPL2026/paper/sections/lambda-calculus.tex
start section 7. lambda calculus w/ explicit substitution

\section{Application --- $\lambda^\sim_{\text{lsub}}$-calculus}

In this section, we show the utility of our theory by demonstrating how it can accommodate the \emph{linear substitution calculus} $\lambda_{\text{lsub}}$ of~\citet[§ 2]{accattoli2014nonstandard}.
We summarise the linear substitution calculus as the $\lambda$-calculus with an additional term former representing explicit substitution; i.e.\ terms are generated by the grammar
\[
	t \coloneqq x \mid t t \mid \lambda x. t \mid t [x/t],
\]
where $x$ is a variable drawn from some countably infinite set, $t u$ is an application, and $\lambda x. t$ is an abstraction, as in standard $\lambda$-calculus.
The term $t [x / u]$ is an explicit substitution of the variable $x$ for the term $u$ in the term $t$.

The notions of context $C[-]$, bound and free variable are as usual, with the addition that $\textsf{fv} (t [x/u]) \coloneq \textsf{fv} (t) \setminus \{ x \} \cup \textsf{fv} (u)$.
Write $C \llbracket t \rrbracket$ for the term given by substitution of the hole in the context $C[-]$ for the term $t$, under the additional restriction that the free variables of $t$ are not captured (via abstraction or explicit substitution) by $C$.
In addition to the $\beta$- and $\eta$-reduction rules of standard $\lambda$-calculus, there are reduction rules to account for this explicit substitution which \enquote{perform} the substitution or commute it with $\beta$-reduction:
\begin{align*}
	(\lambda x. t) L u              \quad & \leadsto \quad t[x/u] L,                                                    \\
	C \llbracket x \rrbracket [x/t] \quad & \leadsto \quad C \llbracket u \rrbracket [x/t],                             \\
	t [x/u]                         \quad & \leadsto \quad t,                               & x \notin \textsf{fv} (t),
\end{align*}
where $L$ is a finite list of substitutions of the form $[x/u] [y/v] \ldots$.

In programming terms, this explicit substitution can be used to model \textbf{let}-bindings and sharing.
For instance, the ML-style term $\textbf{let} \; x = 1 \; \textbf{in} \; \text{plus} (x, x)$ is formally represented as the term $\text{plus} (x, x) [x / 1]$ in this calculus\footnote{Technically speaking, in this case we mean the extension of this calculus with flat natural numbers equipped with a binary operation $\text{plus}$.}.
In this case, the variable $x$ is valued in the constant $1$, but one could imagine instead that $x$ is to be substituted for the result of some expensive computation which we do not desire to repeat twice unnecessarily --- in this sense, the sharing of both operands of the operation $\text{plus}$ is captured by the explicit substitution.

\emph{Graphical equivalence} $\sim$ of $\lambda_{\text{lsub}}$-terms is augmentation of $\alpha$-equivalence of terms with the additional stipulations that
\begin{align*}
	t [x/u] [y/v] \quad        & \sim \quad t [y/v] [x/u],        & x \notin \textsf{fv} (v) \land y \notin \textsf{fv} (u), \\
	(\lambda y. t) [x/u] \quad & \sim \quad \lambda y. (t [x/u]), & y \notin \textsf{fv} (u),                                \\
	(t v) [x/u] \quad          & \sim \quad t [x/u] v,            & x \notin \textsf{fv} (u),
\end{align*}
closed under contexts, transitivity, symmetry, and reflexivity.
Put simply, it is the stipulation that \textbf{let}-bindings may be freely reordered without changing the meaning of the term\footnote{Recall that our setting is \emph{pure}, in the sense that no sequencing of effects is meaningful.}.
The $\lambda^\sim_{\text{lsub}}$-calculus is given by these graphical equivalence classes of terms with the reduction relation defined by that of the $\lambda_{\text{lsub}}$-calculus modulo graphical equivalence.

This calculus is explicitly designed to capture the representation of terms as proof nets: for two terms $t, u$ in the $\lambda_{\text{lsub}}$-calculus, $t \sim u$ if and only if both map to the same proof net~\cite{?}.
It turns out that the same fact is true for the representation of terms as string diagrams~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} instead of proof terms, and in this fact we can exploit the string diagram setting of e-hypergraphs to absorb the reductions arising due to graphical equivalence into our representation rather than explicitly providing them as equations attached to a symmetric monoidal theory presenting the $\lambda^\sim_{\text{lsub}}$-calculus.

Before we proceed to demonstrate this, we first wish to discuss the relevance of this calculus in the domain of e-graphs.
E-graphs excel as a toolkit for compiler optimisation, facilitating the rewriting of programs into more efficient forms by equality saturation.
The theoretical study of this is the field of term rewriting, where programs are often simplistically represented as (the abstract syntax of) terms in the $\lambda$-calculus.
However, in a modern optimising compiler (e.g.\ LLVM~\cite{lattnerLLVMCompilationFramework2004}, GNU Compiler Collection~\cite{merrillGenericGimpleNew2003}), programs are represented as some sort of graph-based intermediate representation (IR) rather than a syntax tree.
Here, basic blocks of instructions\footnote{Maximal sequences of instructions which are unconditionally executed together determined by program control flow.} form the nodes, and terminator instructions (e.g.\ jumps, branches, and other control flow operations) form the edges.
This allows for optimisations and analyses based on graphical structure, e.g.\ reachability analysis to perform dead-code elimination, or data-flow analysis to perform constant folding.
In essence, this is much closer to $\lambda^\sim_{\text{lsub}}$-calculus: the graphical equivalence there is abstracting away from the particular syntax of a term, in much the same way that LLVM IR abstracts away from the raw syntax of the source program.
Thus, an understanding of e-graphs in this context is valuable for the applicability of equality saturation techniques with respect to graph-based IRs which dominate the landscape of modern compiler design.

We will define a monoidal signature whose closed symmetric monoidal theory is the $\lambda^\sim_{\text{lsub}}$-calculus extended by tupling of terms and flat natural numbers equipped with basic arithmetic operations.
Let $\Sigma \coloneq (\{ \mathbf{N} \}, \{ \comonoid, \counit \} \cup \mathbb{N} \cup \{ \text{plus}, \text{mult} \}, t)$ where
\begin{align*}
	t\colon \{ \comonoid, \counit \} \cup \mathbb{N} \cup \{ \text{plus}, \text{mult} \} & \to {\{ \mathbf{N} \}}^* \times {\{ \mathbf{N} \}}^*                                                    \\
	t(n)                                                                                 & = ((), (\mathbf{N})),                                & n \in \mathbb{N},                                \\
	t(o)                                                                                 & = ((\mathbf{N}, \mathbf{N}), (\mathbf{N})),          & o \in \{ \comonoid, \text{plus}, \text{mult} \}, \\
	t(\counit)                                                                           & = ((\mathbf{N}), ()).
\end{align*}
The operations that the signature $\Sigma$ avails are copy and delete, denoted by $\comonoid$ and $\counit$ respectively, constants for each natural number, and the binary arithmetic operations $+$ and $\times$.
We interpret the $\Sigma$-terms $\llbracket \tau \rrbracket$ as follows: fully applied operations as themselves, $\id_{I}$ and $\id_{\mathbf{N}}$ as the identity function $\lambda x. x$ on the empty tuple $()$ and $\mathbf{N}$ respectively, $\sym_{\mathbf{N}, \mathbf{N}}$ as the swap function $\lambda (x, y). (y, x)$, composition $\tau_1 ; \tau_2$ as the composition function $\lambda x. \llbracket \tau_2 \rrbracket \left(\llbracket \tau_1 \rrbracket x\right)$, and $\tau_1 \otimes \tau_2$ as tupling $\left(\llbracket \tau_1 \rrbracket, \llbracket \tau_2 \rrbracket\right)$.

Note that, in contrast to traditional presentations of the $\lambda$-calculus as a language of (slotted~\cite[Example 1]{slotted-egraphs}) e-graphs~\cite[Figure 10]{EggPaper}, here there is no encoding of the term formers for application, abstraction, and explicit substitution.
Instead, these constructs are handled ambiently by our semantic setting: any closed symmetric monoidal category (and in particular, $\catname{CS}(\Sigma, \mathcal{E})$ for any set of equations $\mathcal{E}$) admits an internal language in the multiplicative fragment of intuitionistic linear logic (a particular type of linear $\lambda$-calculus)~\cite[§ 1.7.3]{abramskyIntroductionCategoriesCategorical2010}.
In other words, with respect to~\autoref{fig:egraph-strings}, this allows us to \emph{natively} represent abstraction by the rounded boxes of ($\Lambda(f)$), and application by the $\textsf{ev}$ operation with $\beta$- and $\eta$-conversion arising via equality of morphisms in the category (that is, without having to provide them as data in the form of equations in $\mathcal{E}$).
For explicit substitution, we note that the linearity is not desirable; this is why the copy-delete comonoid $(\comonoid, \counit)$ is present: by Fox's theorem~\cite{foxCoalgebrasCartesianCategories1976}, a cartesian category is precisely a monoidal one admitting such a \emph{natural} copy-delete structure.
Thus, we must endow $\mathcal{E}$ with equations stating that $(\comonoid, \counit)$ forms a comonoid, and moreover that it is natural: $\comonoid$ behaves uniformly as a copy map, and $\counit$ behaves as deletion.

### 2025-07-08 POPL2026/paper/sections/soundness.tex
start section 7. lambda calculus w/ explicit substitution

@@
 \end{figure}
-
-\section{Conclusion}
-In this work we introduced e-graphs with bindings formalised as morphisms of a closed $\catname{SLat}$-SMC.
-We then provided a combinatorial representation for such morphisms in terms of hierarchical hypergraphs (e-hypergraphs) with a suitable notion of sound and complete DPO rewriting, making it possible for an e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
-An important future work is exploring whether such combinatorial structures may lead to better algorithmics for doing equality saturation for $\lambda$-calculus.

### 2025-07-09 POPL2026/paper/sections/lambda-calculus.tex
mostly finish lambda calculus section

@@
 	(\lambda y. t) [x/u] \quad & \sim \quad \lambda y. (t [x/u]), & y \notin \textsf{fv} (u),                                \\
-	(t v) [x/u] \quad          & \sim \quad t [x/u] v,            & x \notin \textsf{fv} (u),
+	(t v) [x/u] \quad          & \sim \quad t [x/u] v,            & x \notin \textsf{fv} (v),
 \end{align*}
@@
 This calculus is explicitly designed to capture the representation of terms as proof nets: for two terms $t, u$ in the $\lambda_{\text{lsub}}$-calculus, $t \sim u$ if and only if both map to the same proof net~\cite{?}.
-It turns out that the same fact is true for the representation of terms as string diagrams~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} instead of proof terms, and in this fact we can exploit the string diagram setting of e-hypergraphs to absorb the reductions arising due to graphical equivalence into our representation rather than explicitly providing them as equations attached to a symmetric monoidal theory presenting the $\lambda^\sim_{\text{lsub}}$-calculus.
+It turns out that the same fact is true for the representation of terms as string diagrams~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} instead of proof nets, and in this fact we can exploit the string diagram setting of e-hypergraphs to absorb the reductions arising due to graphical equivalence into our representation rather than explicitly providing them as equations attached to a closed symmetric monoidal theory presenting the $\lambda^\sim_{\text{lsub}}$-calculus.
 
@@
 The operations that the signature $\Sigma$ avails are copy and delete, denoted by $\comonoid$ and $\counit$ respectively, constants for each natural number, and the binary arithmetic operations $+$ and $\times$.
-We interpret the $\Sigma$-terms $\llbracket \tau \rrbracket$ as follows: fully applied operations as themselves, $\id_{I}$ and $\id_{\mathbf{N}}$ as the identity function $\lambda x.
+We interpret the $\Sigma$-terms $\llbracket \tau \rrbracket$ as follows: operations as themselves, $\id_{I}$ and $\id_{\mathbf{N}}$ as the identity function $\lambda x.
 x$ on the empty tuple $()$ and $\mathbf{N}$ respectively, $\sym_{\mathbf{N}, \mathbf{N}}$ as the swap function $\lambda (x, y).
@@
 Thus, we must endow $\mathcal{E}$ with equations stating that $(\comonoid, \counit)$ forms a comonoid, and moreover that it is natural: $\comonoid$ behaves uniformly as a copy map, and $\counit$ behaves as deletion.
+
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.3\linewidth}
+		\begin{lstlisting}[language=ML]
+let x = 1 in
+let y = 2 in
+plus(x, y)
+                \end{lstlisting}
+		\hrule
+		\begin{lstlisting}[language=ML]
+let y = 2 in
+let x = 1 in
+plus(x, y)
+                \end{lstlisting}
+		\hrule
+		\begin{lstlisting}[language=ML]
+plus(1, 2)
+                \end{lstlisting}
+		\caption{Three programs for $1 + 2$}%
+		\label{fig:plus-1-2-programs}
+	\end{subfigure}
+	\hspace{2em}
+	\begin{subfigure}{0.2\linewidth}
+		\begin{align*}
+			     & \text{plus}(x, y)[y/2][x/1] \\
+			\sim &                             \\
+			     & \text{plus}(x, y)[x/1][y/2] \\
+			\sim &                             \\
+			     & \text{plus}(1, 2)
+		\end{align*}
+		\caption{As terms in $\lambda_{\text{lsub}}$}%
+		\label{fig:plus-1-2-terms}
+	\end{subfigure}
+	\hspace{2em}
+	\begin{subfigure}{0.2\linewidth}
+		\centering
+		\tikzfig{figures/plus-1-2}
+		\caption{String diagrammatic representation}%
+		\label{fig:plus-1-2-string-diagram}
+	\end{subfigure}
+	\caption{Graphically equivalent terms}%
+	\label{fig:plus-1-2}
+\end{figure}
+
+\autoref{fig:plus-1-2-programs} shows some programs which are graphically and semantically equivalent but syntactically distinct.
+As alluded to in~\autoref{sec:vs-e-graphs-with-binding}, the traditional (slotted) e-graph approach to modelling the graphical equivalence of these terms (\autoref{fig:plus-1-2-terms}) relies on explicitly encoding substitution operations as e-graph nodes, and their behaviour (the graphical equivalence relation $\sim$) as rewrite rules, as in~\autoref{fig:e-graph-substitution} --- not only is this more complicated, but it also leads to a blow-up in the state space required to perform equality saturation in any extension of the $\lambda^\sim_{\text{lsub}}$-calculus, because we are really rewriting $\lambda_{\text{lsub}}$-terms and computational resources are spent rewriting between graphically equivalent terms.
+\textcolor{red}{Aleksei, can you say something quantitative here if you ran some tests?}
+In contrast, in our framework, $\lambda^\sim_{\text{lsub}}$-terms are directly modelled as string diagrams (\autoref{fig:plus-1-2-string-diagram}\footnote{Note that dashed boxes are elided for convenience here because no rewriting has occurred.}).
+We do however note that the traditional approach is more flexible, in the sense that it may model more theories with binders (e.g.\ $\pi$-calculus), whereas our approach is essentially specialised to $\lambda$-calculus, relying heavily on the metatheoretic internal language of the ambient setting.
+
+As a second example, we may wish to incorporate the optimisation ${\text{plus} (x, x) \leadsto \text{mult} (x, 2)}$ in our toy theory.
+This corresponds to endowing $\mathcal{E}$ with the equation of string diagrams:
+\[
+	\tikzfig{figures/plus-copy}
+	\quad
+	=
+	\quad
+	\tikzfig{figures/mult-2}
+	\;
+	.
+\]
+Now, in our semantic model, this rewrite can be non-destructively applied to model one step of e-graph saturation:
+\[
+	\tikzfig{figures/plus-copy-rewrite-1}
+	\quad
+	\equiv
+	\quad
+	\tikzfig{figures/plus-copy-rewrite-2}
+	\quad
+	\to
+	\quad
+	\tikzfig{figures/plus-copy-rewrite-3}
+	\;
+	.
+\]
+Here, $\equiv$ is a definitional equality arising from the idempotence in the Hom semilattice, and $\to$ is an application of DPO rewriting using the previous equation along the right branch.
+In isolation, this isn't particularly interesting, but the important observation is that this rewriting all happens \emph{locally} in the context of a bigger diagram.
+
+From a practical perspective, the encoding of string diagrams as (e-)hypergraphs provides a normal form for graphical equivalence classes of terms.
+As a data structure, this provides an efficient representation upon which to build an implementation of this theory as a fully fledged \enquote{e-graphs for string diagram rewriting}, using DPO rewriting as a primitive to implement the standard e-graph operations.
+We leave this as future work, noting its applications to not only $\lambda^\sim_{\text{lsub}}$-calculus, but also other monoidal theories where there are two levels of rewriting:
+\begin{enumerate}
+	\item \enquote{bureaucratic} reassociation and reparenthesisation, absorbed into isotopy of string diagrams (here, graphical equivalence of $\lambda_{\text{lsub}}$-terms);
+	\item algebraic identities arising from the semantics of the theory (e.g. ${\text{plus} (x, x) \leadsto \text{mult} (x, 2)}$ for arithmetic).
+\end{enumerate}
+One such theory is the ZX calculus~\cite{coeckeInteractingQuantumObservables2011}.

### 2025-07-09 POPL2026/notes/SlottedNotes.tex
comparison, semantics, enrichment fix, appendix

@@
 \begin{table}
-\begin{tabular}{lcccc}
-  Set of equations & built-in subst & explicit subst & explicit subst & explicit subst\\
-  &&& + interchange&  + interchange\\
-  &&&&  + comm\\
-  \# of e-nodes & 40 & 161 & 289 & 356
+\begin{tabular}{lccccc}
+  Set of equations & built-in subst & explicit subst & explicit subst & explicit subst & explicit subst\\
+  &&& + comm& + interchange&  + interchange\\
+  &&&&&  + comm\\
+  \# of e-nodes & 40 & 161 & 288 & 289 & 356
 \end{tabular}
@@
 
+\begin{remark}
+
+The rules arise from linear substitution lambda calculus which is as follows
+
+\begin{align*}
+(\lambda\;x\;t)\;L\;u \mapsto_{db} t[x / u]L\\
+C\llbracket x \rrbracket [x / u] \mapsto_{ls} C \llbracket u \rrbracket [x / u]\\
+t[x/u] \mapsto t, \text{if } x \not \in fv(t)
+\end{align*}
+
+The contexts $C$ are defined as
+
+\[
+C \Coloneqq \square \;|\; \lambda x . C \; | \; C\;t \;|\; t\;C\;|\;C[x / t]\;|\;t[x / C]
+\]
+
+and in particular in $\lambda x . C$ and $C[x/t]$ $x$ is bound.
+
+\end{remark}
+
 \bibliographystyle{acm}

### 2025-07-09 POPL2026/paper/sections/categorical_semantics.tex
comparison, semantics, enrichment fix, appendix

--- 2025-07-07 POPL2026/paper/sections/categorical_semantics.tex
@@
 
-    \begin{proposition}(A specialised case of Proposition 6.4.7~\cite{Borceux_1994})
+    \begin{proposition}(A specialised case of Proposition 6.4.7~\cite{Borceux_1994} and Theorem 5.7.1~\cite{cruttwell2008normed})
         There is a 2-adjunction 
@@
         \]
-        that is induced by the usual free-forgetful adjunction
+        that is induced by the usual (symmetric monoidal) free-forgetful adjunction
 
@@
         \[
-        F \dashv U : \catname{SLat} \to \catname{Set}
+        F \dashv U : \catname{SLat} \to \catname{Set}.
         \vspace{-1mm}
         \]
+        Moreover, every symmetric monoidal category $\mathbb{C}$ in $\catname{Cat}$ (respectively, $\catname{SLat}\text{--}\catname{Cat}$) is taken to a symmetric monoidal category $\mathbb{C}'$ in $\catname{SLat}\text{--}\catname{Cat}$ (respectively, $\catname{Cat}$) by a 2-functor $\mathcal{F}$ (respectively, $\mathcal{U}$).
     \end{proposition}
 In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice enriched $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
-Being a 2-functor, it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{Slat}$-adjunction.
-Thus, applying the functor to $\catname{CS}(\Sigma)$ we get a closed symmetric monoidal $\catname{Slat}$-category $\catname{CS}(\Sigma)^{+}$ by lifting the adjunction $(- \otimes B, B \multimap{} -, \eta, \varepsilon)$.
+Being a 2-functor, it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{SLat}$-adjunction.
+Thus, applying the 2-functor to $\catname{CS}(\Sigma)$ we get a closed symmetric monoidal $\catname{SLat}$-category $\catname{CS}(\Sigma)^{+}$ by lifting the adjunction $(- \otimes B, B \multimap{} -, \eta, \varepsilon)$.
 We can similarly derive $\catname{SLat}$-enriched versions of  $\Lambda_{A,B,C}$ and $\textbf{ev}_{A,B}$ morphisms using the lifted adjunction as $\textbf{ev}_{A,B} = \varepsilon_{A}$ and $\Lambda_{A,B,C}(f) = \eta; (B \multimap{} f)$.
@@
 \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
-Note the orientation of string diagrams are now from top to bottom as opposed to what was used in~\nameref{sec:introduction}: the latter was mimicking the orientation of e-graphs, \textit{i.e.}, traditional orientation for expression trees.
+
+% rotated the diagrams
+% Note the orientation of string diagrams are now from top to bottom as opposed to what was used in~\nameref{sec:introduction}: the latter was mimicking the orientation of e-graphs, \textit{i.e.}, traditional orientation for expression trees.
+
 % Fig. \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
@@
 \adjustbox{scale=0.65}{
-\tikzfig{./figures/egraph-strings}
+\tikzfig{./figures/egraph-strings-2}
 }
@@
 \end{figure*}
+
+\subsection{Categorical foundations for slotted e-graphs}
+
+Slotted e-graphs~\cite{slotted-egraphs} extended the traditional signature employed by e-graphs by an explicit construct for binding ($\text{bind}\; \$x\; tc$) and variables (slots).
+\begin{align*}
+&\text{functional symbols}\; f, g, \ldots\\
+&\text{slots}\;  \$x, \$y, \ldots\\
+&\text{term children}\; tc  \Coloneqq t \;|\; tc_{1}, \ldots, tc_{n} \;|\; \text{bind}\; \$x\; tc \;|\; \$x\\ 
+&\text{terms}\;  \Coloneqq f \;|\; f(tc)
+\end{align*}
+
+The traditional congruence relation maintained be e-graphs for a set of equations $E$ that are given below (plus obvious rules for transitivity, symmetry and reflexivity of $\cong$)
+
+\begin{mathpar}
+\inferrule*[right=start]
+    {a = b \in E}
+    {E \vdash a \cong b}
+    \inferrule*[right=cong. (variadic)]
+      {E \vdash tc_i \cong tc'_i,\; i = 1, \ldots, k}
+      {E \vdash tc_1, \ldots, tc_k \cong tc'_1, \ldots, tc'_k}
+    \inferrule*[right=cong. (f)]
+    {E \vdash tc \cong tc'}
+    {E \vdash f(tc_1) \cong f(tc')}
+\end{mathpar}
+
+\noindent
+are extended by the rule
+
+\begin{mathpar}
+
+\inferrule*[right=cong. (bind)]
+    {E \vdash tc \cong tc'}
+    {E \vdash \text{bind}\; \$x\; tc \cong \text{bind}\; \$x\; tc'},
+\end{mathpar}
+
+\noindent
+plus the rules for equivalence of terms up to renaming of slots.
+Recalling that we use $+$ to encode this congruence of terms, the first three rules (plus reflexivity, symmetry, and transitivity) correspond to idempotence, commutativity and associativity of $+$ and the fact that $+$ respect the composition $(;)$.
+%  and tensoring $\otimes$ (traditional e-graphs do not have tensoring as they are defined for algebraic signatures).
+While the other rule is \textit{natural} to enforce when developing e-graphs with binding, we see that it is indeed the \textit{right} rule to enforce as we can see it follow by a purely categorical argument by the free enrichment as per the law \ref{law:distributivity}.
+We do not enforce any rules for renaming as string diagrams are essentially a nameless representation of terms.
+Moreover, some renaming rules are not consistent with string diagrams, as slotted e-graph try to represent, \textit{e.g.}, terms $(a + b)$ and $(c + d)$ as a single entity (up to renaming of free variables), which is not the case for string diagrams (unless we do closure conversion).
+
+% The above distributivity law closely correspond to one of the congruences maintained by a slotted e-graph, namely
+
+% \[
+% \text{bind}\;s_1\; x \cong \text{bind}\;s_2\;y \;\; \text{if}\; x \cong_{\alpha} y
+% \]
+% The congruence intuitively says that if $x$ and $y$ are congruent (up to $\alpha$-equivalence), then the enclosing binders are also congruent.
+% While this is a \textit{natural} congruence to define when constructing an e-graph with support for binders, in our framework it follows from the free enrichment by a purely categorical argument justifying that it is the \textit{right} congruence to define.

### 2025-07-09 POPL2026/paper/sections/introduction.tex
mostly finish lambda calculus section
comparison, semantics, enrichment fix, appendix

--- 2025-07-08 POPL2026/paper/sections/introduction.tex
@@
 
-\subsection{Conventional vs. categorical e-graphs with binding}
+\subsection{Conventional vs.\ categorical e-graphs with binding}%
+\label{sec:vs-e-graphs-with-binding}
 
@@
 A recent work of~\cite{slotted-egraphs} on so-called \textit{slotted} e-graphs has addressed the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by making variables into first-class citizens of an e-graph and made binding a built-in operation for an e-graph.
-An example of a slotted e-graph that represents a term $\lambda x . (a + b) + (x + c)$ can be seen in Figure~\ref{fig:slotted}(left).
+They also solved the problem of explicit variable names guaranteeing freshness inside binders explicitly differentiating between bound and free variables.
+An example of a slotted e-graph that represents a term $\lambda x . (a + b) + (x + c)$ can be seen in Figure~\ref{fig:slotted} (left).
 
@@
 If we were to add another e-node such as $e + d$, there would be a reference added to an already existing e-class such that $s_1 \mapsto e$, $s_2 \mapsto d$.
-We can achieve the same degree of sharing (in the sense we have the same number of nodes for $+$) using string diagrams by doing a closure conversion of $+$ (Figure~\ref{fig:slotted}(right)).
-However, the addition of these parameters and the corresponding alternations in the algorithmics makes the deeper correspondence between string diagrams and slotted e-graphs obscure.
-Despite this, there are applications where the use of string diagrams may be more beneficial than that of (slotted) e-graphs, for example in the case of lambda calculus with explicit substitution that we will describe next.
-Also, the categorical model that we built can help to explain the congruences maintained by slotted e-graphs which is further elaborated on in Section~\ref{sec:categorical}.
+We can achieve the same degree of sharing (in the sense that we have the same number of nodes for $+$) using string diagrams by performing a closure conversion of $+$ (Figure~\ref{fig:slotted} (right)).
+However, the addition of these parameters and the corresponding alterations in the algorithmics makes the deeper correspondence between string diagrams and slotted e-graphs obscure.
+In fact, slotted e-graph appears to be a more general construction in a sense that it supports equality saturation for any signature with binders, such as, for example, the $\pi$-calculus, while our approach is tailored towards structures that arise from (Cartesian) closed symmetric monoidal categories.
+Therefore, the use of string diagrams may be more advantageous than that of (slotted) e-graphs in the case of the latter.
+We defer a case study to Section~\ref{sec:application}.
+Finally, the categorical model that we develop can help justify the congruences maintained by slotted e-graphs which is further elaborated on in Section~\ref{sec:categorical}.
 
@@
 
-\subsection{Lambda calculus with explicit substitutions and graphical equivalence of terms}
+% Do we need this section?
+% \subsection{Lambda calculus with explicit substitutions and graphical equivalence of terms}
 

### 2025-07-09 POPL2026/paper/sections/lambda-calculus.tex
comparison, semantics, enrichment fix, appendix
finish up lambda calculus section

@@
-\section{Application --- $\lambda^\sim_{\text{lsub}}$-calculus}
+\section{Application --- $\lambda^\approx_{\text{lsub}}$-calculus}
+\label{sec:application}
 
@@
 \begin{align*}
-	(\lambda x. t) L u              \quad & \leadsto \quad t[x/u] L,                                                    \\
-	C \llbracket x \rrbracket [x/t] \quad & \leadsto \quad C \llbracket u \rrbracket [x/t],                             \\
-	t [x/u]                         \quad & \leadsto \quad t,                               & x \notin \textsf{fv} (t),
+	(\lambda x. t) L u              \quad & \to_{d\beta} \quad t[x/u] L,                                                \\
+	C \llbracket x \rrbracket [x/t] \quad & \to_{ls} \quad C \llbracket u \rrbracket [x/t],                             \\
+	t [x/u]                         \quad & \to_{gc} \quad t,                               & x \notin \textsf{fv} (t),
 \end{align*}
@@
 \emph{Graphical equivalence} $\sim$ of $\lambda_{\text{lsub}}$-terms is augmentation of $\alpha$-equivalence of terms with the additional stipulations that
-\begin{align*}
-	t [x/u] [y/v] \quad        & \sim \quad t [y/v] [x/u],        & x \notin \textsf{fv} (v) \land y \notin \textsf{fv} (u), \\
-	(\lambda y. t) [x/u] \quad & \sim \quad \lambda y. (t [x/u]), & y \notin \textsf{fv} (u),                                \\
-	(t v) [x/u] \quad          & \sim \quad t [x/u] v,            & x \notin \textsf{fv} (v),
-\end{align*}
+\begin{align}
+	t [x/u] [y/v] \quad        & \sim \quad t [y/v] [x/u],        & x \notin \textsf{fv} (v) \land y \notin \textsf{fv} (u), \nonumber \\
+	\label{eq:subst-lambda}
+	(\lambda y. t) [x/u] \quad & \sim \quad \lambda y. (t [x/u]), & y \notin \textsf{fv} (u), \tag{$\ast$}                             \\
+	(t v) [x/u] \quad          & \sim \quad t [x/u] v,            & x \notin \textsf{fv} (v), \nonumber
+\end{align}
 closed under contexts, transitivity, symmetry, and reflexivity.
@@
 This calculus is explicitly designed to capture the representation of terms as proof nets: for two terms $t, u$ in the $\lambda_{\text{lsub}}$-calculus, $t \sim u$ if and only if both map to the same proof net~\cite{?}.
-It turns out that the same fact is true for the representation of terms as string diagrams~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} instead of proof nets, and in this fact we can exploit the string diagram setting of e-hypergraphs to absorb the reductions arising due to graphical equivalence into our representation rather than explicitly providing them as equations attached to a closed symmetric monoidal theory presenting the $\lambda^\sim_{\text{lsub}}$-calculus.
+By further extending the equivalence relation $\sim$, taking the symmetric closure by $\to_{ls}$ and $\to_{gc}$ and thus identifying terms under elimination of explicit substitution, we determine a \emph{string diagram equivalence} $\approx$ of $\lambda_{\text{lsub}}$-terms, determining a $\lambda^\approx_{\text{lsub}}$-calculus.
+Predictably, $t \approx u$ if and only if both map to the same string diagram~\cite[§ 5.1.1]{ghica2024stringdiagramslambdacalculifunctional}; this allows us to exploit the string diagram setting of e-hypergraphs to absorb many\footnote{The reason this is not \emph{all} reductions is because e-hypergraphs quotient away only the \emph{structural} isomorphisms of symmetric monoidal closed $\catname{SLat}$-categories (e.g.\ associators, unitors, etc.) which generate the \enquote{isotopy of string diagrams}; even in the graphical equivalence relation $\sim$, while~\autoref{eq:subst-lambda} is validated in every symmetric monoidal closed $\catname{SLat}$-category, it is not so by isotopy, but rather the adjunction $A \multimap - \dashv - \otimes A$.
+In other words, terms floating freely through the boundaries of rounded boxes is validated by our semantics, but produces concretely different e-hypergraphs --- this is the purpose of rectification by quotienting along $\mathcal{S}$ in~\autoref{prop:quotient-structural}, and manifests as additional DPO rewrites to be imposed.} of the reductions arising due to the equivalence $\approx$ into our representation rather than explicitly providing them as equations attached to a closed symmetric monoidal theory presenting the $\lambda^\approx_{\text{lsub}}$-calculus.
 
@@
 This allows for optimisations and analyses based on graphical structure, e.g.\ reachability analysis to perform dead-code elimination, or data-flow analysis to perform constant folding.
-In essence, this is much closer to $\lambda^\sim_{\text{lsub}}$-calculus: the graphical equivalence there is abstracting away from the particular syntax of a term, in much the same way that LLVM IR abstracts away from the raw syntax of the source program.
+In essence, this is much closer to $\lambda^\approx_{\text{lsub}}$-calculus: the string diagram equivalence there is abstracting away from the particular syntax of a term, in much the same way that LLVM IR abstracts away from the raw syntax of the source program.
 Thus, an understanding of e-graphs in this context is valuable for the applicability of equality saturation techniques with respect to graph-based IRs which dominate the landscape of modern compiler design.
 
-We will define a monoidal signature whose closed symmetric monoidal theory is the $\lambda^\sim_{\text{lsub}}$-calculus extended by tupling of terms and flat natural numbers equipped with basic arithmetic operations.
+We will define a monoidal signature whose closed symmetric monoidal theory is the $\lambda^\approx_{\text{lsub}}$-calculus extended by tupling of terms and flat natural numbers equipped with basic arithmetic operations.
 Let $\Sigma \coloneq (\{ \mathbf{N} \}, \{ \comonoid, \counit \} \cup \mathbb{N} \cup \{ \text{plus}, \text{mult} \}, t)$ where
@@
 			        & \text{plus}(x, y)[y/2][x/1] \\
-			\sim &                             \\
+			\approx &                             \\
 			        & \text{plus}(x, y)[x/1][y/2] \\
-			\sim &                             \\
+			\approx &                             \\
 			        & \text{plus}(1, 2)
@@
 	\end{subfigure}
-	\caption{Graphically equivalent terms}%
+	\caption{$\approx$-equivalent terms}%
 	\label{fig:plus-1-2}
@@
 
-\autoref{fig:plus-1-2-programs} shows some programs which are graphically and semantically equivalent but syntactically distinct.
-As alluded to in~\autoref{sec:vs-e-graphs-with-binding}, the traditional (slotted) e-graph approach to modelling the graphical equivalence of these terms (\autoref{fig:plus-1-2-terms}) relies on explicitly encoding substitution operations as e-graph nodes, and their behaviour (the graphical equivalence relation $\sim$) as rewrite rules, as in~\autoref{fig:e-graph-substitution} --- not only is this more complicated, but it also leads to a blow-up in the state space required to perform equality saturation in any extension of the $\lambda^\sim_{\text{lsub}}$-calculus, because we are really rewriting $\lambda_{\text{lsub}}$-terms and computational resources are spent rewriting between graphically equivalent terms.
-\textcolor{red}{Aleksei, can you say something quantitative here if you ran some tests?}
-In contrast, in our framework, $\lambda^\sim_{\text{lsub}}$-terms are directly modelled as string diagrams (\autoref{fig:plus-1-2-string-diagram}\footnote{Note that dashed boxes are elided for convenience here because no rewriting has occurred.}).
+\autoref{fig:plus-1-2-programs} shows some programs which are semantically equivalent but syntactically distinct.
+As alluded to in~\autoref{sec:vs-e-graphs-with-binding}, the traditional (slotted) e-graph approach to modelling the equivalence of these terms (\autoref{fig:plus-1-2-terms}) relies on explicitly encoding substitution operations as e-graph nodes, and their behaviour (the string diagram equivalence relation $\approx$) as rewrite rules, as in~\autoref{fig:e-graph-substitution} --- not only is this more complicated, but it also leads to a blow-up in the state space required to perform equality saturation in any extension of the $\lambda^\approx_{\text{lsub}}$-calculus, because we are really rewriting $\lambda_{\text{lsub}}$-terms and computational resources are spent rewriting between equivalent terms.
+\textcolor{cyan}{To illustrate this, let's consider the following equations induced by the graphical equivalence and the $\lambda_{\text{lsub}}$ reduction rules}
+
+\begin{tabular}{lcc}
+	\texttt{beta}                  & $(\lambda x . t) u = t[x / u]$                       &                                                                    \\
+	\texttt{subst\_same\_var}      & $x[x / u] = u$                                       &                                                                    \\
+	\texttt{subst\_diff\_var}      & $y[x / u] = y$                                       & if $x \not \in \mathcal{F}(u)$                                     \\
+	\texttt{subst\_application}    & $(uu')[x / v] = u[x / v]u'[x / v]$                   &                                                                    \\
+	\texttt{subst\_lambda}         & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$    & if $x \not \in \mathcal{F}(v)$                                     \\
+	\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x' / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$             \\
+	\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])$          & if $x' \not \in \mathcal{F}(u)$                                    \\
+	\texttt{subst\_comm}           & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$          & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)~.$ \\
+\end{tabular}
+
+\textcolor{cyan}{These can be encoded in a framework of slotted e-graphs.
+	To demonstrate the effect that a particular set of rewrites has on the number of e-nodes of an e-graph, we saturate the lambda term $plus 2 2$ where $plus$ and $2$ are encoded using Church encoding.
+	The numbers of e-nodes after the e-graph is saturated for each set of rewrite rules are presented in \autoref{tbl:slotted}.
+}
+
+\begin{table}
+	\begin{tabular}{lcccc}
+		Set of equations & explicit subst & explicit subst & explicit subst & explicit subst \\
+		                 &                & + comm         & + interchange  & + interchange  \\
+		                 &                &                &                & + comm         \\
+		\# of e-nodes    & 161            & 288            & 289            & 356
+	\end{tabular}
+	\caption{Slotted e-graphs benchmarks}
+	\label{tbl:slotted}
+\end{table}
+
+\textcolor{cyan}{The \texttt{explicit subst} set of rewrite rules consist of the first five rules above.
+	We can see that additionally saturating the e-graph with respect to interchange and commutativity rules increases the number of e-nodes by a factor of 2.
+	These are the exact equations that are absorbed by using string-diagrammatic representation of $\lambda$-terms.
+	Notably, string diagrams (and (e-)hypergraphs) do not absorb the equation \texttt{subst\_lambda} but we believe it is practically feasible to do rewriting modulo this rule by traversing through the boxes when searching for a redex.
+	Furthermore, (cospans of) e-hypergraphs, as presented in \autoref{sec:closed-e-hyp}, do not absorb the equations for associativity, commutativity and unitality for $\comonoid$ and $\counit$, but the constraints on the cospans and the boundary complement may be relaxed to absorb those as well by following the work of~\cite{zanassi_comonoid} with which our approach is completely compatible.}
+
+In contrast, in our framework, $\lambda^\approx_{\text{lsub}}$-terms are directly modelled as string diagrams (\autoref{fig:plus-1-2-string-diagram}\footnote{Note that dashed boxes are elided for convenience here because no rewriting has occurred.}).
 We do however note that the traditional approach is more flexible, in the sense that it may model more theories with binders (e.g.\ $\pi$-calculus), whereas our approach is essentially specialised to $\lambda$-calculus, relying heavily on the metatheoretic internal language of the ambient setting.
@@
 
-From a practical perspective, the encoding of string diagrams as (e-)hypergraphs provides a normal form for graphical equivalence classes of terms.
+From a practical perspective, the encoding of string diagrams as (e-)hypergraphs provides a normal form for string diagram equivalence classes of terms.
 As a data structure, this provides an efficient representation upon which to build an implementation of this theory as a fully fledged \enquote{e-graphs for string diagram rewriting}, using DPO rewriting as a primitive to implement the standard e-graph operations.
-We leave this as future work, noting its applications to not only $\lambda^\sim_{\text{lsub}}$-calculus, but also other monoidal theories where there are two levels of rewriting:
+We leave this as future work, noting its applications to not only $\lambda^\approx_{\text{lsub}}$-calculus, but also other monoidal theories where there are two levels of rewriting:
 \begin{enumerate}
-	\item \enquote{bureaucratic} reassociation and reparenthesisation, absorbed into isotopy of string diagrams (here, graphical equivalence of $\lambda_{\text{lsub}}$-terms);
+	\item \enquote{bureaucratic} reassociation and reparenthesisation, absorbed into isotopy of string diagrams (here, a fragment of string diagram equivalence of $\lambda_{\text{lsub}}$-terms);
 	\item algebraic identities arising from the semantics of the theory (e.g. ${\text{plus} (x, x) \leadsto \text{mult} (x, 2)}$ for arithmetic).

### 2025-07-09 POPL2026/paper/sections/soundness.tex
comparison, semantics, enrichment fix, appendix
finish up lambda calculus section

@@
 \begin{figure}
-    \begin{subfigure}{0.45\linewidth}
+    \begin{subfigure}[c]{0.45\linewidth}
 \[
 \adjustbox{scale=0.6}{
-    \tikzfig{./figures/ev_interpretation}
+    \tikzfig{./figures/ev_interpretation_2}
 }
@@
     \hfill
-    \begin{subfigure}{0.45\linewidth}
+    \begin{subfigure}[c]{0.45\linewidth}
         \[
         \adjustbox{scale=0.6}{
-            \tikzfig{./figures/lambda_interpretation}
+            \tikzfig{./figures/lambda_interpretation_2}
         }
@@
 \adjustbox{width=0.7\linewidth}{
-\tikzfig{./figures/semilattice_rule_1}~.
+\tikzfig{./figures/semilattice_rule_1_2}~.
 }
@@
 \adjustbox{width=\linewidth}{
-\tikzfig{./figures/semilattice_rule_1}~.
+\tikzfig{./figures/semilattice_rule_1_2}~.
 }
@@
 
-We collect all such rewrite schemas into a set $\mathcal{S}$ and then a category $\WellTypedMdaEcospans / \mathcal{S}$ is a closed $\catname{SLat}$ SMC.
+We collect all such rewrite schemas into a set $\mathcal{S}$ and then a category $\WellTypedMdaEcospans / \mathcal{S}$ is an $\catname{SLat}$ SMC.
+Note that it is not a \textit{closed} $\catname{SLat}$ SMC.
+To turn it into a \textit{closed} $\catname{SLat}$ SMC, we would need to also quotient by the DPO rewrites that arise from the equations in Def~\ref{def:closed}.
+We do not perform such quotienting as these equations bear operational meaning, \textit{e.g.}, they model $\beta$-reduction when interpreting $\lambda$-calculus.
 
@@
 
-\begin{proposition}
+\begin{proposition}%
+	\label{prop:quotient-structural}
 	For two closed $\Sigma^{+}$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
@@
         \adjustbox{scale=0.5}{
-            \tikzfig{./figures/f_plus_g_new}
+            \tikzfig{./figures/f_plus_g_new_2}
         }

### 2025-07-09 POPL2026/paper/sections/combinatorial_semantics.tex
notes and popl template
comparison, semantics, enrichment fix, appendix
figures and benchmarks

--- 2025-06-12 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
@@
     \end{subfigure}
-    \hfill
-    \captionsetup{belowskip=-3ex}
     \caption{DPO and DPOI squares}
@@
 \section{Closed E-hypergraphs}
+\label{sec:closed-e-hyp}
 Prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
@@
   \item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are \textit{label} functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
-  \item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times V + E$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times V + E $ with $\textcolor{e-color}{E}, \textcolor{closed-color} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
+  \item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V + E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V + E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
         For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{<^{\mu}} y_{i+1}\; \}$.
@@
           \item each $x$ has at most one immediate predecessor;
-          \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{<^{\mu}} x$ or $e \textcolor{e-color}{<^{\mu}} x$ are labelled;
-          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{e-color}{<^{\mu}} e$ iff $e' \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<}$.
+          \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{<^{\mu}} x$ or $e \textcolor{closed-color}{<^{\mu}} x$ are labelled;
+          \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{e-color}{<^{\mu}} e$ iff $e' \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<^{\mu}}$.
         \end{enumerate}
@@
 $\textcolor{e-color}{E}$ and $\textcolor{closed-color}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$.
-Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
+When writing $x \textcolor{closed-color}{<} y$ (or $x \textcolor{e-color}{<} y$) we may refer to $x$ as a predecessor of $y$ and refer to $y$ as a successor of $x$.
+% we do not use this
+% Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
 \end{definition}
@@
 \[
-\vspace{-1ex}
 \begin{array}{cccc}
@@
 \]
+
 and, in particular, we have $[e_{5}) = \{e_2, e_1\}$.
@@
 \adjustbox{scale=0.6}{
-\tikzfig{./figures/closed_iso_example_2}
+\tikzfig{./figures/closed_iso_example_2_2}
 }
@@
 All non-black vertices form strictly internal interfaces.
+We follow the same orientation that we chose for string diagrams --- the input interface of the cospan is at the bottom and the output interface is at the top.
 

### 2025-07-09 POPL2026/paper/sections/lambda-calculus.tex
figures and benchmarks

--- 2025-07-09 POPL2026/paper/sections/lambda-calculus.tex
@@
 \section{Application --- $\lambda^\sim_{\text{lsub}}$-calculus}
+\label{sec:application}
 
@@
 As alluded to in~\autoref{sec:vs-e-graphs-with-binding}, the traditional (slotted) e-graph approach to modelling the graphical equivalence of these terms (\autoref{fig:plus-1-2-terms}) relies on explicitly encoding substitution operations as e-graph nodes, and their behaviour (the graphical equivalence relation $\sim$) as rewrite rules, as in~\autoref{fig:e-graph-substitution} --- not only is this more complicated, but it also leads to a blow-up in the state space required to perform equality saturation in any extension of the $\lambda^\sim_{\text{lsub}}$-calculus, because we are really rewriting $\lambda_{\text{lsub}}$-terms and computational resources are spent rewriting between graphically equivalent terms.
+\textcolor{cyan}{To illustrate this, let's consider the following equations induced by the graphical equivalence and the $\lambda_{\text{lsub}}$ reduction rules}
+
+\begin{tabular}{lcc}
+	\texttt{beta} & $(\lambda x . t) u = t[x / u]$&\\
+	\texttt{subst\_same\_var} & $x[x / u] = u$ &\\
+	\texttt{subst\_diff\_var} & $y[x / u] = y$ & if $x \not \in \mathcal{F}(u)$\\
+	\texttt{subst\_application} & $(uu')[x / v] = u[x / v]u'[x / v]$ & \\
+	\texttt{subst\_lambda} & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$ & if $x \not \in \mathcal{F}(v)$ \\
+	\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x' / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$\\
+	\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])$ & if $x' \not \in \mathcal{F}(u)$\\
+	\texttt{subst\_comm} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$ & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)~.$\\
+\end{tabular}
+
+\textcolor{cyan}{These can be encoded in a framework of slotted e-graphs.
+To demonstrate the effect that a particular set of rewrites has on the number of e-nodes of an e-graph, we saturate the lambda term $plus 2 2$ where $plus$ and $2$ are encoded using Church encoding.
+The numbers of e-nodes after the e-graph is saturated for each set of rewrite rules are presented in \autoref{tbl:slotted}.
+}
+
+\begin{table}
+	\begin{tabular}{lcccc}
+	  Set of equations & explicit subst & explicit subst & explicit subst & explicit subst\\
+	  && + comm& + interchange&  + interchange\\
+	  &&&&  + comm\\
+	  \# of e-nodes & 161 & 288 & 289 & 356
+	\end{tabular}
+	\caption{Slotted e-graphs benchmarks}
+	\label{tbl:slotted}
+\end{table}
+
+\textcolor{cyan}{The \texttt{explicit subst} set of rewrite rules consist of the first five rules above.
+We can see that additionally saturating the e-graph with respect to interchange and commutativity rules increases the number of e-nodes by a factor of 2.
+These are the exact equations that are absorbed by using string-diagrammatic representation of $\lambda$-terms.
+Notably, string diagrams (and (e-)hypergraphs) do not absorb the equation \texttt{subst\_lambda} but we believe it is practically feasible to do rewriting modulo this rule by traversing through the boxes when searching for a redex.
+Furthermore, (cospans of) e-hypergraphs, as presented in \autoref{sec:closed-e-hyp}, do not absorb the equations for associativity, commutativity and unitality for $\comonoid$ and $\counit$, but the constraints on the cospans and the boundary complement may be relaxed to absorb those as well by following the work of~\cite{zanassi_comonoid} with which our approach is completely compatible.}
+
 \textcolor{red}{Aleksei, can you say something quantitative here if you ran some tests?}

### 2025-07-09 POPL2026/paper/sections/soundness.tex
figures and benchmarks

@@
 
-\begin{proposition}%
-	\label{prop:quotient-structural}
+\begin{proposition}
     For two closed $\Sigma^{+}$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.

### 2025-07-10 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
mostly finish lambda calculus section
comparison, semantics, enrichment fix, appendix
fixup introduction

@@
 \usepackage{algorithmic}
+\usepackage{listings}
 \usepackage{graphicx}
@@
 \usepackage{microtype}
+
+% inference rules
+\usepackage{mathpartir}
+
+\newcommand{\ONECOLUMN}{}
 
@@
 	However, they struggle in domains like $\lambda$-calculus that involve variable binding, due to a lack of native support for bindings.
-	Building on recent work interpreting e-graphs categorically as morphisms in semilattice-enriched symmetric monoidal categories, we extend this framework to closed symmetric monoidal categories to handle bindings.
-	We provide a concrete combinatorial representation using hierarchical hypergraphs and introduce a corresponding double-pushout (DPO) rewriting mechanism. 
-	Finally, we establish the equivalence of term rewriting and DPO rewriting, with the key property that the combinatorial representation absorbs the equations of the symmetric monoidal category.
+	Building on recent work interpreting e-graphs categorically as morphisms in symmetric monoidal semilattice-enriched categories, we extend this framework to closed symmetric monoidal semilattice-enriched categories to handle bindings.
+	We provide a concrete combinatorial representation using an extension of hierarchical hypergraphs and introduce a corresponding double-pushout (DPO) rewriting mechanism.
+	Finally, we establish the equivalence of term rewriting and DPO rewriting, with the key property that the combinatorial representation absorbs the equations of the symmetric monoidal structure.
 \end{abstract}
@@
 
-% \input{sections/appendix}
+\input{sections/appendix}
 

### 2025-07-10 POPL2026/paper/sections/introduction.tex
figures and benchmarks
fixup introduction

--- 2025-07-08 POPL2026/paper/sections/introduction.tex
@@
-\section{Introduction}
+\section{Introduction}%
 \label{sec:introduction}
@@
 
-E-graphs (short for \emph{equality graphs})~\cite{EggPaper} are a data structure incorporating equivalence representation and used in automated theorem proving, compiler optimisation, and program analysis.
-Using e-graphs mitigates the so-called \emph{phase-ordering problem} of term rewriting, that is the fact that the way rewrites are scheduled may enable or block further rewrites due to introduction or elimination or redexes.
-This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule, instead of replacing the left-hand side by the right-hand side, adds the latter to the graph so that all pre-existing redexes are protected and new ones are introduced.
-Explosive growth of the graph is avoided by the clever use of sharing of vertices.
+E-graphs (short for \emph{equality graphs})~\cite{EggPaper} are an efficient data structure for non-destructive rewriting, incorporating equivalence representation, used in automated theorem proving, compiler optimisation, and program analysis.
+They mitigate the \emph{phase-ordering problem} of term rewriting: the way rewrites are scheduled may enable or block further rewrites due to introduction or elimination or redexes, resulting in a suboptimal result.
+This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule, instead of replacing the left-hand side by the right-hand side, adds the latter to the e-graph so that all pre-existing redexes are protected and new ones are introduced.
+Explosive growth of the e-graph is avoided by the clever use of sharing of vertices.
 
-E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using so-called technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint is reached.
+E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using the technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint (or timeout) is reached.
 The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
-Of particular interest to us is using e-graphs in the context of functional programming languages, which builds upon an internal representation of the $\lambda$-calculus. 
-Pioneering work in this direction has been done by~\cite{koehler2022sketchguided} where
-the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise applying equality saturation to $\lambda$-calculus e-graphs.
+We are interested in using e-graphs in the context of functional programming languages, which builds upon an internal representation of the $\lambda$-calculus.
+Pioneering work in this direction has been done by~\citet{koehler2022sketchguided}, where the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when applying equality saturation to an encoding of the $\lambda$-calculus in e-graphs.
 
 In the textual representation of $\lambda$-terms, $\alpha$-equivalent terms are distinct and thus cannot be shared, which adds a layer of bookkeeping and complexity.
-The traditional approach to handle this is by using de Bruijn indices~\cite{de1972lambda}, for which $\alpha$-equivalent terms are syntactically identified; however, this is also problematic because the same variable can be assigned different indices depending on its scope with respect to that instance of the variable.
-This again interferes with sharing. 
+This is traditionally handled by using de Bruijn indices~\cite{de1972lambda}, for which $\alpha$-equivalent terms are syntactically identified; however, this creates issues because the same variable can be assigned different indices depending on its scope with respect to the instance of that variable, again interfering with sharing.
 As a simple example, consider the term $\lambda f . f ((\lambda x . f x) 2)$ which, with de Bruijn indices,  is encoded as $\lambda . \%0 ((\lambda . \%1 \%0) 2)$.
@@
 
-Another distinct challenge is presented by $\beta$-reduction, which is handled using \emph{explicit substitutions}. 
-The implementations of $\beta$-reduction in e-graphs do not go into much detail regarding explicit substitution, and in particular do not include rules like $[x\!\mapsto\!1][y\!\mapsto\!2]f\!=\![y\!\mapsto\!2][x\!\mapsto\!1]f$ for non-interfering substitutions, while such rules are absorbed by graph equivalence~\cite{accattoli2014nonstandard}.
+A simple extension of $\lambda$-calculus is given by adding \emph{explicit substitution}, which models \textbf{let}-bindings in programming languages.
+Here, a term equipped with explicit substitutions $t[y/v][x/u]$ (modelling $\textbf{let} \; x = u \; \textbf{in} \; \textbf{let} \; y = v \; \textbf{in} \; t$) can be identified with the corresponding term where the order of substitution is reversed $t[x/u][y/v]$, supposing the substitutions are non-interacting.
+This imposes a \emph{graphical equivalence}~\cite{accattoli2014nonstandard} of terms.
+Modelling this in the conventional language of e-graphs requires the introduction of additional nodes to represent the explicit substitutions, and additional rewrite rules which capture graphical equivalence.
+This pollutes the state space of equality saturation for these languages: the interesting rewriting is that of the user-specified algebraic theory one wishes to model, not between semantically equivalent but syntactically distinct forms in the language.
 
 All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
-\cite{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
+\citet{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using symmetric monoidal semilattice-enriched categories, which allows e-graphs to be reconstructed as string diagrams.
 In this paper, we extend this technique to e-graphs with bindings, lifting to monoidal \emph{closed} categories.
@@
 
-We build on the technology of string diagrams, by now well-developed in a number of areas. 
 String diagrams are a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
@@
 
-The general background literature for string diagrams is summarised in~\cite{piedeleu2023introductionstringdiagramscomputer}.
-The particular context of string diagrams with bindings is given in~\cite{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+The general background literature for string diagrams and their relationship to formal languages is summarised by~\citet{piedeleu2023introductionstringdiagramscomputer}.
+The particular context of string diagrams with bindings is given in~\citet{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}.
 
 String diagrams provide a topological framework for understanding monoidal categories, representing objects as \emph{strings} and morphisms as \emph{nodes}, where input strings flow in and output strings flow out.
-Key structures in monoidal categories---such as the monoidal product, tensor unit, braiding, and dualities---are represented graphically, through operations like juxtaposing, weaving, or bending strings.
-This visual approach makes complex operations such as the trace intuitive and reveals insights sometimes obscured by traditional symbolic notation. 
-Originating from notations by Hotz~\cite{hotzsd} (for automata) and Penrose~\cite{penrose1984spinors} (for tensor networks), string diagrams generalise to any monoidal category and even bicategories, and provide a mathematically rigorous diagrammatic calculus~\cite{joyal_geometry_1991}.
+In this article, our string diagrams are oriented left-to-right, bottom-to-top to make the pictures look more similar to those commonly drawn for e-graphs.
+The structure of symmetric monoidal categories is absorbed into the graphical syntax, with the composition and tensoring represented by juxtaposition, and symmetries as weaving of strings.
+This visual approach makes hides the bureacracy in complex operations and reveals insights sometimes obscured by traditional one-dimensional symbolic notation.
+Originating from notations by~\citet{hotzsd} (for automata) and~\citet{penrose1984spinors} (for tensor networks), extensions of string diagrammatic calculi exist for various forms of monoidal categories and even bicategories~\cite{Selinger_2010}, and provide a mathematically rigorous reasoning toolkit~\cite{joyal_geometry_1991}.
 
@@
 Specifically, we require a concrete encoding of string diagrams, analogous to the de Bruijn index representation of $\lambda$-terms.
-For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics for this purpose.
-For instance, hypergraphs admit the property that two string diagrams are equal if and only if they ``look the same'' (\emph{i.e.} they admit a graph isomorphism); secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.
+For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics.
+For instance, hypergraph isomorphism captures when string diagrams \enquote{look the same} --- topologically, that one may be deformed continuously into the other without obstruction\footnote{This is the informal notion of the \enquote{isotopy of string diagrams}.}; secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting over these hypergraphs using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.
 
-Finally, for the diagrammatical representation of two concepts essential to both programming languages (binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
-The diagrammatic boxes require the introduction of further structure to (hyper)graphs in the form of a hierarchical layering of the graphs, as studied in \emph{loc. cit.}.
+Finally, for the diagrammatic representation of two concepts essential to both programming languages (variable binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
+The diagrammatic boxes require the introduction of further structure to hypergraphs in the form of a hierarchical layering of the graphs (\emph{hierarchical hypergraphs})~\cite{fscd}.
 
@@
 
-Revisiting the simple earlier example $\lambda f . f ((\lambda x . f x) 2)$, in Fig.~\ref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
-$\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire attached to the frame,
-the application node a half-circle
+Revisiting the earlier example $\lambda f . f ((\lambda x . f x) 2)$, in~\autoref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
+$\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire exiting the frame along its vertical boundary, the application node a half-circle
 \adjustbox{scale=0.15}{\begin{tikzpicture}
@@
 },
-and sharing a \emph{contraction node} 
-\scalebox{0.3}{
-	\begin{tikzpicture}
-		\begin{pgfonlayer}{nodelayer}
-			\node [style=vertex] (0) at (1.5, 3.5) {};
-			\node [style=none] (1) at (1.5, 3) {};
-			\node [style=none] (2) at (1, 4) {};
-			\node [style=none] (3) at (2, 4) {};
-		\end{pgfonlayer}
-		\begin{pgfonlayer}{edgelayer}
-			\draw (0) to (1.center);
-			\draw [bend right=45] (0) to (3.center);
-			\draw [bend left=45] (0) to (2.center);
-		\end{pgfonlayer}
-	\end{tikzpicture}
-}
-.
+and sharing a \emph{copy node} \comonoid.
 
@@
 E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
-The key concepts of e-graphs can be intuitively grasped from some simple (but non-trivial) examples, as shown in Fig.~\ref{fig:e-graph-example}, where it is shown how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
+The key concepts of e-graphs can be intuitively grasped from some simple examples: \autoref{fig:e-graph-example} shows how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
 
@@
 
-Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node.
+Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the \emph{e-hypergraphs} of~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}.
+For the initial term, $(a*2)/2$, which is already represented efficiently in the e-graph by sharing the node $2$, the string diagram version shares this node explicitly, using a copy node.
 
-Nodes of an e-graph are encapsulated in dashed boxes indicating equivalence classes of nodes, or, more generally, of subgraphs rooted at these boxes.
-A term graph is embedded as an e-graph by making each equivalence class have a single node.
-The first rewrite rule, replacing multiplication by \emph{shift-left} creates the first non-unitary equivalence class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) nodes,  now sharing the node `$a$'.
+E-graphs consist of two types of nodes: \emph{e-classes}, representing equivalence classes of terms and denoted by dashed boxes, containing \emph{e-nodes}, representing term formers.
+E-nodes reference e-classes instead of other e-nodes, in contrast to a traditional term graph.
+By making a sequence of choices for e-nodes in e-classes, we can extract terms from the graph; intuitively, a different sequence of choices results in an equivalent term.
+A term graph is embedded as an e-graph by an e-graph where each e-class has a single e-node.
+The first rewrite rule, replacing multiplication by \emph{shift-left} creates the first non-singleton e-class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) nodes, now sharing the node $a$.
 This second e-graph illustrates the other representational difference, besides sharing, between conventional e-graphs and e-hypergraphs.
-In the former, edges connect directly to nodes inside the equivalence class, whereas in the latter edges connect to the equivalence class itself.
-Explicit discarding (weakening) nodes are depicted with black dots with a single wire
+In the former, edges connect directly to e-nodes inside the e-class, whereas in the latter edges connect to the equivalence class itself.
+Explicit deletion nodes are depicted with black dots with a single incoming wire
 \scalebox{0.4}{
@@
 }
-inside the dashed boxes, indicating which of the class-level edges are connected to which nodes inside the class.
+inside the dashed boxes, indicating which of the equivalence-class-level edges are connected to which nodes inside the equivalence class.
 
-The other steps, (c) to (d), represent further elaborations of the e-graph, or e-hypergraph, via the application of more rewrite rules.
+The other steps represent further elaborations of the e-graph, or e-hypergraph, via the application of more rewrite rules.
 
-A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
-The auxiliary nature of these nodes is revealed by the fact that, after rewriting, they may be removed from the graph. 
-We will see how in the categorical representation these nodes are not necessary because explicit substitution is simply modelled categorically as composition, a significant simplification.
-
-\subsection{Conventional vs. categorical e-graphs with binding}
+\subsection{Conventional vs.\ slotted vs.\ categorical e-graphs with binding}%
+\label{sec:vs-e-graphs-with-binding}
 
 First, let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
-In Fig.~\ref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
+
+A common way of dealing with $\beta$-reduction for the $\lambda$-calculus in e-graphs is by representing the rules of $\lambda$-calculus explicitly, using special nodes for explicit substitution and associating them with particular rewrite rules~\cite{EggPaper,koehler2022sketchguided}.
+The auxiliary nature of these nodes is evident in that, after rewriting, they may be removed from the graph.
+We will see how in our categorical representation these nodes are not necessary, because explicit substitution is simply modelled categorically as composition, a significant simplification.
+
+In~\autoref{fig:e-graph-substitution} we show successive iterations of a conventional e-graph for $(\lambda x . (y + y) + x) 1$ following $\beta$-reductions.
 A subterm for an explicit substitution must be introduced, which is deemed equivalent to the top-level application term (hence they are inside the same dashed box).
@@
 At the end, all explicit substitution nodes can be removed.
-The same example is then given again in Fig.~\ref{fig:e-string-substitution}, this time using string diagrams.
+The same example is then given again in~\autoref{fig:e-string-substitution}, this time using string diagrams.
 
 Using string diagrams, $\beta$-reduction is simply the re-wiring of the diagram where an application node meets an abstraction, and explicit substitution simplifies into mere composition.
-This is appears to be simpler just by comparing diagrams, but the technical advantages of using string diagrams run deep.
-The explicit occurrences of variable names in the older approach raise obvious questions of scope, binding and explicit substitution, whereas the use of terms inside the explicit substitution nodes also raises many technical difficulties which makes proving various mathematical properties essentially intractable.
+The explicit occurrences of variable names in the conventional approach raise obvious questions of scope, binding and explicit substitution, whereas the use of terms inside the explicit substitution nodes also raises many technical difficulties which makes proving various mathematical properties essentially intractable.
 Additionally, explicit substitutions need to be propagated recursively throughout the graph, which adds another layer of technical and conceptual overhead.
-Using de Bruijn indices when performing $\beta$-reduction would hardly solve this problem, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
-A recent work of~\cite{slotted-egraphs} on so-called \textit{slotted} e-graphs has addressed the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by making variables into first-class citizens of an e-graph and made binding a built-in operation for an e-graph.
-An example of a slotted e-graph that represents a term $\lambda x . (a + b) + (x + c)$ can be seen in Figure~\ref{fig:slotted}(left).
+Using de Bruijn indices instead when performing $\beta$-reduction would only push the problem elsewhere, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
+Recent work on \textit{slotted} e-graphs~\cite{slotted-egraphs} uses nominal techniques to address the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by incorporating variables as first-class data of an e-graph and making binding a built-in operation.
+They also solved the problem of explicit variable names, guaranteeing freshness inside binders, explicitly differentiating between bound and free variables.
+An example of a slotted e-graph is given in~\autoref{fig:slotted}.
 
@@
 		\]
+		\caption{Slotted e-graph}%
+		\label{fig:slotted}
 	\end{subfigure}
@@
 		\]
-
+		\caption{Closed e-hypergraph}%
+		\label{fig:closed}
 	\end{subfigure}
-	\caption{}
-	\label{fig:slotted}
+	\caption{Extensions of e-graphs to handle binding in $\lambda x . (a + b) + (x + c)$.}%
+	\label{fig:extended-egraphs}
 \end{figure}
 
-A notable feature of a slotted e-graph is that every e-class is parameterised by a set of free variables of the terms represented by this e-class.
-The references then additionally convey the information about how variables are instantiated.
-For example, an e-class containing $+$ that is parameterised by variable names $s_2$ and $s_3$ gets intantiated once with $a,b$ (via $s_4 \mapsto a, s_2 \mapsto s_4$ and $s_5 \mapsto b, s_3 \mapsto s_5$) and once  with $x,c$.
+In slotted e-graphs, every e-class is parameterised by a set of free variables of the terms it represents.
+The edges then additionally convey the information about how variables are instantiated.
+For example, in~\autoref{fig:slotted}, an e-class containing $+$ that is parameterised by variable names $s_2$ and $s_3$ gets instantiated once with $a,b$ (via $s_4 \mapsto a, s_2 \mapsto s_4$ and $s_5 \mapsto b, s_3 \mapsto s_5$) and once with $x,c$.
 The fact that the variable $x$ is bound is evident from the corresponding e-class being parameterised by only three variables.
-If we were to add another e-node such as $e + d$, there would be a reference added to an already existing e-class such that $s_1 \mapsto e$, $s_2 \mapsto d$.
-We can achieve the same degree of sharing (in the sense we have the same number of nodes for $+$) using string diagrams by doing a closure conversion of $+$ (Figure~\ref{fig:slotted}(right)).
-However, the addition of these parameters and the corresponding alternations in the algorithmics makes the deeper correspondence between string diagrams and slotted e-graphs obscure.
-Despite this, there are applications where the use of string diagrams may be more beneficial than that of (slotted) e-graphs, for example in the case of lambda calculus with explicit substitution that we will describe next.
-Also, the categorical model that we built can help to explain the congruences maintained by slotted e-graphs which is further elaborated on in Section~\ref{sec:categorical}.
+If we were to add another e-node such as $e + d$, there would be an edge added to an already existing e-class such that $s_1 \mapsto e$, $s_2 \mapsto d$.
+We can achieve a similar degree of sharing (in the sense that we have the same number of nodes for $+$) using string diagrams by performing a closure conversion of $+$ (\autoref{fig:closed}).
+Fundamentally, there is a difference in approach here: slotted e-graphs solve the problem of variable binding generically using nominal techniques, in a way that can be applied to arbitrary formal languages with binders (e.g.
+$\pi$-calculus), while ours utilises an embedding into the internal language of our categorical model, which resembles a kind of metatheoretic $\lambda$-calculus, for which we use the metatheoretic $\lambda$-abstraction to handle the binding of variables.
+We defer a case study to~\autoref{sec:application}, where we argue that our approach has advantages if one is particularly interested in e-graphs for (variations on) $\lambda$-calculus, and, by extension, programming languages.
+However, these constructions are not unrelated: our categorical model can help justify the congruences maintained by slotted e-graphs, which we further elaborate on in~\autoref{sec:categorical}.
+We also note that there is also a difference in the scopes of these related works: \citet{slotted-egraphs} focus on the e-graph as a data structure, providing a full implementation of their slotted e-graphs, while we focus on the foundational categorical semantics of e-graphs with binding and their combinatorial representation.
 
-% Notably, a work of~\cite{slotted-egraphs} has emerged recently which introduced a formulation of e-graphs with native support for binding, \textit{i.e.,} where binding is a first-class operation along with a substitution.
-% It significantly complicates the formulation of an e-graph and corresponding algorithmics and based on \textit{nominal} techniques~\cite{Pitts_2013} unlike the approach presented in this work.
-% We believe the approach used in this work strikes with its simplicity although a more detailed comparison with the work of remains a future work.
-
-\subsection{Lambda calculus with explicit substitutions and graphical equivalence of terms}
+% Do we need this section?
+% \subsection{Lambda calculus with explicit substitutions and graphical equivalence of terms}
 
@@
 
-In~\cite{ghica2024stringdiagramslambdacalculifunctional} it was shown that e-graphs over signature $\Sigma$ can be represented as morphisms of a semilattice-enriched free Cartesian symmetric monoidal category over the same signature.
-We argue that restricting the categorical domain to free semilattice-enriched closed symmetric monoidal categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
-That is, morphisms (string diagrams) of such category can both encode variable binding and equivalence between subterms (sub-diagrams).
-We call these string diagrams, or more precisely, the combinatorial representation of them, e-graphs with bindings.
-
-We then define a combinatorial representation for such string diagrams in terms of hierarchical hypergraphs with a suitable notion of DPO rewriting.
-Finally, we show that this notion of rewriting is sound and complete with respect to rewriting of morphism terms.
+\citet{ghica2024equivalencehypergraphsegraphsmonoidal} shows that e-graphs over signature $\Sigma$ can be represented as morphisms of a free cartesian symmetric monoidal semilattice-enriched category over the same signature.
+We argue that restricting the categorical domain to free \textbf{closed} symmetric monoidal semilattice-enriched categories gives rise to structures that naturally support binding and equivalence classes of morphisms.
+That is, morphisms (string diagrams) of such category can both encode variable binding (using the closed structure) and equivalence (using the semilattice enrichment) between subterms (sub-diagrams).
+We call these string diagrams, or more precisely, the combinatorial representation of them, \emph{closed e-hypergraphs}.
+We equip closed e-hypergraphs with a suitable notion of DPO rewriting, and show that this notion of rewriting is sound and complete with respect to rewriting of morphism terms.
 

### 2025-07-10 POPL2026/paper/sections/lambda-calculus.tex
fixup introduction

--- 2025-07-09 POPL2026/paper/sections/lambda-calculus.tex
@@
 By further extending the equivalence relation $\sim$, taking the symmetric closure by $\to_{ls}$ and $\to_{gc}$ and thus identifying terms under elimination of explicit substitution, we determine a \emph{string diagram equivalence} $\approx$ of $\lambda_{\text{lsub}}$-terms, determining a $\lambda^\approx_{\text{lsub}}$-calculus.
-Predictably, $t \approx u$ if and only if both map to the same string diagram~\cite[§ 5.1.1]{ghica2024stringdiagramslambdacalculifunctional}; this allows us to exploit the string diagram setting of e-hypergraphs to absorb many\footnote{The reason this is not \emph{all} reductions is because e-hypergraphs quotient away only the \emph{structural} isomorphisms of symmetric monoidal closed $\catname{SLat}$-categories (e.g.\ associators, unitors, etc.) which generate the \enquote{isotopy of string diagrams}; even in the graphical equivalence relation $\sim$, while~\autoref{eq:subst-lambda} is validated in every symmetric monoidal closed $\catname{SLat}$-category, it is not so by isotopy, but rather the adjunction $A \multimap - \dashv - \otimes A$.
+Predictably, $t \approx u$ if and only if both map to the same string diagram~\cite[§ 5.1.1]{ghica2024stringdiagramslambdacalculifunctional}; this allows us to exploit the string diagram setting of e-hypergraphs to absorb many\footnote{The reason this is not \emph{all} reductions is because e-hypergraphs quotient away only the \emph{structural} isomorphisms of symmetric monoidal closed $\catname{SLat}$-categories (e.g.\ associators, unitors, etc.) which generate the \enquote{isotopy of string diagrams}; even in the graphical equivalence relation $\sim$, while~\autoref{eq:subst-lambda} is validated in every symmetric monoidal closed $\catname{SLat}$-category, it is not so by isotopy, but rather the adjunction $ - \otimes A \dashv A \multimap -$.
 In other words, terms floating freely through the boundaries of rounded boxes is validated by our semantics, but produces concretely different e-hypergraphs --- this is the purpose of rectification by quotienting along $\mathcal{S}$ in~\autoref{prop:quotient-structural}, and manifests as additional DPO rewrites to be imposed.} of the reductions arising due to the equivalence $\approx$ into our representation rather than explicitly providing them as equations attached to a closed symmetric monoidal theory presenting the $\lambda^\approx_{\text{lsub}}$-calculus.
@@
 
-Note that, in contrast to traditional presentations of the $\lambda$-calculus as a language of (slotted~\cite[Example 1]{slotted-egraphs}) e-graphs~\cite[Figure 10]{EggPaper}, here there is no encoding of the term formers for application, abstraction, and explicit substitution.
+Note that, in contrast to conventional presentations of the $\lambda$-calculus as a language of (slotted~\cite[Example 1]{slotted-egraphs}) e-graphs~\cite[Figure 10]{EggPaper}, here there is no encoding of the term formers for application, abstraction, and explicit substitution.
 Instead, these constructs are handled ambiently by our semantic setting: any closed symmetric monoidal category (and in particular, $\catname{CS}(\Sigma, \mathcal{E})$ for any set of equations $\mathcal{E}$) admits an internal language in the multiplicative fragment of intuitionistic linear logic (a particular type of linear $\lambda$-calculus)~\cite[§ 1.7.3]{abramskyIntroductionCategoriesCategorical2010}.
@@
 In contrast, in our framework, $\lambda^\approx_{\text{lsub}}$-terms are directly modelled as string diagrams (\autoref{fig:plus-1-2-string-diagram}\footnote{Note that dashed boxes are elided for convenience here because no rewriting has occurred.}).
-We do however note that the traditional approach is more flexible, in the sense that it may model more theories with binders (e.g.\ $\pi$-calculus), whereas our approach is essentially specialised to $\lambda$-calculus, relying heavily on the metatheoretic internal language of the ambient setting.
+We do however note that the conventional approach is more flexible, in the sense that it may model more theories with binders (e.g.\ $\pi$-calculus), whereas our approach is essentially specialised to $\lambda$-calculus, relying heavily on the metatheoretic internal language of the ambient setting.
 

### 2025-07-10 POPL2026/paper/sections/categorical_semantics.tex
mass format

@@
 	\begin{aligned}
-\Lambda_{A,B,C}(f + g) &= \eta;(B \multimap (f + g))\\
-                       &= \eta;(B \multimap (f) + B \multimap (g))\\
-                       &= \eta;(B \multimap (f)) + \eta;(B \multimap (g))\\
-                       &= \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g)
+		\Lambda_{A,B,C}(f + g) & = \eta;(B \multimap (f + g))                      \\
+		                       & = \eta;(B \multimap (f) + B \multimap (g))        \\
+		                       & = \eta;(B \multimap (f)) + \eta;(B \multimap (g)) \\
+		                       & = \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g)
 	\end{aligned}
@@
 	\begin{align*}
-&\text{functional symbols}\; f, g, \ldots\\
-&\text{slots}\;  \$x, \$y, \ldots\\
-&\text{term children}\; tc  \Coloneqq t \;|\; tc_{1}, \ldots, tc_{n} \;|\; \text{bind}\; \$x\; tc \;|\; \$x\\ 
-&\text{terms}\;  \Coloneqq f \;|\; f(tc)
+		 & \text{functional symbols}\; f, g, \ldots                                                                   \\
+		 & \text{slots}\;  \$x, \$y, \ldots                                                                           \\
+		 & \text{term children}\; tc  \Coloneqq t \;|\; tc_{1}, \ldots, tc_{n} \;|\; \text{bind}\; \$x\; tc \;|\; \$x \\
+		 & \text{terms}\;  \Coloneqq f \;|\; f(tc)
 	\end{align*}

### 2025-07-10 POPL2026/paper/sections/combinatorial_semantics.tex
mass format

@@
 \begin{align*}
-n \xrightarrow{f} &\mathcal{F} \xleftarrow{g} m\\
-&\times \qquad \quad \Coloneqq n + n' \xrightarrow{f + f'} \mathcal{F} + \mathcal{G} \xleftarrow{g + g'} m + m'\\
-n' \xrightarrow{f'} &\mathcal{G} \xleftarrow{g'} m'
+	n \xrightarrow{f}   & \mathcal{F} \xleftarrow{g} m                                                                                   \\
+	                    & \times \qquad \quad \Coloneqq n + n' \xrightarrow{f + f'} \mathcal{F} + \mathcal{G} \xleftarrow{g + g'} m + m' \\
+	n' \xrightarrow{f'} & \mathcal{G} \xleftarrow{g'} m'
 \end{align*}
@@
 	\begin{array}{cccc}
-  v_{4} \textcolor{e-color}{<^{\mu}} e_{1} & v_{7} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{4} \consistency e_{2} & v_{4} \not \consistency v_{10}\\
-  e_{3} \textcolor{e-color}{<^{\mu}} e_{1} & v_{8} \textcolor{closed-color}{<^{\mu}} e_{2} &  e_{2} \consistency u_{1} & e_{2} \not \consistency e_{6}\\
-  e_{2} \textcolor{e-color}{<^{\mu}} e_{1} & e_{5} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency e_{4} & e_{4} \not \consistency e_{7}\\  
+		v_{4} \textcolor{e-color}{<^{\mu}} e_{1}  & v_{7} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{4} \consistency e_{2}   & v_{4} \not \consistency v_{10} \\
+		e_{3} \textcolor{e-color}{<^{\mu}} e_{1}  & v_{8} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency u_{1}   & e_{2} \not \consistency e_{6}  \\
+		e_{2} \textcolor{e-color}{<^{\mu}} e_{1}  & e_{5} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency e_{4}   & e_{4} \not \consistency e_{7}  \\
 		\ldots                                    & \ldots                                        & \ldots                     & \ldots                         \\

### 2025-07-10 POPL2026/paper/sections/appendix.tex
notes and popl template
comparison, semantics, enrichment fix, appendix
figures and benchmarks
fixed figures, updated contributions

--- 2025-01-27 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
@@
-\subsection{$\catname{SLat}$}
+\section{$\catname{SLat}$}
 \label{sec:appendix:slat}
@@
 
-\subsection{Pushout in $\catname{EHyp}(\Sigma)$}
+\section{Pushout in $\catname{EHyp}(\Sigma)$}
 \label{sec:appendix:pushout}
@@
     \begin{align*}    
-    S_{V} &= \{\\
-          &\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+    S_{V} &= \{
+          \;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
           &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
+          \}\\
           &\cup\\
-          &\{\\
-          &\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
+          &\{
+          \;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
           &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
-          &\}\\
+          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
+          \}\\
           &\cup\\
-          &\{\\
-          &\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}\\
-          &\}\\
+          &\{
+          \;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}
+          \}\\
         \\
@@
 \begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = (&\\
-            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
+        \sfrac{X + Y}{\sim (R_{V},R_{E})} = &(
+            \sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
             &\sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},\\
@@
             &\textcolor{e-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-            &\textcolor{closed-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-        &)    
+            &\textcolor{closed-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}})    
 \end{align*}
@@
 In particular, a coproduct of two feet contains\begin{tikzpicture}
+    \adjustbox{scale=0.8}{
 	\begin{pgfonlayer}{nodelayer}
-		\node [style=node, label={above:$v_1$}] (0) at (-5.5, 3.5) {};
+		\node [style=node, label={above:$v_1$}] (0) at (-5, 3.5) {};
 		\node [style=node, label={above:$v_2$}] (1) at (-4.5, 3.5) {};
-		\node [style=node, label={above:$v_3$}] (2) at (-3.5, 3.5) {};
-		\node [style=node, label={above:$v'_1$}] (3) at (-2.5, 3.5) {};
-		\node [style=node, label={above:$v'_2$}] (4) at (-1.5, 3.5) {};
-		\node [style=node, label={above:$v'_3$}] (5) at (-0.5, 3.5) {};
+		\node [style=node, label={above:$v_3$}] (2) at (-4, 3.5) {};
+		\node [style=node, label={above:$v'_1$}] (3) at (-3.5, 3.5) {};
+		\node [style=node, label={above:$v'_2$}] (4) at (-3, 3.5) {};
+		\node [style=node, label={above:$v'_3$}] (5) at (-2.5, 3.5) {};
 	\end{pgfonlayer}
+    }
 \end{tikzpicture}
@@
 
-\subsection{Interpretation of $\Sigma$-terms as cospans of hypergraphs}
+\section{Interpretation of $\Sigma$-terms as cospans of hypergraphs}
 \label{sec:appendix:interpretation}
@@
 
-The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ is given in the Fig.~\ref{fig:base_cases}.
+The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ are given in \autoref{fig:base_cases}.
 Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.
 
-\subsection{DPOI rewriting}
+\section{DPOI rewriting}
 \label{sec:appendix:dpoi}
 
-Consider an example of DPOI rewriting in Fig.~\ref{fig:dpoi-example} as defined in Definition~\ref{def:dpoi-e}.
+Consider an example of DPOI rewriting in \autoref{fig:dpoi-example} as defined in Definition~\ref{def:dpoi-e}.
 The rule in the top half of the diagram is an instance of a $\beta$-reduction rule as defined for string diagrams~\cite{ghica2024stringdiagramslambdacalculifunctional}.
@@
 Note how we removed strict internal interfaces of $\mathcal{L}$ from internal interfaces of $\mathcal{G}$, yielding $[u_1,v_1,v_2]$ and similarly for output interfaces.
-After computing the pushout for the second half of the diagram we add strict internal interfaces of $\mathcal{R}$ but since there are none, we are done as per condition (2) of Definition~\ref{def:dpoi-e}.
+After computing the pushout for the second half of the diagram we add strict internal interfaces of $\mathcal{R}$ but since there are none, we are done as per condition (2) of \autoref{def:dpoi-e}.
 
 \begin{figure}
+    \ifdefined\ONECOLUMN
+    \[
+    \adjustbox{width=0.6\linewidth}{
+    \tikzfig{./figures/dpo_example}
+    }
+    \]
+    \else
     \[
@@
 \]
+    \fi
     \caption{$\beta$-rule DPOI example.}

### 2025-07-10 POPL2026/paper/sections/combinatorial_semantics.tex
fixed figures, updated contributions

@@
 	$\textcolor{e-color}{E}$ and $\textcolor{closed-color}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$.
-	When writing $x \textcolor{closed-color}{<} y$ (or $x \textcolor{e-color}{<} y$) we may refer to $x$ as a predecessor of $y$ and refer to $y$ as a successor of $x$.
+	When writing $x \textcolor{closed-color}{<} y$ (or $x \textcolor{e-color}{<} y$) we may refer to $x$ as the predecessor of $y$ and refer to $y$ as a successor of $x$.
 	% we do not use this

### 2025-07-10 POPL2026/paper/sections/introduction.tex
fixed figures, updated contributions

@@
 We call these string diagrams, or more precisely, the combinatorial representation of them, \emph{closed e-hypergraphs}.
-We equip closed e-hypergraphs with a suitable notion of DPO rewriting, and show that this notion of rewriting is sound and complete with respect to rewriting of morphism terms.
+First, we define closed enriched $\Sigma$-terms which generate a theory with equivalences and bindings.
+Then we define what e-hypergraphs are and provide the interpretation of closed enriched $\Sigma$-terms as cospans of e-hypergraphs (and by doing this we also formalise the corresponding string diagrams as concrete combinatorial objects).
+Next we formulate the DPO rewriting for e-hypergraphs and show that this notion of rewriting is sound and complete with respect to rewriting of closed enriched $\Sigma$-terms modulo SMC equations.
+We finish the paper with a description of what benefits the e-hypergraph representation may bring for equality saturation in the context of $\lambda$-calculus with explicit substitution compared to conventional (slotted) e-graphs.
 

### 2025-07-10 POPL2026/paper/sections/soundness.tex
mass format
fixed figures, updated contributions

--- 2025-07-09 POPL2026/paper/sections/soundness.tex
@@
 First we recall the rewriting of $\Sigma$-terms modulo SMC laws.
+
+\begin{definition}[(Closed) $\Sigma$-term rewrite rule]
+A $\Sigma$-term rewrite rules is a pair $\langle l, r \rangle$ where $l$ and $r$ are $\Sigma$ terms with matching inputs and outputs (as lists of input and output types).
+When $l$ and $r$ are closed $\Sigma$-terms, we call such a rewrite rule closed.
+\end{definition}
 
@@
 \end{definition}
-This notion generalises to rewriting of closed $\Sigma$-terms
+This notion generalises to rewriting of closed $\Sigma$-terms.
+
 \begin{definition}
-    For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle} g$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
+	For closed-$\Sigma$-terms $f,g$, and a closed rewrite rule $\langle l, r \rangle$, $f \leadsto_{\langle l, r \rangle} g$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as
 	$
@@
 \begin{definition}
-    For closed $\Sigma^{+}$-term $f,g$, $f \leadsto_{\langle l, r \rangle}$ SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
+	For closed $\Sigma^{+}$-term $f,g$, and a closed $\Sigma$-term rewrite rule $\langle l, r \rangle$ $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
 	$
@@
 \end{definition}
+Note that the rewrite rule consist of closed $\Sigma$-terms and not closed $\Sigma^{+}$-terms, i.e. neither $l$ nor $r$ contain any $+$.
+This is intentional as there rewrite rules are supposed to be domain-specific (like a rule $\langle 2 \otimes id;times, id \otimes 1; <\;\;< \rangle$) and not related to the free enrichment.
+
 Next, to formulate the correspondence between rewriting systems we first define interpretation $\llbracket - \rrbracket$ of closed $\Sigma^{+}$ terms in $\WellTypedMdaEcospans$.
@@
 
+Practically, this quotienting means that we perform DPO rewrites modulo the equations of SMC and modulo the law~\ref{law:distributivity}.
+We can either rewrite the normal forms as mentioned above that contain all possible redexes, or we can genuinely search for redexes modulo these laws by traversing the e-hypergraph through both types of boxes when searching for a redex, which makes this representation practically viable even though it does not absorb all the possible equations.
+
 \begin{figure}

### 2025-07-10 LICS/sections/introduction.tex
fixed lines 450-452

@@
 The morphisms of the latter can also be depicted using \emph{string diagrams}~\cite{Selinger_2010,joyal_geometry_1991, mellies_functorial_2006}: a two-dimensional syntax that encodes equivalence classes of morphism terms.
-Enriched categories generalise plain categories by imposing more structure making the categorical language more expressive~\cite{Borceux_1994,Kelly2022BASICCO} which will be useful for modelling equivalence between terms as encountered in e-graphs.
-The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022,bonchi_string_2022-1,bonchi_string_2022-2,ghica_rewriting_2023,ghica_hierarchical_2023}.
+Enriched categories~\cite{Borceux_1994,Kelly2022BASICCO} generalise plain categories by imposing more structure making the categorical language more expressive~\cite{Borceux_1994,Kelly2022BASICCO} which will be useful for modelling equivalence between terms as encountered in e-graphs.
+The approach we take is guided by the established methodology of creating a correspondence between string diagrams and graph rewriting, for which there are existing precedents in the literature~\cite{bonchi_string_2022-1,bonchi_string_2022-2,bonchi_string_2022-3,fscd,ghica-zanassi2023string}.
 The key ingredient of our categorical approach is the semilattice-enrichment.
@@
 % E-graphs
-Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}.
+Although e-graphs were first developed in the 80s \cite{nelson1980techniques}, there has recently been an explosion of interest in them for the purpose of building program optimisers and synthesisers, especially due to recent work on the equality saturation technique \cite{10.1145/1594834.1480915, flatt2022small, EggPaper,flatt_small_2022}.
 This includes interest in various extensions of the e-graph formalism, for example to account for conditional rewriting \cite{singher2023colored},  to reason about rewriting for the $\lambda$-calculus \cite{koehler2022sketchguided},  and to combine techniques of e-graphs and database queries (``relational e-matching'') \cite{zhang_relational_2022}.

### 2025-07-10 POPL2026/paper/sections/categorical_semantics.tex
fixed lines 450-452

@@
 	plus the rules for equivalence of terms up to renaming of slots.
-	Recalling that we use $+$ to encode this congruence of terms, the first three rules (plus reflexivity, symmetry, and transitivity) correspond to idempotence, commutativity and associativity of $+$ and the fact that $+$ respect the composition $(;)$.
+	The earlier work of~\citet{ghica2024equivalencehypergraphsegraphsmonoidal} that introduced free semilattice enrichment to describe e-graphs using morphism terms used the $+$ of a semilattice to denote the equivalence of morphism subterms.
+	For example, the e-graph (b) from \autoref{fig:e-graph-example} can be written as a term as 
+	\[(a \otimes ((2;\comonoid) \otimes (1;id));\sym);((* \otimes \counit + (id \otimes \counit \otimes id);<\!\!<) \otimes id);/~,\] where $+$ encodes the equivalence of $*$ and $<\!\!<$
+	We can see that the first three rules (plus reflexivity, symmetry, and transitivity) above correspond to idempotence, commutativity and associativity of $+$ and the fact that $+$ respect the composition $(;)$.
 	%  and tensoring $\otimes$ (traditional e-graphs do not have tensoring as they are defined for algebraic signatures).
-	While the other rule is \textit{natural} to enforce when developing e-graphs with binding, we see that it is indeed the \textit{right} rule to enforce as we can see it follow by a purely categorical argument by the free enrichment as per the law \ref{law:distributivity}.
+	Then, while the binding congruence rule is \textit{natural} to enforce when developing e-graphs with binding, we see that it is indeed the \textit{right} rule to enforce as we can see it follow by a purely categorical argument by the free enrichment as per the law \ref{law:distributivity}.
 	We do not enforce any rules for renaming as string diagrams are essentially a nameless representation of terms.
-	Moreover, some renaming rules are not consistent with string diagrams, as slotted e-graph try to represent, \textit{e.g.}, terms $(a + b)$ and $(c + d)$ as a single entity (up to renaming of free variables), which is not the case for string diagrams (unless we do closure conversion).
+	Moreover, some renaming rules are not consistent with string diagrams, as slotted e-graph try to represent, \textit{e.g.}, terms $(a + b)$ and $(c + d)$ as a single entity (up to renaming of free variables), which in the case for string diagrams require an additional step of closure conversion.
 

### 2025-07-10 POPL2026/paper/sections/introduction.tex
fixed lines 450-452

@@
 
-E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using the technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, griggio_proceedings_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint (or timeout) is reached.
+E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using the technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, flatt2022_small, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint (or timeout) is reached.
 The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
@@
 		\adjustbox{scale=0.6}{
-			\tikzfig{figures/e-graph-example}
+			\tikzfig{figures/e-graph-example-2}
 		}

### 2025-07-10 slides/LICS_2025/poster/poster.tex
fixed lines 450-452

%!LW recipe=lualatex
% Unofficial University of Cambridge Poster Template
% https://github.com/andiac/gemini-cam
% a fork of https://github.com/anishathalye/gemini
% also refer to https://github.com/k4rtik/uchicago-poster

\documentclass[final]{beamer}

% ====================
% Packages
% ====================

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[size=custom,width=84.1,height=59.4,scale=1.0]{beamerposter}
\usetheme{gemini}
\usecolortheme{cam}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[numbers]{natbib}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usepackage{anyfontsize}
% \usepackage{biblatex}
% ====================
% Lengths
% ====================

% If you have N columns, choose \sepwidth and \colwidth such that
% (N+1)*\sepwidth + N*\colwidth = \paperwidth
\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.025\paperwidth}
\setlength{\colwidth}{0.3\paperwidth}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

% ====================
% Title
% ====================

\title{Equivalence Hypergraphs: DPO Rewriting for Monoidal E-Graphs}

\author{Aleksei Tiurin \inst{1} \and Chris Barrett \inst{1} \and Dan Ghica \inst{1} \inst{2} \and Nick Hu \inst{1}}

\institute[shortinst]{\inst{1} University of Birmigham \samelineand \inst{2} Huawei Central Software Institute}

% ====================
% Footer (optional)
% ====================

\footercontent{
  University of Birmingham \hfill
  LICS 2025, Singapore\hfill
  \href{mailto:axt257@student.bham.ac.uk}{axt257@student.bham.ac.uk}}
% (can be left out to remove footer)

% ====================
% Logo (optional)
% ====================

% use this to include logos on the left and/or right side of the header:
% \logoright{\includegraphics[height=7cm]{logo1.pdf}}
% \logoleft{\includegraphics[height=7cm]{logo2.pdf}}

% ====================
% Body
% ====================

\begin{document}

% Refer to https://github.com/k4rtik/uchicago-poster
% logo: https://www.cam.ac.uk/brand-resources/about-the-logo/logo-downloads
\addtobeamertemplate{headline}{}
{
    \begin{tikzpicture}[remember picture,overlay]
      \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=1.0cm]current page.north west)
      {\includegraphics[height=4.5cm]{../figures/birmingham_logo.png}}; 
    \end{tikzpicture}
}

\begin{frame}[t]
\begin{columns}[t]
\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{E-graphs}

    Data structure to represent an equivalence relation on a set of terms over \alert{algebraic} signature.
    \begin{itemize}
      \item $\Sigma_{\mathbf{Mon}} = \{\mathbb{S},\mu : 2 \to 1, i: 0 \to 1\}$
      \item $i$, $\mu(s_1,s_2), s_i \in \mathbb{S}$, $\ldots$
    \end{itemize}
    The relation may be refined by applying equations to a given e-graph in non-destructive fashion to avoid \textit{phase ordering} problem.
    \begin{itemize}
      \item $\mu(x,i) \equiv x$, $\mu(x,y) \equiv \mu(y,x)$, $\ldots$
    \end{itemize}

  \end{block}

  \begin{alertblock}{What we do}

    We provide \alert{categorical interpretation} of e-graphs in terms of Cartesian symmetric monoidal categories freely enriched over a category of semilattices

    \begin{itemize}
      \item Provides a new perspective on e-graph operations
      \item Opens up new avenues towards applications by freely enriching different symmetric monoidal categories
    \end{itemize}
  \end{alertblock}

  \begin{block}{A block containing a list}

    Nam vulputate nunc felis, non condimentum lacus porta ultrices. Nullam sed
    sagittis metus. Etiam consectetur gravida urna quis suscipit.

    \begin{itemize}
      \item \textbf{Mauris tempor} risus nulla, sed ornare
      \item \textbf{Libero tincidunt} a duis congue vitae
      \item \textbf{Dui ac pretium} morbi justo neque, ullamcorper
    \end{itemize}

    Eget augue porta, bibendum venenatis tortor.

  \end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{A block containing an enumerated list}

    Vivamus congue volutpat elit non semper. Praesent molestie nec erat ac
    interdum. In quis suscipit erat. \textbf{Phasellus mauris felis, molestie
    ac pharetra quis}, tempus nec ante. Donec finibus ante vel purus mollis
    fermentum. Sed felis mi, pharetra eget nibh a, feugiat eleifend dolor. Nam
    mollis condimentum purus quis sodales. Nullam eu felis eu nulla eleifend
    bibendum nec eu lorem. Vivamus felis velit, volutpat ut facilisis ac,
    commodo in metus.

    \begin{enumerate}
      \item \textbf{Morbi mauris purus}, egestas at vehicula et, convallis
        accumsan orci. Orci varius natoque penatibus et magnis dis parturient
        montes, nascetur ridiculus mus.
      \item \textbf{Cras vehicula blandit urna ut maximus}. Aliquam blandit nec
        massa ac sollicitudin. Curabitur cursus, metus nec imperdiet bibendum,
        velit lectus faucibus dolor, quis gravida metus mauris gravida turpis.
      \item \textbf{Vestibulum et massa diam}. Phasellus fermentum augue non
        nulla accumsan, non rhoncus lectus condimentum.
    \end{enumerate}

  \end{block}

  \begin{block}{Fusce aliquam magna velit}

    Et rutrum ex euismod vel. Pellentesque ultricies, velit in fermentum
    vestibulum, lectus nisi pretium nibh, sit amet aliquam lectus augue vel
    velit. Suspendisse rhoncus massa porttitor augue feugiat molestie. Sed
    molestie ut orci nec malesuada. Sed ultricies feugiat est fringilla
    posuere.

\vspace{1em}

\begin{columns}
\begin{column}{0.4\textwidth}
\begin{center}
      \begin{figure}
      \begin{tikzpicture}
        \begin{axis}[
            scale only axis,
            no markers,
            domain=0:2*pi,
            samples=100,
            axis lines=center,
            axis line style={-},
            ticks=none]
          \addplot[red] {sin(deg(x))};
          \addplot[blue] {cos(deg(x))};
        \end{axis}
      \end{tikzpicture}
      \caption{Another figure caption.}
    \end{figure}
   \end{center}
\end{column}
\begin{column}{0.6\textwidth}  %%<--- here
\justify
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam vel dapibus erat. Morbi quis leo congue, lobortis augue bibendum, malesuada neque. Duis ullamcorper quis orci sed consequat. Nam pellentesque ullamcorper tempor. Duis eget nulla blandit, vulputate orci vitae, ullamcorper ligula. Mauris a urna ac massa dignissim scelerisque sed et augue. Donec eget urna vitae neque elementum pellentesque et eget enim. Praesent a fermentum nibh. Nullam eu nibh neque. 
\end{column}
\end{columns}

  \end{block}

  \begin{block}{Nam cursus consequat egestas}

    Nulla eget sem quam. Ut aliquam volutpat nisi vestibulum convallis. Nunc a
    lectus et eros facilisis hendrerit eu non urna. Interdum et malesuada fames
    ac ante \textit{ipsum primis} in faucibus. Etiam sit amet velit eget sem
    euismod tristique. Praesent enim erat, porta vel mattis sed, pharetra sed
    ipsum. Morbi commodo condimentum massa, \textit{tempus venenatis} massa
    hendrerit quis. Maecenas sed porta est. Praesent mollis interdum lectus,
    sit amet sollicitudin risus tincidunt non.

    Etiam sit amet tempus lorem, aliquet condimentum velit. Donec et nibh
    consequat, sagittis ex eget, dictum orci. Etiam quis semper ante. Ut eu
    mauris purus. Proin nec consectetur ligula. Mauris pretium molestie
    ullamcorper. Integer nisi neque, aliquet et odio non, sagittis porta justo.

    \begin{itemize}
      \item \textbf{Sed consequat} id ante vel efficitur. Praesent congue massa
        sed est scelerisque, elementum mollis augue iaculis.
        \begin{itemize}
          \item In sed est finibus, vulputate
            nunc gravida, pulvinar lorem. In maximus nunc dolor, sed auctor eros
            porttitor quis.
          \item Fusce ornare dignissim nisi. Nam sit amet risus vel lacus
            tempor tincidunt eu a arcu.
          \item Donec rhoncus vestibulum erat, quis aliquam leo
            gravida egestas.
        \end{itemize}
      \item \textbf{Sed luctus, elit sit amet} dictum maximus, diam dolor
        faucibus purus, sed lobortis justo erat id turpis.
      \item \textbf{Pellentesque facilisis dolor in leo} bibendum congue.
        Maecenas congue finibus justo, vitae eleifend urna facilisis at.
    \end{itemize}

  \end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{exampleblock}{A highlighted block containing some math}

    A different kind of highlighted block.

    $$
    \int_{-\infty}^{\infty} e^{-x^2}\,dx = \sqrt{\pi}
    $$

    Interdum et malesuada fames $\{1, 4, 9, \ldots\}$ ac ante ipsum primis in
    faucibus. Cras eleifend dolor eu nulla suscipit suscipit. Sed lobortis non
    felis id vulputate.

    \heading{A heading inside a block}

    Praesent consectetur mi $x^2 + y^2$ metus, nec vestibulum justo viverra
    nec. Proin eget nulla pretium, egestas magna aliquam, mollis neque. Vivamus
    dictum $\mathbf{u}^\intercal\mathbf{v}$ sagittis odio, vel porta erat
    congue sed. Maecenas ut dolor quis arcu auctor porttitor.

    \heading{Another heading inside a block}

    Sed augue erat, scelerisque a purus ultricies, placerat porttitor neque.
    Donec $P(y \mid x)$ fermentum consectetur $\nabla_x P(y \mid x)$ sapien
    sagittis egestas. Duis eget leo euismod nunc viverra imperdiet nec id
    justo.

  \end{exampleblock}

  \begin{block}{Nullam vel erat at velit convallis laoreet}

    Class aptent taciti sociosqu ad litora torquent per conubia nostra, per
    inceptos himenaeos. Phasellus libero enim, gravida sed erat sit amet,
    scelerisque congue diam. Fusce dapibus dui ut augue pulvinar iaculis.

    \begin{table}
      \centering
      \begin{tabular}{l r r c}
        \toprule
        \textbf{First column} & \textbf{Second column} & \textbf{Third column} & \textbf{Fourth} \\
        \midrule
        Foo & 13.37 & 384,394 & $\alpha$ \\
        Bar & 2.17 & 1,392 & $\beta$ \\
        Baz & 3.14 & 83,742 & $\delta$ \\
        Qux & 7.59 & 974 & $\gamma$ \\
        \bottomrule
      \end{tabular}
      \caption{A table caption.}
    \end{table}

    Donec quis posuere ligula. Nunc feugiat elit a mi malesuada consequat. Sed
    imperdiet augue ac nibh aliquet tristique. Aenean eu tortor vulputate,
    eleifend lorem in, dictum urna. Proin auctor ante in augue tincidunt
    tempor. Proin pellentesque vulputate odio, ac gravida nulla posuere
    efficitur. Aenean at velit vel dolor blandit molestie. Mauris laoreet
    commodo quam, non luctus nibh ullamcorper in. Class aptent taciti sociosqu
    ad litora torquent per conubia nostra, per inceptos himenaeos.

    Nulla varius finibus volutpat. Mauris molestie lorem tincidunt, iaculis
    libero at, gravida ante. Phasellus at felis eu neque suscipit suscipit.
    Integer ullamcorper, dui nec pretium ornare, urna dolor consequat libero,
    in feugiat elit lorem euismod lacus. Pellentesque sit amet dolor mollis,
    auctor urna non, tempus sem.

  \end{block}

  \begin{block}{References}

    \nocite{*}
    \footnotesize{\bibliographystyle{plainnat}\bibliography{poster}}

  \end{block}

\end{column}

\separatorcolumn
\end{columns}
\end{frame}

\end{document}

### 2025-07-10 LICS_2025/e_graphs_with_bindings/sections/introduction.tex
Fix bibliography

@@
 For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics for this purpose.
-For instance, hypergraphs admit the property that two string diagrams are equal if and only if they ``look the same'' (\emph{i.e.} they admit a graph isomorphism); secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.
+For instance, hypergraphs admit the property that two string diagrams are equal if and only if they ``look the same'' (\emph{i.e.} they admit a graph isomorphism); secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting using the technique of double-pushout rewriting (DPO)~\cite{bonchi_string_2022-1}.
 

### 2025-07-10 POPL2026/paper/sections/introduction.tex
Fix bibliography

@@
 For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics.
-For instance, hypergraph isomorphism captures when string diagrams \enquote{look the same} --- topologically, that one may be deformed continuously into the other without obstruction\footnote{This is the informal notion of the \enquote{isotopy of string diagrams}.}; secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting over these hypergraphs using the technique of double-pushout rewriting (DPO)~\cite{1011453502719}.
+For instance, hypergraph isomorphism captures when string diagrams \enquote{look the same} --- topologically, that one may be deformed continuously into the other without obstruction\footnote{This is the informal notion of the \enquote{isotopy of string diagrams}.}; secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting over these hypergraphs using the technique of double-pushout rewriting (DPO)~\cite{bonchi_string_2022-1}.
 

### 2025-07-10 LICS_2025/e_graphs_with_bindings/sections/appendix.tex
one more bib fix

@@
 
-    This construction is analogous to how a pushout was constructed in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and we refer the reader for more detailed proofs there.
+    This construction is analogous to how a pushout was constructed in~\cite{tiurin2025equivalencehypergraphsdporewriting} and we refer the reader for more detailed proofs there.
     The only addition is $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$ that is constructed analogously to $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
@@
 \begin{figure}
+    \ifdefined\ONECOLUMN
+    \[
+    \adjustbox{width=0.6\linewidth}{
+    \tikzfig{./figures/dpo_example}
+    }
+    \]
+    \else
     \[

### 2025-07-10 LICS_2025/e_graphs_with_bindings/sections/categorical_semantics.tex
one more bib fix

@@
 
-To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator \textit{join} $+$ as presented in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator \textit{join} $+$ as presented in~\cite{tiurin2025equivalencehypergraphsdporewriting}.
 This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix~\ref{sec:appendix:slat}).
@@
 It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting and in interpreting $\lambda$-terms.
-Figure \ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
+Fig.
+\ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{tiurin2025equivalencehypergraphsdporewriting}; the first component denotes an empty diagram.
 Note the orientation of string diagrams are now from top to bottom as opposed to what was used in~\nameref{sec:introduction}: the latter was mimicking the orientation of e-graphs, \textit{i.e.}, traditional orientation for expression trees.
-% Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
+% Fig. \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
 % The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.

### 2025-07-10 LICS_2025/e_graphs_with_bindings/sections/combinatorial_semantics.tex
one more bib fix

@@
 \section{Closed E-hypergraphs}
-Prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
+Prior work of~\cite{tiurin2025equivalencehypergraphsdporewriting} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
 The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established body of literature on hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
 The first work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
-Notably, the laws of a semilattice enrichment in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
+Notably, the laws of a semilattice enrichment in~\cite{tiurin2025equivalencehypergraphsdporewriting} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
 
-We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness, and when evoking a term `e-hypergraph' we will mean the structure defined as follows.
+We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{tiurin2025equivalencehypergraphsdporewriting}.
+We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{tiurin2025equivalencehypergraphsdporewriting} to accommodate new structure brought by closedness, and when evoking a term `e-hypergraph' we will mean the structure defined as follows.
 
@@
 In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
-$\textcolor{closed-color}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{e-color}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+$\textcolor{closed-color}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{e-color}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{tiurin2025equivalencehypergraphsdporewriting}.
 
@@
 Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments, defined in this section.
-The construction in this section closely follows the one introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} with some adjustments to account for $\textcolor{closed-color}{<}$.
+The construction in this section closely follows the one introduced in~\cite{tiurin2025equivalencehypergraphsdporewriting} with some adjustments to account for $\textcolor{closed-color}{<}$.
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites. 

### 2025-07-10 LICS_2025/e_graphs_with_bindings/sections/introduction.tex
one more bib fix

@@
 All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
-\cite{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
+\cite{tiurin2025equivalencehypergraphsdporewriting}~shows how e-graphs (without binding) can be given a categorical semantics using semilattice-enriched symmetric monoidal categories, which allows e-graphs to be reconstructed as string diagrams.
 In this paper, we extend this technique to e-graphs with bindings, lifting to monoidal \emph{closed} categories.
@@
 The general background literature for string diagrams is summarised in~\cite{piedeleu2023introductionstringdiagramscomputer}.
-The particular context of string diagrams with bindings is given in~\cite{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+The particular context of string diagrams with bindings is given in~\cite{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\cite{tiurin2025equivalencehypergraphsdporewriting}.
 
@@
 \]
-\caption{E-graph example (top) and its equivalent string diagram representation (bottom)~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}}
+\caption{E-graph example (top) and its equivalent string diagram representation (bottom)~\cite{tiurin2025equivalencehypergraphsdporewriting}}
 \label{fig:e-graph-example}
@@
 
-Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
+Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the approach in~\cite{tiurin2025equivalencehypergraphsdporewriting}.
 Note that for the initial term, $(a*2)/2$ (subfigure (a)), which is already represented efficiently in the e-graph by sharing the node `2', the string diagram version shares this node explicitly, using a contraction node.

### 2025-07-10 POPL2026/paper/sections/categorical_semantics.tex
one more bib fix

@@
 Fig.
-\ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
+\ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{tiurin2025equivalencehypergraphsdporewriting}; the first component denotes an empty diagram.
 
@@
 	plus the rules for equivalence of terms up to renaming of slots.
-	The earlier work of~\citet{ghica2024equivalencehypergraphsegraphsmonoidal} that introduced free semilattice enrichment to describe e-graphs using morphism terms used the $+$ of a semilattice to denote the equivalence of morphism subterms.
+	The earlier work of~\citet{tiurin2025equivalencehypergraphsdporewriting} that introduced free semilattice enrichment to describe e-graphs using morphism terms used the $+$ of a semilattice to denote the equivalence of morphism subterms.
 	For example, the e-graph (b) from \autoref{fig:e-graph-example} can be written as a term as 

### 2025-07-10 POPL2026/paper/sections/combinatorial_semantics.tex
one more bib fix

@@
 \label{sec:closed-e-hyp}
-Prior work of~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
+Prior work of~\cite{tiurin2025equivalencehypergraphsdporewriting} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
 The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established body of literature on hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
 The first work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
-Notably, the laws of a semilattice enrichment in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
+Notably, the laws of a semilattice enrichment in~\cite{tiurin2025equivalencehypergraphsdporewriting} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
 
-We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness, and when evoking a term `e-hypergraph' we will mean the structure defined as follows.
+We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{tiurin2025equivalencehypergraphsdporewriting}.
+We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{tiurin2025equivalencehypergraphsdporewriting} to accommodate new structure brought by closedness, and when evoking a term `e-hypergraph' we will mean the structure defined as follows.
 
@@
 Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments, defined in this section.
-The construction in this section closely follows the one introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} with some adjustments to account for $\textcolor{closed-color}{<}$.
+The construction in this section closely follows the one introduced in~\cite{tiurin2025equivalencehypergraphsdporewriting} with some adjustments to account for $\textcolor{closed-color}{<}$.
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites.

### 2025-07-10 POPL2026/paper/sections/introduction.tex
one more bib fix

@@
 All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
-\citet{ghica2024equivalencehypergraphsegraphsmonoidal}~shows how e-graphs (without binding) can be given a categorical semantics using symmetric monoidal semilattice-enriched categories, which allows e-graphs to be reconstructed as string diagrams.
+\citet{tiurin2025equivalencehypergraphsdporewriting}~shows how e-graphs (without binding) can be given a categorical semantics using symmetric monoidal semilattice-enriched categories, which allows e-graphs to be reconstructed as string diagrams.
 In this paper, we extend this technique to e-graphs with bindings, lifting to monoidal \emph{closed} categories.
@@
 The general background literature for string diagrams and their relationship to formal languages is summarised by~\citet{piedeleu2023introductionstringdiagramscomputer}.
-The particular context of string diagrams with bindings is given in~\citet{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}.
+The particular context of string diagrams with bindings is given in~\citet{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\citet{tiurin2025equivalencehypergraphsdporewriting}.
 
@@
 	\]
-	\caption{E-graph example (top) and its equivalent string diagram representation (bottom)~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}}
+	\caption{E-graph example (top) and its equivalent string diagram representation (bottom)~\cite{tiurin2025equivalencehypergraphsdporewriting}}
 	\label{fig:e-graph-example}
@@
 
-Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the \emph{e-hypergraphs} of~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}.
+Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the \emph{e-hypergraphs} of~\citet{tiurin2025equivalencehypergraphsdporewriting}.
 For the initial term, $(a*2)/2$, which is already represented efficiently in the e-graph by sharing the node $2$, the string diagram version shares this node explicitly, using a copy node.
@@
 
-\citet{ghica2024equivalencehypergraphsegraphsmonoidal} shows that e-graphs over signature $\Sigma$ can be represented as morphisms of a free cartesian symmetric monoidal semilattice-enriched category over the same signature.
+\citet{tiurin2025equivalencehypergraphsdporewriting} shows that e-graphs over signature $\Sigma$ can be represented as morphisms of a free cartesian symmetric monoidal semilattice-enriched category over the same signature.
 We argue that restricting the categorical domain to free \textbf{closed} symmetric monoidal semilattice-enriched categories gives rise to structures that naturally support binding and equivalence classes of morphisms.

### 2025-07-10 closed_monoidal/closed_monoidal.tex
one more bib fix

@@
 
-We can extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{ghica2024equivalencehypergraphsegraphsmonoidal} to accommodate new structure brought by closedness.
+We can extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{tiurin2025equivalencehypergraphsdporewriting} to accommodate new structure brought by closedness.
 

### 2025-07-10 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
mass format
lots of polish

@@
-\documentclass[acmsmall,screen, nonacm, review, anonymous, authoryear]{acmart}
+\documentclass[acmsmall,screen, nonacm, review, anonymous]{acmart}
 
@@
 \newcommand\MdaCospans{\textbf{MHypI}(\Sigma)}
-\newcommand{\Ecospans}{{\catname{EHypI(\Sigma)}}}
-\newcommand{\MdaEcospans}{{\catname{MEHypI(\Sigma)}}}
-\newcommand{\WellTypedMdaEcospans}{{\catname{MEHypI(\Sigma)}}}
+\newcommand{\Ecospans}{{\catname{CEHypI(\Sigma)}}}
+\newcommand{\MdaEcospans}{{\catname{MCEHypI(\Sigma)}}}
+\newcommand{\WellTypedMdaEcospans}{{\catname{MCEHypI(\Sigma)}}}
 % conflict
@@
 \section{Conclusion}
-In this work we introduced e-graphs with bindings formalised as morphisms of a closed $\catname{SLat}$-SMC.
+In this work we introduced e-graphs with bindings formalised as morphisms of a closed symmetric monoidal $\catname{SLat}$-category.
 We then provided a combinatorial representation for such morphisms in terms of hierarchical hypergraphs (e-hypergraphs) with a suitable notion of sound and complete DPO rewriting, making it possible for an e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
@@
 
-\bibliographystyle{acm}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{../../bibliography}

### 2025-07-10 POPL2026/paper/sections/appendix.tex
lots of polish

@@
 	\begin{align*}
-      \catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) &\cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B))\\
-                                                         &\cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B))\\
-                                                         &\cong \catname{Set}(A, U(B))
+		\catname{SLat}(\coprod_{A}(I_{\catname{SLat}}), B) & \cong \prod_{A}(\catname{SLat}(I_{\catname{SLat}}, B))       \\
+		                                                   & \cong \catname{Set}(A,\catname{SLat}(I_{\catname{SLat}}, B)) \\
+		                                                   & \cong \catname{Set}(A, U(B))
 	\end{align*}
@@
 	\begin{align*}
-      F(A) \otimes F(B) &\cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}})\\
-            &\cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})\\
-            &\cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}}))\\
-            &\cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})\\
-            &\cong \coprod_{A \times B} I_{\catname{SLat}}\\
-            &\cong F(A \times B)
+		F(A) \otimes F(B) & \cong (\coprod_{A} I_{\catname{SLat}}) \otimes (\coprod_{B} I_{\catname{SLat}}) \\
+		                  & \cong \coprod_{A} (I_{\catname{SLat}} \otimes \coprod_{B} I_{\catname{SLat}})   \\
+		                  & \cong \coprod_{A} (\coprod_{B} (I_{\catname{SLat}} \otimes I_{\catname{SLat}})) \\
+		                  & \cong \coprod_{A} (\coprod_{B} I_{\catname{SLat}})                              \\
+		                  & \cong \coprod_{A \times B} I_{\catname{SLat}}                                   \\
+		                  & \cong F(A \times B)
 	\end{align*}
@@
 	\begin{align*}
-    S_{V} &= \{
+		S_{V} & = \{
 		\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |                                               \\
-          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
-          \}\\
-          &\cup\\
-          &\{
+		      & \;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z) \\
+		      & \;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
+		\}                                                                                                       \\
+		      & \cup                                                                                             \\
+		      & \{
 		\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |                                               \\
-          &\;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
-          &\;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
-          \}\\
-          &\cup\\
-          &\{
+		      & \;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z) \\
+		      & \;\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }
+		\}                                                                                                       \\
+		      & \cup                                                                                             \\
+		      & \{
 		\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}
-          \}\\
+		\}                                                                                                       \\
 		\\
-    S_{E} &= \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
+		S_{E} & = \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
 	\end{align*}
@@
 	\begin{align*}
-        \sfrac{X + Y}{\sim (R_{V},R_{E})} = &(
-            \sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
-            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})},\\
-            &\consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},\\
-            &\textcolor{e-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
-            &\textcolor{closed-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}})    
+		\sfrac{X + Y}{\sim (R_{V},R_{E})} = & (
+		\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},                                                             \\
+		                                    & \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},                       \\
+		                                    & \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},                             \\
+		                                    & \sfrac{t_{X,Y}}{\sim (R_{V},R_{E})},                             \\
+		                                    & \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},                  \\
+		                                    & \textcolor{e-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}       \\
+		                                    & \textcolor{closed-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}})
 	\end{align*}
@@
 			      \begin{align*}
-                &\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))\\
-                &=\\
-                &[[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+				       & \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))           \\
+				       & =                                                                                               \\
+				       & [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
 			      \end{align*}
@@
 
-    This construction is analogous to how a pushout was constructed in~\cite{tiurin2025equivalencehypergraphsdporewriting} and we refer the reader for more detailed proofs there.
+	This construction is analogous to how a pushout was constructed by~\citet{tiurin2025equivalencehypergraphsdporewriting} and we refer the reader for more detailed proofs there.
 	The only addition is $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$ that is constructed analogously to $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
 \end{proof}
-The construction can be illustrated with the example in Fig.~\ref{fig:pushout_example}.
+The construction can be illustrated with the example in~\autoref{fig:pushout_example}.
 Vertices and edges that are being identified as they have a common preimage inherit the relations from the vertex or edge that has it defined either directly or via a path.
@@
 Therefore, we assign $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [\textcolor{closed-color}{<_{X+Y}}(\iota_{V_{X} + V_{Y}}(v_1))]$ and so on.
-This motivates the requirement for conditions (2), (3), and (4) of Theorem~\ref{th:existence_of_pushouts}, otherwise such inheritance would be ambiguous.
+This motivates the requirement for conditions (2), (3), and (4) of~\autoref{th:existence_of_pushouts}, otherwise such inheritance would be ambiguous.
 
@@
 
-The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ are given in \autoref{fig:base_cases}.
+The base cases for interpretation function $[-] : \textbf{SMT}(\Sigma) \to \MdaCospans$ are given in~\autoref{fig:base_cases}.
 Then it is defined inductively by letting $[f \otimes g] \Coloneqq [f] \otimes [g]$ and $[f;g] \Coloneqq [f];[g]$ and by freeness induces the corresponding functor.
 
-\section{DPOI rewriting}
+\section{EDPOI rewriting}%
 \label{sec:appendix:dpoi}
 
-Consider an example of DPOI rewriting in \autoref{fig:dpoi-example} as defined in Definition~\ref{def:dpoi-e}.
+Consider an example of EDPOI rewriting in~\autoref{fig:dpoi-example} as defined in~\autoref{def:dpoi-e}.
 The rule in the top half of the diagram is an instance of a $\beta$-reduction rule as defined for string diagrams~\cite{ghica2024stringdiagramslambdacalculifunctional}.
@@
 Note how we removed strict internal interfaces of $\mathcal{L}$ from internal interfaces of $\mathcal{G}$, yielding $[u_1,v_1,v_2]$ and similarly for output interfaces.
-After computing the pushout for the second half of the diagram we add strict internal interfaces of $\mathcal{R}$ but since there are none, we are done as per condition (2) of \autoref{def:dpoi-e}.
+After computing the pushout for the second half of the diagram we add strict internal interfaces of $\mathcal{R}$ but since there are none, we are done as per condition (2) of~\autoref{def:dpoi-e}.
 
@@
 	\fi
-    \caption{$\beta$-rule DPOI example.}
+	\caption{$\beta$-rule EDPOI example.}
 	\label{fig:dpoi-example}

### 2025-07-10 POPL2026/paper/sections/applications.tex
lots of polish

@@
 Even more so, explicit substitution becomes mandatory if we want to encode lambda calculus with explicit substitution in an e-graph.
-Below is the Table~\ref{tbl:slotted} that features the number of e-nodes when performing equality saturation for a $\beta$-reduction using a lambda term $plus 2 2$ using Church encoding of arithmetic.
+Below is the~\autoref{tbl:slotted} that features the number of e-nodes when performing equality saturation for a $\beta$-reduction using a lambda term $plus 2 2$ using Church encoding of arithmetic.
 
@@
 	\begin{tabular}{lcccc}
-      Set of equations & built-in subst & explicit subst & explicit subst & explicit subst\\
-      &&& + interchange&  + interchange\\
-      &&&&  + comm\\
+		Set of equations & built-in subst & explicit subst & explicit subst & explicit subst \\
+		                 &                &                & + interchange  & + interchange  \\
+		                 &                &                &                & + comm         \\
 		\# of e-nodes    & 40             & 161            & 289            & 356

### 2025-07-10 POPL2026/paper/sections/categorical_semantics.tex
lots of polish

--- 2025-07-10 POPL2026/paper/sections/categorical_semantics.tex
@@
-\section{Categorical semantics of E-Graphs with Bindings}
+\section{Categorical semantics of E-Graphs with Bindings}%
 \label{sec:categorical}
 
-This section introduces preliminaries on semilattice-enriched symmetric monoidal categories generated by monoidal theories, and the string diagram formalism used to represent them.
-We also show a specific case of semilattice-enriched closed monoidal categories and how the closed structure interacts with the enrichment structure.
+This section introduces preliminaries on symmetric monoidal semilattice-enriched categories generated by monoidal theories, and the string diagram formalism used to represent them.
+We also show a specific case of closed monoidal semilattice-enriched categories and how the closed structure interacts with the enrichment.
 
-Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding hom-set.
+Given a category $\mathbb{C}$  with objects $A,B \in \mathbb{C}$ we denote by $\mathbb{C}(A,B)$ the corresponding Hom set.
 We write the identity morphism on $A$ as $\id_A$.
@@
 Composition in the usual order is written $g \circ f$.
-We denote the tensor product of an SMC $\mathbb{C}$ by $\otimes$,  its unit by $I$, and its symmetry natural transformation as $\sym$ \cite{maclane}.
+We denote the tensor product of an symmetric monoidal category $\mathbb{C}$ by $\otimes$,  its unit by $I$, and its symmetry natural transformation as $\sym$ \cite{maclane}.
 We adopt the convention that $\otimes$ binds more tightly than $(;\!)$.
 We elide all associativity and unit isomorphisms associated with monoidal categories, and often omit subscripts on identities and natural transformations where it can be inferred.
-Given two adjoint functors $F : \mathbb{A} \to \mathbb{B}$ and $G : \mathbb{B} \to \mathbb{A}$ with $F$ being left-adjoint, we write $F \dashv G : \mathbb{B} \to \mathbb{A}$ for this adjunction.
+Given two adjoint functors $F\colon \mathbb{A} \to \mathbb{B}$ and $G\colon \mathbb{B} \to \mathbb{A}$ with $F$ being left-adjoint, we write $F \dashv G\colon \mathbb{B} \to \mathbb{A}$ for this adjunction.
 
@@
 \begin{definition}
-	A \textit{monoidal} \textit{signature} $\Sigma = (\Sigma_{O}, \Sigma_{M}, t)$ is given by a set of objects (called types) $\Sigma_{O}$, a set of generators (operations) $\Sigma_{M}$ and a type assignment function $t : \Sigma_{M} \to \Sigma_{O}^{*} \times \Sigma_{O}^{*}$ assigning lists of input and output types to a given operation.
+	A monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M}, t)$ is given by a set of objects (called types) $\Sigma_{O}$, a set of generators (operations) $\Sigma_{M}$ and a type assignment function $t\colon \Sigma_{M} \to \Sigma_{O}^{*} \times \Sigma_{O}^{*}$ assigning lists of input and output types to a given operation.
 \end{definition}
@@
 
-Identities and symmetries of complex types, \textit{e.g.}, $\id_{A \otimes B}$ and $\sym_{A \otimes B, C}$, are given by the composition and tensoring of $\id_{A}$ and $\sym_{A,B}$ as, $\textit{e.g.}$, $\id_{A \otimes B} = \id_{A} \otimes \id_{B}$ and $\sym_{A \otimes B, C} = (\id_{A} \otimes \sym_{B,C});(\sym_{A,C} \otimes \id_{B})$.
+Identities and symmetries of complex types, e.g.\ $\id_{A \otimes B}$ and $\sym_{A \otimes B, C}$, are given by the composition and tensoring of $\id_{A}$ and $\sym_{A,B}$ as, e.g.\ $\id_{A \otimes B} = \id_{A} \otimes \id_{B}$ and $\sym_{A \otimes B, C} = (\id_{A} \otimes \sym_{B,C});(\sym_{A,C} \otimes \id_{B})$.
 
@@
 
-We can similarly consider \textit{closed} terms that give rise to a free \textit{closed} symmetric monoidal category.
+We can similarly consider \emph{closed} terms that give rise to a free \emph{closed} symmetric monoidal category.
 
-\begin{definition}
+\begin{definition}%
 	\label{def:closed}
-	A monoidal (right) closed category is a monoidal category $\catname{C}$ satisfying that for
+	A (right) closed monoidal category is a monoidal category $\catname{C}$ satisfying that for
 	every pair of objects $B,C$ there is an object $B \multimap C$ and a morphism $\textbf{ev}_{B,C} : (B \multimap C) \otimes B \to
@@
 
-This definition is equivalent to defining a (right) closed category through a pair of adjoint functors $- \otimes B$ and $B \multimap{} -$.
-If the tensor is symmetric, then the category is said to be simply \textit{closed}.
+This is equivalent to defining a (right) closed monoidal category as one admitting an adjunction $- \otimes B \dashv B \multimap{} -$ for every object $B$.
+If the tensor is symmetric, then the monoidal category is said to be simply \emph{closed}.
 
 \begin{definition}
-	\textit{Closed} $\Sigma$-terms are defined inductively by first defining a family of objects $obj_{\Sigma_{O}}$.
+	Closed $\Sigma$-terms are defined inductively by first defining a family of objects $obj_{\Sigma_{O}}$.
 	\begin{itemize}
@@
 	\end{itemize}
-	A family of closed $\Sigma$ terms is then defined as
+	A family of closed $\Sigma$ terms is then defined by:
 	\begin{itemize}
-		\item $\phi : A \to B \in \Sigma_{M}$ is a closed $\Sigma$-term.
-		\item If $A$ and $B$ are $obj_{\Sigma_{O}}$, then $\id_{A}$ and $\sym_{A,B}$ are closed $\Sigma$-terms.
-		\item If $f : A \to B$ and $g : B \to C$ are closed $\Sigma$-terms, then so is $f;g : A \to C$.
-		\item If $f : A \to B$ and $g : C \to D$ are closed $\Sigma$-terms, then so is $f \otimes g : A \otimes C \to B \otimes D$.
-		\item If $A$ and $B$ are $obj_{\Sigma_{O}}$, then $\textbf{ev}_{A,B} : A \multimap B \otimes A \to B$ is a closed $\Sigma$-term.
-		\item If $h : X \otimes A \to B$ is a closed $\Sigma$-term, then $\Lambda_{X,A,B}(h) : X \to (A \multimap B)$ is a closed $\Sigma$-term.
+		\item $\phi : A \to B \in \Sigma_{M}$ is a closed $\Sigma$-term;
+		\item if $A$ and $B$ are $obj_{\Sigma_{O}}$, then $\id_{A}$ and $\sym_{A,B}$ are closed $\Sigma$-terms;
+		\item if $f : A \to B$ and $g : B \to C$ are closed $\Sigma$-terms, then so is $f;g : A \to C$;
+		\item if $f : A \to B$ and $g : C \to D$ are closed $\Sigma$-terms, then so is $f \otimes g : A \otimes C \to B \otimes D$;
+		\item if $A$ and $B$ are $obj_{\Sigma_{O}}$, then $\textbf{ev}_{A,B} : A \multimap B \otimes A \to B$ is a closed $\Sigma$-term;
+		\item if $h : X \otimes A \to B$ is a closed $\Sigma$-term, then $\Lambda_{X,A,B}(h) : X \to (A \multimap B)$ is a closed $\Sigma$-term.
 	\end{itemize}
@@
 \begin{definition}
-	The free closed symmetric monoidal category $\catname{CS}(\Sigma)$ is given by a set of closed $\Sigma$-terms quotiented by the laws of a symmetric monoidal category and the equations from Definition~\ref{def:closed}.
-	We can similarly consider a free closed symmetric monoidal category that is induced by an $\catname{SMT}(\Sigma, \mathcal{E})$ over closed $\Sigma$-terms which we will denote with $\catname{CS}(\Sigma, \mathcal{E})$.
+	The free closed symmetric monoidal category $\catname{CS}(\Sigma)$ is given by a set of closed $\Sigma$-terms quotiented by the laws of a symmetric monoidal category and the equations from~\autoref{def:closed}.
+	We can similarly consider a free closed symmetric monoidal category that is induced by a $\catname{SMT}(\Sigma, \mathcal{E})$ over closed $\Sigma$-terms which we will denote with $\catname{CS}(\Sigma, \mathcal{E})$.
 \end{definition}
 
-To encode equivalences between morphisms (or $\Sigma$-terms), the hom-sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator \textit{join} $+$ as presented in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}.
-This endowment turns such categories into $\catname{SLat}$-enriched categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see Appendix~\ref{sec:appendix:slat}).
+To encode equivalences between morphisms (or $\Sigma$-terms), the Hom sets of the corresponding free categories can be endowed with the additional structure of a semilattice induced by an idempotent, commutative and associative operator \textit{join} $+$ as presented by~\citet{tiurin2025equivalencehypergraphsdporewriting}.
+This is modelled by enrichment in semilattices, yielding $\catname{SLat}$-categories where $\catname{SLat}$ is a category of \textit{unbounded} semilattices and homomorphisms between them (see~\autoref{sec:appendix:slat}).
 
-\begin{definition}[Semilattice-enriched category~\cite{Borceux_1994}]
+\begin{definition}[$\catname{SLat}$-category]%
 	\label{def:slat-cat}
-	A semilattice-enriched category $\mathbb{C}$ is defined by the following data:
+	A $\catname{SLat}$-category $\mathbb{C}$ is defined by the following data:
 	\begin{itemize}
-		\item a set of objects $\mathbb{C}$
-		\item for every pair of objects $A,B \in \mathbb{C}$ --- a hom-object (or, hom-semilattice) $\mathbb{C}(A,B) \in \catname{SLat}$
+		\item a set of objects $\mathbb{C}$;
+		\item for every pair of objects $A,B \in \mathbb{C}$ --- a Hom semilattice $\mathbb{C}(A,B) \in \catname{SLat}$;
 		\item for every triple of objects $A,B,C \in \mathbb{C}$ --- a composition morphism
-		      $\circ: \mathbb{C}(B,C) \otimes \mathbb{C}(A,B) \to \mathbb{C}(A,C)$
-		\item for every object $A \in C$ --- a unit morphism
-		      $u_{A} : I \to \mathbb{C}(A,A)$, where $I$ is the monoidal unit for $\catname{SLat}$.
+		      $\circ\colon \mathbb{C}(B,C) \otimes \mathbb{C}(A,B) \to \mathbb{C}(A,C)$ of $\catname{SLat}$;
+		\item for every object $A \in \mathbb{C}$ --- a unit morphism
+		      $u_{A}\colon I \to \mathbb{C}(A,A)$, where $I$ is the monoidal unit $\{*\}$ for $\catname{SLat}$\footnote{Semilattice homomorphisms $\{*\} \to P$ are in bijection with the elements of $P$ as a set, so this is an element of the underlying set of $\mathbb{C}(A,A)$.}.
 	\end{itemize}
-	such that particular coherence diagrams commute.
+	such that particular coherence diagrams enforcing that composition is associative and unital commute.
 \end{definition}
-As the composition is a semilattice homomorphism (morphism of $\catname{SLat}$), we require that
-$
-	(f + g); h = f;h + g;h$ and $f;(g+h) = f;g + f;h
-$.
+$\catname{SLat}$ is sufficiently close to $\catname{Set}$ that Hom semilattices can be thought of as the ordinary notion of Hom sets with additional semilattice structure; unit morphisms are as usual, and composition works similarly to composition in categories with an extra condition of respecting the join: as it is a semilattice homomorphism, we require that $(f + g); h = f;h + g;h$ and $f;(g+h) = f;g + f;h$.
 
@@
 	Let $\mathbb{C}$ and $\mathbb{D}$ be two $\catname{SLat}$-categories.
-	An $\catname{SLat}$-functor $F : \mathbb{C} \to \mathbb{D}$ is defined by the following data.
+	An $\catname{SLat}$-functor $F\colon \mathbb{C} \to \mathbb{D}$ is defined by the following data:
 	\begin{itemize}
-		\item A mapping $F : \mathbb{C} \to \mathbb{D}$
-		\item An object-indexed family of morphisms in $\catname{SLat}$ $F_{A,B} : \mathbb{C}(A,B) \to \mathbb{D}(FA,FB)$
+		\item a mapping on objects $A$ of $\mathbb{C}$ to $F A$ of $\mathbb{D}$;
+		\item for each pair of objects $A, B$ of $\mathbb{C}$, a morphism in $\catname{SLat}$ $F_{A,B}\colon \mathbb{C}(A,B) \to \mathbb{D}(FA,FB)$;
 	\end{itemize}
-	such that certain coherence diagrams commute~\cite{Borceux_1994}.
-	$F$ being a morphism in $\catname{SLat}$ has a property $F(f + g) = F(f) + F(g)$.
+	such that certain coherence diagrams, enforcing the preservation of composition and units, commute.
 \end{definition}
+As before, this is like an ordinary functor, except with an additional requirement that its action on morphisms respects the join: $F(f + g) = F(f) + F(g)$.
 
-\begin{definition}[Semilattice-enriched SMC~\cite{enriched_monoidal}]\label{def:enriched-prop}
-	A \textit{semilattice enriched strict SMC}  $\mathbb{C}$ is given by a semilattice-enriched category $\mathbb{C}$ as above that additionally has
+The collection of $\catname{SLat}$-categories and $\catname{SLat}$-functors form an $\catname{SLat}$-category $\catname{SLat}--\catname{Cat}$\footnote{More generally, if we were to define $\catname{SLat}$-natural transformations, we could view $\catname{SLat}--\catname{Cat}$ as a symmetric monoidal 2-category~\cite{Kelly2022BASICCO}.}, and for any pair of $\catname{SLat}$-categories $\mathbb{C}$ and $\mathbb{D}$, the product $\catname{SLat}$-category $\mathbb{C} \times \mathbb{D}$ is defined analogously to categories.
+
+\begin{definition}[Symmetric monoidal $\catname{SLat}$-category]%
+	\label{def:enriched-prop}
+	A symmetric monoidal $\catname{SLat}$-category is given by a $\catname{SLat}$-category $\mathbb{C}$ that additionally has
 	\begin{itemize}
-		\item a unit object $I_{C} \in \mathbb{C}$
-		\item for every pair of objects $A,B \in \mathbb{C}$ --- an object $A \otimes B \in \mathbb{C}$
-		\item for all $A,B,C,D$ --- a tensor morphism $ - \otimes_{C} - : \mathbb{C}(A,C) \otimes \mathbb{C}(B,D) \to \mathbb{C}(A \otimes B, C \otimes D)$
+		\item a unit object $I \in \mathbb{C}$;
+		\item a $\catname{SLat}$-functor tensor $\otimes\colon \mathbb{C} \times \mathbb{C} \to \mathbb{C}$;
 	\end{itemize}
-	such that particular coherence diagrams commute.
+	such that particular coherence diagrams analogous to those for symmetric monoidal categories commute.
 \end{definition}
-The latter morphism being in $\catname{SLat}$ implies that for $\otimes$ we have
-$
-	f \otimes (g+h) = f \otimes g + f \otimes h $ and $ (f+g) \otimes h = f \otimes h + g \otimes h
-$.
+This is completely analogous to a symmetric monoidal category, with the fact that the tensor is an $\catname{SLat}$-functor ensuring compatibility with the join: $f \otimes (g+h) = f \otimes g + f \otimes h $ and $ (f+g) \otimes h = f \otimes h + g \otimes h$.
 As before, we take $\otimes$ to bind more tightly than $+$.
-
-Then, the axioms of $\catname{SLat}$-SMC consist of equations in Definitions~\ref{def:enriched-prop}~\ref{def:slat-cat} and associativity, commutativity, and idempotence equations for $+$.
 
@@
 	\]
-	Moreover, every symmetric monoidal category $\mathbb{C}$ in $\catname{Cat}$ (respectively, $\catname{SLat}\text{--}\catname{Cat}$) is taken to a symmetric monoidal category $\mathbb{C}'$ in $\catname{SLat}\text{--}\catname{Cat}$ (respectively, $\catname{Cat}$) by a 2-functor $\mathcal{F}$ (respectively, $\mathcal{U}$).
+	Moreover, $\mathcal{F}$ and $\mathcal{U}$ preserve symmetric monoidal structure, sending symmetric monoidal categories to symmetric monoidal $\catname{SLat}$-categories and vice versa.
 \end{proposition}
-In particular, 2-functor $\mathcal{F}$ turns every category $\mathbb{C'} \in \catname{Cat}$ into a free semilattice enriched $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every hom-set of $\mathbb{C'}$ into a free semilattice on this set.
-Being a 2-functor, it also lifts an adjunction in $\catname{Cat}$ $F \dashv G$ into an $\catname{SLat}$-adjunction.
-Thus, applying the 2-functor to $\catname{CS}(\Sigma)$ we get a closed symmetric monoidal $\catname{SLat}$-category $\catname{CS}(\Sigma)^{+}$ by lifting the adjunction $(- \otimes B, B \multimap{} -, \eta, \varepsilon)$.
-We can similarly derive $\catname{SLat}$-enriched versions of  $\Lambda_{A,B,C}$ and $\textbf{ev}_{A,B}$ morphisms using the lifted adjunction as $\textbf{ev}_{A,B} = \varepsilon_{A}$ and $\Lambda_{A,B,C}(f) = \eta; (B \multimap{} f)$.
-The following distributivity law can then be derived, forming the basis of encoding equivalences inside abstractions.
+In particular, the 2-functor $\mathcal{F}$ turns every category $\mathbb{C}_0 \in \catname{Cat}$ freely into a semilattice-enriched category $\mathbb{C} \in \catname{SLat}\text{--}\catname{Cat}$ by making every Hom set of $\mathbb{C}_0$ into the free semilattice on that set along $F$.
+Being a 2-functor, it preserves dualities (in the 2-categorical sense): it lifts an adjunction in between categories in $\catname{Cat}$ into an $\catname{SLat}$-adjunction between their free semilattice-enriched correspondants.
+We take closure, in the sense of~\autoref{def:closed}, to be defined generally for $\catname{SLat}$-categories by an analogous $\catname{SLat}$-adjunction $- \otimes B \dashv B \multimap -$ for every object $B$.
+Thus, applying the 2-functor $\mathcal{F}$ to $\catname{CS}(\Sigma)$ we get a closed symmetric monoidal $\catname{SLat}$-category $\catname{CS}(\Sigma)^{+}$ by this lifting.
+We can similarly derive $\catname{SLat}$-enriched versions of  $\Lambda_{A,B,C}$ and $\textbf{ev}_{A,B}$ morphisms using the lifted adjunction as $\textbf{ev}_{A,B} = \varepsilon_{A}$ and $\Lambda_{A,B,C}(f) = \eta_A; (B \multimap{} f)$.
+The following distributivity law can then be derived, forming the basis of encoding equivalences inside abstractions:
 \begin{equation}
 	\begin{aligned}
-		\Lambda_{A,B,C}(f + g) & = \eta;(B \multimap (f + g))                      \\
-		                       & = \eta;(B \multimap (f) + B \multimap (g))        \\
-		                       & = \eta;(B \multimap (f)) + \eta;(B \multimap (g)) \\
-		                       & = \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g)
-	\end{aligned}
+		\Lambda_{A,B,C}(f + g) & = \eta_A;(B \multimap (f + g))                    \\
+		                       & = \eta_A;(B \multimap f + B \multimap g)          \\
+		                       & = \eta_A;(B \multimap f) + \eta_A;(B \multimap g) \\
+		                       & = \Lambda_{A,B,C}(f) + \Lambda_{A,B,C}(g).
+	\end{aligned}%
+	\label{law:distributivity}
 \end{equation}
-\label{law:distributivity}
 
-Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of closed $\Sigma^{+}$-terms, the latter being defined as follows.
+Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of closed $\Sigma^{+}$-terms as follows.
 Every closed $\Sigma$-term is also a $\Sigma^{+}$-term.
-Given two closed $\Sigma^{+}$-terms $f : A \to B$ and $g : A \to B$, $f + g : A \to B$ is also a closed $\Sigma^{+}$-term.
-Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed $\catname{SLat}$-SMC, including the distributivity law~\ref{law:distributivity}.
+Given two closed $\Sigma^{+}$-terms $f\colon A \to B$ and $g\colon A \to B$, $f + g\colon A \to B$ is also a closed $\Sigma^{+}$-term.
+Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed symmetric monoidal $\catname{SLat}$-categories, including the distributivity law (\autoref{law:distributivity}).
 
-To aid reasoning, we introduce a new language of string diagrams for closed $\catname{SLat}$-SMCs, using a hierarchical ``box'' structure to capture the join operation on morphisms as well as another hierarchical ``box'' for lambda abstraction.
-It was already discussed in the introduction how this is  used in the translation of equivalence classes from the e-graph to the string diagrammatic setting and in interpreting $\lambda$-terms.
-Fig.
-\ref{fig:egraph-strings} displays the generators of this language which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for lambda abstraction as established in~\cite{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice-enrichment as introduced in~\cite{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
+To aid reasoning, we introduce a new language of string diagrams for closed symmetric monoidal $\catname{SLat}$-categories, using a hierarchical \enquote{dashed box} structure to capture the join operation on morphisms (as was done by~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}) as well as another hierarchical \enquote{rounded box} for $\lambda$-abstraction.
+\autoref{fig:egraph-strings} displays the generators of this language, which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for $\lambda$-abstraction as established by~\citet{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice enrichment as introduced by~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
 
@@
 
-% Fig. \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
-% The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
+% \autoref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard equations of symmetric monoidal categories. 
+% The first four equations are those displayed in\autoref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.
 We overload the binary notation $+$ for our $n$-ary notation.
@@
 	\captionsetup{skip=0pt, belowskip=-2ex}
-	\caption{String diagrams for closed semilattice-enriched symmetric monoidal categories.}
+	\caption{String diagrams for closed symmetric monoidal $\catname{SLat}$-categories.}
 	\label{fig:egraph-strings}
@@
 
-Slotted e-graphs~\cite{slotted-egraphs} extended the traditional signature employed by e-graphs by an explicit construct for binding ($\text{bind}\; \$x\; tc$) and variables (slots).
-	\begin{align*}
-		 & \text{functional symbols}\; f, g, \ldots                                                                   \\
-		 & \text{slots}\;  \$x, \$y, \ldots                                                                           \\
-		 & \text{term children}\; tc  \Coloneqq t \;|\; tc_{1}, \ldots, tc_{n} \;|\; \text{bind}\; \$x\; tc \;|\; \$x \\
-		 & \text{terms}\;  \Coloneqq f \;|\; f(tc)
-	\end{align*}
-
-	The traditional congruence relation maintained be e-graphs for a set of equations $E$ that are given below (plus obvious rules for transitivity, symmetry and reflexivity of $\cong$)
-
+Slotted e-graphs~\cite{slotted-egraphs} extend the traditional signature employed by e-graphs by an explicit construct for binding ($\text{bind}\; \$x\; tc$) and variables (\emph{slots}).
+	Their syntax is generated by the following grammar:
+	\begin{alignat*}{2}
+		\text{(functional symbols)} &        &                            & \; f, g, \ldots ;                                                                       \\
+		\text{(slots)}              &        &                            & \; \$x, \$y, \ldots ;                                                                   \\
+		\text{(term children)}      &        & tc               \Coloneqq & \; t \mid tc_{1}, \ldots, tc_{n} \mid \text{bind}\; \$x\; tc \mid \$x ; \qquad n \geq 1 \\
+		\text{(terms)}              & \qquad & t                \Coloneqq & \; f \mid f(tc) .
+	\end{alignat*}
+	The traditional congruence relation maintained be e-graphs for a set of equations $E$ are given below (plus the obvious rules for transitivity, symmetry and reflexivity of $\cong$):
 	\begin{mathpar}
@@
 		{E \vdash a \cong b}
+		\quad
 		\inferrule*[right=cong. (variadic)]
@@
 		{E \vdash tc_1, \ldots, tc_k \cong tc'_1, \ldots, tc'_k}
+		\quad
 		\inferrule*[right=cong. (f)]
@@
 	\end{mathpar}
-
-	\noindent
-	are extended by the rule
-
-	\begin{mathpar}
-
+	Slotted e-graphs extend these with the following:
+	\begin{equation}%
+		\label{eq:cong-bind}
 		\inferrule*[right=cong. (bind)]
@@
 		{E \vdash \text{bind}\; \$x\; tc \cong \text{bind}\; \$x\; tc'},
-	\end{mathpar}
-
-	\noindent
+	\end{equation}
 	plus the rules for equivalence of terms up to renaming of slots.
-	Recalling that we use $+$ to encode this congruence of terms, the first three rules (plus reflexivity, symmetry, and transitivity) correspond to idempotence, commutativity and associativity of $+$ and the fact that $+$ respect the composition $(;)$.
+	The earlier work of~\citet{tiurin2025equivalencehypergraphsdporewriting} that introduced free semilattice enrichment to describe e-graphs using morphism terms used the join operation $+$ of a semilattice to denote the equivalence of morphism subterms.
+	For example, the e-graph (b) from~\autoref{fig:e-graph-example} can be written as a term as
+	\[(a \otimes ((2;\comonoid) \otimes (1;id));\sym);((* \otimes \counit + (id \otimes \counit \otimes id);<\!\!<) \otimes id);/~,\] where $+$ encodes the equivalence of the subterms rooted at $*$ and $<\!\!<$
+	We can see that the first three rules (plus reflexivity, symmetry, and transitivity) above correspond to idempotence, commutativity and associativity of $+$ and the fact that $+$ respect the composition $(;)$.
 	%  and tensoring $\otimes$ (traditional e-graphs do not have tensoring as they are defined for algebraic signatures).
-	While the other rule is \textit{natural} to enforce when developing e-graphs with binding, we see that it is indeed the \textit{right} rule to enforce as we can see it follow by a purely categorical argument by the free enrichment as per the law \ref{law:distributivity}.
-	We do not enforce any rules for renaming as string diagrams are essentially a nameless representation of terms.
-	Moreover, some renaming rules are not consistent with string diagrams, as slotted e-graph try to represent, \textit{e.g.}, terms $(a + b)$ and $(c + d)$ as a single entity (up to renaming of free variables), which is not the case for string diagrams (unless we do closure conversion).
+	Moreover, we can also justify~\autoref{eq:cong-bind} from a purely categorical perspective, by the free enrichment as per~\autoref{law:distributivity}.
+	We do not require the additional rules regarding renaming as string diagrams are essentially a nameless representation of terms.
+	There is a distinction here though, as some renaming rules are not consistent with string diagrams: slotted e-graph try to represent, e.g., terms $(a + b)$ and $(c + d)$ as a single entity (up to renaming of free variables), which is not the case for string diagrams (unless we do closure conversion --- see~\autoref{fig:extended-egraphs}).
 

### 2025-07-10 POPL2026/paper/sections/combinatorial_semantics.tex
lots of polish

@@
-\section{String Diagram Rewrite Theory}\label{sec:combinatorial-semantics}
+\section{String Diagram Rewrite Theory}%
+\label{sec:combinatorial-semantics}
 
-We recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories,  recapitulating the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs,  as established in~\cite{bonchi_string_2022-1, bonchi_string_2022-2}.
+We recall the fundamental definitions and results of string diagram rewrite theory for symmetric monoidal categories, recapitulating the correspondence between string diagram rewriting and an appropriate notion of double pushout (DPO) rewriting of certain hypergraphs, as established in~\citet{bonchi_string_2022-1, bonchi_string_2022-2}.
 
-First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over \textsf{Set}.
-We extend this to the case of relations $R \subseteq V \times V$,  so that we denote by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $R$ to a relation on ordered sequences.
-We will sometimes use functional notation for relations,  so that given $R \subseteq V \times W$ and $v \in V$,  $R(v) \subseteq W$.
-We elide associativity and unit isomorphisms associated with coproducts $(+)$,  and denote by $\iota_j: X_{j} \rightarrow X_{1} + \ldots X_{j} + \ldots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]: A_1 + A_2 \to B$ the co-pairing of two morphisms $f: A_1 \to B$ and $g:A_2 \to B$.
-We denote by $A +_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
-We will also refer to $A,B$ as \textit{feet} of the cospan, and to $C$ as \textit{carrier}.
-% Given an ordered sequence of labels from some alphabet $\Sigma$, $[l_1, \ldots, l_n]$ we will denote with $w([l_1, \ldots, l_n])$ the word concatenated from these labels using $\otimes$, \textit{e.g.}, $w([A,B]) = A \otimes B$.
+First,  we fix some notation.  Let $(-)^*$ be the free monoid monad over $\catname{Set}$.
+We extend this to the case of relations $R \subseteq V \times V$,  denoting by $R^{*} \subseteq V^* \times V^*$ the element-wise extension of $R$ to a relation on ordered sequences.
+We will sometimes use functional notation for relations, so that given $R \subseteq V \times W$ and $v \in V$, we write $R(v) \in W$.
+We elide associativity and unit isomorphisms associated with coproducts $\sqcup$,  and denote by $\iota_j\colon X_{j} \rightarrow X_{1} \sqcup \dots X_{j} \sqcup \dots X_{n}$ the $j^{\text{th}}$ injection,  and $[f,g]\colon A_1 \sqcup A_2 \to B$ the co-pairing of two morphisms $f\colon A_1 \to B$ and $g\colon A_2 \to B$.
+We denote by $A \sqcup_{f,g} B$ the pushout of the span $A \xleftarrow{f} C \xrightarrow{g} B$.
+We will also refer to $A,B$ as \emph{feet} of the cospan, and to $C$ as \emph{carrier}.
+% Given an ordered sequence of labels from some alphabet $\Sigma$, $[l_1, \ldots, l_n]$ we will denote with $w([l_1, \ldots, l_n])$ the word concatenated from these labels using $\otimes$, e.g.\ $w([A,B]) = A \otimes B$.
 
@@
 
-In ~\cite{bonchi_string_2022-2}, certain class of hypergraphs \textit{with interfaces} and their DPO rewriting are shown to be sound and complete for categorical presentations of arbitrary SMTs.
-Hypergraphs generalise graphs by allowing edges to have multiple sources and targets.
-When interpreting string diagrams as hypergraphs,  morphisms are interpreted as edges,  and wires as vertices.
-However,  there is also a need to specify which vertices are to be considered as \textit{input} and \textit{outputs},  corresponding to the dangling wires at the bottom and top of a string diagram.
-This gives rise to the definition of a hypergraph with interfaces.
-% Subsequent work \cite{bonchi_string_2022-2} builds on this correspondence to give a restriction of hypergraphs with interfaces (called \textit{monogamous} directed acyclic),  leading to a correspondence between symmetric monoidal string diagrams (\textit{i.e.}, without requiring a Frobenius algebra) and the category of (cospans of) such hypergraphs.
-We then evoke the definition of hypergraphs and corresponding cospans as introduced in~\cite{bonchi_string_2022-1,bonchi_string_2022-2}.
+Hypergraphs generalise directed graphs by allowing edges to have multiple (ordered) sources and targets.
+\citet{bonchi_string_2022-2} show that a certain class of hypergraphs \emph{with interfaces} and their DPO rewriting are sound and complete for categorical presentations of arbitrary symmetric monoidal theories.
+Moreover, they develop the use of hypergraphs as combinatorial representations of string diagrams for symmetric monoidal categories, whereby morphisms (nodes) are represented as hyperedges, and objects (wires) as vertices.
+We briefly present this correspondence, but the interested reader should consult the work of~\citet{bonchi_string_2022-1,bonchi_string_2022-2}.
 
 \begin{definition}[Category of hypergraphs]\label{def:hypergraph}
-	A \emph{hypergraph $\mathcal{G}$ over monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$} is a tuple $(V,E,s,t,l)$,  where $V$ is a finite set of vertices, $E$ is a finite set of edges, $s : E \to V^{*}$ is a source function, $t : E \to V^{*}$ is a target function,  and $l = (l_{V} : V \to \Sigma_{O},l_{E} : E \to \Sigma_{M})$ is a pair of labelling functions that assigns each edge and vertex an operator and a type from monoidal signature $\Sigma$ respectively.
-	The labelling function must respect the typing of the operator: for all edges $e$, such that $l(e) = c : A \to B$,  $l_{V}^{*}(s(e)) = A$ and $l_{V}^{*}(t(e)) = B$.
-	We call a hypergraph \textit{discrete} if its set of edges is empty.
-	A \emph{hypergraph homomorphism} $\phi: \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold.
-	\begin{enumerate}
-		\item $\phi_V^*(s_{\mathcal{F}}(e)) = s_{\mathcal{G}}(\phi_E(e)) \qquad \phi_V^*(t_{\mathcal{F}}(e)) = t_{\mathcal{G}}(\phi_E(e))$
-		\item $l_{V,\mathcal{F}}(v) = l_{V,\mathcal{G}}(\phi_V(v)) \qquad\;\; l_{E,\mathcal{F}}(e) = l_{E,\mathcal{G}}(\phi_E(e))$
-	\end{enumerate}
+	Over a monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$, a hypergraph $\mathcal{G}$ is a tuple $(V,E,s,t,l)$, where $V$ is a finite set of vertices, $E$ is a finite set of hyperedges, $s\colon E \to V^{*}$ is a source function, $t\colon E \to V^{*}$ is a target function,  and $l = (l_{V}\colon V \to \Sigma_{O},l_{E}\colon E \to \Sigma_{M})$ is a pair of labelling functions that assigns each hyperedge and vertex an operator and a type from monoidal signature $\Sigma$ respectively.
+
+	The labelling function must respect the typing of the operator:
+	\[
+		\forall e \in E. \, l(e) = c : A \to B \implies l_{V}^{*}(s(e)) = A \text{ and } l_{V}^{*}(t(e)) = B.
+	\]
+	We call a hypergraph discrete if its set of hyperedges is empty.
+
+	A \emph{hypergraph homomorphism} $\phi\colon \mathcal{F} \to \mathcal{G}$ is given by a pair of functions $\phi_V\colon V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E\colon E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that the following hold:
+	\begin{align*}
+		\phi_V^*(s_{\mathcal{F}}(e)) & = s_{\mathcal{G}}(\phi_E(e)),   & \phi_V^*(t_{\mathcal{F}}(e)) & = t_{\mathcal{G}}(\phi_E(e)),   \\
+		l_{V,\mathcal{F}}(v)         & = l_{V,\mathcal{G}}(\phi_V(v)), & l_{E,\mathcal{F}}(e)         & = l_{E,\mathcal{G}}(\phi_E(e)).
+	\end{align*}
+
+	We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms.
 \end{definition}
-We denote by $\catname{Hyp(\Sigma)}$ the category of hypergraphs over $\Sigma$ and hypergraph homomorphisms.
-$\catname{Hyp(\Sigma)}$ has all finite colimits and, in particular,  the coproduct is given by the disjoint union of hypergraphs.
-The initial object is given by the empty hypergraph.
+$\catname{Hyp(\Sigma)}$ is equivalently characterised as the category of (finite) presheaves over the category $\mathbf{I}$ generated by objects pairs $(k, l) \in \mathbb{N} \times \mathbb{N}$ and an additional object $\star$, with $k + l$ morphisms: $\{ i\colon \star \to (k, l) \mid 0 \leq i < k + l \}$.
+As such, it has all finite colimits; the initial object is given by the empty hypergraph, and coproducts are given by the disjoint union of hypergraphs.
 
-To properly interpret morphisms of a symmetric monoidal category we need to account for the dangling wires at the diagram's top and bottom that represent inputs and outputs of a morphism.
-Such dangling wires are represented as interfaces for a given hypergraph giving rise to a category of cospans of hypergraphs as defined next.
+However, there is also a need to specify which vertices are to be considered as \emph{input} and \emph{outputs} to the diagram as a whole, corresponding to the dangling wires exiting at the top and bottom of a string diagram.
+This gives rise to the definition of a hypergraph with interfaces via a construction on cospans.
 
@@
 % A cospan from $X$ to $Y$ is a pair of arrows $X \xrightarrow{} A \xleftarrow{} Y$  in $\mathbb{C}$.
-This \textit{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
+This \emph{category of cospans of $\mathbb{C}$},  denoted $\catname{Csp(\mathbb{C})}$,  has as objects those of $\mathbb{C}$ and as arrows isomorphism classes of cospans.
 
-% Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A +_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
+% Composition of cospans $X \xrightarrow{f} A \xleftarrow{g} Y$ and $Y \xrightarrow{h} B \xleftarrow{i} Z$ is given by pushout: $X \xrightarrow{f} A \sqcup_{g,h} B \xleftarrow{i} Z$,  with identity given by the cospan of identities.
 Its monoidal product is given by the coproduct of $\mathbb{C}$ as follows:
 % \begin{multline*}
-%     (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A' \xrightarrow{f'} \mathcal{G'} \xleftarrow{g'} B') \coloneq\\ A + A' \xrightarrow{f+f'} \mathcal{G} + \mathcal{G'} \xleftarrow{g+g'} B + B'  ,
+%     (A \xrightarrow{f} \mathcal{G} \xleftarrow{g} B) \otimes (A^\prime \xrightarrow{f^\prime} \mathcal{G^\prime} \xleftarrow{g^\prime} B^\prime) \coloneq\\ A \sqcup A^\prime \xrightarrow{f \sqcup f^\prime} \mathcal{G} \sqcup \mathcal{G^\prime} \xleftarrow{g \sqcup g^\prime} B \sqcup B^\prime  ,
 % \end{multline*}
@@
 % Then, morphisms of $\catname{S}(\Sigma)$ can be interpreted as particular cospans of discrete hypergraphs,  which we call hypergraphs with interfaces~\cite{bonchi_string_2022-2}.
-\begin{definition}[Hypergraphs with interfaces]
+\begin{definition}[Hypergraphs with interfaces]%
 	\label{def:cspd}
-	A hypergraph with interfaces is a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\catname{Hyp}(\Sigma)$ where $n$, $m$ are discrete ordered hypergraphs that denote interfaces.
-	Define the category of \emph{hypergraphs with interfaces} $\catname{HypI(\Sigma)}$ as the category with discrete \textit{ordered} hypergraphs as objects and isomorphism classes of hypergraphs with interfaces as morphisms.
+	A hypergraph with interfaces is a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\catname{Hyp}(\Sigma)$ where $n$, $m$ are discrete hypergraphs denoting input and output interfaces.
+	Write $\catname{HypI}(\Sigma)$ for the full subcategory of the quotient category of cospans\footnote{This is the 1-categorical version of the bicategory of cospans: for a bicategory, its quotient category is given by the same objects but by taking \emph{isomorphism classes} of morphisms instead.} of $\catname{Hyp}(\Sigma)$ whose objects are discrete hypergraphs and morphisms are isomorphism classes of cospans of hypergraphs\footnote{By abuse of notation, we ignore the distinction between a cospan and its isomorphism class --- this is validated by replacing \enquote{pushout} (which is only determined up to unique isomorphism) by \enquote{chosen pushouts}, which any concrete setting (e.g.\ an implementation of DPO rewriting) would satisfy.}.
 \end{definition}
-$\HypI{\Sigma}$ is a symmetric monoidal category with the tensor product given by coproduct in $\catname{Hyp}(\Sigma)$:
+$\HypI{\Sigma}$ is a symmetric monoidal category:
+\begin{enumerate}
+	\item the tensor product given by coproduct in $\catname{Hyp}(\Sigma)$:
 	      \begin{align*}
 		      n \xrightarrow{f}               & \mathcal{F} \xleftarrow{g} m                                                                                                                                       \\
-	                    & \times \qquad \quad \Coloneqq n + n' \xrightarrow{f + f'} \mathcal{F} + \mathcal{G} \xleftarrow{g + g'} m + m' \\
-	n' \xrightarrow{f'} & \mathcal{G} \xleftarrow{g'} m'
+		                                      & \otimes \qquad \quad \Coloneqq n \sqcup n^\prime \xrightarrow{f \sqcup f^\prime} \mathcal{F} \sqcup \mathcal{G} \xleftarrow{g \sqcup g^\prime} m \sqcup m^\prime ; \\
+		      n^\prime \xrightarrow{f^\prime} & \mathcal{G} \xleftarrow{g^\prime} m^\prime
 	      \end{align*}
-Unit is given by the initial object of $\catname{Hyp}(\Sigma)$.
-Symmetry is inherited from coproduct in $\catname{Hyp}(\Sigma)$~\cite{MonoidalCoproduct}.
+	\item the monoidal unit is given by the initial object of $\catname{Hyp}(\Sigma)$;
+	\item symmetry is inherited from coproduct in $\catname{Hyp}(\Sigma)$~\cite{MonoidalCoproduct}.
+\end{enumerate}
 
-Composition of two cospans $n \xrightarrow{f} \mathcal{F} \xleftarrow{g} m;m \xrightarrow{f'} \mathcal{G} \xleftarrow{g'} k$ is given by the pushout
+Composition of two cospans $n \xrightarrow{f} \mathcal{F} \xleftarrow{g} m;m \xrightarrow{f^\prime} \mathcal{G} \xleftarrow{g^\prime} k$ is given by the pushout
 $
-	n \xrightarrow{f} \mathcal{F} +_{g,f'} \mathcal{G} \xleftarrow{g'} k
+	n \xrightarrow{f} \mathcal{F} \sqcup_{g,f^\prime} \mathcal{G} \xleftarrow{g^\prime} k
 $,
-computed by identifying vertices in $\mathcal{F} + \mathcal{G}$ that share pre-image in $m$.
-As pushout is only defined up to isomorphism, it justifies the morphisms in this category being isomorphism classes of cospans.
-To make such cospans complete with respect to representing morphisms of $\catname{S}(\Sigma)$, we need to impose some restrictions on their form.
+computed by identifying vertices in $\mathcal{F} \sqcup \mathcal{G}$ that share pre-image in $m$.
+
+As it stands, every string diagram in a symmetric monoidal category can be represented by a hypergraph with interfaces, but the converse is not true.
+In order to rectify this, we impose additional constraints on the hypergraphs under consideration.
 \begin{definition}[Degree,  path,  acyclic]
-	The \emph{in-degree} of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.
-	Similarly, the out-degree of $v$ is the number of edges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
-	A \emph{path} from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of edges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
-	A hypergraph is \emph{directed acyclic} if it contains no paths of length $n > 0$ that start and end in the same edge.
+	The in-degree of a vertex $v$ in a hypergraph $\mathcal{G}$ is the number of hyperedges $e \in {E_\mathcal{G}}$ such that $v \in t(e)$.
+	Similarly, the out-degree of $v$ is the number of hyperedges $e \in E_\mathcal{G}$ such that $v \in s(e)$.
+	A path from $e_0$ to $e_{n-1}$ of length $n$ in a hypergraph $\mathcal{G}$ is an ordered sequence of hyperedges $[e_0, e_1, \ldots, e_{n-1}] \in E_\mathcal{G}^*$ such that for all $i < n - 1$ some vertex $v \in t(e_i)$ is also in $s(e_{i+1})$.
+	A hypergraph is directed acyclic if it contains no paths of length $n > 0$ that start and end in the same hyperedge.
 \end{definition}
 
-\begin{definition}[Category of MDA Hypergraphs with Interfaces~\cite{bonchi_string_2022-2}]
+\begin{definition}[Category of MDA Hypergraphs with Interfaces~\cite{bonchi_string_2022-2}]%
 	\label{def:monogamy_hyp}
-	We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ \emph{monogamous directed acyclic (MDA)} if:
+	We call a cospan $n \xrightarrow{f} \mathcal{G} \xleftarrow{g} m$ in $\HypI{\Sigma}$ monogamous if:
 	\begin{enumerate}
-		\item $\mathcal{G}$ is a directed acyclic hypergraph;
 		\item $f$ and $g$ are monomorphisms;
@@
 	\end{enumerate}
-	We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of MDA cospans.
+	We denote by $\MdaCospans$ the subcategory of $\HypI{\Sigma}$ of monogamous directed acyclic (MDA) cospans.
 \end{definition}
-Note that $\MdaCospans$ remains a symmetric monoidal category with the same equipment as before.  We now recall from
-\cite{bonchi_string_2022-2} the following theorem,  which says the category $\MdaCospans$ is in fact \textit{equivalent} to $\catname{S}(\Sigma)$.
+
+\autoref{fig:hypergraph-with-interfaces} gives an example of an MDA cospan, which is such because it \enquote{looks like} the string diagram \comonoid.
+Analogously, we can see that~\autoref{fig:hypergraph} (equipped with the interfaces choosing $\{ v_1 \}$ and $\{ v_4, v_5, v_6, v_7 \}$) fails to form an MDA cospan --- it is not monogamous for many reasons: $v_1$ has out-degree 0, $v_1, v_5, v_7$ have in-degree 2, and $v_3$ has in-degree 0; additionally, it fails be be directed acyclic due to the path along $e_4$ from $v_7$ to itself\footnote{Recall that loops are not part of the syntax of string diagrams for symmetric monoidal categories in the absence of a trace --- a relaxation of monogamy to allow for this is explored by~\citet{ghica_rewriting_2023}.}.
+
+\begin{figure}
+	\begin{subfigure}[c]{0.3\linewidth}
+		\centering
+		\begin{tikzpicture}[
+				vertex/.style={draw, circle, fill, inner sep=1.5pt},
+				hyperedge/.style={draw, rectangle}
+			]
+			\node[vertex, label=left:{$v_4$}] (v1) at (0,0) {};
+			\node[vertex, label=left:{$v_5$}] (v2) at (1,0) {};
+			\node[vertex, label=left:{$v_6$}] (v3) at (2,0) {};
+			\node[vertex, label=left:{$v_7$}] (v4) at (3,0) {};
+			\node[vertex, label=left:{$v_2$}] (v5) at (0.5,-2) {};
+			\node[vertex, label=left:{$v_3$}] (v6) at (2,-2) {};
+			\node[vertex, label=left:{$v_1$}] (v7) at (2,-3) {};
+
+			\node[hyperedge] (e1) at (0.5,-1) {$e_3$};
+			\node[hyperedge] (e2) at (2,-1) {$e_4$};
+			\node[hyperedge] (e3) at (0.5,-3) {$e_1$};
+			\node[hyperedge] (e4) at (2,-2.5) {$e_2$};
+
+			\draw (v1) -- (e1);
+			\draw (v2) -- (e1);
+			\draw (v5) -- (e1);
+			\draw (v5) edge[bend right] (e3);
+			\draw (v5) edge[bend left] (e3);
+			\draw (v2) edge[in=130,out=-90] (e2);
+			\draw (v3) -- (e2);
+			\draw (v4) edge[in=50,out=-90] (e2);
+			\draw (v4) edge[in=-50,out=-90] (e2);
+			\draw (v6) -- (e2);
+		\end{tikzpicture}
+		\caption{Hypergraph with 7 vertices and 4 hyperedges, all directed bottom-to-top; source/target vertices are ordered left-to-right.}%
+		\label{fig:hypergraph}
+	\end{subfigure}
+	\begin{subfigure}[c]{0.3\linewidth}
+		\centering
+		\begin{tikzpicture}[
+				vertex/.style={draw, circle, fill, inner sep=1.5pt},
+				hyperedge/.style={draw, rectangle}
+			]
+			\node[vertex, label=left:{$v_3$}] (v1) at (0,0) {};
+			\node[vertex, label=left:{$v_3$}] (v1') at (0,1.5) {};
+			\node[vertex, label=left:{$v_2$}] (v2) at (1,0) {};
+			\node[vertex, label=left:{$v_2$}] (v2') at (1,1.5) {};
+			\node[vertex, label=left:{$v_1$}] (v3) at (0.5,-2) {};
+			\node[vertex, label=left:{$v_1$}] (v3') at (0.5,-3.5) {};
+
+			\node[hyperedge] (e1) at (0.5,-1) {$e_1$};
+
+			\draw (v1) -- (e1);
+			\draw (v2) -- (e1);
+			\draw (v3) -- (e1);
+
+			\node[draw, rounded corners, fit={(v1'.west) (v2'.west)}, inner sep=12pt] (S) {};
+			\node[draw, rounded corners, fit={(v3'.west)}, inner sep=12pt] (T) {};
+			\node[draw, rounded corners, fit={(v1.west) (v2.west) (v3.west)}, inner sep=12pt] (H) {};
+			\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (S) -- node[left] {$g$} (H);
+			\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (T) -- node[left] {$f$} (H);
+		\end{tikzpicture}
+		\caption{Hypergraph with interfaces}%
+		\label{fig:hypergraph-with-interfaces}
+	\end{subfigure}
+	\caption{Example hypergraphs}%
+	\label{fig:hypergraph-examples}
+\end{figure}
+
+Note that $\MdaCospans$ remains a symmetric monoidal category with the same structure as before.
+We now recall the following theorem~\cite{bonchi_string_2022-2}, which says there is a symmetric monoidal equivalence which witnesses that MDA hypergraphs with interfaces are precisely string diagrams.
 \begin{theorem}[Corollary 26 of~\cite{bonchi_string_2022-2}]\label{thm:prop-equiv}
@@
 
-The practical meaning of the above theorem is that two $\Sigma$-terms $f$ and $g$ are equal modulo SMC equations if and only if they are represented as the same hypergraph with interfaces.
-The theorem defines a full and faithful functor that is induced by the interpretation function of morphism terms that renders the latter as cospans of hypergraphs and is given in Appendix~\ref{sec:appendix:interpretation}.
-Term rewriting and hence equality modulo $\mathcal{E}$ is then implemented as DPO rewrites over such hypergraphs with interfaces that we recall next.
+The practical meaning of this is that two $\Sigma$-terms $f$ and $g$ are equal modulo symmetric monoidal category equations if and only if they are represented as the same hypergraph with interfaces.
+The theorem defines a full and faithful functor that is induced by the interpretation function of morphism terms that renders the latter as cospans of hypergraphs and is given in~\autoref{sec:appendix:interpretation}.
+Term rewriting and hence equality modulo $\mathcal{E}$ is then implemented as DPO rewriting over such hypergraphs with interfaces, which we recall next.
 
@@
 DPO rewriting is a standard technique for formalising graph transformations categorically.
-% In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \textit{convex} rewriting.
-The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to recover the rewritten graph $\mathcal{H}$.
-In DPO rewriting,  such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs,  where $\mathcal K$ is some invariant that should be maintained during rewriting.
-A typical DPO square is depicted in Fig.~\ref{fig:dpo_dpoi}~(left).
+% In this section we recall standard results of double-pushout (DPO) rewriting of hypergraphs,  as well as the specific notion of DPO rewriting which is sound and complete for arbitrary SMTs,  which is called \emph{convex} rewriting.
+The intuitive notion of (hyper-)graph rewriting is  as expected: given a rewrite rule $\mathcal L \rightsquigarrow \mathcal R$ and a graph $\mathcal{G}$,  we expect to find an occurrence of $\mathcal L$ within $\mathcal{G}$,  remove it,  and then plug $\mathcal R$ into the hole to determine the rewritten graph $\mathcal{H}$.
+In DPO rewriting, such a rewrite rule is formalised as a span $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$ in some appropriate category of graphs, where $\mathcal{K}$ is some invariant graph fragment that should be maintained during rewriting.
+A typical DPO square is depicted in~\autoref{fig:dpo_dpoi}~(left).
 
-The morphism $\mathcal L \to \mathcal G$ is called a \textit{match},  defined whenever there is a corresponding \textit{pushout complement}---the pair of morphisms $\mathcal K \to \mathcal \mathcal{L}^{\bot} \to \mathcal G$ completing the top half of the pushout square defined by $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
-When the morphisms involved in the rewrite rule are monomorphisms,  as is usually required,  constructing $\mathcal{L}^{\bot}$ can be intuitively understood as finding an occurrence of $\mathcal L$ in $\mathcal G$ and removing from this occurrence everything that has no pre-image in $\mathcal K$.
-The final graph $\mathcal H$ is then recovered by inserting into the hole everything from $\mathcal R$ that has no pre-image in $\mathcal K$ --- hence the two pushouts: one for deletion and one for insertion.
-In this case,  we say that $\mathcal G$ is rewritten into $\mathcal H$ via the rewrite rule $\mathcal L \xleftarrow{} \mathcal K \xrightarrow{} \mathcal R$.
+The morphism $\mathcal{L} \to \mathcal{G}$ is called a \emph{match}, defined whenever there is a corresponding \emph{pushout complement} --- the pair of morphisms $\mathcal{K} \to \mathcal{L}^{\bot} \to \mathcal{G}$ completing the top half of the pushout square with $\mathcal K \xrightarrow{} \mathcal L \xrightarrow{} \mathcal G$.
+When the morphisms involved in the rewrite rule are monomorphisms, as is usually required, constructing $\mathcal{L}^{\bot}$ can be intuitively understood as finding an occurrence of $\mathcal{L}$ in $\mathcal{G}$ and removing from this occurrence everything that has no preimage in $\mathcal{K}$.
+The final graph $\mathcal H$ is then obtained by inserting into the hole everything from $\mathcal{R}$ that has no preimage in $\mathcal{K}$ --- hence the two pushouts: one for deletion and one for insertion.
+In this case, we say that $\mathcal{G}$ is rewritten into $\mathcal{H}$ via the rewrite rule $\mathcal{L} \xleftarrow{} \mathcal{K} \xrightarrow{} \mathcal{R}$.
 
-To express the rewriting of (hyper-)graphs with interfaces,  \textit{i.e.},  the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \textit{interfaces}  (DPOI) as shown in Fig.~\ref{fig:dpo_dpoi}~(right).
-In the case of $ \catname{Hyp(\Sigma)}$,  we wish for a rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal L \xleftarrow{} m$, $n \xrightarrow{} \mathcal R \xleftarrow{} m$.
-By noting that a cospan $0 \xrightarrow{} \mathcal L \xleftarrow{[f,g]} n + m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ (Remark 3.14~\cite{bonchi_string_2022-1}) we can formulate a rewrite rule as a single span $\mathcal L \xleftarrow{} n+m \xrightarrow{} \mathcal R$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal G$ and $\mathcal H$.
+To express the rewriting of (hyper-)graphs with interfaces,  i.e.\ the ones that represent morphisms in $\HypI{\Sigma}$, the DPO formalism has been extended to double-pushout rewriting with \emph{interfaces} (DPOI) as shown in~\autoref{fig:dpo_dpoi}~(right).
+In the case of $ \catname{Hyp(\Sigma)}$,  we wish for a rewrite rules to be pairs of discrete cospans of hypergraphs with matching interfaces: $n \xrightarrow{} \mathcal{L} \xleftarrow{} m$, $n \xrightarrow{} \mathcal{R} \xleftarrow{} m$.
+By noting that a cospan $0 \xrightarrow{} \mathcal{L} \xleftarrow{[f,g]} n \sqcup m$ encodes the same data as a cospan $n \xrightarrow{f} \mathcal L \xleftarrow{g} m$ \cite[Remark 3.14]{bonchi_string_2022-1} we can formulate a rewrite rule as a single span $\mathcal{L} \xleftarrow{} n \sqcup m \xrightarrow{} \mathcal{R}$,  and we can do similarly for the discrete cospans giving the interface to $\mathcal{G}$ and $\mathcal{H}$.
 We now make this identification throughout without further mention.
-Thus we reach the following variant of DPO rewriting.
-\begin{definition}[DPOI rewriting]\label{def:dpoi}
-	Given a span of morphisms $\mathcal G \xleftarrow{} n+m \xrightarrow{} \mathcal H$ in $\textbf{Hyp}(\Sigma)$,  we say \textit{$\mathcal G$ rewrites to $\mathcal H$ with interface $n+m$ via rewrite rule $\mathcal L \xleftarrow{} i+j \xrightarrow{} \mathcal R$} if there exists an object $\mathcal C$ and morphisms which complete the rightmost commutative diagram in Fig.~\ref{fig:dpo_dpoi} such that the two marked squares are pushouts.
+Thus, we reach the following variant of DPO rewriting.
+\begin{definition}[DPOI rewriting]%
+	\label{def:dpoi}
+	Given a span of morphisms $\mathcal{G} \xleftarrow{} n \sqcup m \xrightarrow{} \mathcal{H}$ in $\catname{Hyp}(\Sigma)$,  we say $\mathcal{G}$ rewrites to $\mathcal{H}$ with interface $n \sqcup m$ via rewrite rule $\mathcal{L} \xleftarrow{} i \sqcup j \xrightarrow{} \mathcal{R}$ if there exists an object $\mathcal{C}$ and morphisms which complete the rightmost commutative diagram in~\autoref{fig:dpo_dpoi} such that the two marked squares are pushouts.
 
@@
 				\begin{tikzcd}
-					{\mathcal{L}} & {i+j} & {\mathcal{R}} \\
+					{\mathcal{L}} & {i \sqcup j} & {\mathcal{R}} \\
 					{\mathcal{G}} & {\mathcal{L}^{\bot}} & {\mathcal{H}} \\
-					& {n+m}
+					& {n \sqcup m}
 					\arrow["f"', from=1-1, to=2-1]
@@
 \end{definition}
-To make DPOI rewriting sound and complete with respect to rewriting of $\Sigma$-terms modulo SMC laws and equations $\mathcal{E}$ the notion of both pushout complement and match must be restricted: otherwise there might be DPO rewrites that are not derivable from SMC laws and $\mathcal{E}$-equations.
-This gives rise to the following notion of a \textit{boundary complement} and \textit{convex} DPOI rewriting.
+To make DPOI rewriting sound and complete with respect to rewriting of $\Sigma$-terms modulo the laws of symmetric monoidal categories and equations $\mathcal{E}$, the notion of both pushout complement and match must be restricted: otherwise, there might be DPO rewrites that are not derivable.
+This gives rise to the following notion of a \emph{boundary complement} and \emph{convex} DPOI rewriting.
 
-\begin{definition}[Boundary complement]
+\begin{definition}[Boundary complement]%
 	\label{def:boundary_original}
-	For MDA cospans $i \xrightarrow{a_1} L \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal G \xleftarrow{b_2} m$
-	and mono $f : \mathcal L \to \mathcal G$, a pushout complement $i + j \to \mathcal{L}^\bot \to \mathcal G$ as depicted in the square below
+	For MDA cospans $i \xrightarrow{a_1} \mathcal{L} \xleftarrow{a_2} j$ and $n \xrightarrow{b_1} \mathcal{G} \xleftarrow{b_2} m$ and monomorphism $f\colon \mathcal L \to \mathcal G$, a pushout complement $i \sqcup j \to \mathcal{L}^\bot \to \mathcal{G}$ as depicted in the square below
 	\[
@@
 		}\]
-	is called a boundary complement if $[c_1, c_2]$ is mono and there exist $d_1 : n \to \mathcal{L}^{\bot}$ and $d_2 : m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
+	is called a boundary complement if $[c_1, c_2]$ is a monomorphism and there exist $d_1\colon n \to \mathcal{L}^{\bot}$ and $d_2\colon m \to \mathcal{L}^{\bot}$ making the above triangle commute and such that
 	$
-		n + j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m + i
+		n \sqcup j \xrightarrow{[d_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[d_2,c_1]} m \sqcup i
 	$
@@
 \begin{definition}[Convex match]
-	A subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ is a \emph{convex subgraph} if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
-	A match $f : \mathcal L \to \mathcal G$ is a \emph{convex match} if it is mono and its image is a convex subgraph of $\mathcal G$.
-
+	A subgraph $\mathcal{H}$ of a hypergraph $\mathcal{G}$ is a convex subgraph if for all $v_i, v_j \in V_{\mathcal{H}}$ every path from $v_i$ to $v_j$ is also in $\mathcal{H}$.
+	A match $f\colon \mathcal{L} \to \mathcal{G}$ is a convex match if it is a monomorphism and its image is a convex subgraph of $\mathcal{G}$.
 \end{definition}
 
-\begin{definition}[Convex DPOI rewriting]
+\begin{definition}[Convex DPOI rewriting]%
 	\label{def:convex_dpo}
 	Let $\mathfrak{R}$ be a set of DPOI rewrite rules.
-	Then, given $\mathcal G \xleftarrow{} n+m$ and $\mathcal H \xleftarrow{} n + m$ in $\catname{Hyp(\Sigma)}$, $\mathcal G$ rewrites convexly into $\mathcal H$ with interface $n + m$ --- notation $(\mathcal G \xleftarrow{} n + m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n + m )$ --- if there exist rule $\mathcal L \xleftarrow{} i + j \xrightarrow{} \mathcal R$ in $\mathfrak{R}$ and object $\mathcal{L}^{\bot}$ and cospan arrows $i+j \xrightarrow{} \mathcal{L}^{\bot} \xleftarrow{} n+m$ such that the DPOI diagram of Definition \ref{def:dpoi} commutes and its marked squares are pushouts
-	and the following conditions hold
-	\begin{itemize}
-		\item $f : L \to \mathcal G$ is a convex match;
-		\item $i + j \to \mathcal{L}^{\bot} \to \mathcal G$ is a boundary complement in the leftmost pushout.
-	\end{itemize}
+	Then, given $\mathcal{G} \xleftarrow{} n \sqcup m$ and $\mathcal{H} \xleftarrow{} n \sqcup m$ in $\catname{Hyp(\Sigma)}$, $\mathcal{G}$ rewrites convexly into $\mathcal{H}$ with interface $n \sqcup m$, and we write $(\mathcal G \xleftarrow{} n \sqcup m ) \Rrightarrow_{\mathfrak{R}}  (\mathcal H \xleftarrow{} n \sqcup m )$, if
+	\begin{enumerate}
+		\item there exists a rule $\mathcal{L} \xleftarrow{} i \sqcup j \xrightarrow{} \mathcal{R}$ in $\mathfrak{R}$ and object $\mathcal{L}^{\bot}$ and cospan arrows $i \sqcup j \xrightarrow{} \mathcal{L}^{\bot} \xleftarrow{} n \sqcup m$ such that the DPOI diagram of~\autoref{def:dpoi} commutes and its marked squares are pushouts;
+		\item $f\colon L \to \mathcal G$ is a convex match;
+		\item $i \sqcup j \to \mathcal{L}^{\bot} \to \mathcal G$ is a boundary complement in the leftmost pushout.
+	\end{enumerate}
 \end{definition}
 
-Then, the authors of~\cite{bonchi_string_2022-2} showed that this enhanced notion of DPOI rewriting is sound and complete in the following sense.
-Any two $\Sigma$-terms $f$ and $g$ are equal modulo SMC laws and equations $\mathcal{E}$ if and only if $\llbracket f \rrbracket \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} \llbracket g \rrbracket$ where $\llbracket - \rrbracket$ is the functor giving the equivalence in Theorem~\ref{thm:prop-equiv} and $\llbracket \mathcal{E} \rrbracket = \{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle, \langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle \text{ for } l=r \in \mathcal{E} \}$.
-We then proceed with building a combinatorial representation for closed $\Sigma^{+}$-terms to obtain a notion of DPOI rewriting that would also subsume equality modulo SMC laws.
+\citet{bonchi_string_2022-2} goes on to show that this enhanced notion of DPOI rewriting is sound and complete in the following sense: any two $\Sigma$-terms $f$ and $g$ are equal modulo the laws of symmetric monoidal categories and equations $\mathcal{E}$ if and only if $\llbracket f \rrbracket \Rrightarrow{}_{\llbracket \mathcal{E} \rrbracket} \llbracket g \rrbracket$ where $\llbracket - \rrbracket$ is the functor giving the equivalence in~\autoref{thm:prop-equiv} and $\llbracket \mathcal{E} \rrbracket = \{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle, \langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle \mid l=r \in \mathcal{E} \}$.
+We then proceed with building a combinatorial representation for closed $\Sigma^{+}$-terms to obtain a notion of DPOI rewriting that would also subsume equality modulo the laws of closed symmetric monoidal $\catname{SLat}$-categories.
 
-\section{Closed E-hypergraphs}
+\section{Closed E-Hypergraphs}%
 \label{sec:closed-e-hyp}
-Prior work of~\cite{tiurin2025equivalencehypergraphsdporewriting} introduced a notion of \textit{e-hypergraph} --- a combinatorial representation of morphism terms for free semilattice-enriched symmetric monoidal category with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
-The semilattice enrichment structure was encoded using hierarchical edges in a hypergraph following the approach of a well-established body of literature on hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
-The first work in \textit{loc. cit.}, in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
-Notably, the laws of a semilattice enrichment in~\cite{tiurin2025equivalencehypergraphsdporewriting} and the laws of Definition~\ref{def:closed} in~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
+Prior work of~\citet{tiurin2025equivalencehypergraphsdporewriting} introduced a notion of \emph{e-hypergraph} --- a combinatorial representation of morphism terms for free symmetric monoidal $\catname{SLat}$-categories with a modified notion of DPOI rewriting to make it sound and complete for equalities between morphism terms.
+The $\catname{SLat}$-enrichment structure was encoded using hierarchical hyperedges in a hypergraph following the approach of a well-established body of literature on hierarchical hypergraphs~\cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical, fscd}.
+The first work in loc.\ cit., in particular, introduced a combinatorial representation of closed $\Sigma$-terms (without enrichment).
+Notably, the laws of a $\catname{SLat}$-enrichment used by~\citet{tiurin2025equivalencehypergraphsdporewriting} and the laws of~\autoref{def:closed}~\cite{fscd} were not absorbed by the combinatorial representation and instead were implemented on top as structural DPOI rewrites.
 
-We follow an analogous approach by combining combinatorial representations of~\cite{fscd} and~\cite{tiurin2025equivalencehypergraphsdporewriting}.
-We next extend the definition of e-hypergraphs for $\catname{SLat}$-enriched symmetric monoidal categories~\cite{tiurin2025equivalencehypergraphsdporewriting} to accommodate new structure brought by closedness, and when evoking a term `e-hypergraph' we will mean the structure defined as follows.
+We follow an analogous approach by combining combinatorial representations of~\citet{fscd} and~\citet{tiurin2025equivalencehypergraphsdporewriting}.
+We next extend the definition of e-hypergraphs for symmetric monoidal $\catname{SLat}$-categories~\cite{tiurin2025equivalencehypergraphsdporewriting} to accommodate new structure brought by closure.
 
 \begin{definition}
-	We define an e-hypergraph $\mathcal{G}$ over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{e-color}{<}, \textcolor{closed-color}{<},\consistency)$ where
+	Over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$, we define a closed e-hypergraph $\mathcal{G}$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{e-color}{<}, \textcolor{closed-color}{<},\consistency)$ where
 	\begin{itemize}
 		\item $V$ is a set of vertices.
-		\item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that is formed of three disjoint sets of hyperedges defined below.
-		\item $s,t : E \to V^{*}$ are source and target functions.
-		\item $l_{E} : E \to \Sigma_{M} + 1$, $l_{V} : V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are \textit{label} functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
-		\item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V + E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V + E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
-		      For $x \in V + E$, the set of predecessors of $x$ is denoted as $[x) = \{x' ~|~ \exists y_{1} = x', \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{<^{\mu}} y_{i+1}\; \}$.
-		      We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ \textit{top-level}.
-		      Each of these relations should satisfy the following
+		\item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that is formed of three disjoint sets of hyperedges defined as follows.
+		\item $s,t\colon E \to V^{*}$ are source and target functions.
+		\item $l_{E}\colon E \to \Sigma_{M} \sqcup 1$, $l_{V}\colon V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are label functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
+		\item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V \sqcup E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V \sqcup E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
+		      For $x \in V \sqcup E$, the set of predecessors of $x$ is denoted as $[x) \coloneq \{x^\prime ~|~ \exists y_{1} = x^\prime, \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{<^{\mu}} y_{i + 1}\; \}$.
+		      We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ top-level.
+		      Each of these relations should satisfy the following:
 		      \begin{enumerate}
@@
 			      \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{<^{\mu}} x$ or $e \textcolor{closed-color}{<^{\mu}} x$ are labelled;
-			      \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e' \textcolor{e-color}{<^{\mu}} e$ iff $e' \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<^{\mu}}$.
+			      \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e^\prime \textcolor{e-color}{<^{\mu}} e$ if and only if $e^\prime \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<^{\mu}}$.
 		      \end{enumerate}
-		\item $\consistency$ is a \textit{consistency relation} which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V + E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, \textit{i.e.}, if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{<^{\mu}}(v)$ and $p \textcolor{e-color}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
-		      We require that $\consistency_{p} \not = (E_{p} + V_{p}) \times (E_{p} + V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{<^{\mu}} e\}$.
+		\item $\consistency$ is a consistency relation which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V \sqcup E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, i.e.\ if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{<^{\mu}}(v)$ and $p \textcolor{e-color}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
+		      We require that $\consistency_{p} \not = (E_{p} \sqcup V_{p}) \times (E_{p} \sqcup V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{<^{\mu}} e\}$.
 	\end{itemize}
@@
 	% we do not use this
-	% Given an e-hypergraph $\mathcal{G}$ we can consider a corresponding \textit{underlying} hypergraph $\mathcal{G}'$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
+	% Given a closed e-hypergraph $\mathcal{G}$ we can consider a corresponding \emph{underlying} hypergraph $\mathcal{G}^\prime$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
 \end{definition}
 
-Intuitively, edges from $\textcolor{e-color}{E}$ encode equivalence classes, \emph{i.e.}, e-boxes, while edges from $\textcolor{closed-color}{E}$ encode lambda-abstraction boxes.
+Intuitively, edges from $\textcolor{e-color}{E}$ encode equivalence classes, i.e.\ e-classes, while edges from $\textcolor{closed-color}{E}$ encode $\lambda$-abstraction boxes.
 Also note the typing on the $l_{V}$ function.
-We work with freely generated closed semilattice-enriched categories, and to avoid unnecessary complications we limit the labelling of vertices to such a type.
-In particular, it prohibits vertices to be labelled with, \emph{e.g.}, $A \otimes B$, but allows them to be labelled with $A \otimes B \multimap C \otimes D$.
-$\textcolor{closed-color}{<}$ defines closed monoidal structure as defined in~\cite{fscd} and $(\textcolor{e-color}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\cite{tiurin2025equivalencehypergraphsdporewriting}.
+We work with freely generated closed $\catname{SLat}$-categories, and to avoid unnecessary complications we limit the labelling of vertices to such a type.
+In particular, it prohibits the labelling of vertices by e.g.\ $A \otimes B$, but allows them to be labelled by $A \otimes B \multimap C \otimes D$.
+$\textcolor{closed-color}{<}$ defines closed monoidal structure as defined by~\citet{fscd} and $(\textcolor{e-color}{<},\consistency)$ defines $\catname{SLat}$ structure as per~\citet{tiurin2025equivalencehypergraphsdporewriting}.
 
-Consider an example in the middle part of Fig.~\ref{fig:e-cospan-example}.
-The e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{e-color}{E} = \{e_{1}\}$, $\textcolor{closed-color}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
+Consider an example in the middle part of~\autoref{fig:e-cospan-example}.
+The closed e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{e-color}{E} = \{e_{1}\}$, $\textcolor{closed-color}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
 \[
@@
 $\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
-Edges in $\textcolor{e-color}{E}$ are depicted with a dashed box, while edges in $\textcolor{closed-color}{E}$ --- with a round solid box.
+We depict edges in $\textcolor{e-color}{E}$ with a dashed box, and edges in $\textcolor{closed-color}{E}$ with a round box.
 \begin{figure}
@@
 	\captionsetup{skip=0pt, belowskip=-1ex}
-	\caption{Cospan of e-hypergraphs example}
+	\caption{Cospan of closed e-hypergraphs example}
 	\label{fig:e-cospan-example}
 \end{figure}
-\begin{definition}[E-hypergraph homomorphism]
-	An \textit{e-hypergraph homomorphism} $\phi:\mathcal{F}\to\mathcal{G}$ between e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V : V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E : E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that
+\begin{definition}[Closed e-hypergraph homomorphism]
+	A homomorphism $\phi\colon \mathcal{F}\to\mathcal{G}$ between closed e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V\colon  V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E\colon E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that:
 	\begin{enumerate}
@@
 		\item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
-		      \begin{align*}
-			      \text{if } e \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(v)
+		      \begin{equation*}
+			      \text{if } e \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(v),
 			      \qquad
-			      \text{if } e \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(v)
-		      \end{align*}
+			      \text{if } e \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(v),
+		      \end{equation*}
 		      and for $e_1, e_2 \in E_{\mathcal{F}}$
-		      \begin{align*}
-			      \text{if } e_1 \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
+		      \begin{equation*}
+			      \text{if } e_1 \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2),
 			      \qquad
-			      \text{if } e_1 \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2)
-		      \end{align*}
+			      \text{if } e_1 \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2);
+		      \end{equation*}
 		\item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$.
 	\end{enumerate}
-	% If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \textit{strict}.
+	% If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \emph{strict}.
 \end{definition}
-The conditions on e-hypergraph homomorphisms require preserving immediate predecessors and the consistency relation.
-\begin{definition}[Category of e-hypergraphs]
-	The \emph{category of e-hypergraphs},  denoted  $\catname{EHyp(\Sigma)}$,  has e-hypergraphs as objects and e-hypergraph homomorphisms as morphisms.
+The conditions on closed e-hypergraph homomorphisms require preserving immediate predecessors and the consistency relation.
+\begin{definition}[Category of closed e-hypergraphs]
+	The category of closed e-hypergraphs, denoted  $\catname{CEHyp(\Sigma)}$,  has closed e-hypergraphs as objects and closed e-hypergraph homomorphisms as morphisms.
 \end{definition}
 
-The category of e-hypergraphs has coproducts,  given by the disjoint union of e-hypergraphs,  and an initial object given by the empty e-hypergraph.
-A concrete description of the pushout of two morphisms in this category is given in Appendix~\ref{sec:appendix:pushout}.
-Generally,  the pushout of two e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied.
+The category of closed e-hypergraphs has coproducts, given by the disjoint union of closed e-hypergraphs, and an initial object given by the empty closed e-hypergraph.
+A concrete description of the pushout of two morphisms in this category is given in~\autoref{sec:appendix:pushout}.
+Generally,  the pushout of two closed e-hypergraph homomorphisms need not exist,  but it does when certain technical conditions are satisfied.
 However, a pushout for the composition of two cospans with discrete feet always exists, as described below.
 
-As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, e-hypergraphs require interfaces in order to model string diagrams with dedicated inputs and outputs.
-Because string diagrams for closed semilattice-enriched symmetric monoidal categories are nested, containing substring diagrams inside e-boxes and lambda abstraction boxes, we need to account for the interfaces in these nested contexts as well.
-Nested inputs and outputs of a string diagram do not participate in composition of string diagrams, hence they need to be distinguished from the ones that participate in the composition.
-We thus propose \textit{extended} cospans of discrete e-hypergraphs and use the notation $n \setminus m$ to denote the discrete e-hypergraph with $n \setminus m$ vertices, in particular
-with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub-e-hypergraph of $n$.
+As in the case of $\catname{S}(\Sigma)$ and $\MdaCospans$, closed e-hypergraphs require interfaces in order to model string diagrams with dedicated inputs and outputs.
+Because string diagrams for closed  symmetric monoidal $\catname{SLat}$-categories are nested, containing substring diagrams inside e-class boxes and $\lambda$-abstraction boxes, we need to account for the interfaces in these nested contexts as well.
+Nested inputs and outputs of a string diagram do not participate in composition of string diagrams, hence they need to be distinguished from the ones which do.
+We thus propose \emph{extended} cospans of discrete closed e-hypergraphs and use the notation $n \setminus m$ to denote the discrete closed e-hypergraph with $n \setminus m$ vertices, in particular
+with the vertices of $m$ removed from $n$ when vertices of $m$ is a sub closed e-hypergraph of $n$.
 
-\begin{definition}[Category of e-hypergraphs with interfaces]
-	The category of \textit{e-hypergraphs with extended interfaces} $\Ecospans$ has discrete \textit{ordered} e-hypergraphs as objects,  with hom-sets $\Ecospans(n,m)$ consisting of isomorphism classes (defined below) of \textit{extended cospans}, defined as follows:
+\begin{definition}[Category of closed e-hypergraphs with interfaces]
+	The category of closed e-hypergraphs with extended interfaces $\Ecospans$ has discrete \emph{ordered} closed e-hypergraphs as objects,  with Hom sets $\Ecospans(n,m)$ consisting of isomorphism classes (defined below) of extended cospans, defined as follows:
 	\[
-		n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m' \xleftarrow{g_{ext}} m
+		n \xrightarrow{f_{ext}} n^\prime \xrightarrow{f_{int}} \mathcal{G} \xleftarrow{g_{int}} m^\prime \xleftarrow{g_{ext}} m,
 	\]
-	where $\mathcal{G}$ is an e-hypergraph,  and $n, n', m, m'$ are discrete ordered e-hypergraphs,  $f_{ext},g_{ext}$ are monomorphisms in $\catname{EHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
+	where $\mathcal{G}$ is a closed e-hypergraph,  and $n, n^\prime, m, m^\prime$ are discrete ordered closed e-hypergraphs,  $f_{ext},g_{ext}$ are monomorphisms in $\catname{CEHyp(\Sigma)}$,  and the image of $f_{ext};f_{int}$ and of $g_{ext};g_{int}$ consist exclusively of top-level vertices,  and such that vertices in the strictly internal interface (defined below) are not top-level.
 	We sometimes write $\mathcal G$ to denote the extended cospan,  where it is clear from context $\mathcal G$ is equipped with extended interfaces.
@@
 \end{definition}
-We call $n$  \textit{external input interfaces}, $n'$ \textit{internal input interfaces},
-and $n' \setminus f_{ext}(n)$  the \textit{strictly internal input interfaces}.
-We do analogously for the \emph{output interfaces},  with respect to $m$,  $m'$ and $m' \setminus g_{ext}(m)$.
-We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n'$ and $m'$,  and likewise $m, m, n', m'$ with their images in $\mathcal{G}$.
-Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \textit{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
-Graphically, we depict such extended cospans as in Fig.~\ref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
+We call $n$  \emph{external input interfaces}, $n^\prime$ \emph{internal input interfaces},
+and $n^\prime \setminus f_{ext}(n)$  the \emph{strictly internal input interfaces}.
+We do analogously for the \emph{output interfaces},  with respect to $m$,  $m^\prime$ and $m^\prime \setminus g_{ext}(m)$.
+We occasionally conflate $f_{ext}$ with $f_{ext};f_{int}$ when it is clear from context,  and also conflate $n$ and $m$ with their images in $n^\prime$ and $m^\prime$,  and likewise $m, m, n^\prime, m^\prime$ with their images in $\mathcal{G}$.
+Given an edge $e \in E_\mathcal{G}$ such that $l(e) = \bot$,  we call the \emph{inputs of $e$} the intersection of the strictly internal input interface of $\mathcal{G}$ with the immediate successors of $e$,  and analogously for the \emph{outputs of $e$}.
+Graphically, we depict such extended cospans as in~\autoref{fig:e-cospan-example}: feet of the cospan are depicted within shaded regions.
 All non-black vertices form strictly internal interfaces.
@@
 	\fi
-	for $x, y \in n'$.
+	for $x, y \in n^\prime$.
 	And let $S$ be its reflexive closure.
-	The latter partitions $n'$ into non-empty subsets $\{p_{j}\}^{k}_{j=1}$.
-	We get an analogous partition for $m'$.
+	The latter partitions $n^\prime$ into non-empty subsets $\{p_{j}\}^{k}_{j=1}$.
+	We get an analogous partition for $m^\prime$.
 
@@
 
-In Fig.~\ref{fig:e-cospan-example} the partitions for input and output internal interfaces are given with colours.
+In~\autoref{fig:e-cospan-example} the partitions for input and output internal interfaces are given with colours.
 
 Intuitively, such notion of isomorphism makes it possible to freely rearrange interfaces for disjoint components.
-For example, in the input interface of cospan in Fig.~\ref{fig:e-cospan-example}, we can swap
+For example, in the input interface of cospan in~\autoref{fig:e-cospan-example}, we can swap
 \begin{tikzpicture}
@@
 $
-	n \xrightarrow{f_{ext}} n' \xrightarrow{f_{int}} \mathcal{F} \xleftarrow{f'_{int}} m' \xleftarrow{f'_{ext}} m
+	n \xrightarrow{f_{ext}} n^\prime \xrightarrow{f_{int}} \mathcal{F} \xleftarrow{f^\prime_{int}} m^\prime \xleftarrow{f^\prime_{ext}} m
 $ and $
-	m \xrightarrow{g_{ext}} m'\!' \xrightarrow{g_{int}} \mathcal{G} \xleftarrow{g'_{int}} k' \xleftarrow{g'_{ext}} k
+	m \xrightarrow{g_{ext}} m^\prime\!^\prime \xrightarrow{g_{int}} \mathcal{G} \xleftarrow{g^\prime_{int}} k^\prime \xleftarrow{g^\prime_{ext}} k
 $ is computed in two stages.
@@
 \[
-	n \xrightarrow{f_{ext};\iota_1} n' + (m'' \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k' + (m' \setminus m) \xleftarrow{g'_{ext};\iota_1} k
+	n \xrightarrow{f_{ext};\iota_1} n^\prime \sqcup (m^{\prime\prime} \setminus m) \xrightarrow{h_{1}} \mathcal{H} \xleftarrow{h_2} k^\prime \sqcup (m^\prime \setminus m) \xleftarrow{g^\prime_{ext};\iota_1} k
 \]
-where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete e-hypergraph.
+where $h_i$ are defined as follows,  using $|$ to denote the restriction of a function on a discrete closed e-hypergraph.
 \begin{align*}
-	h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m'' \setminus m} ] \; h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
+	h_1 = [ f_{int};p_1, ~(g_{int};p_2)|_{m^{\prime\prime} \setminus m} ] \; h_2 = [ g^\prime_{int};p_2, ~(f^\prime_{int};p_1)|_{m^\prime \setminus m} ]
 	% \\
-	% h_2 = [ g'_{int};p_2, ~(f'_{int};p_1)|_{m' \setminus m} ]
+	% h_2 = [ g^\prime_{int};p_2, ~(f^\prime_{int};p_1)|_{m^\prime \setminus m} ]
 \end{align*}
@@
 % Y_{ext} \arrow[d] & 0 \arrow[l] \arrow[d]    \\
-% Y'          & Y' \setminus Y_{ext} \arrow[l]
+% Y^\prime          & Y^\prime \setminus Y_{ext} \arrow[l]
 % \end{tikzcd}
@@
 % and then the disjoint union is just a coproduct. 
-% where $g'$ is the restriction of $g$ to ${Y' \setminus Y_{ext}}$ and $f'$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
+% where $g^\prime$ is the restriction of $g$ to ${Y^\prime \setminus Y_{ext}}$ and $f^\prime$ the restriction of $f$ to $Y \setminus Y_{ext}$. 
 The identity of composition is given by the obvious extended cospan of identities.
-$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{EHyp({\Sigma}})$,  analogously to Definition~\ref{def:cspd}.
+$\Ecospans$ inherits a symmetric monoidal structure from the coproduct (and initial object) structure of $\catname{CEHyp({\Sigma}})$,  analogously to~\autoref{def:cspd}.
 As in the standard cospan construction,  it is necessary to consider isomorphism classes of cospans since composition (defined by pushout) is associative only up-to isomorphism (since pushouts are unique only up-to isomorphism).
@@
 	$
-		n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+		n \xrightarrow{f_{\text{ext}}} n^\prime \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g^\prime_{\text{int}}} m^\prime \xleftarrow{g_{\text{ext}}} m
 	$
@@
 		\item in- and out- degrees of every vertex is at most 1;
-		\item $f_{\text{int}}$ and $g_{\text{int}}$ are monos;
+		\item $f_{\text{int}}$ and $g_{\text{int}}$ are monomorphisms;
 		\item vertices with in-degree (respectively, out-degree) of 0 are precisely the image of $f_{\text{int}}$ (respectively, $g_{\text{int}}$).
@@
 
-	MDA cospans and discrete e-hypergraphs form a category $\MdaEcospans$.
+	MDA cospans and discrete closed e-hypergraphs form a category $\MdaEcospans$.
 \end{definition}
@@
 	\[
-		n \xrightarrow{f_{\text{ext}}} n' \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g'_{\text{int}}} m' \xleftarrow{g_{\text{ext}}} m
+		n \xrightarrow{f_{\text{ext}}} n^\prime \xrightarrow{f_{\text{int}}} \mathcal{G} \xleftarrow{g^\prime_{\text{int}}} m^\prime \xleftarrow{g_{\text{ext}}} m
 	\]
-	\textit{well-typed} if all hierarchical edges of $\mathcal{G}$ are well-typed in the sense below.
+	well-typed if all hierarchical edges of $\mathcal{G}$ are well-typed in following sense:
 	\begin{itemize}
-		\item For each $e \in \textcolor{e-color}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$.
-		      $e$ is well-typed if for each element $S$ of the partition of $I$ $w(S) = w(s(e))$ and similarly for $O$ and $t(e)$.
-		\item For each $e \in \textcolor{closed-color}{E}$ consider sets $I$ and $O$ of its input and output vertices.
+		\item for each $e \in \textcolor{e-color}{E}$ consider sets $I$ and $O$ of its input and output vertices partitioned according to $\consistency_{e}$;
+		      $e$ is well-typed if for each element $S$ of the partition of $I$ $w(S) = w(s(e))$ and similarly for $O$ and $t(e)$;
+		\item for each $e \in \textcolor{closed-color}{E}$ consider sets $I$ and $O$ of its input and output vertices;
 		      $e$ is well-typed if there exists an object $B \in \text{obj}_{\Sigma_{O}}$ such that $[w(s(e)), B] = w(I)$ and $B \multimap w(O) = w(t(e))$.
@@
 
-Consider, again, Fig. ~\ref{fig:e-cospan-example}.
-To make edge $e_{1}$ well-typed, it is necessary that $l_{V}^{*}([v_{4},v_{5},v_{6}]) = l_{V}^{*}([v_{10},v_{11},v_{12}]) = l_{V}^{*}([v_1,v_2,v_3])$ and $l_{V}^{*}([w_{4},w_{5},w_{6}]) = l_{V}^{*}([w_{10},v_{12},v_{11}]) = l_{V}^{*}([w_1,w_2,w_3])$ (note the change of ordering in the output of $e_{1}$).
+Again, considering~\autoref{fig:e-cospan-example}, to make edge $e_{1}$ well-typed, it is necessary that $l_{V}^{*}([v_{4},v_{5},v_{6}]) = l_{V}^{*}([v_{10},v_{11},v_{12}]) = l_{V}^{*}([v_1,v_2,v_3])$ and $l_{V}^{*}([w_{4},w_{5},w_{6}]) = l_{V}^{*}([w_{10},v_{12},v_{11}]) = l_{V}^{*}([w_1,w_2,w_3])$ (note the change of ordering in the output of $e_{1}$).
 To make edge $e_{2}$ well-typed, it is necessary that $l_{V}^{*}([v_4,v_5,v_9]) = l_{V}^{*}([v_7,v_8,v_9])$ and $l_{V}([u_1]) = l_{V}([v_9]) \multimap l_{V}^{*}([v_8,v_7,w_9])$.
 
-We then proceed with definition a notion of DPOI rewriting for such e-hypergraphs.
+~Wepautoroceed with defining a notion of DPOI rewriting for such closed e-hypergraphs.
 
-\section{DPOI-Rewriting for E-Hypergraphs}
+\section{DPOI-Rewriting for Closed E-Hypergraphs}
 
-Because extended cospans have a more general notion of interface,  including \textit{internal} vertices,  DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments, defined in this section.
-The construction in this section closely follows the one introduced in~\cite{tiurin2025equivalencehypergraphsdporewriting} with some adjustments to account for $\textcolor{closed-color}{<}$.
+Because extended cospans have a more general notion of interface,  including \emph{internal} vertices,  DPOI rewriting as presented in~\autoref{sec:combinatorial-semantics} needs some adjustments, which we define in this section.
+The construction in this section closely follows the one introduced by~\citet{tiurin2025equivalencehypergraphsdporewriting} with some adjustments to account for $\textcolor{closed-color}{<}$.
 We do not expect internal interfaces to be preserved during rewriting: for example,  when the semilattice equations are considered as rewrites.
-Thus,  we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching \textit{external} (but not necessarily internal) interfaces,  as follows.
+Thus,  we wish a rewrite rule to be a pair of \emph{extended} cospans of closed e-hypergraphs with matching \emph{external} (but not necessarily internal) interfaces,  as follows:
 \[
-	n \xrightarrow{} n' \xrightarrow{} \mathcal{L} \xleftarrow{} m' \xleftarrow{} m
+	n \xrightarrow{} n^\prime \xrightarrow{} \mathcal{L} \xleftarrow{} m^\prime \xleftarrow{} m,
 	\qquad
-	n \xrightarrow{} n'' \xrightarrow{} \mathcal{R} \xleftarrow{} m'' \xleftarrow{} m
+	n \xrightarrow{} n^{\prime\prime} \xrightarrow{} \mathcal{R} \xleftarrow{} m^{\prime\prime} \xleftarrow{} m.
 \]
-Analogously to Section \ref{sec:combinatorial-semantics}, observing that  cospan
+Analogously to~\autoref{sec:combinatorial-semantics}, observing that the cospan
 $
-	0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n'' + m'' \xleftarrow{} n + m$ expresses the same data as $
-	0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n' + m' \xleftarrow{} n + m
+	0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} n^{\prime\prime} \sqcup m^{\prime\prime} \xleftarrow{} n \sqcup m$ expresses the same data as $
+	0 \xrightarrow{} 0 \xrightarrow{} \mathcal{R} \xleftarrow{} n^\prime \sqcup m^\prime \xleftarrow{} n \sqcup m
 $
@@
 $
-	\mathcal{L} \xleftarrow{} n' + m' \xleftarrow{} n + m \xrightarrow{} n'' + m'' \xrightarrow{} \mathcal{R},
+	\mathcal{L} \xleftarrow{} n^\prime \sqcup m^\prime \xleftarrow{} n \sqcup m \xrightarrow{} n^{\prime\prime} \sqcup m^{\prime\prime} \xrightarrow{} \mathcal{R},
 $
@@
 
-% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an e-hypergraph we also modify this e-hypergraph's interfaces.
-% Consider a hypothetical DPO square below in Fig.~\ref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
+% Another concern is when we delete the occurrence of the left-hand side of a rewrite rule from an closed e-hypergraph we also modify this closed e-hypergraph's interfaces.
+% Consider a hypothetical DPO square below in~\autoref{fig:interface_change_example} where long arrows denote matching between carriers of cospans.
 
-% The first row of the diagram shows the span for a rewrite rule of $\langle f + g, f \rangle$ which is like a projection rule.
+% The first row of the diagram shows the span for a rewrite rule of $\langle f \sqcup g, f \rangle$ which is like a projection rule.
 % Note once again the difference between the interfaces of the left-hand and right-hand sides of the rewrite rule.
 % The entry in the middle of the span depicts the outermost interfaces which are preserved.
-% The second row depicts the e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
-% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
-% Note that the interfaces of the initial e-hypergraph and the residual e-hypergraph are different. 
-% Here, the outermost and the whole interfaces for the eventual e-hypergraph coincide, but if the nested-ness of the initial e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial e-hypergraph.
+% The second row depicts the closed e-hypergraph to be rewritten on the left and the result of the rewrite on the right.
+% The result is obtained by first deleting the image of the left-hand side of the rewrite rule from the initial closed e-hypergraph (depicted in the middle) and then gluing the right-hand side of the rewrite rule into it.
+% Note that the interfaces of the initial closed e-hypergraph and the residual closed e-hypergraph are different. 
+% Here, the outermost and the whole interfaces for the eventual closed e-hypergraph coincide, but if the nested-ness of the initial closed e-hypergraph was more intense the result could have inner interfaces as well as we would remove the interfaces of the inner edges when deleting the image of the left-hand side part of the rule from the initial closed e-hypergraph.
 
-Before introducing \textit{extended} DPOI rewriting,  the definition of boundary complement must guarantee that rewriting yields a monogamous directed acyclic e-hypergraph.
-First,  we introduce the  ancillary notion of a \textit{down-closed} graph:
+Before introducing \emph{extended} DPOI rewriting,  the definition of boundary complement must guarantee that rewriting yields a monogamous directed acyclic closed e-hypergraph.
+First,  we introduce the  ancillary notion of a \emph{down-closed} graph:
 \begin{definition}[Down-closed subgraph]
-	%A \textit{subgraph} of $\mathcal G $ an e-hypergraph $\mathcal H$ 
-	%is an e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
+	%A \emph{subgraph} of $\mathcal G $ a closed e-hypergraph $\mathcal H$ 
+	%is a closed e-hypergraph whose edge and vertex sets are subsets of those of $\mathcal H$, 
 	%and whose remaining equipment is a restriction of that of $\mathcal H$ to those sets. 
-	We call  a sub-e-hypergraph $\mathcal G $ of $\mathcal H$ \emph{down-closed} if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
+	We call  a sub closed e-hypergraph $\mathcal G $ of $\mathcal H$ down-closed if for all $e \in E_{\mathcal{G}}$,   all children of $e$ are also in $\mathcal{G}$.
 \end{definition}
@@
 	$
-		i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
+		i \xrightarrow{} i^\prime \xrightarrow{} \mathcal{L} \xleftarrow{} j^\prime \xleftarrow{} j
 	$ and $
-		n \xrightarrow{} n' \xrightarrow{} \mathcal{G} \xleftarrow{} k' \xleftarrow{} k
-	$ and mono $m : \mathcal{L} \to \mathcal{G}$ in $\catname{EHyp(\Sigma)}$, a pushout complement $i + j \to \mathcal{L}^{\bot} \to \mathcal{G}$
+		n \xrightarrow{} n^\prime \xrightarrow{} \mathcal{G} \xleftarrow{} k^\prime \xleftarrow{} k
+	$ and mono $m : \mathcal{L} \to \mathcal{G}$ in $\catname{CEHyp(\Sigma)}$, a pushout complement $i \sqcup j \to \mathcal{L}^{\bot} \to \mathcal{G}$
 	as depicted in the square below
@@
 	\]
-	is a \textit{boundary complement} if
+	is a \emph{boundary complement} if:
 	\begin{enumerate}
-		\item $m(\mathcal L)$ is a convex down-closed e-hypergraph;
-		\item $[c_1,c_2]$ is mono;
-		\item for all $v,w$ in $\mathcal{G}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
+		\item $m(\mathcal L)$ is a convex down-closed closed e-hypergraph;
+		\item $[c_1,c_2]$ is a monomorphism;
+		\item for all $v,w$ in $\mathcal{G}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
 $v,w \textcolor{closed-color}{<^{\mu}} e$;
-		\item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i + j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
+		\item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
 $v,w \textcolor{closed-color}{<^{\mu}} e$;
 		\item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and
-		\item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \textit{well-typed} extended MDA cospan
+		\item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of top-level vertices of $\mathcal{G}$ then there exists a \emph{well-typed} extended MDA cospan
 		      \[
-			      \hspace{-2em}n + j \xrightarrow{f_1 + id_{j}} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2 + id_{i}} k + i
+			      \hspace{-2em}n \sqcup j \xrightarrow{f_1 \sqcup id_{j}} n^\prime \setminus (i^\prime \setminus i) \sqcup j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k^\prime \setminus (j^\prime \setminus j) \sqcup i \xleftarrow{f_2 \sqcup id_{i}} k \sqcup i
 		      \]
-		\item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \textit{not necessarily well-typed} extended MDA cospan
+		\item if the image of $\mathcal{L}$'s external interfaces under $m$ consists exclusively of not top-level vertices of $\mathcal{G}$ then there exists a \emph{not necessarily well-typed} extended MDA cospan
 		      \ifdefined\ONECOLUMN
 			      \[
-				      n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+				      n \xrightarrow{f_1} n^\prime \setminus (i^\prime \setminus i) \sqcup j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k^\prime \setminus (j^\prime \setminus j) \sqcup i \xleftarrow{f_2} k
 			      \]
@@
 			      \begin{multline*}
-				      n \xrightarrow{f_1} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \\ \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{f_2} k
+				      n \xrightarrow{f_1} n^\prime \setminus (i^\prime \setminus i) \sqcup j \xrightarrow{[g_1,c_2]} \\ \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k^\prime \setminus (j^\prime \setminus j) \sqcup i \xleftarrow{f_2} k
 			      \end{multline*}
@@
 	where $f_i$ and $g_i$ are defined as follows.
-	%because the image of $m$ is a convex down-closed e-hypergraph, 
-	The strictly internal interface $i' \setminus i$ of $\mathcal L$ is mapped to the internal interface $n'$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i' \setminus i$ in  $n'$.
-	Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n' \setminus (i' \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n' \setminus (i' \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n' \setminus (i' \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n'$ of $\mathcal G$ minus $i' \setminus i$.
+	%because the image of $m$ is a convex down-closed closed e-hypergraph, 
+	The strictly internal interface $i^\prime \setminus i$ of $\mathcal L$ is mapped to the internal interface $n^\prime$ of $\mathcal G$,  since $\mathcal L$ is a down-closed subgraph of $\mathcal G$,  inducing an identification of $i^\prime \setminus i$ in  $n^\prime$.
+	Then map $f_1$ is given by $g_{ext}$ with its codomain restricted to $n^\prime \setminus (i^\prime \setminus i)$.\footnote{Noting the image of $g_{ext}$ indeed lies within $n^\prime \setminus (i^\prime \setminus i)$}  The map $g_1$ is derived from the restriction of $g_{int}$ to type $n^\prime \setminus (i^\prime \setminus i) \to \mathcal G$ by further observing that $\mathcal L^\bot$ can be identified within $\mathcal G$ --- and in particular has the internal interface $n^\prime$ of $\mathcal G$ minus $i^\prime \setminus i$.
 The maps $f_2$ and $g_2$ are defined similarly.
-	%$i' \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n'$ and therefore we can construct a
-	%morphism $f' : i' \setminus i \to n'$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
-	%Then we define $n' \setminus (i' \setminus i)$ as the pushout complement in the diagram below
+	%$i^\prime \setminus i$, which is a strict internal interface of $\mathcal{L}$, is necessarily mapped to vertices in the strict internal interface of $\mathcal{G}$ which are also in the image of $n^\prime$ and therefore we can construct a
+	%morphism $f^\prime : i^\prime \setminus i \to n^\prime$ by first following $m$ and then composing with the partial inverse of $g_{int}$ which exists because it is mono. 
+	%Then we define $n^\prime \setminus (i^\prime \setminus i)$ as the pushout complement in the diagram below
 	%
@@
 	%\]
-	%and $g_1$ is obtained by composition $g';g_{int}$ and then composing with the inverse of $l$.
+	%and $g_1$ is obtained by composition $g^\prime;g_{int}$ and then composing with the inverse of $l$.
 	%Similarly for $g_2$.
 	% CHRIS: TRY TO DEFINE G1, G2 AS SIMPLY AS POSSIBLE, AND POSSIBLY ADD AN INFORMAL EXPLANATION BELOW THE DEFINITION. 
-	% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i + j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
+	% Because $\mathcal{L}^{\bot}$ is essentially $\mathcal{G}$ with $\mathcal{L}$ removed, apart from the part that has a pre-image in $i \sqcup j$, there is an obvious morphism from $l \to \mathcal{L}^{\bot}$ which is $g_1 = g_{int} |_{D}$ where
 	% \[
-	%     D = \{ v \in n' \text{ such that } \not \exists u \in (i' \setminus i) ~ . ~ g_{int}(v) = f'_{int};f(u) \}
+	%     D = \{ v \in n^\prime \text{ such that } \not \exists u \in (i^\prime \setminus i) ~ . ~ g_{int}(v) = f^\prime_{int};f(u) \}
 	% \]
 	% Similarly for $r \to \mathcal{L}^{\bot}$, which we denote as $g_2$.
-	% Then there exists a morphism $n \to l$, because there is a morphism $n \to n'$ and the external interfaces of $n'$ and $l$ are the same.
+	% Then there exists a morphism $n \to l$, because there is a morphism $n \to n^\prime$ and the external interfaces of $n^\prime$ and $l$ are the same.
 	% That is, it is a morphism $f: n \to l$ such that $d_1 = f;g_1$.
@@
 
-%The notion of a boundary complement above is essentially the same as the one from Section~\ref{sec:combinatorial-semantics}.
+%The notion of a boundary complement above is essentially the same as the one from~\autoref{sec:combinatorial-semantics}.
 %CHRIS: THE FOLLOWING SENTENCE IS UNCLEAR. 
@@
 %    \scalebox{0.5}{\tikzfig{combinatorial_semantics/interfaces_change_example}}\]
-%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(EHyp_{\Sigma})}$}
+%    \caption{Hypothetical DPO square for $\catname{MACsp_{D}(CEHyp_{\Sigma})}$}
 %    \label{fig:interface_change_example}
@@
 %     \[
-%     n \xrightarrow{} n' \setminus (i' \setminus i) + j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k' \setminus (j' \setminus j) + i \xleftarrow{} k
+%     n \xrightarrow{} n^\prime \setminus (i^\prime \setminus i) \sqcup j \xrightarrow{[g_1,c_2]} \mathcal{L}^{\bot} \xleftarrow{[g_2,c_1]} k^\prime \setminus (j^\prime \setminus j) \sqcup i \xleftarrow{} k
 %     \]
@@
 
-\begin{proposition}
+\begin{proposition}%
 	\label{prop:boundary_unique}
-	The boundary complement in~\ref{def:boundary_new} when it exists is unique.
+	The boundary complement of~\autoref{def:boundary_new} when it exists is unique.
 \end{proposition}
 
-We are now ready to define \textit{convex extended DPOI (EDPOI) rewriting} for $\catname{EHyp({\Sigma})}$.
-It is analogous to Definition~\ref{def:convex_dpo},  except we must construct internal interfaces explicitly.
-More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting e-hypergraph should be modified to \textit{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
-Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting e-hypergraph should be modified to \textit{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
+We are now ready to define \emph{convex extended DPOI (EDPOI) rewriting} for $\catname{CEHyp({\Sigma})}$.
+It is analogous to~\autoref{def:convex_dpo},  except we must construct internal interfaces explicitly.
+More precisely, when removing the occurrence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of the resulting closed e-hypergraph should be modified to \emph{exclude} the vertices corresponding to $\mathcal{L}$'s strictly internal interfaces (since the vertices they map to have been removed).
+Then, when gluing $\mathcal{R}$ into the hole,  the internal interfaces of the resulting closed e-hypergraph should be modified to \emph{include} the strictly internal interfaces of $\mathcal{R}$ (since new internal interfaces for them to map to have been added).
 \begin{definition}[Convex EDPOI rewriting]\label{def:dpoi-e}
@@
 	$
-		\mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
+		\mathcal{G} \xleftarrow{} n^\prime \sqcup k^\prime \xleftarrow{} n \sqcup k \xrightarrow{} n^{\prime\prime} \sqcup k^{\prime\prime} \xrightarrow{} \mathcal{H}
 	$
-	in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule}
+	in $\catname{CEHyp}(\Sigma)$, we say $\mathcal{G}$ \emph{rewrites (convexly) to} $\mathcal{H}$ (\emph{with external interface} $n \sqcup k$ and \emph{taking internal interface} $n^\prime \sqcup k^\prime$ \emph{to} $n^{\prime\prime} \sqcup k^{\prime\prime}$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \emph{via a rewrite rule}
 	$
-		\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
+		\mathcal{L} \xleftarrow{} i^\prime \sqcup j^\prime \xleftarrow{} i \sqcup j \xrightarrow{} i^{\prime\prime} \sqcup j^{\prime\prime} \xrightarrow{} \mathcal{R}
 	$
@@
 		\label{dpoi-e:assumptions}
-		% \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
-		%\item $m(L)$ is a down-closed e-hypergraph;
-		\item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
+		% \item $m\colon \mathcal{L} \to \mathcal{G}$ is a convex match;
+		%\item $m(L)$ is a down-closed closed e-hypergraph;
+		\item $i \sqcup j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
 		\item the internal interfaces of $\mathcal H$ are such that:
 		      $
-			      n'' = n' \setminus (i' \setminus i) + (i'' \setminus i) $ and $ k'' = k' \setminus (j' \setminus j) +  (j'' \setminus j)
+			      n^{\prime\prime} = n^\prime \setminus (i^\prime \setminus i) \sqcup (i^{\prime\prime} \setminus i) $ and $ k^{\prime\prime} = k^\prime \setminus (j^\prime \setminus j) \sqcup  (j^{\prime\prime} \setminus j)
 		      $
-		\item the map $f_1 = [g_1,h_1]: n'' \to \mathcal H$ in the diagram above consists of $g_1$ as defined in Definition \ref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i'' \to \mathcal R \to \mathcal H$ to $i'' \setminus i$,  and similarly for the map $f_2: k'' \to \mathcal H$.
+		\item the map $f_1 = [g_1,h_1]\colon n^{\prime\prime} \to \mathcal H$ in the diagram above consists of $g_1$ as defined in~\autoref{def:boundary_new} of boundary complement,  and $h_1$ which is the restriction of the composite $i^{\prime\prime} \to \mathcal R \to \mathcal H$ to $i^{\prime\prime} \setminus i$,  and similarly for the map $f_2\colon k^{\prime\prime} \to \mathcal H$.
 	\end{enumerate}
-	% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
+	% The internal interfaces of the resulting closed e-hypergraph $\mathcal{H}$ from the cospan
 	%  \[
-	%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
+	%     \mathcal{H} \xleftarrow{} n^{\prime\prime} \sqcup k^{\prime\prime} \xleftarrow{} n \sqcup k
 	% \]
-	Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ for its reflexive, transitive and symmetric closure.
+	Given  a set $\mathfrak{R}$ of EDPOI rewrite rules and closed e-hypergraphs with extended interfaces $\mathcal G$ and $\mathcal H$,  we write $\mathcal G \Rrightarrow_\mathfrak{R} \mathcal H$ if there exists a EDPOI rewrite in $\mathfrak R$ such that via it $\mathcal G$ rewrites to $\mathcal H$,  and we write $\mathcal G \Rrightarrow^*_\mathfrak{R} \mathcal H$ for its reflexive, transitive and symmetric closure.
 \end{definition}
-% Together, the conditions ensure the result of a rewrite is a well-typed MDA e-hypergraph.
+% Together, the conditions ensure the result of a rewrite is a well-typed MDA closed e-hypergraph.
 
-An example of EDPOI rewrite can be seen in Appendix~\ref{sec:appendix:dpoi}.
-The enhanced definition of extended DPOI ensures that the combinatorial representation of closed semilattice-enriched SMCs is sound and complete for rewriting of closed $\Sigma^{+}$-terms.
-We use the term DPOI to mean EDPOI when it is clear from the context.
+An example of EDPOI rewrite can be seen in~\autoref{sec:appendix:dpoi}.
+The enhanced definition of extended DPOI ensures that the combinatorial representation of morphisms in closed symmetric monoidal $\catname{SLat}$-categories is sound and complete for rewriting of closed $\Sigma^{+}$-terms.

### 2025-07-10 POPL2026/paper/sections/introduction.tex
lots of polish

@@
 
-E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \textit{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
+E-graphs can efficiently simultaneously represent several equivalent terms of an algebraic theory, generalising the \emph{directed acyclic graph} (DAG) representation of terms to include equivalence classes of subterms (or subgraphs).
 The key concepts of e-graphs can be intuitively grasped from some simple examples: \autoref{fig:e-graph-example} shows how a series of rewrite rules can be non-destructively applied to an e-graph~\cite{EggPaper}.
@@
 Using de Bruijn indices instead when performing $\beta$-reduction would only push the problem elsewhere, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
-Recent work on \textit{slotted} e-graphs~\cite{slotted-egraphs} uses nominal techniques to address the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by incorporating variables as first-class data of an e-graph and making binding a built-in operation.
+Recent work on \emph{slotted} e-graphs~\cite{slotted-egraphs} uses nominal techniques to address the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by incorporating variables as first-class data of an e-graph and making binding a built-in operation.
 They also solved the problem of explicit variable names, guaranteeing freshness inside binders, explicitly differentiating between bound and free variables.

### 2025-07-10 POPL2026/paper/sections/soundness.tex
lots of polish

@@
 
-In this section we explore soundness and completeness of DPOI rewriting introduced in previous section with respect to rewriting of closed $\Sigma^{+}$-terms.
-First we recall the rewriting of $\Sigma$-terms modulo SMC laws.
+In this section we explore soundness and completeness of EDPOI rewriting with respect to rewriting of closed $\Sigma^{+}$-terms.
+First we recall the rewriting of $\Sigma$-terms modulo the laws of symmetric monoidal categories.
 
-\begin{definition}[(Closed) $\Sigma$-term rewrite rule]
+\begin{definition}[Closed $\Sigma$-term rewrite rule]
 	A $\Sigma$-term rewrite rules is a pair $\langle l, r \rangle$ where $l$ and $r$ are $\Sigma$ terms with matching inputs and outputs (as lists of input and output types).
@@
 
-\begin{definition}
+\begin{definition}%
 	\label{def:rewrite}
-	We say that a $\Sigma$-term $f$ rewrites to a $\Sigma$-term $g$ modulo SMC laws via a rewrite rule $\langle l, r \rangle$ (notation, $\leadsto_{\langle l, r \rangle}$) if they are representable as
+	We say that a $\Sigma$-term $f$ rewrites to a $\Sigma$-term $g$ modulo the laws of symmetric monoidal categories via a rewrite rule $\langle l, r \rangle$, writing $\leadsto_{\langle l, r \rangle}$, if they are representable as
 	$
@@
 \begin{definition}
-	For closed-$\Sigma$-terms $f,g$, and a closed rewrite rule $\langle l, r \rangle$, $f \leadsto_{\langle l, r \rangle} g$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as
+	For closed-$\Sigma$-terms $f,g$, and a closed rewrite rule $\langle l, r \rangle$, $f \leadsto_{\langle l, r \rangle} g$ modulo the laws of symmetric monoidal categories if they are representable as either in~\autoref{def:rewrite}, or as
 	$
@@
 		f_{1} + \ldots + f_{n}
-	$ modulo SMC laws and $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} where none of $f_{i}$ contain the join operator.
+	$ modulo the laws of symmetric monoidal $\catname{SLat}$-categories and distributivity law (\autoref{law:distributivity}) where none of $f_{i}$ contain the join operator.
 \end{lemma}
@@
 % Such forms are given by orienting $\catname{SLat}$-equations such that $+$ gets propagated to the top, as, for example $f \otimes (g + h) = f \otimes g + g \otimes h$.
-% Most notably, we can also escape lambda abstractions by using distributivity law~\ref{law:distributivity}.
+% Most notably, we can also escape lambda abstractions by using the distributivity law (\autoref{law:distributivity}).
 % \end{proof}
@@
 \begin{definition}
-	For closed $\Sigma^{+}$-term $f,g$, and a closed $\Sigma$-term rewrite rule $\langle l, r \rangle$ $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
+	For closed $\Sigma^{+}$-term $f,g$, and a closed $\Sigma$-term rewrite rule $\langle l, r \rangle$ $f \leadsto_{\langle l, r \rangle}$ modulo the laws of symmetric monoidal $\catname{SLat}$-categories and distributivity law (\autoref{law:distributivity}) if they are representable as
 	$
@@
 Next, to formulate the correspondence between rewriting systems we first define interpretation $\llbracket - \rrbracket$ of closed $\Sigma^{+}$ terms in $\WellTypedMdaEcospans$.
-It follows by first defining the interpretation of operators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in Appendix~\ref{sec:appendix:interpretation}.
-The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in Fig.~\ref{fig:ev_and_lambda}.
+It follows by first defining the interpretation of operators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in~\autoref{sec:appendix:interpretation}.
+The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in~\autoref{fig:ev_and_lambda}.
 
@@
 
-Then we can make $\WellTypedMdaEcospans$ into $\catname{SLat}$-category by defining a join of two cospans and introducing DPOI rewrite rules to satisfy the laws of $\catname{SLat}$ symmetric monoidal category.
-We also add a structural rewrite schema rule corresponding to distributivity law~\ref{law:distributivity}.
-The join of two cospans is defined as in Fig.~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules, one for each axiom.
+Then we can make $\WellTypedMdaEcospans$ into $\catname{SLat}$-category by defining a join of two cospans and introducing EDPOI rewrite rules to satisfy the laws of symmetric monoidal $\catname{SLat}$-categories.
+We also add a structural rewrite schema rule corresponding to the distributivity law (\autoref{law:distributivity}).
+The join of two cospans is defined as in~\autoref{fig:f+g} and the quotienting is done through EDPOI rewrite schema rules, one for each axiom.
 For example, the distributivity rule for $\otimes$ is given by the following schema rule
@@
 
-We collect all such rewrite schemas into a set $\mathcal{S}$ and then a category $\WellTypedMdaEcospans / \mathcal{S}$ is an $\catname{SLat}$ SMC.
-Note that it is not a \textit{closed} $\catname{SLat}$ SMC.
-To turn it into a \textit{closed} $\catname{SLat}$ SMC, we would need to also quotient by the DPO rewrites that arise from the equations in Def~\ref{def:closed}.
-We do not perform such quotienting as these equations bear operational meaning, \textit{e.g.}, they model $\beta$-reduction when interpreting $\lambda$-calculus.
+We collect all such rewrite schemas into a set $\mathcal{S}$ and then determine a symmetric monoidal $\catname{SLat}$-category $\WellTypedMdaEcospans / \mathcal{S}$ by this quotienting.
+Note that it is not a \emph{closed} symmetric monoidal $\catname{SLat}$-category.
+To make it so, we would need to also quotient by the DPO rewrites that arise from the equations in~\autoref{def:closed}.
+We do not perform such quotienting as these equations bear operational meaning, i.e.\ they model $\beta$-reduction when interpreting $\lambda$-calculus.
 
 \begin{definition}[Quotient by rewrites]
-	Given a set of DPOI rewrite rules $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the  relation
-	$
-		f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\mathcal{E}} g
-	$.
+	Given a set of EDPOI rewrite rules $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the $\catname{SLat}$-category $\WellTypedMdaEcospans$ quotiented by the relation $f \sim g \quad \text{if} \quad f \Rrightarrow^{*}_{\mathcal{E}} g$.
 \end{definition}
 
-This explicit quotienting allows us to only consider cospans of the form $f_1 + \ldots + f_{n}$, similar to Lemma~\ref{lemma:normal_form}, such that carrier of $f_{i}$ does not contain any hierarchical edges from $\textcolor{e-color}{E}$.
+This explicit quotienting allows us to only consider cospans of the form $f_1 + \ldots + f_{n}$, similar to~\autoref{lemma:normal_form}, such that carrier of $f_{i}$ does not contain any hierarchical edges from $\textcolor{e-color}{E}$.
 
@@
 	\label{prop:fscd}
-	For two closed $\Sigma$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$, where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
+	For two closed $\Sigma$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$, where the latter is defined in a sub $\catname{SLat}$-category of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
 \end{proposition}
@@
 	such that $f_{i} \leadsto f_{j}$ for some $i, j$.
-	This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and, by noting that every $f_{i}$ is a $\Sigma$-term, and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{e-color}{E}$, we can apply Proposition~\ref{prop:fscd}.
+	This yields two cospans of the form $\llbracket f_{1} \rrbracket + \ldots + \llbracket f_{n} \rrbracket$ for $\llbracket f \rrbracket$ and $\llbracket g \rrbracket$ and, by noting that every $f_{i}$ is a $\Sigma$-term, and carriers in every $\llbracket f_{i} \rrbracket$ have no edges from $\textcolor{e-color}{E}$, we can apply~\autoref{prop:fscd}.
 \end{proof}

### 2025-07-11 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
touch up conclusion, fix broken references, move empirical analysis to appendix

@@
 In this work we introduced e-graphs with bindings formalised as morphisms of a closed symmetric monoidal $\catname{SLat}$-category.
-We then provided a combinatorial representation for such morphisms in terms of hierarchical hypergraphs (e-hypergraphs) with a suitable notion of sound and complete DPO rewriting, making it possible for an e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
-An important future work is exploring whether such combinatorial structures may lead to better algorithmics for doing equality saturation for $\lambda$-calculus.
+In doing so, we are able to leverage its internal language to provide variable binding in $\lambda$-calculus-like languages by metatheoretic $\lambda$-abstraction.
+We then provided a combinatorial representation for such morphisms in terms of closed e-hypergraphs with a suitable notion of sound and complete DPO rewriting, making it possible for a closed e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
+We argued for the utility of our techniques in the representation of programming languages (which inherently require variable binding) as e-graphs, contrasting with the more general-purpose slotted e-graphs, and advocating for the benefits of string-diagrammatic syntax.
+In future, we hope to implement an efficient version of e-graphs for hypergraphs as a data structure using DPO rewriting as its foundation to implement e-graph operations, with the theory we have developed here acting as a roadmap.
 

### 2025-07-11 POPL2026/paper/sections/appendix.tex
touch up conclusion, fix broken references, move empirical analysis to appendix

@@
 \end{figure}
+
+\section{Modelling $\lambda_{\text{lsub}}$-calculus in slotted e-graphs}%
+\label{sec:e-graph-explicit-subst}
+
+\begin{table}
+	\begin{tabular}{lcc}
+		\texttt{beta}                  & $(\lambda x . t) u = t[x / u]$                       &                                                                    \\
+		\texttt{subst\_same\_var}      & $x[x / u] = u$                                       &                                                                    \\
+		\texttt{subst\_diff\_var}      & $y[x / u] = y$                                       & if $x \not \in \mathcal{F}(u)$                                     \\
+		\texttt{subst\_application}    & $(uu')[x / v] = u[x / v]u'[x / v]$                   &                                                                    \\
+		\texttt{subst\_lambda}         & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$    & if $x \not \in \mathcal{F}(v)$                                     \\
+		\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x' / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$             \\
+		\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])$          & if $x' \not \in \mathcal{F}(u)$                                    \\
+		\texttt{subst\_comm}           & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$          & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)~.$ \\
+	\end{tabular}
+	\caption{Encoding of string diagram equivalence and explicit substitution for $\lambda_{\text{lsub}}$}
+	\label{tbl:e-graph-subst-rules}
+\end{table}
+
+Consider the equational presentation of string diagram equivalence and the $\lambda_{\text{lsub}}$ reduction rules given by~\autoref{tbl:e-graph-subst-rules}.
+These can be encoded in the framework of slotted e-graphs.
+To demonstrate the effect that a particular set of rewrites has on the number of e-nodes of an e-graph, we saturate the lambda term $\text{plus}(2, 2)$ where $\text{plus}$ and $2$ are encoded using Church encoding.
+The numbers of e-nodes after the e-graph is saturated for each set of rewrite rules are presented in~\autoref{tbl:slotted}.
+
+\begin{table}
+	\begin{tabular}{lcccc}
+		Set of equations & explicit subst & explicit subst & explicit subst & explicit subst \\
+		                 &                & + comm         & + interchange  & + interchange  \\
+		                 &                &                &                & + comm         \\
+		\# of e-nodes    & 161            & 288            & 289            & 356
+	\end{tabular}
+	\caption{Slotted e-graphs benchmarks}
+	\label{tbl:slotted}
+\end{table}
+
+The \enquote{explicit subst} set of rewrite rules consist of the first five rules of~\autoref{tbl:e-graph-subst-rules}.
+We can see that additionally saturating the e-graph with respect to interchange and commutativity rules increases the number of e-nodes by a factor of 2.
+These are the exact equations that are absorbed by using string-diagrammatic representation of $\lambda_{\text{lsub}}$-terms.
+Notably, string diagrams (and (closed e-)hypergraphs) do not absorb the equation \texttt{subst\_lambda} but we believe it is practically feasible to do rewriting modulo this rule by traversing through the boxes when searching for a redex.
+Furthermore, (cospans of) closed e-hypergraphs, as presented in~\autoref{sec:closed-e-hyp}, do not absorb the equations for associativity, commutativity and unitality for $\comonoid$ and $\counit$, but the constraints on the cospans and the boundary complement may be relaxed to absorb those as well by following the work of~\cite{zanassi_comonoid} with which our approach is completely compatible.

### 2025-07-11 POPL2026/paper/sections/categorical_semantics.tex
touch up conclusion, fix broken references, move empirical analysis to appendix

@@
 
-To aid reasoning, we introduce a new language of string diagrams for closed symmetric monoidal $\catname{SLat}$-categories, using a hierarchical \enquote{dashed box} structure to capture the join operation on morphisms (as was done by~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}) as well as another hierarchical \enquote{rounded box} for $\lambda$-abstraction.
-\autoref{fig:egraph-strings} displays the generators of this language, which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for $\lambda$-abstraction as established by~\citet{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice enrichment as introduced by~\citet{ghica2024equivalencehypergraphsegraphsmonoidal}; the first component denotes an empty diagram.
+To aid reasoning, we introduce a new language of string diagrams for closed symmetric monoidal $\catname{SLat}$-categories, using a hierarchical \enquote{dashed box} structure to capture the join operation on morphisms (as was done by~\citet{tiurin2025equivalencehypergraphsdporewriting}) as well as another hierarchical \enquote{rounded box} for $\lambda$-abstraction.
+\autoref{fig:egraph-strings} displays the generators of this language, which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for $\lambda$-abstraction as established by~\citet{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice enrichment as introduced by~\citet{tiurin2025equivalencehypergraphsdporewriting}; the first component denotes an empty diagram.
 

### 2025-07-11 POPL2026/paper/sections/introduction.tex
touch up conclusion, fix broken references, move empirical analysis to appendix

@@
 This is achieved by making rewrites on e-graphs non-destructive: an application of a rewrite rule, instead of replacing the left-hand side by the right-hand side, adds the latter to the e-graph so that all pre-existing redexes are protected and new ones are introduced.
-Explosive growth of the e-graph is avoided by the clever use of sharing of vertices.
+Exponential growth of the e-graph is avoided by the clever use of sharing of vertices.
 
-E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using the technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, flatt2022_small, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint (or timeout) is reached.
+E-graphs were first introduced in the automated theorem proving literature in the 1980s~\cite{nelson1980techniques}, but recent years have seen renewed interest, particularly for optimisation and synthesis using the technique of \emph{equality saturation}~\cite{10.1145/1594834.1480915, flatt_small_2022, EggPaper,flatt_small_2022}, wherein rewrites to an e-graph are repeated until a fixpoint (or timeout) is reached.
 The wide range of application domains led to extensions of the e-graph formalism, for instance to account for conditional rewriting~\cite{singher2023colored} or  to combine e-graphs with database queries (\emph{relational e-matching})~\cite{zhang_relational_2022}.
@@
 
-First, let us compare how substitution works in conventional e-graphs as compared to our proposed approach.
+First, let us compare how substitution works in conventional e-graphs, which lack specialised support for binding, as compared to our proposed approach.
 

### 2025-07-11 POPL2026/paper/sections/lambda-calculus.tex
touch up conclusion, fix broken references, move empirical analysis to appendix

@@
 
-This calculus is explicitly designed to capture the representation of terms as proof nets: for two terms $t, u$ in the $\lambda_{\text{lsub}}$-calculus, $t \sim u$ if and only if both map to the same proof net~\cite{?}.
+This calculus is explicitly designed to capture the representation of terms as proof nets: for two terms $t, u$ in the $\lambda_{\text{lsub}}$-calculus, $t \sim u$ if and only if both map to the same proof net~\cite{accattoliJumpingBoxGraphical2011}.
 By further extending the equivalence relation $\sim$, taking the symmetric closure by $\to_{ls}$ and $\to_{gc}$ and thus identifying terms under elimination of explicit substitution, we determine a \emph{string diagram equivalence} $\approx$ of $\lambda_{\text{lsub}}$-terms, determining a $\lambda^\approx_{\text{lsub}}$-calculus.
@@
 \autoref{fig:plus-1-2-programs} shows some programs which are semantically equivalent but syntactically distinct.
-As alluded to in~\autoref{sec:vs-e-graphs-with-binding}, the traditional (slotted) e-graph approach to modelling the equivalence of these terms (\autoref{fig:plus-1-2-terms}) relies on explicitly encoding substitution operations as e-graph nodes, and their behaviour (the string diagram equivalence relation $\approx$) as rewrite rules, as in~\autoref{fig:e-graph-substitution} --- not only is this more complicated, but it also leads to a blow-up in the state space required to perform equality saturation in any extension of the $\lambda^\approx_{\text{lsub}}$-calculus, because we are really rewriting $\lambda_{\text{lsub}}$-terms and computational resources are spent rewriting between equivalent terms.
-\textcolor{cyan}{To illustrate this, let's consider the following equations induced by the graphical equivalence and the $\lambda_{\text{lsub}}$ reduction rules}
-
-\begin{tabular}{lcc}
-	\texttt{beta}                  & $(\lambda x . t) u = t[x / u]$                       &                                                                    \\
-	\texttt{subst\_same\_var}      & $x[x / u] = u$                                       &                                                                    \\
-	\texttt{subst\_diff\_var}      & $y[x / u] = y$                                       & if $x \not \in \mathcal{F}(u)$                                     \\
-	\texttt{subst\_application}    & $(uu')[x / v] = u[x / v]u'[x / v]$                   &                                                                    \\
-	\texttt{subst\_lambda}         & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$    & if $x \not \in \mathcal{F}(v)$                                     \\
-	\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x' / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$             \\
-	\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])$          & if $x' \not \in \mathcal{F}(u)$                                    \\
-	\texttt{subst\_comm}           & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$          & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)~.$ \\
-\end{tabular}
-
-\textcolor{cyan}{These can be encoded in a framework of slotted e-graphs.
-	To demonstrate the effect that a particular set of rewrites has on the number of e-nodes of an e-graph, we saturate the lambda term $plus 2 2$ where $plus$ and $2$ are encoded using Church encoding.
-	The numbers of e-nodes after the e-graph is saturated for each set of rewrite rules are presented in \autoref{tbl:slotted}.
-}
-
-\begin{table}
-	\begin{tabular}{lcccc}
-		Set of equations & explicit subst & explicit subst & explicit subst & explicit subst \\
-		                 &                & + comm         & + interchange  & + interchange  \\
-		                 &                &                &                & + comm         \\
-		\# of e-nodes    & 161            & 288            & 289            & 356
-	\end{tabular}
-	\caption{Slotted e-graphs benchmarks}
-	\label{tbl:slotted}
-\end{table}
-
-\textcolor{cyan}{The \texttt{explicit subst} set of rewrite rules consist of the first five rules above.
-	We can see that additionally saturating the e-graph with respect to interchange and commutativity rules increases the number of e-nodes by a factor of 2.
-	These are the exact equations that are absorbed by using string-diagrammatic representation of $\lambda$-terms.
-	Notably, string diagrams (and (e-)hypergraphs) do not absorb the equation \texttt{subst\_lambda} but we believe it is practically feasible to do rewriting modulo this rule by traversing through the boxes when searching for a redex.
-	Furthermore, (cospans of) e-hypergraphs, as presented in \autoref{sec:closed-e-hyp}, do not absorb the equations for associativity, commutativity and unitality for $\comonoid$ and $\counit$, but the constraints on the cospans and the boundary complement may be relaxed to absorb those as well by following the work of~\cite{zanassi_comonoid} with which our approach is completely compatible.}
-
+As alluded to in~\autoref{sec:vs-e-graphs-with-binding}, the traditional (slotted) e-graph approach to modelling the equivalence of these terms (\autoref{fig:plus-1-2-terms}) relies on explicitly encoding substitution operations as e-graph nodes, and their behaviour (the string diagram equivalence relation $\approx$) as rewrite rules, as in~\autoref{fig:e-graph-substitution} --- not only is this more complicated, but it also leads to a blow-up in the state space required to perform equality saturation in any extension of the $\lambda^\approx_{\text{lsub}}$-calculus, because we are really rewriting $\lambda_{\text{lsub}}$-terms and computational resources are spent rewriting between equivalent terms (an empirical analysis is provided in~\autoref{sec:e-graph-explicit-subst}).
 In contrast, in our framework, $\lambda^\approx_{\text{lsub}}$-terms are directly modelled as string diagrams (\autoref{fig:plus-1-2-string-diagram}\footnote{Note that dashed boxes are elided for convenience here because no rewriting has occurred.}).
-We do however note that the conventional approach is more flexible, in the sense that it may model more theories with binders (e.g.\ $\pi$-calculus), whereas our approach is essentially specialised to $\lambda$-calculus, relying heavily on the metatheoretic internal language of the ambient setting.
+We do however note that the slotted approach is more flexible, in the sense that it may model more theories with binders (e.g.\ $\pi$-calculus), whereas our approach is essentially specialised to $\lambda$-calculus, relying heavily on the metatheoretic internal language of the ambient setting.
 

### 2025-07-11 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
save the trees
cosmetic changes

@@
 \usepackage{hyperref}
+\renewcommand{\sectionautorefname}{Section}
+\renewcommand{\subsectionautorefname}{Subsection}
+
+\usepackage{cleveref}
 \usepackage{stmaryrd}
@@
 
+\usepackage{aliascnt}
+
+% space saving
+\usepackage[moderate,bibbreaks=normal,bibnotes=normal,bibliography=normal,leading=normal]{savetrees}
+
 \newcommand{\ONECOLUMN}{}
@@
 \newtheorem{theorem}{Theorem}[section]
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{lemma}[theorem]{Lemma}
+
+% definition
+\newaliascnt{definition}{theorem}
+\newtheorem{definition}[definition]{Definition}
+\aliascntresetthe{definition}
+\providecommand*{\definitionautorefname}{Definition}
+
+\newaliascnt{example}{theorem}
+\newtheorem{example}[example]{Example}
+\aliascntresetthe{example}
+\providecommand*{\exampleautorefname}{Example}
+
+\newaliascnt{proposition}{theorem}
+\newtheorem{proposition}[proposition]{Proposition}
+\aliascntresetthe{proposition}
+\providecommand*{\propositionautorefname}{Proposition}
+
+\newaliascnt{remark}{theorem}
+\newtheorem{remark}[remark]{Remark}
+\aliascntresetthe{remark}
+\providecommand*{\remarkautorefname}{Remark}
+
+\newaliascnt{lemma}{theorem}
+\newtheorem{lemma}[lemma]{Lemma}
+\aliascntresetthe{lemma}
+\providecommand*{\lemmaautorefname}{Lemma}
 
@@
 \begin{abstract}
-	Equality saturation, a technique for program optimisation and reasoning, has gained attention due to the resurgence of equality graphs (e-graphs).
-	E-graphs represent equivalence classes of terms under rewrite rules, enabling simultaneous rewriting across a family of terms.
-	However, they struggle in domains like $\lambda$-calculus that involve variable binding, due to a lack of native support for bindings.
-	Building on recent work interpreting e-graphs categorically as morphisms in symmetric monoidal semilattice-enriched categories, we extend this framework to closed symmetric monoidal semilattice-enriched categories to handle bindings.
-	We provide a concrete combinatorial representation using an extension of hierarchical hypergraphs and introduce a corresponding double-pushout (DPO) rewriting mechanism.
-	Finally, we establish the equivalence of term rewriting and DPO rewriting, with the key property that the combinatorial representation absorbs the equations of the symmetric monoidal structure.
+	% Equality saturation, a technique for program optimisation and reasoning, has gained attention due to the resurgence of equality graphs (e-graphs).
+	% E-graphs represent equivalence classes of terms under rewrite rules, enabling simultaneous rewriting across a family of terms.
+	% However, they struggle in domains like $\lambda$-calculus that involve variable binding, due to a lack of native support for bindings.
+	% Building on recent work interpreting e-graphs categorically as morphisms in symmetric monoidal semilattice-enriched categories, we extend this framework to closed symmetric monoidal semilattice-enriched categories to handle bindings.
+	% We provide a concrete combinatorial representation using an extension of hierarchical hypergraphs and introduce a corresponding double-pushout (DPO) rewriting mechanism.
+	% Finally, we establish the equivalence of term rewriting and DPO rewriting, with the key property that the combinatorial representation absorbs the equations of the symmetric monoidal structure.
+    Equality saturation is a powerful technique for program optimisation and reasoning, driven by the use of equivalence classes of terms under rewrite rules called equality graphs (e-graphs). 
+    Despite their numerous advantages, until recently e-graphs lacked native support for variable binding, limiting their applicability to programming languages. 
+    We propose to address this problem from a categorical perspective, interpreting e-graphs as string diagrams, namely morphisms in symmetric monoidal semilattice-enriched categories with a monoidal closed structure.
+    We further define a concrete representation using extended hierarchical hypergraphs, and introduce a corresponding double-pushout (DPO) rewriting system. 
+    Finally, we establish the equivalence between term rewriting and DPO rewriting, with the combinatorial model inherently absorbing the equations of the symmetric monoidal structure. 
+    Our approach, specifically designed for lambda calculi, is compared and contrasted with slotted e-graphs—a recent, more generic method for incorporating variable binding.
 \end{abstract}

### 2025-07-11 POPL2026/paper/sections/appendix.tex
cosmetic changes

@@
 
+\section{Hypergraphs}
+\label{sec:appendix:hyp}
+\autoref{fig:hypergraph-with-interfaces} gives an example of an MDA cospan, which is such because it \enquote{looks like} the string diagram \comonoid.
+Analogously, we can see that~\autoref{fig:hypergraph} (equipped with the interfaces choosing $\{ v_1 \}$ and $\{ v_4, v_5, v_6, v_7 \}$) fails to form an MDA cospan --- it is not monogamous for many reasons: $v_1$ has out-degree 0, $v_1, v_5, v_7$ have in-degree 2, and $v_3$ has in-degree 0; additionally, it fails be be directed acyclic due to the path along $e_4$ from $v_7$ to itself\footnote{Recall that loops are not part of the syntax of string diagrams for symmetric monoidal categories in the absence of a trace --- a relaxation of monogamy to allow for this is explored by~\citet{ghica_rewriting_2023}.}.
+
+\begin{figure}
+	\begin{subfigure}[c]{0.3\linewidth}
+		\centering
+		% \begin{tikzpicture}[
+		% 		vertex/.style={draw, circle, fill, inner sep=1.5pt},
+		% 		hyperedge/.style={draw, rectangle}
+		% 	]
+		% 	\node[vertex, label=left:{$v_4$}] (v1) at (0,0) {};
+		% 	\node[vertex, label=left:{$v_5$}] (v2) at (1,0) {};
+		% 	\node[vertex, label=left:{$v_6$}] (v3) at (2,0) {};
+		% 	\node[vertex, label=left:{$v_7$}] (v4) at (3,0) {};
+		% 	\node[vertex, label=left:{$v_2$}] (v5) at (0.5,-2) {};
+		% 	\node[vertex, label=left:{$v_3$}] (v6) at (2,-2) {};
+		% 	\node[vertex, label=left:{$v_1$}] (v7) at (2,-3) {};
+
+		% 	\node[hyperedge] (e1) at (0.5,-1) {$e_3$};
+		% 	\node[hyperedge] (e2) at (2,-1) {$e_4$};
+		% 	\node[hyperedge] (e3) at (0.5,-3) {$e_1$};
+		% 	\node[hyperedge] (e4) at (2,-2.5) {$e_2$};
+
+		% 	\draw (v1) -- (e1);
+		% 	\draw (v2) -- (e1);
+		% 	\draw (v5) -- (e1);
+		% 	\draw (v5) edge[bend right] (e3);
+		% 	\draw (v5) edge[bend left] (e3);
+		% 	\draw (v2) edge[in=130,out=-90] (e2);
+		% 	\draw (v3) -- (e2);
+		% 	\draw (v4) edge[in=50,out=-90] (e2);
+		% 	\draw (v4) edge[in=-50,out=-90] (e2);
+		% 	\draw (v6) -- (e2);
+		% \end{tikzpicture}
+        \[
+        \adjustbox{scale=0.8}{
+        \tikzfig{figures/hypergraph_example}
+        }
+        \]
+		\caption{Hypergraph with 7 vertices and 4 hyperedges, all directed bottom-to-top; source/target vertices are ordered left-to-right.}%
+		\label{fig:hypergraph}
+	\end{subfigure}
+	\begin{subfigure}[c]{0.3\linewidth}
+		\centering
+		% \begin{tikzpicture}[
+		% 		vertex/.style={draw, circle, fill, inner sep=1.5pt},
+		% 		hyperedge/.style={draw, rectangle}
+		% 	]
+		% 	\node[vertex, label=left:{$v_3$}] (v1) at (0,0) {};
+		% 	\node[vertex, label=left:{$v_3$}] (v1') at (0,1.5) {};
+		% 	\node[vertex, label=left:{$v_2$}] (v2) at (1,0) {};
+		% 	\node[vertex, label=left:{$v_2$}] (v2') at (1,1.5) {};
+		% 	\node[vertex, label=left:{$v_1$}] (v3) at (0.5,-2) {};
+		% 	\node[vertex, label=left:{$v_1$}] (v3') at (0.5,-3.5) {};
+
+		% 	\node[hyperedge] (e1) at (0.5,-1) {$e_1$};
+
+		% 	\draw (v1) -- (e1);
+		% 	\draw (v2) -- (e1);
+		% 	\draw (v3) -- (e1);
+
+		% 	\node[draw, rounded corners, fit={(v1'.west) (v2'.west)}, inner sep=12pt] (S) {};
+		% 	\node[draw, rounded corners, fit={(v3'.west)}, inner sep=12pt] (T) {};
+		% 	\node[draw, rounded corners, fit={(v1.west) (v2.west) (v3.west)}, inner sep=12pt] (H) {};
+		% 	\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (S) -- node[left] {$g$} (H);
+		% 	\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (T) -- node[left] {$f$} (H);
+		% \end{tikzpicture}
+        \[
+        \adjustbox{scale=0.8}{
+        \tikzfig{figures/cospan_hypergraph_example}
+        }
+        \]
+		\caption{Hypergraph with interfaces}%
+		\label{fig:hypergraph-with-interfaces}
+	\end{subfigure}
+	\caption{Example hypergraphs}%
+	\label{fig:hypergraph-examples}
+\end{figure}
+
 \section{Pushout in $\catname{EHyp}(\Sigma)$}
@@
 \begin{remark}
-	$\textcolor{e-color}{<}$, $\textcolor{closed-color}{<}$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
-	To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
-	For example, an immediate successor of a vertex $x$ can be written functionally as $\textcolor{e-color}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))$.
+	$\textcolor{e-color}{<}$, $\textcolor{closed-color}{<}$ and $\consistency$ can be considered as (partial) functions defined on $V_{\mathcal{F}} \sqcup  E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
+	To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} \sqcup  E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} \sqcup  E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
+	For example, the immediate predecessor of a vertex $x$ can be written functionally as $\textcolor{e-color}{<_{\mathcal{F}}^{\mu}}(\iota_{V_{\mathcal{F}}}(x))$.
 \end{remark}
@@
 		      \]
-		      where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
+		      where $\phi_{V};\iota_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} \sqcup  E_{\mathcal{G}}$, and similarly for $\phi_{E};\iota_{E_\mathcal{G}}$ so that $[\phi_{V};\iota_{V_{\mathcal{G}}}, \phi_{E};\iota_{E_{\mathcal{G}}}] : V_{\mathcal{F}} \sqcup  E_{\mathcal{F}} \to  V_{\mathcal{G}} \sqcup  E_{\mathcal{G}}$.
 		\item When $x \in V_{\mathcal{F}}$
@@
 	\end{enumerate}
-	then the pushout $X +_{f,g} Y$ exists.
+	then the pushout $X \sqcup _{f,g} Y$ exists.
 \end{theorem}
@@
 				Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
-				Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
+				Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X\sqcup Y}{\sim R} \arrow[rd, "u"]                              &   \\
 				&                                                                  & Q
@@
 	Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
-	First, a coproduct of $X+Y$ is computed which is
+	First, a coproduct of $X\sqcup Y$ is computed which is
 	\[
-		X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, \textcolor{e-color}{<_{X+Y}}, \textcolor{closed-color}{<_{X+Y}}\}
+		X \sqcup  Y = \{V_{X} \sqcup  V_{Y}, E_{X} \sqcup  E_{Y}, s_{X\sqcup Y}, t_{X\sqcup Y}, \consistency_{X\sqcup Y}, \textcolor{e-color}{<_{X\sqcup Y}}, \textcolor{closed-color}{<_{X\sqcup Y}}\}
 	\]
-	where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, \textcolor{e-color}{<_{X+Y}}, \textcolor{closed-color}{<_{X+Y}}, \consistency_{X+Y}$.
+	where $s_{X\sqcup Y} : E_{X} \sqcup  E_{Y} \to (V_{X} \sqcup  V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} \sqcup  E_{Y} \to (V_{X} \sqcup  V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} \sqcup  V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} \sqcup  V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X\sqcup Y}, \textcolor{e-color}{<_{X\sqcup Y}}, \textcolor{closed-color}{<_{X\sqcup Y}}, \consistency_{X\sqcup Y}$.
 	We will omit labels as they are irrelevant to pushout construction.
@@
 		S_{V} & = \{
-		\;(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |                                               \\
+		\;(x_i,y_j) \in (V_{X} \sqcup  V_{Y}) \times (V_{X} \sqcup  V_{Y})\; |                                               \\
 		      & \;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z) \\
@@
 		      & \{
-		\;(y_j,x_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; |                                               \\
+		\;(y_j,x_j) \in (V_{X} \sqcup  V_{Y}) \times (V_{X} \sqcup  V_{Y})\; |                                               \\
 		      & \;\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z) \\
@@
 		      & \{
-		\;(x,x)\;\text{where}\; x \in V_{X} + V_{Y}
+		\;(x,x)\;\text{where}\; x \in V_{X} \sqcup  V_{Y}
 		\}                                                                                                       \\
 		\\
-		S_{E} & = \{(x,x) \text{ where } x \in E_{X} + E_{Y}\}
+		S_{E} & = \{(x,x) \text{ where } x \in E_{X} \sqcup  E_{Y}\}
 	\end{align*}
@@
 
-	We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
+	We then quotient the set vertices and edges in $X \sqcup  Y$ by $R_{V}$ and $R_{E}$ respectively.
 
 	\begin{align*}
-		\sfrac{X + Y}{\sim (R_{V},R_{E})} = & (
-		\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})},                                                             \\
-		                                    & \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})},                       \\
-		                                    & \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},                             \\
+		\sfrac{X \sqcup  Y}{\sim (R_{V},R_{E})} = & (
+		\sfrac{V_{X} \sqcup  V_{Y}}{\sim (R_{V},R_{E})},                                                             \\
+		                                    & \sfrac{E_{X} \sqcup  E_{Y}}{\sim (R_{V},R_{E})},                       \\
+		                                    & \sfrac{s_{X\sqcup Y}}{\sim (R_{V},R_{E})},                             \\
 		                                    & \sfrac{t_{X,Y}}{\sim (R_{V},R_{E})},                             \\
-		                                    & \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}},                  \\
-		                                    & \textcolor{e-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}       \\
-		                                    & \textcolor{closed-color}{<}_{\sfrac{X + Y}{\sim (R_{V},R_{E})}})
+		                                    & \consistency_{\sfrac{X\sqcup Y}{\sim (R_{V},R_{E})}},                  \\
+		                                    & \textcolor{e-color}{<_{\sfrac{X \sqcup  Y}{\sim (R_{V},R_{E})}}}       \\
+		                                    & \textcolor{closed-color}{<_{\sfrac{X \sqcup  Y}{\sim (R_{V},R_{E})}}})
 	\end{align*}
 
-	We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
+	We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{e.g.}, by writing $\sfrac{V_{X} \sqcup  V_{Y}}{\sim}$ and concrete relation will be clear from the context.
 	Where necessary we will refer to $S_{V}$ and $S_{R}$ as $\sim_{S}$.
@@
 	\[
-		\sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
+		\sfrac{s_{X\sqcup Y}}{\sim} : \sfrac{E_{X} \sqcup  E_{Y}}{\sim} \to (\sfrac{V_{X} \sqcup  V_{Y}}{\sim})^{*}
 	\]
-	and there is an obvious surjective function $[]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
-	There is also $[] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
+	and there is an obvious surjective function $[]_{V} : (V_{X} \sqcup  V_{Y}) \to (\sfrac{V_{X} \sqcup  V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[]_{V}^{*}$ is its extension to sequences.
+	There is also $[] : E_{X} \sqcup  E_{Y} \to \sfrac{E_{X} \sqcup  E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
 	We then define
 	\[
-		\sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[]^{*}(e) = [s_{X+Y}(e)]^{*}
+		\sfrac{s_{X\sqcup Y}}{\sim}([e]) = s_{X\sqcup Y};[]^{*}(e) = [s_{X\sqcup Y}(e)]^{*}
 	\]
-	We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
+	We will also use subscripts when it is important to tell if an element of $E_{X} \sqcup  E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
 	Similarly, we will write $v_{x}$ to refer to a vertex with a pre-image in $V_{X}$.
@@
 	\[
-		t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
+		t_{\sfrac{X\sqcup Y}{\sim}}([e]) = [t_{X\sqcup Y}(e)]^{*}
 	\]
@@
 	Recall that $\textcolor{e-color}{<}$ is essentially a transitive closure of $\textcolor{e-color}{<^{\mu}}$ and hence we can first define
-	\[\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
+	\[\textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}[e])
 	\] and
 	\[
-		\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u])
-	\] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
+		\textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}[u])
+	\] where $\iota_*$ are injections into $\sfrac{V_{X} \sqcup  V_{Y}}{\sim} \sqcup  \sfrac{E_{X} \sqcup  E_{Y}}{\sim}$.
 	We will consider several cases.
-	First, assume that $u$ is an element of $V_{X+Y}$ and
+	First, assume that $u$ is an element of $V_{X\sqcup Y}$ and
 	\begin{enumerate}
-		\item If there exists $v$ such that $\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))$ is defined and $u \sim v$
+		\item If there exists $v$ such that $\textcolor{e-color}{<_{X\sqcup Y}^{\mu}}(\iota_{V_{X} \sqcup  V_{Y}}(v))$ is defined and $u \sim v$
 		      we let
 		      \[
-			      \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = [\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v))]
+			      \textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([u])) = [\textcolor{e-color}{<_{X\sqcup Y}^{\mu}}(\iota_{V_{X} \sqcup  V_{Y}}(v))]
 		      \]
-		\item \label{def:child_respects_connectivity} $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
-		      If there exists $v$ such that $\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
+		\item \label{def:child_respects_connectivity} $u$ has no pre-image in $V_{Z}$ and $<_{X\sqcup Y}^{\mu}(\iota_{V_{X} \sqcup  V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$).
+		      If there exists $v$ such that $\textcolor{e-color}{<_{X\sqcup Y}^{\mu}}(\iota_{V_{X} \sqcup  V_{Y}}(v)) = e'$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
 		      \[
-			      \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+			      \textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}[u]) = <_{\sfrac{X\sqcup Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}[v])
 		      \]
@@
 	% If $v' = v$ then it would mean that $[u) \not = \varnothing$ since $[v) \not = \varnothing$ and there is a path from $u$ to $v$.
-	% There is no case when $u$ has a pre-image in $V_{Z}$ and yet $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e$ and there is a path from $[u]$ to $[v]$.
+	% There is no case when $u$ has a pre-image in $V_{Z}$ and yet $<_{X\sqcup Y}^{\mu}(\iota_{V_{X} \sqcup  V_{Y}}(u))$ is undefined and there exists $v$ such that $<_{X\sqcup Y}^{\mu}(\iota_{V_{X} \sqcup  V_{Y}}(v)) = e$ and there is a path from $[u]$ to $[v]$.
 	% The existence of a path would mean that there is $u' \sim u$ and $v' \sim v$ such that there is a path from $u'$ to $v'$ and since $u$ has a pre-image in $V_{Z}$, $u = f_{V};\iota_{1,V}(z_1)$.
@@
 
-	Otherwise, we leave $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))$ undefined.
+	Otherwise, we leave $\textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([u]))$ undefined.
 	% \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
-	Now, assume that $e$ is an element of $E_{X+Y}$.
+	Now, assume that $e$ is an element of $E_{X\sqcup Y}$.
 	Consider two cases.
 	\begin{enumerate}
-		\item  $\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e)) = e'$.
+		\item  $\textcolor{e-color}{<_{X\sqcup Y}^{\mu}}(\iota_{E_{X} \sqcup  E_{Y}}(e)) = e'$.
 		      Then we define
 		      \[
-			      \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = [\textcolor{e-color}{<_{X+Y}^{\mu}}(\iota_{E_{X} + E_{Y}}(e))]
+			      \textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}[e]) = [\textcolor{e-color}{<_{X\sqcup Y}^{\mu}}(\iota_{E_{X} \sqcup  E_{Y}}(e))]
 		      \]
-		\item $[e) = \varnothing$ and there exists $v$ such that $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(v)) = e'$ and such that there is an undirected path from $[e]$ to $[v]$, we define
+		\item $[e) = \varnothing$ and there exists $v$ such that $<_{X\sqcup Y}^{\mu}(\iota_{V_{X} \sqcup  V_{Y}}(v)) = e'$ and such that there is an undirected path from $[e]$ to $[v]$, we define
 		      \[
-			      \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[e]) = \textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
+			      \textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}[e]) = \textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}[v])
 		      \]
 	\end{enumerate}
-	Otherwise we leave $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}(\iota_{E_{X} + E_{Y}}{\sim}([e]))$ undefined.
+	Otherwise we leave $\textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}(\iota_{E_{X} \sqcup  E_{Y}}{\sim}([e]))$ undefined.
 	% We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
@@
 	Clearly, all the cases above are disjoint.
-	We similarly define $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
+	We similarly define $\textcolor{closed-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}$.
 
-	Let's now define $\consistency_{\sfrac{X+Y}{\sim}}$.
+	Let's now define $\consistency_{\sfrac{X\sqcup Y}{\sim}}$.
 	We can consider the consistency relation from the coproduct as a function
 	\[
-		\consistency_{X+Y} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}
+		\consistency_{X\sqcup Y} : (V_{X} \sqcup  V_{Y}) \sqcup  (E_{X} \sqcup  E_{Y}) \to 2^{(V_{X} \sqcup  V_{Y}) \sqcup  (E_{X} \sqcup  E_{Y})}
 	\]
@@
 	\[
-		\consistency_{X+Y}' : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{(E_{X} + E_{Y})}{\sim})}
+		\consistency_{X\sqcup Y}' : V_{X} \sqcup  V_{Y} \sqcup  E_{X} \sqcup  E_{Y} \to 2^{(\sfrac{V_{X} \sqcup  V_{Y}}{\sim} \sqcup  \sfrac{(E_{X} \sqcup  E_{Y})}{\sim})}
 	\]
-	which is essentially $[ []_{V};\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X+Y}$.
-	We then first define an auxiliary relation $\consistency^{\hashtag}$ similarly to how $<^{\mu}_{\sfrac{X+Y}{\sim}}$ was defined.
-We begin with defining $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ for vertices.
+	which is essentially $[ []_{V};\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}, []_{E};\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}]^{*}$ (a copairing extended to sequences that we will further denote as $[ []_{V}^{\consistency} []_{E}^{\consistency}]$) applied to the return value of $\consistency_{X\sqcup Y}$.
+	We then first define an auxiliary relation $\consistency^{\hashtag}$ similarly to how $<^{\mu}_{\sfrac{X\sqcup Y}{\sim}}$ was defined.
+We begin with defining $\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}$ for vertices.
 
 	\begin{enumerate}
-		\item If there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and $u \sim v$, we let
+		\item If there exists $v$ such that $\consistency_{X\sqcup Y}(\iota_{V_{X} \sqcup  V_{Y}}(v)) \not = \varnothing$ and $u \sim v$, we let
 		      \ifdefined \ONECOLUMN
 			      \[
-				      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))
+				      \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([u]))
 				      =
-				      [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+				      [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X\sqcup Y}(\iota_{V_{X} \sqcup  V_{Y}}(v)))
 			      \]
@@
 			      \begin{align*}
-				       & \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u]))           \\
+				       & \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([u]))           \\
 				       & =                                                                                               \\
-				       & [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)))
+				       & [[]_{V}^{\consistency},[]_{E}^{\consistency}]^{*}(\consistency_{X\sqcup Y}(\iota_{V_{X} \sqcup  V_{Y}}(v)))
 			      \end{align*}
 		      \fi
-		\item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
+		\item $u$ has no pre-image in $V_{Z}$ and $\consistency_{X\sqcup Y}(\iota_{V_{X} \sqcup  V_{Y}}(u)) = \varnothing$ and there exists $v$ such that $\consistency_{X\sqcup Y}(\iota_{V_{X} \sqcup  V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[u]$ to $[v]$.
 		      Then we let
 		      \[
-			      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([u])) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+			      \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([u])) = \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([v]))
 		      \]
@@
 
-	Next we define $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}$ for edges.
+	Next we define $\consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}$ for edges.
 
 	\begin{enumerate}
-		\item If $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
+		\item If $\consistency_{X\sqcup Y}(\iota_{E_{X} \sqcup  E_{Y}}(e)) \not = \varnothing$, then
 		      \begin{align*}
-			      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
-			      [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)))
+			      \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}([e]_{E})) =
+			      [[]_{V}^{\consistency}, []_{E}^{\consistency}]^{*}(\consistency_{X\sqcup Y}(\iota_{E_{X} \sqcup  E_{Y}}(e)))
 		      \end{align*}
-		\item $\consistency_{X+Y}(\iota_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $v$ such that $\consistency_{X+Y}(\iota_{V_{X} + V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[e]$ to $[v]$
+		\item $\consistency_{X\sqcup Y}(\iota_{E_{X} \sqcup  E_{Y}}(e)) = \varnothing$ and there exists $v$ such that $\consistency_{X\sqcup Y}(\iota_{V_{X} \sqcup  V_{Y}}(v)) \not = \varnothing$ and such that there is an undirected path from $[e]$ to $[v]$
 		      then
 		      \[
-			      \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(\iota_{V_{X} + V_{Y}}([v]))
+			      \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X\sqcup Y}{\sim}}^{\hashtag}(\iota_{V_{X} \sqcup  V_{Y}}([v]))
 		      \]
@@
 
-	The well-definedness of this construction follows by the same argument as the well-definedness of $<_{\sfrac{X+Y}{\sim}}^{\mu}$.
+	The well-definedness of this construction follows by the same argument as the well-definedness of $<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}$.
 	Then we define
@@
 		\[
-			\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) \qquad \text{and} \qquad \consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
+			\consistency_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([v])) \qquad \text{and} \qquad \consistency_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}([e]))
 		\]
@@
 		\[
-			\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))
+			\consistency_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([v]))
 		\]
@@
 		\[
-			\consistency_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]))
+			\consistency_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}([e]))
 		\]
@@
 
-	as closures of $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}$ as below.
+	as closures of $\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}$ as below.
 	\[
-		(\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}
+		(\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}
 	\]
-	where $c$ denotes a closure and $\iota_{\sfrac{*}{\sim}}$ is $\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}$ depending on whether $[x_i]$ comes from $V_{X} + V_{Y}$ or $E_{X} + E_{Y}$, is the smallest set such that
+	where $c$ denotes a closure and $\iota_{\sfrac{*}{\sim}}$ is $\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}$ or $\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}$ depending on whether $[x_i]$ comes from $V_{X} \sqcup  V_{Y}$ or $E_{X} \sqcup  E_{Y}$, is the smallest set such that
 	\begin{itemize}
-		\item $[x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$
-		\item if $[y] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x]))$ then
+		\item $[x] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])))^{c}$ if $\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x])) \not = \varnothing$
+		\item if $[y] \in \consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x]))$ then
 		      \[
-			      [x] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([y])))^{c}
+			      [x] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([y])))^{c}
 		      \]
-		      % \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ then 
+		      % \item if $[e] \in \consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}([v]))$ then 
 		      % \[
-		      %     [v] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}
+		      %     [v] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{E_{X} \sqcup  E_{Y}}{\sim}}([e])))^{c}
 		      % \]
@@
 		      \ifdefined \ONECOLUMN
-			      \[[x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}])) \qquad \text{or} \qquad [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))\]
+			      \[[x_i] \in \consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i\sqcup 1}])) \qquad \text{or} \qquad [x_{i\sqcup 1}] \in \consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))\]
 		      \else
 			      \[
-				      [x_i] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i+1}]))
+				      [x_i] \in \consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i\sqcup 1}]))
 			      \] or
 			      \[
-				      [x_{i+1}] \in \consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))
+				      [x_{i\sqcup 1}] \in \consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{i}]))
 			      \]
@@
 			      \[
-				      [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c} \qquad \text{and} \qquad [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
+				      [x_1] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c} \qquad \text{and} \qquad [x_n] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
 			      \]
@@
 			      \[
-				      [x_1] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c}
+				      [x_1] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{n}])))^{c}
 			      \]
@@
 			      \[
-				      [x_n] \in (\consistency^{\hashtag}_{\sfrac{X+Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
+				      [x_n] \in (\consistency^{\hashtag}_{\sfrac{X\sqcup Y}{\sim}}(\iota_{\sfrac{*}{\sim}}([x_{1}])))^{c}
 			      \]
@@
 	This construction is analogous to how a pushout was constructed by~\citet{tiurin2025equivalencehypergraphsdporewriting} and we refer the reader for more detailed proofs there.
-	The only addition is $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$ that is constructed analogously to $\textcolor{e-color}{<_{\sfrac{X+Y}{\sim}}^{\mu}}$.
+	The only addition is $\textcolor{closed-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}$ that is constructed analogously to $\textcolor{e-color}{<_{\sfrac{X\sqcup Y}{\sim}}^{\mu}}$.
 \end{proof}
@@
 \end{tikzpicture}
-such that $\textcolor{closed-color}{<_{X+Y}}$ is only defined for $v_1,v_2,v_3$.
-Therefore, we assign $\textcolor{closed-color}{<_{\sfrac{X+Y}{\sim}}}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v_1']) = [\textcolor{closed-color}{<_{X+Y}}(\iota_{V_{X} + V_{Y}}(v_1))]$ and so on.
+such that $\textcolor{closed-color}{<_{X\sqcup Y}}$ is only defined for $v_1,v_2,v_3$.
+Therefore, we assign $\textcolor{closed-color}{<_{\sfrac{X\sqcup Y}{\sim}}}(\iota_{\sfrac{V_{X} \sqcup  V_{Y}}{\sim}}[v_1']) = [\textcolor{closed-color}{<_{X\sqcup Y}}(\iota_{V_{X} \sqcup  V_{Y}}(v_1))]$ and so on.
 This motivates the requirement for conditions (2), (3), and (4) of~\autoref{th:existence_of_pushouts}, otherwise such inheritance would be ambiguous.
@@
 \[
-	[u_1] \to [u_1, v_1, v_2] + [v_3] \to \mathcal{L}^{\bot} \xleftarrow{} [v_3] + [v_1,v_2] \xleftarrow{} [u_2] ~.
+	[u_1] \to [u_1, v_1, v_2] \sqcup  [v_3] \to \mathcal{L}^{\bot} \xleftarrow{} [v_3] \sqcup  [v_1,v_2] \xleftarrow{} [u_2] ~.
 \]

### 2025-07-11 POPL2026/paper/sections/categorical_semantics.tex
cosmetic changes

@@
 
-The collection of $\catname{SLat}$-categories and $\catname{SLat}$-functors form an $\catname{SLat}$-category $\catname{SLat}--\catname{Cat}$\footnote{More generally, if we were to define $\catname{SLat}$-natural transformations, we could view $\catname{SLat}--\catname{Cat}$ as a symmetric monoidal 2-category~\cite{Kelly2022BASICCO}.}, and for any pair of $\catname{SLat}$-categories $\mathbb{C}$ and $\mathbb{D}$, the product $\catname{SLat}$-category $\mathbb{C} \times \mathbb{D}$ is defined analogously to categories.
+The collection of $\catname{SLat}$-categories and $\catname{SLat}$-functors form an $\catname{SLat}$-category $\catname{SLat}\text{--}\catname{Cat}$\footnote{More generally, if we were to define $\catname{SLat}$-natural transformations, we could view $\catname{SLat}\text{--}\catname{Cat}$ as a symmetric monoidal 2-category~\cite{Kelly2022BASICCO}.}, and for any pair of $\catname{SLat}$-categories $\mathbb{C}$ and $\mathbb{D}$, the product $\catname{SLat}$-category $\mathbb{C} \times \mathbb{D}$ is defined analogously to categories.
 
@@
 Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed symmetric monoidal $\catname{SLat}$-categories, including the distributivity law (\autoref{law:distributivity}).
+The same applies in the presence of equations for $\catname{CS}(\Sigma,\mathcal{E})^{+}$.
 

### 2025-07-11 POPL2026/paper/sections/combinatorial_semantics.tex
cosmetic changes

@@
 
-\autoref{fig:hypergraph-with-interfaces} gives an example of an MDA cospan, which is such because it \enquote{looks like} the string diagram \comonoid.
-Analogously, we can see that~\autoref{fig:hypergraph} (equipped with the interfaces choosing $\{ v_1 \}$ and $\{ v_4, v_5, v_6, v_7 \}$) fails to form an MDA cospan --- it is not monogamous for many reasons: $v_1$ has out-degree 0, $v_1, v_5, v_7$ have in-degree 2, and $v_3$ has in-degree 0; additionally, it fails be be directed acyclic due to the path along $e_4$ from $v_7$ to itself\footnote{Recall that loops are not part of the syntax of string diagrams for symmetric monoidal categories in the absence of a trace --- a relaxation of monogamy to allow for this is explored by~\citet{ghica_rewriting_2023}.}.
+An example and a non-example of an MDA cospan can be found in~\autoref{sec:appendix:hyp}.
+% \autoref{fig:hypergraph-with-interfaces} gives an example of an MDA cospan, which is such because it \enquote{looks like} the string diagram \comonoid.
+% Analogously, we can see that~\autoref{fig:hypergraph} (equipped with the interfaces choosing $\{ v_1 \}$ and $\{ v_4, v_5, v_6, v_7 \}$) fails to form an MDA cospan --- it is not monogamous for many reasons: $v_1$ has out-degree 0, $v_1, v_5, v_7$ have in-degree 2, and $v_3$ has in-degree 0; additionally, it fails be be directed acyclic due to the path along $e_4$ from $v_7$ to itself\footnote{Recall that loops are not part of the syntax of string diagrams for symmetric monoidal categories in the absence of a trace --- a relaxation of monogamy to allow for this is explored by~\citet{ghica_rewriting_2023}.}.
 
-\begin{figure}
-	\begin{subfigure}[c]{0.3\linewidth}
-		\centering
-		\begin{tikzpicture}[
-				vertex/.style={draw, circle, fill, inner sep=1.5pt},
-				hyperedge/.style={draw, rectangle}
-			]
-			\node[vertex, label=left:{$v_4$}] (v1) at (0,0) {};
-			\node[vertex, label=left:{$v_5$}] (v2) at (1,0) {};
-			\node[vertex, label=left:{$v_6$}] (v3) at (2,0) {};
-			\node[vertex, label=left:{$v_7$}] (v4) at (3,0) {};
-			\node[vertex, label=left:{$v_2$}] (v5) at (0.5,-2) {};
-			\node[vertex, label=left:{$v_3$}] (v6) at (2,-2) {};
-			\node[vertex, label=left:{$v_1$}] (v7) at (2,-3) {};
-
-			\node[hyperedge] (e1) at (0.5,-1) {$e_3$};
-			\node[hyperedge] (e2) at (2,-1) {$e_4$};
-			\node[hyperedge] (e3) at (0.5,-3) {$e_1$};
-			\node[hyperedge] (e4) at (2,-2.5) {$e_2$};
-
-			\draw (v1) -- (e1);
-			\draw (v2) -- (e1);
-			\draw (v5) -- (e1);
-			\draw (v5) edge[bend right] (e3);
-			\draw (v5) edge[bend left] (e3);
-			\draw (v2) edge[in=130,out=-90] (e2);
-			\draw (v3) -- (e2);
-			\draw (v4) edge[in=50,out=-90] (e2);
-			\draw (v4) edge[in=-50,out=-90] (e2);
-			\draw (v6) -- (e2);
-		\end{tikzpicture}
-		\caption{Hypergraph with 7 vertices and 4 hyperedges, all directed bottom-to-top; source/target vertices are ordered left-to-right.}%
-		\label{fig:hypergraph}
-	\end{subfigure}
-	\begin{subfigure}[c]{0.3\linewidth}
-		\centering
-		\begin{tikzpicture}[
-				vertex/.style={draw, circle, fill, inner sep=1.5pt},
-				hyperedge/.style={draw, rectangle}
-			]
-			\node[vertex, label=left:{$v_3$}] (v1) at (0,0) {};
-			\node[vertex, label=left:{$v_3$}] (v1') at (0,1.5) {};
-			\node[vertex, label=left:{$v_2$}] (v2) at (1,0) {};
-			\node[vertex, label=left:{$v_2$}] (v2') at (1,1.5) {};
-			\node[vertex, label=left:{$v_1$}] (v3) at (0.5,-2) {};
-			\node[vertex, label=left:{$v_1$}] (v3') at (0.5,-3.5) {};
-
-			\node[hyperedge] (e1) at (0.5,-1) {$e_1$};
-
-			\draw (v1) -- (e1);
-			\draw (v2) -- (e1);
-			\draw (v3) -- (e1);
-
-			\node[draw, rounded corners, fit={(v1'.west) (v2'.west)}, inner sep=12pt] (S) {};
-			\node[draw, rounded corners, fit={(v3'.west)}, inner sep=12pt] (T) {};
-			\node[draw, rounded corners, fit={(v1.west) (v2.west) (v3.west)}, inner sep=12pt] (H) {};
-			\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (S) -- node[left] {$g$} (H);
-			\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (T) -- node[left] {$f$} (H);
-		\end{tikzpicture}
-		\caption{Hypergraph with interfaces}%
-		\label{fig:hypergraph-with-interfaces}
-	\end{subfigure}
-	\caption{Example hypergraphs}%
-	\label{fig:hypergraph-examples}
-\end{figure}
+% \begin{figure}
+% 	\begin{subfigure}[c]{0.3\linewidth}
+% 		\centering
+%         \[
+%         \adjustbox{scale=0.8}{
+%         \tikzfig{figures/hypergraph_example}
+%         }
+%         \]
+% 		\caption{Hypergraph with 7 vertices and 4 hyperedges, all directed bottom-to-top; source/target vertices are ordered left-to-right.}%
+% 		\label{fig:hypergraph}
+% 	\end{subfigure}
+% 	\begin{subfigure}[c]{0.3\linewidth}
+% 		\centering
+%         \[
+%         \adjustbox{scale=0.8}{
+%         \tikzfig{figures/cospan_hypergraph_example}
+%         }
+%         \]
+% 		\caption{Hypergraph with interfaces}%
+% 		\label{fig:hypergraph-with-interfaces}
+% 	\end{subfigure}
+% 	\caption{Example hypergraphs}%
+% 	\label{fig:hypergraph-examples}
+% \end{figure}
 
@@
 		\item $l_{E}\colon E \to \Sigma_{M} \sqcup 1$, $l_{V}\colon V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are label functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
-		\item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V \sqcup E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V \sqcup E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
-		      For $x \in V \sqcup E$, the set of predecessors of $x$ is denoted as $[x) \coloneq \{x^\prime ~|~ \exists y_{1} = x^\prime, \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{<^{\mu}} y_{i + 1}\; \}$.
+		\item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V \sqcup E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V \sqcup E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{\;<^{\mu}\;}$ and $\textcolor{closed-color}{\;<^{\mu}\;}$.
+		      For $x \in V \sqcup E$, the set of predecessors of $x$ is denoted as $[x) \coloneq \{x^\prime ~|~ \exists y_{1} = x^\prime, \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{\;<^{\mu}\;} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{\;<^{\mu}\;} y_{i + 1}\; \}$.
 		      We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ top-level.
@@
 			      \item each $x$ has at most one immediate predecessor;
-			      \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{<^{\mu}} x$ or $e \textcolor{closed-color}{<^{\mu}} x$ are labelled;
-			      \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e^\prime \textcolor{e-color}{<^{\mu}} e$ if and only if $e^\prime \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<^{\mu}}$.
+			      \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{\;<^{\mu}\;} x$ or $e \textcolor{closed-color}{\;<^{\mu}\;} x$ are labelled;
+			      \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e^\prime \textcolor{e-color}{\;<^{\mu}\;} e$ if and only if $e^\prime \textcolor{e-color}{\;<^{\mu}\;} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{\;<^{\mu}\;}$.
 		      \end{enumerate}
-		\item $\consistency$ is a consistency relation which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V \sqcup E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, i.e.\ if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{<^{\mu}}(v)$ and $p \textcolor{e-color}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
-		      We require that $\consistency_{p} \not = (E_{p} \sqcup V_{p}) \times (E_{p} \sqcup V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{<^{\mu}} e\}$.
+		\item $\consistency$ is a consistency relation which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V \sqcup E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, i.e.\ if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{\;<^{\mu}\;}(v)$ and $p \textcolor{e-color}{\;<^{\mu}\;}(e)$ then $v \consistency_{p} e$.
+		      We require that $\consistency_{p} \not = (E_{p} \sqcup V_{p}) \times (E_{p} \sqcup V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{\;<^{\mu}\;} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{\;<^{\mu}\;} e\}$.
 	\end{itemize}
 	$\textcolor{e-color}{E}$ and $\textcolor{closed-color}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$.
-	When writing $x \textcolor{closed-color}{<} y$ (or $x \textcolor{e-color}{<} y$) we may refer to $x$ as the predecessor of $y$ and refer to $y$ as a successor of $x$.
+	When writing $x \textcolor{closed-color}{\;<\;} y$ (or $x \textcolor{e-color}{\;<\;} y$) we may refer to $x$ as the predecessor of $y$ and refer to $y$ as a successor of $x$.
 	% we do not use this
@@
 	\begin{array}{cccc}
-		v_{4} \textcolor{e-color}{<^{\mu}} e_{1}  & v_{7} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{4} \consistency e_{2}   & v_{4} \not \consistency v_{10} \\
-		e_{3} \textcolor{e-color}{<^{\mu}} e_{1}  & v_{8} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency u_{1}   & e_{2} \not \consistency e_{6}  \\
-		e_{2} \textcolor{e-color}{<^{\mu}} e_{1}  & e_{5} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency e_{4}   & e_{4} \not \consistency e_{7}  \\
+		v_{4} \textcolor{e-color}{\;<^{\mu}\;} e_{1}  & v_{7} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & v_{4} \consistency e_{2}   & v_{4} \not \consistency v_{10} \\
+		e_{3} \textcolor{e-color}{\;<^{\mu}\;} e_{1}  & v_{8} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & e_{2} \consistency u_{1}   & e_{2} \not \consistency e_{6}  \\
+		e_{2} \textcolor{e-color}{\;<^{\mu}\;} e_{1}  & e_{5} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & e_{2} \consistency e_{4}   & e_{4} \not \consistency e_{7}  \\
 		\ldots                                    & \ldots                                        & \ldots                     & \ldots                         \\
-		v_{12} \textcolor{e-color}{<^{\mu}} e_{1} & w_{9} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
+		v_{12} \textcolor{e-color}{\;<^{\mu}\;} e_{1} & w_{9} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
 	\end{array}
@@
 		      \begin{equation*}
-			      \text{if } e \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(v),
+			      \text{if } e \textcolor{e-color}{\;<_{\mathcal{F}}^{\mu}\;} v \text{ then } \phi(e) \textcolor{e-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(v),
 			      \qquad
-			      \text{if } e \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(v),
+			      \text{if } e \textcolor{closed-color}{\;<_{\mathcal{F}}^{\mu}\;} v \text{ then } \phi(e) \textcolor{closed-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(v),
 		      \end{equation*}
@@
 		      \begin{equation*}
-			      \text{if } e_1 \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2),
+			      \text{if } e_1 \textcolor{e-color}{\;<_{\mathcal{F}}^{\mu}\;} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(e_2),
 			      \qquad
-			      \text{if } e_1 \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2);
+			      \text{if } e_1 \textcolor{closed-color}{\;<_{\mathcal{F}}^{\mu}\;} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(e_2);
 		      \end{equation*}
@@
 		\[
-			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(y)\bigr\}
+			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(y)\bigr\}
 		\]
@@
 		\begin{multline*}
-			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(y)\bigr\}
+			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(y)\bigr\}
 		\end{multline*}
@@
 		\item for all $v,w$ in $\mathcal{G}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
-$v,w \textcolor{closed-color}{<^{\mu}} e$;
+$v,w \textcolor{closed-color}{\;<^{\mu}\;} e$;
 		\item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t.
-$v,w \textcolor{closed-color}{<^{\mu}} e$;
+$v,w \textcolor{closed-color}{\;<^{\mu}\;} e$;
 		\item there exist $d_1 : n \to \mathcal{L}^\bot$ and $d_2 : k \to \mathcal{L}^\bot$ making the above triangle commute; and

### 2025-07-11 POPL2026/paper/sections/introduction.tex
cosmetic changes

@@
 The structure of symmetric monoidal categories is absorbed into the graphical syntax, with the composition and tensoring represented by juxtaposition, and symmetries as weaving of strings.
-This visual approach makes hides the bureacracy in complex operations and reveals insights sometimes obscured by traditional one-dimensional symbolic notation.
+This visual approach hides the bureacracy in complex operations and reveals insights sometimes obscured by traditional one-dimensional symbolic notation.
 Originating from notations by~\citet{hotzsd} (for automata) and~\citet{penrose1984spinors} (for tensor networks), extensions of string diagrammatic calculi exist for various forms of monoidal categories and even bicategories~\cite{Selinger_2010}, and provide a mathematically rigorous reasoning toolkit~\cite{joyal_geometry_1991}.
@@
 
-\citet{tiurin2025equivalencehypergraphsdporewriting} shows that e-graphs over signature $\Sigma$ can be represented as morphisms of a free cartesian symmetric monoidal semilattice-enriched category over the same signature.
+\citet{tiurin2025equivalencehypergraphsdporewriting} shows that e-graphs over signature $\Sigma$ can be represented as morphisms of a free Cartesian symmetric monoidal semilattice-enriched category over the same signature.
 We argue that restricting the categorical domain to free \textbf{closed} symmetric monoidal semilattice-enriched categories gives rise to structures that naturally support binding and equivalence classes of morphisms.

### 2025-07-11 POPL2026/paper/sections/lambda-calculus.tex
cosmetic changes

@@
 In other words, with respect to~\autoref{fig:egraph-strings}, this allows us to \emph{natively} represent abstraction by the rounded boxes of ($\Lambda(f)$), and application by the $\textsf{ev}$ operation with $\beta$- and $\eta$-conversion arising via equality of morphisms in the category (that is, without having to provide them as data in the form of equations in $\mathcal{E}$).
-For explicit substitution, we note that the linearity is not desirable; this is why the copy-delete comonoid $(\comonoid, \counit)$ is present: by Fox's theorem~\cite{foxCoalgebrasCartesianCategories1976}, a cartesian category is precisely a monoidal one admitting such a \emph{natural} copy-delete structure.
+For explicit substitution, we note that the linearity is not desirable; this is why the copy-delete comonoid $(\comonoid, \counit)$ is present: by Fox's theorem~\cite{foxCoalgebrasCartesianCategories1976}, a Cartesian category is precisely a monoidal one admitting such a \emph{natural} copy-delete structure.
 Thus, we must endow $\mathcal{E}$ with equations stating that $(\comonoid, \counit)$ forms a comonoid, and moreover that it is natural: $\comonoid$ behaves uniformly as a copy map, and $\counit$ behaves as deletion.
@@
 Now, in our semantic model, this rewrite can be non-destructively applied to model one step of e-graph saturation:
-\[
-	\tikzfig{figures/plus-copy-rewrite-1}
+\[\adjustbox{scale=0.8}{
+	\tikzfig{figures/plus-copy-rewrite-1-2}
+}
 	\quad
@@
 	\quad
-	\tikzfig{figures/plus-copy-rewrite-2}
+	\adjustbox{scale=0.8}{
+	\tikzfig{figures/plus-copy-rewrite-2-2}
+	}
 	\quad
@@
 	\quad
-	\tikzfig{figures/plus-copy-rewrite-3}
+	\adjustbox{scale=0.8}{
+	\tikzfig{figures/plus-copy-rewrite-3-2}
+	}
 	\;

### 2025-07-11 POPL2026/paper/sections/soundness.tex
cosmetic changes

@@
 \end{definition}
+
+In practice such rewrite rules are generated by the equations of the corresponding theory.
 
@@
 Note that the rewrite rule consist of closed $\Sigma$-terms and not closed $\Sigma^{+}$-terms, i.e. neither $l$ nor $r$ contain any $+$.
-This is intentional as there rewrite rules are supposed to be domain-specific (like a rule $\langle 2 \otimes id;times, id \otimes 1; <\;\;< \rangle$) and not related to the free enrichment.
+This is intentional as there rewrite rules are supposed to be domain-specific (like a rule $\langle 2 \otimes id;*, id \otimes 1; <\!\!< \rangle$) and not related to the free enrichment.
 
@@
 	\label{prop:fscd}
-	For two closed $\Sigma$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$, where the latter is defined in a sub $\catname{SLat}$-category of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
+	For two closed $\Sigma$-terms $f$ and $g$ and a closed rewrite rule $\langle l, r \rangle$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$, where the latter is defined in a sub-category of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
 \end{proposition}
@@
 	\label{prop:quotient-structural}
-	For two closed $\Sigma^{+}$-terms $f$ and $g$ and an $\mathcal{E}$ equation $l = r$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
+	For two closed $\Sigma^{+}$-terms $f$ and $g$ and a closed rewrite rule $\langle l, r \rangle$,  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$.
 \end{proposition}
@@
 
+The proposition above allows us to reason about theories of closed $\Sigma^{+}$-terms $\catname{CS}(\Sigma,\mathcal{E})^{+}$.
+Given a set of equations $\mathcal{E}$ of closed $\Sigma$-terms, we construct $\llbracket \mathcal{E} \rrbracket = \{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle, \langle \llbracket r \rrbracket, \llbracket l \rrbracket \rangle \;|\; l = r \;\in\; \mathcal{E}\}$ and then we can perform DPO rewriting in $\WellTypedMdaEcospans / \mathcal{S}$ which formalises, \textit{e.g.}, the example in~\autoref{fig:e-string-substitution}.
 \begin{figure}

### 2025-07-11 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
some polish

@@
 \section{Conclusion}
-In this work we introduced e-graphs with bindings formalised as morphisms of a closed symmetric monoidal $\catname{SLat}$-category.
+In this work we introduced e-graphs with bindings derived as morphisms of a closed symmetric monoidal $\catname{SLat}$-category.
 In doing so, we are able to leverage its internal language to provide variable binding in $\lambda$-calculus-like languages by metatheoretic $\lambda$-abstraction.
 We then provided a combinatorial representation for such morphisms in terms of closed e-hypergraphs with a suitable notion of sound and complete DPO rewriting, making it possible for a closed e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
-We argued for the utility of our techniques in the representation of programming languages (which inherently require variable binding) as e-graphs, contrasting with the more general-purpose slotted e-graphs, and advocating for the benefits of string-diagrammatic syntax.
-In future, we hope to implement an efficient version of e-graphs for hypergraphs as a data structure using DPO rewriting as its foundation to implement e-graph operations, with the theory we have developed here acting as a roadmap.
+We argued for the utility of our techniques in the representation of programming languages (which inherently require variable binding) as e-graphs, contrasting with the more general-purpose slotted e-graphs, and advocating for the benefits of string-diagrammatic syntax, in particular in the context of programming languages.
+In the future, we hope to implement an efficient version of e-graphs for hypergraphs as a data structure using DPO rewriting as its foundation to implement e-graph operations, with the theory we have developed here acting as a roadmap.
 

### 2025-07-11 POPL2026/paper/sections/categorical_semantics.tex
some polish

@@
 To aid reasoning, we introduce a new language of string diagrams for closed symmetric monoidal $\catname{SLat}$-categories, using a hierarchical \enquote{dashed box} structure to capture the join operation on morphisms (as was done by~\citet{tiurin2025equivalencehypergraphsdporewriting}) as well as another hierarchical \enquote{rounded box} for $\lambda$-abstraction.
-\autoref{fig:egraph-strings} displays the generators of this language, which is the usual string diagrammatic syntax~\cite{Selinger_2010} plus the syntax for $\lambda$-abstraction as established by~\citet{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice enrichment as introduced by~\citet{tiurin2025equivalencehypergraphsdporewriting}; the first component denotes an empty diagram.
+\autoref{fig:egraph-strings} displays the generators of this language, which is the usual string-diagrammatic syntax~\cite{Selinger_2010} plus the syntax for $\lambda$-abstraction as established by~\citet{ghica2024stringdiagramslambdacalculifunctional} and syntax for semilattice enrichment as introduced by~\citet{tiurin2025equivalencehypergraphsdporewriting}; the first component denotes an empty diagram.
 

### 2025-07-11 POPL2026/paper/sections/introduction.tex
some polish

@@
 This visual approach hides the bureacracy in complex operations and reveals insights sometimes obscured by traditional one-dimensional symbolic notation.
-Originating from notations by~\citet{hotzsd} (for automata) and~\citet{penrose1984spinors} (for tensor networks), extensions of string diagrammatic calculi exist for various forms of monoidal categories and even bicategories~\cite{Selinger_2010}, and provide a mathematically rigorous reasoning toolkit~\cite{joyal_geometry_1991}.
+Originating from notations by~\citet{hotzsd} (for automata) and~\citet{penrose1984spinors} (for tensor networks), extensions of string-diagrammatic calculi exist for various forms of monoidal categories and even bicategories~\cite{Selinger_2010}, and provide a mathematically rigorous reasoning toolkit~\cite{joyal_geometry_1991}.
 
@@
 Using de Bruijn indices instead when performing $\beta$-reduction would only push the problem elsewhere, as it greatly increases the complexity of representing the reduction as a rewrite rule, and special shifting operators need to be introduced on the right-hand side of a rewrite rule, which is a meta-syntactic operation~\cite{koehler2022sketchguided}.
+
 Recent work on \emph{slotted} e-graphs~\cite{slotted-egraphs} uses nominal techniques to address the sharing issues when representing terms that are equivalent up to renaming of variables (even free ones) by incorporating variables as first-class data of an e-graph and making binding a built-in operation.
@@
 		\]
-		\subcaption{String diagrammatic substitution example.}\label{fig:e-string-substitution}
+		\subcaption{String-diagrammatic substitution example.}\label{fig:e-string-substitution}
 	\end{subfigure}
+	\caption{Conventional e-graphs vs.\ categorical e-graphs with binding.}
 \end{figure*}
@@
 % 	\captionsetup{skip=0pt}
-%     \caption{String diagrammatic substitution example.}
+%     \caption{String-diagrammatic substitution example.}
 %     \label{fig:e-string-substitution}
@@
 
+Our contribution is the construction of a form of e-graphs that support binding, developed through a principled categorical approach.
+Specifically, we extend a prior categorical semantics for e-graphs without binding by incorporating a monoidal closed structure.
 \citet{tiurin2025equivalencehypergraphsdporewriting} shows that e-graphs over signature $\Sigma$ can be represented as morphisms of a free Cartesian symmetric monoidal semilattice-enriched category over the same signature.

### 2025-07-11 POPL2026/notes/SlottedNotes.tex
fixed figures, updated contributions
some fixes

@@
 
-Lambda calculus with explicit substitution (also $\lambda_{\sigma}$-calculus) made the substitution operator a first-class element of the calculus (rather than a meta-level operation) to bridge the gap between lambda calculus and its concrete implementations~\cite{Abadi_Cardelli_Curien_Lévy_1991}.
+Lambda calculus with explicit substitution (also $\lambda_{\sigma}$-calculus) made the substitution operator a first-class element of the calculus (rather than a meta-level operation) to bridge the gap between lambda calculus and its concrete implementations~\cite{AbadiExplicitSubst}.
 However, the existence of unique normal forms turned out to be a delicate matter and to achieve uniqueness, lambda terms must be quotiented by \textit{graphical} equivalence in addition to $\alpha$-equivalence~\cite{Standardisation}.
@@
 \begin{tabular}{lcc}
+\texttt{beta} & $(\lambda x . t) u = t[x / u]$&\\
 \texttt{subst\_same\_var} & $x[x / u] = u$ &\\
@@
 \texttt{subst\_lambda} & $(\lambda x . u)[y / v] = \lambda x . (u[y / v])$ & if $x \not \in \mathcal{F}(v)$ \\
-\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$\\
-\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])[x / v[x / v']]$ & if ???\\
+\texttt{subst\_interchange\_1} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v[x' / v']]$ & if $x' \in \mathcal{F}(v)$ and $x \in \mathcal{F}(v')$\\
+\texttt{subst\_interchange\_2} & $(u[x / v])[x' / v'] = (u[x / v[x' / v']])$ & if $x' \not \in \mathcal{F}(u)$\\
 \texttt{subst\_comm} & $(u[x / v])[x' / v'] = (u[x' / v'])[x / v]$ & if $x \not \in \mathcal{F}(v')$ and $x' \not \in \mathcal{F}(v)$\\

### 2025-07-11 POPL2026/paper/sections/introduction.tex
some fixes

--- 2025-07-11 POPL2026/paper/sections/introduction.tex
@@
 We are interested in using e-graphs in the context of functional programming languages, which builds upon an internal representation of the $\lambda$-calculus.
-Pioneering work in this direction has been done by~\citet{koehler2022sketchguided}, where the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when applying equality saturation to an encoding of the $\lambda$-calculus in e-graphs.
+% Pioneering work in this direction has been done by~\citet{koehler2022sketchguided}, where the authors outlined the two interconnected challenges of binding and capture-avoiding substitution that arise when applying equality saturation to an encoding of the $\lambda$-calculus in e-graphs.
+\citet{koehler2022sketchguided} pioneered work in this area by identifying two key, interconnected challenges—handling binding and ensuring capture-avoiding substitution—when applying equality saturation to an encoding of the $\lambda$-calculus in e-graphs.
 
@@
 
-A simple extension of $\lambda$-calculus is given by adding \emph{explicit substitution}, which models \textbf{let}-bindings in programming languages.
-Here, a term equipped with explicit substitutions $t[y/v][x/u]$ (modelling $\textbf{let} \; x = u \; \textbf{in} \; \textbf{let} \; y = v \; \textbf{in} \; t$) can be identified with the corresponding term where the order of substitution is reversed $t[x/u][y/v]$, supposing the substitutions are non-interacting.
-This imposes a \emph{graphical equivalence}~\cite{accattoli2014nonstandard} of terms.
+% A simple extension of $\lambda$-calculus is given by adding \emph{explicit substitution}, which models \textbf{let}-bindings in programming languages.
+A simple but theoretically~\cite{AbadiExplicitSubst} and practically~\cite{anf} important extension of the $\lambda$-calculus consists of adding let-bindings or explicit substitutions to the syntax.
+% Here, a term equipped with explicit substitutions $t[y/v][x/u]$ (modelling $\textbf{let} \; x = u \; \textbf{in} \; \textbf{let} \; y = v \; \textbf{in} \; t$) can be identified with the corresponding term where the order of substitution is reversed $t[x/u][y/v]$, supposing the substitutions are non-interacting.
+Terms are usually identified up to permutations of non-interacting substitutions, which is sometimes called \enquote{graph equivalence}~\cite{accattoli2014nonstandard} since such identification becomes obvious in certain diagrammatic representations.
+% This imposes a \emph{graphical equivalence}~\cite{accattoli2014nonstandard} of terms.
 Modelling this in the conventional language of e-graphs requires the introduction of additional nodes to represent the explicit substitutions, and additional rewrite rules which capture graphical equivalence.
-This pollutes the state space of equality saturation for these languages: the interesting rewriting is that of the user-specified algebraic theory one wishes to model, not between semantically equivalent but syntactically distinct forms in the language.
-
+This pollutes the state space of equality saturation for these languages: the interesting rewriting is that of the user-specified algebraic theory one wishes to model, 
+% not between semantically equivalent but syntactically distinct forms in the language.
+not the bureaucracy of managing the congruences of semantically equivalent yet syntactically distinct forms in the language.
 All these issues can be resolved by the use of \emph{string diagrams} in place of textual syntax.
@@
 String diagrams are a direct bridge from the language of category theory, which in turn can be used to give semantic models to a variety of programming languages and logics, to the concrete language of graphs, which are a fundamental data structure in algorithm design.
-These categorical techniques provide insight in how to solve the problems of binding and capture-avoiding substitution in e-graphs, as we develop in this paper.
+These categorical techniques provide methodological guidance in how to solve the problems of binding and capture-avoiding substitution in e-graphs, as we develop in this paper.
 
 The general background literature for string diagrams and their relationship to formal languages is summarised by~\citet{piedeleu2023introductionstringdiagramscomputer}.
-The particular context of string diagrams with bindings is given in~\citet{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\citet{tiurin2025equivalencehypergraphsdporewriting}.
+The particular context of string diagrams for $\lambda$-calculi is given in~\citet{ghica2024stringdiagramslambdacalculifunctional}, while string diagrams for e-graphs without bindings are defined in~\citet{tiurin2025equivalencehypergraphsdporewriting}.
 
@@
 The structure of symmetric monoidal categories is absorbed into the graphical syntax, with the composition and tensoring represented by juxtaposition, and symmetries as weaving of strings.
-This visual approach hides the bureacracy in complex operations and reveals insights sometimes obscured by traditional one-dimensional symbolic notation.
+This visual approach hides the bureacracy of complex operations and reveals insights sometimes obscured by traditional one-dimensional (textual) syntax.
 Originating from notations by~\citet{hotzsd} (for automata) and~\citet{penrose1984spinors} (for tensor networks), extensions of string diagrammatic calculi exist for various forms of monoidal categories and even bicategories~\cite{Selinger_2010}, and provide a mathematically rigorous reasoning toolkit~\cite{joyal_geometry_1991}.
@@
 Specifically, we require a concrete encoding of string diagrams, analogous to the de Bruijn index representation of $\lambda$-terms.
-For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have nice syntactic characteristics.
-For instance, hypergraph isomorphism captures when string diagrams \enquote{look the same} --- topologically, that one may be deformed continuously into the other without obstruction\footnote{This is the informal notion of the \enquote{isotopy of string diagrams}.}; secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting over these hypergraphs using the technique of double-pushout rewriting (DPO)~\cite{bonchi_string_2022-1}.
-
+For this, we use \emph{hypergraphs}~\cite{bonchi_string_2022-1}, which have suitable theoretical properties.
+For instance, hypergraph isomorphism rigorously expresses what it means for string diagrams to \enquote{look the same} --- topologically, that one may be deformed continuously into the other without obstruction\footnote{This is the informal notion of the \enquote{isomorphism of string diagrams}.}; secondly, string-diagrammatic equational reasoning can be formalised as (categorical) graph rewriting over these hypergraphs using the technique of double-pushout rewriting (DPO)~\cite{bonchi_string_2022-1}.
 Finally, for the diagrammatic representation of two concepts essential to both programming languages (variable binding) and to e-graphs (equivalence classes) we rely on the concept of \emph{functorial boxes} as introduced by Melli\`es~\cite{10.1007/11874683_1}.
@@
 Revisiting the earlier example $\lambda f . f ((\lambda x . f x) 2)$, in~\autoref{fig:de-brujin-string}, observe that the two occurrences of $f$ are now shared.
-$\lambda$-abstraction is represented by a rounded box with the bound variable a dangling wire exiting the frame along its vertical boundary, the application node a half-circle
+The scope of a $\lambda$-abstraction is represented by a rounded box with the bound variable represented as a wire attached to the frame, the application node a half-circle
 \adjustbox{scale=0.15}{\begin{tikzpicture}
@@
 
-Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the \emph{e-hypergraphs} of~\citet{tiurin2025equivalencehypergraphsdporewriting}.
+Each column represents a term (or term rewrite rule), its conventional e-graph, and the equivalent string diagram representation using the \emph{e-hypergraphs} used by~\citet{tiurin2025equivalencehypergraphsdporewriting}.
 For the initial term, $(a*2)/2$, which is already represented efficiently in the e-graph by sharing the node $2$, the string diagram version shares this node explicitly, using a copy node.
@@
 The first rewrite rule, replacing multiplication by \emph{shift-left} creates the first non-singleton e-class, which includes the multiplication ($*$) and shift-left ($<\!\!<$) nodes, now sharing the node $a$.
-This second e-graph illustrates the other representational difference, besides sharing, between conventional e-graphs and e-hypergraphs.
+This second e-graph illustrates the other representational difference\footnote{Despite the \enquote{difference}, these representations are isomorphic is a sense that there is a non-ambiguous translation from one representation to another and vice versa.}, besides sharing, between conventional e-graphs and e-hypergraphs.
 In the former, edges connect directly to e-nodes inside the e-class, whereas in the latter edges connect to the equivalence class itself.
@@
 We can achieve a similar degree of sharing (in the sense that we have the same number of nodes for $+$) using string diagrams by performing a closure conversion of $+$ (\autoref{fig:closed}).
+% One can consider the example from~\autoref{fig:e-graph-substitution} as an instance of a slotted e-graph as well (with headers omitted), the explicit substitution works similarly between the two versions of e-graphs with the slotted version providing sharing up-to renaming.
+A slotted e-graph version of~\autoref{fig:e-graph-substitution} would be able to introduce sharing for the variable names $x$ and $y$ but must still represent abstraction, application, and substitution explicitly, and so would look much the same.
 Fundamentally, there is a difference in approach here: slotted e-graphs solve the problem of variable binding generically using nominal techniques, in a way that can be applied to arbitrary formal languages with binders (e.g.

### 2025-09-19 term-calculus/cat_logic_notes.tex
notes on cat logic

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%

\documentclass[acmsmall,screen, nonacm, anonymous]{acmart}

\usepackage{stmaryrd}
\newcommand{\parr}{\mathbin{\bindnasrepma}}
% \usepackage{amssymb}
\usepackage{mathpartir}

\newcommand*{\ONECOLUMN}{}

\input{../preamble}
\input{../macros}
\input{../sample.tikzstyles}
\input{../hypergraph.tikzstyles}
\input{../hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\usepackage{minted}
\usepackage{csquotes}

% type derivations
\usepackage{mathpartir}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

\settopmatter{printacmref=false}
\authorsaddresses{}

\begin{document}

\title{Notes on categorical logic and term calculi}

\author{Anonymous}

\maketitle

\section{Linear logic}

\textit{Linear logic} frequently pops up in the context of term calculus and categorical semantics and in fact is an internal logic of closed symmetric monoidal categories (with varied additional structures).
This section will briefly introduce what linear logic is and sketch its categorical semantics.
The corresponding term calculus is system L or $\overline{\lambda}\mu\widetilde{\mu}$-calculus.

Semantics is given by a symmetric monoidal category with enough tensors to support logical connectives.

\begin{center}
\begin{tabular}{c|cc}
{}& conjunction & disjunction\\
\hline
multiplicative & $\otimes$ & $\parr$\\
additive & $\&$ & $\oplus$\\
\end{tabular}
\end{center}

The difference between additive and multiplicative connectives is in how they treat contexts, \textit{e.g.}, multiplicative connectives split context into disjoint bits.
Additive conjunction is modelled by a Cartesian product and the multiplicative one is given by a tensor.
The modality $\!$ is interpreted by a comonad. 
Then we need to require that the comonad interacts with the tensor.

\section{Differential lambda calculus by ~\citet{DiffLambdaCalc}}

This is the calculus such that the model of the lambda terms are differentiable functions between vectors spaces.
The calculus is closely related to the notion of linearity and linear substitution.
In ordinary $\lambda$-calculus, application $(s) t$ substitutes $t$ into $s$.
In differential $\lambda$-calculus, we have two kinds of application:
\begin{itemize}

\item Nonlinear application (ordinary application) --- can duplicate arguments.

\item Linear application --- written as $D_{k} t \cdot u$ (term $t$ differentiated with respect to its k-th argument linearly applied to $u$)

\end{itemize}
The derivative $D t \cdot u$ says: \enquote{The linear effect on the result of $s$ when we change its input in the direction $u$.}
This is analogous to a directional derivative in calculus --- except `change' is in a syntactic/linear resource sense.

\begin{example}
Consider a term $(x) x$.
Its derivative (applied to $u$) $D (x) x \cdot u$ is $(u) x + x (u)$ (if we ignore linearity issues) where $+$ is a nondeterministic choice.
That is, the change is either propagated to the first $x$ or to the second.
\end{example}

The paper then proceeds to formally define $D$, $\partial x$ and the interaction of these with the $\beta$-reduction.
The calculus along with the reduction rules is shown to be strongly normalizing and confluent --- that is, equality of terms is straightforwardly decidable.

\section{Categorical Type Theory}

This is the study of the relation between \textit{algebraic type theory} and categories with finite products.

\begin{definition}

An algebraic signature $\Sigma$ is given by:
\begin{itemize}

\item a collection of types;
\item a collection of functional symbols with given arity (possibly 0) but the co-arity is 1
\item a typing for each functional symbol $f$ which is a list of types $[\alpha_1, \ldots, \alpha_n, \alpha]$ such that $f : \alpha_1, \ldots, \alpha_{n} \to \alpha$
\end{itemize}

The raw terms are then given by
\[
 M \Coloneqq x \;|\; k \;|\; f(M_1, \ldots, M_{n})
\]
where $x$ is a variable from a countably infinite set of variables and $k$ is a functional symbol of arity 0.
\end{definition}

Given an algebraic theory $\Sigma$, an \textit{equation-in-context} is $\Gamma \vdash M = M' : \alpha$.

\begin{definition}

An algebraic theory is a pair $(\Sigma, \mathcal{A})$ where $\Sigma$ is a signature and $\mathcal{A}$ is a set of equations in context called the set of axioms.
\end{definition}

The theorems of a theory can then be derived from the axioms using reflexivity, symmetry, transitivity, substitution and context operations.
Categorical semantics then assigns objects to types and morphisms to type judgements.
In particular, substitution becomes morphism composition as (in the case of one variable)
\[
\inferrule*[right=Subst]
  { x : \alpha \vdash M : \beta  \\ y : \beta \vdash N : \gamma }
  { x : \alpha \vdash N[M / y] : \gamma }
\]

\[
\llbracket x : \alpha \vdash N[M / y] : \beta \rrbracket =  \llbracket y : \beta \vdash N : \gamma \rrbracket \circ \llbracket x : \alpha \vdash M : \beta \rrbracket
\]

Formally, letting 
\begin{align*}
\llbracket \alpha \rrbracket \in \text{obj}(C)\\
\llbracket k : \alpha \rrbracket = 1 \to \llbracket \alpha \rrbracket\\
\llbracket f : \alpha_1, \ldots, \alpha_n \to \alpha  \rrbracket = \llbracket \alpha_1 \rrbracket \otimes \ldots \llbracket \alpha_n \rrbracket \to \llbracket \alpha \rrbracket
\end{align*}

we get the lemma

\begin{lemma}
\[
\llbracket \Gamma \vdash N[\vec{M} / \vec{x}] : \beta \rrbracket = \llbracket \Gamma \vdash N : \beta \rrbracket \circ \langle \llbracket \Gamma \vdash M_{1} : \alpha_{1} \rrbracket, \ldots, \llbracket \Gamma \vdash M_{n} : \alpha_{n} \rrbracket \rangle
\]

\end{lemma}

The assignment of $\llbracket - \rrbracket$ above is called a structure.
A structure is then \textit{models} the theory if all equations-in-context are equal morphisms.
The interpretation is sound if the structure also satisfies all the theorems.

\begin{definition}
  Given two models of a theory $Th$ in a category $\catname{C}$ we can define a homomorphism between two models such that diagrams for the interpretations commute.
  This gives rise to a category $Mod(Th,\catname{C})$.  
\end{definition}

\begin{definition}
  $\mathcal{FP}(C,D)$ for categories with products $C,D$ has as objects product-preserving functors and as morphisms it has natural transformations.
\end{definition}

The following claim is then can be made that for a given algebraic theory $Th$ we can find a category $Cl(Th)$ such that
\[
\mathcal{FP}(Cl(Th),D) \cong Mod(Th,D)
\]
That is, a model can be regarded as a functor. 
The equivalence above means there are functors between the two categories such that the composition of the functors are naturally isomorphic to identity functor.

Given a product preserving functor $F : C \to D$ we can lift it to $F_{*} : Mod(Th, C) \to Mod(Th,D)$.
Let $M$ be an object of $Mod(Th,C)$.
$F_{*}(M)$ is defined by putting $\llbracket \alpha \rrbracket_{F_{*}, M} = F\llbracket \alpha \rrbracket_{M}$ where $\alpha$ is the type of $Th$ and letting
$\llbracket f \rrbracket_{F_{*},M}$ to be given by composition
\[
  \llbracket \alpha_{1} \rrbracket_{F_{*}, M} \times \ldots \times \llbracket \alpha_{n} \rrbracket_{F_{*}, M} \cong F(\llbracket \alpha_{1} \rrbracket \times \ldots \times \llbracket \alpha_{n} \rrbracket) \xrightarrow{F(\llbracket f \rrbracket_{M})} F(\llbracket \beta \rrbracket_{M})
\]
For a morphism $h : M \to N$ in $Mod(Th,C)$, $F_{*}(h) = Fh$.

If $\phi : F \to G$ is a natural transformation between product preserving functors, we can define a homomorphism
\[
\phi_{*}M : FM \to GM\\
\phi_{\llbracket \alpha \rrbracket}M : F\llbracket \alpha \rrbracket \to G\llbracket \alpha \rrbracket_{M}
\]
of models.
Then we will define a family of modelling functors
\[
\mathcal{A}_{p_{M}} : \mathcal{FP}(C,D) \to Mod(Th,D)
\]
by setting $\mathcal{A}_{p_{M}}(F) = F_{*}(M)$ and $\mathcal{A}_{p_{M}}(\phi) = \phi_{*}M$.
A category (with finite products) $Cl(Th)$ is called a \emph{classifying} category of $Th$ if there is a model $G$ of $Th$ in $Cl(Th)$ such that for any other category $D$ the modelling functor
\[
\mathcal{A}_{p_{G}} : \mathcal{FP}(Cl(Th),D) \to Mod(Th,D)
\]
is an equivalence.
Any such $Cl(Th)$ is unique up to isomorphism.

\section{Categorical logic (as per Michael Shulman)}

The idea is to construct a free structure (with some nice syntactic properties, not every free structure will do) and then prove properties about non-free structures using the free structure and the homomorphism from this free structure.
For example, one could show that a property that holds for all groups by showing that it holds for a free group.

\begin{example}

Let a conjugation of $h$ by $g$, $h^{g}$ by $ghg^{-1}$.
For any $h,g,k$ we have $(hk)^{g} = h^{g}k^{g}$.

This statement can be proved by constructing a free group over a set of 3 elements $F_{3}$ and then by tracing homomorphisms to any other group.
\end{example}

\begin{example}

Another example is a type theory of posets (of pre-orders actually) that is a syntactical representation of morphisms in $\mathbf{Poset}$.
We can formulate the following question, is it true that for any objects $A,B,C,D,E$ satisfying

\[
A \leq B \qquad A \leq C \qquad D \leq A \qquad B \leq E \qquad D \leq C~,
\]
it is the case that $D \leq E$.

This can be answered by consulting the adjunction $\mathcal{F} \vdash U : \mathbf{Poset} \to \mathbf{RelGr}$ where the latter is a category of sets with binary relations.
Given a set $X$ in $\mathbf{RelGr}$, $\mathcal{F}X$ is a free poset on this set (i.e. its reflexive and transitive closure).
We can then say that $\mathcal{F}X$ has the same objects as $X$ and the morphisms are the judgements derivable from $X$ in a type theory of posets.
That is, given any other poset $A$ and a morphism $P : X \to UA$ there is a morphism $\mathcal{F}X \to A$ whose action on objects is given by $P$ and we can show by induction on the structure of the derivation that such morphism preserves the order.
\end{example}

\begin{example}

Another example stems from the adjunction $\mathcal{F} \vdash U : \mathbf{Cat} \to \mathbf{Graph}$ where the latter is the category of directed graphs.
To relate morphisms and derivations we will need to label the derivations.
For example, given two edges $f$ and $g$ in $\catname{Graph}(A,B)$, we get two primitive derivations

\[
\inferrule*[right=$f$]{\hspace{1em}}{A \vdash B} \qquad \inferrule*[right=$g$]{\hspace{1em}}{A \vdash B}
\]

Then we will need rules for composition, \textit{e.g.},

\[
\inferrule*[right=$\circ$]{
  \inferrule*{}{
    \inferrule*[right=$h$]{\;}{B \vdash C}
  }{
    \inferrule*[right=$g$]{\;}{A \vdash B}
  }
}{A \vdash C}
\]

Then, to make these derivations into proper categorical morphisms we need to quotient them by $\equiv$ (that is by associativity of $\circ$ and unitality of identity morphisms).
The claim is then that these data define a free category $\mathcal{F}G$ on a directed graph $G$.
The approach above (with quotienting) is being referred to as \textbf{cut-ful type theory}.

Another approach is to build the composition into axioms, \textit{i.e.}, for each $f \in \catname{Graph}(A,B)$ we have
\[
\inferrule*[right=$f$]{X \vdash A}{X \vdash B}
\]
Intuitively, the above rule is a post-composition with $f$.
This approach is called \textbf{cut-free type theory}.
We then can have a theorem that
\begin{theorem}
  Given a derivation $A \vdash B$ and $B \vdash C$ we can construct a derivation $A \vdash C$.
\end{theorem}
The theorem essentially states that cut-rule is admissible in this type theory.
The inference rules can be labelled with terms as follows (for $f : A \to B$)
\[
\inferrule*[right=$id$]{\;}{id_{A} : A \vdash A} \qquad \inferrule*[right=$f$]{\phi : X \vdash A}{f \circ \phi : A \vdash B}
\]
and we can identify terms with type derivations
\[
\inferrule*[right=$h$]{
\inferrule*[right=$g$]{
  \inferrule*[right=$f$]{
    \inferrule*[right=$id$]{\;}{id_{A} : A \vdash A}
  }{f \circ (id_{A}) : A \vdash B}
}{g \circ (f \circ id_{A}) : A \vdash C}
}{h \circ (g \circ (f \circ id_{A})) : A \vdash D}
\]

The syntax can be improved by introducing \textit{formal} variables for the left-hand sides of $\vdash$, so that the above term can be written as
\[
x : A \vdash h(g(f(x))) : D
\]
The identity and generator rules can now be written as
\[
\inferrule*[right=$id$]{\;}{x : A \vdash x : A} \qquad \inferrule*[right=$f$]{x : X \vdash M : A \hspace{2em} f \in \mathcal{G}(A,B)}{x : X \vdash f(M) : B}
\]
We can prove the cut admissibility using this notation as follows.
Given $x : A \vdash M : B$ and $y : B \vdash N : C$ we need to construct $x : A \vdash K : C$.
\begin{itemize}
  \item if $N = y$, then $C = B$ and $x : A \vdash M : B$ is the required term.
  \item if $N = f(N')$ for $N' : D$ and $f : D \to C$ we get a term $A \to D$ using the inductive hypothesis and after applying $f$ to it we get the required $A \to C$.
\end{itemize}

We can then say that a free category $\mathcal{F}(G)$ for a directed graph $G$ has the same objects and the morphisms are given by derivations in the type theory with composition given by substitution.
Substitution being associative and having unitality implies this is indeed a category and we can then define a functor $\mathcal{F}(G) \to \mathcal{A}$ for any category $\mathcal{A}$ by induction on the derivation using the morphism $G \to \mathcal{A}$.
This way no quotienting is required.
\end{example}

When we move to more complex theories, \textit{e.g.} the ones that have more complicated objects, such as meets in semilattices, we need to utilize categories with products.
Then, for the derivations we have two options (just like we had for a composition) regarding the projections $A \times B \to A$ and $A \times B \to B$.
\[
\inferrule*{A \vdash C}{A \times B \vdash C} \qquad \inferrule*{C \vdash A \times B}{C \vdash A}
\]

where the first is essentially precomposing a projection $\pi_1$ to $f: A \to C$ : $A \times B \xrightarrow{\pi_1} A \xrightarrow{f} C$.
The other one is postcomposition for $g : C \to A \times B$: $C \xrightarrow{g} A \times B \xrightarrow{\pi_1} A$.
Utilizing the first rule gives rise to \textit{sequent calculus} style type theories while the second one gives rise to \textit{natural deduction}.

In the case of a meet semilattice these rules get instantiated like this

\[
\inferrule*[right=$\land L_{1}$]{A \vdash C \quad \vdash B\;\text{type}}{A \land B \vdash C} \qquad \inferrule*[right=$\land L_{2}$]{B \vdash C \quad \vdash A\;\text{type}}{A \land B \vdash C} \qquad \inferrule*[right=$\land R$]{A \vdash B \quad A \vdash C}{A \vdash B \land C}
\]

and

\[
\inferrule*[right=$E_{1}$]{X \vdash A \land B}{X \vdash A} \qquad \inferrule*[right=$E_{2}$]{X \vdash A \land B}{X \vdash B} \qquad \inferrule*[right=$\land I$]{X \vdash B \quad X \vdash C}{X \vdash B \land C}
\]

Such construction applies to (unary) type theories for categories with products and co-products where the version with variables is more naturally presented in natural deduction style.

\begin{example}

An unary type theory for a category with coproducts has the following inference rules.
For objects

\[
\inferrule*[]{X \in \mathcal{G}}{\vdash X\;\text{type}} \quad \inferrule*[]{\;}{\vdash \mathbf{0}\; \text{type}} \quad \inferrule*[]{\vdash A\;\text{type} \quad \vdash B\;\text{type}}{\vdash A + B\;\text{type}}
\]

For morphisms
\[
\inferrule*[right=$id$]{\vdash X\;\text{type}}{x : X \vdash x : X} \quad \inferrule*[right=$fI$]{f \in \mathcal{G}(A,B) \quad x : X \vdash M : A}{x : X \vdash f(M) : B}
\]
\[
\inferrule*[right=$\mathbf{0}E$]{x : X \vdash M : \mathbf{0} \quad \vdash C\;\text{type}}{x : X \vdash \text{match}_{0}(M) : C}
\]
\[
\inferrule*[right=$+I1$]{x : X \vdash M : A}{x : X \vdash inl(M) : A + B} \quad \inferrule*[right=$+I2$]{x : X \vdash M : B}{x : X \vdash inr(M) : A + B}
\]
\[
\inferrule*[right=$+E$]{x : X \vdash M : A + B \quad u : A \vdash P : C \quad v : B \vdash Q : C}{x : X \vdash \text{match}_{A+B}(M,u.P,v.Q) : C}
\]
noting that in the very last rule $v$ and $u$ are bound in $u.P$ and $v.Q$.
\end{example}

To enforce the universal property of the coproduct we need to introduce the following equations
\[
\text{match}(inl(y), u.P, v.Q) \equiv P[inl(y) / u]
\]
\[
\text{match}(inr(y), u.P, v.Q) \equiv Q[inl(y) / v]
\]
and
\[
\text{match}(y, u.P[inl(u)/y], v.P[inr(v)/y]) \equiv P
\]
where $P : A + B \to C$ and the rule basically says that pattern matching is redundant here.
And finally the rule (uniqueness of the arrow from the initial object)
\[
\text{match}_{\mathbf{0}}(y) \equiv P, \forall P
\]
\textcolor{red}{I am not sure what the process of coming up with such rules should be. Do I start with introduction/elimination rules and then try and fix the issues that break universal properties?}

\subsection{Simple Type Theories}
To reason about operations with multiple inputs we need to have multiple entities in the antecedent.
The right categorical setting to do this is \textit{multicategories}.

\begin{definition}

A multigraph $\mathcal{G}$ consists of a set of objects $\mathcal{G}_{0}$ together with for each object $B$ and every list of objects $(A_1,\ldots,A_n)$ a set of arrows $\mathcal{G}(A_1,\ldots,A_n;B)$.
\end{definition}

\begin{definition}
A multicategory $\mathcal{M}$ is a multigraph with the following structure and properties
\begin{itemize}
  \item For each object $A$, an identify arrow $id_{A} \in \mathcal{M}(A,A)$
  \item For any object $C$ and a list of objects $(B_1, \ldots, B_m)$ and $(A_{i1}, \ldots, A_{i n_{i}})$ for $1 \leq i \leq m$, a composition operation
  \begin{align*}
    \mathcal{M}(B_1, \ldots, B_m; C) \times \prod_{i=1}^{m}\mathcal{M}(A_{i1}, \ldots, A_{in_{i}};B_{i}) &\to \mathcal{M}(A_{11}, \ldots, A_{m n_{m}};C)\\
    (g, (f_1, \ldots, f_m)) &\mapsto g \circ (f_1, \ldots, f_m)
  \end{align*}
  \item For any $f \in \mathcal{M}(A_1, \ldots, A_n; B)$ we have
    \[
      id_{B} \circ (f) = f \qquad f \circ (id_{A_1}, \ldots, id_{A_n}) = f
    \] 
  \item For any $h, g_i, f_{ij}$ we have
    \[
      (h \circ (g_1, \ldots, g_m)) \circ (f_{11}, \ldots, f_{mn_m}) = h \circ (g_1 \circ (f_{11}, \ldots, f_{1 n_{1}}), \ldots, g_{m} \circ (f_{m 1}, \ldots, f_{m n_{m}}))
    \]
\end{itemize}
\end{definition}

A good example is the category of vector spaces and multilinear maps.
We can also introduce an indexed composition $g \circ_i f_i$ as a surag for $g \circ (f_1, \ldots, f_n)$ when all $f_j$ for $i \not = j$ are identities and $g \in \mathcal{M}(A_1, \ldots, A_n; B)$.
A bit more formally we have
\[
\circ_{i} : \mathcal{M}(B_1, \ldots, B_n; C) \times \mathcal{M}(A_1, \ldots, A_m) \to \mathcal{M}(B_1, \ldots, B_{i-1}, A_{1}, \ldots, A_{m}, B_{i + 1}, \ldots, B_n; C)
\]
\[
\tikzfig{./figures/single_comp}
\]
that satisfy the following properties
\begin{itemize}
  \item $id_{B} \circ_{1} f = f$
  \item $f \circ_{i} id_{B_{i}} = f$
  \item If $h$ is $n$-ary, $g$ is $m$-ary, and $f$ is $k$-ary, then
  \[
    (h \circ_i g) \circ_j f = 
  \begin{cases}
    (h \circ_j f) \circ_{i + k - 1} g \qquad &\text{if}\; j < i\\
    (h \circ_i (g \circ_{j - i + 1} f)) \qquad &\text{if}\; i \leq j < i + m\\
    (h \circ_{j-m+1} f) \circ_i g \qquad &j \geq i + m
  \end{cases}
  \]
\end{itemize}

\begin{definition}

A tensor product in $\mathcal{M}$ is an object $\bigotimes_{i}A_{i}$ together with a morphism $\chi \in \mathcal{M}(A_1,\ldots,A_n; \bigotimes_{i}A_{i})$ such that all $(- \circ_{i} \chi_{i})$ are bijections.
\[
\mathcal{M}(B_{1},\ldots,B_{n},A_1,\ldots,A_m,C_{1},\ldots,C_{k};D) \xrightarrow{\sim} \mathcal{M}(B_{1},\ldots,B_{n},\bigotimes_{i}A_i,C_{1},\ldots,C_{k};D)
\]
\end{definition}

Here are some extensions with additional structures

\paragraph*{Symmetric monoidal structure} If $\mathcal{C}$ is a symmetric monoidal category, we have symmetry isomorphisms $A_1 \otimes \ldots \otimes A_n \xrightarrow{\sim} A_{\sigma 1} \otimes \ldots \otimes A_{\sigma n}$ for any permutation $\sigma \in S_{n}$.
Thus, by precomposing with these isomorphisms, we obtain functions between multicategorical hom-sets
\[
\sigma^* : \mathcal{C}(A_{\sigma 1}, \ldots, A_{\sigma n}; B) \to \mathcal{C}(A_1,\ldots,A_n;B)
\]
That is, suppose we have a multimorphism $f : A_{\sigma 1}, \ldots, A_{\sigma n} \to B$. We can precompose it with the isomorphism above to get
\[
\sigma^{*}(f) : A_1 \otimes \ldots \otimes A_n \xrightarrow{\sim} A_{\sigma 1} \otimes \ldots \otimes A_{\sigma n} \xrightarrow{f} B
\]
\textcolor{red}{I do not really see what is the point of these multicategories}
\textcolor{cyan}{My take on this is that is more general as it does not impose any structure on the list in the antecedent and one could make it into a tensor if they want but it also works if the multicategory is not representable.
It also does not require to invoke any associativity as the antecedent is a literal list.
Additional structure is imposed by universal properties.
}
\paragraph*{Cartesian structure} If $\mathcal{C}$ is a cartesian monoidal category, we have symmetries in addition to dual maps $A \to A \times A$ and projections $A \times B \to B$.
In general for any function $\sigma : \{1, \ldots, m\} \to \{1, \ldots, n\}$ we have a unique morphism
\[
A_1 \times \ldots \times A_n \to A_{\sigma 1} \times \ldots A_{\sigma m}
\]
whose component $A_1 \times \ldots \times A_n \to A_{\sigma k}$ is $\sigma k$-th projection.
Precomposition with those gives
\[
\sigma^{*} : \mathcal{C}(A_{\sigma 1}, \ldots, A_{\sigma m}; B) \to \mathcal{C}(A_{1}, \ldots, A_n; B)
\]

\begin{definition}
  Let $\mathfrak{N}$ be a full subcategory of $\catname{Set}$ whose objects are sets ${1, \ldots, n}$ for all integers $n \geq 0$.
  We regard it as a cocartesian strict monoidal category, under the disjoint union operation $\{1, \ldots, n\} \sqcup \{1,\ldots,m\} = \{1, \ldots, n + m\}$.
  Moreover, for any $\sigma : \{1, \ldots, m\} \to \{1, \ldots, n\}$ and $k_1, \ldots, k_n$, let $\sigma \wr (k_1, \ldots, k_n)$ denote the composite function
  \[
  \{1, \ldots, \sum_{i=1}^{m}k_{\sigma i}\} \xrightarrow{\sim} \bigsqcup_{i=1}^{m}\{1, \ldots, k_{\sigma i}\} \xrightarrow{\hat{\sigma}} \bigsqcup_{j=1}^{n}\{1, \ldots, k_j\} \xrightarrow{\sim} \{1, \ldots, \sum_{j=1}^{n} k_j\}
  \]
  where $\hat{\sigma}$ acts as an identity from the $i$-th summand to the $(\sigma i)^{\text{th}}$ summand.
  A \textbf{faithful cartesian club} is a category $\mathfrak{S} \subseteq \mathfrak{N}$ such that 
  \begin{enumerate}
    \item $\mathfrak{S}$ contains all the objects of $\mathfrak{N}$.
    \item $\mathfrak{S}$ is closed under the cocartesian structure, i.e. if $\sigma$ and $\tau$ are morphisms of $\mathfrak{S}$ then so is $\sigma \sqcup \tau$.
    \item $\mathfrak{S}$ is closed under $\wr$, i.e. whenever it contains $\sigma$ it also contains $\sigma \wr (k_1, \ldots, k_n)$
  \end{enumerate}
\end{definition}

The strange-looking $\sigma \wr (k_1, \ldots, k_n)$. 
Here's what's happening: Start with a function $\sigma : \{1, \ldots, m\} \to \{1, \ldots, n\}$ which says that each $i$ gets assigned one of the $n$ slots.
Now, instead of just $n$ slots, suppose each slot $j$ is \enquote{blown up} into $k_j$ copies.
The construction $\sigma \wr (k_1, \ldots, k_n)$ gives a new function
\[
\{1, \ldots, \sum_{i=1}^{m}k_{\sigma i}\} \to \{1, \ldots, \sum_{j=1}^{n}k_j\}
\]
which is what you get when you first group the inputs according to $\sigma$, and then send them into the right block of the \enquote{expanded} target.

We get symmetric monoidal category when $\mathfrak{S}$ consists of all bijections and cartesian structure when it consists of all functions.

\begin{definition}
  Let $\mathfrak{S}$ be a faithful cartesian club. An $\mathfrak{S}$-multicategory is a multicategory $\mathcal{M}$ together with operations
  \[
  \mathcal{M}(A_{\sigma 1}, \ldots, A_{\sigma m};B) \to \mathcal{M}(A_1, \ldots, A_n;B)
  \]
  \[
  f \mapsto f \circ \sigma^{*}
  \]
  for all functions $\sigma : \{1, \ldots, m\} \to \{1, \ldots, n\}$ in $\mathfrak{S}$, satisfying the following axioms:
  \begin{enumerate}
    \item $f\sigma^{*}\tau^{*} = f(\tau \sigma)^{*}$
    \item $f(id_n) = f$
    \item $g \circ (f_1\sigma^{*}_{1}, \ldots, f_{n}\sigma^{*}_n) = (g \circ (f_1, \ldots, f_n))(\sigma_1 \sqcup \ldots \sqcup)^{*}$
    \item $g \sigma^{*} \circ (f_1, \ldots, f_n) = (g \circ (f_{\sigma 1}, \ldots, f_{\sigma_m}))(\sigma \wr (k_1, \ldots, k_n))^{*}$, where $k_i$ is arity of $f_i$.
  \end{enumerate}
\end{definition}

Finally, we observe that closedness can be naturally characterized multicategorically.
Suppose for simplicity that $\mathfrak{S}$ contains at least all bijections. 
Then we say an $\mathfrak{S}$-multicategory is closed if for each pair of objects $A$ and $B$ there is a
specified object $A \multimap B$ and a morphism $\chi : A \multimap B \to B $ postcomposition
with which defines bijections
\[
(\chi \circ -) \mathcal{M}(C_1, \ldots, C_n; A \multimap B) \to \mathcal{M}(C_1, \ldots, C_n, A; B)
\]

With all of this in hand we can define type theory for symmetric monoidal closed categories (in natural deduction style)

\[
\inferrule*[right=$id$]{\vdash A\;\text{type}}{x : A \vdash x: A}
\]
\[
\inferrule*[right=$fI$]{f \in \mathcal{G}(A_1, \ldots, A_n; B)\;\Gamma_1 \vdash A_1, \ldots, \Gamma_n \vdash A_n\;\text{Shuf}(\Gamma_1,\ldots,\Gamma_n;\Phi)}{\Phi \vdash f(M_1, \ldots, M_n) : B}
\]
\[
\inferrule*[right=$\otimes I$]{\Gamma \vdash A \qquad \Delta \vdash B \qquad \text{Shuf}(\Gamma,\Delta;\Phi)}{\Phi \vdash \{A \otimes B\}}
\]
\textcolor{red}{The rule above seems to imply the tensor is not strict.}
\textcolor{red}{Can the below rule be made admissible? It looks like a composition.}
\[
\inferrule*[right=$\otimes E$]{\Psi \vdash M : A \otimes B \quad \Gamma, x : A, y : B \vdash N : C \quad \text{Shuf}(\Gamma,\Psi;\Phi)}{\Phi \vdash \text{match}_{A \otimes B}(M, xy.N) : C}
\]
\[
\inferrule*[right=$\multimap I$]{\Gamma, x : A \vdash M : B}{\Gamma \vdash \lambda x . M : A \multimap B}
\]
\textcolor{red}{Should there be a symmetry involved in the rule above?}
\[
\inferrule*[right=$\multimap E$]{\Gamma \vdash M : A \multimap B \quad \Delta \vdash N : A \quad \text{Shuf}(\Gamma, \Delta; \Phi)}{\Phi \vdash M N : B}
\]

To implement the universal properties, we need the following equations for the tensor
\[
\text{match}_{\otimes}(\{M,N\}, xy.P) \equiv P[M / x, N /y]
\]
\[
\text{match}_{\otimes}(M, xy.N[\{x,y\}/ u]) \equiv N[M / u]
\]

The above rules enjoy the properties of exchange being admissible, substitution being admissible, terms representing derivations.
In particular, substitution is associative and has an interchange property.
The initiality theorem is then that a free closed symmetric monoidal category on a multigraph is given by this type theory.
\textcolor{cyan}{What we might want to do is augment the above rules for the type theory to generate a semilattice enriched symmetric monoidal category and formulate the corresponding initiality theorem.
Terms calculus may provide an interesting angle on the differences between cartesian and non-cartesian case, for example in the former we may disregard some inputs inside one of the components of $f_1 + \ldots + f_n$ while in non-cartesian case we can not and therefore all the variables used by $f_i$ must be used by $f_j$ for $i \not = j$.}

\begin{remark}
  A term uniquely determines a derivation only when paired with its context.
  For example, below are two different derivations (in the presence of weakening).
  \[
  \inferrule*[]{
    \inferrule*[]{
      \vdash A\;\text{type}\;(x : A) \in \Gamma
      }
      {x : A, z : C, y : B \vdash x : A
      }
     \quad 
      \inferrule*[]{
        \vdash B\;\text{type}\;(y : B) \in \Gamma
        }
        {x : A, z : C, y : B \vdash y : B
        }
        }
      {x : A, z : C, y : B \vdash \{x,y\} : A \otimes B}
  \]

  \[
  \inferrule*[]{
    \inferrule*[]{
      \vdash A\;\text{type}\;(x : A) \in \Gamma
      }
      {x : A, y : B \vdash x : A
      }
     \quad 
      \inferrule*[]{
        \vdash B\;\text{type}\;(y : B) \in \Gamma
        }
        {x : A, y : B \vdash y : B
        }
        }
        {x : A, y : B \vdash \{x,y\} : A \otimes B}
  \]
\end{remark}

So far we can say that the context correspond to the input interface of our hypergraphs and the pushout along the interfaces correspond to a substitution.
It is unclear how to make a claim that a subgraph correspond to a term as they are defined using different structure.
\bibliographystyle{ACM-Reference-Format}
\bibliography{../bibliography}

\end{document}

### 2025-09-25 term-calculus/cat_logic_notes.tex
some updates

@@
 The calculus along with the reduction rules is shown to be strongly normalizing and confluent --- that is, equality of terms is straightforwardly decidable.
+
+\section{Term graphs}
+
+\textit{Term graphs} are essentially abstract syntax trees with sharing.
+The consequence of this is that unlike our hypergraphs, they have to be rooted.
+They share the same vocabulary with hypergraphs with some minor tweaks.
+\begin{definition}
+  A path between vertices $v_0$ and $v_n$ is a sequence $\langle v_0, i_1, v_1, \ldots, i_n, v_n \rangle$ of nodes and positive integers such that for $j = 1 \ldots n$, if $e$ is the unique edge such that $res(e) = v_{j-1}$, then $v_j$ is $i^{\text{th}}_j$ input to $e$.
+\end{definition}
+
+\begin{definition}
+  A (hyper-) graph $G$ is a \textit{term graph} if
+  \begin{enumerate}
+    \item there is a vertex $root_{G}$
+    \item $G$ is acyclic
+    \item each node is the result of a unique edge (i.e. every edge represents a functional symbol with co-artity of 1)
+  \end{enumerate}
+\end{definition}
+
+There is a canonical representative of an isomorphism class of term graphs given by a particular numbering of the vertices.
+Given a vertex $v$ an \textit{access} path to $v$ is a sequence $\langle i_1, \ldots, i_n \rangle$ such that there is a path $\langle v_0, i_1, v_1, \ldots, i_n, v_n \rangle$ where $v_0 = root_G$ and $v_n = v$.
+The set of access paths of $v$ is $Acc(v)$.
+The numbering is then given by
+\begin{enumerate}
+  \item $v = Acc(v)$
+  \item $e = res(e)$
+\end{enumerate}
+
+\textcolor{cyan}{We can do the same canonicalization for our (e-) hypergraphs by topologically sorting them. 
+This gives a simple procedure for deciding equality modulo SMC equations.}
+
+\[
+\tikzfig{./figures/canonical_term_graph_example}
+\]
+
+\begin{remark}
+  Several canonicalized term graphs can represent the same term (these differ in sharing/de-sharing of subtrees).
+  Graphs ordered by sharing/de-sharing relation form a lattice where the least element (with no sharing) is denoted $\Delta G$ and the maximal element (with everything shared) is denoted by $\nabla G$.
+\end{remark}
+
+People then study the relation between term rewriting and term-graph rewriting.
+For example, they answer the question of if term rewriting is confluent is the corresponding term-graph rewriting confluent?
+What is interesting is that soundness and completeness are formulated the other way around.
+Soundness is a statement of the form
+\begin{proposition}
+  For all term graphs $G$ and $H$, $G \xRightarrow{} H$ implies $term(G) \xrightarrow{+} term(H)$.
+\end{proposition}
+
+The other direction is then completeness.
+To get completeness we need to rewrite term-graphs modulo bisimilarity (collapsing and copying).
 
@@
 \]
+A unit $\mathbf{1}$ is given by a nullary tensor product.
 \end{definition}
+
+We call a multicategory representable if it is equipped with a chosed unit object and for every pair of objects with their tensor.
 
@@
 
-With all of this in hand we can define type theory for symmetric monoidal closed categories (in natural deduction style)
+\subsection{Type theory for (closed) symmetric monoidal categories}
+
+With all of this in hand we can define type theory for symmetric monoidal closed categories (in natural deduction style).
+Below $\text{Shuf}(\Gamma_1, \ldots, \Gamma_n; \Phi)$ says that $\Phi$ is a shuffle of $\Gamma_1, \ldots, \Gamma_n$, i.e. $\Phi$ is obtained by rifle-shuffling.
 
@@
 \[
-\inferrule*[right=$fI$]{f \in \mathcal{G}(A_1, \ldots, A_n; B)\;\Gamma_1 \vdash A_1, \ldots, \Gamma_n \vdash A_n\;\text{Shuf}(\Gamma_1,\ldots,\Gamma_n;\Phi)}{\Phi \vdash f(M_1, \ldots, M_n) : B}
+\inferrule*[right=$fI$]{f \in \mathcal{G}(A_1, \ldots, A_n; B)\;\Gamma_1 \vdash M_1 : A_1, \ldots, \Gamma_n \vdash M_n : A_n\;\text{Shuf}(\Gamma_1,\ldots,\Gamma_n;\Phi)}{\Phi \vdash f(M_1, \ldots, M_n) : B}
 \]
 \[
-\inferrule*[right=$\otimes I$]{\Gamma \vdash A \qquad \Delta \vdash B \qquad \text{Shuf}(\Gamma,\Delta;\Phi)}{\Phi \vdash \{A \otimes B\}}
+\inferrule*[right=$\otimes I$]{\Gamma \vdash N : A \qquad \Delta \vdash M : B \qquad \text{Shuf}(\Gamma,\Delta;\Phi)}{\Phi \vdash \{N \otimes M\} : A \otimes B}
 \]
@@
 \inferrule*[right=$\otimes E$]{\Psi \vdash M : A \otimes B \quad \Gamma, x : A, y : B \vdash N : C \quad \text{Shuf}(\Gamma,\Psi;\Phi)}{\Phi \vdash \text{match}_{A \otimes B}(M, xy.N) : C}
+\]
+\[
+\inferrule*[right=$\mathbf{1}I$]{\;}{() \vdash \star : \mathbf{1}} \qquad \inferrule*[right=$\mathbf{1}E$]{\Psi \vdash M : \mathbf{1} \qquad \Gamma,\Delta \vdash N : A}{\Gamma, \Psi.
+\Delta \vdash \text{match}_{\mathbf{1}}(M,N) : A}
 \]
@@
 
-To implement the universal properties, we need the following equations for the tensor
+To implement the universal properties, we need the following equations for the tensor ($\beta$ and $\eta$ laws)
 \[
@@
 \text{match}_{\otimes}(M, xy.N[\{x,y\}/ u]) \equiv N[M / u]
+\]
+
+\[
+\text{match}_{\mathbf{1}}(\star, N) \equiv N \qquad \text{match}_{\mathbf{1}}(M,N[\star / u]) \equiv N[M / u]
 \]
@@
 It is unclear how to make a claim that a subgraph correspond to a term as they are defined using different structure.
+
+The bottom line is that the above syntax is a \enquote{nice} syntax for morphisms in symmetric monoidal categories.
+For example, consider the below morphism, first represented as a string diagram, then represented as a raw term and finally represented using the syntax above.
+
+\[
+\tikzfig{./figures/string-diagram-example}
+\]
+
+\[
+(f \otimes g);\sigma_{D,E};h
+\]
+which is not unique as it is equal e.g. to the following term
+\[
+(f;id_{D} \otimes g);\sigma_{D,E};\sigma_{E,D};\sigma_{D,E};h
+\]
+
+\[
+\inferrule*{
+  \inferrule*{
+    \inferrule*{f \in \mathcal{G} \quad x : A \vdash x : A \quad y : B \vdash y : B}{x : A, y : B \vdash f(x,y) : D}
+    \qquad
+    \inferrule*{g \in \mathcal{G} \quad z : C \vdash z : C}{z : C \vdash g(z) : E}
+  }{x : A, y : B, z : C \vdash \{f(x,y), g(z)\} : D \otimes E}
+  \qquad
+  \inferrule*{h \in \mathcal{G} \quad v : E \vdash v : E \quad u : D \vdash u : D}
+  {u : D, v : E \vdash h(v,u)}
+}{x : A, y : B, z : C \vdash \text{match}(\{f(x,y), g(z)\},uv.h(v,u))}
+\]
+
+\begin{figure}
+\begin{align*}
+x[M/x] &= M\\
+y[M/x] &= y\\
+f(N_1, \ldots, N_n)[M/x] &= f(N_1, \ldots, N_i[M/x], \ldots, N_n) \quad \text{if $x$ occurs in $N_i$}\\
+\{N_1 \otimes N_2\}[M/x] &= \{N_1[M/x] \otimes N_2\} \quad \text{if $x$ occurs in $N_1$}\\
+\{N_1 \otimes N_2\}[M/x] &= \{N_1 \otimes N_2[M/x]\} \quad \text{if $x$ occurs in $N_2$}\\
+\text{match}(N_1,xy.N_2)[M/x] &= \text{match}(N_1[M/x], xy.N_2) \quad \text{if $x$ occurs in $N_1$}\\
+\text{match}(N_1,xy.N_2)[M/x] &= \text{match}(N_1, xy.N_2[M/x]) \quad \text{if $x$ occurs in $N_2$}\\
+(\lambda y . N)[M / x] &= \lambda y . N[M/x]\\
+(N_1 N_2)[M / x] &= (N_1[M/x])N_2 \quad \text{if $x$ occurs in $N_1$}\\
+(N_1 N_2)[M / x] &= (N_1)N_2[M/x] \quad \text{if $x$ occurs in $N_2$}
+\end{align*}
+\caption{Substitution definition}
+\label{fig:sub-def}
+\end{figure}
+
+\begin{lemma}[Substitution (Cut rule) is admissible]
+  Given derivations $\Gamma \vdash M : A$ and $\Delta, x : A, \Phi \vdash N : B$ we can derive
+  \[
+  \Gamma, \Delta, \Phi \vdash N[M / x]
+  \]
+  Substitution is defined in \cref{fig:sub-def}.
+  Note, in particular, that since contexts are disjoint (every variable is used once) we only recurse in a single subterm.
+\end{lemma}
+\begin{proof}
+  We will induct on the derivation of $\Delta, x : A, \Phi \vdash N : B$.
+  \begin{itemize}
+    \item If it is the $id$ rule then $\Delta = \varnothing = \Phi$ and $N : B = x : A$ and $B = A$. Then we have $M[x/x] = M$.
+    \item  
+    \[
+    \inferrule*{f \in \mathcal{G} \quad \Gamma_1 \vdash N_1, \ldots, \Gamma_n \vdash N_n}{\Delta , x : A, \Phi \vdash f(N_1, \ldots, N_n) : B = N : B}
+    \]
+    Then it must be that there exists $1 \leq i \leq n$ such that $x \in \Gamma_i$ and we can apply inductive hypothesis to $\Gamma \vdash M : A$ and $\Gamma', x : A, \Gamma'' \vdash N_i : A_i$ to obtain $N_i[M/x] : A_i$.
+    By applying the same derivation rule again we get $f(N_1, \ldots, N_i[M/x], \ldots, N_n)$.
+    \item The other rules are analogous.
+  \end{itemize}
+\end{proof}
+
+With the lemma above we can prove, for example, that $A \otimes \mathbf{1}$ is isomorphic to $A$.
+For this we need to construct two morphisms $A \otimes \mathbf{1} \to A$ and $A \to A \otimes \mathbf{1}$ and show that their composition is identity.
+First, let's build a morphism from $A \to A \otimes \mathbf{1}$
+
+\[
+\inferrule*{
+  x : A \vdash x : A \qquad () \vdash \star : \mathbf{1}
+}{x : A \vdash \{x,\star\} : A \otimes \mathbf{1}}
+\]
+the morphism $A \otimes \mathbf{1} \to A$ is then
+\[
+\inferrule*{
+\inferrule*{}{z : A \otimes \mathbf{1} \vdash z : A \otimes \mathbf{1}}
+\qquad
+\inferrule*{y : \mathbf{1} \vdash y : \mathbf{1} \qquad x : A \vdash x : A}{x : A, y : \mathbf{1} \vdash \text{match}_{\mathbf{1}}(y,x) : A}
+}{z : A \otimes \mathbf{1} \vdash \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)) : A}
+\]
+Then we have
+\begin{itemize}
+  \item \[
+  \inferrule*{
+  \inferrule*{
+  \inferrule*{        
+  \inferrule*{
+            \inferrule*{
+  w : A \vdash w : A \qquad () \vdash \star : \mathbf{1}
+}{w : A \vdash \{w,\star\} : A \otimes \mathbf{1}}
+\qquad
+\inferrule*{
+\inferrule*{}{z : A \otimes \mathbf{1} \vdash z : A \otimes \mathbf{1}}
+\qquad
+\inferrule*{y : \mathbf{1} \vdash y : \mathbf{1} \qquad x : A \vdash x : A}{x : A, y : \mathbf{1} \vdash \text{match}_{\mathbf{1}}(y,x) : A}
+}{z : A \otimes \mathbf{1} \vdash \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)) : A}
+}{w : A \vdash \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x))[\{w, \star\}/z]}}{
+w : A \vdash \text{match}_{A \otimes \mathbf{1}}(\{w, \star\}, xy.\text{match}_{\mathbf{1}}(y,x))}}{
+w : A \vdash \text{match}_{\mathbf{1}}(y,x)[w / x,  \star / y]}}{
+w : A \vdash \text{match}_{\mathbf{1}}(\star,w) \equiv w
+}
+      \]
+  \item The other direction is substitution 
+  \[
+  \inferrule*{
+  \inferrule*{z : A \otimes \mathbf{1} \vdash \{w, \star\}[\text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)) / w]}
+{z : A \otimes \mathbf{1} \vdash \{\text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)), \star\}}}
+{z : A \otimes \mathbf{1} \vdash \text{\textcolor{red}{What's next?}}}
+  \]
+
+\end{itemize}
+
+\begin{lemma}[Substitution is associative and interchanging]
+  \begin{enumerate}
+  \item If $\Gamma \vdash M : A$ and $\Delta, x : A, \Delta' \vdash N : B$ and $\Psi, y : B, \Psi' \vdash P : C$, then
+  \[
+  (P[M/x])[N/y] = P[M[N/y]/x]
+  \]
+  \item If $\Gamma \vdash M : A$ and $\Delta \vdash N : B$ and $\Psi, x : A, \Psi', y : B, \Psi'' \vdash P : C$, then
+  \[
+  (P[M/x][N/y]) = P[N/y][M/x]
+  \]
+  \end{enumerate}
+\end{lemma}
+
+\begin{proposition}
+  A free (closed) symmetric monoidal category on a (hyper-) multigraph $\mathcal{G}$ is given by the type theory above.
+  The objects are types and the morphisms are derivations (terms).
+\end{proposition}
+
+\subsection{Semilattice-enriched symmetric monoidal categories}
+
+We want to extend the rules of the previous section with
+
+\[
+\inferrule*{\Gamma \vdash M_1 : A, \ldots, \Gamma \vdash M_n : A}{
+  \Gamma \vdash M_1 + \ldots + M_n
+}
+\]
+
+$M_1, \ldots, M_n$ look different depending on what structural rules we have.
+In the most general case they all should use the same variables, but in a cartesian case they can ignore some of them.
+
+and prove the analogous statements, i.e. admissibility of substitution, associativity of substitution etc.
+Then we can say that a free semilattice-enriched symmetric monoidal category is given by appropriate type theory.
+
+\begin{example}
+  Recall the raw morphism representation for our e-graph example
+  \[(a \otimes ((2;\comonoid) \otimes (1;id));\sym);((* \otimes \counit + (id \otimes \counit \otimes id);<\!\!<) \otimes id);/~,\]
+  We can render it using the  \enquote{enhanced} term syntax.
+  \[
+  \inferrule*{
+  \inferrule*{
+  \inferrule*{\vdash a : Int \quad \vdash 2 : Int}{\vdash a * 2}
+  \quad
+  \inferrule*{\vdash a : Int \quad \vdash 1 : Int}{\vdash a <\!\!< 1}
+  }{\vdash (a * 2) + (a <\!\!< 1)}
+  \quad
+  \inferrule*{}{\vdash 1 : Int}
+  }
+  {((a * 2) + (a <\!\!< 1)) / 2}
+  \]
+\end{example}
+
 \bibliographystyle{ACM-Reference-Format}

### 2025-10-02 term-calculus/cat_logic_notes.tex
wip

@@
 \]
-\textcolor{red}{The rule above seems to imply the tensor is not strict.}
-\textcolor{red}{Can the below rule be made admissible? It looks like a composition.}
+% \textcolor{red}{The rule above seems to imply the tensor is not strict.}
+% \textcolor{red}{Can the below rule be made admissible? It looks like a composition.}
 \[
@@
 \]
-\textcolor{red}{Should there be a symmetry involved in the rule above?}
+% \textcolor{red}{Should there be a symmetry involved in the rule above?}
 \[
@@
 \[
-\text{match}_{\otimes}(\{M,N\}, xy.P) \equiv P[M / x, N /y]
+\text{match}_{\otimes}(\{M,N\}, xy.P) \equiv_{\beta_{\otimes}} P[M / x, N /y]
 \]
 \[
-\text{match}_{\otimes}(M, xy.N[\{x,y\}/ u]) \equiv N[M / u]
+\text{match}_{\otimes}(M, xy.N[\{x,y\}/ u]) \equiv_{\eta_{\otimes}} N[M / u]
 \]
@@
 \[
-\text{match}_{\mathbf{1}}(\star, N) \equiv N \qquad \text{match}_{\mathbf{1}}(M,N[\star / u]) \equiv N[M / u]
+\text{match}_{\mathbf{1}}(\star, N) \equiv_{\beta_{\mathbf{1}}} N \qquad \text{match}_{\mathbf{1}}(M,N[\star / u]) \equiv_{\eta_{\mathbf{1}}} N[M / u]
 \]
@@
 
+\begin{lemma}
+
+Each term paired with its context uniquely determines the derivation.
+\end{lemma}
+
+\begin{example}
+Let's show that the term $f(g(h(x)))$ has a unique derivation.
+
+\[
+\inferrule*{
+  \inferrule*{h \in \mathcal{G} \quad x : A \vdash x : A}{
+  \inferrule*{g \in \mathcal{G} \quad x : A \vdash h(x) : B}{f \in \mathcal{G} \quad x : A \vdash g(h(x)) : C}}}
+{x : A \vdash f(g(h(x))) : D}
+\]
+
+it might be tempting to write down another derivation like
+\[
+\inferrule*{f \in \mathcal{G} \quad y : B \vdash g(y) : C}{y : B \vdash f(g(y)) : D}
+\]
+but then you get stuck since the only way to apply this to $h(x)$ is by unraveling $f(g(y))$.
+\end{example}
+
 \begin{lemma}[Substitution (Cut rule) is admissible]
@@
             \inferrule*{
-  \inferrule*{        
-  \inferrule*{
-            \inferrule*{
   w : A \vdash w : A \qquad () \vdash \star : \mathbf{1}
@@
 }{z : A \otimes \mathbf{1} \vdash \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)) : A}
-}{w : A \vdash \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x))[\{w, \star\}/z]}}{
-w : A \vdash \text{match}_{A \otimes \mathbf{1}}(\{w, \star\}, xy.\text{match}_{\mathbf{1}}(y,x))}}{
-w : A \vdash \text{match}_{\mathbf{1}}(y,x)[w / x,  \star / y]}}{
-w : A \vdash \text{match}_{\mathbf{1}}(\star,w) \equiv w
-}
+}{w : A \vdash \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x))[\{w, \star\}/z]}
 \]
+
+\begin{align*}
+  \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x))[\{w, \star\}/z] &\equiv \text{match}_{A \otimes \mathbf{1}}(\{w, \star\}, xy.\text{match}_{\mathbf{1}}(y,x))\\
+  &\equiv_{\beta_{\otimes}} \text{match}_{\mathbf{1}}(\star,w)\\
+  &\equiv_{\beta_{\mathbf{1}}} w
+\end{align*}
+
   \item The other direction is substitution 
   \[
-  \inferrule*{
   \inferrule*{z : A \otimes \mathbf{1} \vdash \{w, \star\}[\text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)) / w]}
-{z : A \otimes \mathbf{1} \vdash \{\text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)), \star\}}}
-{z : A \otimes \mathbf{1} \vdash \text{\textcolor{red}{What's next?}}}
+{z : A \otimes \mathbf{1} \vdash \{\text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,x)), \star\}}
   \]
+
+\begin{align*}
+  \underbrace{\{\text{match}_{A \otimes \mathbf{1}}(\underbrace{z}_{M}, xy.\text{match}_{\mathbf{1}}(y,x)), \star\}}_{N} &\equiv_{\eta_{\otimes}} \text{match}_{A \otimes \mathbf{1}}(z, xy.\{\text{match}_{A \otimes \mathbf{1}}(\{x,y\}, x'y'.\text{match}_{\mathbf{1}}(y',x')), \star\})\\
+  &\equiv_{\beta_{\otimes}} \text{match}_{A \otimes \mathbf{1}}(z,xy.\{{\text{match}_{\mathbf{1}}(y,x), \star}\})\\
+  &\equiv_{\eta_{\mathbf{1}}}\text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(y,\{\text{match}_{\mathbf{1}}(\star,x), \star\}))\\
+  &\equiv_{\beta_{\mathbf{1}}} \text{match}_{A \otimes \mathbf{1}}(z, xy.\text{match}_{\mathbf{1}}(\underbrace{y}_{M},\underbrace{\{x, \star\}}_{N[\star / u]}))\\
+  &\equiv_{\beta_{\mathbf{1}}} \text{match}_{A \otimes \mathbf{1}}(z, xy.\{x, y\})\\
+  &\equiv_{\eta_{\otimes}} z\\
+\end{align*}
+
+The above proof uses a more general statement that 
+\begin{align*}
+\{\text{match}_{\otimes}(M, xy.N), P\} &\equiv_{\eta_{\otimes}} \text{match}_{\otimes}(M,xy.\{\text{match}_{\otimes}(\{x,y\}, x'y'.N), P\})\\
+&\equiv_{\beta_{\otimes}} \text{match}(M,xy.\{N[x/x',y/y'], P\})\\
+&\equiv_{\alpha} \text{match}_{\otimes}(M,xy.\{N,P\})
+\end{align*}
+
+\textcolor{red}{What's the intuition behind these rules?}
 
@@
 
-and prove the analogous statements, i.e. admissibility of substitution, associativity of substitution etc.
+We can then aim to prove the analogous statements, i.e. admissibility of substitution, associativity of substitution etc.
 Then we can say that a free semilattice-enriched symmetric monoidal category is given by appropriate type theory.
+
+We will have the following equations
+\begin{align*}
+  \{N_1 + \ldots + N_n, M\} &\equiv \{N_1, M\} + \ldots + \{N_n, M\}\\
+  \{M, N_1, \ldots, N_n\} &\equiv \{M, N_1\} + \ldots + \{M, N_n\}\\
+  N_1 + \ldots + (M_1 + \ldots + M_m)  + \ldots + N_n &\equiv N_1 + \ldots + M_1 + \ldots + M_m + \ldots + N_n\\
+  M + N &\equiv N + M\\
+  N &\equiv N + N
+\end{align*}
+
+and the substitution is (it should follow from admissibility)
+\begin{align*}
+(N_1 + \ldots + N_n)[M/x] &= (N_1[M/x] + \ldots + N_n[M/x])\\
+N[(M_1 + \ldots + M_n)/x] &= N[M_1/x] + \ldots + N[M_n/x]
+\end{align*}
+
+\begin{lemma}[Substitution is admissible]
+  If $\Gamma \vdash M : A$ and $\Delta, x : A, \Phi \vdash N$ are derivable, then $\Delta, \Gamma, \Phi \vdash N[M/x]$ is derivable.
+\end{lemma}
+\begin{proof}
+  Since the substitution is admissible in the original type theory, we only need to consider the extra rule that we introduced.
+  Suppose $N \Coloneqq N_1 + \ldots + N_n$ and therefore comes from the rule
+  \[
+  \inferrule[]{\inferrule[]{\vdots}{\Gamma \vdash M} \quad \inferrule[]{\Delta, x : A, \Phi \vdash N_1, \ldots, \Delta, x : A, \Phi \vdash N_n}{\Delta, x : A, \Phi \vdash N}}{\;}
+  \]
+  By IH we have derivations
+  \[
+    \Delta, \Gamma, \Phi \vdash N_1[M/x], \ldots, \Delta, \Gamma, \Phi \vdash N_n[M/x]
+  \]
+  and therefore can get a derivation 
+  \[
+  \Delta, \Gamma, \Phi \vdash N_1[M/x] + \ldots + N_n[M/x]
+  \]
+
+  The second rule is actually is not given by the lemma above, so we need to build it into the rules \textcolor{red}{How?
+I think if I extend every rule with a $+$ in the premise, it will absorb the rules above.}
+  % \[
+  % \inferrule[]{\Gamma \vdash N_1 + \ldots + N_n \quad f \in \mathcal{G}}{f(N_1) + \ldots + f(N_n)}
+  % \]
+  % then we have the following case.
+
+\end{proof}
+
+Consider the revised rules
+\begin{align*}
+  \inferrule[]{\vdash A\;\text{type}}{x : A \vdash x + \ldots + x : A}\\
+  \inferrule[]{f \in G \quad \Gamma_1 \vdash M_{11} + \ldots + M_{1 i_1} \ldots \Gamma_n \vdash M_{n 1} + \ldots + M_{n i_n}}{\Phi \vdash f(M_{1 1}, \ldots M_{n 1}) + \ldots + f(M_{1 i_1}, \ldots, M_{n i_n})}
+\end{align*}
+
+Then we can merge the substitution rules into a single rule
+\[
+(N_1 + \ldots + N_n)[M_1 + \ldots + M_m / x] = N_1[M_1 / x] + \ldots + N_n[M_1 / x] + \ldots + N_1[M_m / x] + \ldots + N_n[M_m / x]
+\]
 
@@
   \quad
-  \inferrule*{}{\vdash 1 : Int}
+  \inferrule*{}{\vdash 2 : Int}
   }
-  {((a * 2) + (a <\!\!< 1)) / 2}
+  {\vdash ((a * 2) + (a <\!\!< 1)) / 2}
   \]
+  Recall the string diagram
+  \[
+  \tikzfig{./figures/e-graph-example-string}
+  \]
+  We can linerise it as
+  \begin{minted}{haskell}
+    let (x_1, x_2, x_3) = (a, 2, 1) in
+      let y_1 = {(x_1 * x_2), (x_1 << x_3)} in
+        y1 / x_2
+  \end{minted}
+  \textcolor{red}{Why would one be preferable over the other?
+One syntax provides a reasoning framework about monoidal e-graphs while the other is just a linearisation of hypergraphs?
+Equality procedure for hypergraphs?}
 \end{example}

### 2025-10-08 term-calculus/cat_logic_notes.tex
slat type theory

@@
 x[M/x] &= M\\
-y[M/x] &= y\\
+% y[M/x] &= y\\
 f(N_1, \ldots, N_n)[M/x] &= f(N_1, \ldots, N_i[M/x], \ldots, N_n) \quad \text{if $x$ occurs in $N_i$}\\
@@
 
-We will have the following equations
+We will have the following equations (plus the usual $\beta/\eta$ for $\otimes$) 
 \begin{align*}
@@
 \begin{lemma}[Substitution is admissible]
-  If $\Gamma \vdash M : A$ and $\Delta, x : A, \Phi \vdash N$ are derivable, then $\Delta, \Gamma, \Phi \vdash N[M/x]$ is derivable.
+  If $\Gamma \vdash M_1 + \ldots + M_m : A$ and $\Delta, x : A, \Phi \vdash N$ are derivable, then $\Delta, \Gamma, \Phi \vdash N[M/x]$ is derivable.
 \end{lemma}
@@
   \[
-    \Delta, \Gamma, \Phi \vdash N_1[M/x], \ldots, \Delta, \Gamma, \Phi \vdash N_n[M/x]
-  \]
-  and therefore can get a derivation 
-  \[
-  \Delta, \Gamma, \Phi \vdash N_1[M/x] + \ldots + N_n[M/x]
-  \]
-
+    \Delta, \Gamma, \Phi \vdash N_1[M_1 + \ldots + M_m /x], \ldots, \Delta, \Gamma, \Phi \vdash N_n[M_1 + \ldots + M_m /x]
+  \]
+  or,
+  \[
+    \Delta, \Gamma, \Phi \vdash N_1[M_1/ x] + \ldots + N_1[M_m /x], \ldots, \Delta, \Gamma, \Phi \vdash N_n[M_1 / x] + \ldots + N_n[M_m /x]
+  \]
+  and therefore we can get a derivation 
+  \[
+  \Delta, \Gamma, \Phi \vdash (N_1[M_1/ x] + \ldots + N_1[M_m /x]) + \ldots + (N_n[M_1/ x] + \ldots + N_n[M_m /x])
+  \]
+
+  Now consider the case of
+
+  \[
+  \inferrule*{f \in \mathcal{G} \quad \Gamma_1 \vdash N_1, \ldots, \Gamma_n \vdash N_n}{\Delta , x : A, \Phi \vdash f(N_1, \ldots, N_n) : B}
+  \]
+
+  and there exists $1 \leq j \leq n$ such that $\Gamma_j = \Gamma'_j, x : A, \Gamma''_j \vdash N_j$.
+  Then, by IH we have a derivation $\Gamma'_j,\Gamma,\Gamma''_j \vdash N_j[M_1 + \ldots + M_m / x] = N_j[M_1 / x] + \ldots N_j[M_m / x]$.
+  By applying the same rule we only get 
+  \[
+  f(N_1, \ldots, (N_j[M_1/x] + \ldots + N_j[M_m / x]), \ldots, N_n)
+  \]
+  \textcolor{red}{which is not enough.}
   The second rule is actually is not given by the lemma above, so we need to build it into the rules \textcolor{red}{How?
 I think if I extend every rule with a $+$ in the premise, it will absorb the rules above.}
-  % \[
-  % \inferrule[]{\Gamma \vdash N_1 + \ldots + N_n \quad f \in \mathcal{G}}{f(N_1) + \ldots + f(N_n)}
-  % \]
-  % then we have the following case.
-
 \end{proof}
 
-Consider the revised rules
-\begin{align*}
-  \inferrule[]{\vdash A\;\text{type}}{x : A \vdash x + \ldots + x : A}\\
-  \inferrule[]{f \in G \quad \Gamma_1 \vdash M_{11} + \ldots + M_{1 i_1} \ldots \Gamma_n \vdash M_{n 1} + \ldots + M_{n i_n}}{\Phi \vdash f(M_{1 1}, \ldots M_{n 1}) + \ldots + f(M_{1 i_1}, \ldots, M_{n i_n})}
-\end{align*}
+Consider the revised rules (forget about the symmetry for a moment) in \autoref{fig:slat-typetheory}.
+\begin{figure}
+\begin{gather*}
+  % \inferrule*[right=$id$]{\vdash A\;\text{type}}{x : A \vdash x + \ldots + x : A}\\
+  \inferrule*[right=$id$]{\vdash A\;\text{type}}{x : A}\\
+  \inferrule*[right=$f$]{f \in \mathcal{G}(A_1,\ldots,A_n;B) \quad \Gamma_1 \vdash \sum_{j=1}^{i_1} M_{1j} : A_1 \ldots \Gamma_n \vdash \sum_{j=1}^{i_n} M_{nj} : A_n}{\Gamma_1, \ldots, \Gamma_n \vdash \sum_{k_1 = 1}^{j_1} \ldots \sum_{k_n=1}^{i_n} f(M_{1 k_1}, \ldots, M_{n k_n}) : B}\\
+  \inferrule*[right=$\otimes I$]{\Gamma \vdash \sum_{i=1}^{m}M_i : A \quad \Delta \vdash \sum_{i=1}^{n} N_i : B}{\Gamma, \Delta \vdash \sum_{i=1}^{m}\sum_{j=1}^{n}\{M_i, M_j\} : A \otimes B}\\
+  \inferrule*[right=$\otimes E$]{\Gamma \vdash \sum_{i=1}^{m}M_i : A \otimes B \qquad \Delta_1, x : A, y : B, \Delta_2 \vdash \sum_{i=1}^{n}N_i : C}{\Delta_1, \Gamma, \Delta_2 \vdash \sum_{1}^{i=m}\sum_{j=1}^{n}\text{match}_{A \otimes B}(M_i, xy.
+N_j) : C}\\
+  \inferrule*[right=$\mathbf{1}I$]{\;}{() \vdash \star : \mathbf{1}}\\
+  \inferrule*[right=$\mathbf{1}E$]{\Gamma \vdash \sum_{i=1}^{m} M_i : \mathbf{1} \qquad \Delta_1,\Delta_2 \vdash \sum_{i=1}^{n} N_i : A}{\Delta_1,\Gamma,\Delta_2 \vdash \sum_{i=1}^{m}\sum_{j=1}^{n} \text{match}_{\mathbf{1}}(M_i,N_j) : A}\\
+  \inferrule*[right=$+I$]{\Gamma \vdash \sum_{i=1}^{i_1} M_{1i} \quad \ldots \quad \Gamma \vdash \sum_{i=1}^{i_m} M_{mi}}{\Gamma \vdash \sum_{i=1}^{m}\sum_{j=1}^{i_j} M_{ij}}
+\end{gather*}
+\caption{Type theory for semilattice-enriched monoidal categories}
+\label{fig:slat-typetheory}
+\end{figure}
 
@@
 \]
+or, using a sum notation,
+\[
+(\sum_{i=1}^{n}N_i)[M_1 + \ldots + M_m / x] = \sum_{i=1}^{n}\sum_{j=1}^{m}N_i[M_j / x]
+\]
+the rest substitution rules are as in \autoref{fig:sub-def}.
+
+\begin{remark}
+  With this type theory we lose the property of the uniqueness of derivation. For example, the following term can be derived in two ways.
+  \[
+  \inferrule*[right=$fI$]{f \in \mathcal{G}(A;B) \qquad \inferrule*[right=$+I$]{x : A \vdash x : A \quad x : A \vdash x : A}{x : A \vdash x + x : A}}{x : A \vdash f(x) + f(x) : B}
+  \]
+  \[
+  \inferrule*[right=$+I$]{\inferrule*[right=$fI$]{f \in \mathcal{G}(A;B) \quad x : A \vdash x : A}{x : A \vdash f(x)} \qquad \inferrule*[right=$fI$]{f \in \mathcal{G}(A;B) \quad x : A \vdash x : A}{x : A \vdash f(x)}}{x : A \vdash f(x) + f(x) : B}
+  \]
+\end{remark}
+
+\begin{lemma}[Substitution is admissible]
+  If $\Gamma \vdash M_1 + \ldots + M_m : A$ and $\Delta, x : A, \Phi \vdash N : B$ are derivable, then $\Delta, \Gamma, \Phi \vdash N[M/x]$ is derivable.
+\end{lemma}
+\begin{proof}
+  \begin{enumerate}
+    \item Suppose $N \Coloneqq x : A \vdash x : A$, then $A = B$ and we can just use $M_1 + \ldots + M_m$ which is $x[M_1 + \ldots + M_n / x] = x[M_1/x] + \ldots + x[M_m / x] = M_1 + \ldots + M_m$
+    \item Suppose $N$ is the conclusion of the rule
+    \[
+      \inferrule*[right=$fI$]{f \in \mathcal{G}(A_1,\ldots,A_n;B) \quad \Gamma_1 \vdash \sum_{j=1}^{i_1} M_{1j} : A_1 \ldots \Gamma_n \vdash \sum_{j=1}^{i_n} M_{nj} : A_n}{\Gamma_1, \ldots, \Gamma_n \vdash \sum_{k_1 = 1}^{j_1} \ldots \sum_{k_n=1}^{i_n} f(M_{1 k_1}, \ldots, M_{n k_n}) : B}
+    \] 
+    then there exists $1 \leq j \leq n$ such that $x : A \in \Gamma_j$ and $\Gamma'_j, x : A, \Gamma''_j \vdash \sum_{i=1}^{i_j}N_{ji}$ and by IH we have a derivation
+    \[
+    (\sum_{i=1}^{i_j}N_{ji})[M_1 + \ldots + M_m / x] = \sum_{i=1}^{i_j}\sum_{k=1}^{m}N_{ji}[M_k / x]
+    \]
+    \begin{gather*}
+    \Gamma'_j,\Gamma,\Gamma''_j \vdash (N_{j 1} + \ldots + N_{j i_j})[M_1 + \ldots + M_m / x] =\\
+      = N_{j 1}[M_1 / x] + \ldots + N_{j 1}[M_m /x] + \ldots + N_{j i_j}[M_1 / x] + \ldots + N_{j i_j}[M_m /x]
+    \end{gather*}
+    and by applying the rule above we have
+    \begin{gather*}
+     \Gamma_1, \ldots, \Gamma'_j,\Gamma,\Gamma''_j, \ldots, \Gamma_n \vdash \sum_{k_1=1}^{i_1} \ldots \sum_{k_j=1}^{i_j}\sum_{l=1}^{m} \ldots \sum_{k_n=1}^{i_n} f(N_{1k_1}, \ldots, N_{jk_j}[M_{l} / x], \ldots, N_{nk_n}) =\\
+     = \sum_{k_1=1}^{i_1}\ldots\sum_{k_n=1}^{i_n}f(N_{1k_1}, \ldots, N_{j,k_j}[M_1 + \ldots + M_m], \ldots, N_{nk_n})\\
+     = \sum_{k_1=1}^{i_1}\ldots\sum_{k_n=1}^{i_n}f(N_{1k_1}, \ldots, N_{nk_n})[M_1 + \ldots + M_m]
+    \end{gather*}
+  \item Suppose $N$ is the conclusion of the rule
+  \[
+    \inferrule*[right=$\otimes I$]{\Delta_1 \vdash \sum_{i=1}^{p}P_i : A \quad \Delta_2 \vdash \sum_{i=1}^{n} N_i : B}{\Delta_1, \Delta_2 \vdash \sum_{i=1}^{p}\sum_{j=1}^{n}\{P_i, M_j\} : A \otimes B}
+  \]
+  and w.l.o.g suppose $x : A \in \Delta_1$, then by IH we have a derivation 
+  \[
+    \Delta'_1, \Gamma, \Delta''_1 \vdash (\sum_{i=1}^{p}P_i)[M_1 + \ldots + M_m] = (\sum_{i=1}^{p} \sum_{j=1}^{m}P_i[M_j / x]) : A
+  \]
+  and by applying the same rule we get
+  \begin{gather*}
+    \Delta'_1,\Gamma,\Delta''_1,\Delta_2 \vdash \sum_{i=1}^{p}\sum_{k=1}^{m}\sum_{j=1}^{n}\{P_i[M_j / x],N_j\} = \sum_{i=1}^{p}\sum_{j=1}^{n}\{P_i[M_1 + \ldots + M_m / x],N_j\} = \\
+    = \sum_{i=1}^{p}\sum_{j=1}^{n}\{P_i,N_j\}[M_1 + \ldots + M_m / x]
+  \end{gather*}
+  \item Suppose $N$ is the conclusion of the rule
+  \[
+    \inferrule*[right=$+I$]{\Delta', x : A, \Delta'' \vdash \sum_{i=1}^{i_1} N_{1i} \quad \ldots \quad \Delta', x : A, \Delta'' \vdash \sum_{i=1}^{i_n} N_{ni}}{\Delta', x : A, \Delta'' \vdash \sum_{i=1}^{n}\sum_{j=1}^{i_j} N_{ij}}
+  \]
+  then by IH we have 
+  \begin{gather*}
+  \Delta',\Gamma,\Delta'' \vdash (\sum_{i=1}^{i_1}N_{1i})[M_1 + \ldots + M_m] = (\sum_{i=1}^{i_1}N_{1i}[M_1 + \ldots + M_m]) = (\sum_{i=1}^{i_1}\sum_{j=1}^{m}N_{1i}[M_j / x])\\
+  \vdots\\
+  \Delta',\Gamma,\Delta'' \vdash (\sum_{i=1}^{i_n}N_{ni})[M_1 + \ldots + M_m] = (\sum_{i=1}^{i_1}N_{ni}[M_1 + \ldots + M_m]) = (\sum_{i=1}^{i_1}\sum_{j=1}^{m}N_{ni}[M_j / x])\\
+  \end{gather*}
+  and again by applying the same rule we get
+  \begin{gather*}
+    \sum_{k=1}^{n}\sum_{i=1}^{i_k}\sum_{j=1}^{m}(N_{ki}[M_j / x]) = \sum_{k=1}^{n}\sum_{i=1}^{i_k}(N_{ki}[M_1 + \ldots + M_m / x]) = \sum_{k=1}^{n}\sum_{i=1}^{i_k}(N_{ki})[M_1 + \ldots + M_m / x]
+  \end{gather*}
+  \end{enumerate}
+  The rest of the cases are analogous.
+  \textcolor{red}{Some metastatement that substitution is robust with respect to the ambiguity in derivations?}
+\end{proof}
+
+\begin{lemma}[Substitution is associative and interchanging]
+  TODO
+\end{lemma}
 

### 2025-10-24 term-calculus/cat_logic_notes.tex
type theory update

@@
 \newcommand{\parr}{\mathbin{\bindnasrepma}}
-% \usepackage{amssymb}
-\usepackage{mathpartir}
 
@@
 \input{../hypergraph.tikzdefs}
+\renewcommand*{\itemautorefname}{}
 
@@
 Then we need to require that the comonad interacts with the tensor.
+
+In more detail, below is two different flavours of conjunction rules in a logic where all structural rules are allowed.
+
+\[
+\inferrule*[right=$\land L$]{\Gamma, A \vdash \Delta}{\Gamma, A \land B \vdash \Delta} \qquad \inferrule*[right=$\land R$]{\Gamma, B \vdash \Delta}{\Gamma, A \land B \vdash \Delta} \qquad \inferrule*[]{\Gamma \vdash A, \Delta \quad \Gamma \vdash B, \Delta}{\Gamma \vdash A \land B, \Delta}
+\]
+which is an \textit{additive} version, and
+\[
+\inferrule*[right=$\land$]{\Gamma, A, B \vdash \Delta}{\Gamma, A \land B \vdash \Delta} \qquad \inferrule*[]{\Gamma_1 \vdash A, \Delta_1 \quad \Gamma_2 \vdash B, \Delta_2}{\Gamma_1, \Gamma_2 \vdash A \land B, \Delta_1, \Delta_2}
+\]
+which is a \textit{multiplicative} version.
+
+These presentations are equivalent in the presence of structural rules. For example, we can derive multiplicative $\land$ using the additive $\land L$ and $\land R$:
+
+\[
+  \inferrule{\Gamma, A, B \vdash \Delta}
+  {\inferrule{\Gamma, B, A \land B \vdash \Delta}
+    {\inferrule{
+      \Gamma, A \land B, A \land B \vdash \Delta
+      }{
+        \Gamma, A \land B \vdash \Delta
+        }}}
+\]
+and the other direction
+\[
+\inferrule[]{\Gamma, A \vdash \Delta}{
+  \inferrule{\Gamma, A, B \vdash \Delta}{\Gamma, A \land B \vdash \Delta}}
+\]
+
+If we remove contraction and weakening, then the presentations are no longer equivalent and hence define different connectives, $\otimes$ and $\&$, respectively called \enquote{tensor} and \enquote{with} and also two connectives for multiplicative and additive disjunction $\parr$ and $\oplus$.
+
+\[
+\begin{array}{cc}
+\textbf{Identity rules} & \inferrule*[right=$id$]{\;}{A, A^{\bot}} \qquad \inferrule*[right=cut]{\Gamma,A \quad A^{\bot},\Delta}{\Gamma, \Delta}\\
+\textbf{Multiplicative rules} &  \inferrule*[right=$\parr$]{\Gamma, A, B}{\Gamma, A \parr B} \qquad \inferrule*[right=$\otimes$]{\Gamma,A \quad \Delta,B}{\Gamma,\Delta,A \otimes B}\\
+\textbf{Additive rules} & \inferrule*[right=$\oplus_0$]{\Gamma,A}{\Gamma, A \oplus B} \qquad \inferrule*[right=$\oplus_1$]{\Gamma,B}{\Gamma, A \oplus B} \qquad \inferrule*[right=$\&$]{\Gamma,A \quad \Gamma, B}{\Gamma, A \& B}
+\end{array}
+\]  
+Here's a useful intuition behind $\otimes$ and $\parr$.
+Suppose you are given two resources, a \textit{saw} and a \textit{hammer}, then you can construct $\text{saw} \otimes \text{hammer}$ which is a set of a saw and a hammer and you can give a saw to one person and the hammer to another.
+In contrast, $\text{saw} \parr \text{hammer}$ is like a multitool, you can give it to a single person only.
+The difference between $\&$ and $\oplus$ is a bit more subtle. 
+Consider the menu below
+
+\begin{minipage}{0.45\linewidth}
+  \begin{tabular}{c}
+    \textbf{Menu (price 17 Euros)}\\
+    \textit{Quiche or Salad}\\
+    \textit{Chicken or Fish}\\
+    \textit{Banana or \enquote{Surprise du Chef}}\textsubscript{*}\\
+    {}\\
+    \textit{(*) either \enquote{Profiteroles} or \enquote{Tarte Tatin}}
+
+  \end{tabular}
+\end{minipage}
+\hfill
+\begin{minipage}{0.45\linewidth}
+  \[
+  \begin{aligned}
+  17E \vdash 
+  \left\{ \begin{array}{c}
+    (Q \& S)\\
+    \otimes\\
+    (C \& F)\\
+    \otimes\\
+    (B \& (P \oplus T))
+  \end{array}
+  \right.
+\end{aligned}
+  \]
+\end{minipage}
+
+So let us start again from the beginning, considering a play between the restaurant manager (the Player) and the customer (the Opponent). 
+It is the Player's responsibility to split the $17E$ into three parts, corresponding to the cost of the three parts of the meal. 
+May be, this is done as follows:
+\[
+\inferrule*{5E \vdash Q \& S \quad 8E \vdash C \& F \quad 4E \vdash B \& (P \oplus T)}{17E \vdash (Q \& S) \otimes (C \& F) \otimes (B \& (P \oplus T))}
+\]
+Now let the Opponent challenge $5E \vdash Q \& S$:
+\[
+\inferrule*[]{5E \vdash Q \qquad 5E \vdash S}{5E \vdash Q \& S}
+\]
+
+which reads as: both Quiche and Salad are available to the customer, but he can get only one, and it is his choice of picking one of the antecedents and to order, say, a Quiche. 
+Thus the additive conjunction can be understood as a $\ldots$ disjunction embodying a notion of external choice (remember that in our example the customer is the Opponent, or the context, or the environment).
+Let us now analyse a proof of $4E \vdash B \& (P \otimes T)$:
+
+\[
+\inferrule*{4E \vdash B  \qquad \inferrule*[]{4E \vdash T}{4E \vdash P \oplus T}}{4E \vdash B \& (P \oplus T)}
+\]
+
+Suppose that the Opponent chooses the Surprise. 
+Then it is the Player's turn, who justifies $4E \vdash O \oplus T$ using the right $\oplus$ rule. 
+So, the Opponent will get a Tarte Tatin, but the choice was in the Player's hands. 
+Thus $\oplus$ has an associated meaning of internal choice. 
+In summary, two forms of choice, external and internal, are modelled by N and $\oplus$, respectively. 
+In the case of $\oplus$, whether $A$ or $B$ is chosen is controlled by the rule, that is, by the Player. 
+In the case of $\&$, the choice between $A$ or $B$ is a choice of one of the antecedents of the rule, and is in the hands of the Opponent.
+
+All these oppositions confirm a fundamental polarity: by convention, we shall term $\&$ and $\parr$ as negative, and $\otimes$ and $\oplus$ as positive.
+
+\begin{remark}
+
+Linear logic does not have the property that terms uniquely determine derivations.
+Consider the following two derivations for example
+\[
+\inferrule*[right=$id$]{\;}{A \oplus B, A^{\bot} \;\&\; B^{\bot}} \qquad \inferrule*[]{\inferrule*[]{\inferrule*[right=$id$]{\;}{A,A^{\bot}}}{A \oplus B, A^{\bot}} \quad \inferrule*[]{\inferrule*[right=$id$]{\;}{B,B^{\bot}}}{A \oplus B, B^{\bot}}}{A \oplus B, A^{\bot} \;\&\; B^{\bot}}
+\]
+
+as negation is defined as
+\begin{gather*}
+(A \;\&\; B)^{\bot} = A^{\bot} \oplus B^{\bot}\\
+(A \otimes B)^{\bot} = A^{\bot} \parr B^{\bot}
+\end{gather*}
+
+Therefore, such derivations need to be identified.
+Quotienting by the identity above and a bunch of others defines $\equiv_{\eta}$.
+Quotienting by \texttt{cut}  defines $\equiv_\beta$.
+\end{remark}
+
+\begin{remark}
+We also have that some rules $r_1,r_2$ commute in the sense that they can be applied in any order without loss of provability.
+For example,
+\[
+\inferrule*[right=$\&$]{\inferrule*[right=$\parr$]{\Gamma, A, B, C \quad \Gamma, A, B, D}{\Gamma, A, B, C \;\&\; D}}{\Gamma, A \parr B, C \;\&\; D} 
+\qquad
+\inferrule*[right=$\&$]{\inferrule*[right=$\parr$]{\Gamma, A, B, C}{\Gamma, A \parr B, C} \quad \inferrule*[right=$\parr$]{\Gamma, A, B, D}{\Gamma, A \parr B, D}}{\Gamma, A \parr B, C \;\&\;D}
+\]
+\end{remark}
 
@@
 \begin{lemma}[Substitution is associative and interchanging]
+
   \begin{enumerate}
@@
 
-\begin{lemma}[Substitution is admissible]
+\begin{lemma}[Substitution]
   If $\Gamma \vdash M_1 + \ldots + M_m : A$ and $\Delta, x : A, \Phi \vdash N$ are derivable, then $\Delta, \Gamma, \Phi \vdash N[M/x]$ is derivable.
@@
   \inferrule*[right=$\mathbf{1}E$]{\Gamma \vdash \sum_{i=1}^{m} M_i : \mathbf{1} \qquad \Delta_1,\Delta_2 \vdash \sum_{i=1}^{n} N_i : A}{\Delta_1,\Gamma,\Delta_2 \vdash \sum_{i=1}^{m}\sum_{j=1}^{n} \text{match}_{\mathbf{1}}(M_i,N_j) : A}\\
-  \inferrule*[right=$+I$]{\Gamma \vdash \sum_{i=1}^{i_1} M_{1i} \quad \ldots \quad \Gamma \vdash \sum_{i=1}^{i_m} M_{mi}}{\Gamma \vdash \sum_{i=1}^{m}\sum_{j=1}^{i_j} M_{ij}}
+  \inferrule*[right=$+I$]{\Gamma \vdash \sum_{i=1}^{i_1} M_{1i} : A \quad \ldots \quad \Gamma \vdash \sum_{i=1}^{i_m} M_{mi} : A}{\Gamma \vdash \sum_{i=1}^{m}\sum_{j=1}^{i_j} M_{ij} : A}
 \end{gather*}
@@
 
-\begin{lemma}[Substitution is admissible]
-  If $\Gamma \vdash M_1 + \ldots + M_m : A$ and $\Delta, x : A, \Phi \vdash N : B$ are derivable, then $\Delta, \Gamma, \Phi \vdash N[M/x]$ is derivable.
-\end{lemma}
-\begin{proof}
+\begin{definition}[Substitution]
+  Given derivations $\Gamma \vdash M_1 + \ldots + M_m : A$ and $\Delta_1, x : A, \Delta_2 \vdash N$,
+  we define $\Delta_1, \Gamma, \Delta_2 \vdash N[M_1 + \ldots + M_m / x]$ by induction on derivation of $N$.
+  Moreover, we have the property that $(N_1 + \ldots + N_n)[M_1 + \ldots + M_m / x] = \sum_{i=1}^{n}\sum_{j=1}^{m}N_i[M_j / x]$.
   \begin{enumerate}
-    \item Suppose $N \Coloneqq x : A \vdash x : A$, then $A = B$ and we can just use $M_1 + \ldots + M_m$ which is $x[M_1 + \ldots + M_n / x] = x[M_1/x] + \ldots + x[M_m / x] = M_1 + \ldots + M_m$
+    \item Suppose $N \Coloneqq x : A \vdash x : A$, then $A = B$ and we can just use $M_1 + \ldots + M_m$ and define $\Gamma \vdash x[M_1 + \ldots + M_m / x] = x[M_1 / x] + \ldots + x[M_m / x] = M_1 + \ldots + M_m$.
     \item Suppose $N$ is the conclusion of the rule
@@
   The rest of the cases are analogous.
-  \textcolor{red}{Some metastatement that substitution is robust with respect to the ambiguity in derivations?}
+\end{definition}
+
+Then the admissibility follows from the above construction.
+\begin{lemma}
+  If we have derivations $\Gamma \vdash M_1 + \ldots + M_n : A$ and $\Delta_1, x : A, \Delta_2 \vdash N : B$, then we can derive $\Delta_1, \Gamma, \Delta_2 \vdash N[M_1 + \ldots + M_m / x]$.
+\end{lemma}
+
+The ambiguity technically means that we can not recursively define the action of a substitution on terms, however, we can see that applying the substitution to different derivations produces the same result
+\[
+  \inferrule*[right=$fI$]{f \in \mathcal{G}(A;B) \qquad \inferrule*[right=$+I$]{x : A \vdash x[M_1 + M_2 / x] = M_1 + M_2 : A \quad x : A \vdash x[M_1 + M_2 / x] = M_1 + M_2 : A}{x : A \vdash (x + x)[M_1 + M_2 / x] = (M_1 + M_2) + (M_1 + M_2) \equiv M_1 + M_2 + M_1 + M_2 : A}}{x : A \vdash (f(x) + f(x))[M_1 + M_2 / x] = f(M_1 + M_2) + f(M_1 + M_2) \equiv f(M_1) + f(M_2) + f(M_1) + f(M_2) : B}
+\]
+\[
+  \inferrule*[right=$+I$]{\inferrule*[right=$fI$]{f \in \mathcal{G}(A;B) \quad x : A \vdash x[M_1 + M_2 / x] = M_1 + M_2 : A}{x : A \vdash f(x)[M_1 + M_2 / x] = f(M_1) + f(M_2)} \quad \inferrule*[right=$fI$]{f \in \mathcal{G}(A;B) \quad x : A \vdash x[M_1 + M_2 / x] = M_1 + M_2 : A}{x : A \vdash f(x)[M_1 + M_2 / x] = f(M_1) + f(M_2)}}{x : A \vdash (f(x) + f(x))[M_1 + M_2 / x] = f(M_1) + f(M_2) + f(M_1) + f(M_2) : B}
+\]
+\begin{lemma}[Substitution is associative and interchanging]
+  We have the following properties for the substitution.
+  \begin{enumerate}[label=(\alph*)]
+    \item\label{subst_prop:a} $(N_1 + \ldots + N_n)[M_1 + \ldots + M_m / x] = \sum_{i=1}^{n}\sum_{j=1}^{m}N_i[M_j / x]$ (shown above)
+    \item\label{subst_prop:b} If $\Gamma \vdash N : B$ and $\Delta, y : B, \Delta' \vdash M : A$ and $\Psi, x : A, \Psi' \vdash P : C$, then
+      \[
+      (P[M/x])[N/y] = P[M[N/y]/x]
+      \]
+    \item\label{subst_prop:c} If $\Gamma \vdash M : A$ and $\Delta \vdash N : B$ and $\Psi, x : A, \Psi', y : B, \Psi'' \vdash P : C$, then
+      \[
+      (P[M/x][N/y]) = P[N/y][M/x]
+      \]
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  \leavevmode\\
+  We will first prove~\autoref{subst_prop:b}.
+  \begin{itemize}
+    \item Suppose $P \Coloneqq x : A \vdash x : A$, then the result of $P[M/x] = M$ and therefore $(P[M/x])[N / y] = P[M[N/y]/x] = M[N/y]$
+    \item Suppose 
+    \begin{gather*}
+      P \Coloneqq \inferrule[]{
+        f \in \mathcal{G}(A_1, \ldots, A_n;C) \quad \Gamma_1 \vdash \sum_{j=1}^{i_1} M_{1j} : A_1 \ldots \Gamma_n \vdash \sum_{j=1}^{i_n}M_{nj} : A_n
+        }{
+          \Gamma_1, \ldots, \Gamma_n \vdash \sum_{k_1=1}^{j_1}\ldots\sum_{k_n=1}^{i_n}f(M_{1k_1},\ldots,M_{nk_{n}}): C
+        }
+    \end{gather*}
+    then 
+    \[
+    \Psi, \Delta, y : B, \Delta', \Psi' \vdash P[M/x] = \sum_{k_1=1}^{j_1}\ldots\sum_{k_n=1}^{i_n}f(M_{1k_1},\ldots, M_{ik_i}[M/x] ,\ldots,M_{nk_{n}})
+    \]
+    and then 
+    \[
+    \Psi,\Delta,\Gamma,\Delta',\Psi' \vdash (P[M/x])[N/y] = \sum_{k_1=1}^{j_1}\ldots\sum_{k_n=1}^{i_n}f(M_{1k_1},\ldots, (M_{ik_i}[M/x])[N/y] ,\ldots,M_{nk_{n}})
+    \]
+    by IH we have that
+    \[
+      (M_{i,k_i}[M/x])[N/y] = M_{i,k_i}[M[N/y]/x]
+    \]
+    and
+    \[
+      \Psi,\Delta,\Gamma,\Delta',\Psi' \vdash (P[M[N/y]/x]) = \sum_{k_1=1}^{j_1}\ldots\sum_{k_n=1}^{i_n}f(M_{1k_1},\ldots, (M_{ik_i}[M[N/y] / x]) ,\ldots,M_{nk_{n}})
+    \]
+    \item Suppose 
+    \begin{gather*}
+    P \Coloneqq \inferrule[]{\Gamma \vdash \sum_{i=1}^{i_1}P_{1i} : C \ldots \Gamma \vdash \sum_{i=1}^{i_n}P_{ni} : C}{\Gamma \vdash \sum_{i=1}^{n}\sum_{j=1}^{i_j}P_{ij} : C}
+    \end{gather*}
+    and $\Gamma \Coloneqq \Psi, x : A, \Psi'$.
+    We have that 
+    \[
+    (P[M/x])[N/y] = \sum_{i=1}^{n}\sum_{j=1}^{j_i}(P_{ij}[M/x])[N/y]
+    \] 
+    and by IH we have that $(P_{ij}[M/x])[N/y] = P_{ij}[M[N/y]/x]$ and
+    \[
+      (P[M/x])[N/y] = \sum_{i=1}^{n}\sum_{j=1}^{j_i}(P_{ij}[M[N/y]/x]) = P[M[N/y]/x]
+    \]
+    \item The other cases are analogous
+  \end{itemize}
+  Now for the~\autoref{subst_prop:c}.
+  \begin{itemize}
+    \item Suppose
+      \begin{gather*}
+        P \Coloneqq \inferrule[]{
+          f \in \mathcal{G}(A_1, \ldots, A_n;C) \quad \Gamma_1 \vdash \sum_{j=1}^{i_1} P_{1j} : A_1 \ldots \Gamma_n \vdash \sum_{j=1}^{i_n}P_{nj} : A_n
+          }{
+            \Gamma_1, \ldots, \Gamma_n \vdash \sum_{k_1=1}^{j_1}\ldots\sum_{k_n=1}^{i_n}f(P_{1k_1},\ldots,P_{nk_{n}}): C
+          }
+      \end{gather*}
+      It must be that $x : A$ appears in $\Gamma_w$ and $y : B$ appears in $\Gamma_u$. If $w = u$ we simply apply IH to $\Gamma_w \vdash \sum_{j=1}^{i_w}P_{wj}$.
+      Otherwise, we have
+      \begin{align*}
+      P[M/x][N/y] &= \Gamma_1, \ldots, \Gamma_n \vdash \sum_{k_1=1}^{j_1}\ldots\sum_{k_n=1}^{i_n}f(P_{1k_1},\ldots, P_{wk_w}[M/x],\ldots, P_{uk_{w}}[N/y] ,\ldots,P_{nk_{n}}): C\\
+                  &= P[N/y][M/x]
+      \end{align*}
+    \item Suppose
+    \begin{gather*}
+      P \Coloneqq \inferrule[]{\Gamma \vdash \sum_{i=1}^{i_1}P_{1i} : C \ldots \Gamma \vdash \sum_{i=1}^{i_n}P_{ni} : C}{\Gamma \vdash \sum_{i=1}^{n}\sum_{j=1}^{i_j}P_{ij} : C}
+    \end{gather*}
+    and we have that 
+    \begin{gather*}
+      P[M/x][N/y] \Coloneqq \inferrule[]{\Gamma \vdash \sum_{i=1}^{i_1}P_{1i}[M/x][N/y] : C \ldots \Gamma \vdash \sum_{i=1}^{i_n}P_{ni}[M/x][N/y] : C}{\Gamma \vdash \sum_{i=1}^{n}\sum_{j=1}^{i_j}P_{ij}[M/x][N/y] : C}
+    \end{gather*}
+    and by IH we have that $P_{ji}[M/x][N/y] = P_{ji}[N/y][M/x]$ and 
+    \[
+      \Gamma \vdash \sum_{i=1}^{n}\sum_{j=1}^{i_j}P_{ij}[M/x][N/y] = \Gamma \vdash \sum_{i=1}^{n}\sum_{j=1}^{i_j}P_{ij}[N/y][M/x]
+    \] 
+    \item The rest are analogous.
+  \end{itemize}
 \end{proof}
 
-\begin{lemma}[Substitution is associative and interchanging]
-  TODO
-\end{lemma}
+% $M_1 + M_2 $ $N$
+% ------------------------
+% $\{M_1, N\} + \{M_2, N\}$
+% -------------------------
+% $\{M_1,N\}$, $\{M_2,N\}$
+% -----------------------
+% $\{M_1,N\} + \{M_2,N\}$
+
+\textcolor{red}{TODO: The same issue with non-unique derivations also occurs in linear logic where such rules are said to commute.
+While this does not pose any problems to substitution, we would need to make sure that in the proof of initiality theorem the equivalence between derivations is respected.
+Prove that substitution respects $\equiv$ (This is because substitution does not alter the order of derivation rules).
+}
 
@@
   \]
-  We can linerise it as
+  We can linearise it as
   \begin{minted}{haskell}

### 2025-10-31 term-calculus/cat_logic_notes.tex
type theory update

@@
 
+Pictorially these properties can be displayed as
+\[
+\tikzfig{./figures/composition_assoc}
+\]
+
+and therefore we have that $1 \leq i \leq k$ and $1 \leq j \leq k + m$.
+Suppose, $j < i$ in which case we insert $f$ into $h$ avoiding $g$ and get $(h \circ_j f) \circ_{i + k - 1} g$.
+If $i \leq j \leq i + m$, we have the same diagram just with different parentheses $h \circ_i (g \circ_{j - i + 1} f)$.
+Similarly, we have that when $j \geq i + m$ then $(h \circ_{j - m + 1} f) \circ_{i} g$.
+When we phrase the composition in terms of a substitution, these become the associativity and the interchange properties.
 \begin{definition}
@@
 
-We call a multicategory representable if it is equipped with a chosed unit object and for every pair of objects with their tensor.
+We call a multicategory representable if it is equipped with a chosen unit object and for every pair of objects with their tensor.
 
@@
 
+\begin{theorem}
+There is an equivalence $\catname{RepMCat} \simeq \catname{MonCat}$.
+\textcolor{red}{Can I define \textit{semilattice-enriched} $\catname{RepMCat}$ and restate the theorem above?}
+\end{theorem}  
+\begin{remark}
+  We will eventually need to show that the type theory above defines a suitable multicategory and that this multicategory is representable (and therefore is equivalent to some monoidal category).
+  The latter follows by the following argument. 
+  We have a unit given by $\star$ and for each pair of objects we have their tensor product given by $\{-,-\}$.
+  For the latter to be a tensor we need to make sure that we have a morphism $\chi : (A_1, \ldots, A_n) \to \otimes_i A_i$.
+  It is given by the following derivation
+  \[
+  \inferrule*[right=$\chi$]{x_1 : A_1 \vdash x_1 : A_1 \ldots x_n : A_n \vdash x_n : A_n}{x_1 : A_1, \ldots, x_n : A_n \vdash \{x_1,\ldots,x_n\} : A_1 \otimes \ldots \otimes A_n}
+  \]
+  We need to show that
+  \[
+  \mathcal{M}(B_1,\ldots,B_k, \otimes_i A_i, C_1, \ldots, C_m; D) \xrightarrow{\sim} \mathcal{M}(B_1, \ldots, B_k,A_1, \ldots, A_n,C_1,\ldots, C_m ; D)
+  \]
+  Suppose we have a morphism from the left-hand side that is given by the derivation
+  \[
+  \inferrule*[right=$f$]{\ldots}{B_1, \ldots, B_k, z : \otimes_i A_i, C_1, \ldots, C_m} \vdash D
+  \]
+  and we do $f[\chi / z] : B_1, \ldots, B_k, A_1, \ldots, A_n, C_1, \ldots, C_m \vdash D$.
+  In the other direction we do
+  \[
+  \inferrule*[]{\ldots}{B_1,\ldots,B_k, \otimes_i A_i, C_1, \ldots, C_m \vdash \text{match}(z, x_1\;\ldots\;x_n . D(x_1, \ldots, x_n))}
+  \]
+  These are inverses by $\beta/\eta$ laws.
+\end{remark}
+
 \begin{example}

### 2025-11-12 POPL2026/paper/sections/categorical_semantics.tex
moved all notes to a the same folder

@@
 	A set of $\Sigma$-terms generated by monoidal signature $\Sigma$ is given by the following grammar:
-	\[
-		\tau \coloneqq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2},
-	\]
+	$\tau \coloneqq \phi \;|\; \id_{I} \;|\; \id_{A} \;|\; \sym_{A,B} \;|\; \tau_{1};\tau_{2} \;|\; \tau_{1} \otimes \tau_{2}$,
 	where $\phi$ ranges over the set of operators $\Sigma_{M}$ and $A,B$ range over the set of objects $\Sigma_{O}$.
@@
 	\label{def:enriched-prop}
-	A symmetric monoidal $\catname{SLat}$-category is given by a $\catname{SLat}$-category $\mathbb{C}$ that additionally has
-	\begin{itemize}
-		\item a unit object $I \in \mathbb{C}$;
-		\item a $\catname{SLat}$-functor tensor $\otimes\colon \mathbb{C} \times \mathbb{C} \to \mathbb{C}$;
-	\end{itemize}
-	such that particular coherence diagrams analogous to those for symmetric monoidal categories commute.
+	A symmetric monoidal $\catname{SLat}$-category is given by a $\catname{SLat}$-category $\mathbb{C}$ that additionally has: a unit object $I \in \mathbb{C}$, an $\catname{SLat}$-functor tensor $\otimes\colon \mathbb{C} \times \mathbb{C} \to \mathbb{C}$, such that particular coherence diagrams analogous to those for symmetric monoidal categories commute.
 \end{definition}

### 2025-11-12 POPL2026/paper/sections/combinatorial_semantics.tex
moved all notes to a the same folder

@@
 	\label{def:dpoi}
-	Given a span of morphisms $\mathcal{G} \xleftarrow{} n \sqcup m \xrightarrow{} \mathcal{H}$ in $\catname{Hyp}(\Sigma)$,  we say $\mathcal{G}$ rewrites to $\mathcal{H}$ with interface $n \sqcup m$ via rewrite rule $\mathcal{L} \xleftarrow{} i \sqcup j \xrightarrow{} \mathcal{R}$ if there exists an object $\mathcal{C}$ and morphisms which complete the rightmost commutative diagram in~\autoref{fig:dpo_dpoi} such that the two marked squares are pushouts.
+	Given a span of morphisms $\mathcal{G} \xleftarrow{} n \sqcup m \xrightarrow{} \mathcal{H}$ in $\catname{Hyp}(\Sigma)$,  we say $\mathcal{G}$ rewrites to $\mathcal{H}$ with interface $n \sqcup m$ via rewrite rule $\mathcal{L} \xleftarrow{} i \sqcup j \xrightarrow{} \mathcal{R}$ if there exists an object $\mathcal{L}^{\bot}$ and morphisms which complete the rightmost commutative diagram in~\autoref{fig:dpo_dpoi} such that the two marked squares are pushouts.
 

### 2025-11-12 notes/SlottedEGraphs/SlottedNotes.tex
moved all notes to a the same folder

--- 2025-07-11 POPL2026/notes/SlottedNotes.tex
@@
 An example could be seen in Figure~\ref{fig:f_x_y_string_traced}.
+
+\subsection{Shape}
+
+A \textit{shape} is a pair $(e, \sigma)$ where $e$ is an e-node and $\sigma$ is a slot map for this particular e-node.
+The e-node $e$ can be thought of a canonicalized version of the e-node $\sigma(e)$ (i.e. slot map applied to $e$). The \textit{hashcons} data structure maps such canonicalized e-nodes to e-class ids.
+This is how non-duplication is maintained.
+Below is the code for computing the shape.
+In plain words, it takes an e-node $c$ and replaces all the slots in $c$ with canonical numerical slots $\$1, \$2$ etc..
+For example, given an e-node \mintinline{rust}{Var(Slot::named("x"))} (a code for $\texttt{var \$x}$ e-node), the canonicalized version will be $\texttt{var \$1}$ (or, \mintinline{rust}{Var(Slot::numeric(1))}).
+Now, to give a more complex example, suppose we were to add $\lambda x . x$ and $\lambda y . y$ expressions to a slotted e-graph.
+In code this would be 
+\begin{minted}{rust}
+  Lambda::Lam(Bind { 
+    slot: Slot::named("x"),
+    elem: AppliedId { id: Id(0), m: {"f0" -> "x"}}
+  })
+\end{minted}
+
+Where \mintinline{rust}{AppliedId { id: Id(0), m: {"f0" -> "x"}}} is the e-class id for the variable $x$ inside $\lambda x . x$.
+We get the same e-node for the $\lambda y . y$ with "x" replaced by "y".
+The slots of this e-node are the two occurrences of slot $x$ (one for the binder and another one from the values of the \mintinline{rust}{AppliedId}).
+Therefore, the first component of the shape will be
+
+\begin{minted}{rust}
+  Lambda::Lam(Bind { 
+    slot: Slot::numeric(1),
+    elem: AppliedId { id: Id(0), m: {"f0" -> 1}}
+  })
+\end{minted}
+
+and we get the same shape for $\lambda y . y$.
+
+\begin{listing}
+  \begin{minted}[linenos, fontsize=\small, breaklines]{rust}
+    fn weak_shape(&self) -> (Self, Bijection) {
+      let mut c = self.clone();
+      let mut m = SlotMap::new();
+      let mut i = 0;
+
+      for x in c.all_slot_occurrences_mut() {
+          let x_val = *x;
+          if !m.contains_key(x_val) {
+              let new_slot = Slot::numeric(i);
+              i += 1;
+
+              m.insert(x_val, new_slot);
+          }
+
+          *x = m[x_val];
+      }
+
+      let m = m.inverse();
+
+      let public = c.slots();
+      let m: SlotMap = m.iter().filter(|(x, _)| public.contains(x)).collect();
+
+      (c, m)
+  }
+  \end{minted}
+\end{listing}
 

### 2025-11-12 notes/category_theory/category_theory.tex
moved all notes to a the same folder

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%

\documentclass[acmsmall,screen, nonacm, anonymous]{acmart}

\newcommand*{\ONECOLUMN}{}

\input{../../preamble}
\input{../../macros}
\input{../../sample.tikzstyles}
\input{../../hypergraph.tikzstyles}
\input{../../hypergraph.tikzdefs}

\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\definecolor{americanrose}{rgb}{1.0, 0.01, 0.24}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}

% These are for comments
\newcommand\question[1]{{\color{azure}#1}}
\newcommand\update[1]{{\color{americanrose}#1}}

\usepackage{minted}
\usepackage{csquotes}

% type derivations
\usepackage{mathpartir}

\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

\settopmatter{printacmref=false}
\authorsaddresses{}

\begin{document}

\title{Notes on category theory}

\author{Anonymous}

\maketitle

\section{Yoneda Lemma}

The lemma~\cite{maclane} sometimes pops up in various proofs of category theory statements.

\begin{remark}

The following notation can be used for a set of natural transformations between two functors
\[
F,G : \mathbf{C} \to \mathbf{D}
\]

\[
\alpha : \Pi_{A : |\mathbf{C}|} \mathbf{D}(FA,GA)
\]
or,
\[
\int_{A : \catname{C}} \catname{D}(FA,GA)
\]

\end{remark}

\begin{lemma}[Yoneda]

Let $F : \catname{C}^{\text{op}} \to \catname{Set}$ be a contravariant functor to $\catname{Set}$ (i.e. a \textit{presheaf}).
Recalling that $\catname{C}(-, A)$ is a contravariant functor:
\begin{itemize}
\item $\catname{C}(-,A)B = \catname{C}(A,B)$
\item $\catname{C}(-,A)f : \catname{C}(D,A) \to \catname{C}(C,A) = g \circ f$ for $f : \catname{C}(C,D)$ and $g : \catname{C}(D,A)$
\end{itemize}

We can define a functor $Y : \catname{C} \to \catname{PShC}$ by letting
\begin{itemize}
\item $YA \coloneq \catname{C}(-,A)$
\item $Yf : \catname{C}(-,A) \to \catname{C}(-,B) \coloneq \int_{X : \catname{C}}\catname{C}(X,A) \to \catname{C}(X,B)$ for $f : \catname{C}(A,B)$, $(Yf)_{X}g = f \circ g$
\end{itemize}

Then we have for every contravariant functor $F : \catname{C}^{\text{op}} \to \catname{Set}$
\[
\int_{Y : \catname{C}}(YX)Y \to FY \cong FX
\]
or,
\[
\int_{Y : \catname{C}}\catname{C}(Y,X) \to FY \cong FX
\]
\end{lemma}

\begin{proof}

First let's define 
\[
\psi_{X} : (\int_{Y : \catname{C}}\catname{C}(Y,X) \to FY\;) \to FX
\]

as
\[
\psi_{X} \alpha \coloneq \alpha_{X} \circ id_{X}\;,
\]
i.e. each natural transformation $\alpha$ is mapped to $\alpha_X \circ id_{X}$ (into its component in $X$ applied to $id_{X}$)

and then
\[
\phi_{X} : FX \to (\int_{Y : \catname{C}}\catname{C}(Y,X) \to FY \;)
\]
as
\[
\phi_{X}\; x \coloneq \lambda f : Y \to X . (Ff x)\;.
\]
Let's check if these are the inverses of each other.
\begin{align*}
(\phi_{X};\psi_{X}) x &= \psi_{X}(\phi_{X} x)\\ 
 &= (F\;id_{X}) x\\
 &= id_{F_{X}} x\\
 &= x
\end{align*}

\begin{align*}
 (\psi_{X};\phi_{X}) \alpha &= \lambda f . (Ff) (\alpha_{X} \circ id_{X})\\
  &\overset{\text{naturality of } \alpha}{=} \lambda f . \alpha_{Y} \circ (\catname{C}(f,X) \circ id_{X})\\
  &= \lambda f . \alpha_{Y} \circ (id_{X} \circ f)\\
  &= \lambda f . \alpha_{Y} \circ f\\
  &= \alpha
\end{align*}
\end{proof}

\begin{corollary}
    The functor $Y : \catname{C} \to \catname{PShC}$ is full and faithful.
\end{corollary}
\begin{proof}

Using the lemma by plugging $Y$:
\[
\int_{Z : |\catname{C}|} \catname{C}(Z,X) \to YAZ \cong YAX
\]
or
\[
    \int_{Z : |\catname{C}|} \catname{C}(Z,X) \to \catname{C}(Z,A) \cong \catname{C}(X,A)
\]
We can note that the left-hand side is the action of $Y$ on morphisms of $\catname{C}$.
\end{proof}

To show that $f : \catname{C}(A,B)$ is an isomorphism we can check if $YA \cong YB$, \textit{i.e.}, if $\catname{C}(-,A) \cong \catname{C}(-,B)$ in a sense $\int_{X : |\catname{C}|} \catname{C}(X,A) \cong \catname{C}(X,B)$.

\begin{remark}
    The above is the contravariant version of the lemma.
\end{remark}

Recall the free-forgetful adjunction between $\catname{Mon}$ and $\catname{Set}$ where the free functor $F$ maps a set $X$ to a monoid $FX$ consisting of strings $x_1,x_2,\ldots,x_n$ and the monoidal multiplication is given by string concatenation.
Functor $U$ is then maps any given monoid $M$ to the set of its underlying elements $UM$.
We then have that for every set-morphism $X \to UM$ there is a unique monoid-morphism $FX \to M$.
\[
\catname{Mon}(FX, M) \cong \catname{Set}(X, UM)
\]
For example, let's take a monoid of natural numbers $\mathbb{N}$ and a set $X = \{a,b,c\}$.
We can define a function $f : \{a,b,c\} \to \mathbb{N}$ by $\{a \mapsto 1, b \mapsto 2, c \mapsto 3 \}$.
A free monoid over $\{a,b,c\}$ is $FX = \{\varepsilon,a,b,c,aa,bb,cc,\ldots\}$.
This induces a function $g : FX \to M$ by defining its action on generators $\{a \mapsto 1, b \mapsto 2, c \mapsto 3, \varepsilon \mapsto \varepsilon\}$ and then $g(aa) = g(a)\cdot g(a)$ and so on.

We can analogously define a similar adjunction between $\catname{Cat}$ and $\catname{Graph}$ (a directed graph is given by a set of vertices and edges such that for every edge we have a source and a target vertex).
Every category $\catname{C}$ defines a graph $U\catname{C}$ with the same objects and arrows forgetting which are composites and which are identities.
Let an $O$-graph be a graph over a set of vertices $O$.
A morphism of two $O$-graphs is a function that is identity on vertices.
Any $O$-graph can be used to generate a category $C$ on the same set of objects where the morphisms are all `strings' of composable arrows of the graph.

\begin{theorem}

Given a small graph $G = \{A \rightrightarrows O\}$.
There is a small category $\catname{C}$ with objects $O$ and a morphism $P : G \to U\catname{C}$ of graphs with the following property.
Given a category $\catname{B}$  and any morphism $D : G \to U\catname{B}$ there is a unique functor $D' : \catname{C} \to \catname{B}$ with $(U D') \circ P = D$.
In that sense $P$ is `universal' among the morphisms $G \to U\catname{C}$.
\end{theorem}
\begin{proof}

Let the objects of $\catname{C}$ be the ones of $\catname{G}$ and the arrows be the finite paths $a_1 \xrightarrow{f_1} a_2 \xrightarrow{} \ldots \xrightarrow{f_{n-1}} a_{n}$.
The desired morphism $P : G \to U\catname{C}$ is identity on objects and sends an arrow $a_1 \xrightarrow{f_{1}} a_2$ of $G$ to $\langle a_1, f_1, a_{2} \rangle$ of $U\catname{C}$.
That is, $\catname{C}$ is a free category over $G$.
Now, consider any other morphism $D : G \to U\catname{B}$ for a category $\catname{B}$, such that there is a functor $D' : \catname{C} \to \catname{B}$.
If $(UD') \circ P = D$, then $D_{O}(G) = D'_{O}(C)$ and $D'\langle a_{1}, f_{1}, a_{2} \rangle = Df_{1}$. 
\end{proof}
\begin{example}
    \item When a graph $G$ is given by a single vertex $a$ and a single arrow $a \xrightarrow{f} a$, the morphisms of $FG = \catname{C}$ are $1, f, f^2, \ldots$ (i.e. $\langle a \rangle, \langle a , f, a \rangle, \langle a f a f a \rangle, \ldots$)
\end{example}

\begin{definition}

If $S : \catname{D} \to \catname{C}$ is a functor and $c$ is an object of $\catname{C}$, a \textit{universal} arrow from $c$ to $S$ is a pair $\langle r, u \rangle$ consisting of an object $r$ of $\catname{D}$ and an arrow $u : c \to Sr$ of $\catname{C}$ such that to every pair $\langle d, f \rangle$ with $d$ an object of $\catname{D}$ and $f : c \to Sd$ there is a unique arrow $f' : r \to d$ such that $Sf' \circ u = f$.
That is, every arrow $f'$ is factorized uniquely through $u$.
\end{definition}
\begin{example}

Take $U : \catname{Cat} \to \catname{Graph}$ and let $c$ be an object of $\catname{Graph}$.
The universal morphism is then $\langle r, u \rangle$ where $r = Fc$ and $u : c \to U(Fc)$.
Then, given any other $\langle d, f \rangle$, where $f : c \to Ud$, we have that there exists $j : Fc \to d$ such that $f = u \circ Uj$
\end{example}

\begin{definition}[$T$-algebra]

Given a monad $\langle T : X \to X, \eta : I \to T, \mu : T \circ T \to T \rangle$, a $T$-algebra is given by a pair $\langle x, h \rangle$ where $h : Tx \to x$ and $x \in X$, such that
\[
    \begin{tikzcd}
        T^2x \arrow[r, "Th"] \arrow[d, "\mu_{x}"'] & Tx \arrow[d, "h"] & x \arrow[r, "\eta_{x}"] \arrow[rd, no head, Rightarrow] & Tx \arrow[d, "h"] \\
        Tx \arrow[r, "h"]                          & x                 &                                                         & x                
    \end{tikzcd}
\]
A morphism between two $T$-algebras $\langle x , h \rangle$ and $\langle y, h' \rangle$ is given by a morphism $f : x \to y$ such that $h' \circ Tf = f \circ h$.
\end{definition}

\begin{theorem}[Every monad is defined by its $T$-algebra]
If $\langle T, \mu, \eta \rangle$ is a monad on $X$, then the set of $T$-algebras and their morphisms forms a category $X^T$ and there is an adjunction
\[
\langle F^T, G^T, \eta^{T}, \varepsilon^{T} \rangle : X \to X^{T}
\], where the functors are given by
\[
    \begin{tikzcd}
        {\langle x, h \rangle} \arrow[r, "{G^{T}\langle x, h \rangle}"] \arrow[d, "f"] & x \arrow[d, "G^{T}f"] & x \arrow[r, "F^{T}x"] \arrow[d, "f"'] & {\langle Tx, \mu_{x} \rangle} \arrow[d, "F^{T}f"] \\
        {\langle y, h' \rangle} \arrow[r, "{G^{T}\langle y, h \rangle}"]               & y                     & y \arrow[r, "F^{T}y"]                 & {\langle Ty, \mu_{y} \rangle}                   
    \end{tikzcd}
\]
and $\eta^{T} = \eta : I \to G^{T} \circ F^{T}$ and $\varepsilon^{T}\langle x, h \rangle = h$.
Recall the typing $\varepsilon : F^{T} \circ G^{T} \to I$, or, component-wise $\varepsilon \;\langle F^T \circ G^T \rangle \langle x,h \rangle = \langle x,h \rangle$, $\varepsilon \;\langle Tx, \mu_{x}\rangle = \langle x,h \rangle$.
That is, the co-unit is the morphism between $T$-algebras $\langle Tx, \mu_{x} \rangle$ and $\langle x, h \rangle$.
That is, this is just the morphism between the carriers $Tx \to x$, i.e. $h$.
The monad defined by this adjunction is the given monad $T$.
First, let's check if the triangle equations hold
\[
    \begin{tikzcd}
        F \arrow[r, "F\eta"] \arrow[rd, no head, Rightarrow] & F \circ G \circ F \arrow[d, "\varepsilon_{F}"] &  & G \arrow[r, "\eta_{G}"] \arrow[rd, no head, Rightarrow] & G \circ F \circ G \arrow[d, "G\varepsilon"] \\
                                                             & F                                              &  &                                                         & G                                          
    \end{tikzcd}
\]
We get
\[
\begin{tikzcd}
        {Fx = \langle Tx, \mu_{x} \rangle} \arrow[rr, "F\eta"] \arrow[rd, no head, Rightarrow] &                               & {F(Tx) = \langle TTx, \mu_{Tx} \rangle} \arrow[ld, "\varepsilon_{F}"] &  & {G\langle x, h \rangle = x} \arrow[r, "\eta_{G}"] \arrow[rd, no head, Rightarrow] & Tx \arrow[d, "G_{\varepsilon}"] \\
                                                                                               & {\langle Tx, \mu_{x} \rangle} &                                                                       &  &                                                                                   & x                              
\end{tikzcd}
\]
The right-most arrow of the right-hand diagram is given by $\varepsilon_{\langle x, h \rangle} = h : Tx \to x$, i.e. the morphism $\langle Tx, \mu_{x} \rangle \to \langle x, h \rangle$ and $G\varepsilon : Tx \to x$.
The endofunctor $(G^{T} \circ F^{T})x = G^{T}(\langle Tx, \mu_{x} \rangle) = Tx$ defines the original monad $T$.
The unit of the monad is $\eta$ and the join $\mu\;x : TTx \to Tx$ is given by $G^{T}\varepsilon F^{T}\;x = G^{T}\varepsilon(\langle Tx, \mu_{x}\rangle) = G^{T}\mu_{x} = \mu_{x}$ ($G^{T}$ just `lifts' $\mu_{x}$ from homomorphism of $T$-algebras to a plain morphism in $X$).
\end{theorem}

\begin{theorem}[Comparison of adjunctions with algebras]

Given an adjunction $X \to A$, $\langle F, G, \eta, \epsilon \rangle$ and a monad on $X$ given by $\langle G \circ F, \eta, G\varepsilon F \rangle$.
Then there is a unique functor $K : A \to X^{T}$ such that $G^{T}K = G$ and $KF = F^{T}$, that is the following diagram commutes
\[
\begin{tikzcd}
        A \arrow[r, "K", dashed] \arrow[d, "G", shift left=2] & X^{T} \arrow[d, "G^{T}", shift left] \\
        X \arrow[u, "F"] \arrow[r, no head, Rightarrow]       & X \arrow[u, "F^{T}", shift left]    
\end{tikzcd}
\]
\end{theorem}
\begin{proof}
 The co-unit of the adjunction gives us the morphism $G\varepsilon_{a} : GFGa \to Ga$ and the arrow may be considered as the evaluation map of the corresponding $T$-algebra $\langle Ga, G\varepsilon_{a} \rangle$.
 This is a valid $T$-algebra by checking that the necessary diagrams commute.
 We define $Ka = \langle Ga, G\varepsilon_{a} \rangle$ and $Kf : \langle Ga, G\varepsilon_{a} \rangle \to \langle Gb, G\varepsilon_{b} \rangle$.
 Now we can check, \textit{e.g.}, that $KF = F^{T}$, $KFa = \langle GFa, G\varepsilon_{Fa} \rangle$, $F^{T}a = \langle GFa, \mu_{a}\rangle$. Recalling that $\mu_{a} = G\varepsilon_{Fa}$ we get the equality.
\end{proof}

\begin{definition}

When the functor $K$ is an isomorphism, we say that $G$ is \textit{monadic}.
\end{definition}

\begin{example}

Consider a free-forgetful adjunction $\langle F,G,\eta,\varepsilon \rangle : \catname{Set} \to \catname{SmGrp}$.
$FX$ is defined by turning a set $X$ into a free semigroup, i.e. $x \mapsto \langle x \rangle$ where the multiplication is given by words juxtaposition.
The forgetful functor simply forgets the multiplication and maps $\langle x \rangle \mapsto x$.
$\eta : X \to GFX$ sends each $x$ to one-letter word $x$.
The co-unit $\varepsilon : FGS \to S$, the composite $G\varepsilon \circ \eta_{G}(S) : GS \to GFGS \to GS$ is then the identity, in other words $\varepsilon$ maps generators.
For example, let $X = \{a,b,c\}$, $FX$ is then $\{\langle a \rangle, \langle b \rangle, \langle c \rangle, \langle aa \rangle, \ldots\}$.
$GFX = \{a,b,c,aa,\ldots\}$ and $\eta$ maps $a \mapsto a, b \mapsto b, c \mapsto c$.
Given $S = \{\langle s \rangle, \langle ss \rangle, \langle sss \rangle, \ldots\}$, $GS = \{s, ss, sss, \ldots\}$, $FGS = \{\langle s \rangle, \langle ss \rangle, \langle sss \rangle, \ldots\}$ and $\varepsilon$ just maps $\langle s \rangle \mapsto \langle s \rangle$ etc.
In other words $\varepsilon(\langle s_1 \rangle \ldots \langle s_n \rangle) = s_1 \cdot \ldots \cdot s_n$.

The induced monad $T = GF$ such that $TX = \coprod_{n=1}^{\inf}X^{n}$, $\eta_{x} = \langle x \rangle$, 
\[
\mu = G\varepsilon_{F}(\langle \langle x_{11} \rangle, \ldots, \langle x_{1,n_1} \rangle, \ldots, \langle \langle x_{k,1}\rangle, \ldots, \langle x_{k,n_{k}} \rangle \rangle) = \langle x_{1,1}\rangle \ldots \langle x_{k,n_{k}} \rangle~.
\]

For the above monad the algebras are $\langle S, v \rangle$ where $v : TS \to S$ i.e. $v : \coprod S^{n} \to S$ which is given by $\coprod_{1}^{\inf} v^{n}$.
The law $v\eta_{x} = id$ implies that $v_{1} = id$.
We also have that $v_{k}(v_{n_1} \times \ldots \times v_{n_k}) = v_{n_{1} + \ldots n_{k}} : S^{n_{1} + \ldots + n_{k}} \to S$, i.e. applying $k$ operations to a list of length $k$ that consists of lists of lengths $n_1$ to $n_k$ is the same as applying the operation to a list of length $n_1 + \ldots + n_{k}$.

The isomorphism $K : \catname{SmGrp} \to \catname{Set}^{T}$ is then given by $\langle S, v \rangle \mapsto \langle S, v_1, v_2, \ldots \rangle$ where $v_n$ is the iterated version of $v$.
\end{example}

\begin{example}

Consider now the adjunction $\langle F, G, \eta, \varepsilon \rangle : \catname{Graph} \to \catname{Cat}$.
The corresponding $T$-algebra for a monad $GF$ is $\langle Tx, \mu_{x}$ where $T$ adds identity paths and all other finite paths to a graph via designated edges.
The functor $K : \catname{Cat} \to \catname{Graph}^{T}$ maps a category $\catname{C}$ to $\langle G\catname{C}, h : GFG\catname{C} \to G\catname{C} = G\varepsilon_{C}\rangle$.
Given a category $\catname{C}$, $G\catname{C}$ is a graph with edges generated by morphisms of $\catname{C}$, then, $FG\catname{C}$ is a category where morphisms are all the finite paths of $G\catname{C}$ where each edge in the path is induced by a morphism in $\catname{C}$ and the whole path is given by the composition of morphisms, therefore we can map each path to a composition of the morphisms.
This defines the action of $\varepsilon$.
This is carried to the morphism of the respective underlying graphs and defines the structural map $h$.
Note that not any graph can be a component of a $T$-algebra.
In particular, it should be such graph $X$ that $TX \to X$ is definable, i.e. there must be an edge in $X$ for every finite path in $X$.
This means that there is a corresponding category for $X$ in $\catname{Cat}$ that has the same objects and morphisms are edges.
Since a graph has an edge for every path, it means that there is a composite morphism for any other two morphisms and the composition is associative.
\end{example}

\begin{definition}[Exponentials]
    In a category $\catname{C}$ with binary products we say there are \textit{exponentials} if for all objects $A$ $B$ there exist an object $A \multimap B$ and a map $ev_{A,B} : A \multimap B \times A \to B$ such that for any $g : C \times A \to B$ there is a unique morphism $\Lambda(g) : C \to (A \multimap B)$ such that the following diagram commutes
    \[
    \begin{tikzcd}
        A \multimap B \times A \arrow[r, "{ev_{A,B}}"]               & B \\
        C \times A \arrow[u, "\Lambda(g) \times id"] \arrow[ru, "g"] &  
    \end{tikzcd}
    \]
    We also have $\Lambda(ev_{A,B} \circ h \times id_{A}) = h$ for $h : C \to (A \multimap B)$.
    $h \times id_{A} : C \times A \to (A \multimap B) \times A$, $ev_{A,V} \circ h \times id_{A} : C \times A \to B$ and $\Lambda(ev_{A,B} \circ h \times id_{A}) : C \to A \multimap B = h$ by uniqueness.
\end{definition}

\bibliographystyle{ACM-Reference-Format}
\bibliography{../../bibliography}

\end{document}
